### 2017-02-14 r_scripts/oregon_clean.rmd
Create oregon_clean.rmd

---
title: "Oregon_Wildfire_Claim"
author: "Jingyang Liu"
date: "February 14, 2017"
output: html_document
---

## Explore Data Structure and Cleaning

The Oregon data contains 77069320 obsversations in total in year of 2013, which is a big data set. We want to figure out the structure for this data set and do some basic cleaning.

```{r, echo = F}
library(dplyr)
library(data.table)

# import data ------------------------------------------------------------------
read_path <- paste0("../data/gan_episodes_of_care.txt")

start_time <- Sys.time()
oregon_full <- fread(read_path, sep = "|", showProgress = T) # not change class
stop_time <-  Sys.time() - start_time 
# time it took
stop_time # 7.771713 mins

# Explore data structure
summary(as.factor(oregon_full$clmstatus))

# Clean data
oregon_full_df <- oregon_full %>%
  # filter non-missing value of gender, race 
  filter (gender!='U') %>%
  filter (is.na(race)==F)

# filter dx1, no ndc
oregon_full_dx1 <- oregon_full_df %>%
  filter (dx1!='*NULL*')

# filter ndc, no dx1
oregon_full_ndc <- oregon_full_df %>%
  filter (ndc!="")

fwrite(oregon_full_dx1, "../data/oregon_full_dx1.csv")
fwrite(oregon_full_ndc, "../data/oregon_full_ndc.csv")

write_path1 <- paste0('../data/','oregon_full_dx1.csv')
write_csv(oregon_full_dx1, write_path1)

write_path2 <- paste0('../data/','oregon_full_ndc.csv')
write_csv(oregon_full_ndc, write_path2)

```

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2017-02-15 r_scripts/oregon_clean.rmd
Update oregon_clean.rmd
reduced data set r markdown...needs improving

@@
 ---
-title: "Oregon_Wildfire_Claim"
+title: "Oregon_Wildfire_reduced"
 author: "Jingyang Liu"
-date: "February 14, 2017"
+date: "February 15, 2017"
 output: html_document
@@
 
-```{r, echo = F}
+```{r}
 library(dplyr)
@@
 # import data ------------------------------------------------------------------
-read_path <- paste0("../data/gan_episodes_of_care.txt")
+# Read in first 2000000 lines for exploration.
 
+# read_path <- paste0("./data/gan_episodes_of_care.txt")
+# oregon_df <- fread(read_path, sep = "|", nrows = 200000, showProgress = T)
+# fwrite(oregon_df, "./data/oregon_epis_care_reduced.csv")
+
+# I wrote a reduced dataframe of 2000,000 rows to get an idea of data structure
+
+# import the reduced dataframe
 start_time <- Sys.time()
-oregon_full <- fread(read_path, sep = "|", showProgress = T) # not change class
+
+oregon_df <- fread("./data/oregon_epis_care_reduced.csv", sep = ",")
+
 stop_time <-  Sys.time() - start_time
+
 # time it took
-stop_time # 7.771713 mins
-
-# Explore data structure
-summary(as.factor(oregon_full$clmstatus))
-
-# Clean data
-oregon_full_df <- oregon_full %>%
-  # filter non-missing value of gender, race 
-  filter (gender!='U') %>%
-  filter (is.na(race)==F)
-
-# filter dx1, no ndc
-oregon_full_dx1 <- oregon_full_df %>%
-  filter (dx1!='*NULL*')
-
-# filter ndc, no dx1
-oregon_full_ndc <- oregon_full_df %>%
-  filter (ndc!="")
-
-fwrite(oregon_full_dx1, "../data/oregon_full_dx1.csv")
-fwrite(oregon_full_ndc, "../data/oregon_full_ndc.csv")
-
-write_path1 <- paste0('../data/','oregon_full_dx1.csv')
-write_csv(oregon_full_dx1, write_path1)
-
-write_path2 <- paste0('../data/','oregon_full_ndc.csv')
-write_csv(oregon_full_ndc, write_path2)
+stop_time # 15.83603 secs
 
@@
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+#  The variable "clmstatus" is ..... It is a binary
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
+```{r claim status}
 
-```{r cars}
-summary(cars)
+# Explore data structure, try like clmstatus
+summary(oregon_df)
+summary(as.factor(oregon_df$clmstatus))
+
 ```
 
-## Including Plots
+## Clean data
 
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+```{r }
+# Clean data, 76986701 left (maybe later)
+#oregon_full_df <- oregon_full %>%
+  # filter non-missing value of year of birth, gender, race, STATE
+#  filter (is.na(yob)==F) %>%
+#  filter (gender!='U') %>%
+#  filter (is.na(race)==F) %>%
+#  filter (STATE!='')
+
+# Explore no dx1 and no ndc, see if the dx2 or other reasons, 1042984 obs
+# oregon_full_miss <- oregon_full %>%
+#  filter (dx1=='*NULL*') %>%
+#  filter (ndc=="")
+
+# fwrite(oregon_full_miss, "../data/oregon_full_miss.csv")
+
+# read_path1 <- paste0("./data/oregon_full_miss.csv")
+# oregon_miss <- fread(read_path1, sep = ",")
+
+# oregon_miss_df <- oregon_miss %>%
+#   filter (dx2!='*NULL*') # 981 obs
+
+# filter dx1, no ndc, 52490460
+# oregon_full_dx1 <- oregon_full_df %>%
+#   filter (dx1!='*NULL*')
+
+# filter ndc, no dx1, 23453379
+# oregon_full_ndc <- oregon_full_df %>%
+#  filter (ndc!="")
+
+# fwrite(oregon_full_dx1, "../data/oregon_full_dx1.csv")
+# fwrite(oregon_full_ndc, "../data/oregon_full_ndc.csv")
+
+```

### 2017-02-16 r_scripts/oregon_clean.rmd
Update oregon_clean.rmd
Compiling data exploring r markdown

@@
 
-## Explore Data Structure and Cleaning
+# Explore Data Structure and Cleaning
 
@@
 library(data.table)
+getwd()
 
@@
 
-# read_path <- paste0("./data/gan_episodes_of_care.txt")
-# oregon_df <- fread(read_path, sep = "|", nrows = 200000, showProgress = T)
-# fwrite(oregon_df, "./data/oregon_epis_care_reduced.csv")
+read_path <- paste0("../data/gan_episodes_of_care.txt")
+oregon_df <- fread(read_path, sep = "|", nrows = 200000, showProgress = T)
+fwrite(oregon_df, "../data/oregon_epis_care_reduced.csv")
 
@@
 start_time <- Sys.time()
-
-oregon_df <- fread("./data/oregon_epis_care_reduced.csv", sep = ",")
-
+oregon_df <- fread("../data/oregon_epis_care_reduced.csv", sep = ",")
 stop_time <-  Sys.time() - start_time
@@
 # time it took
-stop_time # 15.83603 secs
+stop_time # 1.968219 secs
 
@@
 
-#  The variable "clmstatus" is ..... It is a binary
+## The overall summary  of data set
+72 variables in total.
+```{r overall summary}
 
+# Explore data structure
+summary(oregon_df)
+```
+
+## *personkey*, Unique person identifier
+person's ID, maximum is 14381640
+numeric; length: 8
+```{r personkey}
+summary(as.factor(oregon_df$personkey))
+min(oregon_df$personkey)
+max(oregon_df$personkey) # seems length is 8
+```
+
+## *year*, Calendar year in which claim was incurred
+All is 2013
+integer (numeric); length: 4
+```{r year}
+summary(as.factor(oregon_df$year)) # all 2013, 77069320
+```
+
+## *clmid*, Claim ID
+numeric; length: 8
+```{r claim id}
+summary(as.factor(oregon_df$clmid))
+```
+
+## *line*, Claim line (Not Sure what this means)
+numeric; range: 0-9999
+```{r claim line}
+unique(as.numeric(oregon_df$line)) # 1362 levels, 0 - 9999
+```
+
+## *clmtatus*, Claim status
+D: 8676643 (11.26%), E: 3324496 (4.31%), P: 65068181 (84.43%)
+character
 ```{r claim status}
+summary(as.factor(oregon_df$clmstatus))
+#       D        E        P
+# 8676643  3324496 65068181
 
-# Explore data structure, try like clmstatus
-summary(oregon_df)
-summary(as.factor(oregon_df$clmstatus))
+summary(as.factor(oregon_df$clmstatus))/77069320 # percentage
+#          D          E          P
+# 0.11258232 0.04313644 0.84428124
+```
 
+## *cob*, COB status (Coordination of benefits) 
+blank: 23517978 (30.52%); N: 44438688 (57.66%); Y: 9112654 (11.82%)
+character
+``` {r cob}
+summary(as.factor(oregon_df$cob)) 
+#                 N        Y
+# 23517978 44438688  9112654
+summary(as.factor(oregon_df$cob))/77069320 # percentage
+#                   N         Y
+# 0.3051536 0.5766067 0.1182397
+```
+
+## *paytype*, Pay type
+C (58.50%), D (31.77%), G (2.08%), P (3.53%), T (4.12%)
+character
+``` {r pay type}
+summary(as.factor(oregon_df$paytype)) #C, D, G, P, T
+summary(as.factor(oregon_df$paytype))/77069320 # percentage
+#         C          D          G          P          T
+# 0.58498296 0.31771349 0.02081743 0.03532560 0.04116052
+```
+
+## *prod*, Product code
+CHP, etc. Insurance type/product code [web](https://ushik.ahrq.gov/ViewItemDetails?&system=apcd&itemKey=194604000)
+character
+```{r product code}
+summary(as.factor(oregon_df$prod))
+```
+
+## *payer*, APAC payer 
+All Payer All Claims [web](https://www.oregon.gov/oha/analytics/Pages/All-Payer-All-Claims.aspx)
+character
+```{r payer}
+summary(as.factor(oregon_df$payer))
+```
+
+## *megflag*, Medical coverage flag
+The medical coverage flag indicates whether this member is covered for medical expenses [web](https://ushik.ahrq.gov/ViewItemDetails?&system=apcd&itemKey=196401000)
+0: (7.39%); 1: (91.57%); NA: (1.04%)
+numeric; range: 0-1
+```{r medical coverage flage}
+summary(as.factor(oregon_df$medflag))
+#       0        1     NA's
+# 5697342 70573441   798537
+summary(as.factor(oregon_df$medflag))/77069320 # percentage
+#          0          1       NA's
+# 0.07392490 0.91571381 0.01036128
+```
+
+## *rxflag*, Pharmacy coverage flag
+This field indicates whether or not the member has prescription drug coverage. 
+[web](https://ushik.ahrq.gov/ViewItemDetails?&system=apcd&itemKey=196401000)
+0: (10.07%); 1: (88.89%); NA: (1.04%)
+numeric; range: 0-1
+```{r pharmacy converage flag}
+summary(as.factor(oregon_df$rxflag))
+#       0        1     NA's
+# 7761348 68509435   798537
+summary(as.factor(oregon_df$rxflag))/77069320 # percentage
+#          0          1       NA's
+# 0.10070607 0.88893265 0.01036128
+```
+
+## *ohvmhflag*, Oregon High value medical home flag
+Required for participants in OHLC high value medical home initiative. 
+[web](https://ushik.ahrq.gov/ViewItemDetails?&system=apcd&itemKey=194764000&enableAsynchronousLoading=true)
+N: (69.15%); Y: (0.24%); NULL: (30.61%)
+character
+```{r Oregon High value medical home flag}
+summary(as.factor(oregon_df$ohvmhflag))
+#        N   *NULL*        Y
+# 53291284 23589431   188605
+summary(as.factor(oregon_df$ohvmhflag))/77069320 # percentage
+#           N      *NULL*           Y
+# 0.691472093 0.306080695 0.002447212
+```
+
+## *pebb*, PEBB flag (NOT Provided)
+0: (95.95%); 1: (4.05%)
+numeric; range: 0-1
+```{r pebb flag}
+summary(as.factor(oregon_df$pebb))
+#        0        1
+# 73949154  3120166
+summary(as.factor(oregon_df$pebb))/77069320 # percentage
+#          0          1
+# 0.95951481 0.04048519
+```
+
+## *oebb*, OEBB flag (NOT Provided)
+0: (96.39%); 1: (3.61%)
+numeric; range: 0-1
+```{r oebb flag}
+summary(as.factor(oregon_full$oebb))
+#        0        1
+# 74285962  2783358
+summary(as.factor(oregon_full$oebb))/77069320 # percentage
+#          0          1
+# 0.96388501 0.03611499
+```
+
+## *patid*, Encrypted patient ID
+numeric
+```{r Encrypted patient ID}
+summary(as.factor(oregon_full$patid))
+```
+
+## *gender*, Gender
+F: (60.56%); M: (39.38%); U (for unknown): (0.06%)
+character
+```{r gender}
+summary(as.factor(oregon_full$gender))
+#        F        M        U
+# 46672307 30350771    46242
+summary(as.factor(oregon_full$gender))/77069320 # percentage
+#            F            M            U
+# 0.6055886700 0.3938113247 0.0006000053
+```
+
+## *yob*, Birth year
+old:1901 (116 years) young:2016 (1 year)
+numeric
+```{r birth year}
+summary(as.factor(oregon_full$yob)) # old:1901 young:2016
+```
+
+## *race*, Race (Not sure the category means)
+1:(31%); 2: (2%); 3: (0.8%); 4: (0.3%); 5: (0.1%); 6: (0.3%); 9: (65%); NA: (0.03%)
+numeric; range: 1-9
+```{r race}
+summary(as.factor(oregon_full$race))
+#        1        2        3        4        5        6        9     NA's
+# 24132190  1611193   620854   203550    85465   248416 50141889    25763
+summary(as.factor(oregon_full$race))/77069320 # percentage
+#            1            2            3            4            5            6
+# 0.3131231727 0.0209057638 0.0080557867 0.0026411288 0.0011089367 0.0032232800
+#            9         NA's
+# 0.6506076478 0.0003342835
+```
+
+## *ethn*, Ethnicity (Not sure the category means)
+1:(0.76%); 2: (9.60%); 3: (89.61%); NA: (0.03%)
+numeric; range: 1-3
+```{r ethnicity}
+summary(as.factor(oregon_full$ethn))
+#        1        2        3     NA's
+#  583948  7398053 69061589    25730
+summary(as.factor(oregon_full$ethn))/77069320 # percentage
+#            1            2            3         NA's
+# 0.0075769191 0.0959921925 0.8960970332 0.0003338553
+```
+
+## *lang*, Primary spoken language 
+Not sure the category means. This web [web](http://www.niso.org/topics/ccm/ccmstandards/) might help.
+1:(47.70%); 2: (1.34%); 3: (0.001%); 4: (0.26%); 8: (0.65%); 9: (49.31%); NA: (0.74%)
+numeric; range: 1-9
+```{r language}
+summary(as.factor(oregon_full$lang))
+#        1        2        3        4        8        9     NA's
+# 36761368  1031059      587   203152   501575 38002063   569516
+round(summary(as.factor(oregon_full$lang))/77069320, 5)  # percentage
+#       1       2       3       4       8       9    NA's
+# 0.47699 0.01338 0.00001 0.00264 0.00651 0.49309 0.00739
+```
+
+## *MSA*, Metropolitan and Micropolitan Statistical Areas Main???
+numeric
+```{r MSA}
+summary(as.factor(oregon_full$MSA))
+```
+
+## *STATE*, Standard two character abbreviation
+OR: (99.33%); blank: (0.1%)
+character
+```{r State}
+summary(as.factor(oregon_full$STATE)) # OR: 76555225; blank: 78646
+76555225/77069320
+```
+
+## *ZIP*, Zip code
+Freely available in the public domain. NA: (0.1%)
+numeric
+```{r ZIP code}
+summary(as.factor(oregon_full$ZIP)) # NA: 78646
+```
+
+## *fromdate*, From date
+YYYY-MM-DD
+character
+```{r from date}
+summary(as.factor(oregon_full$fromdate))
+unique(as.factor(oregon_full$fromdate))
+```
+
+## *todate*, To date
+YYYY-MM-DD
+character
+```{r to date}
+summary(as.factor(oregon_full$todate))
+```
+
+## *paydate*, Pay date
+YYYY-MM-DD
+character
+```{r pay date}
+summary(as.factor(oregon_full$paydate))
+```
+
+## *paid*, Total payment
+NA: (0.36%)
+numeric
+```{r total payment}
+summary(as.factor(oregon_full$paid)) # NA: 279811
+```
+
+## *tob*, Type of bill
+blank: (52.61%); NULL: (25.12%)
+character
+```{r type of bill}
+summary(as.factor(oregon_full$tob)) # blank: 40545960 NULL: 19358510 
+```
+
+## *pos*, Place of service code
+Not sure. [web](https://www.cms.gov/Medicare/Coding/place-of-service-codes/Place_of_Service_Code_Set.html) migh help.
+NULL: (9.22%)
+character
+```{r place of servis code}
+summary(as.factor(oregon_full$pos)) # NULL: 7106719
+```
+
+## *dx1*, Principal diagnosis (For ICD9 code)
+[web](https://www.cms.gov/Medicare/Coding/ICD9ProviderDiagnosticCodes/codes.html)
+NULL: (31.81%)
+character
+```{r dx1}
+summary(as.factor(oregon_full$dx1)) # NULL: 24513232
+```
+
+## *dx2*, Diagnosis 2
+NULL: (59.31%)
+character
+``` {r dx2}
+summary(as.factor(oregon_full$dx2)) # NULL: 45707452
+```
+
+## *dx3*, Diagnosis 3
+NULL: (72.67%)
+character
+``` {r dx3}
+summary(as.factor(oregon_full$dx3)) # NULL: 56005306
+```
+
+## *dx4*, Diagnosis 4
+NULL: (81.46%)
+character
+``` {r dx4}
+summary(as.factor(oregon_full$dx4)) # NULL: 62777450
+```
+
+## *dx5*, Diagnosis 5
+NULL: (93.42%)
+character
+``` {r dx5}
+summary(as.factor(oregon_full$dx5)) # NULL: 71995284
+```
+
+## *dx6*, Diagnosis 6
+NULL: (95.23%)
+character
+``` {r dx6}
+summary(as.factor(oregon_full$dx6)) # NULL: 73390409
+```
+
+## *dx7*, Diagnosis 7
+NULL: (96.31%)
+character
+```{r dx7}
+summary(as.factor(oregon_full$dx7)) # NULL: 74225796
+```
+
+## *dx8*, Diagnosis 8
+NULL: (97.00%)
+character
+```{r dx8}
+summary(as.factor(oregon_full$dx8)) # NULL: 74759988
+```
+
+## *dx9*, Diagnosis 9
+NULL: (98.55%)
+character
+```{r dx9}
+summary(as.factor(oregon_full$dx9)) # NULL: 75949570
+```
+
+## *dx10*, Diagnosis 10
+NULL: (98.81%)
+character
+```{r dx10}
+summary(as.factor(oregon_full$dx10)) # NULL: 76151124
+```
+
+## *dx11*, Diagnosis 11
+blank: (98.35%)
+character
+``` {r dx11}
+summary(as.factor(oregon_full$dx11)) # blank: 75801079
+```
+
+## *dx12*, Diagnosis 12
+blank: (98.61%)
+character
+```{r dx12}
+summary(as.factor(oregon_full$dx12)) # blank: 76000733
+```
+
+## *dx13*, Diagnosis 13
+blank: (98.86%)
+character
+```{r dx13}
+summary(as.factor(oregon_full$dx13)) # blank: 76193335
+```
+
+## *poa1*, POA code 1
+N: (0.07%); Y: (5.94%); blank: (93.99%)
+character
+```{r poa1} 
+summary(as.factor(oregon_full$poa1))
+#                 N        Y
+# 72434455    58842  4576023
+summary(as.factor(oregon_full$poa1))/77069320 # percentage
+#                         N            Y
+# 0.9398610887 0.0007634945 0.0593754168
+```
+
+## *poa2*, POA code 2
+N: (0.34%); Y: (3.15%); blank: (96.51%)
+character
+```{r poa2}
+summary(as.factor(oregon_full$poa2))
+#                 N        Y
+# 74380115   261029  2428176
+summary(as.factor(oregon_full$poa2))/77069320 # percentage
+#                       N           Y
+# 0.965106673 0.003386938 0.031506389
+```
+
+## *poa3*, POA code 3
+N: (0.34%); Y: (2.75%); blank: (96.91%)
+character
+```{r poa3}
+summary(as.factor(oregon_full$poa3))
+#                 N        Y
+# 74688528   264287  2116505
+summary(as.factor(oregon_full$poa3))/77069320 # percentage
+#                       N           Y
+# 0.969108434 0.003429212 0.027462355
+```
+
+## *px1*, Principal inpt procedure
+blank: (52.61%); NULL: (45.38%)
+character
+```{r px1}
+summary(as.factor(oregon_full$px1)) # blank: 40545960; NULL: 34975994
+```
+
+## *px2*, Procedure 2
+blank: (52.61%); NULL: (46.06%)
+character
+```{r px2}
+summary(as.factor(oregon_full$px2)) # blank: 40545960 NULL: 35502262
+```
+
+## *px3*, Procedure 3
+blank: (52.61%); NULL: (45.38%)
+character
+```{r px3}
+summary(as.factor(oregon_full$px3)) # blank: 40545960; NULL: 34975994
+```
+
+## *proccode*, CPT or HCPCS procedure code
+Current Procedural Terminology (CPT) Healthcare Common Procedure Coding System (HCPCS) 
+[web1](https://www.cms.gov/Medicare/Coding/MedHCPCSGenInfo/index.html?redirect=/medhcpcsgenin)
+[web2](https://www.cms.gov/Medicare/Coding/HCPCSReleaseCodeSets/index.html?redirect=/hcpcsrelea)
+NULL: (34.88%)
+character
+```{r procedure ccode}
+summary(as.factor(oregon_full$proccode)) # NULL: 26880271
+```
+
+```{r}
+summary(as.factor(oregon_full$mod1))
+```
+
+```{r}
+summary(as.factor(oregon_full$mod2))
+```
+
+```{r}
+summary(as.factor(oregon_full$megcode))
+```
+
+```{r}
+summary(as.factor(oregon_full$megdesc))
+```
+
+```{r}
+summary(as.factor(oregon_full$megbodysys))
+```
+
+```{r}
+summary(as.factor(oregon_full$megstage))
+```
+
+```{r}
+summary(as.factor(oregon_full$megtype))
+#                   9 NOT-CATEGORIZED             Acute           Chronic
+#              1751           6979372          34062467          31197648
+#         Well Care
+#           4828082
+```
+
+```{r}
+summary(as.factor(oregon_full$megcomplete))
+#        0        1     NA's
+# 10090656 66976913     1751
+```
+
+```{r}
+summary(as.factor(oregon_full$megnum))
+```
+
+```{r}
+summary(as.factor(oregon_full$megdays))
+```
+
+```{r}
+summary(as.factor(oregon_full$megprorate))
+```
+
+```{r}
+summary(as.factor(oregon_full$megoutlier))
+#        0        1
+# 54675387 22393933
+```
+
+```{r}
+summary(as.factor(oregon_full$meglow))
+#    FALSE     TRUE
+# 75145616  1923704
+```
+
+```{r}
+summary(as.factor(oregon_full$meghigh))
+#    FALSE     TRUE
+# 56599091 20470229
+```
+
+``` {r}
+# ndc
+summary(as.numeric(oregon_full$ndc)) # blank: 53599072
+```
+
+```{r}
+summary(as.factor(oregon_full$rxclass))
+```
+
+```{r}
+summary(as.factor(as.character(oregon_full$qtydisp)))
+```
+
+```{r}
+summary(as.factor(oregon_full$rxdays))
+```
+
+```{r}
+summary(as.factor(oregon_full$daw))
+```
+
+```{r}
+summary(as.factor(oregon_full$ptstatus))
+```
+
+```{r}
+summary(as.factor(oregon_full$los))
+```
+
+```{r}
+summary(as.factor(oregon_full$msdrg))
+```
+
+```{r}
+summary(as.factor(oregon_full$icd_ver))
+#    FALSE     NA's
+# 52556088 24513232
 ```

### 2017-02-17 r_scripts/oregon_clean.rmd
data structure overview
data variable summary

--- 2017-02-15 r_scripts/oregon_clean.rmd
@@
 
-#  The variable "clmstatus" is ..... It is a binary
+## The overall summary  of data set
+72 variables in total.
+```{r overall summary}
 
+# Explore data structure
+summary(oregon_df)
+```
+
+## 1) *personkey*, Unique person identifier
+person's ID, maximum is 14381640
+numeric; length: 8
+```{r personkey}
+summary(as.factor(oregon_df$personkey))
+min(oregon_df$personkey)
+max(oregon_df$personkey) # seems length is 8
+```
+
+## 2) *year*, Calendar year in which claim was incurred
+All is 2013
+integer (numeric); length: 4
+```{r year}
+summary(as.factor(oregon_df$year)) # all 2013, 77069320
+```
+
+## 3) *clmid*, Claim ID
+numeric; length: 8
+```{r claim id}
+summary(as.factor(oregon_df$clmid))
+```
+
+## 4) *line*, Claim line (Not Sure what this means)
+numeric; range: 0-9999
+```{r claim line}
+unique(as.numeric(oregon_df$line)) # 1362 levels, 0 - 9999
+```
+
+## 5) *clmtatus*, Claim status
+D: 8676643 (11.26%), E: 3324496 (4.31%), P: 65068181 (84.43%)
+character
 ```{r claim status}
+summary(as.factor(oregon_df$clmstatus))
+#       D        E        P
+# 8676643  3324496 65068181
 
-# Explore data structure, try like clmstatus
-summary(oregon_df)
-summary(as.factor(oregon_df$clmstatus))
+summary(as.factor(oregon_df$clmstatus))/77069320 # percentage
+#          D          E          P
+# 0.11258232 0.04313644 0.84428124
+```
+
+## 6) *cob*, COB status (Coordination of benefits) 
+blank: 23517978 (30.52%); N: 44438688 (57.66%); Y: 9112654 (11.82%)
+character
+``` {r cob}
+summary(as.factor(oregon_df$cob)) 
+#                 N        Y
+# 23517978 44438688  9112654
+summary(as.factor(oregon_df$cob))/77069320 # percentage
+#                   N         Y
+# 0.3051536 0.5766067 0.1182397
+```
+
+## 7) *paytype*, Pay type
+C (58.50%), D (31.77%), G (2.08%), P (3.53%), T (4.12%)
+character
+``` {r pay type}
+summary(as.factor(oregon_df$paytype)) #C, D, G, P, T
+summary(as.factor(oregon_df$paytype))/77069320 # percentage
+#         C          D          G          P          T
+# 0.58498296 0.31771349 0.02081743 0.03532560 0.04116052
+```
+
+## 8) *prod*, Product code
+CHP, etc. Insurance type/product code [web](https://ushik.ahrq.gov/ViewItemDetails?&system=apcd&itemKey=194604000)
+character
+```{r product code}
+summary(as.factor(oregon_df$prod))
+```
+
+## 9) *payer*, APAC payer 
+All Payer All Claims [web](https://www.oregon.gov/oha/analytics/Pages/All-Payer-All-Claims.aspx)
+character
+```{r payer}
+summary(as.factor(oregon_df$payer))
+```
+
+## 10) *megflag*, Medical coverage flag
+The medical coverage flag indicates whether this member is covered for medical expenses [web](https://ushik.ahrq.gov/ViewItemDetails?&system=apcd&itemKey=196401000)
+0: (7.39%); 1: (91.57%); NA: (1.04%)
+numeric; range: 0-1
+```{r medical coverage flage}
+summary(as.factor(oregon_df$medflag))
+#       0        1     NA's
+# 5697342 70573441   798537
+summary(as.factor(oregon_df$medflag))/77069320 # percentage
+#          0          1       NA's
+# 0.07392490 0.91571381 0.01036128
+```
+
+## 11) *rxflag*, Pharmacy coverage flag
+This field indicates whether or not the member has prescription drug coverage. 
+[web](https://ushik.ahrq.gov/ViewItemDetails?&system=apcd&itemKey=196401000)
+0: (10.07%); 1: (88.89%); NA: (1.04%)
+numeric; range: 0-1
+```{r pharmacy converage flag}
+summary(as.factor(oregon_df$rxflag))
+#       0        1     NA's
+# 7761348 68509435   798537
+summary(as.factor(oregon_df$rxflag))/77069320 # percentage
+#          0          1       NA's
+# 0.10070607 0.88893265 0.01036128
+```
+
+## 12) *ohvmhflag*, Oregon High value medical home flag
+Required for participants in OHLC high value medical home initiative. 
+[web](https://ushik.ahrq.gov/ViewItemDetails?&system=apcd&itemKey=194764000&enableAsynchronousLoading=true)
+N: (69.15%); Y: (0.24%); NULL: (30.61%)
+character
+```{r Oregon High value medical home flag}
+summary(as.factor(oregon_df$ohvmhflag))
+#        N   *NULL*        Y
+# 53291284 23589431   188605
+summary(as.factor(oregon_df$ohvmhflag))/77069320 # percentage
+#           N      *NULL*           Y
+# 0.691472093 0.306080695 0.002447212
+```
+
+## 13) *pebb*, PEBB flag (NOT Provided)
+0: (95.95%); 1: (4.05%)
+numeric; range: 0-1
+```{r pebb flag}
+summary(as.factor(oregon_df$pebb))
+#        0        1
+# 73949154  3120166
+summary(as.factor(oregon_df$pebb))/77069320 # percentage
+#          0          1
+# 0.95951481 0.04048519
+```
+
+## 14) *oebb*, OEBB flag (NOT Provided)
+0: (96.39%); 1: (3.61%)
+numeric; range: 0-1
+```{r oebb flag}
+summary(as.factor(oregon_df$oebb))
+#        0        1
+# 74285962  2783358
+summary(as.factor(oregon_df$oebb))/77069320 # percentage
+#          0          1
+# 0.96388501 0.03611499
+```
+
+## 15) *patid*, Encrypted patient ID
+numeric
+```{r Encrypted patient ID}
+summary(as.factor(oregon_df$patid))
+```
+
+## 16) *gender*, Gender
+F: (60.56%); M: (39.38%); U (for unknown): (0.06%)
+character
+```{r gender}
+summary(as.factor(oregon_df$gender))
+#        F        M        U
+# 46672307 30350771    46242
+summary(as.factor(oregon_df$gender))/77069320 # percentage
+#            F            M            U
+# 0.6055886700 0.3938113247 0.0006000053
+```
+
+## 17) *yob*, Birth year
+old:1901 (116 years) young:2016 (1 year)
+numeric
+```{r birth year}
+summary(as.factor(oregon_df$yob)) # old:1901 young:2016
+```
+
+## 18) *race*, Race (Not sure the category means)
+1:(31%); 2: (2%); 3: (0.8%); 4: (0.3%); 5: (0.1%); 6: (0.3%); 9: (65%); NA: (0.03%)
+numeric; range: 1-9
+```{r race}
+summary(as.factor(oregon_df$race))
+#        1        2        3        4        5        6        9     NA's
+# 24132190  1611193   620854   203550    85465   248416 50141889    25763
+summary(as.factor(oregon_df$race))/77069320 # percentage
+#            1            2            3            4            5            6
+# 0.3131231727 0.0209057638 0.0080557867 0.0026411288 0.0011089367 0.0032232800
+#            9         NA's
+# 0.6506076478 0.0003342835
+```
+
+## 19) *ethn*, Ethnicity (Not sure the category means)
+1:(0.76%); 2: (9.60%); 3: (89.61%); NA: (0.03%)
+numeric; range: 1-3
+```{r ethnicity}
+summary(as.factor(oregon_df$ethn))
+#        1        2        3     NA's
+#  583948  7398053 69061589    25730
+summary(as.factor(oregon_df$ethn))/77069320 # percentage
+#            1            2            3         NA's
+# 0.0075769191 0.0959921925 0.8960970332 0.0003338553
+```
+
+## 20) *lang*, Primary spoken language 
+Not sure the category means. This web [web](http://www.niso.org/topics/ccm/ccmstandards/) might help.
+1:(47.70%); 2: (1.34%); 3: (0.001%); 4: (0.26%); 8: (0.65%); 9: (49.31%); NA: (0.74%)
+numeric; range: 1-9
+```{r language}
+summary(as.factor(oregon_df$lang))
+#        1        2        3        4        8        9     NA's
+# 36761368  1031059      587   203152   501575 38002063   569516
+round(summary(as.factor(oregon_df$lang))/77069320, 5)  # percentage
+#       1       2       3       4       8       9    NA's
+# 0.47699 0.01338 0.00001 0.00264 0.00651 0.49309 0.00739
+```
+
+## 21) *MSA*, Metropolitan and Micropolitan Statistical Areas Main???
+numeric
+```{r MSA}
+summary(as.factor(oregon_df$MSA))
+```
+
+## 22) *STATE*, Standard two character abbreviation
+OR: (99.33%); blank: (0.1%)
+character
+```{r State}
+summary(as.factor(oregon_df$STATE)) # OR: 76555225; blank: 78646
+76555225/77069320
+```
+
+## 23) *ZIP*, Zip code
+Freely available in the public domain. NA: (0.1%)
+numeric
+```{r ZIP code}
+summary(as.factor(oregon_df$ZIP)) # NA: 78646
+```
+
+## 24) *fromdate*, From date
+YYYY-MM-DD
+character
+```{r from date}
+summary(as.factor(oregon_df$fromdate))
+unique(as.factor(oregon_df$fromdate))
+```
+
+## 25) *todate*, To date
+YYYY-MM-DD
+character
+```{r to date}
+summary(as.factor(oregon_df$todate))
+```
+
+## 26) *paydate*, Pay date
+YYYY-MM-DD
+character
+```{r pay date}
+summary(as.factor(oregon_df$paydate))
+```
+
+## 27) *paid*, Total payment
+NA: (0.36%)
+numeric
+```{r total payment}
+summary(as.factor(oregon_df$paid)) # NA: 279811
+```
+
+## 28) *tob*, Type of bill
+blank: (52.61%); NULL: (25.12%)
+character
+```{r type of bill}
+summary(as.factor(oregon_df$tob)) # blank: 40545960 NULL: 19358510 
+```
+
+## 29) *pos*, Place of service code
+Not sure. [web](https://www.cms.gov/Medicare/Coding/place-of-service-codes/Place_of_Service_Code_Set.html) migh help.
+NULL: (9.22%)
+character
+```{r place of servis code}
+summary(as.factor(oregon_df$pos)) # NULL: 7106719
+```
+
+## 30) *dx1*, Principal diagnosis (For ICD9 code)
+[web](https://www.cms.gov/Medicare/Coding/ICD9ProviderDiagnosticCodes/codes.html)
+NULL: (31.81%)
+character
+```{r dx1}
+summary(as.factor(oregon_df$dx1)) # NULL: 24513232
+```
+
+## 31) *dx2*, Diagnosis 2
+NULL: (59.31%)
+character
+``` {r dx2}
+summary(as.factor(oregon_df$dx2)) # NULL: 45707452
+```
+
+## 32) *dx3*, Diagnosis 3
+NULL: (72.67%)
+character
+``` {r dx3}
+summary(as.factor(oregon_df$dx3)) # NULL: 56005306
+```
+
+## 33) *dx4*, Diagnosis 4
+NULL: (81.46%)
+character
+``` {r dx4}
+summary(as.factor(oregon_df$dx4)) # NULL: 62777450
+```
+
+## 34) *dx5*, Diagnosis 5
+NULL: (93.42%)
+character
+``` {r dx5}
+summary(as.factor(oregon_df$dx5)) # NULL: 71995284
+```
+
+## 35) *dx6*, Diagnosis 6
+NULL: (95.23%)
+character
+``` {r dx6}
+summary(as.factor(oregon_df$dx6)) # NULL: 73390409
+```
+
+## 36) *dx7*, Diagnosis 7
+NULL: (96.31%)
+character
+```{r dx7}
+summary(as.factor(oregon_df$dx7)) # NULL: 74225796
+```
+
+## *37) dx8*, Diagnosis 8
+NULL: (97.00%)
+character
+```{r dx8}
+summary(as.factor(oregon_df$dx8)) # NULL: 74759988
+```
+
+## 38) *dx9*, Diagnosis 9
+NULL: (98.55%)
+character
+```{r dx9}
+summary(as.factor(oregon_df$dx9)) # NULL: 75949570
+```
+
+## 39) *dx10*, Diagnosis 10
+NULL: (98.81%)
+character
+```{r dx10}
+summary(as.factor(oregon_df$dx10)) # NULL: 76151124
+```
+
+## 40) *dx11*, Diagnosis 11
+blank: (98.35%)
+character
+``` {r dx11}
+summary(as.factor(oregon_df$dx11)) # blank: 75801079
+```
+
+## 41) *dx12*, Diagnosis 12
+blank: (98.61%)
+character
+```{r dx12}
+summary(as.factor(oregon_df$dx12)) # blank: 76000733
+```
+
+## 42) *dx13*, Diagnosis 13
+blank: (98.86%)
+character
+```{r dx13}
+summary(as.factor(oregon_df$dx13)) # blank: 76193335
+```
+
+## 43) *poa1*, POA code 1
+N: (0.07%); Y: (5.94%); blank: (93.99%)
+character
+```{r poa1} 
+summary(as.factor(oregon_df$poa1))
+#                 N        Y
+# 72434455    58842  4576023
+summary(as.factor(oregon_df$poa1))/77069320 # percentage
+#                         N            Y
+# 0.9398610887 0.0007634945 0.0593754168
+```
+
+## 44) *poa2*, POA code 2
+N: (0.34%); Y: (3.15%); blank: (96.51%)
+character
+```{r poa2}
+summary(as.factor(oregon_df$poa2))
+#                 N        Y
+# 74380115   261029  2428176
+summary(as.factor(oregon_df$poa2))/77069320 # percentage
+#                       N           Y
+# 0.965106673 0.003386938 0.031506389
+```
+
+## 45) *poa3*, POA code 3
+N: (0.34%); Y: (2.75%); blank: (96.91%)
+character
+```{r poa3}
+summary(as.factor(oregon_df$poa3))
+#                 N        Y
+# 74688528   264287  2116505
+summary(as.factor(oregon_df$poa3))/77069320 # percentage
+#                       N           Y
+# 0.969108434 0.003429212 0.027462355
+```
+
+## 46) *px1*, Principal inpt procedure
+blank: (52.61%); NULL: (45.38%)
+character
+```{r px1}
+summary(as.factor(oregon_df$px1)) # blank: 40545960; NULL: 34975994
+```
+
+## 47) *px2*, Procedure 2
+blank: (52.61%); NULL: (46.06%)
+character
+```{r px2}
+summary(as.factor(oregon_df$px2)) # blank: 40545960 NULL: 35502262
+```
+
+## 48) *px3*, Procedure 3
+blank: (52.61%); NULL: (45.38%)
+character
+```{r px3}
+summary(as.factor(oregon_df$px3)) # blank: 40545960; NULL: 34975994
+```
+
+## 49) *proccode*, CPT or HCPCS procedure code
+Current Procedural Terminology (CPT) Healthcare Common Procedure Coding System (HCPCS) 
+[web1](https://www.cms.gov/Medicare/Coding/MedHCPCSGenInfo/index.html?redirect=/medhcpcsgenin)
+[web2](https://www.cms.gov/Medicare/Coding/HCPCSReleaseCodeSets/index.html?redirect=/hcpcsrelea)
+NULL: (34.88%)
+character
+```{r procedure ccode}
+summary(as.factor(oregon_df$proccode)) # NULL: 26880271
+```
+
+## 50) *mod1*, Prodcure code modifier 1
+[web](https://www.cms.gov/Medicare/Coding/HCPCSReleaseCodeSets/index.html?redirect=/hcpcsrelea)
+NULL: (82.31%)
+character
+```{r Prodcure code modifier 1}
+summary(as.factor(oregon_df$mod1)) # NULL: 63437297
+```
+
+## 51) *mod2*, Prodcure code modifier 2
+NULL: (97.94%)
+character 
+```{r Prodcure code modifier 2}
+summary(as.factor(oregon_df$mod2)) # NULL: 75483898
+```
+
+## 52) *megcode*, code???
+NA: (0.002%)
+numeric
+```{r megcode}
+summary(as.factor(oregon_df$megcode)) # NA: 1751
+```
+
+## 53) *megdesc*, description
+character
+```{r megdesc}
+summary(as.factor(oregon_df$megdesc))
+```
+
+## 54) *megbodysys*, short name for body part???
+Ex: Oral and Dental (Unknown: 9.06%)
+character
+```{r megbodysys}
+summary(as.factor(oregon_df$megbodysys)) # Unknown: 6979372
+```
+
+## 55) *megstate*, stage???
+Unknown: (41.62%)
+character
+```{r megstate}
+summary(as.factor(oregon_df$megstage)) # Unknown: 32074926
+```
+
+## 56) *megtype*, type???
+9 NOT-CATEGPROZED: 9.06%; Acute: 44.20%; Chronic: 40.48%; Well Care: 6.26; blank:0.00%
+  character
+```{r megtype}
+summary(as.factor(oregon_df$megtype))
+#                   9 NOT-CATEGORIZED             Acute           Chronic
+#              1751           6979372          34062467          31197648
+#         Well Care
+#           4828082
+round(summary(as.factor(oregon_df$megtype))/77069320,4) # percentage
+#                   9 NOT-CATEGORIZED             Acute           Chronic
+#            0.0000            0.0906            0.4420            0.4048
+#         Well Care
+#            0.0626
+
+```
+
+## 57) *megcomplete*, complete?
+0: (13.09%); 1: (86.90%); NA: (22.72%)
+numeric; range: 0-1
+```{r megcomplete}
+summary(as.factor(oregon_df$megcomplete))
+#        0        1     NA's
+# 10090656 66976913     1751
+summary(as.factor(oregon_df$megcomplete))/77069320 # percentage
+#            0            1         NA's
+# 1.309296e-01 8.690477e-01 2.271981e-05
+```
+
+## 58) *megnum*, number???
+numeric; range: 147-103100055
+```{r megnum}
+summary(as.factor(oregon_df$megnum))
+max(oregon_df$megnum)
+min(oregon_df$megnum)
+```
+
+## 59) *megdays*, days???
+numeric; range: 1-42452
+```{r megdays}
+summary(as.factor(oregon_df$megdays))
+max(oregon_df$megdays)
+min(oregon_df$megdays)
+```
+
+## 60) *megproprate*, proportion???
+numeric; range: 0-1
+```{r megproprate}
+summary(as.factor(oregon_df$megprorate))
+```
+
+## 61) *megoutlier*, outliers?
+0: (7.94%); 1: (29.06%)
+numeric; range: 0-1
+```{r megoutlier}
+summary(as.factor(oregon_df$megoutlier))
+#        0        1
+# 54675387 22393933
+summary(as.factor(oregon_df$megoutlier))/77069320 # percentage
+#         0         1
+# 0.7094313 0.2905687
+```
+
+## 62) *meglow*, low?
+False: (97.50%); True: (2.50%)
+logical
+```{r meglow}
+summary(as.factor(oregon_df$meglow)) 
+#    FALSE     TRUE
+# 75145616  1923704
+summary(as.factor(oregon_df$meglow))/77069320 # percentage
+#     FALSE      TRUE
+# 0.9750393 0.0249607
+```
+
+## 63) *meghigh*, high???
+False: (73.44%); True: (26.56%)
+logical
+```{r meghigh}
+summary(as.factor(oregon_df$meghigh))
+#    FALSE     TRUE
+# 56599091 20470229
+summary(as.factor(oregon_df$meghigh))/77069320 # percentage
+#    FALSE     TRUE
+# 0.734392 0.265608
+```
+
+## 64) *ndc*, National Drug Code
+[web](http://www.fda.gov/Drugs/InformationOnDrugs/ucm142438.htm)
+blank: (69.55%)
+character
+``` {r ndc}
+# ndc
+summary(as.numeric(oregon_df$ndc)) # blank: 53599072
+```
+
+## 65) *rxclass*, NDC therapeutic class
+blank: (69.55%); NULL: (2.31%)
+character
+```{r NDC therapeutic class}
+summary(as.factor(oregon_df$rxclass)) # blank: 53599072; NULL: 177765
+```
+
+## 66) *qtydisp*, Quantity dispensed
+NA: (2.37%)
+numeric
+```{r Quantity dispensed}
+summary(as.factor(as.character(oregon_df$qtydisp))) # NA: 182736
+```
+
+## 67) *rxdays*, Days supply
+NA: (69.57%)
+numeric
+```{r rxdays}
+summary(as.factor(oregon_df$rxdays)) # NA: 53619826
+```
+
+## 68) *daw*, Dispense as written code
+NA: (69.55%)
+numeric; range: 0-9
+```{r daw}
+summary(as.factor(oregon_df$daw)) # NA: 53599072
+```
+
+## 69) *ptstatus*, Discharge status
+blank: (52.61%); NULL: (35.39%)
+character
+```{r ptstatus}
+summary(as.factor(oregon_df$ptstatus)) # blank: 40545960; NULL: 27273631 
+```
+
+## 70) *los*, Length of stay
+NULL: (52.61%)
+numeric
+```{r los}
+summary(as.factor(oregon_df$los)) # NULL: 40545960
+```
+
+## 71) *msdrg*, MS-DRG
+blank: (52.98%); NULL: (44.04%)
+character
+```{r msdrg}
+summary(as.factor(oregon_df$msdrg)) # blank: 40828376; NULL: 33942845
+```
+
+## 72) *icd_ver*, icd version
+False: (68.19%); NA: (31.81%)
+logical
+```{r icd_ver}
+summary(as.factor(oregon_df$icd_ver))
+#    FALSE     NA's
+# 52556088 24513232
+summary(as.factor(oregon_df$icd_ver))/77069320 # percentage
+#     FALSE      NA's
+# 0.6819327 0.3180673
 

### 2017-02-20 r_scripts/oregon_clean.rmd
Update oregon_clean.rmd

@@
 date: "February 15, 2017"
-output: html_document
 ---
@@
 library(data.table)
+library(knitr)
 getwd()
@@
 # import data ------------------------------------------------------------------
-# Read in first 2000000 lines for exploration.
+# Read in first 200,000 lines for exploration.
 
@@
 
-# I wrote a reduced dataframe of 2000,000 rows to get an idea of data structure
+# I wrote a reduced dataframe of 200,000 rows to get an idea of data structure
 
@@
 start_time <- Sys.time()
-oregon_df <- fread("../data/oregon_epis_care_reduced.csv", sep = ",")
+oregon_df <- fread("../data/oregon_epis_care_reduced.csv", sep = ",", 
+                   colClasses = rep("character", 72))
 stop_time <-  Sys.time() - start_time
@@
 ```{r overall summary}
+# Check data structure
+# summary(oregon_df)
+```
 
-# Explore data structure
-summary(oregon_df)
+## convert variable type to appropriate functional form 
+(i.e. zipcode leave as character, age as numeric, date as date, dx leave as character, but *NULL* == NA)
+``` {r convert variable}
+oregon_df$yob <- as.numeric(oregon_df$yob)
+
+oregon_df[oregon_df=="*NULL*"]<- NA
+```
+
+# summary statistics
+
+## 1. personkey
+```{r}
+personkey2 <- oregon_df %>% 
+  select(personkey, gender) %>%
+  filter(gender=="U") %>%
+  mutate(id_ind =ifelse(is.na(personkey)==0 , 1, 
+                        ifelse(is.na(personkey)==1, 0, NA))
+   )
+
+# table ----
+knitr::kable(personkey2)
+
+# preserve order of group
+personkey2$group <- reorder(personkey2$id_ind, personkey2$personkey)
+
+```
+
+## Age
+```{r}
+oregon_df$yob <- as.numeric(oregon_df$yob)
+age2 <- oregon_df %>%
+  select(yob) %>%
+  mutate(age = 2013-yob) %>%
+  mutate(age_ind = ifelse(age < 15, 0,
+                   ifelse(age >= 15 & age < 65, 1,
+                   ifelse(age >= 65 & age <=120, 2, NA)))
+  ) %>%
+  group_by(age_ind) %>% 
+  summarise(n = n()) %>% 
+  mutate(total_n = sum(n),
+         percent = round(n/total_n,3)*100,
+         age_cat = ifelse(age_ind == 0, "Age <15",
+                   ifelse(age_ind == 1, "Age 15 to 65",
+                   ifelse(age_ind == 2, "Age >= 65",
+                   ifelse(is.na(age_ind), "Missing", NA))))) %>% 
+  select(age_cat, age_ind, n, percent)
+
+age2[4,1] <- age2[4,2] <- "Missing"
+
+knitr::kable(age2, caption = "Age")
+
+```
+
+## summary statistics key variables for study 
+(date, dx, need to know urgent and emergency department visits, sex, age, patientid, state of residence, zipcode, county, race)
+```{r}
+summary(as.factor(oregon_df$personkey))
+summary(as.factor(oregon_df$patid))
+summary(as.factor(oregon_df$gender))
+summary(as.factor(2013-oregon_df$yob)) # birth year -> age
+summary(as.factor(oregon_df$race))
+summary(as.factor(oregon_df$ethn))
+summary(as.factor(oregon_df$STATE))
+summary(as.factor(oregon_df$ZIP))
+summary(as.factor(oregon_df$MSA)) 
+summary(as.factor(oregon_df$fromdate))
+summary(as.factor(oregon_df$todate))
+summary(as.factor(oregon_df$pos)) # 23
+summary(as.factor(oregon_df$megtype)) # Acute for emergency?
+summary(as.factor(oregon_df$los))
+
 ```
@@
 person's ID, maximum is 14381640
-numeric; length: 8
+length: 8
 ```{r personkey}

### 2017-02-21 r_scripts/oregon_clean.rmd
Update oregon_clean.rmd

--- 2017-02-17 r_scripts/oregon_clean.rmd
@@
 ```{r overall summary}
+# Check data structure
+# summary(oregon_df)
+```
 
-# Explore data structure
-summary(oregon_df)
+## convert variable type to appropriate functional form 
+(i.e. zipcode leave as character, age as numeric, date as date, dx leave as character, but *NULL* == NA)
+```{r}
+oregon_df$yob <- as.numeric(oregon_df$yob)
+oregon_df[oregon_df=="*NULL*"] <- NA
+```
+
+# summary statistics
+
+## 1. Personkey ?? not shown?
+```{r}
+personkey2 <- oregon_df %>% 
+  select(personkey) %>%
+  mutate(personkey_cat = ifelse(!is.na(personkey), "Non-Missing", 
+                         ifelse(is.na(personkey), "Missing", NA))) %>% 
+  group_by(personkey_cat) %>%
+  summarise(n = n()) %>%
+  mutate(total_n =sum(n),
+         percent = round(n/total_n, 4)*100) %>%
+      select(personkey_cat, n, percent)
+
+# table ----
+knitr::kable(personkey2, caption = "Patient ID")
+
+```
+
+## 2. Patid ???
+```{r}
+patid2 <- oregon_df %>% 
+  select(patid) %>%
+  mutate(patid_cat = ifelse(!is.na(patid), "Non-Missing", 
+                         ifelse(is.na(patid), "Missing", NA))) %>% 
+  group_by(patid_cat) %>%
+  summarise(n = n()) %>%
+  mutate(total_n =sum(n),
+         percent = round(n/total_n, 4)*100) %>%
+      select(patid_cat, n, percent)
+
+# table ----
+knitr::kable(patid2, align = "c", caption = "Encrypted Patient ID")
+
+```
+
+## 3. Gender
+```{r}
+gender2 <- oregon_df %>%
+  select(gender) %>%
+  group_by(gender) %>%
+  summarise(n = n()) %>%
+  mutate(total_n = sum(n),
+         percent = round(n/total_n, 4)*100,
+         gender_cat = ifelse(gender == "F", "Female",
+                      ifelse(gender == "M", "Male", 
+                      ifelse(gender == "U", "Unknown", NA)))) %>%
+    select(gender, gender_cat, n, percent)
+
+knitr::kable(gender2, align = "c", caption = "Gender")
+
+```
+
+## Age
+```{r}
+oregon_df$yob <- as.numeric(oregon_df$yob)
+age2 <- oregon_df %>%
+  select(yob) %>%
+  mutate(age = 2013-yob) %>%
+  mutate(age_ind = ifelse(age < 15, 0,
+                   ifelse(age >= 15 & age < 65, 1,
+                   ifelse(age >= 65 & age <=120, 2, NA)))
+  ) %>%
+  group_by(age_ind) %>% 
+  summarise(n = n()) %>% 
+  mutate(total_n = sum(n),
+         percent = round(n/total_n, 4)*100,
+         age_cat = ifelse(age_ind == 0, "Age <15",
+                   ifelse(age_ind == 1, "Age 15 to 65",
+                   ifelse(age_ind == 2, "Age >= 65",
+                   ifelse(is.na(age_ind), "Missing", NA))))) %>% 
+  select(age_cat, age_ind, n, percent)
+
+age2[4,1] <- age2[4,2] <- "Missing" # change NA to "Missing"
+
+knitr::kable(age2, align = "c", caption = "Age")
+
+```
+
+## Race
+```{r}
+race2 <- oregon_df %>%
+  select(race) %>%
+  group_by(race) %>%
+  summarise(n = n()) %>%
+  mutate(total_n = sum(n),
+         percent = round(n/total_n, 4)*100,
+         race_cat = ifelse(race == 1, "White",
+                    ifelse(race == 2, "Black or African American",
+                    ifelse(race == 3, "American Indian or Alaska Native",
+                    ifelse(race == 4, "Asian",
+                    ifelse(race == 5, "Native Hawaiian or Pacific Islander",
+                    ifelse(race == 6, "Some other race",
+                    ifelse(race == 7, "Two or more races",
+                    ifelse(race == 9, "Unknown",
+                    ifelse(race == "", "Missing", NA)))))))))) %>%
+    select(race, race_cat, n, percent)
+
+knitr::kable(race2, align = "c", caption = "Race")
+
+```
+
+## Enthicity
+```{r}
+ethn2 <- oregon_df %>%
+  select(ethn) %>%
+  group_by(ethn) %>%
+  summarise(n = n()) %>%
+  mutate(total_n = sum(n),
+         percent = round(n/total_n, 4)*100,
+         ethn_cat = ifelse(ethn == 1, "Hispanic",
+                    ifelse(ethn == 2, "Not Hispanic",
+                    ifelse(ethn == 3, "Unknown",
+                    ifelse(ethn == "", "Missing", NA))))) %>%
+    select(ethn, ethn_cat, n, percent)
+
+knitr::kable(ethn2, align = "c", caption = "Ethnicity")
+
+```
+
+## State
+```{r}
+state2 <- oregon_df %>%
+  select(STATE) %>%
+  group_by(STATE) %>%
+  summarise(n = n()) %>%
+  mutate(total_n = sum(n),
+         percent = round(n/total_n, 4)*100,
+         state_cat = ifelse(STATE == "OR", "Oregon",
+                     ifelse(STATE != "OR"&is.na(STATE)==F, "Not Oregon",NA))) %>%
+    select(STATE, state_cat, n, percent)
+
+knitr::kable(state2, align = "c", caption = "State")
+
+```
+
+## ZIP
+```{r}
+zip2 <- oregon_df %>%
+  select(ZIP) %>%
+  group_by(ZIP) %>%
+  summarise(n = n()) %>%
+  mutate(total_n = sum(n),
+         percent = round(n/total_n, 4)*100,
+         zip_cat = ifelse(is.na(ZIP) == 0, "Non-Missing",
+                   ifelse(is.na(ZIP) != 0, "Missing", NA))) %>%
+    select(ZIP, zip_cat, n, percent)
+
+knitr::kable(zip2, align = "c", caption = "ZIP COde")
+
+```
+
+## MSA
+```{r}
+msa2 <- oregon_df %>%
+  select(MSA) %>%
+  mutate(msa_cat = ifelse(MSA != "", "Non-Missing",
+                   ifelse(MSA == "", "Missing", NA))) %>% 
+  group_by(msa_cat) %>%
+  summarise(n = n()) %>%
+  mutate(total_n = sum(n),
+         percent = round(n/total_n, 4)*100) %>%
+  select(msa_cat, n, percent)
+
+summary(as.factor(oregon_df$MSA))
+
+knitr::kable(msa2, align = "c", caption = "MSA")
+
+```
+
+## Pos
+```{r}
+pos2 <- oregon_df %>%
+  select(pos) %>%
+  group_by(pos) %>%
+  summarise(n = n()) %>%
+  mutate(total_n = sum(n),
+         percent = round(n/total_n, 4)*100,
+         Place_of_service = ifelse(pos == "20", "Urgent Care Facility",
+                            ifelse(pos == "23", "Emergency Room – Hospital", 
+                            ifelse(is.na(pos) == 0, "Not Urgent", 
+                            ifelse(is.na(pos) ==1, "Missing", NA))))) %>%
+    select(pos, Place_of_service, n, percent)
+
+knitr::kable(pos2, align = "c", caption = "Place of Service")
+
+```
+
+## from date
+```{r}
+fromdate2 <- oregon_df %>%
+  select(fromdate) %>%
+  mutate(dates = as.Date(fromdate2$fromdate, "%Y-%m-%d"))
+
+summary(fromdate2)
+hist(fromdate2$dates, "days", freq = TRUE)
+
+```
+
+## to date
+```{r}
+todate2 <- oregon_df %>%
+  select(todate) %>%
+  mutate(ddd = as.Date(todate2$todate, "%Y-%m-%d"))
+
+todate3 <- oregon_df %>%
+  filter (todate > "2014-12-31")
+todate3
+
+which(oregon_df$todate=="2299-12-31") # 129192
+
+summary(todate2)
+
+hist(todate2$ddd, "days", freq = TRUE)
+
+```
+
+## Length of stay
+```{r}
+
+```
+
+## summary statistics key variables for study 
+(date, dx, need to know urgent and emergency department visits, sex, age, patientid, state of residence, zipcode, county, race)
+```{r}
+summary(as.factor(oregon_df$personkey))
+summary(as.factor(oregon_df$patid))
+summary(as.factor(oregon_df$gender))
+summary(as.factor(2013-oregon_df$yob)) # birth year -> age
+summary(as.factor(oregon_df$race))
+summary(as.factor(oregon_df$ethn))
+summary(as.factor(oregon_df$STATE))
+summary(as.factor(oregon_df$ZIP))
+summary(as.factor(oregon_df$MSA)) 
+summary(as.factor(oregon_df$fromdate))
+summary(as.factor(oregon_df$todate))
+summary(as.factor(oregon_df$pos)) # 23&20?
+summary(as.factor(oregon_df$los))
+
 ```

### 2017-02-22 r_scripts/oregon_clean.rmd
Update oregon_clean.rmd
make categories and make table, check dates

@@
 
-## 1. Personkey ?? not shown?
+## 1. Personkey 
 ```{r}
@@
 # table ----
-knitr::kable(personkey2, caption = "Patient ID")
+knitr::kable(personkey2, align = "c", caption = "Patient ID")
 
@@
 
-## 2. Patid ???
+## 2. Patid
 ```{r}
@@
 
-## State
+## State ???
 ```{r}
@@
          percent = round(n/total_n, 4)*100,
-         state_cat = ifelse(STATE == "OR", "Oregon",
-                     ifelse(STATE != "OR"&is.na(STATE)==F, "Not Oregon",NA))) %>%
+         state_cat = ifelse(STATE == "", "Missing",
+                     ifelse(STATE == "OR", "Oregon",
+                     ifelse(STATE != "OR", "Not Oregon",NA)))) %>%
     select(STATE, state_cat, n, percent)
@@
 ```{r}
+# summary(as.factor(oregon_df$ZIP))
+
+# which(oregon_df$ZIP == "")
+
 zip2 <- oregon_df %>% 
   select(ZIP) %>%
-  group_by(ZIP) %>%
+  mutate(zip_cat = ifelse(ZIP == "", "Missing", 
+                   ifelse(!is.na(ZIP), "Non-Missing", NA))) %>% 
+  group_by(zip_cat) %>%
   summarise(n = n()) %>%
-  mutate(total_n = sum(n),
-         percent = round(n/total_n, 4)*100,
-         zip_cat = ifelse(is.na(ZIP) == 0, "Non-Missing",
-                   ifelse(is.na(ZIP) != 0, "Missing", NA))) %>%
-    select(ZIP, zip_cat, n, percent)
+  mutate(total_n =sum(n),
+         percent = round(n/total_n, 4)*100) %>%
+      select(zip_cat, n, percent)
 
@@
 
-## MSA
+## MSA ??? missing 58
 ```{r}
@@
   select(MSA) %>%
-  mutate(msa_cat = ifelse(MSA != "", "Non-Missing",
-                   ifelse(MSA == "", "Missing", NA))) %>% 
+  mutate(msa_cat = ifelse(!is.na(MSA), "Non-Missing", 
+                   ifelse(MSA=="     ", "Missing", NA))) %>% 
   group_by(msa_cat) %>%
   summarise(n = n()) %>%
-  mutate(total_n = sum(n),
+  mutate(total_n =sum(n),
          percent = round(n/total_n, 4)*100) %>%
@@
 
-## Pos
+## Pos NA: 2974
 ```{r}
+library(readxl)
+pos_key <- read_excel("../instructions/LT_pos.xlsx")
+summary(pos_key)
+
 pos2 <- oregon_df %>%
-  select(pos) %>%
+  select(pos)
+
+which(pos2$pos=="0S") # 14553
+as.numeric(pos2$pos[14553])
+
+pos2$pos <- as.numeric(pos2$pos)
+
+pos2 <- pos2 %>%
   group_by(pos) %>%
@@
          percent = round(n/total_n, 4)*100,
-         Place_of_service = ifelse(pos == "20", "Urgent Care Facility",
-                            ifelse(pos == "23", "Emergency Room – Hospital", 
-                            ifelse(is.na(pos) == 0, "Not Urgent", 
-                            ifelse(is.na(pos) ==1, "Missing", NA))))) %>%
+         Place_of_service = ifelse(pos == pos_key$Code[1], pos_key$Value[1],
+                            ifelse(pos == pos_key$Code[2], pos_key$Value[2],
+                            ifelse(pos == pos_key$Code[3], pos_key$Value[3],       
+                            ifelse(pos == pos_key$Code[4], pos_key$Value[4],       
+                            ifelse(pos == pos_key$Code[5], pos_key$Value[5],       
+                            ifelse(pos == pos_key$Code[6], pos_key$Value[6],        
+                            ifelse(pos == pos_key$Code[7], pos_key$Value[7],
+                            ifelse(pos == pos_key$Code[8], pos_key$Value[8],     
+                            ifelse(pos == pos_key$Code[9], pos_key$Value[9],      
+                            ifelse(pos == pos_key$Code[10], pos_key$Value[10],      
+                            ifelse(pos == pos_key$Code[11], pos_key$Value[11],      
+                            ifelse(pos == pos_key$Code[12], pos_key$Value[12],    
+                            ifelse(pos == pos_key$Code[13], pos_key$Value[13],     
+                            ifelse(pos == pos_key$Code[14], pos_key$Value[14],        
+                            ifelse(pos == pos_key$Code[15], pos_key$Value[15],      
+                            ifelse(pos == pos_key$Code[16], pos_key$Value[16],     
+                            ifelse(pos == pos_key$Code[17], pos_key$Value[17],      
+                            ifelse(pos == pos_key$Code[18], pos_key$Value[18],     
+                            ifelse(pos == pos_key$Code[19], pos_key$Value[19],      
+                            ifelse(pos == pos_key$Code[20], pos_key$Value[20],       
+                            ifelse(pos == pos_key$Code[21], pos_key$Value[21],     
+                            ifelse(pos == pos_key$Code[22], pos_key$Value[22],      
+                            ifelse(pos == pos_key$Code[23], pos_key$Value[23],      
+                            ifelse(pos == pos_key$Code[24], pos_key$Value[24],      
+                            ifelse(pos == pos_key$Code[25], pos_key$Value[25],     
+                            ifelse(pos == pos_key$Code[26], pos_key$Value[26],      
+                            ifelse(pos == pos_key$Code[27], pos_key$Value[27],      
+                            ifelse(pos == pos_key$Code[28], pos_key$Value[28],      
+                            ifelse(pos == pos_key$Code[29], pos_key$Value[29],      
+                            ifelse(pos == pos_key$Code[30], pos_key$Value[30],     
+                            ifelse(pos == pos_key$Code[31], pos_key$Value[31],     
+                            ifelse(pos == pos_key$Code[32], pos_key$Value[32],      
+                            ifelse(pos == pos_key$Code[33], pos_key$Value[33],
+                            ifelse(pos == pos_key$Code[34], pos_key$Value[34],
+                            ifelse(pos == pos_key$Code[35], pos_key$Value[35],
+                            ifelse(pos == pos_key$Code[36], pos_key$Value[36],                                   
+                            ifelse(pos == pos_key$Code[37], pos_key$Value[37],
+                            ifelse(pos == pos_key$Code[38], pos_key$Value[38],                                   
+                            ifelse(pos == pos_key$Code[39], pos_key$Value[39],                                   
+                            ifelse(pos == pos_key$Code[40], pos_key$Value[40],                                   
+                            ifelse(pos == pos_key$Code[41], pos_key$Value[41],                                   
+                            ifelse(pos == pos_key$Code[42], pos_key$Value[42],                                   
+                            ifelse(pos == pos_key$Code[43], pos_key$Value[43],                                   
+                            ifelse(pos == pos_key$Code[44], pos_key$Value[44],                                  
+                            ifelse(pos == pos_key$Code[45], pos_key$Value[45],                                   
+                            ifelse(pos == pos_key$Code[46], pos_key$Value[46],                                   
+                            ifelse(is.na(pos), "Missing", NA
+                                   )))))))))))))))))))))))))))))))))))))))))))))))) %>%
     select(pos, Place_of_service, n, percent)
+
+pos2$Place_of_service[which(is.na(pos2$pos))] <- "Missing"
+pos2$pos[which(is.na(pos2$pos))] <- "Missing"
 
@@
 
-## from date
+## from date (All data is in 2013)
 ```{r}
@@
   select(fromdate) %>%
-  mutate(dates = as.Date(fromdate2$fromdate, "%Y-%m-%d"))
+  mutate(dates = as.Date(fromdate, "%Y-%m-%d"))
+
+fromdate_check <- oregon_df %>%
+  filter((fromdate > "2014-12-31"| fromdate < "2013-01-01"))
 
@@
 
-## to date
+## to date (min: 1913-02-17; max: 2299-12-31)
 ```{r}
@@
   select(todate) %>%
-  mutate(ddd = as.Date(todate2$todate, "%Y-%m-%d"))
+  mutate(ddd = as.Date(todate, "%Y-%m-%d"))
 
 todate3 <- oregon_df %>%
-  filter (todate > "2014-12-31")
+  filter (todate > "2014-12-31"| todate < "2013-01-01")
 todate3
@@
 which(oregon_df$todate=="2299-12-31") # 129192
+which(oregon_df$todate=="1913-10-25") # 178869
 
@@
 
-hist(todate2$ddd, "days", freq = TRUE)
+hist(todate2$ddd[c(-129192,-178869)], "days", freq = TRUE)
 
@@
 ```{r}
+summary(as.factor(oregon_df$los))
 

### 2017-02-23 r_scripts/oregon_clean.rmd
Update oregon_clean.rmd
Data category table for important variable exploring

@@
 library(knitr)
+library(readxl)
 getwd()
@@
 
-# summary statistics
+# summary statistics key variables for study 
+date, dx, need to know urgent and emergency department visits, sex, age, patientid, state of residence, zipcode, county, race
 
@@
   select(personkey) %>%
-  mutate(personkey_cat = ifelse(!is.na(personkey), "Non-Missing", 
-                         ifelse(is.na(personkey), "Missing", NA))) %>% 
+  mutate(personkey_cat = ifelse(is.na(personkey), "Missing",
+                         ifelse(!is.na(personkey), "Non-Missing", NA))) %>% 
   group_by(personkey_cat) %>%
@@
   select(patid) %>%
-  mutate(patid_cat = ifelse(!is.na(patid), "Non-Missing", 
-                         ifelse(is.na(patid), "Missing", NA))) %>% 
+  mutate(patid_cat = ifelse(is.na(patid), "Missing",
+                     ifelse(!is.na(patid), "Non-Missing", NA))) %>% 
   group_by(patid_cat) %>%
@@
 
-## Age
+## 4. Age
 ```{r}
@@
          percent = round(n/total_n, 4)*100,
-         age_cat = ifelse(age_ind == 0, "Age <15",
+         age_cat = ifelse(is.na(age_ind), "Missing",
+                   ifelse(age_ind == 0, "Age <15",
                    ifelse(age_ind == 1, "Age 15 to 65",
-                   ifelse(age_ind == 2, "Age >= 65",
-                   ifelse(is.na(age_ind), "Missing", NA))))) %>% 
+                   ifelse(age_ind == 2, "Age >= 65",NA))))) %>% 
   select(age_cat, age_ind, n, percent)
-
-age2[4,1] <- age2[4,2] <- "Missing" # change NA to "Missing"
 
@@
 
-## Race
+## 5. Race
 ```{r}
@@
 
-## Enthicity
+## 6. Enthicity
 ```{r}
@@
 
-## State ???
+## 7. State 
 ```{r}
@@
                      ifelse(STATE != "OR", "Not Oregon",NA)))) %>%
+  arrange(desc(state_cat)) %>%
   select(STATE, state_cat, n, percent)
@@
 
-## ZIP 
+## 8. ZIP 
 ```{r}
@@
 
-## MSA ??? missing 58
+## 9. MSA (Metropolitan and Micropolitan Statistical Areas Main)
 ```{r}
@@
   select(MSA) %>%
-  mutate(msa_cat = ifelse(!is.na(MSA), "Non-Missing", 
-                   ifelse(MSA=="     ", "Missing", NA))) %>% 
+  mutate(msa_cat = ifelse(MSA=="", "Missing",
+                   ifelse(!is.na(MSA), "Non-Missing", 
+                   NA))) %>% 
   group_by(msa_cat) %>%
@@
 
-## Pos NA: 2974
+## 10. Pos (Urgent and Emergency) NA: 2974
 ```{r}
-library(readxl)
 pos_key <- read_excel("../instructions/LT_pos.xlsx")
-summary(pos_key)
 
@@
          percent = round(n/total_n, 4)*100,
-         Place_of_service = ifelse(pos == pos_key$Code[1], pos_key$Value[1],
+         Place_of_service = ifelse(is.na(pos), "Missing",
+                            ifelse(pos == pos_key$Code[1], pos_key$Value[1],
                             ifelse(pos == pos_key$Code[2], pos_key$Value[2],
@@
                             ifelse(pos == pos_key$Code[45], pos_key$Value[45],                            
-                            ifelse(pos == pos_key$Code[46], pos_key$Value[46],                                   
-                            ifelse(is.na(pos), "Missing", NA
+                            ifelse(pos == pos_key$Code[46], pos_key$Value[46], NA
                                    )))))))))))))))))))))))))))))))))))))))))))))))) %>%
   select(pos, Place_of_service, n, percent)
-
-pos2$Place_of_service[which(is.na(pos2$pos))] <- "Missing"
-pos2$pos[which(is.na(pos2$pos))] <- "Missing"
 
@@
 
-## from date (All data is in 2013)
+## 11. from date (All data is in 2013)
 ```{r}
@@
 
+# filter the dates of 2013
 fromdate_check <- oregon_df %>%
-  filter((fromdate > "2014-12-31"| fromdate < "2013-01-01"))
+  filter((fromdate > "2013-12-31"| fromdate < "2013-01-01")) # All is 2013
 
 summary(fromdate2)
-hist(fromdate2$dates, "days", freq = TRUE)
+hist(fromdate2$dates, "days", freq = TRUE, xlab="fromdate", main="Histogram of from dates")
 
@@
 
-## to date (min: 1913-02-17; max: 2299-12-31)
+## 12. to date (min: 1913-02-17; max: 2299-12-31) 
 ```{r}
@@
   select(todate) %>%
-  mutate(ddd = as.Date(todate, "%Y-%m-%d"))
+  mutate(dates = as.Date(todate, "%Y-%m-%d"))
 
-todate3 <- oregon_df %>%
-  filter (todate > "2014-12-31"| todate < "2013-01-01")
-todate3
+# filter the appropriate to dates (2013-01-01 to 2014-06-30)
+todate3 <- todate2 %>%
+  filter(dates<=as.Date("2014-06-30")&dates>=as.Date("2013-01-01"))
 
-which(oregon_df$todate=="2299-12-31") # 129192
-which(oregon_df$todate=="1913-10-25") # 178869
-
-summary(todate2)
-
-hist(todate2$ddd[c(-129192,-178869)], "days", freq = TRUE)
+hist(todate3$dates, "days", freq = TRUE, xlab="todate", main="Histogram of to dates")
 
@@
 
-## Length of stay
+## 13. Length of stay
 ```{r}
 summary(as.factor(oregon_df$los))
+summary(as.numeric(oregon_df$los)) # min: -2; max: 364
+
+# check length of stay overall range
+los2 <- oregon_df %>%
+  select(los) %>%
+  group_by(los) %>%
+  summarise(n = n()) %>%
+  mutate(total_n = sum(n),
+         percent = round(n/total_n, 4)*100) %>%
+  select(los, n, percent)
+
+knitr::kable(los2, align = "c", caption = "Length of Stay Range")
+
+# density for length of stay
+hist(as.numeric(los2$los), freq = T, xlab = "los", main = "Length of Stay")
+
+# Check if fromdate and todate match los
+los3 <- oregon_df %>%
+  select(fromdate, todate, los) %>%
+  mutate(dif = as.numeric(as.Date(todate)-as.Date(fromdate))) %>%
+  select(los, dif) %>%
+  mutate(los_cat = ifelse(los=="", "Missing",
+                   ifelse(los==dif, "Right", 
+                   ifelse(los!=dif, "Wrong", NA)))) %>%
+  group_by(los_cat) %>%
+  summarise(n=n()) %>%
+  mutate(total_n = sum(n),
+         percent = round(n/total_n, 4)*100) %>%
+  select(los_cat, n, percent)
+
+knitr::kable(los3, align = "c", caption = "Length of Stay Range Match From date and To Date")
 
@@
 
-## summary statistics key variables for study 
-(date, dx, need to know urgent and emergency department visits, sex, age, patientid, state of residence, zipcode, county, race)
+## dx1
 ```{r}
-summary(as.factor(oregon_df$personkey))
-summary(as.factor(oregon_df$patid))
-summary(as.factor(oregon_df$gender))
-summary(as.factor(2013-oregon_df$yob)) # birth year -> age
-summary(as.factor(oregon_df$race))
-summary(as.factor(oregon_df$ethn))
-summary(as.factor(oregon_df$STATE))
-summary(as.factor(oregon_df$ZIP))
-summary(as.factor(oregon_df$MSA)) 
-summary(as.factor(oregon_df$fromdate))
-summary(as.factor(oregon_df$todate))
-summary(as.factor(oregon_df$pos)) # 23&20?
-summary(as.factor(oregon_df$los))
+dx1_2 <- oregon_df %>% 
+  select(dx1) %>%
+  mutate(dx1_cat = ifelse(is.na(dx1), "Missing",
+                     ifelse(!is.na(dx1), "Non-Missing", NA))) %>% 
+  group_by(dx1_cat) %>%
+  summarise(n = n()) %>%
+  mutate(total_n =sum(n),
+         percent = round(n/total_n, 4)*100) %>%
+      select(dx1_cat, n, percent)
+
+# table ----
+knitr::kable(dx1_2, align = "c", caption = "dx1")
 
@@
 
-## 1) *personkey*, Unique person identifier
-person's ID, maximum is 14381640
-length: 8
-```{r personkey}
-summary(as.factor(oregon_df$personkey))
-min(oregon_df$personkey)
-max(oregon_df$personkey) # seems length is 8
-```
+## dx2
+```{r}
+dx2_2 <- oregon_df %>% 
+  select(dx2) %>%
+  mutate(dx2_cat = ifelse(is.na(dx2), "Missing",
+                     ifelse(!is.na(dx2), "Non-Missing", NA))) %>% 
+  group_by(dx2_cat) %>%
+  summarise(n = n()) %>%
+  mutate(total_n =sum(n),
+         percent = round(n/total_n, 4)*100) %>%
+      select(dx2_cat, n, percent)
 
-## 2) *year*, Calendar year in which claim was incurred
-All is 2013
-integer (numeric); length: 4
-```{r year}
-summary(as.factor(oregon_df$year)) # all 2013, 77069320
-```
-
-## 3) *clmid*, Claim ID
-numeric; length: 8
-```{r claim id}
-summary(as.factor(oregon_df$clmid))
-```
-
-## 4) *line*, Claim line (Not Sure what this means)
-numeric; range: 0-9999
-```{r claim line}
-unique(as.numeric(oregon_df$line)) # 1362 levels, 0 - 9999
-```
-
-## 5) *clmtatus*, Claim status
-D: 8676643 (11.26%), E: 3324496 (4.31%), P: 65068181 (84.43%)
-character
-```{r claim status}
-summary(as.factor(oregon_df$clmstatus))
-#       D        E        P
-# 8676643  3324496 65068181
-
-summary(as.factor(oregon_df$clmstatus))/77069320 # percentage
-#          D          E          P
-# 0.11258232 0.04313644 0.84428124
-```
-
-## 6) *cob*, COB status (Coordination of benefits) 
-blank: 23517978 (30.52%); N: 44438688 (57.66%); Y: 9112654 (11.82%)
-character
-``` {r cob}
-summary(as.factor(oregon_df$cob)) 
-#                 N        Y
-# 23517978 44438688  9112654
-summary(as.factor(oregon_df$cob))/77069320 # percentage
-#                   N         Y
-# 0.3051536 0.5766067 0.1182397
-```
-
-## 7) *paytype*, Pay type
-C (58.50%), D (31.77%), G (2.08%), P (3.53%), T (4.12%)
-character
-``` {r pay type}
-summary(as.factor(oregon_df$paytype)) #C, D, G, P, T
-summary(as.factor(oregon_df$paytype))/77069320 # percentage
-#         C          D          G          P          T
-# 0.58498296 0.31771349 0.02081743 0.03532560 0.04116052
-```
-
-## 8) *prod*, Product code
-CHP, etc. Insurance type/product code [web](https://ushik.ahrq.gov/ViewItemDetails?&system=apcd&itemKey=194604000)
-character
-```{r product code}
-summary(as.factor(oregon_df$prod))
-```
-
-## 9) *payer*, APAC payer 
-All Payer All Claims [web](https://www.oregon.gov/oha/analytics/Pages/All-Payer-All-Claims.aspx)
-character
-```{r payer}
-summary(as.factor(oregon_df$payer))
-```
-
-## 10) *megflag*, Medical coverage flag
-The medical coverage flag indicates whether this member is covered for medical expenses [web](https://ushik.ahrq.gov/ViewItemDetails?&system=apcd&itemKey=196401000)
-0: (7.39%); 1: (91.57%); NA: (1.04%)
-numeric; range: 0-1
-```{r medical coverage flage}
-summary(as.factor(oregon_df$medflag))
-#       0        1     NA's
-# 5697342 70573441   798537
-summary(as.factor(oregon_df$medflag))/77069320 # percentage
-#          0          1       NA's
-# 0.07392490 0.91571381 0.01036128
-```
-
-## 11) *rxflag*, Pharmacy coverage flag
-This field indicates whether or not the member has prescription drug coverage. 
-[web](https://ushik.ahrq.gov/ViewItemDetails?&system=apcd&itemKey=196401000)
-0: (10.07%); 1: (88.89%); NA: (1.04%)
-numeric; range: 0-1
-```{r pharmacy converage flag}
-summary(as.factor(oregon_df$rxflag))
-#       0        1     NA's
-# 7761348 68509435   798537
-summary(as.factor(oregon_df$rxflag))/77069320 # percentage
-#          0          1       NA's
-# 0.10070607 0.88893265 0.01036128
-```
-
-## 12) *ohvmhflag*, Oregon High value medical home flag
-Required for participants in OHLC high value medical home initiative. 
-[web](https://ushik.ahrq.gov/ViewItemDetails?&system=apcd&itemKey=194764000&enableAsynchronousLoading=true)
-N: (69.15%); Y: (0.24%); NULL: (30.61%)
-character
-```{r Oregon High value medical home flag}
-summary(as.factor(oregon_df$ohvmhflag))
-#        N   *NULL*        Y
-# 53291284 23589431   188605
-summary(as.factor(oregon_df$ohvmhflag))/77069320 # percentage
-#           N      *NULL*           Y
-# 0.691472093 0.306080695 0.002447212
-```
-
-## 13) *pebb*, PEBB flag (NOT Provided)
-0: (95.95%); 1: (4.05%)
-numeric; range: 0-1
-```{r pebb flag}
-summary(as.factor(oregon_df$pebb))
-#        0        1
-# 73949154  3120166
-summary(as.factor(oregon_df$pebb))/77069320 # percentage
-#          0          1
-# 0.95951481 0.04048519
-```
-
-## 14) *oebb*, OEBB flag (NOT Provided)
-0: (96.39%); 1: (3.61%)
-numeric; range: 0-1
-```{r oebb flag}
-summary(as.factor(oregon_df$oebb))
-#        0        1
-# 74285962  2783358
-summary(as.factor(oregon_df$oebb))/77069320 # percentage
-#          0          1
-# 0.96388501 0.03611499
-```
-
-## 15) *patid*, Encrypted patient ID
-numeric
-```{r Encrypted patient ID}
-summary(as.factor(oregon_df$patid))
-```
-
-## 16) *gender*, Gender
-F: (60.56%); M: (39.38%); U (for unknown): (0.06%)
-character
-```{r gender}
-summary(as.factor(oregon_df$gender))
-#        F        M        U
-# 46672307 30350771    46242
-summary(as.factor(oregon_df$gender))/77069320 # percentage
-#            F            M            U
-# 0.6055886700 0.3938113247 0.0006000053
-```
-
-## 17) *yob*, Birth year
-old:1901 (116 years) young:2016 (1 year)
-numeric
-```{r birth year}
-summary(as.factor(oregon_df$yob)) # old:1901 young:2016
-```
-
-## 18) *race*, Race (Not sure the category means)
-1:(31%); 2: (2%); 3: (0.8%); 4: (0.3%); 5: (0.1%); 6: (0.3%); 9: (65%); NA: (0.03%)
-numeric; range: 1-9
-```{r race}
-summary(as.factor(oregon_df$race))
-#        1        2        3        4        5        6        9     NA's
-# 24132190  1611193   620854   203550    85465   248416 50141889    25763
-summary(as.factor(oregon_df$race))/77069320 # percentage
-#            1            2            3            4            5            6
-# 0.3131231727 0.0209057638 0.0080557867 0.0026411288 0.0011089367 0.0032232800
-#            9         NA's
-# 0.6506076478 0.0003342835
-```
-
-## 19) *ethn*, Ethnicity (Not sure the category means)
-1:(0.76%); 2: (9.60%); 3: (89.61%); NA: (0.03%)
-numeric; range: 1-3
-```{r ethnicity}
-summary(as.factor(oregon_df$ethn))
-#        1        2        3     NA's
-#  583948  7398053 69061589    25730
-summary(as.factor(oregon_df$ethn))/77069320 # percentage
-#            1            2            3         NA's
-# 0.0075769191 0.0959921925 0.8960970332 0.0003338553
-```
-
-## 20) *lang*, Primary spoken language 
-Not sure the category means. This web [web](http://www.niso.org/topics/ccm/ccmstandards/) might help.
-1:(47.70%); 2: (1.34%); 3: (0.001%); 4: (0.26%); 8: (0.65%); 9: (49.31%); NA: (0.74%)
-numeric; range: 1-9
-```{r language}
-summary(as.factor(oregon_df$lang))
-#        1        2        3        4        8        9     NA's
-# 36761368  1031059      587   203152   501575 38002063   569516
-round(summary(as.factor(oregon_df$lang))/77069320, 5)  # percentage
-#       1       2       3       4       8       9    NA's
-# 0.47699 0.01338 0.00001 0.00264 0.00651 0.49309 0.00739
-```
-
-## 21) *MSA*, Metropolitan and Micropolitan Statistical Areas Main???
-numeric
-```{r MSA}
-summary(as.factor(oregon_df$MSA))
-```
-
-## 22) *STATE*, Standard two character abbreviation
-OR: (99.33%); blank: (0.1%)
-character
-```{r State}
-summary(as.factor(oregon_df$STATE)) # OR: 76555225; blank: 78646
-76555225/77069320
-```
-
-## 23) *ZIP*, Zip code
-Freely available in the public domain. NA: (0.1%)
-numeric
-```{r ZIP code}
-summary(as.factor(oregon_df$ZIP)) # NA: 78646
-```
-
-## 24) *fromdate*, From date
-YYYY-MM-DD
-character
-```{r from date}
-summary(as.factor(oregon_df$fromdate))
-unique(as.factor(oregon_df$fromdate))
-```
-
-## 25) *todate*, To date
-YYYY-MM-DD
-character
-```{r to date}
-summary(as.factor(oregon_df$todate))
-```
-
-## 26) *paydate*, Pay date
-YYYY-MM-DD
-character
-```{r pay date}
-summary(as.factor(oregon_df$paydate))
-```
-
-## 27) *paid*, Total payment
-NA: (0.36%)
-numeric
-```{r total payment}
-summary(as.factor(oregon_df$paid)) # NA: 279811
-```
-
-## 28) *tob*, Type of bill
-blank: (52.61%); NULL: (25.12%)
-character
-```{r type of bill}
-summary(as.factor(oregon_df$tob)) # blank: 40545960 NULL: 19358510 
-```
-
-## 29) *pos*, Place of service code
-Not sure. [web](https://www.cms.gov/Medicare/Coding/place-of-service-codes/Place_of_Service_Code_Set.html) migh help.
-NULL: (9.22%)
-character
-```{r place of servis code}
-summary(as.factor(oregon_df$pos)) # NULL: 7106719
-```
-
-## 30) *dx1*, Principal diagnosis (For ICD9 code)
-[web](https://www.cms.gov/Medicare/Coding/ICD9ProviderDiagnosticCodes/codes.html)
-NULL: (31.81%)
-character
-```{r dx1}
-summary(as.factor(oregon_df$dx1)) # NULL: 24513232
-```
-
-## 31) *dx2*, Diagnosis 2
-NULL: (59.31%)
-character
-``` {r dx2}
-summary(as.factor(oregon_df$dx2)) # NULL: 45707452
-```
-
-## 32) *dx3*, Diagnosis 3
-NULL: (72.67%)
-character
-``` {r dx3}
-summary(as.factor(oregon_df$dx3)) # NULL: 56005306
-```
-
-## 33) *dx4*, Diagnosis 4
-NULL: (81.46%)
-character
-``` {r dx4}
-summary(as.factor(oregon_df$dx4)) # NULL: 62777450
-```
-
-## 34) *dx5*, Diagnosis 5
-NULL: (93.42%)
-character
-``` {r dx5}
-summary(as.factor(oregon_df$dx5)) # NULL: 71995284
-```
-
-## 35) *dx6*, Diagnosis 6
-NULL: (95.23%)
-character
-``` {r dx6}
-summary(as.factor(oregon_df$dx6)) # NULL: 73390409
-```
-
-## 36) *dx7*, Diagnosis 7
-NULL: (96.31%)
-character
-```{r dx7}
-summary(as.factor(oregon_df$dx7)) # NULL: 74225796
-```
-
-## *37) dx8*, Diagnosis 8
-NULL: (97.00%)
-character
-```{r dx8}
-summary(as.factor(oregon_df$dx8)) # NULL: 74759988
-```
-
-## 38) *dx9*, Diagnosis 9
-NULL: (98.55%)
-character
-```{r dx9}
-summary(as.factor(oregon_df$dx9)) # NULL: 75949570
-```
-
-## 39) *dx10*, Diagnosis 10
-NULL: (98.81%)
-character
-```{r dx10}
-summary(as.factor(oregon_df$dx10)) # NULL: 76151124
-```
-
-## 40) *dx11*, Diagnosis 11
-blank: (98.35%)
-character
-``` {r dx11}
-summary(as.factor(oregon_df$dx11)) # blank: 75801079
-```
-
-## 41) *dx12*, Diagnosis 12
-blank: (98.61%)
-character
-```{r dx12}
-summary(as.factor(oregon_df$dx12)) # blank: 76000733
-```
-
-## 42) *dx13*, Diagnosis 13
-blank: (98.86%)
-character
-```{r dx13}
-summary(as.factor(oregon_df$dx13)) # blank: 76193335
-```
-
-## 43) *poa1*, POA code 1
-N: (0.07%); Y: (5.94%); blank: (93.99%)
-character
-```{r poa1} 
-summary(as.factor(oregon_df$poa1))
-#                 N        Y
-# 72434455    58842  4576023
-summary(as.factor(oregon_df$poa1))/77069320 # percentage
-#                         N            Y
-# 0.9398610887 0.0007634945 0.0593754168
-```
-
-## 44) *poa2*, POA code 2
-N: (0.34%); Y: (3.15%); blank: (96.51%)
-character
-```{r poa2}
-summary(as.factor(oregon_df$poa2))
-#                 N        Y
-# 74380115   261029  2428176
-summary(as.factor(oregon_df$poa2))/77069320 # percentage
-#                       N           Y
-# 0.965106673 0.003386938 0.031506389
-```
-
-## 45) *poa3*, POA code 3
-N: (0.34%); Y: (2.75%); blank: (96.91%)
-character
-```{r poa3}
-summary(as.factor(oregon_df$poa3))
-#                 N        Y
-# 74688528   264287  2116505
-summary(as.factor(oregon_df$poa3))/77069320 # percentage
-#                       N           Y
-# 0.969108434 0.003429212 0.027462355
-```
-
-## 46) *px1*, Principal inpt procedure
-blank: (52.61%); NULL: (45.38%)
-character
-```{r px1}
-summary(as.factor(oregon_df$px1)) # blank: 40545960; NULL: 34975994
-```
-
-## 47) *px2*, Procedure 2
-blank: (52.61%); NULL: (46.06%)
-character
-```{r px2}
-summary(as.factor(oregon_df$px2)) # blank: 40545960 NULL: 35502262
-```
-
-## 48) *px3*, Procedure 3
-blank: (52.61%); NULL: (45.38%)
-character
-```{r px3}
-summary(as.factor(oregon_df$px3)) # blank: 40545960; NULL: 34975994
-```
-
-## 49) *proccode*, CPT or HCPCS procedure code
-Current Procedural Terminology (CPT) Healthcare Common Procedure Coding System (HCPCS) 
-[web1](https://www.cms.gov/Medicare/Coding/MedHCPCSGenInfo/index.html?redirect=/medhcpcsgenin)
-[web2](https://www.cms.gov/Medicare/Coding/HCPCSReleaseCodeSets/index.html?redirect=/hcpcsrelea)
-NULL: (34.88%)
-character
-```{r procedure ccode}
-summary(as.factor(oregon_df$proccode)) # NULL: 26880271
-```
-
-## 50) *mod1*, Prodcure code modifier 1
-[web](https://www.cms.gov/Medicare/Coding/HCPCSReleaseCodeSets/index.html?redirect=/hcpcsrelea)
-NULL: (82.31%)
-character
-```{r Prodcure code modifier 1}
-summary(as.factor(oregon_df$mod1)) # NULL: 63437297
-```
-
-## 51) *mod2*, Prodcure code modifier 2
-NULL: (97.94%)
-character 
-```{r Prodcure code modifier 2}
-summary(as.factor(oregon_df$mod2)) # NULL: 75483898
-```
-
-## 52) *megcode*, code???
-NA: (0.002%)
-numeric
-```{r megcode}
-summary(as.factor(oregon_df$megcode)) # NA: 1751
-```
-
-## 53) *megdesc*, description
-character
-```{r megdesc}
-summary(as.factor(oregon_df$megdesc))
-```
-
-## 54) *megbodysys*, short name for body part???
-Ex: Oral and Dental (Unknown: 9.06%)
-character
-```{r megbodysys}
-summary(as.factor(oregon_df$megbodysys)) # Unknown: 6979372
-```
-
-## 55) *megstate*, stage???
-Unknown: (41.62%)
-character
-```{r megstate}
-summary(as.factor(oregon_df$megstage)) # Unknown: 32074926
-```
-
-## 56) *megtype*, type???
-9 NOT-CATEGPROZED: 9.06%; Acute: 44.20%; Chronic: 40.48%; Well Care: 6.26; blank:0.00%
-  character
-```{r megtype}
-summary(as.factor(oregon_df$megtype))
-#                   9 NOT-CATEGORIZED             Acute           Chronic
-#              1751           6979372          34062467          31197648
-#         Well Care
-#           4828082
-round(summary(as.factor(oregon_df$megtype))/77069320,4) # percentage
-#                   9 NOT-CATEGORIZED             Acute           Chronic
-#            0.0000            0.0906            0.4420            0.4048
-#         Well Care
-#            0.0626
+# table ----
+knitr::kable(dx2_2, align = "c", caption = "dx2")
 
@@
 
-## 57) *megcomplete*, complete?
-0: (13.09%); 1: (86.90%); NA: (22.72%)
-numeric; range: 0-1
-```{r megcomplete}
-summary(as.factor(oregon_df$megcomplete))
-#        0        1     NA's
-# 10090656 66976913     1751
-summary(as.factor(oregon_df$megcomplete))/77069320 # percentage
-#            0            1         NA's
-# 1.309296e-01 8.690477e-01 2.271981e-05
-```
+## dx3
+```{r}
+dx3_2 <- oregon_df %>% 
+  select(dx3) %>%
+  mutate(dx3_cat = ifelse(is.na(dx3), "Missing",
+                     ifelse(!is.na(dx3), "Non-Missing", NA))) %>% 
+  group_by(dx3_cat) %>%
+  summarise(n = n()) %>%
+  mutate(total_n =sum(n),
+         percent = round(n/total_n, 4)*100) %>%
+      select(dx3_cat, n, percent)
 
-## 58) *megnum*, number???
-numeric; range: 147-103100055
-```{r megnum}
-summary(as.factor(oregon_df$megnum))
-max(oregon_df$megnum)
-min(oregon_df$megnum)
-```
-
-## 59) *megdays*, days???
-numeric; range: 1-42452
-```{r megdays}
-summary(as.factor(oregon_df$megdays))
-max(oregon_df$megdays)
-min(oregon_df$megdays)
-```
-
-## 60) *megproprate*, proportion???
-numeric; range: 0-1
-```{r megproprate}
-summary(as.factor(oregon_df$megprorate))
-```
-
-## 61) *megoutlier*, outliers?
-0: (7.94%); 1: (29.06%)
-numeric; range: 0-1
-```{r megoutlier}
-summary(as.factor(oregon_df$megoutlier))
-#        0        1
-# 54675387 22393933
-summary(as.factor(oregon_df$megoutlier))/77069320 # percentage
-#         0         1
-# 0.7094313 0.2905687
-```
-
-## 62) *meglow*, low?
-False: (97.50%); True: (2.50%)
-logical
-```{r meglow}
-summary(as.factor(oregon_df$meglow)) 
-#    FALSE     TRUE
-# 75145616  1923704
-summary(as.factor(oregon_df$meglow))/77069320 # percentage
-#     FALSE      TRUE
-# 0.9750393 0.0249607
-```
-
-## 63) *meghigh*, high???
-False: (73.44%); True: (26.56%)
-logical
-```{r meghigh}
-summary(as.factor(oregon_df$meghigh))
-#    FALSE     TRUE
-# 56599091 20470229
-summary(as.factor(oregon_df$meghigh))/77069320 # percentage
-#    FALSE     TRUE
-# 0.734392 0.265608
-```
-
-## 64) *ndc*, National Drug Code
-[web](http://www.fda.gov/Drugs/InformationOnDrugs/ucm142438.htm)
-blank: (69.55%)
-character
-``` {r ndc}
-# ndc
-summary(as.numeric(oregon_df$ndc)) # blank: 53599072
-```
-
-## 65) *rxclass*, NDC therapeutic class
-blank: (69.55%); NULL: (2.31%)
-character
-```{r NDC therapeutic class}
-summary(as.factor(oregon_df$rxclass)) # blank: 53599072; NULL: 177765
-```
-
-## 66) *qtydisp*, Quantity dispensed
-NA: (2.37%)
-numeric
-```{r Quantity dispensed}
-summary(as.factor(as.character(oregon_df$qtydisp))) # NA: 182736
-```
-
-## 67) *rxdays*, Days supply
-NA: (69.57%)
-numeric
-```{r rxdays}
-summary(as.factor(oregon_df$rxdays)) # NA: 53619826
-```
-
-## 68) *daw*, Dispense as written code
-NA: (69.55%)
-numeric; range: 0-9
-```{r daw}
-summary(as.factor(oregon_df$daw)) # NA: 53599072
-```
-
-## 69) *ptstatus*, Discharge status
-blank: (52.61%); NULL: (35.39%)
-character
-```{r ptstatus}
-summary(as.factor(oregon_df$ptstatus)) # blank: 40545960; NULL: 27273631 
-```
-
-## 70) *los*, Length of stay
-NULL: (52.61%)
-numeric
-```{r los}
-summary(as.factor(oregon_df$los)) # NULL: 40545960
-```
-
-## 71) *msdrg*, MS-DRG
-blank: (52.98%); NULL: (44.04%)
-character
-```{r msdrg}
-summary(as.factor(oregon_df$msdrg)) # blank: 40828376; NULL: 33942845
-```
-
-## 72) *icd_ver*, icd version
-False: (68.19%); NA: (31.81%)
-logical
-```{r icd_ver}
-summary(as.factor(oregon_df$icd_ver))
-#    FALSE     NA's
-# 52556088 24513232
-summary(as.factor(oregon_df$icd_ver))/77069320 # percentage
-#     FALSE      NA's
-# 0.6819327 0.3180673
+# table ----
+knitr::kable(dx3_2, align = "c", caption = "dx3")
 
@@
 
-## Clean data
+## dx4
+```{r}
+dx4_2 <- oregon_df %>% 
+  select(dx4) %>%
+  mutate(dx4_cat = ifelse(is.na(dx4), "Missing",
+                     ifelse(!is.na(dx4), "Non-Missing", NA))) %>% 
+  group_by(dx4_cat) %>%
+  summarise(n = n()) %>%
+  mutate(total_n =sum(n),
+         percent = round(n/total_n, 4)*100) %>%
+      select(dx4_cat, n, percent)
 
-```{r }
-# Clean data, 76986701 left (maybe later)
-#oregon_full_df <- oregon_full %>%
-# filter non-missing value of year of birth, gender, race, STATE
-#  filter (is.na(yob)==F) %>%
-#  filter (gender!='U') %>%
-#  filter (is.na(race)==F) %>%
-#  filter (STATE!='')
-
-# Explore no dx1 and no ndc, see if the dx2 or other reasons, 1042984 obs
-# oregon_full_miss <- oregon_full %>%
-#  filter (dx1=='*NULL*') %>%
-#  filter (ndc=="")
-
-# fwrite(oregon_full_miss, "../data/oregon_full_miss.csv")
-
-# read_path1 <- paste0("./data/oregon_full_miss.csv")
-# oregon_miss <- fread(read_path1, sep = ",")
-
-# oregon_miss_df <- oregon_miss %>%
-#   filter (dx2!='*NULL*') # 981 obs
-
-# filter dx1, no ndc, 52490460
-# oregon_full_dx1 <- oregon_full_df %>%
-#   filter (dx1!='*NULL*')
-
-# filter ndc, no dx1, 23453379
-# oregon_full_ndc <- oregon_full_df %>%
-#  filter (ndc!="")
-
-# fwrite(oregon_full_dx1, "../data/oregon_full_dx1.csv")
-# fwrite(oregon_full_ndc, "../data/oregon_full_ndc.csv")
+# table ----
+knitr::kable(dx4_2, align = "c", caption = "dx4")
 
 ```
+
+## dx5
+```{r}
+dx5_2 <- oregon_df %>% 
+  select(dx5) %>%
+  mutate(dx5_cat = ifelse(is.na(dx5), "Missing",
+                     ifelse(!is.na(dx5), "Non-Missing", NA))) %>% 
+  group_by(dx5_cat) %>%
+  summarise(n = n()) %>%
+  mutate(total_n =sum(n),
+         percent = round(n/total_n, 4)*100) %>%
+      select(dx5_cat, n, percent)
+
+# table ----
+knitr::kable(dx5_2, align = "c", caption = "dx5")
+
+```

### 2017-02-27 r_scripts/oregon_clean.rmd
Update oregon_clean.rmd

@@
 knitr::kable(personkey2, align = "c", caption = "Patient ID")
+
+```
+
+## How many visits does unique personkey have?
+```{r}
+visit <- oregon_df %>%
+  select(personkey, clmid) %>%
+  mutate(num_visit = NA)
+
+for (i in 1: length(oregon_df$personkey)){
+  visit$num_visit[i] <- length(unique(visit$clmid[which(visit$personkey==visit$personkey[i])]))
+
+}
+
+a<-unique(oregon_df$personkey)
+
+visit2 <- visit %>%
+  select(personkey, num_visit) %>%
+  distinct(personkey, num_visit)
+
+visit3 <- visit2 %>%
+  group_by(num_visit) %>%
+  summarise(n = n()) %>%
+  mutate(total_n =sum(n),
+  percent = round(n/total_n, 4)*100) %>% 
+  select(num_visit, n, percent)
+
+m <- mean(visit2$num_visit)
+median(visit2$num_visit)
+m # 4.61
+
+knitr::kable(visit3, align = "c", caption = "Visits")
+
+hist(visit3$num_visit, br= 20)
+
+ggplot(visit3, aes(x=num_visit)) +
+  geom_density()
 

### 2017-03-01 r_scripts/oregon_clean.rmd

--- 2017-02-23 r_scripts/oregon_clean.rmd
@@
 library(readxl)
+library(ggplot2)
 getwd()
@@
 knitr::kable(personkey2, align = "c", caption = "Patient ID")
+
+```
+
+## How many visits does unique personkey have?
+```{r}
+# pick up personkey and claim id column, make new column of unique number of visit
+visit2 <- oregon_df %>%
+  select(personkey, clmid) %>%
+  distinct(clmid,personkey) %>%
+  select(personkey, clmid) %>%
+  group_by(personkey) %>%
+  summarise(num_visit=n())
+
+visit3 <- visit2 %>%
+  group_by(num_visit) %>%
+  summarise(n = n()) %>%
+  mutate(total_n =sum(n),
+  percent = round(n/total_n, 4)*100) %>% 
+  select(num_visit, n, percent)
+
+m <- cbind(sum(visit2$num_visit)/n_distinct(oregon_df$personkey), median(visit2$num_visit), max(visit2$num_visit))
+colnames(m) <- c("Mean", "Median", "Max")
+
+knitr::kable(m, align = "c", caption = "Statistics")
+
+hist(visit3$num_visit, freq = T, xlab = "number of visit", main = "Visiting number for each personkey")
 

### 2017-03-02 r_scripts/oregon_clean.rmd
updated summary markdown file to use whole dataset
change to whole data

@@
 # import data ------------------------------------------------------------------
-# Read in first 200,000 lines for exploration.
 
 read_path <- paste0("../data/gan_episodes_of_care.txt")
-oregon_df <- fread(read_path, sep = "|", nrows = 200000, showProgress = T)
-fwrite(oregon_df, "../data/oregon_epis_care_reduced.csv")
 
-# I wrote a reduced dataframe of 200,000 rows to get an idea of data structure
-
-# import the reduced dataframe
+# import the full dataframe, change column class to character
+read_path <- paste0("../data/gan_episodes_of_care.txt")
 start_time <- Sys.time()
-oregon_df <- fread("../data/oregon_epis_care_reduced.csv", sep = ",", 
+oregon_df <- fread(read_path, sep = "|", showProgress = T, 
                    colClasses = rep("character", 72)) 
 stop_time <-  Sys.time() - start_time 
-
 # time it took
-stop_time # 1.968219 secs
+stop_time # 20 mins
 

### 2017-03-02 r_scripts/oregon_clean.rmd
make new sentence

@@
 ---
-title: "Oregon_Wildfire_reduced"
+title: "Oregon_Wildfire_full"
 author: "Jingyang Liu"
@@
 # time it took
-stop_time # 20 mins
+stop_time # about 20 mins
 
@@
 72 variables in total.
+
 ```{r overall summary}
@@
 ```
+
+### Output html?

### 2017-03-02 r_scripts/oregon_markdown_excute.R
Script for excuting R markdown file

@@ -1,0 +1,12 @@
+# ------------------------------------------------------------------------------
+# Title: Script to excute descriptive summary R markdown file
+# Author: Jingyang Liu
+# Date: 3/2/2016
+# ------------------------------------------------------------------------------
+
+
+# call libraries
+library(rmarkdown)
+
+rmarkdown::render('./oregon_clean.rmd')
+

### 2017-03-02 r_scripts/oregon_clean.rmd
defined new file path
change directory

@@
 
-read_path <- paste0("../data/gan_episodes_of_care.txt")
+read_path <- paste0("../../../data/data_original/gan_episodes_of_care.txt")
 
@@
 ```{r}
-pos_key <- read_excel("../instructions/LT_pos.xlsx")
+pos_key <- read_excel("../LT_pos.xlsx")
 

### 2017-03-03 r_scripts/oregon_clean.rmd
modified markdown file. fread function was stopping due to missing | seperator. Added fill = T. Also added overall time to run program

@@
 
-read_path <- paste0("../../../data/data_original/gan_episodes_of_care.txt")
+read_path <- paste0("../data/gan_episodes_of_care.txt")
 
@@
 start_time <- Sys.time()
-oregon_df <- fread(read_path, sep = "|", showProgress = T, 
+oregon_df <- fread(read_path, sep = "|", showProgress = T,fill = T, 
                    colClasses = rep("character", 72)) 
 stop_time <-  Sys.time() - start_time
-# time it took
+
+# Note RG: fill = T allows fread to fill in any missing data
+
+# import time
 stop_time # about 20 mins
@@
 
-### Output html?
+## Total Time Markdown File Ran
+```{r}
+stop_time2 <-  Sys.time() - start_time 
+stop_time2
+```

### 2017-03-03 r_scripts/oregon_markdown_excute.R
updated r script to execute markdown file

@@ -8,5 +8,7 @@
 # call libraries
 library(rmarkdown)
 
+# call markdown doc.
 rmarkdown::render('./oregon_clean.rmd')
+
 

### 2017-03-03 r_scripts/oregon_data_wrangle_separate.R
Create oregon_data_wrangle_separate.R
small data clean;
small data separate into 12 diseases

@@ -1,0 +1,147 @@
+# ------------------------------------------------------------------------------
+# Title: Basic data wrangling  for 2013 oregon data
+# Author: Jingyang Liu
+# Date Created: Mar 03, 2016
+# R Version: 3.3.2
+# ------------------------------------------------------------------------------
+
+library(dplyr)
+library(data.table)
+library(readxl) 
+library(readr)
+
+# import data ------------------------------------------------------------------
+
+# Read in first 200000 lines for exploration.
+
+read_path <- paste0("./data/gan_episodes_of_care.txt")
+oregon_df <- fread(read_path, sep = "|", nrows = 200000, showProgress = T)
+fwrite(oregon_df, "./data/oregon_epis_care_reduced.csv")
+
+# I wrote a reduced dataframe of 200,000 rows to get an idea of data structure
+
+
+# Define read paths ------------ -----------------------------------------------
+# import r list of outcome-specific icd9 codes
+load_path <- paste0("./data/outcome_list.RData")
+load(load_path)
+
+# path and file name of data
+read_path <- paste0("./data/oregon_epis_care_reduced.csv")
+start_time <- Sys.time()
+oregon_df <- fread(read_path, sep = ",", 
+                   colClasses = rep("character", 72)) 
+stop_time <-  Sys.time() - start_time 
+# time it took
+stop_time # 20 mins
+
+## 
+
+oregon_df$yob <- as.numeric(oregon_df$yob)
+oregon_df[oregon_df=="*NULL*"] <- NA
+
+### data rough cleaning (remove some NA and blank values) and add index---------
+# at least a primary ICD9 code present, urgent or ED visit
+# whole data still have 1514387 observations after cleaning
+# small data 2926 left
+or_hosp_w_outcome_df <- oregon_df %>% filter(!is.na(dx1)) %>% 
+  # filter sex is known
+  filter(gender!="U") %>%
+  # add age, filter age not missing
+  mutate(age = 2013-yob) %>%
+  filter(!is.na(age)) %>%
+  # filter race
+  filter(race!="") %>%
+  # filter enthnicity
+  filter(ethn!="") %>%
+  # filter State is Oregon
+  filter(STATE=="OR") %>%
+  # filter ZIP non missing
+  filter(!is.na(ZIP)) %>%
+  # filter MSA
+  filter(!is.na(MSA)) %>%
+  # filter POS
+  filter(!is.na(pos)) %>%
[…]

### 2017-03-08 r_scripts/oregon_casecrossover.R
Create oregon_casecrossover.R

@@ -1,0 +1,232 @@
+# ------------------------------------------------------------------------------
+# Title: Case crossover for diseases (Oregon State) for server
+# Author: Jingyang Liu
+# Date Created: Feb 24, 2017
+# R version: 3.3.2
+# ------------------------------------------------------------------------------
+
+#Creating case crossover dataframes
+library(tidyverse)
+library(data.table)
+library(lubridate)
+getwd()
+setwd("C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire/data_disease")
+
+
+read_path <- paste0('../data_new/or_hosp_w_outcome_df.csv')
+disease <- read_csv(read_path) 
+
+disease <- disease %>%
+  # add new transverted from date
+  mutate(dates = as.Date(fromdate, "%m/%d/%Y"))
+
+
+
+### ---------------------------------------------------------------------
+var_list <- c('resp', 'asthma', 'pneum',  'acute_bronch', 'copd', 'cvd', 
+              'ihd', 'arrhythmia', 'hf', 'cereb_vas', 'mi', 'broken_arm')
+
+
+for (i in var_list){
+  i <- c('asthma')
+  read_path <- paste('oregon', i, 'claims_2013.csv', sep='_')
+  or_disease <- read_csv( read_path)
+    
+  or_disease <- or_disease %>% 
+    group_by(personkey) %>% 
+    arrange(personkey, fromdate, line) %>% 
+    mutate(num_visit = dense_rank(fromdate)) %>%
+    arrange(personkey, num_visit) %>% 
+    select(personkey, clmid, num_visit) %>% 
+    unique() %>% 
+    full_join(or_disease, by = c("personkey", "clmid")) %>% 
+    arrange(personkey, clmid, line, fromdate) %>% 
+    filter(row_number() == 1) %>%
+    mutate(dates = as.Date(fromdate, "%m/%d/%Y")) %>%
+    filter(dates >= '2013-07-01' & 
+           dates <= '2013-10-31')
+  
+ 
+  
+}
+
+
+
+# Asthma1 ----------------------------------------------------------------------
+disease_asthma1 <- disease %>%
+  # filter asthma1
+  filter(asthma1==1) %>%
+  # filter date in July 1, 2012 and Oct 31, 2012
+  filter(dates >= '2013-07-01' & 
[…]

### 2017-03-08 r_scripts/oregon_clean.rmd
Path to read excel file for place of service code was redefined
Update oregon_clean.rmd
change to right density plot

--- 2017-03-02 r_scripts/oregon_clean.rmd
@@
 read_path <- paste0("../../../data/data_original/gan_episodes_of_care.txt")
-
-# import the full dataframe, change column class to character
-read_path <- paste0("../data/gan_episodes_of_care.txt")
 start_time <- Sys.time()
@@
 
-hist(visit3$num_visit, freq = T, xlab = "number of visit", main = "Visiting number for each personkey")
+ggplot(visit3, aes(n)) + geom_density()
 
@@
 ```
-
-### Output html?

### 2017-03-08 r_scripts/oregon_data_wrangle_separate.R
Update oregon_data_wrangle_separate.R
change filtering for only dx1

@@ -12,82 +12,58 @@ library(readr)
 
 # import data ------------------------------------------------------------------
 
-# Read in first 200000 lines for exploration.
+read_path <- paste0("../../../data/data_original/gan_episodes_of_care.txt")
+start_time <- Sys.time()
+oregon_df <- fread(read_path, sep = "|", showProgress = T) 
+stop_time <-  Sys.time() - start_time 
+# time it took
+stop_time # 7 mins
 
-read_path <- paste0("./data/gan_episodes_of_care.txt")
-oregon_df <- fread(read_path, sep = "|", nrows = 200000, showProgress = T)
-fwrite(oregon_df, "./data/oregon_epis_care_reduced.csv")
+# basic cleaning ---------------------------------------------------------------
+start_time <- Sys.time()
+oregon_df <- oregon_df %>% filter(!is.na(dx1)) %>% 
+  # filter State is Oregon
+  filter(STATE=="OR") %>%
+  # filter urgent or ED visits (pos = 20 or 23)
+  filter(pos=="20"|pos=="23") 
+stop_time <-  Sys.time() - start_time 
+# time it took
+stop_time # 2.5 mins
 
-# I wrote a reduced dataframe of 200,000 rows to get an idea of data structure
+write_path <- paste0('../../../data/data_original/',
+                     'oregon_epis_care_reduced.csv')
+write_csv(oregon_df, write_path)
+
 
 
 # Define read paths ------------ -----------------------------------------------
 # import r list of outcome-specific icd9 codes
-load_path <- paste0("./data/outcome_list.RData")
+# 1514642 obs
+load_path <- paste0("../../../data/data_original/outcome_list.RData")
 load(load_path)
 
 # path and file name of data
-read_path <- paste0("./data/oregon_epis_care_reduced.csv")
+read_path <- paste0("../../../data/data_original/oregon_epis_care_reduced.csv")
 start_time <- Sys.time()
 oregon_df <- fread(read_path, sep = ",", 
                    colClasses = rep("character", 72)) 
 stop_time <-  Sys.time() - start_time 
 # time it took
-stop_time # 20 mins
+stop_time # 14 sec
 
-## 
-
-oregon_df$yob <- as.numeric(oregon_df$yob)
-oregon_df[oregon_df=="*NULL*"] <- NA
-
-### data rough cleaning (remove some NA and blank values) and add index---------
[…]

### 2017-03-08 r_scripts/oregon_casecrossover.R
Update oregon_casecrossover.R
case crossover for one sample, filter data set

@@ -8,7 +8,7 @@
 #Creating case crossover dataframes
 library(tidyverse)
 library(data.table)
-library(lubridate)
+library(readxl)
 getwd()
 setwd("C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire/data_disease")
 
@@ -23,12 +23,12 @@ disease <- disease %>%
 
 
 ### ---------------------------------------------------------------------
-var_list <- c('resp', 'asthma', 'pneum',  'acute_bronch', 'copd', 'cvd', 
-              'ihd', 'arrhythmia', 'hf', 'cereb_vas', 'mi', 'broken_arm')
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
 
-
+start <- Sys.time()
 for (i in var_list){
-  i <- c('asthma')
   read_path <- paste('oregon', i, 'claims_2013.csv', sep='_')
   or_disease <- read_csv( read_path)
     
@@ -44,33 +44,137 @@ for (i in var_list){
     filter(row_number() == 1) %>%
     mutate(dates = as.Date(fromdate, "%m/%d/%Y")) %>%
     filter(dates >= '2013-07-01' & 
-           dates <= '2013-10-31')
+           dates <= '2013-10-31') %>%
+    mutate(outcome = 1)
   
- 
+  # Create a permanent case-cross over dataset
+  file_name <- paste('oregon', i, 'jul_to_oct_claim.csv', sep = '_')
+  
+  # write permanent dataset
+  write_csv(or_disease, paste0("../data_new/", file_name))
   
 }
+total_time <- Sys.time() - start
+total_time # Time difference of 5 mins
+
+setwd("../data_new/")
+
+## Case crossover for each disease
+var_list2 <- c('resp1', 'asthma1', 'pneum1',  'acute_bronch1', 'copd1', 'cvd1', 
+              'ihd1', 'arrhythmia1', 'hf1', 'cereb_vas1', 'mi1', 'broken_arm1')
+
+start <- Sys.time()
+for(m in var_list2){ # begin first loop of variable names (outcomes)
+  
+  # Case-Crossover loop --------------------------------------------------------
[…]

### 2017-03-09 r_scripts/oregon_casecrossover.R
Update oregon_casecrossover.R
write full loop case crossover study over all diseases

@@ -12,16 +12,6 @@ library(readxl)
 getwd()
 setwd("C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire/data_disease")
 
-
-read_path <- paste0('../data_new/or_hosp_w_outcome_df.csv')
-disease <- read_csv(read_path) 
-
-disease <- disease %>%
-  # add new transverted from date
-  mutate(dates = as.Date(fromdate, "%m/%d/%Y"))
-
-
-
 ### ---------------------------------------------------------------------
 var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
               'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
@@ -42,10 +32,10 @@ for (i in var_list){
     full_join(or_disease, by = c("personkey", "clmid")) %>% 
     arrange(personkey, clmid, line, fromdate) %>% 
     filter(row_number() == 1) %>%
+    # add new transverted from date
     mutate(dates = as.Date(fromdate, "%m/%d/%Y")) %>%
     filter(dates >= '2013-07-01' & 
-           dates <= '2013-10-31') %>%
-    mutate(outcome = 1)
+           dates <= '2013-10-31') 
   
   # Create a permanent case-cross over dataset
   file_name <- paste('oregon', i, 'jul_to_oct_claim.csv', sep = '_')
@@ -59,43 +49,30 @@ total_time # Time difference of 5 mins
 
 setwd("../data_new/")
 
+
+
 ## Case crossover for each disease
-var_list2 <- c('resp1', 'asthma1', 'pneum1',  'acute_bronch1', 'copd1', 'cvd1', 
-              'ihd1', 'arrhythmia1', 'hf1', 'cereb_vas1', 'mi1', 'broken_arm1')
 
 start <- Sys.time()
-for(m in var_list2){ # begin first loop of variable names (outcomes)
-  
-  # Case-Crossover loop --------------------------------------------------------
-  outcome_col <- which(colnames(disease) == m) # use to keep outcome var
+for(m in var_list){ # begin first loop of variable names (outcomes)
+
+  read_path <- paste('oregon', m, 'jul_to_oct_claim.csv', sep='_')
+  disease <- read_csv( read_path)
   
   outcome_id <- disease %>%
-    # filter total outcome for jth disease
[…]

### 2017-03-22 r_scripts/oregon_proportion_server.R
Run mcmapply in server, save time about 13 mins

@@ -12,8 +12,9 @@ library(dplyr)
 library(data.table)
 library(readxl)
 # parallel computing libraries
-library(foreach) 
+# library(foreach) 
 library(doParallel)
+library(parallel)
 library(readr)
 
 getwd()
@@ -26,7 +27,7 @@ grid_dir <- paste0('../../../data/data_new/or_shapefile/oregon_grid.shp')
 smoke_grid <- readOGR(dsn = grid_dir, layer = 'oregon_grid') 
 
 summary(smoke_grid) 
-summary(smoke_grid@data$WRFGRID_ID)
+# summary(smoke_grid@data$WRFGRID_ID)
 
 # Zipcode shapefile
 shp_dir <- paste0('../../../data/data_original/', 
@@ -42,12 +43,8 @@ read_path <- paste0('../../../data/data_original/oregon_zipcode.csv')
 or_zip <- read_csv(read_path)
 names(or_zip) <- c('zip','type','city','county','area')
 
-# check on zip code in colorado range
-# removing anyways
-
-# limit to just Colorado state zipcodes
+# limit to just Oregon state zipcodes
 or_zip_map <- us_zip_2013[us_zip_2013$ZCTA5CE10 %in% or_zip$zip,]
-
 
 # output zipcodes from washington zipcode map to bind values to
 # plot map to check
@@ -58,13 +55,12 @@ plot(or_zip_map)
 summary(or_zip_map)
 or_zip <- as.character(sort(or_zip_map@data$ZCTA5CE10))
 
-
 # save the Oregon zips to a shapefile to use later
 # create save path
+# !! will show error when overlay the writing file, so please write one time
 save_path <- paste0('../../../data/data_new/or_zip_2013_shape_files')
 
 writeOGR(or_zip_map, layer = 'or_zip_2013_shape_files', save_path, driver = "ESRI Shapefile")
-
 
 # Set coordinate reference system for smoke gird
 nad83 <- '+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0'
@@ -76,10 +72,12 @@ plot(or_zip_map, add=T)
 # looks like they overlay pretty well, same projections
 summary(smoke_grid)
 
+
+
 # Test code to figure out proportion calculations in each WRF-Grid -------------
 # Trying 'over' function in sp package
 # limit to a specific zip code
[…]

### 2017-03-23 r_scripts/parallel_presentation.Rmd
Add files via upload
Presentation for Parallel

---
title: "Parallel for Oregon Zip and WRF_Grid Intersection Proportion"
author: "Jingyang Liu"
date: "March 23, 2017"
output: slidy_presentation
---

```{r library setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(rgdal) # package for shape files
library(sp)
library(rgeos) # rgeos package contains the intersect and area commands I need
library(dplyr)
library(data.table)
library(readxl)
# parallel computing libraries
# library(foreach) 
library(doParallel)
library(parallel)
library(readr)
```

```{r shapefile prepare, include=FALSE}
## Imoport WRF Grid file of Oregon
setwd("C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire")
grid_dir <- paste0('./shapefile/oregon_grid.shp')
smoke_grid <- readOGR(dsn = grid_dir, layer = 'oregon_grid') # wrf_grid

## Import U.S. shapefile
shp_dir <- paste0('./shapefile/tl_2013_us_zcta510/tl_2013_us_zcta510.shp')
us_zip_2013 <- readOGR(dsn = shp_dir, layer = 'tl_2013_us_zcta510')

## Filter the U.S. shapefile only to Oregon State Zip Code
read_path2 <- paste0('C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire/', 
                     'oregon_zipcode.csv')
or_zip <- read_csv(read_path2)
names(or_zip) <- c('zip','type','city','county','area')
or_zip_map <- us_zip_2013[us_zip_2013$ZCTA5CE10 %in% or_zip$zip,]
or_zip <- as.character(sort(or_zip_map@data$ZCTA5CE10))

nad83 <- '+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0'
proj4string(smoke_grid) <- nad83
```

## Shapefile preparation: 2013 Oregon Zip code Shapefile and WRF Grid
The shapefile of Oregon can be subset from the whole U.S. shapefile by filtering the zip code in Oregon State.
The whole U.S shapefile can be found from United States Census website:
<br/>https://www.census.gov/cgi-bin/geo/shapefiles/index.php?year=2013&layergroup=ZIP+Code+Tabulation+Areas
<br/>The Oregon zip code can also be found from: 
<br/>https://www.unitedstateszipcodes.org/or/

The plot below shows a Oregon State shapefile and an overlay of projected WRF Grid and Oregon shapefile.
```{r overlay}
plot(or_zip_map)
plot(smoke_grid)
plot(or_zip_map, add=T)
```

The Oregon shapefile contains some lack or extra, which may because the zip code changes (but U.S. shapefile is for 2013).

## Steps

## Why using for loop?

<ul class="incremental"> 
  <li>One choice: <br/>  simple calculation; copy and paste</li> 
  <li>Another choice: <br/>  for loop</li> 
</ul>

## Apply Family

- The apply() Function
- The lapply() Function
- The sapply() Function
- The tapply() Function
- The mapply() Function

## Parallel on server

```{r cars, echo = TRUE}
summary(cars)
```

### 2017-03-24 r_scripts/parallel_presentation.Rmd
Update parallel_presentation.Rmd

@@
 ## Steps
+Choose zip code = 94705. WRF Grid = 261 and 262.
+```{r test}
+test_zip <- c(97405)
+test_zip_map <- or_zip_map[or_zip_map$ZCTA5CE10 %in% test_zip,]
+
+plot(test_zip_map)
+plot(smoke_grid, add = T)
+invisible(text(getSpPPolygonsLabptSlots(smoke_grid), 
+               labels=as.character(smoke_grid$WRFGRID_ID)))
+
+```
+
+```{r}
+test_grid <- over(smoke_grid, test_zip_map)
+summary(test_grid) # 14 grids over the zip code
+test_grid2 <- test_grid %>% filter(!is.na(ZCTA5CE10))
+test_grid2
+
+# test zip map over zip grid
+plot(smoke_grid)
+plot(test_zip_map, add = T)
+zip_over_grid <- over(test_zip_map, smoke_grid)
+summary(zip_over_grid) # this way retains the values
+
+shape_zip <- SpatialPolygons(test_zip_map@polygons)
+shape_grid <- SpatialPolygons(smoke_grid@polygons)
+plot(shape_zip)
+plot(shape_grid, add = T)
+# try and plot the values for this zipcode for 9/21
+invisible(text(getSpPPolygonsLabptSlots(smoke_grid), 
+               labels=as.character(smoke_grid$WRFGRID_ID)))
+
+# output the WRF Grid 261 and 262 for test
+wrf_grid_261 <- smoke_grid[smoke_grid@data$WRFGRID_ID == 261, ]
+plot(wrf_grid_261)
+
+wrf_grid_262 <- smoke_grid[smoke_grid@data$WRFGRID_ID == 262, ]
+plot(wrf_grid_262)
+
+# calculate the area of grid 297
+gArea(SpatialPolygons(wrf_grid_261@polygons))
+gArea(SpatialPolygons(wrf_grid_262@polygons))
+
+# Subset to one zip code and two different WRF grids ---------------------------
+# plot zip and 2 grids
+plot(test_zip_map)
+#invisible(text(getSpPPolygonsLabptSlots(test_zip_map), 
+#               labels=as.character(test_zip_map$ZCTA5CE10)))
+plot(wrf_grid_261, add = T)
+invisible(text(getSpPPolygonsLabptSlots(wrf_grid_261), 
+               labels=as.character(wrf_grid_261$WRFGRID_ID)))
+plot(wrf_grid_262, add=T)
+invisible(text(getSpPPolygonsLabptSlots(wrf_grid_262), 
+               labels=as.character(wrf_grid_262$WRFGRID_ID)))
+
+# Start with the intersection with wrf grid 297 and zipcode
+# first I need to convert the spatial polygon to just polygon
+poly_261 <- SpatialPolygons(wrf_grid_261@polygons)
+shape_zip <- SpatialPolygons(test_zip_map@polygons)
+
+zip_261_int <- gIntersection(poly_261, shape_zip)
+
+plot(zip_261_int)
+prop_261_int <- gArea(zip_261_int)/gArea(poly_261)
+prop_261_int # 41.8% of grid is covered by zip
+
+# now what about grid 296
+poly_262 <- SpatialPolygons(wrf_grid_262@polygons)
+shape_zip <- SpatialPolygons(test_zip_map@polygons)
+
+zip_262_int <- gIntersection(poly_262, shape_zip)
+
+plot(zip_262_int)
+prop_int_262 <- gArea(zip_296_int)/gArea(poly_262)
+prop_int_262 # 6.0% of grid is covered by zip
+
+```
 

### 2017-03-27 r_scripts/parallel_presentation.Rmd
Update parallel_presentation.Rmd

--- 2017-03-23 r_scripts/parallel_presentation.Rmd
@@
 The shapefile of Oregon can be subset from the whole U.S. shapefile by filtering the zip code in Oregon State.
-The whole U.S shapefile can be found from United States Census website:
+<br/>The whole U.S shapefile can be found from United States Census website:
 <br/>https://www.census.gov/cgi-bin/geo/shapefiles/index.php?year=2013&layergroup=ZIP+Code+Tabulation+Areas
@@
 
-## Steps
+## Steps of calculating the proportion of a specific zip code polygon and WRF Grid.
+1. Choose zip code = 94705. WRF Grid = 261 and 262. (6 WRF Grid in total. Only 
+two are chosen for test)
+```{r test, warning= FALSE}
+
+test_zip <- c(97405)
+test_zip_map <- or_zip_map[or_zip_map$ZCTA5CE10 %in% test_zip,]
+
+plot(test_zip_map)
+plot(smoke_grid, add = T)
+invisible(text(getSpPPolygonsLabptSlots(smoke_grid), 
+               labels=as.character(smoke_grid$WRFGRID_ID)))
+
+```
+
+```{r include = FALSE}
+test_grid <- over(smoke_grid, test_zip_map)
+summary(test_grid) # 6 grids over the zip code
+test_grid2 <- test_grid %>% filter(!is.na(ZCTA5CE10))
+test_grid2
+```
+
+2. Show the actual location in the map.
+```{r}
+# test zip map over zip grid
+plot(smoke_grid)
+plot(test_zip_map, add = T)
+zip_over_grid <- over(test_zip_map, smoke_grid)
+# summary(zip_over_grid) # this way retains the values
+```
+
+```{r include = FALSE}
+# test_zip <- c(97405)
+# test_zip_map <- or_zip_map[or_zip_map$ZCTA5CE10 %in% test_zip,]
+
+# plot(test_zip_map)
+# plot(smoke_grid, add = T)
+# invisible(text(getSpPPolygonsLabptSlots(smoke_grid), 
+#                labels=as.character(smoke_grid$WRFGRID_ID)))
+
+shape_zip <- SpatialPolygons(test_zip_map@polygons)
+shape_grid <- SpatialPolygons(smoke_grid@polygons)
+plot(shape_zip)
+plot(shape_grid, add = T)
+# try and plot the values for this zipcode
+invisible(text(getSpPPolygonsLabptSlots(smoke_grid), 
+               labels=as.character(smoke_grid$WRFGRID_ID))) # grid width wider than previous map
+```
+
+```{r include = FALSE}
+# output the WRF Grid 261 and 262 for test
+wrf_grid_261 <- smoke_grid[smoke_grid@data$WRFGRID_ID == 261, ]
+plot(wrf_grid_261)
+
+wrf_grid_262 <- smoke_grid[smoke_grid@data$WRFGRID_ID == 262, ]
+plot(wrf_grid_262)
+```
+
+```{r include = FALSE}
+# calculate the area of grid 261 and 262
+gArea(SpatialPolygons(wrf_grid_261@polygons))
+gArea(SpatialPolygons(wrf_grid_262@polygons))
+```
+
+## Steps (Continued)
+3. Show the intersection of chosen WRF Grid and zip code.
+```{r show, warning=FALSE}
+# Subset to one zip code and two different WRF grids ---------------------------
+# plot zip and 2 grids
+plot(test_zip_map)
+#invisible(text(getSpPPolygonsLabptSlots(test_zip_map), 
+#               labels=as.character(test_zip_map$ZCTA5CE10)))
+plot(wrf_grid_261, add = T)
+invisible(text(getSpPPolygonsLabptSlots(wrf_grid_261), 
+               labels=as.character(wrf_grid_261$WRFGRID_ID)))
+plot(wrf_grid_262, add=T)
+invisible(text(getSpPPolygonsLabptSlots(wrf_grid_262), 
+               labels=as.character(wrf_grid_262$WRFGRID_ID)))
+```
+
+4. The proportion of two WRF Grid.
+```{r calculation, warning=FALSE}
+# Start with the intersection with wrf grid 261 and zipcode
+# first I need to convert the spatial polygon to just polygon
+poly_261 <- SpatialPolygons(wrf_grid_261@polygons)
+shape_zip <- SpatialPolygons(test_zip_map@polygons)
+
+zip_261_int <- gIntersection(poly_261, shape_zip)
+plot(zip_261_int)
+plot(wrf_grid_261, add = T)
+invisible(text(getSpPPolygonsLabptSlots(wrf_grid_261), 
+               labels=as.character(wrf_grid_261$WRFGRID_ID)))
+
+prop_261_int <- gArea(zip_261_int)/gArea(poly_261)
+prop_261_int # 41.8% of grid is covered by zip
+
+# now what about grid 262
+poly_262 <- SpatialPolygons(wrf_grid_262@polygons)
+shape_zip <- SpatialPolygons(test_zip_map@polygons)
+
+zip_262_int <- gIntersection(poly_262, shape_zip)
+
+plot(zip_262_int)
+plot(wrf_grid_262, add = T)
+invisible(text(getSpPPolygonsLabptSlots(wrf_grid_261), 
+               labels=as.character(wrf_grid_261$WRFGRID_ID)))
+
+prop_int_262 <- gArea(zip_262_int)/gArea(poly_262)
+prop_int_262 # 6.0% of grid is covered by zip
+```
 
@@
 <ul class="incremental"> 
-  <li>One choice: <br/>  simple calculation; copy and paste</li> 
-  <li>Another choice: <br/>  for loop</li> 
+  <li> One choice: <br/>  simple calculation; copy and paste</li> 
+  <li> Another choice: for loop
+  <br/> When analyzing big data set (like millions of observations), it is 
+  unrealistic to copy so many rows of similar or same code.</li><br/>
+  <li> And... <br/> Why using apply function? </li>
+  <li> Code Readability <br/>  Understand the purporse more directly.</li> 
+  <li> Working Efficiency <br/>  When analyzing big data set (like millions of 
+  observations), it is unrealistic to copy so many rows of similar or same code.</li> 
 </ul>
@@
 - The apply() Function
+<br/> Apply functions over array margins.<br/>
+
 - The lapply() Function
+<br/> Apply a function over a list or vector.<br/> 
+Apply a given function to every element of a list and obtain a list as result.<br/>
+
 - The sapply() Function
+<br/> Apply a function over a list or vector.<br/>
+Works like lapply(), but it tries to simplify the output to the most elementary data structure that is possible.
+
 - The tapply() Function
+<br/> Apply a function over a ragged array.
+
 - The mapply() Function
+<br/> Apply a function to multiple list or vector arguments.
 
 ## Parallel on server
+One core vs Multiple cores
 
-```{r cars, echo = TRUE}
-summary(cars)
+```{r}
+
+for(i in 1:length(or_zip_name)){
+  # output value of zipcode
+  zipcode <- as.character(or_zip_name[i]) 
+  # limit shapefile to particular zipcode
+  zip_shape <- or_zip_map[or_zip_map$ZCTA5CE10 %in% zipcode, ]
+  # convert to polygon
+  zip_poly <-SpatialPolygons(zip_shape@polygons)
+
+  # now I can create the second loop that finds the proportion of the area of
+  # the zipcode polygon that overlaps with each WRF-Grid
+  for(j in 1:length(wrf_grid_name)){
+    # output each grid and create a polygon
+    wrf_grid <- smoke_grid[smoke_grid@data$WRFGRID_ID == j, ]
+    # now what about grid 719; should be much less
+    wrf_poly <- SpatialPolygons(wrf_grid@polygons)
+
+    zip_wrf_intersect <- gIntersection(wrf_poly, zip_poly)
+    # if empty, then set to 0, else find the proportion
+    grid_prop <- ifelse(is.null(zip_wrf_intersect),
+                        0, gArea(zip_wrf_intersect)/gArea(wrf_poly))
+    # populate the matrix based on i position and j position
+    zip_wrf_proportion[[i,j]] <- grid_prop
+  }
+}
+
 ```
+
+```{r, eval=FALSE}
+f1 <- function(x){
+  zipcode <- as.character(x) 
+  # limit shapefile to particular zipcode
+  zip_shape <- or_zip_map[or_zip_map$ZCTA5CE10 %in% zipcode, ]
+  # convert to polygon
+  zip_poly <-SpatialPolygons(zip_shape@polygons)
+}
+zip_polygon <- lapply(or_zip_name, f1) # list
+# b <- zip_polygon[1] # list
+# c <- zip_polygon[[1]] # spatial polygon
+
+vars1 <- c(1:1610)
+f2 <- function(n){
+  wrf_grid <- smoke_grid[smoke_grid@data$WRFGRID_ID == n,] # sp df
+  # now what about grid 719; should be much less
+  wrf_poly <- SpatialPolygons(wrf_grid@polygons) # sp
+}
+wrf_polygon <- lapply(vars1, f2) # list of 1610 wrf_grid
+
+f3 <- function(x,y){
+  zip_wrf_intersect <- gIntersection(wrf_polygon[y][[1]], zip_polygon[x][[1]])
+  grid_prop <- ifelse(is.null(zip_wrf_intersect),
+                      0, gArea(zip_wrf_intersect)/gArea(wrf_polygon[y][[1]]))
+
+}
+proportion <- mcmapply(f3, rep(1:417, each=1610), rep(1:1610,417))
+
+```

### 2017-03-28 r_scripts/parallel_presentation.Rmd
Update parallel_presentation.Rmd

@@
 
-```{r library setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
+```{r librarysetup, include=FALSE}
 library(rgdal) # package for shape files
@@
 ## Shapefile preparation: 2013 Oregon Zip code Shapefile and WRF Grid
-The shapefile of Oregon can be subset from the whole U.S. shapefile by filtering the zip code in Oregon State.
-<br/>The whole U.S shapefile can be found from United States Census website:
+
+Some libraries for spatial data analysis and parallel computing:
+```{r, eval = FALSE}
+library(rgdal)     # package for shape files
+library(sp)        # spatial polygon
+library(rgeos)     # rgeos package contains the intersect and area commands
+library(parallel)  # parallel computing libraries
+```
+
+The **shapefile of Oregon** can be subset from the whole U.S. shapefile by filtering the zip code in Oregon State.
+
+<br/>The **whole U.S shapefile** can be found from United States Census website:
 <br/>https://www.census.gov/cgi-bin/geo/shapefiles/index.php?year=2013&layergroup=ZIP+Code+Tabulation+Areas
-<br/>The Oregon zip code can also be found from: 
+
+<br/>The **Oregon zip code** can also be found from: 
 <br/>https://www.unitedstateszipcodes.org/or/
 
-The plot below shows a Oregon State shapefile and an overlay of projected WRF Grid and Oregon shapefile.
-```{r overlay}
+The plot below shows an Oregon State shapefile and an overlay of projected WRF Grid and Oregon shapefile.
+```{r overlay, echo=FALSE}
 plot(or_zip_map)
@@
 
-The Oregon shapefile contains some lack or extra, which may because the zip code changes (but U.S. shapefile is for 2013).
-
-## Steps of calculating the proportion of a specific zip code polygon and WRF Grid.
-1. Choose zip code = 94705. WRF Grid = 261 and 262. (6 WRF Grid in total. Only 
-two are chosen for test)
-```{r test, warning= FALSE}
-
+## Steps of calculating the proportion of intersection of a specific zip code polygon and WRF Grid.
+1. Choose a zip code (94705 here) for testing. WRF Grid = 261 and 262. (6 WRF 
+Grids in total. Only two are chosen for convenience)
+```{r test, warning= FALSE, echo=FALSE}
 test_zip <- c(97405)
@@
 
-```{r include = FALSE}
-test_grid <- over(smoke_grid, test_zip_map)
-summary(test_grid) # 6 grids over the zip code
-test_grid2 <- test_grid %>% filter(!is.na(ZCTA5CE10))
-test_grid2
-```
-
-2. Show the actual location in the map.
-```{r}
+2. The actual location in the map.
+```{r, echo=FALSE}
 # test zip map over zip grid
@@
 3. Show the intersection of chosen WRF Grid and zip code.
-```{r show, warning=FALSE}
+```{r show, warning=FALSE, echo=FALSE}
 # Subset to one zip code and two different WRF grids ---------------------------
@@
 4. The proportion of two WRF Grid.
-```{r calculation, warning=FALSE}
+```{r calculation, warning=FALSE, echo=FALSE}
 # Start with the intersection with wrf grid 261 and zipcode
@@
 <ul class="incremental"> 
-  <li> One choice: <br/>  simple calculation; copy and paste</li> 
+  <li> One choice: <br/>  simple calculation; copy and paste codes.</li> 
   <li> Another choice: for loop
-  <br/> When analyzing big data set (like millions of observations), it is 
-  unrealistic to copy so many rows of similar or same code.</li><br/>
+  <br/> When analyzing big data set (for this data set, there are 417 zip codes 
+  and 1610 WRF Grids in total), it is unrealistic to copy so many rows 
+  (417*1610 = 671,370 times) of similar or same code.</li><br/>
   <li> And... <br/> Why using apply function? </li>
-  <li> Code Readability <br/>  Understand the purporse more directly.</li> 
-  <li> Working Efficiency <br/>  When analyzing big data set (like millions of 
-  observations), it is unrealistic to copy so many rows of similar or same code.</li> 
+  <li> **Working Efficiency** <br/> **Vectorization.** Faster than for loop. It 
+  works on margins of an array or matrix, not like for loop.</li> 
+  <li> **Code Readability** <br/>  Understand the purporse more directly.</li> 
 </ul>
@@
 ## Apply Family
+>- Returns a vector or array or list of values obtained by applying a function to 
+margins of an array or matrix.
 
-- The apply() Function
+>- The apply() Function
 <br/> Apply functions over array margins.<br/>
 
-- The lapply() Function
+>- The lapply() Function
 <br/> Apply a function over a list or vector.<br/> 
@@
 
-- The sapply() Function
+>- The sapply() Function
 <br/> Apply a function over a list or vector.<br/>
@@
 
-- The tapply() Function
+>- The tapply() Function
 <br/> Apply a function over a ragged array.
 
-- The mapply() Function
+>- The mapply() Function
 <br/> Apply a function to multiple list or vector arguments.
@@
 ## Parallel on server
-One core vs Multiple cores
+For loop vs Apply
 
-```{r}
-
+```{r, eval=FALSE}
 for(i in 1:length(or_zip_name)){
@@
 ```
+
+One core vs Multiple cores

### 2017-03-29 r_scripts/parallel_presentation_excute.R
Add files via upload
Run rmarkdown

@@ -1,0 +1,14 @@
+# ------------------------------------------------------------------------------
+# Title: Script to excute descriptive summary R markdown file
+# Author: Jingyang Liu
+# Date: Mar 29, 2017
+# ------------------------------------------------------------------------------
+
+
+# call libraries
+library(rmarkdown)
+
+# call markdown doc.
+rmarkdown::render('./parallel_presentation.Rmd')
+
+

### 2017-03-29 r_scripts/oregon_proportion_server.R
revise for wrong

@@ -242,7 +242,7 @@ write_csv(zip_proportion_new_df, paste0(write_path))
 start <- Sys.time()
 
 # first I want to subset out each zipcode shapefile
-for (i in 201:210) %dopar% {
+for (i in 201:210) {
   # output value of zipcode
   zipcode <- as.character(or_zip_name[i]) 
   # limit shapefile to particular zipcode

### 2017-03-29 r_scripts/parallel_presentation.Rmd
smaller sample

@@
 <br/> Apply a function to multiple list or vector arguments.
-
-## Parallel on server
-For loop vs Apply
-
-```{r, eval=FALSE}
-for(i in 1:length(or_zip_name)){
-  # output value of zipcode
-  zipcode <- as.character(or_zip_name[i]) 
-  # limit shapefile to particular zipcode
-  zip_shape <- or_zip_map[or_zip_map$ZCTA5CE10 %in% zipcode, ]
-  # convert to polygon
-  zip_poly <-SpatialPolygons(zip_shape@polygons)
-
-  # now I can create the second loop that finds the proportion of the area of
-  # the zipcode polygon that overlaps with each WRF-Grid
-  for(j in 1:length(wrf_grid_name)){
-    # output each grid and create a polygon
-    wrf_grid <- smoke_grid[smoke_grid@data$WRFGRID_ID == j, ]
-    # now what about grid 719; should be much less
-    wrf_poly <- SpatialPolygons(wrf_grid@polygons)
-
-    zip_wrf_intersect <- gIntersection(wrf_poly, zip_poly)
-    # if empty, then set to 0, else find the proportion
-    grid_prop <- ifelse(is.null(zip_wrf_intersect),
-                        0, gArea(zip_wrf_intersect)/gArea(wrf_poly))
-    # populate the matrix based on i position and j position
-    zip_wrf_proportion[[i,j]] <- grid_prop
-  }
-}
-
-```
-
-```{r, eval=FALSE}
-f1 <- function(x){
-  zipcode <- as.character(x) 
-  # limit shapefile to particular zipcode
-  zip_shape <- or_zip_map[or_zip_map$ZCTA5CE10 %in% zipcode, ]
-  # convert to polygon
-  zip_poly <-SpatialPolygons(zip_shape@polygons)
-}
-zip_polygon <- lapply(or_zip_name, f1) # list
-# b <- zip_polygon[1] # list
-# c <- zip_polygon[[1]] # spatial polygon
-
-vars1 <- c(1:1610)
-f2 <- function(n){
-  wrf_grid <- smoke_grid[smoke_grid@data$WRFGRID_ID == n,] # sp df
-  # now what about grid 719; should be much less
-  wrf_poly <- SpatialPolygons(wrf_grid@polygons) # sp
-}
-wrf_polygon <- lapply(vars1, f2) # list of 1610 wrf_grid
-
-f3 <- function(x,y){
-  zip_wrf_intersect <- gIntersection(wrf_polygon[y][[1]], zip_polygon[x][[1]])
-  grid_prop <- ifelse(is.null(zip_wrf_intersect),
-                      0, gArea(zip_wrf_intersect)/gArea(wrf_polygon[y][[1]]))
-
-}
-proportion <- mcmapply(f3, rep(1:417, each=1610), rep(1:1610,417))
-
-```
-
-One core vs Multiple cores

### 2017-03-29 r_scripts/parallel_presentation.Rmd
try parallel

@@
 ## Imoport WRF Grid file of Oregon
-setwd("C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire")
-grid_dir <- paste0('./shapefile/oregon_grid.shp')
+grid_dir <- paste0('../../../data/data_new/or_shapefile/oregon_grid.shp')
 smoke_grid <- readOGR(dsn = grid_dir, layer = 'oregon_grid') # wrf_grid
@@
 ## Import U.S. shapefile
-shp_dir <- paste0('./shapefile/tl_2013_us_zcta510/tl_2013_us_zcta510.shp')
+shp_dir <- paste0('../../../data/data_original/', 
+                 'tl_2013_us_zcta510/tl_2013_us_zcta510.shp')
 us_zip_2013 <- readOGR(dsn = shp_dir, layer = 'tl_2013_us_zcta510')
@@
 ## Filter the U.S. shapefile only to Oregon State Zip Code
-read_path2 <- paste0('C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire/', 
-                     'oregon_zipcode.csv')
+read_path2 <- paste0('../../../data/data_original/oregon_zipcode.csv')
 or_zip <- read_csv(read_path2)
@@
 <br/> Apply a function to multiple list or vector arguments.
+
+## Parallel on server
+For loop vs Apply
+
+```{r prepare, include = FALSE}
+or_zip_name <- or_zip
+length(or_zip_name)
+wrf_grid_name <- as.character(smoke_grid@data$WRFGRID_ID)
+length(wrf_grid_name)
+# empty matrix
+zip_wrf_proportion <- matrix(nrow = 417, ncol = 1610, byrow = T,
+                             dimnames = list(or_zip_name, wrf_grid_name))
+
+# Setup for parallel computing before for loop ---------------------------------
+cores <- detectCores() # 48
+cl <- makeCluster(cores) # use half the cores on the vet cluster
+registerDoParallel(cl)
+# load packages on each cluster
+clusterCall(cl, function() library(rgdal))
+clusterCall(cl, function() library(sp))
+clusterCall(cl, function() library(rgeos))
+
+clusterExport(cl, "or_zip_map", envir = .GlobalEnv)
+clusterExport(cl, "or_zip_name", envir = .GlobalEnv)
+clusterExport(cl, "smoke_grid", envir = .GlobalEnv)
+clusterExport(cl, "wrf_grid_name", envir = .GlobalEnv)
+clusterExport(cl, "zip_polygon", envir = .GlobalEnv)
+clusterExport(cl, "wrf_polygon", envir = .GlobalEnv)
+
+```
+1. Forloop (one core, or not parallel computing)
+
+```{r forloop, echo = FALSE}
+start <- proc.time()
+for (i in 201:310) {
+  # output value of zipcode
+  zipcode <- as.character(or_zip_name[i]) 
+  # limit shapefile to particular zipcode
+  zip_shape <- or_zip_map[or_zip_map$ZCTA5CE10 %in% zipcode, ]
+  # convert to polygon
+  zip_poly <-SpatialPolygons(zip_shape@polygons)
+for(j in 201:300){
+    # output each grid and create a polygon
+    wrf_grid <- smoke_grid[smoke_grid@data$WRFGRID_ID == j, ]
+    # now what about grid 719; should be much less
+    wrf_poly <- SpatialPolygons(wrf_grid@polygons)
+
+    zip_wrf_intersect <- gIntersection(wrf_poly, zip_poly)
+    # if empty, then set to 0, else find the proportion
+    grid_prop <- ifelse(is.null(zip_wrf_intersect),
+                        0, gArea(zip_wrf_intersect)/gArea(wrf_poly))
+    # populate the matrix based on i position and j position
+    zip_wrf_proportion[[i,j]] <- grid_prop
+  }
+}
+stop <- proc.time() - start
+stop 
+```
+
+2. Mapply function (one core)
+
+```{r mapply, echo = FALSE}
+start <- proc.time()
+f1 <- function(x){
+  zipcode <- as.character(x) 
+  # limit shapefile to particular zipcode
+  zip_shape <- or_zip_map[or_zip_map$ZCTA5CE10 %in% zipcode, ]
+  # convert to polygon
+  zip_poly <-SpatialPolygons(zip_shape@polygons)
+}
+zip_polygon <- lapply(or_zip_name, f1) # list
+
+vars1 <- c(1:1610)
+f2 <- function(n){
+  wrf_grid <- smoke_grid[smoke_grid@data$WRFGRID_ID == n,] # sp df
+  # now what about grid 719; should be much less
+  wrf_poly <- SpatialPolygons(wrf_grid@polygons) # sp
+}
+wrf_polygon <- lapply(vars1, f2) # list of 1610 wrf_grid
+
+start <- proc.time()
+f3 <- function(x,y){
+  zip_wrf_intersect <- gIntersection(wrf_polygon[y][[1]], zip_polygon[x][[1]])
+  grid_prop <- ifelse(is.null(zip_wrf_intersect),
+                      0, gArea(zip_wrf_intersect)/gArea(wrf_polygon[y][[1]]))
+
+}
+proportion <- mapply(f3, rep(201:300, each=100), rep(201:300,100))
+stop <- proc.time() - start
+stop
+
+```
+
+3. Mcmapply function in parallel computing (multiple cores)
+
+```{r mcmapply, echo=FALSE}
+start <- proc.time()
+
+f1 <- function(x){
+  zipcode <- as.character(x) 
+  # limit shapefile to particular zipcode
+  zip_shape <- or_zip_map[or_zip_map$ZCTA5CE10 %in% zipcode, ]
+  # convert to polygon
+  zip_poly <-SpatialPolygons(zip_shape@polygons)
+}
+zip_polygon <- lapply(or_zip_name, f1) # list
+
+vars1 <- c(1:1610)
+f2 <- function(n){
+  wrf_grid <- smoke_grid[smoke_grid@data$WRFGRID_ID == n,] # sp df
+  # now what about grid 719; should be much less
+  wrf_poly <- SpatialPolygons(wrf_grid@polygons) # sp
+}
+wrf_polygon <- lapply(vars1, f2) # list of 1610 wrf_grid
+
+f3 <- function(x,y){
+  zip_wrf_intersect <- gIntersection(wrf_polygon[y][[1]], zip_polygon[x][[1]])
+  grid_prop <- ifelse(is.null(zip_wrf_intersect),
+                      0, gArea(zip_wrf_intersect)/gArea(wrf_polygon[y][[1]]))
+
+}
+proportion <- mcmapply(f3, rep(201:300, each=100), rep(201:300,100))
+stop <- proc.time() - start
+stop
+
+stopCluster(cl)
+
+```

### 2017-03-29 r_scripts/parallel_presentation.Rmd
revise

@@
 clusterExport(cl, "wrf_grid_name", envir = .GlobalEnv)
-clusterExport(cl, "zip_polygon", envir = .GlobalEnv)
-clusterExport(cl, "wrf_polygon", envir = .GlobalEnv)
+clusterExport(cl, "zip_wrf_proportion", envir = .GlobalEnv)
 

### 2017-03-29 r_scripts/parallel_presentation.Rmd
revise

@@
 ## Parallel on server
-For loop vs Apply
+**For loop vs Apply**
 
@@
 ```
-1. Forloop (one core, or not parallel computing)
+**1. Forloop** (one core, or not parallel computing)
 
@@
 start <- proc.time()
-for (i in 201:310) {
+for (i in 1:417) {
   # output value of zipcode
@@
   zip_poly <-SpatialPolygons(zip_shape@polygons)
-for(j in 201:300){
+for(j in 1:1610){
     # output each grid and create a polygon
@@
 
-2. Mapply function (one core)
+**2. Mapply function** (one core)
 
@@
 }
-proportion <- mapply(f3, rep(201:300, each=100), rep(201:300,100))
+proportion1 <- mapply(f3, rep(1:417, each=1610), rep(1:1610,417))
 stop <- proc.time() - start
@@
 
-3. Mcmapply function in parallel computing (multiple cores)
+**One core vs Multiple cores**
+
+**3. Mcmapply function** in parallel computing (multiple cores)
 
@@
 }
-proportion <- mcmapply(f3, rep(201:300, each=100), rep(201:300,100))
+proportion <- mcmapply(f3, rep(1:417, each=1610), rep(1:1610,417))
 stop <- proc.time() - start
@@
 ```
+
+Another important thing:
+Check if they are identical!
+```{r, echo=FALSE}
+## write matrix of for loop formed
+zip_proportion_df <- data.frame(zip_wrf_proportion)
+
+write_path <- paste0('../../../data/data_new/',
+                     'zip_wrf_proportion2.csv')
+write_csv(zip_proportion_df, paste0(write_path))
+
+## write matrix of apply function formed
+zip_wrf_proportion_new <- matrix(proportion, nrow = 417, ncol = 1610, byrow = T,
+                                    dimnames = list(or_zip_name, wrf_grid_name))
+
+zip_proportion_new_df <- data.frame(zip_wrf_proportion_new)
+write_path <- paste0('../../../data/data_new/',
+                     'zip_wrf_proportion_new2.csv')
+write_csv(zip_proportion_new_df, paste0(write_path))
+
+read_path_1 <- paste0('../../../data/data_new/',
+                     'zip_wrf_proportion2.csv')
+zip_wrf_proportion_for <- read_csv(read_path_1)
+read_path_2 <- paste0('../../../data/data_new/',
+                     'zip_wrf_proportion_new2.csv')
+zip_wrf_proportion_apply <- read_csv(read_path_2)
+
+identical(zip_wrf_proportion_for, zip_wrf_proportion_apply)
+
+```

### 2017-03-29 r_scripts/parallel_presentation_excute.R
convert to slidy format

@@ -8,7 +8,7 @@
 # call libraries
 library(rmarkdown)
 
-# call markdown doc.
-rmarkdown::render('./parallel_presentation.Rmd')
-
+# call markdown doc
+# specify an option for incremental rendering
+render("parallel_presentation.Rmd", slidy_presentation(incremental = TRUE))
 

### 2017-03-29 r_scripts/parallel_presentation.Rmd
add parallel

@@
 start <- proc.time()
-for (i in 1:417) {
+for (i in 201:300) {
   # output value of zipcode
@@
   zip_poly <-SpatialPolygons(zip_shape@polygons)
-for(j in 1:1610){
+for(j in 201:300){
     # output each grid and create a polygon
@@
 }
-proportion1 <- mapply(f3, rep(1:417, each=1610), rep(1:1610,417))
+proportion1 <- mapply(f3, rep(201:300, each=100), rep(201:300,100))
 stop <- proc.time() - start
@@
 }
-proportion <- mcmapply(f3, rep(1:417, each=1610), rep(1:1610,417))
+proportion <- mcmapply(f3, rep(201:300, each=100), rep(201:300,100))
 stop <- proc.time() - start
@@
 Check if they are identical!
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning = FALSE, include = FALSE}
 ## write matrix of for loop formed

### 2017-03-30 r_scripts/parallel_presentation_excute.R
slidy

@@ -8,7 +8,6 @@
 # call libraries
 library(rmarkdown)
 
-# call markdown doc
 # specify an option for incremental rendering
 render("parallel_presentation.Rmd", slidy_presentation(incremental = TRUE))
 

### 2017-03-30 r_scripts/parallel_presentation.Rmd
small

@@
 
+vars1 <- c(1:1610)
+z <- rep(201:300, each=100)
+w <- rep(201:300,100)
+
 # Setup for parallel computing before for loop ---------------------------------
@@
 clusterExport(cl, "zip_wrf_proportion", envir = .GlobalEnv)
+clusterExport(cl, "z", envir = .GlobalEnv)
+clusterExport(cl, "w", envir = .GlobalEnv)
+clusterExport(cl, "vars1", envir = .GlobalEnv)
 
@@
   # output value of zipcode
-  zipcode <- as.character(or_zip_name[i]) 
+  zipcode <- or_zip_name[i]
   # limit shapefile to particular zipcode
@@
     wrf_grid <- smoke_grid[smoke_grid@data$WRFGRID_ID == j, ]
-    # now what about grid 719; should be much less
     wrf_poly <- SpatialPolygons(wrf_grid@polygons)
@@
 
-vars1 <- c(1:1610)
 f2 <- function(n){
@@
 }
-proportion1 <- mapply(f3, rep(201:300, each=100), rep(201:300,100))
+proportion1 <- mapply(f3, z, w)
 stop <- proc.time() - start
@@
 }
-proportion <- mcmapply(f3, rep(201:300, each=100), rep(201:300,100))
+proportion <- mcmapply(f3, z, w)
 stop <- proc.time() - start

### 2017-03-30 r_scripts/parallel_presentation.Rmd
background

@@
 
-## Shapefile preparation: 2013 Oregon Zip code Shapefile and WRF Grid
+## Background
 
-Some libraries for spatial data analysis and parallel computing:
-```{r, eval = FALSE}
-library(rgdal)     # package for shape files
-library(sp)        # spatial polygon
-library(rgeos)     # rgeos package contains the intersect and area commands
-library(parallel)  # parallel computing libraries
-```
-
-The **shapefile of Oregon** can be subset from the whole U.S. shapefile by filtering the zip code in Oregon State.
-
-<br/>The **whole U.S shapefile** can be found from United States Census website:
-<br/>https://www.census.gov/cgi-bin/geo/shapefiles/index.php?year=2013&layergroup=ZIP+Code+Tabulation+Areas
-
-<br/>The **Oregon zip code** can also be found from: 
-<br/>https://www.unitedstateszipcodes.org/or/
-
-The plot below shows an Oregon State shapefile and an overlay of projected WRF Grid and Oregon shapefile.
+**Goal:** Estimate population-weighted PM2.5 for each ZIP code.
+<br/>
+**Problem:** The population data is for WRF Grid, not ZIP code.
+<br/>
+**Preparation:** 2013 Oregon Zip code Shapefile and WRF Grid
+<br/>
+**First Step:** Calculate the proportion of intersection of ZIP code and grid box
+*The plot below shows an Oregon State shapefile and an overlay of projected WRF Grid and Oregon shapefile.*
 ```{r overlay, echo=FALSE}
@@
 plot(or_zip_map, add=T)
+```
+
+**Example:** *Intersection for small area (Portland 97229); and big area (97910).*
+
+```{r, echo = FALSE, warning=FALSE}
+test_zip <- c(97229)
+test_zip_map <- or_zip_map[or_zip_map$ZCTA5CE10 %in% test_zip,]
+
+plot(test_zip_map)
+plot(smoke_grid, add = T)
+invisible(text(getSpPPolygonsLabptSlots(smoke_grid), 
+               labels=as.character(smoke_grid$WRFGRID_ID)))
+
+test_zip <- c(97910)
+test_zip_map <- or_zip_map[or_zip_map$ZCTA5CE10 %in% test_zip,]
+
+plot(test_zip_map)
+plot(smoke_grid, add = T)
+
 ```

### 2017-03-31 r_scripts/parallel_presentation.Rmd
add pic

@@
 <br/>
+**Preparation:** 2013 Oregon Zip code Shapefile and WRF Grid
+<br/>
 **Problem:** The population data is for WRF Grid, not ZIP code.
-<br/>
-**Preparation:** 2013 Oregon Zip code Shapefile and WRF Grid
 <br/>
@@
 ## Parallel on server
+![serial to solve problems](./serialProblem.jpg)
+
+![Parallel to solve problems](./parallelProblem.jpg)
+
+## Comparisons
 **For loop vs Apply**
 
-```{r prepare, include = FALSE}
+```{r prepare}
 or_zip_name <- or_zip
@@
 w <- rep(201:300,100)
+```
 
+```{r parallel set}
 # Setup for parallel computing before for loop ---------------------------------
@@
 ```
+
 **1. Forloop** (one core, or not parallel computing)
 
-```{r forloop, echo = FALSE}
+```{r forloop}
 start <- proc.time()
@@
 
-```{r mapply, echo = FALSE}
+```{r mapply}
 start <- proc.time()
@@
 
-Another important thing:
+## Another important thing:
 Check if they are identical!
-```{r, echo=FALSE, warning = FALSE, include = FALSE}
+```{r, echo=FALSE, warning = FALSE, include = FALSE, message = FALSE}
 ## write matrix of for loop formed

### 2017-03-31 r_scripts/parallel_presentation.Rmd
modify

@@
 
-```{r parallel set}
+```{r parallel set, echo = FALSE}
 # Setup for parallel computing before for loop ---------------------------------
@@
 Check if they are identical!
-```{r, echo=FALSE, warning = FALSE, include = FALSE, message = FALSE}
+```{r, warning = FALSE, message = FALSE, echo = FALSE}
 ## write matrix of for loop formed

### 2017-04-03 r_scripts/parallel_presentation.Rmd
Update parallel_presentation.Rmd
change some illustration

@@
   works on margins of an array or matrix, not like for loop.</li> 
-  <li> **Code Readability** <br/>  Understand the purporse more directly.</li> 
 </ul>
@@
 
+An example: 
+```{r, include = FALSE}
+A <- matrix(1:4, 2, 2)
+B <- matrix(5:10, 2, 3)
+```
+
+```{r}
+L <- list(A, B)
+lapply(L, "[", , 2)
+```
+
 >- The sapply() Function
@@
 Works like lapply(), but it tries to simplify the output to the most elementary data structure that is possible.
+
+```{r}
+L <- list(A, B)
+sapply(L, function(x) x[ , 2])
+```
 
@@
 
-```{r parallel set, echo = FALSE}
+```{r parallel set, eval = FALSE}
+# Setup for parallel computing before for loop ---------------------------------
+cores <- detectCores() # 48
+cl <- makeCluster(cores) # use half the cores on the vet cluster
+registerDoParallel(cl)
+# load packages on each cluster
+clusterCall(cl, function() library(rgdal))
+clusterCall(cl, function() library(sp))
+clusterCall(cl, function() library(rgeos))
+
+clusterExport(cl, "or_zip_map", envir = .GlobalEnv)
+clusterExport(cl, "or_zip_name", envir = .GlobalEnv)
+clusterExport(cl, "smoke_grid", envir = .GlobalEnv)
+clusterExport(cl, "wrf_grid_name", envir = .GlobalEnv)
+clusterExport(cl, "zip_wrf_proportion", envir = .GlobalEnv)
+clusterExport(cl, "z", envir = .GlobalEnv)
+clusterExport(cl, "w", envir = .GlobalEnv)
+clusterExport(cl, "vars1", envir = .GlobalEnv)
+
+```
+
+```{r parallel set, include = FALSE}
 # Setup for parallel computing before for loop ---------------------------------
@@
 
-```{r mcmapply, echo=FALSE}
+```{r mcmapply}
 start <- proc.time()

### 2017-04-03 r_scripts/oregon_proportion_server.R
change to new shapefile wrf grid

@@ -22,11 +22,11 @@ getwd()
 
 # Import Shapefiles  -----------------------------------------------------------
 # WRF Grid
-grid_dir <- paste0('../../../data/data_new/or_shapefile/oregon_grid.shp')
+grid_dir <- paste0('../../../data/data_original/oregon_new_grid.shp')
 
-smoke_grid <- readOGR(dsn = grid_dir, layer = 'oregon_grid') 
+smoke_grid <- readOGR(dsn = grid_dir, layer = 'oregon_new_grid') 
 
-summary(smoke_grid) 
+# summary(smoke_grid) 
 # summary(smoke_grid@data$WRFGRID_ID)
 
 # Zipcode shapefile
@@ -58,9 +58,9 @@ or_zip <- as.character(sort(or_zip_map@data$ZCTA5CE10))
 # save the Oregon zips to a shapefile to use later
 # create save path
 # !! will show error when overlay the writing file, so please write one time
-save_path <- paste0('../../../data/data_new/or_zip_2013_shape_files')
+save_path <- paste0('../../../data/data_new/or_zip_2013_shape_files_new')
 
-writeOGR(or_zip_map, layer = 'or_zip_2013_shape_files', save_path, driver = "ESRI Shapefile")
+writeOGR(or_zip_map, layer = 'or_zip_2013_shape_files_new', save_path, driver = "ESRI Shapefile")
 
 # Set coordinate reference system for smoke gird
 nad83 <- '+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0'
@@ -71,7 +71,6 @@ plot(smoke_grid)
 plot(or_zip_map, add=T)
 # looks like they overlay pretty well, same projections
 summary(smoke_grid)
-
 
 
 # Test code to figure out proportion calculations in each WRF-Grid -------------
@@ -109,16 +108,16 @@ plot(shape_grid, add = T)
 invisible(text(getSpPPolygonsLabptSlots(smoke_grid), 
                labels=as.character(smoke_grid$WRFGRID_ID)))
 
-# output the WRF Grid 261 and 262 for test
-wrf_grid_261 <- smoke_grid[smoke_grid@data$WRFGRID_ID == 261, ]
-plot(wrf_grid_261)
+# output the WRF Grid 297 and 298 for test
[…]

### 2017-04-03 r_scripts/zip_population_weighted_estimate_or_2013.R
Create zip_population_weighted_estimate_or_2013.R

@@ -1,0 +1,42 @@
+#-------------------------------------------------------------------------------
+# Title: Washington population-weighted smoke PM2.5 by day and zip
+# Author: Jingyang Liu                                                                   
+# Date Created: April 3, 2017                                                  
+# R version: 3.3.3                                                       
+#-------------------------------------------------------------------------------
+
+# Note: This is general code that could be submitted in batches to handle 
+#       multiple imputs (dataframes)
+
+# Note 8/8/16: I think the matrix approach actually runs pretty quick; doesn't
+#              need to be submitted in batches
+
+# load libraries ---------------------------------------------------------------
+library(tidyverse)
+
+
+# Setting Working Directory ----------------------------------------------------
+#dir <- paste0("C:/Users/RGan/Google Drive/CSU/wild_fire/washington/",
+#               "smoke_data/created_pm_estimates")
+
+# relative path 
+dir <- paste0("./data/Oregon/")
+setwd(dir)
+getwd()
+list.files()
+
+# Import and prepare smoke PM 2.5 data for loop --------------------------------
+
+# These files never change and can take place outside the loop
+# Import population weight csv 
+pop_path <- paste0('./popdensity_oregon.csv')
+pop_grid <- read_csv(pop_path)
+
+# rename missing column
+colnames(pop_grid) <- c('WRFGRID_ID', 'Longitude', 'Latitude', 'pop_density')
+summary(pop_grid)
+
+# calculate the estimated population for each grid (multiply density by grid area)
+population_grid <- pop_grid %>% 
+  mutate(pop_n = pop_density *(15^2)) ##??
+summary(population_grid)

### 2017-04-11 check_explore.R
Create check_explore.R

@@ -1,0 +1,312 @@
+#-------------------------------------------------------------------------------
+# Title: Check and exploring the unreasonable plot
+# Author: Jingyang Liu                                                                   
+# Date Created: April 11, 2017                                                  
+# R version: 3.3.3                                                       
+#-------------------------------------------------------------------------------
+
+library(tidyverse)
+library(data.table)
+library(survival)
+library(htmlTable)
+
+### Limit to only ER visits
+## pos = 23
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+
+N <- NA
+n <- 1
+
+for (i in var_list){
+  read_path <- paste('./data_new/update/ER/or', i, 'may_to_sep_time_strat_casecross.csv', sep = "_")
+  or_disease <- read_csv(read_path)
+  
+  or_disease <- or_disease %>% 
+    filter(pos=="23")
+  
+  # Create a permanent case-cross over dataset
+  file_name <- paste('./data_new/update/ER/or', i, 'may_to_sep_time_strat_casecross_er.csv', sep = "_")
+  
+  # write permanent dataset
+  write_csv(or_disease, paste0(file_name))
+  N[n] = length(unique(or_disease$personkey))
+  n = n+1
+}
+
+
+num_person <- data.frame(rbind(var_list, N))
+
+
+## Analysis
+path <- paste0("./data_new/update/ER")
+
+resp_casecross <- read_csv(paste(path, 
+                                 "or_respiratory_may_to_sep_time_strat_casecross_er.csv", 
+                                 sep = "/"))
+
+# asthma
+asthma_casecross <- read_csv(paste(path, 
+                                   "or_asthma_may_to_sep_time_strat_casecross_er.csv", 
+                                   sep = "/")) 
+
+# copd 
+copd_casecross <- read_csv(paste(path, 
+                                 "or_copd_may_to_sep_time_strat_casecross_er.csv", 
+                                 sep = "/"))
+
+# pneum or bronchitis
+pneum_casecross <- read_csv(paste(path, 
+                                  "or_pneumonia_may_to_sep_time_strat_casecross_er.csv",
[…]

### 2017-04-12 r_scripts/case_crossever_oregon.R
Update case_crossever_oregon.R

@@ -1,5 +1,5 @@
 # ------------------------------------------------------------------------------
-# Title: Case crossover for diseases (Oregon State)
+# Title: Case crossover for diseases of Oregon State (Practice)
 # Author: Jingyang Liu
 # Date Created: Feb 24, 2017
 # R version: 3.3.2

### 2017-04-12 r_scripts/data_clean/oregon_data_clean_server.R
Create oregon_data_clean_server.R

@@ -1,0 +1,473 @@
+---
+title: "Oregon_Wildfire_explore_full_server"
+author: "Jingyang Liu"
+date: "February 15, 2017"
+---
+  
+# Explore Data Structure and Cleaning
+  
+The Oregon data contains 77069320 obsversations in total in year of 2013, which is a big data set. We want to figure out the structure for this data set and do some basic cleaning. 
+
+```{r}
+library(dplyr)
+library(data.table)
+library(knitr)
+library(readxl)
+library(ggplot2)
+getwd()
+
+# import data ------------------------------------------------------------------
+
+read_path <- paste0("../../../data/data_original/gan_episodes_of_care.txt")
+start_time <- Sys.time()
+oregon_df <- fread(read_path, sep = "|", showProgress = T, 
+                   colClasses = rep("character", 72)) 
+stop_time <-  Sys.time() - start_time 
+# time it took
+stop_time # about 20 mins
+
+```
+
+
+## The overall summary  of data set
+72 variables in total.
+
+```{r overall summary}
+# Check data structure
+# summary(oregon_df)
+```
+
+## convert variable type to appropriate functional form 
+(i.e. zipcode leave as character, age as numeric, date as date, dx leave as character, but *NULL* == NA)
+```{r}
+oregon_df$yob <- as.numeric(oregon_df$yob)
+oregon_df[oregon_df=="*NULL*"] <- NA
+```
+
+# summary statistics key variables for study 
+date, dx, need to know urgent and emergency department visits, sex, age, patientid, state of residence, zipcode, county, race
+
+## 1. Personkey 
+```{r}
+personkey2 <- oregon_df %>% 
+  select(personkey) %>%
+  mutate(personkey_cat = ifelse(is.na(personkey), "Missing",
+                         ifelse(!is.na(personkey), "Non-Missing", NA))) %>% 
+  group_by(personkey_cat) %>%
+  summarise(n = n()) %>%
+  mutate(total_n =sum(n),
+         percent = round(n/total_n, 4)*100) %>%
+  select(personkey_cat, n, percent)
+
+# table ----
+knitr::kable(personkey2, align = "c", caption = "Patient ID")
+
+```
+
+
+## How many visits does unique personkey have?
+```{r}
+# pick up personkey and claim id column, make new column of unique number of visit
+visit2 <- oregon_df %>%
+  select(personkey, clmid) %>%
[…]

### 2017-04-12 r_scripts/data_management/try_case_crossever_oregon.R
Create try_case_crossever_oregon.R

@@ -1,0 +1,203 @@
+# ------------------------------------------------------------------------------
+# Title: Case crossover for diseases of Oregon State (Practice)
+# Author: Jingyang Liu
+# Date Created: Feb 24, 2017
+# R version: 3.3.2
+# ------------------------------------------------------------------------------
+
+#Creating case crossover dataframes
+library(tidyverse)
+library(data.table)
+library(dplyr)
+library(readr)
+library(lubridate)
+getwd()
+read_path <- paste0('./data1/or_hosp_w_outcome_df.csv')
+disease <- read_csv(read_path) 
+
+disease <- disease %>%
+  # add new transverted from date
+  mutate(dates = as.Date(fromdate, "%m/%d/%Y"))
+
+ggplot(disease, aes(x = disease$dates)) + geom_density()
+
+
+# Asthma1 ----------------------------------------------------------------------
+disease_asthma1 <- disease %>%
+  # filter asthma1
+  filter(asthma1==1) %>%
+  # filter ZIP=80526
+  filter(ZIP==80526) %>%
+  # filter date in July 1, 2012 and Oct 31, 2012
+  filter(dates >= '2012-07-01' & 
+           dates <= '2012-10-31') %>%
+  # add ID
+  mutate(id = seq(1, nrow(.), by = 1),
+         # add outcome and new admit date
+         outcome_asthma1 = 1)
+
+one_asthma1_case <- sample_n(disease_asthma1, 1) %>% 
+  select(cdpheid, admit, dates, dx1, asthma1, ZIP, county_final, WRFGRID_ID,
+         RACE, sex_ind, age_ind, id)
+
+one_asthma1_case
+
+# find the replicate times of weeks
+dates <- one_asthma1_case[[1,3]] 
+n1 <- 0
+d=as.Date("2012-07-01")
+i=1
+while (dates >= "2012-07-01"){
+  dates <- dates - 7
+  d[i] = dates
+  i = i+1
+  n1 = n1+1
+}
+d[1:n1-1] # shows character(0) when the first week
+n1-1
+
+dates <- one_asthma1_case[[1,3]] 
+n2=0
+e=as.Date("2012-10-31")
+j=1
+while (dates <= "2012-10-31"){
+  dates <- dates + 7
+  e[j]=dates
+  j=j+1
+  n2 = n2 + 1
+}
+e[1:n2-1] # shows character(0) when the last week
+n2-1
+
+# replicate covariates length of counterfactual dates
[…]

### 2017-04-12 r_scripts/data_management/try_oregon_data_cleaning.R
Create try_oregon_data_cleaning.R

@@ -1,0 +1,177 @@
+# ------------------------------------------------------------------------------
+# Title: Oregon 2013 claims data for MobaXterm (Practice)
+# Author: Jingyang Liu
+# Date Created: Feb 9th, 2017
+# R Version: 3.3.2
+# ------------------------------------------------------------------------------
+
+getwd() # "/home/jyliu/wildfire/r_scripts"
+
+library(dplyr)
+library(data.table)
+
+read_path <- paste0("../data/gan_episodes_of_care.txt")
+oregon_df <- fread(read_path, sep = "|", nrows = 2000000, showProgress = T)
+fwrite(oregon_df, "../data/oregon_epis_care_reduced.csv")
+
+# I wrote a reduced dataframe of 2000,000 rows to get an idea of data structure
+
+# import the reduced dataframe
+start_time <- Sys.time()
+
+oregon_df <- fread("../data/oregon_epis_care_reduced.csv", sep = ",", 
+                   colClasses = rep("character", 72))
+
+stop_time <-  Sys.time() - start_time 
+# time it took
+stop_time # 15.83603 secs
+
+
+start_time <- Sys.time()
+oregon_full <- fread(read_path, sep = "|", showProgress = T) # not change class
+stop_time <-  Sys.time() - start_time 
+# time it took
+stop_time # 7.771713 mins
+
+summary(as.factor(oregon_full$year)) # all 2013, 77069320
+summary(as.factor(oregon_full$personkey))
+min(as.numeric(oregon_full$personkey))
+max(as.numeric(oregon_full$personkey)) # seems length is 8
+unique(as.numeric(oregon_full$line)) # 1362 levels, 0 - 9999
+summary(as.factor(oregon_full$clmstatus))
+#       D        E        P
+# 8676643  3324496 65068181
+summary(as.factor(oregon_full$cob)) 
+#                 N        Y
+# 23517978 44438688  9112654
+summary(as.factor(oregon_full$paytype)) #C, D, G, P, T
+summary(as.factor(oregon_full$prod))
+summary(as.factor(oregon_full$payer))
+summary(as.factor(oregon_full$medflag))
+#       0        1     NA's
+# 5697342 70573441   798537
+summary(as.factor(oregon_full$rxflag))
+#       0        1     NA's
+# 7761348 68509435   798537
+summary(as.factor(oregon_full$ohvmhflag))
[…]

### 2017-04-12 r_scripts/data_management/try_oregon_data_import_wrangling.R
Create try_oregon_data_import_wrangling.R

@@ -1,0 +1,472 @@
+# ------------------------------------------------------------------------------
+# Title: Small data set case crossover studies (Practice)
+# Author: Jingyang Liu
+# Date Created: Feb 24, 2017
+# R Version: 3.3.2
+# ------------------------------------------------------------------------------
+
+library(tidyverse)
+library(data.table)
+library(dplyr)
+library(readxl) 
+getwd()
+
+### import data ----------------------------------------------------------------
+# Read in first 200,000 lines for exploration.
+
+read_path <- paste0("./data/gan_episodes_of_care.txt") # R Markdown is ".." because working directory is r_script
+oregon_df <- fread(read_path, sep = "|", nrows = 200000, showProgress = T)
+fwrite(oregon_df, "./data/oregon_epis_care_reduced.csv")
+
+# I wrote a reduced dataframe of 200,000 rows to get an idea of data structure
+
+# import the reduced dataframe
+start_time <- Sys.time()
+oregon_df <- fread("./data/oregon_epis_care_reduced.csv", sep = ",", 
+                   colClasses = rep("character", 72))
+stop_time <-  Sys.time() - start_time 
+
+# time it took
+stop_time # 1.968219 secs
+
+# convert variable type to appropriate functional form 
+oregon_df$yob <- as.numeric(oregon_df$yob)
+oregon_df[oregon_df=="*NULL*"] <- NA
+
+### data rough cleaning (remove some NA and blank values) and add index---------
+# at least a primary ICD9 code present
+# whole data still have 76552034 observations
+or_hosp_w_outcome_df <- oregon_df %>% filter(!is.na(dx1)) %>% 
+  # filter sex is known
+  filter(gender!="U") %>%
+  # add age, filter age not missing
+  mutate(age = 2013-yob) %>%
+  filter(!is.na(age)) %>%
+  # filter race
+  filter(race!="") %>%
+  # filter enthnicity
+  filter(ethn!="") %>%
+  # filter State is Oregon
+  filter(STATE=="OR") %>%
+  # filter ZIP non missing
[…]

### 2017-04-12 r_scripts/zip_population_weighted_estimate_or_2013.R
Delete zip_population_weighted_estimate_or_2013.R

@@ -1,322 +1,0 @@
-#-------------------------------------------------------------------------------
-# Title: Washington population-weighted smoke PM2.5 by day and zip
-# Author: Jingyang Liu                                                                   
-# Date Created: April 3, 2017                                                  
-# R version: 3.3.3                                                       
-#-------------------------------------------------------------------------------
-
-# Note: This is general code that could be submitted in batches to handle 
-#       multiple imputs (dataframes)
-
-# Note 8/8/16: I think the matrix approach actually runs pretty quick; doesn't
-#              need to be submitted in batches
-
-# load libraries ---------------------------------------------------------------
-library(tidyverse)
-library(dplyr)
-
-# Setting Working Directory ----------------------------------------------------
-#dir <- paste0("C:/Users/RGan/Google Drive/CSU/wild_fire/washington/",
-#               "smoke_data/created_pm_estimates")
-
-# relative path 
-dir <- paste0("./data/Oregon/")
-setwd(dir)
-getwd()
-list.files()
-
-# Import and prepare smoke PM 2.5 data for loop --------------------------------
-
-# These files never change and can take place outside the loop
-# Import population weight csv 
-pop_path <- paste0('./popdensity_oregon.csv')
-pop_grid <- read_csv(pop_path)
-
-# rename missing column
-colnames(pop_grid) <- c('WRFGRID_ID', 'Longitude', 'Latitude', 'pop_density')
-summary(pop_grid)
-
-# calculate the estimated population for each grid (multiply density by grid area)
-population_grid <- pop_grid %>% 
-  mutate(pop_n = pop_density *(15^2)) ##??
-summary(population_grid)
-
-
-# Import the file that contains the proportion of each grid that overlaps a zip
-zip_grid_path <- paste0('zip_wrf_proportion.csv')
-zip_grid_proportion <- read_csv(zip_grid_path)
-
-# Importing files that I would like to loop through
-# Geo weighted regresssion (Atmos model of choice)
-geo_wt_path <- paste0("oregon_gwrPM_2013.csv")
-geo_wt <- read_csv(geo_wt_path)
-
-# Kriging 
[…]

### 2017-04-12 r_scripts/smoke/oregon_proportion_zip.R
Create oregon_proportion_zip.R

@@ -1,0 +1,264 @@
+# ------------------------------------------------------------------------------
+# Title: Oregon dataframe of proportion of each zipcode in each WRF-Grid cell 
+# Author: Jingyang Liu
+# Date Created: Mar 10, 2017
+# R version: 3.3.2
+# ------------------------------------------------------------------------------
+
+library(rgdal) # package for shape files
+library(sp)
+library(rgeos) # rgeos package contains the intersect and area commands I need
+library(dplyr)
+library(data.table)
+library(readxl)
+# parallel computing libraries
+library(foreach) 
+library(doParallel)
+
+getwd()
+
+# Import Shapefiles  -----------------------------------------------------------
+# WRF Grid
+grid_dir <- paste0('./shapefile/oregon_grid.shp')
+
+smoke_grid <- readOGR(dsn = grid_dir, layer = 'oregon_grid') 
+
+summary(smoke_grid) 
+plot(smoke_grid) 
+invisible(text(getSpPPolygonsLabptSlots(smoke_grid), 
+               labels=as.character(smoke_grid$WRFGRID_ID)))
+
+summary(smoke_grid@data$WRFGRID_ID)
+
+# Zipcode shapefile
+shp_dir <- paste0('./shapefile/tl_2013_us_zcta510/tl_2013_us_zcta510.shp')
+
+us_zip_2013 <- readOGR(dsn = shp_dir, layer = 'tl_2013_us_zcta510')
+summary(us_zip_2013)
+# plot(us_zip_2013)
+
+## import data
+# because it use overall ZIP, so I filter the full data set with same requirements 
+# of separate disease.
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+
+setwd("C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire/data_new")
+
+or_zip_2013 <- data_frame()
+n <- 0
+
+start <- Sys.time()
+for(m in var_list){ # begin first loop of variable names (outcomes)
+  
+  read_path <- paste('oregon', m, 'jul_to_oct_claim.csv', sep='_')
+  or_disease <- read_csv( read_path)
+  
+  or_disease$dx11 <- as.character(or_disease$dx11)
[…]

### 2017-04-12 r_scripts/smoke/oregon_proportion_zip_server.R
Create oregon_proportion_zip_server.R

@@ -1,0 +1,304 @@
+# ------------------------------------------------------------------------------
+# Title: Oregon dataframe of proportion of each zipcode in each WRF-Grid cell 
+# Author: Jingyang Liu
+# Date Created: Mar 10, 2017
+# R version: 3.3.2
+# ------------------------------------------------------------------------------
+
+library(rgdal) # package for shape files
+library(sp)
+library(rgeos) # rgeos package contains the intersect and area commands I need
+library(dplyr)
+library(data.table)
+library(readxl)
+# parallel computing libraries
+# library(foreach) 
+library(doParallel)
+library(parallel)
+library(readr)
+
+getwd()
+# "/home/jyliu/wildfire/local_git_repo/oregon_wildfire/r_scripts"
+
+# Import Shapefiles  -----------------------------------------------------------
+# WRF Grid
+grid_dir <- paste0('../../../data/data_original/oregon_new_grid.shp')
+
+smoke_grid <- readOGR(dsn = grid_dir, layer = 'oregon_new_grid') 
+
+# summary(smoke_grid) 
+# summary(smoke_grid@data$WRFGRID_ID)
+
+# Zipcode shapefile
+shp_dir <- paste0('../../../data/data_original/', 
+                 'tl_2013_us_zcta510/tl_2013_us_zcta510.shp')
+
+us_zip_2013 <- readOGR(dsn = shp_dir, layer = 'tl_2013_us_zcta510')
+summary(us_zip_2013)
+# plot(us_zip_2013)
+
+
+### whole oregon zip code
+read_path <- paste0('../../../data/data_original/oregon_zipcode.csv')
+or_zip <- read_csv(read_path)
+names(or_zip) <- c('zip','type','city','county','area')
+
+# limit to just Oregon state zipcodes
+or_zip_map <- us_zip_2013[us_zip_2013$ZCTA5CE10 %in% or_zip$zip,]
+
+# output zipcodes from washington zipcode map to bind values to
+# plot map to check
+plot(or_zip_map)
+
+# saving wash shapefile ----
+# save shapefile to use in future work
+summary(or_zip_map)
+or_zip <- as.character(sort(or_zip_map@data$ZCTA5CE10))
+
+or_zip_df <- data.frame(or_zip)
+
+write_path <- paste0('../../../data/data_new/',
+                     'or_zip.csv')
[…]

### 2017-04-12 r_scripts/data_management/oregon_data_wrangle_separate_server.R
Create oregon_data_wrangle_separate_server.R

@@ -1,0 +1,106 @@
+# ------------------------------------------------------------------------------
+# Title: Basic data wrangling  for 2013 oregon data
+# Author: Jingyang Liu
+# Date Created: Mar 03, 2016
+# R Version: 3.3.2
+# ------------------------------------------------------------------------------
+
+library(dplyr)
+library(data.table)
+library(readxl) 
+library(readr)
+
+# import data ------------------------------------------------------------------
+
+read_path <- paste0("../../../data/data_original/gan_episodes_of_care.txt")
+start_time <- Sys.time()
+oregon_df <- fread(read_path, sep = "|", showProgress = T) 
+stop_time <-  Sys.time() - start_time 
+# time it took
+stop_time # 7 mins
+
+# basic cleaning ---------------------------------------------------------------
+start_time <- Sys.time()
+oregon_df <- oregon_df %>% filter(!is.na(dx1)) %>% 
+  # filter State is Oregon
+  filter(STATE=="OR") %>%
+  # filter urgent or ED visits (pos = 20 or 23)
+  filter(pos=="20"|pos=="23") 
+stop_time <-  Sys.time() - start_time 
+# time it took
+stop_time # 2.5 mins
+
+write_path <- paste0('../../../data/data_original/',
+                     'oregon_epis_care_reduced.csv')
+write_csv(oregon_df, write_path)
+
+
+
+# Define read paths ------------ -----------------------------------------------
+# import r list of outcome-specific icd9 codes
+# 1514642 obs
+load_path <- paste0("../../../data/data_original/outcome_list.RData")
+load(load_path)
+
+# path and file name of data
+read_path <- paste0("../../../data/data_original/oregon_epis_care_reduced.csv")
+start_time <- Sys.time()
+oregon_df <- fread(read_path, sep = ",", 
+                   colClasses = rep("character", 72)) 
+stop_time <-  Sys.time() - start_time 
+# time it took
+stop_time # 14 sec
+
+
+
+# now try to work the two chunks together to identify records that contain icd9s
+# of a certain outcome and then output those rows in chunks of data.
+
+# I need a dataframe of start row and length to read after
+n_start<- seq(from = 1, to = length(oregon_df$personkey), by = 10000) 
+
+# create dataframe
+n_read <- data_frame(n_start) %>% # turn the sequence in to a dataframe
+  mutate(n_length = ifelse(row_number()!=n(), 10000,
[…]

### 2017-04-12 r_scripts/oregon_markdown_excute_server.R
Update oregon_markdown_excute_server.R

@@ -9,6 +9,6 @@
 library(rmarkdown)
 
 # call markdown doc.
-rmarkdown::render('./oregon_clean.rmd')
+rmarkdown::render('./oregon_data_explore_server.rmd')
 
 

### 2017-04-12 r_scripts/data_management/oregon_markdown_excute_server.R
Create oregon_markdown_excute_server.R

@@ -1,0 +1,14 @@
+# ------------------------------------------------------------------------------
+# Title: Script to excute descriptive summary R markdown file
+# Author: Jingyang Liu
+# Date: 3/2/2016
+# ------------------------------------------------------------------------------
+
+
+# call libraries
+library(rmarkdown)
+
+# call markdown doc.
+rmarkdown::render('./oregon_data_explore_server.rmd')
+
+

### 2017-04-12 r_scripts/data_management/oregon_casecrossover.R
Create oregon_casecrossover.R

@@ -1,0 +1,214 @@
+# ------------------------------------------------------------------------------
+# Title: Case crossover for diseases (Oregon State)
+# Author: Jingyang Liu
+# Date Created: Feb 24, 2017
+# R version: 3.3.2
+# ------------------------------------------------------------------------------
+
+#Creating case crossover dataframes
+library(tidyverse)
+library(data.table)
+library(readxl)
+getwd()
+setwd("C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire/data_disease")
+
+### ---------------------------------------------------------------------
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+
+start <- Sys.time()
+for (i in var_list){
+  read_path <- paste('oregon', i, 'claims_2013.csv', sep='_')
+  or_disease <- read_csv(read_path)
+    
+  or_disease <- or_disease %>% 
+    group_by(personkey) %>% 
+    arrange(personkey, fromdate, line) %>% 
+    mutate(num_visit = dense_rank(fromdate)) %>%
+    arrange(personkey, num_visit) %>% 
+    select(personkey, clmid, num_visit) %>% 
+    unique() %>% 
+    full_join(or_disease, by = c("personkey", "clmid")) %>% 
+    arrange(personkey, clmid, line, fromdate) %>% 
+    filter(row_number() == 1) %>%
+    # add new transverted from date
+    mutate(dates = as.Date(fromdate, "%m/%d/%Y")) %>%
+    filter(dates >= '2013-05-01' & 
+           dates <= '2013-09-30') 
+  
+  # Create a permanent case-cross over dataset
+  file_name <- paste('oregon', i, 'may_to_sep_claim.csv', sep = '_')
+  
+  # write permanent dataset
+  write_csv(or_disease, paste0("../data_new/", file_name))
+  
+}
+total_time <- Sys.time() - start
+total_time # Time difference of 5.154874 mins
+
+setwd("../data_new/update")
+
+
+
+## Case crossover for each disease
+
+start <- Sys.time()
[…]

### 2017-04-12 r_scripts/check_analysis.R
Delete check_analysis.R

@@ -1,312 +1,0 @@
-#-------------------------------------------------------------------------------
-# Title: Check and exploring the unreasonable plot
-# Author: Jingyang Liu                                                                   
-# Date Created: April 11, 2017                                                  
-# R version: 3.3.3                                                       
-#-------------------------------------------------------------------------------
-
-library(tidyverse)
-library(data.table)
-library(survival)
-library(htmlTable)
-
-### Limit to only ER visits
-## pos = 23
-var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
-              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
-              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
-
-N <- NA
-n <- 1
-
-for (i in var_list){
-  read_path <- paste('./data_new/update/ER/or', i, 'may_to_sep_time_strat_casecross.csv', sep = "_")
-  or_disease <- read_csv(read_path)
-  
-  or_disease <- or_disease %>% 
-    filter(pos=="23")
-  
-  # Create a permanent case-cross over dataset
-  file_name <- paste('./data_new/update/ER/or', i, 'may_to_sep_time_strat_casecross_er.csv', sep = "_")
-  
-  # write permanent dataset
-  write_csv(or_disease, paste0(file_name))
-  N[n] = length(unique(or_disease$personkey))
-  n = n+1
-}
-
-
-num_person <- data.frame(rbind(var_list, N))
-
-
-## Analysis
-path <- paste0("./data_new/update/ER")
-
-resp_casecross <- read_csv(paste(path, 
-                                 "or_respiratory_may_to_sep_time_strat_casecross_er.csv", 
-                                 sep = "/"))
-
-# asthma
-asthma_casecross <- read_csv(paste(path, 
-                                   "or_asthma_may_to_sep_time_strat_casecross_er.csv", 
-                                   sep = "/")) 
-
-# copd 
-copd_casecross <- read_csv(paste(path, 
-                                 "or_copd_may_to_sep_time_strat_casecross_er.csv", 
-                                 sep = "/"))
-
-# pneum or bronchitis
-pneum_casecross <- read_csv(paste(path, 
-                                  "or_pneumonia_may_to_sep_time_strat_casecross_er.csv",
[…]

### 2017-04-12 r_scripts/analysis/or_zip_result.Rmd
Create or_zip_result.Rmd

---
title: "Oregon: Zip-Level Wildfire Smoke and Health Outcomes for Manuscript"
author: "Jingyang Liu"
date: "April 7, 2017"
output: html_document
---

```{r library calls, include=FALSE}
# loading libraries used
library(tidyverse) # import tidyverse
library(survival) # for conditional logistic regression
library(htmlTable) # table
library(lme4) # random-effects model
library(broom) # broom for tidy data from stats models
library(cowplot) # for multi plot image

# spatial packages
library(ggmap) # map package
library(rgdal) # set coord ref system for shape files
library(rgeos) # need it when converting shape to dataframe
library(maptools) # convert shape to dataframe

```

## Overview

This document contains results for the association between wildfire smoke PM~2.5~ and cardiovascular (CVD) and respiratory health outcomes. This particular markdown document focuses on the zip code-level analysis. Note there is a county-level analysis that will be paired with CDC mortality data as well. Herein, I compare different PM~2.5~ estimation methods described in further detail below.

Note, epidemiological methods that Sheryl and I think can improve our estimation and reduce some bias around the point estimates will not be the focus of the manuscript.

```{r data import, include = F, echo = FALSE}
# Set working directory and read in files --------------------------------------
# relative path
# set working directory if you want to run seperate chunks
# setwd("./r_scripts")
getwd()

path <- paste0("../data_new/health_data")
# Infile case-crossover dataframes ---------------------------------------------
# Dataframes made in 'chars_3month_binary_smoke_may2016 script
# resp exacerbations
resp_casecross <- read_csv(paste(path, 
                        "respiratory_may_to_sep_time_strat_casecross.csv", 
                        sep = "/"))

# asthma
asthma_casecross <- read_csv(paste(path, 
                          "asthma_may_to_sep_time_strat_casecross.csv", 
                          sep = "/"))

# copd 
copd_casecross <- read_csv(paste(path, 
                        "copd_may_to_sep_time_strat_casecross.csv", 
                        sep = "/"))

# pneum or bronchitis
pneum_casecross <- read_csv(paste(path, 
                        "copd_may_to_sep_time_strat_casecross.csv",
                        sep="/"))
# acute bronchitis
acute_bronch_casecross <- read_csv(paste(path, 
                                "acute_bronch_may_to_sep_time_strat_casecross.csv",
                                sep = "/"))
# cvd
cvd_casecross <- read_csv(paste(path, 
                      "cvd_may_to_sep_time_strat_casecross.csv",
                      sep="/"))
# arrhythmia
arrhythmia_casecross <- read_csv(paste(path, 
                              "arrhythmia_may_to_sep_time_strat_casecross.csv",
                               sep="/"))
# cerebral vascular
cereb_vas_casecross <- read_csv(paste(path, 
                              "cerbrovas_dis_may_to_sep_time_strat_casecross.csv",
                              sep="/"))
# heart failure
hf_casecross <- read_csv(paste(path, 
                      "heart_failure_may_to_sep_time_strat_casecross.csv", 
                      sep="/"))
# ischemic heart disease
ihd_casecross <- read_csv(paste(path, 
                       "isch_heart_dis_may_to_sep_time_strat_casecross.csv",
                       sep="/"))
# myo infarc
mi_casecross <- read_csv(paste(path, 
                      "myocardial_infarc_may_to_sep_time_strat_casecross.csv", sep="/"))

# broken arm
broken_arm_casecross <- read_csv(paste(path, 
                              "broken_arm_may_to_sep_time_strat_casecross.csv",
                              sep="/"))

# read zip pm for counts by zip
smoke_path <- paste0("../data/Oregon/")

zip_pm <- read_csv(paste0(smoke_path, "zip_pm_to_merge_with_chars.csv"))

```

```{r, eval = FALSE}
## 
# read locations of fires
fire_loc <- read_csv(paste0(smoke_path, "washington_fires_locations201209.csv"))

# import the zip-level met data files Rish sent
zip_met <- read_csv(paste0(smoke_path, "wa_zip_exp_0710_2012_ryan.csv")) %>% 
  # define date variable; original character looked like: 01JUL2012
  mutate(date = as.Date(date, "%d%B%Y"))

# note 12/4/16: read_csv is pretty slow on pc, fread doesn't import date variable 
# that well. Solution may be to use fread but convert to date to use later

```

## Methods Description

In these comparisons, we examine various methods of smoke/PM~2.5~ estimations and associations with health outcomes using a time-stratified case-crossover study design. Health outcomes for a patient with a primary diagnosis of cardiopulmonary health outcomes and their date of admission  (index time) were identified. We then created counterfactual observations for each patient for the same day of the week for the entire wildfire season (MAy 1st to Sep 30th, 2013). We further limited our analyses to claims that were coded as emergency or urgent visits to eliminate bias from patients going in for elective/planned procedures. For patients who have a index time closer to May 1, 2013 or closer to September 30, 2013, their referent observations before or after these dates will be excluded as I will not be able to assign estimates of PM~2.5~ to these referent observations. However, this is not a big issue for a time-stratified design as other referent values throughout the time period average out.

The **health outcomes** of interest are all respiratory, asthma, COPD, pneumonia, acute bronchitis, cardiovascular disease, heart failure, ischemic heart disease, and myocardial infarction. I also included broken arm, which I hypothesize to not be associated with wildfire smoke exposure and use as a check.

As for **exposure methods**, there are four main estimation methods for PM~2.5~: WRF-Chem Smoke (which substracts the WRF-Chem no fire emission from WRF-chem). For the WRF-Chem variable with the 'smoke' designator, this is WRF-Chem - WRF-Chem no fire. For Global Regression, Geo-Weighted Regression, and Kriging with 'smk' designator, I subtracted off the 'Background' estimates of smoke, which I believe are the monthly averages of PM~2.5~ for a given grid.

The **analytic method** is the conditional logistic regression model using the *survival* package in R. Each conditional logistic regression model accounts for the subject, and adjusts for temperature from the WRF-Chem model. The conditional logistic regression model \beta represents the change in risk of an event associated with a short-term unit increase in exposure, and can be calculated as an average difference between exposure at the index time and a weighted average of exposure at all times in the referent window. See Janes et al. 2005, **Statistics in Medicine**.

I've run a couple different scenarios for selecting referent periods, each of which has some paper that may justify the benefit of the design, relative to other methods.

### Time-Series of Smoke PM~2.5~ Ranges by Ecoregion

Figures below are for each smoke PM~2.5~ estimation method. The figures are the daily range (min and max) for each zip code in a specific region of the state that local offices of the Washington Department of Ecology has jurisdiction over. The figure first starts out by showing the map of the state, divided in to eco regions used by the state for organization purposes, and then presents the time-series ranges for WRF-Chem, krig, geo-weighted estimation methods.

```{r smk time series range by eco region, echo = F, results='asis', message=F}

## ???
# creation of map and identifying zipcodes in ecoregions----
# import the regional shapefile downloaded from state website
shp_dir <- paste0('../../data/shape_files/WAECY__Ecology_Regions')
shp_dir <- paste0('./data/shape_files/WAECY__Ecology_Regions')

eco_region <- readOGR(dsn = shp_dir, layer = 'WAECY__Ecology_Regions')
# set coord ref system to WGS84 to match projections of ggmap
eco_region <- spTransform(eco_region, "+proj=longlat +datum=WGS84")

# summary(eco_region) # projection is GRS80, may be off

# ggmap and rgdal needed
# Set base map boundary layer
bbox <- c(-125, 45, -116.1, 49.5)

# call black and white map
washington_map <- get_map(location= bbox, source = 'stamen',
                          maptype = 'toner', crop = T)

# call google terrain map
#wash_map_terrain <- get_map(location = "washington state", source = 'google',
#                             maptype = 'terrain', crop = F)

# fortify washington shape data in to dataframe (need maptools package)
eco_df <- tidy(eco_region, region = "ECY_REGION")

# summary(wash_zip_df)

# creation of eco region map (may change colors)
eco_region_map <- ggmap(washington_map) + 
  geom_polygon(data = eco_region, aes(x=long, y=lat, group = group, 
    fill = group), color = "black", alpha = 0.8) +
  scale_fill_manual(name = "Ecology Regions",
                    values = c("#2c7bb6","#abd9e9","#d7191c","#fdae61"), 
                    labels = c("Southwest", "Northwest", "Central", "Eastern"),
                    guide = guide_legend(title.position = "top", order = 1)) +
  geom_point(data = fire_loc, aes(x = Longitude, y = Latitude, shape = "Fire Locations"), 
             color = "black", fill = "red") +
  scale_shape_manual(name = "", values = 24, 
                     guide = guide_legend(title.position = "top", order = 2)) +
  ggtitle("Washington State Ecology \n Regional Office Jurisdiction") +
  xlab("Longitude") +
  ylab("Latitude") +
  theme(#legend.position = "bottom",
        #legend.box = "horizontal"
    )

print(eco_region_map)

# import washington zip shapefile ----
#zip_path <- paste0('../../data/shape_files/wash_zip_2012_shapefile')
zip_path <- paste0('./data/shape_files/wash_zip_2012_shapefile')

wash_zip_map <- readOGR(dsn = zip_path, layer = "wash_zip_2012_shapefile")

# change projection to match eco region
wash_zip_map <- spTransform(wash_zip_map, "+proj=longlat +datum=WGS84")

#summary(eco_region)
#summary(wash_zip_map)

# check overlay visually 
#plot(eco_region, lwd = 5)
#plot(wash_zip_map, add = T)

# find the eco regions that are over a zipcode
zip_eco_region_overlap <- over(wash_zip_map, eco_region)
# output zip IDs from zip shapefile
ZIPCODE <- as.character(wash_zip_map@data$ZCTA5CE10)
# join eco region ids with zip (can be used to indicate eco region)
zip_eco_region <- cbind(ZIPCODE, zip_eco_region_overlap) %>% 
  mutate(eco_region = ifelse(ECY_REGION == "CRO", "Central",
                      ifelse(ECY_REGION == "ERO", "Eastern",
                      ifelse(ECY_REGION == "NWRO", "Northwestern",
                      ifelse(ECY_REGION == "SWRO", "Southwestern", NA))))) %>%  
  select(-OBJECTID, -ECY_REGION)

# range of PM values for each day in each eco region for each method ----
# find the min, median, and max values for each date
#summary(filter(zip_pm, date == "2012-07-9")) # quick sum of first zip obs 
# zip_pm zipcode needs to be character to join with region
zip_pm$ZIPCODE <- as.factor(zip_pm$ZIPCODE)

# glimpse(zip_pm)
# these joins should really be done in a loop, but...
# each column would need to be specified for each pm val, 
# which I'd have to think about more than just each one manually
# wrf smoke
wrf_range <- zip_pm %>% 
  full_join(zip_eco_region, by = "ZIPCODE") %>%  
  group_by(date, eco_region) %>% 
  summarise(min = min(wrf_smk_pm, na.rm = T), 
            med = median(wrf_smk_pm, na.rm = T), 
            max = max(wrf_smk_pm, na.rm = T)) %>% 
  # variable for small multiples
  mutate(est_method = "wrf_smk")

# krig smoke
krig_range <- zip_pm %>% 
  full_join(zip_eco_region, by = "ZIPCODE") %>%  
  group_by(date, eco_region) %>% 
  summarise(min = min(krig_smk_pm, na.rm = T), 
            med = median(krig_smk_pm, na.rm = T), 
            max = max(krig_smk_pm, na.rm = T)) %>% 
  mutate(est_method = "krig_smk")

# geo smoke
geo_range <- zip_pm %>% 
  full_join(zip_eco_region, by = "ZIPCODE") %>%  
  group_by(date, eco_region) %>% 
  summarise(min = min(geo_smk_pm, na.rm = T), 
            med = median(geo_smk_pm, na.rm = T), 
            max = max(geo_smk_pm, na.rm = T)) %>% 
  mutate(est_method = "geo_smk")

# loop plots though

file_name <- c('wrf', 'krig', 'geo')
title_name <- c('WRF-Chem Smoke', 'Kriging Smoke', 'GWR Smoke')
ts_list <- list(wrf_range, krig_range, geo_range)

for(i in 1:3){

df_range <- ts_list[[i]]  
gg_title <- title_name[i]

# create plot for method range
plot_print <- ggplot(df_range, aes(x = date)) +
  geom_ribbon(aes(ymin = min, ymax = max, group = eco_region, 
                  fill=eco_region), alpha = 0.7) +
  scale_fill_manual(name = "Washington State Ecology \n Regional Office Jurisdiction",
                    values = c("#d7191c", "#fdae61", "#abd9e9", "#2c7bb6"),
                    labels = c("Central", "Eastern","Northwest", "Southwest"),
                    guide = F) +
  ylim(0,500) +
  #geom_line(aes(y=med, group=eco_region, color=eco_region)) +
  ggtitle(gg_title) +
  ylab(expression('Smoke PM2.5 µg/m'^3)) +
  xlab('Date') + 
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    axis.title.y = element_text(angle = 90, size = 12),
    axis.title.x = element_text(size = 14), 
    strip.background = element_rect(fill = 'white'))

# file name
name <- file_name[i]

assign(paste0(name,"_plot"), plot_print)

}

# Layout needs to be edited and I would like to change colors
# but this should be good enough for a review by CDC
map_timeseries <- ggdraw() +
  draw_plot(eco_region_map, x = 0, y = 0.5, width = 0.5, height = 0.5) +
  draw_plot(wrf_plot, x = 0.5, y = 0.5, width = 0.5, height = 0.5) +
  draw_plot(krig_plot, x = 0, y = 0, width = 0.5, height = 0.5) +
  draw_plot(geo_plot, x = 0.5, y = 0, width = 0.5, height = 0.5)

plot(map_timeseries)
# save 6 inch by 4 inch
ggsave("./fig1.pdf", plot = map_timeseries, 
       width = 12, height = 8, units = "in")

```

## Days of Smoke In Washington

Here is a figure of the days from May 1st to September 30 2013 where PM~2.5~ exceeded >10.

```{r smoke map, message = F, warning = F, echo = F, results='asis'}

# fire location has a weird date format, but I think is only september

# zip code count of smoke ------------------------------------------------------
zip_smoke_days <- zip_pm %>% 
  # create binary indicators of days with smoke
  mutate(wrf_smoke = if_else(wrf_smk_pm > 0, 1, 0),
         wrf_smoke5 = if_else(wrf_smk_pm > 5, 1, 0),
         wrf_smoke10 = if_else(wrf_smk_pm > 10, 1, 0),
         geo_smoke = if_else(geo_smk_pm > 0, 1, 0),
         geo_smoke5 = if_else(geo_smk_pm > 5, 1, 0),
         geo_smoke10 = if_else(geo_smk_pm > 10, 1, 0)) %>% 
  group_by(ZIPCODE) %>% 
  summarise(wrf_smk_days = sum(wrf_smoke), wrf_smk5_days = sum(wrf_smoke5), 
            wrf_smk10_days = sum(wrf_smoke10), geo_smk_days = sum(geo_smoke), 
            geo_smk_days = sum(geo_smoke), geo_smk5_days = sum(geo_smoke5),
            geo_smk10_days = sum(geo_smoke10)) %>% 
  # change zipcode to character
  mutate(ZIPCODE = as.character(ZIPCODE))

# check smoke days
#summary(zip_smoke_days)

# tidy from broom package now works
wash_zip_df <- tidy(wash_zip_map, region = "ZCTA5CE10") %>% 
  # merge in smoke days
  full_join(zip_smoke_days, by = c("id" = "ZIPCODE"))

# map
print_map <- ggmap(washington_map) + 
  geom_polygon(data = wash_zip_df, 
    aes(x=long, y=lat, group = group, fill = geo_smk10_days), alpha = 0.7) +
  scale_fill_gradient(expression("Number of \nSmoke Days"),
    low = '#2c7bb6', high = '#d7191c') + 
  geom_point(data = fire_loc, aes(x = Longitude, y = Latitude, shape = "Fire Locations"), 
             colour = "black", fill = "red") +
  scale_shape_manual(name = "", values = 24) + 
  xlab("Longitude") +
  ylab("Latitude") +
  # control legend/guide lable order
  guides(fill = guide_colourbar(title.position = "top", order = 1),
         shape = guide_legend(title.position = "top", order = 2))

print(print_map)

ggsave("./fig2.pdf", plot = print_map, 
       width = 12, height = 8, units = "in")

```

## Population-Weighted PM~2.5~ Assigned by Zip Code

### Descriptives

Note for following tables: I wrote out 'Greater than rather' than using  '>' as it was increasing the following font and looked weird.

```{r descriptive table, echo = F, warning = F, results='asis'}

# dataframe list
df_list <- list(resp_casecross, asthma_casecross, copd_casecross, 
  pneum_casecross, acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
  cereb_vas_casecross, hf_casecross, ihd_casecross, mi_casecross, 
  broken_arm_casecross)

outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 
                  'Pneumonia', 'Acute Bronchitis', 'Cardiovascular Disease',
                  'Arrhythmia', 'Cerebrovascular Disease', 'Heart Failure',
                  'Ischemic Heart Disease', 'Myocardial Infarction',
                  'Broken Arm')

method_list <- c('WRF-Chem Smoke', 'Geo-Weighted Smoke', 'Kriging Smoke')

# create an empty list to row bind dataframes together
datalist <- list()

# set list and do cross tabs to find values
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){

  i = 2
# dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- colnames(df_to_loop[76])
  # outcome name
  outcome_name <- outcome_list[i]

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- df_to_loop %>%
    # only look at outcomes
    filter(outcome == 1) %>%
    # add another row that makes sure there is a person <15 in the dataframe
    # tricking xtabs to produce a 0 cell for the outcome for age <15
    add_row(outcome = 0, age_ind = 0)

  # cross tabs
  outcome_n <- xtabs(~ outcome, df_analysis)
  cross_tab_age <- xtabs(~ outcome + age_ind, df_analysis)
  cross_tab_sex <- xtabs(~ outcome + sex_ind, df_analysis)
  # empty matrix
  point_estimates <- matrix(nrow = 1, ncol = 7, byrow = T)

  colnames(point_estimates) <- c("outcome", "n", "age_15", "age_15_65", 
                                 "age_65", "female", "male")

  # fill in the outcome name for the dataframe before the loop
  point_estimates[, 1] <- outcome_name
  # fill n
  point_estimates[, 2] <- outcome_n[2] # second element of the 1 dimension vector
  # age <15
  point_estimates[, 3] <- cross_tab_age[2, 1]
  # age 15 to 65
  point_estimates[, 4] <- cross_tab_age[2, 2]
  # age >65
  point_estimates[, 5] <- cross_tab_age[2, 3]
  # male == 0
  point_estimates[, 7] <- cross_tab_sex[1, 1]
  # female == 1
  point_estimates[, 6] <- cross_tab_sex[1, 2]

  # save point estimates as a dataframe
  point_est_df <- as_data_frame(point_estimates)

  # combine previous values in dataframe that has all outcome/methods comparisons
  datalist[[i]] <- point_est_df

} # end of loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist) %>% 
  # find proportions/percents for each strata in a row
  mutate(age_15_pr = as.character(round((as.numeric(age_15)/as.numeric(n))*100,1)),
         age_15_65_pr = as.character(round((as.numeric(age_15_65)/as.numeric(n))*100,1)),
         age_65_pr = as.character(round((as.numeric(age_65)/as.numeric(n))*100,1)),
         female_pr = as.character(round((as.numeric(female)/as.numeric(n))*100,1)),
         male_pr = as.character(round((as.numeric(male)/as.numeric(n))*100,1))) %>% 
  select(outcome, n, age_15, age_15_pr, age_15_65, age_15_65_pr, age_65,
         age_65_pr, female, female_pr, male, male_pr)

#str(combined_point_est_df)

tab <- htmlTable(txtRound(combined_point_est_df, digits = 1), 
           caption = "Number of cases for each outcome observed from July 1st to October 31st, 2012",
           # column headers
           header = c("Outcome", "Cases n", "Less than 15", "(%)", "15 to 65", "(%)", "Greater than 65",
                      "(%)", "Female", "(%)", "Male", "(%)"),
           # column spanner
           cgroup = c("","Age Category", "Sex"), 
           n.cgroup = c(2, 6, 4),
           padding.rgroup = "&nbsp;&nbsp;",
           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
           align = "llccccc" # column alignment,
            ) # end table

print(tab)

```

### Background on Time-Stratified Case-Crossover Design

The original time-stratified case-crossover design suggested picking referent periods based on any potential time-varying confounding, *a priori*. For example, if day of the week and the month of the exposure sequence is of concern, one might pick referents in the same day of the week and the same day of the month. The papers that discuss and support the use of a time-stratified design are the Janes et al. 2005 **Epidemiology paper**, *Case-Crossover Analyses of Air Pollution Exposure Data; Referent Selection Strategies and Their Implication for Bias*.

The authors state that with the time-stratified design, the *a priori* selection of the referent period on the same day of the week within the same month has the following benefits:

1. Localizable - Meaning that the likelihood of index times conditional on the referent windows contains information about \beta, which I think means the relative difference in the exposed vs unexposed, conditioned on the exposure (air pollution in their example), is unbiased.

2. Non-ignroable - Meaning that time-varying factors can be ignored when using a conditional logistic regression; controlled through design (matching on day of week within the same season).

More referent periods is better for statistical efficiency, but there is a tradeoff, where additional referent periods may introduce what they call 'overlap bias', which they mathmatically define in Janes 2005 **Statistics in Medicine** paper, *Overlap Bias in the Case-Crossover Design, with Applications to Air Pollution Exposures*. Basically this bias occurs when the exposure assigned in the referent period is conditional on the strategy used to define the referent period.

As Janes et al. state in their 2005 **Statistics in Medicine** paper:

  > *Two of the main strengths of the case-crossover design in the air pollution context are
that it controls fixed confounders and allows for control over time-dependent confounders
by design. Choosing referents that are restricted to the same day of the week and season
as the index time will control for these effects. Though we have not examined the degree
to which various referent selection strategies control bias due to confounding, several points
are clear. Referents that are closer to the index time will achieve greater control of seasonal
confounding. However, sampling referents closer to the index time will also result in a loss
of power due to less heterogeneity in the exposure series, and a smaller number of referents
will also result in lower power.*

I will use both Janes 2005 papers in **Epidemiology** and **Statistics in Medicine** to justify the use of the time-stratified design.

This design has an advantage over other case-crossover designs I've explored in that you can use more referent times, and as the exposure at these times are averaged for each subject over a set time period, this balances out some of the more extreme values of PM~2.5~ that may persist over a couple weeks due to wildfire smoke.

### Time-Stratified Case-Crossover (referents same day of week within fire season July - October)

For these analyses and for the final results in the paper, I used referents within the entire wildfire season (July to October). This time-stratified referent strategy allows even more referent observations to average to 'smooth' out some of those extreme PM~2.5~ concentrations in some referent periods. For my sensitivity analysis of varying referent periods, please see other scripts.

Adjusting for CDC Met Variables: Rish has provided zip-level meteorologic data on wind speed, precipitation, relative humidity, and temperature. Note a slight difference that will likely need to be addressed in discussion is that these measures are weighted based on census tract, not on the WRF grid like our measures of smoke.

```{r data wrangle with zip met vars, warning = F, echo = F, message = F}

# create dataset that can be merged to case-cross over dataframes
daily_zip_met <- zip_met %>%
  # limit to new mean met variables (getting rid of smoke variables I already have)
  select(1:7) %>% 
  rename(ZIPCODE = ZIP) %>% arrange(ZIPCODE, date) %>%
  group_by(ZIPCODE) %>% 
  # create lag variables
  mutate(
    # absolute temp
    meanAT_lag1 = lag(daily_meanAT, 1, order_by = ZIPCODE), 
    meanAT_lag2 = lag(daily_meanAT, 2, order_by = ZIPCODE),
    meanAT_lag3 = lag(daily_meanAT, 3, order_by = ZIPCODE),
    meanAT_lag4 = lag(daily_meanAT, 4, order_by = ZIPCODE),
    meanAT_lag5 = lag(daily_meanAT, 5, order_by = ZIPCODE),
    # relative humidity
    meanRH_lag1 = lag(daily_meanRH, 1, order_by = ZIPCODE),
    meanRH_lag2 = lag(daily_meanRH, 2, order_by = ZIPCODE),
    meanRH_lag3 = lag(daily_meanRH, 3, order_by = ZIPCODE),
    meanRH_lag4 = lag(daily_meanRH, 4, order_by = ZIPCODE),
    meanRH_lag5 = lag(daily_meanRH, 5, order_by = ZIPCODE),
    # temp
    meanT_lag1 = lag(daily_meanT, 1, order_by = ZIPCODE),
    meanT_lag2 = lag(daily_meanT, 2, order_by = ZIPCODE),
    meanT_lag3 = lag(daily_meanT, 3, order_by = ZIPCODE),
    meanT_lag4 = lag(daily_meanT, 4, order_by = ZIPCODE),
    meanT_lag5 = lag(daily_meanT, 5, order_by = ZIPCODE),
    # precip
    meanPrcp_lag1 = lag(daily_meanPrcp, 1, order_by = ZIPCODE),
    meanPrcp_lag2 = lag(daily_meanPrcp, 2, order_by = ZIPCODE),
    meanPrcp_lag3 = lag(daily_meanPrcp, 3, order_by = ZIPCODE),
    meanPrcp_lag4 = lag(daily_meanPrcp, 4, order_by = ZIPCODE),
    meanPrcp_lag5 = lag(daily_meanPrcp, 5, order_by = ZIPCODE),
    # wind speed
    meanWS_lag1 = lag(daily_meanWS, 1, order_by = ZIPCODE),
    meanWS_lag2 = lag(daily_meanWS, 2, order_by = ZIPCODE),
    meanWS_lag3 = lag(daily_meanWS, 3, order_by = ZIPCODE),
    meanWS_lag4 = lag(daily_meanWS, 4, order_by = ZIPCODE),
    meanWS_lag5 = lag(daily_meanWS, 5, order_by = ZIPCODE)) %>% 
  # sort so all like variables are together
  select(1:3, 8:12, 4, 13:17, 5, 18:22, 6, 23:27, 7, 28:32)

#which(colnames(daily_zip_met)=="meanAT_lag5")
#glimpse(daily_zip_met)

```

```{r time strat within fire season adj for cdc met, warning = F, echo = F, results='asis'} 
# note above, results = 'asis' needed to print htmlTables

# I removed copd exacerbation and rheuamtoid arthritis

# dataframe list
df_list <- list(resp_casecross, asthma_casecross, copd_casecross, 
  pneum_casecross, acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
  cereb_vas_casecross, hf_casecross, ihd_casecross, mi_casecross, 
  broken_arm_casecross)

outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 
                  'Pneumonia', 'Acute Bronchitis', 'Cardiovascular Disease',
                  'Arrhythmia', 'Cerebrovascular Disease', 'Heart Failure',
                  'Ischemic Heart Disease', 'Myocardial Infarction',
                  'Broken Arm')

method_list <- c('WRF-Chem Smoke', 'Geo-Weighted Smoke', 'Kriging Smoke')

# create an empty list to row bind dataframes together
datalist <- list()

# data wrangling ----
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){

# dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- colnames(df_to_loop[76])
  # outcome name
  outcome_name <- outcome_list[i]

  # extract covariates from dataframe
  covariates_df <- df_to_loop[, c(1:26, 71, 74, 76:85)]

  # extract pm values and divide by 10 and ordered
  which(colnames(df_to_loop)=="background_pm_zip") # code to find column numbers, 91
  pm_estimates_df <- df_to_loop[, c(88, 92:94)]/10  # create 10 unit increases

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
    # remove missing pm values
    filter(!is.na(wrf_smk_pm_zip)) %>%

    # join in cdc met data
    # full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
    # the following code makes sure that the counterfactual values retained are 
    # symetric in that number of obs before = number of obs after
    mutate(obs_diff_admission = (fromdate - date)/7) 
    # dataframe is already for the entire fire season, so I don't need to subset anymore

  # empty df for table
  table_df <- data.frame()

  # empty matrix
  point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)

  colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 'odds_ratio', 
                                 'lower95', 'upper95', 'se', 'p_val')

  # fill in the outcome name for the dataframe before the loop
  point_estimates[, 1] <- outcome_name

  # second loop to run a model for each pm estimation method
    for(j in 39:41){

      # variable to model 
      var_name <- colnames(df_analysis[j])

      # conditional logistic regression model
      mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)

      # some model/DAG checks of data rish provided

      #glimpse(df_analysis)

      # adjusting for admit month is not possible as it's the same var for the subject.
      # adjusting for month smoke is a collider and should not be adjusted for.
      #summary(clogit(outcome ~ geo_smk_pm_zip + daily_meanRH + daily_meanT + daily_meanWS + daily_meanPrcp +
      #                as.factor(month_smk) + strata(PATIENTID), df_analysis))
      #outocme association: daily_meanRH,

      # check for confoudning following DAG assumptions
      #summary(clogit(outcome ~ wrf_temp_zip + strata(PATIENTID), df_analysis))
      #summary(clogit(outcome ~ wrf_temp_zip + strata(PATIENTID), df_analysis))
      # smoke predictor association: daily_meanRH, 
      #summary(lmer(geo_smk_pm_zip ~ daily_meanWS + (1 | PATIENTID), df_analysis ))
      #summary(lmer(wrf_temp_zip ~ daily_meanT + (1 | PATIENTID), df_analysis ))
      #cor(df_analysis$wrf_temp_county, df_analysis$daily_meanT)

      # populate matrix
      row_n <- j-38

      point_estimates[row_n, 2] <- method_list[row_n]
      point_estimates[row_n, 3] <- mod$n
      point_estimates[row_n, 4] <- mod$nevent
      # odds ratio
      point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)

      # 95% lower bound
      point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -
                                        1.96*(summary(mod)$coefficient[1,3])), 3)
      # 95% upper bound
      point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
                                        1.96*(summary(mod)$coefficient[1,3])), 3)
      # standard error
      point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
      # p val
      point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)

      # save point estimates as a dataframe
      point_est_df <- as_data_frame(point_estimates)

    }

# combine previous values in dataframe that has all outcome/methods comparisons
datalist[[i]] <- point_est_df

} # end of loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist)

# subset columns I want to put in to the table
table_df <- combined_point_est_df %>% select(2, 3:7)

  tab <- htmlTable(txtRound(table_df, digits = 3, 1:3), 
           caption = "Association between a 10 ug/m^3 in PM2.5 and Health Outcomes",
           # row group by outcome
           rgroup = outcome_list,
           n.rgroup = c(rep(3, 12)), # 4 rows for each method for each outcome
           # column headers
           header = c("Method", "Obs.", "Events",
                      "OR&dagger;", "Lower", "Upper"),
           # column spanner
           cgroup = c("", "95% CI"), 
           n.cgroup = c(4, 2),
           padding.rgroup = "&nbsp;&nbsp;",
           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
           align = "llccccc", # column alignment,
           tfoot="&dagger; Adjusted for CDC temperature, relatively humidity, wind speed, and precipitation; accounting for subject. Time-stratified: referent periods matched to events on same day of week within July to October fire season."
            ) # end table

  print(tab)

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
combined_point_est_df$outcome <- factor(combined_point_est_df$outcome, 
                                        levels = unique(combined_point_est_df$outcome))

combined_point_est_df$pm_method <- factor(combined_point_est_df$pm_method, 
                                          levels = unique(combined_point_est_df$pm_method))

combined_point_est_df$n <- as.numeric(combined_point_est_df$n)
combined_point_est_df$n_events <- as.numeric(combined_point_est_df$n_events)
combined_point_est_df$odds_ratio <- as.numeric(combined_point_est_df$odds_ratio)
combined_point_est_df$lower95 <- as.numeric(combined_point_est_df$lower95)
combined_point_est_df$upper95 <- as.numeric(combined_point_est_df$upper95)
combined_point_est_df$se <- as.numeric(combined_point_est_df$se)
combined_point_est_df$p_val <- as.numeric(combined_point_est_df$p_val)

# subset out global method
combined_point_est_df <- combined_point_est_df

## ggplot
  print_plot <- ggplot(combined_point_est_df, 
                       aes(x = pm_method, y = odds_ratio, colour = pm_method)) +
    # custom color 
    scale_color_manual(name = "Smoke-Estimation Method", 
                       values = c("red", "blue", "#32115C"),
                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
    geom_point() + #geom_text(vjust = 0, nudge_x = 0.3) +
    geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
    facet_wrap(~outcome, nrow = 3) +
    geom_hline(yintercept = 1, linetype=2) +
#    ggtitle('Association Between PM2.5 \n from Wildfire Smoke on Hospitalizations') +
    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
    #xlab('Smoke Estimation Method') +
    # plot theme
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # strip element
    strip.background = element_rect(colour=NA, fill=NA),
    panel.border = element_rect(fill = NA, color = "black"),
    # facet text size
    #strip.text = element_text(size = 8),
    # axis element
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(angle = 90),
    # legend elements
    legend.position = "bottom")
    #legend.text = element_text(size = 8))

  print(print_plot)
  # save figure
  ggsave("./fig4.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

# figures for research day
  # plot of only respiratory outcomes

  xtabs(~ combined_point_est_df$outcome)
# subset out global method and respiratory outcomes ----
resp_point_est_df <- combined_point_est_df %>%

  filter(outcome == "All Respiratory" |
         outcome == "Asthma" |
         outcome == "COPD" |
         outcome == "Pneumonia" |
         outcome == "Acute Bronchitis" |
         outcome == "Broken Arm")

## ggplot
  resp_plot <- ggplot(resp_point_est_df, 
                      aes(x = pm_method, y = odds_ratio, colour = pm_method)) +
    # custom color 
    scale_color_manual(name = "Smoke-Estimation Method", 
                       values = c("red", "blue", "#32115C"),
                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
    geom_point() + #geom_text(vjust = 0, nudge_x = 0.3) +
    geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
    facet_wrap(~outcome, nrow = 2) +
    geom_hline(yintercept = 1, linetype=2) +
#    ggtitle('Association Between PM2.5 \n from Wildfire Smoke on Hospitalizations') +
    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
    #xlab('Smoke Estimation Method') +
    # plot theme
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # strip element
    strip.background = element_rect(colour=NA, fill=NA),
    panel.border = element_rect(fill = NA, color = "black"),
    # facet text size
    #strip.text = element_text(size = 8),
    # axis element
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(angle = 90),
    # legend elements
    legend.position = "bottom")
    #legend.text = element_text(size = 8))

  print(resp_plot)

# subset out global method and respiratory outcomes ----
cvd_point_est_df <- combined_point_est_df %>% 
   filter(outcome == "Cardiovascular Disease" |
         outcome == "Arrhythmia" |
         outcome == "Cerebrovascular Disease" |
         outcome == "Heart Failure" |
         outcome == "Ischemic Heart Disease" |
         outcome == "Myocardial Infarction")

## ggplot
  cvd_plot <- ggplot(cvd_point_est_df, 
                      aes(x = pm_method, y = odds_ratio, colour = pm_method)) +
    # custom color 
    scale_color_manual(name = "Smoke-Estimation Method", 
                       values = c("red", "blue", "#32115C"),
                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
    geom_point() + #geom_text(vjust = 0, nudge_x = 0.3) +
    geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
    facet_wrap(~outcome, nrow = 2) +
    geom_hline(yintercept = 1, linetype=2) +
#    ggtitle('Association Between PM2.5 \n from Wildfire Smoke on Hospitalizations') +
    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
    #xlab('Smoke Estimation Method') +
    # plot theme
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # strip element
    strip.background = element_rect(colour=NA, fill=NA),
    panel.border = element_rect(fill = NA, color = "black"),
    # facet text size
    #strip.text = element_text(size = 8),
    # axis element
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(angle = 90),
    # legend elements
    legend.position = "bottom")
    #legend.text = element_text(size = 8))

  print(cvd_plot)

```

#### Result Summary: Time-Stratified Case-Crossover Design (same day of week within the season)

A general interpretation is as follows:
A 10 $\mu$g/m^3^ increase in PM~2.5~ attributed to wildfire smoke was associated with a X% increase in the risk for a health event. Where X% is the odds ratio - 1 (e.g. OR of 1.20 - 1 = 20% increase). 
Note that the odds ratio approximates the relative risk in this case due to the rare outcome assumption in that the calcuation for an odds ratio is similar to risk ratio calcuation.

We see that all exposure methods are significantly associated with an increase in all respiratory outcomes and asthma. The WRF-Chem smoke method is associated with cerebrovascular disease as well. As for the other models, Kriging, Global Regression, and Geo-Weighted Regression, those are associated with COPD. I believe this is the most justifiable design approach that would average any extreme values of PM~2.5~ across the wildfire season for the vector of referent values for a particular subject.

When adjusting for the CDC met variables, the results are slightly attenuated when compared to my adjustments with just WRF-Temp, but my overall conclusions do not change much.

Here is the same table, but within season instead of fire season and adjusting for CDC met variables. Rish is worried about time-variant confounding within a subject that may bias the overall results. This might include seasonal differences in baseline risk of asthma.

### Time-Stratified by Age Categories

Following table and figure look at some outcomes stratified by age category. Note that kids <=15 likely won't have many outcomes like COPD or MI. I have restricted the dataset to just the geo smoke method to reduce the number of rows/points on the tables and figures to make it easier to read for now. Code can easily be changed to have all PM~2.5~ methods.

For some outcomes, there may be some evidence of effect modificaiton, like COPD, where >65 age group is at significant risk. However, the point estimates are relatively similar for the 15 to 65 strata, just not significant. Some error bars are not shown as they are too wide; usually <15 age category where there are only a handful of outcomes.

```{r time stratified age strata cdc met, warning =F, echo = F, results='asis'} 
# note above, results = 'asis' needed to print htmlTables

# For age categories, I want a loop for each age category. 
# Presenting just WRF and GWR

# dataframe list
df_list <- list(resp_casecross, asthma_casecross, copd_casecross, pneum_casecross,
                acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
                cereb_vas_casecross, hf_casecross, ihd_casecross, mi_casecross, 
                broken_arm_casecross)

outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 'Pneumonia', 
                  'Acute Bronchitis', 'Cardiovascular Disease', 'Arrhythmia', 
                  'Cerebrovascular Disease', 'Heart Failure',
                  'Ischemic Heart Disease', 'Myocardial Infarction', 'Broken Arm')

# looking only at wrf chem, kriging, and geo-weighted; taking out 'Global Smoke' method
method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Global Smoke', 'GWR Smoke')

# age category list
age_cat_list <- c(0,1,2)

# create an empty list to row bind dataframes together
datalist1 <- list()
datalist2 <- list()

# data wrangling ----
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){

# dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- colnames(df_to_loop[3])
  # outcome name
  outcome_name <- outcome_list[i]

  # extract covariates from dataframe
  covariates_df <- df_to_loop[, c(1:16, 27:28, 151:157)]

  # extract pm values and divide by 10 and ordered
  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
  pm_estimates_df <- df_to_loop[, c(19, 26, 25, 24)]/10  # create 10 unit increases

  # new loop for age categories
  for(k in 0:2){

    # empty matrix (12 x 10 matrix)
    point_estimates <- matrix(nrow = 4, ncol = 10, byrow = T)

    colnames(point_estimates) <- c('outcome', 'age_cat', 'pm_method', 'n', 'n_events', 
                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')

    # fill in the outcome namedataframe before method loop
    point_estimates[, 1] <- outcome_name
    # repeat the age category 4 times for each pm method
    point_estimates[, 2] <- k

    # dataframe for analysis creation
    # bind columns back together 
    df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
      # remove missing pm values
      filter(!is.na(wrf_smk_pm_zip)) %>% 
      # join in cdc met data
      full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
      # limit to emergency or urgent care
      filter(ADM_TYPE == 1 | ADM_TYPE == 2) %>%
      # limit to specific age category
      filter(age_cat == k) %>% 
      # the following code makes sure that the counterfactual values retained are 
      # symetric in that number of obs before = number of obs after
      mutate(obs_diff_admission = (date - date_admit)/7) 
      # dataframe is already for the entire fire season, so I don't need to subset anymore

  # second loop to run a model for each pm estimation method
    for(j in 26:29){

      # variable to model 
      var_name <- colnames(df_analysis[j])

      # set row number to fill
      row_n <- j-25

      # only run the model if the dataframe has observations
      if(nrow(df_analysis) != 0){
      # conditional logistic regression model
      mod <- clogit(outcome ~ df_analysis[[j]] + daily_meanRH + daily_meanT + daily_meanWS + 
                    daily_meanPrcp + strata(PATIENTID), df_analysis)

      # populate matrix
      point_estimates[row_n, 3] <- method_list[row_n]
      point_estimates[row_n, 4] <- mod$n
      point_estimates[row_n, 5] <- mod$nevent
      # odds ratio
      point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)

      # 95% lower bound
      point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
                                        1.96*(summary(mod)$coefficient[1,3])), 3)
      # 95% upper bound
      point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
                                        1.96*(summary(mod)$coefficient[1,3])), 3)
      # standard error
      point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
      # p val
      point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)

      # create else statement that fills matrix with missing so I still have the row
      # in the final dataframe
      } else {point_estimates[row_n, 3] <- method_list[row_n]
              point_estimates[row_n, 4] <- 0
              point_estimates[row_n, 5] <- 0
              point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement

    } # end methods loop

    # save point estimates as a dataframe
    point_est_df <- as_data_frame(point_estimates)

    # combine previous values in dataframe that has all outcome/methods comparisons
    datalist1[[k+1]] <- point_est_df
  } # end age category loop

  # bind rows of age category estimates together
 age_est_df <- bind_rows(datalist1)

 # populate second dataframe list
 datalist2[[i]] <- age_est_df

} # end of outcome loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist2)

wrf_geo_age <- combined_point_est_df %>% 
  # filter columns to just geo-smk
  filter(pm_method == "GWR Smoke" |
         pm_method == "WRF-Chem Smoke") %>% 
  mutate(age_cat2 = ifelse(age_cat == 0, "Less than 15", 
                    ifelse(age_cat == 1, "15-65",
                    ifelse(age_cat == 2, "Greater than 65", NA)))) %>% 
  # subset columns I want to put in to the table
  select(3, 11, 5:8) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
         lower95 = ifelse(lower95 == 99, "--", lower95),
         upper95 = ifelse(upper95 == 99, "--", upper95))

  tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by age",
           # row group by outcome
           rgroup = outcome_list,
           n.rgroup = c(rep(6, 12)), # 6 rows for each age cat for each outcome
           # column headers
           header = c("Est Method", "Age Group", "Events",
                      "OR&dagger;", "Lower", "Upper"),
           # column spanner
           cgroup = c("", "95% CI"), 
           n.cgroup = c(4, 2),
           padding.rgroup = "&nbsp;&nbsp;",
           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
           align = "llcccc", # column alignment,
           tfoot="&dagger; Adjusted for CDC temperature, wind speed, relative humidity, precipitation, and accounting for subject. Time-stratified: referent periods matched to events on same day of week within July to October fire season."
            ) # end table

  print(tab)

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
wrf_geo_age_plot <- combined_point_est_df %>% 
  # filter columns to just geo-smk
  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
  # do not plot results with less than 15 events, create missing vals
  # in mutate
  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
                             99, odds_ratio),
         lower95 = ifelse(as.numeric(n_events) < 15, 
                             99, lower95),
         upper95 = ifelse(as.numeric(n_events) < 15, 
                             99, upper95),
         age_cat2 = ifelse(age_cat == 0, "<15", 
                    ifelse(age_cat == 1, "15-65",
                    ifelse(age_cat == 2, ">65", NA)))) %>% 
  # subset columns I want to put in to the table
  select(1, 3, 11, 5:8) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
         lower95 = ifelse(lower95 == 99, NA, lower95),
         upper95 = ifelse(upper95 == 99, NA, upper95))

# change characters to numeric and factor  
wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
                              levels = unique(wrf_geo_age_plot$outcome))

wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
                                levels = unique(wrf_geo_age_plot$pm_method))

wrf_geo_age_plot$age_cat2 <- factor(wrf_geo_age_plot$age_cat2,
                                levels = unique(wrf_geo_age_plot$age_cat2))

wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)

## ggplot 
  print_plot <- ggplot(wrf_geo_age_plot,
    aes(x = age_cat2, y = odds_ratio, color = pm_method), na.rm = T) +
    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
    # custom color 
    scale_color_manual(name = "Smoke-Estimation Method", 
                       values = c("red", "#32115C"),
                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
    facet_wrap(~outcome, nrow = 3, scales = "free") +
    geom_hline(yintercept = 1, linetype=2) +
    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
    #ylim(0, 2) +
    xlab('Age Category ') +
    # plot theme
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # strip element
    strip.background = element_rect(colour=NA, fill=NA),
    panel.border = element_rect(fill = NA, color = "black"),
    # facet text size
    strip.text = element_text(size = 10),
    # axis element
    #axis.text.x = element_blank(),
    #axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(angle = 90),
    # legend elements
    legend.position = "bottom")
    #legend.text = element_text(size = 8))

  print(print_plot)
  # save figure
  ggsave("./fig5.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

```

### Seasonal Time-Stratified by Sex

Association with smoke PM~2.5~ and health outcomes stratified by sex. May be some evidence of effect modification by sex, particularly where males appear to be at greater risk for certain CVD outcomes with increasing smoke. However, only marginally significant. Also, broken arm estimate seems a little strange, suggesting females at greater risk too. Larger sample size may help to answer this.

```{r sex strata cdc met adj, warning = F, echo = F, results='asis'} 
# note above, results = 'asis' needed to print htmlTables

# For age categories, I want a loop for each age category.

# dataframe list
df_list <- list(resp_casecross, asthma_casecross, copd_casecross, pneum_casecross,
                acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
                cereb_vas_casecross, hf_casecross, ihd_casecross, mi_casecross, 
                broken_arm_casecross)

outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 'Pneumonia', 
                  'Acute Bronchitis', 'Cardiovascular Disease', 'Arrhythmia', 
                  'Cerebrovascular Disease', 'Heart Failure',
                  'Ischemic Heart Disease', 'Myocardial Infarction', 'Broken Arm')

# looking only at wrf chem, kriging, and geo-weighted; taking out 'Global Smoke' method
method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Global Smoke', 'GWR Smoke')

# sex category list
sex_strata_list <- c(0,1)

# create an empty list to row bind dataframes together
datalist1 <- list()
datalist2 <- list()

# data wrangling ----
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){

# dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- colnames(df_to_loop[3])
  # outcome name
  outcome_name <- outcome_list[i]

  # extract covariates from dataframe
  covariates_df <- df_to_loop[, c(1:16, 27:28, 151:157)]

  # extract pm values and divide by 10 and ordered
  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
  pm_estimates_df <- df_to_loop[, c(19, 26, 25, 24)]/10  # create 10 unit increases

  # new loop for sex categories
  for(k in 0:1){

    # empty matrix (12 x 10 matrix)
    point_estimates <- matrix(nrow = 4, ncol = 10, byrow = T)

    colnames(point_estimates) <- c('outcome', 'sex', 'pm_method', 'n', 'n_events', 
                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')

    # fill in the outcome namedataframe before method loop
    point_estimates[, 1] <- outcome_name
    # repeat the sex category 4 times for each pm method
    point_estimates[, 2] <- k

    # dataframe for analysis creation
    # bind columns back together 
    df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
      # remove missing pm values
      filter(!is.na(wrf_smk_pm_zip)) %>% 
      # join in cdc met data
      full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
      # limit to emergency or urgent care
      filter(ADM_TYPE == 1 | ADM_TYPE == 2) %>%
      # limit to specific sex category
      filter(sex_num == k) %>% 
      # the following code makes sure that the counterfactual values retained are 
      # symetric in that number of obs before = number of obs after
      mutate(obs_diff_admission = (date - date_admit)/7) 
      # dataframe is already for the entire fire season, so I don't need to subset anymore

  # second loop to run a model for each pm estimation method
    for(j in 26:29){

      # variable to model 
      var_name <- colnames(df_analysis[j])

      # set row number to fill
      row_n <- j-25

      # only run the model if the dataframe has observations
      if(nrow(df_analysis) != 0){
      # conditional logistic regression model
      mod <- clogit(outcome ~ df_analysis[[j]] + daily_meanRH + daily_meanT + daily_meanWS + 
                    daily_meanPrcp + strata(PATIENTID), df_analysis)

      # populate matrix
      point_estimates[row_n, 3] <- method_list[row_n]
      point_estimates[row_n, 4] <- mod$n
      point_estimates[row_n, 5] <- mod$nevent
      # odds ratio
      point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)

      # 95% lower bound
      point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
                                        1.96*(summary(mod)$coefficient[1,3])), 3)
      # 95% upper bound
      point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
                                        1.96*(summary(mod)$coefficient[1,3])), 3)
      # standard error
      point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
      # p val
      point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)

      # create else statement that fills matrix with missing so I still have the row
      # in the final dataframe
      } else {point_estimates[row_n, 3] <- method_list[row_n]
              point_estimates[row_n, 4] <- 0
              point_estimates[row_n, 5] <- 0
              point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement

    } # end methods loop

    # save point estimates as a dataframe
    point_est_df <- as_data_frame(point_estimates)

    # combine previous values in dataframe that has all outcome/methods comparisons
    datalist1[[k+1]] <- point_est_df
  } # end sex category loop

  # bind rows of sex category estimates together
 sex_est_df <- bind_rows(datalist1)

 # populate second dataframe list
 datalist2[[i]] <- sex_est_df

} # end of outcome loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist2)

wrf_geo_sex <- combined_point_est_df %>% 
  # filter columns to just geo-smk
  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
  mutate(sex_cat = ifelse(sex == 0, "Male", 
               ifelse(sex == 1, "Female", NA))) %>% 
  # subset columns I want to put in to the table
  select(3, 11, 5:8)%>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
         lower95 = ifelse(lower95 == 99, "--", lower95),
         upper95 = ifelse(upper95 == 99, "--", upper95))

  tab <- htmlTable(txtRound(wrf_geo_sex, digits = 3, 1:3), 
           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by sex",
           # row group by outcome
           rgroup = outcome_list,
           n.rgroup = c(rep(4, 12)), # 2 rows for each sex strata for each outcome
           # column headers
           header = c("Est Method", "Sex Strata", "Events",
                      "OR&dagger;", "Lower", "Upper"),
           # column spanner
           cgroup = c("", "95% CI"), 
           n.cgroup = c(4, 2),
           padding.rgroup = "&nbsp;&nbsp;",
           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
           align = "llcccc", # column alignment,
           tfoot="&dagger; Adjusted for temperature, accounting for subject. Time-stratified: referent periods matched to events on same day of week within July to October fire season."
            ) # end table

  print(tab)

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
wrf_geo_sex_plot <- combined_point_est_df %>% 
  # filter columns to just geo-smk
  filter(pm_method == "GWR Smoke" | 
         pm_method == "WRF-Chem Smoke") %>% 
  mutate(sex_cat = ifelse(sex == 0, "Male", 
               ifelse(sex == 1, "Female", NA))) %>%
  # subset columns I want to put in to the table
  select(1, 3, 11, 5:8) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
         lower95 = ifelse(lower95 == 99, NA, lower95),
         upper95 = ifelse(upper95 == 99, NA, upper95))

# change characters to numeric and factor  
wrf_geo_sex_plot$outcome <- factor(wrf_geo_sex_plot$outcome,
                              levels = unique(wrf_geo_sex_plot$outcome))

wrf_geo_sex_plot$pm_method <- factor(wrf_geo_sex_plot$pm_method,
                                levels = unique(wrf_geo_sex_plot$pm_method))

wrf_geo_sex_plot$sex_cat <- factor(wrf_geo_sex_plot$sex_cat,
                                levels = unique(wrf_geo_sex_plot$sex_cat))

wrf_geo_sex_plot$n_events <- as.numeric(wrf_geo_sex_plot$n_events)
wrf_geo_sex_plot$odds_ratio <- as.numeric(wrf_geo_sex_plot$odds_ratio)
wrf_geo_sex_plot$lower95 <- as.numeric(wrf_geo_sex_plot$lower95)
wrf_geo_sex_plot$upper95 <- as.numeric(wrf_geo_sex_plot$upper95)

## ggplot
  print_plot <- ggplot(wrf_geo_sex_plot,
    aes(x = sex_cat, y = odds_ratio, color = pm_method), na.rm = T) +
    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
    # custom color 
    scale_color_manual(name = "Smoke-Estimation Method", 
                       values = c("red", "#32115C"),
                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
    facet_wrap(~outcome, nrow = 3, scales = "free") +
    geom_hline(yintercept = 1, linetype=2) +
    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
    #ylim(0, 2) +
    xlab('Sex') +
    # plot theme
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # strip element
    strip.background = element_rect(colour=NA, fill=NA),
    panel.border = element_rect(fill = NA, color = "black"),
    # facet text size
    strip.text = element_text(size = 10),
    # axis element
    #axis.text.x = element_blank(),
    #axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(angle = 90),
    # legend elements
    legend.position = "bottom")
    #legend.text = element_text(size = 8))

  print(print_plot)
  # save figure
  ggsave("./fig6.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

```

### Lag Periods

Lag period adjusted for CDC met variables.

```{r lags with cdc met adj, warning = F, echo = F, results='asis'} 
# note above, results = 'asis' needed to print htmlTables

# dataframe list
df_list <- list(resp_casecross, asthma_casecross, copd_casecross, pneum_casecross,
                acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
                cereb_vas_casecross, hf_casecross, ihd_casecross, mi_casecross, 
                broken_arm_casecross)

outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 'Pneumonia', 
                  'Acute Bronchitis', 'Cardiovascular Disease', 'Arrhythmia', 
                  'Cerebrovascular Disease', 'Heart Failure',
                  'Ischemic Heart Disease', 'Myocardial Infarction', 'Broken Arm')

# looking only at wrf chem, kriging, and geo-weighted; taking out 'Global Smoke' method
method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Global Smoke', 'GWR Smoke')

# create an empty list to row bind dataframes together
datalist1 <- list()
datalist2 <- list()

# data wrangling ----
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){

# dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- colnames(df_to_loop[3])
  # outcome name
  outcome_name <- outcome_list[i]
  # extract covariates from dataframe
  covariates_df <- df_to_loop[, c(1:16, 27,79:83, 28, 151:157)]

  # extract pm values and divide by 10 and ordered
  #which(colnames(df_to_loop)=="wrf_temp_lag1_zip") # code to find column numbers
  #which(colnames(df_to_loop)=="wrf_temp_zip") # code to find column numbers
  pm_lag_list <- list(df_to_loop[, c(19, 39:43)]/10, df_to_loop[, c(26, 74:78)]/10,
                  df_to_loop[, c(25, 69:73)]/10, df_to_loop[, c(24, 64:68)]/10)  # create 10 unit increases

# need to think if I need a methods section (probably not since this is it)

  # new loop for each element of the lag list 
  for(k in 1:4){

    # empty matrix (12 x 10 matrix)
    point_estimates <- matrix(nrow = 6, ncol = 10, byrow = T)

    colnames(point_estimates) <- c('outcome', 'pm_method', 'lag', 'n', 'n_events', 
                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')

    # fill in the outcome namedataframe before method loop
    point_estimates[, 1] <- outcome_name
    # repeat the sex category 4 times for each pm method
    point_estimates[, 2] <- method_list[k]

    pm_test <- pm_lag_list[[k]]

    # dataframe for analysis creation
    # bind columns back together 
    df_analysis <- cbind(covariates_df, pm_lag_list[[k]]) %>% 
      # limit to emergency or urgent care
      filter(ADM_TYPE == 1 | ADM_TYPE == 2) %>%
      # join in cdc met data
      full_join(daily_zip_met, by = c("ZIPCODE", "date")) 
      # dataframe is already for the entire fire season, so I don't need to subset anymore
#glimpse(df_analysis)

  # loop to run a model for each lag 0-5
    for(j in 0:5){

      # finding locations of columns in dataframe
      which(colnames(df_analysis)=="daily_meanWS")

      # set column number to evaluate for the pm lag method
      pm_lag_colnum <- j + 31
      # set column number of cdc temp lag to adjust for
      temp_lag_colnum <- j + 49
      # set col num for abs temp
      ab_temp_lag_colnum <- j + 37
      # set col num for relative humidity
      rh_lag_colnum <- j + 43
      # precip
      pr_lag_colnum <- j + 55
      # wind speed
      ws_lag_colnum <- j + 61
      # set row number to fill
      row_n <- j+1

      # only run the model if the dataframe has observations
      if(nrow(df_analysis) != 0){
      # conditional logistic regression model
      mod <- clogit(outcome ~ df_analysis[[pm_lag_colnum]] + df_analysis[[temp_lag_colnum]] + 
                    df_analysis[[rh_lag_colnum]] + df_analysis[[pr_lag_colnum]] + 
                    df_analysis[[ws_lag_colnum]] + strata(PATIENTID), df_analysis)

      # populate matrix
      point_estimates[row_n, 3] <- j
      point_estimates[row_n, 4] <- mod$n
      point_estimates[row_n, 5] <- mod$nevent
      # odds ratio
      point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
      # 95% lower bound
      point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
                                        1.96*(summary(mod)$coefficient[1,3])), 3)
      # 95% upper bound
      point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
                                        1.96*(summary(mod)$coefficient[1,3])), 3)
      # standard error
      point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
      # p val
      point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)

      # create else statement that fills matrix with missing so I still have the row
      # in the final dataframe
      } else {point_estimates[row_n, 3] <- method_list[row_n]
              point_estimates[row_n, 4] <- 0
              point_estimates[row_n, 5] <- 0
              point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement

    } # end methods loop

    # save point estimates as a dataframe
    point_est_df <- as_data_frame(point_estimates)

    # combine previous values in dataframe that has all outcome/methods comparisons
    datalist1[[k+1]] <- point_est_df
  } # end sex category loop

  # bind rows of sex category estimates together
 lag_est_df <- bind_rows(datalist1)

 # populate second dataframe list
 datalist2[[i]] <- lag_est_df

} # end of outcome loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist2)

#glimpse(combined_point_est_df)
# subset to just geo-smoke estimates for now to reduce number of rows
wrf_geo_lag <- combined_point_est_df %>% 
  # filter columns to just geo-smk
  filter(pm_method == "GWR Smoke" |
         pm_method == "WRF-Chem Smoke") %>% 
  # subset columns I want to put in to the table
  select(2:3, 5:8) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
         lower95 = ifelse(lower95 == 99, "--", lower95),
         upper95 = ifelse(upper95 == 99, "--", upper95))

  tab <- htmlTable(txtRound(wrf_geo_lag, digits = 3, 1:3), 
           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes by lagged days",
           # row group by outcome
           rgroup = outcome_list,
           n.rgroup = c(rep(12, 12)), # 6 rows for each lag period for each outcome
           # column headers
           header = c("Est Method", "Lagged Days", "Events",
                      "OR&dagger;", "Lower", "Upper"),
           # column spanner
           cgroup = c("", "95% CI"), 
           n.cgroup = c(4, 2),
           padding.rgroup = "&nbsp;&nbsp;",
           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
           align = "lccccc", # column alignment,
           tfoot="&dagger; Adjusted for lagged CDC temperature, windspeed, relative humidity, and precipitation, accounting for subject."
            ) # end table

  print(tab)

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
wrf_geo_lag_plot <- combined_point_est_df %>% 
  # filter columns to just geo-smk
  filter(pm_method == "GWR Smoke" |
         pm_method == "WRF-Chem Smoke") %>% 
  # subset columns I want to put in to the table
  select(1:3, 5:8)

# change characters to numeric and factor  
wrf_geo_lag_plot$outcome <- factor(wrf_geo_lag_plot$outcome,
                              levels = unique(wrf_geo_lag_plot$outcome))

wrf_geo_lag_plot$pm_method <- factor(wrf_geo_lag_plot$pm_method,
                                levels = unique(wrf_geo_lag_plot$pm_method))

wrf_geo_lag_plot$lag <- factor(wrf_geo_lag_plot$lag,
                                levels = unique(wrf_geo_lag_plot$lag))

wrf_geo_lag_plot$n_events <- as.numeric(wrf_geo_lag_plot$n_events)
wrf_geo_lag_plot$odds_ratio <- as.numeric(wrf_geo_lag_plot$odds_ratio)
wrf_geo_lag_plot$lower95 <- as.numeric(wrf_geo_lag_plot$lower95)
wrf_geo_lag_plot$upper95 <- as.numeric(wrf_geo_lag_plot$upper95)

## ggplot
  print_plot <- ggplot(wrf_geo_lag_plot,
                  aes(x = lag, y = odds_ratio, color = pm_method), na.rm = T) +
    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
    # custom color 
    scale_color_manual(name = "Smoke-Estimation Method", 
                       values = c("red", "#32115C"),
                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
    facet_wrap(~outcome, nrow = 3, scales = "free") +
    geom_hline(yintercept = 1, linetype=2) +
    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
    #ylim(0.7, 1.4) +
    xlab('Lag Days') +
    # plot theme
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # strip element
    strip.background = element_rect(colour=NA, fill=NA),
    panel.border = element_rect(fill = NA, color = "black"),
    # facet text size
    strip.text = element_text(size = 10),
    # axis element
    #axis.text.x = element_blank(),
    #axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(angle = 90),
    # legend elements
    legend.position = "bottom")
    #legend.text = element_text(size = 8))

  print(print_plot)

  # save figure
  ggsave("./fig4.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

```

### Overall Summary of Time-Stratified Results (Zip Code-Level)

In some of my other documents, I tried different referent selection schemes for the case-crossover design, specifically the symmetric bi-directional design, where referent periods were matched to the same day of the week. In some cases, this produced similar estimates to the time-stratified results within season and wildfire season presented herein. However, as demonstrated in the Janes 2005 **Statistics in Medicine** paper, the symmetric bi-directional still has a bias in that the \beta is not necissarily unbiased, and can be a slight overestimation (which is evident in these results compared to some of my other results using that design).

### 2017-04-12 r_scripts/data_management/or_smoke_jul_oct_time_stratified_case_crossover_server.R
Delete or_smoke_jul_oct_time_stratified_case_crossover_server.R

@@ -1,176 +1,0 @@
-# ------------------------------------------------------------------------------
-# Title: Oregon Case-crossover dataframe creation (only patients with single obs)
-#        and joining with and zip-level PM2.5 values. 
-# Author: Jingyang Liu
-# Date Created: 04/05/2017     
-# R version: 3.3.3
-# ------------------------------------------------------------------------------
-
-# Notes ----
-# Notes 8/16/16: I updated smoke data with population weighted averages for zip
-# This code also assumed only one visit, so I remove the people with multiple
-# admissions for the outcome
-
-# Notes 10/3/16: I'm retaining the county assignment of the observation so I
-# can join the county population weighted averages as well for comparison
-# and to analyze morbidity similar to Rish
-
-# Libraries ----
-# load libraries
-#library(tidyverse) # tidyverse not on the server
-# if tidyverse is not available, use dplyr and readr
-library(dplyr) 
-library(readr)
-library(lubridate) # working with date
-# parallel computing libraries
-# library(foreach) 
-# library(doParallel)
-
-# Read in smoke data created in loop -------------------------------------------
-
-# read in zipcode level populatoin-weighted pm
-read_path <- paste0('../../../data/data_new/zip_pm_to_merge_with_chars.csv')
-
-zip_smoke <- read_csv(read_path) # 63801 rows
-
-# descriptives of the two smoke datasets
-summary(zip_smoke)
-
-# Zipcode PM2.5 estimates
-# create lag variables that take smoke values from n previous days for zipcodes
-zip_smoke_w_lag <- zip_smoke %>% arrange(ZIPCODE, date) %>%
-  # group by zipcode
-  group_by(ZIPCODE) %>% 
-  # wrf
-  mutate(wrf_f_pm_lag1 = lag(wrf_f_pm, 1, order_by = ZIPCODE), 
-         wrf_f_pm_lag2 = lag(wrf_f_pm, 2, order_by = ZIPCODE),
-         wrf_f_pm_lag3 = lag(wrf_f_pm, 3, order_by = ZIPCODE),
[…]

### 2017-04-13 r_scripts/analysis/or_result_zip_manuscript.Rmd

--- 2017-04-12 r_scripts/analysis/or_zip_result.Rmd
@@
 
-```
-
-```{r, eval = FALSE}
-## 
+## ???
 # read locations of fires
@@
 # dataframe list
-df_list <- list(resp_casecross, asthma_casecross, copd_casecross, 
-  pneum_casecross, acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
+df_list <- list(resp_casecross, asthma_casecross, copd_casecross, pneum_casecross,
+                acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
                 cereb_vas_casecross, hf_casecross, ihd_casecross, mi_casecross, 
@@
 
-outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 
-                  'Pneumonia', 'Acute Bronchitis', 'Cardiovascular Disease',
-                  'Arrhythmia', 'Cerebrovascular Disease', 'Heart Failure',
-                  'Ischemic Heart Disease', 'Myocardial Infarction',
-                  'Broken Arm')
-
-method_list <- c('WRF-Chem Smoke', 'Geo-Weighted Smoke', 'Kriging Smoke')
+outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 'Pneumonia', 
+                  'Acute Bronchitis', 'Cardiovascular Disease', 'Arrhythmia', 
+                  'Cerebrovascular Disease', 'Heart Failure',
+                  'Ischemic Heart Disease', 'Myocardial Infarction', 'Broken Arm')
 
@@
 
-  i = 2
 # dataframe to loop through
@@
   # indication of column
-  outcome <- colnames(df_to_loop[76])
+  outcome <- colnames(df_to_loop[3])
   # outcome name
@@
   # bind columns back together 
-  df_analysis <- df_to_loop %>%
+  df_analysis <- df_to_loop %>% select(c(1:16, 19, 27:28, 151:155)) %>% 
     # only look at outcomes
     filter(outcome == 1) %>%
+    # left in only observations with estimates of smoke
+    filter(!is.na(wrf_smk_pm_zip)) %>% 
+    # limit to emergency or urgent care
+    filter(ADM_TYPE == 1 | ADM_TYPE == 2) %>% 
     # add another row that makes sure there is a person <15 in the dataframe
     # tricking xtabs to produce a 0 cell for the outcome for age <15
-    add_row(outcome = 0, age_ind = 0)
+    add_row(outcome = 0, age_cat = 0)
 
@@
   outcome_n <- xtabs(~ outcome, df_analysis)
-  cross_tab_age <- xtabs(~ outcome + age_ind, df_analysis)
-  cross_tab_sex <- xtabs(~ outcome + sex_ind, df_analysis)
+  cross_tab_age <- xtabs(~ outcome + age_cat, df_analysis)
+  cross_tab_sex <- xtabs(~ outcome + sex_num, df_analysis)
   # empty matrix
@@
 
-outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 
-                  'Pneumonia', 'Acute Bronchitis', 'Cardiovascular Disease',
-                  'Arrhythmia', 'Cerebrovascular Disease', 'Heart Failure',
-                  'Ischemic Heart Disease', 'Myocardial Infarction',
-                  'Broken Arm')
-
-method_list <- c('WRF-Chem Smoke', 'Geo-Weighted Smoke', 'Kriging Smoke')
+outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 'Pneumonia', 
+                  'Acute Bronchitis', 'Cardiovascular Disease', 'Arrhythmia', 
+                  'Cerebrovascular Disease', 'Heart Failure',
+                  'Ischemic Heart Disease', 'Myocardial Infarction', 'Broken Arm')
+
+# looking only at wrf chem, kriging, and geo-weighted; taking out 'Global Smoke' method
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Global Smoke', 'GWR Smoke')
 
@@
   # extract covariates from dataframe
-  covariates_df <- df_to_loop[, c(1:26, 71, 74, 76:85)]
+  covariates_df <- df_to_loop[, c(1:16, 27:28, 151:157)]
 
   # extract pm values and divide by 10 and ordered
-  which(colnames(df_to_loop)=="background_pm_zip") # code to find column numbers, 91
-  pm_estimates_df <- df_to_loop[, c(88, 92:94)]/10  # create 10 unit increases
+  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+  pm_estimates_df <- df_to_loop[, c(19, 26, 25, 24)]/10  # create 10 unit increases
 
@@
     filter(!is.na(wrf_smk_pm_zip)) %>% 
-
     # join in cdc met data
-    # full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
+    full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
+    # limit to emergency or urgent care
+    filter(ADM_TYPE == 1 | ADM_TYPE == 2) %>%
     # the following code makes sure that the counterfactual values retained are 
     # symetric in that number of obs before = number of obs after
-    mutate(obs_diff_admission = (fromdate - date)/7) 
+    mutate(obs_diff_admission = (date - date_admit)/7) 
     # dataframe is already for the entire fire season, so I don't need to subset anymore
@@
   # empty matrix
-  point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)
+  point_estimates <- matrix(nrow = 4, ncol = 9, byrow = T)
 
@@
   # second loop to run a model for each pm estimation method
-    for(j in 39:41){
+    for(j in 26:29){
 
@@
       # conditional logistic regression model
-      mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+      mod <- clogit(outcome ~ df_analysis[[j]] + daily_meanRH + daily_meanT + daily_meanWS + 
+                      daily_meanPrcp + strata(PATIENTID), df_analysis)
 
@@
       # populate matrix
-      row_n <- j-38
+      row_n <- j-25
 
@@
 # subset columns I want to put in to the table
-table_df <- combined_point_est_df %>% select(2, 3:7)
+table_df <- combined_point_est_df %>% select(2, 3:7) %>% 
+  # exclude global smoke
+  filter(pm_method != "Global Smoke")
 
@@
 # subset out global method
-combined_point_est_df <- combined_point_est_df
+combined_point_est_df <- combined_point_est_df %>% 
+  filter(pm_method != "Global Smoke")
 
@@
   # save figure
-  ggsave("./fig4.pdf", plot = print_plot, 
+  ggsave("./fig3.pdf", plot = print_plot, 
        width = 12, height = 8, units = "in")
@@
 resp_point_est_df <- combined_point_est_df %>% 
-
+  filter(pm_method != "Global Smoke") %>% 
   filter(outcome == "All Respiratory" |
@@
 cvd_point_est_df <- combined_point_est_df %>% 
+  filter(pm_method != "Global Smoke") %>% 
   filter(outcome == "Cardiovascular Disease" |

### 2017-04-13 r_scripts/parallel_presentation.Rmd

@@
 
-```{r parallel set, eval = FALSE}
+```{r parallel, eval = FALSE}
 # Setup for parallel computing before for loop ---------------------------------

### 2017-04-17 r_scripts/analysis/or_result_zip_smoke.Rmd
Create or_result_zip_smoke.Rmd

--- 2017-04-12 r_scripts/analysis/or_zip_result.Rmd
@@
 ---
-title: "Oregon: Zip-Level Wildfire Smoke and Health Outcomes for Manuscript"
+title: "Oregon: Zip-Level Wildfire Smoke and Health Outcomes (open in wildfire, not script)"
 author: "Jingyang Liu"
@@
 
-```{r library calls, include=FALSE}
-# loading libraries used
-library(tidyverse) # import tidyverse
-library(survival) # for conditional logistic regression
-library(htmlTable) # table
-library(lme4) # random-effects model
-library(broom) # broom for tidy data from stats models
-library(cowplot) # for multi plot image
-
-# spatial packages
-library(ggmap) # map package
-library(rgdal) # set coord ref system for shape files
-library(rgeos) # need it when converting shape to dataframe
-library(maptools) # convert shape to dataframe
-
-```
-
 ## Overview
@@
 This particular markdown document focuses on the zip code-level analysis.
-Note there is a county-level analysis that will be paired with CDC mortality data as well.
 Herein, I compare different PM~2.5~ estimation methods described in further detail below.
 
-Note, epidemiological methods that Sheryl and I think can improve our estimation and reduce some bias around the point estimates will not be the focus of the manuscript.
+```{r include=FALSE}
+library(tidyverse)
+library(data.table)
+library(survival)
+library(htmlTable)
+library(readr)
+```
 
-```{r data import, include = F, echo = FALSE}
-# Set working directory and read in files --------------------------------------
-# relative path
-# set working directory if you want to run seperate chunks
-# setwd("./r_scripts")
-getwd()
+```{r data import, include = F, eval = FALSE}
 
-path <- paste0("../data_new/health_data")
-# Infile case-crossover dataframes ---------------------------------------------
-# Dataframes made in 'chars_3month_binary_smoke_may2016 script
-# resp exacerbations
+### Limit to only ER visits
+## pos = 23
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+
+N <- NA
+n <- 1
+
+for (i in var_list){
+  read_path <- paste('../data_new/update/ER/or', i, 'may_to_sep_time_strat_casecross.csv', sep = "_")
+  or_disease <- read_csv(read_path)
+
+  or_disease <- or_disease %>% 
+    filter(pos=="23") %>%
+    # filter the unknown gender
+    filter(gender!= "U")
+
+  # Create a permanent case-cross over dataset
+  file_name <- paste('../data_new/update/ER/or', i, 'may_to_sep_time_strat_casecross_er.csv', sep = "_")
+
+  # write permanent dataset
+  write_csv(or_disease, paste0(file_name))
+  N[n] = length(unique(or_disease$personkey))
+  n = n+1
+}
+
+num_person <- data.frame(rbind(var_list, N))
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+## Analysis
+path <- paste0("../data_new/update/ER")
+
 resp_casecross <- read_csv(paste(path, 
-                        "respiratory_may_to_sep_time_strat_casecross.csv", 
+                                 "or_respiratory_may_to_sep_time_strat_casecross_er.csv", 
                                  sep = "/"))
@@
 asthma_casecross <- read_csv(paste(path, 
-                          "asthma_may_to_sep_time_strat_casecross.csv", 
+                                   "or_asthma_may_to_sep_time_strat_casecross_er.csv", 
                                    sep = "/"))
@@
 copd_casecross <- read_csv(paste(path, 
-                        "copd_may_to_sep_time_strat_casecross.csv", 
+                                 "or_copd_may_to_sep_time_strat_casecross_er.csv", 
                                  sep = "/"))
@@
 pneum_casecross <- read_csv(paste(path, 
-                        "copd_may_to_sep_time_strat_casecross.csv",
+                                  "or_pneumonia_may_to_sep_time_strat_casecross_er.csv",
                                   sep="/"))
@@
 acute_bronch_casecross <- read_csv(paste(path, 
-                                "acute_bronch_may_to_sep_time_strat_casecross.csv",
+                                         "or_acute_bronch_may_to_sep_time_strat_casecross_er.csv",
                                          sep = "/"))
@@
 cvd_casecross <- read_csv(paste(path, 
-                      "cvd_may_to_sep_time_strat_casecross.csv",
+                                "or_cvd_may_to_sep_time_strat_casecross_er.csv",
                                 sep="/"))
@@
 arrhythmia_casecross <- read_csv(paste(path, 
-                              "arrhythmia_may_to_sep_time_strat_casecross.csv",
+                                       "or_arrhythmia_may_to_sep_time_strat_casecross_er.csv",
                                        sep="/"))
@@
 cereb_vas_casecross <- read_csv(paste(path, 
-                              "cerbrovas_dis_may_to_sep_time_strat_casecross.csv",
+                                      "or_cerbrovas_dis_may_to_sep_time_strat_casecross_er.csv",
                                       sep="/"))
@@
 hf_casecross <- read_csv(paste(path, 
-                      "heart_failure_may_to_sep_time_strat_casecross.csv", 
+                               "or_heart_failure_may_to_sep_time_strat_casecross_er.csv", 
                                sep="/"))
@@
 ihd_casecross <- read_csv(paste(path, 
-                       "isch_heart_dis_may_to_sep_time_strat_casecross.csv",
+                                "or_isch_heart_dis_may_to_sep_time_strat_casecross_er.csv",
                                 sep="/"))
@@
 mi_casecross <- read_csv(paste(path, 
-                      "myocardial_infarc_may_to_sep_time_strat_casecross.csv", sep="/"))
+                               "or_myocardial_infarc_may_to_sep_time_strat_casecross_er.csv", sep="/"))
 
@@
 broken_arm_casecross <- read_csv(paste(path, 
-                              "broken_arm_may_to_sep_time_strat_casecross.csv",
+                                       "or_broken_arm_may_to_sep_time_strat_casecross_er.csv",
                                        sep="/"))
@@
 smoke_path <- paste0("../data/Oregon/")
-
 zip_pm <- read_csv(paste0(smoke_path, "zip_pm_to_merge_with_chars.csv"))
-
-```
-
-```{r, eval = FALSE}
-## 
-# read locations of fires
-fire_loc <- read_csv(paste0(smoke_path, "washington_fires_locations201209.csv"))
-
-# import the zip-level met data files Rish sent
-zip_met <- read_csv(paste0(smoke_path, "wa_zip_exp_0710_2012_ryan.csv")) %>% 
-  # define date variable; original character looked like: 01JUL2012
-  mutate(date = as.Date(date, "%d%B%Y"))
-
-# note 12/4/16: read_csv is pretty slow on pc, fread doesn't import date variable 
-# that well. Solution may be to use fread but convert to date to use later
-
 ```
@@
 We further limited our analyses to claims that were coded as emergency or urgent visits to eliminate bias from patients going in for elective/planned procedures.
+However, we found the filtered population is much larger than the population we chose in Washington, so we choose only the emegency room (hospital) visits.
 For patients who have a index time closer to May 1, 2013 or closer to September 30, 2013, their referent observations before or after these dates will be excluded as I will not be able to assign estimates of PM~2.5~ to these referent observations.
@@
 
-As for **exposure methods**, there are four main estimation methods for PM~2.5~: WRF-Chem Smoke (which substracts the WRF-Chem no fire emission from WRF-chem).
+As for **exposure methods**, there are three main estimation methods for PM~2.5~: WRF-Chem Smoke (which substracts the WRF-Chem no fire emission from WRF-chem).
 For the WRF-Chem variable with the 'smoke' designator, this is WRF-Chem - WRF-Chem no fire.
-For Global Regression, Geo-Weighted Regression, and Kriging with 'smk' designator, I subtracted off the 'Background' estimates of smoke, which I believe are the monthly averages of PM~2.5~ for a given grid.
+For Geo-Weighted Regression, and Kriging with 'smk' designator, I subtracted off the 'Background' estimates of smoke, which I believe are the monthly averages of PM~2.5~ for a given grid.
+However, the 'Background' is the mean of total period from May 1st to Sep 30th 2013 so I have to assign the mean for every day.
 
@@
 I've run a couple different scenarios for selecting referent periods, each of which has some paper that may justify the benefit of the design, relative to other methods.
-
-### Time-Series of Smoke PM~2.5~ Ranges by Ecoregion
-
-Figures below are for each smoke PM~2.5~ estimation method.
-The figures are the daily range (min and max) for each zip code in a specific region of the state that local offices of the Washington Department of Ecology has jurisdiction over.
-The figure first starts out by showing the map of the state, divided in to eco regions used by the state for organization purposes, and then presents the time-series ranges for WRF-Chem, krig, geo-weighted estimation methods.
-
-```{r smk time series range by eco region, echo = F, results='asis', message=F}
-
-## ???
-# creation of map and identifying zipcodes in ecoregions----
-# import the regional shapefile downloaded from state website
-shp_dir <- paste0('../../data/shape_files/WAECY__Ecology_Regions')
-shp_dir <- paste0('./data/shape_files/WAECY__Ecology_Regions')
-
-eco_region <- readOGR(dsn = shp_dir, layer = 'WAECY__Ecology_Regions')
-# set coord ref system to WGS84 to match projections of ggmap
-eco_region <- spTransform(eco_region, "+proj=longlat +datum=WGS84")
-
-# summary(eco_region) # projection is GRS80, may be off
-
-# ggmap and rgdal needed
-# Set base map boundary layer
-bbox <- c(-125, 45, -116.1, 49.5)
-
-# call black and white map
-washington_map <- get_map(location= bbox, source = 'stamen',
-                          maptype = 'toner', crop = T)
-
-# call google terrain map
-#wash_map_terrain <- get_map(location = "washington state", source = 'google',
-#                             maptype = 'terrain', crop = F)
-
-# fortify washington shape data in to dataframe (need maptools package)
-eco_df <- tidy(eco_region, region = "ECY_REGION")
-
-# summary(wash_zip_df)
-
-# creation of eco region map (may change colors)
-eco_region_map <- ggmap(washington_map) + 
-  geom_polygon(data = eco_region, aes(x=long, y=lat, group = group, 
-    fill = group), color = "black", alpha = 0.8) +
-  scale_fill_manual(name = "Ecology Regions",
-                    values = c("#2c7bb6","#abd9e9","#d7191c","#fdae61"), 
-                    labels = c("Southwest", "Northwest", "Central", "Eastern"),
-                    guide = guide_legend(title.position = "top", order = 1)) +
-  geom_point(data = fire_loc, aes(x = Longitude, y = Latitude, shape = "Fire Locations"), 
-             color = "black", fill = "red") +
-  scale_shape_manual(name = "", values = 24, 
-                     guide = guide_legend(title.position = "top", order = 2)) +
-  ggtitle("Washington State Ecology \n Regional Office Jurisdiction") +
-  xlab("Longitude") +
-  ylab("Latitude") +
-  theme(#legend.position = "bottom",
-        #legend.box = "horizontal"
-    )
-
-print(eco_region_map)
-
-# import washington zip shapefile ----
-#zip_path <- paste0('../../data/shape_files/wash_zip_2012_shapefile')
-zip_path <- paste0('./data/shape_files/wash_zip_2012_shapefile')
-
-wash_zip_map <- readOGR(dsn = zip_path, layer = "wash_zip_2012_shapefile")
-
-# change projection to match eco region
-wash_zip_map <- spTransform(wash_zip_map, "+proj=longlat +datum=WGS84")
-
-#summary(eco_region)
-#summary(wash_zip_map)
-
-# check overlay visually 
-#plot(eco_region, lwd = 5)
-#plot(wash_zip_map, add = T)
-
-# find the eco regions that are over a zipcode
-zip_eco_region_overlap <- over(wash_zip_map, eco_region)
-# output zip IDs from zip shapefile
-ZIPCODE <- as.character(wash_zip_map@data$ZCTA5CE10)
-# join eco region ids with zip (can be used to indicate eco region)
-zip_eco_region <- cbind(ZIPCODE, zip_eco_region_overlap) %>% 
-  mutate(eco_region = ifelse(ECY_REGION == "CRO", "Central",
-                      ifelse(ECY_REGION == "ERO", "Eastern",
-                      ifelse(ECY_REGION == "NWRO", "Northwestern",
-                      ifelse(ECY_REGION == "SWRO", "Southwestern", NA))))) %>%  
-  select(-OBJECTID, -ECY_REGION)
-
-# range of PM values for each day in each eco region for each method ----
-# find the min, median, and max values for each date
-#summary(filter(zip_pm, date == "2012-07-9")) # quick sum of first zip obs 
-# zip_pm zipcode needs to be character to join with region
-zip_pm$ZIPCODE <- as.factor(zip_pm$ZIPCODE)
-
-# glimpse(zip_pm)
-# these joins should really be done in a loop, but...
-# each column would need to be specified for each pm val, 
-# which I'd have to think about more than just each one manually
-# wrf smoke
-wrf_range <- zip_pm %>% 
-  full_join(zip_eco_region, by = "ZIPCODE") %>%  
-  group_by(date, eco_region) %>% 
-  summarise(min = min(wrf_smk_pm, na.rm = T), 
-            med = median(wrf_smk_pm, na.rm = T), 
-            max = max(wrf_smk_pm, na.rm = T)) %>% 
-  # variable for small multiples
-  mutate(est_method = "wrf_smk")
-
-# krig smoke
-krig_range <- zip_pm %>% 
-  full_join(zip_eco_region, by = "ZIPCODE") %>%  
-  group_by(date, eco_region) %>% 
-  summarise(min = min(krig_smk_pm, na.rm = T), 
-            med = median(krig_smk_pm, na.rm = T), 
-            max = max(krig_smk_pm, na.rm = T)) %>% 
-  mutate(est_method = "krig_smk")
-
-# geo smoke
-geo_range <- zip_pm %>% 
-  full_join(zip_eco_region, by = "ZIPCODE") %>%  
-  group_by(date, eco_region) %>% 
-  summarise(min = min(geo_smk_pm, na.rm = T), 
-            med = median(geo_smk_pm, na.rm = T), 
-            max = max(geo_smk_pm, na.rm = T)) %>% 
-  mutate(est_method = "geo_smk")
-
-# loop plots though
-
-file_name <- c('wrf', 'krig', 'geo')
-title_name <- c('WRF-Chem Smoke', 'Kriging Smoke', 'GWR Smoke')
-ts_list <- list(wrf_range, krig_range, geo_range)
-
-for(i in 1:3){
-
-df_range <- ts_list[[i]]  
-gg_title <- title_name[i]
-
-# create plot for method range
-plot_print <- ggplot(df_range, aes(x = date)) +
-  geom_ribbon(aes(ymin = min, ymax = max, group = eco_region, 
-                  fill=eco_region), alpha = 0.7) +
-  scale_fill_manual(name = "Washington State Ecology \n Regional Office Jurisdiction",
-                    values = c("#d7191c", "#fdae61", "#abd9e9", "#2c7bb6"),
-                    labels = c("Central", "Eastern","Northwest", "Southwest"),
-                    guide = F) +
-  ylim(0,500) +
-  #geom_line(aes(y=med, group=eco_region, color=eco_region)) +
-  ggtitle(gg_title) +
-  ylab(expression('Smoke PM2.5 µg/m'^3)) +
-  xlab('Date') + 
-  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    axis.text.x = element_text(size = 12),
-    axis.text.y = element_text(size = 12),
-    axis.title.y = element_text(angle = 90, size = 12),
-    axis.title.x = element_text(size = 14), 
-    strip.background = element_rect(fill = 'white'))
-
-# file name
-name <- file_name[i]
-
-assign(paste0(name,"_plot"), plot_print)
-
-}
-
-# Layout needs to be edited and I would like to change colors
-# but this should be good enough for a review by CDC
-map_timeseries <- ggdraw() +
-  draw_plot(eco_region_map, x = 0, y = 0.5, width = 0.5, height = 0.5) +
-  draw_plot(wrf_plot, x = 0.5, y = 0.5, width = 0.5, height = 0.5) +
-  draw_plot(krig_plot, x = 0, y = 0, width = 0.5, height = 0.5) +
-  draw_plot(geo_plot, x = 0.5, y = 0, width = 0.5, height = 0.5)
-
-plot(map_timeseries)
-# save 6 inch by 4 inch
-ggsave("./fig1.pdf", plot = map_timeseries, 
-       width = 12, height = 8, units = "in")
-
-```
-
-## Days of Smoke In Washington
-
-Here is a figure of the days from May 1st to September 30 2013 where PM~2.5~ exceeded >10.
-
-```{r smoke map, message = F, warning = F, echo = F, results='asis'}
-
-# fire location has a weird date format, but I think is only september
-
-# zip code count of smoke ------------------------------------------------------
-zip_smoke_days <- zip_pm %>% 
-  # create binary indicators of days with smoke
-  mutate(wrf_smoke = if_else(wrf_smk_pm > 0, 1, 0),
-         wrf_smoke5 = if_else(wrf_smk_pm > 5, 1, 0),
-         wrf_smoke10 = if_else(wrf_smk_pm > 10, 1, 0),
-         geo_smoke = if_else(geo_smk_pm > 0, 1, 0),
-         geo_smoke5 = if_else(geo_smk_pm > 5, 1, 0),
-         geo_smoke10 = if_else(geo_smk_pm > 10, 1, 0)) %>% 
-  group_by(ZIPCODE) %>% 
-  summarise(wrf_smk_days = sum(wrf_smoke), wrf_smk5_days = sum(wrf_smoke5), 
-            wrf_smk10_days = sum(wrf_smoke10), geo_smk_days = sum(geo_smoke), 
-            geo_smk_days = sum(geo_smoke), geo_smk5_days = sum(geo_smoke5),
-            geo_smk10_days = sum(geo_smoke10)) %>% 
-  # change zipcode to character
-  mutate(ZIPCODE = as.character(ZIPCODE))
-
-# check smoke days
-#summary(zip_smoke_days)
-
-# tidy from broom package now works
-wash_zip_df <- tidy(wash_zip_map, region = "ZCTA5CE10") %>% 
-  # merge in smoke days
-  full_join(zip_smoke_days, by = c("id" = "ZIPCODE"))
-
-# map
-print_map <- ggmap(washington_map) + 
-  geom_polygon(data = wash_zip_df, 
-    aes(x=long, y=lat, group = group, fill = geo_smk10_days), alpha = 0.7) +
-  scale_fill_gradient(expression("Number of \nSmoke Days"),
-    low = '#2c7bb6', high = '#d7191c') + 
-  geom_point(data = fire_loc, aes(x = Longitude, y = Latitude, shape = "Fire Locations"), 
-             colour = "black", fill = "red") +
-  scale_shape_manual(name = "", values = 24) + 
-  xlab("Longitude") +
-  ylab("Latitude") +
-  # control legend/guide lable order
-  guides(fill = guide_colourbar(title.position = "top", order = 1),
-         shape = guide_legend(title.position = "top", order = 2))
-
-print(print_map)
-
-ggsave("./fig2.pdf", plot = print_map, 
-       width = 12, height = 8, units = "in")
-
-```
 
@@
 
-### Background on Time-Stratified Case-Crossover Design
-
-The original time-stratified case-crossover design suggested picking referent periods based on any potential time-varying confounding, *a priori*.
-For example, if day of the week and the month of the exposure sequence is of concern, one might pick referents in the same day of the week and the same day of the month.
-The papers that discuss and support the use of a time-stratified design are the Janes et al.
-2005 **Epidemiology paper**, *Case-Crossover Analyses of Air Pollution Exposure Data; Referent Selection Strategies and Their Implication for Bias*.
-
-The authors state that with the time-stratified design, the *a priori* selection of the referent period on the same day of the week within the same month has the following benefits:
-
-1.
-Localizable - Meaning that the likelihood of index times conditional on the referent windows contains information about \beta, which I think means the relative difference in the exposed vs unexposed, conditioned on the exposure (air pollution in their example), is unbiased.
-
-2. Non-ignroable - Meaning that time-varying factors can be ignored when using a conditional logistic regression; controlled through design (matching on day of week within the same season).
-
-More referent periods is better for statistical efficiency, but there is a tradeoff, where additional referent periods may introduce what they call 'overlap bias', which they mathmatically define in Janes 2005 **Statistics in Medicine** paper, *Overlap Bias in the Case-Crossover Design, with Applications to Air Pollution Exposures*.
-Basically this bias occurs when the exposure assigned in the referent period is conditional on the strategy used to define the referent period.
-
-As Janes et al. state in their 2005 **Statistics in Medicine** paper:
-
-  > *Two of the main strengths of the case-crossover design in the air pollution context are
-that it controls fixed confounders and allows for control over time-dependent confounders
-by design. Choosing referents that are restricted to the same day of the week and season
-as the index time will control for these effects. Though we have not examined the degree
-to which various referent selection strategies control bias due to confounding, several points
-are clear. Referents that are closer to the index time will achieve greater control of seasonal
-confounding. However, sampling referents closer to the index time will also result in a loss
-of power due to less heterogeneity in the exposure series, and a smaller number of referents
-will also result in lower power.*
-
-I will use both Janes 2005 papers in **Epidemiology** and **Statistics in Medicine** to justify the use of the time-stratified design.
-
-This design has an advantage over other case-crossover designs I've explored in that you can use more referent times, and as the exposure at these times are averaged for each subject over a set time period, this balances out some of the more extreme values of PM~2.5~ that may persist over a couple weeks due to wildfire smoke.
-
 ### Time-Stratified Case-Crossover (referents same day of week within fire season July - October)
@@
 
-```{r data wrangle with zip met vars, warning = F, echo = F, message = F}
+```{r echo = F, warning = F, results='asis'}
 
-# create dataset that can be merged to case-cross over dataframes
-daily_zip_met <- zip_met %>%
-  # limit to new mean met variables (getting rid of smoke variables I already have)
-  select(1:7) %>% 
-  rename(ZIPCODE = ZIP) %>% arrange(ZIPCODE, date) %>%
-  group_by(ZIPCODE) %>% 
-  # create lag variables
-  mutate(
-    # absolute temp
-    meanAT_lag1 = lag(daily_meanAT, 1, order_by = ZIPCODE), 
-    meanAT_lag2 = lag(daily_meanAT, 2, order_by = ZIPCODE),
-    meanAT_lag3 = lag(daily_meanAT, 3, order_by = ZIPCODE),
-    meanAT_lag4 = lag(daily_meanAT, 4, order_by = ZIPCODE),
-    meanAT_lag5 = lag(daily_meanAT, 5, order_by = ZIPCODE),
-    # relative humidity
-    meanRH_lag1 = lag(daily_meanRH, 1, order_by = ZIPCODE),
-    meanRH_lag2 = lag(daily_meanRH, 2, order_by = ZIPCODE),
-    meanRH_lag3 = lag(daily_meanRH, 3, order_by = ZIPCODE),
-    meanRH_lag4 = lag(daily_meanRH, 4, order_by = ZIPCODE),
-    meanRH_lag5 = lag(daily_meanRH, 5, order_by = ZIPCODE),
-    # temp
-    meanT_lag1 = lag(daily_meanT, 1, order_by = ZIPCODE),
-    meanT_lag2 = lag(daily_meanT, 2, order_by = ZIPCODE),
-    meanT_lag3 = lag(daily_meanT, 3, order_by = ZIPCODE),
-    meanT_lag4 = lag(daily_meanT, 4, order_by = ZIPCODE),
-    meanT_lag5 = lag(daily_meanT, 5, order_by = ZIPCODE),
-    # precip
-    meanPrcp_lag1 = lag(daily_meanPrcp, 1, order_by = ZIPCODE),
-    meanPrcp_lag2 = lag(daily_meanPrcp, 2, order_by = ZIPCODE),
-    meanPrcp_lag3 = lag(daily_meanPrcp, 3, order_by = ZIPCODE),
-    meanPrcp_lag4 = lag(daily_meanPrcp, 4, order_by = ZIPCODE),
-    meanPrcp_lag5 = lag(daily_meanPrcp, 5, order_by = ZIPCODE),
-    # wind speed
-    meanWS_lag1 = lag(daily_meanWS, 1, order_by = ZIPCODE),
-    meanWS_lag2 = lag(daily_meanWS, 2, order_by = ZIPCODE),
-    meanWS_lag3 = lag(daily_meanWS, 3, order_by = ZIPCODE),
-    meanWS_lag4 = lag(daily_meanWS, 4, order_by = ZIPCODE),
-    meanWS_lag5 = lag(daily_meanWS, 5, order_by = ZIPCODE)) %>% 
-  # sort so all like variables are together
-  select(1:3, 8:12, 4, 13:17, 5, 18:22, 6, 23:27, 7, 28:32)
-
-#which(colnames(daily_zip_met)=="meanAT_lag5")
-#glimpse(daily_zip_met)
-
-```
-
-```{r time strat within fire season adj for cdc met, warning = F, echo = F, results='asis'} 
-# note above, results = 'asis' needed to print htmlTables
-
-# I removed copd exacerbation and rheuamtoid arthritis
-
-# dataframe list
 df_list <- list(resp_casecross, asthma_casecross, copd_casecross, 
@@
 
-method_list <- c('WRF-Chem Smoke', 'Geo-Weighted Smoke', 'Kriging Smoke')
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
 
@@
   which(colnames(df_to_loop)=="background_pm_zip") # code to find column numbers, 91
-  pm_estimates_df <- df_to_loop[, c(88, 92:94)]/10  # create 10 unit increases
+  pm_estimates_df <- df_to_loop[, c(88, 93, 92, 94)]/10  # create 10 unit increases
 
@@
 # save figure
-  ggsave("./fig4.pdf", plot = print_plot, 
+ggsave("./fig_only_er.pdf", plot = print_plot, 
        width = 12, height = 8, units = "in")
-
-# figures for research day
-  # plot of only respiratory outcomes
-
-  xtabs(~ combined_point_est_df$outcome)
-# subset out global method and respiratory outcomes ----
-resp_point_est_df <- combined_point_est_df %>%
-
-  filter(outcome == "All Respiratory" |
-         outcome == "Asthma" |
-         outcome == "COPD" |
-         outcome == "Pneumonia" |
-         outcome == "Acute Bronchitis" |
-         outcome == "Broken Arm")
-
-## ggplot
-  resp_plot <- ggplot(resp_point_est_df, 
-                      aes(x = pm_method, y = odds_ratio, colour = pm_method)) +
-    # custom color 
-    scale_color_manual(name = "Smoke-Estimation Method", 
-                       values = c("red", "blue", "#32115C"),
-                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
-    geom_point() + #geom_text(vjust = 0, nudge_x = 0.3) +
-    geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
-    facet_wrap(~outcome, nrow = 2) +
-    geom_hline(yintercept = 1, linetype=2) +
-#    ggtitle('Association Between PM2.5 \n from Wildfire Smoke on Hospitalizations') +
-    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
-    #xlab('Smoke Estimation Method') +
-    # plot theme
-    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    # strip element
-    strip.background = element_rect(colour=NA, fill=NA),
-    panel.border = element_rect(fill = NA, color = "black"),
-    # facet text size
-    #strip.text = element_text(size = 8),
-    # axis element
-    axis.text.x = element_blank(),
-    axis.title.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.title.y = element_text(angle = 90),
-    # legend elements
-    legend.position = "bottom")
-    #legend.text = element_text(size = 8))
-
-  print(resp_plot)
-
-# subset out global method and respiratory outcomes ----
-cvd_point_est_df <- combined_point_est_df %>% 
-   filter(outcome == "Cardiovascular Disease" |
-         outcome == "Arrhythmia" |
-         outcome == "Cerebrovascular Disease" |
-         outcome == "Heart Failure" |
-         outcome == "Ischemic Heart Disease" |
-         outcome == "Myocardial Infarction")
-
-## ggplot
-  cvd_plot <- ggplot(cvd_point_est_df, 
-                      aes(x = pm_method, y = odds_ratio, colour = pm_method)) +
-    # custom color 
-    scale_color_manual(name = "Smoke-Estimation Method", 
-                       values = c("red", "blue", "#32115C"),
-                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
-    geom_point() + #geom_text(vjust = 0, nudge_x = 0.3) +
-    geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
-    facet_wrap(~outcome, nrow = 2) +
-    geom_hline(yintercept = 1, linetype=2) +
-#    ggtitle('Association Between PM2.5 \n from Wildfire Smoke on Hospitalizations') +
-    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
-    #xlab('Smoke Estimation Method') +
-    # plot theme
-    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    # strip element
-    strip.background = element_rect(colour=NA, fill=NA),
-    panel.border = element_rect(fill = NA, color = "black"),
-    # facet text size
-    #strip.text = element_text(size = 8),
-    # axis element
-    axis.text.x = element_blank(),
-    axis.title.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.title.y = element_text(angle = 90),
-    # legend elements
-    legend.position = "bottom")
-    #legend.text = element_text(size = 8))
-
-  print(cvd_plot)
-
 ```
@@
 I believe this is the most justifiable design approach that would average any extreme values of PM~2.5~ across the wildfire season for the vector of referent values for a particular subject.
-
-When adjusting for the CDC met variables, the results are slightly attenuated when compared to my adjustments with just WRF-Temp, but my overall conclusions do not change much.
-
-Here is the same table, but within season instead of fire season and adjusting for CDC met variables.
-Rish is worried about time-variant confounding within a subject that may bias the overall results.
-This might include seasonal differences in baseline risk of asthma.
-
-### Time-Stratified by Age Categories
-
-Following table and figure look at some outcomes stratified by age category.
-Note that kids <=15 likely won't have many outcomes like COPD or MI.
-I have restricted the dataset to just the geo smoke method to reduce the number of rows/points on the tables and figures to make it easier to read for now.
-Code can easily be changed to have all PM~2.5~ methods.
-
-For some outcomes, there may be some evidence of effect modificaiton, like COPD, where >65 age group is at significant risk.
-However, the point estimates are relatively similar for the 15 to 65 strata, just not significant.
-Some error bars are not shown as they are too wide; usually <15 age category where there are only a handful of outcomes.
-
-```{r time stratified age strata cdc met, warning =F, echo = F, results='asis'} 
-# note above, results = 'asis' needed to print htmlTables
-
-# For age categories, I want a loop for each age category. 
-# Presenting just WRF and GWR
-
-# dataframe list
-df_list <- list(resp_casecross, asthma_casecross, copd_casecross, pneum_casecross,
-                acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
-                cereb_vas_casecross, hf_casecross, ihd_casecross, mi_casecross, 
-                broken_arm_casecross)
-
-outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 'Pneumonia', 
-                  'Acute Bronchitis', 'Cardiovascular Disease', 'Arrhythmia', 
-                  'Cerebrovascular Disease', 'Heart Failure',
-                  'Ischemic Heart Disease', 'Myocardial Infarction', 'Broken Arm')
-
-# looking only at wrf chem, kriging, and geo-weighted; taking out 'Global Smoke' method
-method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Global Smoke', 'GWR Smoke')
-
-# age category list
-age_cat_list <- c(0,1,2)
-
-# create an empty list to row bind dataframes together
-datalist1 <- list()
-datalist2 <- list()
-
-# data wrangling ----
-# Producing conditional logit model estimates loop 
-for(i in 1:length(df_list)){
-
-# dataframe to loop through
-  df_to_loop <- data.frame(df_list[i])
-  # indication of column
-  outcome <- colnames(df_to_loop[3])
-  # outcome name
-  outcome_name <- outcome_list[i]
-
-  # extract covariates from dataframe
-  covariates_df <- df_to_loop[, c(1:16, 27:28, 151:157)]
-
-  # extract pm values and divide by 10 and ordered
-  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
-  pm_estimates_df <- df_to_loop[, c(19, 26, 25, 24)]/10  # create 10 unit increases
-
-  # new loop for age categories
-  for(k in 0:2){
-
-    # empty matrix (12 x 10 matrix)
-    point_estimates <- matrix(nrow = 4, ncol = 10, byrow = T)
-
-    colnames(point_estimates) <- c('outcome', 'age_cat', 'pm_method', 'n', 'n_events', 
-                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
-
-    # fill in the outcome namedataframe before method loop
-    point_estimates[, 1] <- outcome_name
-    # repeat the age category 4 times for each pm method
-    point_estimates[, 2] <- k
-
-    # dataframe for analysis creation
-    # bind columns back together 
-    df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
-      # remove missing pm values
-      filter(!is.na(wrf_smk_pm_zip)) %>% 
-      # join in cdc met data
-      full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
-      # limit to emergency or urgent care
-      filter(ADM_TYPE == 1 | ADM_TYPE == 2) %>%
-      # limit to specific age category
-      filter(age_cat == k) %>% 
-      # the following code makes sure that the counterfactual values retained are 
-      # symetric in that number of obs before = number of obs after
-      mutate(obs_diff_admission = (date - date_admit)/7) 
-      # dataframe is already for the entire fire season, so I don't need to subset anymore
-
-  # second loop to run a model for each pm estimation method
-    for(j in 26:29){
-
-      # variable to model 
-      var_name <- colnames(df_analysis[j])
-
-      # set row number to fill
-      row_n <- j-25
-
-      # only run the model if the dataframe has observations
-      if(nrow(df_analysis) != 0){
-      # conditional logistic regression model
-      mod <- clogit(outcome ~ df_analysis[[j]] + daily_meanRH + daily_meanT + daily_meanWS + 
-                    daily_meanPrcp + strata(PATIENTID), df_analysis)
-
-      # populate matrix
-      point_estimates[row_n, 3] <- method_list[row_n]
-      point_estimates[row_n, 4] <- mod$n
-      point_estimates[row_n, 5] <- mod$nevent
-      # odds ratio
-      point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
-
-      # 95% lower bound
-      point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
-                                        1.96*(summary(mod)$coefficient[1,3])), 3)
-      # 95% upper bound
-      point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
-                                        1.96*(summary(mod)$coefficient[1,3])), 3)
-      # standard error
-      point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
-      # p val
-      point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
-
-      # create else statement that fills matrix with missing so I still have the row
-      # in the final dataframe
-      } else {point_estimates[row_n, 3] <- method_list[row_n]
-              point_estimates[row_n, 4] <- 0
-              point_estimates[row_n, 5] <- 0
-              point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
-
-    } # end methods loop
-
-    # save point estimates as a dataframe
-    point_est_df <- as_data_frame(point_estimates)
-
-    # combine previous values in dataframe that has all outcome/methods comparisons
-    datalist1[[k+1]] <- point_est_df
-  } # end age category loop
-
-  # bind rows of age category estimates together
- age_est_df <- bind_rows(datalist1)
-
- # populate second dataframe list
- datalist2[[i]] <- age_est_df
-
-} # end of outcome loop
-
-# combine each outcome dataframe itteration in to a big dataset
-combined_point_est_df <- bind_rows(datalist2)
-
-wrf_geo_age <- combined_point_est_df %>% 
-  # filter columns to just geo-smk
-  filter(pm_method == "GWR Smoke" |
-         pm_method == "WRF-Chem Smoke") %>% 
-  mutate(age_cat2 = ifelse(age_cat == 0, "Less than 15", 
-                    ifelse(age_cat == 1, "15-65",
-                    ifelse(age_cat == 2, "Greater than 65", NA)))) %>% 
-  # subset columns I want to put in to the table
-  select(3, 11, 5:8) %>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
-         lower95 = ifelse(lower95 == 99, "--", lower95),
-         upper95 = ifelse(upper95 == 99, "--", upper95))
-
-  tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
-           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by age",
-           # row group by outcome
-           rgroup = outcome_list,
-           n.rgroup = c(rep(6, 12)), # 6 rows for each age cat for each outcome
-           # column headers
-           header = c("Est Method", "Age Group", "Events",
-                      "OR&dagger;", "Lower", "Upper"),
-           # column spanner
-           cgroup = c("", "95% CI"), 
-           n.cgroup = c(4, 2),
-           padding.rgroup = "&nbsp;&nbsp;",
-           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
-           align = "llcccc", # column alignment,
-           tfoot="&dagger; Adjusted for CDC temperature, wind speed, relative humidity, precipitation, and accounting for subject.
-Time-stratified: referent periods matched to events on same day of week within July to October fire season."
-            ) # end table
-
-  print(tab)
-
-# ggplot of odds ratios, facet wrapped by outcomes -----
-# convert variables from character to either numeric or factor
-# factor preserves the order of the variable
-wrf_geo_age_plot <- combined_point_est_df %>% 
-  # filter columns to just geo-smk
-  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
-  # do not plot results with less than 15 events, create missing vals
-  # in mutate
-  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
-                             99, odds_ratio),
-         lower95 = ifelse(as.numeric(n_events) < 15, 
-                             99, lower95),
-         upper95 = ifelse(as.numeric(n_events) < 15, 
-                             99, upper95),
-         age_cat2 = ifelse(age_cat == 0, "<15", 
-                    ifelse(age_cat == 1, "15-65",
-                    ifelse(age_cat == 2, ">65", NA)))) %>% 
-  # subset columns I want to put in to the table
-  select(1, 3, 11, 5:8) %>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
-         lower95 = ifelse(lower95 == 99, NA, lower95),
-         upper95 = ifelse(upper95 == 99, NA, upper95))
-
-# change characters to numeric and factor  
-wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
-                              levels = unique(wrf_geo_age_plot$outcome))
-
-wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
-                                levels = unique(wrf_geo_age_plot$pm_method))
-
-wrf_geo_age_plot$age_cat2 <- factor(wrf_geo_age_plot$age_cat2,
-                                levels = unique(wrf_geo_age_plot$age_cat2))
-
-wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
-wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
-wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
-wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)
-
-## ggplot 
-  print_plot <- ggplot(wrf_geo_age_plot,
-    aes(x = age_cat2, y = odds_ratio, color = pm_method), na.rm = T) +
-    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
-    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
-                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
-    # custom color 
-    scale_color_manual(name = "Smoke-Estimation Method", 
-                       values = c("red", "#32115C"),
-                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
-    facet_wrap(~outcome, nrow = 3, scales = "free") +
-    geom_hline(yintercept = 1, linetype=2) +
-    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
-    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
-    #ylim(0, 2) +
-    xlab('Age Category ') +
-    # plot theme
-    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    # strip element
-    strip.background = element_rect(colour=NA, fill=NA),
-    panel.border = element_rect(fill = NA, color = "black"),
-    # facet text size
-    strip.text = element_text(size = 10),
-    # axis element
-    #axis.text.x = element_blank(),
-    #axis.title.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.title.y = element_text(angle = 90),
-    # legend elements
-    legend.position = "bottom")
-    #legend.text = element_text(size = 8))
-
-  print(print_plot)
-  # save figure
-  ggsave("./fig5.pdf", plot = print_plot, 
-       width = 12, height = 8, units = "in")
-
-```
-
-### Seasonal Time-Stratified by Sex
-
-Association with smoke PM~2.5~ and health outcomes stratified by sex.
-May be some evidence of effect modification by sex, particularly where males appear to be at greater risk for certain CVD outcomes with increasing smoke.
-However, only marginally significant.
-Also, broken arm estimate seems a little strange, suggesting females at greater risk too.
-Larger sample size may help to answer this.
-
-```{r sex strata cdc met adj, warning = F, echo = F, results='asis'} 
-# note above, results = 'asis' needed to print htmlTables
-
-# For age categories, I want a loop for each age category.
-
-# dataframe list
-df_list <- list(resp_casecross, asthma_casecross, copd_casecross, pneum_casecross,
-                acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
-                cereb_vas_casecross, hf_casecross, ihd_casecross, mi_casecross, 
-                broken_arm_casecross)
-
-outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 'Pneumonia', 
-                  'Acute Bronchitis', 'Cardiovascular Disease', 'Arrhythmia', 
-                  'Cerebrovascular Disease', 'Heart Failure',
-                  'Ischemic Heart Disease', 'Myocardial Infarction', 'Broken Arm')
-
-# looking only at wrf chem, kriging, and geo-weighted; taking out 'Global Smoke' method
-method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Global Smoke', 'GWR Smoke')
-
-# sex category list
-sex_strata_list <- c(0,1)
-
-# create an empty list to row bind dataframes together
-datalist1 <- list()
-datalist2 <- list()
-
-# data wrangling ----
-# Producing conditional logit model estimates loop 
-for(i in 1:length(df_list)){
-
-# dataframe to loop through
-  df_to_loop <- data.frame(df_list[i])
-  # indication of column
-  outcome <- colnames(df_to_loop[3])
-  # outcome name
-  outcome_name <- outcome_list[i]
-
-  # extract covariates from dataframe
-  covariates_df <- df_to_loop[, c(1:16, 27:28, 151:157)]
-
-  # extract pm values and divide by 10 and ordered
-  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
-  pm_estimates_df <- df_to_loop[, c(19, 26, 25, 24)]/10  # create 10 unit increases
-
-  # new loop for sex categories
-  for(k in 0:1){
-
-    # empty matrix (12 x 10 matrix)
-    point_estimates <- matrix(nrow = 4, ncol = 10, byrow = T)
-
-    colnames(point_estimates) <- c('outcome', 'sex', 'pm_method', 'n', 'n_events', 
-                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
-
-    # fill in the outcome namedataframe before method loop
-    point_estimates[, 1] <- outcome_name
-    # repeat the sex category 4 times for each pm method
-    point_estimates[, 2] <- k
-
-    # dataframe for analysis creation
-    # bind columns back together 
-    df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
-      # remove missing pm values
-      filter(!is.na(wrf_smk_pm_zip)) %>% 
-      # join in cdc met data
-      full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
-      # limit to emergency or urgent care
-      filter(ADM_TYPE == 1 | ADM_TYPE == 2) %>%
-      # limit to specific sex category
-      filter(sex_num == k) %>% 
-      # the following code makes sure that the counterfactual values retained are 
-      # symetric in that number of obs before = number of obs after
-      mutate(obs_diff_admission = (date - date_admit)/7) 
-      # dataframe is already for the entire fire season, so I don't need to subset anymore
-
-  # second loop to run a model for each pm estimation method
-    for(j in 26:29){
-
-      # variable to model 
-      var_name <- colnames(df_analysis[j])
-
-      # set row number to fill
-      row_n <- j-25
-
-      # only run the model if the dataframe has observations
-      if(nrow(df_analysis) != 0){
-      # conditional logistic regression model
-      mod <- clogit(outcome ~ df_analysis[[j]] + daily_meanRH + daily_meanT + daily_meanWS + 
-                    daily_meanPrcp + strata(PATIENTID), df_analysis)
-
-      # populate matrix
-      point_estimates[row_n, 3] <- method_list[row_n]
-      point_estimates[row_n, 4] <- mod$n
-      point_estimates[row_n, 5] <- mod$nevent
-      # odds ratio
-      point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
-
-      # 95% lower bound
-      point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
-                                        1.96*(summary(mod)$coefficient[1,3])), 3)
-      # 95% upper bound
-      point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
-                                        1.96*(summary(mod)$coefficient[1,3])), 3)
-      # standard error
-      point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
-      # p val
-      point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
-
-      # create else statement that fills matrix with missing so I still have the row
-      # in the final dataframe
-      } else {point_estimates[row_n, 3] <- method_list[row_n]
-              point_estimates[row_n, 4] <- 0
-              point_estimates[row_n, 5] <- 0
-              point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
-
-    } # end methods loop
-
-    # save point estimates as a dataframe
-    point_est_df <- as_data_frame(point_estimates)
-
-    # combine previous values in dataframe that has all outcome/methods comparisons
-    datalist1[[k+1]] <- point_est_df
-  } # end sex category loop
-
-  # bind rows of sex category estimates together
- sex_est_df <- bind_rows(datalist1)
-
- # populate second dataframe list
- datalist2[[i]] <- sex_est_df
-
-} # end of outcome loop
-
-# combine each outcome dataframe itteration in to a big dataset
-combined_point_est_df <- bind_rows(datalist2)
-
-wrf_geo_sex <- combined_point_est_df %>% 
-  # filter columns to just geo-smk
-  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
-  mutate(sex_cat = ifelse(sex == 0, "Male", 
-               ifelse(sex == 1, "Female", NA))) %>% 
-  # subset columns I want to put in to the table
-  select(3, 11, 5:8)%>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
-         lower95 = ifelse(lower95 == 99, "--", lower95),
-         upper95 = ifelse(upper95 == 99, "--", upper95))
-
-  tab <- htmlTable(txtRound(wrf_geo_sex, digits = 3, 1:3), 
-           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by sex",
-           # row group by outcome
-           rgroup = outcome_list,
-           n.rgroup = c(rep(4, 12)), # 2 rows for each sex strata for each outcome
-           # column headers
-           header = c("Est Method", "Sex Strata", "Events",
-                      "OR&dagger;", "Lower", "Upper"),
-           # column spanner
-           cgroup = c("", "95% CI"), 
-           n.cgroup = c(4, 2),
-           padding.rgroup = "&nbsp;&nbsp;",
-           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
-           align = "llcccc", # column alignment,
-           tfoot="&dagger; Adjusted for temperature, accounting for subject. Time-stratified: referent periods matched to events on same day of week within July to October fire season."
-            ) # end table
-
-  print(tab)
-
-# ggplot of odds ratios, facet wrapped by outcomes -----
-# convert variables from character to either numeric or factor
-# factor preserves the order of the variable
-wrf_geo_sex_plot <- combined_point_est_df %>% 
-  # filter columns to just geo-smk
-  filter(pm_method == "GWR Smoke" | 
-         pm_method == "WRF-Chem Smoke") %>% 
-  mutate(sex_cat = ifelse(sex == 0, "Male", 
-               ifelse(sex == 1, "Female", NA))) %>%
-  # subset columns I want to put in to the table
-  select(1, 3, 11, 5:8) %>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
-         lower95 = ifelse(lower95 == 99, NA, lower95),
-         upper95 = ifelse(upper95 == 99, NA, upper95))
-
-# change characters to numeric and factor  
-wrf_geo_sex_plot$outcome <- factor(wrf_geo_sex_plot$outcome,
-                              levels = unique(wrf_geo_sex_plot$outcome))
-
-wrf_geo_sex_plot$pm_method <- factor(wrf_geo_sex_plot$pm_method,
-                                levels = unique(wrf_geo_sex_plot$pm_method))
-
-wrf_geo_sex_plot$sex_cat <- factor(wrf_geo_sex_plot$sex_cat,
-                                levels = unique(wrf_geo_sex_plot$sex_cat))
-
-wrf_geo_sex_plot$n_events <- as.numeric(wrf_geo_sex_plot$n_events)
-wrf_geo_sex_plot$odds_ratio <- as.numeric(wrf_geo_sex_plot$odds_ratio)
-wrf_geo_sex_plot$lower95 <- as.numeric(wrf_geo_sex_plot$lower95)
-wrf_geo_sex_plot$upper95 <- as.numeric(wrf_geo_sex_plot$upper95)
-
-## ggplot
-  print_plot <- ggplot(wrf_geo_sex_plot,
-    aes(x = sex_cat, y = odds_ratio, color = pm_method), na.rm = T) +
-    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
-    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
-                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
-    # custom color 
-    scale_color_manual(name = "Smoke-Estimation Method", 
-                       values = c("red", "#32115C"),
-                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
-    facet_wrap(~outcome, nrow = 3, scales = "free") +
-    geom_hline(yintercept = 1, linetype=2) +
-    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
-    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
-    #ylim(0, 2) +
-    xlab('Sex') +
-    # plot theme
-    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    # strip element
-    strip.background = element_rect(colour=NA, fill=NA),
-    panel.border = element_rect(fill = NA, color = "black"),
-    # facet text size
-    strip.text = element_text(size = 10),
-    # axis element
-    #axis.text.x = element_blank(),
-    #axis.title.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.title.y = element_text(angle = 90),
-    # legend elements
-    legend.position = "bottom")
-    #legend.text = element_text(size = 8))
-
-  print(print_plot)
-  # save figure
-  ggsave("./fig6.pdf", plot = print_plot, 
-       width = 12, height = 8, units = "in")
-
-```
-
-### Lag Periods
-
-Lag period adjusted for CDC met variables.
-
-```{r lags with cdc met adj, warning = F, echo = F, results='asis'} 
-# note above, results = 'asis' needed to print htmlTables
-
-# dataframe list
-df_list <- list(resp_casecross, asthma_casecross, copd_casecross, pneum_casecross,
-                acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
-                cereb_vas_casecross, hf_casecross, ihd_casecross, mi_casecross, 
-                broken_arm_casecross)
-
-outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 'Pneumonia', 
-                  'Acute Bronchitis', 'Cardiovascular Disease', 'Arrhythmia', 
-                  'Cerebrovascular Disease', 'Heart Failure',
-                  'Ischemic Heart Disease', 'Myocardial Infarction', 'Broken Arm')
-
-# looking only at wrf chem, kriging, and geo-weighted; taking out 'Global Smoke' method
-method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Global Smoke', 'GWR Smoke')
-
-# create an empty list to row bind dataframes together
-datalist1 <- list()
-datalist2 <- list()
-
-# data wrangling ----
-# Producing conditional logit model estimates loop 
-for(i in 1:length(df_list)){
-
-# dataframe to loop through
-  df_to_loop <- data.frame(df_list[i])
-  # indication of column
-  outcome <- colnames(df_to_loop[3])
-  # outcome name
-  outcome_name <- outcome_list[i]
-  # extract covariates from dataframe
-  covariates_df <- df_to_loop[, c(1:16, 27,79:83, 28, 151:157)]
-
-  # extract pm values and divide by 10 and ordered
-  #which(colnames(df_to_loop)=="wrf_temp_lag1_zip") # code to find column numbers
-  #which(colnames(df_to_loop)=="wrf_temp_zip") # code to find column numbers
-  pm_lag_list <- list(df_to_loop[, c(19, 39:43)]/10, df_to_loop[, c(26, 74:78)]/10,
-                  df_to_loop[, c(25, 69:73)]/10, df_to_loop[, c(24, 64:68)]/10)  # create 10 unit increases
-
-# need to think if I need a methods section (probably not since this is it)
-
-  # new loop for each element of the lag list 
-  for(k in 1:4){
-
-    # empty matrix (12 x 10 matrix)
-    point_estimates <- matrix(nrow = 6, ncol = 10, byrow = T)
-
-    colnames(point_estimates) <- c('outcome', 'pm_method', 'lag', 'n', 'n_events', 
-                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
-
-    # fill in the outcome namedataframe before method loop
-    point_estimates[, 1] <- outcome_name
-    # repeat the sex category 4 times for each pm method
-    point_estimates[, 2] <- method_list[k]
-
-    pm_test <- pm_lag_list[[k]]
-
-    # dataframe for analysis creation
-    # bind columns back together 
-    df_analysis <- cbind(covariates_df, pm_lag_list[[k]]) %>% 
-      # limit to emergency or urgent care
-      filter(ADM_TYPE == 1 | ADM_TYPE == 2) %>%
-      # join in cdc met data
-      full_join(daily_zip_met, by = c("ZIPCODE", "date")) 
-      # dataframe is already for the entire fire season, so I don't need to subset anymore
-#glimpse(df_analysis)
-
-  # loop to run a model for each lag 0-5
-    for(j in 0:5){
-
-      # finding locations of columns in dataframe
-      which(colnames(df_analysis)=="daily_meanWS")
-
-      # set column number to evaluate for the pm lag method
-      pm_lag_colnum <- j + 31
-      # set column number of cdc temp lag to adjust for
-      temp_lag_colnum <- j + 49
-      # set col num for abs temp
-      ab_temp_lag_colnum <- j + 37
-      # set col num for relative humidity
-      rh_lag_colnum <- j + 43
-      # precip
-      pr_lag_colnum <- j + 55
-      # wind speed
-      ws_lag_colnum <- j + 61
-      # set row number to fill
-      row_n <- j+1
-
-      # only run the model if the dataframe has observations
-      if(nrow(df_analysis) != 0){
-      # conditional logistic regression model
-      mod <- clogit(outcome ~ df_analysis[[pm_lag_colnum]] + df_analysis[[temp_lag_colnum]] + 
-                    df_analysis[[rh_lag_colnum]] + df_analysis[[pr_lag_colnum]] + 
-                    df_analysis[[ws_lag_colnum]] + strata(PATIENTID), df_analysis)
-
-      # populate matrix
-      point_estimates[row_n, 3] <- j
-      point_estimates[row_n, 4] <- mod$n
-      point_estimates[row_n, 5] <- mod$nevent
-      # odds ratio
-      point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
-      # 95% lower bound
-      point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
-                                        1.96*(summary(mod)$coefficient[1,3])), 3)
-      # 95% upper bound
-      point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
-                                        1.96*(summary(mod)$coefficient[1,3])), 3)
-      # standard error
-      point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
-      # p val
-      point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
-
-      # create else statement that fills matrix with missing so I still have the row
-      # in the final dataframe
-      } else {point_estimates[row_n, 3] <- method_list[row_n]
-              point_estimates[row_n, 4] <- 0
-              point_estimates[row_n, 5] <- 0
-              point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
-
-    } # end methods loop
-
-    # save point estimates as a dataframe
-    point_est_df <- as_data_frame(point_estimates)
-
-    # combine previous values in dataframe that has all outcome/methods comparisons
-    datalist1[[k+1]] <- point_est_df
-  } # end sex category loop
-
-  # bind rows of sex category estimates together
- lag_est_df <- bind_rows(datalist1)
-
- # populate second dataframe list
- datalist2[[i]] <- lag_est_df
-
-} # end of outcome loop
-
-# combine each outcome dataframe itteration in to a big dataset
-combined_point_est_df <- bind_rows(datalist2)
-
-#glimpse(combined_point_est_df)
-# subset to just geo-smoke estimates for now to reduce number of rows
-wrf_geo_lag <- combined_point_est_df %>% 
-  # filter columns to just geo-smk
-  filter(pm_method == "GWR Smoke" |
-         pm_method == "WRF-Chem Smoke") %>% 
-  # subset columns I want to put in to the table
-  select(2:3, 5:8) %>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
-         lower95 = ifelse(lower95 == 99, "--", lower95),
-         upper95 = ifelse(upper95 == 99, "--", upper95))
-
-  tab <- htmlTable(txtRound(wrf_geo_lag, digits = 3, 1:3), 
-           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes by lagged days",
-           # row group by outcome
-           rgroup = outcome_list,
-           n.rgroup = c(rep(12, 12)), # 6 rows for each lag period for each outcome
-           # column headers
-           header = c("Est Method", "Lagged Days", "Events",
-                      "OR&dagger;", "Lower", "Upper"),
-           # column spanner
-           cgroup = c("", "95% CI"), 
-           n.cgroup = c(4, 2),
-           padding.rgroup = "&nbsp;&nbsp;",
-           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
-           align = "lccccc", # column alignment,
-           tfoot="&dagger; Adjusted for lagged CDC temperature, windspeed, relative humidity, and precipitation, accounting for subject."
-            ) # end table
-
-  print(tab)
-
-# ggplot of odds ratios, facet wrapped by outcomes -----
-# convert variables from character to either numeric or factor
-# factor preserves the order of the variable
-wrf_geo_lag_plot <- combined_point_est_df %>% 
-  # filter columns to just geo-smk
-  filter(pm_method == "GWR Smoke" |
-         pm_method == "WRF-Chem Smoke") %>% 
-  # subset columns I want to put in to the table
-  select(1:3, 5:8)
-
-# change characters to numeric and factor  
-wrf_geo_lag_plot$outcome <- factor(wrf_geo_lag_plot$outcome,
-                              levels = unique(wrf_geo_lag_plot$outcome))
-
-wrf_geo_lag_plot$pm_method <- factor(wrf_geo_lag_plot$pm_method,
-                                levels = unique(wrf_geo_lag_plot$pm_method))
-
-wrf_geo_lag_plot$lag <- factor(wrf_geo_lag_plot$lag,
-                                levels = unique(wrf_geo_lag_plot$lag))
-
-wrf_geo_lag_plot$n_events <- as.numeric(wrf_geo_lag_plot$n_events)
-wrf_geo_lag_plot$odds_ratio <- as.numeric(wrf_geo_lag_plot$odds_ratio)
-wrf_geo_lag_plot$lower95 <- as.numeric(wrf_geo_lag_plot$lower95)
-wrf_geo_lag_plot$upper95 <- as.numeric(wrf_geo_lag_plot$upper95)
-
-## ggplot
-  print_plot <- ggplot(wrf_geo_lag_plot,
-                  aes(x = lag, y = odds_ratio, color = pm_method), na.rm = T) +
-    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
-    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
-                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
-    # custom color 
-    scale_color_manual(name = "Smoke-Estimation Method", 
-                       values = c("red", "#32115C"),
-                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
-    facet_wrap(~outcome, nrow = 3, scales = "free") +
-    geom_hline(yintercept = 1, linetype=2) +
-    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
-    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
-    #ylim(0.7, 1.4) +
-    xlab('Lag Days') +
-    # plot theme
-    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    # strip element
-    strip.background = element_rect(colour=NA, fill=NA),
-    panel.border = element_rect(fill = NA, color = "black"),
-    # facet text size
-    strip.text = element_text(size = 10),
-    # axis element
-    #axis.text.x = element_blank(),
-    #axis.title.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.title.y = element_text(angle = 90),
-    # legend elements
-    legend.position = "bottom")
-    #legend.text = element_text(size = 8))
-
-  print(print_plot)
-
-  # save figure
-  ggsave("./fig4.pdf", plot = print_plot, 
-       width = 12, height = 8, units = "in")
-
-```
-
-### Overall Summary of Time-Stratified Results (Zip Code-Level)
-
-In some of my other documents, I tried different referent selection schemes for the case-crossover design, specifically the symmetric bi-directional design, where referent periods were matched to the same day of the week.
-In some cases, this produced similar estimates to the time-stratified results within season and wildfire season presented herein.
-However, as demonstrated in the Janes 2005 **Statistics in Medicine** paper, the symmetric bi-directional still has a bias in that the \beta is not necissarily unbiased, and can be a slight overestimation (which is evident in these results compared to some of my other results using that design).

### 2017-04-24 r_scripts/data_management/or_county_health.R
Update or_county_health.R
check the county with list

@@ -169,17 +169,58 @@ stop # 33.17 secs
 not_match <- which((zip_county_proportion > 0)&(zip_county_proportion < 1), arr.ind=T )
 dim(not_match) # 582
 
-not_match_value <- zip_county_proportion[not_match[,1], not_match[,2]]
+zip_county_prop_df <- data.frame(zip_county_proportion)
 
-# 1st col
-# rownames(not_match)
-
-# 2nd col
-# colnames(zip_county_proportion[,not_match[,2]])
-
-# 3rd col
+zip_county_prop_df2 <- zip_county_prop_df %>%
+  mutate(county_order = NA) %>%
+  mutate(county_name = NA)
 
 
+for (i in 1:length(zip_county_prop_df[,1])){
+  find_order <- which(zip_county_prop_df[i,]==(max(zip_county_prop_df[i,], na.rm=TRUE)))
+  zip_county_prop_df2$county_order[i] <- find_order
+  zip_county_prop_df2$county_name[i] <- colnames(zip_county_prop_df2[zip_county_prop_df2$county_order[i]])
+}
 
+zip_county_prop_df2 <- zip_county_prop_df2 %>%
+  mutate(zip=rownames(zip_county_prop_df))
+
+### Check using the zip county file from website
+or_check_county <- read_csv('./oregon_zip_county.csv')
+head(or_check_county)
+names(or_check_county) <- c('zip','city','county','area')
+
+or_check_county_df <- data.frame(or_check_county)
+
+or_check <- zip_county_prop_df2 %>%
+  select(zip, county_name) 
+
+or_check2 <- or_check_county %>%
+  select(zip, city, county)
+  
+or_check3 <- or_check2%>%
+  filter(zip %in% or_check$zip)
+  
+
+or_check4 <- or_check %>%
+  filter(zip %in% or_check3$zip)
+
+or_check5 <- or_check%>%
+  filter(!zip %in% or_check3$zip) # 97471
+
+a <- which(or_check4$county_name=="Hood.River") # 27, 36
+or_check4$county_name[a] <- "Hood River"
+
+or_check_combine <- or_check %>%
+  left_join(or_check, or_check2, by = "zip")
+
+b <- which(or_check3$county!=or_check4$county_name)
+# 11 147 160 162 167 168 184 346 347 348 349
+
+or_check3$county[b]
+or_check4$county[b]
+
+or_check3$zip[b]
+#  [1] 97014 97326 97347 97350 97358 97360 97378 97758 97759 97760 97761
 
 

### 2017-04-25 r_scripts/analysis/or_result_zip_manuscript.Rmd

--- 2017-04-12 r_scripts/analysis/or_zip_result.Rmd
@@
 This particular markdown document focuses on the zip code-level analysis.
-Note there is a county-level analysis that will be paired with CDC mortality data as well.
 Herein, I compare different PM~2.5~ estimation methods described in further detail below.
 
-Note, epidemiological methods that Sheryl and I think can improve our estimation and reduce some bias around the point estimates will not be the focus of the manuscript.
-
-```{r data import, include = F, echo = FALSE}
-# Set working directory and read in files --------------------------------------
-# relative path
-# set working directory if you want to run seperate chunks
-# setwd("./r_scripts")
-getwd()
-
-path <- paste0("../data_new/health_data")
-# Infile case-crossover dataframes ---------------------------------------------
-# Dataframes made in 'chars_3month_binary_smoke_may2016 script
-# resp exacerbations
+### Time-Stratified Case-Crossover (referents same day of week within fire season May - September)
+
+```{r include=FALSE}
+library(tidyverse)
+library(data.table)
+library(survival)
+library(htmlTable)
+library(readr)
+```
+
+```{r data import, include = F, eval = FALSE}
+
+### Limit to only ER visits
+## pos = 23
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+
+N <- NA
+n <- 1
+
+for (i in var_list){
+  read_path <- paste('../data_new/update/ER/or', i, 'may_to_sep_time_strat_casecross.csv', sep = "_")
+  or_disease <- read_csv(read_path)
+
+  or_disease <- or_disease %>% 
+    filter(pos=="23") %>%
+    # filter the unknown gender
+    filter(gender!= "U")
+
+  # Create a permanent case-cross over dataset
+  file_name <- paste('../data_new/update/ER/or', i, 'may_to_sep_time_strat_casecross_er.csv', sep = "_")
+
+  # write permanent dataset
+  write_csv(or_disease, paste0(file_name))
+  N[n] = length(unique(or_disease$personkey))
+  n = n+1
+}
+
+num_person <- data.frame(rbind(var_list, N))
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+## Analysis
+path <- paste0("../data_new/update/ER")
+
 resp_casecross <- read_csv(paste(path, 
-                        "respiratory_may_to_sep_time_strat_casecross.csv", 
+                                 "or_respiratory_may_to_sep_time_strat_casecross_er.csv", 
                                  sep = "/"))
@@
 
-### Time-Series of Smoke PM~2.5~ Ranges by Ecoregion
-
-Figures below are for each smoke PM~2.5~ estimation method.
-The figures are the daily range (min and max) for each zip code in a specific region of the state that local offices of the Washington Department of Ecology has jurisdiction over.
-The figure first starts out by showing the map of the state, divided in to eco regions used by the state for organization purposes, and then presents the time-series ranges for WRF-Chem, krig, geo-weighted estimation methods.
-
-```{r smk time series range by eco region, echo = F, results='asis', message=F}
-
-## ???
-# creation of map and identifying zipcodes in ecoregions----
-# import the regional shapefile downloaded from state website
-shp_dir <- paste0('../../data/shape_files/WAECY__Ecology_Regions')
-shp_dir <- paste0('./data/shape_files/WAECY__Ecology_Regions')
-
-eco_region <- readOGR(dsn = shp_dir, layer = 'WAECY__Ecology_Regions')
-# set coord ref system to WGS84 to match projections of ggmap
-eco_region <- spTransform(eco_region, "+proj=longlat +datum=WGS84")
-
-# summary(eco_region) # projection is GRS80, may be off
-
-# ggmap and rgdal needed
-# Set base map boundary layer
-bbox <- c(-125, 45, -116.1, 49.5)
-
-# call black and white map
-washington_map <- get_map(location= bbox, source = 'stamen',
-                          maptype = 'toner', crop = T)
-
-# call google terrain map
-#wash_map_terrain <- get_map(location = "washington state", source = 'google',
-#                             maptype = 'terrain', crop = F)
-
-# fortify washington shape data in to dataframe (need maptools package)
-eco_df <- tidy(eco_region, region = "ECY_REGION")
-
-# summary(wash_zip_df)
-
-# creation of eco region map (may change colors)
-eco_region_map <- ggmap(washington_map) + 
-  geom_polygon(data = eco_region, aes(x=long, y=lat, group = group, 
-    fill = group), color = "black", alpha = 0.8) +
-  scale_fill_manual(name = "Ecology Regions",
-                    values = c("#2c7bb6","#abd9e9","#d7191c","#fdae61"), 
-                    labels = c("Southwest", "Northwest", "Central", "Eastern"),
-                    guide = guide_legend(title.position = "top", order = 1)) +
-  geom_point(data = fire_loc, aes(x = Longitude, y = Latitude, shape = "Fire Locations"), 
-             color = "black", fill = "red") +
-  scale_shape_manual(name = "", values = 24, 
-                     guide = guide_legend(title.position = "top", order = 2)) +
-  ggtitle("Washington State Ecology \n Regional Office Jurisdiction") +
-  xlab("Longitude") +
-  ylab("Latitude") +
-  theme(#legend.position = "bottom",
-        #legend.box = "horizontal"
-    )
-
-print(eco_region_map)
-
-# import washington zip shapefile ----
-#zip_path <- paste0('../../data/shape_files/wash_zip_2012_shapefile')
-zip_path <- paste0('./data/shape_files/wash_zip_2012_shapefile')
-
-wash_zip_map <- readOGR(dsn = zip_path, layer = "wash_zip_2012_shapefile")
-
-# change projection to match eco region
-wash_zip_map <- spTransform(wash_zip_map, "+proj=longlat +datum=WGS84")
-
-#summary(eco_region)
-#summary(wash_zip_map)
-
-# check overlay visually 
-#plot(eco_region, lwd = 5)
-#plot(wash_zip_map, add = T)
-
-# find the eco regions that are over a zipcode
-zip_eco_region_overlap <- over(wash_zip_map, eco_region)
-# output zip IDs from zip shapefile
-ZIPCODE <- as.character(wash_zip_map@data$ZCTA5CE10)
-# join eco region ids with zip (can be used to indicate eco region)
-zip_eco_region <- cbind(ZIPCODE, zip_eco_region_overlap) %>% 
-  mutate(eco_region = ifelse(ECY_REGION == "CRO", "Central",
-                      ifelse(ECY_REGION == "ERO", "Eastern",
-                      ifelse(ECY_REGION == "NWRO", "Northwestern",
-                      ifelse(ECY_REGION == "SWRO", "Southwestern", NA))))) %>%  
-  select(-OBJECTID, -ECY_REGION)
-
-# range of PM values for each day in each eco region for each method ----
-# find the min, median, and max values for each date
-#summary(filter(zip_pm, date == "2012-07-9")) # quick sum of first zip obs 
-# zip_pm zipcode needs to be character to join with region
-zip_pm$ZIPCODE <- as.factor(zip_pm$ZIPCODE)
-
-# glimpse(zip_pm)
-# these joins should really be done in a loop, but...
-# each column would need to be specified for each pm val, 
-# which I'd have to think about more than just each one manually
-# wrf smoke
-wrf_range <- zip_pm %>% 
-  full_join(zip_eco_region, by = "ZIPCODE") %>%  
-  group_by(date, eco_region) %>% 
-  summarise(min = min(wrf_smk_pm, na.rm = T), 
-            med = median(wrf_smk_pm, na.rm = T), 
-            max = max(wrf_smk_pm, na.rm = T)) %>% 
-  # variable for small multiples
-  mutate(est_method = "wrf_smk")
-
-# krig smoke
-krig_range <- zip_pm %>% 
-  full_join(zip_eco_region, by = "ZIPCODE") %>%  
-  group_by(date, eco_region) %>% 
-  summarise(min = min(krig_smk_pm, na.rm = T), 
-            med = median(krig_smk_pm, na.rm = T), 
-            max = max(krig_smk_pm, na.rm = T)) %>% 
-  mutate(est_method = "krig_smk")
-
-# geo smoke
-geo_range <- zip_pm %>% 
-  full_join(zip_eco_region, by = "ZIPCODE") %>%  
-  group_by(date, eco_region) %>% 
-  summarise(min = min(geo_smk_pm, na.rm = T), 
-            med = median(geo_smk_pm, na.rm = T), 
-            max = max(geo_smk_pm, na.rm = T)) %>% 
-  mutate(est_method = "geo_smk")
-
-# loop plots though
-
-file_name <- c('wrf', 'krig', 'geo')
-title_name <- c('WRF-Chem Smoke', 'Kriging Smoke', 'GWR Smoke')
-ts_list <- list(wrf_range, krig_range, geo_range)
-
-for(i in 1:3){
-
-df_range <- ts_list[[i]]  
-gg_title <- title_name[i]
-
-# create plot for method range
-plot_print <- ggplot(df_range, aes(x = date)) +
-  geom_ribbon(aes(ymin = min, ymax = max, group = eco_region, 
-                  fill=eco_region), alpha = 0.7) +
-  scale_fill_manual(name = "Washington State Ecology \n Regional Office Jurisdiction",
-                    values = c("#d7191c", "#fdae61", "#abd9e9", "#2c7bb6"),
-                    labels = c("Central", "Eastern","Northwest", "Southwest"),
-                    guide = F) +
-  ylim(0,500) +
-  #geom_line(aes(y=med, group=eco_region, color=eco_region)) +
-  ggtitle(gg_title) +
-  ylab(expression('Smoke PM2.5 µg/m'^3)) +
-  xlab('Date') + 
-  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    axis.text.x = element_text(size = 12),
-    axis.text.y = element_text(size = 12),
-    axis.title.y = element_text(angle = 90, size = 12),
-    axis.title.x = element_text(size = 14), 
-    strip.background = element_rect(fill = 'white'))
-
-# file name
-name <- file_name[i]
-
-assign(paste0(name,"_plot"), plot_print)
-
-}
-
-# Layout needs to be edited and I would like to change colors
-# but this should be good enough for a review by CDC
-map_timeseries <- ggdraw() +
-  draw_plot(eco_region_map, x = 0, y = 0.5, width = 0.5, height = 0.5) +
-  draw_plot(wrf_plot, x = 0.5, y = 0.5, width = 0.5, height = 0.5) +
-  draw_plot(krig_plot, x = 0, y = 0, width = 0.5, height = 0.5) +
-  draw_plot(geo_plot, x = 0.5, y = 0, width = 0.5, height = 0.5)
-
-plot(map_timeseries)
-# save 6 inch by 4 inch
-ggsave("./fig1.pdf", plot = map_timeseries, 
-       width = 12, height = 8, units = "in")
-
-```
-
-## Days of Smoke In Washington
-
-Here is a figure of the days from May 1st to September 30 2013 where PM~2.5~ exceeded >10.
-
-```{r smoke map, message = F, warning = F, echo = F, results='asis'}
-
-# fire location has a weird date format, but I think is only september
-
-# zip code count of smoke ------------------------------------------------------
-zip_smoke_days <- zip_pm %>% 
-  # create binary indicators of days with smoke
-  mutate(wrf_smoke = if_else(wrf_smk_pm > 0, 1, 0),
-         wrf_smoke5 = if_else(wrf_smk_pm > 5, 1, 0),
-         wrf_smoke10 = if_else(wrf_smk_pm > 10, 1, 0),
-         geo_smoke = if_else(geo_smk_pm > 0, 1, 0),
-         geo_smoke5 = if_else(geo_smk_pm > 5, 1, 0),
-         geo_smoke10 = if_else(geo_smk_pm > 10, 1, 0)) %>% 
-  group_by(ZIPCODE) %>% 
-  summarise(wrf_smk_days = sum(wrf_smoke), wrf_smk5_days = sum(wrf_smoke5), 
-            wrf_smk10_days = sum(wrf_smoke10), geo_smk_days = sum(geo_smoke), 
-            geo_smk_days = sum(geo_smoke), geo_smk5_days = sum(geo_smoke5),
-            geo_smk10_days = sum(geo_smoke10)) %>% 
-  # change zipcode to character
-  mutate(ZIPCODE = as.character(ZIPCODE))
-
-# check smoke days
-#summary(zip_smoke_days)
-
-# tidy from broom package now works
-wash_zip_df <- tidy(wash_zip_map, region = "ZCTA5CE10") %>% 
-  # merge in smoke days
-  full_join(zip_smoke_days, by = c("id" = "ZIPCODE"))
-
-# map
-print_map <- ggmap(washington_map) + 
-  geom_polygon(data = wash_zip_df, 
-    aes(x=long, y=lat, group = group, fill = geo_smk10_days), alpha = 0.7) +
-  scale_fill_gradient(expression("Number of \nSmoke Days"),
-    low = '#2c7bb6', high = '#d7191c') + 
-  geom_point(data = fire_loc, aes(x = Longitude, y = Latitude, shape = "Fire Locations"), 
-             colour = "black", fill = "red") +
-  scale_shape_manual(name = "", values = 24) + 
-  xlab("Longitude") +
-  ylab("Latitude") +
-  # control legend/guide lable order
-  guides(fill = guide_colourbar(title.position = "top", order = 1),
-         shape = guide_legend(title.position = "top", order = 2))
-
-print(print_map)
-
-ggsave("./fig2.pdf", plot = print_map, 
-       width = 12, height = 8, units = "in")
-
-```
-
 ## Population-Weighted PM~2.5~ Assigned by Zip Code
@@
 
-Note for following tables: I wrote out 'Greater than rather' than using  '>' as it was increasing the following font and looked weird.
+For all of 2013, there were a total of millons of all claims records in the APAC dataset.
+During the May 1st to September 30th wildfire season in Oregon state, there were 47413 emergency visits for the specified cardiopulmonary codes.
+The number of persons (cases), as well as age-specific and sex-specific strate.
 
@@
 for(i in 1:length(df_list)){
-
-  i = 2
 # dataframe to loop through
@@
 tab <- htmlTable(txtRound(combined_point_est_df, digits = 1), 
-           caption = "Number of cases for each outcome observed from July 1st to October 31st, 2012",
+           caption = "Number of cases for each outcome observed from May 1st to September 30st, 2013",
            # column headers
@@
 
-### Background on Time-Stratified Case-Crossover Design
-
-The original time-stratified case-crossover design suggested picking referent periods based on any potential time-varying confounding, *a priori*.
-For example, if day of the week and the month of the exposure sequence is of concern, one might pick referents in the same day of the week and the same day of the month.
-The papers that discuss and support the use of a time-stratified design are the Janes et al.
-2005 **Epidemiology paper**, *Case-Crossover Analyses of Air Pollution Exposure Data; Referent Selection Strategies and Their Implication for Bias*.
-
-The authors state that with the time-stratified design, the *a priori* selection of the referent period on the same day of the week within the same month has the following benefits:
-
-1.
-Localizable - Meaning that the likelihood of index times conditional on the referent windows contains information about \beta, which I think means the relative difference in the exposed vs unexposed, conditioned on the exposure (air pollution in their example), is unbiased.
-
-2. Non-ignroable - Meaning that time-varying factors can be ignored when using a conditional logistic regression; controlled through design (matching on day of week within the same season).
-
-More referent periods is better for statistical efficiency, but there is a tradeoff, where additional referent periods may introduce what they call 'overlap bias', which they mathmatically define in Janes 2005 **Statistics in Medicine** paper, *Overlap Bias in the Case-Crossover Design, with Applications to Air Pollution Exposures*.
-Basically this bias occurs when the exposure assigned in the referent period is conditional on the strategy used to define the referent period.
-
-As Janes et al. state in their 2005 **Statistics in Medicine** paper:
-
-  > *Two of the main strengths of the case-crossover design in the air pollution context are
-that it controls fixed confounders and allows for control over time-dependent confounders
-by design. Choosing referents that are restricted to the same day of the week and season
-as the index time will control for these effects. Though we have not examined the degree
-to which various referent selection strategies control bias due to confounding, several points
-are clear. Referents that are closer to the index time will achieve greater control of seasonal
-confounding. However, sampling referents closer to the index time will also result in a loss
-of power due to less heterogeneity in the exposure series, and a smaller number of referents
-will also result in lower power.*
-
-I will use both Janes 2005 papers in **Epidemiology** and **Statistics in Medicine** to justify the use of the time-stratified design.
-
-This design has an advantage over other case-crossover designs I've explored in that you can use more referent times, and as the exposure at these times are averaged for each subject over a set time period, this balances out some of the more extreme values of PM~2.5~ that may persist over a couple weeks due to wildfire smoke.
-
-### Time-Stratified Case-Crossover (referents same day of week within fire season July - October)
-
-For these analyses and for the final results in the paper, I used referents within the entire wildfire season (July to October).
-This time-stratified referent strategy allows even more referent observations to average to 'smooth' out some of those extreme PM~2.5~ concentrations in some referent periods.
-For my sensitivity analysis of varying referent periods, please see other scripts.
-
-Adjusting for CDC Met Variables: Rish has provided zip-level meteorologic data on wind speed, precipitation, relative humidity, and temperature.
-Note a slight difference that will likely need to be addressed in discussion is that these measures are weighted based on census tract, not on the WRF grid like our measures of smoke.
-
-```{r data wrangle with zip met vars, warning = F, echo = F, message = F}
-
-# create dataset that can be merged to case-cross over dataframes
-daily_zip_met <- zip_met %>%
-  # limit to new mean met variables (getting rid of smoke variables I already have)
-  select(1:7) %>% 
-  rename(ZIPCODE = ZIP) %>% arrange(ZIPCODE, date) %>%
-  group_by(ZIPCODE) %>% 
-  # create lag variables
-  mutate(
-    # absolute temp
-    meanAT_lag1 = lag(daily_meanAT, 1, order_by = ZIPCODE), 
-    meanAT_lag2 = lag(daily_meanAT, 2, order_by = ZIPCODE),
-    meanAT_lag3 = lag(daily_meanAT, 3, order_by = ZIPCODE),
-    meanAT_lag4 = lag(daily_meanAT, 4, order_by = ZIPCODE),
-    meanAT_lag5 = lag(daily_meanAT, 5, order_by = ZIPCODE),
-    # relative humidity
-    meanRH_lag1 = lag(daily_meanRH, 1, order_by = ZIPCODE),
-    meanRH_lag2 = lag(daily_meanRH, 2, order_by = ZIPCODE),
-    meanRH_lag3 = lag(daily_meanRH, 3, order_by = ZIPCODE),
-    meanRH_lag4 = lag(daily_meanRH, 4, order_by = ZIPCODE),
-    meanRH_lag5 = lag(daily_meanRH, 5, order_by = ZIPCODE),
-    # temp
-    meanT_lag1 = lag(daily_meanT, 1, order_by = ZIPCODE),
-    meanT_lag2 = lag(daily_meanT, 2, order_by = ZIPCODE),
-    meanT_lag3 = lag(daily_meanT, 3, order_by = ZIPCODE),
-    meanT_lag4 = lag(daily_meanT, 4, order_by = ZIPCODE),
-    meanT_lag5 = lag(daily_meanT, 5, order_by = ZIPCODE),
-    # precip
-    meanPrcp_lag1 = lag(daily_meanPrcp, 1, order_by = ZIPCODE),
-    meanPrcp_lag2 = lag(daily_meanPrcp, 2, order_by = ZIPCODE),
-    meanPrcp_lag3 = lag(daily_meanPrcp, 3, order_by = ZIPCODE),
-    meanPrcp_lag4 = lag(daily_meanPrcp, 4, order_by = ZIPCODE),
-    meanPrcp_lag5 = lag(daily_meanPrcp, 5, order_by = ZIPCODE),
-    # wind speed
-    meanWS_lag1 = lag(daily_meanWS, 1, order_by = ZIPCODE),
-    meanWS_lag2 = lag(daily_meanWS, 2, order_by = ZIPCODE),
-    meanWS_lag3 = lag(daily_meanWS, 3, order_by = ZIPCODE),
-    meanWS_lag4 = lag(daily_meanWS, 4, order_by = ZIPCODE),
-    meanWS_lag5 = lag(daily_meanWS, 5, order_by = ZIPCODE)) %>% 
-  # sort so all like variables are together
-  select(1:3, 8:12, 4, 13:17, 5, 18:22, 6, 23:27, 7, 28:32)
-
-#which(colnames(daily_zip_met)=="meanAT_lag5")
-#glimpse(daily_zip_met)
-
-```
-
-```{r time strat within fire season adj for cdc met, warning = F, echo = F, results='asis'} 
-# note above, results = 'asis' needed to print htmlTables
-
-# I removed copd exacerbation and rheuamtoid arthritis
-
-# dataframe list
+```{r echo = F, warning = F, results='asis'}
+
 df_list <- list(resp_casecross, asthma_casecross, copd_casecross, 
@@
 # save figure
-  ggsave("./fig4.pdf", plot = print_plot, 
+ggsave("./fig_only_er.pdf", plot = print_plot, 
        width = 12, height = 8, units = "in")
-
-# figures for research day
-  # plot of only respiratory outcomes
-
-  xtabs(~ combined_point_est_df$outcome)
-# subset out global method and respiratory outcomes ----
-resp_point_est_df <- combined_point_est_df %>%
-
-  filter(outcome == "All Respiratory" |
-         outcome == "Asthma" |
-         outcome == "COPD" |
-         outcome == "Pneumonia" |
-         outcome == "Acute Bronchitis" |
-         outcome == "Broken Arm")
-
-## ggplot
-  resp_plot <- ggplot(resp_point_est_df, 
-                      aes(x = pm_method, y = odds_ratio, colour = pm_method)) +
-    # custom color 
-    scale_color_manual(name = "Smoke-Estimation Method", 
-                       values = c("red", "blue", "#32115C"),
-                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
-    geom_point() + #geom_text(vjust = 0, nudge_x = 0.3) +
-    geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
-    facet_wrap(~outcome, nrow = 2) +
-    geom_hline(yintercept = 1, linetype=2) +
-#    ggtitle('Association Between PM2.5 \n from Wildfire Smoke on Hospitalizations') +
-    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
-    #xlab('Smoke Estimation Method') +
-    # plot theme
-    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    # strip element
-    strip.background = element_rect(colour=NA, fill=NA),
-    panel.border = element_rect(fill = NA, color = "black"),
-    # facet text size
-    #strip.text = element_text(size = 8),
-    # axis element
-    axis.text.x = element_blank(),
-    axis.title.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.title.y = element_text(angle = 90),
-    # legend elements
-    legend.position = "bottom")
-    #legend.text = element_text(size = 8))
-
-  print(resp_plot)
-
-# subset out global method and respiratory outcomes ----
-cvd_point_est_df <- combined_point_est_df %>% 
-   filter(outcome == "Cardiovascular Disease" |
-         outcome == "Arrhythmia" |
-         outcome == "Cerebrovascular Disease" |
-         outcome == "Heart Failure" |
-         outcome == "Ischemic Heart Disease" |
-         outcome == "Myocardial Infarction")
-
-## ggplot
-  cvd_plot <- ggplot(cvd_point_est_df, 
-                      aes(x = pm_method, y = odds_ratio, colour = pm_method)) +
-    # custom color 
-    scale_color_manual(name = "Smoke-Estimation Method", 
-                       values = c("red", "blue", "#32115C"),
-                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
-    geom_point() + #geom_text(vjust = 0, nudge_x = 0.3) +
-    geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
-    facet_wrap(~outcome, nrow = 2) +
-    geom_hline(yintercept = 1, linetype=2) +
-#    ggtitle('Association Between PM2.5 \n from Wildfire Smoke on Hospitalizations') +
-    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
-    #xlab('Smoke Estimation Method') +
-    # plot theme
-    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    # strip element
-    strip.background = element_rect(colour=NA, fill=NA),
-    panel.border = element_rect(fill = NA, color = "black"),
-    # facet text size
-    #strip.text = element_text(size = 8),
-    # axis element
-    axis.text.x = element_blank(),
-    axis.title.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.title.y = element_text(angle = 90),
-    # legend elements
-    legend.position = "bottom")
-    #legend.text = element_text(size = 8))
-
-  print(cvd_plot)
-
 ```
-
-#### Result Summary: Time-Stratified Case-Crossover Design (same day of week within the season)
-
-A general interpretation is as follows:
-A 10 $\mu$g/m^3^ increase in PM~2.5~ attributed to wildfire smoke was associated with a X% increase in the risk for a health event.
-Where X% is the odds ratio - 1 (e.g.
-OR of 1.20 - 1 = 20% increase).
-Note that the odds ratio approximates the relative risk in this case due to the rare outcome assumption in that the calcuation for an odds ratio is similar to risk ratio calcuation.
-
-We see that all exposure methods are significantly associated with an increase in all respiratory outcomes and asthma.
-The WRF-Chem smoke method is associated with cerebrovascular disease as well.
-As for the other models, Kriging, Global Regression, and Geo-Weighted Regression, those are associated with COPD.
-I believe this is the most justifiable design approach that would average any extreme values of PM~2.5~ across the wildfire season for the vector of referent values for a particular subject.
-
-When adjusting for the CDC met variables, the results are slightly attenuated when compared to my adjustments with just WRF-Temp, but my overall conclusions do not change much.
-
-Here is the same table, but within season instead of fire season and adjusting for CDC met variables.
-Rish is worried about time-variant confounding within a subject that may bias the overall results.
-This might include seasonal differences in baseline risk of asthma.
 
@@
 # looking only at wrf chem, kriging, and geo-weighted; taking out 'Global Smoke' method
-method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Global Smoke', 'GWR Smoke')
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'GWR Smoke')
 
@@
 
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+
 # data wrangling ----
@@
   # indication of column
-  outcome <- colnames(df_to_loop[3])
+  outcome <- var_list[i]
   # outcome name
@@
   # extract covariates from dataframe
-  covariates_df <- df_to_loop[, c(1:16, 27:28, 151:157)]
+  covariates_df <- df_to_loop[, c(1:26, 71, 74, 76:85)]
 
@@
   #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
-  pm_estimates_df <- df_to_loop[, c(19, 26, 25, 24)]/10  # create 10 unit increases
+  pm_estimates_df <- df_to_loop[, c(88, 93, 92, 94)]/10  # create 10 unit increases
 
@@
     # empty matrix (12 x 10 matrix)
-    point_estimates <- matrix(nrow = 4, ncol = 10, byrow = T)
+    point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)
 
@@
       filter(!is.na(wrf_smk_pm_zip)) %>% 
-      # join in cdc met data
-      full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
-      # limit to emergency or urgent care
-      filter(ADM_TYPE == 1 | ADM_TYPE == 2) %>%
       # limit to specific age category
-      filter(age_cat == k) %>% 
+      filter(age_ind == k) %>% 
       # the following code makes sure that the counterfactual values retained are 
       # symetric in that number of obs before = number of obs after
-      mutate(obs_diff_admission = (date - date_admit)/7) 
+      mutate(obs_diff_admission = (fromdate - date)/7) 
       # dataframe is already for the entire fire season, so I don't need to subset anymore
@@
   # second loop to run a model for each pm estimation method
-    for(j in 26:29){
+    for(j in 39:41){
 
@@
       # set row number to fill
-      row_n <- j-25
+      row_n <- j-38
 
@@
       # conditional logistic regression model
-      mod <- clogit(outcome ~ df_analysis[[j]] + daily_meanRH + daily_meanT + daily_meanWS + 
-                    daily_meanPrcp + strata(PATIENTID), df_analysis)
+      mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
 
@@
   # save figure
-  ggsave("./fig5.pdf", plot = print_plot, 
+  ggsave("./age_fig.pdf", plot = print_plot, 
        width = 12, height = 8, units = "in")
@@
       filter(!is.na(wrf_smk_pm_zip)) %>% 
-      # join in cdc met data
-      full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
-      # limit to emergency or urgent care
-      filter(ADM_TYPE == 1 | ADM_TYPE == 2) %>%
       # limit to specific sex category
-      filter(sex_num == k) %>% 
+      filter(sex_ind == k) %>% 
       # the following code makes sure that the counterfactual values retained are 
       # symetric in that number of obs before = number of obs after
-      mutate(obs_diff_admission = (date - date_admit)/7) 
+      mutate(obs_diff_admission = (fromdate - date)/7) 
       # dataframe is already for the entire fire season, so I don't need to subset anymore
@@
   # save figure
-  ggsave("./fig6.pdf", plot = print_plot, 
+  ggsave("./sex_fig.pdf", plot = print_plot, 
        width = 12, height = 8, units = "in")
@@
 ```
-
-### Lag Periods
-
-Lag period adjusted for CDC met variables.
-
-```{r lags with cdc met adj, warning = F, echo = F, results='asis'} 
-# note above, results = 'asis' needed to print htmlTables
-
-# dataframe list
-df_list <- list(resp_casecross, asthma_casecross, copd_casecross, pneum_casecross,
-                acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
-                cereb_vas_casecross, hf_casecross, ihd_casecross, mi_casecross, 
-                broken_arm_casecross)
-
-outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 'Pneumonia', 
-                  'Acute Bronchitis', 'Cardiovascular Disease', 'Arrhythmia', 
-                  'Cerebrovascular Disease', 'Heart Failure',
-                  'Ischemic Heart Disease', 'Myocardial Infarction', 'Broken Arm')
-
-# looking only at wrf chem, kriging, and geo-weighted; taking out 'Global Smoke' method
-method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Global Smoke', 'GWR Smoke')
-
-# create an empty list to row bind dataframes together
-datalist1 <- list()
-datalist2 <- list()
-
-# data wrangling ----
-# Producing conditional logit model estimates loop 
-for(i in 1:length(df_list)){
-
-# dataframe to loop through
-  df_to_loop <- data.frame(df_list[i])
-  # indication of column
-  outcome <- colnames(df_to_loop[3])
-  # outcome name
-  outcome_name <- outcome_list[i]
-  # extract covariates from dataframe
-  covariates_df <- df_to_loop[, c(1:16, 27,79:83, 28, 151:157)]
-
-  # extract pm values and divide by 10 and ordered
-  #which(colnames(df_to_loop)=="wrf_temp_lag1_zip") # code to find column numbers
-  #which(colnames(df_to_loop)=="wrf_temp_zip") # code to find column numbers
-  pm_lag_list <- list(df_to_loop[, c(19, 39:43)]/10, df_to_loop[, c(26, 74:78)]/10,
-                  df_to_loop[, c(25, 69:73)]/10, df_to_loop[, c(24, 64:68)]/10)  # create 10 unit increases
-
-# need to think if I need a methods section (probably not since this is it)
-
-  # new loop for each element of the lag list 
-  for(k in 1:4){
-
-    # empty matrix (12 x 10 matrix)
-    point_estimates <- matrix(nrow = 6, ncol = 10, byrow = T)
-
-    colnames(point_estimates) <- c('outcome', 'pm_method', 'lag', 'n', 'n_events', 
-                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
-
-    # fill in the outcome namedataframe before method loop
-    point_estimates[, 1] <- outcome_name
-    # repeat the sex category 4 times for each pm method
-    point_estimates[, 2] <- method_list[k]
-
-    pm_test <- pm_lag_list[[k]]
-
-    # dataframe for analysis creation
-    # bind columns back together 
-    df_analysis <- cbind(covariates_df, pm_lag_list[[k]]) %>% 
-      # limit to emergency or urgent care
-      filter(ADM_TYPE == 1 | ADM_TYPE == 2) %>%
-      # join in cdc met data
-      full_join(daily_zip_met, by = c("ZIPCODE", "date")) 
-      # dataframe is already for the entire fire season, so I don't need to subset anymore
-#glimpse(df_analysis)
-
-  # loop to run a model for each lag 0-5
-    for(j in 0:5){
-
-      # finding locations of columns in dataframe
-      which(colnames(df_analysis)=="daily_meanWS")
-
-      # set column number to evaluate for the pm lag method
-      pm_lag_colnum <- j + 31
-      # set column number of cdc temp lag to adjust for
-      temp_lag_colnum <- j + 49
-      # set col num for abs temp
-      ab_temp_lag_colnum <- j + 37
-      # set col num for relative humidity
-      rh_lag_colnum <- j + 43
-      # precip
-      pr_lag_colnum <- j + 55
-      # wind speed
-      ws_lag_colnum <- j + 61
-      # set row number to fill
-      row_n <- j+1
-
-      # only run the model if the dataframe has observations
-      if(nrow(df_analysis) != 0){
-      # conditional logistic regression model
-      mod <- clogit(outcome ~ df_analysis[[pm_lag_colnum]] + df_analysis[[temp_lag_colnum]] + 
-                    df_analysis[[rh_lag_colnum]] + df_analysis[[pr_lag_colnum]] + 
-                    df_analysis[[ws_lag_colnum]] + strata(PATIENTID), df_analysis)
-
-      # populate matrix
-      point_estimates[row_n, 3] <- j
-      point_estimates[row_n, 4] <- mod$n
-      point_estimates[row_n, 5] <- mod$nevent
-      # odds ratio
-      point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
-      # 95% lower bound
-      point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
-                                        1.96*(summary(mod)$coefficient[1,3])), 3)
-      # 95% upper bound
-      point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
-                                        1.96*(summary(mod)$coefficient[1,3])), 3)
-      # standard error
-      point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
-      # p val
-      point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
-
-      # create else statement that fills matrix with missing so I still have the row
-      # in the final dataframe
-      } else {point_estimates[row_n, 3] <- method_list[row_n]
-              point_estimates[row_n, 4] <- 0
-              point_estimates[row_n, 5] <- 0
-              point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
-
-    } # end methods loop
-
-    # save point estimates as a dataframe
-    point_est_df <- as_data_frame(point_estimates)
-
-    # combine previous values in dataframe that has all outcome/methods comparisons
-    datalist1[[k+1]] <- point_est_df
-  } # end sex category loop
-
-  # bind rows of sex category estimates together
- lag_est_df <- bind_rows(datalist1)
-
- # populate second dataframe list
- datalist2[[i]] <- lag_est_df
-
-} # end of outcome loop
-
-# combine each outcome dataframe itteration in to a big dataset
-combined_point_est_df <- bind_rows(datalist2)
-
-#glimpse(combined_point_est_df)
-# subset to just geo-smoke estimates for now to reduce number of rows
-wrf_geo_lag <- combined_point_est_df %>% 
-  # filter columns to just geo-smk
-  filter(pm_method == "GWR Smoke" |
-         pm_method == "WRF-Chem Smoke") %>% 
-  # subset columns I want to put in to the table
-  select(2:3, 5:8) %>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
-         lower95 = ifelse(lower95 == 99, "--", lower95),
-         upper95 = ifelse(upper95 == 99, "--", upper95))
-
-  tab <- htmlTable(txtRound(wrf_geo_lag, digits = 3, 1:3), 
-           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes by lagged days",
-           # row group by outcome
-           rgroup = outcome_list,
-           n.rgroup = c(rep(12, 12)), # 6 rows for each lag period for each outcome
-           # column headers
-           header = c("Est Method", "Lagged Days", "Events",
-                      "OR&dagger;", "Lower", "Upper"),
-           # column spanner
-           cgroup = c("", "95% CI"), 
-           n.cgroup = c(4, 2),
-           padding.rgroup = "&nbsp;&nbsp;",
-           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
-           align = "lccccc", # column alignment,
-           tfoot="&dagger; Adjusted for lagged CDC temperature, windspeed, relative humidity, and precipitation, accounting for subject."
-            ) # end table
-
-  print(tab)
-
-# ggplot of odds ratios, facet wrapped by outcomes -----
-# convert variables from character to either numeric or factor
-# factor preserves the order of the variable
-wrf_geo_lag_plot <- combined_point_est_df %>% 
-  # filter columns to just geo-smk
-  filter(pm_method == "GWR Smoke" |
-         pm_method == "WRF-Chem Smoke") %>% 
-  # subset columns I want to put in to the table
-  select(1:3, 5:8)
-
-# change characters to numeric and factor  
-wrf_geo_lag_plot$outcome <- factor(wrf_geo_lag_plot$outcome,
-                              levels = unique(wrf_geo_lag_plot$outcome))
-
-wrf_geo_lag_plot$pm_method <- factor(wrf_geo_lag_plot$pm_method,
-                                levels = unique(wrf_geo_lag_plot$pm_method))
-
-wrf_geo_lag_plot$lag <- factor(wrf_geo_lag_plot$lag,
-                                levels = unique(wrf_geo_lag_plot$lag))
-
-wrf_geo_lag_plot$n_events <- as.numeric(wrf_geo_lag_plot$n_events)
-wrf_geo_lag_plot$odds_ratio <- as.numeric(wrf_geo_lag_plot$odds_ratio)
-wrf_geo_lag_plot$lower95 <- as.numeric(wrf_geo_lag_plot$lower95)
-wrf_geo_lag_plot$upper95 <- as.numeric(wrf_geo_lag_plot$upper95)
-
-## ggplot
-  print_plot <- ggplot(wrf_geo_lag_plot,
-                  aes(x = lag, y = odds_ratio, color = pm_method), na.rm = T) +
-    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
-    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
-                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
-    # custom color 
-    scale_color_manual(name = "Smoke-Estimation Method", 
-                       values = c("red", "#32115C"),
-                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
-    facet_wrap(~outcome, nrow = 3, scales = "free") +
-    geom_hline(yintercept = 1, linetype=2) +
-    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
-    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
-    #ylim(0.7, 1.4) +
-    xlab('Lag Days') +
-    # plot theme
-    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    # strip element
-    strip.background = element_rect(colour=NA, fill=NA),
-    panel.border = element_rect(fill = NA, color = "black"),
-    # facet text size
-    strip.text = element_text(size = 10),
-    # axis element
-    #axis.text.x = element_blank(),
-    #axis.title.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.title.y = element_text(angle = 90),
-    # legend elements
-    legend.position = "bottom")
-    #legend.text = element_text(size = 8))
-
-  print(print_plot)
-
-  # save figure
-  ggsave("./fig4.pdf", plot = print_plot, 
-       width = 12, height = 8, units = "in")
-
-```
-
-### Overall Summary of Time-Stratified Results (Zip Code-Level)
-
-In some of my other documents, I tried different referent selection schemes for the case-crossover design, specifically the symmetric bi-directional design, where referent periods were matched to the same day of the week.
-In some cases, this produced similar estimates to the time-stratified results within season and wildfire season presented herein.
-However, as demonstrated in the Janes 2005 **Statistics in Medicine** paper, the symmetric bi-directional still has a bias in that the \beta is not necissarily unbiased, and can be a slight overestimation (which is evident in these results compared to some of my other results using that design).

### 2017-04-25 r_scripts/parallel_presentation.Rmd

@@
 vars1 <- c(1:1610)
-z <- rep(201:300, each=100)
-w <- rep(201:300,100)
+z <- rep(1:417, each=1610)
+w <- rep(1:1610,417)
 ```
@@
 start <- proc.time()
-for (i in 201:300) {
+for (i in 1:417) {
   # output value of zipcode

### 2017-04-25 r_scripts/data_management/try_case_crossever_oregon.R
delete trial code

@@ -1,203 +1,0 @@
-# ------------------------------------------------------------------------------
-# Title: Case crossover for diseases of Oregon State (Practice)
-# Author: Jingyang Liu
-# Date Created: Feb 24, 2017
-# R version: 3.3.2
-# ------------------------------------------------------------------------------
-
-#Creating case crossover dataframes
-library(tidyverse)
-library(data.table)
-library(dplyr)
-library(readr)
-library(lubridate)
-getwd()
-read_path <- paste0('./data1/or_hosp_w_outcome_df.csv')
-disease <- read_csv(read_path) 
-
-disease <- disease %>%
-  # add new transverted from date
-  mutate(dates = as.Date(fromdate, "%m/%d/%Y"))
-
-ggplot(disease, aes(x = disease$dates)) + geom_density()
-
-
-# Asthma1 ----------------------------------------------------------------------
-disease_asthma1 <- disease %>%
-  # filter asthma1
-  filter(asthma1==1) %>%
-  # filter ZIP=80526
-  filter(ZIP==80526) %>%
-  # filter date in July 1, 2012 and Oct 31, 2012
-  filter(dates >= '2012-07-01' & 
-           dates <= '2012-10-31') %>%
-  # add ID
-  mutate(id = seq(1, nrow(.), by = 1),
-         # add outcome and new admit date
-         outcome_asthma1 = 1)
-
-one_asthma1_case <- sample_n(disease_asthma1, 1) %>% 
-  select(cdpheid, admit, dates, dx1, asthma1, ZIP, county_final, WRFGRID_ID,
-         RACE, sex_ind, age_ind, id)
-
-one_asthma1_case
-
-# find the replicate times of weeks
-dates <- one_asthma1_case[[1,3]] 
-n1 <- 0
-d=as.Date("2012-07-01")
-i=1
-while (dates >= "2012-07-01"){
-  dates <- dates - 7
-  d[i] = dates
-  i = i+1
-  n1 = n1+1
-}
-d[1:n1-1] # shows character(0) when the first week
-n1-1
-
-dates <- one_asthma1_case[[1,3]] 
-n2=0
-e=as.Date("2012-10-31")
-j=1
-while (dates <= "2012-10-31"){
-  dates <- dates + 7
-  e[j]=dates
-  j=j+1
-  n2 = n2 + 1
-}
-e[1:n2-1] # shows character(0) when the last week
-n2-1
-
-# replicate covariates length of counterfactual dates
[…]

### 2017-04-25 r_scripts/data_management/try_oregon_data_cleaning.R
delete primary trial summary code

@@ -1,177 +1,0 @@
-# ------------------------------------------------------------------------------
-# Title: Oregon 2013 claims data for MobaXterm (Practice)
-# Author: Jingyang Liu
-# Date Created: Feb 9th, 2017
-# R Version: 3.3.2
-# ------------------------------------------------------------------------------
-
-getwd() # "/home/jyliu/wildfire/r_scripts"
-
-library(dplyr)
-library(data.table)
-
-read_path <- paste0("../data/gan_episodes_of_care.txt")
-oregon_df <- fread(read_path, sep = "|", nrows = 2000000, showProgress = T)
-fwrite(oregon_df, "../data/oregon_epis_care_reduced.csv")
-
-# I wrote a reduced dataframe of 2000,000 rows to get an idea of data structure
-
-# import the reduced dataframe
-start_time <- Sys.time()
-
-oregon_df <- fread("../data/oregon_epis_care_reduced.csv", sep = ",", 
-                   colClasses = rep("character", 72))
-
-stop_time <-  Sys.time() - start_time 
-# time it took
-stop_time # 15.83603 secs
-
-
-start_time <- Sys.time()
-oregon_full <- fread(read_path, sep = "|", showProgress = T) # not change class
-stop_time <-  Sys.time() - start_time 
-# time it took
-stop_time # 7.771713 mins
-
-summary(as.factor(oregon_full$year)) # all 2013, 77069320
-summary(as.factor(oregon_full$personkey))
-min(as.numeric(oregon_full$personkey))
-max(as.numeric(oregon_full$personkey)) # seems length is 8
-unique(as.numeric(oregon_full$line)) # 1362 levels, 0 - 9999
-summary(as.factor(oregon_full$clmstatus))
-#       D        E        P
-# 8676643  3324496 65068181
-summary(as.factor(oregon_full$cob)) 
-#                 N        Y
-# 23517978 44438688  9112654
-summary(as.factor(oregon_full$paytype)) #C, D, G, P, T
-summary(as.factor(oregon_full$prod))
-summary(as.factor(oregon_full$payer))
-summary(as.factor(oregon_full$medflag))
-#       0        1     NA's
-# 5697342 70573441   798537
-summary(as.factor(oregon_full$rxflag))
-#       0        1     NA's
-# 7761348 68509435   798537
-summary(as.factor(oregon_full$ohvmhflag))
[…]

### 2017-04-25 r_scripts/data_management/or_county_health.R
Update or_county_health.R

@@ -1,5 +1,5 @@
 #-------------------------------------------------------------------------------
-# Title: Check and exploring the unreasonable plot
+# Title: Zip Code Assigned in County and Check
 # Author: Jingyang Liu                                                                   
 # Date Created: April 11, 2017                                                  
 # R version: 3.3.3                                                       

### 2017-05-04 r_scripts/data_management/or_2013_county_er_time_series.R
ER visit time series

@@ -1,0 +1,110 @@
+# ------------------------------------------------------------------------------
+# Title: Creation of county-level time-series only er visit dataframe for Oregon 2013
+# Author: Jingyang Liu
+# Date Created: May 2 2017
+# ------------------------------------------------------------------------------
+
+# Load libraries -----
+library(tidyverse)
+
+# Import dataframe with outcome ---- -------------------------------------------
+
+# Note: file is saved on my work computer or the secured CSU server
+# create read path
+getwd()
+setwd("C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire_new/data_disease")
+
+### ---------------------------------------------------------------------
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+or_list <- list()
+
+start <- Sys.time()
+for (i in var_list){
+  read_path <- paste('oregon', i, 'claims_2013.csv', sep='_')
+  or_disease <- read_csv(read_path)
+  
+  or_disease1 <- or_disease %>%
+    filter(pos == 23) %>%
+    group_by(personkey) %>% 
+    arrange(personkey, fromdate, line) %>% 
+    mutate(num_visit = dense_rank(fromdate)) %>%
+    group_by(personkey, fromdate) %>%
+    select(personkey, fromdate, num_visit, clmid, line, ZIP) %>% 
+    filter(row_number() == 1) %>%
+    mutate(dates = as.Date(fromdate, "%m/%d/%Y")) %>%
+    mutate(outcome = 1)
+  
+  colnames(or_disease1)[colnames(or_disease1)=="outcome"] <- i
+  or_list[[which(var_list==i)]] <- or_disease1
+
+}
+total_time <- Sys.time() - start
+total_time # Time difference of 12 sec
+
+summary(or_list)
+
+or_2013_disease <- or_list[[1]]%>%
+  full_join(or_list[[2]], by=c("personkey", "fromdate", "num_visit", "clmid", "line", "ZIP", "dates")) 
+
+for(j in 3:12){
+  or_2013_disease <- or_2013_disease %>%
+    full_join(data.frame(or_list[[j]]), by=c("personkey", "fromdate", "num_visit", "clmid", "line", "ZIP", "dates")) 
+}
+
[…]

### 2017-05-05 r_scripts/data_management/or_2013_county_er_time_series.R
Save as write csv file name

@@ -105,6 +105,6 @@ summary(or_2013_county_er_time_series)
 
 # write permanent dataframe
 
-write_path <- paste0("../data_new/county_data/or_asthma_county_time_series.csv")
+write_path <- paste0("../data_new/county_data/or_2013_county_er_time_series.csv")
 
 write_csv(or_2013_county_er_time_series, write_path)

### 2017-05-05 r_scripts/analysis/or_result_zip_smoke.Rmd
Update or_result_zip_smoke.Rmd
updated to the latest version with plots and tables are readable.

@@
 ---
-title: "Oregon: Zip-Level Wildfire Smoke and Health Outcomes (open in wildfire, not script)"
+title: "Oregon: Zip-Level Wildfire Smoke and Health Outcomes (open in script)"
 author: "Jingyang Liu"
@@
 Herein, I compare different PM~2.5~ estimation methods described in further detail below.
+
+### Time-Stratified Case-Crossover (referents same day of week within fire season May - September)
 
@@
 for (i in var_list){
-  read_path <- paste('../data_new/update/ER/or', i, 'may_to_sep_time_strat_casecross.csv', sep = "_")
+  read_path <- paste('../data_new/casecrossover/time_strat/or', i, 'may_to_sep_time_strat_casecross.csv', sep = "_")
   or_disease <- read_csv(read_path)
@@
   # Create a permanent case-cross over dataset
-  file_name <- paste('../data_new/update/ER/or', i, 'may_to_sep_time_strat_casecross_er.csv', sep = "_")
+  file_name <- paste('../data_new/casecrossover/ER/or', i, 'may_to_sep_time_strat_casecross_er.csv', sep = "_")
 
@@
 ## Analysis
-path <- paste0("../data_new/update/ER")
+path <- paste0("../data_new/casecrossover/ER")
 
@@
 # read zip pm for counts by zip
-smoke_path <- paste0("../data/Oregon/")
-zip_pm <- read_csv(paste0(smoke_path, "zip_pm_to_merge_with_chars.csv"))
+smoke_path <- paste0("../data/Oregon_PM/")
+zip_pm <- read_csv(paste0(smoke_path, "zip_pm_to_merge_with_acap.csv"))
 ```
@@
   df_analysis <- df_to_loop %>%
+    filter(!is.na(wrf_smk_pm_zip)) %>% 
     # only look at outcomes
@@
 ```
-
-### Time-Stratified Case-Crossover (referents same day of week within fire season July - October)
-
-For these analyses and for the final results in the paper, I used referents within the entire wildfire season (July to October).
-This time-stratified referent strategy allows even more referent observations to average to 'smooth' out some of those extreme PM~2.5~ concentrations in some referent periods.
-For my sensitivity analysis of varying referent periods, please see other scripts.
-
-Adjusting for CDC Met Variables: Rish has provided zip-level meteorologic data on wind speed, precipitation, relative humidity, and temperature.
-Note a slight difference that will likely need to be addressed in discussion is that these measures are weighted based on census tract, not on the WRF grid like our measures of smoke.
 
@@
 # save figure
-ggsave("./fig_only_er.pdf", plot = print_plot, 
+ggsave("../plot_new/fig_only_er.pdf", plot = print_plot, 
        width = 12, height = 8, units = "in")
@@
 I believe this is the most justifiable design approach that would average any extreme values of PM~2.5~ across the wildfire season for the vector of referent values for a particular subject.
+
+#### Wildfire Smoke and Oregon All Payer All Claims Results
+We found a  10 $\mu$g/m^3^ increase in PM~2.5~ was significantly associated with an 2.8% increase for respiratory oucomes  using method of WRF-Chem smoke (OR: 1.028 95% Confidence Interval [CI]: 1.001-1.055), and no significant association for kriging smoke method (OR: 1.018, 95% CI: 0.998-1.038), and also no significant association for GWR smoke method (OR: 1.021, 95% CI: 0.999-1.043)
+A 10 $\mu$g/m^3^ increase in PM~2.5~ was significantly associated with an 9.0% increase for asthma outcomes using  WRF-Chem smoke method (OR: 1.090, 95% Confidence Interval [CI]: 1.014-1.171), and 7.9% increase for kriging smoke method (OR: 1.079, 95% CI: 1.036-1.123), and 8.5% increase for GWR smoke method (OR: 1.085, 95% CI: 1.038-1.134)
+
+```{r lags with cdc met adj, warning = F, echo = F, results='asis', eval=FALSE, include = F} 
+# note above, results = 'asis' needed to print htmlTables
+
+# dataframe list
+df_list <- list(resp_casecross, asthma_casecross, copd_casecross, pneum_casecross,
+                acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
+                cereb_vas_casecross, hf_casecross, ihd_casecross, mi_casecross, 
+                broken_arm_casecross)
+
+outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 'Pneumonia', 
+                  'Acute Bronchitis', 'Cardiovascular Disease', 'Arrhythmia', 
+                  'Cerebrovascular Disease', 'Heart Failure',
+                  'Ischemic Heart Disease', 'Myocardial Infarction', 'Broken Arm')
+
+# looking only at wrf chem, kriging, and geo-weighted; taking out 'Global Smoke' method
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Global Smoke', 'GWR Smoke')
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# create an empty list to row bind dataframes together
+datalist1 <- list()
+datalist2 <- list()
+
+# data wrangling ----
+# Producing conditional logit model estimates loop 
+for(i in 1:length(df_list)){
+ i = 2
+
+# dataframe to loop through
+  df_to_loop <- data.frame(df_list[i])
+  # indication of column
+  outcome <- colnames(df_to_loop[76])
+  # outcome name
+  outcome_name <- outcome_list[i]
+
+  # extract covariates from dataframe (Oregon)
+  covariates_df <- df_to_loop[, c(1:26, 71, 74, 76:85)]
+
+  # extract pm values and divide by 10 and ordered
+  # which(colnames(df_to_loop)=="wrf_smk_pm_lag1_zip") # code to find column numbers (105)
+  # which(colnames(df_to_loop)=="wrf_smk_pm_zip") # code to find column numbers (88)
+  # which(colnames(df_to_loop)=="krig_smk_pm_lag1_zip") # code to find column numbers (130)
+  # which(colnames(df_to_loop)=="krig_smk_pm_zip") # code to find column numbers (93)
+  # which(colnames(df_to_loop)=="geo_smk_pm_lag1_zip") # code to find column numbers (125)
+  # which(colnames(df_to_loop)=="geo_smk_pm_zip") # code to find column numbers (92)
+  # which(colnames(df_to_loop)=="wrf_temp_lag1_zip") # code to find column numbers (135)
+  # which(colnames(df_to_loop)=="wrf_temp_zip") # code to find column numbers (94)
+
+  pm_lag_list <- list(df_to_loop[, c(88, 105:109)]/10, df_to_loop[, c(93, 130:134)]/10,
+                  df_to_loop[, c(92, 125:129)]/10, df_to_loop[, c(94, 135:139)]/10)  # create 10 unit increases
+
+# need to think if I need a methods section (probably not since this is it)
+
+  # new loop for each element of the lag list 
+  for(k in 1:4){
+    k = 1
+    # empty matrix (12 x 10 matrix)
+    point_estimates <- matrix(nrow = 6, ncol = 10, byrow = T)
+
+    colnames(point_estimates) <- c('outcome', 'pm_method', 'lag', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+    # fill in the outcome namedataframe before method loop
+    point_estimates[, 1] <- outcome_name
+    # repeat the sex category 4 times for each pm method
+    point_estimates[, 2] <- method_list[k]
+
+    pm_test <- pm_lag_list[[k]]
+
+    # dataframe for analysis creation
+    # bind columns back together 
+    df_analysis <- cbind(covariates_df, pm_lag_list[[k]])
+#glimpse(df_analysis)
+
+  # loop to run a model for each lag 0-5
+    for(j in 0:5){
+
+      j = 0
+
+      # finding locations of columns in dataframe
+      # which(colnames(df_analysis)=="daily_meanWS")
+
+      # set column number to evaluate for the pm lag method
+      #pm_lag_colnum <- j + 31
+      # set column number of cdc temp lag to adjust for
+      #temp_lag_colnum <- j + 49
+      # set col num for abs temp
+      #ab_temp_lag_colnum <- j + 37
+      # set col num for relative humidity
+      #rh_lag_colnum <- j + 43
+      # precip
+      #pr_lag_colnum <- j + 55
+      # wind speed
+      #ws_lag_colnum <- j + 61
+      # set row number to fill
+      #row_n <- j+1
+
+      # only run the model if the dataframe has observations
+      if(nrow(df_analysis) != 0){
+      # conditional logistic regression model
+     #  mod <- clogit(outcome ~  df_analysis[[pm_lag_colnum]] + df_analysis[[temp_lag_colnum]] + 
+     #                df_analysis[[rh_lag_colnum]] + df_analysis[[pr_lag_colnum]] + 
+     #                df_analysis[[ws_lag_colnum]] +strata(personkey), df_analysis)
+
+        mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+      # populate matrix
+      point_estimates[row_n, 3] <- j
+      point_estimates[row_n, 4] <- mod$n
+      point_estimates[row_n, 5] <- mod$nevent
+      # odds ratio
+      point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+      # 95% lower bound
+      point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
+                                        1.96*(summary(mod)$coefficient[1,3])), 3)
+      # 95% upper bound
+      point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                        1.96*(summary(mod)$coefficient[1,3])), 3)
+      # standard error
+      point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
+      # p val
+      point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
+
+      # create else statement that fills matrix with missing so I still have the row
+      # in the final dataframe
+      } else {point_estimates[row_n, 3] <- method_list[row_n]
+              point_estimates[row_n, 4] <- 0
+              point_estimates[row_n, 5] <- 0
+              point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
+
+    } # end methods loop
+
+    # save point estimates as a dataframe
+    point_est_df <- as_data_frame(point_estimates)
+
+    # combine previous values in dataframe that has all outcome/methods comparisons
+    datalist1[[k+1]] <- point_est_df
+  } # end sex category loop
+
+  # bind rows of sex category estimates together
+ lag_est_df <- bind_rows(datalist1)
+
+ # populate second dataframe list
+ datalist2[[i]] <- lag_est_df
+
+} # end of outcome loop
+
+# combine each outcome dataframe itteration in to a big dataset
+combined_point_est_df <- bind_rows(datalist2)
+
+#glimpse(combined_point_est_df)
+# subset to just geo-smoke estimates for now to reduce number of rows
+wrf_geo_lag <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  filter(pm_method == "GWR Smoke" |
+         pm_method == "WRF-Chem Smoke") %>% 
+  # subset columns I want to put in to the table
+  select(2:3, 5:8) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+  tab <- htmlTable(txtRound(wrf_geo_lag, digits = 3, 1:3), 
+           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes by lagged days",
+           # row group by outcome
+           rgroup = outcome_list,
+           n.rgroup = c(rep(12, 12)), # 6 rows for each lag period for each outcome
+           # column headers
+           header = c("Est Method", "Lagged Days", "Events",
+                      "OR&dagger;", "Lower", "Upper"),
+           # column spanner
+           cgroup = c("", "95% CI"), 
+           n.cgroup = c(4, 2),
+           padding.rgroup = "&nbsp;&nbsp;",
+           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+           align = "lccccc", # column alignment,
+           tfoot="&dagger; Adjusted for lagged CDC temperature, windspeed, relative humidity, and precipitation, accounting for subject."
+            ) # end table
+
+  print(tab)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_lag_plot <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  filter(pm_method == "GWR Smoke" |
+         pm_method == "WRF-Chem Smoke") %>% 
+  # subset columns I want to put in to the table
+  select(1:3, 5:8)
+
+# change characters to numeric and factor  
+wrf_geo_lag_plot$outcome <- factor(wrf_geo_lag_plot$outcome,
+                              levels = unique(wrf_geo_lag_plot$outcome))
+
+wrf_geo_lag_plot$pm_method <- factor(wrf_geo_lag_plot$pm_method,
+                                levels = unique(wrf_geo_lag_plot$pm_method))
+
+wrf_geo_lag_plot$lag <- factor(wrf_geo_lag_plot$lag,
+                                levels = unique(wrf_geo_lag_plot$lag))
+
+wrf_geo_lag_plot$n_events <- as.numeric(wrf_geo_lag_plot$n_events)
+wrf_geo_lag_plot$odds_ratio <- as.numeric(wrf_geo_lag_plot$odds_ratio)
+wrf_geo_lag_plot$lower95 <- as.numeric(wrf_geo_lag_plot$lower95)
+wrf_geo_lag_plot$upper95 <- as.numeric(wrf_geo_lag_plot$upper95)
+
+## ggplot
+  print_plot <- ggplot(wrf_geo_lag_plot,
+                  aes(x = lag, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0.7, 1.4) +
+    xlab('Lag Days') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10),
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+
+  # save figure
+  ggsave("../plot_new/fig4.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+### Time-Stratified by Age Categories
+
+Following table and figure look at some outcomes stratified by age category.
+Note that kids <=15 likely won't have many outcomes like COPD or MI.
+I have restricted the dataset to just the geo smoke method to reduce the number of rows/points on the tables and figures to make it easier to read for now.
+Code can easily be changed to have all PM~2.5~ methods.
+
+For some outcomes, there may be some evidence of effect modificaiton, like COPD, where >65 age group is at significant risk.
+However, the point estimates are relatively similar for the 15 to 65 strata, just not significant.
+Some error bars are not shown as they are too wide; usually <15 age category where there are only a handful of outcomes.
+
+```{r time stratified age strata cdc met, warning =F, echo = F, results='asis'} 
+# note above, results = 'asis' needed to print htmlTables
+
+# For age categories, I want a loop for each age category. 
+# Presenting just WRF and GWR
+
+# dataframe list
+df_list <- list(resp_casecross, asthma_casecross, copd_casecross, pneum_casecross,
+                acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
+                cereb_vas_casecross, hf_casecross, ihd_casecross, mi_casecross, 
+                broken_arm_casecross)
+
+outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 'Pneumonia', 
+                  'Acute Bronchitis', 'Cardiovascular Disease', 'Arrhythmia', 
+                  'Cerebrovascular Disease', 'Heart Failure',
+                  'Ischemic Heart Disease', 'Myocardial Infarction', 'Broken Arm')
+
+# looking only at wrf chem, kriging, and geo-weighted; taking out 'Global Smoke' method
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'GWR Smoke')
+
+# age category list
+age_cat_list <- c(0,1,2)
+
+# create an empty list to row bind dataframes together
+datalist1 <- list()
+datalist2 <- list()
+
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+
+# data wrangling ----
+# Producing conditional logit model estimates loop 
+for(i in 1:length(df_list)){
+
+# dataframe to loop through
+  df_to_loop <- data.frame(df_list[i])
+  # indication of column
+  outcome <- var_list[i]
+  # outcome name
+  outcome_name <- outcome_list[i]
+
+  # extract covariates from dataframe
+  covariates_df <- df_to_loop[, c(1:26, 71, 74, 76:85)]
+
+  # extract pm values and divide by 10 and ordered
+  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+  pm_estimates_df <- df_to_loop[, c(88, 93, 92, 94)]/10  # create 10 unit increases
+
+  # new loop for age categories
+  for(k in 0:2){
+
+    # empty matrix (12 x 10 matrix)
+    point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)
+
+    colnames(point_estimates) <- c('outcome', 'age_cat', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+    # fill in the outcome namedataframe before method loop
+    point_estimates[, 1] <- outcome_name
+    # repeat the age category 4 times for each pm method
+    point_estimates[, 2] <- k
+
+    # dataframe for analysis creation
+    # bind columns back together 
+    df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+      # remove missing pm values
+      filter(!is.na(wrf_smk_pm_zip)) %>% 
+      # limit to specific age category
+      filter(age_ind == k) %>% 
+      # the following code makes sure that the counterfactual values retained are 
+      # symetric in that number of obs before = number of obs after
+      mutate(obs_diff_admission = (fromdate - date)/7) 
+      # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+  # second loop to run a model for each pm estimation method
+    for(j in 39:41){
+
+      # variable to model 
+      var_name <- colnames(df_analysis[j])
+
+      # set row number to fill
+      row_n <- j-38
+
+      # only run the model if the dataframe has observations
+      if(nrow(df_analysis) != 0){
+      # conditional logistic regression model
+      mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+      # populate matrix
+      point_estimates[row_n, 3] <- method_list[row_n]
+      point_estimates[row_n, 4] <- mod$n
+      point_estimates[row_n, 5] <- mod$nevent
+      # odds ratio
+      point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+      # 95% lower bound
+      point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
+                                        1.96*(summary(mod)$coefficient[1,3])), 3)
+      # 95% upper bound
+      point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                        1.96*(summary(mod)$coefficient[1,3])), 3)
+      # standard error
+      point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
+      # p val
+      point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
+
+      # create else statement that fills matrix with missing so I still have the row
+      # in the final dataframe
+      } else {point_estimates[row_n, 3] <- method_list[row_n]
+              point_estimates[row_n, 4] <- 0
+              point_estimates[row_n, 5] <- 0
+              point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
+
+    } # end methods loop
+
+    # save point estimates as a dataframe
+    point_est_df <- as_data_frame(point_estimates)
+
+    # combine previous values in dataframe that has all outcome/methods comparisons
+    datalist1[[k+1]] <- point_est_df
+  } # end age category loop
+
+  # bind rows of age category estimates together
+ age_est_df <- bind_rows(datalist1)
+
+ # populate second dataframe list
+ datalist2[[i]] <- age_est_df
+
+} # end of outcome loop
+
+# combine each outcome dataframe itteration in to a big dataset
+combined_point_est_df <- bind_rows(datalist2)
+
+wrf_geo_age <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" |
+    #      pm_method == "WRF-Chem Smoke") %>% 
+  mutate(age_cat2 = ifelse(age_cat == 0, "Less than 15", 
+                    ifelse(age_cat == 1, "15-65",
+                    ifelse(age_cat == 2, "Greater than 65", NA)))) %>% 
+  # subset columns I want to put in to the table
+  select(3, 11, 5:8) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+  tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
+           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by age",
+           # row group by outcome
+           rgroup = outcome_list,
+           n.rgroup = c(rep(9, 12)), # 6 rows for each age cat for each outcome
+           # column headers
+           header = c("Est Method", "Age Group", "Events",
+                      "OR&dagger;", "Lower", "Upper"),
+           # column spanner
+           cgroup = c("", "95% CI"), 
+           n.cgroup = c(4, 2),
+           padding.rgroup = "&nbsp;&nbsp;",
+           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+           align = "llcccc", # column alignment,
+           tfoot="&dagger; Adjusted for CDC temperature, wind speed, relative humidity, precipitation, and accounting for subject.
+Time-stratified: referent periods matched to events on same day of week within July to October fire season."
+            ) # end table
+
+  print(tab)
+
+combined_point_est_df_new <- combined_point_est_df %>%
+  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
+                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
+  select(-n_events) %>%
+  rename(n_events = n_events_new)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_age_plot <- combined_point_est_df_new %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  # do not plot results with less than 15 events, create missing vals
+  # in mutate
+  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
+                             99, odds_ratio),
+         lower95 = ifelse(as.numeric(n_events) < 15, 
+                             99, lower95),
+         upper95 = ifelse(as.numeric(n_events) < 15, 
+                             99, upper95),
+         age_cat2 = ifelse(age_cat == 0, "<15", 
+                    ifelse(age_cat == 1, "15-65",
+                    ifelse(age_cat == 2, ">65", NA)))) %>% 
+  # subset columns I want to put in to the table
+  select(1, 3, 11, 10, 5:7) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
+                              levels = unique(wrf_geo_age_plot$outcome))
+
+wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
+                                levels = unique(wrf_geo_age_plot$pm_method))
+
+wrf_geo_age_plot$age_cat2 <- factor(wrf_geo_age_plot$age_cat2,
+                                levels = unique(wrf_geo_age_plot$age_cat2))
+
+wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
+wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
+wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
+wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)
+
+## ggplot 
+  print_plot <- ggplot(wrf_geo_age_plot,
+    aes(x = age_cat2, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red",  "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Age Category ') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10),
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
+  ggsave("../plot_new/age_fig_three.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+### Seasonal Time-Stratified by Sex
+
+Association with smoke PM~2.5~ and health outcomes stratified by sex.
+May be some evidence of effect modification by sex, particularly where males appear to be at greater risk for certain CVD outcomes with increasing smoke.
+However, only marginally significant.
+Also, broken arm estimate seems a little strange, suggesting females at greater risk too.
+Larger sample size may help to answer this.
+
+```{r sex strata cdc met adj, warning = F, echo = F, results='asis'} 
+# note above, results = 'asis' needed to print htmlTables
+
+# For age categories, I want a loop for each age category.
+
+# dataframe list
+df_list <- list(resp_casecross, asthma_casecross, copd_casecross, pneum_casecross,
+                acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
+                cereb_vas_casecross, hf_casecross, ihd_casecross, mi_casecross, 
+                broken_arm_casecross)
+
+outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 'Pneumonia', 
+                  'Acute Bronchitis', 'Cardiovascular Disease', 'Arrhythmia', 
+                  'Cerebrovascular Disease', 'Heart Failure',
+                  'Ischemic Heart Disease', 'Myocardial Infarction', 'Broken Arm')
+
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+
+# looking only at wrf chem, kriging, and geo-weighted; taking out 'Global Smoke' method
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'GWR Smoke')
+
+# sex category list
+sex_strata_list <- c(0,1)
+
+# create an empty list to row bind dataframes together
+datalist1 <- list()
+datalist2 <- list()
+
+# data wrangling ----
+# Producing conditional logit model estimates loop 
+for(i in 1:length(df_list)){
+
+  # dataframe to loop through
+  df_to_loop <- data.frame(df_list[i])
+  # indication of column
+  outcome <- var_list[i]
+  # outcome name
+  outcome_name <- outcome_list[i]
+
+  # extract covariates from dataframe
+  covariates_df <- df_to_loop[, c(1:26, 71, 74, 76:85)]
+
+  # extract pm values and divide by 10 and ordered
+  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+  pm_estimates_df <- df_to_loop[, c(88, 93, 92, 94)]/10  # create 10 unit increases
+
+  # new loop for sex categories
+  for(k in 0:1){
+
+    # empty matrix (12 x 10 matrix)
+    point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)
+
+    colnames(point_estimates) <- c('outcome', 'sex', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+    # fill in the outcome namedataframe before method loop
+    point_estimates[, 1] <- outcome_name
+    # repeat the sex category 4 times for each pm method
+    point_estimates[, 2] <- k
+
+    # dataframe for analysis creation
+    # bind columns back together 
+    df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+      # remove missing pm values
+      filter(!is.na(wrf_smk_pm_zip)) %>% 
+      # limit to specific sex category
+      filter(sex_ind == k) %>% 
+      # the following code makes sure that the counterfactual values retained are 
+      # symetric in that number of obs before = number of obs after
+      mutate(obs_diff_admission = (fromdate - date)/7) 
+      # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+  # second loop to run a model for each pm estimation method
+    for(j in 39:41){
+
+      # variable to model 
+      var_name <- colnames(df_analysis[j])
+
+      # set row number to fill
+      row_n <- j-38
+
+      # only run the model if the dataframe has observations
+      if(nrow(df_analysis) != 0){
+      # conditional logistic regression model
+      mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+      # populate matrix
+      point_estimates[row_n, 3] <- method_list[row_n]
+      point_estimates[row_n, 4] <- mod$n
+      point_estimates[row_n, 5] <- mod$nevent
+      # odds ratio
+      point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+      # 95% lower bound
+      point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
+                                        1.96*(summary(mod)$coefficient[1,3])), 3)
+      # 95% upper bound
+      point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                        1.96*(summary(mod)$coefficient[1,3])), 3)
+      # standard error
+      point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
+      # p val
+      point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
+
+      # create else statement that fills matrix with missing so I still have the row
+      # in the final dataframe
+      } else {point_estimates[row_n, 3] <- method_list[row_n]
+              point_estimates[row_n, 4] <- 0
+              point_estimates[row_n, 5] <- 0
+              point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
+
+    } # end methods loop
+
+    # save point estimates as a dataframe
+    point_est_df <- as_data_frame(point_estimates)
+
+    # combine previous values in dataframe that has all outcome/methods comparisons
+    datalist1[[k+1]] <- point_est_df
+  } # end sex category loop
+
+  # bind rows of sex category estimates together
+ sex_est_df <- bind_rows(datalist1)
+
+ # populate second dataframe list
+ datalist2[[i]] <- sex_est_df
+
+} # end of outcome loop
+
+# combine each outcome dataframe itteration in to a big dataset
+combined_point_est_df <- bind_rows(datalist2)
+
+wrf_geo_sex <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  mutate(sex_cat = ifelse(sex == 0, "Male", 
+               ifelse(sex == 1, "Female", NA))) %>% 
+  # subset columns I want to put in to the table
+  select(3, 11, 5:8)%>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+  tab <- htmlTable(txtRound(wrf_geo_sex, digits = 3, 1:3), 
+           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by sex",
+           # row group by outcome
+           rgroup = outcome_list,
+           n.rgroup = c(rep(6, 12)), # 2 rows for each sex strata for each outcome
+           # column headers
+           header = c("Est Method", "Sex Strata", "Events",
+                      "OR&dagger;", "Lower", "Upper"),
+           # column spanner
+           cgroup = c("", "95% CI"), 
+           n.cgroup = c(4, 2),
+           padding.rgroup = "&nbsp;&nbsp;",
+           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+           align = "llcccc", # column alignment,
+           tfoot="&dagger; Adjusted for temperature, accounting for subject. Time-stratified: referent periods matched to events on same day of week within July to October fire season."
+            ) # end table
+
+  print(tab)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_sex_plot <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | 
+   #       pm_method == "WRF-Chem Smoke") %>% 
+  mutate(sex_cat = ifelse(sex == 0, "Male", 
+               ifelse(sex == 1, "Female", NA))) %>%
+  # subset columns I want to put in to the table
+  select(1, 3, 11, 5:8) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_sex_plot$outcome <- factor(wrf_geo_sex_plot$outcome,
+                              levels = unique(wrf_geo_sex_plot$outcome))
+
+wrf_geo_sex_plot$pm_method <- factor(wrf_geo_sex_plot$pm_method,
+                                levels = unique(wrf_geo_sex_plot$pm_method))
+
+wrf_geo_sex_plot$sex_cat <- factor(wrf_geo_sex_plot$sex_cat,
+                                levels = unique(wrf_geo_sex_plot$sex_cat))
+
+wrf_geo_sex_plot$n_events <- as.numeric(wrf_geo_sex_plot$n_events)
+wrf_geo_sex_plot$odds_ratio <- as.numeric(wrf_geo_sex_plot$odds_ratio)
+wrf_geo_sex_plot$lower95 <- as.numeric(wrf_geo_sex_plot$lower95)
+wrf_geo_sex_plot$upper95 <- as.numeric(wrf_geo_sex_plot$upper95)
+
+## ggplot
+  print_plot <- ggplot(wrf_geo_sex_plot,
+    aes(x = sex_cat, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red", "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Sex') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10),
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
+  ggsave("../plot_new/sex_fig_three.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```

### 2017-05-05 r_scripts/database/wildfire_database.R
create database and put tables in it

@@ -6,31 +6,83 @@
 
 library(RSQLite)
 library(dplyr)
+library(readr)
 
 # getwd()  "/home/jyliu/wildfire/local_git_repo/oregon_wildfire/r_scripts"
 
-read_path <- paste0("../../../../../magzamen/wildfire_database/wa_2012_county_er_time_series.csv")
-wa_2012_county_ts <- read_csv(read_path)
+### Import csv files
+read_path <- paste0("../../../../../magzamen/ryan_data/wa_2012_county_er_time_series.csv")
+wa_2012_county_er_ts <- read_csv(read_path)
 
-read_path2 <- paste0("../../../../../magzamen/wildfire_database/wa_county_pop_wt_pm.csv")
+read_path2 <- paste0("../../../../../magzamen/ryan_data/wa_county_pop_wt_pm.csv")
 wa_county_pop_wt_pm <- read_csv(read_path2)
 
+read_path3 <- paste0("../../../../../magzamen/ryan_data/or_2013_county_er_time_series.csv")
+or_2013_county_er_ts <- read_csv(read_path3)
+
+read_path4 <- paste0("../../../../../magzamen/ryan_data/or_county_pop_wt_pm.csv")
+or_county_pop_wt_pm <- read_csv(read_path4)
+
+read_path5 <- paste0("../../../../../magzamen/ryan_data/colorado_county_timeseries_2010_2015.csv")
+co_2010_2015_county_ts <- read_csv(read_path5)
+
+### Create wildfire database
 wildfire_DB <- "../../../../../magzamen/wildfire_database/wildfire_db.db"
 wf_Conn <- dbConnect(drv = SQLite(), dbname= wildfire_DB)
 
-start_time <- Sys.time()
+### Write tables in database
 dbListTables(wf_Conn)
-dbWriteTable(wf_Conn, name = "health_wa_2012_county_er_time_series",wa_2012_county_ts)
-
-stop_time <-  Sys.time() - start_time 
-# time it took
-stop_time # 18.47879 mins
-
-start_time <- Sys.time()
-dbListTables(wf_Conn)
+dbWriteTable(wf_Conn, name = "health_wa_2012_county_er_ts",or_2013_county_er_ts)
 dbWriteTable(wf_Conn, name = "smoke_wa_county_pop_wt_pm",wa_county_pop_wt_pm)
[…]

### 2017-05-10 r_scripts/analysis/county_timeseries.Rmd
Create county_timeseries.Rmd

---
title: "Oregon 2013 Time-Series Analysis"
author: "Jingyang Liu"
date: "May 10, 2017"
output: html_document
---

```{r }
library(tidyverse)
library(readr)
library(lme4)
library(broom)
library(htmlTable)

# load time series dataframe
or_ts_df <- read_csv("../data_new/county_data/or_county_timeseries.csv")

# analysis dataframe
or_may_sep_df <- or_ts_df %>% 
  # create day of the week variable
  mutate(day = as.factor(weekdays(date))) %>% 
  # set missing outcome values to 0 since missing indicates no ER or urgent care 
  # visits on that date (which is reasonable in sparsely populated counties)
  mutate_each(funs(wo_miss = ifelse(is.na(.), 0, .)), n_obs:broken_arm_n) %>% 
  # create a binary smoke indicator variable based smk variables >5, >10, and >15 units 
  mutate(wrf_smk5 = ifelse(wrf_smk_pm >= 5, 1, 0), 
         wrf_smk10 = ifelse(wrf_smk_pm >= 10, 1, 0),
         wrf_smk15 = ifelse(wrf_smk_pm >= 15, 1, 0), 
         geo_smk5 = ifelse(geo_smk_pm >= 5, 1, 0),
         geo_smk10 = ifelse(geo_smk_pm >= 10, 1, 0),
         geo_smk15 = ifelse(geo_smk_pm >= 15, 1, 0),
         krig_smk5 = ifelse(krig_smk_pm >= 5, 1, 0),
         krig_smk10 = ifelse(krig_smk_pm >= 10, 1, 0),
         krig_smk15 = ifelse(krig_smk_pm >= 15, 1, 0),
         season = ifelse(date >= "2013-05-01" & date <= "2013-06-21", "spring", 
                  ifelse(date >= "2013-06-22" & date <= "2013-09-22", "summer",
                  ifelse(date >= "2013-09-23" & date <= "2013-12-21", "fall",
                                "other")))) %>% 
  # rescale pm variables to 10 units
  mutate_each(funs(unit10 = ./10), wrf_pm:krig_smk_pm)
  # select(1:2, 31:33, 36, 34:35, 38:44, 28:29, 30, 55, 58, 
  #        65, 64, 63) # subset and order variables used in analysis

# outcomes to loop through
outcome_list <- c('All ER', 'All Respiratory', 'Asthma', 'Pneumonia', 
                  'Acute Bronchitis','COPD', 'Cardiovascular Disease',
                  'Ischemic Heart Disease', 'Arrhythmia', 'Heart Failure', 
                  'Cerebrovascular Disease', 'Myocardial Infarction',
                  'Broken Arm')

# looking only at wrf chem, kriging, and geo-weighted; taking out 'Global Smoke' method
method_list <- c('Geo-Weighted Smoke', 'Kriging Smoke', 'WRF-Chem Smoke')

# create an empty list to row bind dataframes together
datalist <- list()

#glimpse(wash_jul_oct_df)
# finding columns to retain
#which(colnames(wash_jul_oct_df)=='geo_smk_pm_unit10')
#summary(wash_jul_oct_df)
# loop through outcomes
for(i in 6:18){
  i = 8
  # empty matrix for table
  point_estimates <- matrix(nrow = 3, ncol = 5, byrow = T)

  colnames(point_estimates) <- c('outcome', 'pm_method', 'risk_ratio', 
                                 'lower95', 'upper95')

  # variable to model 
  outcome_name <- outcome_list[i-5]
  # fill outcome name
  point_estimates[, 1] <- outcome_name

  # loop through pm estimation methods
  for(j in 25:27){
    j = 25
    var_name <- method_list[j-22]

    # random effects model where county is treated as a random intercept 
    mixed_mod_adj <- glmer(or_may_sep_df[[i]] ~ or_may_sep_df[[j]] + day + season +
                             wrf_temp + (1|county), or_may_sep_df, family = "poisson")

    model_df <-tidy(mixed_mod_adj)

    # populate matrix
    row_n <- j-22
    # method
    point_estimates[row_n, 2] <- var_name
    # rate ratio
    point_estimates[row_n, 3] <- round(exp(model_df[2,2]), 3)
    # 95% lower bound
    point_estimates[row_n, 4] <- round(exp((model_df[2,2]) -
                                             1.96*(model_df[2,3])), 3)
    # 95% upper bound
    point_estimates[row_n, 5] <- round(exp((model_df[2,2]) +
                                             1.96*(model_df[2,3])), 3)

    # save point estimates as a dataframe
    point_est_df <- as_data_frame(point_estimates)

  } # end methods loop

  # combine previous values in dataframe that has all outcome/methods comparisons
  datalist[[i]] <- point_est_df

} # end outcome loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist)

# subset columns I want to put in to the table
table_df <- combined_point_est_df %>% select(2:5)

tab <- htmlTable(txtRound(table_df, digits = 3, 1:3), 
                 caption = paste0("County-Level Time-Series: Association between",
                                  " a 10 ug/m^3 in PM2.5 and Health Outcomes"),
                 # row group by outcome
                 rgroup = outcome_list,
                 n.rgroup = c(rep(3, 12)), # 4 rows for each method for each outcome
                 # column headers
                 header = c("Method", "RR&dagger;", "Lower", "Upper"),
                 # column spanner
                 cgroup = c("", "95% CI"), 
                 n.cgroup = c(2, 2),
                 padding.rgroup = "&nbsp;&nbsp;",
                 css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
                 align = "llccccc", # column alignment,
                 tfoot="&dagger; Adjusted for temperature, day of the week, and season."
) # end table

print(tab)

### 2017-05-22 r_scripts/analysis/oregon_medication.Rmd
make overall, zip level age and sex categories

---
title: "Oregon 2013 Medication for Asthma"
author: "Jingyang Liu"
date: "May 17, 2017"
output: html_document
---

```{r library, include=FALSE, echo=FALSE}
library(dplyr)
library(data.table)
library(readxl) 
library(readr)
library(survival)
library(ggplot2)
library(htmlTable)
```

## Overview
For Oregon All Payer All Claims

Focus on GWR

## Data cleaning (Previous prepare work and code omitted)
I filter the observations in whole data set with only ndc (National Drug Code) exists. Also I limit the state to Oregon State.

```{r import original data, eval=FALSE, echo=FALSE}
### Import whole data set ------------------------------------------------------
read_path <- paste0("../../../data/data_original/gan_episodes_of_care.txt")
start_time <- Sys.time()
oregon_df <- fread(read_path, sep = "|", showProgress = T) 
stop_time <-  Sys.time() - start_time 
# time it took
stop_time # 7 mins

## Basic cleaning
start_time <- Sys.time()
oregon_df2 <- oregon_df %>% 
  filter((!ndc=="")&(!ndc=="*NULL*")) %>% 
  # filter State is Oregon
  filter(STATE=="OR")
stop_time <-  Sys.time() - start_time 
# time it took
stop_time # 40 secs

write_path <- paste0('../../../data/data_new/',
                     'oregon_ndc.csv')
write_csv(oregon_df2, write_path)
```

### Choose ndc
This analysis researches on asthma medication of Oregon All Claims All Payers data set.

I choose the inhalation medicine beta 2 agonists to filter the sample for asthma.

```{r ndc, eval=FALSE, echo=FALSE}
### Import NDC data set --------------------------------------------------------
read_path <- paste0("../../../data/data_new/oregon_ndc.csv")
oregon_ndc <- read_csv(read_path)
dim(oregon_ndc) # 23330714 obs

which(oregon_ndc$ndc=="23490502001")

## Import Asthma NDC table
read_path2 <- paste0("../../../data/data_new/ndc_2013_beta_2_agonists.csv")
asthma_ndc <- read_csv(read_path2, col_names = FALSE)

colnames(asthma_ndc) <- c("ndc_code", "brand_name", "generic_product_name", 
                          "route", "category", "drug_id")

# convert to vector
ndc_codes <- as.vector(as.matrix(asthma_ndc$ndc_code))

oregon_asthma <- oregon_ndc %>%
  filter(ndc %in% ndc_codes) # 232665

# limit to the unique personkey different days' first visit
oregon_asthma_df <- oregon_asthma %>% 
  group_by(personkey) %>% 
  arrange(personkey, fromdate, line) %>% 
  mutate(num_visit = dense_rank(fromdate)) %>%
  arrange(personkey, num_visit) %>% 
  select(personkey, clmid, num_visit) %>% 
  unique() %>% 
  full_join(oregon_asthma, by = c("personkey", "clmid")) %>% 
  arrange(personkey, clmid, line, fromdate) %>% 
  filter(row_number() == 1) %>%
  # add new transverted from date
  mutate(dates = as.Date(fromdate, "%m/%d/%Y")) %>%
  filter(dates >= '2013-05-01' & 
         dates <= '2013-09-30') # 24538

write_path2 <- paste0("../../../data/data_new/oregon_asthma_ndc.csv")
write_csv(oregon_asthma_df, write_path2)

```

## Casecrossover study

I also make the "false" dates for each claim.

```{r casecrossover, eval=FALSE, echo=FALSE}
### Casecrossover study
read_path3 <- paste0("../../../data/data_new/medication/oregon_asthma_ndc.csv")
oregon_asthma_df <- read_csv(read_path3)

outcome_id <- oregon_asthma_df %>%
  # arrange with dates
  arrange(dates) %>%
  mutate(id = seq(1, nrow(.), by = 1))

# create dataset to populate
id_date_df <- data_frame()

# begin second loop to create counterfactual observations for each case subject
for (k in 1:nrow(outcome_id)){

  # find the replicate times of weeks
  dates_l <- outcome_id[[k,74]] 
  n1 <- 0
  d=as.Date("2013-05-01")
  i=1
  while (dates_l >= "2013-05-01"){
    dates_l <- dates_l - 7
    d[i] = dates_l
    i = i+1
    n1 = n1+1
  }
  d[1:n1-1] # shows character(0) when the first week
  n1-1

  dates_l <- outcome_id[[k,74]] 
  n2=0
  e=as.Date("2013-09-30")
  j=1
  while (dates_l <= "2013-09-30"){
    dates_l <- dates_l + 7
    e[j]=dates_l
    j=j+1
    n2 = n2 + 1
  }
  e[1:n2-1] # shows character(0) when the last week
  n2-1

  # replicate covariates length of counterfactual dates
  # and make conuterfactual dates
  if (n1==1){
    cov_df <- do.call("bind_rows", replicate(n1+n2-1, outcome_id[k,],simplify = F))
    cov_df$dates <- c(outcome_id[[k,74]], e[1:n2-1])
  } else if (n2==1){
    cov_df <- do.call("bind_rows", replicate(n1+n2-1, outcome_id[k,],simplify = F))
    cov_df$dates <- c(outcome_id[[k,74]], d[1:(n1-1)])
  }else{
    cov_df <- do.call("bind_rows", replicate(n1+n2-1, outcome_id[k,],simplify = F))
    cov_df$dates <- c(outcome_id[[k,74]], d[1:(n1-1)], e[1:n2-1])
  }

  # bind unique id and date of the year with covariates
  id_date <- bind_cols(cov_df)
  # iteration which binds rows of unique ids
  id_date_df <- bind_rows(id_date_df, id_date)
}

outcome_casecross <- id_date_df %>%
  mutate(outcome = ifelse(dates == as.Date(fromdate, "%m/%d/%Y"), 1, 0)) %>%
  arrange(id, dates) # order by id and date

outcome_casecross <- id_date_df %>%
  mutate(outcome = ifelse(dates == as.Date(fromdate, "%m/%d/%Y"), 1, 0)) %>%
  arrange(personkey, dates) # order by id and date

write_path2 <- paste0("../../../data/data_new/oregon_ndc_casecrossover.csv")
write_csv(outcome_casecross, write_path2)

```

## Time-Stratified Case-Crossover (referents same day of week within fire season May - September)

** The following analysis are done in local R.**

```{r time stratified, eval=FALSE, echo=FALSE}
### Time stratified ------------------------------------------------------------
# read in zipcode level populatoin-weighted pm
read_path5 <- paste0('C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire_new/data/Oregon_PM/zip_pm_to_merge_with_acap.csv')

zip_smoke <- read_csv(read_path5) # 63801 rows

# descriptives of the two smoke datasets
summary(zip_smoke)

# Zipcode PM2.5 estimates
# create lag variables that take smoke values from n previous days for zipcodes
zip_smoke_w_lag <- zip_smoke %>% arrange(ZIPCODE, date) %>%
  # group by zipcode
  group_by(ZIPCODE) %>% 
  # wrf
  mutate(wrf_f_pm_lag1 = lag(wrf_f_pm, 1, order_by = ZIPCODE), 
         wrf_f_pm_lag2 = lag(wrf_f_pm, 2, order_by = ZIPCODE),
         wrf_f_pm_lag3 = lag(wrf_f_pm, 3, order_by = ZIPCODE),
         wrf_f_pm_lag4 = lag(wrf_f_pm, 4, order_by = ZIPCODE),
         wrf_f_pm_lag5 = lag(wrf_f_pm, 5, order_by = ZIPCODE),
         # wrf no fire lag
         wrf_nf_pm_lag1 = lag(wrf_nf_pm, 1, order_by = ZIPCODE),
         wrf_nf_pm_lag2 = lag(wrf_nf_pm, 2, order_by = ZIPCODE),
         wrf_nf_pm_lag3 = lag(wrf_nf_pm, 3, order_by = ZIPCODE),
         wrf_nf_pm_lag4 = lag(wrf_nf_pm, 4, order_by = ZIPCODE),
         wrf_nf_pm_lag5 = lag(wrf_nf_pm, 5, order_by = ZIPCODE),
         # wrf_smk_pm
         wrf_smk_pm_lag1 = lag(wrf_smk_pm, 1, order_by = ZIPCODE),
         wrf_smk_pm_lag2 = lag(wrf_smk_pm, 2, order_by = ZIPCODE),
         wrf_smk_pm_lag3 = lag(wrf_smk_pm, 3, order_by = ZIPCODE),
         wrf_smk_pm_lag4 = lag(wrf_smk_pm, 4, order_by = ZIPCODE),
         wrf_smk_pm_lag5 = lag(wrf_smk_pm, 5, order_by = ZIPCODE),
         # geo weighted pm
         geo_wt_pm_lag1 = lag(geo_wt_pm, 1, order_by = ZIPCODE),
         geo_wt_pm_lag2 = lag(geo_wt_pm, 2, order_by = ZIPCODE),
         geo_wt_pm_lag3 = lag(geo_wt_pm, 3, order_by = ZIPCODE),
         geo_wt_pm_lag4 = lag(geo_wt_pm, 4, order_by = ZIPCODE),
         geo_wt_pm_lag5 = lag(geo_wt_pm, 5, order_by = ZIPCODE),
         # krig pm
         krig_pm_lag1 = lag(krig_pm, 1, order_by = ZIPCODE),
         krig_pm_lag2 = lag(krig_pm, 2, order_by = ZIPCODE),
         krig_pm_lag3 = lag(krig_pm, 3, order_by = ZIPCODE),
         krig_pm_lag4 = lag(krig_pm, 4, order_by = ZIPCODE),
         krig_pm_lag5 = lag(krig_pm, 5, order_by = ZIPCODE),   
         # background pm
         background_pm_lag1 = lag(background_pm, 1, order_by = ZIPCODE),
         background_pm_lag2 = lag(background_pm, 2, order_by = ZIPCODE),
         background_pm_lag3 = lag(background_pm, 3, order_by = ZIPCODE),
         background_pm_lag4 = lag(background_pm, 4, order_by = ZIPCODE),
         background_pm_lag5 = lag(background_pm, 5, order_by = ZIPCODE),   
         # geo_smk_pm 
         geo_smk_pm_lag1 = lag(geo_smk_pm, 1, order_by = ZIPCODE),
         geo_smk_pm_lag2 = lag(geo_smk_pm, 2, order_by = ZIPCODE),
         geo_smk_pm_lag3 = lag(geo_smk_pm, 3, order_by = ZIPCODE),
         geo_smk_pm_lag4 = lag(geo_smk_pm, 4, order_by = ZIPCODE),
         geo_smk_pm_lag5 = lag(geo_smk_pm, 5, order_by = ZIPCODE),
         # krig smk pm
         krig_smk_pm_lag1 = lag(krig_smk_pm, 1, order_by = ZIPCODE),
         krig_smk_pm_lag2 = lag(krig_smk_pm, 2, order_by = ZIPCODE),
         krig_smk_pm_lag3 = lag(krig_smk_pm, 3, order_by = ZIPCODE),
         krig_smk_pm_lag4 = lag(krig_smk_pm, 4, order_by = ZIPCODE),
         krig_smk_pm_lag5 = lag(krig_smk_pm, 5, order_by = ZIPCODE),
         # temp
         wrf_temp_lag1 = lag(wrf_temp, 1, order_by = ZIPCODE),
         wrf_temp_lag2 = lag(wrf_temp, 2, order_by = ZIPCODE),
         wrf_temp_lag3 = lag(wrf_temp, 3, order_by = ZIPCODE),
         wrf_temp_lag4 = lag(wrf_temp, 4, order_by = ZIPCODE),
         wrf_temp_lag5 = lag(wrf_temp, 5, order_by = ZIPCODE)) %>% 
  # ungroup by zip
  ungroup(ZIPCODE) %>% 
  # attach a zip indicator for each smoke variable
  setNames(paste(colnames(.), "zip", sep="_")) %>% 
  # remove the '_zip' from the zipcode and date variable 
  rename(ZIPCODE = ZIPCODE_zip, date = date_zip)

read_path4 <- paste0("../data_new/medication/oregon_ndc_casecrossover.csv")
# read_path4 <- paste0("../../../data/data_new/oregon_ndc_casecrossover.csv") # for server
or_disease <- read_csv(read_path4)

### try join
colnames(or_disease)[24] <- c("ZIPCODE")
colnames(or_disease)[74] <- c("date")

outcome_casecross <- or_disease %>%
  # indicator for male=0, female=1, unknown = 2
  mutate(age = 2013-yob,
         sex_ind =ifelse(gender == "F", 1, 
                         ifelse(gender == "M", 0, 2)),
         age_ind = ifelse(age < 15, 0,
                          ifelse(age >= 15 & age < 65, 1,
                                 ifelse(age >= 65 & age <=105, 2, NA)))
  ) %>% # end of mutate 
  # create variables
  mutate(day = as.factor(weekdays(fromdate)),
         day_admit = as.factor(weekdays(date)),
         month_smk = month(fromdate),
         month_admit = month(date),
         season_smk = ifelse(fromdate >= "2013-03-20" &  
                               fromdate <= "2013-06-21", "spring",
                             ifelse(fromdate >= "2013-06-22" &  
                                      fromdate <= "2013-09-22", "summer",
                                    ifelse(fromdate >= "2013-09-23" & 
                                             fromdate <= "2013-12-21", "fall", "other"))),
         season_admit = ifelse(date >= "2013-03-20" &  
                                 date <= "2013-06-21", "spring",
                               ifelse(date >= "2013-06-22" &  
                                        date <= "2013-09-22", "summer",
                                      ifelse(date >= "2013-09-23" & 
                                               date <= "2013-12-21", "fall", "other")))) %>%
  # join with zip-level pm estimates
  left_join(zip_smoke_w_lag, by = c("date", "ZIPCODE"))%>%
  arrange(personkey, fromdate) # order by id and fromdate

write_path4 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover.csv")
write_csv(outcome_casecross, write_path4)

```

## Population-Weighted PM~2.5~ Assigned by Zip Code

### Descriptives

For all of 2013, there were a total of millons of all claims records in the APAC dataset. During the May 1st to September 30th wildfire season in Oregon state, there were 24217 pharmacy visits for the beta 2 agonists with ndc. The number of persons (cases), as well as age-specific and sex-specific strate.

```{r descriptive table, echo = F, message = F, warning = F, results='asis'}
read_path6 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover.csv")
asthma_ndc <- read_csv(read_path6)

# dataframe list
method_list <- c('WRF-Chem Smoke', 'Geo-Weighted Smoke', 'Kriging Smoke')

# dataframe to loop through
asthma_ndc_df <- data.frame(asthma_ndc)

# outcome name
outcome_name <- "asthma"

# dataframe for analysis creation
# bind columns back together 
df_analysis <- asthma_ndc_df %>%
  filter(!is.na(wrf_smk_pm_zip)) %>% 
  # only look at outcomes
  filter(outcome == 1) %>%
  # add another row that makes sure there is a person <15 in the dataframe
  # tricking xtabs to produce a 0 cell for the outcome for age <15
  add_row(outcome = 0, age_ind = 0)

# cross tabs
outcome_n <- xtabs(~ outcome, df_analysis)
cross_tab_age <- xtabs(~ outcome + age_ind, df_analysis)
cross_tab_sex <- xtabs(~ outcome + sex_ind, df_analysis)
# empty matrix
point_estimates <- matrix(nrow = 1, ncol = 7, byrow = T)

colnames(point_estimates) <- c("outcome", "n", "age_15", "age_15_65", 
                               "age_65", "female", "male")

# fill in the outcome name for the dataframe before the loop
point_estimates[, 1] <- outcome_name
# fill n
point_estimates[, 2] <- outcome_n[2] # second element of the 1 dimension vector
# age <15
point_estimates[, 3] <- cross_tab_age[2, 1]
# age 15 to 65
point_estimates[, 4] <- cross_tab_age[2, 2]
# age >65
point_estimates[, 5] <- cross_tab_age[2, 3]
# male == 0
point_estimates[, 7] <- cross_tab_sex[1, 1]
# female == 1
point_estimates[, 6] <- cross_tab_sex[1, 2]

# save point estimates as a dataframe
point_est_df <- as_data_frame(point_estimates)

# combine each outcome dataframe itteration in to a big dataset
asthma_point_est_df <- point_est_df %>% 
  # find proportions/percents for each strata in a row
  mutate(age_15_pr = as.character(round((as.numeric(age_15)/as.numeric(n))*100,1)),
         age_15_65_pr = as.character(round((as.numeric(age_15_65)/as.numeric(n))*100,1)),
         age_65_pr = as.character(round((as.numeric(age_65)/as.numeric(n))*100,1)),
         female_pr = as.character(round((as.numeric(female)/as.numeric(n))*100,1)),
         male_pr = as.character(round((as.numeric(male)/as.numeric(n))*100,1))) %>% 
  select(outcome, n, age_15, age_15_pr, age_15_65, age_15_65_pr, age_65,
         age_65_pr, female, female_pr, male, male_pr)

# str(asthma_point_est_df)

tab <- htmlTable(txtRound(asthma_point_est_df, digits = 1), 
           caption = "Number of cases for each outcome observed from May 1st to September 30st, 2013",
           # column headers
           header = c("Outcome", "Cases n", "Less than 15", "(%)", "15 to 65", "(%)", "Greater than 65",
                      "(%)", "Female", "(%)", "Male", "(%)"),
           # column spanner
           cgroup = c("","Age Category", "Sex"), 
           n.cgroup = c(2, 6, 4),
           padding.rgroup = "&nbsp;&nbsp;",
           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
           align = "llccccc" # column alignment,
            ) # end table

print(tab)

```

### Overall Age Categories

```{r time stratified age strata cdc met, warning =F, echo = F, results='asis'} 
# dataframe list

method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')

# dataframe to loop through
asthma_ndc_df <- data.frame(asthma_ndc)

# outcome name
outcome <- "Asthma Inhalers"
outcome_name <- "Asthma Inhalers"

# data wrangling ----
# Producing conditional logit model estimates loop

# extract covariates from dataframe
covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]

# extract pm values and divide by 10 and ordered
#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
pm_estimates_df <- asthma_ndc_df[, c(87, 92, 91, 93)]/10  # create 10 unit increases

# new loop for age categories

# empty matrix (12 x 10 matrix)
point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)

colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 
                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')

# fill in the outcome namedataframe before method loop
point_estimates[, 1] <- outcome_name

# dataframe for analysis creation
# bind columns back together 
df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
  # remove missing pm values
  filter(!is.na(wrf_smk_pm_zip)) %>% 
  # the following code makes sure that the counterfactual values retained are 
  # symetric in that number of obs before = number of obs after
  mutate(obs_diff_admission = (fromdate - date)/7) 
  # dataframe is already for the entire fire season, so I don't need to subset anymore

# second loop to run a model for each pm estimation method
for(j in 38:40){

  # variable to model 
  var_name <- colnames(df_analysis[j])

  # set row number to fill
  row_n <- j-37

  # only run the model if the dataframe has observations
  if(nrow(df_analysis) != 0){
  # conditional logistic regression model
  mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)

  # populate matrix
  point_estimates[row_n, 2] <- method_list[row_n]
  point_estimates[row_n, 3] <- mod$n
  point_estimates[row_n, 4] <- mod$nevent
  # odds ratio
  point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)

  # 95% lower bound
  point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
  # 95% upper bound
  point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
                                        1.96*(summary(mod)$coefficient[1,3])), 3)
  # standard error
  point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
  # p val
  point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)

  # create else statement that fills matrix with missing so I still have the row
  # in the final dataframe
  } else {point_estimates[row_n, 2] <- method_list[row_n]
          point_estimates[row_n, 3] <- 0
          point_estimates[row_n, 4] <- 0
          point_estimates[row_n, c(5:9)] <- 99 } # end 'if else' statement
  } # end methods loop

# save point estimates as a dataframe
combined_point_est_df <- as_data_frame(point_estimates)

wrf_geo_age <- combined_point_est_df %>% 
  # filter columns to just geo-smk
  # filter(pm_method == "GWR Smoke" |
    #      pm_method == "WRF-Chem Smoke") %>% 
  # subset columns I want to put in to the table
  select(2, 4:7) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
         lower95 = ifelse(lower95 == 99, "--", lower95),
         upper95 = ifelse(upper95 == 99, "--", upper95))

tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
         caption = "Association between a 10 ug/m^3 in PM2.5 three smoke method and health outcomes",
         # row group by outcome
         rgroup = "Asthma",
         n.rgroup = c(rep(3, 1)), # 6 rows for each age cat for each outcome
         # column headers
         header = c("Est Method", "Events",
                    "OR&dagger;", "Lower", "Upper"),
         # column spanner
         cgroup = c("", "95% CI"), 
         n.cgroup = c(3, 2),
         padding.rgroup = "&nbsp;&nbsp;",
         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
         align = "llcccc", # column alignment,
         tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within May to September fire season."
         ) # end table

print(tab)

combined_point_est_df_new <- combined_point_est_df %>%
  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
  select(-n_events) %>%
  rename(n_events = n_events_new)

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
wrf_geo_age_plot <- combined_point_est_df_new %>% 
  # filter columns to just geo-smk
 #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
  # do not plot results with less than 15 events, create missing vals
  # in mutate
  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
                             99, odds_ratio),
         lower95 = ifelse(as.numeric(n_events) < 15, 
                             99, lower95),
         upper95 = ifelse(as.numeric(n_events) < 15, 
                             99, upper95)) %>% 
  # subset columns I want to put in to the table
  select(1, 2, 9, 4:6) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
         lower95 = ifelse(lower95 == 99, NA, lower95),
         upper95 = ifelse(upper95 == 99, NA, upper95))

# change characters to numeric and factor  
wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
                              levels = unique(wrf_geo_age_plot$outcome))

wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
                                levels = unique(wrf_geo_age_plot$pm_method))

wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)

## ggplot 
  print_plot <- ggplot(wrf_geo_age_plot,
    aes(x = pm_method, y = odds_ratio, color = pm_method), na.rm = T) +
    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
    # custom color 
    scale_color_manual(name = "Smoke-Estimation Method", 
                       values = c("red",  "blue", "#32115C"),
                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
    facet_wrap(~outcome, nrow = 3, scales = "free") +
    geom_hline(yintercept = 1, linetype=2) +
    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
    #ylim(0, 2) +
    xlab('Smoke Estimation Method') +
    # plot theme
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # strip element
    strip.background = element_rect(colour=NA, fill=NA),
    panel.border = element_rect(fill = NA, color = "black"),
    # facet text size
    strip.text = element_text(size = 10),
    # axis element
    #axis.text.x = element_blank(),
    #axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(angle = 90),
    # legend elements
    legend.position = "bottom")
    #legend.text = element_text(size = 8))

  print(print_plot)
  # save figure
  ggsave("../plot_new/ndc_age_fig_three.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

```

### Time-Stratified by Age Categories

Following table and figure look at some outcomes stratified by age category.

For some outcomes, like Age > 65 in WRF-Chem method, the CI is the only insignificant one in the plot.

```{r time stratified age strata cdc met, warning =F, echo = F, results='asis'} 
# dataframe list

method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')

# dataframe to loop through
asthma_ndc_df <- data.frame(asthma_ndc)

# outcome name
outcome <- "asthma"
outcome_name <- "Asthma"

# age category list
age_cat_list <- c(0,1,2)

# create an empty list to row bind dataframes together
datalist1 <- list()

# data wrangling ----
# Producing conditional logit model estimates loop

# extract covariates from dataframe
covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]

# extract pm values and divide by 10 and ordered
#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
pm_estimates_df <- asthma_ndc_df[, c(87, 92, 91, 93)]/10  # create 10 unit increases

# new loop for age categories
for(k in 0:2){
  # empty matrix (12 x 10 matrix)
  point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)

  colnames(point_estimates) <- c('outcome', 'age_cat', 'pm_method', 'n', 'n_events', 
                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')

  # fill in the outcome namedataframe before method loop
  point_estimates[, 1] <- outcome_name
  # repeat the age category 4 times for each pm method
  point_estimates[, 2] <- k

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
    # remove missing pm values
    filter(!is.na(wrf_smk_pm_zip)) %>% 
    # limit to specific age category
    filter(age_ind == k) %>% 
    # the following code makes sure that the counterfactual values retained are 
    # symetric in that number of obs before = number of obs after
    mutate(obs_diff_admission = (fromdate - date)/7) 
    # dataframe is already for the entire fire season, so I don't need to subset anymore

  # second loop to run a model for each pm estimation method
  for(j in 38:40){

    # variable to model 
    var_name <- colnames(df_analysis[j])

    # set row number to fill
    row_n <- j-37

    # only run the model if the dataframe has observations
    if(nrow(df_analysis) != 0){
    # conditional logistic regression model
    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)

    # populate matrix
    point_estimates[row_n, 3] <- method_list[row_n]
    point_estimates[row_n, 4] <- mod$n
    point_estimates[row_n, 5] <- mod$nevent
    # odds ratio
    point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)

    # 95% lower bound
    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
    # 95% upper bound
    point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
                                        1.96*(summary(mod)$coefficient[1,3])), 3)
    # standard error
    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
    # p val
    point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)

    # create else statement that fills matrix with missing so I still have the row
    # in the final dataframe
    } else {point_estimates[row_n, 3] <- method_list[row_n]
            point_estimates[row_n, 4] <- 0
            point_estimates[row_n, 5] <- 0
            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
    } # end methods loop

  # save point estimates as a dataframe
  point_est_df <- as_data_frame(point_estimates)
  datalist1[[k+1]] <- point_est_df
} # end age category loop

# bind rows of age category estimates together
combined_point_est_df <- bind_rows(datalist1)

wrf_geo_age <- combined_point_est_df %>% 
  # filter columns to just geo-smk
  # filter(pm_method == "GWR Smoke" |
    #      pm_method == "WRF-Chem Smoke") %>% 
  mutate(age_cat2 = ifelse(age_cat == 0, "Less than 15", 
                    ifelse(age_cat == 1, "15-65",
                    ifelse(age_cat == 2, "Greater than 65", NA)))) %>% 
  # subset columns I want to put in to the table
  select(3, 11, 5:8) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
         lower95 = ifelse(lower95 == 99, "--", lower95),
         upper95 = ifelse(upper95 == 99, "--", upper95))

tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
         caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by age",
         # row group by outcome
         rgroup = "Asthma",
         n.rgroup = c(rep(9, 1)), # 6 rows for each age cat for each outcome
         # column headers
         header = c("Est Method", "Age Group", "Events",
                    "OR&dagger;", "Lower", "Upper"),
         # column spanner
         cgroup = c("", "95% CI"), 
         n.cgroup = c(4, 2),
         padding.rgroup = "&nbsp;&nbsp;",
         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
         align = "llcccc", # column alignment,
         tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within May to September fire season."
         ) # end table

print(tab)

combined_point_est_df_new <- combined_point_est_df %>%
  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
  select(-n_events) %>%
  rename(n_events = n_events_new)

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
wrf_geo_age_plot <- combined_point_est_df_new %>% 
  # filter columns to just geo-smk
 #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
  # do not plot results with less than 15 events, create missing vals
  # in mutate
  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
                             99, odds_ratio),
         lower95 = ifelse(as.numeric(n_events) < 15, 
                             99, lower95),
         upper95 = ifelse(as.numeric(n_events) < 15, 
                             99, upper95),
         age_cat2 = ifelse(age_cat == 0, "<15", 
                    ifelse(age_cat == 1, "15-65",
                    ifelse(age_cat == 2, ">65", NA)))) %>% 
  # subset columns I want to put in to the table
  select(1, 3, 11, 10, 5:7) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
         lower95 = ifelse(lower95 == 99, NA, lower95),
         upper95 = ifelse(upper95 == 99, NA, upper95))

# change characters to numeric and factor  
wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
                              levels = unique(wrf_geo_age_plot$outcome))

wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
                                levels = unique(wrf_geo_age_plot$pm_method))

wrf_geo_age_plot$age_cat2 <- factor(wrf_geo_age_plot$age_cat2,
                                levels = unique(wrf_geo_age_plot$age_cat2))

wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)

## ggplot 
  print_plot <- ggplot(wrf_geo_age_plot,
    aes(x = age_cat2, y = odds_ratio, color = pm_method), na.rm = T) +
    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
    # custom color 
    scale_color_manual(name = "Smoke-Estimation Method", 
                       values = c("red",  "blue", "#32115C"),
                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
    facet_wrap(~outcome, nrow = 3, scales = "free") +
    geom_hline(yintercept = 1, linetype=2) +
    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
    #ylim(0, 2) +
    xlab('Age Category ') +
    # plot theme
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # strip element
    strip.background = element_rect(colour=NA, fill=NA),
    panel.border = element_rect(fill = NA, color = "black"),
    # facet text size
    strip.text = element_text(size = 10),
    # axis element
    #axis.text.x = element_blank(),
    #axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(angle = 90),
    # legend elements
    legend.position = "bottom")
    #legend.text = element_text(size = 8))

  print(print_plot)
  # save figure
  ggsave("../plot_new/ndc_age_fig_three.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

```

### Seasonal Time-Stratified by Sex

Association with smoke PM~2.5~ and health outcomes stratified by sex.

All the CI do not include 1, which means the use of asthma medication (beta 2 agonists) increases about 5% when smoke increase 10 units. So the risk for getting asthma will increase about 5% ~ 8% when the smoke PM2.5 increase 10 units.

```{r sex strata cdc met adj, warning = F, echo = F, results='asis'} 
method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')

# dataframe to loop through
asthma_ndc_df <- data.frame(asthma_ndc)

# outcome name
outcome <- "asthma"
outcome_name <- "Asthma"

# create an empty list to row bind dataframes together
datalist1 <- list()

# sex category list
sex_strata_list <- c(0,1)

# data wrangling ----
# Producing conditional logit model estimates loop

# extract covariates from dataframe
covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]

# extract pm values and divide by 10 and ordered
#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
pm_estimates_df <- asthma_ndc_df[, c(87, 92, 91, 93)]/10  # create 10 unit increases

# new loop for sex categories
for(k in 0:1){

  # empty matrix (12 x 10 matrix)
  point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)

  colnames(point_estimates) <- c('outcome', 'sex', 'pm_method', 'n', 'n_events', 
                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')

  # fill in the outcome namedataframe before method loop
  point_estimates[, 1] <- outcome_name
  # repeat the sex category 4 times for each pm method
  point_estimates[, 2] <- k

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
    # remove missing pm values
    filter(!is.na(wrf_smk_pm_zip)) %>% 
    # limit to specific sex category
    filter(sex_ind == k) %>% 
    # the following code makes sure that the counterfactual values retained are 
    # symetric in that number of obs before = number of obs after
    mutate(obs_diff_admission = (fromdate - date)/7) 
    # dataframe is already for the entire fire season, so I don't need to subset anymore

  # second loop to run a model for each pm estimation method
  for(j in 38:40){

    # variable to model 
    var_name <- colnames(df_analysis[j])

    # set row number to fill
    row_n <- j-37

    # only run the model if the dataframe has observations
    if(nrow(df_analysis) != 0){
    # conditional logistic regression model
    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)

    # populate matrix
    point_estimates[row_n, 3] <- method_list[row_n]
    point_estimates[row_n, 4] <- mod$n
    point_estimates[row_n, 5] <- mod$nevent
    # odds ratio
    point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)

    # 95% lower bound
    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
                                      1.96*(summary(mod)$coefficient[1,3])), 3)
    # 95% upper bound
    point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
                                      1.96*(summary(mod)$coefficient[1,3])), 3)
    # standard error
    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
    # p val
    point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)

    # create else statement that fills matrix with missing so I still have the row
    # in the final dataframe
    } else {point_estimates[row_n, 3] <- method_list[row_n]
            point_estimates[row_n, 4] <- 0
            point_estimates[row_n, 5] <- 0
            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
    } # end methods loop

  # save point estimates as a dataframe
  point_est_df <- as_data_frame(point_estimates)

  # combine previous values in dataframe that has all outcome/methods comparisons
  datalist1[[k+1]] <- point_est_df
} # end sex category loop

# bind rows of sex category estimates together
combined_point_est_df <- bind_rows(datalist1)

wrf_geo_sex <- combined_point_est_df %>% 
  # filter columns to just geo-smk
  # filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
  mutate(sex_cat = ifelse(sex == 0, "Male", 
               ifelse(sex == 1, "Female", NA))) %>% 
  # subset columns I want to put in to the table
  select(3, 11, 5:8)%>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
         lower95 = ifelse(lower95 == 99, "--", lower95),
         upper95 = ifelse(upper95 == 99, "--", upper95))

  tab <- htmlTable(txtRound(wrf_geo_sex, digits = 3, 1:3), 
           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by sex",
           # row group by outcome
           rgroup = "Asthma",
           n.rgroup = c(rep(6, 1)), # 2 rows for each sex strata for each outcome
           # column headers
           header = c("Est Method", "Sex Strata", "Events",
                      "OR&dagger;", "Lower", "Upper"),
           # column spanner
           cgroup = c("", "95% CI"), 
           n.cgroup = c(4, 2),
           padding.rgroup = "&nbsp;&nbsp;",
           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
           align = "llcccc", # column alignment,
           tfoot="&dagger; Adjusted for temperature, accounting for subject. Time-stratified: referent periods matched to events on same day of week within July to October fire season."
            ) # end table

  print(tab)

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
wrf_geo_sex_plot <- combined_point_est_df %>% 
  # filter columns to just geo-smk
 #  filter(pm_method == "GWR Smoke" | 
   #       pm_method == "WRF-Chem Smoke") %>% 
  mutate(sex_cat = ifelse(sex == 0, "Male", 
               ifelse(sex == 1, "Female", NA))) %>%
  # subset columns I want to put in to the table
  select(1, 3, 11, 5:8) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
         lower95 = ifelse(lower95 == 99, NA, lower95),
         upper95 = ifelse(upper95 == 99, NA, upper95))

# change characters to numeric and factor  
wrf_geo_sex_plot$outcome <- factor(wrf_geo_sex_plot$outcome,
                              levels = unique(wrf_geo_sex_plot$outcome))

wrf_geo_sex_plot$pm_method <- factor(wrf_geo_sex_plot$pm_method,
                                levels = unique(wrf_geo_sex_plot$pm_method))

wrf_geo_sex_plot$sex_cat <- factor(wrf_geo_sex_plot$sex_cat,
                                levels = unique(wrf_geo_sex_plot$sex_cat))

wrf_geo_sex_plot$n_events <- as.numeric(wrf_geo_sex_plot$n_events)
wrf_geo_sex_plot$odds_ratio <- as.numeric(wrf_geo_sex_plot$odds_ratio)
wrf_geo_sex_plot$lower95 <- as.numeric(wrf_geo_sex_plot$lower95)
wrf_geo_sex_plot$upper95 <- as.numeric(wrf_geo_sex_plot$upper95)

## ggplot
  print_plot <- ggplot(wrf_geo_sex_plot,
    aes(x = sex_cat, y = odds_ratio, color = pm_method), na.rm = T) +
    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
    # custom color 
    scale_color_manual(name = "Smoke-Estimation Method", 
                       values = c("red", "blue", "#32115C"),
                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
    facet_wrap(~outcome, nrow = 3, scales = "free") +
    geom_hline(yintercept = 1, linetype=2) +
    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
    #ylim(0, 2) +
    xlab('Sex') +
    # plot theme
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # strip element
    strip.background = element_rect(colour=NA, fill=NA),
    panel.border = element_rect(fill = NA, color = "black"),
    # facet text size
    strip.text = element_text(size = 10),
    # axis element
    #axis.text.x = element_blank(),
    #axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(angle = 90),
    # legend elements
    legend.position = "bottom")
    #legend.text = element_text(size = 8))

  print(print_plot)
  # save figure
  ggsave("../plot_new/ndc_sex_fig_three.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

```

## County

### Time series

### 2017-05-25 r_scripts/analysis/oregon_medication.Rmd
update some change of time series and compiling

@@
 ## Overview
-For Oregon All Payer All Claims
-
-Focus on GWR
-
-## Data cleaning (Previous prepare work and code omitted)
-I filter the observations in whole data set with only ndc (National Drug Code) exists. Also I limit the state to Oregon State.
+For this research I have the Oregon All Payer All Claims (APAC) data set in 2013.
+I want to research on the effects of wildfire smoke on pharmacy health outcome of asthma.
+NDC (National Drug Code) is the code that the data set offered.
+The standard NDC I use the link below, which is a ndc list in 2013, which fits my original data set.
+(http://ww2.ncqa.org/hedis-quality-measurement/hedis-measures/hedis-2013/hedis-2013-final-ndc-lists)
+
+The data cleaning and casecrossover study are done with server, the following are done with local R.
+
+## Preparation
+1. Data cleaning
+
+There are about 77,000,000 claims in total data set.
+Some have the NDC but another are null or missing.
+So I choose the claims in Oregon State with ndc, which is 23,330,714 claims.
+Also, the POS (Place of service code) all show "pharmacy" type.
+
+For ndc, I choose the people who use the beta 2 agonists (one inhaler) for the day's first visit to form the sample.
+Because the ndc csv file has more information than what I need, I filter the ndc file by SQLite Studio (code and steps not attached) to a file only containing the beta 2 agonists medication.
+
+After choosing the claims with beta 2 agonists, the data set is reduced to 232,665. After choosing one day's first visit, the data set is the final data set we research on, containing 24,538 claims.
+
+2. Smoke model
+
+In this research, we foocus on GWR method (geographically weighted ridge regression).
+There are also WRF-Chem model (Weather Research and Forecasting with Chemistry) and Krig model.
+To reduce the occurance error I will compare all three models.
 
@@
 
-### Overall Age Categories
-
-```{r time stratified age strata cdc met, warning =F, echo = F, results='asis'} 
+### Overall plot
+
+```{r overall, warning =F, echo = F, results='asis'} 
 # dataframe list
@@
 ### Time series
+
+The time series for the number of pharmacy in each day (and separated with different counties).
+The overall time series plot shows a
+
+!!! In claim data I make a proportion table of county and zip code because we need to make shapefiles. In this pharmacy I choose a different data set
+
+```{r time series, warning=FALSE, message=FALSE}
+read_path7 <- paste0("../data_new/medication/oregon_asthma_ndc.csv")
+or_ndc_claim <- read_csv(read_path7)
+
+or_ndc_asthma <- or_ndc_claim %>%
+  select(personkey, num_visit, clmid, line, ZIP, dates)
+
+# Join data with names of Oregon counties 
+oregon_fips <- read_csv('../instructions/oregon_FIPS.csv')
+# remove the "County" character
+# factor(oregon_fips$county)
+oregon_fips$county <- gsub(" County", "", as.character(factor(oregon_fips$county)))
+
+oregon_fips <- oregon_fips %>%
+  mutate(st_county_fips = with(oregon_fips, paste0(st_code, fips)))
+
+### Join county with zip
+read_path <- paste0('../data_new/update/or_zip_county_prop.csv') 
+or_zip_county <- read_csv(read_path)
+
+or_zip_county <- or_zip_county %>%
+  select(zip, county_name) %>%
+  rename(county = county_name) %>%
+  rename(ZIP = zip) %>%
+  full_join(oregon_fips, by = "county") %>%
+  select(ZIP, county, state, st_county_fips) %>%
+  rename(fips = st_county_fips)
+
+or_zip_county$county[which(or_zip_county$county=="Hood.River")] <- "Hood River"
+
+# Import population weighted county PM2.5
+pm_path <- paste0("../data_new/county_data/or_county_pop_wt_pm.csv")
+county_pm <- read_csv(pm_path)
+
+# create a working dataset 'or_2013'
+or_ndc_county_time_series <- or_ndc_asthma %>% 
+  # join in the washington county name based on CHARS county identifier
+  left_join(or_zip_county, by = "ZIP") %>% 
+  # rename 'admit_date_impute' to 'date'
+  rename(date = dates) %>%
+  # rename(date = admit_date_impute) %>% 
+  # group by county of residence and date
+  group_by(county, date) %>% 
+  # sum up each primary diagnosis for each outcome for each day for each county
+  summarise(n_obs = n()) %>% 
+  # join daily county estimates of pm2.5
+  left_join(county_pm, by = c('date', 'county'))
+
+# which(is.na(or_ndc_county_time_series$county))
+# or_ndc_county_time_series[113,]
+# which(or_zip_county$ZIP=="97078")
+
+summary(or_ndc_asthma)
+summary(or_ndc_county_time_series)
+summary(as.factor(or_ndc_county_time_series$county)) # 125 is missing out of 3592 (3.16%)
+
+# remove NA values
+which(is.na(or_ndc_county_time_series$county)) # 3468-3592
+or_ndc_county_time_series <- or_ndc_county_time_series[-c(3468:3592),]
+
+# write permanent dataframe
+
+write_path5 <- paste0("../data_new/medication/or_ndc_county_time_series.csv")
+write_csv(or_ndc_county_time_series, write_path5)
+
+## Check for missing values in sparsely populated counties. set to 0 (But no)
+# check_missing <- filter(or_ndc_county_time_series, is.na(n_obs))
+# mutate each to fill in missing values (Still no)
+# check2 <- check_missing %>% mutate_each(funs(new = ifelse(is.na(.), 0, .)), 
+#                                         n_obs:wrf_temp)
+
+plot_county <- ggplot(or_ndc_county_time_series, aes(y =n_obs, x = date)) +
+               geom_jitter() + facet_wrap(~county)
+
+plot_county_line <- ggplot(or_ndc_county_time_series, aes(y =n_obs, x = date)) +
+                    facet_wrap(~county) + geom_line()
+
+print(plot_county)
+print(plot_county_line)
+# save figure
+ggsave("../plot_new/ndc_county_ts.pdf", plot = plot_county, 
+       width = 12, height = 8, units = "in")
+ggsave("../plot_new/ndc_county_ts_line.pdf", plot = plot_county_line, 
+       width = 12, height = 8, units = "in")
+
+plot_ts <- ggplot(or_ndc_county_time_series, aes(y =n_obs, x = date)) +
+           geom_jitter()
+
+print(plot_ts)
+# save figure
+ggsave("../plot_new/ndc_ts.pdf", plot = plot_ts, 
+       width = 12, height = 8, units = "in")
+
+```
+
+Because we could see a large and distinct strata dotted time series in some big counties (such as Multnomah), I hope to find if the weekday and weekend have the different observation numbers/
+
+```{r}
+or_ndc_county_time_series_week <- or_ndc_county_time_series%>%
+  mutate(day_admit = as.factor(weekdays(date)),
+         month_admit = month(date)) %>%
+  mutate(outcome = 1)
+
+plot_ts_week <- ggplot(or_ndc_county_time_series_week, aes(y =n_obs, x = date)) +
+           geom_jitter() + facet_wrap(~day_admit)
+
+print(plot_ts_week)
+ggsave("../plot_new/ndc_ts_week.pdf", plot = plot_ts_week, 
+       width = 12, height = 8, units = "in")
+
+# Also check for Multnomah
+multnomah_df <- or_ndc_county_time_series_week %>%
+  filter(county == "Multnomah")
+
+plot_mult_week <- ggplot(multnomah_df, aes(y =n_obs, x = date)) +
+           geom_jitter() + facet_wrap(~day_admit)
+
+print(plot_mult_week)
+
+ggsave("../plot_new/multnomah_ts_week.pdf", plot = plot_mult_week, 
+       width = 12, height = 8, units = "in")
+
+plot_smoke_ts_wrf <- ggplot(or_ndc_county_time_series_week, aes(y =wrf_smk_pm, x = date)) +
+           geom_jitter(color="red")
+# print(plot_smoke_ts_wrf)
+
+plot_smoke_ts_krig <- ggplot(or_ndc_county_time_series_week, aes(y =krig_smk_pm, x = date)) +
+           geom_jitter() 
+# print(plot_smoke_ts_krig)
+
+plot_smoke_ts_gwr <- ggplot(or_ndc_county_time_series_week, aes(y =geo_smk_pm, x = date)) +
+           geom_jitter() 
+# print(plot_smoke_ts_gwr)
+
+plot_smk <- ggplot(or_ndc_county_time_series_week, aes(x = date)) +
+  geom_jitter(aes(y =wrf_smk_pm, x = date), color='green', size = 2) + 
+  geom_jitter(aes(y =krig_smk_pm, x = date), color='blue', size = 1) + 
+  geom_jitter(aes(y =wrf_smk_pm, x = date), color="red", size = 1)  +
+  ylab("Smoke PM 2.5")
+
+print(plot_smk)
+
+```

### 2017-05-31 r_scripts/analysis/oregon_medication.Rmd
update the manuscript

@@
 ---
-title: "Oregon 2013 Medication for Asthma"
+title: "Oregon 2013 Medication for Asthma Inhalers"
 author: "Jingyang Liu"
@@
 For this research I have the Oregon All Payer All Claims (APAC) data set in 2013.
-I want to research on the effects of wildfire smoke on pharmacy health outcome of asthma.
+I want to research on the effects of wildfire smoke on pharmacy outcome of asthma.
 NDC (National Drug Code) is the code that the data set offered.
@@
 
-The data cleaning and casecrossover study are done with server, the following are done with local R.
-
-## Preparation
-1. Data cleaning
-
-There are about 77,000,000 claims in total data set.
-Some have the NDC but another are null or missing.
-So I choose the claims in Oregon State with ndc, which is 23,330,714 claims.
-Also, the POS (Place of service code) all show "pharmacy" type.
+The data cleaning and casecrossover study are huge computation tasks, so they are done with server, the following work is done with local R.
+
+## Method Description
+
+In these comparisons, we examine various methods of smoke PM2.5 estimations and associations with health outcomes using a time-stratified case-crossover study design.
+Pharmacy outcomes for a patient with a primary diagnosis of cardiopulmonary health outcomes and their date of admission (index time) were identified.
+We then created counterfactual observations for each patient on the same day of the week for the entire wildfire season (May 1st to Sep 30th, 2013).
+For patients who have a index time before May 1, 2013 or after September 30, 2013, their referent observations before or after these dates will be excluded as I will not be able to assign estimates of PM2.5 to these referent observations.
+
+1. Data cleaning for health outcomes
+
+There are about 77 million claims in the total data set.
+Some have the NDC but another are null or missing because people are going for pharmacy or seeing the doctor.
+So I choose the claims in Oregon State with ndc existing, which is about 23 million claims.
+Also, the POS (Place of service code) all show "pharmacy" type, comparing to the previous health disease outcomes we limit the "emergency visit" type or "urgent care" type.
 
 For ndc, I choose the people who use the beta 2 agonists (one inhaler) for the day's first visit to form the sample.
-Because the ndc csv file has more information than what I need, I filter the ndc file by SQLite Studio (code and steps not attached) to a file only containing the beta 2 agonists medication.
-
-After choosing the claims with beta 2 agonists, the data set is reduced to 232,665. After choosing one day's first visit, the data set is the final data set we research on, containing 24,538 claims.
-
-2. Smoke model
-
-In this research, we foocus on GWR method (geographically weighted ridge regression).
-There are also WRF-Chem model (Weather Research and Forecasting with Chemistry) and Krig model.
-To reduce the occurance error I will compare all three models.
+Because the ndc csv file has more information than what I need, I filter the ndc file by SQLite Studio (code and steps not attached) to a file only containing the beta 2 agonists medication information.
+
+After choosing the claims with beta 2 agonists, the data set is reduced to 232,665.
+After choosing people's each day's first visit and limit the wildfire season dates (May 1st, 2013 to September 30th, 2013), the data set is the final data set we research on, containing  only 24,538 claims.
+
+2. Exposure methods for smoke model
+
+In this research, we focus on the GWR method (geographically weighted ridge regression).
+There are also WRF-Chem model (which substracts WRF-chem from the Weather Research and Forecasting with Chemistry no fire emission).
+For the WRF-Chem variable with the 'smoke' designator, this is WRF-Chem - WRF-Chem no fire.
+For Geo-Weighted Regression and Kriging model, with 'smk' designator, I subtracted off the 'Background' estimates of smoke, which I believe are the monthly averages of PM2.5 for a given grid.
+However, the 'Background' is the mean of total period from May 1st to Sep 30th 2013 so I have to assign the mean for every day.
+To reduce the occurance error and compare the differences I will analyze all three models.
+
+3. Analytic method
+
+We use the conditional logistic regression model using the survival package in R.
+Each conditional logistic regression model accounts for the subject, and adjusts for temperature from the WRF-Chem model.
+The conditional logistic regression model \beta represents the change in risk of an pharmacy event associated with a short-term unit increase in exposure, and can be calculated as an average difference between exposure at the index time and a weighted average of exposure at all times in the referent window.
 
@@
 
-### Choose ndc
-This analysis researches on asthma medication of Oregon All Claims All Payers data set.
-
-I choose the inhalation medicine beta 2 agonists to filter the sample for asthma.
-
-```{r ndc, eval=FALSE, echo=FALSE}
+```{r choose ndc, eval=FALSE, echo=FALSE}
 ### Import NDC data set --------------------------------------------------------
@@
 
-## Casecrossover study
-
-I also make the "false" dates for each claim.
+## Time-Stratified Case-Crossover (Population-Weighted PM~2.5~ Assigned by Zip Code)
+*referents same day of week within fire season May - September*
+
+Because we want the I also make the counterfactual dates for each claim, which means that for each claim, I choose the day the claim shows, to indicate the outcome for this day is "1".
+Then I make the dates week by week before this date, and after this date in the whole wildfire season, and indicate these dates' outcomes are "0".
+Then I can use these outcomes for the following model building.
 
@@
 
-## Time-Stratified Case-Crossover (referents same day of week within fire season May - September)
-
-** The following analysis are done in local R.**
+To get the time stratified data set, I create the lag variables that take smoke values from the previous days for zip codes, join the zipcode level populatoin-weighted pm to the data set, and create age, sex, month and season index.
 
@@
 
-## Population-Weighted PM~2.5~ Assigned by Zip Code
-
 ### Descriptives
@@
 For all of 2013, there were a total of millons of all claims records in the APAC dataset.
-During the May 1st to September 30th wildfire season in Oregon state, there were 24217 pharmacy visits for the beta 2 agonists with ndc.
-The number of persons (cases), as well as age-specific and sex-specific strate.
+During the May 1st to September 30th wildfire season in Oregon state, there were 24217 pharmacy visits for the ndc of beta 2 agonists.
+The number of cases, as well as age-specific and sex-specific strate are shown below.
 
@@
 # outcome name
-outcome_name <- "asthma"
+outcome_name <- "Inhalers"
 
@@
 tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
-         caption = "Association between a 10 ug/m^3 in PM2.5 three smoke method and health outcomes",
+         caption = "Association between a 10 ug/m^3 in PM2.5 three smoke method and pharmacy outcomes",
          # row group by outcome
-         rgroup = "Asthma",
+         rgroup = "Asthma Inhalers",
          n.rgroup = c(rep(3, 1)), # 6 rows for each age cat for each outcome
@@
          align = "llcccc", # column alignment,
-         tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within May to September fire season."
+         tfoot="&dagger; Referent periods matched to events on same day of week within May to September fire season."
          ) # end table
@@
 
-For some outcomes, like Age > 65 in WRF-Chem method, the CI is the only insignificant one in the plot.
+For some outcomes, such like Age > 65 in WRF-Chem method, the CI is the only insignificant one in the plot.
 
@@
 # outcome name
-outcome <- "asthma"
-outcome_name <- "Asthma"
+outcome <- "asthma inhalers"
+outcome_name <- "Asthma inhalers"
 
@@
          # row group by outcome
-         rgroup = "Asthma",
+         rgroup = "Asthma inhalers",
          n.rgroup = c(rep(9, 1)), # 6 rows for each age cat for each outcome
@@
   tab <- htmlTable(txtRound(wrf_geo_sex, digits = 3, 1:3), 
-           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by sex",
+           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and pharmacy outcomes stratified by sex",
            # row group by outcome
-           rgroup = "Asthma",
+           rgroup = "Asthma inhalers",
            n.rgroup = c(rep(6, 1)), # 2 rows for each sex strata for each outcome
@@
            align = "llcccc", # column alignment,
-           tfoot="&dagger; Adjusted for temperature, accounting for subject. Time-stratified: referent periods matched to events on same day of week within July to October fire season."
+           tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within July to October fire season."
             ) # end table
@@
 
-## County
-
-### Time series
-
-The time series for the number of pharmacy in each day (and separated with different counties).
-The overall time series plot shows a
-
-!!! In claim data I make a proportion table of county and zip code because we need to make shapefiles. In this pharmacy I choose a different data set
-
-```{r time series, warning=FALSE, message=FALSE}
+## County-level time series
+
+### Overall pharmacy visit time series
+
+The time series for the number of pharmacy in each day and separated with different counties. For convenience and representativeness, I choose some of the largest counties to show the visit features.
+
+The overall time series plot shows a time series plot.
+
+```{r time series, warning=FALSE, message=FALSE, echo=FALSE}
 read_path7 <- paste0("../data_new/medication/oregon_asthma_ndc.csv")
@@
 
-summary(or_ndc_asthma)
-summary(or_ndc_county_time_series)
-summary(as.factor(or_ndc_county_time_series$county)) # 125 is missing out of 3592 (3.16%)
+# summary(or_ndc_asthma)
+# summary(or_ndc_county_time_series)
+# summary(as.factor(or_ndc_county_time_series$county)) # 125 is missing out of 3592 (3.16%)
 
 # remove NA values
-which(is.na(or_ndc_county_time_series$county)) # 3468-3592
+# which(is.na(or_ndc_county_time_series$county)) # 3468-3592
 or_ndc_county_time_series <- or_ndc_county_time_series[-c(3468:3592),]
@@
 #                                         n_obs:wrf_temp)
+
+plot_ts <- ggplot(or_ndc_county_time_series, aes(y =n_obs, x = date)) +
+           geom_jitter()
+
+print(plot_ts)
+# save figure
+ggsave("../plot_new/ndc_ts.pdf", plot = plot_ts, 
+       width = 12, height = 8, units = "in")
+```
+
+### County time series
+
+The dotted time series plot of pharmacy visits by counties shows a large and distinct strata dotted time series in some big counties (such as Multnomah), I guess it may be the difference of visits between weekday and weekend.
+The lined time series plot also shows fluctuation.
+So we choose some larger counties (with more visits) to research on the effect of weekend and weekay.
+
+```{r, warning=FALSE, message=FALSE,echo=FALSE}
 
@@
 
-plot_ts <- ggplot(or_ndc_county_time_series, aes(y =n_obs, x = date)) +
-           geom_jitter()
-
-print(plot_ts)
-# save figure
-ggsave("../plot_new/ndc_ts.pdf", plot = plot_ts, 
-       width = 12, height = 8, units = "in")
-
 ```
 
-Because we could see a large and distinct strata dotted time series in some big counties (such as Multnomah), I hope to find if the weekday and weekend have the different observation numbers/
-
-```{r}
+### Time series of weekday and weekend in larger counties.
+We check the time series for Multnomah, and it indeed shows more visits of weekdays than that in weekends (except some holidays such like Labor Day, Independent Day, etc.) We may choose the new covariates for weekday and weekend.
+
+```{r echo = FALSE, message=FALSE, warning=FALSE}
 or_ndc_county_time_series_week <- or_ndc_county_time_series%>%
@@
 
-plot_ts_week <- ggplot(or_ndc_county_time_series_week, aes(y =n_obs, x = date)) +
-           geom_jitter() + facet_wrap(~day_admit)
-
-print(plot_ts_week)
-ggsave("../plot_new/ndc_ts_week.pdf", plot = plot_ts_week, 
-       width = 12, height = 8, units = "in")
-
 # Also check for Multnomah
@@
        width = 12, height = 8, units = "in")
+```
+
+### Time series of smoke exposure
+The plot shows the times series of smoke PM 2.5 exposure.
+It seems that at the beginning of August or at the end of July the smoke has a peak.
+Also during the August the smoke is very high.
+The green one shows the WRF-Chem method; The bule one shows the Krigging method; And the red one, which we focus on, shows the GWR method.
+
+```{r echo = FALSE, message=FALSE, warning=FALSE}
 
@@
 plot_smk <- ggplot(or_ndc_county_time_series_week, aes(x = date)) +
-  geom_jitter(aes(y =wrf_smk_pm, x = date), color='green', size = 2) + 
+  geom_jitter(aes(y =wrf_smk_pm, x = date), color='green', size = 1) + 
   geom_jitter(aes(y =krig_smk_pm, x = date), color='blue', size = 1) + 
-  geom_jitter(aes(y =wrf_smk_pm, x = date), color="red", size = 1)  +
+  geom_jitter(aes(y =geo_smk_pm, x = date), color="red", size = 1)  +
   ylab("Smoke PM 2.5")

### 2017-06-02 r_scripts/analysis/oregon_medication.Rmd
make indicator of disease; unfinished

@@
 
-```{r choose ndc, eval=FALSE, echo=FALSE}
+```{r import original data, eval=FALSE, echo=FALSE}
+### Import whole data set ------------------------------------------------------
+read_path <- paste0("../../../data/data_original/gan_episodes_of_care.txt")
+start_time <- Sys.time()
+oregon_df <- fread(read_path, sep = "|", showProgress = T) 
+stop_time <-  Sys.time() - start_time 
+# time it took
+stop_time # 7 mins
+
 ### Import NDC data set --------------------------------------------------------
@@
 
-which(oregon_ndc$ndc=="23490502001")
-
 ## Import Asthma NDC table
@@
 oregon_asthma <- oregon_ndc %>%
-  filter(ndc %in% ndc_codes) # 232665
+  filter(ndc %in% ndc_codes) # 232665, unique personkey 83787
+
+oregon_asthma_df <- oregon_asthma %>% 
+  group_by(personkey) %>% 
+  arrange(personkey, fromdate, line) %>% 
+  mutate(num_visit = dense_rank(fromdate)) %>%
+  arrange(personkey, num_visit) %>% 
+  select(personkey, clmid, num_visit) %>% 
+  unique() %>% 
+  full_join(oregon_asthma, by = c("personkey", "clmid")) %>% 
+  arrange(personkey, clmid, line, fromdate) %>% 
+  filter(row_number() == 1) %>%
+  # add new transverted from date
+  mutate(dates = as.Date(fromdate, "%m/%d/%Y")) # 83787 (same)
+
+oregon_df_ndc <- oregon_df %>%
+  filter(personkey %in% oregon_asthma$personkey) # 8909892
+
+### Import ICD9 code
+icd9_key <- read_excel("../../../data/data_original/CMS32_DESC_LONG_SHORT_DX.xlsx") %>% 
+  # rename the terrible variable names
+  select(dx_code = 1, long_desc = 2, short_desc = 3)
+
+# sort by icd9 code add row variable 
+icd9_key$X <- NULL
+icd9_key <- arrange(icd9_key, dx_code) %>%
+  mutate(n = as.numeric(row.names(icd9_key)))
+
+### asthma
+which(icd9_key$dx_code == '49300') # start of asthma is row 5206
+which(icd9_key$dx_code == '49392') # end of asthma is row 5219
+
+# limit just to asthma code and just the diagnosis column
+icd9_check <- icd9_key %>%
+  filter(n >= 5206 & n <= 5219) 
+icd9_check
+
+asthma_icd9 <- filter(icd9_key, n >= 5206 & n <= 5219) %>%
+  select(dx_code)
+# convert to vector
+asthma_icd9 <- as.vector(as.matrix(asthma_icd9))
+
+# now can I make a new variable, asthma1, that indicates an asthma claim?
+or_ndc_asthma_df <- oregon_df_ndc %>%
+  filter(dx1 %in% asthma_icd9 |
+         dx2 %in% asthma_icd9 |
+         dx3 %in% asthma_icd9 |
+         dx4 %in% asthma_icd9 |
+         dx5 %in% asthma_icd9 |
+         dx6 %in% asthma_icd9 |
+         dx7 %in% asthma_icd9 |
+         dx8 %in% asthma_icd9 |
+         dx9 %in% asthma_icd9 |
+         dx10 %in% asthma_icd9 |
+         dx11 %in% asthma_icd9 |
+         dx12 %in% asthma_icd9 |
+         dx13 %in% asthma_icd9) # 375451
+
+### COPD
+copd_icd9 <- c('490', '4910','4911','49120','49121','49122','4918','4919', '4920',
+               '4928', '4940', '4941', '496')
+
+# now can I make a new variable, asthma1, that indicates an asthma claim?
+or_ndc_copd_df <- oregon_df_ndc %>%
+  filter(dx1 %in% copd_icd9 |
+         dx2 %in% copd_icd9 |
+         dx3 %in% copd_icd9 |
+         dx4 %in% copd_icd9 |
+         dx5 %in% copd_icd9 |
+         dx6 %in% copd_icd9 |
+         dx7 %in% copd_icd9 |
+         dx8 %in% copd_icd9 |
+         dx9 %in% copd_icd9 |
+         dx10 %in% copd_icd9 |
+         dx11 %in% copd_icd9 |
+         dx12 %in% copd_icd9 |
+         dx13 %in% copd_icd9) # 361738
+
+or_ndc_asthma_copd <- oregon_df_ndc %>%
+  filter(dx1 %in% asthma_icd9 |
+         dx2 %in% asthma_icd9 |
+         dx3 %in% asthma_icd9 |
+         dx4 %in% asthma_icd9 |
+         dx5 %in% asthma_icd9 |
+         dx6 %in% asthma_icd9 |
+         dx7 %in% asthma_icd9 |
+         dx8 %in% asthma_icd9 |
+         dx9 %in% asthma_icd9 |
+         dx10 %in% asthma_icd9 |
+         dx11 %in% asthma_icd9 |
+         dx12 %in% asthma_icd9 |
+         dx13 %in% asthma_icd9 |
+         dx1 %in% copd_icd9 |
+         dx2 %in% copd_icd9 |
+         dx3 %in% copd_icd9 |
+         dx4 %in% copd_icd9 |
+         dx5 %in% copd_icd9 |
+         dx6 %in% copd_icd9 |
+         dx7 %in% copd_icd9 |
+         dx8 %in% copd_icd9 |
+         dx9 %in% copd_icd9 |
+         dx10 %in% copd_icd9 |
+         dx11 %in% copd_icd9 |
+         dx12 %in% copd_icd9 | 
+         dx13 %in% copd_icd9) # 722107
+
+or_ndc_asthma_copd_sum <- or_ndc_asthma_copd %>%
+  mutate(asthma1 = ifelse(dx1 %in% asthma_icd9, 1, 0),
+         asthma2 = ifelse(dx2 %in% asthma_icd9, 1, 0),
+         asthma3 = ifelse(dx3 %in% asthma_icd9, 1, 0),
+         asthma4 = ifelse(dx4 %in% asthma_icd9, 1, 0),
+         asthma5 = ifelse(dx5 %in% asthma_icd9, 1, 0),
+         asthma6 = ifelse(dx6 %in% asthma_icd9, 1, 0),
+         asthma7 = ifelse(dx7 %in% asthma_icd9, 1, 0),
+         asthma8 = ifelse(dx8 %in% asthma_icd9, 1, 0),
+         asthma9 = ifelse(dx9 %in% asthma_icd9, 1, 0),
+         asthma10 = ifelse(dx10 %in% asthma_icd9, 1, 0),
+         asthma11 = ifelse(dx11 %in% asthma_icd9, 1, 0),
+         asthma12 = ifelse(dx12 %in% asthma_icd9, 1, 0),
+         asthma13 = ifelse(dx13 %in% asthma_icd9, 1, 0),
+         # sum up the asthma indicators
+         asthma_sum = (asthma1 + asthma2 + asthma3 + asthma4 + asthma5 + asthma6 + asthma7 + asthma8 + asthma9 + asthma10 + asthma11 + asthma12 + asthma13),
+         asthma_dx = ifelse(asthma_sum>0, asthma_sum, 0)
+         ) %>% # end of mutate asthma
+  mutate(copd1 = ifelse(dx1 %in% copd_icd9, 1, 0),
+         copd2 = ifelse(dx2 %in% copd_icd9, 1, 0),
+         copd3 = ifelse(dx3 %in% copd_icd9, 1, 0),
+         copd4 = ifelse(dx4 %in% copd_icd9, 1, 0),
+         copd5 = ifelse(dx5 %in% copd_icd9, 1, 0),
+         copd6 = ifelse(dx6 %in% copd_icd9, 1, 0),
+         copd7 = ifelse(dx7 %in% copd_icd9, 1, 0),
+         copd8 = ifelse(dx8 %in% copd_icd9, 1, 0),
+         copd9 = ifelse(dx9 %in% copd_icd9, 1, 0),
+         copd10 = ifelse(dx10 %in% copd_icd9, 1, 0),
+         copd11 = ifelse(dx11 %in% copd_icd9, 1, 0),
+         copd12 = ifelse(dx12 %in% copd_icd9, 1, 0),
+         copd13 = ifelse(dx13 %in% copd_icd9, 1, 0),
+         # sum up the pneumonia indicators
+         copd_sum = (copd1 + copd2 + copd3 + copd4 +copd5 + copd6 + copd7 + copd8 + copd9 +copd10 + copd11 + copd12 + copd13),
+         copd_dx = ifelse(copd_sum>0, copd_sum, 0)
+  ) %>% #end of mutate of copd
+  select(-c(73:85), c(88:100))
+
+or_index <- or_ndc_asthma_copd_sum %>%
+  mutate(index = ifelse(asthma_dx > copd_dx, "asthma", 
+                 ifelse(asthma_dx < copd_dx, "copd", 
+                 ifelse(asthma_dx == copd_dx, "both", NA)))) # 56512 unique personkey
+# asthma   both   copd
+# 360631  14152 347324
+
+write_path <- paste0('../../../data/data_new/',
+                     'oregon_index.csv')
+write_csv(or_index, write_path)
+
+```
+
+```{r choose ndc, eval=FALSE, echo=FALSE}
 

### 2017-06-05 r_scripts/analysis/oregon_medication.Rmd
make 4 categories index

--- 2017-05-31 r_scripts/analysis/oregon_medication.Rmd
@@
 
-```{r choose ndc, eval=FALSE, echo=FALSE}
+```{r import filtered original data, eval=FALSE, echo=FALSE}
+### Import whole data set ------------------------------------------------------
+read_path <- paste0("../../../data/data_original/gan_episodes_of_care.txt")
+start_time <- Sys.time()
+oregon_df <- fread(read_path, sep = "|", showProgress = T) 
+stop_time <-  Sys.time() - start_time 
+# time it took
+stop_time # 7 mins
+
 ### Import NDC data set --------------------------------------------------------
@@
 
-oregon_asthma <- oregon_ndc %>%
-  filter(ndc %in% ndc_codes) # 232665
+oregon_beta <- oregon_ndc %>%
+  filter(ndc %in% ndc_codes) # 232665, unique personkey 83787
+
+oregon_beta_df <- oregon_beta %>% 
+  group_by(personkey) %>% 
+  arrange(personkey, fromdate, line) %>% 
+  mutate(num_visit = dense_rank(fromdate)) %>%
+  arrange(personkey, num_visit) %>% 
+  select(personkey, clmid, num_visit) %>% 
+  unique() %>% 
+  full_join(oregon_beta, by = c("personkey", "clmid")) %>% 
+  arrange(personkey, clmid, line, fromdate) %>% 
+  filter(row_number() == 1) %>%
+  # add new transverted from date
+  mutate(dates = as.Date(fromdate, "%m/%d/%Y")) # 83787 (same)
+
+oregon_df_ndc <- oregon_df %>%
+  filter(personkey %in% oregon_beta$personkey) # 8909892
+
+### Import ICD9 code
+icd9_key <- read_excel("../../../data/data_original/CMS32_DESC_LONG_SHORT_DX.xlsx") %>% 
+  # rename the terrible variable names
+  select(dx_code = 1, long_desc = 2, short_desc = 3)
+
+# sort by icd9 code add row variable 
+icd9_key$X <- NULL
+icd9_key <- arrange(icd9_key, dx_code) %>%
+  mutate(n = as.numeric(row.names(icd9_key)))
+
+### without ICD-9
+or_ndc_no_icd <- oregon_df_ndc %>%
+  filter(dx1=="*NULL*" &
+         dx2=="*NULL*" &
+         dx3=="*NULL*" &
+         dx4=="*NULL*" &
+         dx5=="*NULL*" &
+         dx6=="*NULL*" &
+         dx7=="*NULL*" &
+         dx8=="*NULL*" &
+         dx9=="*NULL*" &
+         dx10=="*NULL*" &
+         dx11=="" &
+         dx12=="" &
+         dx13=="" )
+
+length(or_ndc_no_icd$personkey) # 3656720 claims
+length(unique(or_ndc_no_icd$personkey)) # 83787
+
+oregon_ndc_dx <- oregon_ndc %>%
+    filter(dx1!="*NULL*" |
+         dx2!="*NULL*" |
+         dx3!="*NULL*" |
+         dx4!="*NULL*" |
+         dx5!="*NULL*" |
+         dx6!="*NULL*" |
+         dx7!="*NULL*" |
+         dx8!="*NULL*" |
+         dx9!="*NULL*" |
+         dx10!="*NULL*" |
+         dx11!="" |
+         dx12!="" |
+         dx13!="" ) # 0 claims
+
+### asthma
+which(icd9_key$dx_code == '49300') # start of asthma is row 5206
+which(icd9_key$dx_code == '49392') # end of asthma is row 5219
+
+# limit just to asthma code and just the diagnosis column
+icd9_check <- icd9_key %>%
+  filter(n >= 5206 & n <= 5219) 
+icd9_check
+
+asthma_icd9 <- filter(icd9_key, n >= 5206 & n <= 5219) %>%
+  select(dx_code)
+# convert to vector
+asthma_icd9 <- as.vector(as.matrix(asthma_icd9))
+
+# now can I make a new variable, asthma1, that indicates an asthma claim?
+or_ndc_asthma_df <- oregon_df_ndc %>%
+  filter(dx1 %in% asthma_icd9 |
+         dx2 %in% asthma_icd9 |
+         dx3 %in% asthma_icd9 |
+         dx4 %in% asthma_icd9 |
+         dx5 %in% asthma_icd9 |
+         dx6 %in% asthma_icd9 |
+         dx7 %in% asthma_icd9 |
+         dx8 %in% asthma_icd9 |
+         dx9 %in% asthma_icd9 |
+         dx10 %in% asthma_icd9 |
+         dx11 %in% asthma_icd9 |
+         dx12 %in% asthma_icd9 |
+         dx13 %in% asthma_icd9) # 375451, unique personkey: 45165
+
+### COPD
+copd_icd9 <- c('490', '4910','4911','49120','49121','49122','4918','4919', '4920',
+               '4928', '4940', '4941', '496')
+
+# now can I make a new variable, asthma1, that indicates an asthma claim?
+or_ndc_copd_df <- oregon_df_ndc %>%
+  filter(dx1 %in% copd_icd9 |
+         dx2 %in% copd_icd9 |
+         dx3 %in% copd_icd9 |
+         dx4 %in% copd_icd9 |
+         dx5 %in% copd_icd9 |
+         dx6 %in% copd_icd9 |
+         dx7 %in% copd_icd9 |
+         dx8 %in% copd_icd9 |
+         dx9 %in% copd_icd9 |
+         dx10 %in% copd_icd9 |
+         dx11 %in% copd_icd9 |
+         dx12 %in% copd_icd9 |
+         dx13 %in% copd_icd9) # 361738, unique personkey: 18626
+
+### asthma or copd
+or_or_asthma_copd <- oregon_df_ndc %>%
+  filter(dx1 %in% asthma_icd9 |
+         dx2 %in% asthma_icd9 |
+         dx3 %in% asthma_icd9 |
+         dx4 %in% asthma_icd9 |
+         dx5 %in% asthma_icd9 |
+         dx6 %in% asthma_icd9 |
+         dx7 %in% asthma_icd9 |
+         dx8 %in% asthma_icd9 |
+         dx9 %in% asthma_icd9 |
+         dx10 %in% asthma_icd9 |
+         dx11 %in% asthma_icd9 |
+         dx12 %in% asthma_icd9 |
+         dx13 %in% asthma_icd9 |
+         dx1 %in% copd_icd9 |
+         dx2 %in% copd_icd9 |
+         dx3 %in% copd_icd9 |
+         dx4 %in% copd_icd9 |
+         dx5 %in% copd_icd9 |
+         dx6 %in% copd_icd9 |
+         dx7 %in% copd_icd9 |
+         dx8 %in% copd_icd9 |
+         dx9 %in% copd_icd9 |
+         dx10 %in% copd_icd9 |
+         dx11 %in% copd_icd9 |
+         dx12 %in% copd_icd9 | 
+         dx13 %in% copd_icd9) # 722107 , unique personkey: 56512
+
+# 45165+ 18626 - 56512
+# [1] 7279 on both asthma and copd
+
+### Both asthma and copd
+length(unique(or_or_asthma_copd$personkey))
+length(unique(or_ndc_asthma_df$personkey))
+length(unique(or_ndc_copd_df$personkey))
+
+or_with_icd <- oregon_df_ndc %>%
+  filter(!dx1=="*NULL*" |
+         !dx2=="*NULL*" |
+         !dx3=="*NULL*" |
+         !dx4=="*NULL*" |
+         !dx5=="*NULL*" |
+         !dx6=="*NULL*" |
+         !dx7=="*NULL*" |
+         !dx8=="*NULL*" |
+         !dx9=="*NULL*" |
+         !dx10=="*NULL*" |
+         !dx11=="" |
+         !dx12=="" |
+         !dx13=="" )
+
+asthma_copd_both_id <- or_ndc_asthma_df %>%
+  filter(personkey %in% unique(or_ndc_copd_df$personkey)) %>% # 7279 person
+  select(personkey) %>%
+  unique() %>%
+  mutate(index="both")
+
+### Only asthma
+asthma_only_id <- or_ndc_asthma_df %>%
+  filter(!personkey %in% asthma_copd_both_id$personkey) %>% # 37886 person
+  select(personkey) %>%
+  unique() %>%
+  mutate(index="asthma")
+
+### Only copd
+copd_only_id <- or_ndc_copd_df %>%
+  filter(!personkey %in% asthma_copd_both_id$personkey) %>% # 11347 person
+  select(personkey) %>%
+  unique() %>%
+  mutate(index="copd")
+
+### Other
+other_disease_id <- oregon_df_ndc %>%
+  filter(!personkey %in% or_or_asthma_copd$personkey) %>% # 27275 person
+  select(personkey) %>%
+  unique() %>%
+  mutate(index="other")
+
+id_index <- bind_rows(asthma_copd_both_id, asthma_only_id, copd_only_id, other_disease_id)
+
+## Join with original data and indicator for 4 categories
+oregon_ndc_index <- oregon_df_ndc %>%
+  left_join(id_index, by="personkey")
+summary(as.factor(oregon_ndc_index$index))
+
+### Check
+or_ndc_asthma_copd_sum <- oregon_ndc_index %>%
+  mutate(asthma1 = ifelse(dx1 %in% asthma_icd9, 1, 0),
+         asthma2 = ifelse(dx2 %in% asthma_icd9, 1, 0),
+         asthma3 = ifelse(dx3 %in% asthma_icd9, 1, 0),
+         asthma4 = ifelse(dx4 %in% asthma_icd9, 1, 0),
+         asthma5 = ifelse(dx5 %in% asthma_icd9, 1, 0),
+         asthma6 = ifelse(dx6 %in% asthma_icd9, 1, 0),
+         asthma7 = ifelse(dx7 %in% asthma_icd9, 1, 0),
+         asthma8 = ifelse(dx8 %in% asthma_icd9, 1, 0),
+         asthma9 = ifelse(dx9 %in% asthma_icd9, 1, 0),
+         asthma10 = ifelse(dx10 %in% asthma_icd9, 1, 0),
+         asthma11 = ifelse(dx11 %in% asthma_icd9, 1, 0),
+         asthma12 = ifelse(dx12 %in% asthma_icd9, 1, 0),
+         asthma13 = ifelse(dx13 %in% asthma_icd9, 1, 0),
+         # sum up the asthma indicators
+         asthma_sum = (asthma1 + asthma2 + asthma3 + asthma4 + asthma5 + asthma6 + asthma7 + asthma8 + asthma9 + asthma10 + asthma11 + asthma12 + asthma13),
+         asthma_dx = ifelse(asthma_sum>0, asthma_sum, 0)
+         ) %>% # end of mutate asthma
+  mutate(copd1 = ifelse(dx1 %in% copd_icd9, 1, 0),
+         copd2 = ifelse(dx2 %in% copd_icd9, 1, 0),
+         copd3 = ifelse(dx3 %in% copd_icd9, 1, 0),
+         copd4 = ifelse(dx4 %in% copd_icd9, 1, 0),
+         copd5 = ifelse(dx5 %in% copd_icd9, 1, 0),
+         copd6 = ifelse(dx6 %in% copd_icd9, 1, 0),
+         copd7 = ifelse(dx7 %in% copd_icd9, 1, 0),
+         copd8 = ifelse(dx8 %in% copd_icd9, 1, 0),
+         copd9 = ifelse(dx9 %in% copd_icd9, 1, 0),
+         copd10 = ifelse(dx10 %in% copd_icd9, 1, 0),
+         copd11 = ifelse(dx11 %in% copd_icd9, 1, 0),
+         copd12 = ifelse(dx12 %in% copd_icd9, 1, 0),
+         copd13 = ifelse(dx13 %in% copd_icd9, 1, 0),
+         # sum up the pneumonia indicators
+         copd_sum = (copd1 + copd2 + copd3 + copd4 +copd5 + copd6 + copd7 + copd8 + copd9 +copd10 + copd11 + copd12 + copd13),
+         copd_dx = ifelse(copd_sum>0, copd_sum, 0)
+  ) #end of mutate of copd
+
+or_ndc_check <- or_ndc_asthma_copd_sum %>%
+  select(personkey, index, asthma_dx, copd_dx)
+summary(as.factor(or_ndc_check$asthma_dx))
+
+or_asthma <- or_ndc_asthma_copd_sum %>%
+  filter(index=="asthma")
+or_copd <- or_ndc_asthma_copd_sum %>%
+  filter(index=="copd")
+or_both <- or_ndc_asthma_copd_sum %>%
+  filter(index=="both")
+or_other <- or_ndc_asthma_copd_sum %>%
+  filter(index=="other")
+
+summary(as.factor(or_other$dx1))
+
+## check for one sample
+summary(as.factor(or_asthma$personkey))
+summary(as.factor(or_copd$personkey))
+summary(as.factor(or_both$personkey))
+summary(as.factor(or_other$personkey))
+
+or_asthma_sample <- or_asthma %>%
+  filter(personkey == 10101244) 
+which(or_asthma_sample$asthma_dx!=0)
+
+or_copd_sample  <- or_copd %>%
+  filter(personkey == 13564804)
+
+or_both_sample  <- or_both %>%
+  filter(personkey == 12979399)
+
+or_other_sample  <- or_other %>%
+  filter(personkey == 10289787)
+
+write_path <- paste0('../../../data/data_new/',
+                     'oregon_asthma_sample.csv')
+write_csv(or_asthma_sample, write_path)
+
+write_path2 <- paste0('../../../data/data_new/',
+                     'oregon_ndc_with_index.csv')
+
+write_csv(or_ndc_asthma_copd_sum, write_path2)
+
+```
+
+```{r choose ndc, eval=FALSE, echo=FALSE}
 

### 2017-06-06 r_scripts/analysis/oregon_medication.Rmd
try to figure out missing values

@@
 ```
+
+### Make category index for beta-2 agonists.
+First I filtered the whole data set with ndc existing and only Oregon state.
+Then there are about 24 million claims left.
+Then I filtered the data set with only beta-2-agonists and found about 233 thousand claims (83787 unique person).
+By using these unique personkey, I found the claims of them in the total data set, which is about 8.9 million claims.
+This the data set that I use for the category classification.
 
@@
 length(unique(or_ndc_no_icd$personkey)) # 83787
+
+or_no_ndc_no_icd <- oregon_df %>%
+  filter(dx1=="*NULL*" &
+         dx2=="*NULL*" &
+         dx3=="*NULL*" &
+         dx4=="*NULL*" &
+         dx5=="*NULL*" &
+         dx6=="*NULL*" &
+         dx7=="*NULL*" &
+         dx8=="*NULL*" &
+         dx9=="*NULL*" &
+         dx10=="*NULL*" &
+         dx11=="" &
+         dx12=="" &
+         dx13=="") %>%
+  filter(ndc=="" |
+         ndc=="*NULL*") # 1057022 missing in total data set
+
+or_ndc_all_miss <- oregon_df_ndc %>%
+  filter(dx1=="*NULL*" &
+         dx2=="*NULL*" &
+         dx3=="*NULL*" &
+         dx4=="*NULL*" &
+         dx5=="*NULL*" &
+         dx6=="*NULL*" &
+         dx7=="*NULL*" &
+         dx8=="*NULL*" &
+         dx9=="*NULL*" &
+         dx10=="*NULL*" &
+         dx11=="" &
+         dx12=="" &
+         dx13=="") %>%
+  filter(ndc=="" |
+         ndc=="*NULL*") # 20269 missing in total data set
+
+or_ndc_all_miss_id <- or_ndc_asthma_df %>%
+  filter(personkey %in% unique(or_ndc_copd_df$personkey)) %>% # 7279 person
+  select(personkey)
 
@@
 id_index <- bind_rows(asthma_copd_both_id, asthma_only_id, copd_only_id, other_disease_id)
+
+## Join the index with beta 2 ndc claims
+oregon_beta_index <- oregon_beta %>%
+  left_join(id_index, by="personkey")
+summary(as.factor(oregon_beta_index$index))
+
+write_path3 <- paste0('../../../data/data_new/',
+                     'oregon_beta_ndc_index.csv')
+write_csv(oregon_beta_index, write_path3)
 
@@
   filter(personkey == 10289787)
+
+or_asthma_miss <- or_asthma %>%
+  filter(dx1=="*NULL*" &
+         dx2=="*NULL*" &
+         dx3=="*NULL*" &
+         dx4=="*NULL*" &
+         dx5=="*NULL*" &
+         dx6=="*NULL*" &
+         dx7=="*NULL*" &
+         dx8=="*NULL*" &
+         dx9=="*NULL*" &
+         dx10=="*NULL*" &
+         dx11=="" &
+         dx12=="" &
+         dx13=="") %>%
+  filter(ndc=="" |
+         ndc=="*NULL*") # 20269 missing in total data set
+
+or_copd_miss <- or_copd %>%
+  filter(dx1=="*NULL*" &
+         dx2=="*NULL*" &
+         dx3=="*NULL*" &
+         dx4=="*NULL*" &
+         dx5=="*NULL*" &
+         dx6=="*NULL*" &
+         dx7=="*NULL*" &
+         dx8=="*NULL*" &
+         dx9=="*NULL*" &
+         dx10=="*NULL*" &
+         dx11=="" &
+         dx12=="" &
+         dx13=="") %>%
+  filter(ndc=="" |
+         ndc=="*NULL*") # 20269 missing in total data set
+
+or_both_miss <- or_both %>%
+  filter(dx1=="*NULL*" &
+         dx2=="*NULL*" &
+         dx3=="*NULL*" &
+         dx4=="*NULL*" &
+         dx5=="*NULL*" &
+         dx6=="*NULL*" &
+         dx7=="*NULL*" &
+         dx8=="*NULL*" &
+         dx9=="*NULL*" &
+         dx10=="*NULL*" &
+         dx11=="" &
+         dx12=="" &
+         dx13=="") %>%
+  filter(ndc=="" |
+         ndc=="*NULL*") # 20269 missing in total data set
+
+or_other_miss <- or_other %>%
+  filter(dx1=="*NULL*" &
+         dx2=="*NULL*" &
+         dx3=="*NULL*" &
+         dx4=="*NULL*" &
+         dx5=="*NULL*" &
+         dx6=="*NULL*" &
+         dx7=="*NULL*" &
+         dx8=="*NULL*" &
+         dx9=="*NULL*" &
+         dx10=="*NULL*" &
+         dx11=="" &
+         dx12=="" &
+         dx13=="") %>%
+  filter(ndc=="" |
+         ndc=="*NULL*") # 20269 missing in total data set
+
+dim(or_asthma_miss)
+dim(or_copd_miss)
+dim(or_both_miss)
+dim(or_other_miss)
+
+sum(or_ndc_all_miss_id %in% unique(or_other$personkey))
 
@@
 ```
+
+### NDC claims with index
+
+```{r echo = FALSE, message=FALSE, warning=FALSE}
+read_path7 <- paste0("../data_new/medication/oregon_beta_ndc_index.csv")
+or_ndc_index <- read_csv(read_path7)
+
+oregon_beta_df <- or_ndc_index %>% 
+  group_by(personkey) %>% 
+  arrange(personkey, fromdate, line) %>% 
+  mutate(num_visit = dense_rank(fromdate)) %>%
+  arrange(personkey, num_visit) %>% 
+  select(personkey, clmid, num_visit) %>% 
+  unique() %>% 
+  full_join(or_ndc_index, by = c("personkey", "clmid")) %>% 
+  arrange(personkey, clmid, line, fromdate) %>% 
+  filter(row_number() == 1) %>%
+  # add new transverted from date
+  mutate(dates = as.Date(fromdate, "%m/%d/%Y")) %>%
+  filter(dates >= '2013-05-01' & 
+         dates <= '2013-09-30') # 24538
+
+summary(as.factor(oregon_beta_df$index))
+
+```

### 2017-06-09 r_scripts/analysis/oregon_medication.Rmd
add four disease category -need modify explanation

@@
 library(htmlTable)
+library(knitr)
 ```
@@
   filter(ndc=="" |
-         ndc=="*NULL*") # 20269 missing in total data set
+         ndc=="*NULL*") # 89482 claims missing in asthma data set
 
@@
   filter(ndc=="" |
-         ndc=="*NULL*") # 20269 missing in total data set
+         ndc=="*NULL*") # 18941 claims missing in copd data set
 
@@
   filter(ndc=="" |
-         ndc=="*NULL*") # 20269 missing in total data set
+         ndc=="*NULL*") # 17963 claims missing in both data set
 
@@
   filter(ndc=="" |
-         ndc=="*NULL*") # 20269 missing in total data set
+         ndc=="*NULL*") # 46971 claims missing in other data set (5848 personkey)
+
+or_other_not_miss <- or_other %>%
+  filter(dx1!="*NULL*" |
+         dx2!="*NULL*" |
+         dx3!="*NULL*" |
+         dx4!="*NULL*" |
+         dx5!="*NULL*" |
+         dx6!="*NULL*" |
+         dx7!="*NULL*" |
+         dx8!="*NULL*" |
+         dx9!="*NULL*" |
+         dx10!="*NULL*" |
+         dx11!="" |
+         dx12!="" |
+         dx13!="" |
+         ndc!="" |
+         ndc!="*NULL*") # 20269 missing in total data set
+
+length(unique(or_other_not_miss$personkey)) # [1] 27275
 
@@
   # save figure
-  ggsave("../plot_new/ndc_age_fig_three.pdf", plot = print_plot, 
+  ggsave("../plot_new/ndc_overall_three.pdf", plot = print_plot, 
        width = 12, height = 8, units = "in")
@@
 
-### NDC claims with index
+## NDC claims with index
+
+Check the claims in each category (Asthma, COPD, Both or Other).
 
@@
 ```
+
+### Descriptives
+
+For all of 2013, there were a total of millons of all claims records in the APAC dataset.
+During the May 1st to September 30th wildfire season in Oregon state, there were 24217 pharmacy visits for the ndc of beta 2 agonists.
+The number of cases, as well as age-specific and sex-specific strate are shown below.
+
+```{r ndc descriptive table, echo = F, message = F, warning = F, results='asis'}
+read_path6 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover.csv")
+beta_ndc <- read_csv(read_path6)
+
+# join personkey with index
+oregon_beta_index_df <- oregon_beta_df %>%
+  select(personkey, index)
+
+beta_ndc_with_index <- beta_ndc %>%
+  left_join(oregon_beta_index_df, by="personkey")
+
+# dataframe to loop through
+beta_ndc_with_index_df <- data.frame(beta_ndc_with_index)
+
+beta_asthma <- beta_ndc_with_index_df %>%
+  filter(index=="asthma")
+
+beta_copd <- beta_ndc_with_index_df %>%
+  filter(index=="copd")
+
+beta_both <- beta_ndc_with_index_df %>%
+  filter(index=="both")
+
+beta_other <- beta_ndc_with_index_df %>%
+  filter(index=="other")
+
+df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)
+
+outcome_list <- c("Asthma", "COPD", "Both", "Other")
+
+# create an empty list to row bind dataframes together
+datalist <- list()
+
+# dataframe list
+method_list <- c('WRF-Chem Smoke', 'Geo-Weighted Smoke', 'Kriging Smoke')
+
+# outcome name
+outcome_name <- c("Asthma", "COPD", "Both", "Other")
+
+for(i in 1:length(df_list)){
+  # dataframe to loop through
+  df_to_loop <- data.frame(df_list[i])
+  # indication of column
+  outcome <- colnames(df_to_loop[75])
+  # outcome name
+  outcome_name <- outcome_list[i]
+
+  # dataframe for analysis creation
+  # bind columns back together 
+  df_analysis <- df_to_loop %>%
+    filter(!is.na(wrf_smk_pm_zip)) %>% 
+    # only look at outcomes
+    filter(outcome == 1) %>%
+    # add another row that makes sure there is a person <15 in the dataframe
+    # tricking xtabs to produce a 0 cell for the outcome for age <15
+    add_row(outcome = 0, age_ind = 0)
+
+  # cross tabs
+  outcome_n <- xtabs(~ outcome, df_analysis)
+  cross_tab_age <- xtabs(~ outcome + age_ind, df_analysis)
+  cross_tab_sex <- xtabs(~ outcome + sex_ind, df_analysis)
+  # empty matrix
+  point_estimates <- matrix(nrow = 1, ncol = 7, byrow = T)
+
+  colnames(point_estimates) <- c("outcome", "n", "age_15", "age_15_65", 
+                                 "age_65", "female", "male")
+
+  # fill in the outcome name for the dataframe before the loop
+  point_estimates[, 1] <- outcome_name
+  # fill n
+  point_estimates[, 2] <- outcome_n[2] # second element of the 1 dimension vector
+  # age <15
+  point_estimates[, 3] <- cross_tab_age[2, 1]
+  # age 15 to 65
+  point_estimates[, 4] <- cross_tab_age[2, 2]
+  # age >65
+  point_estimates[, 5] <- cross_tab_age[2, 3]
+  # male == 0
+  point_estimates[, 7] <- cross_tab_sex[1, 1]
+  # female == 1
+  point_estimates[, 6] <- cross_tab_sex[1, 2]
+
+  # save point estimates as a dataframe
+  point_est_df <- as_data_frame(point_estimates)
+
+  # combine previous values in dataframe that has all outcome/methods comparisons
+  datalist[[i]] <- point_est_df
+
+} # end of loop
+
+# combine each outcome dataframe itteration in to a big dataset
+combined_point_est_df <- bind_rows(datalist) %>% 
+  # find proportions/percents for each strata in a row
+  mutate(age_15_pr = as.character(round((as.numeric(age_15)/as.numeric(n))*100,1)),
+         age_15_65_pr = as.character(round((as.numeric(age_15_65)/as.numeric(n))*100,1)),
+         age_65_pr = as.character(round((as.numeric(age_65)/as.numeric(n))*100,1)),
+         female_pr = as.character(round((as.numeric(female)/as.numeric(n))*100,1)),
+         male_pr = as.character(round((as.numeric(male)/as.numeric(n))*100,1))) %>% 
+  select(outcome, n, age_15, age_15_pr, age_15_65, age_15_65_pr, age_65,
+         age_65_pr, female, female_pr, male, male_pr)
+
+#str(combined_point_est_df)
+
+tab <- htmlTable(txtRound(combined_point_est_df, digits = 1), 
+           caption = "Number of cases for each outcome observed from May 1st to September 30st, 2013",
+           # column headers
+           header = c("Outcome", "Cases n", "Less than 15", "(%)", "15 to 65", "(%)", "Greater than 65",
+                      "(%)", "Female", "(%)", "Male", "(%)"),
+           # column spanner
+           cgroup = c("","Age Category", "Sex"), 
+           n.cgroup = c(2, 6, 4),
+           padding.rgroup = "&nbsp;&nbsp;",
+           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+           align = "llccccc" # column alignment,
+            ) # end table
+
+print(tab)
+
+```
+
+### Overall plot
+
+```{r ndc overall try, warning =F, echo = F, results='asis'}
+df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)
+
+outcome_list <- c("Asthma", "COPD", "Both", "Other")
+
+# create an empty list to row bind dataframes together
+datalist <- list()
+
+# dataframe list
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# outcome name
+outcome_name <- c("Asthma", "COPD", "Both", "Other")
+
+# data wrangling ----
+# Producing conditional logit model estimates loop 
+for(i in 1:length(df_list)){
+  # dataframe to loop through
+  df_to_loop <- data.frame(df_list[i])
+  # indication of column
+  outcome <- colnames(df_to_loop[75])
+  # outcome name
+  outcome_name <- outcome_list[i]
+
+  # extract covariates from dataframe
+  covariates_df <- df_to_loop[, c(1:26, 74:84)]
+
+  # extract pm values and divide by 10 and ordered
+  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+  pm_estimates_df <- df_to_loop[, c(87, 92, 91, 93)]/10  # create 10 unit increases
+
+  # dataframe for analysis creation
+  # bind columns back together 
+  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+    # remove missing pm values
+    filter(!is.na(wrf_smk_pm_zip)) %>%
+
+    # join in cdc met data
+    # full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
+    # the following code makes sure that the counterfactual values retained are 
+    # symetric in that number of obs before = number of obs after
+    mutate(obs_diff_admission = (fromdate - date)/7) 
+  # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+  # empty df for table
+  table_df <- data.frame()
+
+  # empty matrix
+  point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)
+
+  colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 'odds_ratio', 
+                                 'lower95', 'upper95', 'se', 'p_val')
+
+  # fill in the outcome name for the dataframe before the loop
+  point_estimates[, 1] <- outcome_name
+
+  # second loop to run a model for each pm estimation method
+
+  for(j in 38:40){
+
+    # variable to model 
+    var_name <- colnames(df_analysis[j])
+
+    # set row number to fill
+    row_n <- j-37
+
+    # only run the model if the dataframe has observations
+    if(nrow(df_analysis) != 0){
+      # conditional logistic regression model
+      mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+      # populate matrix
+      point_estimates[row_n, 2] <- method_list[row_n]
+      point_estimates[row_n, 3] <- mod$n
+      point_estimates[row_n, 4] <- mod$nevent
+      # odds ratio
+      point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+      # 95% lower bound
+      point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
+      # 95% upper bound
+      point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                               1.96*(summary(mod)$coefficient[1,3])), 3)
+      # standard error
+      point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
+      # p val
+      point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)
+
+      # save point estimates as a dataframe
+      point_est_df <- as_data_frame(point_estimates)
+
+      # create else statement that fills matrix with missing so I still have the row
+      # in the final dataframe
+    } else {point_estimates[row_n, 2] <- method_list[row_n]
+    point_estimates[row_n, 3] <- 0
+    point_estimates[row_n, 4] <- 0
+    point_estimates[row_n, c(5:9)] <- 99 } # end 'if else' statement
+  } # end methods loop
+
+  # combine previous values in dataframe that has all outcome/methods comparisons
+  datalist[[i]] <- point_est_df
+
+} # end of loop
+
+# combine each outcome dataframe itteration in to a big dataset
+combined_point_est_df <- bind_rows(datalist)
+
+# subset columns I want to put in to the table
+table_df <- combined_point_est_df %>% select(2, 3:7)
+
+tab <- htmlTable(txtRound(table_df, digits = 3, 1:3), 
+                 caption = "Association between a 10 ug/m^3 in PM2.5 and Health Outcomes",
+                 # row group by outcome
+                 rgroup = outcome_list,
+                 n.rgroup = c(rep(3, 4)), # 4 rows for each method for each outcome
+                 # column headers
+                 header = c("Method", "Obs.", "Events",
+                            "OR&dagger;", "Lower", "Upper"),
+                 # column spanner
+                 cgroup = c("", "95% CI"), 
+                 n.cgroup = c(4, 2),
+                 padding.rgroup = "&nbsp;&nbsp;",
+                 css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+                 align = "llccccc", # column alignment,
+                 tfoot="&dagger; Adjusted for CDC temperature, relatively humidity, wind speed, and precipitation; accounting for subject.
+Time-stratified: referent periods matched to events on same day of week within July to October fire season."
+) # end table
+
+print(tab)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+combined_point_est_df$outcome <- factor(combined_point_est_df$outcome, 
+                                        levels = unique(combined_point_est_df$outcome))
+
+combined_point_est_df$pm_method <- factor(combined_point_est_df$pm_method, 
+                                          levels = unique(combined_point_est_df$pm_method))
+
+combined_point_est_df$n <- as.numeric(combined_point_est_df$n)
+combined_point_est_df$n_events <- as.numeric(combined_point_est_df$n_events)
+combined_point_est_df$odds_ratio <- as.numeric(combined_point_est_df$odds_ratio)
+combined_point_est_df$lower95 <- as.numeric(combined_point_est_df$lower95)
+combined_point_est_df$upper95 <- as.numeric(combined_point_est_df$upper95)
+combined_point_est_df$se <- as.numeric(combined_point_est_df$se)
+combined_point_est_df$p_val <- as.numeric(combined_point_est_df$p_val)
+
+# subset out global method
+combined_point_est_df <- combined_point_est_df
+
+## ggplot
+print_plot <- ggplot(combined_point_est_df, 
+                     aes(x = pm_method, y = odds_ratio, colour = pm_method)) +
+  # custom color 
+  scale_color_manual(name = "Smoke-Estimation Method", 
+                     values = c("red", "blue", "#32115C"),
+                     guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+  geom_point() + #geom_text(vjust = 0, nudge_x = 0.3) +
+  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
+  facet_wrap(~outcome, nrow = 3) +
+  geom_hline(yintercept = 1, linetype=2) +
+  #    ggtitle('Association Between PM2.5 \n from Wildfire Smoke on Hospitalizations') +
+  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+  #xlab('Smoke Estimation Method') +
+  # plot theme
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        # strip element
+        strip.background = element_rect(colour=NA, fill=NA),
+        panel.border = element_rect(fill = NA, color = "black"),
+        # facet text size
+        #strip.text = element_text(size = 8),
+        # axis element
+        axis.text.x = element_blank(),
+        axis.title.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        axis.title.y = element_text(angle = 90),
+        # legend elements
+        legend.position = "bottom")
+#legend.text = element_text(size = 8))
+
+print(print_plot)
+# save figure
+ggsave("../plot_new/fig_ndc_overall_four.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+### Time-Stratified by Age Categories
+
+Following table and figure look at some outcomes stratified by age category.
+
+For some outcomes, such like Age > 65 in WRF-Chem method, the CI is the only insignificant one in the plot.
+
+```{r ndc time stratified age strata, warning =F, echo = F, results='asis'} 
+df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)
+
+outcome_list <- c("Asthma", "COPD", "Both", "Other")
+
+# create an empty list to row bind dataframes together
+datalist <- list()
+
+# outcome name
+outcome_name <- c("Asthma", "COPD", "Both", "Other")
+
+# age category list
+age_cat_list <- c(0,1,2)
+
+# looking only at wrf chem, kriging, and geo-weighted; taking out 'Global Smoke' method
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# create an empty list to row bind dataframes together
+datalist1 <- list()
+datalist2 <- list()
+
+for(i in 1:length(df_list)){
+
+  # dataframe to loop through
+  df_to_loop <- data.frame(df_list[i])
+  # indication of column
+  outcome <- df_list[i]
+  # outcome name
+  outcome_name <- outcome_list[i]
+
+  covariates_df <- df_to_loop[, c(1:26, 74:84)]
+
+  # extract pm values and divide by 10 and ordered
+  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+  pm_estimates_df <- df_to_loop[, c(87, 92, 91, 93)]/10  # create 10 unit increases
+
+  # new loop for age categories
+  for(k in 0:2){
+
+    point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)
+
+    colnames(point_estimates) <- c('outcome', 'age_cat', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+    # fill in the outcome namedataframe before method loop
+    point_estimates[, 1] <- outcome_name
+    # repeat the age category 4 times for each pm method
+    point_estimates[, 2] <- k
+
+    # dataframe for analysis creation
+    # bind columns back together 
+    df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+      # remove missing pm values
+      filter(!is.na(wrf_smk_pm_zip)) %>% 
+      # limit to specific age category
+      filter(age_ind == k) %>% 
+      # the following code makes sure that the counterfactual values retained are 
+      # symetric in that number of obs before = number of obs after
+      mutate(obs_diff_admission = (fromdate - date)/7)
+
+    for(j in 38:40){
+
+      # variable to model 
+      var_name <- colnames(df_analysis[j])
+
+      # set row number to fill
+      row_n <- j-37
+
+      # only run the model if the dataframe has observations
+      if(nrow(df_analysis) != 0){
+        # conditional logistic regression model
+        mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+        # populate matrix
+        point_estimates[row_n, 3] <- method_list[row_n]
+        point_estimates[row_n, 4] <- mod$n
+        point_estimates[row_n, 5] <- mod$nevent
+        # odds ratio
+        point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+        # 95% lower bound
+        point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
+        # 95% upper bound
+        point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                                 1.96*(summary(mod)$coefficient[1,3])), 3)
+        # standard error
+        point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
+        # p val
+        point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
+
+        # create else statement that fills matrix with missing so I still have the row
+        # in the final dataframe
+      } else {point_estimates[row_n, 3] <- method_list[row_n]
+      point_estimates[row_n, 4] <- 0
+      point_estimates[row_n, 5] <- 0
+      point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
+    } # end methods loop
+
+    # save point estimates as a dataframe
+    point_est_df <- as_data_frame(point_estimates)
+
+    # combine previous values in dataframe that has all outcome/methods comparisons
+    datalist1[[k+1]] <- point_est_df
+  } # end age category loop
+
+  # bind rows of age category estimates together
+  age_est_df <- bind_rows(datalist1)
+
+  # populate second dataframe list
+  datalist2[[i]] <- age_est_df
+
+} # end of outcome loop
+
+# combine each outcome dataframe itteration in to a big dataset
+combined_point_est_df <- bind_rows(datalist2)
+
+wrf_geo_age <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" |
+    #      pm_method == "WRF-Chem Smoke") %>% 
+  mutate(age_cat2 = ifelse(age_cat == 0, "Less than 15", 
+                    ifelse(age_cat == 1, "15-65",
+                    ifelse(age_cat == 2, "Greater than 65", NA)))) %>% 
+  # subset columns I want to put in to the table
+  select(3, 11, 5:8) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
+         caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by age",
+         # row group by outcome
+         rgroup = "Asthma inhalers",
+         n.rgroup = c(rep(9, 1)), # 6 rows for each age cat for each outcome
+         # column headers
+         header = c("Est Method", "Age Group", "Events",
+                    "OR&dagger;", "Lower", "Upper"),
+         # column spanner
+         cgroup = c("", "95% CI"), 
+         n.cgroup = c(4, 2),
+         padding.rgroup = "&nbsp;&nbsp;",
+         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+         align = "llcccc", # column alignment,
+         tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within May to September fire season."
+         ) # end table
+
+print(tab)
+
+combined_point_est_df_new <- combined_point_est_df %>%
+  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
+                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
+  select(-n_events) %>%
+  rename(n_events = n_events_new)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_age_plot <- combined_point_est_df_new %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  # do not plot results with less than 15 events, create missing vals
+  # in mutate
+  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
+                             99, odds_ratio),
+         lower95 = ifelse(as.numeric(n_events) < 15, 
+                             99, lower95),
+         upper95 = ifelse(as.numeric(n_events) < 15, 
+                             99, upper95),
+         age_cat2 = ifelse(age_cat == 0, "<15", 
+                    ifelse(age_cat == 1, "15-65",
+                    ifelse(age_cat == 2, ">65", NA)))) %>% 
+  # subset columns I want to put in to the table
+  select(1, 3, 11, 10, 5:7) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
+                              levels = unique(wrf_geo_age_plot$outcome))
+
+wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
+                                levels = unique(wrf_geo_age_plot$pm_method))
+
+wrf_geo_age_plot$age_cat2 <- factor(wrf_geo_age_plot$age_cat2,
+                                levels = unique(wrf_geo_age_plot$age_cat2))
+
+wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
+wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
+wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
+wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)
+
+## ggplot 
+  print_plot <- ggplot(wrf_geo_age_plot,
+    aes(x = age_cat2, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red",  "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Age Category ') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10), 
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
+  ggsave("../plot_new/ndc_age_fig_four.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+### Seasonal Time-Stratified by Sex
+
+Association with smoke PM~2.5~ and health outcomes stratified by sex.
+
+All the CI do not include 1, which means the use of asthma medication (beta 2 agonists) increases about 5% when smoke increase 10 units.
+So the risk for getting asthma will increase about 5% ~ 8% when the smoke PM2.5 increase 10 units.
+
+```{r ndc sex strata, warning = F, echo = F, results='asis'} 
+df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)
+
+outcome_list <- c("Asthma", "COPD", "Both", "Other")
+
+# outcome name
+outcome_name <- c("Asthma", "COPD", "Both", "Other")
+
+# sex category list
+sex_strata_list <- c(0,1)
+
+datalist1 <- list()
+datalist2 <- list()
+
+# data wrangling ----
+# Producing conditional logit model estimates loop 
+for(i in 1:length(df_list)){
+
+  # dataframe to loop through
+  df_to_loop <- data.frame(df_list[i])
+  # indication of column
+  outcome <- df_list[i]
+  # outcome name
+  outcome_name <- outcome_list[i]
+
+  covariates_df <- df_to_loop[, c(1:26, 74:84)]
+
+  # extract pm values and divide by 10 and ordered
+  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+  pm_estimates_df <- df_to_loop[, c(87, 92, 91, 93)]/10  # create 10 unit increases
+
+  # new loop for sex categories
+  for(k in 0:1){
+
+    # empty matrix (12 x 10 matrix)
+    point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)
+
+    colnames(point_estimates) <- c('outcome', 'sex', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+    # fill in the outcome namedataframe before method loop
+    point_estimates[, 1] <- outcome_name
+    # repeat the sex category 4 times for each pm method
+    point_estimates[, 2] <- k
+
+    # dataframe for analysis creation
+    # bind columns back together 
+    df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+      # remove missing pm values
+      filter(!is.na(wrf_smk_pm_zip)) %>% 
+      # limit to specific sex category
+      filter(sex_ind == k) %>% 
+      # the following code makes sure that the counterfactual values retained are 
+      # symetric in that number of obs before = number of obs after
+      mutate(obs_diff_admission = (fromdate - date)/7) 
+    # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+    for(j in 38:40){
+
+      # variable to model 
+      var_name <- colnames(df_analysis[j])
+
+      # set row number to fill
+      row_n <- j-37
+
+      # only run the model if the dataframe has observations
+      if(nrow(df_analysis) != 0){
+        # conditional logistic regression model
+        mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+        # populate matrix
+        point_estimates[row_n, 3] <- method_list[row_n]
+        point_estimates[row_n, 4] <- mod$n
+        point_estimates[row_n, 5] <- mod$nevent
+        # odds ratio
+        point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+        # 95% lower bound
+        point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
+                                                 1.96*(summary(mod)$coefficient[1,3])), 3)
+        # 95% upper bound
+        point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                                 1.96*(summary(mod)$coefficient[1,3])), 3)
+        # standard error
+        point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
+        # p val
+        point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
+
+        # create else statement that fills matrix with missing so I still have the row
+        # in the final dataframe
+      } else {point_estimates[row_n, 3] <- method_list[row_n]
+      point_estimates[row_n, 4] <- 0
+      point_estimates[row_n, 5] <- 0
+      point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
+    } # end methods loop
+
+    # save point estimates as a dataframe
+    point_est_df <- as_data_frame(point_estimates)
+
+    # combine previous values in dataframe that has all outcome/methods comparisons
+    datalist1[[k+1]] <- point_est_df
+  } # end sex category loop
+
+  # bind rows of sex category estimates together
+  sex_est_df <- bind_rows(datalist1)
+
+  # populate second dataframe list
+  datalist2[[i]] <- sex_est_df
+
+} # end of outcome loop
+
+# combine each outcome dataframe itteration in to a big dataset
+combined_point_est_df <- bind_rows(datalist2)
+
+wrf_geo_sex <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  mutate(sex_cat = ifelse(sex == 0, "Male", 
+               ifelse(sex == 1, "Female", NA))) %>% 
+  # subset columns I want to put in to the table
+  select(3, 11, 5:8)%>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+  tab <- htmlTable(txtRound(wrf_geo_sex, digits = 3, 1:3), 
+           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and pharmacy outcomes stratified by sex",
+           # row group by outcome
+           rgroup = "Asthma inhalers",
+           n.rgroup = c(rep(6, 1)), # 2 rows for each sex strata for each outcome
+           # column headers
+           header = c("Est Method", "Sex Strata", "Events",
+                      "OR&dagger;", "Lower", "Upper"),
+           # column spanner
+           cgroup = c("", "95% CI"), 
+           n.cgroup = c(4, 2),
+           padding.rgroup = "&nbsp;&nbsp;",
+           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+           align = "llcccc", # column alignment,
+           tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within July to October fire season."
+            ) # end table
+
+  print(tab)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_sex_plot <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | 
+   #       pm_method == "WRF-Chem Smoke") %>% 
+  mutate(sex_cat = ifelse(sex == 0, "Male", 
+               ifelse(sex == 1, "Female", NA))) %>%
+  # subset columns I want to put in to the table
+  select(1, 3, 11, 5:8) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_sex_plot$outcome <- factor(wrf_geo_sex_plot$outcome,
+                              levels = unique(wrf_geo_sex_plot$outcome))
+
+wrf_geo_sex_plot$pm_method <- factor(wrf_geo_sex_plot$pm_method,
+                                levels = unique(wrf_geo_sex_plot$pm_method))
+
+wrf_geo_sex_plot$sex_cat <- factor(wrf_geo_sex_plot$sex_cat,
+                                levels = unique(wrf_geo_sex_plot$sex_cat))
+
+wrf_geo_sex_plot$n_events <- as.numeric(wrf_geo_sex_plot$n_events)
+wrf_geo_sex_plot$odds_ratio <- as.numeric(wrf_geo_sex_plot$odds_ratio)
+wrf_geo_sex_plot$lower95 <- as.numeric(wrf_geo_sex_plot$lower95)
+wrf_geo_sex_plot$upper95 <- as.numeric(wrf_geo_sex_plot$upper95)
+
+## ggplot
+  print_plot <- ggplot(wrf_geo_sex_plot,
+    aes(x = sex_cat, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red", "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Sex') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10),
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
+  ggsave("../plot_new/ndc_sex_fig_four.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```

### 2017-06-13 r_scripts/analysis/or_dist_lag_model.Rmd
DLM-unfinished

---
title: "Oregon Distributed Lag Model"
author: "Jingyang Liu"
date: "June 12, 2017"
output: html_document
---

```{r library, include=FALSE, echo=FALSE}
library(readr)
library(dplyr)
library(lubridate)
library(splines)
library(ggplot2)
library(survival)
library(htmlTable)

```

### Set up for distributed lag model

```{r set up,  include=FALSE, echo=FALSE, eval=FALSE}
read_path0 <- paste0('C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire_new/data/Oregon_PM/zip_pm_to_merge_with_acap.csv')
zip_smoke <- read_csv(read_path0) # 63801 rows

zip_smoke_w_lag <- zip_smoke %>% arrange(ZIPCODE, date) %>%
  # group by zipcode
  group_by(ZIPCODE) %>% 
  # wrf
  mutate(wrf_f_pm_lag1 = lag(wrf_f_pm, 1, order_by = ZIPCODE), 
         wrf_f_pm_lag2 = lag(wrf_f_pm, 2, order_by = ZIPCODE),
         wrf_f_pm_lag3 = lag(wrf_f_pm, 3, order_by = ZIPCODE),
         wrf_f_pm_lag4 = lag(wrf_f_pm, 4, order_by = ZIPCODE),
         wrf_f_pm_lag5 = lag(wrf_f_pm, 5, order_by = ZIPCODE),
         wrf_f_pm_lag6 = lag(wrf_f_pm, 6, order_by = ZIPCODE),
         wrf_f_pm_lag7 = lag(wrf_f_pm, 7, order_by = ZIPCODE),
         # wrf no fire lag
         wrf_nf_pm_lag1 = lag(wrf_nf_pm, 1, order_by = ZIPCODE),
         wrf_nf_pm_lag2 = lag(wrf_nf_pm, 2, order_by = ZIPCODE),
         wrf_nf_pm_lag3 = lag(wrf_nf_pm, 3, order_by = ZIPCODE),
         wrf_nf_pm_lag4 = lag(wrf_nf_pm, 4, order_by = ZIPCODE),
         wrf_nf_pm_lag5 = lag(wrf_nf_pm, 5, order_by = ZIPCODE),
         wrf_nf_pm_lag6 = lag(wrf_nf_pm, 6, order_by = ZIPCODE),
         wrf_nf_pm_lag7 = lag(wrf_nf_pm, 7, order_by = ZIPCODE),
         # wrf_smk_pm
         wrf_smk_pm_lag1 = lag(wrf_smk_pm, 1, order_by = ZIPCODE),
         wrf_smk_pm_lag2 = lag(wrf_smk_pm, 2, order_by = ZIPCODE),
         wrf_smk_pm_lag3 = lag(wrf_smk_pm, 3, order_by = ZIPCODE),
         wrf_smk_pm_lag4 = lag(wrf_smk_pm, 4, order_by = ZIPCODE),
         wrf_smk_pm_lag5 = lag(wrf_smk_pm, 5, order_by = ZIPCODE),
         wrf_smk_pm_lag6 = lag(wrf_smk_pm, 6, order_by = ZIPCODE),
         wrf_smk_pm_lag7 = lag(wrf_smk_pm, 7, order_by = ZIPCODE),
         # geo weighted pm
         geo_wt_pm_lag1 = lag(geo_wt_pm, 1, order_by = ZIPCODE),
         geo_wt_pm_lag2 = lag(geo_wt_pm, 2, order_by = ZIPCODE),
         geo_wt_pm_lag3 = lag(geo_wt_pm, 3, order_by = ZIPCODE),
         geo_wt_pm_lag4 = lag(geo_wt_pm, 4, order_by = ZIPCODE),
         geo_wt_pm_lag5 = lag(geo_wt_pm, 5, order_by = ZIPCODE),
         geo_wt_pm_lag6 = lag(geo_wt_pm, 6, order_by = ZIPCODE),
         geo_wt_pm_lag7 = lag(geo_wt_pm, 7, order_by = ZIPCODE),
         # krig pm
         krig_pm_lag1 = lag(krig_pm, 1, order_by = ZIPCODE),
         krig_pm_lag2 = lag(krig_pm, 2, order_by = ZIPCODE),
         krig_pm_lag3 = lag(krig_pm, 3, order_by = ZIPCODE),
         krig_pm_lag4 = lag(krig_pm, 4, order_by = ZIPCODE),
         krig_pm_lag5 = lag(krig_pm, 5, order_by = ZIPCODE), 
         krig_pm_lag6 = lag(krig_pm, 6, order_by = ZIPCODE),
         krig_pm_lag7 = lag(krig_pm, 7, order_by = ZIPCODE), 
         # background pm
         background_pm_lag1 = lag(background_pm, 1, order_by = ZIPCODE),
         background_pm_lag2 = lag(background_pm, 2, order_by = ZIPCODE),
         background_pm_lag3 = lag(background_pm, 3, order_by = ZIPCODE),
         background_pm_lag4 = lag(background_pm, 4, order_by = ZIPCODE),
         background_pm_lag5 = lag(background_pm, 5, order_by = ZIPCODE), 
         background_pm_lag6 = lag(background_pm, 6, order_by = ZIPCODE),
         background_pm_lag7 = lag(background_pm, 7, order_by = ZIPCODE), 
         # geo_smk_pm 
         geo_smk_pm_lag1 = lag(geo_smk_pm, 1, order_by = ZIPCODE),
         geo_smk_pm_lag2 = lag(geo_smk_pm, 2, order_by = ZIPCODE),
         geo_smk_pm_lag3 = lag(geo_smk_pm, 3, order_by = ZIPCODE),
         geo_smk_pm_lag4 = lag(geo_smk_pm, 4, order_by = ZIPCODE),
         geo_smk_pm_lag5 = lag(geo_smk_pm, 5, order_by = ZIPCODE),
         geo_smk_pm_lag6 = lag(geo_smk_pm, 6, order_by = ZIPCODE),
         geo_smk_pm_lag7 = lag(geo_smk_pm, 7, order_by = ZIPCODE),
         # krig smk pm
         krig_smk_pm_lag1 = lag(krig_smk_pm, 1, order_by = ZIPCODE),
         krig_smk_pm_lag2 = lag(krig_smk_pm, 2, order_by = ZIPCODE),
         krig_smk_pm_lag3 = lag(krig_smk_pm, 3, order_by = ZIPCODE),
         krig_smk_pm_lag4 = lag(krig_smk_pm, 4, order_by = ZIPCODE),
         krig_smk_pm_lag5 = lag(krig_smk_pm, 5, order_by = ZIPCODE),
         krig_smk_pm_lag6 = lag(krig_smk_pm, 6, order_by = ZIPCODE),
         krig_smk_pm_lag7 = lag(krig_smk_pm, 7, order_by = ZIPCODE),
         # temp
         wrf_temp_lag1 = lag(wrf_temp, 1, order_by = ZIPCODE),
         wrf_temp_lag2 = lag(wrf_temp, 2, order_by = ZIPCODE),
         wrf_temp_lag3 = lag(wrf_temp, 3, order_by = ZIPCODE),
         wrf_temp_lag4 = lag(wrf_temp, 4, order_by = ZIPCODE),
         wrf_temp_lag5 = lag(wrf_temp, 5, order_by = ZIPCODE), 
         wrf_temp_lag6 = lag(wrf_temp, 6, order_by = ZIPCODE),
         wrf_temp_lag7 = lag(wrf_temp, 7, order_by = ZIPCODE)) %>% 
  # ungroup by zip
  ungroup(ZIPCODE) %>% 
  # attach a zip indicator for each smoke variable
  setNames(paste(colnames(.), "zip", sep="_"))  %>% 
  # remove the '_zip' from the zipcode and date variable 
  rename(ZIPCODE = ZIPCODE_zip, date = date_zip)

# set variable list
var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')

start <- Sys.time()
for(j in var_list) { # begin first loop of variable names (outcomes)

  # standard for statement without parallel computing
  #for(j in 1:length(var_list)){
  read_path <- paste('../data_new/casecrossover/or', j, 'may_to_sep_casecross.csv', sep='_')
  or_disease <- read_csv(read_path)

  ### try join
  colnames(or_disease)[24] <- c("ZIPCODE")
  colnames(or_disease)[74] <- c("date")

  outcome_casecross <- or_disease %>%
    # indicator for male=0, female=1, unknown = 2
    mutate(age = 2013-yob,
           sex_ind =ifelse(gender == "F", 1, 
                           ifelse(gender == "M", 0, 2)),
           age_ind = ifelse(age < 15, 0,
                            ifelse(age >= 15 & age < 65, 1,
                                   ifelse(age >= 65 & age <=105, 2, NA)))
    ) %>% # end of mutate 
    # create variables
    mutate(day = as.factor(weekdays(fromdate)),
           day_admit = as.factor(weekdays(date)),
           month_smk = month(fromdate),
           month_admit = month(date),
           season_smk = ifelse(fromdate >= "2013-03-20" &  
                                 fromdate <= "2013-06-21", "spring",
                               ifelse(fromdate >= "2013-06-22" &  
                                        fromdate <= "2013-09-22", "summer",
                                      ifelse(fromdate >= "2013-09-23" & 
                                               fromdate <= "2013-12-21", "fall", "other"))),
           season_admit = ifelse(date >= "2013-03-20" &  
                                   date <= "2013-06-21", "spring",
                                 ifelse(date >= "2013-06-22" &  
                                          date <= "2013-09-22", "summer",
                                        ifelse(date >= "2013-09-23" & 
                                                 date <= "2013-12-21", "fall", "other")))) %>%
    # join with zip-level pm estimates
    left_join(zip_smoke_w_lag, by = c("date", "ZIPCODE"))%>%
    arrange(personkey, fromdate) # order by id and fromdate

  # checks
  # glimpse(outcome_casecross)
  # which(colnames(outcome_casecross)=='geo_wt_pm_zip')
  # check <- outcome_casecross[, c(1:16, 20, 89, 111:133)]

  # Create a permanent case-cross over datasets
  file_name <- paste0('../data_new/casecrossover/dist_lag/or_', j, '_may_to_sep_dist_lag_casecross.csv')

  # write permanent dataset
  write_csv(outcome_casecross, file_name)

} # End of the overall loop

total_time <- Sys.time() - start
total_time #  9.131646 mins
```

### Distributed Lag Model

For asthma outcomes with 7 lags totally. Use GWR method.

```{r DLM}
path <- paste0("../data_new/casecrossover/dist_lag")

# asthma
asthma_casecross <- read_csv(paste(path, 
                                   "or_asthma_may_to_sep_dist_lag_casecross.csv", 
                                   sep = "/"))

asthma_lag <- asthma_casecross %>%
  filter(gender!="U") %>%
  filter(outcome==1) %>%
  select(89, 116:122) # GWR-method

which(colnames(asthma_casecross)=="geo_wt_pm_zip") # 89
which(colnames(asthma_casecross)=="geo_wt_pm_lag1_zip") #116

lag_max <- 7
n <- length(asthma_lag[,1])

asthma_lag <- as.matrix(asthma_lag)

# true DL function
DLfunc <- exp(-(0:lag_max)/2)

# simulate outcomes
y <- drop(asthma_lag%*%DLfunc) # + rnorm(n)

B <- ns(0:lag_max, df =4, intercept=TRUE)

# the nrow B should equal ncol(pm)
dim(B)
dim(asthma_lag)

asthma_lag_B <- asthma_lag%*%B
fit <- lm(y~asthma_lag_B)
coef(fit)

dlparms <- grep("asthma_lag_B",names(coef(fit)))
DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms])
DLvar <- B%*%vcov(fit)[dlparms,dlparms]%*%t(B)
DLestimate$SE <- sqrt(diag(DLvar))
DLestimate$lower <- DLestimate$estimate + DLestimate$SE * qt(1-0.975, df=fit$df.residual)
DLestimate$upper <- DLestimate$estimate + DLestimate$SE * qt(0.975, df=fit$df.residual)
DLestimate

DLestimate$lag <- 0:lag_max
DLestimate$truth <- DLfunc
p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
p <- p + geom_line(aes(x=lag, y=truth), linetype=2, size=2)
p <- p + theme_bw()
p

cumulative <- sum(DLestimate$estimate)
cumulative
## [1] 2.489717
cumulative_se <- sqrt(sum(DLvar))
cumulative_se
## [1] 0.0007685608
cumulative_CI <- cumulative + cumulative_se * qt(c(1-0.975,0.975), df=fit$df.residual)
cumulative_CI
## [1] 2.488209 2.491224

```

### Analysis

```{r load data, include=FALSE, echo=FALSE,}
resp_casecross <- read_csv(paste(path, 
                                 "or_respiratory_may_to_sep_dist_lag_casecross.csv", 
                                 sep = "/"))

# asthma
# asthma_casecross <- read_csv(paste(path, 
#                                    "or_asthma_may_to_sep_time_strat_casecross_both.csv", 
#                                    sep = "/"))

# copd 
copd_casecross <- read_csv(paste(path, 
                                 "or_copd_may_to_sep_dist_lag_casecross.csv", 
                                 sep = "/"))

# pneum or bronchitis
pneum_casecross <- read_csv(paste(path, 
                                  "or_pneumonia_may_to_sep_dist_lag_casecross.csv",
                                  sep="/"))
 # acute bronchitis
acute_bronch_casecross <- read_csv(paste(path, 
                                         "or_acute_bronch_may_to_sep_dist_lag_casecross.csv",
                                         sep = "/"))
# cvd
cvd_casecross <- read_csv(paste(path, 
                                "or_cvd_may_to_sep_dist_lag_casecross.csv",
                                sep="/"))
# arrhythmia
arrhythmia_casecross <- read_csv(paste(path, 
                                       "or_arrhythmia_may_to_sep_dist_lag_casecross.csv",
                                       sep="/"))
# cerebral vascular
cereb_vas_casecross <- read_csv(paste(path, 
                                      "or_cerbrovas_dis_may_to_sep_dist_lag_casecross.csv",
                                      sep="/"))
# heart failure
hf_casecross <- read_csv(paste(path, 
                               "or_heart_failure_may_to_sep_dist_lag_casecross.csv", 
                               sep="/"))
# ischemic heart disease
ihd_casecross <- read_csv(paste(path, 
                                "or_isch_heart_dis_may_to_sep_dist_lag_casecross.csv",
                                sep="/"))
# myo infarc
mi_casecross <- read_csv(paste(path, 
                               "or_myocardial_infarc_may_to_sep_dist_lag_casecross.csv", sep="/"))

# broken arm
broken_arm_casecross <- read_csv(paste(path, 
                                       "or_broken_arm_may_to_sep_dist_lag_casecross.csv",
                                       sep="/"))

```

```{r lag1}
df_list <- list(resp_casecross, asthma_casecross, copd_casecross, 
                pneum_casecross, acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
                cereb_vas_casecross, hf_casecross, ihd_casecross, mi_casecross, 
                broken_arm_casecross)

outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 
                  'Pneumonia', 'Acute Bronchitis', 'Cardiovascular Disease',
                  'Arrhythmia', 'Cerebrovascular Disease', 'Heart Failure',
                  'Ischemic Heart Disease', 'Myocardial Infarction',
                  'Broken Arm')

method_list <- c('WRF-Chem Smoke Lag 1', 'Kriging Smoke Lag 1', 'Geo-Weighted Smoke Lag 1')

# create an empty list to row bind dataframes together
datalist <- list()

# data wrangling ----
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){

  # dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- colnames(df_to_loop[76])
  # outcome name
  outcome_name <- outcome_list[i]

  # extract covariates from dataframe
  covariates_df <- df_to_loop[, c(1:26, 71, 74, 76:85)]

  # extract pm values and divide by 10 and ordered
  which(colnames(df_to_loop)=="wrf_smk_pm_lag1_zip") # 109
  which(colnames(df_to_loop)=="krig_pm_lag1_zip") # 123
  which(colnames(df_to_loop)=="geo_wt_pm_lag1_zip") # code to find column numbers, 116
  pm_estimates_df <- df_to_loop[, c(109,123,116, 94)]/10  # create 10 unit increases

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- cbind(covariates_df, pm_estimates_df) %>%
    # remove missing pm values
    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
    filter(!is.na(krig_pm_lag1_zip)) %>% 
    filter(!is.na(geo_wt_pm_lag1_zip)) %>%

    # join in cdc met data
    # full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
    # the following code makes sure that the counterfactual values retained are 
    # symetric in that number of obs before = number of obs after
    mutate(obs_diff_admission = (fromdate - date)/7) 
  # dataframe is already for the entire fire season, so I don't need to subset anymore

  # empty df for table
  table_df <- data.frame()

  # empty matrix
  point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)

  colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 'odds_ratio', 
                                 'lower95', 'upper95', 'se', 'p_val')

  # fill in the outcome name for the dataframe before the loop
  point_estimates[, 1] <- outcome_name

  # second loop to run a model for each pm estimation method
  for(j in 39:41){

    # variable to model 
    var_name <- colnames(df_analysis[j])

    # conditional logistic regression model
    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)

    # some model/DAG checks of data rish provided

    #glimpse(df_analysis)

    # adjusting for admit month is not possible as it's the same var for the subject.
    # adjusting for month smoke is a collider and should not be adjusted for.
    #summary(clogit(outcome ~ geo_smk_pm_zip + daily_meanRH + daily_meanT + daily_meanWS + daily_meanPrcp +
    #                as.factor(month_smk) + strata(PATIENTID), df_analysis))
    #outocme association: daily_meanRH,

    # check for confoudning following DAG assumptions
    #summary(clogit(outcome ~ wrf_temp_zip + strata(PATIENTID), df_analysis))
    #summary(clogit(outcome ~ wrf_temp_zip + strata(PATIENTID), df_analysis))
    # smoke predictor association: daily_meanRH, 
    #summary(lmer(geo_smk_pm_zip ~ daily_meanWS + (1 | PATIENTID), df_analysis ))
    #summary(lmer(wrf_temp_zip ~ daily_meanT + (1 | PATIENTID), df_analysis ))
    #cor(df_analysis$wrf_temp_county, df_analysis$daily_meanT)

    # populate matrix
    row_n <- j-38

    point_estimates[row_n, 2] <- method_list[row_n]
    point_estimates[row_n, 3] <- mod$n
    point_estimates[row_n, 4] <- mod$nevent
    # odds ratio
    point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)

    # 95% lower bound
    point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -
                                             1.96*(summary(mod)$coefficient[1,3])), 3)
    # 95% upper bound
    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
                                             1.96*(summary(mod)$coefficient[1,3])), 3)
    # standard error
    point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
    # p val
    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)

    # save point estimates as a dataframe
    point_est_df <- as_data_frame(point_estimates)

  }

  # combine previous values in dataframe that has all outcome/methods comparisons
  datalist[[i]] <- point_est_df

} # end of loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist)

# subset columns I want to put in to the table
table_df <- combined_point_est_df %>% select(2, 3:7)

tab <- htmlTable(txtRound(table_df, digits = 3, 1:3), 
                 caption = "Association between a 10 ug/m^3 in PM2.5 and Health Outcomes",
                 # row group by outcome
                 rgroup = outcome_list,
                 n.rgroup = c(rep(3, 12)), # 4 rows for each method for each outcome
                 # column headers
                 header = c("Method", "Obs.", "Events",
                            "OR&dagger;", "Lower", "Upper"),
                 # column spanner
                 cgroup = c("", "95% CI"), 
                 n.cgroup = c(4, 2),
                 padding.rgroup = "&nbsp;&nbsp;",
                 css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
                 align = "llccccc", # column alignment,
                 tfoot="&dagger; Adjusted for CDC temperature, relatively humidity, wind speed, and precipitation; accounting for subject. Time-stratified: referent periods matched to events on same day of week within July to October fire season."
) # end table

print(tab)

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
combined_point_est_df$outcome <- factor(combined_point_est_df$outcome, 
                                        levels = unique(combined_point_est_df$outcome))

combined_point_est_df$pm_method <- factor(combined_point_est_df$pm_method, 
                                          levels = unique(combined_point_est_df$pm_method))

combined_point_est_df$n <- as.numeric(combined_point_est_df$n)
combined_point_est_df$n_events <- as.numeric(combined_point_est_df$n_events)
combined_point_est_df$odds_ratio <- as.numeric(combined_point_est_df$odds_ratio)
combined_point_est_df$lower95 <- as.numeric(combined_point_est_df$lower95)
combined_point_est_df$upper95 <- as.numeric(combined_point_est_df$upper95)
combined_point_est_df$se <- as.numeric(combined_point_est_df$se)
combined_point_est_df$p_val <- as.numeric(combined_point_est_df$p_val)

# subset out global method
combined_point_est_df <- combined_point_est_df

## ggplot
print_plot <- ggplot(combined_point_est_df, 
                     aes(x = pm_method, y = odds_ratio, colour = pm_method)) +
  # custom color 
  scale_color_manual(name = "Smoke-Estimation Method", 
                     values = c("red", "blue", "#32115C"),
                     guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
  geom_point() + #geom_text(vjust = 0, nudge_x = 0.3) +
  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
  facet_wrap(~outcome, nrow = 3) +
  geom_hline(yintercept = 1, linetype=2) +
  #    ggtitle('Association Between PM2.5 \n from Wildfire Smoke on Hospitalizations') +
  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
  #xlab('Smoke Estimation Method') +
  # plot theme
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        #strip.text = element_text(size = 8),
        # axis element
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(angle = 90),
        # legend elements
        legend.position = "bottom")
#legend.text = element_text(size = 8))

print(print_plot)
# save figure
ggsave("../plot_lag/fig_lag1.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

```

```{r lag2}
df_list <- list(resp_casecross, asthma_casecross, copd_casecross, 
                pneum_casecross, acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
                cereb_vas_casecross, hf_casecross, ihd_casecross, mi_casecross, 
                broken_arm_casecross)

outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 
                  'Pneumonia', 'Acute Bronchitis', 'Cardiovascular Disease',
                  'Arrhythmia', 'Cerebrovascular Disease', 'Heart Failure',
                  'Ischemic Heart Disease', 'Myocardial Infarction',
                  'Broken Arm')

method_list <- c('WRF-Chem Smoke Lag 2', 'Kriging Smoke Lag 2', 'Geo-Weighted Smoke Lag 2')

# create an empty list to row bind dataframes together
datalist <- list()

# data wrangling ----
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){

  # dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- colnames(df_to_loop[76])
  # outcome name
  outcome_name <- outcome_list[i]

  # extract covariates from dataframe
  covariates_df <- df_to_loop[, c(1:26, 71, 74, 76:85)]

  # extract pm values and divide by 10 and ordered
  which(colnames(df_to_loop)=="wrf_smk_pm_lag2_zip") # 109
  which(colnames(df_to_loop)=="krig_pm_lag2_zip") # 123
  which(colnames(df_to_loop)=="geo_wt_pm_lag2_zip") # code to find column numbers, 116
  pm_estimates_df <- df_to_loop[, c(110,124,117, 94)]/10  # create 10 unit increases

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- cbind(covariates_df, pm_estimates_df) %>%
    # remove missing pm values
    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
    filter(!is.na(krig_pm_lag1_zip)) %>% 
    filter(!is.na(geo_wt_pm_lag1_zip)) %>%

    # join in cdc met data
    # full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
    # the following code makes sure that the counterfactual values retained are 
    # symetric in that number of obs before = number of obs after
    mutate(obs_diff_admission = (fromdate - date)/7) 
  # dataframe is already for the entire fire season, so I don't need to subset anymore

  # empty df for table
  table_df <- data.frame()

  # empty matrix
  point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)

  colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 'odds_ratio', 
                                 'lower95', 'upper95', 'se', 'p_val')

  # fill in the outcome name for the dataframe before the loop
  point_estimates[, 1] <- outcome_name

  # second loop to run a model for each pm estimation method
  for(j in 39:41){

    # variable to model 
    var_name <- colnames(df_analysis[j])

    # conditional logistic regression model
    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)

    # some model/DAG checks of data rish provided

    #glimpse(df_analysis)

    # adjusting for admit month is not possible as it's the same var for the subject.
    # adjusting for month smoke is a collider and should not be adjusted for.
    #summary(clogit(outcome ~ geo_smk_pm_zip + daily_meanRH + daily_meanT + daily_meanWS + daily_meanPrcp +
    #                as.factor(month_smk) + strata(PATIENTID), df_analysis))
    #outocme association: daily_meanRH,

    # check for confoudning following DAG assumptions
    #summary(clogit(outcome ~ wrf_temp_zip + strata(PATIENTID), df_analysis))
    #summary(clogit(outcome ~ wrf_temp_zip + strata(PATIENTID), df_analysis))
    # smoke predictor association: daily_meanRH, 
    #summary(lmer(geo_smk_pm_zip ~ daily_meanWS + (1 | PATIENTID), df_analysis ))
    #summary(lmer(wrf_temp_zip ~ daily_meanT + (1 | PATIENTID), df_analysis ))
    #cor(df_analysis$wrf_temp_county, df_analysis$daily_meanT)

    # populate matrix
    row_n <- j-38

    point_estimates[row_n, 2] <- method_list[row_n]
    point_estimates[row_n, 3] <- mod$n
    point_estimates[row_n, 4] <- mod$nevent
    # odds ratio
    point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)

    # 95% lower bound
    point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -
                                             1.96*(summary(mod)$coefficient[1,3])), 3)
    # 95% upper bound
    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
                                             1.96*(summary(mod)$coefficient[1,3])), 3)
    # standard error
    point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
    # p val
    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)

    # save point estimates as a dataframe
    point_est_df <- as_data_frame(point_estimates)

  }

  # combine previous values in dataframe that has all outcome/methods comparisons
  datalist[[i]] <- point_est_df

} # end of loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist)

# subset columns I want to put in to the table
table_df <- combined_point_est_df %>% select(2, 3:7)

tab <- htmlTable(txtRound(table_df, digits = 3, 1:3), 
                 caption = "Association between a 10 ug/m^3 in PM2.5 and Health Outcomes",
                 # row group by outcome
                 rgroup = outcome_list,
                 n.rgroup = c(rep(3, 12)), # 4 rows for each method for each outcome
                 # column headers
                 header = c("Method", "Obs.", "Events",
                            "OR&dagger;", "Lower", "Upper"),
                 # column spanner
                 cgroup = c("", "95% CI"), 
                 n.cgroup = c(4, 2),
                 padding.rgroup = "&nbsp;&nbsp;",
                 css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
                 align = "llccccc", # column alignment,
                 tfoot="&dagger; Adjusted for CDC temperature, relatively humidity, wind speed, and precipitation; accounting for subject. Time-stratified: referent periods matched to events on same day of week within July to October fire season."
) # end table

print(tab)

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
combined_point_est_df$outcome <- factor(combined_point_est_df$outcome, 
                                        levels = unique(combined_point_est_df$outcome))

combined_point_est_df$pm_method <- factor(combined_point_est_df$pm_method, 
                                          levels = unique(combined_point_est_df$pm_method))

combined_point_est_df$n <- as.numeric(combined_point_est_df$n)
combined_point_est_df$n_events <- as.numeric(combined_point_est_df$n_events)
combined_point_est_df$odds_ratio <- as.numeric(combined_point_est_df$odds_ratio)
combined_point_est_df$lower95 <- as.numeric(combined_point_est_df$lower95)
combined_point_est_df$upper95 <- as.numeric(combined_point_est_df$upper95)
combined_point_est_df$se <- as.numeric(combined_point_est_df$se)
combined_point_est_df$p_val <- as.numeric(combined_point_est_df$p_val)

# subset out global method
combined_point_est_df <- combined_point_est_df

## ggplot
print_plot <- ggplot(combined_point_est_df, 
                     aes(x = pm_method, y = odds_ratio, colour = pm_method)) +
  # custom color 
  scale_color_manual(name = "Smoke-Estimation Method", 
                     values = c("red", "blue", "#32115C"),
                     guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
  geom_point() + #geom_text(vjust = 0, nudge_x = 0.3) +
  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
  facet_wrap(~outcome, nrow = 3) +
  geom_hline(yintercept = 1, linetype=2) +
  #    ggtitle('Association Between PM2.5 \n from Wildfire Smoke on Hospitalizations') +
  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
  #xlab('Smoke Estimation Method') +
  # plot theme
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        #strip.text = element_text(size = 8),
        # axis element
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(angle = 90),
        # legend elements
        legend.position = "bottom")
#legend.text = element_text(size = 8))

print(print_plot)
# save figure
ggsave("../plot_lag/fig_lag1.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

```

```{r lag3}
df_list <- list(resp_casecross, asthma_casecross, copd_casecross, 
                pneum_casecross, acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
                cereb_vas_casecross, hf_casecross, ihd_casecross, mi_casecross, 
                broken_arm_casecross)

outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 
                  'Pneumonia', 'Acute Bronchitis', 'Cardiovascular Disease',
                  'Arrhythmia', 'Cerebrovascular Disease', 'Heart Failure',
                  'Ischemic Heart Disease', 'Myocardial Infarction',
                  'Broken Arm')

method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')

# create an empty list to row bind dataframes together
datalist <- list()

# data wrangling ----
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){

  # dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- colnames(df_to_loop[76])
  # outcome name
  outcome_name <- outcome_list[i]

  # extract covariates from dataframe
  covariates_df <- df_to_loop[, c(1:26, 71, 74, 76:85)]

  # extract pm values and divide by 10 and ordered
  which(colnames(df_to_loop)=="wrf_smk_pm_lag1_zip") # 109
  which(colnames(df_to_loop)=="krig_pm_lag1_zip") # 123
  which(colnames(df_to_loop)=="geo_wt_pm_lag1_zip") # code to find column numbers, 116
  pm_estimates_df <- df_to_loop[, c(109,123,116, 94)]/10  # create 10 unit increases

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- cbind(covariates_df, pm_estimates_df) %>%
    # remove missing pm values
    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
    filter(!is.na(krig_pm_lag1_zip)) %>% 
    filter(!is.na(geo_wt_pm_lag1_zip)) %>%

    # join in cdc met data
    # full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
    # the following code makes sure that the counterfactual values retained are 
    # symetric in that number of obs before = number of obs after
    mutate(obs_diff_admission = (fromdate - date)/7) 
  # dataframe is already for the entire fire season, so I don't need to subset anymore

  # empty df for table
  table_df <- data.frame()

  # empty matrix
  point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)

  colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 'odds_ratio', 
                                 'lower95', 'upper95', 'se', 'p_val')

  # fill in the outcome name for the dataframe before the loop
  point_estimates[, 1] <- outcome_name

  # second loop to run a model for each pm estimation method
  for(j in 39:41){

    # variable to model 
    var_name <- colnames(df_analysis[j])

    # conditional logistic regression model
    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)

    # some model/DAG checks of data rish provided

    #glimpse(df_analysis)

    # adjusting for admit month is not possible as it's the same var for the subject.
    # adjusting for month smoke is a collider and should not be adjusted for.
    #summary(clogit(outcome ~ geo_smk_pm_zip + daily_meanRH + daily_meanT + daily_meanWS + daily_meanPrcp +
    #                as.factor(month_smk) + strata(PATIENTID), df_analysis))
    #outocme association: daily_meanRH,

    # check for confoudning following DAG assumptions
    #summary(clogit(outcome ~ wrf_temp_zip + strata(PATIENTID), df_analysis))
    #summary(clogit(outcome ~ wrf_temp_zip + strata(PATIENTID), df_analysis))
    # smoke predictor association: daily_meanRH, 
    #summary(lmer(geo_smk_pm_zip ~ daily_meanWS + (1 | PATIENTID), df_analysis ))
    #summary(lmer(wrf_temp_zip ~ daily_meanT + (1 | PATIENTID), df_analysis ))
    #cor(df_analysis$wrf_temp_county, df_analysis$daily_meanT)

    # populate matrix
    row_n <- j-38

    point_estimates[row_n, 2] <- method_list[row_n]
    point_estimates[row_n, 3] <- mod$n
    point_estimates[row_n, 4] <- mod$nevent
    # odds ratio
    point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)

    # 95% lower bound
    point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -
                                             1.96*(summary(mod)$coefficient[1,3])), 3)
    # 95% upper bound
    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
                                             1.96*(summary(mod)$coefficient[1,3])), 3)
    # standard error
    point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
    # p val
    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)

    # save point estimates as a dataframe
    point_est_df <- as_data_frame(point_estimates)

  }

  # combine previous values in dataframe that has all outcome/methods comparisons
  datalist[[i]] <- point_est_df

} # end of loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist)

# subset columns I want to put in to the table
table_df <- combined_point_est_df %>% select(2, 3:7)

tab <- htmlTable(txtRound(table_df, digits = 3, 1:3), 
                 caption = "Association between a 10 ug/m^3 in PM2.5 and Health Outcomes",
                 # row group by outcome
                 rgroup = outcome_list,
                 n.rgroup = c(rep(3, 12)), # 4 rows for each method for each outcome
                 # column headers
                 header = c("Method", "Obs.", "Events",
                            "OR&dagger;", "Lower", "Upper"),
                 # column spanner
                 cgroup = c("", "95% CI"), 
                 n.cgroup = c(4, 2),
                 padding.rgroup = "&nbsp;&nbsp;",
                 css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
                 align = "llccccc", # column alignment,
                 tfoot="&dagger; Adjusted for CDC temperature, relatively humidity, wind speed, and precipitation; accounting for subject. Time-stratified: referent periods matched to events on same day of week within July to October fire season."
) # end table

print(tab)

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
combined_point_est_df$outcome <- factor(combined_point_est_df$outcome, 
                                        levels = unique(combined_point_est_df$outcome))

combined_point_est_df$pm_method <- factor(combined_point_est_df$pm_method, 
                                          levels = unique(combined_point_est_df$pm_method))

combined_point_est_df$n <- as.numeric(combined_point_est_df$n)
combined_point_est_df$n_events <- as.numeric(combined_point_est_df$n_events)
combined_point_est_df$odds_ratio <- as.numeric(combined_point_est_df$odds_ratio)
combined_point_est_df$lower95 <- as.numeric(combined_point_est_df$lower95)
combined_point_est_df$upper95 <- as.numeric(combined_point_est_df$upper95)
combined_point_est_df$se <- as.numeric(combined_point_est_df$se)
combined_point_est_df$p_val <- as.numeric(combined_point_est_df$p_val)

# subset out global method
combined_point_est_df <- combined_point_est_df

## ggplot
print_plot <- ggplot(combined_point_est_df, 
                     aes(x = pm_method, y = odds_ratio, colour = pm_method)) +
  # custom color 
  scale_color_manual(name = "Smoke-Estimation Method", 
                     values = c("red", "blue", "#32115C"),
                     guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
  geom_point() + #geom_text(vjust = 0, nudge_x = 0.3) +
  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
  facet_wrap(~outcome, nrow = 3) +
  geom_hline(yintercept = 1, linetype=2) +
  #    ggtitle('Association Between PM2.5 \n from Wildfire Smoke on Hospitalizations') +
  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
  #xlab('Smoke Estimation Method') +
  # plot theme
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        #strip.text = element_text(size = 8),
        # axis element
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(angle = 90),
        # legend elements
        legend.position = "bottom")
#legend.text = element_text(size = 8))

print(print_plot)
# save figure
ggsave("../plot_lag/fig_lag1.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

```

```{r lag4}
df_list <- list(resp_casecross, asthma_casecross, copd_casecross, 
                pneum_casecross, acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
                cereb_vas_casecross, hf_casecross, ihd_casecross, mi_casecross, 
                broken_arm_casecross)

outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 
                  'Pneumonia', 'Acute Bronchitis', 'Cardiovascular Disease',
                  'Arrhythmia', 'Cerebrovascular Disease', 'Heart Failure',
                  'Ischemic Heart Disease', 'Myocardial Infarction',
                  'Broken Arm')

method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')

# create an empty list to row bind dataframes together
datalist <- list()

# data wrangling ----
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){

  # dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- colnames(df_to_loop[76])
  # outcome name
  outcome_name <- outcome_list[i]

  # extract covariates from dataframe
  covariates_df <- df_to_loop[, c(1:26, 71, 74, 76:85)]

  # extract pm values and divide by 10 and ordered
  which(colnames(df_to_loop)=="wrf_smk_pm_lag1_zip") # 109
  which(colnames(df_to_loop)=="krig_pm_lag1_zip") # 123
  which(colnames(df_to_loop)=="geo_wt_pm_lag1_zip") # code to find column numbers, 116
  pm_estimates_df <- df_to_loop[, c(109,123,116, 94)]/10  # create 10 unit increases

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- cbind(covariates_df, pm_estimates_df) %>%
    # remove missing pm values
    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
    filter(!is.na(krig_pm_lag1_zip)) %>% 
    filter(!is.na(geo_wt_pm_lag1_zip)) %>%

    # join in cdc met data
    # full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
    # the following code makes sure that the counterfactual values retained are 
    # symetric in that number of obs before = number of obs after
    mutate(obs_diff_admission = (fromdate - date)/7) 
  # dataframe is already for the entire fire season, so I don't need to subset anymore

  # empty df for table
  table_df <- data.frame()

  # empty matrix
  point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)

  colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 'odds_ratio', 
                                 'lower95', 'upper95', 'se', 'p_val')

  # fill in the outcome name for the dataframe before the loop
  point_estimates[, 1] <- outcome_name

  # second loop to run a model for each pm estimation method
  for(j in 39:41){

    # variable to model 
    var_name <- colnames(df_analysis[j])

    # conditional logistic regression model
    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)

    # some model/DAG checks of data rish provided

    #glimpse(df_analysis)

    # adjusting for admit month is not possible as it's the same var for the subject.
    # adjusting for month smoke is a collider and should not be adjusted for.
    #summary(clogit(outcome ~ geo_smk_pm_zip + daily_meanRH + daily_meanT + daily_meanWS + daily_meanPrcp +
    #                as.factor(month_smk) + strata(PATIENTID), df_analysis))
    #outocme association: daily_meanRH,

    # check for confoudning following DAG assumptions
    #summary(clogit(outcome ~ wrf_temp_zip + strata(PATIENTID), df_analysis))
    #summary(clogit(outcome ~ wrf_temp_zip + strata(PATIENTID), df_analysis))
    # smoke predictor association: daily_meanRH, 
    #summary(lmer(geo_smk_pm_zip ~ daily_meanWS + (1 | PATIENTID), df_analysis ))
    #summary(lmer(wrf_temp_zip ~ daily_meanT + (1 | PATIENTID), df_analysis ))
    #cor(df_analysis$wrf_temp_county, df_analysis$daily_meanT)

    # populate matrix
    row_n <- j-38

    point_estimates[row_n, 2] <- method_list[row_n]
    point_estimates[row_n, 3] <- mod$n
    point_estimates[row_n, 4] <- mod$nevent
    # odds ratio
    point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)

    # 95% lower bound
    point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -
                                             1.96*(summary(mod)$coefficient[1,3])), 3)
    # 95% upper bound
    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
                                             1.96*(summary(mod)$coefficient[1,3])), 3)
    # standard error
    point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
    # p val
    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)

    # save point estimates as a dataframe
    point_est_df <- as_data_frame(point_estimates)

  }

  # combine previous values in dataframe that has all outcome/methods comparisons
  datalist[[i]] <- point_est_df

} # end of loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist)

# subset columns I want to put in to the table
table_df <- combined_point_est_df %>% select(2, 3:7)

tab <- htmlTable(txtRound(table_df, digits = 3, 1:3), 
                 caption = "Association between a 10 ug/m^3 in PM2.5 and Health Outcomes",
                 # row group by outcome
                 rgroup = outcome_list,
                 n.rgroup = c(rep(3, 12)), # 4 rows for each method for each outcome
                 # column headers
                 header = c("Method", "Obs.", "Events",
                            "OR&dagger;", "Lower", "Upper"),
                 # column spanner
                 cgroup = c("", "95% CI"), 
                 n.cgroup = c(4, 2),
                 padding.rgroup = "&nbsp;&nbsp;",
                 css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
                 align = "llccccc", # column alignment,
                 tfoot="&dagger; Adjusted for CDC temperature, relatively humidity, wind speed, and precipitation; accounting for subject. Time-stratified: referent periods matched to events on same day of week within July to October fire season."
) # end table

print(tab)

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
combined_point_est_df$outcome <- factor(combined_point_est_df$outcome, 
                                        levels = unique(combined_point_est_df$outcome))

combined_point_est_df$pm_method <- factor(combined_point_est_df$pm_method, 
                                          levels = unique(combined_point_est_df$pm_method))

combined_point_est_df$n <- as.numeric(combined_point_est_df$n)
combined_point_est_df$n_events <- as.numeric(combined_point_est_df$n_events)
combined_point_est_df$odds_ratio <- as.numeric(combined_point_est_df$odds_ratio)
combined_point_est_df$lower95 <- as.numeric(combined_point_est_df$lower95)
combined_point_est_df$upper95 <- as.numeric(combined_point_est_df$upper95)
combined_point_est_df$se <- as.numeric(combined_point_est_df$se)
combined_point_est_df$p_val <- as.numeric(combined_point_est_df$p_val)

# subset out global method
combined_point_est_df <- combined_point_est_df

## ggplot
print_plot <- ggplot(combined_point_est_df, 
                     aes(x = pm_method, y = odds_ratio, colour = pm_method)) +
  # custom color 
  scale_color_manual(name = "Smoke-Estimation Method", 
                     values = c("red", "blue", "#32115C"),
                     guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
  geom_point() + #geom_text(vjust = 0, nudge_x = 0.3) +
  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
  facet_wrap(~outcome, nrow = 3) +
  geom_hline(yintercept = 1, linetype=2) +
  #    ggtitle('Association Between PM2.5 \n from Wildfire Smoke on Hospitalizations') +
  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
  #xlab('Smoke Estimation Method') +
  # plot theme
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        #strip.text = element_text(size = 8),
        # axis element
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(angle = 90),
        # legend elements
        legend.position = "bottom")
#legend.text = element_text(size = 8))

print(print_plot)
# save figure
ggsave("../plot_lag/fig_lag1.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

```

```{r lag5}
df_list <- list(resp_casecross, asthma_casecross, copd_casecross, 
                pneum_casecross, acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
                cereb_vas_casecross, hf_casecross, ihd_casecross, mi_casecross, 
                broken_arm_casecross)

outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 
                  'Pneumonia', 'Acute Bronchitis', 'Cardiovascular Disease',
                  'Arrhythmia', 'Cerebrovascular Disease', 'Heart Failure',
                  'Ischemic Heart Disease', 'Myocardial Infarction',
                  'Broken Arm')

method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')

# create an empty list to row bind dataframes together
datalist <- list()

# data wrangling ----
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){

  # dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- colnames(df_to_loop[76])
  # outcome name
  outcome_name <- outcome_list[i]

  # extract covariates from dataframe
  covariates_df <- df_to_loop[, c(1:26, 71, 74, 76:85)]

  # extract pm values and divide by 10 and ordered
  which(colnames(df_to_loop)=="wrf_smk_pm_lag1_zip") # 109
  which(colnames(df_to_loop)=="krig_pm_lag1_zip") # 123
  which(colnames(df_to_loop)=="geo_wt_pm_lag1_zip") # code to find column numbers, 116
  pm_estimates_df <- df_to_loop[, c(109,123,116, 94)]/10  # create 10 unit increases

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- cbind(covariates_df, pm_estimates_df) %>%
    # remove missing pm values
    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
    filter(!is.na(krig_pm_lag1_zip)) %>% 
    filter(!is.na(geo_wt_pm_lag1_zip)) %>%

    # join in cdc met data
    # full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
    # the following code makes sure that the counterfactual values retained are 
    # symetric in that number of obs before = number of obs after
    mutate(obs_diff_admission = (fromdate - date)/7) 
  # dataframe is already for the entire fire season, so I don't need to subset anymore

  # empty df for table
  table_df <- data.frame()

  # empty matrix
  point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)

  colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 'odds_ratio', 
                                 'lower95', 'upper95', 'se', 'p_val')

  # fill in the outcome name for the dataframe before the loop
  point_estimates[, 1] <- outcome_name

  # second loop to run a model for each pm estimation method
  for(j in 39:41){

    # variable to model 
    var_name <- colnames(df_analysis[j])

    # conditional logistic regression model
    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)

    # some model/DAG checks of data rish provided

    #glimpse(df_analysis)

    # adjusting for admit month is not possible as it's the same var for the subject.
    # adjusting for month smoke is a collider and should not be adjusted for.
    #summary(clogit(outcome ~ geo_smk_pm_zip + daily_meanRH + daily_meanT + daily_meanWS + daily_meanPrcp +
    #                as.factor(month_smk) + strata(PATIENTID), df_analysis))
    #outocme association: daily_meanRH,

    # check for confoudning following DAG assumptions
    #summary(clogit(outcome ~ wrf_temp_zip + strata(PATIENTID), df_analysis))
    #summary(clogit(outcome ~ wrf_temp_zip + strata(PATIENTID), df_analysis))
    # smoke predictor association: daily_meanRH, 
    #summary(lmer(geo_smk_pm_zip ~ daily_meanWS + (1 | PATIENTID), df_analysis ))
    #summary(lmer(wrf_temp_zip ~ daily_meanT + (1 | PATIENTID), df_analysis ))
    #cor(df_analysis$wrf_temp_county, df_analysis$daily_meanT)

    # populate matrix
    row_n <- j-38

    point_estimates[row_n, 2] <- method_list[row_n]
    point_estimates[row_n, 3] <- mod$n
    point_estimates[row_n, 4] <- mod$nevent
    # odds ratio
    point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)

    # 95% lower bound
    point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -
                                             1.96*(summary(mod)$coefficient[1,3])), 3)
    # 95% upper bound
    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
                                             1.96*(summary(mod)$coefficient[1,3])), 3)
    # standard error
    point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
    # p val
    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)

    # save point estimates as a dataframe
    point_est_df <- as_data_frame(point_estimates)

  }

  # combine previous values in dataframe that has all outcome/methods comparisons
  datalist[[i]] <- point_est_df

} # end of loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist)

# subset columns I want to put in to the table
table_df <- combined_point_est_df %>% select(2, 3:7)

tab <- htmlTable(txtRound(table_df, digits = 3, 1:3), 
                 caption = "Association between a 10 ug/m^3 in PM2.5 and Health Outcomes",
                 # row group by outcome
                 rgroup = outcome_list,
                 n.rgroup = c(rep(3, 12)), # 4 rows for each method for each outcome
                 # column headers
                 header = c("Method", "Obs.", "Events",
                            "OR&dagger;", "Lower", "Upper"),
                 # column spanner
                 cgroup = c("", "95% CI"), 
                 n.cgroup = c(4, 2),
                 padding.rgroup = "&nbsp;&nbsp;",
                 css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
                 align = "llccccc", # column alignment,
                 tfoot="&dagger; Adjusted for CDC temperature, relatively humidity, wind speed, and precipitation; accounting for subject. Time-stratified: referent periods matched to events on same day of week within July to October fire season."
) # end table

print(tab)

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
combined_point_est_df$outcome <- factor(combined_point_est_df$outcome, 
                                        levels = unique(combined_point_est_df$outcome))

combined_point_est_df$pm_method <- factor(combined_point_est_df$pm_method, 
                                          levels = unique(combined_point_est_df$pm_method))

combined_point_est_df$n <- as.numeric(combined_point_est_df$n)
combined_point_est_df$n_events <- as.numeric(combined_point_est_df$n_events)
combined_point_est_df$odds_ratio <- as.numeric(combined_point_est_df$odds_ratio)
combined_point_est_df$lower95 <- as.numeric(combined_point_est_df$lower95)
combined_point_est_df$upper95 <- as.numeric(combined_point_est_df$upper95)
combined_point_est_df$se <- as.numeric(combined_point_est_df$se)
combined_point_est_df$p_val <- as.numeric(combined_point_est_df$p_val)

# subset out global method
combined_point_est_df <- combined_point_est_df

## ggplot
print_plot <- ggplot(combined_point_est_df, 
                     aes(x = pm_method, y = odds_ratio, colour = pm_method)) +
  # custom color 
  scale_color_manual(name = "Smoke-Estimation Method", 
                     values = c("red", "blue", "#32115C"),
                     guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
  geom_point() + #geom_text(vjust = 0, nudge_x = 0.3) +
  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
  facet_wrap(~outcome, nrow = 3) +
  geom_hline(yintercept = 1, linetype=2) +
  #    ggtitle('Association Between PM2.5 \n from Wildfire Smoke on Hospitalizations') +
  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
  #xlab('Smoke Estimation Method') +
  # plot theme
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        #strip.text = element_text(size = 8),
        # axis element
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(angle = 90),
        # legend elements
        legend.position = "bottom")
#legend.text = element_text(size = 8))

print(print_plot)
# save figure
ggsave("../plot_lag/fig_lag1.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

```

```{r lag6}
df_list <- list(resp_casecross, asthma_casecross, copd_casecross, 
                pneum_casecross, acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
                cereb_vas_casecross, hf_casecross, ihd_casecross, mi_casecross, 
                broken_arm_casecross)

outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 
                  'Pneumonia', 'Acute Bronchitis', 'Cardiovascular Disease',
                  'Arrhythmia', 'Cerebrovascular Disease', 'Heart Failure',
                  'Ischemic Heart Disease', 'Myocardial Infarction',
                  'Broken Arm')

method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')

# create an empty list to row bind dataframes together
datalist <- list()

# data wrangling ----
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){

  # dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- colnames(df_to_loop[76])
  # outcome name
  outcome_name <- outcome_list[i]

  # extract covariates from dataframe
  covariates_df <- df_to_loop[, c(1:26, 71, 74, 76:85)]

  # extract pm values and divide by 10 and ordered
  which(colnames(df_to_loop)=="wrf_smk_pm_lag1_zip") # 109
  which(colnames(df_to_loop)=="krig_pm_lag1_zip") # 123
  which(colnames(df_to_loop)=="geo_wt_pm_lag1_zip") # code to find column numbers, 116
  pm_estimates_df <- df_to_loop[, c(109,123,116, 94)]/10  # create 10 unit increases

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- cbind(covariates_df, pm_estimates_df) %>%
    # remove missing pm values
    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
    filter(!is.na(krig_pm_lag1_zip)) %>% 
    filter(!is.na(geo_wt_pm_lag1_zip)) %>%

    # join in cdc met data
    # full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
    # the following code makes sure that the counterfactual values retained are 
    # symetric in that number of obs before = number of obs after
    mutate(obs_diff_admission = (fromdate - date)/7) 
  # dataframe is already for the entire fire season, so I don't need to subset anymore

  # empty df for table
  table_df <- data.frame()

  # empty matrix
  point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)

  colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 'odds_ratio', 
                                 'lower95', 'upper95', 'se', 'p_val')

  # fill in the outcome name for the dataframe before the loop
  point_estimates[, 1] <- outcome_name

  # second loop to run a model for each pm estimation method
  for(j in 39:41){

    # variable to model 
    var_name <- colnames(df_analysis[j])

    # conditional logistic regression model
    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)

    # some model/DAG checks of data rish provided

    #glimpse(df_analysis)

    # adjusting for admit month is not possible as it's the same var for the subject.
    # adjusting for month smoke is a collider and should not be adjusted for.
    #summary(clogit(outcome ~ geo_smk_pm_zip + daily_meanRH + daily_meanT + daily_meanWS + daily_meanPrcp +
    #                as.factor(month_smk) + strata(PATIENTID), df_analysis))
    #outocme association: daily_meanRH,

    # check for confoudning following DAG assumptions
    #summary(clogit(outcome ~ wrf_temp_zip + strata(PATIENTID), df_analysis))
    #summary(clogit(outcome ~ wrf_temp_zip + strata(PATIENTID), df_analysis))
    # smoke predictor association: daily_meanRH, 
    #summary(lmer(geo_smk_pm_zip ~ daily_meanWS + (1 | PATIENTID), df_analysis ))
    #summary(lmer(wrf_temp_zip ~ daily_meanT + (1 | PATIENTID), df_analysis ))
    #cor(df_analysis$wrf_temp_county, df_analysis$daily_meanT)

    # populate matrix
    row_n <- j-38

    point_estimates[row_n, 2] <- method_list[row_n]
    point_estimates[row_n, 3] <- mod$n
    point_estimates[row_n, 4] <- mod$nevent
    # odds ratio
    point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)

    # 95% lower bound
    point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -
                                             1.96*(summary(mod)$coefficient[1,3])), 3)
    # 95% upper bound
    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
                                             1.96*(summary(mod)$coefficient[1,3])), 3)
    # standard error
    point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
    # p val
    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)

    # save point estimates as a dataframe
    point_est_df <- as_data_frame(point_estimates)

  }

  # combine previous values in dataframe that has all outcome/methods comparisons
  datalist[[i]] <- point_est_df

} # end of loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist)

# subset columns I want to put in to the table
table_df <- combined_point_est_df %>% select(2, 3:7)

tab <- htmlTable(txtRound(table_df, digits = 3, 1:3), 
                 caption = "Association between a 10 ug/m^3 in PM2.5 and Health Outcomes",
                 # row group by outcome
                 rgroup = outcome_list,
                 n.rgroup = c(rep(3, 12)), # 4 rows for each method for each outcome
                 # column headers
                 header = c("Method", "Obs.", "Events",
                            "OR&dagger;", "Lower", "Upper"),
                 # column spanner
                 cgroup = c("", "95% CI"), 
                 n.cgroup = c(4, 2),
                 padding.rgroup = "&nbsp;&nbsp;",
                 css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
                 align = "llccccc", # column alignment,
                 tfoot="&dagger; Adjusted for CDC temperature, relatively humidity, wind speed, and precipitation; accounting for subject. Time-stratified: referent periods matched to events on same day of week within July to October fire season."
) # end table

print(tab)

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
combined_point_est_df$outcome <- factor(combined_point_est_df$outcome, 
                                        levels = unique(combined_point_est_df$outcome))

combined_point_est_df$pm_method <- factor(combined_point_est_df$pm_method, 
                                          levels = unique(combined_point_est_df$pm_method))

combined_point_est_df$n <- as.numeric(combined_point_est_df$n)
combined_point_est_df$n_events <- as.numeric(combined_point_est_df$n_events)
combined_point_est_df$odds_ratio <- as.numeric(combined_point_est_df$odds_ratio)
combined_point_est_df$lower95 <- as.numeric(combined_point_est_df$lower95)
combined_point_est_df$upper95 <- as.numeric(combined_point_est_df$upper95)
combined_point_est_df$se <- as.numeric(combined_point_est_df$se)
combined_point_est_df$p_val <- as.numeric(combined_point_est_df$p_val)

# subset out global method
combined_point_est_df <- combined_point_est_df

## ggplot
print_plot <- ggplot(combined_point_est_df, 
                     aes(x = pm_method, y = odds_ratio, colour = pm_method)) +
  # custom color 
  scale_color_manual(name = "Smoke-Estimation Method", 
                     values = c("red", "blue", "#32115C"),
                     guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
  geom_point() + #geom_text(vjust = 0, nudge_x = 0.3) +
  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
  facet_wrap(~outcome, nrow = 3) +
  geom_hline(yintercept = 1, linetype=2) +
  #    ggtitle('Association Between PM2.5 \n from Wildfire Smoke on Hospitalizations') +
  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
  #xlab('Smoke Estimation Method') +
  # plot theme
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        #strip.text = element_text(size = 8),
        # axis element
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(angle = 90),
        # legend elements
        legend.position = "bottom")
#legend.text = element_text(size = 8))

print(print_plot)
# save figure
ggsave("../plot_lag/fig_lag1.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

```

```{r lag7}
df_list <- list(resp_casecross, asthma_casecross, copd_casecross, 
                pneum_casecross, acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
                cereb_vas_casecross, hf_casecross, ihd_casecross, mi_casecross, 
                broken_arm_casecross)

outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 
                  'Pneumonia', 'Acute Bronchitis', 'Cardiovascular Disease',
                  'Arrhythmia', 'Cerebrovascular Disease', 'Heart Failure',
                  'Ischemic Heart Disease', 'Myocardial Infarction',
                  'Broken Arm')

method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')

# create an empty list to row bind dataframes together
datalist <- list()

# data wrangling ----
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){

  # dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- colnames(df_to_loop[76])
  # outcome name
  outcome_name <- outcome_list[i]

  # extract covariates from dataframe
  covariates_df <- df_to_loop[, c(1:26, 71, 74, 76:85)]

  # extract pm values and divide by 10 and ordered
  which(colnames(df_to_loop)=="wrf_smk_pm_lag1_zip") # 109
  which(colnames(df_to_loop)=="krig_pm_lag1_zip") # 123
  which(colnames(df_to_loop)=="geo_wt_pm_lag1_zip") # code to find column numbers, 116
  pm_estimates_df <- df_to_loop[, c(109,123,116, 94)]/10  # create 10 unit increases

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- cbind(covariates_df, pm_estimates_df) %>%
    # remove missing pm values
    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
    filter(!is.na(krig_pm_lag1_zip)) %>% 
    filter(!is.na(geo_wt_pm_lag1_zip)) %>%

    # join in cdc met data
    # full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
    # the following code makes sure that the counterfactual values retained are 
    # symetric in that number of obs before = number of obs after
    mutate(obs_diff_admission = (fromdate - date)/7) 
  # dataframe is already for the entire fire season, so I don't need to subset anymore

  # empty df for table
  table_df <- data.frame()

  # empty matrix
  point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)

  colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 'odds_ratio', 
                                 'lower95', 'upper95', 'se', 'p_val')

  # fill in the outcome name for the dataframe before the loop
  point_estimates[, 1] <- outcome_name

  # second loop to run a model for each pm estimation method
  for(j in 39:41){

    # variable to model 
    var_name <- colnames(df_analysis[j])

    # conditional logistic regression model
    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)

    # some model/DAG checks of data rish provided

    #glimpse(df_analysis)

    # adjusting for admit month is not possible as it's the same var for the subject.
    # adjusting for month smoke is a collider and should not be adjusted for.
    #summary(clogit(outcome ~ geo_smk_pm_zip + daily_meanRH + daily_meanT + daily_meanWS + daily_meanPrcp +
    #                as.factor(month_smk) + strata(PATIENTID), df_analysis))
    #outocme association: daily_meanRH,

    # check for confoudning following DAG assumptions
    #summary(clogit(outcome ~ wrf_temp_zip + strata(PATIENTID), df_analysis))
    #summary(clogit(outcome ~ wrf_temp_zip + strata(PATIENTID), df_analysis))
    # smoke predictor association: daily_meanRH, 
    #summary(lmer(geo_smk_pm_zip ~ daily_meanWS + (1 | PATIENTID), df_analysis ))
    #summary(lmer(wrf_temp_zip ~ daily_meanT + (1 | PATIENTID), df_analysis ))
    #cor(df_analysis$wrf_temp_county, df_analysis$daily_meanT)

    # populate matrix
    row_n <- j-38

    point_estimates[row_n, 2] <- method_list[row_n]
    point_estimates[row_n, 3] <- mod$n
    point_estimates[row_n, 4] <- mod$nevent
    # odds ratio
    point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)

    # 95% lower bound
    point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -
                                             1.96*(summary(mod)$coefficient[1,3])), 3)
    # 95% upper bound
    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
                                             1.96*(summary(mod)$coefficient[1,3])), 3)
    # standard error
    point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
    # p val
    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)

    # save point estimates as a dataframe
    point_est_df <- as_data_frame(point_estimates)

  }

  # combine previous values in dataframe that has all outcome/methods comparisons
  datalist[[i]] <- point_est_df

} # end of loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist)

# subset columns I want to put in to the table
table_df <- combined_point_est_df %>% select(2, 3:7)

tab <- htmlTable(txtRound(table_df, digits = 3, 1:3), 
                 caption = "Association between a 10 ug/m^3 in PM2.5 and Health Outcomes",
                 # row group by outcome
                 rgroup = outcome_list,
                 n.rgroup = c(rep(3, 12)), # 4 rows for each method for each outcome
                 # column headers
                 header = c("Method", "Obs.", "Events",
                            "OR&dagger;", "Lower", "Upper"),
                 # column spanner
                 cgroup = c("", "95% CI"), 
                 n.cgroup = c(4, 2),
                 padding.rgroup = "&nbsp;&nbsp;",
                 css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
                 align = "llccccc", # column alignment,
                 tfoot="&dagger; Adjusted for CDC temperature, relatively humidity, wind speed, and precipitation; accounting for subject. Time-stratified: referent periods matched to events on same day of week within July to October fire season."
) # end table

print(tab)

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
combined_point_est_df$outcome <- factor(combined_point_est_df$outcome, 
                                        levels = unique(combined_point_est_df$outcome))

combined_point_est_df$pm_method <- factor(combined_point_est_df$pm_method, 
                                          levels = unique(combined_point_est_df$pm_method))

combined_point_est_df$n <- as.numeric(combined_point_est_df$n)
combined_point_est_df$n_events <- as.numeric(combined_point_est_df$n_events)
combined_point_est_df$odds_ratio <- as.numeric(combined_point_est_df$odds_ratio)
combined_point_est_df$lower95 <- as.numeric(combined_point_est_df$lower95)
combined_point_est_df$upper95 <- as.numeric(combined_point_est_df$upper95)
combined_point_est_df$se <- as.numeric(combined_point_est_df$se)
combined_point_est_df$p_val <- as.numeric(combined_point_est_df$p_val)

# subset out global method
combined_point_est_df <- combined_point_est_df

## ggplot
print_plot <- ggplot(combined_point_est_df, 
                     aes(x = pm_method, y = odds_ratio, colour = pm_method)) +
  # custom color 
  scale_color_manual(name = "Smoke-Estimation Method", 
                     values = c("red", "blue", "#32115C"),
                     guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
  geom_point() + #geom_text(vjust = 0, nudge_x = 0.3) +
  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
  facet_wrap(~outcome, nrow = 3) +
  geom_hline(yintercept = 1, linetype=2) +
  #    ggtitle('Association Between PM2.5 \n from Wildfire Smoke on Hospitalizations') +
  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
  #xlab('Smoke Estimation Method') +
  # plot theme
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        #strip.text = element_text(size = 8),
        # axis element
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(angle = 90),
        # legend elements
        legend.position = "bottom")
#legend.text = element_text(size = 8))

print(print_plot)
# save figure
ggsave("../plot_lag/fig_lag1.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

```

### Medication Distributed Lag

### 2017-06-14 r_scripts/analysis/oregon_medication.Rmd
DML-unfinished

@@
 library(knitr)
+library(splines)
 ```
@@
          wrf_f_pm_lag5 = lag(wrf_f_pm, 5, order_by = ZIPCODE),
+         wrf_f_pm_lag6 = lag(wrf_f_pm, 6, order_by = ZIPCODE),
+         wrf_f_pm_lag7 = lag(wrf_f_pm, 7, order_by = ZIPCODE),
          # wrf no fire lag
@@
          wrf_nf_pm_lag5 = lag(wrf_nf_pm, 5, order_by = ZIPCODE),
+         wrf_nf_pm_lag6 = lag(wrf_nf_pm, 6, order_by = ZIPCODE),
+         wrf_nf_pm_lag7 = lag(wrf_nf_pm, 7, order_by = ZIPCODE),
          # wrf_smk_pm
@@
          wrf_smk_pm_lag5 = lag(wrf_smk_pm, 5, order_by = ZIPCODE),
+         wrf_smk_pm_lag6 = lag(wrf_smk_pm, 6, order_by = ZIPCODE),
+         wrf_smk_pm_lag7 = lag(wrf_smk_pm, 7, order_by = ZIPCODE),
          # geo weighted pm
@@
          geo_wt_pm_lag5 = lag(geo_wt_pm, 5, order_by = ZIPCODE),
+         geo_wt_pm_lag6 = lag(geo_wt_pm, 6, order_by = ZIPCODE),
+         geo_wt_pm_lag7 = lag(geo_wt_pm, 7, order_by = ZIPCODE),
          # krig pm
@@
          krig_pm_lag5 = lag(krig_pm, 5, order_by = ZIPCODE), 
+         krig_pm_lag6 = lag(krig_pm, 6, order_by = ZIPCODE),
+         krig_pm_lag7 = lag(krig_pm, 7, order_by = ZIPCODE), 
          # background pm
@@
          background_pm_lag5 = lag(background_pm, 5, order_by = ZIPCODE), 
+         background_pm_lag6 = lag(background_pm, 6, order_by = ZIPCODE),
+         background_pm_lag7 = lag(background_pm, 7, order_by = ZIPCODE), 
          # geo_smk_pm 
@@
          geo_smk_pm_lag5 = lag(geo_smk_pm, 5, order_by = ZIPCODE),
+         geo_smk_pm_lag6 = lag(geo_smk_pm, 6, order_by = ZIPCODE),
+         geo_smk_pm_lag7 = lag(geo_smk_pm, 7, order_by = ZIPCODE),
          # krig smk pm
@@
          krig_smk_pm_lag5 = lag(krig_smk_pm, 5, order_by = ZIPCODE),
+         krig_smk_pm_lag6 = lag(krig_smk_pm, 6, order_by = ZIPCODE),
+         krig_smk_pm_lag7 = lag(krig_smk_pm, 7, order_by = ZIPCODE),
          # temp
@@
          wrf_temp_lag4 = lag(wrf_temp, 4, order_by = ZIPCODE),
-         wrf_temp_lag5 = lag(wrf_temp, 5, order_by = ZIPCODE)) %>% 
+         wrf_temp_lag5 = lag(wrf_temp, 5, order_by = ZIPCODE), 
+         wrf_temp_lag6 = lag(wrf_temp, 6, order_by = ZIPCODE),
+         wrf_temp_lag7 = lag(wrf_temp, 7, order_by = ZIPCODE)) %>% 
   # ungroup by zip
@@
 
-## NDC claims with index
-
-Check the claims in each category (Asthma, COPD, Both or Other).
-
-```{r echo = FALSE, message=FALSE, warning=FALSE}
-read_path7 <- paste0("../data_new/medication/oregon_beta_ndc_index.csv")
-or_ndc_index <- read_csv(read_path7)
-
-oregon_beta_df <- or_ndc_index %>% 
-  group_by(personkey) %>% 
-  arrange(personkey, fromdate, line) %>% 
-  mutate(num_visit = dense_rank(fromdate)) %>%
-  arrange(personkey, num_visit) %>% 
-  select(personkey, clmid, num_visit) %>% 
-  unique() %>% 
-  full_join(or_ndc_index, by = c("personkey", "clmid")) %>% 
-  arrange(personkey, clmid, line, fromdate) %>% 
-  filter(row_number() == 1) %>%
-  # add new transverted from date
-  mutate(dates = as.Date(fromdate, "%m/%d/%Y")) %>%
-  filter(dates >= '2013-05-01' & 
-         dates <= '2013-09-30') # 24538
-
-summary(as.factor(oregon_beta_df$index))
+## Distributed Lag Model
+
+```{r DLM, echo = FALSE, message=FALSE, warning=FALSE}
+
+which(colnames(asthma_ndc_df)=="geo_wt_pm_zip") # 88
+which(colnames(asthma_ndc_df)=="geo_wt_pm_lag1_zip") #115
+
+asthma_ndc_lag <- asthma_ndc_df %>%
+  filter(gender!="U") %>%
+  filter(outcome==1) %>%
+  select(88, 115:121) # GWR-method
+
+lag_max <- 7
+n <- length(asthma_ndc_lag[,1])
+
+asthma_ndc_lag <- as.matrix(asthma_ndc_lag)
+
+# true DL function
+DLfunc <- exp(-(0:lag_max)/2)
+
+# simulate outcomes
+y <- drop(asthma_ndc_lag%*%DLfunc) # + rnorm(n)
+
+B <- ns(0:lag_max, df =4, intercept=TRUE)
+
+# the nrow B should equal ncol(pm)
+dim(B)
+dim(asthma_ndc_lag)
+
+asthma_ndc_lag_B <- asthma_ndc_lag%*%B
+fit <- lm(y~asthma_ndc_lag_B)
+coef(fit)
+
+dlparms <- grep("asthma_ndc_lag_B",names(coef(fit)))
+DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms])
+DLvar <- B%*%vcov(fit)[dlparms,dlparms]%*%t(B)
+DLestimate$SE <- sqrt(diag(DLvar))
+DLestimate$lower <- DLestimate$estimate + DLestimate$SE * qt(1-0.975, df=fit$df.residual)
+DLestimate$upper <- DLestimate$estimate + DLestimate$SE * qt(0.975, df=fit$df.residual)
+DLestimate
+
+DLestimate$lag <- 0:lag_max
+DLestimate$truth <- DLfunc
+p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
+p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
+p <- p + geom_line(aes(x=lag, y=truth), linetype=2, size=2)
+p <- p + theme_bw()
+p
+
+cumulative <- sum(DLestimate$estimate)
+cumulative
+## [1] 2.485711
+cumulative_se <- sqrt(sum(DLvar))
+cumulative_se
+## [1] 0.0001887442
+cumulative_CI <- cumulative + cumulative_se * qt(c(1-0.975,0.975), df=fit$df.residual)
+cumulative_CI
+## [1] 2.485341 2.486080
 
@@
 
-### Descriptives
-
-For all of 2013, there were a total of millons of all claims records in the APAC dataset.
-During the May 1st to September 30th wildfire season in Oregon state, there were 24217 pharmacy visits for the ndc of beta 2 agonists.
-The number of cases, as well as age-specific and sex-specific strate are shown below.
-
-```{r ndc descriptive table, echo = F, message = F, warning = F, results='asis'}
-read_path6 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover.csv")
-beta_ndc <- read_csv(read_path6)
-
-# join personkey with index
-oregon_beta_index_df <- oregon_beta_df %>%
-  select(personkey, index)
-
-beta_ndc_with_index <- beta_ndc %>%
-  left_join(oregon_beta_index_df, by="personkey")
+```{r overall lag1, warning =F, echo = F, results='asis'} 
+# dataframe list
+
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
 
 # dataframe to loop through
-beta_ndc_with_index_df <- data.frame(beta_ndc_with_index)
-
-beta_asthma <- beta_ndc_with_index_df %>%
-  filter(index=="asthma")
-
-beta_copd <- beta_ndc_with_index_df %>%
-  filter(index=="copd")
-
-beta_both <- beta_ndc_with_index_df %>%
-  filter(index=="both")
-
-beta_other <- beta_ndc_with_index_df %>%
-  filter(index=="other")
-
-df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)
-
-outcome_list <- c("Asthma", "COPD", "Both", "Other")
-
-# create an empty list to row bind dataframes together
-datalist <- list()
-
-# dataframe list
-method_list <- c('WRF-Chem Smoke', 'Geo-Weighted Smoke', 'Kriging Smoke')
+# asthma_ndc_df <- data.frame(asthma_ndc)
 
 # outcome name
-outcome_name <- c("Asthma", "COPD", "Both", "Other")
-
-for(i in 1:length(df_list)){
-  # dataframe to loop through
-  df_to_loop <- data.frame(df_list[i])
-  # indication of column
-  outcome <- colnames(df_to_loop[75])
-  # outcome name
-  outcome_name <- outcome_list[i]
-
-  # dataframe for analysis creation
-  # bind columns back together 
-  df_analysis <- df_to_loop %>%
-    filter(!is.na(wrf_smk_pm_zip)) %>% 
-    # only look at outcomes
-    filter(outcome == 1) %>%
-    # add another row that makes sure there is a person <15 in the dataframe
-    # tricking xtabs to produce a 0 cell for the outcome for age <15
-    add_row(outcome = 0, age_ind = 0)
-
-  # cross tabs
-  outcome_n <- xtabs(~ outcome, df_analysis)
-  cross_tab_age <- xtabs(~ outcome + age_ind, df_analysis)
-  cross_tab_sex <- xtabs(~ outcome + sex_ind, df_analysis)
-  # empty matrix
-  point_estimates <- matrix(nrow = 1, ncol = 7, byrow = T)
-
-  colnames(point_estimates) <- c("outcome", "n", "age_15", "age_15_65", 
-                                 "age_65", "female", "male")
-
-  # fill in the outcome name for the dataframe before the loop
-  point_estimates[, 1] <- outcome_name
-  # fill n
-  point_estimates[, 2] <- outcome_n[2] # second element of the 1 dimension vector
-  # age <15
-  point_estimates[, 3] <- cross_tab_age[2, 1]
-  # age 15 to 65
-  point_estimates[, 4] <- cross_tab_age[2, 2]
-  # age >65
-  point_estimates[, 5] <- cross_tab_age[2, 3]
-  # male == 0
-  point_estimates[, 7] <- cross_tab_sex[1, 1]
-  # female == 1
-  point_estimates[, 6] <- cross_tab_sex[1, 2]
-
-  # save point estimates as a dataframe
-  point_est_df <- as_data_frame(point_estimates)
-
-  # combine previous values in dataframe that has all outcome/methods comparisons
-  datalist[[i]] <- point_est_df
-
-} # end of loop
-
-# combine each outcome dataframe itteration in to a big dataset
-combined_point_est_df <- bind_rows(datalist) %>% 
-  # find proportions/percents for each strata in a row
-  mutate(age_15_pr = as.character(round((as.numeric(age_15)/as.numeric(n))*100,1)),
-         age_15_65_pr = as.character(round((as.numeric(age_15_65)/as.numeric(n))*100,1)),
-         age_65_pr = as.character(round((as.numeric(age_65)/as.numeric(n))*100,1)),
-         female_pr = as.character(round((as.numeric(female)/as.numeric(n))*100,1)),
-         male_pr = as.character(round((as.numeric(male)/as.numeric(n))*100,1))) %>% 
-  select(outcome, n, age_15, age_15_pr, age_15_65, age_15_65_pr, age_65,
-         age_65_pr, female, female_pr, male, male_pr)
-
-#str(combined_point_est_df)
-
-tab <- htmlTable(txtRound(combined_point_est_df, digits = 1), 
-           caption = "Number of cases for each outcome observed from May 1st to September 30st, 2013",
-           # column headers
-           header = c("Outcome", "Cases n", "Less than 15", "(%)", "15 to 65", "(%)", "Greater than 65",
-                      "(%)", "Female", "(%)", "Male", "(%)"),
-           # column spanner
-           cgroup = c("","Age Category", "Sex"), 
-           n.cgroup = c(2, 6, 4),
-           padding.rgroup = "&nbsp;&nbsp;",
-           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
-           align = "llccccc" # column alignment,
-            ) # end table
-
-print(tab)
-
-```
-
-### Overall plot
-
-```{r ndc overall try, warning =F, echo = F, results='asis'}
-df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)
-
-outcome_list <- c("Asthma", "COPD", "Both", "Other")
-
-# create an empty list to row bind dataframes together
-datalist <- list()
-
-# dataframe list
-method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
-
-# outcome name
-outcome_name <- c("Asthma", "COPD", "Both", "Other")
+outcome <- "Asthma Inhalers"
+outcome_name <- "Asthma Inhalers"
 
@@
 # Producing conditional logit model estimates loop
-for(i in 1:length(df_list)){
-  # dataframe to loop through
-  df_to_loop <- data.frame(df_list[i])
-  # indication of column
-  outcome <- colnames(df_to_loop[75])
-  # outcome name
-  outcome_name <- outcome_list[i]
 
 # extract covariates from dataframe
-  covariates_df <- df_to_loop[, c(1:26, 74:84)]
+covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]
 
 # extract pm values and divide by 10 and ordered
-  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
-  pm_estimates_df <- df_to_loop[, c(87, 92, 91, 93)]/10  # create 10 unit increases
+which(colnames(asthma_ndc_df)=="wrf_smk_pm_lag1_zip") # 108
+which(colnames(asthma_ndc_df)=="krig_pm_lag1_zip") # 122
+which(colnames(asthma_ndc_df)=="geo_wt_pm_lag1_zip") # code to find column numbers, 115
+pm_estimates_df <- asthma_ndc_df[, c(108, 122, 115, 93)]/10  # create 10 unit increases
+
+# new loop for age categories
+
+# empty matrix (12 x 10 matrix)
+point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)
+
+colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+# fill in the outcome namedataframe before method loop
+point_estimates[, 1] <- outcome_name
 
@@
   # remove missing pm values
-    filter(!is.na(wrf_smk_pm_zip)) %>%
-
-    # join in cdc met data
-    # full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
+  filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
+  filter(!is.na(krig_pm_lag1_zip)) %>% 
+  filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
   # the following code makes sure that the counterfactual values retained are 
@@
 
-  # empty df for table
-  table_df <- data.frame()
-
-  # empty matrix
-  point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)
-
-  colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 'odds_ratio', 
-                                 'lower95', 'upper95', 'se', 'p_val')
-
-  # fill in the outcome name for the dataframe before the loop
-  point_estimates[, 1] <- outcome_name
-
 # second loop to run a model for each pm estimation method
-
 for(j in 38:40){
@@
 
-      # save point estimates as a dataframe
-      point_est_df <- as_data_frame(point_estimates)
-
   # create else statement that fills matrix with missing so I still have the row
@@
 
-  # combine previous values in dataframe that has all outcome/methods comparisons
-  datalist[[i]] <- point_est_df
-
-} # end of loop
-
-# combine each outcome dataframe itteration in to a big dataset
-combined_point_est_df <- bind_rows(datalist)
-
+# save point estimates as a dataframe
+combined_point_est_df <- as_data_frame(point_estimates)
+
+wrf_geo_age <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" |
+    #      pm_method == "WRF-Chem Smoke") %>% 
   # subset columns I want to put in to the table
-table_df <- combined_point_est_df %>% select(2, 3:7)
-
-tab <- htmlTable(txtRound(table_df, digits = 3, 1:3), 
-                 caption = "Association between a 10 ug/m^3 in PM2.5 and Health Outcomes",
+  select(2, 4:7) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
+         caption = "Association between a 10 ug/m^3 in PM2.5 three smoke method and pharmacy outcomes",
          # row group by outcome
-                 rgroup = outcome_list,
-                 n.rgroup = c(rep(3, 4)), # 4 rows for each method for each outcome
+         rgroup = "Asthma Inhalers",
+         n.rgroup = c(rep(3, 1)), # 6 rows for each age cat for each outcome
          # column headers
-                 header = c("Method", "Obs.", "Events",
+         header = c("Est Method", "Events",
                     "OR&dagger;", "Lower", "Upper"),
@@
          cgroup = c("", "95% CI"), 
-                 n.cgroup = c(4, 2),
+         n.cgroup = c(3, 2),
          padding.rgroup = "&nbsp;&nbsp;",
          css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
-                 align = "llccccc", # column alignment,
-                 tfoot="&dagger; Adjusted for CDC temperature, relatively humidity, wind speed, and precipitation; accounting for subject.
-Time-stratified: referent periods matched to events on same day of week within July to October fire season."
+         align = "llcccc", # column alignment,
+         tfoot="&dagger; Referent periods matched to events on same day of week within May to September fire season."
          ) # end table
@@
 print(tab)
+
+combined_point_est_df_new <- combined_point_est_df %>%
+  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
+                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
+  select(-n_events) %>%
+  rename(n_events = n_events_new)
 
@@
 # factor preserves the order of the variable
-combined_point_est_df$outcome <- factor(combined_point_est_df$outcome, 
-                                        levels = unique(combined_point_est_df$outcome))
-
-combined_point_est_df$pm_method <- factor(combined_point_est_df$pm_method, 
-                                          levels = unique(combined_point_est_df$pm_method))
-
-combined_point_est_df$n <- as.numeric(combined_point_est_df$n)
-combined_point_est_df$n_events <- as.numeric(combined_point_est_df$n_events)
-combined_point_est_df$odds_ratio <- as.numeric(combined_point_est_df$odds_ratio)
-combined_point_est_df$lower95 <- as.numeric(combined_point_est_df$lower95)
-combined_point_est_df$upper95 <- as.numeric(combined_point_est_df$upper95)
-combined_point_est_df$se <- as.numeric(combined_point_est_df$se)
-combined_point_est_df$p_val <- as.numeric(combined_point_est_df$p_val)
-
-# subset out global method
-combined_point_est_df <- combined_point_est_df
+wrf_geo_age_plot <- combined_point_est_df_new %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  # do not plot results with less than 15 events, create missing vals
+  # in mutate
+  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
+                             99, odds_ratio),
+         lower95 = ifelse(as.numeric(n_events) < 15, 
+                             99, lower95),
+         upper95 = ifelse(as.numeric(n_events) < 15, 
+                             99, upper95)) %>% 
+  # subset columns I want to put in to the table
+  select(1, 2, 9, 4:6) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
+                              levels = unique(wrf_geo_age_plot$outcome))
+
+wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
+                                levels = unique(wrf_geo_age_plot$pm_method))
+
+wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
+wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
+wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
+wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)
 
 ## ggplot 
-print_plot <- ggplot(combined_point_est_df, 
-                     aes(x = pm_method, y = odds_ratio, colour = pm_method)) +
+  print_plot <- ggplot(wrf_geo_age_plot,
+    aes(x = pm_method, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
     # custom color 
@@
                        guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
-  geom_point() + #geom_text(vjust = 0, nudge_x = 0.3) +
-  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
-  facet_wrap(~outcome, nrow = 3) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
     geom_hline(yintercept = 1, linetype=2) +
-  #    ggtitle('Association Between PM2.5 \n from Wildfire Smoke on Hospitalizations') +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
     ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
-  #xlab('Smoke Estimation Method') +
+    #ylim(0, 2) +
+    xlab('Smoke Estimation Method') +
     # plot theme
@@
     # facet text size
-        #strip.text = element_text(size = 8),
+    strip.text = element_text(size = 10),
     # axis element
-        axis.text.x = element_blank(),
-        axis.title.x = element_blank(),
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
     axis.ticks.x = element_blank(),
@@
   # save figure
-ggsave("../plot_new/fig_ndc_overall_four.pdf", plot = print_plot, 
+  ggsave("../plot_lag/ndc_overall_lag1_three.pdf", plot = print_plot, 
        width = 12, height = 8, units = "in")
@@
 
-### Time-Stratified by Age Categories
-
-Following table and figure look at some outcomes stratified by age category.
-
-For some outcomes, such like Age > 65 in WRF-Chem method, the CI is the only insignificant one in the plot.
-
-```{r ndc time stratified age strata, warning =F, echo = F, results='asis'} 
-df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)
-
-outcome_list <- c("Asthma", "COPD", "Both", "Other")
-
-# create an empty list to row bind dataframes together
-datalist <- list()
+```{r time stratified age strata lag 1, warning =F, echo = F, results='asis'} 
+# dataframe list
+
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# dataframe to loop through
+asthma_ndc_df <- data.frame(asthma_ndc)
 
 # outcome name
-outcome_name <- c("Asthma", "COPD", "Both", "Other")
+outcome <- "asthma inhalers"
+outcome_name <- "Asthma inhalers"
 
@@
 
-# looking only at wrf chem, kriging, and geo-weighted; taking out 'Global Smoke' method
-method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
-
 # create an empty list to row bind dataframes together
 datalist1 <- list()
-datalist2 <- list()
-
-for(i in 1:length(df_list)){
-
-  # dataframe to loop through
-  df_to_loop <- data.frame(df_list[i])
-  # indication of column
-  outcome <- df_list[i]
-  # outcome name
-  outcome_name <- outcome_list[i]
-
-  covariates_df <- df_to_loop[, c(1:26, 74:84)]
+
+# data wrangling ----
+# Producing conditional logit model estimates loop
+
+# extract covariates from dataframe
+covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]
 
@@
 #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
-  pm_estimates_df <- df_to_loop[, c(87, 92, 91, 93)]/10  # create 10 unit increases
+pm_estimates_df <- asthma_ndc_df[, c(108, 122, 115, 93)]/10  # create 10 unit increases
 
@@
 for(k in 0:2){
-
+  # empty matrix (12 x 10 matrix)
   point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)
@@
 
-    # dataframe for analysis creation
-    # bind columns back together 
   df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
     # remove missing pm values
-      filter(!is.na(wrf_smk_pm_zip)) %>% 
+    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
+    filter(!is.na(krig_pm_lag1_zip)) %>% 
+    filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
     # limit to specific age category
@@
     mutate(obs_diff_admission = (fromdate - date)/7) 
-
+    # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+  # second loop to run a model for each pm estimation method
   for(j in 38:40){
@@
   point_est_df <- as_data_frame(point_estimates)
-
-    # combine previous values in dataframe that has all outcome/methods comparisons
   datalist1[[k+1]] <- point_est_df
@@
 # bind rows of age category estimates together
-  age_est_df <- bind_rows(datalist1)
-
-  # populate second dataframe list
-  datalist2[[i]] <- age_est_df
-
-} # end of outcome loop
-
-# combine each outcome dataframe itteration in to a big dataset
-combined_point_est_df <- bind_rows(datalist2)
+combined_point_est_df <- bind_rows(datalist1)
 
@@
   # save figure
-  ggsave("../plot_new/ndc_age_fig_four.pdf", plot = print_plot, 
+  ggsave("../plot_lag/ndc_age_fig_three_lag1.pdf", plot = print_plot, 
        width = 12, height = 8, units = "in")
@@
 
-### Seasonal Time-Stratified by Sex
-
-Association with smoke PM~2.5~ and health outcomes stratified by sex.
-
-All the CI do not include 1, which means the use of asthma medication (beta 2 agonists) increases about 5% when smoke increase 10 units.
-So the risk for getting asthma will increase about 5% ~ 8% when the smoke PM2.5 increase 10 units.
-
-```{r ndc sex strata, warning = F, echo = F, results='asis'} 
-df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)
-
-outcome_list <- c("Asthma", "COPD", "Both", "Other")
+```{r sex strata cdc met  lag 1, warning = F, echo = F, results='asis'} 
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# dataframe to loop through
+asthma_ndc_df <- data.frame(asthma_ndc)
 
 # outcome name
-outcome_name <- c("Asthma", "COPD", "Both", "Other")
+outcome <- "asthma inhalers"
+outcome_name <- "Asthma inhalers"
+
+# create an empty list to row bind dataframes together
+datalist1 <- list()
 
@@
 
-datalist1 <- list()
-datalist2 <- list()
-
 # data wrangling ----
 # Producing conditional logit model estimates loop
-for(i in 1:length(df_list)){
-
-  # dataframe to loop through
-  df_to_loop <- data.frame(df_list[i])
-  # indication of column
-  outcome <- df_list[i]
-  # outcome name
-  outcome_name <- outcome_list[i]
-
-  covariates_df <- df_to_loop[, c(1:26, 74:84)]
+
+# extract covariates from dataframe
+covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]
 
@@
     # remove missing pm values
-      filter(!is.na(wrf_smk_pm_zip)) %>% 
-      # limit to specific sex category
+    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
+    filter(!is.na(krig_pm_lag1_zip)) %>% 
+    filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
+    # limit to specific age category
     filter(sex_ind == k) %>% 
@@
 
+  # second loop to run a model for each pm estimation method
   for(j in 38:40){
@@
 # bind rows of sex category estimates together
-  sex_est_df <- bind_rows(datalist1)
-
-  # populate second dataframe list
-  datalist2[[i]] <- sex_est_df
-
-} # end of outcome loop
-
-# combine each outcome dataframe itteration in to a big dataset
-combined_point_est_df <- bind_rows(datalist2)
+combined_point_est_df <- bind_rows(datalist1)
 
@@
   # save figure
+  ggsave("../plot_lag/ndc_sex_fig_three_lag1.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+## NDC claims with index
+
+Check the claims in each category (Asthma, COPD, Both or Other).
+
+```{r echo = FALSE, message=FALSE, warning=FALSE}
+read_path7 <- paste0("../data_new/medication/oregon_beta_ndc_index.csv")
+or_ndc_index <- read_csv(read_path7)
+
+oregon_beta_df <- or_ndc_index %>% 
+  group_by(personkey) %>% 
+  arrange(personkey, fromdate, line) %>% 
+  mutate(num_visit = dense_rank(fromdate)) %>%
+  arrange(personkey, num_visit) %>% 
+  select(personkey, clmid, num_visit) %>% 
+  unique() %>% 
+  full_join(or_ndc_index, by = c("personkey", "clmid")) %>% 
+  arrange(personkey, clmid, line, fromdate) %>% 
+  filter(row_number() == 1) %>%
+  # add new transverted from date
+  mutate(dates = as.Date(fromdate, "%m/%d/%Y")) %>%
+  filter(dates >= '2013-05-01' & 
+         dates <= '2013-09-30') # 24538
+
+summary(as.factor(oregon_beta_df$index))
+
+```
+
+### Descriptives
+
+For all of 2013, there were a total of millons of all claims records in the APAC dataset.
+During the May 1st to September 30th wildfire season in Oregon state, there were 24217 pharmacy visits for the ndc of beta 2 agonists.
+The number of cases, as well as age-specific and sex-specific strate are shown below.
+
+```{r ndc descriptive table, echo = F, message = F, warning = F, results='asis'}
+read_path6 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover.csv")
+beta_ndc <- read_csv(read_path6)
+
+# join personkey with index
+oregon_beta_index_df <- oregon_beta_df %>%
+  select(personkey, index)
+
+beta_ndc_with_index <- beta_ndc %>%
+  left_join(oregon_beta_index_df, by="personkey")
+
+# dataframe to loop through
+beta_ndc_with_index_df <- data.frame(beta_ndc_with_index)
+
+beta_asthma <- beta_ndc_with_index_df %>%
+  filter(index=="asthma")
+
+beta_copd <- beta_ndc_with_index_df %>%
+  filter(index=="copd")
+
+beta_both <- beta_ndc_with_index_df %>%
+  filter(index=="both")
+
+beta_other <- beta_ndc_with_index_df %>%
+  filter(index=="other")
+
+df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)
+
+outcome_list <- c("Asthma", "COPD", "Both", "Other")
+
+# create an empty list to row bind dataframes together
+datalist <- list()
+
+# dataframe list
+method_list <- c('WRF-Chem Smoke', 'Geo-Weighted Smoke', 'Kriging Smoke')
+
+# outcome name
+outcome_name <- c("Asthma", "COPD", "Both", "Other")
+
+for(i in 1:length(df_list)){
+  # dataframe to loop through
+  df_to_loop <- data.frame(df_list[i])
+  # indication of column
+  outcome <- colnames(df_to_loop[75])
+  # outcome name
+  outcome_name <- outcome_list[i]
+
+  # dataframe for analysis creation
+  # bind columns back together 
+  df_analysis <- df_to_loop %>%
+    filter(!is.na(wrf_smk_pm_zip)) %>% 
+    # only look at outcomes
+    filter(outcome == 1) %>%
+    # add another row that makes sure there is a person <15 in the dataframe
+    # tricking xtabs to produce a 0 cell for the outcome for age <15
+    add_row(outcome = 0, age_ind = 0)
+
+  # cross tabs
+  outcome_n <- xtabs(~ outcome, df_analysis)
+  cross_tab_age <- xtabs(~ outcome + age_ind, df_analysis)
+  cross_tab_sex <- xtabs(~ outcome + sex_ind, df_analysis)
+  # empty matrix
+  point_estimates <- matrix(nrow = 1, ncol = 7, byrow = T)
+
+  colnames(point_estimates) <- c("outcome", "n", "age_15", "age_15_65", 
+                                 "age_65", "female", "male")
+
+  # fill in the outcome name for the dataframe before the loop
+  point_estimates[, 1] <- outcome_name
+  # fill n
+  point_estimates[, 2] <- outcome_n[2] # second element of the 1 dimension vector
+  # age <15
+  point_estimates[, 3] <- cross_tab_age[2, 1]
+  # age 15 to 65
+  point_estimates[, 4] <- cross_tab_age[2, 2]
+  # age >65
+  point_estimates[, 5] <- cross_tab_age[2, 3]
+  # male == 0
+  point_estimates[, 7] <- cross_tab_sex[1, 1]
+  # female == 1
+  point_estimates[, 6] <- cross_tab_sex[1, 2]
+
+  # save point estimates as a dataframe
+  point_est_df <- as_data_frame(point_estimates)
+
+  # combine previous values in dataframe that has all outcome/methods comparisons
+  datalist[[i]] <- point_est_df
+
+} # end of loop
+
+# combine each outcome dataframe itteration in to a big dataset
+combined_point_est_df <- bind_rows(datalist) %>% 
+  # find proportions/percents for each strata in a row
+  mutate(age_15_pr = as.character(round((as.numeric(age_15)/as.numeric(n))*100,1)),
+         age_15_65_pr = as.character(round((as.numeric(age_15_65)/as.numeric(n))*100,1)),
+         age_65_pr = as.character(round((as.numeric(age_65)/as.numeric(n))*100,1)),
+         female_pr = as.character(round((as.numeric(female)/as.numeric(n))*100,1)),
+         male_pr = as.character(round((as.numeric(male)/as.numeric(n))*100,1))) %>% 
+  select(outcome, n, age_15, age_15_pr, age_15_65, age_15_65_pr, age_65,
+         age_65_pr, female, female_pr, male, male_pr)
+
+#str(combined_point_est_df)
+
+tab <- htmlTable(txtRound(combined_point_est_df, digits = 1), 
+           caption = "Number of cases for each outcome observed from May 1st to September 30st, 2013",
+           # column headers
+           header = c("Outcome", "Cases n", "Less than 15", "(%)", "15 to 65", "(%)", "Greater than 65",
+                      "(%)", "Female", "(%)", "Male", "(%)"),
+           # column spanner
+           cgroup = c("","Age Category", "Sex"), 
+           n.cgroup = c(2, 6, 4),
+           padding.rgroup = "&nbsp;&nbsp;",
+           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+           align = "llccccc" # column alignment,
+            ) # end table
+
+print(tab)
+
+```
+
+### Overall plot
+
+```{r ndc overall try, warning =F, echo = F, results='asis'}
+df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)
+
+outcome_list <- c("Asthma", "COPD", "Both", "Other")
+
+# create an empty list to row bind dataframes together
+datalist <- list()
+
+# dataframe list
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# outcome name
+outcome_name <- c("Asthma", "COPD", "Both", "Other")
+
+# data wrangling ----
+# Producing conditional logit model estimates loop 
+for(i in 1:length(df_list)){
+  # dataframe to loop through
+  df_to_loop <- data.frame(df_list[i])
+  # indication of column
+  outcome <- colnames(df_to_loop[75])
+  # outcome name
+  outcome_name <- outcome_list[i]
+
+  # extract covariates from dataframe
+  covariates_df <- df_to_loop[, c(1:26, 74:84)]
+
+  # extract pm values and divide by 10 and ordered
+  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+  pm_estimates_df <- df_to_loop[, c(87, 92, 91, 93)]/10  # create 10 unit increases
+
+  # dataframe for analysis creation
+  # bind columns back together 
+  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+    # remove missing pm values
+    filter(!is.na(wrf_smk_pm_zip)) %>%
+
+    # join in cdc met data
+    # full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
+    # the following code makes sure that the counterfactual values retained are 
+    # symetric in that number of obs before = number of obs after
+    mutate(obs_diff_admission = (fromdate - date)/7) 
+  # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+  # empty df for table
+  table_df <- data.frame()
+
+  # empty matrix
+  point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)
+
+  colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 'odds_ratio', 
+                                 'lower95', 'upper95', 'se', 'p_val')
+
+  # fill in the outcome name for the dataframe before the loop
+  point_estimates[, 1] <- outcome_name
+
+  # second loop to run a model for each pm estimation method
+
+  for(j in 38:40){
+
+    # variable to model 
+    var_name <- colnames(df_analysis[j])
+
+    # set row number to fill
+    row_n <- j-37
+
+    # only run the model if the dataframe has observations
+    if(nrow(df_analysis) != 0){
+      # conditional logistic regression model
+      mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+      # populate matrix
+      point_estimates[row_n, 2] <- method_list[row_n]
+      point_estimates[row_n, 3] <- mod$n
+      point_estimates[row_n, 4] <- mod$nevent
+      # odds ratio
+      point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+      # 95% lower bound
+      point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
+      # 95% upper bound
+      point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                               1.96*(summary(mod)$coefficient[1,3])), 3)
+      # standard error
+      point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
+      # p val
+      point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)
+
+      # save point estimates as a dataframe
+      point_est_df <- as_data_frame(point_estimates)
+
+      # create else statement that fills matrix with missing so I still have the row
+      # in the final dataframe
+    } else {point_estimates[row_n, 2] <- method_list[row_n]
+    point_estimates[row_n, 3] <- 0
+    point_estimates[row_n, 4] <- 0
+    point_estimates[row_n, c(5:9)] <- 99 } # end 'if else' statement
+  } # end methods loop
+
+  # combine previous values in dataframe that has all outcome/methods comparisons
+  datalist[[i]] <- point_est_df
+
+} # end of loop
+
+# combine each outcome dataframe itteration in to a big dataset
+combined_point_est_df <- bind_rows(datalist)
+
+# subset columns I want to put in to the table
+table_df <- combined_point_est_df %>% select(2, 3:7)
+
+tab <- htmlTable(txtRound(table_df, digits = 3, 1:3), 
+                 caption = "Association between a 10 ug/m^3 in PM2.5 and Health Outcomes",
+                 # row group by outcome
+                 rgroup = outcome_list,
+                 n.rgroup = c(rep(3, 4)), # 4 rows for each method for each outcome
+                 # column headers
+                 header = c("Method", "Obs.", "Events",
+                            "OR&dagger;", "Lower", "Upper"),
+                 # column spanner
+                 cgroup = c("", "95% CI"), 
+                 n.cgroup = c(4, 2),
+                 padding.rgroup = "&nbsp;&nbsp;",
+                 css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+                 align = "llccccc", # column alignment,
+                 tfoot="&dagger; Adjusted for CDC temperature, relatively humidity, wind speed, and precipitation; accounting for subject.
+Time-stratified: referent periods matched to events on same day of week within July to October fire season."
+) # end table
+
+print(tab)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+combined_point_est_df$outcome <- factor(combined_point_est_df$outcome, 
+                                        levels = unique(combined_point_est_df$outcome))
+
+combined_point_est_df$pm_method <- factor(combined_point_est_df$pm_method, 
+                                          levels = unique(combined_point_est_df$pm_method))
+
+combined_point_est_df$n <- as.numeric(combined_point_est_df$n)
+combined_point_est_df$n_events <- as.numeric(combined_point_est_df$n_events)
+combined_point_est_df$odds_ratio <- as.numeric(combined_point_est_df$odds_ratio)
+combined_point_est_df$lower95 <- as.numeric(combined_point_est_df$lower95)
+combined_point_est_df$upper95 <- as.numeric(combined_point_est_df$upper95)
+combined_point_est_df$se <- as.numeric(combined_point_est_df$se)
+combined_point_est_df$p_val <- as.numeric(combined_point_est_df$p_val)
+
+# subset out global method
+combined_point_est_df <- combined_point_est_df
+
+## ggplot
+print_plot <- ggplot(combined_point_est_df, 
+                     aes(x = pm_method, y = odds_ratio, colour = pm_method)) +
+  # custom color 
+  scale_color_manual(name = "Smoke-Estimation Method", 
+                     values = c("red", "blue", "#32115C"),
+                     guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+  geom_point() + #geom_text(vjust = 0, nudge_x = 0.3) +
+  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
+  facet_wrap(~outcome, nrow = 3) +
+  geom_hline(yintercept = 1, linetype=2) +
+  #    ggtitle('Association Between PM2.5 \n from Wildfire Smoke on Hospitalizations') +
+  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+  #xlab('Smoke Estimation Method') +
+  # plot theme
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        # strip element
+        strip.background = element_rect(colour=NA, fill=NA),
+        panel.border = element_rect(fill = NA, color = "black"),
+        # facet text size
+        #strip.text = element_text(size = 8),
+        # axis element
+        axis.text.x = element_blank(),
+        axis.title.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        axis.title.y = element_text(angle = 90),
+        # legend elements
+        legend.position = "bottom")
+#legend.text = element_text(size = 8))
+
+print(print_plot)
+# save figure
+ggsave("../plot_new/fig_ndc_overall_four.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+### Time-Stratified by Age Categories
+
+Following table and figure look at some outcomes stratified by age category.
+
+For some outcomes, such like Age > 65 in WRF-Chem method, the CI is the only insignificant one in the plot.
+
+```{r ndc time stratified age strata, warning =F, echo = F, results='asis'} 
+df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)
+
+outcome_list <- c("Asthma", "COPD", "Both", "Other")
+
+# create an empty list to row bind dataframes together
+datalist <- list()
+
+# outcome name
+outcome_name <- c("Asthma", "COPD", "Both", "Other")
+
+# age category list
+age_cat_list <- c(0,1,2)
+
+# looking only at wrf chem, kriging, and geo-weighted; taking out 'Global Smoke' method
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# create an empty list to row bind dataframes together
+datalist1 <- list()
+datalist2 <- list()
+
+for(i in 1:length(df_list)){
+
+  # dataframe to loop through
+  df_to_loop <- data.frame(df_list[i])
+  # indication of column
+  outcome <- df_list[i]
+  # outcome name
+  outcome_name <- outcome_list[i]
+
+  covariates_df <- df_to_loop[, c(1:26, 74:84)]
+
+  # extract pm values and divide by 10 and ordered
+  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+  pm_estimates_df <- df_to_loop[, c(87, 92, 91, 93)]/10  # create 10 unit increases
+
+  # new loop for age categories
+  for(k in 0:2){
+
+    point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)
+
+    colnames(point_estimates) <- c('outcome', 'age_cat', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+    # fill in the outcome namedataframe before method loop
+    point_estimates[, 1] <- outcome_name
+    # repeat the age category 4 times for each pm method
+    point_estimates[, 2] <- k
+
+    # dataframe for analysis creation
+    # bind columns back together 
+    df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+      # remove missing pm values
+      filter(!is.na(wrf_smk_pm_zip)) %>% 
+      # limit to specific age category
+      filter(age_ind == k) %>% 
+      # the following code makes sure that the counterfactual values retained are 
+      # symetric in that number of obs before = number of obs after
+      mutate(obs_diff_admission = (fromdate - date)/7)
+
+    for(j in 38:40){
+
+      # variable to model 
+      var_name <- colnames(df_analysis[j])
+
+      # set row number to fill
+      row_n <- j-37
+
+      # only run the model if the dataframe has observations
+      if(nrow(df_analysis) != 0){
+        # conditional logistic regression model
+        mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+        # populate matrix
+        point_estimates[row_n, 3] <- method_list[row_n]
+        point_estimates[row_n, 4] <- mod$n
+        point_estimates[row_n, 5] <- mod$nevent
+        # odds ratio
+        point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+        # 95% lower bound
+        point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
+        # 95% upper bound
+        point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                                 1.96*(summary(mod)$coefficient[1,3])), 3)
+        # standard error
+        point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
+        # p val
+        point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
+
+        # create else statement that fills matrix with missing so I still have the row
+        # in the final dataframe
+      } else {point_estimates[row_n, 3] <- method_list[row_n]
+      point_estimates[row_n, 4] <- 0
+      point_estimates[row_n, 5] <- 0
+      point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
+    } # end methods loop
+
+    # save point estimates as a dataframe
+    point_est_df <- as_data_frame(point_estimates)
+
+    # combine previous values in dataframe that has all outcome/methods comparisons
+    datalist1[[k+1]] <- point_est_df
+  } # end age category loop
+
+  # bind rows of age category estimates together
+  age_est_df <- bind_rows(datalist1)
+
+  # populate second dataframe list
+  datalist2[[i]] <- age_est_df
+
+} # end of outcome loop
+
+# combine each outcome dataframe itteration in to a big dataset
+combined_point_est_df <- bind_rows(datalist2)
+
+wrf_geo_age <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" |
+    #      pm_method == "WRF-Chem Smoke") %>% 
+  mutate(age_cat2 = ifelse(age_cat == 0, "Less than 15", 
+                    ifelse(age_cat == 1, "15-65",
+                    ifelse(age_cat == 2, "Greater than 65", NA)))) %>% 
+  # subset columns I want to put in to the table
+  select(3, 11, 5:8) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
+         caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by age",
+         # row group by outcome
+         rgroup = "Asthma inhalers",
+         n.rgroup = c(rep(9, 1)), # 6 rows for each age cat for each outcome
+         # column headers
+         header = c("Est Method", "Age Group", "Events",
+                    "OR&dagger;", "Lower", "Upper"),
+         # column spanner
+         cgroup = c("", "95% CI"), 
+         n.cgroup = c(4, 2),
+         padding.rgroup = "&nbsp;&nbsp;",
+         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+         align = "llcccc", # column alignment,
+         tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within May to September fire season."
+         ) # end table
+
+print(tab)
+
+combined_point_est_df_new <- combined_point_est_df %>%
+  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
+                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
+  select(-n_events) %>%
+  rename(n_events = n_events_new)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_age_plot <- combined_point_est_df_new %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  # do not plot results with less than 15 events, create missing vals
+  # in mutate
+  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
+                             99, odds_ratio),
+         lower95 = ifelse(as.numeric(n_events) < 15, 
+                             99, lower95),
+         upper95 = ifelse(as.numeric(n_events) < 15, 
+                             99, upper95),
+         age_cat2 = ifelse(age_cat == 0, "<15", 
+                    ifelse(age_cat == 1, "15-65",
+                    ifelse(age_cat == 2, ">65", NA)))) %>% 
+  # subset columns I want to put in to the table
+  select(1, 3, 11, 10, 5:7) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
+                              levels = unique(wrf_geo_age_plot$outcome))
+
+wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
+                                levels = unique(wrf_geo_age_plot$pm_method))
+
+wrf_geo_age_plot$age_cat2 <- factor(wrf_geo_age_plot$age_cat2,
+                                levels = unique(wrf_geo_age_plot$age_cat2))
+
+wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
+wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
+wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
+wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)
+
+## ggplot 
+  print_plot <- ggplot(wrf_geo_age_plot,
+    aes(x = age_cat2, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red",  "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Age Category ') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10), 
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
+  ggsave("../plot_new/ndc_age_fig_four.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+### Seasonal Time-Stratified by Sex
+
+Association with smoke PM~2.5~ and health outcomes stratified by sex.
+
+All the CI do not include 1, which means the use of asthma medication (beta 2 agonists) increases about 5% when smoke increase 10 units.
+So the risk for getting asthma will increase about 5% ~ 8% when the smoke PM2.5 increase 10 units.
+
+```{r ndc sex strata, warning = F, echo = F, results='asis'} 
+df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)
+
+outcome_list <- c("Asthma", "COPD", "Both", "Other")
+
+# outcome name
+outcome_name <- c("Asthma", "COPD", "Both", "Other")
+
+# sex category list
+sex_strata_list <- c(0,1)
+
+datalist1 <- list()
+datalist2 <- list()
+
+# data wrangling ----
+# Producing conditional logit model estimates loop 
+for(i in 1:length(df_list)){
+
+  # dataframe to loop through
+  df_to_loop <- data.frame(df_list[i])
+  # indication of column
+  outcome <- df_list[i]
+  # outcome name
+  outcome_name <- outcome_list[i]
+
+  covariates_df <- df_to_loop[, c(1:26, 74:84)]
+
+  # extract pm values and divide by 10 and ordered
+  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+  pm_estimates_df <- df_to_loop[, c(87, 92, 91, 93)]/10  # create 10 unit increases
+
+  # new loop for sex categories
+  for(k in 0:1){
+
+    # empty matrix (12 x 10 matrix)
+    point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)
+
+    colnames(point_estimates) <- c('outcome', 'sex', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+    # fill in the outcome namedataframe before method loop
+    point_estimates[, 1] <- outcome_name
+    # repeat the sex category 4 times for each pm method
+    point_estimates[, 2] <- k
+
+    # dataframe for analysis creation
+    # bind columns back together 
+    df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+      # remove missing pm values
+      filter(!is.na(wrf_smk_pm_zip)) %>% 
+      # limit to specific sex category
+      filter(sex_ind == k) %>% 
+      # the following code makes sure that the counterfactual values retained are 
+      # symetric in that number of obs before = number of obs after
+      mutate(obs_diff_admission = (fromdate - date)/7) 
+    # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+    for(j in 38:40){
+
+      # variable to model 
+      var_name <- colnames(df_analysis[j])
+
+      # set row number to fill
+      row_n <- j-37
+
+      # only run the model if the dataframe has observations
+      if(nrow(df_analysis) != 0){
+        # conditional logistic regression model
+        mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+        # populate matrix
+        point_estimates[row_n, 3] <- method_list[row_n]
+        point_estimates[row_n, 4] <- mod$n
+        point_estimates[row_n, 5] <- mod$nevent
+        # odds ratio
+        point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+        # 95% lower bound
+        point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
+                                                 1.96*(summary(mod)$coefficient[1,3])), 3)
+        # 95% upper bound
+        point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                                 1.96*(summary(mod)$coefficient[1,3])), 3)
+        # standard error
+        point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
+        # p val
+        point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
+
+        # create else statement that fills matrix with missing so I still have the row
+        # in the final dataframe
+      } else {point_estimates[row_n, 3] <- method_list[row_n]
+      point_estimates[row_n, 4] <- 0
+      point_estimates[row_n, 5] <- 0
+      point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
+    } # end methods loop
+
+    # save point estimates as a dataframe
+    point_est_df <- as_data_frame(point_estimates)
+
+    # combine previous values in dataframe that has all outcome/methods comparisons
+    datalist1[[k+1]] <- point_est_df
+  } # end sex category loop
+
+  # bind rows of sex category estimates together
+  sex_est_df <- bind_rows(datalist1)
+
+  # populate second dataframe list
+  datalist2[[i]] <- sex_est_df
+
+} # end of outcome loop
+
+# combine each outcome dataframe itteration in to a big dataset
+combined_point_est_df <- bind_rows(datalist2)
+
+wrf_geo_sex <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  mutate(sex_cat = ifelse(sex == 0, "Male", 
+               ifelse(sex == 1, "Female", NA))) %>% 
+  # subset columns I want to put in to the table
+  select(3, 11, 5:8)%>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+  tab <- htmlTable(txtRound(wrf_geo_sex, digits = 3, 1:3), 
+           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and pharmacy outcomes stratified by sex",
+           # row group by outcome
+           rgroup = "Asthma inhalers",
+           n.rgroup = c(rep(6, 1)), # 2 rows for each sex strata for each outcome
+           # column headers
+           header = c("Est Method", "Sex Strata", "Events",
+                      "OR&dagger;", "Lower", "Upper"),
+           # column spanner
+           cgroup = c("", "95% CI"), 
+           n.cgroup = c(4, 2),
+           padding.rgroup = "&nbsp;&nbsp;",
+           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+           align = "llcccc", # column alignment,
+           tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within July to October fire season."
+            ) # end table
+
+  print(tab)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_sex_plot <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | 
+   #       pm_method == "WRF-Chem Smoke") %>% 
+  mutate(sex_cat = ifelse(sex == 0, "Male", 
+               ifelse(sex == 1, "Female", NA))) %>%
+  # subset columns I want to put in to the table
+  select(1, 3, 11, 5:8) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_sex_plot$outcome <- factor(wrf_geo_sex_plot$outcome,
+                              levels = unique(wrf_geo_sex_plot$outcome))
+
+wrf_geo_sex_plot$pm_method <- factor(wrf_geo_sex_plot$pm_method,
+                                levels = unique(wrf_geo_sex_plot$pm_method))
+
+wrf_geo_sex_plot$sex_cat <- factor(wrf_geo_sex_plot$sex_cat,
+                                levels = unique(wrf_geo_sex_plot$sex_cat))
+
+wrf_geo_sex_plot$n_events <- as.numeric(wrf_geo_sex_plot$n_events)
+wrf_geo_sex_plot$odds_ratio <- as.numeric(wrf_geo_sex_plot$odds_ratio)
+wrf_geo_sex_plot$lower95 <- as.numeric(wrf_geo_sex_plot$lower95)
+wrf_geo_sex_plot$upper95 <- as.numeric(wrf_geo_sex_plot$upper95)
+
+## ggplot
+  print_plot <- ggplot(wrf_geo_sex_plot,
+    aes(x = sex_cat, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red", "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Sex') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10),
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
   ggsave("../plot_new/ndc_sex_fig_four.pdf", plot = print_plot, 

### 2017-06-16 r_scripts/analysis/oregon_medication.Rmd
two state distributed lag??

@@
 ```{r library, include=FALSE, echo=FALSE}
-library(dplyr)
+library(tidyverse)
 library(data.table)
@@
 library(knitr)
+
+# two-stage distributed lag model libraries
+library(dlnm)
+library(mvmeta)
 library(splines)
+
 ```
@@
 read_path6 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover.csv")
-asthma_ndc <- read_csv(read_path6)
+asthma_ndc <- fread(read_path6, showProgress = T)
 
@@
 print(plot_ts)
+
 # save figure
@@
                     facet_wrap(~county) + geom_line()
+
+state_n_obs <- or_ndc_county_time_series %>% 
+  select(county, date, n_obs) %>% 
+  gather(key = n_obs, value = n, -date, -county)
+
+count_plot <- ggplot(state_n_obs, aes(x = date, y = n)) +
+  geom_point() +
+  facet_wrap(~county+n_obs, scales = "free") +
+  ylab("Outcome Counts") +
+  xlab("Date") +
+  theme_bw()
+
+print(count_plot)
+
+ggsave("../plot_new/ndc_county_ts2.pdf", plot = count_plot, 
+       width = 12, height = 8, units = "in")
 
@@
 
+### County
+or_smk <- or_ndc_county_time_series_week %>% 
+  select(county, date, wrf_smk_pm, krig_smk_pm, geo_smk_pm) %>% 
+  gather(key = smk_method, value = pm, -date, -county)
+
+or_smk_plot <- ggplot(or_smk, aes(x = date, y = pm)) +
+  geom_point(aes(colour = smk_method), alpha = 0.8, size = 0.8) +
+  facet_wrap(~county) +
+  ggtitle("Oregon: 2013 Fire Season Smoke PM2.5") +
+  ylab("Smoke PM2.5 ug/m^3") +
+  xlab("Year: 2013") +
+  theme_bw()
+
+print(or_smk_plot)
+
+ggsave("../plot_new/ndc_smoke_county.pdf", plot = plot_mult_week, 
+       width = 12, height = 8, units = "in")
+
 ```
 
+## Two stage distributed lag
+
+```{r}
+# create vector of state_county
+counties <- as.character(unique(or_ndc_county_time_series$county))
+
+# create dataframe list of each county
+county_list_df <- lapply(counties, function(x) 
+  or_ndc_county_time_series[or_ndc_county_time_series$county==x,])
+# assign each element of the list object the county name
+names(county_list_df) <- counties
+# calculate length of counties (75)
+m <- length(counties)
+
+# GWR PM2.5 ranges
+smk_range <- t(sapply(county_list_df, 
+  function(x) range(x$geo_smk10,na.rm=T)))
+# smk_range
+
+# population range
+pop_range <- t(sapply(county_list_df, 
+  function(x) range(x$population,na.rm=T)))
+# pop_range
+# -----
+
+# function to compute the QAIC (quasi-poisson AIC)
+fqaic <- function(model) {
+  loglik <- sum(dpois(model$y,model$fitted.values,log=TRUE))
+  phi <- summary(model)$dispersion
+  qaic <- -2*loglik + 2*summary(model)$df[3]*phi
+  return(qaic)
+}
+
+# assuming linear relationship between smoke and outcomes
+argvar <- list(fun="lin")
+
+# fitting a natural spline of the lag function
+degree_freedom <- 4
+arglag <- list(fun="ns", df=degree_freedom)
+# I'm not certain on lag yet, but maybe a week makes sense
+lag <- 7
+
+# overall cumulative summaries
+# columns need to match degrees of freedom for lag
+# 4 degrees of freedom for the spline
+cumulative_counties <- matrix(NA, length(county_list_df), 
+  as.numeric(degree_freedom), 
+  dimnames=list(counties,paste("b",seq(degree_freedom),sep="")))
+
+names(cumulative_counties) <- counties
+
+# (co)variance matrices
+counties_cov <- vector("list",length(county_list_df))
+names(counties_cov) <- counties
+
+# I also want to make a matrix of county-level models that fail to converge
+converge <- matrix(NA, length(county_list_df), 1,
+  dimnames=list(counties, "converge"))
+
+# qaic
+qaic_val <- 0
+
+```
+
+### first stage
+
+```{r}
+# first stage ----
+# I'm not certain on time lag yet, but maybe a week makes sense
+lag <- 7
+
+# create outcome list to loop through
+outcome_names <- "Asthma inhalers"
+# copd is having issues; i suspect counties with no events is the problem
+# empty vector to fill with error messages of counties where models can be run
+# mostly due to no variation in the outcome (i.e. all 0s in the time series)
+
+fit_matrix <- matrix(NA, nrow=0, ncol=6)
+colnames(fit_matrix) <- c("outcome", "exposure", "fit_type", "df", "aic",
+                           "n_mod_conv")
+
+# 1st loop: loop through outcomes
+
+# 2nd loop: vary degrees of freedom for lag knots
+# set up empty matrix to assess model fit
+fit_mat <- matrix(NA, nrow = 3, ncol = 6)
+
+  # 2 df will not work for a spline, going from 3 to 6 df
+  for(h in 3:5){ # start loop for df
+
+    # fitting a natural spline for lag with 5 knots?
+    fit_mat[h-2, 4] <- degree_freedom <- h
+    arglag <- list(fun="ns", df=degree_freedom)
+
+    # overall cumulative summaries
+    # columns need to match degrees of freedom for lag
+    # q: do i need to make this object multiple times or just fill it in?
+    cumulative_counties <- matrix(NA, length(county_list_df), 
+      as.numeric(degree_freedom), 
+      dimnames=list(counties,paste("b",seq(degree_freedom),sep="")))
+
+    names(cumulative_counties) <- counties
+
+    # (co)variance matrices
+    counties_cov <- vector("list",length(county_list_df))
+    names(counties_cov) <- counties
+
+    # I also want to make a matrix of county-level models that fail to converge
+    converge <- matrix(NA, length(county_list_df), 1,
+      dimnames=list(counties, "converge"))
+
+    # empty table to log degrees of freedom and AIC for each outcome
+
+    # qaic
+    qaic_val <- 0
+
+    # first stage ----
+    # loop for each county
+    for(i in seq(county_list_df)) {
+      # if error in following code, move on to next county
+      tryCatch({
+      # load county-specific dataframe
+      sub <- county_list_df[[i]]
+
+      # run only if there is at least one observation of outcome in ts
+      #if(max(sub[, which(colnames(sub)==outcome_names[[q]])]) > 2){
+
+      # define the cross-basis ----
+      suppressWarnings({
+      # cross-basis for smoke
+      smk_cb <- crossbasis(sub$geo_smk10,lag=lag,argvar=argvar,arglag=arglag)
+      # cross-basis for temp
+      temp_cb <- crossbasis(sub$wrf_temp, lag = lag, argvar = argvar,
+                            arglag = arglag)
+      })
+
+      # run the first stage county-specific models
+      county_model <- glm(as.formula(paste0(outcome_names[q], 
+        "~ smk_cb + temp_cb + offset(log(population))")), 
+        family = "quasipoisson", sub)
+
+      # alert warning if there are convergence issues
+      converge[i,] <- county_model$converged
+
+      # reduction to summary results ---
+      # prediction summary for smoke 10 ug/m^3 
+      # uses cross reduce function
+      suppressWarnings({
+        cr_counties <- crossreduce(smk_cb, county_model, type = "var", value = 1)
+
+        })
+
+      # store results in matrix ---
+      # daily lag results
+      cumulative_counties[i,] <- coef(cr_counties)
+      counties_cov[[i]] <- vcov(cr_counties)
+
+      # Q-AIC
+      qaic_val[i] <- fqaic(county_model)
+
+        }, error = function(e){
+          # print(names(county_list_df[i]))
+          }
+        ) # end try catch
+
+      } # end loop for county estimates
+
+    # error at hood river; using trycatch to keep going
+    #which(counties[]=="OR_Hood River")
+
+    # fill in the fit matrix ----
+    fit_mat[,1] <- outcome_names[q]
+    fit_mat[,2] <- exposure <- "wrf_smk10"
+    fit_mat[,3] <- "ns"
+    # subset qaic to only models that converged
+    # sum the grand q aic and fill fit matrix
+    fit_mat[h-2,5] <- round(sum(qaic_val[!is.na(converge)[]==T]),2)
+    # fill in number of county models that successfully converge
+    fit_mat[h-2, 6] <- length(which(converge==T))
+    sum(complete.cases(qaic_val))
+
+    } # end model fit loop
+
+ fit_matrix <- rbind(fit_matrix, fit_mat)
+
+# end outcome loop and system time
+
+# now that we have a range of df/knots, I want to filter to the minimum aic
+# for each outcome and print out that value in a table
+
+lag_spline_best_fit <- as_tibble(fit_matrix) %>%  
+  group_by(outcome) %>% 
+  slice(which.min(aic))
+
+# kable
+knitr::kable(lag_spline_best_fit, caption = paste0("Distributed lag spline",
+  "degree of freedom best fit by AIC for two-stage approach"))
+
+```
+
+```{r}
+# set estimation method for multivariate approach
+# reml estimation method
+method <- "reml"
+
+# create a dataframe that contains the distributed lag estimates for each outcome
+dist_lag_df <- as_tibble(matrix(NA,  nrow = 0, ncol = 5))
+colnames(dist_lag_df) <- c("outcome", "lag", "rel_risk", 
+                                  "lower_95", "upper_95")
+
+# produce results using fit matrix
+
+# 1st loop: loop through outcomes
+for(j in 1:length(lag_spline_best_fit$outcome)){
+
+  j = 1
+    # easiest to set out the parameters from the start
+    outcome <- lag_spline_best_fit$outcome[j]
+    degree_freedom <- as.numeric(lag_spline_best_fit$df[j])
+
+    arglag <- list(fun="ns", df=degree_freedom)
+
+    # overall cumulative summaries
+    # columns need to match degrees of freedom for lag
+    # q: do i need to make this object multiple times or just fill it in?
+    cumulative_counties <- matrix(NA, length(county_list_df), 
+      as.numeric(degree_freedom), 
+      dimnames=list(counties,paste("b",seq(degree_freedom),sep="")))
+
+    names(cumulative_counties) <- counties
+
+    # (co)variance matrices
+    counties_cov <- vector("list",length(county_list_df))
+    names(counties_cov) <- counties
+
+    # I also want to make a matrix of county-level models that fail to converge
+    converge <- matrix(NA, length(county_list_df), 1,
+      dimnames=list(counties, "converge"))
+
+    # first stage loop for each county ----
+    for(i in seq(county_list_df)) {
+      # if error in following code, move on to next county using trycatch
+      tryCatch({
+      # load county-specific dataframe
+      sub <- county_list_df[[i]]
+
+      # define the cross-basis
+      suppressWarnings({
+      # cross-basis for smoke
+      smk_cb <- crossbasis(sub$geo_smk10,lag=lag,argvar=argvar,arglag=arglag)
+      # cross-basis for temp
+      temp_cb <- crossbasis(sub$wrf_temp, lag = lag, argvar = argvar,
+                            arglag = arglag)
+      })
+
+      # run the first stage county-specific models
+      county_model <- glm(as.formula(paste0(outcome, 
+        "~ smk_cb + temp_cb + offset(log(population))")), 
+        family = "quasipoisson", sub)
+
+      # alert warning if there are convergence issues
+      converge[i,] <- county_model$converged
+
+      # reduction to summary results ---
+      # prediction summary for smoke 10 ug/m^3 
+      # uses cross reduce function
+      suppressWarnings({
+        cr_counties <- crossreduce(smk_cb, county_model, type = "var", value = 1)
+        })
+
+      # store results in matrix ---
+      # daily lag results
+      cumulative_counties[i,] <- coef(cr_counties)
+      counties_cov[[i]] <- vcov(cr_counties)
+        }, error = function(e){
+          #print(names(county_list_df[i]))
+          }) # end try catch
+
+      } # end loop for county estimates
+
+  # second stage multivariate meta analysis ---- 
+  # For counties with outcomes that will not even run, set their converge
+  # from missing to False for subsetting
+  converge[is.na(converge)] <- FALSE
+
+  # first step is to reduce the county-specific matrix to those that converged
+  meta_counties <- cumulative_counties[converge[]==T, ]
+  # reduce covariance matrix to models that converged
+  meta_cov <- counties_cov[converge[]==T]
+
+  # overall pooled summary for asthma model
+  cumulative_mv_meta <- mvmeta(meta_counties~1, meta_cov, method = method)
+
+  # create bases for prediction ----
+  # bases of smoke lag (7 days with 10 intervals in each)
+  xlag <- 0:70/10
+  blag <- do.call("onebasis",c(list(x=xlag),attr(smk_cb,"arglag")))
+
+  # county-specific first stage summaries
+  county_lag_estimates <- apply(cumulative_counties,1, function(x) exp(blag%*%x))
+
+  # prediction for a grid of lag days
+  # predictor-specific summaries for 10 ug/m^3 of smoke
+  pooled_cp <- crosspred(blag, coef=coef(cumulative_mv_meta), 
+    vcov=vcov(cumulative_mv_meta), model.link="log", at = xlag)
+
+  # create empty dataframe to fill with estimates
+  # fill lag relationship with outcome tibble
+  # empty matrix to fill in distributed lag values
+  outcome_smk_lag_df <- as_tibble(matrix(NA,  nrow = length(xlag), ncol = 5))
+  # assign outcome name
+  colnames(outcome_smk_lag_df) <- c("outcome", "lag", "rel_risk", 
+                                  "lower_95", "upper_95")
+
+  # fill dataframe with values
+  outcome_smk_lag_df[,1] <- outcome
+  outcome_smk_lag_df[,2] <- xlag
+  outcome_smk_lag_df[,3] <- pooled_cp$matRRfit
+  outcome_smk_lag_df[,4] <- pooled_cp$matRRlow
+  outcome_smk_lag_df[,5] <- pooled_cp$matRRhigh
+
+  # bind works better than trying to figure out how to fill many rows
+  dist_lag_df <- rbind(dist_lag_df, outcome_smk_lag_df)
+
+}
+
+ggplot(dist_lag_df, aes(x=lag, y=rel_risk)) +
+  geom_line(colour = "blue") +
+  geom_ribbon(aes(ymin = lower_95, ymax = upper_95), fill = "blue", alpha = 0.3) + 
+  scale_y_continuous(limits = c(0.9, 1.15)) +
+  scale_x_continuous(breaks = c(seq(0, 7,by=1))) +
+  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
+  facet_wrap(~ outcome) +
+  ggtitle(paste0("Distributed lag association between a 10 ug/m^3",
+                 " increase in GWR smoke PM2.5 and cardiopulmonary outcomes")) +
+  ylab("Relative Risk for a 10 ug/m^3 increase in GWR smoke (fixed)") +
+  xlab("Lagged Days") +
+  theme_bw()
+
+```
+
 ## Distributed Lag Model
 
-```{r DLM, echo = FALSE, message=FALSE, warning=FALSE}
+### Lag 1
+
+```{r DLM, echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}
+asthma_ndc_df <- data.frame(asthma_ndc)
+
+asthma_ndc_df <- asthma_ndc_df %>% # GWR
+  filter(!is.na(geo_wt_pm_lag7_zip)) # 503998
 
@@
 asthma_ndc_lag <- asthma_ndc_df %>%
-  filter(gender!="U") %>%
-  filter(outcome==1) %>%
   select(88, 115:121) # GWR-method
@@
 asthma_ndc_lag <- as.matrix(asthma_ndc_lag)
-
-# true DL function
-DLfunc <- exp(-(0:lag_max)/2)
-
-# simulate outcomes
-y <- drop(asthma_ndc_lag%*%DLfunc) # + rnorm(n)
 
@@
 asthma_ndc_lag_B <- asthma_ndc_lag%*%B
-fit <- lm(y~asthma_ndc_lag_B)
+fit <- clogit(outcome ~ asthma_ndc_lag_B + wrf_temp_zip + strata(personkey), asthma_ndc_df)
 coef(fit)
+summary(fit)
 
@@
 DLestimate$SE <- sqrt(diag(DLvar))
-DLestimate$lower <- DLestimate$estimate + DLestimate$SE * qt(1-0.975, df=fit$df.residual)
-DLestimate$upper <- DLestimate$estimate + DLestimate$SE * qt(0.975, df=fit$df.residual)
+DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
+DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
+
+DLestimate$estimate <- exp(DLestimate$estimate)
+DLestimate$lower <- exp(DLestimate$lower)
+DLestimate$upper <- exp(DLestimate$upper)
+
 DLestimate
@@
 DLestimate$lag <- 0:lag_max
-DLestimate$truth <- DLfunc
 p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
 p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
-p <- p + geom_line(aes(x=lag, y=truth), linetype=2, size=2)
+p <- p + geom_line(aes(x=lag), linetype=2, size=2)
 p <- p + theme_bw()
@@
 
+DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms])
+
 cumulative <- sum(DLestimate$estimate)
 cumulative
-## [1] 2.485711
+## [1] 0.009141771
 cumulative_se <- sqrt(sum(DLvar))
 cumulative_se
-## [1] 0.0001887442
-cumulative_CI <- cumulative + cumulative_se * qt(c(1-0.975,0.975), df=fit$df.residual)
+## [1] 0.000955469
+cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
 cumulative_CI
-## [1] 2.485341 2.486080
+## [1] 0.007269052 0.011014490
 
@@
 
-```{r overall lag1, warning =F, echo = F, results='asis'} 
+```{r overall lag 1, warning =F, echo = F, results='asis'} 
 # dataframe list
@@
 # outcome name
-outcome <- "Asthma Inhalers"
-outcome_name <- "Asthma Inhalers"
+outcome <- "Asthma Inhalers Lag 1"
+outcome_name <- "Asthma Inhalers Lag 1"
 
@@
          # row group by outcome
-         rgroup = "Asthma Inhalers",
+         rgroup = "Asthma Inhalers Lag 1",
          n.rgroup = c(rep(3, 1)), # 6 rows for each age cat for each outcome
@@
 # outcome name
-outcome <- "asthma inhalers"
-outcome_name <- "Asthma inhalers"
+outcome <- "asthma inhalers Lag 1"
+outcome_name <- "Asthma inhalers Lag 1"
 
@@
          # row group by outcome
-         rgroup = "Asthma inhalers",
+         rgroup = "Asthma inhalers Lag 1",
          n.rgroup = c(rep(9, 1)), # 6 rows for each age cat for each outcome
@@
            # row group by outcome
-           rgroup = "Asthma inhalers",
+           rgroup = "Asthma inhalers Lag 1",
            n.rgroup = c(rep(6, 1)), # 2 rows for each sex strata for each outcome
@@
 
-## NDC claims with index
-
-Check the claims in each category (Asthma, COPD, Both or Other).
-
-```{r echo = FALSE, message=FALSE, warning=FALSE}
-read_path7 <- paste0("../data_new/medication/oregon_beta_ndc_index.csv")
-or_ndc_index <- read_csv(read_path7)
-
-oregon_beta_df <- or_ndc_index %>% 
-  group_by(personkey) %>% 
-  arrange(personkey, fromdate, line) %>% 
-  mutate(num_visit = dense_rank(fromdate)) %>%
-  arrange(personkey, num_visit) %>% 
-  select(personkey, clmid, num_visit) %>% 
-  unique() %>% 
-  full_join(or_ndc_index, by = c("personkey", "clmid")) %>% 
-  arrange(personkey, clmid, line, fromdate) %>% 
-  filter(row_number() == 1) %>%
-  # add new transverted from date
-  mutate(dates = as.Date(fromdate, "%m/%d/%Y")) %>%
-  filter(dates >= '2013-05-01' & 
-         dates <= '2013-09-30') # 24538
-
-summary(as.factor(oregon_beta_df$index))
-
-```
-
-### Descriptives
-
-For all of 2013, there were a total of millons of all claims records in the APAC dataset.
-During the May 1st to September 30th wildfire season in Oregon state, there were 24217 pharmacy visits for the ndc of beta 2 agonists.
-The number of cases, as well as age-specific and sex-specific strate are shown below.
-
-```{r ndc descriptive table, echo = F, message = F, warning = F, results='asis'}
-read_path6 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover.csv")
-beta_ndc <- read_csv(read_path6)
-
-# join personkey with index
-oregon_beta_index_df <- oregon_beta_df %>%
-  select(personkey, index)
-
-beta_ndc_with_index <- beta_ndc %>%
-  left_join(oregon_beta_index_df, by="personkey")
+### Lag 2
+
+```{r overall lag 2, warning =F, echo = F, results='asis'} 
+# dataframe list
+
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
 
 # dataframe to loop through
-beta_ndc_with_index_df <- data.frame(beta_ndc_with_index)
-
-beta_asthma <- beta_ndc_with_index_df %>%
-  filter(index=="asthma")
-
-beta_copd <- beta_ndc_with_index_df %>%
-  filter(index=="copd")
-
-beta_both <- beta_ndc_with_index_df %>%
-  filter(index=="both")
-
-beta_other <- beta_ndc_with_index_df %>%
-  filter(index=="other")
-
-df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)
-
-outcome_list <- c("Asthma", "COPD", "Both", "Other")
-
-# create an empty list to row bind dataframes together
-datalist <- list()
-
-# dataframe list
-method_list <- c('WRF-Chem Smoke', 'Geo-Weighted Smoke', 'Kriging Smoke')
+# asthma_ndc_df <- data.frame(asthma_ndc)
 
 # outcome name
-outcome_name <- c("Asthma", "COPD", "Both", "Other")
-
-for(i in 1:length(df_list)){
-  # dataframe to loop through
-  df_to_loop <- data.frame(df_list[i])
-  # indication of column
-  outcome <- colnames(df_to_loop[75])
-  # outcome name
-  outcome_name <- outcome_list[i]
-
-  # dataframe for analysis creation
-  # bind columns back together 
-  df_analysis <- df_to_loop %>%
-    filter(!is.na(wrf_smk_pm_zip)) %>% 
-    # only look at outcomes
-    filter(outcome == 1) %>%
-    # add another row that makes sure there is a person <15 in the dataframe
-    # tricking xtabs to produce a 0 cell for the outcome for age <15
-    add_row(outcome = 0, age_ind = 0)
-
-  # cross tabs
-  outcome_n <- xtabs(~ outcome, df_analysis)
-  cross_tab_age <- xtabs(~ outcome + age_ind, df_analysis)
-  cross_tab_sex <- xtabs(~ outcome + sex_ind, df_analysis)
-  # empty matrix
-  point_estimates <- matrix(nrow = 1, ncol = 7, byrow = T)
-
-  colnames(point_estimates) <- c("outcome", "n", "age_15", "age_15_65", 
-                                 "age_65", "female", "male")
-
-  # fill in the outcome name for the dataframe before the loop
-  point_estimates[, 1] <- outcome_name
-  # fill n
-  point_estimates[, 2] <- outcome_n[2] # second element of the 1 dimension vector
-  # age <15
-  point_estimates[, 3] <- cross_tab_age[2, 1]
-  # age 15 to 65
-  point_estimates[, 4] <- cross_tab_age[2, 2]
-  # age >65
-  point_estimates[, 5] <- cross_tab_age[2, 3]
-  # male == 0
-  point_estimates[, 7] <- cross_tab_sex[1, 1]
-  # female == 1
-  point_estimates[, 6] <- cross_tab_sex[1, 2]
-
-  # save point estimates as a dataframe
-  point_est_df <- as_data_frame(point_estimates)
-
-  # combine previous values in dataframe that has all outcome/methods comparisons
-  datalist[[i]] <- point_est_df
-
-} # end of loop
-
-# combine each outcome dataframe itteration in to a big dataset
-combined_point_est_df <- bind_rows(datalist) %>% 
-  # find proportions/percents for each strata in a row
-  mutate(age_15_pr = as.character(round((as.numeric(age_15)/as.numeric(n))*100,1)),
-         age_15_65_pr = as.character(round((as.numeric(age_15_65)/as.numeric(n))*100,1)),
-         age_65_pr = as.character(round((as.numeric(age_65)/as.numeric(n))*100,1)),
-         female_pr = as.character(round((as.numeric(female)/as.numeric(n))*100,1)),
-         male_pr = as.character(round((as.numeric(male)/as.numeric(n))*100,1))) %>% 
-  select(outcome, n, age_15, age_15_pr, age_15_65, age_15_65_pr, age_65,
-         age_65_pr, female, female_pr, male, male_pr)
-
-#str(combined_point_est_df)
-
-tab <- htmlTable(txtRound(combined_point_est_df, digits = 1), 
-           caption = "Number of cases for each outcome observed from May 1st to September 30st, 2013",
-           # column headers
-           header = c("Outcome", "Cases n", "Less than 15", "(%)", "15 to 65", "(%)", "Greater than 65",
-                      "(%)", "Female", "(%)", "Male", "(%)"),
-           # column spanner
-           cgroup = c("","Age Category", "Sex"), 
-           n.cgroup = c(2, 6, 4),
-           padding.rgroup = "&nbsp;&nbsp;",
-           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
-           align = "llccccc" # column alignment,
-            ) # end table
-
-print(tab)
-
-```
-
-### Overall plot
-
-```{r ndc overall try, warning =F, echo = F, results='asis'}
-df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)
-
-outcome_list <- c("Asthma", "COPD", "Both", "Other")
-
-# create an empty list to row bind dataframes together
-datalist <- list()
-
-# dataframe list
-method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
-
-# outcome name
-outcome_name <- c("Asthma", "COPD", "Both", "Other")
+outcome <- "Asthma Inhalers Lag 1"
+outcome_name <- "Asthma Inhalers Lag 1"
 
@@
 
-  # combine previous values in dataframe that has all outcome/methods comparisons
-  datalist[[i]] <- point_est_df
-
-} # end of loop
-
-# combine each outcome dataframe itteration in to a big dataset
-combined_point_est_df <- bind_rows(datalist)
-
+# save point estimates as a dataframe
+combined_point_est_df <- as_data_frame(point_estimates)
+
+wrf_geo_age <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" |
+    #      pm_method == "WRF-Chem Smoke") %>% 
   # subset columns I want to put in to the table
-table_df <- combined_point_est_df %>% select(2, 3:7)
-
-tab <- htmlTable(txtRound(table_df, digits = 3, 1:3), 
-                 caption = "Association between a 10 ug/m^3 in PM2.5 and Health Outcomes",
+  select(2, 4:7) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
+         caption = "Association between a 10 ug/m^3 in PM2.5 three smoke method and pharmacy outcomes",
          # row group by outcome
-                 rgroup = outcome_list,
-                 n.rgroup = c(rep(3, 4)), # 4 rows for each method for each outcome
+         rgroup = "Asthma Inhalers Lag 1",
+         n.rgroup = c(rep(3, 1)), # 6 rows for each age cat for each outcome
          # column headers
-                 header = c("Method", "Obs.", "Events",
+         header = c("Est Method", "Events",
                     "OR&dagger;", "Lower", "Upper"),
@@
 
-### Time-Stratified by Age Categories
-
-Following table and figure look at some outcomes stratified by age category.
-
-For some outcomes, such like Age > 65 in WRF-Chem method, the CI is the only insignificant one in the plot.
-
-```{r ndc time stratified age strata, warning =F, echo = F, results='asis'} 
-df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)
-
-outcome_list <- c("Asthma", "COPD", "Both", "Other")
-
-# create an empty list to row bind dataframes together
-datalist <- list()
+```{r time stratified age strata lag 2, warning =F, echo = F, results='asis'} 
+# dataframe list
+
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# dataframe to loop through
+asthma_ndc_df <- data.frame(asthma_ndc)
 
 # outcome name
-outcome_name <- c("Asthma", "COPD", "Both", "Other")
+outcome <- "asthma inhalers Lag 1"
+outcome_name <- "Asthma inhalers Lag 1"
 
@@
 
-### Seasonal Time-Stratified by Sex
-
-Association with smoke PM~2.5~ and health outcomes stratified by sex.
-
-All the CI do not include 1, which means the use of asthma medication (beta 2 agonists) increases about 5% when smoke increase 10 units.
-So the risk for getting asthma will increase about 5% ~ 8% when the smoke PM2.5 increase 10 units.
-
-```{r ndc sex strata, warning = F, echo = F, results='asis'} 
-df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)
-
-outcome_list <- c("Asthma", "COPD", "Both", "Other")
+```{r sex strata cdc met  lag 2, warning = F, echo = F, results='asis'} 
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# dataframe to loop through
+asthma_ndc_df <- data.frame(asthma_ndc)
 
 # outcome name
-outcome_name <- c("Asthma", "COPD", "Both", "Other")
+outcome <- "asthma inhalers Lag 1"
+outcome_name <- "Asthma inhalers Lag 1"
+
+# create an empty list to row bind dataframes together
+datalist1 <- list()
 
@@
   # save figure
+  ggsave("../plot_lag/ndc_sex_fig_three_lag1.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+### Lag 3
+
+```{r overall lag 3, warning =F, echo = F, results='asis'} 
+# dataframe list
+
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# dataframe to loop through
+# asthma_ndc_df <- data.frame(asthma_ndc)
+
+# outcome name
+outcome <- "Asthma Inhalers Lag 1"
+outcome_name <- "Asthma Inhalers Lag 1"
+
+# data wrangling ----
+# Producing conditional logit model estimates loop
+
+# extract covariates from dataframe
+covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]
+
+# extract pm values and divide by 10 and ordered
+which(colnames(asthma_ndc_df)=="wrf_smk_pm_lag1_zip") # 108
+which(colnames(asthma_ndc_df)=="krig_pm_lag1_zip") # 122
+which(colnames(asthma_ndc_df)=="geo_wt_pm_lag1_zip") # code to find column numbers, 115
+pm_estimates_df <- asthma_ndc_df[, c(108, 122, 115, 93)]/10  # create 10 unit increases
+
+# new loop for age categories
+
+# empty matrix (12 x 10 matrix)
+point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)
+
+colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+# fill in the outcome namedataframe before method loop
+point_estimates[, 1] <- outcome_name
+
+# dataframe for analysis creation
+# bind columns back together 
+df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+  # remove missing pm values
+  filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
+  filter(!is.na(krig_pm_lag1_zip)) %>% 
+  filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
+  # the following code makes sure that the counterfactual values retained are 
+  # symetric in that number of obs before = number of obs after
+  mutate(obs_diff_admission = (fromdate - date)/7) 
+  # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+# second loop to run a model for each pm estimation method
+for(j in 38:40){
+
+  # variable to model 
+  var_name <- colnames(df_analysis[j])
+
+  # set row number to fill
+  row_n <- j-37
+
+  # only run the model if the dataframe has observations
+  if(nrow(df_analysis) != 0){
+  # conditional logistic regression model
+  mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+  # populate matrix
+  point_estimates[row_n, 2] <- method_list[row_n]
+  point_estimates[row_n, 3] <- mod$n
+  point_estimates[row_n, 4] <- mod$nevent
+  # odds ratio
+  point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+  # 95% lower bound
+  point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
+  # 95% upper bound
+  point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                        1.96*(summary(mod)$coefficient[1,3])), 3)
+  # standard error
+  point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
+  # p val
+  point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)
+
+  # create else statement that fills matrix with missing so I still have the row
+  # in the final dataframe
+  } else {point_estimates[row_n, 2] <- method_list[row_n]
+          point_estimates[row_n, 3] <- 0
+          point_estimates[row_n, 4] <- 0
+          point_estimates[row_n, c(5:9)] <- 99 } # end 'if else' statement
+  } # end methods loop
+
+# save point estimates as a dataframe
+combined_point_est_df <- as_data_frame(point_estimates)
+
+wrf_geo_age <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" |
+    #      pm_method == "WRF-Chem Smoke") %>% 
+  # subset columns I want to put in to the table
+  select(2, 4:7) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
+         caption = "Association between a 10 ug/m^3 in PM2.5 three smoke method and pharmacy outcomes",
+         # row group by outcome
+         rgroup = "Asthma Inhalers Lag 1",
+         n.rgroup = c(rep(3, 1)), # 6 rows for each age cat for each outcome
+         # column headers
+         header = c("Est Method", "Events",
+                    "OR&dagger;", "Lower", "Upper"),
+         # column spanner
+         cgroup = c("", "95% CI"), 
+         n.cgroup = c(3, 2),
+         padding.rgroup = "&nbsp;&nbsp;",
+         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+         align = "llcccc", # column alignment,
+         tfoot="&dagger; Referent periods matched to events on same day of week within May to September fire season."
+         ) # end table
+
+print(tab)
+
+combined_point_est_df_new <- combined_point_est_df %>%
+  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
+                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
+  select(-n_events) %>%
+  rename(n_events = n_events_new)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_age_plot <- combined_point_est_df_new %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  # do not plot results with less than 15 events, create missing vals
+  # in mutate
+  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
+                             99, odds_ratio),
+         lower95 = ifelse(as.numeric(n_events) < 15, 
+                             99, lower95),
+         upper95 = ifelse(as.numeric(n_events) < 15, 
+                             99, upper95)) %>% 
+  # subset columns I want to put in to the table
+  select(1, 2, 9, 4:6) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
+                              levels = unique(wrf_geo_age_plot$outcome))
+
+wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
+                                levels = unique(wrf_geo_age_plot$pm_method))
+
+wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
+wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
+wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
+wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)
+
+## ggplot 
+  print_plot <- ggplot(wrf_geo_age_plot,
+    aes(x = pm_method, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red",  "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Smoke Estimation Method') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10),
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
+  ggsave("../plot_lag/ndc_overall_lag1_three.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+```{r time stratified age strata lag 3, warning =F, echo = F, results='asis'} 
+# dataframe list
+
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# dataframe to loop through
+asthma_ndc_df <- data.frame(asthma_ndc)
+
+# outcome name
+outcome <- "asthma inhalers Lag 1"
+outcome_name <- "Asthma inhalers Lag 1"
+
+# age category list
+age_cat_list <- c(0,1,2)
+
+# create an empty list to row bind dataframes together
+datalist1 <- list()
+
+# data wrangling ----
+# Producing conditional logit model estimates loop
+
+# extract covariates from dataframe
+covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]
+
+# extract pm values and divide by 10 and ordered
+#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+pm_estimates_df <- asthma_ndc_df[, c(108, 122, 115, 93)]/10  # create 10 unit increases
+
+# new loop for age categories
+for(k in 0:2){
+  # empty matrix (12 x 10 matrix)
+  point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)
+
+  colnames(point_estimates) <- c('outcome', 'age_cat', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+  # fill in the outcome namedataframe before method loop
+  point_estimates[, 1] <- outcome_name
+  # repeat the age category 4 times for each pm method
+  point_estimates[, 2] <- k
+
+  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+    # remove missing pm values
+    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
+    filter(!is.na(krig_pm_lag1_zip)) %>% 
+    filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
+    # limit to specific age category
+    filter(age_ind == k) %>% 
+    # the following code makes sure that the counterfactual values retained are 
+    # symetric in that number of obs before = number of obs after
+    mutate(obs_diff_admission = (fromdate - date)/7) 
+    # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+  # second loop to run a model for each pm estimation method
+  for(j in 38:40){
+
+    # variable to model 
+    var_name <- colnames(df_analysis[j])
+
+    # set row number to fill
+    row_n <- j-37
+
+    # only run the model if the dataframe has observations
+    if(nrow(df_analysis) != 0){
+    # conditional logistic regression model
+    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+    # populate matrix
+    point_estimates[row_n, 3] <- method_list[row_n]
+    point_estimates[row_n, 4] <- mod$n
+    point_estimates[row_n, 5] <- mod$nevent
+    # odds ratio
+    point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+    # 95% lower bound
+    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
+    # 95% upper bound
+    point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                        1.96*(summary(mod)$coefficient[1,3])), 3)
+    # standard error
+    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
+    # p val
+    point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
+
+    # create else statement that fills matrix with missing so I still have the row
+    # in the final dataframe
+    } else {point_estimates[row_n, 3] <- method_list[row_n]
+            point_estimates[row_n, 4] <- 0
+            point_estimates[row_n, 5] <- 0
+            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
+    } # end methods loop
+
+  # save point estimates as a dataframe
+  point_est_df <- as_data_frame(point_estimates)
+  datalist1[[k+1]] <- point_est_df
+} # end age category loop
+
+# bind rows of age category estimates together
+combined_point_est_df <- bind_rows(datalist1)
+
+wrf_geo_age <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" |
+    #      pm_method == "WRF-Chem Smoke") %>% 
+  mutate(age_cat2 = ifelse(age_cat == 0, "Less than 15", 
+                    ifelse(age_cat == 1, "15-65",
+                    ifelse(age_cat == 2, "Greater than 65", NA)))) %>% 
+  # subset columns I want to put in to the table
+  select(3, 11, 5:8) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
+         caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by age",
+         # row group by outcome
+         rgroup = "Asthma inhalers Lag 1",
+         n.rgroup = c(rep(9, 1)), # 6 rows for each age cat for each outcome
+         # column headers
+         header = c("Est Method", "Age Group", "Events",
+                    "OR&dagger;", "Lower", "Upper"),
+         # column spanner
+         cgroup = c("", "95% CI"), 
+         n.cgroup = c(4, 2),
+         padding.rgroup = "&nbsp;&nbsp;",
+         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+         align = "llcccc", # column alignment,
+         tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within May to September fire season."
+         ) # end table
+
+print(tab)
+
+combined_point_est_df_new <- combined_point_est_df %>%
+  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
+                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
+  select(-n_events) %>%
+  rename(n_events = n_events_new)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_age_plot <- combined_point_est_df_new %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  # do not plot results with less than 15 events, create missing vals
+  # in mutate
+  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
+                             99, odds_ratio),
+         lower95 = ifelse(as.numeric(n_events) < 15, 
+                             99, lower95),
+         upper95 = ifelse(as.numeric(n_events) < 15, 
+                             99, upper95),
+         age_cat2 = ifelse(age_cat == 0, "<15", 
+                    ifelse(age_cat == 1, "15-65",
+                    ifelse(age_cat == 2, ">65", NA)))) %>% 
+  # subset columns I want to put in to the table
+  select(1, 3, 11, 10, 5:7) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
+                              levels = unique(wrf_geo_age_plot$outcome))
+
+wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
+                                levels = unique(wrf_geo_age_plot$pm_method))
+
+wrf_geo_age_plot$age_cat2 <- factor(wrf_geo_age_plot$age_cat2,
+                                levels = unique(wrf_geo_age_plot$age_cat2))
+
+wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
+wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
+wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
+wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)
+
+## ggplot 
+  print_plot <- ggplot(wrf_geo_age_plot,
+    aes(x = age_cat2, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red",  "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Age Category ') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10), 
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
+  ggsave("../plot_lag/ndc_age_fig_three_lag1.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+```{r sex strata cdc met  lag 3, warning = F, echo = F, results='asis'} 
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# dataframe to loop through
+asthma_ndc_df <- data.frame(asthma_ndc)
+
+# outcome name
+outcome <- "asthma inhalers Lag 1"
+outcome_name <- "Asthma inhalers Lag 1"
+
+# create an empty list to row bind dataframes together
+datalist1 <- list()
+
+# sex category list
+sex_strata_list <- c(0,1)
+
+# data wrangling ----
+# Producing conditional logit model estimates loop
+
+# extract covariates from dataframe
+covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]
+
+# extract pm values and divide by 10 and ordered
+#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+pm_estimates_df <- asthma_ndc_df[, c(108, 122, 115, 93)]/10  # create 10 unit increases
+
+# new loop for sex categories
+for(k in 0:1){
+
+  # empty matrix (12 x 10 matrix)
+  point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)
+
+  colnames(point_estimates) <- c('outcome', 'sex', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+  # fill in the outcome namedataframe before method loop
+  point_estimates[, 1] <- outcome_name
+  # repeat the sex category 4 times for each pm method
+  point_estimates[, 2] <- k
+
+  # dataframe for analysis creation
+  # bind columns back together 
+   df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+    # remove missing pm values
+    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
+    filter(!is.na(krig_pm_lag1_zip)) %>% 
+    filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
+    # limit to specific age category
+    filter(sex_ind == k) %>% 
+    # the following code makes sure that the counterfactual values retained are 
+    # symetric in that number of obs before = number of obs after
+    mutate(obs_diff_admission = (fromdate - date)/7) 
+    # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+  # second loop to run a model for each pm estimation method
+  for(j in 38:40){
+
+    # variable to model 
+    var_name <- colnames(df_analysis[j])
+
+    # set row number to fill
+    row_n <- j-37
+
+    # only run the model if the dataframe has observations
+    if(nrow(df_analysis) != 0){
+    # conditional logistic regression model
+    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+    # populate matrix
+    point_estimates[row_n, 3] <- method_list[row_n]
+    point_estimates[row_n, 4] <- mod$n
+    point_estimates[row_n, 5] <- mod$nevent
+    # odds ratio
+    point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+    # 95% lower bound
+    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
+                                      1.96*(summary(mod)$coefficient[1,3])), 3)
+    # 95% upper bound
+    point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                      1.96*(summary(mod)$coefficient[1,3])), 3)
+    # standard error
+    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
+    # p val
+    point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
+
+    # create else statement that fills matrix with missing so I still have the row
+    # in the final dataframe
+    } else {point_estimates[row_n, 3] <- method_list[row_n]
+            point_estimates[row_n, 4] <- 0
+            point_estimates[row_n, 5] <- 0
+            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
+    } # end methods loop
+
+  # save point estimates as a dataframe
+  point_est_df <- as_data_frame(point_estimates)
+
+  # combine previous values in dataframe that has all outcome/methods comparisons
+  datalist1[[k+1]] <- point_est_df
+} # end sex category loop
+
+# bind rows of sex category estimates together
+combined_point_est_df <- bind_rows(datalist1)
+
+wrf_geo_sex <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  mutate(sex_cat = ifelse(sex == 0, "Male", 
+               ifelse(sex == 1, "Female", NA))) %>% 
+  # subset columns I want to put in to the table
+  select(3, 11, 5:8)%>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+  tab <- htmlTable(txtRound(wrf_geo_sex, digits = 3, 1:3), 
+           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and pharmacy outcomes stratified by sex",
+           # row group by outcome
+           rgroup = "Asthma inhalers Lag 1",
+           n.rgroup = c(rep(6, 1)), # 2 rows for each sex strata for each outcome
+           # column headers
+           header = c("Est Method", "Sex Strata", "Events",
+                      "OR&dagger;", "Lower", "Upper"),
+           # column spanner
+           cgroup = c("", "95% CI"), 
+           n.cgroup = c(4, 2),
+           padding.rgroup = "&nbsp;&nbsp;",
+           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+           align = "llcccc", # column alignment,
+           tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within July to October fire season."
+            ) # end table
+
+  print(tab)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_sex_plot <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | 
+   #       pm_method == "WRF-Chem Smoke") %>% 
+  mutate(sex_cat = ifelse(sex == 0, "Male", 
+               ifelse(sex == 1, "Female", NA))) %>%
+  # subset columns I want to put in to the table
+  select(1, 3, 11, 5:8) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_sex_plot$outcome <- factor(wrf_geo_sex_plot$outcome,
+                              levels = unique(wrf_geo_sex_plot$outcome))
+
+wrf_geo_sex_plot$pm_method <- factor(wrf_geo_sex_plot$pm_method,
+                                levels = unique(wrf_geo_sex_plot$pm_method))
+
+wrf_geo_sex_plot$sex_cat <- factor(wrf_geo_sex_plot$sex_cat,
+                                levels = unique(wrf_geo_sex_plot$sex_cat))
+
+wrf_geo_sex_plot$n_events <- as.numeric(wrf_geo_sex_plot$n_events)
+wrf_geo_sex_plot$odds_ratio <- as.numeric(wrf_geo_sex_plot$odds_ratio)
+wrf_geo_sex_plot$lower95 <- as.numeric(wrf_geo_sex_plot$lower95)
+wrf_geo_sex_plot$upper95 <- as.numeric(wrf_geo_sex_plot$upper95)
+
+## ggplot
+  print_plot <- ggplot(wrf_geo_sex_plot,
+    aes(x = sex_cat, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red", "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Sex') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10),
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
+  ggsave("../plot_lag/ndc_sex_fig_three_lag1.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+### Lag 4
+
+```{r overall lag 4, warning =F, echo = F, results='asis'} 
+# dataframe list
+
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# dataframe to loop through
+# asthma_ndc_df <- data.frame(asthma_ndc)
+
+# outcome name
+outcome <- "Asthma Inhalers Lag 1"
+outcome_name <- "Asthma Inhalers Lag 1"
+
+# data wrangling ----
+# Producing conditional logit model estimates loop
+
+# extract covariates from dataframe
+covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]
+
+# extract pm values and divide by 10 and ordered
+which(colnames(asthma_ndc_df)=="wrf_smk_pm_lag1_zip") # 108
+which(colnames(asthma_ndc_df)=="krig_pm_lag1_zip") # 122
+which(colnames(asthma_ndc_df)=="geo_wt_pm_lag1_zip") # code to find column numbers, 115
+pm_estimates_df <- asthma_ndc_df[, c(108, 122, 115, 93)]/10  # create 10 unit increases
+
+# new loop for age categories
+
+# empty matrix (12 x 10 matrix)
+point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)
+
+colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+# fill in the outcome namedataframe before method loop
+point_estimates[, 1] <- outcome_name
+
+# dataframe for analysis creation
+# bind columns back together 
+df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+  # remove missing pm values
+  filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
+  filter(!is.na(krig_pm_lag1_zip)) %>% 
+  filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
+  # the following code makes sure that the counterfactual values retained are 
+  # symetric in that number of obs before = number of obs after
+  mutate(obs_diff_admission = (fromdate - date)/7) 
+  # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+# second loop to run a model for each pm estimation method
+for(j in 38:40){
+
+  # variable to model 
+  var_name <- colnames(df_analysis[j])
+
+  # set row number to fill
+  row_n <- j-37
+
+  # only run the model if the dataframe has observations
+  if(nrow(df_analysis) != 0){
+  # conditional logistic regression model
+  mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+  # populate matrix
+  point_estimates[row_n, 2] <- method_list[row_n]
+  point_estimates[row_n, 3] <- mod$n
+  point_estimates[row_n, 4] <- mod$nevent
+  # odds ratio
+  point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+  # 95% lower bound
+  point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
+  # 95% upper bound
+  point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                        1.96*(summary(mod)$coefficient[1,3])), 3)
+  # standard error
+  point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
+  # p val
+  point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)
+
+  # create else statement that fills matrix with missing so I still have the row
+  # in the final dataframe
+  } else {point_estimates[row_n, 2] <- method_list[row_n]
+          point_estimates[row_n, 3] <- 0
+          point_estimates[row_n, 4] <- 0
+          point_estimates[row_n, c(5:9)] <- 99 } # end 'if else' statement
+  } # end methods loop
+
+# save point estimates as a dataframe
+combined_point_est_df <- as_data_frame(point_estimates)
+
+wrf_geo_age <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" |
+    #      pm_method == "WRF-Chem Smoke") %>% 
+  # subset columns I want to put in to the table
+  select(2, 4:7) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
+         caption = "Association between a 10 ug/m^3 in PM2.5 three smoke method and pharmacy outcomes",
+         # row group by outcome
+         rgroup = "Asthma Inhalers Lag 1",
+         n.rgroup = c(rep(3, 1)), # 6 rows for each age cat for each outcome
+         # column headers
+         header = c("Est Method", "Events",
+                    "OR&dagger;", "Lower", "Upper"),
+         # column spanner
+         cgroup = c("", "95% CI"), 
+         n.cgroup = c(3, 2),
+         padding.rgroup = "&nbsp;&nbsp;",
+         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+         align = "llcccc", # column alignment,
+         tfoot="&dagger; Referent periods matched to events on same day of week within May to September fire season."
+         ) # end table
+
+print(tab)
+
+combined_point_est_df_new <- combined_point_est_df %>%
+  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
+                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
+  select(-n_events) %>%
+  rename(n_events = n_events_new)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_age_plot <- combined_point_est_df_new %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  # do not plot results with less than 15 events, create missing vals
+  # in mutate
+  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
+                             99, odds_ratio),
+         lower95 = ifelse(as.numeric(n_events) < 15, 
+                             99, lower95),
+         upper95 = ifelse(as.numeric(n_events) < 15, 
+                             99, upper95)) %>% 
+  # subset columns I want to put in to the table
+  select(1, 2, 9, 4:6) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
+                              levels = unique(wrf_geo_age_plot$outcome))
+
+wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
+                                levels = unique(wrf_geo_age_plot$pm_method))
+
+wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
+wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
+wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
+wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)
+
+## ggplot 
+  print_plot <- ggplot(wrf_geo_age_plot,
+    aes(x = pm_method, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red",  "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Smoke Estimation Method') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10),
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
+  ggsave("../plot_lag/ndc_overall_lag1_three.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+```{r time stratified age strata lag 4, warning =F, echo = F, results='asis'} 
+# dataframe list
+
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# dataframe to loop through
+asthma_ndc_df <- data.frame(asthma_ndc)
+
+# outcome name
+outcome <- "asthma inhalers Lag 1"
+outcome_name <- "Asthma inhalers Lag 1"
+
+# age category list
+age_cat_list <- c(0,1,2)
+
+# create an empty list to row bind dataframes together
+datalist1 <- list()
+
+# data wrangling ----
+# Producing conditional logit model estimates loop
+
+# extract covariates from dataframe
+covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]
+
+# extract pm values and divide by 10 and ordered
+#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+pm_estimates_df <- asthma_ndc_df[, c(108, 122, 115, 93)]/10  # create 10 unit increases
+
+# new loop for age categories
+for(k in 0:2){
+  # empty matrix (12 x 10 matrix)
+  point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)
+
+  colnames(point_estimates) <- c('outcome', 'age_cat', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+  # fill in the outcome namedataframe before method loop
+  point_estimates[, 1] <- outcome_name
+  # repeat the age category 4 times for each pm method
+  point_estimates[, 2] <- k
+
+  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+    # remove missing pm values
+    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
+    filter(!is.na(krig_pm_lag1_zip)) %>% 
+    filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
+    # limit to specific age category
+    filter(age_ind == k) %>% 
+    # the following code makes sure that the counterfactual values retained are 
+    # symetric in that number of obs before = number of obs after
+    mutate(obs_diff_admission = (fromdate - date)/7) 
+    # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+  # second loop to run a model for each pm estimation method
+  for(j in 38:40){
+
+    # variable to model 
+    var_name <- colnames(df_analysis[j])
+
+    # set row number to fill
+    row_n <- j-37
+
+    # only run the model if the dataframe has observations
+    if(nrow(df_analysis) != 0){
+    # conditional logistic regression model
+    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+    # populate matrix
+    point_estimates[row_n, 3] <- method_list[row_n]
+    point_estimates[row_n, 4] <- mod$n
+    point_estimates[row_n, 5] <- mod$nevent
+    # odds ratio
+    point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+    # 95% lower bound
+    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
+    # 95% upper bound
+    point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                        1.96*(summary(mod)$coefficient[1,3])), 3)
+    # standard error
+    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
+    # p val
+    point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
+
+    # create else statement that fills matrix with missing so I still have the row
+    # in the final dataframe
+    } else {point_estimates[row_n, 3] <- method_list[row_n]
+            point_estimates[row_n, 4] <- 0
+            point_estimates[row_n, 5] <- 0
+            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
+    } # end methods loop
+
+  # save point estimates as a dataframe
+  point_est_df <- as_data_frame(point_estimates)
+  datalist1[[k+1]] <- point_est_df
+} # end age category loop
+
+# bind rows of age category estimates together
+combined_point_est_df <- bind_rows(datalist1)
+
+wrf_geo_age <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" |
+    #      pm_method == "WRF-Chem Smoke") %>% 
+  mutate(age_cat2 = ifelse(age_cat == 0, "Less than 15", 
+                    ifelse(age_cat == 1, "15-65",
+                    ifelse(age_cat == 2, "Greater than 65", NA)))) %>% 
+  # subset columns I want to put in to the table
+  select(3, 11, 5:8) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
+         caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by age",
+         # row group by outcome
+         rgroup = "Asthma inhalers Lag 1",
+         n.rgroup = c(rep(9, 1)), # 6 rows for each age cat for each outcome
+         # column headers
+         header = c("Est Method", "Age Group", "Events",
+                    "OR&dagger;", "Lower", "Upper"),
+         # column spanner
+         cgroup = c("", "95% CI"), 
+         n.cgroup = c(4, 2),
+         padding.rgroup = "&nbsp;&nbsp;",
+         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+         align = "llcccc", # column alignment,
+         tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within May to September fire season."
+         ) # end table
+
+print(tab)
+
+combined_point_est_df_new <- combined_point_est_df %>%
+  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
+                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
+  select(-n_events) %>%
+  rename(n_events = n_events_new)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_age_plot <- combined_point_est_df_new %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  # do not plot results with less than 15 events, create missing vals
+  # in mutate
+  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
+                             99, odds_ratio),
+         lower95 = ifelse(as.numeric(n_events) < 15, 
+                             99, lower95),
+         upper95 = ifelse(as.numeric(n_events) < 15, 
+                             99, upper95),
+         age_cat2 = ifelse(age_cat == 0, "<15", 
+                    ifelse(age_cat == 1, "15-65",
+                    ifelse(age_cat == 2, ">65", NA)))) %>% 
+  # subset columns I want to put in to the table
+  select(1, 3, 11, 10, 5:7) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
+                              levels = unique(wrf_geo_age_plot$outcome))
+
+wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
+                                levels = unique(wrf_geo_age_plot$pm_method))
+
+wrf_geo_age_plot$age_cat2 <- factor(wrf_geo_age_plot$age_cat2,
+                                levels = unique(wrf_geo_age_plot$age_cat2))
+
+wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
+wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
+wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
+wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)
+
+## ggplot 
+  print_plot <- ggplot(wrf_geo_age_plot,
+    aes(x = age_cat2, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red",  "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Age Category ') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10), 
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
+  ggsave("../plot_lag/ndc_age_fig_three_lag1.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+```{r sex strata cdc met  lag 4, warning = F, echo = F, results='asis'} 
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# dataframe to loop through
+asthma_ndc_df <- data.frame(asthma_ndc)
+
+# outcome name
+outcome <- "asthma inhalers Lag 1"
+outcome_name <- "Asthma inhalers Lag 1"
+
+# create an empty list to row bind dataframes together
+datalist1 <- list()
+
+# sex category list
+sex_strata_list <- c(0,1)
+
+# data wrangling ----
+# Producing conditional logit model estimates loop
+
+# extract covariates from dataframe
+covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]
+
+# extract pm values and divide by 10 and ordered
+#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+pm_estimates_df <- asthma_ndc_df[, c(108, 122, 115, 93)]/10  # create 10 unit increases
+
+# new loop for sex categories
+for(k in 0:1){
+
+  # empty matrix (12 x 10 matrix)
+  point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)
+
+  colnames(point_estimates) <- c('outcome', 'sex', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+  # fill in the outcome namedataframe before method loop
+  point_estimates[, 1] <- outcome_name
+  # repeat the sex category 4 times for each pm method
+  point_estimates[, 2] <- k
+
+  # dataframe for analysis creation
+  # bind columns back together 
+   df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+    # remove missing pm values
+    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
+    filter(!is.na(krig_pm_lag1_zip)) %>% 
+    filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
+    # limit to specific age category
+    filter(sex_ind == k) %>% 
+    # the following code makes sure that the counterfactual values retained are 
+    # symetric in that number of obs before = number of obs after
+    mutate(obs_diff_admission = (fromdate - date)/7) 
+    # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+  # second loop to run a model for each pm estimation method
+  for(j in 38:40){
+
+    # variable to model 
+    var_name <- colnames(df_analysis[j])
+
+    # set row number to fill
+    row_n <- j-37
+
+    # only run the model if the dataframe has observations
+    if(nrow(df_analysis) != 0){
+    # conditional logistic regression model
+    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+    # populate matrix
+    point_estimates[row_n, 3] <- method_list[row_n]
+    point_estimates[row_n, 4] <- mod$n
+    point_estimates[row_n, 5] <- mod$nevent
+    # odds ratio
+    point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+    # 95% lower bound
+    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
+                                      1.96*(summary(mod)$coefficient[1,3])), 3)
+    # 95% upper bound
+    point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                      1.96*(summary(mod)$coefficient[1,3])), 3)
+    # standard error
+    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
+    # p val
+    point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
+
+    # create else statement that fills matrix with missing so I still have the row
+    # in the final dataframe
+    } else {point_estimates[row_n, 3] <- method_list[row_n]
+            point_estimates[row_n, 4] <- 0
+            point_estimates[row_n, 5] <- 0
+            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
+    } # end methods loop
+
+  # save point estimates as a dataframe
+  point_est_df <- as_data_frame(point_estimates)
+
+  # combine previous values in dataframe that has all outcome/methods comparisons
+  datalist1[[k+1]] <- point_est_df
+} # end sex category loop
+
+# bind rows of sex category estimates together
+combined_point_est_df <- bind_rows(datalist1)
+
+wrf_geo_sex <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  mutate(sex_cat = ifelse(sex == 0, "Male", 
+               ifelse(sex == 1, "Female", NA))) %>% 
+  # subset columns I want to put in to the table
+  select(3, 11, 5:8)%>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+  tab <- htmlTable(txtRound(wrf_geo_sex, digits = 3, 1:3), 
+           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and pharmacy outcomes stratified by sex",
+           # row group by outcome
+           rgroup = "Asthma inhalers Lag 1",
+           n.rgroup = c(rep(6, 1)), # 2 rows for each sex strata for each outcome
+           # column headers
+           header = c("Est Method", "Sex Strata", "Events",
+                      "OR&dagger;", "Lower", "Upper"),
+           # column spanner
+           cgroup = c("", "95% CI"), 
+           n.cgroup = c(4, 2),
+           padding.rgroup = "&nbsp;&nbsp;",
+           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+           align = "llcccc", # column alignment,
+           tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within July to October fire season."
+            ) # end table
+
+  print(tab)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_sex_plot <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | 
+   #       pm_method == "WRF-Chem Smoke") %>% 
+  mutate(sex_cat = ifelse(sex == 0, "Male", 
+               ifelse(sex == 1, "Female", NA))) %>%
+  # subset columns I want to put in to the table
+  select(1, 3, 11, 5:8) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_sex_plot$outcome <- factor(wrf_geo_sex_plot$outcome,
+                              levels = unique(wrf_geo_sex_plot$outcome))
+
+wrf_geo_sex_plot$pm_method <- factor(wrf_geo_sex_plot$pm_method,
+                                levels = unique(wrf_geo_sex_plot$pm_method))
+
+wrf_geo_sex_plot$sex_cat <- factor(wrf_geo_sex_plot$sex_cat,
+                                levels = unique(wrf_geo_sex_plot$sex_cat))
+
+wrf_geo_sex_plot$n_events <- as.numeric(wrf_geo_sex_plot$n_events)
+wrf_geo_sex_plot$odds_ratio <- as.numeric(wrf_geo_sex_plot$odds_ratio)
+wrf_geo_sex_plot$lower95 <- as.numeric(wrf_geo_sex_plot$lower95)
+wrf_geo_sex_plot$upper95 <- as.numeric(wrf_geo_sex_plot$upper95)
+
+## ggplot
+  print_plot <- ggplot(wrf_geo_sex_plot,
+    aes(x = sex_cat, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red", "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Sex') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10),
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
+  ggsave("../plot_lag/ndc_sex_fig_three_lag1.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+### Lag 5
+
+```{r overall lag 5, warning =F, echo = F, results='asis'} 
+# dataframe list
+
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# dataframe to loop through
+# asthma_ndc_df <- data.frame(asthma_ndc)
+
+# outcome name
+outcome <- "Asthma Inhalers Lag 1"
+outcome_name <- "Asthma Inhalers Lag 1"
+
+# data wrangling ----
+# Producing conditional logit model estimates loop
+
+# extract covariates from dataframe
+covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]
+
+# extract pm values and divide by 10 and ordered
+which(colnames(asthma_ndc_df)=="wrf_smk_pm_lag1_zip") # 108
+which(colnames(asthma_ndc_df)=="krig_pm_lag1_zip") # 122
+which(colnames(asthma_ndc_df)=="geo_wt_pm_lag1_zip") # code to find column numbers, 115
+pm_estimates_df <- asthma_ndc_df[, c(108, 122, 115, 93)]/10  # create 10 unit increases
+
+# new loop for age categories
+
+# empty matrix (12 x 10 matrix)
+point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)
+
+colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+# fill in the outcome namedataframe before method loop
+point_estimates[, 1] <- outcome_name
+
+# dataframe for analysis creation
+# bind columns back together 
+df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+  # remove missing pm values
+  filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
+  filter(!is.na(krig_pm_lag1_zip)) %>% 
+  filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
+  # the following code makes sure that the counterfactual values retained are 
+  # symetric in that number of obs before = number of obs after
+  mutate(obs_diff_admission = (fromdate - date)/7) 
+  # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+# second loop to run a model for each pm estimation method
+for(j in 38:40){
+
+  # variable to model 
+  var_name <- colnames(df_analysis[j])
+
+  # set row number to fill
+  row_n <- j-37
+
+  # only run the model if the dataframe has observations
+  if(nrow(df_analysis) != 0){
+  # conditional logistic regression model
+  mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+  # populate matrix
+  point_estimates[row_n, 2] <- method_list[row_n]
+  point_estimates[row_n, 3] <- mod$n
+  point_estimates[row_n, 4] <- mod$nevent
+  # odds ratio
+  point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+  # 95% lower bound
+  point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
+  # 95% upper bound
+  point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                        1.96*(summary(mod)$coefficient[1,3])), 3)
+  # standard error
+  point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
+  # p val
+  point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)
+
+  # create else statement that fills matrix with missing so I still have the row
+  # in the final dataframe
+  } else {point_estimates[row_n, 2] <- method_list[row_n]
+          point_estimates[row_n, 3] <- 0
+          point_estimates[row_n, 4] <- 0
+          point_estimates[row_n, c(5:9)] <- 99 } # end 'if else' statement
+  } # end methods loop
+
+# save point estimates as a dataframe
+combined_point_est_df <- as_data_frame(point_estimates)
+
+wrf_geo_age <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" |
+    #      pm_method == "WRF-Chem Smoke") %>% 
+  # subset columns I want to put in to the table
+  select(2, 4:7) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
+         caption = "Association between a 10 ug/m^3 in PM2.5 three smoke method and pharmacy outcomes",
+         # row group by outcome
+         rgroup = "Asthma Inhalers Lag 1",
+         n.rgroup = c(rep(3, 1)), # 6 rows for each age cat for each outcome
+         # column headers
+         header = c("Est Method", "Events",
+                    "OR&dagger;", "Lower", "Upper"),
+         # column spanner
+         cgroup = c("", "95% CI"), 
+         n.cgroup = c(3, 2),
+         padding.rgroup = "&nbsp;&nbsp;",
+         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+         align = "llcccc", # column alignment,
+         tfoot="&dagger; Referent periods matched to events on same day of week within May to September fire season."
+         ) # end table
+
+print(tab)
+
+combined_point_est_df_new <- combined_point_est_df %>%
+  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
+                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
+  select(-n_events) %>%
+  rename(n_events = n_events_new)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_age_plot <- combined_point_est_df_new %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  # do not plot results with less than 15 events, create missing vals
+  # in mutate
+  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
+                             99, odds_ratio),
+         lower95 = ifelse(as.numeric(n_events) < 15, 
+                             99, lower95),
+         upper95 = ifelse(as.numeric(n_events) < 15, 
+                             99, upper95)) %>% 
+  # subset columns I want to put in to the table
+  select(1, 2, 9, 4:6) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
+                              levels = unique(wrf_geo_age_plot$outcome))
+
+wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
+                                levels = unique(wrf_geo_age_plot$pm_method))
+
+wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
+wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
+wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
+wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)
+
+## ggplot 
+  print_plot <- ggplot(wrf_geo_age_plot,
+    aes(x = pm_method, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red",  "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Smoke Estimation Method') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10),
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
+  ggsave("../plot_lag/ndc_overall_lag1_three.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+```{r time stratified age strata lag 5, warning =F, echo = F, results='asis'} 
+# dataframe list
+
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# dataframe to loop through
+asthma_ndc_df <- data.frame(asthma_ndc)
+
+# outcome name
+outcome <- "asthma inhalers Lag 1"
+outcome_name <- "Asthma inhalers Lag 1"
+
+# age category list
+age_cat_list <- c(0,1,2)
+
+# create an empty list to row bind dataframes together
+datalist1 <- list()
+
+# data wrangling ----
+# Producing conditional logit model estimates loop
+
+# extract covariates from dataframe
+covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]
+
+# extract pm values and divide by 10 and ordered
+#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+pm_estimates_df <- asthma_ndc_df[, c(108, 122, 115, 93)]/10  # create 10 unit increases
+
+# new loop for age categories
+for(k in 0:2){
+  # empty matrix (12 x 10 matrix)
+  point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)
+
+  colnames(point_estimates) <- c('outcome', 'age_cat', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+  # fill in the outcome namedataframe before method loop
+  point_estimates[, 1] <- outcome_name
+  # repeat the age category 4 times for each pm method
+  point_estimates[, 2] <- k
+
+  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+    # remove missing pm values
+    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
+    filter(!is.na(krig_pm_lag1_zip)) %>% 
+    filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
+    # limit to specific age category
+    filter(age_ind == k) %>% 
+    # the following code makes sure that the counterfactual values retained are 
+    # symetric in that number of obs before = number of obs after
+    mutate(obs_diff_admission = (fromdate - date)/7) 
+    # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+  # second loop to run a model for each pm estimation method
+  for(j in 38:40){
+
+    # variable to model 
+    var_name <- colnames(df_analysis[j])
+
+    # set row number to fill
+    row_n <- j-37
+
+    # only run the model if the dataframe has observations
+    if(nrow(df_analysis) != 0){
+    # conditional logistic regression model
+    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+    # populate matrix
+    point_estimates[row_n, 3] <- method_list[row_n]
+    point_estimates[row_n, 4] <- mod$n
+    point_estimates[row_n, 5] <- mod$nevent
+    # odds ratio
+    point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+    # 95% lower bound
+    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
+    # 95% upper bound
+    point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                        1.96*(summary(mod)$coefficient[1,3])), 3)
+    # standard error
+    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
+    # p val
+    point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
+
+    # create else statement that fills matrix with missing so I still have the row
+    # in the final dataframe
+    } else {point_estimates[row_n, 3] <- method_list[row_n]
+            point_estimates[row_n, 4] <- 0
+            point_estimates[row_n, 5] <- 0
+            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
+    } # end methods loop
+
+  # save point estimates as a dataframe
+  point_est_df <- as_data_frame(point_estimates)
+  datalist1[[k+1]] <- point_est_df
+} # end age category loop
+
+# bind rows of age category estimates together
+combined_point_est_df <- bind_rows(datalist1)
+
+wrf_geo_age <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" |
+    #      pm_method == "WRF-Chem Smoke") %>% 
+  mutate(age_cat2 = ifelse(age_cat == 0, "Less than 15", 
+                    ifelse(age_cat == 1, "15-65",
+                    ifelse(age_cat == 2, "Greater than 65", NA)))) %>% 
+  # subset columns I want to put in to the table
+  select(3, 11, 5:8) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
+         caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by age",
+         # row group by outcome
+         rgroup = "Asthma inhalers Lag 1",
+         n.rgroup = c(rep(9, 1)), # 6 rows for each age cat for each outcome
+         # column headers
+         header = c("Est Method", "Age Group", "Events",
+                    "OR&dagger;", "Lower", "Upper"),
+         # column spanner
+         cgroup = c("", "95% CI"), 
+         n.cgroup = c(4, 2),
+         padding.rgroup = "&nbsp;&nbsp;",
+         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+         align = "llcccc", # column alignment,
+         tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within May to September fire season."
+         ) # end table
+
+print(tab)
+
+combined_point_est_df_new <- combined_point_est_df %>%
+  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
+                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
+  select(-n_events) %>%
+  rename(n_events = n_events_new)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_age_plot <- combined_point_est_df_new %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  # do not plot results with less than 15 events, create missing vals
+  # in mutate
+  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
+                             99, odds_ratio),
+         lower95 = ifelse(as.numeric(n_events) < 15, 
+                             99, lower95),
+         upper95 = ifelse(as.numeric(n_events) < 15, 
+                             99, upper95),
+         age_cat2 = ifelse(age_cat == 0, "<15", 
+                    ifelse(age_cat == 1, "15-65",
+                    ifelse(age_cat == 2, ">65", NA)))) %>% 
+  # subset columns I want to put in to the table
+  select(1, 3, 11, 10, 5:7) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
+                              levels = unique(wrf_geo_age_plot$outcome))
+
+wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
+                                levels = unique(wrf_geo_age_plot$pm_method))
+
+wrf_geo_age_plot$age_cat2 <- factor(wrf_geo_age_plot$age_cat2,
+                                levels = unique(wrf_geo_age_plot$age_cat2))
+
+wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
+wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
+wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
+wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)
+
+## ggplot 
+  print_plot <- ggplot(wrf_geo_age_plot,
+    aes(x = age_cat2, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red",  "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Age Category ') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10), 
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
+  ggsave("../plot_lag/ndc_age_fig_three_lag1.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+```{r sex strata cdc met  lag 5, warning = F, echo = F, results='asis'} 
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# dataframe to loop through
+asthma_ndc_df <- data.frame(asthma_ndc)
+
+# outcome name
+outcome <- "asthma inhalers Lag 1"
+outcome_name <- "Asthma inhalers Lag 1"
+
+# create an empty list to row bind dataframes together
+datalist1 <- list()
+
+# sex category list
+sex_strata_list <- c(0,1)
+
+# data wrangling ----
+# Producing conditional logit model estimates loop
+
+# extract covariates from dataframe
+covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]
+
+# extract pm values and divide by 10 and ordered
+#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+pm_estimates_df <- asthma_ndc_df[, c(108, 122, 115, 93)]/10  # create 10 unit increases
+
+# new loop for sex categories
+for(k in 0:1){
+
+  # empty matrix (12 x 10 matrix)
+  point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)
+
+  colnames(point_estimates) <- c('outcome', 'sex', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+  # fill in the outcome namedataframe before method loop
+  point_estimates[, 1] <- outcome_name
+  # repeat the sex category 4 times for each pm method
+  point_estimates[, 2] <- k
+
+  # dataframe for analysis creation
+  # bind columns back together 
+   df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+    # remove missing pm values
+    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
+    filter(!is.na(krig_pm_lag1_zip)) %>% 
+    filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
+    # limit to specific age category
+    filter(sex_ind == k) %>% 
+    # the following code makes sure that the counterfactual values retained are 
+    # symetric in that number of obs before = number of obs after
+    mutate(obs_diff_admission = (fromdate - date)/7) 
+    # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+  # second loop to run a model for each pm estimation method
+  for(j in 38:40){
+
+    # variable to model 
+    var_name <- colnames(df_analysis[j])
+
+    # set row number to fill
+    row_n <- j-37
+
+    # only run the model if the dataframe has observations
+    if(nrow(df_analysis) != 0){
+    # conditional logistic regression model
+    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+    # populate matrix
+    point_estimates[row_n, 3] <- method_list[row_n]
+    point_estimates[row_n, 4] <- mod$n
+    point_estimates[row_n, 5] <- mod$nevent
+    # odds ratio
+    point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+    # 95% lower bound
+    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
+                                      1.96*(summary(mod)$coefficient[1,3])), 3)
+    # 95% upper bound
+    point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                      1.96*(summary(mod)$coefficient[1,3])), 3)
+    # standard error
+    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
+    # p val
+    point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
+
+    # create else statement that fills matrix with missing so I still have the row
+    # in the final dataframe
+    } else {point_estimates[row_n, 3] <- method_list[row_n]
+            point_estimates[row_n, 4] <- 0
+            point_estimates[row_n, 5] <- 0
+            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
+    } # end methods loop
+
+  # save point estimates as a dataframe
+  point_est_df <- as_data_frame(point_estimates)
+
+  # combine previous values in dataframe that has all outcome/methods comparisons
+  datalist1[[k+1]] <- point_est_df
+} # end sex category loop
+
+# bind rows of sex category estimates together
+combined_point_est_df <- bind_rows(datalist1)
+
+wrf_geo_sex <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  mutate(sex_cat = ifelse(sex == 0, "Male", 
+               ifelse(sex == 1, "Female", NA))) %>% 
+  # subset columns I want to put in to the table
+  select(3, 11, 5:8)%>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+  tab <- htmlTable(txtRound(wrf_geo_sex, digits = 3, 1:3), 
+           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and pharmacy outcomes stratified by sex",
+           # row group by outcome
+           rgroup = "Asthma inhalers Lag 1",
+           n.rgroup = c(rep(6, 1)), # 2 rows for each sex strata for each outcome
+           # column headers
+           header = c("Est Method", "Sex Strata", "Events",
+                      "OR&dagger;", "Lower", "Upper"),
+           # column spanner
+           cgroup = c("", "95% CI"), 
+           n.cgroup = c(4, 2),
+           padding.rgroup = "&nbsp;&nbsp;",
+           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+           align = "llcccc", # column alignment,
+           tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within July to October fire season."
+            ) # end table
+
+  print(tab)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_sex_plot <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | 
+   #       pm_method == "WRF-Chem Smoke") %>% 
+  mutate(sex_cat = ifelse(sex == 0, "Male", 
+               ifelse(sex == 1, "Female", NA))) %>%
+  # subset columns I want to put in to the table
+  select(1, 3, 11, 5:8) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_sex_plot$outcome <- factor(wrf_geo_sex_plot$outcome,
+                              levels = unique(wrf_geo_sex_plot$outcome))
+
+wrf_geo_sex_plot$pm_method <- factor(wrf_geo_sex_plot$pm_method,
+                                levels = unique(wrf_geo_sex_plot$pm_method))
+
+wrf_geo_sex_plot$sex_cat <- factor(wrf_geo_sex_plot$sex_cat,
+                                levels = unique(wrf_geo_sex_plot$sex_cat))
+
+wrf_geo_sex_plot$n_events <- as.numeric(wrf_geo_sex_plot$n_events)
+wrf_geo_sex_plot$odds_ratio <- as.numeric(wrf_geo_sex_plot$odds_ratio)
+wrf_geo_sex_plot$lower95 <- as.numeric(wrf_geo_sex_plot$lower95)
+wrf_geo_sex_plot$upper95 <- as.numeric(wrf_geo_sex_plot$upper95)
+
+## ggplot
+  print_plot <- ggplot(wrf_geo_sex_plot,
+    aes(x = sex_cat, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red", "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Sex') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10),
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
+  ggsave("../plot_lag/ndc_sex_fig_three_lag1.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+### Lag 6
+
+```{r overall lag 6, warning =F, echo = F, results='asis'} 
+# dataframe list
+
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# dataframe to loop through
+# asthma_ndc_df <- data.frame(asthma_ndc)
+
+# outcome name
+outcome <- "Asthma Inhalers Lag 1"
+outcome_name <- "Asthma Inhalers Lag 1"
+
+# data wrangling ----
+# Producing conditional logit model estimates loop
+
+# extract covariates from dataframe
+covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]
+
+# extract pm values and divide by 10 and ordered
+which(colnames(asthma_ndc_df)=="wrf_smk_pm_lag1_zip") # 108
+which(colnames(asthma_ndc_df)=="krig_pm_lag1_zip") # 122
+which(colnames(asthma_ndc_df)=="geo_wt_pm_lag1_zip") # code to find column numbers, 115
+pm_estimates_df <- asthma_ndc_df[, c(108, 122, 115, 93)]/10  # create 10 unit increases
+
+# new loop for age categories
+
+# empty matrix (12 x 10 matrix)
+point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)
+
+colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+# fill in the outcome namedataframe before method loop
+point_estimates[, 1] <- outcome_name
+
+# dataframe for analysis creation
+# bind columns back together 
+df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+  # remove missing pm values
+  filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
+  filter(!is.na(krig_pm_lag1_zip)) %>% 
+  filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
+  # the following code makes sure that the counterfactual values retained are 
+  # symetric in that number of obs before = number of obs after
+  mutate(obs_diff_admission = (fromdate - date)/7) 
+  # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+# second loop to run a model for each pm estimation method
+for(j in 38:40){
+
+  # variable to model 
+  var_name <- colnames(df_analysis[j])
+
+  # set row number to fill
+  row_n <- j-37
+
+  # only run the model if the dataframe has observations
+  if(nrow(df_analysis) != 0){
+  # conditional logistic regression model
+  mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+  # populate matrix
+  point_estimates[row_n, 2] <- method_list[row_n]
+  point_estimates[row_n, 3] <- mod$n
+  point_estimates[row_n, 4] <- mod$nevent
+  # odds ratio
+  point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+  # 95% lower bound
+  point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
+  # 95% upper bound
+  point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                        1.96*(summary(mod)$coefficient[1,3])), 3)
+  # standard error
+  point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
+  # p val
+  point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)
+
+  # create else statement that fills matrix with missing so I still have the row
+  # in the final dataframe
+  } else {point_estimates[row_n, 2] <- method_list[row_n]
+          point_estimates[row_n, 3] <- 0
+          point_estimates[row_n, 4] <- 0
+          point_estimates[row_n, c(5:9)] <- 99 } # end 'if else' statement
+  } # end methods loop
+
+# save point estimates as a dataframe
+combined_point_est_df <- as_data_frame(point_estimates)
+
+wrf_geo_age <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" |
+    #      pm_method == "WRF-Chem Smoke") %>% 
+  # subset columns I want to put in to the table
+  select(2, 4:7) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
+         caption = "Association between a 10 ug/m^3 in PM2.5 three smoke method and pharmacy outcomes",
+         # row group by outcome
+         rgroup = "Asthma Inhalers Lag 1",
+         n.rgroup = c(rep(3, 1)), # 6 rows for each age cat for each outcome
+         # column headers
+         header = c("Est Method", "Events",
+                    "OR&dagger;", "Lower", "Upper"),
+         # column spanner
+         cgroup = c("", "95% CI"), 
+         n.cgroup = c(3, 2),
+         padding.rgroup = "&nbsp;&nbsp;",
+         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+         align = "llcccc", # column alignment,
+         tfoot="&dagger; Referent periods matched to events on same day of week within May to September fire season."
+         ) # end table
+
+print(tab)
+
+combined_point_est_df_new <- combined_point_est_df %>%
+  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
+                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
+  select(-n_events) %>%
+  rename(n_events = n_events_new)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_age_plot <- combined_point_est_df_new %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  # do not plot results with less than 15 events, create missing vals
+  # in mutate
+  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
+                             99, odds_ratio),
+         lower95 = ifelse(as.numeric(n_events) < 15, 
+                             99, lower95),
+         upper95 = ifelse(as.numeric(n_events) < 15, 
+                             99, upper95)) %>% 
+  # subset columns I want to put in to the table
+  select(1, 2, 9, 4:6) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
+                              levels = unique(wrf_geo_age_plot$outcome))
+
+wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
+                                levels = unique(wrf_geo_age_plot$pm_method))
+
+wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
+wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
+wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
+wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)
+
+## ggplot 
+  print_plot <- ggplot(wrf_geo_age_plot,
+    aes(x = pm_method, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red",  "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Smoke Estimation Method') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10),
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
+  ggsave("../plot_lag/ndc_overall_lag1_three.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+```{r time stratified age strata lag 6, warning =F, echo = F, results='asis'} 
+# dataframe list
+
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# dataframe to loop through
+asthma_ndc_df <- data.frame(asthma_ndc)
+
+# outcome name
+outcome <- "asthma inhalers Lag 1"
+outcome_name <- "Asthma inhalers Lag 1"
+
+# age category list
+age_cat_list <- c(0,1,2)
+
+# create an empty list to row bind dataframes together
+datalist1 <- list()
+
+# data wrangling ----
+# Producing conditional logit model estimates loop
+
+# extract covariates from dataframe
+covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]
+
+# extract pm values and divide by 10 and ordered
+#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+pm_estimates_df <- asthma_ndc_df[, c(108, 122, 115, 93)]/10  # create 10 unit increases
+
+# new loop for age categories
+for(k in 0:2){
+  # empty matrix (12 x 10 matrix)
+  point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)
+
+  colnames(point_estimates) <- c('outcome', 'age_cat', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+  # fill in the outcome namedataframe before method loop
+  point_estimates[, 1] <- outcome_name
+  # repeat the age category 4 times for each pm method
+  point_estimates[, 2] <- k
+
+  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+    # remove missing pm values
+    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
+    filter(!is.na(krig_pm_lag1_zip)) %>% 
+    filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
+    # limit to specific age category
+    filter(age_ind == k) %>% 
+    # the following code makes sure that the counterfactual values retained are 
+    # symetric in that number of obs before = number of obs after
+    mutate(obs_diff_admission = (fromdate - date)/7) 
+    # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+  # second loop to run a model for each pm estimation method
+  for(j in 38:40){
+
+    # variable to model 
+    var_name <- colnames(df_analysis[j])
+
+    # set row number to fill
+    row_n <- j-37
+
+    # only run the model if the dataframe has observations
+    if(nrow(df_analysis) != 0){
+    # conditional logistic regression model
+    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+    # populate matrix
+    point_estimates[row_n, 3] <- method_list[row_n]
+    point_estimates[row_n, 4] <- mod$n
+    point_estimates[row_n, 5] <- mod$nevent
+    # odds ratio
+    point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+    # 95% lower bound
+    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
+    # 95% upper bound
+    point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                        1.96*(summary(mod)$coefficient[1,3])), 3)
+    # standard error
+    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
+    # p val
+    point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
+
+    # create else statement that fills matrix with missing so I still have the row
+    # in the final dataframe
+    } else {point_estimates[row_n, 3] <- method_list[row_n]
+            point_estimates[row_n, 4] <- 0
+            point_estimates[row_n, 5] <- 0
+            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
+    } # end methods loop
+
+  # save point estimates as a dataframe
+  point_est_df <- as_data_frame(point_estimates)
+  datalist1[[k+1]] <- point_est_df
+} # end age category loop
+
+# bind rows of age category estimates together
+combined_point_est_df <- bind_rows(datalist1)
+
+wrf_geo_age <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" |
+    #      pm_method == "WRF-Chem Smoke") %>% 
+  mutate(age_cat2 = ifelse(age_cat == 0, "Less than 15", 
+                    ifelse(age_cat == 1, "15-65",
+                    ifelse(age_cat == 2, "Greater than 65", NA)))) %>% 
+  # subset columns I want to put in to the table
+  select(3, 11, 5:8) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
+         caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by age",
+         # row group by outcome
+         rgroup = "Asthma inhalers Lag 1",
+         n.rgroup = c(rep(9, 1)), # 6 rows for each age cat for each outcome
+         # column headers
+         header = c("Est Method", "Age Group", "Events",
+                    "OR&dagger;", "Lower", "Upper"),
+         # column spanner
+         cgroup = c("", "95% CI"), 
+         n.cgroup = c(4, 2),
+         padding.rgroup = "&nbsp;&nbsp;",
+         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+         align = "llcccc", # column alignment,
+         tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within May to September fire season."
+         ) # end table
+
+print(tab)
+
+combined_point_est_df_new <- combined_point_est_df %>%
+  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
+                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
+  select(-n_events) %>%
+  rename(n_events = n_events_new)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_age_plot <- combined_point_est_df_new %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  # do not plot results with less than 15 events, create missing vals
+  # in mutate
+  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
+                             99, odds_ratio),
+         lower95 = ifelse(as.numeric(n_events) < 15, 
+                             99, lower95),
+         upper95 = ifelse(as.numeric(n_events) < 15, 
+                             99, upper95),
+         age_cat2 = ifelse(age_cat == 0, "<15", 
+                    ifelse(age_cat == 1, "15-65",
+                    ifelse(age_cat == 2, ">65", NA)))) %>% 
+  # subset columns I want to put in to the table
+  select(1, 3, 11, 10, 5:7) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
+                              levels = unique(wrf_geo_age_plot$outcome))
+
+wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
+                                levels = unique(wrf_geo_age_plot$pm_method))
+
+wrf_geo_age_plot$age_cat2 <- factor(wrf_geo_age_plot$age_cat2,
+                                levels = unique(wrf_geo_age_plot$age_cat2))
+
+wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
+wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
+wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
+wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)
+
+## ggplot 
+  print_plot <- ggplot(wrf_geo_age_plot,
+    aes(x = age_cat2, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red",  "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Age Category ') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10), 
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
+  ggsave("../plot_lag/ndc_age_fig_three_lag1.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+```{r sex strata cdc met  lag 6, warning = F, echo = F, results='asis'} 
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# dataframe to loop through
+asthma_ndc_df <- data.frame(asthma_ndc)
+
+# outcome name
+outcome <- "asthma inhalers Lag 1"
+outcome_name <- "Asthma inhalers Lag 1"
+
+# create an empty list to row bind dataframes together
+datalist1 <- list()
+
+# sex category list
+sex_strata_list <- c(0,1)
+
+# data wrangling ----
+# Producing conditional logit model estimates loop
+
+# extract covariates from dataframe
+covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]
+
+# extract pm values and divide by 10 and ordered
+#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+pm_estimates_df <- asthma_ndc_df[, c(108, 122, 115, 93)]/10  # create 10 unit increases
+
+# new loop for sex categories
+for(k in 0:1){
+
+  # empty matrix (12 x 10 matrix)
+  point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)
+
+  colnames(point_estimates) <- c('outcome', 'sex', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+  # fill in the outcome namedataframe before method loop
+  point_estimates[, 1] <- outcome_name
+  # repeat the sex category 4 times for each pm method
+  point_estimates[, 2] <- k
+
+  # dataframe for analysis creation
+  # bind columns back together 
+   df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+    # remove missing pm values
+    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
+    filter(!is.na(krig_pm_lag1_zip)) %>% 
+    filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
+    # limit to specific age category
+    filter(sex_ind == k) %>% 
+    # the following code makes sure that the counterfactual values retained are 
+    # symetric in that number of obs before = number of obs after
+    mutate(obs_diff_admission = (fromdate - date)/7) 
+    # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+  # second loop to run a model for each pm estimation method
+  for(j in 38:40){
+
+    # variable to model 
+    var_name <- colnames(df_analysis[j])
+
+    # set row number to fill
+    row_n <- j-37
+
+    # only run the model if the dataframe has observations
+    if(nrow(df_analysis) != 0){
+    # conditional logistic regression model
+    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+    # populate matrix
+    point_estimates[row_n, 3] <- method_list[row_n]
+    point_estimates[row_n, 4] <- mod$n
+    point_estimates[row_n, 5] <- mod$nevent
+    # odds ratio
+    point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+    # 95% lower bound
+    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
+                                      1.96*(summary(mod)$coefficient[1,3])), 3)
+    # 95% upper bound
+    point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                      1.96*(summary(mod)$coefficient[1,3])), 3)
+    # standard error
+    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
+    # p val
+    point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
+
+    # create else statement that fills matrix with missing so I still have the row
+    # in the final dataframe
+    } else {point_estimates[row_n, 3] <- method_list[row_n]
+            point_estimates[row_n, 4] <- 0
+            point_estimates[row_n, 5] <- 0
+            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
+    } # end methods loop
+
+  # save point estimates as a dataframe
+  point_est_df <- as_data_frame(point_estimates)
+
+  # combine previous values in dataframe that has all outcome/methods comparisons
+  datalist1[[k+1]] <- point_est_df
+} # end sex category loop
+
+# bind rows of sex category estimates together
+combined_point_est_df <- bind_rows(datalist1)
+
+wrf_geo_sex <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  mutate(sex_cat = ifelse(sex == 0, "Male", 
+               ifelse(sex == 1, "Female", NA))) %>% 
+  # subset columns I want to put in to the table
+  select(3, 11, 5:8)%>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+  tab <- htmlTable(txtRound(wrf_geo_sex, digits = 3, 1:3), 
+           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and pharmacy outcomes stratified by sex",
+           # row group by outcome
+           rgroup = "Asthma inhalers Lag 1",
+           n.rgroup = c(rep(6, 1)), # 2 rows for each sex strata for each outcome
+           # column headers
+           header = c("Est Method", "Sex Strata", "Events",
+                      "OR&dagger;", "Lower", "Upper"),
+           # column spanner
+           cgroup = c("", "95% CI"), 
+           n.cgroup = c(4, 2),
+           padding.rgroup = "&nbsp;&nbsp;",
+           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+           align = "llcccc", # column alignment,
+           tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within July to October fire season."
+            ) # end table
+
+  print(tab)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_sex_plot <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | 
+   #       pm_method == "WRF-Chem Smoke") %>% 
+  mutate(sex_cat = ifelse(sex == 0, "Male", 
+               ifelse(sex == 1, "Female", NA))) %>%
+  # subset columns I want to put in to the table
+  select(1, 3, 11, 5:8) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_sex_plot$outcome <- factor(wrf_geo_sex_plot$outcome,
+                              levels = unique(wrf_geo_sex_plot$outcome))
+
+wrf_geo_sex_plot$pm_method <- factor(wrf_geo_sex_plot$pm_method,
+                                levels = unique(wrf_geo_sex_plot$pm_method))
+
+wrf_geo_sex_plot$sex_cat <- factor(wrf_geo_sex_plot$sex_cat,
+                                levels = unique(wrf_geo_sex_plot$sex_cat))
+
+wrf_geo_sex_plot$n_events <- as.numeric(wrf_geo_sex_plot$n_events)
+wrf_geo_sex_plot$odds_ratio <- as.numeric(wrf_geo_sex_plot$odds_ratio)
+wrf_geo_sex_plot$lower95 <- as.numeric(wrf_geo_sex_plot$lower95)
+wrf_geo_sex_plot$upper95 <- as.numeric(wrf_geo_sex_plot$upper95)
+
+## ggplot
+  print_plot <- ggplot(wrf_geo_sex_plot,
+    aes(x = sex_cat, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red", "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Sex') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10),
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
+  ggsave("../plot_lag/ndc_sex_fig_three_lag1.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+### Lag 7
+
+```{r overall lag 7, warning =F, echo = F, results='asis'} 
+# dataframe list
+
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# dataframe to loop through
+# asthma_ndc_df <- data.frame(asthma_ndc)
+
+# outcome name
+outcome <- "Asthma Inhalers Lag 1"
+outcome_name <- "Asthma Inhalers Lag 1"
+
+# data wrangling ----
+# Producing conditional logit model estimates loop
+
+# extract covariates from dataframe
+covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]
+
+# extract pm values and divide by 10 and ordered
+which(colnames(asthma_ndc_df)=="wrf_smk_pm_lag1_zip") # 108
+which(colnames(asthma_ndc_df)=="krig_pm_lag1_zip") # 122
+which(colnames(asthma_ndc_df)=="geo_wt_pm_lag1_zip") # code to find column numbers, 115
+pm_estimates_df <- asthma_ndc_df[, c(108, 122, 115, 93)]/10  # create 10 unit increases
+
+# new loop for age categories
+
+# empty matrix (12 x 10 matrix)
+point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)
+
+colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+# fill in the outcome namedataframe before method loop
+point_estimates[, 1] <- outcome_name
+
+# dataframe for analysis creation
+# bind columns back together 
+df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+  # remove missing pm values
+  filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
+  filter(!is.na(krig_pm_lag1_zip)) %>% 
+  filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
+  # the following code makes sure that the counterfactual values retained are 
+  # symetric in that number of obs before = number of obs after
+  mutate(obs_diff_admission = (fromdate - date)/7) 
+  # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+# second loop to run a model for each pm estimation method
+for(j in 38:40){
+
+  # variable to model 
+  var_name <- colnames(df_analysis[j])
+
+  # set row number to fill
+  row_n <- j-37
+
+  # only run the model if the dataframe has observations
+  if(nrow(df_analysis) != 0){
+  # conditional logistic regression model
+  mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+  # populate matrix
+  point_estimates[row_n, 2] <- method_list[row_n]
+  point_estimates[row_n, 3] <- mod$n
+  point_estimates[row_n, 4] <- mod$nevent
+  # odds ratio
+  point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+  # 95% lower bound
+  point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
+  # 95% upper bound
+  point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                        1.96*(summary(mod)$coefficient[1,3])), 3)
+  # standard error
+  point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
+  # p val
+  point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)
+
+  # create else statement that fills matrix with missing so I still have the row
+  # in the final dataframe
+  } else {point_estimates[row_n, 2] <- method_list[row_n]
+          point_estimates[row_n, 3] <- 0
+          point_estimates[row_n, 4] <- 0
+          point_estimates[row_n, c(5:9)] <- 99 } # end 'if else' statement
+  } # end methods loop
+
+# save point estimates as a dataframe
+combined_point_est_df <- as_data_frame(point_estimates)
+
+wrf_geo_age <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" |
+    #      pm_method == "WRF-Chem Smoke") %>% 
+  # subset columns I want to put in to the table
+  select(2, 4:7) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
+         caption = "Association between a 10 ug/m^3 in PM2.5 three smoke method and pharmacy outcomes",
+         # row group by outcome
+         rgroup = "Asthma Inhalers Lag 1",
+         n.rgroup = c(rep(3, 1)), # 6 rows for each age cat for each outcome
+         # column headers
+         header = c("Est Method", "Events",
+                    "OR&dagger;", "Lower", "Upper"),
+         # column spanner
+         cgroup = c("", "95% CI"), 
+         n.cgroup = c(3, 2),
+         padding.rgroup = "&nbsp;&nbsp;",
+         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+         align = "llcccc", # column alignment,
+         tfoot="&dagger; Referent periods matched to events on same day of week within May to September fire season."
+         ) # end table
+
+print(tab)
+
+combined_point_est_df_new <- combined_point_est_df %>%
+  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
+                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
+  select(-n_events) %>%
+  rename(n_events = n_events_new)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_age_plot <- combined_point_est_df_new %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  # do not plot results with less than 15 events, create missing vals
+  # in mutate
+  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
+                             99, odds_ratio),
+         lower95 = ifelse(as.numeric(n_events) < 15, 
+                             99, lower95),
+         upper95 = ifelse(as.numeric(n_events) < 15, 
+                             99, upper95)) %>% 
+  # subset columns I want to put in to the table
+  select(1, 2, 9, 4:6) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
+                              levels = unique(wrf_geo_age_plot$outcome))
+
+wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
+                                levels = unique(wrf_geo_age_plot$pm_method))
+
+wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
+wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
+wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
+wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)
+
+## ggplot 
+  print_plot <- ggplot(wrf_geo_age_plot,
+    aes(x = pm_method, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red",  "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Smoke Estimation Method') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10),
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
+  ggsave("../plot_lag/ndc_overall_lag1_three.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+```{r time stratified age strata lag 7, warning =F, echo = F, results='asis'} 
+# dataframe list
+
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# dataframe to loop through
+asthma_ndc_df <- data.frame(asthma_ndc)
+
+# outcome name
+outcome <- "asthma inhalers Lag 1"
+outcome_name <- "Asthma inhalers Lag 1"
+
+# age category list
+age_cat_list <- c(0,1,2)
+
+# create an empty list to row bind dataframes together
+datalist1 <- list()
+
+# data wrangling ----
+# Producing conditional logit model estimates loop
+
+# extract covariates from dataframe
+covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]
+
+# extract pm values and divide by 10 and ordered
+#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+pm_estimates_df <- asthma_ndc_df[, c(108, 122, 115, 93)]/10  # create 10 unit increases
+
+# new loop for age categories
+for(k in 0:2){
+  # empty matrix (12 x 10 matrix)
+  point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)
+
+  colnames(point_estimates) <- c('outcome', 'age_cat', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+  # fill in the outcome namedataframe before method loop
+  point_estimates[, 1] <- outcome_name
+  # repeat the age category 4 times for each pm method
+  point_estimates[, 2] <- k
+
+  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+    # remove missing pm values
+    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
+    filter(!is.na(krig_pm_lag1_zip)) %>% 
+    filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
+    # limit to specific age category
+    filter(age_ind == k) %>% 
+    # the following code makes sure that the counterfactual values retained are 
+    # symetric in that number of obs before = number of obs after
+    mutate(obs_diff_admission = (fromdate - date)/7) 
+    # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+  # second loop to run a model for each pm estimation method
+  for(j in 38:40){
+
+    # variable to model 
+    var_name <- colnames(df_analysis[j])
+
+    # set row number to fill
+    row_n <- j-37
+
+    # only run the model if the dataframe has observations
+    if(nrow(df_analysis) != 0){
+    # conditional logistic regression model
+    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+    # populate matrix
+    point_estimates[row_n, 3] <- method_list[row_n]
+    point_estimates[row_n, 4] <- mod$n
+    point_estimates[row_n, 5] <- mod$nevent
+    # odds ratio
+    point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+    # 95% lower bound
+    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
+    # 95% upper bound
+    point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                        1.96*(summary(mod)$coefficient[1,3])), 3)
+    # standard error
+    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
+    # p val
+    point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
+
+    # create else statement that fills matrix with missing so I still have the row
+    # in the final dataframe
+    } else {point_estimates[row_n, 3] <- method_list[row_n]
+            point_estimates[row_n, 4] <- 0
+            point_estimates[row_n, 5] <- 0
+            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
+    } # end methods loop
+
+  # save point estimates as a dataframe
+  point_est_df <- as_data_frame(point_estimates)
+  datalist1[[k+1]] <- point_est_df
+} # end age category loop
+
+# bind rows of age category estimates together
+combined_point_est_df <- bind_rows(datalist1)
+
+wrf_geo_age <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" |
+    #      pm_method == "WRF-Chem Smoke") %>% 
+  mutate(age_cat2 = ifelse(age_cat == 0, "Less than 15", 
+                    ifelse(age_cat == 1, "15-65",
+                    ifelse(age_cat == 2, "Greater than 65", NA)))) %>% 
+  # subset columns I want to put in to the table
+  select(3, 11, 5:8) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
+         caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by age",
+         # row group by outcome
+         rgroup = "Asthma inhalers Lag 1",
+         n.rgroup = c(rep(9, 1)), # 6 rows for each age cat for each outcome
+         # column headers
+         header = c("Est Method", "Age Group", "Events",
+                    "OR&dagger;", "Lower", "Upper"),
+         # column spanner
+         cgroup = c("", "95% CI"), 
+         n.cgroup = c(4, 2),
+         padding.rgroup = "&nbsp;&nbsp;",
+         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+         align = "llcccc", # column alignment,
+         tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within May to September fire season."
+         ) # end table
+
+print(tab)
+
+combined_point_est_df_new <- combined_point_est_df %>%
+  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
+                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
+  select(-n_events) %>%
+  rename(n_events = n_events_new)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_age_plot <- combined_point_est_df_new %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  # do not plot results with less than 15 events, create missing vals
+  # in mutate
+  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
+                             99, odds_ratio),
+         lower95 = ifelse(as.numeric(n_events) < 15, 
+                             99, lower95),
+         upper95 = ifelse(as.numeric(n_events) < 15, 
+                             99, upper95),
+         age_cat2 = ifelse(age_cat == 0, "<15", 
+                    ifelse(age_cat == 1, "15-65",
+                    ifelse(age_cat == 2, ">65", NA)))) %>% 
+  # subset columns I want to put in to the table
+  select(1, 3, 11, 10, 5:7) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
+                              levels = unique(wrf_geo_age_plot$outcome))
+
+wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
+                                levels = unique(wrf_geo_age_plot$pm_method))
+
+wrf_geo_age_plot$age_cat2 <- factor(wrf_geo_age_plot$age_cat2,
+                                levels = unique(wrf_geo_age_plot$age_cat2))
+
+wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
+wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
+wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
+wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)
+
+## ggplot 
+  print_plot <- ggplot(wrf_geo_age_plot,
+    aes(x = age_cat2, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red",  "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Age Category ') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10), 
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
+  ggsave("../plot_lag/ndc_age_fig_three_lag1.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+```{r sex strata cdc met  lag 7, warning = F, echo = F, results='asis'} 
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# dataframe to loop through
+asthma_ndc_df <- data.frame(asthma_ndc)
+
+# outcome name
+outcome <- "asthma inhalers Lag 1"
+outcome_name <- "Asthma inhalers Lag 1"
+
+# create an empty list to row bind dataframes together
+datalist1 <- list()
+
+# sex category list
+sex_strata_list <- c(0,1)
+
+# data wrangling ----
+# Producing conditional logit model estimates loop
+
+# extract covariates from dataframe
+covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]
+
+# extract pm values and divide by 10 and ordered
+#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+pm_estimates_df <- asthma_ndc_df[, c(108, 122, 115, 93)]/10  # create 10 unit increases
+
+# new loop for sex categories
+for(k in 0:1){
+
+  # empty matrix (12 x 10 matrix)
+  point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)
+
+  colnames(point_estimates) <- c('outcome', 'sex', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+  # fill in the outcome namedataframe before method loop
+  point_estimates[, 1] <- outcome_name
+  # repeat the sex category 4 times for each pm method
+  point_estimates[, 2] <- k
+
+  # dataframe for analysis creation
+  # bind columns back together 
+   df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+    # remove missing pm values
+    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
+    filter(!is.na(krig_pm_lag1_zip)) %>% 
+    filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
+    # limit to specific age category
+    filter(sex_ind == k) %>% 
+    # the following code makes sure that the counterfactual values retained are 
+    # symetric in that number of obs before = number of obs after
+    mutate(obs_diff_admission = (fromdate - date)/7) 
+    # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+  # second loop to run a model for each pm estimation method
+  for(j in 38:40){
+
+    # variable to model 
+    var_name <- colnames(df_analysis[j])
+
+    # set row number to fill
+    row_n <- j-37
+
+    # only run the model if the dataframe has observations
+    if(nrow(df_analysis) != 0){
+    # conditional logistic regression model
+    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+    # populate matrix
+    point_estimates[row_n, 3] <- method_list[row_n]
+    point_estimates[row_n, 4] <- mod$n
+    point_estimates[row_n, 5] <- mod$nevent
+    # odds ratio
+    point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+    # 95% lower bound
+    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
+                                      1.96*(summary(mod)$coefficient[1,3])), 3)
+    # 95% upper bound
+    point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                      1.96*(summary(mod)$coefficient[1,3])), 3)
+    # standard error
+    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
+    # p val
+    point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
+
+    # create else statement that fills matrix with missing so I still have the row
+    # in the final dataframe
+    } else {point_estimates[row_n, 3] <- method_list[row_n]
+            point_estimates[row_n, 4] <- 0
+            point_estimates[row_n, 5] <- 0
+            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
+    } # end methods loop
+
+  # save point estimates as a dataframe
+  point_est_df <- as_data_frame(point_estimates)
+
+  # combine previous values in dataframe that has all outcome/methods comparisons
+  datalist1[[k+1]] <- point_est_df
+} # end sex category loop
+
+# bind rows of sex category estimates together
+combined_point_est_df <- bind_rows(datalist1)
+
+wrf_geo_sex <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  mutate(sex_cat = ifelse(sex == 0, "Male", 
+               ifelse(sex == 1, "Female", NA))) %>% 
+  # subset columns I want to put in to the table
+  select(3, 11, 5:8)%>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+  tab <- htmlTable(txtRound(wrf_geo_sex, digits = 3, 1:3), 
+           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and pharmacy outcomes stratified by sex",
+           # row group by outcome
+           rgroup = "Asthma inhalers Lag 1",
+           n.rgroup = c(rep(6, 1)), # 2 rows for each sex strata for each outcome
+           # column headers
+           header = c("Est Method", "Sex Strata", "Events",
+                      "OR&dagger;", "Lower", "Upper"),
+           # column spanner
+           cgroup = c("", "95% CI"), 
+           n.cgroup = c(4, 2),
+           padding.rgroup = "&nbsp;&nbsp;",
+           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+           align = "llcccc", # column alignment,
+           tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within July to October fire season."
+            ) # end table
+
+  print(tab)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_sex_plot <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | 
+   #       pm_method == "WRF-Chem Smoke") %>% 
+  mutate(sex_cat = ifelse(sex == 0, "Male", 
+               ifelse(sex == 1, "Female", NA))) %>%
+  # subset columns I want to put in to the table
+  select(1, 3, 11, 5:8) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_sex_plot$outcome <- factor(wrf_geo_sex_plot$outcome,
+                              levels = unique(wrf_geo_sex_plot$outcome))
+
+wrf_geo_sex_plot$pm_method <- factor(wrf_geo_sex_plot$pm_method,
+                                levels = unique(wrf_geo_sex_plot$pm_method))
+
+wrf_geo_sex_plot$sex_cat <- factor(wrf_geo_sex_plot$sex_cat,
+                                levels = unique(wrf_geo_sex_plot$sex_cat))
+
+wrf_geo_sex_plot$n_events <- as.numeric(wrf_geo_sex_plot$n_events)
+wrf_geo_sex_plot$odds_ratio <- as.numeric(wrf_geo_sex_plot$odds_ratio)
+wrf_geo_sex_plot$lower95 <- as.numeric(wrf_geo_sex_plot$lower95)
+wrf_geo_sex_plot$upper95 <- as.numeric(wrf_geo_sex_plot$upper95)
+
+## ggplot
+  print_plot <- ggplot(wrf_geo_sex_plot,
+    aes(x = sex_cat, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red", "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Sex') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10),
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
+  ggsave("../plot_lag/ndc_sex_fig_three_lag1.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+## NDC claims with index
+
+Check the claims in each category (Asthma, COPD, Both or Other).
+
+```{r echo = FALSE, message=FALSE, warning=FALSE}
+read_path7 <- paste0("../data_new/medication/oregon_beta_ndc_index.csv")
+or_ndc_index <- read_csv(read_path7)
+
+oregon_beta_df <- or_ndc_index %>% 
+  group_by(personkey) %>% 
+  arrange(personkey, fromdate, line) %>% 
+  mutate(num_visit = dense_rank(fromdate)) %>%
+  arrange(personkey, num_visit) %>% 
+  select(personkey, clmid, num_visit) %>% 
+  unique() %>% 
+  full_join(or_ndc_index, by = c("personkey", "clmid")) %>% 
+  arrange(personkey, clmid, line, fromdate) %>% 
+  filter(row_number() == 1) %>%
+  # add new transverted from date
+  mutate(dates = as.Date(fromdate, "%m/%d/%Y")) %>%
+  filter(dates >= '2013-05-01' & 
+         dates <= '2013-09-30') # 24538
+
+summary(as.factor(oregon_beta_df$index))
+
+```
+
+### Descriptives
+
+For all of 2013, there were a total of millons of all claims records in the APAC dataset.
+During the May 1st to September 30th wildfire season in Oregon state, there were 24217 pharmacy visits for the ndc of beta 2 agonists.
+The number of cases, as well as age-specific and sex-specific strate are shown below.
+
+```{r ndc descriptive table, echo = F, message = F, warning = F, results='asis'}
+read_path6 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover.csv")
+beta_ndc <- read_csv(read_path6)
+
+# join personkey with index
+oregon_beta_index_df <- oregon_beta_df %>%
+  select(personkey, index)
+
+beta_ndc_with_index <- beta_ndc %>%
+  left_join(oregon_beta_index_df, by="personkey")
+
+# dataframe to loop through
+beta_ndc_with_index_df <- data.frame(beta_ndc_with_index)
+
+beta_asthma <- beta_ndc_with_index_df %>%
+  filter(index=="asthma")
+
+beta_copd <- beta_ndc_with_index_df %>%
+  filter(index=="copd")
+
+beta_both <- beta_ndc_with_index_df %>%
+  filter(index=="both")
+
+beta_other <- beta_ndc_with_index_df %>%
+  filter(index=="other")
+
+df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)
+
+outcome_list <- c("Asthma", "COPD", "Both", "Other")
+
+# create an empty list to row bind dataframes together
+datalist <- list()
+
+# dataframe list
+method_list <- c('WRF-Chem Smoke', 'Geo-Weighted Smoke', 'Kriging Smoke')
+
+# outcome name
+outcome_name <- c("Asthma", "COPD", "Both", "Other")
+
+for(i in 1:length(df_list)){
+  # dataframe to loop through
+  df_to_loop <- data.frame(df_list[i])
+  # indication of column
+  outcome <- colnames(df_to_loop[75])
+  # outcome name
+  outcome_name <- outcome_list[i]
+
+  # dataframe for analysis creation
+  # bind columns back together 
+  df_analysis <- df_to_loop %>%
+    filter(!is.na(wrf_smk_pm_zip)) %>% 
+    # only look at outcomes
+    filter(outcome == 1) %>%
+    # add another row that makes sure there is a person <15 in the dataframe
+    # tricking xtabs to produce a 0 cell for the outcome for age <15
+    add_row(outcome = 0, age_ind = 0)
+
+  # cross tabs
+  outcome_n <- xtabs(~ outcome, df_analysis)
+  cross_tab_age <- xtabs(~ outcome + age_ind, df_analysis)
+  cross_tab_sex <- xtabs(~ outcome + sex_ind, df_analysis)
+  # empty matrix
+  point_estimates <- matrix(nrow = 1, ncol = 7, byrow = T)
+
+  colnames(point_estimates) <- c("outcome", "n", "age_15", "age_15_65", 
+                                 "age_65", "female", "male")
+
+  # fill in the outcome name for the dataframe before the loop
+  point_estimates[, 1] <- outcome_name
+  # fill n
+  point_estimates[, 2] <- outcome_n[2] # second element of the 1 dimension vector
+  # age <15
+  point_estimates[, 3] <- cross_tab_age[2, 1]
+  # age 15 to 65
+  point_estimates[, 4] <- cross_tab_age[2, 2]
+  # age >65
+  point_estimates[, 5] <- cross_tab_age[2, 3]
+  # male == 0
+  point_estimates[, 7] <- cross_tab_sex[1, 1]
+  # female == 1
+  point_estimates[, 6] <- cross_tab_sex[1, 2]
+
+  # save point estimates as a dataframe
+  point_est_df <- as_data_frame(point_estimates)
+
+  # combine previous values in dataframe that has all outcome/methods comparisons
+  datalist[[i]] <- point_est_df
+
+} # end of loop
+
+# combine each outcome dataframe itteration in to a big dataset
+combined_point_est_df <- bind_rows(datalist) %>% 
+  # find proportions/percents for each strata in a row
+  mutate(age_15_pr = as.character(round((as.numeric(age_15)/as.numeric(n))*100,1)),
+         age_15_65_pr = as.character(round((as.numeric(age_15_65)/as.numeric(n))*100,1)),
+         age_65_pr = as.character(round((as.numeric(age_65)/as.numeric(n))*100,1)),
+         female_pr = as.character(round((as.numeric(female)/as.numeric(n))*100,1)),
+         male_pr = as.character(round((as.numeric(male)/as.numeric(n))*100,1))) %>% 
+  select(outcome, n, age_15, age_15_pr, age_15_65, age_15_65_pr, age_65,
+         age_65_pr, female, female_pr, male, male_pr)
+
+#str(combined_point_est_df)
+
+tab <- htmlTable(txtRound(combined_point_est_df, digits = 1), 
+           caption = "Number of cases for each outcome observed from May 1st to September 30st, 2013",
+           # column headers
+           header = c("Outcome", "Cases n", "Less than 15", "(%)", "15 to 65", "(%)", "Greater than 65",
+                      "(%)", "Female", "(%)", "Male", "(%)"),
+           # column spanner
+           cgroup = c("","Age Category", "Sex"), 
+           n.cgroup = c(2, 6, 4),
+           padding.rgroup = "&nbsp;&nbsp;",
+           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+           align = "llccccc" # column alignment,
+            ) # end table
+
+print(tab)
+
+```
+
+### Overall plot
+
+```{r ndc overall try, warning =F, echo = F, results='asis'}
+df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)
+
+outcome_list <- c("Asthma", "COPD", "Both", "Other")
+
+# create an empty list to row bind dataframes together
+datalist <- list()
+
+# dataframe list
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# outcome name
+outcome_name <- c("Asthma", "COPD", "Both", "Other")
+
+# data wrangling ----
+# Producing conditional logit model estimates loop 
+for(i in 1:length(df_list)){
+  # dataframe to loop through
+  df_to_loop <- data.frame(df_list[i])
+  # indication of column
+  outcome <- colnames(df_to_loop[75])
+  # outcome name
+  outcome_name <- outcome_list[i]
+
+  # extract covariates from dataframe
+  covariates_df <- df_to_loop[, c(1:26, 74:84)]
+
+  # extract pm values and divide by 10 and ordered
+  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+  pm_estimates_df <- df_to_loop[, c(87, 92, 91, 93)]/10  # create 10 unit increases
+
+  # dataframe for analysis creation
+  # bind columns back together 
+  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+    # remove missing pm values
+    filter(!is.na(wrf_smk_pm_zip)) %>%
+
+    # join in cdc met data
+    # full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
+    # the following code makes sure that the counterfactual values retained are 
+    # symetric in that number of obs before = number of obs after
+    mutate(obs_diff_admission = (fromdate - date)/7) 
+  # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+  # empty df for table
+  table_df <- data.frame()
+
+  # empty matrix
+  point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)
+
+  colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 'odds_ratio', 
+                                 'lower95', 'upper95', 'se', 'p_val')
+
+  # fill in the outcome name for the dataframe before the loop
+  point_estimates[, 1] <- outcome_name
+
+  # second loop to run a model for each pm estimation method
+
+  for(j in 38:40){
+
+    # variable to model 
+    var_name <- colnames(df_analysis[j])
+
+    # set row number to fill
+    row_n <- j-37
+
+    # only run the model if the dataframe has observations
+    if(nrow(df_analysis) != 0){
+      # conditional logistic regression model
+      mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+      # populate matrix
+      point_estimates[row_n, 2] <- method_list[row_n]
+      point_estimates[row_n, 3] <- mod$n
+      point_estimates[row_n, 4] <- mod$nevent
+      # odds ratio
+      point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+      # 95% lower bound
+      point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
+      # 95% upper bound
+      point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                               1.96*(summary(mod)$coefficient[1,3])), 3)
+      # standard error
+      point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
+      # p val
+      point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)
+
+      # save point estimates as a dataframe
+      point_est_df <- as_data_frame(point_estimates)
+
+      # create else statement that fills matrix with missing so I still have the row
+      # in the final dataframe
+    } else {point_estimates[row_n, 2] <- method_list[row_n]
+    point_estimates[row_n, 3] <- 0
+    point_estimates[row_n, 4] <- 0
+    point_estimates[row_n, c(5:9)] <- 99 } # end 'if else' statement
+  } # end methods loop
+
+  # combine previous values in dataframe that has all outcome/methods comparisons
+  datalist[[i]] <- point_est_df
+
+} # end of loop
+
+# combine each outcome dataframe itteration in to a big dataset
+combined_point_est_df <- bind_rows(datalist)
+
+# subset columns I want to put in to the table
+table_df <- combined_point_est_df %>% select(2, 3:7)
+
+tab <- htmlTable(txtRound(table_df, digits = 3, 1:3), 
+                 caption = "Association between a 10 ug/m^3 in PM2.5 and Health Outcomes",
+                 # row group by outcome
+                 rgroup = outcome_list,
+                 n.rgroup = c(rep(3, 4)), # 4 rows for each method for each outcome
+                 # column headers
+                 header = c("Method", "Obs.", "Events",
+                            "OR&dagger;", "Lower", "Upper"),
+                 # column spanner
+                 cgroup = c("", "95% CI"), 
+                 n.cgroup = c(4, 2),
+                 padding.rgroup = "&nbsp;&nbsp;",
+                 css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+                 align = "llccccc", # column alignment,
+                 tfoot="&dagger; Adjusted for CDC temperature, relatively humidity, wind speed, and precipitation; accounting for subject.
+Time-stratified: referent periods matched to events on same day of week within July to October fire season."
+) # end table
+
+print(tab)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+combined_point_est_df$outcome <- factor(combined_point_est_df$outcome, 
+                                        levels = unique(combined_point_est_df$outcome))
+
+combined_point_est_df$pm_method <- factor(combined_point_est_df$pm_method, 
+                                          levels = unique(combined_point_est_df$pm_method))
+
+combined_point_est_df$n <- as.numeric(combined_point_est_df$n)
+combined_point_est_df$n_events <- as.numeric(combined_point_est_df$n_events)
+combined_point_est_df$odds_ratio <- as.numeric(combined_point_est_df$odds_ratio)
+combined_point_est_df$lower95 <- as.numeric(combined_point_est_df$lower95)
+combined_point_est_df$upper95 <- as.numeric(combined_point_est_df$upper95)
+combined_point_est_df$se <- as.numeric(combined_point_est_df$se)
+combined_point_est_df$p_val <- as.numeric(combined_point_est_df$p_val)
+
+# subset out global method
+combined_point_est_df <- combined_point_est_df
+
+## ggplot
+print_plot <- ggplot(combined_point_est_df, 
+                     aes(x = pm_method, y = odds_ratio, colour = pm_method)) +
+  # custom color 
+  scale_color_manual(name = "Smoke-Estimation Method", 
+                     values = c("red", "blue", "#32115C"),
+                     guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+  geom_point() + #geom_text(vjust = 0, nudge_x = 0.3) +
+  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
+  facet_wrap(~outcome, nrow = 3) +
+  geom_hline(yintercept = 1, linetype=2) +
+  #    ggtitle('Association Between PM2.5 \n from Wildfire Smoke on Hospitalizations') +
+  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+  #xlab('Smoke Estimation Method') +
+  # plot theme
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        # strip element
+        strip.background = element_rect(colour=NA, fill=NA),
+        panel.border = element_rect(fill = NA, color = "black"),
+        # facet text size
+        #strip.text = element_text(size = 8),
+        # axis element
+        axis.text.x = element_blank(),
+        axis.title.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        axis.title.y = element_text(angle = 90),
+        # legend elements
+        legend.position = "bottom")
+#legend.text = element_text(size = 8))
+
+print(print_plot)
+# save figure
+ggsave("../plot_new/fig_ndc_overall_four.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+### Time-Stratified by Age Categories
+
+Following table and figure look at some outcomes stratified by age category.
+
+For some outcomes, such like Age > 65 in WRF-Chem method, the CI is the only insignificant one in the plot.
+
+```{r ndc time stratified age strata, warning =F, echo = F, results='asis'} 
+df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)
+
+outcome_list <- c("Asthma", "COPD", "Both", "Other")
+
+# create an empty list to row bind dataframes together
+datalist <- list()
+
+# outcome name
+outcome_name <- c("Asthma", "COPD", "Both", "Other")
+
+# age category list
+age_cat_list <- c(0,1,2)
+
+# looking only at wrf chem, kriging, and geo-weighted; taking out 'Global Smoke' method
+method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+
+# create an empty list to row bind dataframes together
+datalist1 <- list()
+datalist2 <- list()
+
+for(i in 1:length(df_list)){
+
+  # dataframe to loop through
+  df_to_loop <- data.frame(df_list[i])
+  # indication of column
+  outcome <- df_list[i]
+  # outcome name
+  outcome_name <- outcome_list[i]
+
+  covariates_df <- df_to_loop[, c(1:26, 74:84)]
+
+  # extract pm values and divide by 10 and ordered
+  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+  pm_estimates_df <- df_to_loop[, c(87, 92, 91, 93)]/10  # create 10 unit increases
+
+  # new loop for age categories
+  for(k in 0:2){
+
+    point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)
+
+    colnames(point_estimates) <- c('outcome', 'age_cat', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+    # fill in the outcome namedataframe before method loop
+    point_estimates[, 1] <- outcome_name
+    # repeat the age category 4 times for each pm method
+    point_estimates[, 2] <- k
+
+    # dataframe for analysis creation
+    # bind columns back together 
+    df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+      # remove missing pm values
+      filter(!is.na(wrf_smk_pm_zip)) %>% 
+      # limit to specific age category
+      filter(age_ind == k) %>% 
+      # the following code makes sure that the counterfactual values retained are 
+      # symetric in that number of obs before = number of obs after
+      mutate(obs_diff_admission = (fromdate - date)/7)
+
+    for(j in 38:40){
+
+      # variable to model 
+      var_name <- colnames(df_analysis[j])
+
+      # set row number to fill
+      row_n <- j-37
+
+      # only run the model if the dataframe has observations
+      if(nrow(df_analysis) != 0){
+        # conditional logistic regression model
+        mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+        # populate matrix
+        point_estimates[row_n, 3] <- method_list[row_n]
+        point_estimates[row_n, 4] <- mod$n
+        point_estimates[row_n, 5] <- mod$nevent
+        # odds ratio
+        point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+        # 95% lower bound
+        point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
+        # 95% upper bound
+        point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                                 1.96*(summary(mod)$coefficient[1,3])), 3)
+        # standard error
+        point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
+        # p val
+        point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
+
+        # create else statement that fills matrix with missing so I still have the row
+        # in the final dataframe
+      } else {point_estimates[row_n, 3] <- method_list[row_n]
+      point_estimates[row_n, 4] <- 0
+      point_estimates[row_n, 5] <- 0
+      point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
+    } # end methods loop
+
+    # save point estimates as a dataframe
+    point_est_df <- as_data_frame(point_estimates)
+
+    # combine previous values in dataframe that has all outcome/methods comparisons
+    datalist1[[k+1]] <- point_est_df
+  } # end age category loop
+
+  # bind rows of age category estimates together
+  age_est_df <- bind_rows(datalist1)
+
+  # populate second dataframe list
+  datalist2[[i]] <- age_est_df
+
+} # end of outcome loop
+
+# combine each outcome dataframe itteration in to a big dataset
+combined_point_est_df <- bind_rows(datalist2)
+
+wrf_geo_age <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" |
+    #      pm_method == "WRF-Chem Smoke") %>% 
+  mutate(age_cat2 = ifelse(age_cat == 0, "Less than 15", 
+                    ifelse(age_cat == 1, "15-65",
+                    ifelse(age_cat == 2, "Greater than 65", NA)))) %>% 
+  # subset columns I want to put in to the table
+  select(3, 11, 5:8) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
+         caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by age",
+         # row group by outcome
+         rgroup = "Asthma inhalers",
+         n.rgroup = c(rep(9, 1)), # 6 rows for each age cat for each outcome
+         # column headers
+         header = c("Est Method", "Age Group", "Events",
+                    "OR&dagger;", "Lower", "Upper"),
+         # column spanner
+         cgroup = c("", "95% CI"), 
+         n.cgroup = c(4, 2),
+         padding.rgroup = "&nbsp;&nbsp;",
+         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+         align = "llcccc", # column alignment,
+         tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within May to September fire season."
+         ) # end table
+
+print(tab)
+
+combined_point_est_df_new <- combined_point_est_df %>%
+  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
+                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
+  select(-n_events) %>%
+  rename(n_events = n_events_new)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_age_plot <- combined_point_est_df_new %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  # do not plot results with less than 15 events, create missing vals
+  # in mutate
+  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
+                             99, odds_ratio),
+         lower95 = ifelse(as.numeric(n_events) < 15, 
+                             99, lower95),
+         upper95 = ifelse(as.numeric(n_events) < 15, 
+                             99, upper95),
+         age_cat2 = ifelse(age_cat == 0, "<15", 
+                    ifelse(age_cat == 1, "15-65",
+                    ifelse(age_cat == 2, ">65", NA)))) %>% 
+  # subset columns I want to put in to the table
+  select(1, 3, 11, 10, 5:7) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
+                              levels = unique(wrf_geo_age_plot$outcome))
+
+wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
+                                levels = unique(wrf_geo_age_plot$pm_method))
+
+wrf_geo_age_plot$age_cat2 <- factor(wrf_geo_age_plot$age_cat2,
+                                levels = unique(wrf_geo_age_plot$age_cat2))
+
+wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
+wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
+wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
+wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)
+
+## ggplot 
+  print_plot <- ggplot(wrf_geo_age_plot,
+    aes(x = age_cat2, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red",  "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Age Category ') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10), 
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
+  ggsave("../plot_new/ndc_age_fig_four.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+### Seasonal Time-Stratified by Sex
+
+Association with smoke PM~2.5~ and health outcomes stratified by sex.
+
+All the CI do not include 1, which means the use of asthma medication (beta 2 agonists) increases about 5% when smoke increase 10 units.
+So the risk for getting asthma will increase about 5% ~ 8% when the smoke PM2.5 increase 10 units.
+
+```{r ndc sex strata, warning = F, echo = F, results='asis'} 
+df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)
+
+outcome_list <- c("Asthma", "COPD", "Both", "Other")
+
+# outcome name
+outcome_name <- c("Asthma", "COPD", "Both", "Other")
+
+# sex category list
+sex_strata_list <- c(0,1)
+
+datalist1 <- list()
+datalist2 <- list()
+
+# data wrangling ----
+# Producing conditional logit model estimates loop 
+for(i in 1:length(df_list)){
+
+  # dataframe to loop through
+  df_to_loop <- data.frame(df_list[i])
+  # indication of column
+  outcome <- df_list[i]
+  # outcome name
+  outcome_name <- outcome_list[i]
+
+  covariates_df <- df_to_loop[, c(1:26, 74:84)]
+
+  # extract pm values and divide by 10 and ordered
+  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+  pm_estimates_df <- df_to_loop[, c(87, 92, 91, 93)]/10  # create 10 unit increases
+
+  # new loop for sex categories
+  for(k in 0:1){
+
+    # empty matrix (12 x 10 matrix)
+    point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)
+
+    colnames(point_estimates) <- c('outcome', 'sex', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+    # fill in the outcome namedataframe before method loop
+    point_estimates[, 1] <- outcome_name
+    # repeat the sex category 4 times for each pm method
+    point_estimates[, 2] <- k
+
+    # dataframe for analysis creation
+    # bind columns back together 
+    df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+      # remove missing pm values
+      filter(!is.na(wrf_smk_pm_zip)) %>% 
+      # limit to specific sex category
+      filter(sex_ind == k) %>% 
+      # the following code makes sure that the counterfactual values retained are 
+      # symetric in that number of obs before = number of obs after
+      mutate(obs_diff_admission = (fromdate - date)/7) 
+    # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+    for(j in 38:40){
+
+      # variable to model 
+      var_name <- colnames(df_analysis[j])
+
+      # set row number to fill
+      row_n <- j-37
+
+      # only run the model if the dataframe has observations
+      if(nrow(df_analysis) != 0){
+        # conditional logistic regression model
+        mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+        # populate matrix
+        point_estimates[row_n, 3] <- method_list[row_n]
+        point_estimates[row_n, 4] <- mod$n
+        point_estimates[row_n, 5] <- mod$nevent
+        # odds ratio
+        point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+        # 95% lower bound
+        point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
+                                                 1.96*(summary(mod)$coefficient[1,3])), 3)
+        # 95% upper bound
+        point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                                 1.96*(summary(mod)$coefficient[1,3])), 3)
+        # standard error
+        point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
+        # p val
+        point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
+
+        # create else statement that fills matrix with missing so I still have the row
+        # in the final dataframe
+      } else {point_estimates[row_n, 3] <- method_list[row_n]
+      point_estimates[row_n, 4] <- 0
+      point_estimates[row_n, 5] <- 0
+      point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
+    } # end methods loop
+
+    # save point estimates as a dataframe
+    point_est_df <- as_data_frame(point_estimates)
+
+    # combine previous values in dataframe that has all outcome/methods comparisons
+    datalist1[[k+1]] <- point_est_df
+  } # end sex category loop
+
+  # bind rows of sex category estimates together
+  sex_est_df <- bind_rows(datalist1)
+
+  # populate second dataframe list
+  datalist2[[i]] <- sex_est_df
+
+} # end of outcome loop
+
+# combine each outcome dataframe itteration in to a big dataset
+combined_point_est_df <- bind_rows(datalist2)
+
+wrf_geo_sex <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  mutate(sex_cat = ifelse(sex == 0, "Male", 
+               ifelse(sex == 1, "Female", NA))) %>% 
+  # subset columns I want to put in to the table
+  select(3, 11, 5:8)%>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+  tab <- htmlTable(txtRound(wrf_geo_sex, digits = 3, 1:3), 
+           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and pharmacy outcomes stratified by sex",
+           # row group by outcome
+           rgroup = "Asthma inhalers",
+           n.rgroup = c(rep(6, 1)), # 2 rows for each sex strata for each outcome
+           # column headers
+           header = c("Est Method", "Sex Strata", "Events",
+                      "OR&dagger;", "Lower", "Upper"),
+           # column spanner
+           cgroup = c("", "95% CI"), 
+           n.cgroup = c(4, 2),
+           padding.rgroup = "&nbsp;&nbsp;",
+           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+           align = "llcccc", # column alignment,
+           tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within July to October fire season."
+            ) # end table
+
+  print(tab)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_sex_plot <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | 
+   #       pm_method == "WRF-Chem Smoke") %>% 
+  mutate(sex_cat = ifelse(sex == 0, "Male", 
+               ifelse(sex == 1, "Female", NA))) %>%
+  # subset columns I want to put in to the table
+  select(1, 3, 11, 5:8) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_sex_plot$outcome <- factor(wrf_geo_sex_plot$outcome,
+                              levels = unique(wrf_geo_sex_plot$outcome))
+
+wrf_geo_sex_plot$pm_method <- factor(wrf_geo_sex_plot$pm_method,
+                                levels = unique(wrf_geo_sex_plot$pm_method))
+
+wrf_geo_sex_plot$sex_cat <- factor(wrf_geo_sex_plot$sex_cat,
+                                levels = unique(wrf_geo_sex_plot$sex_cat))
+
+wrf_geo_sex_plot$n_events <- as.numeric(wrf_geo_sex_plot$n_events)
+wrf_geo_sex_plot$odds_ratio <- as.numeric(wrf_geo_sex_plot$odds_ratio)
+wrf_geo_sex_plot$lower95 <- as.numeric(wrf_geo_sex_plot$lower95)
+wrf_geo_sex_plot$upper95 <- as.numeric(wrf_geo_sex_plot$upper95)
+
+## ggplot
+  print_plot <- ggplot(wrf_geo_sex_plot,
+    aes(x = sex_cat, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red", "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Sex') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10),
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
   ggsave("../plot_new/ndc_sex_fig_four.pdf", plot = print_plot, 

### 2017-06-22 r_scripts/analysis/oregon_medication.Rmd
Update oregon_medication.Rmd

@@
 library(knitr)
+library(broom)
 
@@
 read_path6 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover.csv")
-asthma_ndc <- fread(read_path6, showProgress = T)
+asthma_ndc <- read_csv(read_path6)
+# asthma_ndc <- fread(read_path6, showProgress = T) # faster, but seems error in overall plot analysis
 
@@
 
+a <- glm(n_obs ~ geo_smk_pm, data = or_ndc_county_time_series, family = "poisson")
+summary(a)
+
+```
+
+```{r}
+
+# create empty dataframe
+rr_df <- data.frame(matrix(nrow = 3, ncol = 4))
+colnames(rr_df) <- c("outcome", "rel_risk", "lower_95", "upper_95")
+
+# vector of outcomes
+outcome_list <- c("obs_n")
+
+# names 
+outcome_names <- c('Inhalers')
+
+# populate outcome column
+rr_df[,1] <- factor(outcome_names, levels=unique(outcome_names))
+
+meow <- lapply(outcome_list, function(x){
+
+ smk_est <- tidy(glm(as.formula(paste0(x, 
+    "~ geo_smk10 + wrf_temp + county + 
+    day + offset(log(population))")), 
+    or_ndc_county_time_series, family="quasipoisson"))[2, ]
+
+ rr <- c(round(exp(smk_est[1,2]),3),
+         round(exp(smk_est[1,2]-1.96*smk_est[1,3]),3),
+         round(exp(smk_est[1,2]+1.96*smk_est[1,3]),3))
+
+}) # end apply
+
+for(i in 1:nrow(rr_df)){
+  rr_df[i,2:4] <- meow[[i]]
+}
+
+# add in a grouping variable
+rr_df$group <- NA
+rr_df[1:5, 5] <- "Respiratory"
+rr_df[6:11, 5] <- "CVD"
+rr_df[12, 5] <- "Broken Arm"
+
+# tried apply function and it might have been easier to just use a loop
+# as I needed a for loop to fill the list in to the dataframe
+# code is more compact though
+# I should check speed (or make a custom function to do this one day)
+
+ggplot(rr_df, aes(x=outcome, y = rel_risk, colour = group)) +
+  geom_point() +
+  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
+  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
+  theme(
+   panel.background = element_rect(fill = 'white', colour = 'black'),
+   panel.grid.major = element_blank(),
+   panel.grid.minor = element_blank(),
+   axis.text.x = element_text(angle = 90, vjust = .75)) +
+  ylab("Relative Risk") +
+  xlab("Outcome") +
+  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")
+
 ```
@@
 ```
+
+### Time series
+
+```{r, eval=FALSE}
+# Join data with names of Oregon counties 
+oregon_fips <- read_csv('../instructions/oregon_FIPS.csv')
+# remove the "County" character
+# factor(oregon_fips$county)
+oregon_fips$county <- gsub(" County", "", as.character(factor(oregon_fips$county)))
+
+oregon_fips <- oregon_fips %>%
+  mutate(st_county_fips = with(oregon_fips, paste0(st_code, fips)))
+
+### Join county with zip
+read_path <- paste0('../data_new/update/or_zip_county_prop.csv') 
+or_zip_county <- read_csv(read_path)
+
+or_zip_county <- or_zip_county %>%
+  select(zip, county_name) %>%
+  rename(county = county_name) %>%
+  rename(ZIP = zip) %>%
+  full_join(oregon_fips, by = "county") %>%
+  select(ZIP, county, state, st_county_fips) %>%
+  rename(fips = st_county_fips)
+
+or_zip_county$county[which(or_zip_county$county=="Hood.River")] <- "Hood River"
+
+# Import population weighted county PM2.5
+pm_path <- paste0("../data_new/county_data/or_county_pop_wt_pm.csv")
+county_pm <- read_csv(pm_path)
+
+```
+
+```{r time series, warning=FALSE, message=FALSE, echo=FALSE}
+
+oregon_beta_cat_df <- oregon_beta_df %>%
+  select(personkey, num_visit, clmid, line, ZIP, dates, index)
+
+# create a working dataset 'or_2013'
+oregon_county_beta_cat_ts <- oregon_beta_cat_df %>% 
+  # join in the washington county name based on CHARS county identifier
+  left_join(or_zip_county, by = "ZIP") %>% 
+  # rename 'admit_date_impute' to 'date'
+  rename(date = dates) %>%
+  # rename(date = admit_date_impute) %>% 
+  # group by county of residence and date
+  group_by(county, date, index) %>% 
+  # sum up each primary diagnosis for each outcome for each day for each county
+  summarise(n_obs = n()) %>% 
+  # join daily county estimates of pm2.5
+  left_join(county_pm, by = c('date', 'county'))
+
+# which(is.na(oregon_county_beta_cat_ts$county))
+# oregon_county_beta_cat_ts[113,]
+# which(or_zip_county$ZIP=="97078")
+
+# summary(or_ndc_asthma)
+# summary(oregon_county_beta_cat_ts)
+summary(as.factor(oregon_county_beta_cat_ts$county)) # 125 is missing out of 3592 (3.16%)
+
+# remove NA values
+which(is.na(oregon_county_beta_cat_ts$county)) # 7728-7931
+oregon_county_beta_cat_ts <- oregon_county_beta_cat_ts[-c(7728:7931),]
+
+# write permanent dataframe
+
+write_path5 <- paste0("../data_new/medication/or_county_beta_cat_ts.csv")
+write_csv(oregon_county_beta_cat_ts, write_path5)
+
+## Check for missing values in sparsely populated counties. set to 0 (But no)
+# check_missing <- filter(oregon_county_beta_cat_ts, is.na(n_obs))
+# mutate each to fill in missing values (Still no)
+# check2 <- check_missing %>% mutate_each(funs(new = ifelse(is.na(.), 0, .)), 
+#                                         n_obs:wrf_temp)
+
+plot_ts <- ggplot(oregon_county_beta_cat_ts, aes(y =n_obs, x = date)) +
+           facet_wrap(~county, scales = "free") +
+           geom_jitter()
+
+```
+
+```{r}
+
+oregon_beta_rate <- oregon_county_beta_cat_ts %>%
+  group_by(county, index) %>%
+  summarise(population=n()) %>%  # 143 rows
+  mutate
+
+# create empty dataframe
+rr_df <- data.frame(matrix(nrow = 12, ncol = 4))
+colnames(rr_df) <- c("outcome", "rel_risk", "lower_95", "upper_95")
+
+# vector of outcomes
+outcome_list <- colnames(meta_ts[, which(colnames(meta_ts)=="resp_n"):
+    which(colnames(meta_ts)=="broken_arm_n")])
+
+# names 
+outcome_names <- c('All Respiratory', 'Asthma', 'COPD', 'Pneumonia', 
+  'Acute Bronchitis', 'Cardiovascular Disease', 'Arrhythmia', 
+  'Cerebrovascular Disease', 'Heart Failure', 'Ischemic Heart Disease',
+  'Myocardial Infarction', 'Broken Arm')
+
+# populate outcome column
+rr_df[,1] <- factor(outcome_names, levels=unique(outcome_names))
+
+meow <- lapply(outcome_list, function(x){
+
+ smk_est <- tidy(glm(as.formula(paste0(x, 
+    "~ geo_smk10 + wrf_temp + state + county + 
+    day + offset(log(population))")), 
+    meta_ts, family="quasipoisson"))[2, ]
+
+ rr <- c(round(exp(smk_est[1,2]),3),
+         round(exp(smk_est[1,2]-1.96*smk_est[1,3]),3),
+         round(exp(smk_est[1,2]+1.96*smk_est[1,3]),3))
+
+}) # end apply
+
+for(i in 1:nrow(rr_df)){
+  rr_df[i,2:4] <- meow[[i]]
+}
+
+# add in a grouping variable
+rr_df$group <- NA
+rr_df[1:5, 5] <- "Respiratory"
+rr_df[6:11, 5] <- "CVD"
+rr_df[12, 5] <- "Broken Arm"
+
+# tried apply function and it might have been easier to just use a loop
+# as I needed a for loop to fill the list in to the dataframe
+# code is more compact though
+# I should check speed (or make a custom function to do this one day)
+
+ggplot(rr_df, aes(x=outcome, y = rel_risk, colour = group)) +
+  geom_point() +
+  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
+  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
+  theme(
+   panel.background = element_rect(fill = 'white', colour = 'black'),
+   panel.grid.major = element_blank(),
+   panel.grid.minor = element_blank(),
+   axis.text.x = element_text(angle = 90, vjust = .75)) +
+  ylab("Relative Risk") +
+  xlab("Outcome") +
+  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")
+
+```

### 2017-07-07 r_scripts/analysis/oregon_medication.Rmd
doing distributed lag
update revised zipcode and DLM

--- 2017-06-14 r_scripts/analysis/oregon_medication.Rmd
@@
 ```{r library, include=FALSE, echo=FALSE}
-library(dplyr)
-library(data.table)
-library(readxl) 
-library(readr)
+library(tidyverse)
+# library(data.table)
 library(survival)
@@
 library(knitr)
+library(broom)
+library(lme4)
+library(lubridate) # date
+
+# two-stage distributed lag model libraries
+library(dlnm)
+library(mvmeta)
 library(splines)
+
 ```
@@
 read_path2 <- paste0("../../../data/data_new/ndc_2013_beta_2_agonists.csv")
-asthma_ndc <- read_csv(read_path2, col_names = FALSE)
-
-colnames(asthma_ndc) <- c("ndc_code", "brand_name", "generic_product_name", 
+inhaler_ndc <- read_csv(read_path2, col_names = FALSE)
+
+colnames(inhaler_ndc) <- c("ndc_code", "brand_name", "generic_product_name", 
                           "route", "category", "drug_id")
@@
 # convert to vector
-ndc_codes <- as.vector(as.matrix(asthma_ndc$ndc_code))
+ndc_codes <- as.vector(as.matrix(inhaler_ndc$ndc_code))
 
@@
 
-```{r time stratified, eval=FALSE, echo=FALSE}
+```{r time stratified zip for casecrossover, eval=FALSE, echo=FALSE}
 ### Time stratified ------------------------------------------------------------
@@
 
-write_path4 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover.csv")
+write_path4 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover_zip.csv")
 write_csv(outcome_casecross, write_path4)
+
+```
+
+## Time stratified for time series (Population-Weighted PM~2.5~ Assigned by County)
+
+```{r time stratified county for ts, echo=FALSE, include=FALSE, eval=FALSE}
+read_path7 <- paste0("../data_new/medication/oregon_asthma_ndc.csv")
+or_ndc_claim <- read_csv(read_path7)
+
+or_ndc_inhaler <- or_ndc_claim %>%
+  select(personkey, num_visit, clmid, line, ZIP, dates) %>%
+  rename(ZIPCODE = ZIP)
+
+# Join data with names of Oregon counties 
+oregon_fips <- read_csv('../instructions/oregon_FIPS.csv')
+# remove the "County" character
+# factor(oregon_fips$county)
+oregon_fips$county <- gsub(" County", "", as.character(factor(oregon_fips$county)))
+
+oregon_fips <- oregon_fips %>%
+  mutate(st_county_fips = with(oregon_fips, paste0(st_code, fips)))
+
+### Join county with zip
+# read_path <- paste0('../data_new/update/or_zip_county_prop.csv') 
+# or_zip_county <- read_csv(read_path)
+
+read_path <- paste0('../instructions/oregon_zip_county.csv')
+or_zip_county <- read_csv(read_path)
+
+# or_zip_county$county_name[which(or_zip_county$county_name=="Hood.River")] <- "Hood River"
+
+or_zip_county <- or_zip_county %>%
+  select(zip, county_name) %>%
+  rename(county = county_name) %>%
+  rename(ZIPCODE = zip) %>%
+  full_join(oregon_fips, by = "county") %>%
+  select(ZIPCODE, county, state, st_county_fips) %>%
+  rename(fips = st_county_fips) # 484
+
+county_new <- c("Washington", "Douglas", "Lane", "Deschutes")
+
+zip_new <- or_ndc_inhaler %>%
+  filter(!(ZIPCODE %in% or_zip_county$ZIPCODE)) %>%
+  select(ZIPCODE) %>%
+  unique() %>%
+  arrange(ZIPCODE) %>%
+  mutate(county = county_new)
+
+# summary(as.factor(zip_new$ZIPCODE))
+# 97003 97471 97475 97703 
+#    48   138     5    16
+
+or_zip_county_new <- or_zip_county %>%
+  select(ZIPCODE, county) %>%
+  bind_rows(zip_new)
+
+write_path <- paste0("../data_new/county_data/zip_to_county_new.csv")
+write_csv(or_zip_county_new, write_path)
+
+read_path9 <- paste0("../data_new/county_data/or_census_pop_est_county.csv")
+or_pop <- read_csv(read_path9)
+
+or_pop <- or_pop %>%
+  rename(county = GEO_display_label, pop = respop72013) %>%
+  select(county, pop) 
+or_pop$county <- gsub(" County, Oregon", "", as.character(factor(or_pop$county)))
+
+### import county data
+read_path_county_pm <- paste0('C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire_new/data_new/county_data/or_county_pop_wt_pm.csv')
+
+county_smoke <- read_csv(read_path_county_pm) # 63801 rows
+
+# descriptives of the two smoke datasets
+summary(county_smoke)
+
+county_smoke$county[which(county_smoke$county=="Hood.River")] <- "Hood River"
+
+# Zipcode PM2.5 estimates
+# create lag variables that take smoke values from n previous days for zipcodes
+county_smoke_w_lag <- county_smoke %>% arrange(county, date) %>%
+  # group by zipcode
+  group_by(county) %>% 
+  # wrf
+  mutate(wrf_f_pm_lag1 = lag(wrf_pm, 1, order_by = county), 
+         wrf_f_pm_lag2 = lag(wrf_pm, 2, order_by = county),
+         wrf_f_pm_lag3 = lag(wrf_pm, 3, order_by = county),
+         wrf_f_pm_lag4 = lag(wrf_pm, 4, order_by = county),
+         wrf_f_pm_lag5 = lag(wrf_pm, 5, order_by = county),
+         wrf_f_pm_lag6 = lag(wrf_pm, 6, order_by = county),
+         wrf_f_pm_lag7 = lag(wrf_pm, 7, order_by = county),
+         # wrf no fire lag
+         wrf_nf_pm_lag1 = lag(wrf_nf_pm, 1, order_by = county),
+         wrf_nf_pm_lag2 = lag(wrf_nf_pm, 2, order_by = county),
+         wrf_nf_pm_lag3 = lag(wrf_nf_pm, 3, order_by = county),
+         wrf_nf_pm_lag4 = lag(wrf_nf_pm, 4, order_by = county),
+         wrf_nf_pm_lag5 = lag(wrf_nf_pm, 5, order_by = county),
+         wrf_nf_pm_lag6 = lag(wrf_nf_pm, 6, order_by = county),
+         wrf_nf_pm_lag7 = lag(wrf_nf_pm, 7, order_by = county),
+         # wrf_smk_pm
+         wrf_smk_pm_lag1 = lag(wrf_smk_pm, 1, order_by = county),
+         wrf_smk_pm_lag2 = lag(wrf_smk_pm, 2, order_by = county),
+         wrf_smk_pm_lag3 = lag(wrf_smk_pm, 3, order_by = county),
+         wrf_smk_pm_lag4 = lag(wrf_smk_pm, 4, order_by = county),
+         wrf_smk_pm_lag5 = lag(wrf_smk_pm, 5, order_by = county),
+         wrf_smk_pm_lag6 = lag(wrf_smk_pm, 6, order_by = county),
+         wrf_smk_pm_lag7 = lag(wrf_smk_pm, 7, order_by = county),
+         # geo weighted pm
+         geo_wt_pm_lag1 = lag(geo_wt_pm, 1, order_by = county),
+         geo_wt_pm_lag2 = lag(geo_wt_pm, 2, order_by = county),
+         geo_wt_pm_lag3 = lag(geo_wt_pm, 3, order_by = county),
+         geo_wt_pm_lag4 = lag(geo_wt_pm, 4, order_by = county),
+         geo_wt_pm_lag5 = lag(geo_wt_pm, 5, order_by = county),
+         geo_wt_pm_lag6 = lag(geo_wt_pm, 6, order_by = county),
+         geo_wt_pm_lag7 = lag(geo_wt_pm, 7, order_by = county),
+         # krig pm
+         krig_pm_lag1 = lag(krig_pm, 1, order_by = county),
+         krig_pm_lag2 = lag(krig_pm, 2, order_by = county),
+         krig_pm_lag3 = lag(krig_pm, 3, order_by = county),
+         krig_pm_lag4 = lag(krig_pm, 4, order_by = county),
+         krig_pm_lag5 = lag(krig_pm, 5, order_by = county), 
+         krig_pm_lag6 = lag(krig_pm, 6, order_by = county),
+         krig_pm_lag7 = lag(krig_pm, 7, order_by = county), 
+         # background pm
+         background_pm_lag1 = lag(background_pm, 1, order_by = county),
+         background_pm_lag2 = lag(background_pm, 2, order_by = county),
+         background_pm_lag3 = lag(background_pm, 3, order_by = county),
+         background_pm_lag4 = lag(background_pm, 4, order_by = county),
+         background_pm_lag5 = lag(background_pm, 5, order_by = county), 
+         background_pm_lag6 = lag(background_pm, 6, order_by = county),
+         background_pm_lag7 = lag(background_pm, 7, order_by = county), 
+         # geo_smk_pm 
+         geo_smk_pm_lag1 = lag(geo_smk_pm, 1, order_by = county),
+         geo_smk_pm_lag2 = lag(geo_smk_pm, 2, order_by = county),
+         geo_smk_pm_lag3 = lag(geo_smk_pm, 3, order_by = county),
+         geo_smk_pm_lag4 = lag(geo_smk_pm, 4, order_by = county),
+         geo_smk_pm_lag5 = lag(geo_smk_pm, 5, order_by = county),
+         geo_smk_pm_lag6 = lag(geo_smk_pm, 6, order_by = county),
+         geo_smk_pm_lag7 = lag(geo_smk_pm, 7, order_by = county),
+         # krig smk pm
+         krig_smk_pm_lag1 = lag(krig_smk_pm, 1, order_by = county),
+         krig_smk_pm_lag2 = lag(krig_smk_pm, 2, order_by = county),
+         krig_smk_pm_lag3 = lag(krig_smk_pm, 3, order_by = county),
+         krig_smk_pm_lag4 = lag(krig_smk_pm, 4, order_by = county),
+         krig_smk_pm_lag5 = lag(krig_smk_pm, 5, order_by = county),
+         krig_smk_pm_lag6 = lag(krig_smk_pm, 6, order_by = county),
+         krig_smk_pm_lag7 = lag(krig_smk_pm, 7, order_by = county),
+         # temp
+         wrf_temp_lag1 = lag(wrf_temp, 1, order_by = county),
+         wrf_temp_lag2 = lag(wrf_temp, 2, order_by = county),
+         wrf_temp_lag3 = lag(wrf_temp, 3, order_by = county),
+         wrf_temp_lag4 = lag(wrf_temp, 4, order_by = county),
+         wrf_temp_lag5 = lag(wrf_temp, 5, order_by = county), 
+         wrf_temp_lag6 = lag(wrf_temp, 6, order_by = county),
+         wrf_temp_lag7 = lag(wrf_temp, 7, order_by = county)) %>% 
+  # ungroup by zip
+  ungroup(county) %>% 
+  # attach a zip indicator for each smoke variable
+  setNames(paste(colnames(.), "county", sep="_"))  %>% 
+  # remove the '_zip' from the zipcode and date variable 
+  rename(county = county_county, date = date_county, fips = fips_county) %>%
+  select(-1)
+
+or_ndc_county_ts <- or_ndc_inhaler %>%
+  rename(date = dates) %>%
+  mutate(day = ifelse(weekdays(date) == "Monday" |
+                      weekdays(date) == "Tuesday" |
+                      weekdays(date) == "Wednesday"|
+                      weekdays(date) == "Thursday"|
+                      weekdays(date) == "Friday", "weekday",
+               ifelse(weekdays(date) == "Saturday" |
+                      weekdays(date) == "Sunday", "weekend", NA))) %>%
+  # join with county-level population
+  left_join(or_zip_county_new, by = "ZIPCODE") %>% # 321 missing 
+  filter(!is.na(county)) %>% # no missing value
+  left_join(or_pop, by = "county") %>%
+  left_join(county_smoke_w_lag, by = c("date", "county")) %>%
+  arrange(personkey, date) %>% # order by id and fromdate
+  group_by(county, date) %>% 
+  # sum up each primary diagnosis for each outcome for each day for each county
+  summarise(n_obs = n()) %>% 
+  # join daily county estimates of pm2.5
+  left_join(county_smoke_w_lag, by = c('date', 'county'))
+
+# summary(as.factor(or_ndc_county_ts$county))
+
+or_ndc_county_ts <- or_ndc_county_ts %>%
+  left_join(or_pop, by = "county") %>%
+  mutate(geo_smk_pm10 = geo_smk_pm_county/10) %>%
+  mutate(rate_per_100k = (n_obs/pop)*100000) %>%
+  mutate(day = ifelse(weekdays(date) == "Monday" |
+                      weekdays(date) == "Tuesday" |
+                      weekdays(date) == "Wednesday"|
+                      weekdays(date) == "Thursday"|
+                      weekdays(date) == "Friday", "weekday",
+               ifelse(weekdays(date) == "Saturday" |
+                      weekdays(date) == "Sunday", "weekend", NA)))
+
+or_ndc_county_ts$day[which(or_ndc_county_ts$date=="2013-05-27")] <- "weekend"
+or_ndc_county_ts$day[which(or_ndc_county_ts$date=="2013-07-04")] <- "weekend"
+or_ndc_county_ts$day[which(or_ndc_county_ts$date=="2013-09-02")] <- "weekend"
+
+write_path5 <- paste0("../data_new/medication/or_ndc_county_time_series.csv")
+write_csv(or_ndc_county_ts, write_path5)
 
@@
 ```{r descriptive table, echo = F, message = F, warning = F, results='asis'}
-read_path6 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover.csv")
-asthma_ndc <- read_csv(read_path6)
+read_path6 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover_zip.csv")
+inhaler_ndc <- read_csv(read_path6)
+# asthma_ndc <- fread(read_path6, showProgress = T) # faster, but seems error in overall plot analysis
 
@@
 # dataframe to loop through
-asthma_ndc_df <- data.frame(asthma_ndc)
+inhaler_ndc_df <- data.frame(inhaler_ndc)
 
@@
 # bind columns back together 
-df_analysis <- asthma_ndc_df %>%
+df_analysis <- inhaler_ndc_df %>%
   filter(!is.na(wrf_smk_pm_zip)) %>% 
@@
 # extract covariates from dataframe
-covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]
+covariates_df <- inhaler_ndc_df[, c(1:26, 74:84)]
 
@@
 #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
-pm_estimates_df <- asthma_ndc_df[, c(87, 92, 91, 93)]/10  # create 10 unit increases
+pm_estimates_df <- inhaler_ndc_df[, c(87, 92, 91, 93)]/10  # create 10 unit increases
 
@@
 
-### Overall pharmacy visit time series
-
-The time series for the number of pharmacy in each day and separated with different counties. For convenience and representativeness, I choose some of the largest counties to show the visit features.
-
-The overall time series plot shows a time series plot.
+### County Time series plots of rate
+
+### Prepare (already done in previous time stratified work)
+
+1. Make 10 units' geo smoke variable
+2. Make county population variable (to calculate rate)
+3. Make weekend/weekday variables
 
 ```{r time series, warning=FALSE, message=FALSE, echo=FALSE}
-read_path7 <- paste0("../data_new/medication/oregon_asthma_ndc.csv")
-or_ndc_claim <- read_csv(read_path7)
-
-or_ndc_asthma <- or_ndc_claim %>%
-  select(personkey, num_visit, clmid, line, ZIP, dates)
-
-# Join data with names of Oregon counties 
-oregon_fips <- read_csv('../instructions/oregon_FIPS.csv')
-# remove the "County" character
-# factor(oregon_fips$county)
-oregon_fips$county <- gsub(" County", "", as.character(factor(oregon_fips$county)))
-
-oregon_fips <- oregon_fips %>%
-  mutate(st_county_fips = with(oregon_fips, paste0(st_code, fips)))
-
-### Join county with zip
-read_path <- paste0('../data_new/update/or_zip_county_prop.csv') 
-or_zip_county <- read_csv(read_path)
-
-or_zip_county <- or_zip_county %>%
-  select(zip, county_name) %>%
-  rename(county = county_name) %>%
-  rename(ZIP = zip) %>%
-  full_join(oregon_fips, by = "county") %>%
-  select(ZIP, county, state, st_county_fips) %>%
-  rename(fips = st_county_fips)
-
-or_zip_county$county[which(or_zip_county$county=="Hood.River")] <- "Hood River"
-
-# Import population weighted county PM2.5
-pm_path <- paste0("../data_new/county_data/or_county_pop_wt_pm.csv")
-county_pm <- read_csv(pm_path)
-
-# create a working dataset 'or_2013'
-or_ndc_county_time_series <- or_ndc_asthma %>% 
-  # join in the washington county name based on CHARS county identifier
-  left_join(or_zip_county, by = "ZIP") %>% 
-  # rename 'admit_date_impute' to 'date'
-  rename(date = dates) %>%
-  # rename(date = admit_date_impute) %>% 
-  # group by county of residence and date
-  group_by(county, date) %>% 
-  # sum up each primary diagnosis for each outcome for each day for each county
-  summarise(n_obs = n()) %>% 
-  # join daily county estimates of pm2.5
-  left_join(county_pm, by = c('date', 'county'))
-
-# which(is.na(or_ndc_county_time_series$county))
-# or_ndc_county_time_series[113,]
-# which(or_zip_county$ZIP=="97078")
-
-# summary(or_ndc_asthma)
-# summary(or_ndc_county_time_series)
-# summary(as.factor(or_ndc_county_time_series$county)) # 125 is missing out of 3592 (3.16%)
-
-# remove NA values
-# which(is.na(or_ndc_county_time_series$county)) # 3468-3592
-or_ndc_county_time_series <- or_ndc_county_time_series[-c(3468:3592),]
-
-# write permanent dataframe
-
-write_path5 <- paste0("../data_new/medication/or_ndc_county_time_series.csv")
-write_csv(or_ndc_county_time_series, write_path5)
-
-## Check for missing values in sparsely populated counties. set to 0 (But no)
-# check_missing <- filter(or_ndc_county_time_series, is.na(n_obs))
-# mutate each to fill in missing values (Still no)
-# check2 <- check_missing %>% mutate_each(funs(new = ifelse(is.na(.), 0, .)), 
-#                                         n_obs:wrf_temp)
-
-plot_ts <- ggplot(or_ndc_county_time_series, aes(y =n_obs, x = date)) +
+read_path5 <- paste0("../data_new/medication/or_ndc_county_time_series.csv")
+or_ndc_county_ts <- read_csv(read_path5)
+
+plot_ts <- ggplot(or_ndc_county_ts, aes(y =rate_per_100k, x = date)) +
            geom_jitter()
@@
 print(plot_ts)
+
 # save figure
-ggsave("../plot_new/ndc_ts.pdf", plot = plot_ts, 
+ggsave("../plot_new/ndc_rate_ts.pdf", plot = plot_ts, 
        width = 12, height = 8, units = "in")
-```
-
-### County time series
-
-The dotted time series plot of pharmacy visits by counties shows a large and distinct strata dotted time series in some big counties (such as Multnomah), I guess it may be the difference of visits between weekday and weekend.
-The lined time series plot also shows fluctuation.
-So we choose some larger counties (with more visits) to research on the effect of weekend and weekay.
-
-```{r, warning=FALSE, message=FALSE,echo=FALSE}
-
-plot_county <- ggplot(or_ndc_county_time_series, aes(y =n_obs, x = date)) +
+
+plot_county <- ggplot(or_ndc_county_ts, aes(y =rate_per_100k, x = date)) +
                geom_jitter() + facet_wrap(~county)
 
-plot_county_line <- ggplot(or_ndc_county_time_series, aes(y =n_obs, x = date)) +
+plot_county_line <- ggplot(or_ndc_county_ts, aes(y =rate_per_100k, x = date)) +
                     facet_wrap(~county) + geom_line()
@@
 # save figure
-ggsave("../plot_new/ndc_county_ts.pdf", plot = plot_county, 
+ggsave("../plot_new/ndc_county_rate_ts.pdf", plot = plot_county, 
        width = 12, height = 8, units = "in")
-ggsave("../plot_new/ndc_county_ts_line.pdf", plot = plot_county_line, 
+ggsave("../plot_new/ndc_county_rate_ts_line.pdf", plot = plot_county_line, 
        width = 12, height = 8, units = "in")
 
+### check if the time series has the association with get smoke
+a <- glm(rate_per_100k ~ geo_smk_pm10, data = or_ndc_county_ts, family = "poisson")
+summary(a)
+
+plot_health_smk <- ggplot(or_ndc_county_ts, aes(x = geo_smk_pm10, y = log(rate_per_100k+0.1))) +
+  geom_point(aes(colour = c("red")), alpha = 0.2,  size = 1) + 
+  # scale_colour_manual(values = c("red", "blue")) +
+  # truncated y axis
+  #scale_y_continuous(limits = c(0, 1.5)) +
+  # facet_wrap(~county, scale  = "free") +
+  ylab("Log Outcome Rate per 100,000 Persons") +
+  xlab("GWR Smoke PM2.5 per 10*10ug/m^3") +
+  theme_bw()
+
+# unnecessary
+plot_county_health_smk <- ggplot(or_ndc_county_ts, aes(x = geo_smk_pm10, y = log(rate_per_100k+0.1))) +
+  geom_point(aes(colour = county), alpha = 0.2,  size = 1) + 
+  # scale_colour_manual(values = c("red", "blue")) +
+  # truncated y axis
+  #scale_y_continuous(limits = c(0, 1.5)) +
+  facet_wrap(~county, scale  = "free") +
+  ylab("Log Outcome Rate per 100,000 Persons") +
+  xlab("GWR Smoke PM2.5 per 10*10ug/m^3") +
+  theme_bw()
+
+plot(plot_health_smk)
+plot(plot_county_health_smk)
+
+ggsave("../plot_new/ndc_rate_health_smk.pdf", plot = plot_health_smk, 
+       width = 12, height = 8, units = "in")
+
+ggsave("../plot_new/ndc_county_rate_health_smk.pdf", plot = plot_county_health_smk, 
+       width = 12, height = 8, units = "in")
+
 ```
 
-### Time series of weekday and weekend in larger counties.
-We check the time series for Multnomah, and it indeed shows more visits of weekdays than that in weekends (except some holidays such like Labor Day, Independent Day, etc.) We may choose the new covariates for weekday and weekend.
-
-```{r echo = FALSE, message=FALSE, warning=FALSE}
-or_ndc_county_time_series_week <- or_ndc_county_time_series%>%
-  mutate(day_admit = as.factor(weekdays(date)),
-         month_admit = month(date)) %>%
-  mutate(outcome = 1)
-
-# Also check for Multnomah
-multnomah_df <- or_ndc_county_time_series_week %>%
-  filter(county == "Multnomah")
-
-plot_mult_week <- ggplot(multnomah_df, aes(y =n_obs, x = date)) +
-           geom_jitter() + facet_wrap(~day_admit)
-
-print(plot_mult_week)
-
-ggsave("../plot_new/multnomah_ts_week.pdf", plot = plot_mult_week, 
+### Standard Regression Model
+
+Same-day estimates of association between smoke via GWR and outcomes. County treated as a fixed effect covariate.
+
+1. Quasipoisson
+
+```{r regression model q, message=FALSE, echo=FALSE}
+# create empty dataframe
+rr_df <- data.frame(matrix(nrow = 1, ncol = 4))
+colnames(rr_df) <- c("outcome", "rel_risk", "lower_95", "upper_95")
+
+# vector of outcomes
+outcome_list <- "n_obs"
+
+# names 
+outcome_names <- c('Inhalers')
+
+# populate outcome column
+rr_df[,1] <- factor(outcome_names, levels=unique(outcome_names))
+
+smk_est <- tidy(glm(as.formula(paste0(outcome_list, 
+    "~ geo_smk_pm10 + wrf_temp_county + county + 
+    day + offset(log(pop))")), 
+    or_ndc_county_ts, family="quasipoisson"))[2, ]
+
+rr <- c(round(exp(smk_est[1,2]),5),
+        round(exp(smk_est[1,2]-1.96*smk_est[1,3]),5),
+        round(exp(smk_est[1,2]+1.96*smk_est[1,3]),5))
+
+rr_df[1,2:4] <- rr
+
+# tried apply function and it might have been easier to just use a loop
+# as I needed a for loop to fill the list in to the dataframe
+# code is more compact though
+# I should check speed (or make a custom function to do this one day)
+
+rr_df
+
+plot_rr <- ggplot(rr_df, aes(x=outcome, y = rel_risk)) +
+  geom_point() +
+  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
+  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
+  theme(
+   panel.background = element_rect(fill = 'white', colour = 'black'),
+   panel.grid.major = element_blank(),
+   panel.grid.minor = element_blank(),
+   axis.text.x = element_text(angle = 90, vjust = .75)) +
+  ylab("Relative Risk") +
+  xlab("Outcome") +
+  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")
+
+print(plot_rr)
+
+ggsave("../plot_new/ndc_rr.pdf", plot = plot_rr, 
        width = 12, height = 8, units = "in")
+
 ```
 
-### Time series of smoke exposure
-The plot shows the times series of smoke PM 2.5 exposure.
-It seems that at the beginning of August or at the end of July the smoke has a peak.
-Also during the August the smoke is very high.
-The green one shows the WRF-Chem method; The bule one shows the Krigging method; And the red one, which we focus on, shows the GWR method.
-
-```{r echo = FALSE, message=FALSE, warning=FALSE}
-
-plot_smoke_ts_wrf <- ggplot(or_ndc_county_time_series_week, aes(y =wrf_smk_pm, x = date)) +
-           geom_jitter(color="red")
-# print(plot_smoke_ts_wrf)
-
-plot_smoke_ts_krig <- ggplot(or_ndc_county_time_series_week, aes(y =krig_smk_pm, x = date)) +
-           geom_jitter() 
-# print(plot_smoke_ts_krig)
-
-plot_smoke_ts_gwr <- ggplot(or_ndc_county_time_series_week, aes(y =geo_smk_pm, x = date)) +
-           geom_jitter() 
-# print(plot_smoke_ts_gwr)
-
-plot_smk <- ggplot(or_ndc_county_time_series_week, aes(x = date)) +
-  geom_jitter(aes(y =wrf_smk_pm, x = date), color='green', size = 1) + 
-  geom_jitter(aes(y =krig_smk_pm, x = date), color='blue', size = 1) + 
-  geom_jitter(aes(y =geo_smk_pm, x = date), color="red", size = 1)  +
-  ylab("Smoke PM 2.5")
-
-print(plot_smk)
+2. Poisson
+
+```{r regression model,  message=FALSE, echo=FALSE}
+# create empty dataframe
+rr_df <- data.frame(matrix(nrow = 1, ncol = 4))
+colnames(rr_df) <- c("outcome", "rel_risk", "lower_95", "upper_95")
+
+# vector of outcomes
+outcome_list <- "n_obs"
+
+# names 
+outcome_names <- c('Inhalers')
+
+# populate outcome column
+rr_df[,1] <- factor(outcome_names, levels=unique(outcome_names))
+
+smk_est <- tidy(glm(as.formula(paste0(outcome_list, 
+    "~ geo_smk_pm10 + wrf_temp_county + county + 
+    day + offset(log(pop))")), 
+    or_ndc_county_ts, family="poisson"))[2, ]
+
+rr <- c(round(exp(smk_est[1,2]),5),
+        round(exp(smk_est[1,2]-1.96*smk_est[1,3]),5),
+        round(exp(smk_est[1,2]+1.96*smk_est[1,3]),5))
+
+rr_df[1,2:4] <- rr
+
+rr_df
+
+# tried apply function and it might have been easier to just use a loop
+# as I needed a for loop to fill the list in to the dataframe
+# code is more compact though
+# I should check speed (or make a custom function to do this one day)
+
+plot_rr <- ggplot(rr_df, aes(x=outcome, y = rel_risk)) +
+  geom_point() +
+  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
+  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
+  theme(
+   panel.background = element_rect(fill = 'white', colour = 'black'),
+   panel.grid.major = element_blank(),
+   panel.grid.minor = element_blank(),
+   axis.text.x = element_text(angle = 90, vjust = .75)) +
+  ylab("Relative Risk") +
+  xlab("Outcome") +
+  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")
+
+print(plot_rr)
+
+ggsave("../plot_new/ndc_rr_poi.pdf", plot = plot_rr, 
+       width = 12, height = 8, units = "in")
 
@@
 
+### Mixed models
+
+1. Poisson
+
+```{r mixed model,  message=FALSE, echo=FALSE}
+# create empty dataframe
+rr_mix_df <- data.frame(matrix(nrow = 1, ncol = 4))
+colnames(rr_mix_df) <- c("outcome", "rel_risk", "lower_95", "upper_95")
+
+# vector of outcomes
+outcome_list <- "n_obs"
+
+# names 
+outcome_names <- c('Inhalers')
+
+# populate outcome column
+rr_mix_df[,1] <- factor(outcome_names, levels=unique(outcome_names))
+
+smk_est <- tidy(glmer(as.formula(paste0(outcome_list, 
+    "~ geo_smk_pm10 + wrf_temp_county + (1|county) + 
+    day + offset(log(pop))")), 
+    or_ndc_county_ts, family="poisson"))[2, ]
+
+rr <- c(round(exp(smk_est[1,2]),5),
+        round(exp(smk_est[1,2]-1.96*smk_est[1,3]),5),
+        round(exp(smk_est[1,2]+1.96*smk_est[1,3]),5))
+
+rr_mix_df[1,2:4] <- rr
+
+rr_mix_df
+
+plot_mix_rr <- ggplot(rr_mix_df, aes(x=outcome, y = rel_risk)) +
+  geom_point() +
+  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
+  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
+  theme(
+   panel.background = element_rect(fill = 'white', colour = 'black'),
+   panel.grid.major = element_blank(),
+   panel.grid.minor = element_blank(),
+   axis.text.x = element_text(angle = 90, vjust = .75)) +
+  ylab("Relative Risk") +
+  xlab("Outcome") +
+  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")
+
+print(plot_rr)
+
+ggsave("../plot_new/ndc_mix_rr_poi.pdf", plot = plot_mix_rr, 
+       width = 12, height = 8, units = "in")
+
+```
+
 ## Distributed Lag Model
 
+### Lag 1
+
+Remove the missing values for GWR smoke. 
+But I use GWR smoke of zip for the case crossover study and add the county variable.
+
 ```{r DLM, echo = FALSE, message=FALSE, warning=FALSE}
-
-which(colnames(asthma_ndc_df)=="geo_wt_pm_zip") # 88
-which(colnames(asthma_ndc_df)=="geo_wt_pm_lag1_zip") #115
-
-asthma_ndc_lag <- asthma_ndc_df %>%
-  filter(gender!="U") %>%
-  filter(outcome==1) %>%
-  select(88, 115:121) # GWR-method
+read_path8 <- paste0("../data_new/county_data/zip_to_county_new.csv")
+or_zip_to_county <- read_csv(read_path8)
+
+read_path9 <- paste0("../data_new/county_data/or_census_pop_est_county.csv")
+or_pop <- read_csv(read_path9)
+
+or_pop <- or_pop %>%
+  rename(county = GEO_display_label, pop = respop72013) %>%
+  select(county, pop) 
+or_pop$county <- gsub(" County, Oregon", "", as.character(factor(or_pop$county)))
+
+inhaler_ndc_lag_df <- inhaler_ndc_df %>%
+  left_join(or_zip_to_county, by = "ZIPCODE") %>%
+  left_join(or_pop, by = "county") %>%
+  mutate(day = ifelse(weekdays(date) == "Monday" |
+                      weekdays(date) == "Tuesday" |
+                      weekdays(date) == "Wednesday"|
+                      weekdays(date) == "Thursday"|
+                      weekdays(date) == "Friday", "weekday",
+               ifelse(weekdays(date) == "Saturday" |
+                      weekdays(date) == "Sunday", "weekend", NA))) %>%
+  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
+         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
+         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
+         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
+         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
+         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
+         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
+         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10) %>%
+  filter(!is.na(geo_smk_pm10_lag7_zip))
+
+inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-05-27")] <- "weekend"
+inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-07-04")] <- "weekend"
+inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-09-02")] <- "weekend"
+
+which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_zip") # 159
+which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_lag1_zip") # 160
+
+inhaler_ndc_lag <- inhaler_ndc_lag_df %>%
+  select(159, 160:166) # GWR-method
 
 lag_max <- 7
-n <- length(asthma_ndc_lag[,1])
-
-asthma_ndc_lag <- as.matrix(asthma_ndc_lag)
-
-# true DL function
-DLfunc <- exp(-(0:lag_max)/2)
-
-# simulate outcomes
-y <- drop(asthma_ndc_lag%*%DLfunc) # + rnorm(n)
-
-B <- ns(0:lag_max, df =4, intercept=TRUE)
+n <- length(inhaler_ndc_lag[,1])
+
+inhaler_ndc_lag <- as.matrix(inhaler_ndc_lag)
+
+B <- ns(0:lag_max, df =3, intercept=TRUE) # or 3?
 
@@
 dim(B)
-dim(asthma_ndc_lag)
-
-asthma_ndc_lag_B <- asthma_ndc_lag%*%B
-fit <- lm(y~asthma_ndc_lag_B)
+dim(inhaler_ndc_lag)
+
+inhaler_ndc_lag_B <- inhaler_ndc_lag%*%B
+fit <- clogit(outcome ~ inhaler_ndc_lag_B + wrf_temp_zip + strata(personkey), inhaler_ndc_lag_df)
+
 coef(fit)
-
-dlparms <- grep("asthma_ndc_lag_B",names(coef(fit)))
+summary(fit)
+
+dlparms <- grep("inhaler_ndc_lag_B",names(coef(fit)))
 DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms])  
@@
 DLestimate$lag <- 0:lag_max
-DLestimate$truth <- DLfunc
 p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
 p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
-p <- p + geom_line(aes(x=lag, y=truth), linetype=2, size=2)
+p <- p + geom_line(aes(x=lag), linetype=2, size=2)
 p <- p + theme_bw()
-p
+# p
+
+DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms])
 
@@
 cumulative
-## [1] 2.485711
+## [1] 0.009141771
 cumulative_se <- sqrt(sum(DLvar))
 cumulative_se
-## [1] 0.0001887442
-cumulative_CI <- cumulative + cumulative_se * qt(c(1-0.975,0.975), df=fit$df.residual)
+## [1] 0.000955469
+cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
 cumulative_CI
-## [1] 2.485341 2.486080
-
+## [1] 0.007269052 0.011014490
+
+## mixed model
+mod <-glmer(outcome ~  inhaler_ndc_lag_B + wrf_temp_zip + day + (1|county) + offset(log(pop)),   
+            inhaler_ndc_lag_df, family="poisson", control = glmerControl(optimizer = "bobyqa"))
+
+summary(mod)
+# check AIC
+AIC(mod)
+# output distributed lag beta parameters
+dlparms <- mod@beta[2:4]
+# estimate distributed lag values for each day
+dl_estimates <- data.frame(estimate = B %*% dlparms)
+
+# covariance matrix for knots (need to convert to matrix object)
+cov_mat <- as.matrix(vcov(mod)[2:4,2:4])
+
+# estimate variance of splines
+dl_var <- B%*%cov_mat%*%t(B)
+
+# calculate standard error for each lag value
+dl_estimates$stderr <- sqrt(diag(dl_var))
+
+# calculate lower and upper bounds
+dl_estimates$lower_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(1-0.975,
+  df=df.residual(mod))  
+dl_estimates$upper_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(0.975,
+  df=df.residual(mod))
+
+# estimate relative risk and 95%CI from beta
+inhaler_dl_rr <- dl_estimates %>% 
+  mutate(estimate = exp(estimate),
+         lower_bound = exp(lower_bound),
+         upper_bound = exp(upper_bound),
+         lag = 0:7)
+
+# estimate cumulative effect
+cumulative <- sum(dl_estimates$estimate)
+cumulative
+
+# estimate cumulative effect stnd error
+cumulative_se <- sqrt(sum(dl_var))
+cumulative_se
+
+# estimate cumulative CI
+cumulative_ci <- cumulative+cumulative_se*qt(c(1-0.975,0.975), 
+                                             df=df.residual(mod))
+cumulative_ci
+
+exp(cumulative)
+exp(cumulative_ci)
+
+# plot 
+p <- ggplot(inhaler_dl_rr, aes(x=lag, y=estimate, 
+                              ymin=lower_bound, ymax=upper_bound)) +
+  geom_ribbon(alpha = 0.5) +
+  geom_line(size=2)
+
+p
 ```
 
-```{r overall lag1, warning =F, echo = F, results='asis'} 
+### Find the smallest of AIC with df from 3 to 6
+
+```{r, message=FALSE, echo=FALSE}
+# create fit matrix
+fit_mat <- matrix(NA, nrow=4, ncol=6)
+colnames(fit_mat) <- c("outcome", "exposure", "fit_mix", "df", "aic", "fit_fix")
+
+# fill outcome name in first column
+fit_mat[,1] <- "inhalers"
+# fill in exposure (geo smoke)
+fit_mat[,2] <- "geo_smk10"
+# fit type
+fit_mat[,3] <- "ns"
+
+# 2 df will not work for a spline, going from 3 to 6 df
+  for(h in 3:6){ # start loop for df
+
+    # fitting a natural spline for lag with 5 knots?
+    fit_mat[h-2, 4] <- degree_freedom <- h
+
+    # define basis b using natural spline function
+    b <- ns(0:7, df = degree_freedom, intercept = T)
+
+    # multiply lagged pm matrix by basis
+    pm_b <- inhaler_ndc_lag %*% b
+
+    # fit mixed model
+    mod <-glmer(outcome ~  pm_b + (1|county) + offset(log(pop)), inhaler_ndc_lag_df, family="poisson", control = glmerControl(optimizer = "bobyqa"))
+
+    # fixed mod
+    fixed_mod <- glm(outcome ~ pm_b + county + offset(log(pop)), inhaler_ndc_lag_df, 
+                     family = "poisson")
+
+    AIC(mod)
+    AIC(fixed_mod)
+    # fill AIC
+    fit_mat[h-2,5] <- round(AIC(mod),5)
+    fit_mat[h-2,6] <- round(AIC(fixed_mod),5)
+
+    } # end model fit loop
+
+# now that we have a range of df/knots, I want to filter to the minimum aic
+# for each outcome and print out that value in a table
+
+fit_mat
+
+lag_spline_best_fit <- as_tibble(fit_mat) %>%  
+  slice(which.min(aic)) # df of 3 has the smallest AIC
+
+# kable
+knitr::kable(lag_spline_best_fit, caption = paste0("Distributed lag spline",
+  "degree of freedom best fit by AIC for mixed model approach"))
+```
+
+### Compare fixed and mixed models of df 3 and 4
+
+```{r compare mixed and fixed, message=FALSE, echo=FALSE}
+
+## degree of 3 or 4, mixed or fixed
+degree_freedom <- 3
+
+# define basis b using natural spline function
+b <- ns(0:7, df = degree_freedom, intercept = T)
+
+# multiply lagged pm matrix by basis
+pm_b <- inhaler_ndc_lag %*% b
+
+mod <-glmer(outcome ~  pm_b + (1|county) + wrf_temp_zip + offset(log(pop)), inhaler_ndc_lag_df, family="poisson",
+            control = glmerControl(optimizer = "bobyqa"))
+
+dlparms <- mod@beta[2:(1+degree_freedom)]
+# estimate distributed lag values for each day
+dl_estimates <- data.frame(estimate = b %*% dlparms)
+
+# covariance matrix for knots (need to convert to matrix object)
+cov_mat <- as.matrix(vcov(mod)[2:(1+degree_freedom),2:(1+degree_freedom)])
+
+# estimate variance of splines
+dl_var <- b%*%cov_mat%*%t(b)
+
+# calculate standard error for each lag value
+dl_estimates$stderr <- sqrt(diag(dl_var))
+
+# calculate lower and upper bounds
+dl_estimates$lower_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(1-0.975,
+  df=df.residual(mod))  
+dl_estimates$upper_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(0.975,
+  df=df.residual(mod))
+
+# estimate relative risk and 95%CI from beta
+inhaler_dl_rr <- dl_estimates %>% 
+  mutate(estimate = exp(estimate),
+         lower_bound = exp(lower_bound),
+         upper_bound = exp(upper_bound),
+         lag = 0:7)
+
+# estimate cumulative effect
+cumulative <- sum(dl_estimates$estimate)
+cumulative
+
+# estimate cumulative effect stnd error
+cumulative_se <- sqrt(sum(dl_var))
+cumulative_se
+
+# estimate cumulative CI
+cumulative_ci <- cumulative+cumulative_se*qt(c(1-0.975,0.975), 
+                                             df=df.residual(mod))
+cumulative_ci
+
+exp(cumulative)
+exp(cumulative_ci)
+
+# plot 
+p3_mix <- ggplot(inhaler_dl_rr, aes(x=lag, y=estimate, 
+                              ymin=lower_bound, ymax=upper_bound)) +
+  geom_ribbon(alpha = 0.5) +
+  geom_line(size=2)
+
+fixed_mod <- glm(outcome ~ pm_b + county + wrf_temp_zip + offset(log(pop)), inhaler_ndc_lag_df, 
+                     family = "poisson")
+
+dlparms <- fixed_mod$coefficients[2:4]
+# estimate distributed lag values for each day
+dl_estimates <- data.frame(estimate = b %*% dlparms)
+
+# covariance matrix for knots (need to convert to matrix object)
+cov_mat <- as.matrix(vcov(fixed_mod)[2:4,2:4])
+
+# estimate variance of splines
+dl_var <- b%*%cov_mat%*%t(b)
+
+# calculate standard error for each lag value
+dl_estimates$stderr <- sqrt(diag(dl_var))
+
+# calculate lower and upper bounds
+dl_estimates$lower_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(1-0.975,
+  df=df.residual(fixed_mod))  
+dl_estimates$upper_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(0.975,
+  df=df.residual(fixed_mod))
+
+# estimate relative risk and 95%CI from beta
+inhaler_dl_rr <- dl_estimates %>% 
+  mutate(estimate = exp(estimate),
+         lower_bound = exp(lower_bound),
+         upper_bound = exp(upper_bound),
+         lag = 0:7)
+
+# estimate cumulative effect
+cumulative <- sum(dl_estimates$estimate)
+cumulative
+
+# estimate cumulative effect stnd error
+cumulative_se <- sqrt(sum(dl_var))
+cumulative_se
+
+# estimate cumulative CI
+cumulative_ci <- cumulative+cumulative_se*qt(c(1-0.975,0.975), 
+                                             df=df.residual(fixed_mod))
+cumulative_ci
+
+exp(cumulative)
+exp(cumulative_ci)
+
+# plot 
+p3_fix <- ggplot(inhaler_dl_rr, aes(x=lag, y=estimate, 
+                              ymin=lower_bound, ymax=upper_bound)) +
+  geom_ribbon(alpha = 0.5) +
+  geom_line(size=2)
+
+p3_fix
+p3_mix
+
+## df = 4
+degree_freedom <- 4
+
+# define basis b using natural spline function
+b <- ns(0:7, df = degree_freedom, intercept = T)
+
+# multiply lagged pm matrix by basis
+pm_b <- inhaler_ndc_lag %*% b
+
+mod <-glmer(outcome ~  pm_b + (1|county) + wrf_temp_zip + offset(log(pop)), inhaler_ndc_lag_df, family="poisson",
+            control = glmerControl(optimizer = "bobyqa"))
+
+dlparms <- mod@beta[2:5]
+# estimate distributed lag values for each day
+dl_estimates <- data.frame(estimate = b %*% dlparms)
+
+# covariance matrix for knots (need to convert to matrix object)
+cov_mat <- as.matrix(vcov(mod)[2:5,2:5])
+
+# estimate variance of splines
+dl_var <- b%*%cov_mat%*%t(b)
+
+# calculate standard error for each lag value
+dl_estimates$stderr <- sqrt(diag(dl_var))
+
+# calculate lower and upper bounds
+dl_estimates$lower_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(1-0.975,
+  df=df.residual(mod))  
+dl_estimates$upper_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(0.975,
+  df=df.residual(mod))
+
+# estimate relative risk and 95%CI from beta
+inhaler_dl_rr <- dl_estimates %>% 
+  mutate(estimate = exp(estimate),
+         lower_bound = exp(lower_bound),
+         upper_bound = exp(upper_bound),
+         lag = 0:7)
+
+# estimate cumulative effect
+cumulative <- sum(dl_estimates$estimate)
+cumulative
+
+# estimate cumulative effect stnd error
+cumulative_se <- sqrt(sum(dl_var))
+cumulative_se
+
+# estimate cumulative CI
+cumulative_ci <- cumulative+cumulative_se*qt(c(1-0.975,0.975), 
+                                             df=df.residual(mod))
+cumulative_ci
+
+exp(cumulative)
+exp(cumulative_ci)
+
+# plot 
+p4_mix <- ggplot(inhaler_dl_rr, aes(x=lag, y=estimate, 
+                              ymin=lower_bound, ymax=upper_bound)) +
+  geom_ribbon(alpha = 0.5) +
+  geom_line(size=2)
+
+fixed_mod <- glm(outcome ~ pm_b + county + wrf_temp_zip + offset(log(pop)), inhaler_ndc_lag_df, 
+                     family = "poisson")
+
+dlparms <- fixed_mod$coefficients[2:5]
+# estimate distributed lag values for each day
+dl_estimates <- data.frame(estimate = b %*% dlparms)
+
+# covariance matrix for knots (need to convert to matrix object)
+cov_mat <- as.matrix(vcov(fixed_mod)[2:5,2:5])
+
+# estimate variance of splines
+dl_var <- b%*%cov_mat%*%t(b)
+
+# calculate standard error for each lag value
+dl_estimates$stderr <- sqrt(diag(dl_var))
+
+# calculate lower and upper bounds
+dl_estimates$lower_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(1-0.975,
+  df=df.residual(fixed_mod))  
+dl_estimates$upper_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(0.975,
+  df=df.residual(fixed_mod))
+
+# estimate relative risk and 95%CI from beta
+inhaler_dl_rr <- dl_estimates %>% 
+  mutate(estimate = exp(estimate),
+         lower_bound = exp(lower_bound),
+         upper_bound = exp(upper_bound),
+         lag = 0:7)
+
+# estimate cumulative effect
+cumulative <- sum(dl_estimates$estimate)
+cumulative
+
+# estimate cumulative effect stnd error
+cumulative_se <- sqrt(sum(dl_var))
+cumulative_se
+
+# estimate cumulative CI
+cumulative_ci <- cumulative+cumulative_se*qt(c(1-0.975,0.975), 
+                                             df=df.residual(fixed_mod))
+cumulative_ci
+
+exp(cumulative)
+exp(cumulative_ci)
+
+# plot 
+p4_fix <- ggplot(inhaler_dl_rr, aes(x=lag, y=estimate, 
+                              ymin=lower_bound, ymax=upper_bound)) +
+  geom_ribbon(alpha = 0.5) +
+  geom_line(size=2)
+
+p4_fix
+p4_mix
+
+ggsave("../plot_new/ndc_p3_mix.pdf", plot = p3_mix, 
+       width = 12, height = 8, units = "in")
+ggsave("../plot_new/ndc_p4_mix.pdf", plot = p4_mix, 
+       width = 12, height = 8, units = "in")
+
+ggsave("../plot_new/ndc_p3_fix.pdf", plot = p3_fix, 
+       width = 12, height = 8, units = "in")
+ggsave("../plot_new/ndc_p4_fix.pdf", plot = p4_fix, 
+       width = 12, height = 8, units = "in")
+
+```
+
+Note I get some warnings (including max grade convergence issues), slightly different answers in best fit, and it takes about 18 minutes to run. Consider running in parallel.
+I suspect the convergence issues are likely due to the same state/counties with small sample size and few outcome observations and lack of variability between exposure/outcome.
+Going to fit what I have now, take a walk, and see what comes out. First thing to do is find cumulative effect, then lagged day effects.
+
+```{r, message=FALSE, echo=FALSE, include=FALSE, eval=FALSE}
+# create a dataframe that contains cumulative estimates for each outcome
+cumulative_df <- as_tibble(matrix(NA, nrow=0, ncol=4))
+colnames(cumulative_df) <- c("outcome", "rel_risk", "lower_95", "upper_95")
+
+# create a dataframe that contains the distributed lag estimates for each outcome
+dist_lag_df <- as_tibble(matrix(NA,  nrow = 0, ncol = 5))
+colnames(dist_lag_df) <- c("outcome", "lag", "rel_risk", 
+                                  "lower_95", "upper_95")
+
+# loop through best fit df
+
+    # create empty matrix two empty matrices to fill
+    cumul_val <- as_tibble(matrix(NA, nrow=1, ncol=4))
+    colnames(cumul_val) <- c("outcome", "rel_risk", "lower_95", "upper_95")
+
+    dl_val <- as_tibble(matrix(NA, nrow=8, ncol=5))
+    colnames(dl_val) <- c("outcome", "lag", "rel_risk", 
+                                  "lower_95", "upper_95")
+
+    # easiest to set out the parameters from the start
+    outcome <- "inhalers"
+    degree_freedom <- as.numeric(lag_spline_best_fit[4])
+
+    # define basis b using natural spline function
+    b <- ns(0:7, df = degree_freedom, intercept = T)
+
+    # multiply lagged pm matrix by basis
+    pm_b <- inhaler_ndc_lag %*% b
+
+    # fit mixed model
+     mod <-glmer(outcome ~  inhaler_ndc_lag_B + wrf_temp_zip + day + (1|county) + offset(log(pop)), inhaler_ndc_lag_df, family="poisson", control = glmerControl(optimizer = "bobyqa"))
+
+    mod
+    # output distributed lag beta parameters; this will change based on DF
+    dlparms <- mod@beta[2:(1+degree_freedom)]
+    # estimate distributed lag values for each day
+    dl_estimates <- data.frame(estimate = b %*% dlparms)
+
+    # covariance matrix for knots (need to convert to matrix object)
+    cov_mat <- as.matrix(vcov(mod)[2:(1+degree_freedom),2:(1+degree_freedom)]) 
+    # estimate variance of splines
+    dl_var <- b%*%cov_mat%*%t(b)
+    # calculate standard error for each lag value
+    dl_estimates$stderr <- sqrt(diag(dl_var))
+    # calculate lower and upper bounds
+    dl_estimates$lower_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(1-0.975,
+      df=df.residual(mod))  
+    dl_estimates$upper_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(0.975,
+      df=df.residual(mod))
+
+    # estimate cumulative effect ----
+    cumulative <- sum(dl_estimates$estimate)
+    # estimate cumulative effect stnd error
+    cumulative_se <- sqrt(sum(dl_var))
+    # estimate cumulative CI
+    cumulative_ci <- cumulative+cumulative_se*qt(c(1-0.975,0.975), 
+                                                 df=df.residual(mod))
+    # fill tibbles and bind in to larger dataframes ----
+    # outcome
+    cumul_val[1,1] <- outcome
+    # cumulative rel risk
+    cumul_val[1,2] <- exp(cumulative)
+    # cumulative 95% CI
+    cumul_val[1,3:4] <- exp(cumulative_ci)
+    # bind to cumulative dataframe
+    cumulative_df <- rbind(cumulative_df, cumul_val)
+
+    # lag vals (this step is redundant)
+    dl_val[,1] <- outcome
+    dl_val[,2] <- 0:7
+    dl_val[,3] <- exp(dl_estimates[,1])
+    dl_val[,4] <- exp(dl_estimates[,3])
+    dl_val[,5] <- exp(dl_estimates[,4])
+
+    # bind in with overall df
+    dist_lag_df <- rbind(dist_lag_df, dl_val)
+
+# print table of cumulative effects
+knitr::kable(cumulative_df, caption = paste0("Cumulative Effect for a 10 ug/m^3 increase in GWR smoke lagged days 0-7 mixed model approach"))
+
+# plot distributed lag effects
+ggplot(dist_lag_df, aes(x=lag, y=rel_risk)) +
+  geom_line(colour = "blue") +
+  geom_ribbon(aes(ymin = lower_95, ymax = upper_95), fill = "blue", alpha = 0.3) + 
+  scale_y_continuous(limits = c(0.98, 1.06)) +
+  scale_x_continuous(breaks = c(seq(0, 7,by=1))) +
+  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
+  facet_wrap(~ outcome) +
+  ggtitle(paste0("Distributed lag association between a 10 ug/m^3",
+                 " increase in GWR smoke PM2.5 and cardiopulmonary outcomes")) +
+  ylab("Relative Risk for a 10 ug/m^3 increase in GWR smoke (fixed)") +
+  xlab("Lagged Days") +
+  theme_bw()
+
+```
+
+## Distributed Lag Casecrossover Analysis
+
+### Lag 1
+
+### Overall plot
+
+```{r overall, warning =F, echo = F, results='asis'} 
 # dataframe list
-
-method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+method_list <- c('Geo-Weighted Smoke')
 
 # dataframe to loop through
-# asthma_ndc_df <- data.frame(asthma_ndc)
+inhaler_ndc_lag1_df <- data.frame(inhaler_ndc) %>%
+  filter(!is.na(geo_smk_pm_lag1_zip))
 
@@
 # extract covariates from dataframe
-covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]
+covariates_df <- inhaler_ndc_lag1_df[, c(1:26, 74:79, 94:156)]
 
 # extract pm values and divide by 10 and ordered
-which(colnames(asthma_ndc_df)=="wrf_smk_pm_lag1_zip") # 108
-which(colnames(asthma_ndc_df)=="krig_pm_lag1_zip") # 122
-which(colnames(asthma_ndc_df)=="geo_wt_pm_lag1_zip") # code to find column numbers, 115
-pm_estimates_df <- asthma_ndc_df[, c(108, 122, 115, 93)]/10  # create 10 unit increases
+which(colnames(inhaler_ndc_lag1_df)=="geo_smk_pm_lag1_zip") # 136
+pm_estimates_df <- inhaler_ndc_lag1_df[, 136]/10  # create 10 unit increases
 
@@
 # empty matrix (12 x 10 matrix)
-point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)
+point_estimates <- matrix(nrow = 1, ncol = 9, byrow = T)
 
@@
   # remove missing pm values
-  filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
-  filter(!is.na(krig_pm_lag1_zip)) %>% 
-  filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
+  # filter(!is.na(wrf_smk_pm_zip)) %>% 
   # the following code makes sure that the counterfactual values retained are 
@@
 # second loop to run a model for each pm estimation method
-for(j in 38:40){
 
   # variable to model 
-  var_name <- colnames(df_analysis[j])
+  var_name <- colnames(df_analysis[75])
 
   # set row number to fill
-  row_n <- j-37
+  row_n <- 1
 
@@
   # conditional logistic regression model
-  mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+  mod <- clogit(outcome ~ df_analysis[[75]] + wrf_temp_lag1_zip + strata(personkey), df_analysis)
 
@@
           point_estimates[row_n, c(5:9)] <- 99 } # end 'if else' statement
-  } # end methods loop
 
@@
   # save figure
-  ggsave("../plot_lag/ndc_overall_lag1_three.pdf", plot = print_plot, 
+  ggsave("../plot_new/ndc_overall_three.pdf", plot = print_plot, 
        width = 12, height = 8, units = "in")
@@
 
-```{r time stratified age strata lag 1, warning =F, echo = F, results='asis'} 
+### Time-Stratified by Age Categories
+
+Following table and figure look at some outcomes stratified by age category.
+
+For some outcomes, such like Age > 65 in WRF-Chem method, the CI is the only insignificant one in the plot.
+
+```{r time stratified age strata cdc met, warning =F, echo = F, results='asis'} 
 # dataframe list
@@
 #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
-pm_estimates_df <- asthma_ndc_df[, c(108, 122, 115, 93)]/10  # create 10 unit increases
+pm_estimates_df <- inhaler_ndc_df[, c(87, 92, 91, 93)]/10  # create 10 unit increases
 
@@
 
+  # dataframe for analysis creation
+  # bind columns back together 
   df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
     # remove missing pm values
-    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
-    filter(!is.na(krig_pm_lag1_zip)) %>% 
-    filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
+    filter(!is.na(wrf_smk_pm_zip)) %>% 
     # limit to specific age category
@@
   # save figure
-  ggsave("../plot_lag/ndc_age_fig_three_lag1.pdf", plot = print_plot, 
+  ggsave("../plot_new/ndc_age_fig_three.pdf", plot = print_plot, 
        width = 12, height = 8, units = "in")
@@
 
-```{r sex strata cdc met  lag 1, warning = F, echo = F, results='asis'} 
+### Seasonal Time-Stratified by Sex
+
+Association with smoke PM~2.5~ and health outcomes stratified by sex.
+
+All the CI do not include 1, which means the use of asthma medication (beta 2 agonists) increases about 5% when smoke increase 10 units.
+So the risk for getting asthma will increase about 5% ~ 8% when the smoke PM2.5 increase 10 units.
+
+```{r sex strata cdc met adj, warning = F, echo = F, results='asis'} 
 method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
@@
     # remove missing pm values
-    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
-    filter(!is.na(krig_pm_lag1_zip)) %>% 
-    filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
-    # limit to specific age category
+    filter(!is.na(wrf_smk_pm_zip)) %>% 
+    # limit to specific sex category
     filter(sex_ind == k) %>% 
@@
   # save figure
-  ggsave("../plot_lag/ndc_sex_fig_three_lag1.pdf", plot = print_plot, 
+  ggsave("../plot_new/ndc_sex_fig_three.pdf", plot = print_plot, 
        width = 12, height = 8, units = "in")
@@
 ```
-
-## NDC claims with index
-
-Check the claims in each category (Asthma, COPD, Both or Other).
-
-```{r echo = FALSE, message=FALSE, warning=FALSE}
-read_path7 <- paste0("../data_new/medication/oregon_beta_ndc_index.csv")
-or_ndc_index <- read_csv(read_path7)
-
-oregon_beta_df <- or_ndc_index %>% 
-  group_by(personkey) %>% 
-  arrange(personkey, fromdate, line) %>% 
-  mutate(num_visit = dense_rank(fromdate)) %>%
-  arrange(personkey, num_visit) %>% 
-  select(personkey, clmid, num_visit) %>% 
-  unique() %>% 
-  full_join(or_ndc_index, by = c("personkey", "clmid")) %>% 
-  arrange(personkey, clmid, line, fromdate) %>% 
-  filter(row_number() == 1) %>%
-  # add new transverted from date
-  mutate(dates = as.Date(fromdate, "%m/%d/%Y")) %>%
-  filter(dates >= '2013-05-01' & 
-         dates <= '2013-09-30') # 24538
-
-summary(as.factor(oregon_beta_df$index))
-
-```
-
-### Descriptives
-
-For all of 2013, there were a total of millons of all claims records in the APAC dataset.
-During the May 1st to September 30th wildfire season in Oregon state, there were 24217 pharmacy visits for the ndc of beta 2 agonists.
-The number of cases, as well as age-specific and sex-specific strate are shown below.
-
-```{r ndc descriptive table, echo = F, message = F, warning = F, results='asis'}
-read_path6 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover.csv")
-beta_ndc <- read_csv(read_path6)
-
-# join personkey with index
-oregon_beta_index_df <- oregon_beta_df %>%
-  select(personkey, index)
-
-beta_ndc_with_index <- beta_ndc %>%
-  left_join(oregon_beta_index_df, by="personkey")
-
-# dataframe to loop through
-beta_ndc_with_index_df <- data.frame(beta_ndc_with_index)
-
-beta_asthma <- beta_ndc_with_index_df %>%
-  filter(index=="asthma")
-
-beta_copd <- beta_ndc_with_index_df %>%
-  filter(index=="copd")
-
-beta_both <- beta_ndc_with_index_df %>%
-  filter(index=="both")
-
-beta_other <- beta_ndc_with_index_df %>%
-  filter(index=="other")
-
-df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)
-
-outcome_list <- c("Asthma", "COPD", "Both", "Other")
-
-# create an empty list to row bind dataframes together
-datalist <- list()
-
-# dataframe list
-method_list <- c('WRF-Chem Smoke', 'Geo-Weighted Smoke', 'Kriging Smoke')
-
-# outcome name
-outcome_name <- c("Asthma", "COPD", "Both", "Other")
-
-for(i in 1:length(df_list)){
-  # dataframe to loop through
-  df_to_loop <- data.frame(df_list[i])
-  # indication of column
-  outcome <- colnames(df_to_loop[75])
-  # outcome name
-  outcome_name <- outcome_list[i]
-
-  # dataframe for analysis creation
-  # bind columns back together 
-  df_analysis <- df_to_loop %>%
-    filter(!is.na(wrf_smk_pm_zip)) %>% 
-    # only look at outcomes
-    filter(outcome == 1) %>%
-    # add another row that makes sure there is a person <15 in the dataframe
-    # tricking xtabs to produce a 0 cell for the outcome for age <15
-    add_row(outcome = 0, age_ind = 0)
-
-  # cross tabs
-  outcome_n <- xtabs(~ outcome, df_analysis)
-  cross_tab_age <- xtabs(~ outcome + age_ind, df_analysis)
-  cross_tab_sex <- xtabs(~ outcome + sex_ind, df_analysis)
-  # empty matrix
-  point_estimates <- matrix(nrow = 1, ncol = 7, byrow = T)
-
-  colnames(point_estimates) <- c("outcome", "n", "age_15", "age_15_65", 
-                                 "age_65", "female", "male")
-
-  # fill in the outcome name for the dataframe before the loop
-  point_estimates[, 1] <- outcome_name
-  # fill n
-  point_estimates[, 2] <- outcome_n[2] # second element of the 1 dimension vector
-  # age <15
-  point_estimates[, 3] <- cross_tab_age[2, 1]
-  # age 15 to 65
-  point_estimates[, 4] <- cross_tab_age[2, 2]
-  # age >65
-  point_estimates[, 5] <- cross_tab_age[2, 3]
-  # male == 0
-  point_estimates[, 7] <- cross_tab_sex[1, 1]
-  # female == 1
-  point_estimates[, 6] <- cross_tab_sex[1, 2]
-
-  # save point estimates as a dataframe
-  point_est_df <- as_data_frame(point_estimates)
-
-  # combine previous values in dataframe that has all outcome/methods comparisons
-  datalist[[i]] <- point_est_df
-
-} # end of loop
-
-# combine each outcome dataframe itteration in to a big dataset
-combined_point_est_df <- bind_rows(datalist) %>% 
-  # find proportions/percents for each strata in a row
-  mutate(age_15_pr = as.character(round((as.numeric(age_15)/as.numeric(n))*100,1)),
-         age_15_65_pr = as.character(round((as.numeric(age_15_65)/as.numeric(n))*100,1)),
-         age_65_pr = as.character(round((as.numeric(age_65)/as.numeric(n))*100,1)),
-         female_pr = as.character(round((as.numeric(female)/as.numeric(n))*100,1)),
-         male_pr = as.character(round((as.numeric(male)/as.numeric(n))*100,1))) %>% 
-  select(outcome, n, age_15, age_15_pr, age_15_65, age_15_65_pr, age_65,
-         age_65_pr, female, female_pr, male, male_pr)
-
-#str(combined_point_est_df)
-
-tab <- htmlTable(txtRound(combined_point_est_df, digits = 1), 
-           caption = "Number of cases for each outcome observed from May 1st to September 30st, 2013",
-           # column headers
-           header = c("Outcome", "Cases n", "Less than 15", "(%)", "15 to 65", "(%)", "Greater than 65",
-                      "(%)", "Female", "(%)", "Male", "(%)"),
-           # column spanner
-           cgroup = c("","Age Category", "Sex"), 
-           n.cgroup = c(2, 6, 4),
-           padding.rgroup = "&nbsp;&nbsp;",
-           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
-           align = "llccccc" # column alignment,
-            ) # end table
-
-print(tab)
-
-```
-
-### Overall plot
-
-```{r ndc overall try, warning =F, echo = F, results='asis'}
-df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)
-
-outcome_list <- c("Asthma", "COPD", "Both", "Other")
-
-# create an empty list to row bind dataframes together
-datalist <- list()
-
-# dataframe list
-method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
-
-# outcome name
-outcome_name <- c("Asthma", "COPD", "Both", "Other")
-
-# data wrangling ----
-# Producing conditional logit model estimates loop 
-for(i in 1:length(df_list)){
-  # dataframe to loop through
-  df_to_loop <- data.frame(df_list[i])
-  # indication of column
-  outcome <- colnames(df_to_loop[75])
-  # outcome name
-  outcome_name <- outcome_list[i]
-
-  # extract covariates from dataframe
-  covariates_df <- df_to_loop[, c(1:26, 74:84)]
-
-  # extract pm values and divide by 10 and ordered
-  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
-  pm_estimates_df <- df_to_loop[, c(87, 92, 91, 93)]/10  # create 10 unit increases
-
-  # dataframe for analysis creation
-  # bind columns back together 
-  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
-    # remove missing pm values
-    filter(!is.na(wrf_smk_pm_zip)) %>%
-
-    # join in cdc met data
-    # full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
-    # the following code makes sure that the counterfactual values retained are 
-    # symetric in that number of obs before = number of obs after
-    mutate(obs_diff_admission = (fromdate - date)/7) 
-  # dataframe is already for the entire fire season, so I don't need to subset anymore
-
-  # empty df for table
-  table_df <- data.frame()
-
-  # empty matrix
-  point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)
-
-  colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 'odds_ratio', 
-                                 'lower95', 'upper95', 'se', 'p_val')
-
-  # fill in the outcome name for the dataframe before the loop
-  point_estimates[, 1] <- outcome_name
-
-  # second loop to run a model for each pm estimation method
-
-  for(j in 38:40){
-
-    # variable to model 
-    var_name <- colnames(df_analysis[j])
-
-    # set row number to fill
-    row_n <- j-37
-
-    # only run the model if the dataframe has observations
-    if(nrow(df_analysis) != 0){
-      # conditional logistic regression model
-      mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
-
-      # populate matrix
-      point_estimates[row_n, 2] <- method_list[row_n]
-      point_estimates[row_n, 3] <- mod$n
-      point_estimates[row_n, 4] <- mod$nevent
-      # odds ratio
-      point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)
-
-      # 95% lower bound
-      point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
-      # 95% upper bound
-      point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
-                                               1.96*(summary(mod)$coefficient[1,3])), 3)
-      # standard error
-      point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
-      # p val
-      point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)
-
-      # save point estimates as a dataframe
-      point_est_df <- as_data_frame(point_estimates)
-
-      # create else statement that fills matrix with missing so I still have the row
-      # in the final dataframe
-    } else {point_estimates[row_n, 2] <- method_list[row_n]
-    point_estimates[row_n, 3] <- 0
-    point_estimates[row_n, 4] <- 0
-    point_estimates[row_n, c(5:9)] <- 99 } # end 'if else' statement
-  } # end methods loop
-
-  # combine previous values in dataframe that has all outcome/methods comparisons
-  datalist[[i]] <- point_est_df
-
-} # end of loop
-
-# combine each outcome dataframe itteration in to a big dataset
-combined_point_est_df <- bind_rows(datalist)
-
-# subset columns I want to put in to the table
-table_df <- combined_point_est_df %>% select(2, 3:7)
-
-tab <- htmlTable(txtRound(table_df, digits = 3, 1:3), 
-                 caption = "Association between a 10 ug/m^3 in PM2.5 and Health Outcomes",
-                 # row group by outcome
-                 rgroup = outcome_list,
-                 n.rgroup = c(rep(3, 4)), # 4 rows for each method for each outcome
-                 # column headers
-                 header = c("Method", "Obs.", "Events",
-                            "OR&dagger;", "Lower", "Upper"),
-                 # column spanner
-                 cgroup = c("", "95% CI"), 
-                 n.cgroup = c(4, 2),
-                 padding.rgroup = "&nbsp;&nbsp;",
-                 css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
-                 align = "llccccc", # column alignment,
-                 tfoot="&dagger; Adjusted for CDC temperature, relatively humidity, wind speed, and precipitation; accounting for subject.
-Time-stratified: referent periods matched to events on same day of week within July to October fire season."
-) # end table
-
-print(tab)
-
-# ggplot of odds ratios, facet wrapped by outcomes -----
-# convert variables from character to either numeric or factor
-# factor preserves the order of the variable
-combined_point_est_df$outcome <- factor(combined_point_est_df$outcome, 
-                                        levels = unique(combined_point_est_df$outcome))
-
-combined_point_est_df$pm_method <- factor(combined_point_est_df$pm_method, 
-                                          levels = unique(combined_point_est_df$pm_method))
-
-combined_point_est_df$n <- as.numeric(combined_point_est_df$n)
-combined_point_est_df$n_events <- as.numeric(combined_point_est_df$n_events)
-combined_point_est_df$odds_ratio <- as.numeric(combined_point_est_df$odds_ratio)
-combined_point_est_df$lower95 <- as.numeric(combined_point_est_df$lower95)
-combined_point_est_df$upper95 <- as.numeric(combined_point_est_df$upper95)
-combined_point_est_df$se <- as.numeric(combined_point_est_df$se)
-combined_point_est_df$p_val <- as.numeric(combined_point_est_df$p_val)
-
-# subset out global method
-combined_point_est_df <- combined_point_est_df
-
-## ggplot
-print_plot <- ggplot(combined_point_est_df, 
-                     aes(x = pm_method, y = odds_ratio, colour = pm_method)) +
-  # custom color 
-  scale_color_manual(name = "Smoke-Estimation Method", 
-                     values = c("red", "blue", "#32115C"),
-                     guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
-  geom_point() + #geom_text(vjust = 0, nudge_x = 0.3) +
-  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
-  facet_wrap(~outcome, nrow = 3) +
-  geom_hline(yintercept = 1, linetype=2) +
-  #    ggtitle('Association Between PM2.5 \n from Wildfire Smoke on Hospitalizations') +
-  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
-  #xlab('Smoke Estimation Method') +
-  # plot theme
-  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-        panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        # strip element
-        strip.background = element_rect(colour=NA, fill=NA),
-        panel.border = element_rect(fill = NA, color = "black"),
-        # facet text size
-        #strip.text = element_text(size = 8),
-        # axis element
-        axis.text.x = element_blank(),
-        axis.title.x = element_blank(),
-        axis.ticks.x = element_blank(),
-        axis.title.y = element_text(angle = 90),
-        # legend elements
-        legend.position = "bottom")
-#legend.text = element_text(size = 8))
-
-print(print_plot)
-# save figure
-ggsave("../plot_new/fig_ndc_overall_four.pdf", plot = print_plot, 
-       width = 12, height = 8, units = "in")
-
-```
-
-### Time-Stratified by Age Categories
-
-Following table and figure look at some outcomes stratified by age category.
-
-For some outcomes, such like Age > 65 in WRF-Chem method, the CI is the only insignificant one in the plot.
-
-```{r ndc time stratified age strata, warning =F, echo = F, results='asis'} 
-df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)
-
-outcome_list <- c("Asthma", "COPD", "Both", "Other")
-
-# create an empty list to row bind dataframes together
-datalist <- list()
-
-# outcome name
-outcome_name <- c("Asthma", "COPD", "Both", "Other")
-
-# age category list
-age_cat_list <- c(0,1,2)
-
-# looking only at wrf chem, kriging, and geo-weighted; taking out 'Global Smoke' method
-method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
-
-# create an empty list to row bind dataframes together
-datalist1 <- list()
-datalist2 <- list()
-
-for(i in 1:length(df_list)){
-
-  # dataframe to loop through
-  df_to_loop <- data.frame(df_list[i])
-  # indication of column
-  outcome <- df_list[i]
-  # outcome name
-  outcome_name <- outcome_list[i]
-
-  covariates_df <- df_to_loop[, c(1:26, 74:84)]
-
-  # extract pm values and divide by 10 and ordered
-  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
-  pm_estimates_df <- df_to_loop[, c(87, 92, 91, 93)]/10  # create 10 unit increases
-
-  # new loop for age categories
-  for(k in 0:2){
-
-    point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)
-
-    colnames(point_estimates) <- c('outcome', 'age_cat', 'pm_method', 'n', 'n_events', 
-                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
-
-    # fill in the outcome namedataframe before method loop
-    point_estimates[, 1] <- outcome_name
-    # repeat the age category 4 times for each pm method
-    point_estimates[, 2] <- k
-
-    # dataframe for analysis creation
-    # bind columns back together 
-    df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
-      # remove missing pm values
-      filter(!is.na(wrf_smk_pm_zip)) %>% 
-      # limit to specific age category
-      filter(age_ind == k) %>% 
-      # the following code makes sure that the counterfactual values retained are 
-      # symetric in that number of obs before = number of obs after
-      mutate(obs_diff_admission = (fromdate - date)/7)
-
-    for(j in 38:40){
-
-      # variable to model 
-      var_name <- colnames(df_analysis[j])
-
-      # set row number to fill
-      row_n <- j-37
-
-      # only run the model if the dataframe has observations
-      if(nrow(df_analysis) != 0){
-        # conditional logistic regression model
-        mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
-
-        # populate matrix
-        point_estimates[row_n, 3] <- method_list[row_n]
-        point_estimates[row_n, 4] <- mod$n
-        point_estimates[row_n, 5] <- mod$nevent
-        # odds ratio
-        point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
-
-        # 95% lower bound
-        point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
-        # 95% upper bound
-        point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
-                                                 1.96*(summary(mod)$coefficient[1,3])), 3)
-        # standard error
-        point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
-        # p val
-        point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
-
-        # create else statement that fills matrix with missing so I still have the row
-        # in the final dataframe
-      } else {point_estimates[row_n, 3] <- method_list[row_n]
-      point_estimates[row_n, 4] <- 0
-      point_estimates[row_n, 5] <- 0
-      point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
-    } # end methods loop
-
-    # save point estimates as a dataframe
-    point_est_df <- as_data_frame(point_estimates)
-
-    # combine previous values in dataframe that has all outcome/methods comparisons
-    datalist1[[k+1]] <- point_est_df
-  } # end age category loop
-
-  # bind rows of age category estimates together
-  age_est_df <- bind_rows(datalist1)
-
-  # populate second dataframe list
-  datalist2[[i]] <- age_est_df
-
-} # end of outcome loop
-
-# combine each outcome dataframe itteration in to a big dataset
-combined_point_est_df <- bind_rows(datalist2)
-
-wrf_geo_age <- combined_point_est_df %>% 
-  # filter columns to just geo-smk
-  # filter(pm_method == "GWR Smoke" |
-    #      pm_method == "WRF-Chem Smoke") %>% 
-  mutate(age_cat2 = ifelse(age_cat == 0, "Less than 15", 
-                    ifelse(age_cat == 1, "15-65",
-                    ifelse(age_cat == 2, "Greater than 65", NA)))) %>% 
-  # subset columns I want to put in to the table
-  select(3, 11, 5:8) %>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
-         lower95 = ifelse(lower95 == 99, "--", lower95),
-         upper95 = ifelse(upper95 == 99, "--", upper95))
-
-tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
-         caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by age",
-         # row group by outcome
-         rgroup = "Asthma inhalers",
-         n.rgroup = c(rep(9, 1)), # 6 rows for each age cat for each outcome
-         # column headers
-         header = c("Est Method", "Age Group", "Events",
-                    "OR&dagger;", "Lower", "Upper"),
-         # column spanner
-         cgroup = c("", "95% CI"), 
-         n.cgroup = c(4, 2),
-         padding.rgroup = "&nbsp;&nbsp;",
-         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
-         align = "llcccc", # column alignment,
-         tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within May to September fire season."
-         ) # end table
-
-print(tab)
-
-combined_point_est_df_new <- combined_point_est_df %>%
-  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
-                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
-  select(-n_events) %>%
-  rename(n_events = n_events_new)
-
-# ggplot of odds ratios, facet wrapped by outcomes -----
-# convert variables from character to either numeric or factor
-# factor preserves the order of the variable
-wrf_geo_age_plot <- combined_point_est_df_new %>% 
-  # filter columns to just geo-smk
- #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
-  # do not plot results with less than 15 events, create missing vals
-  # in mutate
-  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
-                             99, odds_ratio),
-         lower95 = ifelse(as.numeric(n_events) < 15, 
-                             99, lower95),
-         upper95 = ifelse(as.numeric(n_events) < 15, 
-                             99, upper95),
-         age_cat2 = ifelse(age_cat == 0, "<15", 
-                    ifelse(age_cat == 1, "15-65",
-                    ifelse(age_cat == 2, ">65", NA)))) %>% 
-  # subset columns I want to put in to the table
-  select(1, 3, 11, 10, 5:7) %>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
-         lower95 = ifelse(lower95 == 99, NA, lower95),
-         upper95 = ifelse(upper95 == 99, NA, upper95))
-
-# change characters to numeric and factor  
-wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
-                              levels = unique(wrf_geo_age_plot$outcome))
-
-wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
-                                levels = unique(wrf_geo_age_plot$pm_method))
-
-wrf_geo_age_plot$age_cat2 <- factor(wrf_geo_age_plot$age_cat2,
-                                levels = unique(wrf_geo_age_plot$age_cat2))
-
-wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
-wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
-wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
-wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)
-
-## ggplot 
-  print_plot <- ggplot(wrf_geo_age_plot,
-    aes(x = age_cat2, y = odds_ratio, color = pm_method), na.rm = T) +
-    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
-    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
-                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
-    # custom color 
-    scale_color_manual(name = "Smoke-Estimation Method", 
-                       values = c("red",  "blue", "#32115C"),
-                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
-    facet_wrap(~outcome, nrow = 3, scales = "free") +
-    geom_hline(yintercept = 1, linetype=2) +
-    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
-    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
-    #ylim(0, 2) +
-    xlab('Age Category ') +
-    # plot theme
-    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    # strip element
-    strip.background = element_rect(colour=NA, fill=NA),
-    panel.border = element_rect(fill = NA, color = "black"),
-    # facet text size
-    strip.text = element_text(size = 10), 
-    # axis element
-    #axis.text.x = element_blank(),
-    #axis.title.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.title.y = element_text(angle = 90),
-    # legend elements
-    legend.position = "bottom")
-    #legend.text = element_text(size = 8))
-
-  print(print_plot)
-  # save figure
-  ggsave("../plot_new/ndc_age_fig_four.pdf", plot = print_plot, 
-       width = 12, height = 8, units = "in")
-
-```
-
-### Seasonal Time-Stratified by Sex
-
-Association with smoke PM~2.5~ and health outcomes stratified by sex.
-
-All the CI do not include 1, which means the use of asthma medication (beta 2 agonists) increases about 5% when smoke increase 10 units.
-So the risk for getting asthma will increase about 5% ~ 8% when the smoke PM2.5 increase 10 units.
-
-```{r ndc sex strata, warning = F, echo = F, results='asis'} 
-df_list <- list(beta_asthma, beta_copd, beta_both, beta_other)
-
-outcome_list <- c("Asthma", "COPD", "Both", "Other")
-
-# outcome name
-outcome_name <- c("Asthma", "COPD", "Both", "Other")
-
-# sex category list
-sex_strata_list <- c(0,1)
-
-datalist1 <- list()
-datalist2 <- list()
-
-# data wrangling ----
-# Producing conditional logit model estimates loop 
-for(i in 1:length(df_list)){
-
-  # dataframe to loop through
-  df_to_loop <- data.frame(df_list[i])
-  # indication of column
-  outcome <- df_list[i]
-  # outcome name
-  outcome_name <- outcome_list[i]
-
-  covariates_df <- df_to_loop[, c(1:26, 74:84)]
-
-  # extract pm values and divide by 10 and ordered
-  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
-  pm_estimates_df <- df_to_loop[, c(87, 92, 91, 93)]/10  # create 10 unit increases
-
-  # new loop for sex categories
-  for(k in 0:1){
-
-    # empty matrix (12 x 10 matrix)
-    point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)
-
-    colnames(point_estimates) <- c('outcome', 'sex', 'pm_method', 'n', 'n_events', 
-                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
-
-    # fill in the outcome namedataframe before method loop
-    point_estimates[, 1] <- outcome_name
-    # repeat the sex category 4 times for each pm method
-    point_estimates[, 2] <- k
-
-    # dataframe for analysis creation
-    # bind columns back together 
-    df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
-      # remove missing pm values
-      filter(!is.na(wrf_smk_pm_zip)) %>% 
-      # limit to specific sex category
-      filter(sex_ind == k) %>% 
-      # the following code makes sure that the counterfactual values retained are 
-      # symetric in that number of obs before = number of obs after
-      mutate(obs_diff_admission = (fromdate - date)/7) 
-    # dataframe is already for the entire fire season, so I don't need to subset anymore
-
-    for(j in 38:40){
-
-      # variable to model 
-      var_name <- colnames(df_analysis[j])
-
-      # set row number to fill
-      row_n <- j-37
-
-      # only run the model if the dataframe has observations
-      if(nrow(df_analysis) != 0){
-        # conditional logistic regression model
-        mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
-
-        # populate matrix
-        point_estimates[row_n, 3] <- method_list[row_n]
-        point_estimates[row_n, 4] <- mod$n
-        point_estimates[row_n, 5] <- mod$nevent
-        # odds ratio
-        point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
-
-        # 95% lower bound
-        point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
-                                                 1.96*(summary(mod)$coefficient[1,3])), 3)
-        # 95% upper bound
-        point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
-                                                 1.96*(summary(mod)$coefficient[1,3])), 3)
-        # standard error
-        point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
-        # p val
-        point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
-
-        # create else statement that fills matrix with missing so I still have the row
-        # in the final dataframe
-      } else {point_estimates[row_n, 3] <- method_list[row_n]
-      point_estimates[row_n, 4] <- 0
-      point_estimates[row_n, 5] <- 0
-      point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
-    } # end methods loop
-
-    # save point estimates as a dataframe
-    point_est_df <- as_data_frame(point_estimates)
-
-    # combine previous values in dataframe that has all outcome/methods comparisons
-    datalist1[[k+1]] <- point_est_df
-  } # end sex category loop
-
-  # bind rows of sex category estimates together
-  sex_est_df <- bind_rows(datalist1)
-
-  # populate second dataframe list
-  datalist2[[i]] <- sex_est_df
-
-} # end of outcome loop
-
-# combine each outcome dataframe itteration in to a big dataset
-combined_point_est_df <- bind_rows(datalist2)
-
-wrf_geo_sex <- combined_point_est_df %>% 
-  # filter columns to just geo-smk
-  # filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
-  mutate(sex_cat = ifelse(sex == 0, "Male", 
-               ifelse(sex == 1, "Female", NA))) %>% 
-  # subset columns I want to put in to the table
-  select(3, 11, 5:8)%>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
-         lower95 = ifelse(lower95 == 99, "--", lower95),
-         upper95 = ifelse(upper95 == 99, "--", upper95))
-
-  tab <- htmlTable(txtRound(wrf_geo_sex, digits = 3, 1:3), 
-           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and pharmacy outcomes stratified by sex",
-           # row group by outcome
-           rgroup = "Asthma inhalers",
-           n.rgroup = c(rep(6, 1)), # 2 rows for each sex strata for each outcome
-           # column headers
-           header = c("Est Method", "Sex Strata", "Events",
-                      "OR&dagger;", "Lower", "Upper"),
-           # column spanner
-           cgroup = c("", "95% CI"), 
-           n.cgroup = c(4, 2),
-           padding.rgroup = "&nbsp;&nbsp;",
-           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
-           align = "llcccc", # column alignment,
-           tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within July to October fire season."
-            ) # end table
-
-  print(tab)
-
-# ggplot of odds ratios, facet wrapped by outcomes -----
-# convert variables from character to either numeric or factor
-# factor preserves the order of the variable
-wrf_geo_sex_plot <- combined_point_est_df %>% 
-  # filter columns to just geo-smk
- #  filter(pm_method == "GWR Smoke" | 
-   #       pm_method == "WRF-Chem Smoke") %>% 
-  mutate(sex_cat = ifelse(sex == 0, "Male", 
-               ifelse(sex == 1, "Female", NA))) %>%
-  # subset columns I want to put in to the table
-  select(1, 3, 11, 5:8) %>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
-         lower95 = ifelse(lower95 == 99, NA, lower95),
-         upper95 = ifelse(upper95 == 99, NA, upper95))
-
-# change characters to numeric and factor  
-wrf_geo_sex_plot$outcome <- factor(wrf_geo_sex_plot$outcome,
-                              levels = unique(wrf_geo_sex_plot$outcome))
-
-wrf_geo_sex_plot$pm_method <- factor(wrf_geo_sex_plot$pm_method,
-                                levels = unique(wrf_geo_sex_plot$pm_method))
-
-wrf_geo_sex_plot$sex_cat <- factor(wrf_geo_sex_plot$sex_cat,
-                                levels = unique(wrf_geo_sex_plot$sex_cat))
-
-wrf_geo_sex_plot$n_events <- as.numeric(wrf_geo_sex_plot$n_events)
-wrf_geo_sex_plot$odds_ratio <- as.numeric(wrf_geo_sex_plot$odds_ratio)
-wrf_geo_sex_plot$lower95 <- as.numeric(wrf_geo_sex_plot$lower95)
-wrf_geo_sex_plot$upper95 <- as.numeric(wrf_geo_sex_plot$upper95)
-
-## ggplot
-  print_plot <- ggplot(wrf_geo_sex_plot,
-    aes(x = sex_cat, y = odds_ratio, color = pm_method), na.rm = T) +
-    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
-    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
-                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
-    # custom color 
-    scale_color_manual(name = "Smoke-Estimation Method", 
-                       values = c("red", "blue", "#32115C"),
-                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
-    facet_wrap(~outcome, nrow = 3, scales = "free") +
-    geom_hline(yintercept = 1, linetype=2) +
-    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
-    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
-    #ylim(0, 2) +
-    xlab('Sex') +
-    # plot theme
-    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    # strip element
-    strip.background = element_rect(colour=NA, fill=NA),
-    panel.border = element_rect(fill = NA, color = "black"),
-    # facet text size
-    strip.text = element_text(size = 10),
-    # axis element
-    #axis.text.x = element_blank(),
-    #axis.title.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.title.y = element_text(angle = 90),
-    # legend elements
-    legend.position = "bottom")
-    #legend.text = element_text(size = 8))
-
-  print(print_plot)
-  # save figure
-  ggsave("../plot_new/ndc_sex_fig_four.pdf", plot = print_plot, 
-       width = 12, height = 8, units = "in")
-
-```

### 2017-07-10 r_scripts/analysis/oregon_medication.Rmd
health analysis overall of lags

@@
 
-### Standard Regression Model
-
-Same-day estimates of association between smoke via GWR and outcomes. County treated as a fixed effect covariate.
-
-1. Quasipoisson
+### Standard Regression Model (Relative Risk)
+
+1. Fixed-Quasipoisson
+2. Fixed-Poisson
+3. Mixed-Poisson
 
@@
 # create empty dataframe
-rr_df <- data.frame(matrix(nrow = 1, ncol = 4))
-colnames(rr_df) <- c("outcome", "rel_risk", "lower_95", "upper_95")
+rr_df <- data.frame(matrix(nrow = 3, ncol = 5))
+colnames(rr_df) <- c("method", "outcome", "rel_risk", "lower_95", "upper_95")
 
@@
 
+# methods
+rr_df[,1] <- c("Fixed-Quasipoisson", "Fixed-Poisson", "Mixed-Poisson")
+
 # names 
@@
 # populate outcome column
-rr_df[,1] <- factor(outcome_names, levels=unique(outcome_names))
-
-smk_est <- tidy(glm(as.formula(paste0(outcome_list, 
+rr_df[,2] <- factor(outcome_names, levels=unique(outcome_names))
+
+## quasipoisson
+smk_est_fq <- tidy(glm(as.formula(paste0(outcome_list, 
     "~ geo_smk_pm10 + wrf_temp_county + county + 
@@
 
-rr <- c(round(exp(smk_est[1,2]),5),
-        round(exp(smk_est[1,2]-1.96*smk_est[1,3]),5),
-        round(exp(smk_est[1,2]+1.96*smk_est[1,3]),5))
-
-rr_df[1,2:4] <- rr
-
-# tried apply function and it might have been easier to just use a loop
-# as I needed a for loop to fill the list in to the dataframe
-# code is more compact though
-# I should check speed (or make a custom function to do this one day)
+rr_fq <- c(round(exp(smk_est_fq[1,2]),5),
+         round(exp(smk_est_fq[1,2]-1.96*smk_est_fq[1,3]),5),
+         round(exp(smk_est_fq[1,2]+1.96*smk_est_fq[1,3]),5))
+
+rr_df[1,3:5] <- rr_fq
+
+## poisson
+smk_est_fp <- tidy(glm(as.formula(paste0(outcome_list, 
+    "~ geo_smk_pm10 + wrf_temp_county + county + 
+    day + offset(log(pop))")), 
+    or_ndc_county_ts, family="poisson"))[2, ]
+
+rr_fp <- c(round(exp(smk_est_fp[1,2]),5),
+         round(exp(smk_est_fp[1,2]-1.96*smk_est_fp[1,3]),5),
+         round(exp(smk_est_fp[1,2]+1.96*smk_est_fp[1,3]),5))
+
+rr_df[2,3:5] <- rr_fp
+
+## mixed
+smk_est_mp <- tidy(glmer(as.formula(paste0(outcome_list, 
+    "~ geo_smk_pm10 + wrf_temp_county + (1|county) + 
+    day + offset(log(pop))")), 
+    or_ndc_county_ts, family="poisson"))[2, ]
+
+rr_mp <- c(round(exp(smk_est_mp[1,2]),5),
+         round(exp(smk_est_mp[1,2]-1.96*smk_est_mp[1,3]),5),
+         round(exp(smk_est_mp[1,2]+1.96*smk_est_mp[1,3]),5))
+
+rr_df[3,3:5] <- rr_mp
 
@@
 
-plot_rr <- ggplot(rr_df, aes(x=outcome, y = rel_risk)) +
+plot_rr <- ggplot(rr_df, aes(x=method, y = rel_risk)) +
   geom_point() +
@@
 
-ggsave("../plot_new/ndc_rr.pdf", plot = plot_rr, 
+ggsave("../plot_new/ndc_rr_geo.pdf", plot = plot_rr, 
        width = 12, height = 8, units = "in")
@@
 
-2. Poisson
-
-```{r regression model,  message=FALSE, echo=FALSE}
-# create empty dataframe
-rr_df <- data.frame(matrix(nrow = 1, ncol = 4))
-colnames(rr_df) <- c("outcome", "rel_risk", "lower_95", "upper_95")
-
-# vector of outcomes
-outcome_list <- "n_obs"
-
-# names 
-outcome_names <- c('Inhalers')
-
-# populate outcome column
-rr_df[,1] <- factor(outcome_names, levels=unique(outcome_names))
-
-smk_est <- tidy(glm(as.formula(paste0(outcome_list, 
-    "~ geo_smk_pm10 + wrf_temp_county + county + 
-    day + offset(log(pop))")), 
-    or_ndc_county_ts, family="poisson"))[2, ]
-
-rr <- c(round(exp(smk_est[1,2]),5),
-        round(exp(smk_est[1,2]-1.96*smk_est[1,3]),5),
-        round(exp(smk_est[1,2]+1.96*smk_est[1,3]),5))
-
-rr_df[1,2:4] <- rr
-
-rr_df
-
-# tried apply function and it might have been easier to just use a loop
-# as I needed a for loop to fill the list in to the dataframe
-# code is more compact though
-# I should check speed (or make a custom function to do this one day)
-
-plot_rr <- ggplot(rr_df, aes(x=outcome, y = rel_risk)) +
-  geom_point() +
-  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
-  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
-  theme(
-   panel.background = element_rect(fill = 'white', colour = 'black'),
-   panel.grid.major = element_blank(),
-   panel.grid.minor = element_blank(),
-   axis.text.x = element_text(angle = 90, vjust = .75)) +
-  ylab("Relative Risk") +
-  xlab("Outcome") +
-  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")
-
-print(plot_rr)
-
-ggsave("../plot_new/ndc_rr_poi.pdf", plot = plot_rr, 
-       width = 12, height = 8, units = "in")
-
-```
-
-### Mixed models
-
-1. Poisson
-
-```{r mixed model,  message=FALSE, echo=FALSE}
-# create empty dataframe
-rr_mix_df <- data.frame(matrix(nrow = 1, ncol = 4))
-colnames(rr_mix_df) <- c("outcome", "rel_risk", "lower_95", "upper_95")
-
-# vector of outcomes
-outcome_list <- "n_obs"
-
-# names 
-outcome_names <- c('Inhalers')
-
-# populate outcome column
-rr_mix_df[,1] <- factor(outcome_names, levels=unique(outcome_names))
-
-smk_est <- tidy(glmer(as.formula(paste0(outcome_list, 
-    "~ geo_smk_pm10 + wrf_temp_county + (1|county) + 
-    day + offset(log(pop))")), 
-    or_ndc_county_ts, family="poisson"))[2, ]
-
-rr <- c(round(exp(smk_est[1,2]),5),
-        round(exp(smk_est[1,2]-1.96*smk_est[1,3]),5),
-        round(exp(smk_est[1,2]+1.96*smk_est[1,3]),5))
-
-rr_mix_df[1,2:4] <- rr
-
-rr_mix_df
-
-plot_mix_rr <- ggplot(rr_mix_df, aes(x=outcome, y = rel_risk)) +
-  geom_point() +
-  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
-  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
-  theme(
-   panel.background = element_rect(fill = 'white', colour = 'black'),
-   panel.grid.major = element_blank(),
-   panel.grid.minor = element_blank(),
-   axis.text.x = element_text(angle = 90, vjust = .75)) +
-  ylab("Relative Risk") +
-  xlab("Outcome") +
-  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")
-
-print(plot_rr)
-
-ggsave("../plot_new/ndc_mix_rr_poi.pdf", plot = plot_mix_rr, 
-       width = 12, height = 8, units = "in")
-
-```
-
 ## Distributed Lag Model
@@
 Remove the missing values for GWR smoke. 
-But I use GWR smoke of zip for the case crossover study and add the county variable.
+But I use GWR smoke of zip for the case crossover study.
 
@@
 
-## Distributed Lag Casecrossover Analysis
-
-### Lag 1
-
-### Overall plot
-
-```{r overall, warning =F, echo = F, results='asis'} 
+## Distributed Lag -- Casecrossover Analysis
+
+### Lag
+
+### 1. For Overall Plot
+
+```{r overall lags, warning =F, echo = F, results='asis'} 
 # dataframe list
@@
 
+lag_list <- c("Lag 1", "Lag 2", "Lag 3", "Lag 4", "Lag 5", "Lag 6", "Lag 7")
+
 # dataframe to loop through
-inhaler_ndc_lag1_df <- data.frame(inhaler_ndc) %>%
-  filter(!is.na(geo_smk_pm_lag1_zip))
+inhaler_ndc <- data.frame(inhaler_ndc) 
+#  filter(!is.na(geo_smk_pm_lag1_zip))
 
@@
 outcome <- "Asthma Inhalers"
-outcome_name <- "Asthma Inhalers"
 
@@
 # extract covariates from dataframe
-covariates_df <- inhaler_ndc_lag1_df[, c(1:26, 74:79, 94:156)]
+covariates_df <- inhaler_ndc[, c(1:26, 74:79)] # 94:156 is lag
 
 # extract pm values and divide by 10 and ordered
-which(colnames(inhaler_ndc_lag1_df)=="geo_smk_pm_lag1_zip") # 136
-pm_estimates_df <- inhaler_ndc_lag1_df[, 136]/10  # create 10 unit increases
+which(colnames(inhaler_ndc)=="geo_smk_pm_lag1_zip") # 136
+which(colnames(inhaler_ndc)=="wrf_temp_lag1_zip") # 150
+pm_estimates_df <- inhaler_ndc[, 136:142]/10  # create 10 unit increases
+temp_estimates_df <- inhaler_ndc[, 150:156]  # temperature
 
@@
 # empty matrix (12 x 10 matrix)
-point_estimates <- matrix(nrow = 1, ncol = 9, byrow = T)
-
-colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 
+point_estimates <- matrix(nrow = 7, ncol = 10, byrow = T)
+
+colnames(point_estimates) <- c('outcome', 'pm_method',  'lag', 'n', 'n_events', 
                                    'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
@@
 # fill in the outcome namedataframe before method loop
-point_estimates[, 1] <- outcome_name
+point_estimates[, 1] <- outcome
+
+for (i in 1:7){
+geo_lag <- i + 32 
+temp_lag <- i + 39
 
@@
 # bind columns back together 
-df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+df_analysis <- cbind(covariates_df, pm_estimates_df, temp_estimates_df)
+
+df_analysis <- df_analysis %>% 
   # remove missing pm values
-  # filter(!is.na(wrf_smk_pm_zip)) %>% 
+  filter(!is.na(df_analysis[geo_lag])) %>% 
   # the following code makes sure that the counterfactual values retained are 
@@
   # variable to model 
-  var_name <- colnames(df_analysis[75])
+  var_name <- colnames(df_analysis[geo_lag])
 
   # set row number to fill
-  row_n <- 1
+  row_n <- i
 
@@
   # conditional logistic regression model
-  mod <- clogit(outcome ~ df_analysis[[75]] + wrf_temp_lag1_zip + strata(personkey), df_analysis)
+  mod <- clogit(outcome ~ df_analysis[[geo_lag]] + df_analysis[[temp_lag]] + strata(personkey), df_analysis)
 
   # populate matrix
-  point_estimates[row_n, 2] <- method_list[row_n]
-  point_estimates[row_n, 3] <- mod$n
-  point_estimates[row_n, 4] <- mod$nevent
+  point_estimates[row_n, 2] <- method_list
+  point_estimates[row_n, 3] <- lag_list[row_n]
+  point_estimates[row_n, 4] <- mod$n
+  point_estimates[row_n, 5] <- mod$nevent
   # odds ratio
-  point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+  point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
 
   # 95% lower bound
-  point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
+  point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
   # 95% upper bound
-  point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
+  point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
                                         1.96*(summary(mod)$coefficient[1,3])), 3)
   # standard error
-  point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
+  point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
   # p val
-  point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)
+  point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
 
@@
   # in the final dataframe
-  } else {point_estimates[row_n, 2] <- method_list[row_n]
-          point_estimates[row_n, 3] <- 0
+  } else {point_estimates[row_n, 3] <- method_list[row_n]
           point_estimates[row_n, 4] <- 0
-          point_estimates[row_n, c(5:9)] <- 99 } # end 'if else' statement
-
+          point_estimates[row_n, 5] <- 0
+          point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
+
+}
 # save point estimates as a dataframe
@@
 
-wrf_geo_age <- combined_point_est_df %>% 
-  # filter columns to just geo-smk
-  # filter(pm_method == "GWR Smoke" |
-    #      pm_method == "WRF-Chem Smoke") %>% 
-  # subset columns I want to put in to the table
-  select(2, 4:7) %>% 
+wrf_geo_lag <- combined_point_est_df %>% 
+  select(3, 5:8) %>% 
   mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
@@
 
-tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
+tab <- htmlTable(txtRound(wrf_geo_lag, digits = 3, 1:3), 
          caption = "Association between a 10 ug/m^3 in PM2.5 three smoke method and pharmacy outcomes",
@@
          rgroup = "Asthma Inhalers",
-         n.rgroup = c(rep(3, 1)), # 6 rows for each age cat for each outcome
+         n.rgroup = c(rep(1, 1)), # 6 rows for each age cat for each outcome
          # column headers
-         header = c("Est Method", "Events",
+         header = c("Lag", "Events",
                     "OR&dagger;", "Lower", "Upper"),
@@
 # factor preserves the order of the variable
-wrf_geo_age_plot <- combined_point_est_df_new %>% 
+wrf_geo_lag_plot <- combined_point_est_df_new %>% 
   # filter columns to just geo-smk
@@
   # subset columns I want to put in to the table
-  select(1, 2, 9, 4:6) %>% 
+  select(1, 3, 10, 5:7) %>% 
   mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
@@
 # change characters to numeric and factor  
-wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
-                              levels = unique(wrf_geo_age_plot$outcome))
-
-wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
-                                levels = unique(wrf_geo_age_plot$pm_method))
-
-wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
-wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
-wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
-wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)
+wrf_geo_lag_plot$outcome <- factor(wrf_geo_lag_plot$outcome,
+                              levels = unique(wrf_geo_lag_plot$outcome))
+
+wrf_geo_lag_plot$lag <- factor(wrf_geo_lag_plot$lag,
+                                levels = unique(wrf_geo_lag_plot$lag))
+
+wrf_geo_lag_plot$n_events <- as.numeric(wrf_geo_lag_plot$n_events)
+wrf_geo_lag_plot$odds_ratio <- as.numeric(wrf_geo_lag_plot$odds_ratio)
+wrf_geo_lag_plot$lower95 <- as.numeric(wrf_geo_lag_plot$lower95)
+wrf_geo_lag_plot$upper95 <- as.numeric(wrf_geo_lag_plot$upper95)
 
 ## ggplot 
-  print_plot <- ggplot(wrf_geo_age_plot,
-    aes(x = pm_method, y = odds_ratio, color = pm_method), na.rm = T) +
+  print_plot <- ggplot(wrf_geo_lag_plot,
+    aes(x = lag, y = odds_ratio, color = lag), na.rm = T) +
     geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
@@
     # custom color 
-    scale_color_manual(name = "Smoke-Estimation Method", 
-                       values = c("red",  "blue", "#32115C"),
+    scale_color_manual(name = "Geo-smk Lag", 
+                       values = c("red", "orange", "yellow", "green", "blue", "#32115C", "purple"),
                        guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
-    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    facet_wrap(~outcome, nrow = 7, scales = "free") +
     geom_hline(yintercept = 1, linetype=2) +
@@
   # save figure
-  ggsave("../plot_new/ndc_overall_three.pdf", plot = print_plot, 
+  ggsave("../plot_new/ndc_health_lag.pdf", plot = print_plot, 
        width = 12, height = 8, units = "in")
@@
 
-### Time-Stratified by Age Categories
-
-Following table and figure look at some outcomes stratified by age category.
-
-For some outcomes, such like Age > 65 in WRF-Chem method, the CI is the only insignificant one in the plot.
-
-```{r time stratified age strata cdc met, warning =F, echo = F, results='asis'} 
-# dataframe list
-
-method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
-
-# dataframe to loop through
-inhaler_ndc_df <- data.frame(inhaler_ndc)
-
-# outcome name
-outcome <- "asthma inhalers"
-outcome_name <- "Asthma inhalers"
-
-# age category list
-age_cat_list <- c(0,1,2)
-
-# create an empty list to row bind dataframes together
-datalist1 <- list()
-
-# data wrangling ----
-# Producing conditional logit model estimates loop
-
-# extract covariates from dataframe
-covariates_df <- inhaler_ndc_df[, c(1:26, 74:84)]
-
-# extract pm values and divide by 10 and ordered
-#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
-pm_estimates_df <- inhaler_ndc_df[, c(87, 92, 91, 93)]/10  # create 10 unit increases
-
-# new loop for age categories
-for(k in 0:2){
-  # empty matrix (12 x 10 matrix)
-  point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)
-
-  colnames(point_estimates) <- c('outcome', 'age_cat', 'pm_method', 'n', 'n_events', 
-                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
-
-  # fill in the outcome namedataframe before method loop
-  point_estimates[, 1] <- outcome_name
-  # repeat the age category 4 times for each pm method
-  point_estimates[, 2] <- k
-
-  # dataframe for analysis creation
-  # bind columns back together 
-  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
-    # remove missing pm values
-    filter(!is.na(wrf_smk_pm_zip)) %>% 
-    # limit to specific age category
-    filter(age_ind == k) %>% 
-    # the following code makes sure that the counterfactual values retained are 
-    # symetric in that number of obs before = number of obs after
-    mutate(obs_diff_admission = (fromdate - date)/7) 
-    # dataframe is already for the entire fire season, so I don't need to subset anymore
-
-  # second loop to run a model for each pm estimation method
-  for(j in 38:40){
-
-    # variable to model 
-    var_name <- colnames(df_analysis[j])
-
-    # set row number to fill
-    row_n <- j-37
-
-    # only run the model if the dataframe has observations
-    if(nrow(df_analysis) != 0){
-    # conditional logistic regression model
-    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
-
-    # populate matrix
-    point_estimates[row_n, 3] <- method_list[row_n]
-    point_estimates[row_n, 4] <- mod$n
-    point_estimates[row_n, 5] <- mod$nevent
-    # odds ratio
-    point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
-
-    # 95% lower bound
-    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
-    # 95% upper bound
-    point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
-                                        1.96*(summary(mod)$coefficient[1,3])), 3)
-    # standard error
-    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
-    # p val
-    point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
-
-    # create else statement that fills matrix with missing so I still have the row
-    # in the final dataframe
-    } else {point_estimates[row_n, 3] <- method_list[row_n]
-            point_estimates[row_n, 4] <- 0
-            point_estimates[row_n, 5] <- 0
-            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
-    } # end methods loop
-
-  # save point estimates as a dataframe
-  point_est_df <- as_data_frame(point_estimates)
-  datalist1[[k+1]] <- point_est_df
-} # end age category loop
-
-# bind rows of age category estimates together
-combined_point_est_df <- bind_rows(datalist1)
-
-wrf_geo_age <- combined_point_est_df %>% 
-  # filter columns to just geo-smk
-  # filter(pm_method == "GWR Smoke" |
-    #      pm_method == "WRF-Chem Smoke") %>% 
-  mutate(age_cat2 = ifelse(age_cat == 0, "Less than 15", 
-                    ifelse(age_cat == 1, "15-65",
-                    ifelse(age_cat == 2, "Greater than 65", NA)))) %>% 
-  # subset columns I want to put in to the table
-  select(3, 11, 5:8) %>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
-         lower95 = ifelse(lower95 == 99, "--", lower95),
-         upper95 = ifelse(upper95 == 99, "--", upper95))
-
-tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
-         caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by age",
-         # row group by outcome
-         rgroup = "Asthma inhalers",
-         n.rgroup = c(rep(9, 1)), # 6 rows for each age cat for each outcome
-         # column headers
-         header = c("Est Method", "Age Group", "Events",
-                    "OR&dagger;", "Lower", "Upper"),
-         # column spanner
-         cgroup = c("", "95% CI"), 
-         n.cgroup = c(4, 2),
-         padding.rgroup = "&nbsp;&nbsp;",
-         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
-         align = "llcccc", # column alignment,
-         tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within May to September fire season."
-         ) # end table
-
-print(tab)
-
-combined_point_est_df_new <- combined_point_est_df %>%
-  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
-                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
-  select(-n_events) %>%
-  rename(n_events = n_events_new)
-
-# ggplot of odds ratios, facet wrapped by outcomes -----
-# convert variables from character to either numeric or factor
-# factor preserves the order of the variable
-wrf_geo_age_plot <- combined_point_est_df_new %>% 
-  # filter columns to just geo-smk
- #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
-  # do not plot results with less than 15 events, create missing vals
-  # in mutate
-  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
-                             99, odds_ratio),
-         lower95 = ifelse(as.numeric(n_events) < 15, 
-                             99, lower95),
-         upper95 = ifelse(as.numeric(n_events) < 15, 
-                             99, upper95),
-         age_cat2 = ifelse(age_cat == 0, "<15", 
-                    ifelse(age_cat == 1, "15-65",
-                    ifelse(age_cat == 2, ">65", NA)))) %>% 
-  # subset columns I want to put in to the table
-  select(1, 3, 11, 10, 5:7) %>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
-         lower95 = ifelse(lower95 == 99, NA, lower95),
-         upper95 = ifelse(upper95 == 99, NA, upper95))
-
-# change characters to numeric and factor  
-wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
-                              levels = unique(wrf_geo_age_plot$outcome))
-
-wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
-                                levels = unique(wrf_geo_age_plot$pm_method))
-
-wrf_geo_age_plot$age_cat2 <- factor(wrf_geo_age_plot$age_cat2,
-                                levels = unique(wrf_geo_age_plot$age_cat2))
-
-wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
-wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
-wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
-wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)
-
-## ggplot 
-  print_plot <- ggplot(wrf_geo_age_plot,
-    aes(x = age_cat2, y = odds_ratio, color = pm_method), na.rm = T) +
-    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
-    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
-                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
-    # custom color 
-    scale_color_manual(name = "Smoke-Estimation Method", 
-                       values = c("red",  "blue", "#32115C"),
-                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
-    facet_wrap(~outcome, nrow = 3, scales = "free") +
-    geom_hline(yintercept = 1, linetype=2) +
-    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
-    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
-    #ylim(0, 2) +
-    xlab('Age Category ') +
-    # plot theme
-    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    # strip element
-    strip.background = element_rect(colour=NA, fill=NA),
-    panel.border = element_rect(fill = NA, color = "black"),
-    # facet text size
-    strip.text = element_text(size = 10), 
-    # axis element
-    #axis.text.x = element_blank(),
-    #axis.title.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.title.y = element_text(angle = 90),
-    # legend elements
-    legend.position = "bottom")
-    #legend.text = element_text(size = 8))
-
-  print(print_plot)
-  # save figure
-  ggsave("../plot_new/ndc_age_fig_three.pdf", plot = print_plot, 
-       width = 12, height = 8, units = "in")
-
-```
-
-### Seasonal Time-Stratified by Sex
-
-Association with smoke PM~2.5~ and health outcomes stratified by sex.
-
-All the CI do not include 1, which means the use of asthma medication (beta 2 agonists) increases about 5% when smoke increase 10 units.
-So the risk for getting asthma will increase about 5% ~ 8% when the smoke PM2.5 increase 10 units.
-
-```{r sex strata cdc met adj, warning = F, echo = F, results='asis'} 
-method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
-
-# dataframe to loop through
-inhaler_ndc_df <- data.frame(inhaler_ndc)
-
-# outcome name
-outcome <- "asthma inhalers"
-outcome_name <- "Asthma inhalers"
-
-# create an empty list to row bind dataframes together
-datalist1 <- list()
-
-# sex category list
-sex_strata_list <- c(0,1)
-
-# data wrangling ----
-# Producing conditional logit model estimates loop
-
-# extract covariates from dataframe
-covariates_df <- inhaler_ndc_df[, c(1:26, 74:84)]
-
-# extract pm values and divide by 10 and ordered
-#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
-pm_estimates_df <- inhaler_ndc_df[, c(87, 92, 91, 93)]/10  # create 10 unit increases
-
-# new loop for sex categories
-for(k in 0:1){
-
-  # empty matrix (12 x 10 matrix)
-  point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)
-
-  colnames(point_estimates) <- c('outcome', 'sex', 'pm_method', 'n', 'n_events', 
-                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
-
-  # fill in the outcome namedataframe before method loop
-  point_estimates[, 1] <- outcome_name
-  # repeat the sex category 4 times for each pm method
-  point_estimates[, 2] <- k
-
-  # dataframe for analysis creation
-  # bind columns back together 
-  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
-    # remove missing pm values
-    filter(!is.na(wrf_smk_pm_zip)) %>% 
-    # limit to specific sex category
-    filter(sex_ind == k) %>% 
-    # the following code makes sure that the counterfactual values retained are 
-    # symetric in that number of obs before = number of obs after
-    mutate(obs_diff_admission = (fromdate - date)/7) 
-    # dataframe is already for the entire fire season, so I don't need to subset anymore
-
-  # second loop to run a model for each pm estimation method
-  for(j in 38:40){
-
-    # variable to model 
-    var_name <- colnames(df_analysis[j])
-
-    # set row number to fill
-    row_n <- j-37
-
-    # only run the model if the dataframe has observations
-    if(nrow(df_analysis) != 0){
-    # conditional logistic regression model
-    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
-
-    # populate matrix
-    point_estimates[row_n, 3] <- method_list[row_n]
-    point_estimates[row_n, 4] <- mod$n
-    point_estimates[row_n, 5] <- mod$nevent
-    # odds ratio
-    point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
-
-    # 95% lower bound
-    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
-                                      1.96*(summary(mod)$coefficient[1,3])), 3)
-    # 95% upper bound
-    point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
-                                      1.96*(summary(mod)$coefficient[1,3])), 3)
-    # standard error
-    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
-    # p val
-    point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
-
-    # create else statement that fills matrix with missing so I still have the row
-    # in the final dataframe
-    } else {point_estimates[row_n, 3] <- method_list[row_n]
-            point_estimates[row_n, 4] <- 0
-            point_estimates[row_n, 5] <- 0
-            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
-    } # end methods loop
-
-  # save point estimates as a dataframe
-  point_est_df <- as_data_frame(point_estimates)
-
-  # combine previous values in dataframe that has all outcome/methods comparisons
-  datalist1[[k+1]] <- point_est_df
-} # end sex category loop
-
-# bind rows of sex category estimates together
-combined_point_est_df <- bind_rows(datalist1)
-
-wrf_geo_sex <- combined_point_est_df %>% 
-  # filter columns to just geo-smk
-  # filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
-  mutate(sex_cat = ifelse(sex == 0, "Male", 
-               ifelse(sex == 1, "Female", NA))) %>% 
-  # subset columns I want to put in to the table
-  select(3, 11, 5:8)%>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
-         lower95 = ifelse(lower95 == 99, "--", lower95),
-         upper95 = ifelse(upper95 == 99, "--", upper95))
-
-  tab <- htmlTable(txtRound(wrf_geo_sex, digits = 3, 1:3), 
-           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and pharmacy outcomes stratified by sex",
-           # row group by outcome
-           rgroup = "Asthma inhalers",
-           n.rgroup = c(rep(6, 1)), # 2 rows for each sex strata for each outcome
-           # column headers
-           header = c("Est Method", "Sex Strata", "Events",
-                      "OR&dagger;", "Lower", "Upper"),
-           # column spanner
-           cgroup = c("", "95% CI"), 
-           n.cgroup = c(4, 2),
-           padding.rgroup = "&nbsp;&nbsp;",
-           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
-           align = "llcccc", # column alignment,
-           tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within July to October fire season."
-            ) # end table
-
-  print(tab)
-
-# ggplot of odds ratios, facet wrapped by outcomes -----
-# convert variables from character to either numeric or factor
-# factor preserves the order of the variable
-wrf_geo_sex_plot <- combined_point_est_df %>% 
-  # filter columns to just geo-smk
- #  filter(pm_method == "GWR Smoke" | 
-   #       pm_method == "WRF-Chem Smoke") %>% 
-  mutate(sex_cat = ifelse(sex == 0, "Male", 
-               ifelse(sex == 1, "Female", NA))) %>%
-  # subset columns I want to put in to the table
-  select(1, 3, 11, 5:8) %>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
-         lower95 = ifelse(lower95 == 99, NA, lower95),
-         upper95 = ifelse(upper95 == 99, NA, upper95))
-
-# change characters to numeric and factor  
-wrf_geo_sex_plot$outcome <- factor(wrf_geo_sex_plot$outcome,
-                              levels = unique(wrf_geo_sex_plot$outcome))
-
-wrf_geo_sex_plot$pm_method <- factor(wrf_geo_sex_plot$pm_method,
-                                levels = unique(wrf_geo_sex_plot$pm_method))
-
-wrf_geo_sex_plot$sex_cat <- factor(wrf_geo_sex_plot$sex_cat,
-                                levels = unique(wrf_geo_sex_plot$sex_cat))
-
-wrf_geo_sex_plot$n_events <- as.numeric(wrf_geo_sex_plot$n_events)
-wrf_geo_sex_plot$odds_ratio <- as.numeric(wrf_geo_sex_plot$odds_ratio)
-wrf_geo_sex_plot$lower95 <- as.numeric(wrf_geo_sex_plot$lower95)
-wrf_geo_sex_plot$upper95 <- as.numeric(wrf_geo_sex_plot$upper95)
-
-## ggplot
-  print_plot <- ggplot(wrf_geo_sex_plot,
-    aes(x = sex_cat, y = odds_ratio, color = pm_method), na.rm = T) +
-    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
-    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
-                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
-    # custom color 
-    scale_color_manual(name = "Smoke-Estimation Method", 
-                       values = c("red", "blue", "#32115C"),
-                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
-    facet_wrap(~outcome, nrow = 3, scales = "free") +
-    geom_hline(yintercept = 1, linetype=2) +
-    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
-    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
-    #ylim(0, 2) +
-    xlab('Sex') +
-    # plot theme
-    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    # strip element
-    strip.background = element_rect(colour=NA, fill=NA),
-    panel.border = element_rect(fill = NA, color = "black"),
-    # facet text size
-    strip.text = element_text(size = 10),
-    # axis element
-    #axis.text.x = element_blank(),
-    #axis.title.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.title.y = element_text(angle = 90),
-    # legend elements
-    legend.position = "bottom")
-    #legend.text = element_text(size = 8))
-
-  print(print_plot)
-  # save figure
-  ggsave("../plot_new/ndc_sex_fig_three.pdf", plot = print_plot, 
-       width = 12, height = 8, units = "in")
-
-```
+### 2. For Age Category
+
+### 3. For Sex Category

### 2017-07-11 r_scripts/analysis/oregon_medication.Rmd
distributed casecrossover

@@
 p <- p + theme_bw()
-# p
+p
 
@@
 
-### Lag
-
-### 1. For Overall Plot
+### 1. For Overall Plot (Unconstrained lag model)
 
@@
 
-### 2. For Age Category
-
-### 3. For Sex Category
+### 2. Distributed lag for casecrossover
+
+```{r DLM for casecrossover, echo = FALSE, message=FALSE, warning=FALSE}
+read_path8 <- paste0("../data_new/county_data/zip_to_county_new.csv")
+or_zip_to_county <- read_csv(read_path8)
+
+read_path9 <- paste0("../data_new/county_data/or_census_pop_est_county.csv")
+or_pop <- read_csv(read_path9)
+
+or_pop <- or_pop %>%
+  rename(county = GEO_display_label, pop = respop72013) %>%
+  select(county, pop) 
+or_pop$county <- gsub(" County, Oregon", "", as.character(factor(or_pop$county)))
+
+inhaler_ndc_lag_df <- inhaler_ndc_df %>%
+  left_join(or_zip_to_county, by = "ZIPCODE") %>%
+  left_join(or_pop, by = "county") %>%
+  mutate(day = ifelse(weekdays(date) == "Monday" |
+                      weekdays(date) == "Tuesday" |
+                      weekdays(date) == "Wednesday"|
+                      weekdays(date) == "Thursday"|
+                      weekdays(date) == "Friday", "weekday",
+               ifelse(weekdays(date) == "Saturday" |
+                      weekdays(date) == "Sunday", "weekend", NA))) %>%
+  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
+         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
+         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
+         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
+         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
+         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
+         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
+         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10) %>%
+  filter(!is.na(geo_smk_pm10_lag7_zip))
+
+inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-05-27")] <- "weekend"
+inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-07-04")] <- "weekend"
+inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-09-02")] <- "weekend"
+
+which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_zip") # 159
+which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_lag1_zip") # 160
+which(colnames(inhaler_ndc_lag_df)=="wrf_temp_zip") # 93
+which(colnames(inhaler_ndc_lag_df)=="wrf_temp_lag1_zip") # 150
+
+inhaler_geo_lag <- inhaler_ndc_lag_df %>%
+  select(159, 160:166) # GWR-method
+
+inhaler_temp_lag <- inhaler_ndc_lag_df %>%
+  select(93, 150:156) # GWR-method
+
+lag_max <- 7
+n1 <- length(inhaler_geo_lag[,1])
+n2 <- length(inhaler_temp_lag[,1])
+
+inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
+inhaler_temp_lag <- as.matrix(inhaler_temp_lag)
+
+B <- ns(0:lag_max, df =3, intercept=TRUE) # or 3?
+
+# the nrow B should equal ncol(pm)
+dim(B)
+
+inhaler_geo_lag_B <- inhaler_geo_lag%*%B
+inhaler_temp_lag_B <- inhaler_temp_lag%*%B
+
+fit <- clogit(outcome ~ inhaler_geo_lag_B + inhaler_temp_lag_B + strata(personkey), 
+              inhaler_ndc_lag_df)
+
+coef(fit)
+summary(fit)
+AIC(fit)
+
+D <- cbind(B,B)
+
+# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
+dlparms <- c(1,2,3,4,5,6)
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms]) 
+DLvar <- D%*%vcov(fit)[dlparms,dlparms]%*%t(D)
+DLestimate$SE <- sqrt(diag(DLvar))
+DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
+DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
+
+DLestimate$estimate <- exp(DLestimate$estimate)
+DLestimate$lower <- exp(DLestimate$lower)
+DLestimate$upper <- exp(DLestimate$upper)
+
+DLestimate
+
+DLestimate$lag <- 0:lag_max
+p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
+p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
+p <- p + geom_line(aes(x=lag), linetype=2, size=2)
+p <- p + theme_bw()
+p
+
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms])
+
+cumulative <- sum(DLestimate$estimate)
+cumulative
+
+cumulative_se <- sqrt(sum(DLvar))
+cumulative_se
+
+cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
+cumulative_CI
+
+```

### 2017-07-12 r_scripts/analysis/oregon_medication.Rmd
Update oregon_medication.Rmd

@@
 
-```{r regression model q, message=FALSE, echo=FALSE}
+```{r regression model q, message=FALSE, echo=FALSE, eval=FALSE, warning=FALSE}
 # create empty dataframe
@@
 
-```{r DLM, echo = FALSE, message=FALSE, warning=FALSE}
+```{r DLM, echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}
 read_path8 <- paste0("../data_new/county_data/zip_to_county_new.csv")
@@
 
-```{r, message=FALSE, echo=FALSE}
+```{r, message=FALSE, echo=FALSE, eval=FALSE}
 # create fit matrix
@@
 
-```{r compare mixed and fixed, message=FALSE, echo=FALSE}
+```{r compare mixed and fixed, message=FALSE, echo=FALSE, eval=FALSE}
 
@@
     scale_color_manual(name = "Geo-smk Lag", 
-                       values = c("red", "orange", "yellow", "green", "blue", "#32115C", "purple"),
+                       values = rep(c("black"), 7),
                        guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
@@
          geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
-         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10) %>%
-  filter(!is.na(geo_smk_pm10_lag7_zip))
+         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)
 
@@
 
-B <- ns(0:lag_max, df =3, intercept=TRUE) # or 3?
+B <- ns(0:lag_max, df =3, intercept=TRUE)
 
@@
 ```
+
+```{r DLM for casecrossover temp, echo = FALSE, message=FALSE, warning=FALSE}
+read_path8 <- paste0("../data_new/county_data/zip_to_county_new.csv")
+or_zip_to_county <- read_csv(read_path8)
+
+read_path9 <- paste0("../data_new/county_data/or_census_pop_est_county.csv")
+or_pop <- read_csv(read_path9)
+
+or_pop <- or_pop %>%
+  rename(county = GEO_display_label, pop = respop72013) %>%
+  select(county, pop) 
+or_pop$county <- gsub(" County, Oregon", "", as.character(factor(or_pop$county)))
+
+inhaler_ndc_lag_df <- inhaler_ndc_df %>%
+  left_join(or_zip_to_county, by = "ZIPCODE") %>%
+  left_join(or_pop, by = "county") %>%
+  mutate(day = ifelse(weekdays(date) == "Monday" |
+                      weekdays(date) == "Tuesday" |
+                      weekdays(date) == "Wednesday"|
+                      weekdays(date) == "Thursday"|
+                      weekdays(date) == "Friday", "weekday",
+               ifelse(weekdays(date) == "Saturday" |
+                      weekdays(date) == "Sunday", "weekend", NA))) %>%
+  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
+         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
+         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
+         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
+         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
+         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
+         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
+         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)
+
+inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-05-27")] <- "weekend"
+inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-07-04")] <- "weekend"
+inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-09-02")] <- "weekend"
+
+which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_zip") # 159
+which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_lag1_zip") # 160
+which(colnames(inhaler_ndc_lag_df)=="wrf_temp_zip") # 93
+which(colnames(inhaler_ndc_lag_df)=="wrf_temp_lag1_zip") # 150
+
+inhaler_geo_lag <- inhaler_ndc_lag_df %>%
+  select(159, 160:166) # GWR-method
+
+inhaler_temp_lag <- inhaler_ndc_lag_df %>%
+  select(93, 150:156) # GWR-method
+
+lag_max <- 7
+n1 <- length(inhaler_geo_lag[,1])
+n2 <- length(inhaler_temp_lag[,1])
+
+inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
+inhaler_temp_lag <- as.matrix(inhaler_temp_lag)
+
+B <- ns(0:lag_max, df =3, intercept=TRUE)
+
+# the nrow B should equal ncol(pm)
+dim(B)
+
+inhaler_geo_lag_B <- inhaler_geo_lag%*%B
+inhaler_temp_lag_B <- inhaler_temp_lag%*%B
+
+fit <- clogit(outcome ~  inhaler_temp_lag_B + strata(personkey), 
+              inhaler_ndc_lag_df)
+
+coef(fit)
+summary(fit)
+AIC(fit)
+
+D <- cbind(B,B)
+
+# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
+dlparms <- c(1,2,3)
+DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms]) 
+DLvar <- B%*%vcov(fit)[dlparms,dlparms]%*%t(B)
+DLestimate$SE <- sqrt(diag(DLvar))
+DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
+DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
+
+DLestimate$estimate <- exp(DLestimate$estimate)
+DLestimate$lower <- exp(DLestimate$lower)
+DLestimate$upper <- exp(DLestimate$upper)
+
+DLestimate
+
+DLestimate$lag <- 0:lag_max
+p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
+p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
+p <- p + geom_line(aes(x=lag), linetype=2, size=2)
+p <- p + theme_bw()
+p
+
+DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms])
+
+cumulative <- sum(DLestimate$estimate)
+cumulative
+
+cumulative_se <- sqrt(sum(DLvar))
+cumulative_se
+
+cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
+cumulative_CI
+
+```
+
+### 3. Unconstrained lag for time series
+
+```{r unconstrained ts lag, message=FALSE, echo=FALSE}
+or_ndc_county_ts_df <- or_ndc_county_ts %>%
+  select(1:4, pop, rate_per_100k, day, geo_smk_pm_county,
+         geo_smk_pm_lag1_county:geo_smk_pm_lag7_county, wrf_temp_county,
+         wrf_temp_lag1_county:wrf_temp_lag7_county) %>%
+  mutate(geo_smk_pm10_county = geo_smk_pm_county/10,
+         geo_smk_pm10_lag1_county =geo_smk_pm_lag1_county/10,
+         geo_smk_pm10_lag2_county =geo_smk_pm_lag2_county/10,
+         geo_smk_pm10_lag3_county =geo_smk_pm_lag3_county/10,
+         geo_smk_pm10_lag4_county =geo_smk_pm_lag4_county/10,
+         geo_smk_pm10_lag5_county =geo_smk_pm_lag5_county/10,
+         geo_smk_pm10_lag6_county =geo_smk_pm_lag6_county/10,
+         geo_smk_pm10_lag7_county =geo_smk_pm_lag7_county/10) 
+  # filter(!is.na(geo_smk_pm10_lag7_zip))
+
+# create empty dataframe
+rr_df <- data.frame(matrix(nrow = 7, ncol = 5))
+colnames(rr_df) <- c("method", "lag", "rel_risk", "lower_95", "upper_95")
+
+# vector of outcomes
+outcome_list <- "n_obs"
+
+# methods
+rr_df[,1] <- c("Mixed-Poisson")
+
+# populate outcome column
+rr_df[,2] <- c("Lag 1", "Lag 2", "Lag 3", "Lag 4", "Lag 5", "Lag 6", "Lag 7")
+
+# which(colnames(or_ndc_county_ts_df)=="geo_smk_pm10_lag1_county") #25
+# which(colnames(or_ndc_county_ts_df)=="wrf_temp_lag1_county") #17
+
+for (i in 1:7){
+
+  geo <- i + 24
+  tem <- i +14
+
+## mixed
+smk_est_mp <- tidy(glmer(as.formula(paste0(outcome_list, 
+    "~ or_ndc_county_ts_df[[geo]] + or_ndc_county_ts_df[[tem]] + (1|county) + day + offset(log(pop))")), 
+    or_ndc_county_ts_df, family="poisson"))[2, ]
+
+rr_mp <- c(round(exp(smk_est_mp[1,2]),5),
+         round(exp(smk_est_mp[1,2]-1.96*smk_est_mp[1,3]),5),
+         round(exp(smk_est_mp[1,2]+1.96*smk_est_mp[1,3]),5))
+
+rr_df[i,3:5] <- rr_mp
+}
+rr_df
+
+plot_rr <- ggplot(rr_df, aes(x=lag, y = rel_risk)) +
+  geom_point() +
+  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
+  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
+  theme(
+   panel.background = element_rect(fill = 'white', colour = 'black'),
+   panel.grid.major = element_blank(),
+   panel.grid.minor = element_blank(),
+   axis.text.x = element_text(angle = 90, vjust = .75)) +
+  ylab("Relative Risk") +
+  xlab("Outcome") +
+  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")
+
+print(plot_rr)
+
+ggsave("../plot_new/ndc_rr_geo_lag.pdf", plot = plot_rr, 
+       width = 12, height = 8, units = "in")
+
+```
+
+### 4. Distributed lag for time series
+
+```{r DLM ts, echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}
+
+which(colnames(or_ndc_county_ts_df)=="wrf_temp_lag1_county") # 17
+which(colnames(or_ndc_county_ts_df)=="geo_smk_pm10_lag1_county") # 25
+
+geo_ts_lag <- or_ndc_county_ts_df %>%
+  select(24:31) # GWR-method
+tem_ts_lag <- or_ndc_county_ts_df %>%
+  select(16:23) # temp
+
+lag_max <- 7
+
+geo_ts_lag <- as.matrix(geo_ts_lag)
+tem_ts_lag <- as.matrix(tem_ts_lag)
+
+B <- ns(0:lag_max, df =3, intercept=TRUE) # or 3?
+
+# the nrow B should equal ncol(pm)
+dim(B)
+dim(geo_ts_lag)
+dim(tem_ts_lag)
+
+geo_ts_lag_B <- geo_ts_lag%*%B
+
+tem_ts_lag_B <- tem_ts_lag%*%B
+
+## mixed model
+mod <-glmer(outcome ~  geo_ts_lag_B + tem_ts_lag_B + day + (1|county) + offset(log(pop)),   
+            or_ndc_county_ts_df, family="poisson", control = glmerControl(optimizer = "bobyqa"))
+
+summary(mod)
+# check AIC
+AIC(mod)
+# output distributed lag beta parameters
+dlparms <- mod@beta[2:4]
+# estimate distributed lag values for each day
+dl_estimates <- data.frame(estimate = B %*% dlparms)
+
+# covariance matrix for knots (need to convert to matrix object)
+cov_mat <- as.matrix(vcov(mod)[2:4,2:4])
+
+# estimate variance of splines
+dl_var <- B%*%cov_mat%*%t(B)
+
+# calculate standard error for each lag value
+dl_estimates$stderr <- sqrt(diag(dl_var))
+
+# calculate lower and upper bounds
+dl_estimates$lower_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(1-0.975,
+  df=df.residual(mod))  
+dl_estimates$upper_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(0.975,
+  df=df.residual(mod))
+
+# estimate relative risk and 95%CI from beta
+inhaler_dl_rr <- dl_estimates %>% 
+  mutate(estimate = exp(estimate),
+         lower_bound = exp(lower_bound),
+         upper_bound = exp(upper_bound),
+         lag = 0:7)
+
+# estimate cumulative effect
+cumulative <- sum(dl_estimates$estimate)
+cumulative
+
+# estimate cumulative effect stnd error
+cumulative_se <- sqrt(sum(dl_var))
+cumulative_se
+
+# estimate cumulative CI
+cumulative_ci <- cumulative+cumulative_se*qt(c(1-0.975,0.975), 
+                                             df=df.residual(mod))
+cumulative_ci
+
+exp(cumulative)
+exp(cumulative_ci)
+
+# plot 
+p <- ggplot(inhaler_dl_rr, aes(x=lag, y=estimate, 
+                              ymin=lower_bound, ymax=upper_bound)) +
+  geom_ribbon(alpha = 0.5) +
+  geom_line(size=2)
+
+p
+```

### 2017-07-13 r_scripts/analysis/oregon_medication.Rmd
Update oregon_medication.Rmd

@@
 
-## Time-Stratified Case-Crossover (Population-Weighted PM~2.5~ Assigned by Zip Code)
+### Time-Stratified Case-Crossover (Population-Weighted PM~2.5~ Assigned by Zip Code)
 *referents same day of week within fire season May - September*
@@
 
-## Time stratified for time series (Population-Weighted PM~2.5~ Assigned by County)
+### Time stratified for time series (Population-Weighted PM~2.5~ Assigned by County)
+
+I also make time stratified time series data set.
+The original data set only contains the zip code, but for time series, we want to research on the county level.
+Therefore, I assign each zip code with one county which I found online.
+There are 488 zipcodes in total (4 are missing but I use google to find it).
 
@@
 
+## 1. smoke exposure time series and map of smoke days in Oregon
+
+### Prepare (already done in previous time stratified work)
+
+1. Make 10 units' geo smoke variable
+2. Make county population variable (to calculate rate)
+3. Make weekend/weekday variables
+
+```{r time series, warning=FALSE, message=FALSE, echo=FALSE}
+read_path5 <- paste0("../data_new/medication/or_ndc_county_time_series.csv")
+or_ndc_county_ts <- read_csv(read_path5)
+
+plot_ts <- ggplot(or_ndc_county_ts, aes(y =rate_per_100k, x = date)) +
+           geom_jitter()
+
+print(plot_ts)
+
+# save figure
+ggsave("../plot_new/ndc_rate_ts.pdf", plot = plot_ts, 
+       width = 12, height = 8, units = "in")
+
+plot_county <- ggplot(or_ndc_county_ts, aes(y =rate_per_100k, x = date)) +
+               geom_jitter() + facet_wrap(~county)
+
+plot_county_line <- ggplot(or_ndc_county_ts, aes(y =rate_per_100k, x = date)) +
+                    facet_wrap(~county) + geom_line()
+
+print(plot_county)
+# print(plot_county_line)
+# save figure
+ggsave("../plot_new/ndc_county_rate_ts.pdf", plot = plot_county, 
+       width = 12, height = 8, units = "in")
+ggsave("../plot_new/ndc_county_rate_ts_line.pdf", plot = plot_county_line, 
+       width = 12, height = 8, units = "in")
+
+### check if the time series has the association with get smoke
+a <- glm(rate_per_100k ~ geo_smk_pm10, data = or_ndc_county_ts, family = "poisson")
+summary(a)
+
+plot_health_smk <- ggplot(or_ndc_county_ts, aes(x = geo_smk_pm10, y = log(rate_per_100k+0.1))) +
+  geom_point(aes(colour = c("red")), alpha = 0.2,  size = 1) + 
+  # scale_colour_manual(values = c("red", "blue")) +
+  # truncated y axis
+  #scale_y_continuous(limits = c(0, 1.5)) +
+  # facet_wrap(~county, scale  = "free") +
+  ylab("Log Outcome Rate per 100,000 Persons") +
+  xlab("GWR Smoke PM2.5 per 10*10ug/m^3") +
+  theme_bw()
+
+# unnecessary
+plot_county_health_smk <- ggplot(or_ndc_county_ts, aes(x = geo_smk_pm10, y = log(rate_per_100k+0.1))) +
+  geom_point(aes(colour = county), alpha = 0.2,  size = 1) + 
+  # scale_colour_manual(values = c("red", "blue")) +
+  # truncated y axis
+  #scale_y_continuous(limits = c(0, 1.5)) +
+  facet_wrap(~county, scale  = "free") +
+  ylab("Log Outcome Rate per 100,000 Persons") +
+  xlab("GWR Smoke PM2.5 per 10*10ug/m^3") +
+  theme_bw()
+
+plot(plot_health_smk)
+# plot(plot_county_health_smk)
+
+ggsave("../plot_new/ndc_rate_health_smk.pdf", plot = plot_health_smk, 
+       width = 12, height = 8, units = "in")
+
+ggsave("../plot_new/ndc_county_rate_health_smk.pdf", plot = plot_county_health_smk, 
+       width = 12, height = 8, units = "in")
+
+```
+
+## 2. descriptive table of total ndc inhalhers; and age and sex category; Plots
+
 ### Descriptives
@@
 
+# dataframe to loop through
+inhaler_ndc_df <- data.frame(inhaler_ndc)
+
 # dataframe list
 method_list <- c('WRF-Chem Smoke', 'Geo-Weighted Smoke', 'Kriging Smoke')
-
-# dataframe to loop through
-inhaler_ndc_df <- data.frame(inhaler_ndc)
 
@@
 
-## County-level time series
-
-### County Time series plots of rate
-
-### Prepare (already done in previous time stratified work)
-
-1. Make 10 units' geo smoke variable
-2. Make county population variable (to calculate rate)
-3. Make weekend/weekday variables
-
-```{r time series, warning=FALSE, message=FALSE, echo=FALSE}
-read_path5 <- paste0("../data_new/medication/or_ndc_county_time_series.csv")
-or_ndc_county_ts <- read_csv(read_path5)
-
-plot_ts <- ggplot(or_ndc_county_ts, aes(y =rate_per_100k, x = date)) +
-           geom_jitter()
-
-print(plot_ts)
-
-# save figure
-ggsave("../plot_new/ndc_rate_ts.pdf", plot = plot_ts, 
-       width = 12, height = 8, units = "in")
-
-plot_county <- ggplot(or_ndc_county_ts, aes(y =rate_per_100k, x = date)) +
-               geom_jitter() + facet_wrap(~county)
-
-plot_county_line <- ggplot(or_ndc_county_ts, aes(y =rate_per_100k, x = date)) +
-                    facet_wrap(~county) + geom_line()
-
-print(plot_county)
-print(plot_county_line)
-# save figure
-ggsave("../plot_new/ndc_county_rate_ts.pdf", plot = plot_county, 
-       width = 12, height = 8, units = "in")
-ggsave("../plot_new/ndc_county_rate_ts_line.pdf", plot = plot_county_line, 
-       width = 12, height = 8, units = "in")
-
-### check if the time series has the association with get smoke
-a <- glm(rate_per_100k ~ geo_smk_pm10, data = or_ndc_county_ts, family = "poisson")
-summary(a)
-
-plot_health_smk <- ggplot(or_ndc_county_ts, aes(x = geo_smk_pm10, y = log(rate_per_100k+0.1))) +
-  geom_point(aes(colour = c("red")), alpha = 0.2,  size = 1) + 
-  # scale_colour_manual(values = c("red", "blue")) +
-  # truncated y axis
-  #scale_y_continuous(limits = c(0, 1.5)) +
-  # facet_wrap(~county, scale  = "free") +
-  ylab("Log Outcome Rate per 100,000 Persons") +
-  xlab("GWR Smoke PM2.5 per 10*10ug/m^3") +
-  theme_bw()
-
-# unnecessary
-plot_county_health_smk <- ggplot(or_ndc_county_ts, aes(x = geo_smk_pm10, y = log(rate_per_100k+0.1))) +
-  geom_point(aes(colour = county), alpha = 0.2,  size = 1) + 
-  # scale_colour_manual(values = c("red", "blue")) +
-  # truncated y axis
-  #scale_y_continuous(limits = c(0, 1.5)) +
-  facet_wrap(~county, scale  = "free") +
-  ylab("Log Outcome Rate per 100,000 Persons") +
-  xlab("GWR Smoke PM2.5 per 10*10ug/m^3") +
-  theme_bw()
-
-plot(plot_health_smk)
-plot(plot_county_health_smk)
-
-ggsave("../plot_new/ndc_rate_health_smk.pdf", plot = plot_health_smk, 
-       width = 12, height = 8, units = "in")
-
-ggsave("../plot_new/ndc_county_rate_health_smk.pdf", plot = plot_county_health_smk, 
-       width = 12, height = 8, units = "in")
-
-```
-
-### Standard Regression Model (Relative Risk)
-
-1. Fixed-Quasipoisson
-2. Fixed-Poisson
-3. Mixed-Poisson
-
-```{r regression model q, message=FALSE, echo=FALSE, eval=FALSE, warning=FALSE}
-# create empty dataframe
-rr_df <- data.frame(matrix(nrow = 3, ncol = 5))
-colnames(rr_df) <- c("method", "outcome", "rel_risk", "lower_95", "upper_95")
-
-# vector of outcomes
-outcome_list <- "n_obs"
-
-# methods
-rr_df[,1] <- c("Fixed-Quasipoisson", "Fixed-Poisson", "Mixed-Poisson")
-
-# names 
-outcome_names <- c('Inhalers')
-
-# populate outcome column
-rr_df[,2] <- factor(outcome_names, levels=unique(outcome_names))
-
-## quasipoisson
-smk_est_fq <- tidy(glm(as.formula(paste0(outcome_list, 
-    "~ geo_smk_pm10 + wrf_temp_county + county + 
-    day + offset(log(pop))")), 
-    or_ndc_county_ts, family="quasipoisson"))[2, ]
-
-rr_fq <- c(round(exp(smk_est_fq[1,2]),5),
-         round(exp(smk_est_fq[1,2]-1.96*smk_est_fq[1,3]),5),
-         round(exp(smk_est_fq[1,2]+1.96*smk_est_fq[1,3]),5))
-
-rr_df[1,3:5] <- rr_fq
-
-## poisson
-smk_est_fp <- tidy(glm(as.formula(paste0(outcome_list, 
-    "~ geo_smk_pm10 + wrf_temp_county + county + 
-    day + offset(log(pop))")), 
-    or_ndc_county_ts, family="poisson"))[2, ]
-
-rr_fp <- c(round(exp(smk_est_fp[1,2]),5),
-         round(exp(smk_est_fp[1,2]-1.96*smk_est_fp[1,3]),5),
-         round(exp(smk_est_fp[1,2]+1.96*smk_est_fp[1,3]),5))
-
-rr_df[2,3:5] <- rr_fp
-
-## mixed
-smk_est_mp <- tidy(glmer(as.formula(paste0(outcome_list, 
-    "~ geo_smk_pm10 + wrf_temp_county + (1|county) + 
-    day + offset(log(pop))")), 
-    or_ndc_county_ts, family="poisson"))[2, ]
-
-rr_mp <- c(round(exp(smk_est_mp[1,2]),5),
-         round(exp(smk_est_mp[1,2]-1.96*smk_est_mp[1,3]),5),
-         round(exp(smk_est_mp[1,2]+1.96*smk_est_mp[1,3]),5))
-
-rr_df[3,3:5] <- rr_mp
-
-rr_df
-
-plot_rr <- ggplot(rr_df, aes(x=method, y = rel_risk)) +
-  geom_point() +
-  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
-  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
-  theme(
-   panel.background = element_rect(fill = 'white', colour = 'black'),
-   panel.grid.major = element_blank(),
-   panel.grid.minor = element_blank(),
-   axis.text.x = element_text(angle = 90, vjust = .75)) +
-  ylab("Relative Risk") +
-  xlab("Outcome") +
-  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")
-
-print(plot_rr)
-
-ggsave("../plot_new/ndc_rr_geo.pdf", plot = plot_rr, 
-       width = 12, height = 8, units = "in")
-
-```
-
-## Distributed Lag Model
-
-### Lag 1
+## 3. Distributed Lag Model
 
 Remove the missing values for GWR smoke. 
-But I use GWR smoke of zip for the case crossover study.
-
-```{r DLM, echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}
+But I only use GWR smoke of zip for the case crossover study.
+
+### Find the smallest of AIC with df from 3 to 6
+
+Choose df = 5 because it has the least AIC.
+
+```{r, message=FALSE, echo=FALSE, warning=FALSE}
 read_path8 <- paste0("../data_new/county_data/zip_to_county_new.csv")
@@
 
-which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_zip") # 159
-which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_lag1_zip") # 160
-
-inhaler_ndc_lag <- inhaler_ndc_lag_df %>%
+# which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_zip") # 159
+# which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_lag1_zip") # 160
+# which(colnames(inhaler_ndc_lag_df)=="wrf_temp_zip") # 93
+# which(colnames(inhaler_ndc_lag_df)=="wrf_temp_lag1_zip") # 150
+
+inhaler_geo_lag <- inhaler_ndc_lag_df %>%
   select(159, 160:166) # GWR-method
 
+inhaler_temp_lag <- inhaler_ndc_lag_df %>%
+  select(93, 150:156) # GWR-method
+
+inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
+inhaler_temp_lag <- as.matrix(inhaler_temp_lag)
+
+# create fit matrix
+fit_mat <- matrix(NA, nrow=4, ncol=5)
+colnames(fit_mat) <- c("outcome", "exposure", "fit", "df", "aic")
+
+# fill outcome name in first column
+fit_mat[,1] <- "inhalers"
+# fill in exposure (geo smoke)
+fit_mat[,2] <- "geo_smk10"
+# fit type
+fit_mat[,3] <- "ns"
+
+# 2 df will not work for a spline, going from 3 to 6 df
+  for(h in 3:6){ # start loop for df
+
+    # fitting a natural spline for lag with 5 knots?
+    fit_mat[h-2, 4] <- degree_freedom <- h
+
+    # define basis b using natural spline function
+    B <- ns(0:7, df = degree_freedom, intercept = T)
+
+    # multiply lagged pm matrix by basis
+    inhaler_geo_lag_B <- inhaler_geo_lag%*%B
+    inhaler_temp_lag_B <- inhaler_temp_lag%*%B
+
+    fit <- clogit(outcome ~ inhaler_geo_lag_B + inhaler_temp_lag_B + strata(personkey), 
+              inhaler_ndc_lag_df)
+
+    AIC(fit)
+
+    # fill AIC
+    fit_mat[h-2,5] <- round(AIC(fit),5)
+
+    } # end model fit loop
+
+# now that we have a range of df/knots, I want to filter to the minimum aic
+# for each outcome and print out that value in a table
+
+fit_mat
+
+lag_spline_best_fit <- as_tibble(fit_mat) %>%  
+  slice(which.min(aic)) # df of 3 has the smallest AIC
+
+# kable
+knitr::kable(lag_spline_best_fit, caption = paste0("Distributed lag spline",
+  "degree of freedom best fit by AIC for mixed model approach"))
+```
+
+### Distributed lag 0-7 for casecrossover
+
+```{r DLM for casecrossover, echo = FALSE, message=FALSE, warning=FALSE}
+inhaler_geo_lag <- inhaler_ndc_lag_df %>%
+  select(159, 160:166) # GWR-method
+
+inhaler_temp_lag <- inhaler_ndc_lag_df %>%
+  select(93, 150:156) # GWR-method
+
 lag_max <- 7
-n <- length(inhaler_ndc_lag[,1])
-
-inhaler_ndc_lag <- as.matrix(inhaler_ndc_lag)
-
-B <- ns(0:lag_max, df =3, intercept=TRUE) # or 3?
+n1 <- length(inhaler_geo_lag[,1])
+n2 <- length(inhaler_temp_lag[,1])
+
+inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
+inhaler_temp_lag <- as.matrix(inhaler_temp_lag)
+
+B <- ns(0:lag_max, df =5, intercept=TRUE)
 
@@
 dim(B)
-dim(inhaler_ndc_lag)
-
-inhaler_ndc_lag_B <- inhaler_ndc_lag%*%B
-fit <- clogit(outcome ~ inhaler_ndc_lag_B + wrf_temp_zip + strata(personkey), inhaler_ndc_lag_df)
+
+inhaler_geo_lag_B <- inhaler_geo_lag%*%B
+inhaler_temp_lag_B <- inhaler_temp_lag%*%B
+
+fit <- clogit(outcome ~ inhaler_geo_lag_B + inhaler_temp_lag_B + strata(personkey), 
+              inhaler_ndc_lag_df)
 
@@
 summary(fit)
-
-dlparms <- grep("inhaler_ndc_lag_B",names(coef(fit)))
-DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms])  
-DLvar <- B%*%vcov(fit)[dlparms,dlparms]%*%t(B)
+AIC(fit)
+
+D <- cbind(B,B)
+
+# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
+dlparms <- c(1,2,3,4,5,6,7,8,9,10)
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms]) 
+DLvar <- D%*%vcov(fit)[dlparms,dlparms]%*%t(D)
 DLestimate$SE <- sqrt(diag(DLvar))
@@
 
-DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms])
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms])
 
@@
 cumulative
-## [1] 0.009141771
+
 cumulative_se <- sqrt(sum(DLvar))
 cumulative_se
-## [1] 0.000955469
+
 cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
 cumulative_CI
-## [1] 0.007269052 0.011014490
-
-## mixed model
-mod <-glmer(outcome ~  inhaler_ndc_lag_B + wrf_temp_zip + day + (1|county) + offset(log(pop)),   
-            inhaler_ndc_lag_df, family="poisson", control = glmerControl(optimizer = "bobyqa"))
-
-summary(mod)
-# check AIC
-AIC(mod)
-# output distributed lag beta parameters
-dlparms <- mod@beta[2:4]
-# estimate distributed lag values for each day
-dl_estimates <- data.frame(estimate = B %*% dlparms)
-
-# covariance matrix for knots (need to convert to matrix object)
-cov_mat <- as.matrix(vcov(mod)[2:4,2:4])
-
-# estimate variance of splines
-dl_var <- B%*%cov_mat%*%t(B)
-
-# calculate standard error for each lag value
-dl_estimates$stderr <- sqrt(diag(dl_var))
-
-# calculate lower and upper bounds
-dl_estimates$lower_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(1-0.975,
-  df=df.residual(mod))  
-dl_estimates$upper_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(0.975,
-  df=df.residual(mod))
-
-# estimate relative risk and 95%CI from beta
-inhaler_dl_rr <- dl_estimates %>% 
-  mutate(estimate = exp(estimate),
-         lower_bound = exp(lower_bound),
-         upper_bound = exp(upper_bound),
-         lag = 0:7)
-
-# estimate cumulative effect
-cumulative <- sum(dl_estimates$estimate)
+
+```
+
+```{r DLM for casecrossover temp, echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}
+read_path8 <- paste0("../data_new/county_data/zip_to_county_new.csv")
+or_zip_to_county <- read_csv(read_path8)
+
+read_path9 <- paste0("../data_new/county_data/or_census_pop_est_county.csv")
+or_pop <- read_csv(read_path9)
+
+or_pop <- or_pop %>%
+  rename(county = GEO_display_label, pop = respop72013) %>%
+  select(county, pop) 
+or_pop$county <- gsub(" County, Oregon", "", as.character(factor(or_pop$county)))
+
+inhaler_ndc_lag_df <- inhaler_ndc_df %>%
+  left_join(or_zip_to_county, by = "ZIPCODE") %>%
+  left_join(or_pop, by = "county") %>%
+  mutate(day = ifelse(weekdays(date) == "Monday" |
+                      weekdays(date) == "Tuesday" |
+                      weekdays(date) == "Wednesday"|
+                      weekdays(date) == "Thursday"|
+                      weekdays(date) == "Friday", "weekday",
+               ifelse(weekdays(date) == "Saturday" |
+                      weekdays(date) == "Sunday", "weekend", NA))) %>%
+  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
+         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
+         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
+         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
+         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
+         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
+         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
+         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)
+
+inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-05-27")] <- "weekend"
+inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-07-04")] <- "weekend"
+inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-09-02")] <- "weekend"
+
+which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_zip") # 159
+which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_lag1_zip") # 160
+which(colnames(inhaler_ndc_lag_df)=="wrf_temp_zip") # 93
+which(colnames(inhaler_ndc_lag_df)=="wrf_temp_lag1_zip") # 150
+
+inhaler_geo_lag <- inhaler_ndc_lag_df %>%
+  select(159, 160:166) # GWR-method
+
+inhaler_temp_lag <- inhaler_ndc_lag_df %>%
+  select(93, 150:156) # GWR-method
+
+lag_max <- 7
+n1 <- length(inhaler_geo_lag[,1])
+n2 <- length(inhaler_temp_lag[,1])
+
+inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
+inhaler_temp_lag <- as.matrix(inhaler_temp_lag)
+
+B <- ns(0:lag_max, df =3, intercept=TRUE)
+
+# the nrow B should equal ncol(pm)
+dim(B)
+
+inhaler_geo_lag_B <- inhaler_geo_lag%*%B
+inhaler_temp_lag_B <- inhaler_temp_lag%*%B
+
+fit <- clogit(outcome ~  inhaler_temp_lag_B + strata(personkey), 
+              inhaler_ndc_lag_df)
+
+coef(fit)
+summary(fit)
+AIC(fit)
+
+D <- cbind(B,B)
+
+# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
+dlparms <- c(1,2,3)
+DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms]) 
+DLvar <- B%*%vcov(fit)[dlparms,dlparms]%*%t(B)
+DLestimate$SE <- sqrt(diag(DLvar))
+DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
+DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
+
+DLestimate$estimate <- exp(DLestimate$estimate)
+DLestimate$lower <- exp(DLestimate$lower)
+DLestimate$upper <- exp(DLestimate$upper)
+
+DLestimate
+
+DLestimate$lag <- 0:lag_max
+p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
+p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
+p <- p + geom_line(aes(x=lag), linetype=2, size=2)
+p <- p + theme_bw()
+p
+
+DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms])
+
+cumulative <- sum(DLestimate$estimate)
 cumulative
 
-# estimate cumulative effect stnd error
-cumulative_se <- sqrt(sum(dl_var))
+cumulative_se <- sqrt(sum(DLvar))
 cumulative_se
 
-# estimate cumulative CI
-cumulative_ci <- cumulative+cumulative_se*qt(c(1-0.975,0.975), 
-                                             df=df.residual(mod))
-cumulative_ci
-
-exp(cumulative)
-exp(cumulative_ci)
-
-# plot 
-p <- ggplot(inhaler_dl_rr, aes(x=lag, y=estimate, 
-                              ymin=lower_bound, ymax=upper_bound)) +
-  geom_ribbon(alpha = 0.5) +
-  geom_line(size=2)
-
-p
+cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
+cumulative_CI
+
 ```
 
-### Find the smallest of AIC with df from 3 to 6
-
-```{r, message=FALSE, echo=FALSE, eval=FALSE}
+### Distributed lag -7 to 0 for casecrossover
+
+```{r time stratified zip for casecrossover lag -7 to 0 prepare, eval=FALSE, echo=FALSE}
+### Time stratified ------------------------------------------------------------
+# read in zipcode level populatoin-weighted pm
+read_path5 <- paste0('C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire_new/data/Oregon_PM/zip_pm_to_merge_with_acap.csv')
+
+zip_smoke <- read_csv(read_path5) # 63801 rows
+
+# descriptives of the two smoke datasets
+summary(zip_smoke)
+
+# Zipcode PM2.5 estimates
+# create lag variables that take smoke values from n previous days for zipcodes
+zip_smoke_w_lag <- zip_smoke %>% arrange(ZIPCODE, date) %>%
+  # group by zipcode
+  group_by(ZIPCODE) %>% 
+  # wrf
+  mutate(wrf_f_pm_lag1 = lag(wrf_f_pm, 1, order_by = ZIPCODE), 
+         wrf_f_pm_lag2 = lag(wrf_f_pm, 2, order_by = ZIPCODE),
+         wrf_f_pm_lag3 = lag(wrf_f_pm, 3, order_by = ZIPCODE),
+         wrf_f_pm_lag4 = lag(wrf_f_pm, 4, order_by = ZIPCODE),
+         wrf_f_pm_lag5 = lag(wrf_f_pm, 5, order_by = ZIPCODE),
+         wrf_f_pm_lag6 = lag(wrf_f_pm, 6, order_by = ZIPCODE),
+         wrf_f_pm_lag7 = lag(wrf_f_pm, 7, order_by = ZIPCODE),
+         # wrf no fire lag
+         wrf_nf_pm_lag1 = lag(wrf_nf_pm, 1, order_by = ZIPCODE),
+         wrf_nf_pm_lag2 = lag(wrf_nf_pm, 2, order_by = ZIPCODE),
+         wrf_nf_pm_lag3 = lag(wrf_nf_pm, 3, order_by = ZIPCODE),
+         wrf_nf_pm_lag4 = lag(wrf_nf_pm, 4, order_by = ZIPCODE),
+         wrf_nf_pm_lag5 = lag(wrf_nf_pm, 5, order_by = ZIPCODE),
+         wrf_nf_pm_lag6 = lag(wrf_nf_pm, 6, order_by = ZIPCODE),
+         wrf_nf_pm_lag7 = lag(wrf_nf_pm, 7, order_by = ZIPCODE),
+         # wrf_smk_pm
+         wrf_smk_pm_lag1 = lag(wrf_smk_pm, 1, order_by = ZIPCODE),
+         wrf_smk_pm_lag2 = lag(wrf_smk_pm, 2, order_by = ZIPCODE),
+         wrf_smk_pm_lag3 = lag(wrf_smk_pm, 3, order_by = ZIPCODE),
+         wrf_smk_pm_lag4 = lag(wrf_smk_pm, 4, order_by = ZIPCODE),
+         wrf_smk_pm_lag5 = lag(wrf_smk_pm, 5, order_by = ZIPCODE),
+         wrf_smk_pm_lag6 = lag(wrf_smk_pm, 6, order_by = ZIPCODE),
+         wrf_smk_pm_lag7 = lag(wrf_smk_pm, 7, order_by = ZIPCODE),
+         # geo weighted pm
+         geo_wt_pm_lag1 = lag(geo_wt_pm, 1, order_by = ZIPCODE),
+         geo_wt_pm_lag2 = lag(geo_wt_pm, 2, order_by = ZIPCODE),
+         geo_wt_pm_lag3 = lag(geo_wt_pm, 3, order_by = ZIPCODE),
+         geo_wt_pm_lag4 = lag(geo_wt_pm, 4, order_by = ZIPCODE),
+         geo_wt_pm_lag5 = lag(geo_wt_pm, 5, order_by = ZIPCODE),
+         geo_wt_pm_lag6 = lag(geo_wt_pm, 6, order_by = ZIPCODE),
+         geo_wt_pm_lag7 = lag(geo_wt_pm, 7, order_by = ZIPCODE),
+         # krig pm
+         krig_pm_lag1 = lag(krig_pm, 1, order_by = ZIPCODE),
+         krig_pm_lag2 = lag(krig_pm, 2, order_by = ZIPCODE),
+         krig_pm_lag3 = lag(krig_pm, 3, order_by = ZIPCODE),
+         krig_pm_lag4 = lag(krig_pm, 4, order_by = ZIPCODE),
+         krig_pm_lag5 = lag(krig_pm, 5, order_by = ZIPCODE), 
+         krig_pm_lag6 = lag(krig_pm, 6, order_by = ZIPCODE),
+         krig_pm_lag7 = lag(krig_pm, 7, order_by = ZIPCODE), 
+         # background pm
+         background_pm_lag1 = lag(background_pm, 1, order_by = ZIPCODE),
+         background_pm_lag2 = lag(background_pm, 2, order_by = ZIPCODE),
+         background_pm_lag3 = lag(background_pm, 3, order_by = ZIPCODE),
+         background_pm_lag4 = lag(background_pm, 4, order_by = ZIPCODE),
+         background_pm_lag5 = lag(background_pm, 5, order_by = ZIPCODE), 
+         background_pm_lag6 = lag(background_pm, 6, order_by = ZIPCODE),
+         background_pm_lag7 = lag(background_pm, 7, order_by = ZIPCODE), 
+         # geo_smk_pm 
+         geo_smk_pm_lag1 = lag(geo_smk_pm, 1, order_by = ZIPCODE),
+         geo_smk_pm_lag2 = lag(geo_smk_pm, 2, order_by = ZIPCODE),
+         geo_smk_pm_lag3 = lag(geo_smk_pm, 3, order_by = ZIPCODE),
+         geo_smk_pm_lag4 = lag(geo_smk_pm, 4, order_by = ZIPCODE),
+         geo_smk_pm_lag5 = lag(geo_smk_pm, 5, order_by = ZIPCODE),
+         geo_smk_pm_lag6 = lag(geo_smk_pm, 6, order_by = ZIPCODE),
+         geo_smk_pm_lag7 = lag(geo_smk_pm, 7, order_by = ZIPCODE),
+         # krig smk pm
+         krig_smk_pm_lag1 = lag(krig_smk_pm, 1, order_by = ZIPCODE),
+         krig_smk_pm_lag2 = lag(krig_smk_pm, 2, order_by = ZIPCODE),
+         krig_smk_pm_lag3 = lag(krig_smk_pm, 3, order_by = ZIPCODE),
+         krig_smk_pm_lag4 = lag(krig_smk_pm, 4, order_by = ZIPCODE),
+         krig_smk_pm_lag5 = lag(krig_smk_pm, 5, order_by = ZIPCODE),
+         krig_smk_pm_lag6 = lag(krig_smk_pm, 6, order_by = ZIPCODE),
+         krig_smk_pm_lag7 = lag(krig_smk_pm, 7, order_by = ZIPCODE),
+         # temp
+         wrf_temp_lag1 = lag(wrf_temp, 1, order_by = ZIPCODE),
+         wrf_temp_lag2 = lag(wrf_temp, 2, order_by = ZIPCODE),
+         wrf_temp_lag3 = lag(wrf_temp, 3, order_by = ZIPCODE),
+         wrf_temp_lag4 = lag(wrf_temp, 4, order_by = ZIPCODE),
+         wrf_temp_lag5 = lag(wrf_temp, 5, order_by = ZIPCODE), 
+         wrf_temp_lag6 = lag(wrf_temp, 6, order_by = ZIPCODE),
+         wrf_temp_lag7 = lag(wrf_temp, 7, order_by = ZIPCODE),
+         # geo lead
+         geo_smk_pm_lag_m1 = lead(geo_smk_pm, 1, order_by = ZIPCODE),
+         geo_smk_pm_lag_m2 = lead(geo_smk_pm, 2, order_by = ZIPCODE),
+         geo_smk_pm_lag_m3 = lead(geo_smk_pm, 3, order_by = ZIPCODE),
+         geo_smk_pm_lag_m4 = lead(geo_smk_pm, 4, order_by = ZIPCODE),
+         geo_smk_pm_lag_m5 = lead(geo_smk_pm, 5, order_by = ZIPCODE),
+         geo_smk_pm_lag_m6 = lead(geo_smk_pm, 6, order_by = ZIPCODE),
+         geo_smk_pm_lag_m7 = lead(geo_smk_pm, 7, order_by = ZIPCODE),
+         # temp lead
+         wrf_temp_lag_m1 = lead(wrf_temp, 1, order_by = ZIPCODE),
+         wrf_temp_lag_m2 = lead(wrf_temp, 2, order_by = ZIPCODE),
+         wrf_temp_lag_m3 = lead(wrf_temp, 3, order_by = ZIPCODE),
+         wrf_temp_lag_m4 = lead(wrf_temp, 4, order_by = ZIPCODE),
+         wrf_temp_lag_m5 = lead(wrf_temp, 5, order_by = ZIPCODE), 
+         wrf_temp_lag_m6 = lead(wrf_temp, 6, order_by = ZIPCODE),
+         wrf_temp_lag_m7 = lead(wrf_temp, 7, order_by = ZIPCODE)) %>% 
+  # ungroup by zip
+  ungroup(ZIPCODE) %>% 
+  # attach a zip indicator for each smoke variable
+  setNames(paste(colnames(.), "zip", sep="_"))  %>% 
+  # remove the '_zip' from the zipcode and date variable 
+  rename(ZIPCODE = ZIPCODE_zip, date = date_zip)
+
+read_path4 <- paste0("../data_new/medication/oregon_ndc_casecrossover.csv")
+# read_path4 <- paste0("../../../data/data_new/oregon_ndc_casecrossover.csv") # for server
+or_disease <- read_csv(read_path4)
+
+### try join
+colnames(or_disease)[24] <- c("ZIPCODE")
+colnames(or_disease)[74] <- c("date")
+
+outcome_casecross <- or_disease %>%
+  # indicator for male=0, female=1, unknown = 2
+  mutate(age = 2013-yob,
+         sex_ind =ifelse(gender == "F", 1, 
+                         ifelse(gender == "M", 0, 2)),
+         age_ind = ifelse(age < 15, 0,
+                          ifelse(age >= 15 & age < 65, 1,
+                                 ifelse(age >= 65 & age <=105, 2, NA)))
+  ) %>% # end of mutate 
+  # create variables
+  mutate(day = as.factor(weekdays(fromdate)),
+         day_admit = as.factor(weekdays(date)),
+         month_smk = month(fromdate),
+         month_admit = month(date),
+         season_smk = ifelse(fromdate >= "2013-03-20" &  
+                               fromdate <= "2013-06-21", "spring",
+                             ifelse(fromdate >= "2013-06-22" &  
+                                      fromdate <= "2013-09-22", "summer",
+                                    ifelse(fromdate >= "2013-09-23" & 
+                                             fromdate <= "2013-12-21", "fall", "other"))),
+         season_admit = ifelse(date >= "2013-03-20" &  
+                                 date <= "2013-06-21", "spring",
+                               ifelse(date >= "2013-06-22" &  
+                                        date <= "2013-09-22", "summer",
+                                      ifelse(date >= "2013-09-23" & 
+                                               date <= "2013-12-21", "fall", "other")))) %>%
+  # join with zip-level pm estimates
+  left_join(zip_smoke_w_lag, by = c("date", "ZIPCODE"))%>%
+  arrange(personkey, fromdate) # order by id and fromdate
+
+write_path4 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover_zip_lag15.csv")
+write_csv(outcome_casecross, write_path4)
+
+```
+
+```{r time stratified county for ts lag -7 to 0 prepare, echo=FALSE, include=FALSE, eval=FALSE}
+read_path7 <- paste0("../data_new/medication/oregon_asthma_ndc.csv")
+or_ndc_claim <- read_csv(read_path7)
+
+or_ndc_inhaler <- or_ndc_claim %>%
+  select(personkey, num_visit, clmid, line, ZIP, dates) %>%
+  rename(ZIPCODE = ZIP)
+
+# Join data with names of Oregon counties 
+oregon_fips <- read_csv('../instructions/oregon_FIPS.csv')
+# remove the "County" character
+# factor(oregon_fips$county)
+oregon_fips$county <- gsub(" County", "", as.character(factor(oregon_fips$county)))
+
+oregon_fips <- oregon_fips %>%
+  mutate(st_county_fips = with(oregon_fips, paste0(st_code, fips)))
+
+### Join county with zip
+# read_path <- paste0('../data_new/update/or_zip_county_prop.csv') 
+# or_zip_county <- read_csv(read_path)
+
+read_path <- paste0('../instructions/oregon_zip_county.csv')
+or_zip_county <- read_csv(read_path)
+
+# or_zip_county$county_name[which(or_zip_county$county_name=="Hood.River")] <- "Hood River"
+
+or_zip_county <- or_zip_county %>%
+  select(zip, county_name) %>%
+  rename(county = county_name) %>%
+  rename(ZIPCODE = zip) %>%
+  full_join(oregon_fips, by = "county") %>%
+  select(ZIPCODE, county, state, st_county_fips) %>%
+  rename(fips = st_county_fips) # 484
+
+county_new <- c("Washington", "Douglas", "Lane", "Deschutes")
+
+zip_new <- or_ndc_inhaler %>%
+  filter(!(ZIPCODE %in% or_zip_county$ZIPCODE)) %>%
+  select(ZIPCODE) %>%
+  unique() %>%
+  arrange(ZIPCODE) %>%
+  mutate(county = county_new)
+
+# summary(as.factor(zip_new$ZIPCODE))
+# 97003 97471 97475 97703 
+#    48   138     5    16
+
+or_zip_county_new <- or_zip_county %>%
+  select(ZIPCODE, county) %>%
+  bind_rows(zip_new)
+
+write_path <- paste0("../data_new/county_data/zip_to_county_new.csv")
+write_csv(or_zip_county_new, write_path)
+
+read_path9 <- paste0("../data_new/county_data/or_census_pop_est_county.csv")
+or_pop <- read_csv(read_path9)
+
+or_pop <- or_pop %>%
+  rename(county = GEO_display_label, pop = respop72013) %>%
+  select(county, pop) 
+or_pop$county <- gsub(" County, Oregon", "", as.character(factor(or_pop$county)))
+
+### import county data
+read_path_county_pm <- paste0('C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire_new/data_new/county_data/or_county_pop_wt_pm.csv')
+
+county_smoke <- read_csv(read_path_county_pm) # 63801 rows
+
+# descriptives of the two smoke datasets
+summary(county_smoke)
+
+county_smoke$county[which(county_smoke$county=="Hood.River")] <- "Hood River"
+
+# Zipcode PM2.5 estimates
+# create lag variables that take smoke values from n previous days for zipcodes
+county_smoke_w_lag <- county_smoke %>% arrange(county, date) %>%
+  # group by zipcode
+  group_by(county) %>% 
+  # wrf
+  mutate(wrf_f_pm_lag1 = lag(wrf_pm, 1, order_by = county), 
+         wrf_f_pm_lag2 = lag(wrf_pm, 2, order_by = county),
+         wrf_f_pm_lag3 = lag(wrf_pm, 3, order_by = county),
+         wrf_f_pm_lag4 = lag(wrf_pm, 4, order_by = county),
+         wrf_f_pm_lag5 = lag(wrf_pm, 5, order_by = county),
+         wrf_f_pm_lag6 = lag(wrf_pm, 6, order_by = county),
+         wrf_f_pm_lag7 = lag(wrf_pm, 7, order_by = county),
+         # wrf no fire lag
+         wrf_nf_pm_lag1 = lag(wrf_nf_pm, 1, order_by = county),
+         wrf_nf_pm_lag2 = lag(wrf_nf_pm, 2, order_by = county),
+         wrf_nf_pm_lag3 = lag(wrf_nf_pm, 3, order_by = county),
+         wrf_nf_pm_lag4 = lag(wrf_nf_pm, 4, order_by = county),
+         wrf_nf_pm_lag5 = lag(wrf_nf_pm, 5, order_by = county),
+         wrf_nf_pm_lag6 = lag(wrf_nf_pm, 6, order_by = county),
+         wrf_nf_pm_lag7 = lag(wrf_nf_pm, 7, order_by = county),
+         # wrf_smk_pm
+         wrf_smk_pm_lag1 = lag(wrf_smk_pm, 1, order_by = county),
+         wrf_smk_pm_lag2 = lag(wrf_smk_pm, 2, order_by = county),
+         wrf_smk_pm_lag3 = lag(wrf_smk_pm, 3, order_by = county),
+         wrf_smk_pm_lag4 = lag(wrf_smk_pm, 4, order_by = county),
+         wrf_smk_pm_lag5 = lag(wrf_smk_pm, 5, order_by = county),
+         wrf_smk_pm_lag6 = lag(wrf_smk_pm, 6, order_by = county),
+         wrf_smk_pm_lag7 = lag(wrf_smk_pm, 7, order_by = county),
+         # geo weighted pm
+         geo_wt_pm_lag1 = lag(geo_wt_pm, 1, order_by = county),
+         geo_wt_pm_lag2 = lag(geo_wt_pm, 2, order_by = county),
+         geo_wt_pm_lag3 = lag(geo_wt_pm, 3, order_by = county),
+         geo_wt_pm_lag4 = lag(geo_wt_pm, 4, order_by = county),
+         geo_wt_pm_lag5 = lag(geo_wt_pm, 5, order_by = county),
+         geo_wt_pm_lag6 = lag(geo_wt_pm, 6, order_by = county),
+         geo_wt_pm_lag7 = lag(geo_wt_pm, 7, order_by = county),
+         # krig pm
+         krig_pm_lag1 = lag(krig_pm, 1, order_by = county),
+         krig_pm_lag2 = lag(krig_pm, 2, order_by = county),
+         krig_pm_lag3 = lag(krig_pm, 3, order_by = county),
+         krig_pm_lag4 = lag(krig_pm, 4, order_by = county),
+         krig_pm_lag5 = lag(krig_pm, 5, order_by = county), 
+         krig_pm_lag6 = lag(krig_pm, 6, order_by = county),
+         krig_pm_lag7 = lag(krig_pm, 7, order_by = county), 
+         # background pm
+         background_pm_lag1 = lag(background_pm, 1, order_by = county),
+         background_pm_lag2 = lag(background_pm, 2, order_by = county),
+         background_pm_lag3 = lag(background_pm, 3, order_by = county),
+         background_pm_lag4 = lag(background_pm, 4, order_by = county),
+         background_pm_lag5 = lag(background_pm, 5, order_by = county), 
+         background_pm_lag6 = lag(background_pm, 6, order_by = county),
+         background_pm_lag7 = lag(background_pm, 7, order_by = county), 
+         # geo_smk_pm 
+         geo_smk_pm_lag1 = lag(geo_smk_pm, 1, order_by = county),
+         geo_smk_pm_lag2 = lag(geo_smk_pm, 2, order_by = county),
+         geo_smk_pm_lag3 = lag(geo_smk_pm, 3, order_by = county),
+         geo_smk_pm_lag4 = lag(geo_smk_pm, 4, order_by = county),
+         geo_smk_pm_lag5 = lag(geo_smk_pm, 5, order_by = county),
+         geo_smk_pm_lag6 = lag(geo_smk_pm, 6, order_by = county),
+         geo_smk_pm_lag7 = lag(geo_smk_pm, 7, order_by = county),
+         # krig smk pm
+         krig_smk_pm_lag1 = lag(krig_smk_pm, 1, order_by = county),
+         krig_smk_pm_lag2 = lag(krig_smk_pm, 2, order_by = county),
+         krig_smk_pm_lag3 = lag(krig_smk_pm, 3, order_by = county),
+         krig_smk_pm_lag4 = lag(krig_smk_pm, 4, order_by = county),
+         krig_smk_pm_lag5 = lag(krig_smk_pm, 5, order_by = county),
+         krig_smk_pm_lag6 = lag(krig_smk_pm, 6, order_by = county),
+         krig_smk_pm_lag7 = lag(krig_smk_pm, 7, order_by = county),
+         # temp
+         wrf_temp_lag1 = lag(wrf_temp, 1, order_by = county),
+         wrf_temp_lag2 = lag(wrf_temp, 2, order_by = county),
+         wrf_temp_lag3 = lag(wrf_temp, 3, order_by = county),
+         wrf_temp_lag4 = lag(wrf_temp, 4, order_by = county),
+         wrf_temp_lag5 = lag(wrf_temp, 5, order_by = county), 
+         wrf_temp_lag6 = lag(wrf_temp, 6, order_by = county),
+         wrf_temp_lag7 = lag(wrf_temp, 7, order_by = county),
+          # geo lead
+         geo_smk_pm_lag_m1 = lead(geo_smk_pm, 1, order_by = county),
+         geo_smk_pm_lag_m2 = lead(geo_smk_pm, 2, order_by = county),
+         geo_smk_pm_lag_m3 = lead(geo_smk_pm, 3, order_by = county),
+         geo_smk_pm_lag_m4 = lead(geo_smk_pm, 4, order_by = county),
+         geo_smk_pm_lag_m5 = lead(geo_smk_pm, 5, order_by = county),
+         geo_smk_pm_lag_m6 = lead(geo_smk_pm, 6, order_by = county),
+         geo_smk_pm_lag_m7 = lead(geo_smk_pm, 7, order_by = county),
+         # temp lead
+         wrf_temp_lag_m1 = lead(wrf_temp, 1, order_by = county),
+         wrf_temp_lag_m2 = lead(wrf_temp, 2, order_by = county),
+         wrf_temp_lag_m3 = lead(wrf_temp, 3, order_by = county),
+         wrf_temp_lag_m4 = lead(wrf_temp, 4, order_by = county),
+         wrf_temp_lag_m5 = lead(wrf_temp, 5, order_by = county), 
+         wrf_temp_lag_m6 = lead(wrf_temp, 6, order_by = county),
+         wrf_temp_lag_m7 = lead(wrf_temp, 7, order_by = county)) %>% 
+  # ungroup by zip
+  ungroup(county) %>% 
+  # attach a zip indicator for each smoke variable
+  setNames(paste(colnames(.), "county", sep="_"))  %>% 
+  # remove the '_zip' from the zipcode and date variable 
+  rename(county = county_county, date = date_county, fips = fips_county) %>%
+  select(-1)
+
+or_ndc_county_ts <- or_ndc_inhaler %>%
+  rename(date = dates) %>%
+  mutate(day = ifelse(weekdays(date) == "Monday" |
+                      weekdays(date) == "Tuesday" |
+                      weekdays(date) == "Wednesday"|
+                      weekdays(date) == "Thursday"|
+                      weekdays(date) == "Friday", "weekday",
+               ifelse(weekdays(date) == "Saturday" |
+                      weekdays(date) == "Sunday", "weekend", NA))) %>%
+  # join with county-level population
+  left_join(or_zip_county_new, by = "ZIPCODE") %>% # 321 missing 
+  filter(!is.na(county)) %>% # no missing value
+  left_join(or_pop, by = "county") %>%
+  left_join(county_smoke_w_lag, by = c("date", "county")) %>%
+  arrange(personkey, date) %>% # order by id and fromdate
+  group_by(county, date) %>% 
+  # sum up each primary diagnosis for each outcome for each day for each county
+  summarise(n_obs = n()) %>% 
+  # join daily county estimates of pm2.5
+  left_join(county_smoke_w_lag, by = c('date', 'county'))
+
+# summary(as.factor(or_ndc_county_ts$county))
+
+or_ndc_county_ts <- or_ndc_county_ts %>%
+  left_join(or_pop, by = "county") %>%
+  mutate(geo_smk_pm10 = geo_smk_pm_county/10) %>%
+  mutate(rate_per_100k = (n_obs/pop)*100000) %>%
+  mutate(day = ifelse(weekdays(date) == "Monday" |
+                      weekdays(date) == "Tuesday" |
+                      weekdays(date) == "Wednesday"|
+                      weekdays(date) == "Thursday"|
+                      weekdays(date) == "Friday", "weekday",
+               ifelse(weekdays(date) == "Saturday" |
+                      weekdays(date) == "Sunday", "weekend", NA)))
+
+or_ndc_county_ts$day[which(or_ndc_county_ts$date=="2013-05-27")] <- "weekend"
+or_ndc_county_ts$day[which(or_ndc_county_ts$date=="2013-07-04")] <- "weekend"
+or_ndc_county_ts$day[which(or_ndc_county_ts$date=="2013-09-02")] <- "weekend"
+
+write_path5 <- paste0("../data_new/medication/or_ndc_county_time_series_lag15.csv")
+write_csv(or_ndc_county_ts, write_path5)
+
+```
+
+```{r time stratified zip for casecrossover lag -7 to 0 df, echo=FALSE, include=FALSE}
+read_path10 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover_zip_lag15.csv")
+inhaler_ndc_lag15 <- read_csv(read_path10)
+# asthma_ndc <- fread(read_path6, showProgress = T) # faster, but seems error in overall plot analysis
+
+# dataframe to loop through
+inhaler_ndc_lag15_df <- data.frame(inhaler_ndc_lag15)
+
+inhaler_ndc_lag_15_df <- inhaler_ndc_lag15_df %>%
+  mutate(geo_smk_pm10_lag_m1_zip =geo_smk_pm_lag_m1_zip/10,
+         geo_smk_pm10_lag_m2_zip =geo_smk_pm_lag_m2_zip/10,
+         geo_smk_pm10_lag_m3_zip =geo_smk_pm_lag_m3_zip/10,
+         geo_smk_pm10_lag_m4_zip =geo_smk_pm_lag_m4_zip/10,
+         geo_smk_pm10_lag_m5_zip =geo_smk_pm_lag_m5_zip/10,
+         geo_smk_pm10_lag_m6_zip =geo_smk_pm_lag_m6_zip/10,
+         geo_smk_pm10_lag_m7_zip =geo_smk_pm_lag_m7_zip/10) %>%
+  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
+         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
+         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
+         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
+         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
+         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
+         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
+         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)
+
+# which(colnames(inhaler_ndc_lag_15_df)=="geo_smk_pm10_zip") # 178
+# which(colnames(inhaler_ndc_lag_15_df)=="geo_smk_pm10_lag_m1_zip") # 171
+# which(colnames(inhaler_ndc_lag_15_df)=="geo_smk_pm10_lag1_zip") # 179
+
+# which(colnames(inhaler_ndc_lag_15_df)=="wrf_temp_zip") # 93
+# which(colnames(inhaler_ndc_lag_15_df)=="wrf_temp_lag1_zip") # 150
+# which(colnames(inhaler_ndc_lag_15_df)=="wrf_temp_lag_m1_zip") # 164
+
+inhaler_geo_lag <- inhaler_ndc_lag_15_df %>%
+  select(177:171, 178) # GWR-method
+
+inhaler_temp_lag <- inhaler_ndc_lag_15_df %>%
+  select(170:164, 93) # temp
+
+inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
+inhaler_temp_lag <- as.matrix(inhaler_temp_lag)
+
 # create fit matrix
-fit_mat <- matrix(NA, nrow=4, ncol=6)
-colnames(fit_mat) <- c("outcome", "exposure", "fit_mix", "df", "aic", "fit_fix")
+fit_mat <- matrix(NA, nrow=4, ncol=5)
+colnames(fit_mat) <- c("outcome", "exposure", "fit", "df", "aic")
 
@@
     # define basis b using natural spline function
-    b <- ns(0:7, df = degree_freedom, intercept = T)
-
-    # multiply lagged pm matrix by basis
-    pm_b <- inhaler_ndc_lag %*% b
-
-    # fit mixed model
-    mod <-glmer(outcome ~  pm_b + (1|county) + offset(log(pop)), inhaler_ndc_lag_df, family="poisson", control = glmerControl(optimizer = "bobyqa"))
-
-    # fixed mod
-    fixed_mod <- glm(outcome ~ pm_b + county + offset(log(pop)), inhaler_ndc_lag_df, 
-                     family = "poisson")
-
-    AIC(mod)
-    AIC(fixed_mod)
+    B <- ns(-7:0, df = degree_freedom, intercept = T)
+
+    inhaler_geo_lag_B <- inhaler_geo_lag%*%B
+    inhaler_temp_lag_B <- inhaler_temp_lag%*%B
+
+    fit <- clogit(outcome ~ inhaler_geo_lag_B + inhaler_temp_lag_B + strata(personkey), 
+              inhaler_ndc_lag_15_df)
+
+    AIC(fit)
+
     # fill AIC
-    fit_mat[h-2,5] <- round(AIC(mod),5)
-    fit_mat[h-2,6] <- round(AIC(fixed_mod),5)
+    fit_mat[h-2,5] <- round(AIC(fit),5)
 
@@
   "degree of freedom best fit by AIC for mixed model approach"))
+
 ```
 
-### Compare fixed and mixed models of df 3 and 4
-
-```{r compare mixed and fixed, message=FALSE, echo=FALSE, eval=FALSE}
-
-## degree of 3 or 4, mixed or fixed
-degree_freedom <- 3
-
-# define basis b using natural spline function
-b <- ns(0:7, df = degree_freedom, intercept = T)
-
-# multiply lagged pm matrix by basis
-pm_b <- inhaler_ndc_lag %*% b
-
-mod <-glmer(outcome ~  pm_b + (1|county) + wrf_temp_zip + offset(log(pop)), inhaler_ndc_lag_df, family="poisson",
-            control = glmerControl(optimizer = "bobyqa"))
-
-dlparms <- mod@beta[2:(1+degree_freedom)]
-# estimate distributed lag values for each day
-dl_estimates <- data.frame(estimate = b %*% dlparms)
-
-# covariance matrix for knots (need to convert to matrix object)
-cov_mat <- as.matrix(vcov(mod)[2:(1+degree_freedom),2:(1+degree_freedom)])
-
-# estimate variance of splines
-dl_var <- b%*%cov_mat%*%t(b)
-
-# calculate standard error for each lag value
-dl_estimates$stderr <- sqrt(diag(dl_var))
-
-# calculate lower and upper bounds
-dl_estimates$lower_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(1-0.975,
-  df=df.residual(mod))  
-dl_estimates$upper_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(0.975,
-  df=df.residual(mod))
-
-# estimate relative risk and 95%CI from beta
-inhaler_dl_rr <- dl_estimates %>% 
-  mutate(estimate = exp(estimate),
-         lower_bound = exp(lower_bound),
-         upper_bound = exp(upper_bound),
-         lag = 0:7)
-
-# estimate cumulative effect
-cumulative <- sum(dl_estimates$estimate)
+```{r DLM for casecrossover lag -7 to 0, echo = FALSE, message=FALSE, warning=FALSE}
+inhaler_geo_lag <- inhaler_ndc_lag_15_df %>%
+  select(177:171, 178) # GWR-method
+
+inhaler_temp_lag <- inhaler_ndc_lag_15_df %>%
+  select(170:164, 93) # temp
+
+lag_max <- 0
+n1 <- length(inhaler_geo_lag[,1])
+n2 <- length(inhaler_temp_lag[,1])
+
+inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
+inhaler_temp_lag <- as.matrix(inhaler_temp_lag)
+
+B <- ns(-7:lag_max, df =5, intercept=TRUE)
+
+# the nrow B should equal ncol(pm)
+dim(B)
+
+inhaler_geo_lag_B <- inhaler_geo_lag%*%B
+inhaler_temp_lag_B <- inhaler_temp_lag%*%B
+
+fit <- clogit(outcome ~ inhaler_geo_lag_B + inhaler_temp_lag_B + strata(personkey), 
+              inhaler_ndc_lag_15_df)
+
+coef(fit)
+summary(fit)
+AIC(fit)
+
+D <- cbind(B,B)
+
+# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
+dlparms <- c(1,2,3,4,5,6,7,8,9,10)
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms]) 
+DLvar <- D%*%vcov(fit)[dlparms,dlparms]%*%t(D)
+DLestimate$SE <- sqrt(diag(DLvar))
+DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
+DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
+
+DLestimate$estimate <- exp(DLestimate$estimate)
+DLestimate$lower <- exp(DLestimate$lower)
+DLestimate$upper <- exp(DLestimate$upper)
+
+DLestimate
+
+DLestimate$lag <- -7:lag_max
+p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
+p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
+p <- p + geom_line(aes(x=lag), linetype=2, size=2)
+p <- p + theme_bw()
+p
+
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms])
+
+cumulative <- sum(DLestimate$estimate)
 cumulative
 
-# estimate cumulative effect stnd error
-cumulative_se <- sqrt(sum(dl_var))
+cumulative_se <- sqrt(sum(DLvar))
 cumulative_se
 
-# estimate cumulative CI
-cumulative_ci <- cumulative+cumulative_se*qt(c(1-0.975,0.975), 
-                                             df=df.residual(mod))
-cumulative_ci
-
-exp(cumulative)
-exp(cumulative_ci)
-
-# plot 
-p3_mix <- ggplot(inhaler_dl_rr, aes(x=lag, y=estimate, 
-                              ymin=lower_bound, ymax=upper_bound)) +
-  geom_ribbon(alpha = 0.5) +
-  geom_line(size=2)
-
-fixed_mod <- glm(outcome ~ pm_b + county + wrf_temp_zip + offset(log(pop)), inhaler_ndc_lag_df, 
-                     family = "poisson")
-
-dlparms <- fixed_mod$coefficients[2:4]
-# estimate distributed lag values for each day
-dl_estimates <- data.frame(estimate = b %*% dlparms)
-
-# covariance matrix for knots (need to convert to matrix object)
-cov_mat <- as.matrix(vcov(fixed_mod)[2:4,2:4])
-
-# estimate variance of splines
-dl_var <- b%*%cov_mat%*%t(b)
-
-# calculate standard error for each lag value
-dl_estimates$stderr <- sqrt(diag(dl_var))
-
-# calculate lower and upper bounds
-dl_estimates$lower_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(1-0.975,
-  df=df.residual(fixed_mod))  
-dl_estimates$upper_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(0.975,
-  df=df.residual(fixed_mod))
-
-# estimate relative risk and 95%CI from beta
-inhaler_dl_rr <- dl_estimates %>% 
-  mutate(estimate = exp(estimate),
-         lower_bound = exp(lower_bound),
-         upper_bound = exp(upper_bound),
-         lag = 0:7)
-
-# estimate cumulative effect
-cumulative <- sum(dl_estimates$estimate)
-cumulative
-
-# estimate cumulative effect stnd error
-cumulative_se <- sqrt(sum(dl_var))
-cumulative_se
-
-# estimate cumulative CI
-cumulative_ci <- cumulative+cumulative_se*qt(c(1-0.975,0.975), 
-                                             df=df.residual(fixed_mod))
-cumulative_ci
-
-exp(cumulative)
-exp(cumulative_ci)
-
-# plot 
-p3_fix <- ggplot(inhaler_dl_rr, aes(x=lag, y=estimate, 
-                              ymin=lower_bound, ymax=upper_bound)) +
-  geom_ribbon(alpha = 0.5) +
-  geom_line(size=2)
-
-p3_fix
-p3_mix
-
-## df = 4
-degree_freedom <- 4
-
-# define basis b using natural spline function
-b <- ns(0:7, df = degree_freedom, intercept = T)
-
-# multiply lagged pm matrix by basis
-pm_b <- inhaler_ndc_lag %*% b
-
-mod <-glmer(outcome ~  pm_b + (1|county) + wrf_temp_zip + offset(log(pop)), inhaler_ndc_lag_df, family="poisson",
-            control = glmerControl(optimizer = "bobyqa"))
-
-dlparms <- mod@beta[2:5]
-# estimate distributed lag values for each day
-dl_estimates <- data.frame(estimate = b %*% dlparms)
-
-# covariance matrix for knots (need to convert to matrix object)
-cov_mat <- as.matrix(vcov(mod)[2:5,2:5])
-
-# estimate variance of splines
-dl_var <- b%*%cov_mat%*%t(b)
-
-# calculate standard error for each lag value
-dl_estimates$stderr <- sqrt(diag(dl_var))
-
-# calculate lower and upper bounds
-dl_estimates$lower_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(1-0.975,
-  df=df.residual(mod))  
-dl_estimates$upper_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(0.975,
-  df=df.residual(mod))
-
-# estimate relative risk and 95%CI from beta
-inhaler_dl_rr <- dl_estimates %>% 
-  mutate(estimate = exp(estimate),
-         lower_bound = exp(lower_bound),
-         upper_bound = exp(upper_bound),
-         lag = 0:7)
-
-# estimate cumulative effect
-cumulative <- sum(dl_estimates$estimate)
-cumulative
-
-# estimate cumulative effect stnd error
-cumulative_se <- sqrt(sum(dl_var))
-cumulative_se
-
-# estimate cumulative CI
-cumulative_ci <- cumulative+cumulative_se*qt(c(1-0.975,0.975), 
-                                             df=df.residual(mod))
-cumulative_ci
-
-exp(cumulative)
-exp(cumulative_ci)
-
-# plot 
-p4_mix <- ggplot(inhaler_dl_rr, aes(x=lag, y=estimate, 
-                              ymin=lower_bound, ymax=upper_bound)) +
-  geom_ribbon(alpha = 0.5) +
-  geom_line(size=2)
-
-fixed_mod <- glm(outcome ~ pm_b + county + wrf_temp_zip + offset(log(pop)), inhaler_ndc_lag_df, 
-                     family = "poisson")
-
-dlparms <- fixed_mod$coefficients[2:5]
-# estimate distributed lag values for each day
-dl_estimates <- data.frame(estimate = b %*% dlparms)
-
-# covariance matrix for knots (need to convert to matrix object)
-cov_mat <- as.matrix(vcov(fixed_mod)[2:5,2:5])
-
-# estimate variance of splines
-dl_var <- b%*%cov_mat%*%t(b)
-
-# calculate standard error for each lag value
-dl_estimates$stderr <- sqrt(diag(dl_var))
-
-# calculate lower and upper bounds
-dl_estimates$lower_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(1-0.975,
-  df=df.residual(fixed_mod))  
-dl_estimates$upper_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(0.975,
-  df=df.residual(fixed_mod))
-
-# estimate relative risk and 95%CI from beta
-inhaler_dl_rr <- dl_estimates %>% 
-  mutate(estimate = exp(estimate),
-         lower_bound = exp(lower_bound),
-         upper_bound = exp(upper_bound),
-         lag = 0:7)
-
-# estimate cumulative effect
-cumulative <- sum(dl_estimates$estimate)
-cumulative
-
-# estimate cumulative effect stnd error
-cumulative_se <- sqrt(sum(dl_var))
-cumulative_se
-
-# estimate cumulative CI
-cumulative_ci <- cumulative+cumulative_se*qt(c(1-0.975,0.975), 
-                                             df=df.residual(fixed_mod))
-cumulative_ci
-
-exp(cumulative)
-exp(cumulative_ci)
-
-# plot 
-p4_fix <- ggplot(inhaler_dl_rr, aes(x=lag, y=estimate, 
-                              ymin=lower_bound, ymax=upper_bound)) +
-  geom_ribbon(alpha = 0.5) +
-  geom_line(size=2)
-
-p4_fix
-p4_mix
-
-ggsave("../plot_new/ndc_p3_mix.pdf", plot = p3_mix, 
-       width = 12, height = 8, units = "in")
-ggsave("../plot_new/ndc_p4_mix.pdf", plot = p4_mix, 
-       width = 12, height = 8, units = "in")
-
-ggsave("../plot_new/ndc_p3_fix.pdf", plot = p3_fix, 
-       width = 12, height = 8, units = "in")
-ggsave("../plot_new/ndc_p4_fix.pdf", plot = p4_fix, 
-       width = 12, height = 8, units = "in")
+cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
+cumulative_CI
 
@@
 
-Note I get some warnings (including max grade convergence issues), slightly different answers in best fit, and it takes about 18 minutes to run. Consider running in parallel.
-I suspect the convergence issues are likely due to the same state/counties with small sample size and few outcome observations and lack of variability between exposure/outcome.
-Going to fit what I have now, take a walk, and see what comes out. First thing to do is find cumulative effect, then lagged day effects.
-
-```{r, message=FALSE, echo=FALSE, include=FALSE, eval=FALSE}
-# create a dataframe that contains cumulative estimates for each outcome
-cumulative_df <- as_tibble(matrix(NA, nrow=0, ncol=4))
-colnames(cumulative_df) <- c("outcome", "rel_risk", "lower_95", "upper_95")
-
-# create a dataframe that contains the distributed lag estimates for each outcome
-dist_lag_df <- as_tibble(matrix(NA,  nrow = 0, ncol = 5))
-colnames(dist_lag_df) <- c("outcome", "lag", "rel_risk", 
-                                  "lower_95", "upper_95")
-
-# loop through best fit df
-
-    # create empty matrix two empty matrices to fill
-    cumul_val <- as_tibble(matrix(NA, nrow=1, ncol=4))
-    colnames(cumul_val) <- c("outcome", "rel_risk", "lower_95", "upper_95")
-
-    dl_val <- as_tibble(matrix(NA, nrow=8, ncol=5))
-    colnames(dl_val) <- c("outcome", "lag", "rel_risk", 
-                                  "lower_95", "upper_95")
-
-    # easiest to set out the parameters from the start
-    outcome <- "inhalers"
-    degree_freedom <- as.numeric(lag_spline_best_fit[4])
-
-    # define basis b using natural spline function
-    b <- ns(0:7, df = degree_freedom, intercept = T)
-
-    # multiply lagged pm matrix by basis
-    pm_b <- inhaler_ndc_lag %*% b
-
-    # fit mixed model
-     mod <-glmer(outcome ~  inhaler_ndc_lag_B + wrf_temp_zip + day + (1|county) + offset(log(pop)), inhaler_ndc_lag_df, family="poisson", control = glmerControl(optimizer = "bobyqa"))
-
-    mod
-    # output distributed lag beta parameters; this will change based on DF
-    dlparms <- mod@beta[2:(1+degree_freedom)]
-    # estimate distributed lag values for each day
-    dl_estimates <- data.frame(estimate = b %*% dlparms)
-
-    # covariance matrix for knots (need to convert to matrix object)
-    cov_mat <- as.matrix(vcov(mod)[2:(1+degree_freedom),2:(1+degree_freedom)]) 
-    # estimate variance of splines
-    dl_var <- b%*%cov_mat%*%t(b)
-    # calculate standard error for each lag value
-    dl_estimates$stderr <- sqrt(diag(dl_var))
-    # calculate lower and upper bounds
-    dl_estimates$lower_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(1-0.975,
-      df=df.residual(mod))  
-    dl_estimates$upper_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(0.975,
-      df=df.residual(mod))
-
-    # estimate cumulative effect ----
-    cumulative <- sum(dl_estimates$estimate)
-    # estimate cumulative effect stnd error
-    cumulative_se <- sqrt(sum(dl_var))
-    # estimate cumulative CI
-    cumulative_ci <- cumulative+cumulative_se*qt(c(1-0.975,0.975), 
-                                                 df=df.residual(mod))
-    # fill tibbles and bind in to larger dataframes ----
-    # outcome
-    cumul_val[1,1] <- outcome
-    # cumulative rel risk
-    cumul_val[1,2] <- exp(cumulative)
-    # cumulative 95% CI
-    cumul_val[1,3:4] <- exp(cumulative_ci)
-    # bind to cumulative dataframe
-    cumulative_df <- rbind(cumulative_df, cumul_val)
-
-    # lag vals (this step is redundant)
-    dl_val[,1] <- outcome
-    dl_val[,2] <- 0:7
-    dl_val[,3] <- exp(dl_estimates[,1])
-    dl_val[,4] <- exp(dl_estimates[,3])
-    dl_val[,5] <- exp(dl_estimates[,4])
-
-    # bind in with overall df
-    dist_lag_df <- rbind(dist_lag_df, dl_val)
-
-# print table of cumulative effects
-knitr::kable(cumulative_df, caption = paste0("Cumulative Effect for a 10 ug/m^3 increase in GWR smoke lagged days 0-7 mixed model approach"))
-
-# plot distributed lag effects
-ggplot(dist_lag_df, aes(x=lag, y=rel_risk)) +
-  geom_line(colour = "blue") +
-  geom_ribbon(aes(ymin = lower_95, ymax = upper_95), fill = "blue", alpha = 0.3) + 
-  scale_y_continuous(limits = c(0.98, 1.06)) +
-  scale_x_continuous(breaks = c(seq(0, 7,by=1))) +
-  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
-  facet_wrap(~ outcome) +
-  ggtitle(paste0("Distributed lag association between a 10 ug/m^3",
-                 " increase in GWR smoke PM2.5 and cardiopulmonary outcomes")) +
-  ylab("Relative Risk for a 10 ug/m^3 increase in GWR smoke (fixed)") +
-  xlab("Lagged Days") +
-  theme_bw()
-
-```
-
-## Distributed Lag -- Casecrossover Analysis
-
+## Supplement
 ### 1. For Overall Plot (Unconstrained lag model)
@@
 
-### 2. Distributed lag for casecrossover
-
-```{r DLM for casecrossover, echo = FALSE, message=FALSE, warning=FALSE}
-read_path8 <- paste0("../data_new/county_data/zip_to_county_new.csv")
-or_zip_to_county <- read_csv(read_path8)
-
-read_path9 <- paste0("../data_new/county_data/or_census_pop_est_county.csv")
-or_pop <- read_csv(read_path9)
-
-or_pop <- or_pop %>%
-  rename(county = GEO_display_label, pop = respop72013) %>%
-  select(county, pop) 
-or_pop$county <- gsub(" County, Oregon", "", as.character(factor(or_pop$county)))
-
-inhaler_ndc_lag_df <- inhaler_ndc_df %>%
-  left_join(or_zip_to_county, by = "ZIPCODE") %>%
-  left_join(or_pop, by = "county") %>%
-  mutate(day = ifelse(weekdays(date) == "Monday" |
-                      weekdays(date) == "Tuesday" |
-                      weekdays(date) == "Wednesday"|
-                      weekdays(date) == "Thursday"|
-                      weekdays(date) == "Friday", "weekday",
-               ifelse(weekdays(date) == "Saturday" |
-                      weekdays(date) == "Sunday", "weekend", NA))) %>%
-  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
-         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
-         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
-         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
-         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
-         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
-         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
-         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)
-
-inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-05-27")] <- "weekend"
-inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-07-04")] <- "weekend"
-inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-09-02")] <- "weekend"
-
-which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_zip") # 159
-which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_lag1_zip") # 160
-which(colnames(inhaler_ndc_lag_df)=="wrf_temp_zip") # 93
-which(colnames(inhaler_ndc_lag_df)=="wrf_temp_lag1_zip") # 150
-
-inhaler_geo_lag <- inhaler_ndc_lag_df %>%
-  select(159, 160:166) # GWR-method
-
-inhaler_temp_lag <- inhaler_ndc_lag_df %>%
-  select(93, 150:156) # GWR-method
-
-lag_max <- 7
-n1 <- length(inhaler_geo_lag[,1])
-n2 <- length(inhaler_temp_lag[,1])
-
-inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
-inhaler_temp_lag <- as.matrix(inhaler_temp_lag)
-
-B <- ns(0:lag_max, df =3, intercept=TRUE)
-
-# the nrow B should equal ncol(pm)
-dim(B)
-
-inhaler_geo_lag_B <- inhaler_geo_lag%*%B
-inhaler_temp_lag_B <- inhaler_temp_lag%*%B
-
-fit <- clogit(outcome ~ inhaler_geo_lag_B + inhaler_temp_lag_B + strata(personkey), 
-              inhaler_ndc_lag_df)
-
-coef(fit)
-summary(fit)
-AIC(fit)
-
-D <- cbind(B,B)
-
-# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
-dlparms <- c(1,2,3,4,5,6)
-DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms]) 
-DLvar <- D%*%vcov(fit)[dlparms,dlparms]%*%t(D)
-DLestimate$SE <- sqrt(diag(DLvar))
-DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
-DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
-
-DLestimate$estimate <- exp(DLestimate$estimate)
-DLestimate$lower <- exp(DLestimate$lower)
-DLestimate$upper <- exp(DLestimate$upper)
-
-DLestimate
-
-DLestimate$lag <- 0:lag_max
-p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
-p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
-p <- p + geom_line(aes(x=lag), linetype=2, size=2)
-p <- p + theme_bw()
-p
-
-DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms])
-
-cumulative <- sum(DLestimate$estimate)
-cumulative
-
-cumulative_se <- sqrt(sum(DLvar))
-cumulative_se
-
-cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
-cumulative_CI
-
-```
-
-```{r DLM for casecrossover temp, echo = FALSE, message=FALSE, warning=FALSE}
-read_path8 <- paste0("../data_new/county_data/zip_to_county_new.csv")
-or_zip_to_county <- read_csv(read_path8)
-
-read_path9 <- paste0("../data_new/county_data/or_census_pop_est_county.csv")
-or_pop <- read_csv(read_path9)
-
-or_pop <- or_pop %>%
-  rename(county = GEO_display_label, pop = respop72013) %>%
-  select(county, pop) 
-or_pop$county <- gsub(" County, Oregon", "", as.character(factor(or_pop$county)))
-
-inhaler_ndc_lag_df <- inhaler_ndc_df %>%
-  left_join(or_zip_to_county, by = "ZIPCODE") %>%
-  left_join(or_pop, by = "county") %>%
-  mutate(day = ifelse(weekdays(date) == "Monday" |
-                      weekdays(date) == "Tuesday" |
-                      weekdays(date) == "Wednesday"|
-                      weekdays(date) == "Thursday"|
-                      weekdays(date) == "Friday", "weekday",
-               ifelse(weekdays(date) == "Saturday" |
-                      weekdays(date) == "Sunday", "weekend", NA))) %>%
-  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
-         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
-         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
-         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
-         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
-         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
-         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
-         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)
-
-inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-05-27")] <- "weekend"
-inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-07-04")] <- "weekend"
-inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-09-02")] <- "weekend"
-
-which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_zip") # 159
-which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_lag1_zip") # 160
-which(colnames(inhaler_ndc_lag_df)=="wrf_temp_zip") # 93
-which(colnames(inhaler_ndc_lag_df)=="wrf_temp_lag1_zip") # 150
-
-inhaler_geo_lag <- inhaler_ndc_lag_df %>%
-  select(159, 160:166) # GWR-method
-
-inhaler_temp_lag <- inhaler_ndc_lag_df %>%
-  select(93, 150:156) # GWR-method
-
-lag_max <- 7
-n1 <- length(inhaler_geo_lag[,1])
-n2 <- length(inhaler_temp_lag[,1])
-
-inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
-inhaler_temp_lag <- as.matrix(inhaler_temp_lag)
-
-B <- ns(0:lag_max, df =3, intercept=TRUE)
-
-# the nrow B should equal ncol(pm)
-dim(B)
-
-inhaler_geo_lag_B <- inhaler_geo_lag%*%B
-inhaler_temp_lag_B <- inhaler_temp_lag%*%B
-
-fit <- clogit(outcome ~  inhaler_temp_lag_B + strata(personkey), 
-              inhaler_ndc_lag_df)
-
-coef(fit)
-summary(fit)
-AIC(fit)
-
-D <- cbind(B,B)
-
-# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
-dlparms <- c(1,2,3)
-DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms]) 
-DLvar <- B%*%vcov(fit)[dlparms,dlparms]%*%t(B)
-DLestimate$SE <- sqrt(diag(DLvar))
-DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
-DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
-
-DLestimate$estimate <- exp(DLestimate$estimate)
-DLestimate$lower <- exp(DLestimate$lower)
-DLestimate$upper <- exp(DLestimate$upper)
-
-DLestimate
-
-DLestimate$lag <- 0:lag_max
-p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
-p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
-p <- p + geom_line(aes(x=lag), linetype=2, size=2)
-p <- p + theme_bw()
-p
-
-DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms])
-
-cumulative <- sum(DLestimate$estimate)
-cumulative
-
-cumulative_se <- sqrt(sum(DLvar))
-cumulative_se
-
-cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
-cumulative_CI
-
-```
-
-### 3. Unconstrained lag for time series
-
-```{r unconstrained ts lag, message=FALSE, echo=FALSE}
+### 2. Unconstrained lag for time series
+
+```{r unconstrained ts lag, message=FALSE, echo=FALSE, warning=FALSE}
 or_ndc_county_ts_df <- or_ndc_county_ts %>%
@@
 
-### 4. Distributed lag for time series
+### 3. Distributed lag check for time series
 
 ```{r DLM ts, echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}
-
 which(colnames(or_ndc_county_ts_df)=="wrf_temp_lag1_county") # 17
@@
 ## mixed model
-mod <-glmer(outcome ~  geo_ts_lag_B + tem_ts_lag_B + day + (1|county) + offset(log(pop)),   
+mod <-glmer(n_obs ~  geo_ts_lag_B + tem_ts_lag_B + day + (1|county) + offset(log(pop)),   
             or_ndc_county_ts_df, family="poisson", control = glmerControl(optimizer = "bobyqa"))
@@
 AIC(mod)
+
+D <- cbind(B, B)
 # output distributed lag beta parameters
-dlparms <- mod@beta[2:4]
+dlparms <- mod@beta[2:7]
 # estimate distributed lag values for each day
-dl_estimates <- data.frame(estimate = B %*% dlparms)
+dl_estimates <- data.frame(estimate = D %*% dlparms)
 
 # covariance matrix for knots (need to convert to matrix object)
-cov_mat <- as.matrix(vcov(mod)[2:4,2:4])
+cov_mat <- as.matrix(vcov(mod)[2:7,2:7])
 
 # estimate variance of splines
-dl_var <- B%*%cov_mat%*%t(B)
+dl_var <- D%*%cov_mat%*%t(D)
 
@@
 ```
+
+```{r DLM ts temp, echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}
+which(colnames(or_ndc_county_ts_df)=="wrf_temp_lag1_county") # 17
+which(colnames(or_ndc_county_ts_df)=="geo_smk_pm10_lag1_county") # 25
+
+geo_ts_lag <- or_ndc_county_ts_df %>%
+  select(24:31) # GWR-method
+
+lag_max <- 7
+
+geo_ts_lag <- as.matrix(geo_ts_lag)
+tem_ts_lag <- as.matrix(tem_ts_lag)
+
+B <- ns(0:lag_max, df =3, intercept=TRUE) # or 3?
+
+# the nrow B should equal ncol(pm)
+dim(B)
+dim(tem_ts_lag)
+
+tem_ts_lag_B <- tem_ts_lag%*%B
+
+## mixed model
+mod <-glmer(n_obs ~ tem_ts_lag_B + day + (1|county) + offset(log(pop)),   
+            or_ndc_county_ts_df, family="poisson", control = glmerControl(optimizer = "bobyqa"))
+
+summary(mod)
+# check AIC
+AIC(mod)
+
+# output distributed lag beta parameters
+dlparms <- mod@beta[2:4]
+# estimate distributed lag values for each day
+dl_estimates <- data.frame(estimate = B %*% dlparms)
+
+# covariance matrix for knots (need to convert to matrix object)
+cov_mat <- as.matrix(vcov(mod)[2:4,2:4])
+
+# estimate variance of splines
+dl_var <- B%*%cov_mat%*%t(B)
+
+# calculate standard error for each lag value
+dl_estimates$stderr <- sqrt(diag(dl_var))
+
+# calculate lower and upper bounds
+dl_estimates$lower_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(1-0.975,
+  df=df.residual(mod))  
+dl_estimates$upper_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(0.975,
+  df=df.residual(mod))
+
+# estimate relative risk and 95%CI from beta
+inhaler_dl_rr <- dl_estimates %>% 
+  mutate(estimate = exp(estimate),
+         lower_bound = exp(lower_bound),
+         upper_bound = exp(upper_bound),
+         lag = 0:7)
+
+# estimate cumulative effect
+cumulative <- sum(dl_estimates$estimate)
+cumulative
+
+# estimate cumulative effect stnd error
+cumulative_se <- sqrt(sum(dl_var))
+cumulative_se
+
+# estimate cumulative CI
+cumulative_ci <- cumulative+cumulative_se*qt(c(1-0.975,0.975), 
+                                             df=df.residual(mod))
+cumulative_ci
+
+exp(cumulative)
+exp(cumulative_ci)
+
+# plot 
+p <- ggplot(inhaler_dl_rr, aes(x=lag, y=estimate, 
+                              ymin=lower_bound, ymax=upper_bound)) +
+  geom_ribbon(alpha = 0.5) +
+  geom_line(size=2)
+
+p
+```
+
+### 4. Distributed lag -7 to 0 for time series
+```{r DLM for casecrossover lag -7 to 0 temp, echo = FALSE, message=FALSE, warning=FALSE, eval = FALSE}
+
+inhaler_temp_lag <- inhaler_ndc_lag_15_df %>%
+  select(170:164, 93, 150:156) # temp
+
+lag_max <- 7
+
+inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
+inhaler_temp_lag <- as.matrix(inhaler_temp_lag)
+
+B <- ns(-7:lag_max, df =3, intercept=TRUE)
+
+# the nrow B should equal ncol(pm)
+dim(B)
+
+inhaler_temp_lag_B <- inhaler_temp_lag%*%B
+
+fit <- clogit(outcome ~  inhaler_temp_lag_B + strata(personkey), 
+              inhaler_ndc_lag_15_df)
+
+coef(fit)
+summary(fit)
+AIC(fit)
+
+# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
+dlparms <- c(1,2,3)
+DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms]) 
+DLvar <- B%*%vcov(fit)[dlparms,dlparms]%*%t(B)
+DLestimate$SE <- sqrt(diag(DLvar))
+DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
+DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
+
+DLestimate$estimate <- exp(DLestimate$estimate)
+DLestimate$lower <- exp(DLestimate$lower)
+DLestimate$upper <- exp(DLestimate$upper)
+
+DLestimate
+
+DLestimate$lag <- -7:lag_max
+p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
+p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
+p <- p + geom_line(aes(x=lag), linetype=2, size=2)
+p <- p + theme_bw()
+p
+
+DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms])
+
+cumulative <- sum(DLestimate$estimate)
+cumulative
+
+cumulative_se <- sqrt(sum(DLvar))
+cumulative_se
+
+cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
+cumulative_CI
+
+```
+
+```{r DLM ts -7 to 0, echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}
+read_path11 <- paste0("../data_new/medication/or_ndc_county_time_series_lag15.csv")
+or_ndc_county_ts_lag15 <- read_csv(read_path11)
+
+or_ndc_county_ts_df <- or_ndc_county_ts_lag15 %>%
+  select(1:4, pop, rate_per_100k, day,  geo_smk_pm_lag_m1_county:geo_smk_pm_lag_m7_county, geo_smk_pm_county,
+         geo_smk_pm_lag1_county:geo_smk_pm_lag7_county, wrf_temp_lag_m1_county:wrf_temp_lag_m7_county, 
+         wrf_temp_county, wrf_temp_lag1_county:wrf_temp_lag7_county) %>%
+  mutate(geo_smk_pm10_county = geo_smk_pm_county/10,
+         geo_smk_pm10_lag1_county =geo_smk_pm_lag1_county/10,
+         geo_smk_pm10_lag2_county =geo_smk_pm_lag2_county/10,
+         geo_smk_pm10_lag3_county =geo_smk_pm_lag3_county/10,
+         geo_smk_pm10_lag4_county =geo_smk_pm_lag4_county/10,
+         geo_smk_pm10_lag5_county =geo_smk_pm_lag5_county/10,
+         geo_smk_pm10_lag6_county =geo_smk_pm_lag6_county/10,
+         geo_smk_pm10_lag7_county =geo_smk_pm_lag7_county/10,
+         # lag -7 to -1
+         geo_smk_pm10_lag_m1_county =geo_smk_pm_lag_m1_county/10,
+         geo_smk_pm10_lag_m2_county =geo_smk_pm_lag_m2_county/10,
+         geo_smk_pm10_lag_m3_county =geo_smk_pm_lag_m3_county/10,
+         geo_smk_pm10_lag_m4_county =geo_smk_pm_lag_m4_county/10,
+         geo_smk_pm10_lag_m5_county =geo_smk_pm_lag_m5_county/10,
+         geo_smk_pm10_lag_m6_county =geo_smk_pm_lag_m6_county/10,
+         geo_smk_pm10_lag_m7_county =geo_smk_pm_lag_m7_county/10)
+
+which(colnames(or_ndc_county_ts_df)=="wrf_temp_lag1_county") # 31
+which(colnames(or_ndc_county_ts_df)=="wrf_temp_county") # 30
+which(colnames(or_ndc_county_ts_df)=="wrf_temp_lag_m1_county") # 23
+which(colnames(or_ndc_county_ts_df)=="geo_smk_pm10_lag1_county") # 39
+which(colnames(or_ndc_county_ts_df)=="geo_smk_pm10_lag_m1_county") # 46
+which(colnames(or_ndc_county_ts_df)=="geo_smk_pm10_county") # 38
+
+geo_ts_lag <- or_ndc_county_ts_df %>%
+  select(52:46, 38, 39:45) # GWR-method
+tem_ts_lag <- or_ndc_county_ts_df %>%
+  select(29:23, 30, 31:37) # temp
+
+lag_max <- 7
+
+geo_ts_lag <- as.matrix(geo_ts_lag)
+tem_ts_lag <- as.matrix(tem_ts_lag)
+
+B <- ns(-7:lag_max, df =3, intercept=TRUE) # or 3?
+
+# the nrow B should equal ncol(pm)
+dim(B)
+dim(geo_ts_lag)
+dim(tem_ts_lag)
+
+geo_ts_lag_B <- geo_ts_lag%*%B
+
+tem_ts_lag_B <- tem_ts_lag%*%B
+
+## mixed model
+mod <-glmer(n_obs ~  geo_ts_lag_B + tem_ts_lag_B + day + (1|county) + offset(log(pop)),   
+            or_ndc_county_ts_df, family="poisson", control = glmerControl(optimizer = "bobyqa"))
+
+summary(mod)
+# check AIC
+AIC(mod)
+
+D <- cbind(B, B)
+# output distributed lag beta parameters
+dlparms <- mod@beta[2:7]
+# estimate distributed lag values for each day
+dl_estimates <- data.frame(estimate = D %*% dlparms)
+
+# covariance matrix for knots (need to convert to matrix object)
+cov_mat <- as.matrix(vcov(mod)[2:7,2:7])
+
+# estimate variance of splines
+dl_var <- D%*%cov_mat%*%t(D)
+
+# calculate standard error for each lag value
+dl_estimates$stderr <- sqrt(diag(dl_var))
+
+# calculate lower and upper bounds
+dl_estimates$lower_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(1-0.975,
+  df=df.residual(mod))  
+dl_estimates$upper_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(0.975,
+  df=df.residual(mod))
+
+# estimate relative risk and 95%CI from beta
+inhaler_dl_rr <- dl_estimates %>% 
+  mutate(estimate = exp(estimate),
+         lower_bound = exp(lower_bound),
+         upper_bound = exp(upper_bound),
+         lag = -7:7)
+
+# estimate cumulative effect
+cumulative <- sum(dl_estimates$estimate)
+cumulative
+
+# estimate cumulative effect stnd error
+cumulative_se <- sqrt(sum(dl_var))
+cumulative_se
+
+# estimate cumulative CI
+cumulative_ci <- cumulative+cumulative_se*qt(c(1-0.975,0.975), 
+                                             df=df.residual(mod))
+cumulative_ci
+
+exp(cumulative)
+exp(cumulative_ci)
+
+# plot 
+p <- ggplot(inhaler_dl_rr, aes(x=lag, y=estimate, 
+                              ymin=lower_bound, ymax=upper_bound)) +
+  geom_ribbon(alpha = 0.5) +
+  geom_line(size=2)
+
+p
+
+```

### 2017-07-17 r_scripts/analysis/oregon_medication.Rmd
Update oregon_medication.Rmd

@@
 I want to research on the effects of wildfire smoke on pharmacy outcome of asthma.
-NDC (National Drug Code) is the code that the data set offered.
+NDC (National Drug Code) is the variable in the data set that the data set offered.
 The standard NDC I use the link below, which is a ndc list in 2013, which fits my original data set.
@@
 
-The data cleaning and casecrossover study are huge computation tasks, so they are done with server, the following work is done with local R.
+The data cleaning and creation of the casecross-over dataframes are huge computation tasks, so they are done with server, the following work is done with local R.
 
@@
 
-In these comparisons, we examine various methods of smoke PM2.5 estimations and associations with health outcomes using a time-stratified case-crossover study design.
+In these comparisons, we examine GWR of smoke PM2.5 estimations and associations with health outcomes using a time-stratified case-crossover study design.
 Pharmacy outcomes for a patient with a primary diagnosis of cardiopulmonary health outcomes and their date of admission (index time) were identified.
@@
 There are about 77 million claims in the total data set.
-Some have the NDC but another are null or missing because people are going for pharmacy or seeing the doctor.
+Some have the NDC but others are null or missing because people are going for pharmacy or seeing the doctor.
 So I choose the claims in Oregon State with ndc existing, which is about 23 million claims.
@@
 
-After choosing the claims with beta 2 agonists, the data set is reduced to 232,665.
-After choosing people's each day's first visit and limit the wildfire season dates (May 1st, 2013 to September 30th, 2013), the data set is the final data set we research on, containing  only 24,538 claims.
+After choosing the claims with beta 2 agonists, the data set is reduced to 232,665 observations.
+After first recorded pharmacy fill and limit the wildfire season dates (May 1st, 2013 to September 30th, 2013), the data set is the final data set used for analysis and contained 24,538 beta2 agonist records.
 
@@
 
-In this research, we focus on the GWR method (geographically weighted ridge regression).
+For our research question, we focus on the GWR method (geographically weighted ridge regression).
 There are also WRF-Chem model (which substracts WRF-chem from the Weather Research and Forecasting with Chemistry no fire emission).
@@
 
-Because we want the I also make the counterfactual dates for each claim, which means that for each claim, I choose the day the claim shows, to indicate the outcome for this day is "1".
+The time-stratified casecross-over design uses subjects as their own controls.
+For each pharmacy claim, I also created counterfactual observations on the the same day of the week over the study period for each subject to indicate the outcome for this day is "1".
 Then I make the dates week by week before this date, and after this date in the whole wildfire season, and indicate these dates' outcomes are "0".
@@
            # column headers
-           header = c("Outcome", "Cases n", "Less than 15", "(%)", "15 to 65", "(%)", "Greater than 65",
+           header = c("Medication", "Cases n", "Less than 15", "(%)", "15 to 65", "(%)", "Greater than 65",
                       "(%)", "Female", "(%)", "Male", "(%)"),
@@
 
-```{r time stratified zip for casecrossover lag -7 to 0 prepare, eval=FALSE, echo=FALSE}
+```{r time stratified zip for casecrossover lag -7 to 0 prepare, echo=FALSE, eval=FALSE}
 ### Time stratified ------------------------------------------------------------
@@
 
-```{r time stratified county for ts lag -7 to 0 prepare, echo=FALSE, include=FALSE, eval=FALSE}
+```{r time stratified county for ts lag -7 to 0 prepare, echo=FALSE, include=FALSE}
 read_path7 <- paste0("../data_new/medication/oregon_asthma_ndc.csv")

### 2017-07-18 r_scripts/analysis/oregon_medication.Rmd
Update oregon_medication.Rmd

@@
 library(lme4)
+library(maps)
 library(lubridate) # date
@@
 ## Overview
+
+We had previously shown GWR to be a good estimate of wildfire smoke concentrations, and showed associations with respiratory outcomes (including asthma) in Washington state.
+We want to replicate results for Oregon and also look at beta2 agonist pharmacy fills.
+
 For this research I have the Oregon All Payer All Claims (APAC) data set in 2013.
-I want to research on the effects of wildfire smoke on pharmacy outcome of asthma.
+I want to research on the effects of wildfire smoke on beta-2 agonist pharmacy fill, a medication primarily used as a rescue medication for persons with asthma or COPD.
 NDC (National Drug Code) is the variable in the data set that the data set offered.
@@
 
-In these comparisons, we examine GWR of smoke PM2.5 estimations and associations with health outcomes using a time-stratified case-crossover study design.
-Pharmacy outcomes for a patient with a primary diagnosis of cardiopulmonary health outcomes and their date of admission (index time) were identified.
+In these comparisons, we examine GWR method of smoke PM2.5 estimations and associations with health outcomes using a time-stratified case-crossover study design.
+Pharmacy outcomes for a patient with a primary diagnosis of all pharmacy records from the state, and it just so happens we also have APAC data for people that were billed for a cardiopulmonary event that we can link to the pharmacy claim, and their date of admission (index time) were identified.
 We then created counterfactual observations for each patient on the same day of the week for the entire wildfire season (May 1st to Sep 30th, 2013).
@@
 
-There are about 77 million claims in the total data set.
-Some have the NDC but others are null or missing because people are going for pharmacy or seeing the doctor.
+There are about 77 million claims in the total data set, of which 23 million are pharmacy records denoted by the NDC variable.
 So I choose the claims in Oregon State with ndc existing, which is about 23 million claims.
@@
 
-For ndc, I choose the people who use the beta 2 agonists (one inhaler) for the day's first visit to form the sample.
+For ndc, I choose the people the person was billed for filling a beta2 agonist prescription, which we use as a proxy for use to form the sample.
 Because the ndc csv file has more information than what I need, I filter the ndc file by SQLite Studio (code and steps not attached) to a file only containing the beta 2 agonists medication information.
@@
 After first recorded pharmacy fill and limit the wildfire season dates (May 1st, 2013 to September 30th, 2013), the data set is the final data set used for analysis and contained 24,538 beta2 agonist records.
-
-2. Exposure methods for smoke model
-
-For our research question, we focus on the GWR method (geographically weighted ridge regression).
-There are also WRF-Chem model (which substracts WRF-chem from the Weather Research and Forecasting with Chemistry no fire emission).
-For the WRF-Chem variable with the 'smoke' designator, this is WRF-Chem - WRF-Chem no fire.
-For Geo-Weighted Regression and Kriging model, with 'smk' designator, I subtracted off the 'Background' estimates of smoke, which I believe are the monthly averages of PM2.5 for a given grid.
-However, the 'Background' is the mean of total period from May 1st to Sep 30th 2013 so I have to assign the mean for every day.
-To reduce the occurance error and compare the differences I will analyze all three models.
-
-3. Analytic method
-
-We use the conditional logistic regression model using the survival package in R.
-Each conditional logistic regression model accounts for the subject, and adjusts for temperature from the WRF-Chem model.
-The conditional logistic regression model \beta represents the change in risk of an pharmacy event associated with a short-term unit increase in exposure, and can be calculated as an average difference between exposure at the index time and a weighted average of exposure at all times in the referent window.
-
-```{r import original data, eval=FALSE, echo=FALSE}
-### Import whole data set ------------------------------------------------------
-read_path <- paste0("../../../data/data_original/gan_episodes_of_care.txt")
-start_time <- Sys.time()
-oregon_df <- fread(read_path, sep = "|", showProgress = T) 
-stop_time <-  Sys.time() - start_time 
-# time it took
-stop_time # 7 mins
-
-## Basic cleaning
-start_time <- Sys.time()
-oregon_df2 <- oregon_df %>% 
-  filter((!ndc=="")&(!ndc=="*NULL*")) %>% 
-  # filter State is Oregon
-  filter(STATE=="OR")
-stop_time <-  Sys.time() - start_time 
-# time it took
-stop_time # 40 secs
-
-write_path <- paste0('../../../data/data_new/',
-                     'oregon_ndc.csv')
-write_csv(oregon_df2, write_path)
-```
-
-### Make category index for beta-2 agonists.
-First I filtered the whole data set with ndc existing and only Oregon state.
-Then there are about 24 million claims left.
-Then I filtered the data set with only beta-2-agonists and found about 233 thousand claims (83787 unique person).
-By using these unique personkey, I found the claims of them in the total data set, which is about 8.9 million claims.
-This the data set that I use for the category classification.
 
@@
 
+```
+
+2. Exposure methods for smoke model
+
+For our research question, we focus on the GWR method (geographically weighted ridge regression).
+For Geo-Weighted Regression model, with 'smk' designator, I subtracted off the 'Background' estimates of smoke, which I believe are the monthly averages of PM2.5 for a given grid.
+However, the 'Background' is the mean of total period from May 1st to Sep 30th 2013 so I have to assign the mean for every day.
+
+3. Analytic method
+
+We use the conditional logistic regression model using the survival package in R.
+Each conditional logistic regression model accounts for the subject, and adjusts for temperature from the WRF-Chem model.
+The conditional logistic regression model \beta represents the change in risk of an pharmacy event associated with a short-term unit increase in exposure, and can be calculated as an average difference between exposure at the index time and a weighted average of exposure at all times in the referent window.
+
+```{r import original data, eval=FALSE, echo=FALSE}
+### Import whole data set ------------------------------------------------------
+read_path <- paste0("../../../data/data_original/gan_episodes_of_care.txt")
+start_time <- Sys.time()
+oregon_df <- fread(read_path, sep = "|", showProgress = T) 
+stop_time <-  Sys.time() - start_time 
+# time it took
+stop_time # 7 mins
+
+## Basic cleaning
+start_time <- Sys.time()
+oregon_df2 <- oregon_df %>% 
+  filter((!ndc=="")&(!ndc=="*NULL*")) %>% 
+  # filter State is Oregon
+  filter(STATE=="OR")
+stop_time <-  Sys.time() - start_time 
+# time it took
+stop_time # 40 secs
+
+write_path <- paste0('../../../data/data_new/',
+                     'oregon_ndc.csv')
+write_csv(oregon_df2, write_path)
 ```
@@
 There are 488 zipcodes in total (4 are missing but I use google to find it).
+This is a sensitivity analysis to confirm the casecross over results.
 
@@
 plot_ts <- ggplot(or_ndc_county_ts, aes(y =rate_per_100k, x = date)) +
-           geom_jitter()
+           geom_jitter() +
+           ylab("Rate of 100k of Beta2 agonist fills") +
+           xlab("Year 2013")
 
@@
 plot_county <- ggplot(or_ndc_county_ts, aes(y =rate_per_100k, x = date)) +
-               geom_jitter() + facet_wrap(~county)
+               geom_jitter() + facet_wrap(~county) +
+               ylab("Rate of 100k of Beta2 agonist fills")
 
@@
 
+plot_smk_ts <- ggplot(or_ndc_county_ts, aes(y =geo_smk_pm10, x = date)) +
+           geom_jitter() +
+           ylab("GWR smoke concentration (10ug/m^3)") +
+           xlab("Year 2013")
+
+print(plot_smk_ts)
+
 ### check if the time series has the association with get smoke
 a <- glm(rate_per_100k ~ geo_smk_pm10, data = or_ndc_county_ts, family = "poisson")
-summary(a)
+# summary(a)
 
 plot_health_smk <- ggplot(or_ndc_county_ts, aes(x = geo_smk_pm10, y = log(rate_per_100k+0.1))) +
-  geom_point(aes(colour = c("red")), alpha = 0.2,  size = 1) + 
+  geom_point() + 
+  # geom_point(aes(colour = c("red")), alpha = 0.2,  size = 1) + 
   # scale_colour_manual(values = c("red", "blue")) +
@@
   # facet_wrap(~county, scale  = "free") +
-  ylab("Log Outcome Rate per 100,000 Persons") +
-  xlab("GWR Smoke PM2.5 per 10*10ug/m^3") +
+  ylab("Log Beta 2 agonists Rate per 100,000 Persons") +
+  xlab("GWR Smoke PM2.5 per 10ug/m^3") +
   theme_bw()
@@
 # unnecessary
-plot_county_health_smk <- ggplot(or_ndc_county_ts, aes(x = geo_smk_pm10, y = log(rate_per_100k+0.1))) +
-  geom_point(aes(colour = county), alpha = 0.2,  size = 1) + 
+# plot_county_health_smk <- ggplot(or_ndc_county_ts, aes(x = geo_smk_pm10, y = log(rate_per_100k+0.1))) +
+#   geom_point(aes(colour = county), alpha = 0.2,  size = 1) + 
   # scale_colour_manual(values = c("red", "blue")) +
@@
   #scale_y_continuous(limits = c(0, 1.5)) +
-  facet_wrap(~county, scale  = "free") +
-  ylab("Log Outcome Rate per 100,000 Persons") +
-  xlab("GWR Smoke PM2.5 per 10*10ug/m^3") +
-  theme_bw()
+#   facet_wrap(~county, scale  = "free") +
+#   ylab("Log Outcome Rate per 100,000 Persons") +
+#   xlab("GWR Smoke PM2.5 per 10*10ug/m^3") +
+#   theme_bw()
 
@@
 
-ggsave("../plot_new/ndc_county_rate_health_smk.pdf", plot = plot_county_health_smk, 
-       width = 12, height = 8, units = "in")
+# ggsave("../plot_new/ndc_county_rate_health_smk.pdf", plot = plot_county_health_smk, 
+#        width = 12, height = 8, units = "in")
 
@@
 
+### map of smoke days
+
+```{r}
+meta_ts <- read_csv("../data/meta_timeseries.csv") %>% 
+  # creating 10 unit increase var of geo smoke and day of week
+  mutate(geo_smk10 = geo_smk_pm/10,
+         day = weekdays(date)) %>% 
+  # create smoke and lag variables
+  mutate(smoke0 = ifelse(geo_smk_pm > 0, 1, 0),
+         smoke5 = ifelse(geo_smk_pm > 5, 1, 0),
+         smoke10 = ifelse(geo_smk_pm > 10, 1, 0),
+         smoke15 = ifelse(geo_smk_pm > 15, 1, 0), 
+         # lag variable
+         geo_smk_lag1 = lag(geo_smk10, 1, order_by = county),
+         geo_smk_lag2 = lag(geo_smk10, 2, order_by = county),
+         geo_smk_lag3 = lag(geo_smk10, 3, order_by = county),
+         geo_smk_lag4 = lag(geo_smk10, 4, order_by = county),
+         geo_smk_lag5 = lag(geo_smk10, 5, order_by = county),
+         geo_smk_lag6 = lag(geo_smk10, 6, order_by = county),
+         geo_smk_lag7 = lag(geo_smk10, 7, order_by = county)) %>% 
+  ungroup()
+
+# smoke wave days (at least 2 consecutive days in the 98th percentile of PM)
+quantile(meta_ts$geo_wt_pm, probs = c(0.95,0.98,0.99, 0.999))
+# following Coco's smoke wave day, the 98th percentile for PM2.5 concentrations
+# would be >26.9
+
+smoke_wave <- meta_ts %>% 
+  select(county, date, geo_wt_pm) %>% 
+  mutate(high_pm_day = ifelse(geo_wt_pm > 26.9, 1, 0)) %>% 
+  group_by(county) %>% 
+  # identify consecutive days
+  mutate(smoke_wave = ifelse(lag(high_pm_day, 
+    order_by = county)==1 & high_pm_day ==1, 1, 0)) %>% 
+  select(-high_pm_day, -geo_wt_pm)
+
+# merge in smokewave day
+meta_ts <- meta_ts %>% 
+  right_join(smoke_wave, by = c("county", "date"))
+
+# count of state outcome timeseries
+state_outcome_rate <- meta_ts %>% 
+  group_by(date) %>% 
+  # remove columns I don't want to calculate
+  select(-fips, -county) %>% 
+  # add up outcomes in each county
+  summarise_each(funs(sum), population, n_obs:broken_arm_n) %>% 
+  mutate_each(funs((./population*100000)), n_obs:broken_arm_n)
+
+# average of environmental measures, joined with outcomes
+state_smoke <- meta_ts %>% 
+  group_by(state, date) %>% 
+  select(-fips, -county) %>% 
+  summarise_each(funs(mean, median, min, max), wrf_pm:wrf_temp) %>% 
+  full_join(state_outcome_rate, by = c("date"))
+
+# only 150 smoke wave days
+xtabs(~smoke_wave,  meta_ts)
+# if we compare to our geo smoke > 10, we'd add 362 extra days
+xtabs(~smoke_wave + smoke10, meta_ts)
+# if we used smoke >15 we'd add 164 extra days 
+xtabs(~smoke_wave + smoke15, meta_ts)
+
+# small multiples dataset
+or_smk <- or_ts %>% 
+  select(county, date, wrf_smk_pm, krig_smk_pm, geo_smk_pm) %>% 
+  gather(key = smk_method, value = pm, -date, -county)
+
+or_smk_plot <- ggplot(or_smk, aes(x = date, y = pm)) +
+  geom_point(aes(colour = smk_method), alpha = 0.8, size = 0.8) +
+  facet_wrap(~county) +
+  ggtitle("Oregon: 2013 Fire Season Smoke PM2.5") +
+  ylab("Smoke PM2.5 ug/m^3") +
+  xlab("Year: 2013") +
+  theme_bw()
+
+or_smk_plot
+
+# data wrangle ----
+#first step is to summarise the counts of smoke days in each county 
+or_smk_count <- meta_ts %>% 
+  group_by(county) %>% 
+  # summing up smoke days
+  summarise(smk_wave_n = sum(smoke_wave), smk_day0 = sum(smoke0), 
+            smk_day5 = sum(smoke5), smk_day10 = sum(smoke10),
+            smk_day15 = sum(smoke15)) %>% 
+  # county needs to be lowercase to join to maps dataframe
+  mutate(county = tolower(county))
+
+# maps package will be loaded to pull map data and create a dataframe
+or_county_df <- map_data("county", "oregon")
+
+# join county counts to map df
+or_smk_df <- or_county_df %>% 
+  right_join(or_smk_count, by = c("subregion" = "county")) %>% 
+  # small multiples of different smoke cutoff methods
+  gather(key = smk_cut_method, value = smoke_days, -long:-subregion) %>% 
+  # filter out smoke day 0 since it's not very specific
+  filter(smk_cut_method != "smk_day0") %>% 
+  # rename smoke day cutoffs 
+  mutate(smk_cut_method = 
+    ifelse(smk_cut_method == "smk_day5","Smoke PM2.5 > 5 ug/m^3",
+    ifelse(smk_cut_method == "smk_day10", "Smoke PM2.5 > 10 ug/m^3",
+    ifelse(smk_cut_method == "smk_day15", "Smoke PM2.5 > 15 ug/m^3",
+    ifelse(smk_cut_method == "smk_wave_n", "Smoke Wave", NA)))),
+    # preserve smoke method order for small multiple
+    smk_cut_method = factor(smk_cut_method, 
+      levels = c("Smoke PM2.5 > 5 ug/m^3", "Smoke PM2.5 > 10 ug/m^3", 
+                 "Smoke PM2.5 > 15 ug/m^3", "Smoke Wave")))
+
+# map ----
+
+# smoke days where GWR smoke > 15
+smoke_map <- ggplot(or_smk_df, aes(x=long, y=lat, group=group)) +
+  # fill with number of smoke days
+  geom_polygon(aes(fill = smoke_days), alpha = 0.7) +
+  scale_fill_gradient("Smoke-Impacted Days",
+                      low = "#0b486b", high = "#f56217") +
+  # add county path on top
+  geom_path() +
+  facet_wrap(~smk_cut_method) +
+  xlab("Longitude") +
+  ylab("Latitude") +
+  ggtitle("Oregon: Counties affected by smoke") +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    strip.background = element_rect(colour=NA, fill=NA))
+
+smoke_map
+
+```
+
 ## 2. descriptive table of total ndc inhalhers; and age and sex category; Plots
@@
 
-For all of 2013, there were a total of millons of all claims records in the APAC dataset.
+For all of 2013, there were a total of 77 millon claim records in the APAC dataset.
 During the May 1st to September 30th wildfire season in Oregon state, there were 24217 pharmacy visits for the ndc of beta 2 agonists.
@@
 
-# dataframe list
-method_list <- c('WRF-Chem Smoke', 'Geo-Weighted Smoke', 'Kriging Smoke')
-
 # outcome name
-outcome_name <- "Inhalers"
+outcome_name <- "beta-2 agonist"
 
@@
 tab <- htmlTable(txtRound(asthma_point_est_df, digits = 1), 
-           caption = "Number of cases for each outcome observed from May 1st to September 30st, 2013",
+           caption = "Total number of cases for pharmacy fills of beta-2 agonist observed from May 1st to September 30st, 2013",
            # column headers
@@
 
-### Overall plot
+### Overall plot with GWR method
 
@@
 # outcome name
-outcome <- "Asthma Inhalers"
-outcome_name <- "Asthma Inhalers"
+outcome <- "Beta-2 agonists"
+outcome_name <- "Beta-2 agonists"
 
@@
   # save figure
-  ggsave("../plot_new/ndc_overall_three.pdf", plot = print_plot, 
+  ggsave("../plot_new/ndc_overall_gwr.pdf", plot = print_plot, 
        width = 12, height = 8, units = "in")
@@
 
-For some outcomes, such like Age > 65 in WRF-Chem method, the CI is the only insignificant one in the plot.
+For some outcomes, such like Age > 65 in WRF-Chem method, the CI is the only not significant one in the plot.
+
+The plot shows that overall, the risk of pharmacy fills of beta-2 agonist will increase significantly as the smoke concentrate increases.
+For age group of 15-65, the risk is the highest, for 10 ug/m^3 increase of pm2.5, the risk of filling the beta-2 agonist will increase about 9%.
+For age groups which are less than 15 or more than 65, the risk is similar, both of them will increase about 5.6%.
 
@@
 
-All the CI do not include 1, which means the use of asthma medication (beta 2 agonists) increases about 5% when smoke increase 10 units.
+All the CI do not include 1, which means the pharmacy filling (beta 2 agonists) increases about 5% when smoke increase 10 units.
 So the risk for getting asthma will increase about 5% ~ 8% when the smoke PM2.5 increase 10 units.
+Also, the risk for female is much higher than that for male.
+
+These results show that there may be effect modification by sex, where females might be more likely to fill a beta-2 agonist on the same day it's really smoky for the GWR estimates.
 
@@
 
-Remove the missing values for GWR smoke. 
-But I only use GWR smoke of zip for the case crossover study.
+Remove the missing values for GWR smoke. But I only use GWR smoke of zip for the case crossover study.
 

### 2017-07-18 r_scripts/analysis/oregon_medication.Rmd
Update oregon_medication.Rmd

@@
 ---
-title: "Oregon 2013 Medication for Asthma Inhalers"
+title: "Oregon 2013 Medication for Beta-2 agonists"
 author: "Jingyang Liu"
@@
 
-# two-stage distributed lag model libraries
+# distributed lag model libraries
 library(dlnm)
@@
 We had previously shown GWR to be a good estimate of wildfire smoke concentrations, and showed associations with respiratory outcomes (including asthma) in Washington state.
-We want to replicate results for Oregon and also look at beta2 agonist pharmacy fills.
+We want to replicate results for Oregon and also look at beta-2 agonist pharmacy fills.
 
 For this research I have the Oregon All Payer All Claims (APAC) data set in 2013.
-I want to research on the effects of wildfire smoke on beta-2 agonist pharmacy fill, a medication primarily used as a rescue medication for persons with asthma or COPD.
+I want to research on the effects of wildfire GWR smoke on beta-2 agonist pharmacy fill, a medication primarily used as a rescue medication for persons with asthma or COPD.
 NDC (National Drug Code) is the variable in the data set that the data set offered.
-The standard NDC I use the link below, which is a ndc list in 2013, which fits my original data set.
+The standard NDC I use the link below, which is a ndc list in 2013 and fits my original data set.
 (http://ww2.ncqa.org/hedis-quality-measurement/hedis-measures/hedis-2013/hedis-2013-final-ndc-lists)
@@
 The data cleaning and creation of the casecross-over dataframes are huge computation tasks, so they are done with server, the following work is done with local R.
+We analyze the case crossover results with overall, age strarified and sex stratified categories for the pharmacy fills, and then we also analyze with the distributed lag method, and check with time series analysis.
 
@@
 
-plot_smk_ts <- ggplot(or_ndc_county_ts, aes(y =geo_smk_pm10, x = date)) +
-           geom_jitter() +
-           ylab("GWR smoke concentration (10ug/m^3)") +
-           xlab("Year 2013")
-
-print(plot_smk_ts)
+# plot_smk_ts <- ggplot(or_ndc_county_ts, aes(y =geo_smk_pm10, x = date)) +
+#           geom_jitter() +
+#           ylab("GWR smoke concentration (10ug/m^3)") +
+#           xlab("Year 2013")
+
+# print(plot_smk_ts)
 
@@
 
-```{r}
-meta_ts <- read_csv("../data/meta_timeseries.csv") %>% 
-  # creating 10 unit increase var of geo smoke and day of week
-  mutate(geo_smk10 = geo_smk_pm/10,
-         day = weekdays(date)) %>% 
-  # create smoke and lag variables
-  mutate(smoke0 = ifelse(geo_smk_pm > 0, 1, 0),
-         smoke5 = ifelse(geo_smk_pm > 5, 1, 0),
-         smoke10 = ifelse(geo_smk_pm > 10, 1, 0),
-         smoke15 = ifelse(geo_smk_pm > 15, 1, 0), 
+```{r smoke days and map, warning=FALSE, message=FALSE, echo=FALSE}
+or_ts <- or_ndc_county_ts %>% 
+  mutate(smoke0 = ifelse(geo_smk_pm_county > 0, 1, 0),
+         smoke5 = ifelse(geo_smk_pm_county > 5, 1, 0),
+         smoke10 = ifelse(geo_smk_pm_county > 10, 1, 0),
+         smoke15 = ifelse(geo_smk_pm_county > 15, 1, 0), 
          # lag variable
-         geo_smk_lag1 = lag(geo_smk10, 1, order_by = county),
-         geo_smk_lag2 = lag(geo_smk10, 2, order_by = county),
-         geo_smk_lag3 = lag(geo_smk10, 3, order_by = county),
-         geo_smk_lag4 = lag(geo_smk10, 4, order_by = county),
-         geo_smk_lag5 = lag(geo_smk10, 5, order_by = county),
-         geo_smk_lag6 = lag(geo_smk10, 6, order_by = county),
-         geo_smk_lag7 = lag(geo_smk10, 7, order_by = county)) %>% 
+         geo_smk_lag1 = lag(geo_smk_pm10, 1, order_by = county),
+         geo_smk_lag2 = lag(geo_smk_pm10, 2, order_by = county),
+         geo_smk_lag3 = lag(geo_smk_pm10, 3, order_by = county),
+         geo_smk_lag4 = lag(geo_smk_pm10, 4, order_by = county),
+         geo_smk_lag5 = lag(geo_smk_pm10, 5, order_by = county),
+         geo_smk_lag6 = lag(geo_smk_pm10, 6, order_by = county),
+         geo_smk_lag7 = lag(geo_smk_pm10, 7, order_by = county)) %>% 
   ungroup()
@@
 # smoke wave days (at least 2 consecutive days in the 98th percentile of PM)
-quantile(meta_ts$geo_wt_pm, probs = c(0.95,0.98,0.99, 0.999))
+# quantile(or_ts$geo_wt_pm_county, probs = c(0.95,0.98,0.99, 0.999))
 # following Coco's smoke wave day, the 98th percentile for PM2.5 concentrations
-# would be >26.9
-
-smoke_wave <- meta_ts %>% 
-  select(county, date, geo_wt_pm) %>% 
-  mutate(high_pm_day = ifelse(geo_wt_pm > 26.9, 1, 0)) %>% 
+# would be >16.3
+
+smoke_wave <- or_ts %>% 
+  select(county, date, geo_wt_pm_county) %>% 
+  mutate(high_pm_day = ifelse(geo_wt_pm_county > 16.33, 1, 0)) %>% 
   group_by(county) %>% 
@@
     order_by = county)==1 & high_pm_day ==1, 1, 0)) %>% 
-  select(-high_pm_day, -geo_wt_pm)
+  select(-high_pm_day, -geo_wt_pm_county)
 
 # merge in smokewave day
-meta_ts <- meta_ts %>% 
+or_ts <- or_ts %>% 
   right_join(smoke_wave, by = c("county", "date"))
@@
 # count of state outcome timeseries
-state_outcome_rate <- meta_ts %>% 
+state_outcome_rate <- or_ts %>% 
   group_by(date) %>% 
   # remove columns I don't want to calculate
+  select(-fips, -county)
+
+# average of environmental measures, joined with outcomes
+state_smoke <- or_ts %>% 
+  group_by(date) %>% 
   select(-fips, -county) %>% 
-  # add up outcomes in each county
-  summarise_each(funs(sum), population, n_obs:broken_arm_n) %>% 
-  mutate_each(funs((./population*100000)), n_obs:broken_arm_n)
-
-# average of environmental measures, joined with outcomes
-state_smoke <- meta_ts %>% 
-  group_by(state, date) %>% 
-  select(-fips, -county) %>% 
-  summarise_each(funs(mean, median, min, max), wrf_pm:wrf_temp) %>% 
-  full_join(state_outcome_rate, by = c("date"))
-
-# only 150 smoke wave days
-xtabs(~smoke_wave,  meta_ts)
+  summarise_each(funs(mean, median, min, max), wrf_pm_county:wrf_temp_county) # %>% 
+  # full_join(state_outcome_rate, by = c("date"))
+
+# only 55 smoke wave days
+# xtabs(~smoke_wave,  or_ts)
 # if we compare to our geo smoke > 10, we'd add 362 extra days
-xtabs(~smoke_wave + smoke10, meta_ts)
+# xtabs(~smoke_wave + smoke10, or_ts)
 # if we used smoke >15 we'd add 164 extra days 
-xtabs(~smoke_wave + smoke15, meta_ts)
+# xtabs(~smoke_wave + smoke15, or_ts)
 
@@
 or_smk <- or_ts %>% 
-  select(county, date, wrf_smk_pm, krig_smk_pm, geo_smk_pm) %>% 
+  select(county, date, wrf_smk_pm_county, krig_smk_pm_county, geo_smk_pm_county) %>% 
   gather(key = smk_method, value = pm, -date, -county)
@@
 #first step is to summarise the counts of smoke days in each county 
-or_smk_count <- meta_ts %>% 
+or_smk_count <- or_ts %>% 
   group_by(county) %>% 
@@
 # outcome name
-outcome_name <- "beta-2 agonist"
+outcome_name <- "Beta - 2 agonist"
 
@@
 
+The overall plot shows that when smoke concentration increases 10ug/m^3, the risk to get beta-2 agonist fill will increase about 7.5% significantly.
+
 ```{r overall, warning =F, echo = F, results='asis'} 
@@
 
-method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+method_list <- c('Geo-Weighted Smoke')
 
@@
 # outcome name
-outcome <- "Beta-2 agonists"
+outcome <- "Beta - 2 agonists"
 outcome_name <- "Beta-2 agonists"
@@
 #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
-pm_estimates_df <- inhaler_ndc_df[, c(87, 92, 91, 93)]/10  # create 10 unit increases
+pm_estimates_df <- inhaler_ndc_df[, c(91, 93)]/10  # create 10 unit increases
 
@@
   # remove missing pm values
-  filter(!is.na(wrf_smk_pm_zip)) %>% 
+  filter(!is.na(geo_smk_pm_zip)) %>% 
   # the following code makes sure that the counterfactual values retained are 
@@
 # second loop to run a model for each pm estimation method
-for(j in 38:40){
-
-  # variable to model 
-  var_name <- colnames(df_analysis[j])
+j <- 38
 
@@
   # in the final dataframe
-  } else {point_estimates[row_n, 2] <- method_list[row_n]
+  } else {point_estimates[row_n, 2] <- method_list
           point_estimates[row_n, 3] <- 0
@@
 wrf_geo_age <- combined_point_est_df %>% 
-  # filter columns to just geo-smk
-  # filter(pm_method == "GWR Smoke" |
-    #      pm_method == "WRF-Chem Smoke") %>% 
   # subset columns I want to put in to the table
@@
          # row group by outcome
-         rgroup = "Asthma Inhalers",
-         n.rgroup = c(rep(3, 1)), # 6 rows for each age cat for each outcome
+         rgroup = "Beta - 2 agonist",
+         n.rgroup = c(rep(1, 1)), # 6 rows for each age cat for each outcome
          # column headers
@@
 # outcome name
-outcome <- "asthma inhalers"
-outcome_name <- "Asthma inhalers"
+outcome <- "Beta - 2 agonists"
+outcome_name <- "Beta - 2 agonists"
 
@@
   # empty matrix (12 x 10 matrix)
-  point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)
+  point_estimates <- matrix(nrow = 1, ncol = 10, byrow = T)
 
@@
     # remove missing pm values
-    filter(!is.na(wrf_smk_pm_zip)) %>% 
+    filter(!is.na(geo_smk_pm_zip)) %>% 
     # limit to specific age category
@@
 
-  # second loop to run a model for each pm estimation method
-  for(j in 38:40){
+ j <- 38
 
@@
             point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
-    } # end methods loop
+     # end methods loop
 
@@
          # row group by outcome
-         rgroup = "Asthma inhalers",
-         n.rgroup = c(rep(9, 1)), # 6 rows for each age cat for each outcome
+         rgroup = "Beta - 2 agonists",
+         n.rgroup = c(rep(3, 1)), # 6 rows for each age cat for each outcome
          # column headers
@@
   # save figure
-  ggsave("../plot_new/ndc_age_fig_three.pdf", plot = print_plot, 
+  ggsave("../plot_new/ndc_age_fig_gwr.pdf", plot = print_plot, 
        width = 12, height = 8, units = "in")
@@
 
-Association with smoke PM~2.5~ and health outcomes stratified by sex.
+The association with smoke PM~2.5~ and health outcomes stratified by sex is shown below.
 
@@
 
-These results show that there may be effect modification by sex, where females might be more likely to fill a beta-2 agonist on the same day it's really smoky for the GWR estimates.
+These results show that there might be effect modification by sex, where females might be more likely to fill a beta-2 agonist on the same day it's really smoky for the GWR estimates.
 
 ```{r sex strata cdc met adj, warning = F, echo = F, results='asis'} 
-method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')
+method_list <- c('Geo-Weighted Smoke')
 
@@
     # remove missing pm values
-    filter(!is.na(wrf_smk_pm_zip)) %>% 
+    filter(!is.na(geo_smk_pm_zip)) %>% 
     # limit to specific sex category
@@
 
-  # second loop to run a model for each pm estimation method
-  for(j in 38:40){
-
+  j <- 38
     # variable to model 
@@
             point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
-    } # end methods loop
+    # end methods loop
 
@@
            # row group by outcome
-           rgroup = "Asthma inhalers",
-           n.rgroup = c(rep(6, 1)), # 2 rows for each sex strata for each outcome
+           rgroup = "Beta - 2 agonists",
+           n.rgroup = c(rep(2, 1)), # 2 rows for each sex strata for each outcome
            # column headers
@@
   # save figure
-  ggsave("../plot_new/ndc_sex_fig_three.pdf", plot = print_plot, 
+  ggsave("../plot_new/ndc_sex_fig_gwr.pdf", plot = print_plot, 
        width = 12, height = 8, units = "in")
@@
 
-Remove the missing values for GWR smoke. But I only use GWR smoke of zip for the case crossover study.
+Although the casecrossover study shows a significant association of getting beta-2 agonist pharmacy fills and smoke concentration, we still wonder the association of the pharmacy fills with some lag days.
+For medication analysis, people could get the pharmacy fills one or some days later than the day they find they have the disease.
+Then we make 0 (the original day) to 7 lags (one week) to see association of pharmacy fills with different lagged days.
 
@@
 # fill outcome name in first column
-fit_mat[,1] <- "inhalers"
+fit_mat[,1] <- "Beta - 2 agonists"
 # fill in exposure (geo smoke)
@@
 
-fit_mat
+kable(fit_mat, caption="Distributed lag spline of 3-6 degree of freedom fit by AIC")
 
@@
 # kable
-knitr::kable(lag_spline_best_fit, caption = paste0("Distributed lag spline",
-  "degree of freedom best fit by AIC for mixed model approach"))
+knitr::kable(lag_spline_best_fit, caption = paste0("Distributed lag spline ",
+  "degree of freedom best fit by AIC"))
 ```
@@
 
-```{r DLM for casecrossover, echo = FALSE, message=FALSE, warning=FALSE}
+When choosing the df of 5, the plot shows that the same day of fill with smoke exposure has the highest significant association.
+But other lags shows that pharmacy fills with lagged days do not has the significant association with smoke.
+And the association tends to decrease as lag goes.
+
+```{r DLM for casecrossover df 5, echo = FALSE, message=FALSE, warning=FALSE}
 inhaler_geo_lag <- inhaler_ndc_lag_df %>%
@@
 # the nrow B should equal ncol(pm)
-dim(B)
+# dim(B)
 
@@
 
-coef(fit)
-summary(fit)
-AIC(fit)
+# coef(fit)
+# summary(fit)
+# AIC(fit)
 
@@
 
-DLestimate
+# DLestimate
 
@@
 cumulative <- sum(DLestimate$estimate)
-cumulative
+# cumulative
 
 cumulative_se <- sqrt(sum(DLvar))
-cumulative_se
+# cumulative_se
 
 cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
-cumulative_CI
+# cumulative_CI
+
+cum <- cbind(cumulative, cumulative_se, cumulative_CI[1], cumulative_CI[2])
+colnames(cum) <- c("Cumulative Est", "Cumulative SE", "Lower 95% CI", "Upper 95% CI")
+
+knitr::kable(cum, caption = paste0("Distributed lag 0-7 spline estimates, se and CI"))
 
@@
 ### Distributed lag -7 to 0 for casecrossover
+
+Then we make the reversed lag to check with the same df of 5. The plot shows the same result as previous analysis.
+
+```{r DLM for casecrossover lag -7 to 0 df 5, echo = FALSE, message=FALSE, warning=FALSE}
+inhaler_geo_lag <- inhaler_ndc_lag_df %>%
+  select(166:160, 159) # GWR-method
+
+inhaler_temp_lag <- inhaler_ndc_lag_df %>%
+  select(156:150, 93) # GWR-method
+
+lag_max <- 0
+n1 <- length(inhaler_geo_lag[,1])
+n2 <- length(inhaler_temp_lag[,1])
+
+inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
+inhaler_temp_lag <- as.matrix(inhaler_temp_lag)
+
+B <- ns(-7:0, df =5, intercept=TRUE)
+
+# the nrow B should equal ncol(pm)
+# dim(B)
+
+inhaler_geo_lag_B <- inhaler_geo_lag%*%B
+inhaler_temp_lag_B <- inhaler_temp_lag%*%B
+
+fit <- clogit(outcome ~ inhaler_geo_lag_B + inhaler_temp_lag_B + strata(personkey), 
+              inhaler_ndc_lag_df)
+
+# coef(fit)
+# summary(fit)
+# AIC(fit)
+
+D <- cbind(B,B)
+
+# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
+dlparms <- c(1,2,3,4,5,6,7,8,9,10)
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms]) 
+DLvar <- D%*%vcov(fit)[dlparms,dlparms]%*%t(D)
+DLestimate$SE <- sqrt(diag(DLvar))
+DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
+DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
+
+DLestimate$estimate <- exp(DLestimate$estimate)
+DLestimate$lower <- exp(DLestimate$lower)
+DLestimate$upper <- exp(DLestimate$upper)
+
+# DLestimate
+
+DLestimate$lag <- -7:0
+p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
+p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
+p <- p + geom_line(aes(x=lag), linetype=2, size=2)
+p <- p + theme_bw()
+p
+
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms])
+
+cumulative <- sum(DLestimate$estimate)
+# cumulative
+
+cumulative_se <- sqrt(sum(DLvar))
+# cumulative_se
+
+cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
+# cumulative_CI
+
+cum <- cbind(cumulative, cumulative_se, cumulative_CI[1], cumulative_CI[2])
+colnames(cum) <- c("Cumulative Est", "Cumulative SE", "Lower 95% CI", "Upper 95% CI")
+
+knitr::kable(cum, caption = paste0("Distributed lag -7 to 0 spline estimates, se and CI"))
+
+```
 
@@
 
-```{r time stratified county for ts lag -7 to 0 prepare, echo=FALSE, include=FALSE}
+```{r time stratified county for ts lag -7 to 0 prepare, echo=FALSE, include=FALSE, eval=FALSE}
 read_path7 <- paste0("../data_new/medication/oregon_asthma_ndc.csv")
@@
 
-```{r time stratified zip for casecrossover lag -7 to 0 df, echo=FALSE, include=FALSE}
+```{r time stratified zip for casecrossover lag -7 to 0 df, echo=FALSE, include=FALSE, eval=FALSE}
 read_path10 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover_zip_lag15.csv")
@@
 
-```{r DLM for casecrossover lag -7 to 0, echo = FALSE, message=FALSE, warning=FALSE}
-inhaler_geo_lag <- inhaler_ndc_lag_15_df %>%
-  select(177:171, 178) # GWR-method
-
-inhaler_temp_lag <- inhaler_ndc_lag_15_df %>%
-  select(170:164, 93) # temp
-
-lag_max <- 0
-n1 <- length(inhaler_geo_lag[,1])
-n2 <- length(inhaler_temp_lag[,1])
-
-inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
-inhaler_temp_lag <- as.matrix(inhaler_temp_lag)
-
-B <- ns(-7:lag_max, df =5, intercept=TRUE)
-
-# the nrow B should equal ncol(pm)
-dim(B)
-
-inhaler_geo_lag_B <- inhaler_geo_lag%*%B
-inhaler_temp_lag_B <- inhaler_temp_lag%*%B
-
-fit <- clogit(outcome ~ inhaler_geo_lag_B + inhaler_temp_lag_B + strata(personkey), 
-              inhaler_ndc_lag_15_df)
-
-coef(fit)
-summary(fit)
-AIC(fit)
-
-D <- cbind(B,B)
-
-# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
-dlparms <- c(1,2,3,4,5,6,7,8,9,10)
-DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms]) 
-DLvar <- D%*%vcov(fit)[dlparms,dlparms]%*%t(D)
-DLestimate$SE <- sqrt(diag(DLvar))
-DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
-DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
-
-DLestimate$estimate <- exp(DLestimate$estimate)
-DLestimate$lower <- exp(DLestimate$lower)
-DLestimate$upper <- exp(DLestimate$upper)
-
-DLestimate
-
-DLestimate$lag <- -7:lag_max
-p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
-p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
-p <- p + geom_line(aes(x=lag), linetype=2, size=2)
-p <- p + theme_bw()
-p
-
-DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms])
-
-cumulative <- sum(DLestimate$estimate)
-cumulative
-
-cumulative_se <- sqrt(sum(DLvar))
-cumulative_se
-
-cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
-cumulative_CI
-
-```
-
 ## Supplement
@@
 # outcome name
-outcome <- "Asthma Inhalers"
+outcome <- "Beta - 2 agonists"
 
@@
 # extract pm values and divide by 10 and ordered
-which(colnames(inhaler_ndc)=="geo_smk_pm_lag1_zip") # 136
-which(colnames(inhaler_ndc)=="wrf_temp_lag1_zip") # 150
+# which(colnames(inhaler_ndc)=="geo_smk_pm_lag1_zip") # 136
+# which(colnames(inhaler_ndc)=="wrf_temp_lag1_zip") # 150
 pm_estimates_df <- inhaler_ndc[, 136:142]/10  # create 10 unit increases
@@
          # row group by outcome
-         rgroup = "Asthma Inhalers",
+         rgroup = "Beta - 2 agonists",
          n.rgroup = c(rep(1, 1)), # 6 rows for each age cat for each outcome
@@
 }
-rr_df
+
+kable(rr_df, caption = "Unconstrained lag for time series")
 
@@
 
+(not shown)
+
 ```{r DLM ts, echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}
@@
 ### 4. Distributed lag -7 to 0 for time series
+
+(not shown)
+
 ```{r DLM for casecrossover lag -7 to 0 temp, echo = FALSE, message=FALSE, warning=FALSE, eval = FALSE}

### 2017-07-20 r_scripts/analysis/oregon_medication.Rmd
Update oregon_medication.Rmd

@@
            ylab("Rate of 100k of Beta2 agonist fills") +
-           xlab("Year 2013")
+           xlab("Year 2013") +
+           ggtitle("Oregon: 2013 Fire Season Pharmacy Fills")
 
@@
                geom_jitter() + facet_wrap(~county) +
-               ylab("Rate of 100k of Beta2 agonist fills")
+               ylab("Rate of 100k of Beta2 agonist fills") +
+               ggtitle("Oregon: 2013 Fire Season Pharmacy Fills by County")
 
@@
   xlab("GWR Smoke PM2.5 per 10ug/m^3") +
+  ggtitle("Regression of Pharmacy Fills on GWR Smoke Concentration") +
   theme_bw()
@@
   ylab("Smoke PM2.5 ug/m^3") +
-  xlab("Year: 2013") +
+  xlab("Year 2013") +
   theme_bw()
@@
     #ylim(0, 2) +
-    xlab('Smoke Estimation Method') +
+    xlab('Overall') +
     # plot theme
@@
 p <- p + geom_line(aes(x=lag), linetype=2, size=2)
-p <- p + theme_bw()
+p <- p + theme_bw() + xlab("lagged days of events") + ylab("Odds Ratio for 10ug/m^3 Increase in PM2.5")
 p
@@
    axis.text.x = element_text(angle = 90, vjust = .75)) +
-  ylab("Relative Risk") +
+  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
   xlab("Outcome") +

### 2017-07-27 r_scripts/analysis/oregon_medication.Rmd
Update oregon_medication.Rmd

@@
 ---
-title: "Oregon 2013 Medication for Beta-2 agonists"
+title: "The GWR Smoke's effect on Oregon 2013 Medication of Beta-2 agonists"
 author: "Jingyang Liu"
@@
 library(lubridate) # date
+library(ggrepel)
+library(sp)
 
@@
 
+# server
+library(dplyr)
+library(data.table)
+library(readxl) 
+library(readr)
+
 ```
 
-## Overview
+## Overview (1, 2)
+
+1.
+Intro: We found an association with wildfire smoke and respiratory outcomes in Washington state 2012 using a novel estimate of smoke concentration, geo-graphically weighted ridge regression (GWR).
+We want to know if smoke impacts other health outcomes, specifically respiratory rescue medications (beta 2 agonists) (abbreviate to SABA)
+
+2. Research question: We evaluated the association between smoke concentrations using the GWR method and SABA refills in Oregon state during the 2013 wildfire season.
 
@@
 
-## Method Description
+## Method Description (3)
+
+3.Methods: use case-crossover for all health outcome figures and results. Make sure smoke is assigned via zipcode. For the descriptive of smoke, I think the county-level for now. We may move to zip.
 
@@
 
-```{r import filtered original data, eval=FALSE, echo=FALSE}
+```{r import filtered original data and make index, eval=FALSE, echo=FALSE}
 ### Import whole data set ------------------------------------------------------
@@
 summary(as.factor(oregon_ndc_index$index))
+
+write_path4 <- paste0('../../../data/data_new/',
+                     'oregon_beta_ndc_index_full.csv')
+write_csv(oregon_ndc_index, write_path4)
 
@@
 
-write_path <- paste0('../../../data/data_new/',
-                     'oregon_asthma_sample.csv')
-write_csv(or_asthma_sample, write_path)
-
-write_path2 <- paste0('../../../data/data_new/',
-                     'oregon_ndc_with_index.csv')
-
-write_csv(or_ndc_asthma_copd_sum, write_path2)
+# write_path <- paste0('../../../data/data_new/',
+#                      'oregon_asthma_sample.csv')
+# write_csv(or_asthma_sample, write_path)
+
+# write_path2 <- paste0('../../../data/data_new/',
+#                      'oregon_ndc_with_index.csv')
+
+# write_csv(or_ndc_asthma_copd_sum, write_path2)
 
@@
 
-```{r time stratified county for ts, echo=FALSE, include=FALSE, eval=FALSE}
+```{r time stratified county for ndc ts, echo=FALSE, include=FALSE, eval=FALSE}
 read_path7 <- paste0("../data_new/medication/oregon_asthma_ndc.csv")
@@
   mutate(st_county_fips = with(oregon_fips, paste0(st_code, fips)))
-
-### Join county with zip
-# read_path <- paste0('../data_new/update/or_zip_county_prop.csv') 
-# or_zip_county <- read_csv(read_path)
 
@@
 
+```{r time stratified county smk for ts, echo=FALSE, include=FALSE, eval=FALSE}
+read_path_county_pm <- paste0("../data_new/county_data/or_county_pop_wt_pm.csv")
+county_smoke <- read_csv(read_path_county_pm) # 5508 rows
+
+# descriptives of the two smoke datasets
+summary(county_smoke)
+
+county_smoke_w_lag <- county_smoke %>% arrange(county, date) %>%
+  # group by zipcode
+  group_by(county) %>% 
+  # wrf
+  mutate(wrf_f_pm_lag1 = lag(wrf_pm, 1, order_by = county), 
+         wrf_f_pm_lag2 = lag(wrf_pm, 2, order_by = county),
+         wrf_f_pm_lag3 = lag(wrf_pm, 3, order_by = county),
+         wrf_f_pm_lag4 = lag(wrf_pm, 4, order_by = county),
+         wrf_f_pm_lag5 = lag(wrf_pm, 5, order_by = county),
+         wrf_f_pm_lag6 = lag(wrf_pm, 6, order_by = county),
+         wrf_f_pm_lag7 = lag(wrf_pm, 7, order_by = county),
+         # wrf no fire lag
+         wrf_nf_pm_lag1 = lag(wrf_nf_pm, 1, order_by = county),
+         wrf_nf_pm_lag2 = lag(wrf_nf_pm, 2, order_by = county),
+         wrf_nf_pm_lag3 = lag(wrf_nf_pm, 3, order_by = county),
+         wrf_nf_pm_lag4 = lag(wrf_nf_pm, 4, order_by = county),
+         wrf_nf_pm_lag5 = lag(wrf_nf_pm, 5, order_by = county),
+         wrf_nf_pm_lag6 = lag(wrf_nf_pm, 6, order_by = county),
+         wrf_nf_pm_lag7 = lag(wrf_nf_pm, 7, order_by = county),
+         # wrf_smk_pm
+         wrf_smk_pm_lag1 = lag(wrf_smk_pm, 1, order_by = county),
+         wrf_smk_pm_lag2 = lag(wrf_smk_pm, 2, order_by = county),
+         wrf_smk_pm_lag3 = lag(wrf_smk_pm, 3, order_by = county),
+         wrf_smk_pm_lag4 = lag(wrf_smk_pm, 4, order_by = county),
+         wrf_smk_pm_lag5 = lag(wrf_smk_pm, 5, order_by = county),
+         wrf_smk_pm_lag6 = lag(wrf_smk_pm, 6, order_by = county),
+         wrf_smk_pm_lag7 = lag(wrf_smk_pm, 7, order_by = county),
+         # geo weighted pm
+         geo_wt_pm_lag1 = lag(geo_wt_pm, 1, order_by = county),
+         geo_wt_pm_lag2 = lag(geo_wt_pm, 2, order_by = county),
+         geo_wt_pm_lag3 = lag(geo_wt_pm, 3, order_by = county),
+         geo_wt_pm_lag4 = lag(geo_wt_pm, 4, order_by = county),
+         geo_wt_pm_lag5 = lag(geo_wt_pm, 5, order_by = county),
+         geo_wt_pm_lag6 = lag(geo_wt_pm, 6, order_by = county),
+         geo_wt_pm_lag7 = lag(geo_wt_pm, 7, order_by = county),
+         # krig pm
+         krig_pm_lag1 = lag(krig_pm, 1, order_by = county),
+         krig_pm_lag2 = lag(krig_pm, 2, order_by = county),
+         krig_pm_lag3 = lag(krig_pm, 3, order_by = county),
+         krig_pm_lag4 = lag(krig_pm, 4, order_by = county),
+         krig_pm_lag5 = lag(krig_pm, 5, order_by = county), 
+         krig_pm_lag6 = lag(krig_pm, 6, order_by = county),
+         krig_pm_lag7 = lag(krig_pm, 7, order_by = county), 
+         # background pm
+         background_pm_lag1 = lag(background_pm, 1, order_by = county),
+         background_pm_lag2 = lag(background_pm, 2, order_by = county),
+         background_pm_lag3 = lag(background_pm, 3, order_by = county),
+         background_pm_lag4 = lag(background_pm, 4, order_by = county),
+         background_pm_lag5 = lag(background_pm, 5, order_by = county), 
+         background_pm_lag6 = lag(background_pm, 6, order_by = county),
+         background_pm_lag7 = lag(background_pm, 7, order_by = county), 
+         # geo_smk_pm 
+         geo_smk_pm_lag1 = lag(geo_smk_pm, 1, order_by = county),
+         geo_smk_pm_lag2 = lag(geo_smk_pm, 2, order_by = county),
+         geo_smk_pm_lag3 = lag(geo_smk_pm, 3, order_by = county),
+         geo_smk_pm_lag4 = lag(geo_smk_pm, 4, order_by = county),
+         geo_smk_pm_lag5 = lag(geo_smk_pm, 5, order_by = county),
+         geo_smk_pm_lag6 = lag(geo_smk_pm, 6, order_by = county),
+         geo_smk_pm_lag7 = lag(geo_smk_pm, 7, order_by = county),
+         # krig smk pm
+         krig_smk_pm_lag1 = lag(krig_smk_pm, 1, order_by = county),
+         krig_smk_pm_lag2 = lag(krig_smk_pm, 2, order_by = county),
+         krig_smk_pm_lag3 = lag(krig_smk_pm, 3, order_by = county),
+         krig_smk_pm_lag4 = lag(krig_smk_pm, 4, order_by = county),
+         krig_smk_pm_lag5 = lag(krig_smk_pm, 5, order_by = county),
+         krig_smk_pm_lag6 = lag(krig_smk_pm, 6, order_by = county),
+         krig_smk_pm_lag7 = lag(krig_smk_pm, 7, order_by = county),
+         # temp
+         wrf_temp_lag1 = lag(wrf_temp, 1, order_by = county),
+         wrf_temp_lag2 = lag(wrf_temp, 2, order_by = county),
+         wrf_temp_lag3 = lag(wrf_temp, 3, order_by = county),
+         wrf_temp_lag4 = lag(wrf_temp, 4, order_by = county),
+         wrf_temp_lag5 = lag(wrf_temp, 5, order_by = county), 
+         wrf_temp_lag6 = lag(wrf_temp, 6, order_by = county),
+         wrf_temp_lag7 = lag(wrf_temp, 7, order_by = county)) %>% 
+  # ungroup by zip
+  ungroup(county) %>% 
+  # attach a zip indicator for each smoke variable
+  setNames(paste(colnames(.), "county", sep="_"))  %>% 
+  # remove the '_zip' from the zipcode and date variable 
+  rename(county = county_county, date = date_county, fips = fips_county) %>%
+  select(-1)
+
+or_county_ts <- county_smoke_w_lag %>%
+  mutate(day = ifelse(weekdays(date) == "Monday" |
+                      weekdays(date) == "Tuesday" |
+                      weekdays(date) == "Wednesday"|
+                      weekdays(date) == "Thursday"|
+                      weekdays(date) == "Friday", "weekday",
+               ifelse(weekdays(date) == "Saturday" |
+                      weekdays(date) == "Sunday", "weekend", NA))) %>%
+  group_by(county, date) %>%
+  mutate(geo_smk_pm10 = geo_smk_pm_county/10)
+
+or_county_ts$day[which(or_county_ts$date=="2013-05-27")] <- "weekend"
+or_county_ts$day[which(or_county_ts$date=="2013-07-04")] <- "weekend"
+or_county_ts$day[which(or_county_ts$date=="2013-09-02")] <- "weekend"
+
+write_path5 <- paste0("../data_new/medication/or_county_time_series.csv")
+write_csv(or_county_ts, write_path5)
+
+```
+
 ## 1. smoke exposure time series and map of smoke days in Oregon
@@
 
-print(plot_ts)
+# print(plot_ts)
 
@@
 
-print(plot_county)
+# print(plot_county)
 # print(plot_county_line)
@@
 
-plot(plot_health_smk)
+# plot(plot_health_smk)
 # plot(plot_county_health_smk)
@@
 
-### map of smoke days
+### map of smoke days (4)
 
 ```{r smoke days and map, warning=FALSE, message=FALSE, echo=FALSE}
-or_ts <- or_ndc_county_ts %>% 
+read_path6 <- paste0("../data_new/medication/or_county_time_series.csv")
+or_county_ts <- read_csv(read_path6)
+
+or_ts <- or_county_ts %>% 
   mutate(smoke0 = ifelse(geo_smk_pm_county > 0, 1, 0),
@@
   select(county, date, geo_wt_pm_county) %>% 
-  mutate(high_pm_day = ifelse(geo_wt_pm_county > 16.33, 1, 0)) %>% 
+  mutate(high_pm_day = ifelse(geo_wt_pm_county > 10, 1, 0)) %>% 
   group_by(county) %>% 
@@
 
-# only 55 smoke wave days
-# xtabs(~smoke_wave,  or_ts)
-# if we compare to our geo smoke > 10, we'd add 362 extra days
-# xtabs(~smoke_wave + smoke10, or_ts)
-# if we used smoke >15 we'd add 164 extra days 
-# xtabs(~smoke_wave + smoke15, or_ts)
-
 # small multiples dataset
 or_smk <- or_ts %>% 
-  select(county, date, wrf_smk_pm_county, krig_smk_pm_county, geo_smk_pm_county) %>% 
+  select(county, date, geo_smk_pm_county) %>% 
   gather(key = smk_method, value = pm, -date, -county)
@@
   facet_wrap(~county) +
-  ggtitle("Oregon: 2013 Fire Season Smoke PM2.5") +
-  ylab("Smoke PM2.5 ug/m^3") +
-  xlab("Year 2013") +
-  theme_bw()
+  ggtitle(expression("Oregon: 2013 Fire Season Smoke PM"[2.5])) +
+   theme(legend.position="none") +
+   # theme(axis.text.x=element_blank()) +
+  theme(panel.background = element_rect(fill = 'white', colour = 'white')) +
+  theme(strip.background = element_rect(colour = "white", fill = "white"))+
+  ylab(expression(paste("GWR Smoke PM"[2.5], " Concentration, µg/m^3" ))) +
+  xlab("May 1st to Sep 30th in 2013") +
+  theme(axis.line.x = element_line(color="black"),
+        axis.line.y = element_line(color="black"))+
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
 
@@
   # summing up smoke days
-  summarise(smk_wave_n = sum(smoke_wave), smk_day0 = sum(smoke0), 
-            smk_day5 = sum(smoke5), smk_day10 = sum(smoke10),
-            smk_day15 = sum(smoke15)) %>% 
+  summarise(smk_day10 = sum(smoke10)) %>% 
   # county needs to be lowercase to join to maps dataframe
@@
   mutate(smk_cut_method = 
-    ifelse(smk_cut_method == "smk_day5","Smoke PM2.5 > 5 ug/m^3",
-    ifelse(smk_cut_method == "smk_day10", "Smoke PM2.5 > 10 ug/m^3",
-    ifelse(smk_cut_method == "smk_day15", "Smoke PM2.5 > 15 ug/m^3",
-    ifelse(smk_cut_method == "smk_wave_n", "Smoke Wave", NA)))),
+    ifelse(smk_cut_method == "smk_day10", "Smoke PM2.5 > 10 ug/m^3", NA),
     # preserve smoke method order for small multiple
     smk_cut_method = factor(smk_cut_method, 
-      levels = c("Smoke PM2.5 > 5 ug/m^3", "Smoke PM2.5 > 10 ug/m^3", 
-                 "Smoke PM2.5 > 15 ug/m^3", "Smoke Wave")))
+      levels = c("Smoke PM2.5 > 10 ug/m^3")))
 
 # map ----
+
+# or_county <- or_smk_df %>%
+#   group_by(subregion) %>%
+#   summarise(ave_long = mean(long), ave_lat = mean(lat))
+
+# or_county$subregion <- gsub("(^|[[:space:]])([[:alpha:]])", "\\1\\U\\2", or_county$subregion, perl=TRUE)
+
+getLabelPoint <- # Returns a county-named list of label points
+  function(county) {Polygon(county[c('long', 'lat')])@labpt}
+
+centroids <- by(or_smk_df, or_smk_df$subregion, getLabelPoint)     # Returns list
+centroids <- do.call("rbind.data.frame", centroids)  # Convert to Data Frame
+names(centroids) <- c('long', 'lat')                 # Appropriate Header
+
+rownames(centroids) <- gsub("(^|[[:space:]])([[:alpha:]])", "\\1\\U\\2", rownames(centroids), perl=TRUE)
 
@@
   scale_fill_gradient("Smoke-Impacted Days",
-                      low = "#0b486b", high = "#f56217") +
+                      low = "#3964C3", high = "#dc2430") +
   # add county path on top
@@
   facet_wrap(~smk_cut_method) +
+   theme(strip.background = element_blank(),
+       strip.text.x = element_blank()) +
   xlab("Longitude") +
   ylab("Latitude") +
-  ggtitle("Oregon: Counties affected by smoke") +
+  ggtitle("Oregon: Counties Affected by Smoke PM2.5 > 10µg/m^3") +
   theme(panel.background = element_rect(fill = 'white', colour = 'black'),
@@
     panel.grid.minor = element_blank(),
-    strip.background = element_rect(colour=NA, fill=NA))
-
+    strip.background = element_rect(colour=NA, fill=NA)) +
+  geom_text_repel(data = centroids, aes(long, lat, label = rownames(centroids), group=NULL), size = 2.5)
+
+smoke_map1 <- ggplot(or_smk_df, aes(x=long, y=lat, group=group)) +
+  # fill with number of smoke days
+  geom_polygon(aes(fill = smoke_days), alpha = 0.7) +
+  scale_fill_gradient("Smoke-Impacted Days",
+                      low = "#3964C3", high = "#dc2430") +
+  # add county path on top
+  geom_path() +
+  facet_wrap(~smk_cut_method) +
+   theme(strip.background = element_blank(),
+       strip.text.x = element_blank()) +
+  xlab("Longitude") +
+  ylab("Latitude") +
+  ggtitle("Oregon: Counties Affected by Smoke PM2.5 > 10µg/m^3") +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    strip.background = element_rect(colour=NA, fill=NA)) +
+  geom_text(data = centroids, aes(long, lat, label = rownames(centroids), group=NULL), size = 2)
 smoke_map
+
+smoke_map1
 
@@
 print(tab)
+
+```
+
+### Outcomes
+
+```{r category, echo = F, message = F, warning = F, results='asis', eval=FALSE}
+read_path <- paste0("../data_new/medication/oregon_beta_ndc_index_full.csv")
+inhaler_ndc <- read_csv(read_path)
 
@@
     #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
-    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+    ylab(expression(paste("Odds Ratio for 10Âµg/m"^3, " Increase in PM"[2.5]))) +
     #ylim(0, 2) +
@@
   ggsave("../plot_new/ndc_overall_gwr.pdf", plot = print_plot, 
-       width = 12, height = 8, units = "in")
-
-```
-
-### Time-Stratified by Age Categories
-
-Following table and figure look at some outcomes stratified by age category.
-
-For some outcomes, such like Age > 65 in WRF-Chem method, the CI is the only not significant one in the plot.
-
-The plot shows that overall, the risk of pharmacy fills of beta-2 agonist will increase significantly as the smoke concentrate increases.
-For age group of 15-65, the risk is the highest, for 10 ug/m^3 increase of pm2.5, the risk of filling the beta-2 agonist will increase about 9%.
-For age groups which are less than 15 or more than 65, the risk is similar, both of them will increase about 5.6%.
-
-```{r time stratified age strata cdc met, warning =F, echo = F, results='asis'} 
-# dataframe list
-
-method_list <- c('Geo-Weighted Smoke')
-
-# dataframe to loop through
-inhaler_ndc_df <- data.frame(inhaler_ndc)
-
-# outcome name
-outcome <- "Beta - 2 agonists"
-outcome_name <- "Beta - 2 agonists"
-
-# age category list
-age_cat_list <- c(0,1,2)
-
-# create an empty list to row bind dataframes together
-datalist1 <- list()
-
-# data wrangling ----
-# Producing conditional logit model estimates loop
-
-# extract covariates from dataframe
-covariates_df <- inhaler_ndc_df[, c(1:26, 74:84)]
-
-# extract pm values and divide by 10 and ordered
-#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
-pm_estimates_df <- inhaler_ndc_df[, c(91, 93)]/10  # create 10 unit increases
-
-# new loop for age categories
-for(k in 0:2){
-  # empty matrix (12 x 10 matrix)
-  point_estimates <- matrix(nrow = 1, ncol = 10, byrow = T)
-
-  colnames(point_estimates) <- c('outcome', 'age_cat', 'pm_method', 'n', 'n_events', 
-                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
-
-  # fill in the outcome namedataframe before method loop
-  point_estimates[, 1] <- outcome_name
-  # repeat the age category 4 times for each pm method
-  point_estimates[, 2] <- k
-
-  # dataframe for analysis creation
-  # bind columns back together 
-  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
-    # remove missing pm values
-    filter(!is.na(geo_smk_pm_zip)) %>% 
-    # limit to specific age category
-    filter(age_ind == k) %>% 
-    # the following code makes sure that the counterfactual values retained are 
-    # symetric in that number of obs before = number of obs after
-    mutate(obs_diff_admission = (fromdate - date)/7) 
-    # dataframe is already for the entire fire season, so I don't need to subset anymore
-
- j <- 38
-
-    # variable to model 
-    var_name <- colnames(df_analysis[j])
-
-    # set row number to fill
-    row_n <- j-37
-
-    # only run the model if the dataframe has observations
-    if(nrow(df_analysis) != 0){
-    # conditional logistic regression model
-    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
-
-    # populate matrix
-    point_estimates[row_n, 3] <- method_list[row_n]
-    point_estimates[row_n, 4] <- mod$n
-    point_estimates[row_n, 5] <- mod$nevent
-    # odds ratio
-    point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
-
-    # 95% lower bound
-    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
-    # 95% upper bound
-    point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
-                                        1.96*(summary(mod)$coefficient[1,3])), 3)
-    # standard error
-    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
-    # p val
-    point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
-
-    # create else statement that fills matrix with missing so I still have the row
-    # in the final dataframe
-    } else {point_estimates[row_n, 3] <- method_list[row_n]
-            point_estimates[row_n, 4] <- 0
-            point_estimates[row_n, 5] <- 0
-            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
-     # end methods loop
-
-  # save point estimates as a dataframe
-  point_est_df <- as_data_frame(point_estimates)
-  datalist1[[k+1]] <- point_est_df
-} # end age category loop
-
-# bind rows of age category estimates together
-combined_point_est_df <- bind_rows(datalist1)
-
-wrf_geo_age <- combined_point_est_df %>% 
-  # filter columns to just geo-smk
-  # filter(pm_method == "GWR Smoke" |
-    #      pm_method == "WRF-Chem Smoke") %>% 
-  mutate(age_cat2 = ifelse(age_cat == 0, "Less than 15", 
-                    ifelse(age_cat == 1, "15-65",
-                    ifelse(age_cat == 2, "Greater than 65", NA)))) %>% 
-  # subset columns I want to put in to the table
-  select(3, 11, 5:8) %>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
-         lower95 = ifelse(lower95 == 99, "--", lower95),
-         upper95 = ifelse(upper95 == 99, "--", upper95))
-
-tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
-         caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by age",
-         # row group by outcome
-         rgroup = "Beta - 2 agonists",
-         n.rgroup = c(rep(3, 1)), # 6 rows for each age cat for each outcome
-         # column headers
-         header = c("Est Method", "Age Group", "Events",
-                    "OR&dagger;", "Lower", "Upper"),
-         # column spanner
-         cgroup = c("", "95% CI"), 
-         n.cgroup = c(4, 2),
-         padding.rgroup = "&nbsp;&nbsp;",
-         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
-         align = "llcccc", # column alignment,
-         tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within May to September fire season."
-         ) # end table
-
-print(tab)
-
-combined_point_est_df_new <- combined_point_est_df %>%
-  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
-                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
-  select(-n_events) %>%
-  rename(n_events = n_events_new)
-
-# ggplot of odds ratios, facet wrapped by outcomes -----
-# convert variables from character to either numeric or factor
-# factor preserves the order of the variable
-wrf_geo_age_plot <- combined_point_est_df_new %>% 
-  # filter columns to just geo-smk
- #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
-  # do not plot results with less than 15 events, create missing vals
-  # in mutate
-  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
-                             99, odds_ratio),
-         lower95 = ifelse(as.numeric(n_events) < 15, 
-                             99, lower95),
-         upper95 = ifelse(as.numeric(n_events) < 15, 
-                             99, upper95),
-         age_cat2 = ifelse(age_cat == 0, "<15", 
-                    ifelse(age_cat == 1, "15-65",
-                    ifelse(age_cat == 2, ">65", NA)))) %>% 
-  # subset columns I want to put in to the table
-  select(1, 3, 11, 10, 5:7) %>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
-         lower95 = ifelse(lower95 == 99, NA, lower95),
-         upper95 = ifelse(upper95 == 99, NA, upper95))
-
-# change characters to numeric and factor  
-wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
-                              levels = unique(wrf_geo_age_plot$outcome))
-
-wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
-                                levels = unique(wrf_geo_age_plot$pm_method))
-
-wrf_geo_age_plot$age_cat2 <- factor(wrf_geo_age_plot$age_cat2,
-                                levels = unique(wrf_geo_age_plot$age_cat2))
-
-wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
-wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
-wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
-wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)
-
-## ggplot 
-  print_plot <- ggplot(wrf_geo_age_plot,
-    aes(x = age_cat2, y = odds_ratio, color = pm_method), na.rm = T) +
-    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
-    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
-                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
-    # custom color 
-    scale_color_manual(name = "Smoke-Estimation Method", 
-                       values = c("red",  "blue", "#32115C"),
-                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
-    facet_wrap(~outcome, nrow = 3, scales = "free") +
-    geom_hline(yintercept = 1, linetype=2) +
-    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
-    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
-    #ylim(0, 2) +
-    xlab('Age Category ') +
-    # plot theme
-    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    # strip element
-    strip.background = element_rect(colour=NA, fill=NA),
-    panel.border = element_rect(fill = NA, color = "black"),
-    # facet text size
-    strip.text = element_text(size = 10), 
-    # axis element
-    #axis.text.x = element_blank(),
-    #axis.title.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.title.y = element_text(angle = 90),
-    # legend elements
-    legend.position = "bottom")
-    #legend.text = element_text(size = 8))
-
-  print(print_plot)
-  # save figure
-  ggsave("../plot_new/ndc_age_fig_gwr.pdf", plot = print_plot, 
-       width = 12, height = 8, units = "in")
-
-```
-
-### Seasonal Time-Stratified by Sex
-
-The association with smoke PM~2.5~ and health outcomes stratified by sex is shown below.
-
-All the CI do not include 1, which means the pharmacy filling (beta 2 agonists) increases about 5% when smoke increase 10 units.
-So the risk for getting asthma will increase about 5% ~ 8% when the smoke PM2.5 increase 10 units.
-Also, the risk for female is much higher than that for male.
-
-These results show that there might be effect modification by sex, where females might be more likely to fill a beta-2 agonist on the same day it's really smoky for the GWR estimates.
-
-```{r sex strata cdc met adj, warning = F, echo = F, results='asis'} 
-method_list <- c('Geo-Weighted Smoke')
-
-# dataframe to loop through
-inhaler_ndc_df <- data.frame(inhaler_ndc)
-
-# outcome name
-outcome <- "Beta - 2 agonists"
-outcome_name <- "Beta - 2 agonists"
-
-# create an empty list to row bind dataframes together
-datalist1 <- list()
-
-# sex category list
-sex_strata_list <- c(0,1)
-
-# data wrangling ----
-# Producing conditional logit model estimates loop
-
-# extract covariates from dataframe
-covariates_df <- inhaler_ndc_df[, c(1:26, 74:84)]
-
-# extract pm values and divide by 10 and ordered
-#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
-pm_estimates_df <- inhaler_ndc_df[, c(91, 93)]/10  # create 10 unit increases
-
-# new loop for sex categories
-for(k in 0:1){
-
-  # empty matrix (12 x 10 matrix)
-  point_estimates <- matrix(nrow = 1, ncol = 10, byrow = T)
-
-  colnames(point_estimates) <- c('outcome', 'sex', 'pm_method', 'n', 'n_events', 
-                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
-
-  # fill in the outcome namedataframe before method loop
-  point_estimates[, 1] <- outcome_name
-  # repeat the sex category 4 times for each pm method
-  point_estimates[, 2] <- k
-
-  # dataframe for analysis creation
-  # bind columns back together 
-  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
-    # remove missing pm values
-    filter(!is.na(geo_smk_pm_zip)) %>% 
-    # limit to specific sex category
-    filter(sex_ind == k) %>% 
-    # the following code makes sure that the counterfactual values retained are 
-    # symetric in that number of obs before = number of obs after
-    mutate(obs_diff_admission = (fromdate - date)/7) 
-    # dataframe is already for the entire fire season, so I don't need to subset anymore
-
-  j <- 38
-    # variable to model 
-    var_name <- colnames(df_analysis[j])
-
-    # set row number to fill
-    row_n <- j-37
-
-    # only run the model if the dataframe has observations
-    if(nrow(df_analysis) != 0){
-    # conditional logistic regression model
-    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
-
-    # populate matrix
-    point_estimates[row_n, 3] <- method_list[row_n]
-    point_estimates[row_n, 4] <- mod$n
-    point_estimates[row_n, 5] <- mod$nevent
-    # odds ratio
-    point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
-
-    # 95% lower bound
-    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
-                                      1.96*(summary(mod)$coefficient[1,3])), 3)
-    # 95% upper bound
-    point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
-                                      1.96*(summary(mod)$coefficient[1,3])), 3)
-    # standard error
-    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
-    # p val
-    point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
-
-    # create else statement that fills matrix with missing so I still have the row
-    # in the final dataframe
-    } else {point_estimates[row_n, 3] <- method_list[row_n]
-            point_estimates[row_n, 4] <- 0
-            point_estimates[row_n, 5] <- 0
-            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
-    # end methods loop
-
-  # save point estimates as a dataframe
-  point_est_df <- as_data_frame(point_estimates)
-
-  # combine previous values in dataframe that has all outcome/methods comparisons
-  datalist1[[k+1]] <- point_est_df
-} # end sex category loop
-
-# bind rows of sex category estimates together
-combined_point_est_df <- bind_rows(datalist1)
-
-wrf_geo_sex <- combined_point_est_df %>% 
-  # filter columns to just geo-smk
-  # filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
-  mutate(sex_cat = ifelse(sex == 0, "Male", 
-               ifelse(sex == 1, "Female", NA))) %>% 
-  # subset columns I want to put in to the table
-  select(3, 11, 5:8)%>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
-         lower95 = ifelse(lower95 == 99, "--", lower95),
-         upper95 = ifelse(upper95 == 99, "--", upper95))
-
-  tab <- htmlTable(txtRound(wrf_geo_sex, digits = 3, 1:3), 
-           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and pharmacy outcomes stratified by sex",
-           # row group by outcome
-           rgroup = "Beta - 2 agonists",
-           n.rgroup = c(rep(2, 1)), # 2 rows for each sex strata for each outcome
-           # column headers
-           header = c("Est Method", "Sex Strata", "Events",
-                      "OR&dagger;", "Lower", "Upper"),
-           # column spanner
-           cgroup = c("", "95% CI"), 
-           n.cgroup = c(4, 2),
-           padding.rgroup = "&nbsp;&nbsp;",
-           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
-           align = "llcccc", # column alignment,
-           tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within July to October fire season."
-            ) # end table
-
-  print(tab)
-
-# ggplot of odds ratios, facet wrapped by outcomes -----
-# convert variables from character to either numeric or factor
-# factor preserves the order of the variable
-wrf_geo_sex_plot <- combined_point_est_df %>% 
-  # filter columns to just geo-smk
- #  filter(pm_method == "GWR Smoke" | 
-   #       pm_method == "WRF-Chem Smoke") %>% 
-  mutate(sex_cat = ifelse(sex == 0, "Male", 
-               ifelse(sex == 1, "Female", NA))) %>%
-  # subset columns I want to put in to the table
-  select(1, 3, 11, 5:8) %>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
-         lower95 = ifelse(lower95 == 99, NA, lower95),
-         upper95 = ifelse(upper95 == 99, NA, upper95))
-
-# change characters to numeric and factor  
-wrf_geo_sex_plot$outcome <- factor(wrf_geo_sex_plot$outcome,
-                              levels = unique(wrf_geo_sex_plot$outcome))
-
-wrf_geo_sex_plot$pm_method <- factor(wrf_geo_sex_plot$pm_method,
-                                levels = unique(wrf_geo_sex_plot$pm_method))
-
-wrf_geo_sex_plot$sex_cat <- factor(wrf_geo_sex_plot$sex_cat,
-                                levels = unique(wrf_geo_sex_plot$sex_cat))
-
-wrf_geo_sex_plot$n_events <- as.numeric(wrf_geo_sex_plot$n_events)
-wrf_geo_sex_plot$odds_ratio <- as.numeric(wrf_geo_sex_plot$odds_ratio)
-wrf_geo_sex_plot$lower95 <- as.numeric(wrf_geo_sex_plot$lower95)
-wrf_geo_sex_plot$upper95 <- as.numeric(wrf_geo_sex_plot$upper95)
-
-## ggplot
-  print_plot <- ggplot(wrf_geo_sex_plot,
-    aes(x = sex_cat, y = odds_ratio, color = pm_method), na.rm = T) +
-    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
-    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
-                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
-    # custom color 
-    scale_color_manual(name = "Smoke-Estimation Method", 
-                       values = c("red", "blue", "#32115C"),
-                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
-    facet_wrap(~outcome, nrow = 3, scales = "free") +
-    geom_hline(yintercept = 1, linetype=2) +
-    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
-    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
-    #ylim(0, 2) +
-    xlab('Sex') +
-    # plot theme
-    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    # strip element
-    strip.background = element_rect(colour=NA, fill=NA),
-    panel.border = element_rect(fill = NA, color = "black"),
-    # facet text size
-    strip.text = element_text(size = 10),
-    # axis element
-    #axis.text.x = element_blank(),
-    #axis.title.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.title.y = element_text(angle = 90),
-    # legend elements
-    legend.position = "bottom")
-    #legend.text = element_text(size = 8))
-
-  print(print_plot)
-  # save figure
-  ggsave("../plot_new/ndc_sex_fig_gwr.pdf", plot = print_plot, 
        width = 12, height = 8, units = "in")
@@
 ## Supplement
+
+### Age and Sex category (8)
+### Time-Stratified by Age Categories
+
+Following table and figure look at some outcomes stratified by age category.
+
+For some outcomes, such like Age > 65 in WRF-Chem method, the CI is the only not significant one in the plot.
+
+The plot shows that overall, the risk of pharmacy fills of beta-2 agonist will increase significantly as the smoke concentrate increases.
+For age group of 15-65, the risk is the highest, for 10 ug/m^3 increase of pm2.5, the risk of filling the beta-2 agonist will increase about 9%.
+For age groups which are less than 15 or more than 65, the risk is similar, both of them will increase about 5.6%.
+
+```{r time stratified age strata cdc met, warning =F, echo = F, results='asis'} 
+# dataframe list
+
+method_list <- c('Geo-Weighted Smoke')
+
+# dataframe to loop through
+inhaler_ndc_df <- data.frame(inhaler_ndc)
+
+# outcome name
+outcome <- "Beta - 2 agonists"
+outcome_name <- "Beta - 2 agonists"
+
+# age category list
+age_cat_list <- c(0,1,2)
+
+# create an empty list to row bind dataframes together
+datalist1 <- list()
+
+# data wrangling ----
+# Producing conditional logit model estimates loop
+
+# extract covariates from dataframe
+covariates_df <- inhaler_ndc_df[, c(1:26, 74:84)]
+
+# extract pm values and divide by 10 and ordered
+#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+pm_estimates_df <- inhaler_ndc_df[, c(91, 93)]/10  # create 10 unit increases
+
+# new loop for age categories
+for(k in 0:2){
+  # empty matrix (12 x 10 matrix)
+  point_estimates <- matrix(nrow = 1, ncol = 10, byrow = T)
+
+  colnames(point_estimates) <- c('outcome', 'age_cat', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+  # fill in the outcome namedataframe before method loop
+  point_estimates[, 1] <- outcome_name
+  # repeat the age category 4 times for each pm method
+  point_estimates[, 2] <- k
+
+  # dataframe for analysis creation
+  # bind columns back together 
+  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+    # remove missing pm values
+    filter(!is.na(geo_smk_pm_zip)) %>% 
+    # limit to specific age category
+    filter(age_ind == k) %>% 
+    # the following code makes sure that the counterfactual values retained are 
+    # symetric in that number of obs before = number of obs after
+    mutate(obs_diff_admission = (fromdate - date)/7) 
+    # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+ j <- 38
+
+    # variable to model 
+    var_name <- colnames(df_analysis[j])
+
+    # set row number to fill
+    row_n <- j-37
+
+    # only run the model if the dataframe has observations
+    if(nrow(df_analysis) != 0){
+    # conditional logistic regression model
+    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+    # populate matrix
+    point_estimates[row_n, 3] <- method_list[row_n]
+    point_estimates[row_n, 4] <- mod$n
+    point_estimates[row_n, 5] <- mod$nevent
+    # odds ratio
+    point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+    # 95% lower bound
+    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
+    # 95% upper bound
+    point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                        1.96*(summary(mod)$coefficient[1,3])), 3)
+    # standard error
+    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
+    # p val
+    point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
+
+    # create else statement that fills matrix with missing so I still have the row
+    # in the final dataframe
+    } else {point_estimates[row_n, 3] <- method_list[row_n]
+            point_estimates[row_n, 4] <- 0
+            point_estimates[row_n, 5] <- 0
+            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
+     # end methods loop
+
+  # save point estimates as a dataframe
+  point_est_df <- as_data_frame(point_estimates)
+  datalist1[[k+1]] <- point_est_df
+} # end age category loop
+
+# bind rows of age category estimates together
+combined_point_est_df <- bind_rows(datalist1)
+
+wrf_geo_age <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" |
+    #      pm_method == "WRF-Chem Smoke") %>% 
+  mutate(age_cat2 = ifelse(age_cat == 0, "Less than 15", 
+                    ifelse(age_cat == 1, "15-65",
+                    ifelse(age_cat == 2, "Greater than 65", NA)))) %>% 
+  # subset columns I want to put in to the table
+  select(3, 11, 5:8) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
+         caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by age",
+         # row group by outcome
+         rgroup = "Beta - 2 agonists",
+         n.rgroup = c(rep(3, 1)), # 6 rows for each age cat for each outcome
+         # column headers
+         header = c("Est Method", "Age Group", "Events",
+                    "OR&dagger;", "Lower", "Upper"),
+         # column spanner
+         cgroup = c("", "95% CI"), 
+         n.cgroup = c(4, 2),
+         padding.rgroup = "&nbsp;&nbsp;",
+         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+         align = "llcccc", # column alignment,
+         tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within May to September fire season."
+         ) # end table
+
+print(tab)
+
+combined_point_est_df_new <- combined_point_est_df %>%
+  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
+                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
+  select(-n_events) %>%
+  rename(n_events = n_events_new)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_age_plot <- combined_point_est_df_new %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  # do not plot results with less than 15 events, create missing vals
+  # in mutate
+  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
+                             99, odds_ratio),
+         lower95 = ifelse(as.numeric(n_events) < 15, 
+                             99, lower95),
+         upper95 = ifelse(as.numeric(n_events) < 15, 
+                             99, upper95),
+         age_cat2 = ifelse(age_cat == 0, "<15", 
+                    ifelse(age_cat == 1, "15-65",
+                    ifelse(age_cat == 2, ">65", NA)))) %>% 
+  # subset columns I want to put in to the table
+  select(1, 3, 11, 10, 5:7) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
+                              levels = unique(wrf_geo_age_plot$outcome))
+
+wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
+                                levels = unique(wrf_geo_age_plot$pm_method))
+
+wrf_geo_age_plot$age_cat2 <- factor(wrf_geo_age_plot$age_cat2,
+                                levels = unique(wrf_geo_age_plot$age_cat2))
+
+wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
+wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
+wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
+wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)
+
+## ggplot 
+  print_plot <- ggplot(wrf_geo_age_plot,
+    aes(x = age_cat2, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red",  "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10Âµg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Age Category ') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10), 
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
+  ggsave("../plot_new/ndc_age_fig_gwr.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+### Seasonal Time-Stratified by Sex
+
+The association with smoke PM~2.5~ and health outcomes stratified by sex is shown below.
+
+All the CI do not include 1, which means the pharmacy filling (beta 2 agonists) increases about 5% when smoke increase 10 units.
+So the risk for getting asthma will increase about 5% ~ 8% when the smoke PM2.5 increase 10 units.
+Also, the risk for female is much higher than that for male.
+
+These results show that there might be effect modification by sex, where females might be more likely to fill a beta-2 agonist on the same day it's really smoky for the GWR estimates.
+
+```{r sex strata cdc met adj, warning = F, echo = F, results='asis'} 
+method_list <- c('Geo-Weighted Smoke')
+
+# dataframe to loop through
+inhaler_ndc_df <- data.frame(inhaler_ndc)
+
+# outcome name
+outcome <- "Beta - 2 agonists"
+outcome_name <- "Beta - 2 agonists"
+
+# create an empty list to row bind dataframes together
+datalist1 <- list()
+
+# sex category list
+sex_strata_list <- c(0,1)
+
+# data wrangling ----
+# Producing conditional logit model estimates loop
+
+# extract covariates from dataframe
+covariates_df <- inhaler_ndc_df[, c(1:26, 74:84)]
+
+# extract pm values and divide by 10 and ordered
+#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+pm_estimates_df <- inhaler_ndc_df[, c(91, 93)]/10  # create 10 unit increases
+
+# new loop for sex categories
+for(k in 0:1){
+
+  # empty matrix (12 x 10 matrix)
+  point_estimates <- matrix(nrow = 1, ncol = 10, byrow = T)
+
+  colnames(point_estimates) <- c('outcome', 'sex', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+  # fill in the outcome namedataframe before method loop
+  point_estimates[, 1] <- outcome_name
+  # repeat the sex category 4 times for each pm method
+  point_estimates[, 2] <- k
+
+  # dataframe for analysis creation
+  # bind columns back together 
+  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+    # remove missing pm values
+    filter(!is.na(geo_smk_pm_zip)) %>% 
+    # limit to specific sex category
+    filter(sex_ind == k) %>% 
+    # the following code makes sure that the counterfactual values retained are 
+    # symetric in that number of obs before = number of obs after
+    mutate(obs_diff_admission = (fromdate - date)/7) 
+    # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+  j <- 38
+    # variable to model 
+    var_name <- colnames(df_analysis[j])
+
+    # set row number to fill
+    row_n <- j-37
+
+    # only run the model if the dataframe has observations
+    if(nrow(df_analysis) != 0){
+    # conditional logistic regression model
+    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+    # populate matrix
+    point_estimates[row_n, 3] <- method_list[row_n]
+    point_estimates[row_n, 4] <- mod$n
+    point_estimates[row_n, 5] <- mod$nevent
+    # odds ratio
+    point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+    # 95% lower bound
+    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
+                                      1.96*(summary(mod)$coefficient[1,3])), 3)
+    # 95% upper bound
+    point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                      1.96*(summary(mod)$coefficient[1,3])), 3)
+    # standard error
+    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
+    # p val
+    point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
+
+    # create else statement that fills matrix with missing so I still have the row
+    # in the final dataframe
+    } else {point_estimates[row_n, 3] <- method_list[row_n]
+            point_estimates[row_n, 4] <- 0
+            point_estimates[row_n, 5] <- 0
+            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
+    # end methods loop
+
+  # save point estimates as a dataframe
+  point_est_df <- as_data_frame(point_estimates)
+
+  # combine previous values in dataframe that has all outcome/methods comparisons
+  datalist1[[k+1]] <- point_est_df
+} # end sex category loop
+
+# bind rows of sex category estimates together
+combined_point_est_df <- bind_rows(datalist1)
+
+wrf_geo_sex <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  mutate(sex_cat = ifelse(sex == 0, "Male", 
+               ifelse(sex == 1, "Female", NA))) %>% 
+  # subset columns I want to put in to the table
+  select(3, 11, 5:8)%>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+  tab <- htmlTable(txtRound(wrf_geo_sex, digits = 3, 1:3), 
+           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and pharmacy outcomes stratified by sex",
+           # row group by outcome
+           rgroup = "Beta - 2 agonists",
+           n.rgroup = c(rep(2, 1)), # 2 rows for each sex strata for each outcome
+           # column headers
+           header = c("Est Method", "Sex Strata", "Events",
+                      "OR&dagger;", "Lower", "Upper"),
+           # column spanner
+           cgroup = c("", "95% CI"), 
+           n.cgroup = c(4, 2),
+           padding.rgroup = "&nbsp;&nbsp;",
+           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+           align = "llcccc", # column alignment,
+           tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within July to October fire season."
+            ) # end table
+
+  print(tab)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_sex_plot <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | 
+   #       pm_method == "WRF-Chem Smoke") %>% 
+  mutate(sex_cat = ifelse(sex == 0, "Male", 
+               ifelse(sex == 1, "Female", NA))) %>%
+  # subset columns I want to put in to the table
+  select(1, 3, 11, 5:8) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_sex_plot$outcome <- factor(wrf_geo_sex_plot$outcome,
+                              levels = unique(wrf_geo_sex_plot$outcome))
+
+wrf_geo_sex_plot$pm_method <- factor(wrf_geo_sex_plot$pm_method,
+                                levels = unique(wrf_geo_sex_plot$pm_method))
+
+wrf_geo_sex_plot$sex_cat <- factor(wrf_geo_sex_plot$sex_cat,
+                                levels = unique(wrf_geo_sex_plot$sex_cat))
+
+wrf_geo_sex_plot$n_events <- as.numeric(wrf_geo_sex_plot$n_events)
+wrf_geo_sex_plot$odds_ratio <- as.numeric(wrf_geo_sex_plot$odds_ratio)
+wrf_geo_sex_plot$lower95 <- as.numeric(wrf_geo_sex_plot$lower95)
+wrf_geo_sex_plot$upper95 <- as.numeric(wrf_geo_sex_plot$upper95)
+
+## ggplot
+  print_plot <- ggplot(wrf_geo_sex_plot,
+    aes(x = sex_cat, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red", "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10Âµg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Sex') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10),
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
+  ggsave("../plot_new/ndc_sex_fig_gwr.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
 ### 1. For Overall Plot (Unconstrained lag model)
 
-```{r overall lags, warning =F, echo = F, results='asis'} 
+```{r overall lags, warning =F, echo = F, results='asis', eval=FALSE} 
 # dataframe list
@@
 
-```{r unconstrained ts lag, message=FALSE, echo=FALSE, warning=FALSE}
+```{r unconstrained ts lag, message=FALSE, echo=FALSE, warning=FALSE, eval=FALSE}
 or_ndc_county_ts_df <- or_ndc_county_ts %>%
@@
    axis.text.x = element_text(angle = 90, vjust = .75)) +
-  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+  ylab(expression(paste("Odds Ratio for 10Âµg/m"^3, " Increase in PM"[2.5]))) +
   xlab("Outcome") +

### 2017-08-04 r_scripts/analysis/oregon_medication.Rmd
Update oregon_medication.Rmd

@@
 
+centroids[4,] <- centroids[4,] + c(0, 0.02) # Clatsop
+centroids[5,] <- centroids[5,] + c(0.1, 0) # Columbia
+centroids[29,] <- centroids[29,] + c(-0.1, -0.03) # Tillamook
+centroids[34,] <- centroids[34,] + c(-0.12, -0.02) # Washington
+centroids[26,] <- centroids[26,] + c(0, 0.03) # Multnomah
+centroids[11,] <- centroids[11,] + c(-0.05, -0.12) # Gilliam
+centroids[24,] <- centroids[24,] + c(-0.2, 0) # Marion
+centroids[33,] <- centroids[33,] + c(0, -0.1) # Wasco
+centroids[28,] <- centroids[28,] + c(0, 0.1) # Sherman
+centroids[36,] <- centroids[36,] + c(-0.05, -0.08) # Yamhill
+centroids[8,] <- centroids[8,] + c(0.08, -0.2) # Yamhill
+centroids[17,] <- centroids[17,] + c(-0.05, 0.08) # Yamhill
+centroids[29,] <- centroids[29,] + c(-0.01, -0.1) # Sherman
+
 # smoke days where GWR smoke > 15
@@
   geom_text(data = centroids, aes(long, lat, label = rownames(centroids), group=NULL), size = 2)
-smoke_map
 
@@
 df_analysis <- inhaler_ndc_df %>%
-  filter(!is.na(wrf_smk_pm_zip)) %>% 
+  filter(!is.na(geo_smk_pm_zip)) %>% 
   # only look at outcomes
@@
 
-### Outcomes
-
-```{r category, echo = F, message = F, warning = F, results='asis', eval=FALSE}
-read_path <- paste0("../data_new/medication/oregon_beta_ndc_index_full.csv")
-inhaler_ndc <- read_csv(read_path)
+### Outcomes of four kinds
+
+```{r category, echo = F, message = F, warning = F, results='asis'}
+read_path <- paste0("../data_new/medication/oregon_beta_ndc_index.csv")
+inhaler_index <- read_csv(read_path)
+
+# dataframe to loop through
+inhaler_index_df <- data.frame(inhaler_index) %>%
+  select(personkey, index)
+
+cate <- inhaler_index_df %>%
+  filter(personkey %in% inhaler_ndc$personkey) %>%
+  unique()
+
+inhaler_ndc_index_df <- inhaler_ndc_df %>%
+  left_join(cate, by="personkey")
+
+asthma_df <- inhaler_ndc_index_df %>%
+  filter(index=="asthma")
+
+copd_df <- inhaler_ndc_index_df %>%
+  filter(index=="copd")
+
+both_df <- inhaler_ndc_index_df %>%
+  filter(index=="both")
+
+other_df <- inhaler_ndc_index_df %>%
+  filter(index=="other")
+
+df_list <- list(asthma_df, copd_df, both_df, other_df)
+
+outcome_list <- c('Asthma', 'COPD', 'Both', 'Other')
+
+method_list <- c('Geo-Weighted Smoke')
+
+# create an empty list to row bind dataframes together
+datalist <- list()
+
+# set list and do cross tabs to find values
+# Producing conditional logit model estimates loop 
+for(i in 1:length(df_list)){
+
+# dataframe to loop through
+  df_to_loop <- data.frame(df_list[i])
+  # indication of column
+  outcome <- colnames(df_to_loop[76])
+  # outcome name
+  outcome_name <- outcome_list[i]
+
+  # dataframe for analysis creation
+  # bind columns back together 
+  df_analysis <- df_to_loop %>%
+    filter(!is.na(wrf_smk_pm_zip)) %>% 
+    # only look at outcomes
+    filter(outcome == 1) %>%
+    # add another row that makes sure there is a person <15 in the dataframe
+    # tricking xtabs to produce a 0 cell for the outcome for age <15
+    add_row(outcome = 0, age_ind = 0)
+
+  # cross tabs
+  outcome_n <- xtabs(~ outcome, df_analysis)
+  cross_tab_age <- xtabs(~ outcome + age_ind, df_analysis)
+  cross_tab_sex <- xtabs(~ outcome + sex_ind, df_analysis)
+  # empty matrix
+  point_estimates <- matrix(nrow = 1, ncol = 7, byrow = T)
+
+  colnames(point_estimates) <- c("outcome", "n", "age_15", "age_15_65", 
+                                 "age_65", "female", "male")
+
+  # fill in the outcome name for the dataframe before the loop
+  point_estimates[, 1] <- outcome_name
+  # fill n
+  point_estimates[, 2] <- outcome_n[2] # second element of the 1 dimension vector
+  # age <15
+  point_estimates[, 3] <- cross_tab_age[2, 1]
+  # age 15 to 65
+  point_estimates[, 4] <- cross_tab_age[2, 2]
+  # age >65
+  point_estimates[, 5] <- cross_tab_age[2, 3]
+  # male == 0
+  point_estimates[, 7] <- cross_tab_sex[1, 1]
+  # female == 1
+  point_estimates[, 6] <- cross_tab_sex[1, 2]
+
+  # save point estimates as a dataframe
+  point_est_df <- as_data_frame(point_estimates)
+
+  # combine previous values in dataframe that has all outcome/methods comparisons
+  datalist[[i]] <- point_est_df
+
+} # end of loop
+
+# combine each outcome dataframe itteration in to a big dataset
+combined_point_est_df <- bind_rows(datalist) %>% 
+  # find proportions/percents for each strata in a row
+  mutate(age_15_pr = as.character(round((as.numeric(age_15)/as.numeric(n))*100,1)),
+         age_15_65_pr = as.character(round((as.numeric(age_15_65)/as.numeric(n))*100,1)),
+         age_65_pr = as.character(round((as.numeric(age_65)/as.numeric(n))*100,1)),
+         female_pr = as.character(round((as.numeric(female)/as.numeric(n))*100,1)),
+         male_pr = as.character(round((as.numeric(male)/as.numeric(n))*100,1))) %>% 
+  select(outcome, n, age_15, age_15_pr, age_15_65, age_15_65_pr, age_65,
+         age_65_pr, female, female_pr, male, male_pr)
+
+#str(combined_point_est_df)
+
+tab <- htmlTable(txtRound(combined_point_est_df, digits = 1), 
+           caption = "Number of cases for each outcome observed from May 1st to September 30st, 2013",
+           # column headers
+           header = c("Outcome", "Cases n", "Less than 15", "(%)", "15 to 65", "(%)", "Greater than 65",
+                      "(%)", "Female", "(%)", "Male", "(%)"),
+           # column spanner
+           cgroup = c("","Age Category", "Sex"), 
+           n.cgroup = c(2, 6, 4),
+           padding.rgroup = "&nbsp;&nbsp;",
+           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+           align = "llccccc" # column alignment,
+            ) # end table
+
+print(tab)
 
@@
 
+The plot shows that four different categories
+
+```{r category plot, warning =F, echo = F, results='asis'}
+df_list <- list(asthma_df, copd_df, both_df, other_df)
+
+outcome_list <- c('Asthma', 'COPD', 'Both', 'Other')
+
+method_list <- c('Geo-Weighted Smoke')
+
+# create an empty list to row bind dataframes together
+datalist <- list()
+
+# data wrangling ----
+# Producing conditional logit model estimates loop 
+for(i in 1:length(df_list)){
+
+  # dataframe to loop through
+  df_to_loop <- data.frame(df_list[i])
+  # indication of column
+  outcome <- colnames(df_to_loop[75])
+  # outcome name
+  outcome_name <- outcome_list[i]
+
+  covariates_df <- df_to_loop[, c(1:26, 74:84)]
+
+  # extract pm values and divide by 10 and ordered
+  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+  pm_estimates_df <- df_to_loop[, c(91, 93)]/10  # create 10 unit increases
+
+  # dataframe for analysis creation
+  # bind columns back together 
+  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+    # remove missing pm values
+    filter(!is.na(geo_smk_pm_zip)) %>%
+
+    # join in cdc met data
+    # full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
+    # the following code makes sure that the counterfactual values retained are 
+    # symetric in that number of obs before = number of obs after
+    mutate(obs_diff_admission = (fromdate - date)/7) 
+  # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+  # empty df for table
+  table_df <- data.frame()
+
+  # empty matrix
+  point_estimates <- matrix(nrow = 1, ncol = 9, byrow = T)
+
+  colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 'odds_ratio', 
+                                 'lower95', 'upper95', 'se', 'p_val')
+
+  # fill in the outcome name for the dataframe before the loop
+  point_estimates[, 1] <- outcome_name
+
+# set row number to fill
+row_n <- 1
+
+# only run the model if the dataframe has observations
+if(nrow(df_analysis) != 0){
+  # conditional logistic regression model
+  mod <- clogit(outcome ~ df_analysis[[38]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+      # populate matrix
+    point_estimates[row_n, 2] <- method_list[row_n]
+    point_estimates[row_n, 3] <- mod$n
+    point_estimates[row_n, 4] <- mod$nevent
+    # odds ratio
+    point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+    # 95% lower bound
+    point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -
+                                             1.96*(summary(mod)$coefficient[1,3])), 3)
+    # 95% upper bound
+    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                             1.96*(summary(mod)$coefficient[1,3])), 3)
+    # standard error
+    point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
+    # p val
+    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)
+
+    # save point estimates as a dataframe
+    point_est_df <- as_data_frame(point_estimates)      
+  # create else statement that fills matrix with missing so I still have the row
+  # in the final dataframe
+  } else {point_estimates[row_n, 2] <- method_list
+          point_estimates[row_n, 3] <- 0
+          point_estimates[row_n, 4] <- 0
+          point_estimates[row_n, c(5:9)] <- 99 } # end 'if else' statement
+
+  # combine previous values in dataframe that has all outcome/methods comparisons
+  datalist[[i]] <- point_est_df
+
+} # end of loop
+
+# combine each outcome dataframe itteration in to a big dataset
+combined_point_est_df <- bind_rows(datalist)
+
+# subset columns I want to put in to the table
+table_df <- combined_point_est_df %>% select(2, 3:7)
+
+tab <- htmlTable(txtRound(table_df, digits = 3, 1:3), 
+                 caption = "Association between a 10 ug/m^3 in PM2.5 and Health Outcomes",
+                 # row group by outcome
+                 rgroup = outcome_list,
+                 n.rgroup = c(rep(1, 4)), # 4 rows for each method for each outcome
+                 # column headers
+                 header = c("Method", "Obs.", "Events",
+                            "OR&dagger;", "Lower", "Upper"),
+                 # column spanner
+                 cgroup = c("", "95% CI"), 
+                 n.cgroup = c(4, 2),
+                 padding.rgroup = "&nbsp;&nbsp;",
+                 css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+                 align = "llccccc", # column alignment,
+                 tfoot="&dagger; Adjusted for CDC temperature, relatively humidity, wind speed, and precipitation; accounting for subject.
+Time-stratified: referent periods matched to events on same day of week within July to October fire season."
+) # end table
+
+print(tab)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+combined_point_est_df$outcome <- factor(combined_point_est_df$outcome, 
+                                        levels = unique(combined_point_est_df$outcome))
+
+combined_point_est_df$pm_method <- factor(combined_point_est_df$pm_method, 
+                                          levels = unique(combined_point_est_df$pm_method))
+
+combined_point_est_df$n <- as.numeric(combined_point_est_df$n)
+combined_point_est_df$n_events <- as.numeric(combined_point_est_df$n_events)
+combined_point_est_df$odds_ratio <- as.numeric(combined_point_est_df$odds_ratio)
+combined_point_est_df$lower95 <- as.numeric(combined_point_est_df$lower95)
+combined_point_est_df$upper95 <- as.numeric(combined_point_est_df$upper95)
+combined_point_est_df$se <- as.numeric(combined_point_est_df$se)
+combined_point_est_df$p_val <- as.numeric(combined_point_est_df$p_val)
+
+# subset out global method
+combined_point_est_df <- combined_point_est_df
+
+## ggplot
+print_plot <- ggplot(combined_point_est_df, 
+                     aes(x = pm_method, y = odds_ratio, colour = pm_method)) +
+  # custom color 
+  scale_color_manual(name = "Smoke-Estimation Method", 
+                     values = c("red", "blue", "#32115C"),
+                     guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+  geom_point() + #geom_text(vjust = 0, nudge_x = 0.3) +
+  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
+  facet_wrap(~outcome, nrow = 3) +
+  geom_hline(yintercept = 1, linetype=2) +
+  #    ggtitle('Association Between PM2.5 \n from Wildfire Smoke on Hospitalizations') +
+  ylab(expression(paste("Odds Ratio for 10Âµg/m"^3, " Increase in PM"[2.5]))) +
+  #xlab('Smoke Estimation Method') +
+  # plot theme
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        # strip element
+        strip.background = element_rect(colour=NA, fill=NA),
+        panel.border = element_rect(fill = NA, color = "black"),
+        # facet text size
+        #strip.text = element_text(size = 8),
+        # axis element
+        axis.text.x = element_blank(),
+        axis.title.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        axis.title.y = element_text(angle = 90),
+        # legend elements
+        legend.position = "bottom")
+#legend.text = element_text(size = 8))
+
+print(print_plot)
+# save figure
+# ggsave("../plot_new/fig_both_er_ur.pdf", plot = print_plot, 
+#        width = 12, height = 8, units = "in")
+
+```
+
 ## 3. Distributed Lag Model
@@
 
-### Find the smallest of AIC with df from 3 to 6
-
-Choose df = 5 because it has the least AIC.
-
-```{r, message=FALSE, echo=FALSE, warning=FALSE}
+### Distributed lag 0-7 for casecrossover
+
+When choosing the df of 5, the plot shows that the same day of fill with smoke exposure has the highest significant association.
+But other lags shows that pharmacy fills with lagged days do not has the significant association with smoke.
+And the association tends to decrease as lag goes.
+
+```{r input, echo = FALSE, message=FALSE, warning=FALSE}
 read_path8 <- paste0("../data_new/county_data/zip_to_county_new.csv")
@@
 
-# which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_zip") # 159
-# which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_lag1_zip") # 160
-# which(colnames(inhaler_ndc_lag_df)=="wrf_temp_zip") # 93
-# which(colnames(inhaler_ndc_lag_df)=="wrf_temp_lag1_zip") # 150
-
-inhaler_geo_lag <- inhaler_ndc_lag_df %>%
-  select(159, 160:166) # GWR-method
-
-inhaler_temp_lag <- inhaler_ndc_lag_df %>%
-  select(93, 150:156) # GWR-method
-
-inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
-inhaler_temp_lag <- as.matrix(inhaler_temp_lag)
-
-# create fit matrix
-fit_mat <- matrix(NA, nrow=4, ncol=5)
-colnames(fit_mat) <- c("outcome", "exposure", "fit", "df", "aic")
-
-# fill outcome name in first column
-fit_mat[,1] <- "Beta - 2 agonists"
-# fill in exposure (geo smoke)
-fit_mat[,2] <- "geo_smk10"
-# fit type
-fit_mat[,3] <- "ns"
-
-# 2 df will not work for a spline, going from 3 to 6 df
-  for(h in 3:6){ # start loop for df
-
-    # fitting a natural spline for lag with 5 knots?
-    fit_mat[h-2, 4] <- degree_freedom <- h
-
-    # define basis b using natural spline function
-    B <- ns(0:7, df = degree_freedom, intercept = T)
-
-    # multiply lagged pm matrix by basis
-    inhaler_geo_lag_B <- inhaler_geo_lag%*%B
-    inhaler_temp_lag_B <- inhaler_temp_lag%*%B
-
-    fit <- clogit(outcome ~ inhaler_geo_lag_B + inhaler_temp_lag_B + strata(personkey), 
-              inhaler_ndc_lag_df)
-
-    AIC(fit)
-
-    # fill AIC
-    fit_mat[h-2,5] <- round(AIC(fit),5)
-
-    } # end model fit loop
-
-# now that we have a range of df/knots, I want to filter to the minimum aic
-# for each outcome and print out that value in a table
-
-kable(fit_mat, caption="Distributed lag spline of 3-6 degree of freedom fit by AIC")
-
-lag_spline_best_fit <- as_tibble(fit_mat) %>%  
-  slice(which.min(aic)) # df of 3 has the smallest AIC
-
-# kable
-knitr::kable(lag_spline_best_fit, caption = paste0("Distributed lag spline ",
-  "degree of freedom best fit by AIC"))
 ```
-
-### Distributed lag 0-7 for casecrossover
-
-When choosing the df of 5, the plot shows that the same day of fill with smoke exposure has the highest significant association.
-But other lags shows that pharmacy fills with lagged days do not has the significant association with smoke.
-And the association tends to decrease as lag goes.
 
@@
 cumulative_CI
+
+```
+
+```{r input category, echo = FALSE, message=FALSE, warning=FALSE}
+df_list <- list(asthma_df, copd_df, both_df, other_df)
+
+outcome_list <- c('Asthma', 'COPD', 'Both', 'Other')
+
+method_list <- c('Geo-Weighted Smoke')
+
+# create an empty list to row bind dataframes together
+datalist_dml <- list()
+
+# set list and do cross tabs to find values
+# Producing conditional logit model estimates loop 
+for(i in 1:length(df_list)){
+
+# datalist_dml[[i]] <- data.frame(df_list[i]) %>%
+  asthma_dml <- asthma_df %>%
+  left_join(or_zip_to_county, by = "ZIPCODE") %>%
+  left_join(or_pop, by = "county") %>%
+  mutate(day = ifelse(weekdays(date) == "Monday" |
+                      weekdays(date) == "Tuesday" |
+                      weekdays(date) == "Wednesday"|
+                      weekdays(date) == "Thursday"|
+                      weekdays(date) == "Friday", "weekday",
+               ifelse(weekdays(date) == "Saturday" |
+                      weekdays(date) == "Sunday", "weekend", NA))) %>%
+  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
+         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
+         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
+         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
+         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
+         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
+         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
+         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)
+
+asthma_dml$day[which(asthma_dml$date=="2013-05-27")] <- "weekend"
+asthma_dml$day[which(asthma_dml$date=="2013-07-04")] <- "weekend"
+asthma_dml$day[which(asthma_dml$date=="2013-09-02")] <- "weekend"
+
+datalist_dml[[i]]$day[which(datalist_dml[[i]]$date=="2013-05-27")] <- "weekend"
+datalist_dml[[i]]$day[which(datalist_dml[[i]]$date=="2013-07-04")] <- "weekend"
+datalist_dml[[i]]$day[which(datalist_dml[[i]]$date=="2013-09-02")] <- "weekend"
+}
+
+for(i in 1:4){
+  i <- 1
+
+  inhaler_geo_lag <- asthma_dml %>%
+  select(159, 160:166) # GWR-method
+
+inhaler_temp_lag <- asthma_dml %>%
+  select(93, 150:156) # GWR-method
+
+dml_df <- datalist_dml[[i]]
+
+lag_max <- 7
+n1 <- length(inhaler_geo_lag[,1])
+n2 <- length(inhaler_temp_lag[,1])
+
+inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
+inhaler_temp_lag <- as.matrix(inhaler_temp_lag)
+
+B <- ns(0:lag_max, df =5, intercept=TRUE)
+
+# the nrow B should equal ncol(pm)
+# dim(B)
+
+inhaler_geo_lag_B <- inhaler_geo_lag%*%B
+inhaler_temp_lag_B <- inhaler_temp_lag%*%B
+
+fit <- clogit(outcome ~ inhaler_geo_lag_B + inhaler_temp_lag_B + strata(personkey), 
+              asthma_dml)
+
+# coef(fit)
+# summary(fit)
+# AIC(fit)
+
+D <- cbind(B,B)
+
+# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
+dlparms <- c(1,2,3,4,5,6,7,8,9,10)
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms]) 
+DLvar <- D%*%vcov(fit)[dlparms,dlparms]%*%t(D)
+DLestimate$SE <- sqrt(diag(DLvar))
+DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
+DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
+
+DLestimate$estimate <- exp(DLestimate$estimate)
+DLestimate$lower <- exp(DLestimate$lower)
+DLestimate$upper <- exp(DLestimate$upper)
+
+# DLestimate
+
+DLestimate$lag <- 0:lag_max
+p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
+p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
+p <- p + geom_line(aes(x=lag), linetype=2, size=2)
+p <- p + theme_bw() + xlab("lagged days of events") + ylab("Odds Ratio for 10ug/m^3 Increase in PM2.5")
+p
+
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms])
+
+cumulative <- sum(DLestimate$estimate)
+# cumulative
+
+cumulative_se <- sqrt(sum(DLvar))
+# cumulative_se
+
+cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
+# cumulative_CI
+
+cum <- cbind(cumulative, cumulative_se, cumulative_CI[1], cumulative_CI[2])
+colnames(cum) <- c("Cumulative Est", "Cumulative SE", "Lower 95% CI", "Upper 95% CI")
+
+knitr::kable(cum, caption = paste0("Distributed lag 0-7 spline estimates, se and CI"))
+
+}
 
@@
 
+### Find the smallest of AIC with df from 3 to 6 (10)
+
+Choose df = 5 because it has the least AIC.
+
+```{r, message=FALSE, echo=FALSE, warning=FALSE}
+
+# which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_zip") # 159
+# which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_lag1_zip") # 160
+# which(colnames(inhaler_ndc_lag_df)=="wrf_temp_zip") # 93
+# which(colnames(inhaler_ndc_lag_df)=="wrf_temp_lag1_zip") # 150
+
+inhaler_geo_lag <- inhaler_ndc_lag_df %>%
+  select(159, 160:166) # GWR-method
+
+inhaler_temp_lag <- inhaler_ndc_lag_df %>%
+  select(93, 150:156) # GWR-method
+
+inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
+inhaler_temp_lag <- as.matrix(inhaler_temp_lag)
+
+# create fit matrix
+fit_mat <- matrix(NA, nrow=4, ncol=5)
+colnames(fit_mat) <- c("outcome", "exposure", "fit", "df", "aic")
+
+# fill outcome name in first column
+fit_mat[,1] <- "Beta - 2 agonists"
+# fill in exposure (geo smoke)
+fit_mat[,2] <- "geo_smk10"
+# fit type
+fit_mat[,3] <- "ns"
+
+# 2 df will not work for a spline, going from 3 to 6 df
+  for(h in 3:6){ # start loop for df
+
+    # fitting a natural spline for lag with 5 knots?
+    fit_mat[h-2, 4] <- degree_freedom <- h
+
+    # define basis b using natural spline function
+    B <- ns(0:7, df = degree_freedom, intercept = T)
+
+    # multiply lagged pm matrix by basis
+    inhaler_geo_lag_B <- inhaler_geo_lag%*%B
+    inhaler_temp_lag_B <- inhaler_temp_lag%*%B
+
+    fit <- clogit(outcome ~ inhaler_geo_lag_B + inhaler_temp_lag_B + strata(personkey), 
+              inhaler_ndc_lag_df)
+
+    AIC(fit)
+
+    # fill AIC
+    fit_mat[h-2,5] <- round(AIC(fit),5)
+
+    } # end model fit loop
+
+# now that we have a range of df/knots, I want to filter to the minimum aic
+# for each outcome and print out that value in a table
+
+kable(fit_mat, caption="Distributed lag spline of 3-6 degree of freedom fit by AIC")
+
+lag_spline_best_fit <- as_tibble(fit_mat) %>%  
+  slice(which.min(aic)) # df of 3 has the smallest AIC
+
+# kable
+knitr::kable(lag_spline_best_fit, caption = paste0("Distributed lag spline ",
+  "degree of freedom best fit by AIC"))
+```
+
 ### 1. For Overall Plot (Unconstrained lag model)

### 2017-08-08 r_scripts/analysis/oregon_medication.Rmd
Update oregon_medication.Rmd

@@
 
-```{r input category, echo = FALSE, message=FALSE, warning=FALSE}
-df_list <- list(asthma_df, copd_df, both_df, other_df)
-
-outcome_list <- c('Asthma', 'COPD', 'Both', 'Other')
-
-method_list <- c('Geo-Weighted Smoke')
-
-# create an empty list to row bind dataframes together
-datalist_dml <- list()
-
-# set list and do cross tabs to find values
-# Producing conditional logit model estimates loop 
-for(i in 1:length(df_list)){
-
-# datalist_dml[[i]] <- data.frame(df_list[i]) %>%
-  asthma_dml <- asthma_df %>%
+```{r DLM for disease, }
+
+# set variable list
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+
+start <- Sys.time()
+for(j in var_list) { # begin first loop of variable names (outcomes)
+
+  j <- "broken_arm"
+  # standard for statement without parallel computing
+  #for(j in 1:length(var_list)){
+  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
+  or_disease <- read_csv(read_path)
+
+  or_disease_df <- data.frame(or_disease)
+
+  or_disease_lag_df <- or_disease_df %>%
   left_join(or_zip_to_county, by = "ZIPCODE") %>%
@@
 
-asthma_dml$day[which(asthma_dml$date=="2013-05-27")] <- "weekend"
-asthma_dml$day[which(asthma_dml$date=="2013-07-04")] <- "weekend"
-asthma_dml$day[which(asthma_dml$date=="2013-09-02")] <- "weekend"
-
-datalist_dml[[i]]$day[which(datalist_dml[[i]]$date=="2013-05-27")] <- "weekend"
-datalist_dml[[i]]$day[which(datalist_dml[[i]]$date=="2013-07-04")] <- "weekend"
-datalist_dml[[i]]$day[which(datalist_dml[[i]]$date=="2013-09-02")] <- "weekend"
-}
-
-for(i in 1:4){
-  i <- 1
-
-  inhaler_geo_lag <- asthma_dml %>%
-  select(159, 160:166) # GWR-method
-
-inhaler_temp_lag <- asthma_dml %>%
-  select(93, 150:156) # GWR-method
-
-dml_df <- datalist_dml[[i]]
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-05-27")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-07-04")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-09-02")] <- "weekend"
+
+which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
+which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
+which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
+which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151
+
+or_disease_geo_lag <- or_disease_lag_df %>%
+  select(174, 175:181) # GWR-method
+
+or_disease_temp_lag <- or_disease_lag_df %>%
+  select(94, 151:157) # GWR-method
 
 lag_max <- 7
-n1 <- length(inhaler_geo_lag[,1])
-n2 <- length(inhaler_temp_lag[,1])
-
-inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
-inhaler_temp_lag <- as.matrix(inhaler_temp_lag)
+n1 <- length(or_disease_geo_lag[,1])
+n2 <- length(or_disease_temp_lag[,1])
+
+or_disease_geo_lag <- as.matrix(or_disease_geo_lag)
+or_disease_temp_lag <- as.matrix(or_disease_temp_lag)
 
@@
 
-inhaler_geo_lag_B <- inhaler_geo_lag%*%B
-inhaler_temp_lag_B <- inhaler_temp_lag%*%B
-
-fit <- clogit(outcome ~ inhaler_geo_lag_B + inhaler_temp_lag_B + strata(personkey), 
-              asthma_dml)
+or_disease_geo_lag_B <- or_disease_geo_lag%*%B
+or_disease_geo_lag_B <- or_disease_temp_lag%*%B
+
+fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_geo_lag_B + strata(personkey), 
+              or_disease_lag_df)
 
@@
 
+```
+
+## Supplement
+
+### Age and Sex category (8)
+### Time-Stratified by Age Categories
+
+Following table and figure look at some outcomes stratified by age category.
+
+For some outcomes, such like Age > 65 in WRF-Chem method, the CI is the only not significant one in the plot.
+
+The plot shows that overall, the risk of pharmacy fills of beta-2 agonist will increase significantly as the smoke concentrate increases.
+For age group of 15-65, the risk is the highest, for 10 ug/m^3 increase of pm2.5, the risk of filling the beta-2 agonist will increase about 9%.
+For age groups which are less than 15 or more than 65, the risk is similar, both of them will increase about 5.6%.
+
+```{r time stratified age strata cdc met, warning =F, echo = F, results='asis'} 
+# dataframe list
+
+method_list <- c('Geo-Weighted Smoke')
+
+# dataframe to loop through
+inhaler_ndc_df <- data.frame(inhaler_ndc)
+
+# outcome name
+outcome <- "Beta - 2 agonists"
+outcome_name <- "Beta - 2 agonists"
+
+# age category list
+age_cat_list <- c(0,1,2)
+
+# create an empty list to row bind dataframes together
+datalist1 <- list()
+
+# data wrangling ----
+# Producing conditional logit model estimates loop
+
+# extract covariates from dataframe
+covariates_df <- inhaler_ndc_df[, c(1:26, 74:84)]
+
+# extract pm values and divide by 10 and ordered
+#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+pm_estimates_df <- inhaler_ndc_df[, c(91, 93)]/10  # create 10 unit increases
+
+# new loop for age categories
+for(k in 0:2){
+  # empty matrix (12 x 10 matrix)
+  point_estimates <- matrix(nrow = 1, ncol = 10, byrow = T)
+
+  colnames(point_estimates) <- c('outcome', 'age_cat', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+  # fill in the outcome namedataframe before method loop
+  point_estimates[, 1] <- outcome_name
+  # repeat the age category 4 times for each pm method
+  point_estimates[, 2] <- k
+
+  # dataframe for analysis creation
+  # bind columns back together 
+  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+    # remove missing pm values
+    filter(!is.na(geo_smk_pm_zip)) %>% 
+    # limit to specific age category
+    filter(age_ind == k) %>% 
+    # the following code makes sure that the counterfactual values retained are 
+    # symetric in that number of obs before = number of obs after
+    mutate(obs_diff_admission = (fromdate - date)/7) 
+    # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+ j <- 38
+
+    # variable to model 
+    var_name <- colnames(df_analysis[j])
+
+    # set row number to fill
+    row_n <- j-37
+
+    # only run the model if the dataframe has observations
+    if(nrow(df_analysis) != 0){
+    # conditional logistic regression model
+    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+    # populate matrix
+    point_estimates[row_n, 3] <- method_list[row_n]
+    point_estimates[row_n, 4] <- mod$n
+    point_estimates[row_n, 5] <- mod$nevent
+    # odds ratio
+    point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+    # 95% lower bound
+    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
+    # 95% upper bound
+    point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                        1.96*(summary(mod)$coefficient[1,3])), 3)
+    # standard error
+    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
+    # p val
+    point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
+
+    # create else statement that fills matrix with missing so I still have the row
+    # in the final dataframe
+    } else {point_estimates[row_n, 3] <- method_list[row_n]
+            point_estimates[row_n, 4] <- 0
+            point_estimates[row_n, 5] <- 0
+            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
+     # end methods loop
+
+  # save point estimates as a dataframe
+  point_est_df <- as_data_frame(point_estimates)
+  datalist1[[k+1]] <- point_est_df
+} # end age category loop
+
+# bind rows of age category estimates together
+combined_point_est_df <- bind_rows(datalist1)
+
+wrf_geo_age <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" |
+    #      pm_method == "WRF-Chem Smoke") %>% 
+  mutate(age_cat2 = ifelse(age_cat == 0, "Less than 15", 
+                    ifelse(age_cat == 1, "15-65",
+                    ifelse(age_cat == 2, "Greater than 65", NA)))) %>% 
+  # subset columns I want to put in to the table
+  select(3, 11, 5:8) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
+         caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by age",
+         # row group by outcome
+         rgroup = "Beta - 2 agonists",
+         n.rgroup = c(rep(3, 1)), # 6 rows for each age cat for each outcome
+         # column headers
+         header = c("Est Method", "Age Group", "Events",
+                    "OR&dagger;", "Lower", "Upper"),
+         # column spanner
+         cgroup = c("", "95% CI"), 
+         n.cgroup = c(4, 2),
+         padding.rgroup = "&nbsp;&nbsp;",
+         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+         align = "llcccc", # column alignment,
+         tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within May to September fire season."
+         ) # end table
+
+print(tab)
+
+combined_point_est_df_new <- combined_point_est_df %>%
+  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
+                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
+  select(-n_events) %>%
+  rename(n_events = n_events_new)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_age_plot <- combined_point_est_df_new %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  # do not plot results with less than 15 events, create missing vals
+  # in mutate
+  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
+                             99, odds_ratio),
+         lower95 = ifelse(as.numeric(n_events) < 15, 
+                             99, lower95),
+         upper95 = ifelse(as.numeric(n_events) < 15, 
+                             99, upper95),
+         age_cat2 = ifelse(age_cat == 0, "<15", 
+                    ifelse(age_cat == 1, "15-65",
+                    ifelse(age_cat == 2, ">65", NA)))) %>% 
+  # subset columns I want to put in to the table
+  select(1, 3, 11, 10, 5:7) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
+                              levels = unique(wrf_geo_age_plot$outcome))
+
+wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
+                                levels = unique(wrf_geo_age_plot$pm_method))
+
+wrf_geo_age_plot$age_cat2 <- factor(wrf_geo_age_plot$age_cat2,
+                                levels = unique(wrf_geo_age_plot$age_cat2))
+
+wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
+wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
+wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
+wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)
+
+## ggplot 
+  print_plot <- ggplot(wrf_geo_age_plot,
+    aes(x = age_cat2, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red",  "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10Âµg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Age Category ') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10), 
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
+  ggsave("../plot_new/ndc_age_fig_gwr.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+### Seasonal Time-Stratified by Sex
+
+The association with smoke PM~2.5~ and health outcomes stratified by sex is shown below.
+
+All the CI do not include 1, which means the pharmacy filling (beta 2 agonists) increases about 5% when smoke increase 10 units.
+So the risk for getting asthma will increase about 5% ~ 8% when the smoke PM2.5 increase 10 units.
+Also, the risk for female is much higher than that for male.
+
+These results show that there might be effect modification by sex, where females might be more likely to fill a beta-2 agonist on the same day it's really smoky for the GWR estimates.
+
+```{r sex strata cdc met adj, warning = F, echo = F, results='asis'} 
+method_list <- c('Geo-Weighted Smoke')
+
+# dataframe to loop through
+inhaler_ndc_df <- data.frame(inhaler_ndc)
+
+# outcome name
+outcome <- "Beta - 2 agonists"
+outcome_name <- "Beta - 2 agonists"
+
+# create an empty list to row bind dataframes together
+datalist1 <- list()
+
+# sex category list
+sex_strata_list <- c(0,1)
+
+# data wrangling ----
+# Producing conditional logit model estimates loop
+
+# extract covariates from dataframe
+covariates_df <- inhaler_ndc_df[, c(1:26, 74:84)]
+
+# extract pm values and divide by 10 and ordered
+#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
+pm_estimates_df <- inhaler_ndc_df[, c(91, 93)]/10  # create 10 unit increases
+
+# new loop for sex categories
+for(k in 0:1){
+
+  # empty matrix (12 x 10 matrix)
+  point_estimates <- matrix(nrow = 1, ncol = 10, byrow = T)
+
+  colnames(point_estimates) <- c('outcome', 'sex', 'pm_method', 'n', 'n_events', 
+                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
+
+  # fill in the outcome namedataframe before method loop
+  point_estimates[, 1] <- outcome_name
+  # repeat the sex category 4 times for each pm method
+  point_estimates[, 2] <- k
+
+  # dataframe for analysis creation
+  # bind columns back together 
+  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
+    # remove missing pm values
+    filter(!is.na(geo_smk_pm_zip)) %>% 
+    # limit to specific sex category
+    filter(sex_ind == k) %>% 
+    # the following code makes sure that the counterfactual values retained are 
+    # symetric in that number of obs before = number of obs after
+    mutate(obs_diff_admission = (fromdate - date)/7) 
+    # dataframe is already for the entire fire season, so I don't need to subset anymore
+
+  j <- 38
+    # variable to model 
+    var_name <- colnames(df_analysis[j])
+
+    # set row number to fill
+    row_n <- j-37
+
+    # only run the model if the dataframe has observations
+    if(nrow(df_analysis) != 0){
+    # conditional logistic regression model
+    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
+
+    # populate matrix
+    point_estimates[row_n, 3] <- method_list[row_n]
+    point_estimates[row_n, 4] <- mod$n
+    point_estimates[row_n, 5] <- mod$nevent
+    # odds ratio
+    point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
+
+    # 95% lower bound
+    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
+                                      1.96*(summary(mod)$coefficient[1,3])), 3)
+    # 95% upper bound
+    point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
+                                      1.96*(summary(mod)$coefficient[1,3])), 3)
+    # standard error
+    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
+    # p val
+    point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
+
+    # create else statement that fills matrix with missing so I still have the row
+    # in the final dataframe
+    } else {point_estimates[row_n, 3] <- method_list[row_n]
+            point_estimates[row_n, 4] <- 0
+            point_estimates[row_n, 5] <- 0
+            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
+    # end methods loop
+
+  # save point estimates as a dataframe
+  point_est_df <- as_data_frame(point_estimates)
+
+  # combine previous values in dataframe that has all outcome/methods comparisons
+  datalist1[[k+1]] <- point_est_df
+} # end sex category loop
+
+# bind rows of sex category estimates together
+combined_point_est_df <- bind_rows(datalist1)
+
+wrf_geo_sex <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+  # filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
+  mutate(sex_cat = ifelse(sex == 0, "Male", 
+               ifelse(sex == 1, "Female", NA))) %>% 
+  # subset columns I want to put in to the table
+  select(3, 11, 5:8)%>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
+         lower95 = ifelse(lower95 == 99, "--", lower95),
+         upper95 = ifelse(upper95 == 99, "--", upper95))
+
+  tab <- htmlTable(txtRound(wrf_geo_sex, digits = 3, 1:3), 
+           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and pharmacy outcomes stratified by sex",
+           # row group by outcome
+           rgroup = "Beta - 2 agonists",
+           n.rgroup = c(rep(2, 1)), # 2 rows for each sex strata for each outcome
+           # column headers
+           header = c("Est Method", "Sex Strata", "Events",
+                      "OR&dagger;", "Lower", "Upper"),
+           # column spanner
+           cgroup = c("", "95% CI"), 
+           n.cgroup = c(4, 2),
+           padding.rgroup = "&nbsp;&nbsp;",
+           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
+           align = "llcccc", # column alignment,
+           tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within July to October fire season."
+            ) # end table
+
+  print(tab)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+wrf_geo_sex_plot <- combined_point_est_df %>% 
+  # filter columns to just geo-smk
+ #  filter(pm_method == "GWR Smoke" | 
+   #       pm_method == "WRF-Chem Smoke") %>% 
+  mutate(sex_cat = ifelse(sex == 0, "Male", 
+               ifelse(sex == 1, "Female", NA))) %>%
+  # subset columns I want to put in to the table
+  select(1, 3, 11, 5:8) %>% 
+  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
+         lower95 = ifelse(lower95 == 99, NA, lower95),
+         upper95 = ifelse(upper95 == 99, NA, upper95))
+
+# change characters to numeric and factor  
+wrf_geo_sex_plot$outcome <- factor(wrf_geo_sex_plot$outcome,
+                              levels = unique(wrf_geo_sex_plot$outcome))
+
+wrf_geo_sex_plot$pm_method <- factor(wrf_geo_sex_plot$pm_method,
+                                levels = unique(wrf_geo_sex_plot$pm_method))
+
+wrf_geo_sex_plot$sex_cat <- factor(wrf_geo_sex_plot$sex_cat,
+                                levels = unique(wrf_geo_sex_plot$sex_cat))
+
+wrf_geo_sex_plot$n_events <- as.numeric(wrf_geo_sex_plot$n_events)
+wrf_geo_sex_plot$odds_ratio <- as.numeric(wrf_geo_sex_plot$odds_ratio)
+wrf_geo_sex_plot$lower95 <- as.numeric(wrf_geo_sex_plot$lower95)
+wrf_geo_sex_plot$upper95 <- as.numeric(wrf_geo_sex_plot$upper95)
+
+## ggplot
+  print_plot <- ggplot(wrf_geo_sex_plot,
+    aes(x = sex_cat, y = odds_ratio, color = pm_method), na.rm = T) +
+    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
+    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
+    # custom color 
+    scale_color_manual(name = "Smoke-Estimation Method", 
+                       values = c("red", "blue", "#32115C"),
+                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
+    facet_wrap(~outcome, nrow = 3, scales = "free") +
+    geom_hline(yintercept = 1, linetype=2) +
+    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+    ylab(expression(paste("Odds Ratio for 10Âµg/m"^3, " Increase in PM"[2.5]))) +
+    #ylim(0, 2) +
+    xlab('Sex') +
+    # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10),
+    # axis element
+    #axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90),
+    # legend elements
+    legend.position = "bottom")
+    #legend.text = element_text(size = 8))
+
+  print(print_plot)
+  # save figure
+  ggsave("../plot_new/ndc_sex_fig_gwr.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```
+
+### Find the smallest of AIC with df from 3 to 6 (10)
+
+Choose df = 5 because it has the least AIC.
+
+```{r, message=FALSE, echo=FALSE, warning=FALSE}
+
+# which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_zip") # 159
+# which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_lag1_zip") # 160
+# which(colnames(inhaler_ndc_lag_df)=="wrf_temp_zip") # 93
+# which(colnames(inhaler_ndc_lag_df)=="wrf_temp_lag1_zip") # 150
+
+inhaler_geo_lag <- inhaler_ndc_lag_df %>%
+  select(159, 160:166) # GWR-method
+
+inhaler_temp_lag <- inhaler_ndc_lag_df %>%
+  select(93, 150:156) # GWR-method
+
+inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
+inhaler_temp_lag <- as.matrix(inhaler_temp_lag)
+
+# create fit matrix
+fit_mat <- matrix(NA, nrow=4, ncol=5)
+colnames(fit_mat) <- c("outcome", "exposure", "fit", "df", "aic")
+
+# fill outcome name in first column
+fit_mat[,1] <- "Beta - 2 agonists"
+# fill in exposure (geo smoke)
+fit_mat[,2] <- "geo_smk10"
+# fit type
+fit_mat[,3] <- "ns"
+
+# 2 df will not work for a spline, going from 3 to 6 df
+  for(h in 3:6){ # start loop for df
+
+    # fitting a natural spline for lag with 5 knots?
+    fit_mat[h-2, 4] <- degree_freedom <- h
+
+    # define basis b using natural spline function
+    B <- ns(0:7, df = degree_freedom, intercept = T)
+
+    # multiply lagged pm matrix by basis
+    inhaler_geo_lag_B <- inhaler_geo_lag%*%B
+    inhaler_temp_lag_B <- inhaler_temp_lag%*%B
+
+    fit <- clogit(outcome ~ inhaler_geo_lag_B + inhaler_temp_lag_B + strata(personkey), 
+              inhaler_ndc_lag_df)
+
+    AIC(fit)
+
+    # fill AIC
+    fit_mat[h-2,5] <- round(AIC(fit),5)
+
+    } # end model fit loop
+
+# now that we have a range of df/knots, I want to filter to the minimum aic
+# for each outcome and print out that value in a table
+
+kable(fit_mat, caption="Distributed lag spline of 3-6 degree of freedom fit by AIC")
+
+lag_spline_best_fit <- as_tibble(fit_mat) %>%  
+  slice(which.min(aic)) # df of 3 has the smallest AIC
+
+# kable
+knitr::kable(lag_spline_best_fit, caption = paste0("Distributed lag spline ",
+  "degree of freedom best fit by AIC"))
 ```
@@
 
-## Supplement
-
-### Age and Sex category (8)
-### Time-Stratified by Age Categories
-
-Following table and figure look at some outcomes stratified by age category.
-
-For some outcomes, such like Age > 65 in WRF-Chem method, the CI is the only not significant one in the plot.
-
-The plot shows that overall, the risk of pharmacy fills of beta-2 agonist will increase significantly as the smoke concentrate increases.
-For age group of 15-65, the risk is the highest, for 10 ug/m^3 increase of pm2.5, the risk of filling the beta-2 agonist will increase about 9%.
-For age groups which are less than 15 or more than 65, the risk is similar, both of them will increase about 5.6%.
-
-```{r time stratified age strata cdc met, warning =F, echo = F, results='asis'} 
-# dataframe list
-
-method_list <- c('Geo-Weighted Smoke')
-
-# dataframe to loop through
-inhaler_ndc_df <- data.frame(inhaler_ndc)
-
-# outcome name
-outcome <- "Beta - 2 agonists"
-outcome_name <- "Beta - 2 agonists"
-
-# age category list
-age_cat_list <- c(0,1,2)
-
-# create an empty list to row bind dataframes together
-datalist1 <- list()
-
-# data wrangling ----
-# Producing conditional logit model estimates loop
-
-# extract covariates from dataframe
-covariates_df <- inhaler_ndc_df[, c(1:26, 74:84)]
-
-# extract pm values and divide by 10 and ordered
-#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
-pm_estimates_df <- inhaler_ndc_df[, c(91, 93)]/10  # create 10 unit increases
-
-# new loop for age categories
-for(k in 0:2){
-  # empty matrix (12 x 10 matrix)
-  point_estimates <- matrix(nrow = 1, ncol = 10, byrow = T)
-
-  colnames(point_estimates) <- c('outcome', 'age_cat', 'pm_method', 'n', 'n_events', 
-                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
-
-  # fill in the outcome namedataframe before method loop
-  point_estimates[, 1] <- outcome_name
-  # repeat the age category 4 times for each pm method
-  point_estimates[, 2] <- k
-
-  # dataframe for analysis creation
-  # bind columns back together 
-  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
-    # remove missing pm values
-    filter(!is.na(geo_smk_pm_zip)) %>% 
-    # limit to specific age category
-    filter(age_ind == k) %>% 
-    # the following code makes sure that the counterfactual values retained are 
-    # symetric in that number of obs before = number of obs after
-    mutate(obs_diff_admission = (fromdate - date)/7) 
-    # dataframe is already for the entire fire season, so I don't need to subset anymore
-
- j <- 38
-
-    # variable to model 
-    var_name <- colnames(df_analysis[j])
-
-    # set row number to fill
-    row_n <- j-37
-
-    # only run the model if the dataframe has observations
-    if(nrow(df_analysis) != 0){
-    # conditional logistic regression model
-    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
-
-    # populate matrix
-    point_estimates[row_n, 3] <- method_list[row_n]
-    point_estimates[row_n, 4] <- mod$n
-    point_estimates[row_n, 5] <- mod$nevent
-    # odds ratio
-    point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
-
-    # 95% lower bound
-    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
-    # 95% upper bound
-    point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
-                                        1.96*(summary(mod)$coefficient[1,3])), 3)
-    # standard error
-    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
-    # p val
-    point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
-
-    # create else statement that fills matrix with missing so I still have the row
-    # in the final dataframe
-    } else {point_estimates[row_n, 3] <- method_list[row_n]
-            point_estimates[row_n, 4] <- 0
-            point_estimates[row_n, 5] <- 0
-            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
-     # end methods loop
-
-  # save point estimates as a dataframe
-  point_est_df <- as_data_frame(point_estimates)
-  datalist1[[k+1]] <- point_est_df
-} # end age category loop
-
-# bind rows of age category estimates together
-combined_point_est_df <- bind_rows(datalist1)
-
-wrf_geo_age <- combined_point_est_df %>% 
-  # filter columns to just geo-smk
-  # filter(pm_method == "GWR Smoke" |
-    #      pm_method == "WRF-Chem Smoke") %>% 
-  mutate(age_cat2 = ifelse(age_cat == 0, "Less than 15", 
-                    ifelse(age_cat == 1, "15-65",
-                    ifelse(age_cat == 2, "Greater than 65", NA)))) %>% 
-  # subset columns I want to put in to the table
-  select(3, 11, 5:8) %>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
-         lower95 = ifelse(lower95 == 99, "--", lower95),
-         upper95 = ifelse(upper95 == 99, "--", upper95))
-
-tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
-         caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by age",
-         # row group by outcome
-         rgroup = "Beta - 2 agonists",
-         n.rgroup = c(rep(3, 1)), # 6 rows for each age cat for each outcome
-         # column headers
-         header = c("Est Method", "Age Group", "Events",
-                    "OR&dagger;", "Lower", "Upper"),
-         # column spanner
-         cgroup = c("", "95% CI"), 
-         n.cgroup = c(4, 2),
-         padding.rgroup = "&nbsp;&nbsp;",
-         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
-         align = "llcccc", # column alignment,
-         tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within May to September fire season."
-         ) # end table
-
-print(tab)
-
-combined_point_est_df_new <- combined_point_est_df %>%
-  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
-                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
-  select(-n_events) %>%
-  rename(n_events = n_events_new)
-
-# ggplot of odds ratios, facet wrapped by outcomes -----
-# convert variables from character to either numeric or factor
-# factor preserves the order of the variable
-wrf_geo_age_plot <- combined_point_est_df_new %>% 
-  # filter columns to just geo-smk
- #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
-  # do not plot results with less than 15 events, create missing vals
-  # in mutate
-  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
-                             99, odds_ratio),
-         lower95 = ifelse(as.numeric(n_events) < 15, 
-                             99, lower95),
-         upper95 = ifelse(as.numeric(n_events) < 15, 
-                             99, upper95),
-         age_cat2 = ifelse(age_cat == 0, "<15", 
-                    ifelse(age_cat == 1, "15-65",
-                    ifelse(age_cat == 2, ">65", NA)))) %>% 
-  # subset columns I want to put in to the table
-  select(1, 3, 11, 10, 5:7) %>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
-         lower95 = ifelse(lower95 == 99, NA, lower95),
-         upper95 = ifelse(upper95 == 99, NA, upper95))
-
-# change characters to numeric and factor  
-wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
-                              levels = unique(wrf_geo_age_plot$outcome))
-
-wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
-                                levels = unique(wrf_geo_age_plot$pm_method))
-
-wrf_geo_age_plot$age_cat2 <- factor(wrf_geo_age_plot$age_cat2,
-                                levels = unique(wrf_geo_age_plot$age_cat2))
-
-wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
-wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
-wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
-wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)
-
-## ggplot 
-  print_plot <- ggplot(wrf_geo_age_plot,
-    aes(x = age_cat2, y = odds_ratio, color = pm_method), na.rm = T) +
-    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
-    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
-                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
-    # custom color 
-    scale_color_manual(name = "Smoke-Estimation Method", 
-                       values = c("red",  "blue", "#32115C"),
-                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
-    facet_wrap(~outcome, nrow = 3, scales = "free") +
-    geom_hline(yintercept = 1, linetype=2) +
-    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
-    ylab(expression(paste("Odds Ratio for 10Âµg/m"^3, " Increase in PM"[2.5]))) +
-    #ylim(0, 2) +
-    xlab('Age Category ') +
-    # plot theme
-    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    # strip element
-    strip.background = element_rect(colour=NA, fill=NA),
-    panel.border = element_rect(fill = NA, color = "black"),
-    # facet text size
-    strip.text = element_text(size = 10), 
-    # axis element
-    #axis.text.x = element_blank(),
-    #axis.title.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.title.y = element_text(angle = 90),
-    # legend elements
-    legend.position = "bottom")
-    #legend.text = element_text(size = 8))
-
-  print(print_plot)
-  # save figure
-  ggsave("../plot_new/ndc_age_fig_gwr.pdf", plot = print_plot, 
-       width = 12, height = 8, units = "in")
-
-```
-
-### Seasonal Time-Stratified by Sex
-
-The association with smoke PM~2.5~ and health outcomes stratified by sex is shown below.
-
-All the CI do not include 1, which means the pharmacy filling (beta 2 agonists) increases about 5% when smoke increase 10 units.
-So the risk for getting asthma will increase about 5% ~ 8% when the smoke PM2.5 increase 10 units.
-Also, the risk for female is much higher than that for male.
-
-These results show that there might be effect modification by sex, where females might be more likely to fill a beta-2 agonist on the same day it's really smoky for the GWR estimates.
-
-```{r sex strata cdc met adj, warning = F, echo = F, results='asis'} 
-method_list <- c('Geo-Weighted Smoke')
-
-# dataframe to loop through
-inhaler_ndc_df <- data.frame(inhaler_ndc)
-
-# outcome name
-outcome <- "Beta - 2 agonists"
-outcome_name <- "Beta - 2 agonists"
-
-# create an empty list to row bind dataframes together
-datalist1 <- list()
-
-# sex category list
-sex_strata_list <- c(0,1)
-
-# data wrangling ----
-# Producing conditional logit model estimates loop
-
-# extract covariates from dataframe
-covariates_df <- inhaler_ndc_df[, c(1:26, 74:84)]
-
-# extract pm values and divide by 10 and ordered
-#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
-pm_estimates_df <- inhaler_ndc_df[, c(91, 93)]/10  # create 10 unit increases
-
-# new loop for sex categories
-for(k in 0:1){
-
-  # empty matrix (12 x 10 matrix)
-  point_estimates <- matrix(nrow = 1, ncol = 10, byrow = T)
-
-  colnames(point_estimates) <- c('outcome', 'sex', 'pm_method', 'n', 'n_events', 
-                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
-
-  # fill in the outcome namedataframe before method loop
-  point_estimates[, 1] <- outcome_name
-  # repeat the sex category 4 times for each pm method
-  point_estimates[, 2] <- k
-
-  # dataframe for analysis creation
-  # bind columns back together 
-  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
-    # remove missing pm values
-    filter(!is.na(geo_smk_pm_zip)) %>% 
-    # limit to specific sex category
-    filter(sex_ind == k) %>% 
-    # the following code makes sure that the counterfactual values retained are 
-    # symetric in that number of obs before = number of obs after
-    mutate(obs_diff_admission = (fromdate - date)/7) 
-    # dataframe is already for the entire fire season, so I don't need to subset anymore
-
-  j <- 38
-    # variable to model 
-    var_name <- colnames(df_analysis[j])
-
-    # set row number to fill
-    row_n <- j-37
-
-    # only run the model if the dataframe has observations
-    if(nrow(df_analysis) != 0){
-    # conditional logistic regression model
-    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
-
-    # populate matrix
-    point_estimates[row_n, 3] <- method_list[row_n]
-    point_estimates[row_n, 4] <- mod$n
-    point_estimates[row_n, 5] <- mod$nevent
-    # odds ratio
-    point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
-
-    # 95% lower bound
-    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
-                                      1.96*(summary(mod)$coefficient[1,3])), 3)
-    # 95% upper bound
-    point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
-                                      1.96*(summary(mod)$coefficient[1,3])), 3)
-    # standard error
-    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
-    # p val
-    point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
-
-    # create else statement that fills matrix with missing so I still have the row
-    # in the final dataframe
-    } else {point_estimates[row_n, 3] <- method_list[row_n]
-            point_estimates[row_n, 4] <- 0
-            point_estimates[row_n, 5] <- 0
-            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
-    # end methods loop
-
-  # save point estimates as a dataframe
-  point_est_df <- as_data_frame(point_estimates)
-
-  # combine previous values in dataframe that has all outcome/methods comparisons
-  datalist1[[k+1]] <- point_est_df
-} # end sex category loop
-
-# bind rows of sex category estimates together
-combined_point_est_df <- bind_rows(datalist1)
-
-wrf_geo_sex <- combined_point_est_df %>% 
-  # filter columns to just geo-smk
-  # filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
-  mutate(sex_cat = ifelse(sex == 0, "Male", 
-               ifelse(sex == 1, "Female", NA))) %>% 
-  # subset columns I want to put in to the table
-  select(3, 11, 5:8)%>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
-         lower95 = ifelse(lower95 == 99, "--", lower95),
-         upper95 = ifelse(upper95 == 99, "--", upper95))
-
-  tab <- htmlTable(txtRound(wrf_geo_sex, digits = 3, 1:3), 
-           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and pharmacy outcomes stratified by sex",
-           # row group by outcome
-           rgroup = "Beta - 2 agonists",
-           n.rgroup = c(rep(2, 1)), # 2 rows for each sex strata for each outcome
-           # column headers
-           header = c("Est Method", "Sex Strata", "Events",
-                      "OR&dagger;", "Lower", "Upper"),
-           # column spanner
-           cgroup = c("", "95% CI"), 
-           n.cgroup = c(4, 2),
-           padding.rgroup = "&nbsp;&nbsp;",
-           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
-           align = "llcccc", # column alignment,
-           tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within July to October fire season."
-            ) # end table
-
-  print(tab)
-
-# ggplot of odds ratios, facet wrapped by outcomes -----
-# convert variables from character to either numeric or factor
-# factor preserves the order of the variable
-wrf_geo_sex_plot <- combined_point_est_df %>% 
-  # filter columns to just geo-smk
- #  filter(pm_method == "GWR Smoke" | 
-   #       pm_method == "WRF-Chem Smoke") %>% 
-  mutate(sex_cat = ifelse(sex == 0, "Male", 
-               ifelse(sex == 1, "Female", NA))) %>%
-  # subset columns I want to put in to the table
-  select(1, 3, 11, 5:8) %>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
-         lower95 = ifelse(lower95 == 99, NA, lower95),
-         upper95 = ifelse(upper95 == 99, NA, upper95))
-
-# change characters to numeric and factor  
-wrf_geo_sex_plot$outcome <- factor(wrf_geo_sex_plot$outcome,
-                              levels = unique(wrf_geo_sex_plot$outcome))
-
-wrf_geo_sex_plot$pm_method <- factor(wrf_geo_sex_plot$pm_method,
-                                levels = unique(wrf_geo_sex_plot$pm_method))
-
-wrf_geo_sex_plot$sex_cat <- factor(wrf_geo_sex_plot$sex_cat,
-                                levels = unique(wrf_geo_sex_plot$sex_cat))
-
-wrf_geo_sex_plot$n_events <- as.numeric(wrf_geo_sex_plot$n_events)
-wrf_geo_sex_plot$odds_ratio <- as.numeric(wrf_geo_sex_plot$odds_ratio)
-wrf_geo_sex_plot$lower95 <- as.numeric(wrf_geo_sex_plot$lower95)
-wrf_geo_sex_plot$upper95 <- as.numeric(wrf_geo_sex_plot$upper95)
-
-## ggplot
-  print_plot <- ggplot(wrf_geo_sex_plot,
-    aes(x = sex_cat, y = odds_ratio, color = pm_method), na.rm = T) +
-    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
-    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
-                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
-    # custom color 
-    scale_color_manual(name = "Smoke-Estimation Method", 
-                       values = c("red", "blue", "#32115C"),
-                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
-    facet_wrap(~outcome, nrow = 3, scales = "free") +
-    geom_hline(yintercept = 1, linetype=2) +
-    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
-    ylab(expression(paste("Odds Ratio for 10Âµg/m"^3, " Increase in PM"[2.5]))) +
-    #ylim(0, 2) +
-    xlab('Sex') +
-    # plot theme
-    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    # strip element
-    strip.background = element_rect(colour=NA, fill=NA),
-    panel.border = element_rect(fill = NA, color = "black"),
-    # facet text size
-    strip.text = element_text(size = 10),
-    # axis element
-    #axis.text.x = element_blank(),
-    #axis.title.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.title.y = element_text(angle = 90),
-    # legend elements
-    legend.position = "bottom")
-    #legend.text = element_text(size = 8))
-
-  print(print_plot)
-  # save figure
-  ggsave("../plot_new/ndc_sex_fig_gwr.pdf", plot = print_plot, 
-       width = 12, height = 8, units = "in")
-
-```
-
-### Find the smallest of AIC with df from 3 to 6 (10)
-
-Choose df = 5 because it has the least AIC.
-
-```{r, message=FALSE, echo=FALSE, warning=FALSE}
-
-# which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_zip") # 159
-# which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_lag1_zip") # 160
-# which(colnames(inhaler_ndc_lag_df)=="wrf_temp_zip") # 93
-# which(colnames(inhaler_ndc_lag_df)=="wrf_temp_lag1_zip") # 150
-
-inhaler_geo_lag <- inhaler_ndc_lag_df %>%
-  select(159, 160:166) # GWR-method
-
-inhaler_temp_lag <- inhaler_ndc_lag_df %>%
-  select(93, 150:156) # GWR-method
-
-inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
-inhaler_temp_lag <- as.matrix(inhaler_temp_lag)
-
-# create fit matrix
-fit_mat <- matrix(NA, nrow=4, ncol=5)
-colnames(fit_mat) <- c("outcome", "exposure", "fit", "df", "aic")
-
-# fill outcome name in first column
-fit_mat[,1] <- "Beta - 2 agonists"
-# fill in exposure (geo smoke)
-fit_mat[,2] <- "geo_smk10"
-# fit type
-fit_mat[,3] <- "ns"
-
-# 2 df will not work for a spline, going from 3 to 6 df
-  for(h in 3:6){ # start loop for df
-
-    # fitting a natural spline for lag with 5 knots?
-    fit_mat[h-2, 4] <- degree_freedom <- h
-
-    # define basis b using natural spline function
-    B <- ns(0:7, df = degree_freedom, intercept = T)
-
-    # multiply lagged pm matrix by basis
-    inhaler_geo_lag_B <- inhaler_geo_lag%*%B
-    inhaler_temp_lag_B <- inhaler_temp_lag%*%B
-
-    fit <- clogit(outcome ~ inhaler_geo_lag_B + inhaler_temp_lag_B + strata(personkey), 
-              inhaler_ndc_lag_df)
-
-    AIC(fit)
-
-    # fill AIC
-    fit_mat[h-2,5] <- round(AIC(fit),5)
-
-    } # end model fit loop
-
-# now that we have a range of df/knots, I want to filter to the minimum aic
-# for each outcome and print out that value in a table
-
-kable(fit_mat, caption="Distributed lag spline of 3-6 degree of freedom fit by AIC")
-
-lag_spline_best_fit <- as_tibble(fit_mat) %>%  
-  slice(which.min(aic)) # df of 3 has the smallest AIC
-
-# kable
-knitr::kable(lag_spline_best_fit, caption = paste0("Distributed lag spline ",
-  "degree of freedom best fit by AIC"))
-```
-
 ### 1. For Overall Plot (Unconstrained lag model)

### 2017-09-01 r_scripts/analysis/oregon_medication.Rmd
Update oregon_medication.Rmd

@@
 ---
-title: "The GWR Smoke's effect on Oregon 2013 Medication of Beta-2 agonists"
+title: "The GWR wildfire smoke's association with 2013 Oregon medication of short acting beta agonists"
 author: "Jingyang Liu"
@@
 library(sp)
+library(cowplot)
 
@@
 
-## Overview (1, 2)
-
-1.
-Intro: We found an association with wildfire smoke and respiratory outcomes in Washington state 2012 using a novel estimate of smoke concentration, geo-graphically weighted ridge regression (GWR).
+## Overview
+
+**1. Introduction: **
+We found an association with wildfire smoke and respiratory outcomes in Washington state 2012 using a novel estimate of smoke concentration, geo-graphically weighted ridge regression (GWR).
 We want to know if smoke impacts other health outcomes, specifically respiratory rescue medications (beta 2 agonists) (abbreviate to SABA)
 
-2. Research question: We evaluated the association between smoke concentrations using the GWR method and SABA refills in Oregon state during the 2013 wildfire season.
+**2. Research question: **
+We evaluated the association between smoke concentrations using the GWR method and SABA refills in Oregon state during the 2013 wildfire season.
 
@@
 
-## Method Description (3)
-
-3.Methods: use case-crossover for all health outcome figures and results. Make sure smoke is assigned via zipcode. For the descriptive of smoke, I think the county-level for now. We may move to zip.
+**3.Methods: **
+use case-crossover for all health outcome figures and results. Make sure smoke is assigned via zipcode. For the descriptive of smoke, I think the county-level for now. We may move to zip.
 
@@
 
-```{r DLM for casecrossover temp, echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}
-read_path8 <- paste0("../data_new/county_data/zip_to_county_new.csv")
-or_zip_to_county <- read_csv(read_path8)
-
-read_path9 <- paste0("../data_new/county_data/or_census_pop_est_county.csv")
-or_pop <- read_csv(read_path9)
-
-or_pop <- or_pop %>%
-  rename(county = GEO_display_label, pop = respop72013) %>%
-  select(county, pop) 
-or_pop$county <- gsub(" County, Oregon", "", as.character(factor(or_pop$county)))
-
-inhaler_ndc_lag_df <- inhaler_ndc_df %>%
-  left_join(or_zip_to_county, by = "ZIPCODE") %>%
-  left_join(or_pop, by = "county") %>%
-  mutate(day = ifelse(weekdays(date) == "Monday" |
-                      weekdays(date) == "Tuesday" |
-                      weekdays(date) == "Wednesday"|
-                      weekdays(date) == "Thursday"|
-                      weekdays(date) == "Friday", "weekday",
-               ifelse(weekdays(date) == "Saturday" |
-                      weekdays(date) == "Sunday", "weekend", NA))) %>%
-  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
-         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
-         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
-         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
-         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
-         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
-         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
-         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)
-
-inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-05-27")] <- "weekend"
-inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-07-04")] <- "weekend"
-inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-09-02")] <- "weekend"
-
-which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_zip") # 159
-which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_lag1_zip") # 160
-which(colnames(inhaler_ndc_lag_df)=="wrf_temp_zip") # 93
-which(colnames(inhaler_ndc_lag_df)=="wrf_temp_lag1_zip") # 150
-
-inhaler_geo_lag <- inhaler_ndc_lag_df %>%
-  select(159, 160:166) # GWR-method
-
-inhaler_temp_lag <- inhaler_ndc_lag_df %>%
-  select(93, 150:156) # GWR-method
-
-lag_max <- 7
-n1 <- length(inhaler_geo_lag[,1])
-n2 <- length(inhaler_temp_lag[,1])
-
-inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
-inhaler_temp_lag <- as.matrix(inhaler_temp_lag)
-
-B <- ns(0:lag_max, df =3, intercept=TRUE)
-
-# the nrow B should equal ncol(pm)
-dim(B)
-
-inhaler_geo_lag_B <- inhaler_geo_lag%*%B
-inhaler_temp_lag_B <- inhaler_temp_lag%*%B
-
-fit <- clogit(outcome ~  inhaler_temp_lag_B + strata(personkey), 
-              inhaler_ndc_lag_df)
-
-coef(fit)
-summary(fit)
-AIC(fit)
-
-D <- cbind(B,B)
-
-# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
-dlparms <- c(1,2,3)
-DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms]) 
-DLvar <- B%*%vcov(fit)[dlparms,dlparms]%*%t(B)
-DLestimate$SE <- sqrt(diag(DLvar))
-DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
-DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
-
-DLestimate$estimate <- exp(DLestimate$estimate)
-DLestimate$lower <- exp(DLestimate$lower)
-DLestimate$upper <- exp(DLestimate$upper)
-
-DLestimate
-
-DLestimate$lag <- 0:lag_max
-p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
-p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
-p <- p + geom_line(aes(x=lag), linetype=2, size=2)
-p <- p + theme_bw()
-p
-
-DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms])
-
-cumulative <- sum(DLestimate$estimate)
-cumulative
-
-cumulative_se <- sqrt(sum(DLvar))
-cumulative_se
-
-cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
-cumulative_CI
-
-```
-
-```{r DLM for disease, }
+```{r DLM for disease respiratory, echo = FALSE, message=FALSE, warning=FALSE}
 
@@
 
-start <- Sys.time()
-for(j in var_list) { # begin first loop of variable names (outcomes)
-
-  j <- "broken_arm"
+  j <- c('respiratory')
+
   # standard for statement without parallel computing
@@
 
-which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
-which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
-which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
-which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
+# which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
+# which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151
 
@@
 or_disease_geo_lag_B <- or_disease_geo_lag%*%B
-or_disease_geo_lag_B <- or_disease_temp_lag%*%B
-
-fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_geo_lag_B + strata(personkey), 
+or_disease_temp_lag_B <- or_disease_temp_lag%*%B
+
+fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_temp_lag_B + strata(personkey), 
               or_disease_lag_df)
@@
 DLestimate$lag <- 0:lag_max
-p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
-p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
-p <- p + geom_line(aes(x=lag), linetype=2, size=2)
-p <- p + theme_bw() + xlab("lagged days of events") + ylab("Odds Ratio for 10ug/m^3 Increase in PM2.5")
-p
+p1 <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper)) + 
+  geom_ribbon(alpha=.5) + geom_line(size=2) + 
+  geom_line(aes(x=lag), linetype=2, size=2) + 
+  theme_bw() + 
+  xlab("lagged days of events") + 
+  ylab("OR for 10ug/m^3 Increase") +
+  ggtitle("Respiratory") +
+  geom_hline(yintercept=1, linetype="dashed", color = "red")
 
@@
 
-}
+```
+
+```{r DLM for disease asthma, echo = FALSE, message=FALSE, warning=FALSE}
+
+# set variable list
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+
+  j <- c('asthma')
+
+  # standard for statement without parallel computing
+  #for(j in 1:length(var_list)){
+  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
+  or_disease <- read_csv(read_path)
+
+  or_disease_df <- data.frame(or_disease)
+
+  or_disease_lag_df <- or_disease_df %>%
+  left_join(or_zip_to_county, by = "ZIPCODE") %>%
+  left_join(or_pop, by = "county") %>%
+  mutate(day = ifelse(weekdays(date) == "Monday" |
+                      weekdays(date) == "Tuesday" |
+                      weekdays(date) == "Wednesday"|
+                      weekdays(date) == "Thursday"|
+                      weekdays(date) == "Friday", "weekday",
+               ifelse(weekdays(date) == "Saturday" |
+                      weekdays(date) == "Sunday", "weekend", NA))) %>%
+  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
+         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
+         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
+         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
+         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
+         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
+         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
+         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)
+
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-05-27")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-07-04")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-09-02")] <- "weekend"
+
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
+# which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
+# which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151
+
+or_disease_geo_lag <- or_disease_lag_df %>%
+  select(174, 175:181) # GWR-method
+
+or_disease_temp_lag <- or_disease_lag_df %>%
+  select(94, 151:157) # GWR-method
+
+lag_max <- 7
+n1 <- length(or_disease_geo_lag[,1])
+n2 <- length(or_disease_temp_lag[,1])
+
+or_disease_geo_lag <- as.matrix(or_disease_geo_lag)
+or_disease_temp_lag <- as.matrix(or_disease_temp_lag)
+
+B <- ns(0:lag_max, df =5, intercept=TRUE)
+
+# the nrow B should equal ncol(pm)
+# dim(B)
+
+or_disease_geo_lag_B <- or_disease_geo_lag%*%B
+or_disease_temp_lag_B <- or_disease_temp_lag%*%B
+
+fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_temp_lag_B + strata(personkey), 
+              or_disease_lag_df)
+
+# coef(fit)
+# summary(fit)
+# AIC(fit)
+
+D <- cbind(B,B)
+
+# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
+dlparms <- c(1,2,3,4,5,6,7,8,9,10)
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms]) 
+DLvar <- D%*%vcov(fit)[dlparms,dlparms]%*%t(D)
+DLestimate$SE <- sqrt(diag(DLvar))
+DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
+DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
+
+DLestimate$estimate <- exp(DLestimate$estimate)
+DLestimate$lower <- exp(DLestimate$lower)
+DLestimate$upper <- exp(DLestimate$upper)
+
+# DLestimate
+DLestimate$lag <- 0:lag_max
+
+p2 <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper)) + 
+  geom_ribbon(alpha=.5) + geom_line(size=2) + 
+  geom_line(aes(x=lag), linetype=2, size=2) + 
+  theme_bw() + 
+  xlab("lagged days of events") + 
+  ylab("OR for 10ug/m^3 Increase") +
+  ggtitle("Asthma") +
+  geom_hline(yintercept=1, linetype="dashed", color = "red")
+
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms])
+
+cumulative <- sum(DLestimate$estimate)
+# cumulative
+
+cumulative_se <- sqrt(sum(DLvar))
+# cumulative_se
+
+cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
+# cumulative_CI
+
+cum <- cbind(cumulative, cumulative_se, cumulative_CI[1], cumulative_CI[2])
+colnames(cum) <- c("Cumulative Est", "Cumulative SE", "Lower 95% CI", "Upper 95% CI")
+
+knitr::kable(cum, caption = paste0("Distributed lag 0-7 spline estimates, se and CI"))
+
+```
+
+```{r DLM for disease pneumonia, echo = FALSE, message=FALSE, warning=FALSE}
+
+# set variable list
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+
+  j <- c('pneumonia')
+
+  # standard for statement without parallel computing
+  #for(j in 1:length(var_list)){
+  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
+  or_disease <- read_csv(read_path)
+
+  or_disease_df <- data.frame(or_disease)
+
+  or_disease_lag_df <- or_disease_df %>%
+  left_join(or_zip_to_county, by = "ZIPCODE") %>%
+  left_join(or_pop, by = "county") %>%
+  mutate(day = ifelse(weekdays(date) == "Monday" |
+                      weekdays(date) == "Tuesday" |
+                      weekdays(date) == "Wednesday"|
+                      weekdays(date) == "Thursday"|
+                      weekdays(date) == "Friday", "weekday",
+               ifelse(weekdays(date) == "Saturday" |
+                      weekdays(date) == "Sunday", "weekend", NA))) %>%
+  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
+         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
+         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
+         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
+         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
+         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
+         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
+         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)
+
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-05-27")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-07-04")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-09-02")] <- "weekend"
+
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
+# which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
+# which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151
+
+or_disease_geo_lag <- or_disease_lag_df %>%
+  select(174, 175:181) # GWR-method
+
+or_disease_temp_lag <- or_disease_lag_df %>%
+  select(94, 151:157) # GWR-method
+
+lag_max <- 7
+n1 <- length(or_disease_geo_lag[,1])
+n2 <- length(or_disease_temp_lag[,1])
+
+or_disease_geo_lag <- as.matrix(or_disease_geo_lag)
+or_disease_temp_lag <- as.matrix(or_disease_temp_lag)
+
+B <- ns(0:lag_max, df =5, intercept=TRUE)
+
+# the nrow B should equal ncol(pm)
+# dim(B)
+
+or_disease_geo_lag_B <- or_disease_geo_lag%*%B
+or_disease_temp_lag_B <- or_disease_temp_lag%*%B
+
+fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_temp_lag_B + strata(personkey), 
+              or_disease_lag_df)
+
+# coef(fit)
+# summary(fit)
+# AIC(fit)
+
+D <- cbind(B,B)
+
+# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
+dlparms <- c(1,2,3,4,5,6,7,8,9,10)
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms]) 
+DLvar <- D%*%vcov(fit)[dlparms,dlparms]%*%t(D)
+DLestimate$SE <- sqrt(diag(DLvar))
+DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
+DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
+
+DLestimate$estimate <- exp(DLestimate$estimate)
+DLestimate$lower <- exp(DLestimate$lower)
+DLestimate$upper <- exp(DLestimate$upper)
+
+# DLestimate
+
+DLestimate$lag <- 0:lag_max
+p3 <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper)) + 
+  geom_ribbon(alpha=.5) + geom_line(size=2) + 
+  geom_line(aes(x=lag), linetype=2, size=2) + 
+  theme_bw() + 
+  xlab("lagged days of events") + 
+  ylab("OR for 10ug/m^3 Increase") +
+  ggtitle("Pneumonia") +
+  geom_hline(yintercept=1, linetype="dashed", color = "red")
+
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms])
+
+cumulative <- sum(DLestimate$estimate)
+# cumulative
+
+cumulative_se <- sqrt(sum(DLvar))
+# cumulative_se
+
+cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
+# cumulative_CI
+
+cum <- cbind(cumulative, cumulative_se, cumulative_CI[1], cumulative_CI[2])
+colnames(cum) <- c("Cumulative Est", "Cumulative SE", "Lower 95% CI", "Upper 95% CI")
+
+knitr::kable(cum, caption = paste0("Distributed lag 0-7 spline estimates, se and CI"))
+
+```
+
+```{r DLM for disease acute_bronch, echo = FALSE, message=FALSE, warning=FALSE}
+
+# set variable list
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+
+  j <- c('acute_bronch')
+
+  # standard for statement without parallel computing
+  #for(j in 1:length(var_list)){
+  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
+  or_disease <- read_csv(read_path)
+
+  or_disease_df <- data.frame(or_disease)
+
+  or_disease_lag_df <- or_disease_df %>%
+  left_join(or_zip_to_county, by = "ZIPCODE") %>%
+  left_join(or_pop, by = "county") %>%
+  mutate(day = ifelse(weekdays(date) == "Monday" |
+                      weekdays(date) == "Tuesday" |
+                      weekdays(date) == "Wednesday"|
+                      weekdays(date) == "Thursday"|
+                      weekdays(date) == "Friday", "weekday",
+               ifelse(weekdays(date) == "Saturday" |
+                      weekdays(date) == "Sunday", "weekend", NA))) %>%
+  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
+         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
+         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
+         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
+         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
+         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
+         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
+         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)
+
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-05-27")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-07-04")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-09-02")] <- "weekend"
+
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
+# which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
+# which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151
+
+or_disease_geo_lag <- or_disease_lag_df %>%
+  select(174, 175:181) # GWR-method
+
+or_disease_temp_lag <- or_disease_lag_df %>%
+  select(94, 151:157) # GWR-method
+
+lag_max <- 7
+n1 <- length(or_disease_geo_lag[,1])
+n2 <- length(or_disease_temp_lag[,1])
+
+or_disease_geo_lag <- as.matrix(or_disease_geo_lag)
+or_disease_temp_lag <- as.matrix(or_disease_temp_lag)
+
+B <- ns(0:lag_max, df =5, intercept=TRUE)
+
+# the nrow B should equal ncol(pm)
+# dim(B)
+
+or_disease_geo_lag_B <- or_disease_geo_lag%*%B
+or_disease_temp_lag_B <- or_disease_temp_lag%*%B
+
+fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_temp_lag_B + strata(personkey), 
+              or_disease_lag_df)
+
+# coef(fit)
+# summary(fit)
+# AIC(fit)
+
+D <- cbind(B,B)
+
+# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
+dlparms <- c(1,2,3,4,5,6,7,8,9,10)
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms]) 
+DLvar <- D%*%vcov(fit)[dlparms,dlparms]%*%t(D)
+DLestimate$SE <- sqrt(diag(DLvar))
+DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
+DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
+
+DLestimate$estimate <- exp(DLestimate$estimate)
+DLestimate$lower <- exp(DLestimate$lower)
+DLestimate$upper <- exp(DLestimate$upper)
+
+# DLestimate
+
+DLestimate$lag <- 0:lag_max
+
+p4 <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper)) + 
+  geom_ribbon(alpha=.5) + geom_line(size=2) + 
+  geom_line(aes(x=lag), linetype=2, size=2) + 
+  theme_bw() + 
+  xlab("lagged days of events") + 
+  ylab("OR for 10ug/m^3 Increase") +
+  ggtitle("Acute bronch") +
+  geom_hline(yintercept=1, linetype="dashed", color = "red")
+
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms])
+
+cumulative <- sum(DLestimate$estimate)
+# cumulative
+
+cumulative_se <- sqrt(sum(DLvar))
+# cumulative_se
+
+cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
+# cumulative_CI
+
+cum <- cbind(cumulative, cumulative_se, cumulative_CI[1], cumulative_CI[2])
+colnames(cum) <- c("Cumulative Est", "Cumulative SE", "Lower 95% CI", "Upper 95% CI")
+
+knitr::kable(cum, caption = paste0("Distributed lag 0-7 spline estimates, se and CI"))
+
+```
+
+```{r DLM for disease copd, echo = FALSE, message=FALSE, warning=FALSE}
+
+# set variable list
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+
+  j <- c('copd')
+
+  # standard for statement without parallel computing
+  #for(j in 1:length(var_list)){
+  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
+  or_disease <- read_csv(read_path)
+
+  or_disease_df <- data.frame(or_disease)
+
+  or_disease_lag_df <- or_disease_df %>%
+  left_join(or_zip_to_county, by = "ZIPCODE") %>%
+  left_join(or_pop, by = "county") %>%
+  mutate(day = ifelse(weekdays(date) == "Monday" |
+                      weekdays(date) == "Tuesday" |
+                      weekdays(date) == "Wednesday"|
+                      weekdays(date) == "Thursday"|
+                      weekdays(date) == "Friday", "weekday",
+               ifelse(weekdays(date) == "Saturday" |
+                      weekdays(date) == "Sunday", "weekend", NA))) %>%
+  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
+         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
+         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
+         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
+         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
+         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
+         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
+         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)
+
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-05-27")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-07-04")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-09-02")] <- "weekend"
+
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
+# which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
+# which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151
+
+or_disease_geo_lag <- or_disease_lag_df %>%
+  select(174, 175:181) # GWR-method
+
+or_disease_temp_lag <- or_disease_lag_df %>%
+  select(94, 151:157) # GWR-method
+
+lag_max <- 7
+n1 <- length(or_disease_geo_lag[,1])
+n2 <- length(or_disease_temp_lag[,1])
+
+or_disease_geo_lag <- as.matrix(or_disease_geo_lag)
+or_disease_temp_lag <- as.matrix(or_disease_temp_lag)
+
+B <- ns(0:lag_max, df =5, intercept=TRUE)
+
+# the nrow B should equal ncol(pm)
+# dim(B)
+
+or_disease_geo_lag_B <- or_disease_geo_lag%*%B
+or_disease_temp_lag_B <- or_disease_temp_lag%*%B
+
+fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_temp_lag_B + strata(personkey), 
+              or_disease_lag_df)
+
+# coef(fit)
+# summary(fit)
+# AIC(fit)
+
+D <- cbind(B,B)
+
+# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
+dlparms <- c(1,2,3,4,5,6,7,8,9,10)
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms]) 
+DLvar <- D%*%vcov(fit)[dlparms,dlparms]%*%t(D)
+DLestimate$SE <- sqrt(diag(DLvar))
+DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
+DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
+
+DLestimate$estimate <- exp(DLestimate$estimate)
+DLestimate$lower <- exp(DLestimate$lower)
+DLestimate$upper <- exp(DLestimate$upper)
+
+# DLestimate
+
+DLestimate$lag <- 0:lag_max
+p5 <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper)) + 
+  geom_ribbon(alpha=.5) + geom_line(size=2) + 
+  geom_line(aes(x=lag), linetype=2, size=2) + 
+  theme_bw() + 
+  xlab("lagged days of events") + 
+  ylab("OR for 10ug/m^3 Increase") +
+  ggtitle("COPD") +
+  geom_hline(yintercept=1, linetype="dashed", color = "red")
+
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms])
+
+cumulative <- sum(DLestimate$estimate)
+# cumulative
+
+cumulative_se <- sqrt(sum(DLvar))
+# cumulative_se
+
+cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
+# cumulative_CI
+
+cum <- cbind(cumulative, cumulative_se, cumulative_CI[1], cumulative_CI[2])
+colnames(cum) <- c("Cumulative Est", "Cumulative SE", "Lower 95% CI", "Upper 95% CI")
+
+knitr::kable(cum, caption = paste0("Distributed lag 0-7 spline estimates, se and CI"))
+
+```
+
+```{r DLM for disease cvd, echo = FALSE, message=FALSE, warning=FALSE}
+
+# set variable list
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+
+  j <- c('cvd')
+
+  # standard for statement without parallel computing
+  #for(j in 1:length(var_list)){
+  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
+  or_disease <- read_csv(read_path)
+
+  or_disease_df <- data.frame(or_disease)
+
+  or_disease_lag_df <- or_disease_df %>%
+  left_join(or_zip_to_county, by = "ZIPCODE") %>%
+  left_join(or_pop, by = "county") %>%
+  mutate(day = ifelse(weekdays(date) == "Monday" |
+                      weekdays(date) == "Tuesday" |
+                      weekdays(date) == "Wednesday"|
+                      weekdays(date) == "Thursday"|
+                      weekdays(date) == "Friday", "weekday",
+               ifelse(weekdays(date) == "Saturday" |
+                      weekdays(date) == "Sunday", "weekend", NA))) %>%
+  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
+         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
+         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
+         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
+         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
+         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
+         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
+         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)
+
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-05-27")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-07-04")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-09-02")] <- "weekend"
+
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
+# which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
+# which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151
+
+or_disease_geo_lag <- or_disease_lag_df %>%
+  select(174, 175:181) # GWR-method
+
+or_disease_temp_lag <- or_disease_lag_df %>%
+  select(94, 151:157) # GWR-method
+
+lag_max <- 7
+n1 <- length(or_disease_geo_lag[,1])
+n2 <- length(or_disease_temp_lag[,1])
+
+or_disease_geo_lag <- as.matrix(or_disease_geo_lag)
+or_disease_temp_lag <- as.matrix(or_disease_temp_lag)
+
+B <- ns(0:lag_max, df =5, intercept=TRUE)
+
+# the nrow B should equal ncol(pm)
+# dim(B)
+
+or_disease_geo_lag_B <- or_disease_geo_lag%*%B
+or_disease_temp_lag_B <- or_disease_temp_lag%*%B
+
+fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_temp_lag_B + strata(personkey), 
+              or_disease_lag_df)
+
+# coef(fit)
+# summary(fit)
+# AIC(fit)
+
+D <- cbind(B,B)
+
+# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
+dlparms <- c(1,2,3,4,5,6,7,8,9,10)
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms]) 
+DLvar <- D%*%vcov(fit)[dlparms,dlparms]%*%t(D)
+DLestimate$SE <- sqrt(diag(DLvar))
+DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
+DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
+
+DLestimate$estimate <- exp(DLestimate$estimate)
+DLestimate$lower <- exp(DLestimate$lower)
+DLestimate$upper <- exp(DLestimate$upper)
+
+# DLestimate
+
+DLestimate$lag <- 0:lag_max
+p6 <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper)) + 
+  geom_ribbon(alpha=.5) + geom_line(size=2) + 
+  geom_line(aes(x=lag), linetype=2, size=2) + 
+  theme_bw() + 
+  xlab("lagged days of events") + 
+  ylab("OR for 10ug/m^3 Increase") +
+  ggtitle("cvd") +
+  geom_hline(yintercept=1, linetype="dashed", color = "red")
+
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms])
+
+cumulative <- sum(DLestimate$estimate)
+# cumulative
+
+cumulative_se <- sqrt(sum(DLvar))
+# cumulative_se
+
+cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
+# cumulative_CI
+
+cum <- cbind(cumulative, cumulative_se, cumulative_CI[1], cumulative_CI[2])
+colnames(cum) <- c("Cumulative Est", "Cumulative SE", "Lower 95% CI", "Upper 95% CI")
+
+knitr::kable(cum, caption = paste0("Distributed lag 0-7 spline estimates, se and CI"))
+
+```
+
+```{r DLM for disease isch_heart_dis, echo = FALSE, message=FALSE, warning=FALSE}
+
+# set variable list
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+
+  j <- c('isch_heart_dis')
+
+  # standard for statement without parallel computing
+  #for(j in 1:length(var_list)){
+  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
+  or_disease <- read_csv(read_path)
+
+  or_disease_df <- data.frame(or_disease)
+
+  or_disease_lag_df <- or_disease_df %>%
+  left_join(or_zip_to_county, by = "ZIPCODE") %>%
+  left_join(or_pop, by = "county") %>%
+  mutate(day = ifelse(weekdays(date) == "Monday" |
+                      weekdays(date) == "Tuesday" |
+                      weekdays(date) == "Wednesday"|
+                      weekdays(date) == "Thursday"|
+                      weekdays(date) == "Friday", "weekday",
+               ifelse(weekdays(date) == "Saturday" |
+                      weekdays(date) == "Sunday", "weekend", NA))) %>%
+  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
+         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
+         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
+         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
+         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
+         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
+         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
+         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)
+
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-05-27")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-07-04")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-09-02")] <- "weekend"
+
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
+# which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
+# which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151
+
+or_disease_geo_lag <- or_disease_lag_df %>%
+  select(174, 175:181) # GWR-method
+
+or_disease_temp_lag <- or_disease_lag_df %>%
+  select(94, 151:157) # GWR-method
+
+lag_max <- 7
+n1 <- length(or_disease_geo_lag[,1])
+n2 <- length(or_disease_temp_lag[,1])
+
+or_disease_geo_lag <- as.matrix(or_disease_geo_lag)
+or_disease_temp_lag <- as.matrix(or_disease_temp_lag)
+
+B <- ns(0:lag_max, df =5, intercept=TRUE)
+
+# the nrow B should equal ncol(pm)
+# dim(B)
+
+or_disease_geo_lag_B <- or_disease_geo_lag%*%B
+or_disease_temp_lag_B <- or_disease_temp_lag%*%B
+
+fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_temp_lag_B + strata(personkey), 
+              or_disease_lag_df)
+
+# coef(fit)
+# summary(fit)
+# AIC(fit)
+
+D <- cbind(B,B)
+
+# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
+dlparms <- c(1,2,3,4,5,6,7,8,9,10)
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms]) 
+DLvar <- D%*%vcov(fit)[dlparms,dlparms]%*%t(D)
+DLestimate$SE <- sqrt(diag(DLvar))
+DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
+DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
+
+DLestimate$estimate <- exp(DLestimate$estimate)
+DLestimate$lower <- exp(DLestimate$lower)
+DLestimate$upper <- exp(DLestimate$upper)
+
+# DLestimate
+
+DLestimate$lag <- 0:lag_max
+p7 <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper)) + 
+  geom_ribbon(alpha=.5) + geom_line(size=2) + 
+  geom_line(aes(x=lag), linetype=2, size=2) + 
+  theme_bw() + 
+  xlab("lagged days of events") + 
+  ylab("OR for 10ug/m^3 Increase") +
+  ggtitle("Isch heart disease") +
+  geom_hline(yintercept=1, linetype="dashed", color = "red")
+
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms])
+
+cumulative <- sum(DLestimate$estimate)
+# cumulative
+
+cumulative_se <- sqrt(sum(DLvar))
+# cumulative_se
+
+cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
+# cumulative_CI
+
+cum <- cbind(cumulative, cumulative_se, cumulative_CI[1], cumulative_CI[2])
+colnames(cum) <- c("Cumulative Est", "Cumulative SE", "Lower 95% CI", "Upper 95% CI")
+
+knitr::kable(cum, caption = paste0("Distributed lag 0-7 spline estimates, se and CI"))
+
+```
+
+```{r DLM for disease arrhythmia, echo = FALSE, message=FALSE, warning=FALSE}
+
+# set variable list
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+
+  j <- c('arrhythmia')
+
+  # standard for statement without parallel computing
+  #for(j in 1:length(var_list)){
+  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
+  or_disease <- read_csv(read_path)
+
+  or_disease_df <- data.frame(or_disease)
+
+  or_disease_lag_df <- or_disease_df %>%
+  left_join(or_zip_to_county, by = "ZIPCODE") %>%
+  left_join(or_pop, by = "county") %>%
+  mutate(day = ifelse(weekdays(date) == "Monday" |
+                      weekdays(date) == "Tuesday" |
+                      weekdays(date) == "Wednesday"|
+                      weekdays(date) == "Thursday"|
+                      weekdays(date) == "Friday", "weekday",
+               ifelse(weekdays(date) == "Saturday" |
+                      weekdays(date) == "Sunday", "weekend", NA))) %>%
+  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
+         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
+         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
+         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
+         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
+         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
+         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
+         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)
+
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-05-27")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-07-04")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-09-02")] <- "weekend"
+
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
+# which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
+# which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151
+
+or_disease_geo_lag <- or_disease_lag_df %>%
+  select(174, 175:181) # GWR-method
+
+or_disease_temp_lag <- or_disease_lag_df %>%
+  select(94, 151:157) # GWR-method
+
+lag_max <- 7
+n1 <- length(or_disease_geo_lag[,1])
+n2 <- length(or_disease_temp_lag[,1])
+
+or_disease_geo_lag <- as.matrix(or_disease_geo_lag)
+or_disease_temp_lag <- as.matrix(or_disease_temp_lag)
+
+B <- ns(0:lag_max, df =5, intercept=TRUE)
+
+# the nrow B should equal ncol(pm)
+# dim(B)
+
+or_disease_geo_lag_B <- or_disease_geo_lag%*%B
+or_disease_temp_lag_B <- or_disease_temp_lag%*%B
+
+fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_temp_lag_B + strata(personkey), 
+              or_disease_lag_df)
+
+# coef(fit)
+# summary(fit)
+# AIC(fit)
+
+D <- cbind(B,B)
+
+# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
+dlparms <- c(1,2,3,4,5,6,7,8,9,10)
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms]) 
+DLvar <- D%*%vcov(fit)[dlparms,dlparms]%*%t(D)
+DLestimate$SE <- sqrt(diag(DLvar))
+DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
+DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
+
+DLestimate$estimate <- exp(DLestimate$estimate)
+DLestimate$lower <- exp(DLestimate$lower)
+DLestimate$upper <- exp(DLestimate$upper)
+
+# DLestimate
+
+DLestimate$lag <- 0:lag_max
+p8 <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper)) + 
+  geom_ribbon(alpha=.5) + geom_line(size=2) + 
+  geom_line(aes(x=lag), linetype=2, size=2) + 
+  theme_bw() + 
+  xlab("lagged days of events") + 
+  ylab("Odds Ratio for 10ug/m^3 Increase in PM2.5") +
+  ggtitle("Arrhythmia") +
+  geom_hline(yintercept=1, linetype="dashed", color = "red")
+
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms])
+
+cumulative <- sum(DLestimate$estimate)
+# cumulative
+
+cumulative_se <- sqrt(sum(DLvar))
+# cumulative_se
+
+cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
+# cumulative_CI
+
+cum <- cbind(cumulative, cumulative_se, cumulative_CI[1], cumulative_CI[2])
+colnames(cum) <- c("Cumulative Est", "Cumulative SE", "Lower 95% CI", "Upper 95% CI")
+
+knitr::kable(cum, caption = paste0("Distributed lag 0-7 spline estimates, se and CI"))
+
+```
+
+```{r DLM for disease heart_failure, echo = FALSE, message=FALSE, warning=FALSE}
+
+# set variable list
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+
+  j <- c('heart_failure')
+
+  # standard for statement without parallel computing
+  #for(j in 1:length(var_list)){
+  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
+  or_disease <- read_csv(read_path)
+
+  or_disease_df <- data.frame(or_disease)
+
+  or_disease_lag_df <- or_disease_df %>%
+  left_join(or_zip_to_county, by = "ZIPCODE") %>%
+  left_join(or_pop, by = "county") %>%
+  mutate(day = ifelse(weekdays(date) == "Monday" |
+                      weekdays(date) == "Tuesday" |
+                      weekdays(date) == "Wednesday"|
+                      weekdays(date) == "Thursday"|
+                      weekdays(date) == "Friday", "weekday",
+               ifelse(weekdays(date) == "Saturday" |
+                      weekdays(date) == "Sunday", "weekend", NA))) %>%
+  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
+         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
+         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
+         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
+         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
+         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
+         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
+         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)
+
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-05-27")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-07-04")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-09-02")] <- "weekend"
+
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
+# which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
+# which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151
+
+or_disease_geo_lag <- or_disease_lag_df %>%
+  select(174, 175:181) # GWR-method
+
+or_disease_temp_lag <- or_disease_lag_df %>%
+  select(94, 151:157) # GWR-method
+
+lag_max <- 7
+n1 <- length(or_disease_geo_lag[,1])
+n2 <- length(or_disease_temp_lag[,1])
+
+or_disease_geo_lag <- as.matrix(or_disease_geo_lag)
+or_disease_temp_lag <- as.matrix(or_disease_temp_lag)
+
+B <- ns(0:lag_max, df =5, intercept=TRUE)
+
+# the nrow B should equal ncol(pm)
+# dim(B)
+
+or_disease_geo_lag_B <- or_disease_geo_lag%*%B
+or_disease_temp_lag_B <- or_disease_temp_lag%*%B
+
+fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_temp_lag_B + strata(personkey), 
+              or_disease_lag_df)
+
+# coef(fit)
+# summary(fit)
+# AIC(fit)
+
+D <- cbind(B,B)
+
+# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
+dlparms <- c(1,2,3,4,5,6,7,8,9,10)
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms]) 
+DLvar <- D%*%vcov(fit)[dlparms,dlparms]%*%t(D)
+DLestimate$SE <- sqrt(diag(DLvar))
+DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
+DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
+
+DLestimate$estimate <- exp(DLestimate$estimate)
+DLestimate$lower <- exp(DLestimate$lower)
+DLestimate$upper <- exp(DLestimate$upper)
+
+# DLestimate
+
+DLestimate$lag <- 0:lag_max
+p9 <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper)) + 
+  geom_ribbon(alpha=.5) + geom_line(size=2) + 
+  geom_line(aes(x=lag), linetype=2, size=2) + 
+  theme_bw() + 
+  xlab("lagged days of events") + 
+  ylab("OR for 10ug/m^3 Increase") +
+  ggtitle("Heart failure") +
+  geom_hline(yintercept=1, linetype="dashed", color = "red")
+
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms])
+
+cumulative <- sum(DLestimate$estimate)
+# cumulative
+
+cumulative_se <- sqrt(sum(DLvar))
+# cumulative_se
+
+cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
+# cumulative_CI
+
+cum <- cbind(cumulative, cumulative_se, cumulative_CI[1], cumulative_CI[2])
+colnames(cum) <- c("Cumulative Est", "Cumulative SE", "Lower 95% CI", "Upper 95% CI")
+
+knitr::kable(cum, caption = paste0("Distributed lag 0-7 spline estimates, se and CI"))
+
+```
+
+```{r DLM for disease cerbrovas_dis, echo = FALSE, message=FALSE, warning=FALSE}
+
+# set variable list
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+
+  j <- c('cerbrovas_dis')
+
+  # standard for statement without parallel computing
+  #for(j in 1:length(var_list)){
+  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
+  or_disease <- read_csv(read_path)
+
+  or_disease_df <- data.frame(or_disease)
+
+  or_disease_lag_df <- or_disease_df %>%
+  left_join(or_zip_to_county, by = "ZIPCODE") %>%
+  left_join(or_pop, by = "county") %>%
+  mutate(day = ifelse(weekdays(date) == "Monday" |
+                      weekdays(date) == "Tuesday" |
+                      weekdays(date) == "Wednesday"|
+                      weekdays(date) == "Thursday"|
+                      weekdays(date) == "Friday", "weekday",
+               ifelse(weekdays(date) == "Saturday" |
+                      weekdays(date) == "Sunday", "weekend", NA))) %>%
+  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
+         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
+         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
+         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
+         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
+         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
+         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
+         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)
+
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-05-27")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-07-04")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-09-02")] <- "weekend"
+
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
+# which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
+# which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151
+
+or_disease_geo_lag <- or_disease_lag_df %>%
+  select(174, 175:181) # GWR-method
+
+or_disease_temp_lag <- or_disease_lag_df %>%
+  select(94, 151:157) # GWR-method
+
+lag_max <- 7
+n1 <- length(or_disease_geo_lag[,1])
+n2 <- length(or_disease_temp_lag[,1])
+
+or_disease_geo_lag <- as.matrix(or_disease_geo_lag)
+or_disease_temp_lag <- as.matrix(or_disease_temp_lag)
+
+B <- ns(0:lag_max, df =5, intercept=TRUE)
+
+# the nrow B should equal ncol(pm)
+# dim(B)
+
+or_disease_geo_lag_B <- or_disease_geo_lag%*%B
+or_disease_temp_lag_B <- or_disease_temp_lag%*%B
+
+fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_temp_lag_B + strata(personkey), 
+              or_disease_lag_df)
+
+# coef(fit)
+# summary(fit)
+# AIC(fit)
+
+D <- cbind(B,B)
+
+# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
+dlparms <- c(1,2,3,4,5,6,7,8,9,10)
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms]) 
+DLvar <- D%*%vcov(fit)[dlparms,dlparms]%*%t(D)
+DLestimate$SE <- sqrt(diag(DLvar))
+DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
+DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
+
+DLestimate$estimate <- exp(DLestimate$estimate)
+DLestimate$lower <- exp(DLestimate$lower)
+DLestimate$upper <- exp(DLestimate$upper)
+
+# DLestimate
+
+DLestimate$lag <- 0:lag_max
+p10 <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper)) + 
+  geom_ribbon(alpha=.5) + geom_line(size=2) + 
+  geom_line(aes(x=lag), linetype=2, size=2) + 
+  theme_bw() + 
+  xlab("lagged days of events") + 
+  ylab("OR for 10ug/m^3 Increase") +
+  ggtitle("Cerbrovas disease") +
+  geom_hline(yintercept=1, linetype="dashed", color = "red")
+
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms])
+
+cumulative <- sum(DLestimate$estimate)
+# cumulative
+
+cumulative_se <- sqrt(sum(DLvar))
+# cumulative_se
+
+cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
+# cumulative_CI
+
+cum <- cbind(cumulative, cumulative_se, cumulative_CI[1], cumulative_CI[2])
+colnames(cum) <- c("Cumulative Est", "Cumulative SE", "Lower 95% CI", "Upper 95% CI")
+
+knitr::kable(cum, caption = paste0("Distributed lag 0-7 spline estimates, se and CI"))
+
+```
+
+```{r DLM for disease myocardial_infarc, echo = FALSE, message=FALSE, warning=FALSE}
+
+# set variable list
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+
+  j <- c('myocardial_infarc')
+
+  # standard for statement without parallel computing
+  #for(j in 1:length(var_list)){
+  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
+  or_disease <- read_csv(read_path)
+
+  or_disease_df <- data.frame(or_disease)
+
+  or_disease_lag_df <- or_disease_df %>%
+  left_join(or_zip_to_county, by = "ZIPCODE") %>%
+  left_join(or_pop, by = "county") %>%
+  mutate(day = ifelse(weekdays(date) == "Monday" |
+                      weekdays(date) == "Tuesday" |
+                      weekdays(date) == "Wednesday"|
+                      weekdays(date) == "Thursday"|
+                      weekdays(date) == "Friday", "weekday",
+               ifelse(weekdays(date) == "Saturday" |
+                      weekdays(date) == "Sunday", "weekend", NA))) %>%
+  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
+         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
+         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
+         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
+         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
+         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
+         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
+         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)
+
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-05-27")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-07-04")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-09-02")] <- "weekend"
+
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
+# which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
+# which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151
+
+or_disease_geo_lag <- or_disease_lag_df %>%
+  select(174, 175:181) # GWR-method
+
+or_disease_temp_lag <- or_disease_lag_df %>%
+  select(94, 151:157) # GWR-method
+
+lag_max <- 7
+n1 <- length(or_disease_geo_lag[,1])
+n2 <- length(or_disease_temp_lag[,1])
+
+or_disease_geo_lag <- as.matrix(or_disease_geo_lag)
+or_disease_temp_lag <- as.matrix(or_disease_temp_lag)
+
+B <- ns(0:lag_max, df =5, intercept=TRUE)
+
+# the nrow B should equal ncol(pm)
+# dim(B)
+
+or_disease_geo_lag_B <- or_disease_geo_lag%*%B
+or_disease_temp_lag_B <- or_disease_temp_lag%*%B
+
+fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_temp_lag_B + strata(personkey), 
+              or_disease_lag_df)
+
+# coef(fit)
+# summary(fit)
+# AIC(fit)
+
+D <- cbind(B,B)
+
+# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
+dlparms <- c(1,2,3,4,5,6,7,8,9,10)
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms]) 
+DLvar <- D%*%vcov(fit)[dlparms,dlparms]%*%t(D)
+DLestimate$SE <- sqrt(diag(DLvar))
+DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
+DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
+
+DLestimate$estimate <- exp(DLestimate$estimate)
+DLestimate$lower <- exp(DLestimate$lower)
+DLestimate$upper <- exp(DLestimate$upper)
+
+# DLestimate
+
+DLestimate$lag <- 0:lag_max
+p11 <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper)) + 
+  geom_ribbon(alpha=.5) + geom_line(size=2) + 
+  geom_line(aes(x=lag), linetype=2, size=2) + 
+  theme_bw() + 
+  xlab("lagged days of events") + 
+  ylab("OR for 10ug/m^3 Increase") +
+  ggtitle("Myocardial infarc") +
+  geom_hline(yintercept=1, linetype="dashed", color = "red")
+
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms])
+
+cumulative <- sum(DLestimate$estimate)
+# cumulative
+
+cumulative_se <- sqrt(sum(DLvar))
+# cumulative_se
+
+cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
+# cumulative_CI
+
+cum <- cbind(cumulative, cumulative_se, cumulative_CI[1], cumulative_CI[2])
+colnames(cum) <- c("Cumulative Est", "Cumulative SE", "Lower 95% CI", "Upper 95% CI")
+
+knitr::kable(cum, caption = paste0("Distributed lag 0-7 spline estimates, se and CI"))
+
+```
+
+```{r DLM for disease broken_arm, echo = FALSE, message=FALSE, warning=FALSE}
+
+# set variable list
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+
+  j <- c('broken_arm')
+
+  # standard for statement without parallel computing
+  #for(j in 1:length(var_list)){
+  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
+  or_disease <- read_csv(read_path)
+
+  or_disease_df <- data.frame(or_disease)
+
+  or_disease_lag_df <- or_disease_df %>%
+  left_join(or_zip_to_county, by = "ZIPCODE") %>%
+  left_join(or_pop, by = "county") %>%
+  mutate(day = ifelse(weekdays(date) == "Monday" |
+                      weekdays(date) == "Tuesday" |
+                      weekdays(date) == "Wednesday"|
+                      weekdays(date) == "Thursday"|
+                      weekdays(date) == "Friday", "weekday",
+               ifelse(weekdays(date) == "Saturday" |
+                      weekdays(date) == "Sunday", "weekend", NA))) %>%
+  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
+         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
+         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
+         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
+         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
+         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
+         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
+         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)
+
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-05-27")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-07-04")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-09-02")] <- "weekend"
+
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
+# which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
+# which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151
+
+or_disease_geo_lag <- or_disease_lag_df %>%
+  select(174, 175:181) # GWR-method
+
+or_disease_temp_lag <- or_disease_lag_df %>%
+  select(94, 151:157) # GWR-method
+
+lag_max <- 7
+n1 <- length(or_disease_geo_lag[,1])
+n2 <- length(or_disease_temp_lag[,1])
+
+or_disease_geo_lag <- as.matrix(or_disease_geo_lag)
+or_disease_temp_lag <- as.matrix(or_disease_temp_lag)
+
+B <- ns(0:lag_max, df =5, intercept=TRUE)
+
+# the nrow B should equal ncol(pm)
+# dim(B)
+
+or_disease_geo_lag_B <- or_disease_geo_lag%*%B
+or_disease_temp_lag_B <- or_disease_temp_lag%*%B
+
+fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_temp_lag_B + strata(personkey), 
+              or_disease_lag_df)
+
+# coef(fit)
+# summary(fit)
+# AIC(fit)
+
+D <- cbind(B,B)
+
+# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
+dlparms <- c(1,2,3,4,5,6,7,8,9,10)
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms]) 
+DLvar <- D%*%vcov(fit)[dlparms,dlparms]%*%t(D)
+DLestimate$SE <- sqrt(diag(DLvar))
+DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
+DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
+
+DLestimate$estimate <- exp(DLestimate$estimate)
+DLestimate$lower <- exp(DLestimate$lower)
+DLestimate$upper <- exp(DLestimate$upper)
+
+# DLestimate
+
+DLestimate$lag <- 0:lag_max
+p12 <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper)) + 
+  geom_ribbon(alpha=.5) + geom_line(size=2) + 
+  geom_line(aes(x=lag), linetype=2, size=2) + 
+  theme_bw() + 
+  xlab("lagged days of events") + 
+  ylab("OR for 10ug/m^3 Increase") +
+  ggtitle("Broden arm") +
+  geom_hline(yintercept=1, linetype="dashed", color = "red")
+
+DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms])
+
+cumulative <- sum(DLestimate$estimate)
+# cumulative
+
+cumulative_se <- sqrt(sum(DLvar))
+# cumulative_se
+
+cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
+# cumulative_CI
+
+cum <- cbind(cumulative, cumulative_se, cumulative_CI[1], cumulative_CI[2])
+colnames(cum) <- c("Cumulative Est", "Cumulative SE", "Lower 95% CI", "Upper 95% CI")
+
+knitr::kable(cum, caption = paste0("Distributed lag 0-7 spline estimates, se and CI"))
+
+```
+
+```{r}
+plot_grid(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, ncol=4)
 
@@
 Then we make the reversed lag to check with the same df of 5. The plot shows the same result as previous analysis.
+
+```{r DLM for casecrossover temp, echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}
+read_path8 <- paste0("../data_new/county_data/zip_to_county_new.csv")
+or_zip_to_county <- read_csv(read_path8)
+
+read_path9 <- paste0("../data_new/county_data/or_census_pop_est_county.csv")
+or_pop <- read_csv(read_path9)
+
+or_pop <- or_pop %>%
+  rename(county = GEO_display_label, pop = respop72013) %>%
+  select(county, pop) 
+or_pop$county <- gsub(" County, Oregon", "", as.character(factor(or_pop$county)))
+
+inhaler_ndc_lag_df <- inhaler_ndc_df %>%
+  left_join(or_zip_to_county, by = "ZIPCODE") %>%
+  left_join(or_pop, by = "county") %>%
+  mutate(day = ifelse(weekdays(date) == "Monday" |
+                      weekdays(date) == "Tuesday" |
+                      weekdays(date) == "Wednesday"|
+                      weekdays(date) == "Thursday"|
+                      weekdays(date) == "Friday", "weekday",
+               ifelse(weekdays(date) == "Saturday" |
+                      weekdays(date) == "Sunday", "weekend", NA))) %>%
+  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
+         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
+         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
+         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
+         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
+         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
+         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
+         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)
+
+inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-05-27")] <- "weekend"
+inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-07-04")] <- "weekend"
+inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-09-02")] <- "weekend"
+
+which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_zip") # 159
+which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_lag1_zip") # 160
+which(colnames(inhaler_ndc_lag_df)=="wrf_temp_zip") # 93
+which(colnames(inhaler_ndc_lag_df)=="wrf_temp_lag1_zip") # 150
+
+inhaler_geo_lag <- inhaler_ndc_lag_df %>%
+  select(159, 160:166) # GWR-method
+
+inhaler_temp_lag <- inhaler_ndc_lag_df %>%
+  select(93, 150:156) # GWR-method
+
+lag_max <- 7
+n1 <- length(inhaler_geo_lag[,1])
+n2 <- length(inhaler_temp_lag[,1])
+
+inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
+inhaler_temp_lag <- as.matrix(inhaler_temp_lag)
+
+B <- ns(0:lag_max, df =3, intercept=TRUE)
+
+# the nrow B should equal ncol(pm)
+dim(B)
+
+inhaler_geo_lag_B <- inhaler_geo_lag%*%B
+inhaler_temp_lag_B <- inhaler_temp_lag%*%B
+
+fit <- clogit(outcome ~  inhaler_temp_lag_B + strata(personkey), 
+              inhaler_ndc_lag_df)
+
+coef(fit)
+summary(fit)
+AIC(fit)
+
+D <- cbind(B,B)
+
+# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
+dlparms <- c(1,2,3)
+DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms]) 
+DLvar <- B%*%vcov(fit)[dlparms,dlparms]%*%t(B)
+DLestimate$SE <- sqrt(diag(DLvar))
+DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
+DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
+
+DLestimate$estimate <- exp(DLestimate$estimate)
+DLestimate$lower <- exp(DLestimate$lower)
+DLestimate$upper <- exp(DLestimate$upper)
+
+DLestimate
+
+DLestimate$lag <- 0:lag_max
+p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
+p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
+p <- p + geom_line(aes(x=lag), linetype=2, size=2)
+p <- p + theme_bw()
+p
+
+DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms])
+
+cumulative <- sum(DLestimate$estimate)
+cumulative
+
+cumulative_se <- sqrt(sum(DLvar))
+cumulative_se
+
+cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
+cumulative_CI
+
+```
 
@@
 
-```{r DLM ts, echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}
+```{r DLM ts, echo = FALSE, message=FALSE, warning=FALSE, echo=FALSE}
+or_ndc_county_ts_df <- or_ndc_county_ts %>%
+  select(1:4, pop, rate_per_100k, day, geo_smk_pm_county,
+         geo_smk_pm_lag1_county:geo_smk_pm_lag7_county, wrf_temp_county,
+         wrf_temp_lag1_county:wrf_temp_lag7_county) %>%
+  mutate(geo_smk_pm10_county = geo_smk_pm_county/10,
+         geo_smk_pm10_lag1_county =geo_smk_pm_lag1_county/10,
+         geo_smk_pm10_lag2_county =geo_smk_pm_lag2_county/10,
+         geo_smk_pm10_lag3_county =geo_smk_pm_lag3_county/10,
+         geo_smk_pm10_lag4_county =geo_smk_pm_lag4_county/10,
+         geo_smk_pm10_lag5_county =geo_smk_pm_lag5_county/10,
+         geo_smk_pm10_lag6_county =geo_smk_pm_lag6_county/10,
+         geo_smk_pm10_lag7_county =geo_smk_pm_lag7_county/10)
+
 which(colnames(or_ndc_county_ts_df)=="wrf_temp_lag1_county") # 17

### 2017-09-20 r_scripts/analysis/oregon_medication.Rmd
Update oregon_medication.Rmd

@@
 **1. Introduction: **
-We found an association with wildfire smoke and respiratory outcomes in Washington state 2012 using a novel estimate of smoke concentration, geo-graphically weighted ridge regression (GWR).
+We already found an association with wildfire smoke and respiratory outcomes in Washington state in 2012 wildfire season using a novel estimate of smoke concentration, geo-graphically weighted ridge regression (GWR).
 We want to know if smoke impacts other health outcomes, specifically respiratory rescue medications (beta 2 agonists) (abbreviate to SABA)
+
+For this research we have the Oregon All Payer All Claims (APAC) data set in 2013.
+We want to research on the effects of wildfire smoke on beta-2 agonist pharmacy fill, a medication primarily used as a rescue medication for persons with asthma or COPD.
+NDC (National Drug Code) is the variable in the data set that the data set offered.
+The standard NDC I use is shown in the link below, which is a ndc list in 2013 and fits the original data set after checking.
+(http://ww2.ncqa.org/hedis-quality-measurement/hedis-measures/hedis-2013/hedis-2013-final-ndc-lists)
 
@@
 
-For this research I have the Oregon All Payer All Claims (APAC) data set in 2013.
-I want to research on the effects of wildfire GWR smoke on beta-2 agonist pharmacy fill, a medication primarily used as a rescue medication for persons with asthma or COPD.
-NDC (National Drug Code) is the variable in the data set that the data set offered.
-The standard NDC I use the link below, which is a ndc list in 2013 and fits my original data set.
-(http://ww2.ncqa.org/hedis-quality-measurement/hedis-measures/hedis-2013/hedis-2013-final-ndc-lists)
+**3.Methods: **
 
 The data cleaning and creation of the casecross-over dataframes are huge computation tasks, so they are done with server, the following work is done with local R.
-We analyze the case crossover results with overall, age strarified and sex stratified categories for the pharmacy fills, and then we also analyze with the distributed lag method, and check with time series analysis.
-
-**3.Methods: **
-use case-crossover for all health outcome figures and results. Make sure smoke is assigned via zipcode. For the descriptive of smoke, I think the county-level for now. We may move to zip.
-
-In these comparisons, we examine GWR method of smoke PM2.5 estimations and associations with health outcomes using a time-stratified case-crossover study design.
-Pharmacy outcomes for a patient with a primary diagnosis of all pharmacy records from the state, and it just so happens we also have APAC data for people that were billed for a cardiopulmonary event that we can link to the pharmacy claim, and their date of admission (index time) were identified.
-We then created counterfactual observations for each patient on the same day of the week for the entire wildfire season (May 1st to Sep 30th, 2013).
-For patients who have a index time before May 1, 2013 or after September 30, 2013, their referent observations before or after these dates will be excluded as I will not be able to assign estimates of PM2.5 to these referent observations.
+We analyze the case crossover results with overall pharmacy fills, and analyze the results with age strarified and sex stratified categories for the pharmacy fills in supplement.
+We also analyze with the distributed lag method, and check with time series analysis.
+
+In these comparisons, we examine GWR method of smoke PM2.5 estimations and associations with medication outcomes using a time-stratified case-crossover study design.
+Pharmacy outcomes for a patient with a primary diagnosis of all pharmacy records from Oregon state are recorded, and we also have APAC data for people that were billed for a cardiopulmonary event that we can link to the pharmacy claims, in which their dates of admission (index time) are also identified.
+We then created counterfactual observations for each patient on the same day of the week during the entire wildfire season (May 1st to Sep 30th, 2013).
+For patients who have a index time before May 1, 2013 or after September 30, 2013, their referent observations before or after these dates will be excluded because we are not able to assign estimates of PM2.5 to these observations.
 
@@
 There are about 77 million claims in the total data set, of which 23 million are pharmacy records denoted by the NDC variable.
-So I choose the claims in Oregon State with ndc existing, which is about 23 million claims.
-Also, the POS (Place of service code) all show "pharmacy" type, comparing to the previous health disease outcomes we limit the "emergency visit" type or "urgent care" type.
-
-For ndc, I choose the people the person was billed for filling a beta2 agonist prescription, which we use as a proxy for use to form the sample.
-Because the ndc csv file has more information than what I need, I filter the ndc file by SQLite Studio (code and steps not attached) to a file only containing the beta 2 agonists medication information.
+So we choose the claims in Oregon State with ndc existing, which is about 23 million claims.
+Also, the POS (Place of service code) all show "pharmacy" type, comparing to the previous health disease outcomes we limit the "emergency visit" type and the "urgent care" type.
+
+For ndc, we choose the people who were billed for filling a beta2 agonist prescription, which we use as a proxy for use to form the sample.
+Because the ndc csv file has more information than what we need, we filter the ndc file by SQLite Studio (code and steps not attached) to a file only containing the beta 2 agonists medication information.
 
 After choosing the claims with beta 2 agonists, the data set is reduced to 232,665 observations.
-After first recorded pharmacy fill and limit the wildfire season dates (May 1st, 2013 to September 30th, 2013), the data set is the final data set used for analysis and contained 24,538 beta2 agonist records.
+After first recorded pharmacy fill and limit the wildfire season dates (May 1st, 2013 to September 30th, 2013), the data set is the final data set for analysis, which contains 24,538 beta2 agonist records.
 
@@
 For our research question, we focus on the GWR method (geographically weighted ridge regression).
-For Geo-Weighted Regression model, with 'smk' designator, I subtracted off the 'Background' estimates of smoke, which I believe are the monthly averages of PM2.5 for a given grid.
-However, the 'Background' is the mean of total period from May 1st to Sep 30th 2013 so I have to assign the mean for every day.
+For Geo-Weighted Regression model, with 'smk' designator, we subtracted off the 'Background' estimates of smoke, which are the monthly averages of PM2.5 for a given grid.
+However, the 'Background' is the mean of total period from May 1st to Sep 30th 2013 so we have to assign the mean smoke estimates for every day.
 
@@
 
-We use the conditional logistic regression model using the survival package in R.
-Each conditional logistic regression model accounts for the subject, and adjusts for temperature from the WRF-Chem model.
+To analyze the result, we use the conditional logistic regression model using the survival package in R.
+Each conditional logistic regression model accounts for the subject, and adjusts for temperature from the GWR smoke model.
 The conditional logistic regression model \beta represents the change in risk of an pharmacy event associated with a short-term unit increase in exposure, and can be calculated as an average difference between exposure at the index time and a weighted average of exposure at all times in the referent window.
+(???)
 
@@
 
-### Time-Stratified Case-Crossover (Population-Weighted PM~2.5~ Assigned by Zip Code)
+Time-Stratified Case-Crossover (Population-Weighted PM~2.5~ Assigned by Zip Code)
 *referents same day of week within fire season May - September*
@@
 The time-stratified casecross-over design uses subjects as their own controls.
-For each pharmacy claim, I also created counterfactual observations on the the same day of the week over the study period for each subject to indicate the outcome for this day is "1".
+For each pharmacy claim, we also created counterfactual observations on the the same day of the week over the study period for each subject to indicate the outcome for this day is "1".
 Then I make the dates week by week before this date, and after this date in the whole wildfire season, and indicate these dates' outcomes are "0".
-Then I can use these outcomes for the following model building.
+Then I can use these outcomes for the following regression model building.
 
@@
 
-To get the time stratified data set, I create the lag variables that take smoke values from the previous days for zip codes, join the zipcode level populatoin-weighted pm to the data set, and create age, sex, month and season index.
+Also, to get the time stratified data set, I create the lag variables that take smoke values from the previous days for zip codes, joined with the zipcode level populatoin-weighted pm to the data set, and create age, sex, month and season index.
 
@@
 
-### Time stratified for time series (Population-Weighted PM~2.5~ Assigned by County)
+Time stratified for time series (Population-Weighted PM~2.5~ Assigned by County)
 
@@
 Therefore, I assign each zip code with one county which I found online.
-There are 488 zipcodes in total (4 are missing but I use google to find it).
-This is a sensitivity analysis to confirm the casecross over results.
+There are 488 zipcodes in total (4 are missing but I use google to find the relavent county).
+This is a sensitivity analysis to check and confirm the casecross over results.
 
@@
 
-### map of smoke days (4)
+### map of smoke days
+
+The map shows
 

### 2017-11-03 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
reworking results markdown file.

---
title: "Douglas-complex fire smoke and cardiopulmonary morbidity"
author: "Ryan Gan"
date: "11/3/2017"
output: html_document
---

##1. Introduction

In the summer of 2013, the Douglas-Complex fires occured in southwest Oregon. Oregonians in this part of the state were at risk of exposure to extreme levels of particulate matter (PM). This project aim is to determine if there is an association to smoke from the Douglas-Complex fires and acute cardiopulmonary morbidity in the state of Oregon.

Data came from the Oregon All Payer All Claims Database (APAC) in the year 2013. The APAC records the health care billing data for Oregon's insured populations. APAC include individual billing records for both diagnoses codes (International Classification of Diseases, Clinical Modification (ICD-9-CM) diagnoses codes) and pharmacy codes (National Drug Codes (NDC)).

Our previous research that found an association with wildfire smoke and respiratory outcomes in Washington state in 2012 wildfire season using a novel estimate of smoke concentration, geographically weighted ridge regression (GWR) guided the methodological approaches used in this project. As Oregon contains pharmacy records, we evaluate the association between smoke and respiratory rescue medications (beta 2 agonists) (abbreviate to SABA).

*Research question*
We evaluated the association between smoke concentrations using the GWR method and cardiopulmonary morbidity, including ED/urgent care visits and SABA fills in Oregon state during the 2013 wildfire season.

This markdown document contains the code and results that were used to address this research question.

**Note to Jingyang: I saw all the packages you used. How many do you actually need to run these analyses? It also looks like you need to run this markdown file on the server? I would like to avoid that and I can help you do this.**

Packages used:
```{r setup}
library(tidyverse)
```

### 2017-11-03 r_scripts/analysis/county_timeseries.Rmd

@@
 
-```{r }
+###  County-Level Time-Series for the Fire Season
+This time series analysis is for the Oregon county-level all emergency visit in All Payer All Claims data set.
+
+```{r, echo=FALSE, message=FALSE}
 library(tidyverse)
@@
 library(htmlTable)
+library(knitr)
+```
+
+First filter the date to only 2013 May to 2013 September.
+Rescale the pm variables to 10 units.
+
+```{r, message=FALSE, warning=FALSE}
 
@@
   # visits on that date (which is reasonable in sparsely populated counties)
-  mutate_each(funs(wo_miss = ifelse(is.na(.), 0, .)), n_obs:broken_arm_n) %>% 
+  # mutate_each(funs(wo_miss = ifelse(is.na(.), 0, .)), n_obs:broken_arm_n) %>% 
   # create a binary smoke indicator variable based smk variables >5, >10, and >15 units 
@@
   # rescale pm variables to 10 units
-  mutate_each(funs(unit10 = ./10), wrf_pm:krig_smk_pm)
+  mutate_each(funs(unit10 = ./10), wrf_pm:krig_smk_pm) %>%
+  select(-42, -43, -44)
   # select(1:2, 31:33, 36, 34:35, 38:44, 28:29, 30, 55, 58, 
-  #        65, 64, 63) # subset and order variables used in analysis
+  #        65, 64, 63) # subset and order variables used in analysis # 23 column
+  # wa: 12 pm; 15 disease
 
@@
 # looking only at wrf chem, kriging, and geo-weighted; taking out 'Global Smoke' method
-method_list <- c('Geo-Weighted Smoke', 'Kriging Smoke', 'WRF-Chem Smoke')
+method_list <- c('WRF-Chem Smoke', 'Geo-Weighted Smoke', 'Kriging Smoke')
 
@@
 for(i in 6:18){
-  i = 8
+
   # empty matrix for table
@@
   # loop through pm estimation methods
-  for(j in 25:27){
-    j = 25
-    var_name <- method_list[j-22]
+  for(j in 41:43){
+
+    var_name <- method_list[j-40]
 
@@
     # populate matrix
-    row_n <- j-22
+    row_n <- j-40
     # method
@@
   # combine previous values in dataframe that has all outcome/methods comparisons
-  datalist[[i]] <- point_est_df
+  datalist[[i-5]] <- point_est_df
 
@@
 print(tab)
+
+kable(combined_point_est_df)
+
+# ggplot of odds ratios, facet wrapped by outcomes -----
+# convert variables from character to either numeric or factor
+# factor preserves the order of the variable
+combined_point_est_df$outcome <- factor(combined_point_est_df$outcome, 
+                                        levels = unique(combined_point_est_df$outcome))
+
+combined_point_est_df$pm_method <- factor(combined_point_est_df$pm_method, 
+                                          levels = unique(combined_point_est_df$pm_method))
+
+combined_point_est_df$risk_ratio <- as.numeric(combined_point_est_df$risk_ratio)
+combined_point_est_df$lower95 <- as.numeric(combined_point_est_df$lower95)
+combined_point_est_df$upper95 <- as.numeric(combined_point_est_df$upper95)
+
+## ggplot
+print_plot <- ggplot(combined_point_est_df, 
+                     aes(x = pm_method, y = risk_ratio, colour = pm_method)) +
+  geom_point() + #geom_text(vjust = 0, nudge_x = 0.3) +
+  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
+  facet_wrap(~outcome, nrow = 3) +
+  geom_hline(yintercept = 1, linetype=2) +
+  ggtitle('County-Level Time-Series: Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
+  ylab('Risk Ratio for 10µg/m^3 Increase in PM2.5') +
+  xlab('Time-Stratified Within July to October Fire Season') +
+  scale_colour_discrete(name= "Smoke Method") +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        axis.text.x = element_blank(),
+        axis.title.y = element_text(angle = 90),
+        # facet text size
+        strip.text = element_text(size = 7),
+        legend.text = element_text(size = 7))
+
+print(print_plot)
+
+ggsave("../plot_new/or_county_time_series.pdf", plot = print_plot, 
+       width = 12, height = 8, units = "in")
+
+```

### 2017-11-03 r_scripts/analysis/or_dist_lag_model.Rmd

@@
     # remove missing pm values
-    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
-    filter(!is.na(krig_pm_lag1_zip)) %>% 
-    filter(!is.na(geo_wt_pm_lag1_zip)) %>%
+    filter(!is.na(wrf_smk_pm_lag2_zip)) %>% 
+    filter(!is.na(krig_pm_lag2_zip)) %>% 
+    filter(!is.na(geo_wt_pm_lag2_zip)) %>%
 
@@
 # save figure
-ggsave("../plot_lag/fig_lag1.pdf", plot = print_plot, 
+ggsave("../plot_lag/fig_lag2.pdf", plot = print_plot, 
        width = 12, height = 8, units = "in")
@@
 ```
-
-### Medication Distributed Lag

### 2017-11-03 r_scripts/analysis/or_result_zip_smoke.Rmd

@@
 
-```{r data import, include = F, eval = FALSE}
+```{r ER data import, include = F, eval = FALSE}
 
@@
 
+```{r ER and urgent data import, include = F, eval = FALSE}
+
+### Limit to ER visits and urgent care
+## pos = 23 or pos = 20
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+
+N <- NA
+n <- 1
+
+for (i in var_list){
+  read_path <- paste('../data_new/casecrossover/time_strat/or', i, 'may_to_sep_time_strat_casecross.csv', sep = "_")
+  or_disease <- read_csv(read_path)
+
+  or_disease <- or_disease %>% 
+    # filter the unknown gender
+    filter(gender!= "U")
+
+  # Create a permanent case-cross over dataset
+  file_name <- paste('../data_new/casecrossover/ER_Urgent/or', i, 'may_to_sep_time_strat_casecross_both.csv', sep = "_")
+
+  # write permanent dataset
+  write_csv(or_disease, paste0(file_name))
+  N[n] = length(unique(or_disease$personkey))
+  n = n+1
+}
+
+num_person_both <- data.frame(rbind(var_list, N))
+```
+
+```{r urgent data import, include = F, eval = FALSE}
+
+### Limit to ER visits and urgent care
+## pos = 23 or pos = 20
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+
+N <- NA
+n <- 1
+
+for (i in var_list){
+  read_path <- paste('../data_new/casecrossover/time_strat/or', i, 'may_to_sep_time_strat_casecross.csv', sep = "_")
+  or_disease <- read_csv(read_path)
+
+  or_disease <- or_disease %>% 
+    filter(pos=="20") %>%
+    # filter the unknown gender
+    filter(gender!= "U")
+
+  # Create a permanent case-cross over dataset
+  file_name <- paste('../data_new/casecrossover/Urgent/or', i, 'may_to_sep_time_strat_casecross_ur.csv', sep = "_")
+
+  # write permanent dataset
+  write_csv(or_disease, paste0(file_name))
+  N[n] = length(unique(or_disease$personkey))
+  n = n+1
+}
+
+num_person_ur <- data.frame(rbind(var_list, N))
+```
+
 ```{r echo=FALSE, message=FALSE, warning=FALSE}
 ## Analysis
-path <- paste0("../data_new/casecrossover/ER")
+path <- paste0("../data_new/casecrossover/ER_Urgent")
 
 resp_casecross <- read_csv(paste(path, 
-                                 "or_respiratory_may_to_sep_time_strat_casecross_er.csv", 
+                                 "or_respiratory_may_to_sep_time_strat_casecross_both.csv", 
                                  sep = "/"))
@@
 asthma_casecross <- read_csv(paste(path, 
-                                   "or_asthma_may_to_sep_time_strat_casecross_er.csv", 
+                                   "or_asthma_may_to_sep_time_strat_casecross_both.csv", 
                                    sep = "/"))
@@
 copd_casecross <- read_csv(paste(path, 
-                                 "or_copd_may_to_sep_time_strat_casecross_er.csv", 
+                                 "or_copd_may_to_sep_time_strat_casecross_both.csv", 
                                  sep = "/"))
@@
 pneum_casecross <- read_csv(paste(path, 
-                                  "or_pneumonia_may_to_sep_time_strat_casecross_er.csv",
+                                  "or_pneumonia_may_to_sep_time_strat_casecross_both.csv",
                                   sep="/"))
@@
 acute_bronch_casecross <- read_csv(paste(path, 
-                                         "or_acute_bronch_may_to_sep_time_strat_casecross_er.csv",
+                                         "or_acute_bronch_may_to_sep_time_strat_casecross_both.csv",
                                          sep = "/"))
@@
 cvd_casecross <- read_csv(paste(path, 
-                                "or_cvd_may_to_sep_time_strat_casecross_er.csv",
+                                "or_cvd_may_to_sep_time_strat_casecross_both.csv",
                                 sep="/"))
@@
 arrhythmia_casecross <- read_csv(paste(path, 
-                                       "or_arrhythmia_may_to_sep_time_strat_casecross_er.csv",
+                                       "or_arrhythmia_may_to_sep_time_strat_casecross_both.csv",
                                        sep="/"))
@@
 cereb_vas_casecross <- read_csv(paste(path, 
-                                      "or_cerbrovas_dis_may_to_sep_time_strat_casecross_er.csv",
+                                      "or_cerbrovas_dis_may_to_sep_time_strat_casecross_both.csv",
                                       sep="/"))
@@
 hf_casecross <- read_csv(paste(path, 
-                               "or_heart_failure_may_to_sep_time_strat_casecross_er.csv", 
+                               "or_heart_failure_may_to_sep_time_strat_casecross_both.csv", 
                                sep="/"))
@@
 ihd_casecross <- read_csv(paste(path, 
-                                "or_isch_heart_dis_may_to_sep_time_strat_casecross_er.csv",
+                                "or_isch_heart_dis_may_to_sep_time_strat_casecross_both.csv",
                                 sep="/"))
@@
 mi_casecross <- read_csv(paste(path, 
-                               "or_myocardial_infarc_may_to_sep_time_strat_casecross_er.csv", sep="/"))
+                               "or_myocardial_infarc_may_to_sep_time_strat_casecross_both.csv", sep="/"))
 
@@
 broken_arm_casecross <- read_csv(paste(path, 
-                                       "or_broken_arm_may_to_sep_time_strat_casecross_er.csv",
+                                       "or_broken_arm_may_to_sep_time_strat_casecross_both.csv",
                                        sep="/"))
@@
 We further limited our analyses to claims that were coded as emergency or urgent visits to eliminate bias from patients going in for elective/planned procedures.
-However, we found the filtered population is much larger than the population we chose in Washington, so we choose only the emegency room (hospital) visits.
+However, we found the filtered population is much larger than the population we chose in Washington, so we choose only the emegency room (hospital) visits and urgent care visits.
 For patients who have a index time closer to May 1, 2013 or closer to September 30, 2013, their referent observations before or after these dates will be excluded as I will not be able to assign estimates of PM~2.5~ to these referent observations.
@@
 # save figure
-ggsave("../plot_new/fig_only_er.pdf", plot = print_plot, 
+ggsave("../plot_new/fig_both_er_ur.pdf", plot = print_plot, 
        width = 12, height = 8, units = "in")
@@
   # save figure
-  ggsave("../plot_new/age_fig_three.pdf", plot = print_plot, 
+  ggsave("../plot_new/age_fig_three_both.pdf", plot = print_plot, 
        width = 12, height = 8, units = "in")
@@
   # save figure
-  ggsave("../plot_new/sex_fig_three.pdf", plot = print_plot, 
+  ggsave("../plot_new/sex_fig_three_both.pdf", plot = print_plot, 
        width = 12, height = 8, units = "in")

### 2017-11-03 r_scripts/analysis/oregon_medication_ts.Rmd

---
title: "Oregon 2013 Medication for Inhalers Time Series and Distributed Lag"
author: "Jingyang Liu"
date: "May 17, 2017"
output: html_document
---

```{r library, include=FALSE, echo=FALSE}
library(tidyverse)
# library(readxl) 
library(survival)
library(htmlTable)
library(knitr)
library(broom)
library(lme4)

# two-stage distributed lag model libraries
# library(dlnm)
# library(mvmeta)
# library(splines)

```

### Data Import

```{r import strat casecrossover data, echo = F, message = F, warning = F, results='asis'}
read_path6 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover.csv")
asthma_ndc <- read_csv(read_path6)
# asthma_ndc <- fread(read_path6, showProgress = T) # faster, but seems error in overall plot analysis

# dataframe to loop through
asthma_ndc_df <- data.frame(asthma_ndc)

```

## County-level time series

### County Time series plots

```{r prepare pop and pm, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Join data with names of Oregon counties 
oregon_fips <- read_csv('../instructions/oregon_FIPS.csv')
# remove the "County" character
# factor(oregon_fips$county)
oregon_fips$county <- gsub(" County", "", as.character(factor(oregon_fips$county)))

oregon_fips <- oregon_fips %>%
  mutate(st_county_fips = with(oregon_fips, paste0(st_code, fips)))

### Join county with zip
read_path <- paste0('../data_new/update/or_zip_county_prop.csv') 
or_zip_county <- read_csv(read_path)

or_zip_county <- or_zip_county %>%
  select(zip, county_name) %>%
  rename(county = county_name) %>%
  rename(ZIP = zip) %>%
  full_join(oregon_fips, by = "county") %>%
  select(ZIP, county, state, st_county_fips) %>%
  rename(fips = st_county_fips)

or_zip_county$county[which(or_zip_county$county=="Hood.River")] <- "Hood River"

read_path9 <- paste0("../data_new/county_data/or_census_pop_est_county.csv")
or_pop <- read_csv(read_path9)

or_pop <- or_pop %>%
  rename(county = GEO_display_label, pop = respop72013) %>%
  select(county, pop) 
or_pop$county <- gsub(" County, Oregon", "", as.character(factor(or_pop$county)))

```

### Prepare

1. Make 10 units geo smoke
2. Make county population (to calculate rate)
3. Make weekend/weekday variables

```{r time series, warning=FALSE, message=FALSE, echo=FALSE}
read_path7 <- paste0("../data_new/medication/or_ndc_county_time_series.csv")
or_ndc_county_ts <- read_csv(read_path7)

or_ndc_county_ts <- or_ndc_county_ts %>%
  left_join(or_pop, by = "county") %>%
  mutate(geo_smk_pm10 = 10*geo_smk_pm) %>%
  mutate(rate_per_100k = (n_obs/pop)*100000) %>%
  mutate(day = ifelse(weekdays(date) == "Monday" |
                      weekdays(date) == "Tuesday" |
                      weekdays(date) == "Wednesday"|
                      weekdays(date) == "Thursday"|
                      weekdays(date) == "Friday", "weekday",
               ifelse(weekdays(date) == "Saturday" |
                      weekdays(date) == "Sunday", "weekend", NA)))

or_ndc_county_ts$day[which(or_ndc_county_ts$date=="2013-05-27")] <- "weekend"
or_ndc_county_ts$day[which(or_ndc_county_ts$date=="2013-07-04")] <- "weekend"
or_ndc_county_ts$day[which(or_ndc_county_ts$date=="2013-09-02")] <- "weekend"

plot_ts <- ggplot(or_ndc_county_ts, aes(y =rate_per_100k, x = date)) +
           geom_jitter()

print(plot_ts)

# save figure
ggsave("../plot_new/ndc_rate_ts.pdf", plot = plot_ts, 
       width = 12, height = 8, units = "in")

plot_county <- ggplot(or_ndc_county_ts, aes(y =rate_per_100k, x = date)) +
               geom_jitter() + facet_wrap(~county)

plot_county_line <- ggplot(or_ndc_county_ts, aes(y =rate_per_100k, x = date)) +
                    facet_wrap(~county) + geom_line()

print(plot_county)
print(plot_county_line)
# save figure
ggsave("../plot_new/ndc_county_rate_ts.pdf", plot = plot_county, 
       width = 12, height = 8, units = "in")
ggsave("../plot_new/ndc_county_rate_ts_line.pdf", plot = plot_county_line, 
       width = 12, height = 8, units = "in")

### check if the time series has the association with get smoke
a <- glm(rate_per_100k ~ geo_smk_pm10, data = or_ndc_county_ts, family = "poisson")
summary(a)

plot_health_smk <- ggplot(or_ndc_county_ts, aes(x = geo_smk_pm10, y = log(rate_per_100k+0.1))) +
  geom_point(aes(colour = c("red")), alpha = 0.2,  size = 1) + 
  # scale_colour_manual(values = c("red", "blue")) +
  # truncated y axis
  #scale_y_continuous(limits = c(0, 1.5)) +
  # facet_wrap(~county, scale  = "free") +
  ylab("Log Outcome Rate per 100,000 Persons") +
  xlab("GWR Smoke PM2.5 per 10*10ug/m^3") +
  theme_bw()

plot_county_health_smk <- ggplot(or_ndc_county_ts, aes(x = geo_smk_pm10, y = log(rate_per_100k+0.1))) +
  geom_point(aes(colour = county), alpha = 0.2,  size = 1) + 
  # scale_colour_manual(values = c("red", "blue")) +
  # truncated y axis
  #scale_y_continuous(limits = c(0, 1.5)) +
  facet_wrap(~county, scale  = "free") +
  ylab("Log Outcome Rate per 100,000 Persons") +
  xlab("GWR Smoke PM2.5 per 10*10ug/m^3") +
  theme_bw()

plot(plot_health_smk)
plot(plot_county_health_smk)

ggsave("../plot_new/ndc_rate_health_smk.pdf", plot = plot_health_smk, 
       width = 12, height = 8, units = "in")

ggsave("../plot_new/ndc_county_rate_health_smk.pdf", plot = plot_county_health_smk, 
       width = 12, height = 8, units = "in")

```

### Standard Regression Model

Same-day estimates of association between smoke via GWR and outcomes. County treated as a fixed effect covariate.

1. Quasipoisson

```{r regression model q, message=FALSE, echo=FALSE}
# create empty dataframe
rr_df <- data.frame(matrix(nrow = 1, ncol = 4))
colnames(rr_df) <- c("outcome", "rel_risk", "lower_95", "upper_95")

# vector of outcomes
outcome_list <- "n_obs"

# names 
outcome_names <- c('Inhalers')

# populate outcome column
rr_df[,1] <- factor(outcome_names, levels=unique(outcome_names))

smk_est <- tidy(glm(as.formula(paste0(outcome_list, 
    "~ geo_smk_pm10 + wrf_temp + county + 
    day + offset(log(pop))")), 
    or_ndc_county_ts, family="quasipoisson"))[2, ]

rr <- c(round(exp(smk_est[1,2]),5),
        round(exp(smk_est[1,2]-1.96*smk_est[1,3]),5),
        round(exp(smk_est[1,2]+1.96*smk_est[1,3]),5))

rr_df[1,2:4] <- rr

# tried apply function and it might have been easier to just use a loop
# as I needed a for loop to fill the list in to the dataframe
# code is more compact though
# I should check speed (or make a custom function to do this one day)

rr_df

plot_rr <- ggplot(rr_df, aes(x=outcome, y = rel_risk)) +
  geom_point() +
  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
  theme(
   panel.background = element_rect(fill = 'white', colour = 'black'),
   panel.grid.major = element_blank(),
   panel.grid.minor = element_blank(),
   axis.text.x = element_text(angle = 90, vjust = .75)) +
  ylab("Relative Risk") +
  xlab("Outcome") +
  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")

print(plot_rr)

ggsave("../plot_new/ndc_rr.pdf", plot = plot_rr, 
       width = 12, height = 8, units = "in")

```

2. Poisson

```{r regression model,  message=FALSE, echo=FALSE}
# create empty dataframe
rr_df <- data.frame(matrix(nrow = 1, ncol = 4))
colnames(rr_df) <- c("outcome", "rel_risk", "lower_95", "upper_95")

# vector of outcomes
outcome_list <- "n_obs"

# names 
outcome_names <- c('Inhalers')

# populate outcome column
rr_df[,1] <- factor(outcome_names, levels=unique(outcome_names))

smk_est <- tidy(glm(as.formula(paste0(outcome_list, 
    "~ geo_smk_pm10 + wrf_temp + county + 
    day + offset(log(pop))")), 
    or_ndc_county_ts, family="poisson"))[2, ]

rr <- c(round(exp(smk_est[1,2]),5),
        round(exp(smk_est[1,2]-1.96*smk_est[1,3]),5),
        round(exp(smk_est[1,2]+1.96*smk_est[1,3]),5))

rr_df[1,2:4] <- rr

rr_df

# tried apply function and it might have been easier to just use a loop
# as I needed a for loop to fill the list in to the dataframe
# code is more compact though
# I should check speed (or make a custom function to do this one day)

plot_rr <- ggplot(rr_df, aes(x=outcome, y = rel_risk)) +
  geom_point() +
  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
  theme(
   panel.background = element_rect(fill = 'white', colour = 'black'),
   panel.grid.major = element_blank(),
   panel.grid.minor = element_blank(),
   axis.text.x = element_text(angle = 90, vjust = .75)) +
  ylab("Relative Risk") +
  xlab("Outcome") +
  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")

print(plot_rr)

ggsave("../plot_new/ndc_rr_poi.pdf", plot = plot_rr, 
       width = 12, height = 8, units = "in")

```

### Mixed models

1. Poisson

```{r mixed model,  message=FALSE, echo=FALSE}
# create empty dataframe
rr_mix_df <- data.frame(matrix(nrow = 1, ncol = 4))
colnames(rr_mix_df) <- c("outcome", "rel_risk", "lower_95", "upper_95")

# vector of outcomes
outcome_list <- "n_obs"

# names 
outcome_names <- c('Inhalers')

# populate outcome column
rr_mix_df[,1] <- factor(outcome_names, levels=unique(outcome_names))

smk_est <- tidy(glmer(as.formula(paste0(outcome_list, 
    "~ geo_smk_pm10 + wrf_temp + (1|county) + 
    day + offset(log(pop))")), 
    or_ndc_county_ts, family="poisson"))[2, ]

rr <- c(round(exp(smk_est[1,2]),5),
        round(exp(smk_est[1,2]-1.96*smk_est[1,3]),5),
        round(exp(smk_est[1,2]+1.96*smk_est[1,3]),5))

rr_mix_df[1,2:4] <- rr

rr_mix_df

plot_mix_rr <- ggplot(rr_mix_df, aes(x=outcome, y = rel_risk)) +
  geom_point() +
  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
  theme(
   panel.background = element_rect(fill = 'white', colour = 'black'),
   panel.grid.major = element_blank(),
   panel.grid.minor = element_blank(),
   axis.text.x = element_text(angle = 90, vjust = .75)) +
  ylab("Relative Risk") +
  xlab("Outcome") +
  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")

print(plot_rr)

ggsave("../plot_new/ndc_mix_rr_poi.pdf", plot = plot_mix_rr, 
       width = 12, height = 8, units = "in")

```

```{r, eval = FALSE, include=FALSE}
asthma_ndc_df <- asthma_ndc_df %>%
  rename(ZIP = ZIPCODE) %>%
  left_join(or_zip_county, by = "ZIP") %>% # NA: 7005
  left_join(or_pop, by = "county")

```

```{r, eval=FALSE, include=FALSE}

# Import population weighted county PM2.5
pm_path <- paste0("../data_new/county_data/or_county_pop_wt_pm.csv")
county_pm <- read_csv(pm_path)

# create a working dataset 'or_2013'
or_ndc_county_time_series <- or_ndc_asthma %>% 
  # join in the washington county name based on CHARS county identifier
  left_join(or_zip_county, by = "ZIP") %>% 
  # rename 'admit_date_impute' to 'date'
  rename(date = dates) %>%
  # rename(date = admit_date_impute) %>% 
  # group by county of residence and date
  group_by(county, date) %>% 
  # sum up each primary diagnosis for each outcome for each day for each county
  summarise(n_obs = n()) %>% 
  # join daily county estimates of pm2.5
  left_join(county_pm, by = c('date', 'county'))

# which(is.na(or_ndc_county_time_series$county))
# or_ndc_county_time_series[113,]
# which(or_zip_county$ZIP=="97078")

# summary(or_ndc_asthma)
# summary(or_ndc_county_time_series)
# summary(as.factor(or_ndc_county_time_series$county)) # 125 is missing out of 3592 (3.16%)

# remove NA values
# which(is.na(or_ndc_county_time_series$county)) # 3468-3592
or_ndc_county_time_series <- or_ndc_county_time_series[-c(3468:3592),]

# write permanent dataframe

write_path5 <- paste0("../data_new/medication/or_ndc_county_time_series.csv")
write_csv(or_ndc_county_time_series, write_path5)

## Check for missing values in sparsely populated counties. set to 0 (But no)
# check_missing <- filter(or_ndc_county_time_series, is.na(n_obs))
# mutate each to fill in missing values (Still no)
# check2 <- check_missing %>% mutate_each(funs(new = ifelse(is.na(.), 0, .)), 
#                                         n_obs:wrf_temp)

```

### Time series of weekday and weekend in larger counties.
We check the time series for Multnomah, and it indeed shows more visits of weekdays than that in weekends (except some holidays such like Labor Day, Independent Day, etc.) We may choose the new covariates for weekday and weekend.

```{r echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}
or_ndc_county_time_series_week <- or_ndc_county_time_series%>%
  mutate(day_admit = as.factor(weekdays(date)),
         month_admit = month(date)) %>%
  mutate(outcome = 1)

# Also check for Multnomah
multnomah_df <- or_ndc_county_time_series_week %>%
  filter(county == "Multnomah")

plot_mult_week <- ggplot(multnomah_df, aes(y =n_obs, x = date)) +
           geom_jitter() + facet_wrap(~day_admit)

print(plot_mult_week)

ggsave("../plot_new/multnomah_ts_week.pdf", plot = plot_mult_week, 
       width = 12, height = 8, units = "in")
```

### Time series of smoke exposure
The plot shows the times series of smoke PM 2.5 exposure. It seems that at the beginning of August or at the end of July the smoke has a peak. Also during the August the smoke is very high. The green one shows the WRF-Chem method; The bule one shows the Krigging method; And the red one, which we focus on, shows the GWR method.

```{r echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}

plot_smoke_ts_wrf <- ggplot(or_ndc_county_time_series_week, aes(y =wrf_smk_pm, x = date)) +
           geom_jitter(color="red")
# print(plot_smoke_ts_wrf)

plot_smoke_ts_krig <- ggplot(or_ndc_county_time_series_week, aes(y =krig_smk_pm, x = date)) +
           geom_jitter() 
# print(plot_smoke_ts_krig)

plot_smoke_ts_gwr <- ggplot(or_ndc_county_time_series_week, aes(y =geo_smk_pm, x = date)) +
           geom_jitter() 
# print(plot_smoke_ts_gwr)

plot_smk <- ggplot(or_ndc_county_time_series_week, aes(x = date)) +
  geom_jitter(aes(y =wrf_smk_pm, x = date), color='green', size = 1) + 
  geom_jitter(aes(y =krig_smk_pm, x = date), color='blue', size = 1) + 
  geom_jitter(aes(y =geo_smk_pm, x = date), color="red", size = 1)  +
  ylab("Smoke PM 2.5")

print(plot_smk)

### County
or_smk <- or_ndc_county_time_series_week %>% 
  select(county, date, wrf_smk_pm, krig_smk_pm, geo_smk_pm) %>% 
  gather(key = smk_method, value = pm, -date, -county)

or_smk_plot <- ggplot(or_smk, aes(x = date, y = pm)) +
  geom_point(aes(colour = smk_method), alpha = 0.8, size = 0.8) +
  facet_wrap(~county) +
  ggtitle("Oregon: 2013 Fire Season Smoke PM2.5") +
  ylab("Smoke PM2.5 ug/m^3") +
  xlab("Year: 2013") +
  theme_bw()

print(or_smk_plot)

ggsave("../plot_new/ndc_smoke_county.pdf", plot = plot_mult_week, 
       width = 12, height = 8, units = "in")

```

## Distributed Lag Model

### Lag 1

```{r DLM, echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}
asthma_ndc_df <- data.frame(asthma_ndc)

asthma_ndc_df <- asthma_ndc_df %>% # GWR
  filter(!is.na(geo_wt_pm_lag7_zip)) # 503998

which(colnames(asthma_ndc_df)=="geo_wt_pm_zip") # 88
which(colnames(asthma_ndc_df)=="geo_wt_pm_lag1_zip") #115

asthma_ndc_lag <- asthma_ndc_df %>%
  select(88, 115:121) # GWR-method

lag_max <- 7
n <- length(asthma_ndc_lag[,1])

asthma_ndc_lag <- as.matrix(asthma_ndc_lag)

B <- ns(0:lag_max, df =4, intercept=TRUE)

# the nrow B should equal ncol(pm)
dim(B)
dim(asthma_ndc_lag)

asthma_ndc_lag_B <- asthma_ndc_lag%*%B
fit <- clogit(outcome ~ asthma_ndc_lag_B + wrf_temp_zip + strata(personkey), asthma_ndc_df)
coef(fit)
summary(fit)

dlparms <- grep("asthma_ndc_lag_B",names(coef(fit)))
DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms])
DLvar <- B%*%vcov(fit)[dlparms,dlparms]%*%t(B)
DLestimate$SE <- sqrt(diag(DLvar))
DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96

DLestimate$estimate <- exp(DLestimate$estimate)
DLestimate$lower <- exp(DLestimate$lower)
DLestimate$upper <- exp(DLestimate$upper)

DLestimate

DLestimate$lag <- 0:lag_max
p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
p <- p + geom_line(aes(x=lag), linetype=2, size=2)
p <- p + theme_bw()
p

DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms])

cumulative <- sum(DLestimate$estimate)
cumulative
## [1] 0.009141771
cumulative_se <- sqrt(sum(DLvar))
cumulative_se
## [1] 0.000955469
cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
cumulative_CI
## [1] 0.007269052 0.011014490

```

```{r overall lag 1, warning =F, echo = F, results='asis', eval=FALSE} 
# dataframe list

method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')

# dataframe to loop through
# asthma_ndc_df <- data.frame(asthma_ndc)

# outcome name
outcome <- "Asthma Inhalers Lag 1"
outcome_name <- "Asthma Inhalers Lag 1"

# data wrangling ----
# Producing conditional logit model estimates loop

# extract covariates from dataframe
covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]

# extract pm values and divide by 10 and ordered
which(colnames(asthma_ndc_df)=="wrf_smk_pm_lag1_zip") # 108
which(colnames(asthma_ndc_df)=="krig_pm_lag1_zip") # 122
which(colnames(asthma_ndc_df)=="geo_wt_pm_lag1_zip") # code to find column numbers, 115
pm_estimates_df <- asthma_ndc_df[, c(108, 122, 115, 93)]/10  # create 10 unit increases

# new loop for age categories

# empty matrix (12 x 10 matrix)
point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)

colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 
                               'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')

# fill in the outcome namedataframe before method loop
point_estimates[, 1] <- outcome_name

# dataframe for analysis creation
# bind columns back together 
df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
  # remove missing pm values
  filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
  filter(!is.na(krig_pm_lag1_zip)) %>% 
  filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
  # the following code makes sure that the counterfactual values retained are 
  # symetric in that number of obs before = number of obs after
  mutate(obs_diff_admission = (fromdate - date)/7) 
# dataframe is already for the entire fire season, so I don't need to subset anymore

# second loop to run a model for each pm estimation method
for(j in 38:40){

  # variable to model 
  var_name <- colnames(df_analysis[j])

  # set row number to fill
  row_n <- j-37

  # only run the model if the dataframe has observations
  if(nrow(df_analysis) != 0){
    # conditional logistic regression model
    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)

    # populate matrix
    point_estimates[row_n, 2] <- method_list[row_n]
    point_estimates[row_n, 3] <- mod$n
    point_estimates[row_n, 4] <- mod$nevent
    # odds ratio
    point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)

    # 95% lower bound
    point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
    # 95% upper bound
    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
                                             1.96*(summary(mod)$coefficient[1,3])), 3)
    # standard error
    point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
    # p val
    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)

    # create else statement that fills matrix with missing so I still have the row
    # in the final dataframe
  } else {point_estimates[row_n, 2] <- method_list[row_n]
  point_estimates[row_n, 3] <- 0
  point_estimates[row_n, 4] <- 0
  point_estimates[row_n, c(5:9)] <- 99 } # end 'if else' statement
} # end methods loop

# save point estimates as a dataframe
combined_point_est_df <- as_data_frame(point_estimates)

wrf_geo_age <- combined_point_est_df %>% 
  # filter columns to just geo-smk
  # filter(pm_method == "GWR Smoke" |
  #      pm_method == "WRF-Chem Smoke") %>% 
  # subset columns I want to put in to the table
  select(2, 4:7) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
         lower95 = ifelse(lower95 == 99, "--", lower95),
         upper95 = ifelse(upper95 == 99, "--", upper95))

tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
                 caption = "Association between a 10 ug/m^3 in PM2.5 three smoke method and pharmacy outcomes",
                 # row group by outcome
                 rgroup = "Asthma Inhalers Lag 1",
                 n.rgroup = c(rep(3, 1)), # 6 rows for each age cat for each outcome
                 # column headers
                 header = c("Est Method", "Events",
                            "OR&dagger;", "Lower", "Upper"),
                 # column spanner
                 cgroup = c("", "95% CI"), 
                 n.cgroup = c(3, 2),
                 padding.rgroup = "&nbsp;&nbsp;",
                 css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
                 align = "llcccc", # column alignment,
                 tfoot="&dagger; Referent periods matched to events on same day of week within May to September fire season."
) # end table

print(tab)

combined_point_est_df_new <- combined_point_est_df %>%
  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
                               ifelse(as.numeric(n_events)<100, "0", NA))) %>%
  select(-n_events) %>%
  rename(n_events = n_events_new)

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
wrf_geo_age_plot <- combined_point_est_df_new %>% 
  # filter columns to just geo-smk
  #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
  # do not plot results with less than 15 events, create missing vals
  # in mutate
  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
                             99, odds_ratio),
         lower95 = ifelse(as.numeric(n_events) < 15, 
                          99, lower95),
         upper95 = ifelse(as.numeric(n_events) < 15, 
                          99, upper95)) %>% 
  # subset columns I want to put in to the table
  select(1, 2, 9, 4:6) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
         lower95 = ifelse(lower95 == 99, NA, lower95),
         upper95 = ifelse(upper95 == 99, NA, upper95))

# change characters to numeric and factor  
wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
                                   levels = unique(wrf_geo_age_plot$outcome))

wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
                                     levels = unique(wrf_geo_age_plot$pm_method))

wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)

## ggplot 
print_plot <- ggplot(wrf_geo_age_plot,
                     aes(x = pm_method, y = odds_ratio, color = pm_method), na.rm = T) +
  geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
  geom_errorbar(aes(ymin=lower95, ymax=upper95), 
                position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
  # custom color 
  scale_color_manual(name = "Smoke-Estimation Method", 
                     values = c("red",  "blue", "#32115C"),
                     guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
  facet_wrap(~outcome, nrow = 3, scales = "free") +
  geom_hline(yintercept = 1, linetype=2) +
  #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
  ylab(expression(paste("Odds Ratio for 10Âµg/m"^3, " Increase in PM"[2.5]))) +
  #ylim(0, 2) +
  xlab('Smoke Estimation Method') +
  # plot theme
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        strip.text = element_text(size = 10),
        # axis element
        #axis.text.x = element_blank(),
        #axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(angle = 90),
        # legend elements
        legend.position = "bottom")
#legend.text = element_text(size = 8))

print(print_plot)
# save figure
ggsave("../plot_lag/ndc_overall_lag1_three.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

```

```{r time stratified age strata lag 1, warning =F, echo = F, results='asis', eval=FALSE} 
# dataframe list

method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')

# dataframe to loop through
asthma_ndc_df <- data.frame(asthma_ndc)

# outcome name
outcome <- "asthma inhalers Lag 1"
outcome_name <- "Asthma inhalers Lag 1"

# age category list
age_cat_list <- c(0,1,2)

# create an empty list to row bind dataframes together
datalist1 <- list()

# data wrangling ----
# Producing conditional logit model estimates loop

# extract covariates from dataframe
covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]

# extract pm values and divide by 10 and ordered
#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
pm_estimates_df <- asthma_ndc_df[, c(108, 122, 115, 93)]/10  # create 10 unit increases

# new loop for age categories
for(k in 0:2){
  # empty matrix (12 x 10 matrix)
  point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)

  colnames(point_estimates) <- c('outcome', 'age_cat', 'pm_method', 'n', 'n_events', 
                                 'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')

  # fill in the outcome namedataframe before method loop
  point_estimates[, 1] <- outcome_name
  # repeat the age category 4 times for each pm method
  point_estimates[, 2] <- k

  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
    # remove missing pm values
    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
    filter(!is.na(krig_pm_lag1_zip)) %>% 
    filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
    # limit to specific age category
    filter(age_ind == k) %>% 
    # the following code makes sure that the counterfactual values retained are 
    # symetric in that number of obs before = number of obs after
    mutate(obs_diff_admission = (fromdate - date)/7) 
  # dataframe is already for the entire fire season, so I don't need to subset anymore

  # second loop to run a model for each pm estimation method
  for(j in 38:40){

    # variable to model 
    var_name <- colnames(df_analysis[j])

    # set row number to fill
    row_n <- j-37

    # only run the model if the dataframe has observations
    if(nrow(df_analysis) != 0){
      # conditional logistic regression model
      mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)

      # populate matrix
      point_estimates[row_n, 3] <- method_list[row_n]
      point_estimates[row_n, 4] <- mod$n
      point_estimates[row_n, 5] <- mod$nevent
      # odds ratio
      point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)

      # 95% lower bound
      point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
      # 95% upper bound
      point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
                                               1.96*(summary(mod)$coefficient[1,3])), 3)
      # standard error
      point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
      # p val
      point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)

      # create else statement that fills matrix with missing so I still have the row
      # in the final dataframe
    } else {point_estimates[row_n, 3] <- method_list[row_n]
    point_estimates[row_n, 4] <- 0
    point_estimates[row_n, 5] <- 0
    point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
  } # end methods loop

  # save point estimates as a dataframe
  point_est_df <- as_data_frame(point_estimates)
  datalist1[[k+1]] <- point_est_df
} # end age category loop

# bind rows of age category estimates together
combined_point_est_df <- bind_rows(datalist1)

wrf_geo_age <- combined_point_est_df %>% 
  # filter columns to just geo-smk
  # filter(pm_method == "GWR Smoke" |
  #      pm_method == "WRF-Chem Smoke") %>% 
  mutate(age_cat2 = ifelse(age_cat == 0, "Less than 15", 
                           ifelse(age_cat == 1, "15-65",
                                  ifelse(age_cat == 2, "Greater than 65", NA)))) %>% 
  # subset columns I want to put in to the table
  select(3, 11, 5:8) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
         lower95 = ifelse(lower95 == 99, "--", lower95),
         upper95 = ifelse(upper95 == 99, "--", upper95))

tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
                 caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by age",
                 # row group by outcome
                 rgroup = "Asthma inhalers Lag 1",
                 n.rgroup = c(rep(9, 1)), # 6 rows for each age cat for each outcome
                 # column headers
                 header = c("Est Method", "Age Group", "Events",
                            "OR&dagger;", "Lower", "Upper"),
                 # column spanner
                 cgroup = c("", "95% CI"), 
                 n.cgroup = c(4, 2),
                 padding.rgroup = "&nbsp;&nbsp;",
                 css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
                 align = "llcccc", # column alignment,
                 tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within May to September fire season."
) # end table

print(tab)

combined_point_est_df_new <- combined_point_est_df %>%
  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
                               ifelse(as.numeric(n_events)<100, "0", NA))) %>%
  select(-n_events) %>%
  rename(n_events = n_events_new)

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
wrf_geo_age_plot <- combined_point_est_df_new %>% 
  # filter columns to just geo-smk
  #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
  # do not plot results with less than 15 events, create missing vals
  # in mutate
  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
                             99, odds_ratio),
         lower95 = ifelse(as.numeric(n_events) < 15, 
                          99, lower95),
         upper95 = ifelse(as.numeric(n_events) < 15, 
                          99, upper95),
         age_cat2 = ifelse(age_cat == 0, "<15", 
                           ifelse(age_cat == 1, "15-65",
                                  ifelse(age_cat == 2, ">65", NA)))) %>% 
  # subset columns I want to put in to the table
  select(1, 3, 11, 10, 5:7) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
         lower95 = ifelse(lower95 == 99, NA, lower95),
         upper95 = ifelse(upper95 == 99, NA, upper95))

# change characters to numeric and factor  
wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
                                   levels = unique(wrf_geo_age_plot$outcome))

wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
                                     levels = unique(wrf_geo_age_plot$pm_method))

wrf_geo_age_plot$age_cat2 <- factor(wrf_geo_age_plot$age_cat2,
                                    levels = unique(wrf_geo_age_plot$age_cat2))

wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)

## ggplot 
print_plot <- ggplot(wrf_geo_age_plot,
                     aes(x = age_cat2, y = odds_ratio, color = pm_method), na.rm = T) +
  geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
  geom_errorbar(aes(ymin=lower95, ymax=upper95), 
                position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
  # custom color 
  scale_color_manual(name = "Smoke-Estimation Method", 
                     values = c("red",  "blue", "#32115C"),
                     guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
  facet_wrap(~outcome, nrow = 3, scales = "free") +
  geom_hline(yintercept = 1, linetype=2) +
  #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
  ylab(expression(paste("Odds Ratio for 10Âµg/m"^3, " Increase in PM"[2.5]))) +
  #ylim(0, 2) +
  xlab('Age Category ') +
  # plot theme
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        strip.text = element_text(size = 10), 
        # axis element
        #axis.text.x = element_blank(),
        #axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(angle = 90),
        # legend elements
        legend.position = "bottom")
#legend.text = element_text(size = 8))

print(print_plot)
# save figure
ggsave("../plot_lag/ndc_age_fig_three_lag1.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

```

```{r sex strata cdc met  lag 1, warning = F, echo = F, results='asis', eval=FALSE} 
method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')

# dataframe to loop through
asthma_ndc_df <- data.frame(asthma_ndc)

# outcome name
outcome <- "asthma inhalers Lag 1"
outcome_name <- "Asthma inhalers Lag 1"

# create an empty list to row bind dataframes together
datalist1 <- list()

# sex category list
sex_strata_list <- c(0,1)

# data wrangling ----
# Producing conditional logit model estimates loop

# extract covariates from dataframe
covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]

# extract pm values and divide by 10 and ordered
#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
pm_estimates_df <- asthma_ndc_df[, c(108, 122, 115, 93)]/10  # create 10 unit increases

# new loop for sex categories
for(k in 0:1){

  # empty matrix (12 x 10 matrix)
  point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)

  colnames(point_estimates) <- c('outcome', 'sex', 'pm_method', 'n', 'n_events', 
                                 'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')

  # fill in the outcome namedataframe before method loop
  point_estimates[, 1] <- outcome_name
  # repeat the sex category 4 times for each pm method
  point_estimates[, 2] <- k

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
    # remove missing pm values
    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
    filter(!is.na(krig_pm_lag1_zip)) %>% 
    filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
    # limit to specific age category
    filter(sex_ind == k) %>% 
    # the following code makes sure that the counterfactual values retained are 
    # symetric in that number of obs before = number of obs after
    mutate(obs_diff_admission = (fromdate - date)/7) 
  # dataframe is already for the entire fire season, so I don't need to subset anymore

  # second loop to run a model for each pm estimation method
  for(j in 38:40){

    # variable to model 
    var_name <- colnames(df_analysis[j])

    # set row number to fill
    row_n <- j-37

    # only run the model if the dataframe has observations
    if(nrow(df_analysis) != 0){
      # conditional logistic regression model
      mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)

      # populate matrix
      point_estimates[row_n, 3] <- method_list[row_n]
      point_estimates[row_n, 4] <- mod$n
      point_estimates[row_n, 5] <- mod$nevent
      # odds ratio
      point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)

      # 95% lower bound
      point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
                                               1.96*(summary(mod)$coefficient[1,3])), 3)
      # 95% upper bound
      point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
                                               1.96*(summary(mod)$coefficient[1,3])), 3)
      # standard error
      point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
      # p val
      point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)

      # create else statement that fills matrix with missing so I still have the row
      # in the final dataframe
    } else {point_estimates[row_n, 3] <- method_list[row_n]
    point_estimates[row_n, 4] <- 0
    point_estimates[row_n, 5] <- 0
    point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
  } # end methods loop

  # save point estimates as a dataframe
  point_est_df <- as_data_frame(point_estimates)

  # combine previous values in dataframe that has all outcome/methods comparisons
  datalist1[[k+1]] <- point_est_df
} # end sex category loop

# bind rows of sex category estimates together
combined_point_est_df <- bind_rows(datalist1)

wrf_geo_sex <- combined_point_est_df %>% 
  # filter columns to just geo-smk
  # filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
  mutate(sex_cat = ifelse(sex == 0, "Male", 
                          ifelse(sex == 1, "Female", NA))) %>% 
  # subset columns I want to put in to the table
  select(3, 11, 5:8)%>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
         lower95 = ifelse(lower95 == 99, "--", lower95),
         upper95 = ifelse(upper95 == 99, "--", upper95))

tab <- htmlTable(txtRound(wrf_geo_sex, digits = 3, 1:3), 
                 caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and pharmacy outcomes stratified by sex",
                 # row group by outcome
                 rgroup = "Asthma inhalers Lag 1",
                 n.rgroup = c(rep(6, 1)), # 2 rows for each sex strata for each outcome
                 # column headers
                 header = c("Est Method", "Sex Strata", "Events",
                            "OR&dagger;", "Lower", "Upper"),
                 # column spanner
                 cgroup = c("", "95% CI"), 
                 n.cgroup = c(4, 2),
                 padding.rgroup = "&nbsp;&nbsp;",
                 css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
                 align = "llcccc", # column alignment,
                 tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within July to October fire season."
) # end table

print(tab)

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
wrf_geo_sex_plot <- combined_point_est_df %>% 
  # filter columns to just geo-smk
  #  filter(pm_method == "GWR Smoke" | 
  #       pm_method == "WRF-Chem Smoke") %>% 
  mutate(sex_cat = ifelse(sex == 0, "Male", 
                          ifelse(sex == 1, "Female", NA))) %>%
  # subset columns I want to put in to the table
  select(1, 3, 11, 5:8) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
         lower95 = ifelse(lower95 == 99, NA, lower95),
         upper95 = ifelse(upper95 == 99, NA, upper95))

# change characters to numeric and factor  
wrf_geo_sex_plot$outcome <- factor(wrf_geo_sex_plot$outcome,
                                   levels = unique(wrf_geo_sex_plot$outcome))

wrf_geo_sex_plot$pm_method <- factor(wrf_geo_sex_plot$pm_method,
                                     levels = unique(wrf_geo_sex_plot$pm_method))

wrf_geo_sex_plot$sex_cat <- factor(wrf_geo_sex_plot$sex_cat,
                                   levels = unique(wrf_geo_sex_plot$sex_cat))

wrf_geo_sex_plot$n_events <- as.numeric(wrf_geo_sex_plot$n_events)
wrf_geo_sex_plot$odds_ratio <- as.numeric(wrf_geo_sex_plot$odds_ratio)
wrf_geo_sex_plot$lower95 <- as.numeric(wrf_geo_sex_plot$lower95)
wrf_geo_sex_plot$upper95 <- as.numeric(wrf_geo_sex_plot$upper95)

## ggplot
print_plot <- ggplot(wrf_geo_sex_plot,
                     aes(x = sex_cat, y = odds_ratio, color = pm_method), na.rm = T) +
  geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
  geom_errorbar(aes(ymin=lower95, ymax=upper95), 
                position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
  # custom color 
  scale_color_manual(name = "Smoke-Estimation Method", 
                     values = c("red", "blue", "#32115C"),
                     guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
  facet_wrap(~outcome, nrow = 3, scales = "free") +
  geom_hline(yintercept = 1, linetype=2) +
  #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
  ylab(expression(paste("Odds Ratio for 10Âµg/m"^3, " Increase in PM"[2.5]))) +
  #ylim(0, 2) +
  xlab('Sex') +
  # plot theme
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        strip.text = element_text(size = 10),
        # axis element
        #axis.text.x = element_blank(),
        #axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(angle = 90),
        # legend elements
        legend.position = "bottom")
#legend.text = element_text(size = 8))

print(print_plot)
# save figure
ggsave("../plot_lag/ndc_sex_fig_three_lag1.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

```

### 2017-11-03 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
Cleaning up results code to be more streamlined.

@@
 
-**Note to Jingyang: I saw all the packages you used.
+**Note to Jingyang 11/3/17: I saw all the packages you used.
 How many do you actually need to run these analyses?
@@
 
-Packages used:
+Packages used: tidyverse
 ```{r setup}
@@
 ```
+
+##2. Wildfire smoke descriptive characteristics
+Map of smoke count and smoke time series will go here.
+
+##3. Descriptive table of number of health outcomes
+
+Importing casecross over dataframes.
+**Note to Jingyang 11/3/17: Same issue as above.
+How many of these files needs to be uploaded to produce results?
+A lot of this seems redundant or inefficient.
+I cannot find a lot of outcome files I need.
+I also need the SABA casecrossover dataframe.
+Listing example of two casecross over dataframes for now.
+Let's talk about this.**
+
+```{r health data import}
+# importing case-crossover dataframes 
+# SABA case-crossover will be imported first
+# i can't find a lot of the files
+#resp_path <- paste0("./"
+
+# asthma
+asthma_path <- paste0("./data/health/2017-oregon_casecross_asthma.csv")
+asthma_df <- read_csv(asthma_path)
+
+# cvd 
+cvd_path <- paste0("./data/health/2017-oregon_casecross_cvd.csv")
+cvd_df <- read_csv(cvd_path)
+# had a warning message for cvd dataframe import; should check on this
+
+# import all other dataframes
+
+# I think purrr would be a good way to loop through a list of outcome dataframes
+# bind dataframes in a list
+outcome_df_list <- list(asthma_df, cvd_df)
+```
+
+CVD warning message on import. Have Jingyang check. Maybe why proportion of CVD vs respiratory so different in Oregon?
+*Warning message:In rbind(names(probs), probs_f) :number of columns of result is not a multiple of vector length (arg 1)*
+
+Descriptive table will go here.
+
+```{r descriptive table}
+# name list elements
+names(outcome_df_list) <- c( 'Asthma', 'CVD')
+
+# output row of summary numbers
+test_row <- asthma_df %>% 
+  filter(outcome == 1) %>%
+  group_by(age_ind) %>% 
+  summarise(n = n()) %>%
+  mutate(age_cat = ifelse(age_ind == 0, "age_under15",
+                    ifelse(age_ind == 1, "age_15to65", "age_over65")))
+
+  forcats::as_factor(age_ind, levels = c("age_under15", "age_15to65", "age_over65"))
+
+glimpse(test_ro)
+
+descriptive_table <- outcome_df_list %>% 
+  lmap()
+
+?lmap
+
+mtcars %>%
+  split(.$cyl) %>%
+  map(~ lm(mpg ~ wt, data = .x)) %>%
+  map(summary) %>%
+  map_dbl("r.squared")
+
+# Use map_lgl(), map_dbl(), etc to reduce to a vector.
+# * list
+mtcars %>% map(sum)
+# * vector
+mtcars %>% map_dbl(sum)
+```
+
+##4. Distributed lag cumulative effect results
+
+##5. Distributed lag results
+
+##6. Stratified cumulative results
+
+##7. SABA cumulative effect by underlying respiratory diagnosis
+
+### Comments:
+1. File structure needs to be cleaned up. There are way too many subdirectories.
+2. Remove r code files that are not useful. There are too many files with similar code inside.
+3. Variables in the casecross over datasets raise some questions. For example, what is the difference from the varialbes geo_smk_pm and

### 2017-11-06 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
added table 1 function.

@@
 
-```{r health data import}
+```{r health data import, message=F, warning=F}
 # importing case-crossover dataframes 
@@
 # asthma
-asthma_path <- paste0("./data/health/2017-oregon_casecross_asthma.csv")
+asthma_path <- paste0("../../data/health/2017-oregon_casecross_asthma.csv")
 asthma_df <- read_csv(asthma_path)
@@
 # cvd 
-cvd_path <- paste0("./data/health/2017-oregon_casecross_cvd.csv")
+cvd_path <- paste0("../../data/health/2017-oregon_casecross_cvd.csv")
 cvd_df <- read_csv(cvd_path)
@@
 outcome_df_list <- list(asthma_df, cvd_df)
+
+# name list elements in order I'd like to present in document
+outcome_names <- c("SABA", "Respiratory", "Asthma", "COPD", "Pneumonia", 
+  "Acute Bronchitis", "Cardiovascular Disease", "Arrhythmia", 
+  "Cerebrovascular Disease", "Heart Failure", "Ischemic Heart Disease", 
+  "Myocardial Infarction")
+
+names(outcome_df_list) <- c( 'Asthma', 'CVD')
 ```
@@
 
-Descriptive table will go here.
+Descriptive table will go here. Jingyang, I'll have you add the other dataframes to analyze. I provided an example of Asthma and CVD.
 
 ```{r descriptive table}
-# name list elements
-names(outcome_df_list) <- c( 'Asthma', 'CVD')
+# creating a row calculation function
+table1_fun <- function(data_list){
+  # output of total n
+  total_n <- data_list %>% 
+    filter(outcome == 1) %>%
+    summarise(total_n = n())
 
   # output row of summary numbers
-test_row <- asthma_df %>% 
+  age_cat_n <- data_list %>% 
     filter(outcome == 1) %>%
@@
     mutate(age_cat = ifelse(age_ind == 0, "age_under15",
-                    ifelse(age_ind == 1, "age_15to65", "age_over65")))
+                      ifelse(age_ind == 1, "age_15to65", "age_over65")),
+           # preserve factor order
+           age_cat = parse_factor(age_cat, levels = age_cat)) %>% 
+    select(-age_ind) %>% 
+    spread(age_cat, n)
+  # sex n
+  sex_n <- data_list %>% 
+    filter(outcome == 1) %>% 
+    group_by(gender) %>% 
+    summarise(n = n()) %>% 
+    filter(gender != "U") %>% 
+    spread(gender, n)
+  # bind all rows together
+  outcome_n_row <- bind_cols(total_n, age_cat_n, sex_n) %>% 
+    mutate(perc_15 = round(age_under15/total_n,2), 
+           perc_15to65 = round(age_15to65/total_n,2),
+           perc_65 = round(age_over65/total_n,2),
+           perc_F = round(F/total_n,2),
+           perc_M = round(M/total_n,2)) %>% 
+    select(total_n, age_under15, perc_15, age_15to65, perc_15to65, 
+           age_over65, perc_65, F, perc_F, M, perc_M)
+return(outcome_n_row)
+} # end function
 
-  forcats::as_factor(age_ind, levels = c("age_under15", "age_15to65", "age_over65"))
+# outcome names
+outcome_names <- names(outcome_df_list) %>% 
+  data_frame() %>% 
+  rename(outcome = ".")
 
-glimpse(test_ro)
+# apply custom function to generate descriptive table
+table_vals <- outcome_df_list %>% 
+  # map to dataframe
+  map_dfr(table1_fun) 
+# descriptive table, bind outcome names and values
+descriptive_table <- bind_cols(outcome_names, table_vals)
 
-descriptive_table <- outcome_df_list %>% 
-  lmap()
-
-?lmap
-
-mtcars %>%
-  split(.$cyl) %>%
-  map(~ lm(mpg ~ wt, data = .x)) %>%
-  map(summary) %>%
-  map_dbl("r.squared")
-
-# Use map_lgl(), map_dbl(), etc to reduce to a vector.
-# * list
-mtcars %>% map(sum)
-# * vector
-mtcars %>% map_dbl(sum)
+# output table
+knitr::kable(descriptive_table, 
+  caption = "Descriptive charactersitics of ED/urgent care visits")
 ```

### 2017-11-06 r_scripts/analysis/model_dl_spline_knots.Rmd

--- 2017-08-08 r_scripts/analysis/oregon_medication.Rmd
@@
 ---
-title: "The GWR Smoke's effect on Oregon 2013 Medication of Beta-2 agonists"
+title: "Oregon distributed lag model best fit degree of freedom"
 author: "Jingyang Liu"
-date: "May 17, 2017"
+date: "2017/11/6"
 output: html_document
@@
 library(sp)
+library(cowplot)
+library(geofacet)
 
@@
 
-# server
-library(dplyr)
-library(data.table)
-library(readxl) 
-library(readr)
-
 ```
 
-## Overview (1, 2)
-
-1.
-Intro: We found an association with wildfire smoke and respiratory outcomes in Washington state 2012 using a novel estimate of smoke concentration, geo-graphically weighted ridge regression (GWR).
-We want to know if smoke impacts other health outcomes, specifically respiratory rescue medications (beta 2 agonists) (abbreviate to SABA)
-
-2. Research question: We evaluated the association between smoke concentrations using the GWR method and SABA refills in Oregon state during the 2013 wildfire season.
-
-We had previously shown GWR to be a good estimate of wildfire smoke concentrations, and showed associations with respiratory outcomes (including asthma) in Washington state.
-We want to replicate results for Oregon and also look at beta-2 agonist pharmacy fills.
-
-For this research I have the Oregon All Payer All Claims (APAC) data set in 2013.
-I want to research on the effects of wildfire GWR smoke on beta-2 agonist pharmacy fill, a medication primarily used as a rescue medication for persons with asthma or COPD.
-NDC (National Drug Code) is the variable in the data set that the data set offered.
-The standard NDC I use the link below, which is a ndc list in 2013 and fits my original data set.
-(http://ww2.ncqa.org/hedis-quality-measurement/hedis-measures/hedis-2013/hedis-2013-final-ndc-lists)
-
-The data cleaning and creation of the casecross-over dataframes are huge computation tasks, so they are done with server, the following work is done with local R.
-We analyze the case crossover results with overall, age strarified and sex stratified categories for the pharmacy fills, and then we also analyze with the distributed lag method, and check with time series analysis.
-
-## Method Description (3)
-
-3.Methods: use case-crossover for all health outcome figures and results. Make sure smoke is assigned via zipcode. For the descriptive of smoke, I think the county-level for now. We may move to zip.
-
-In these comparisons, we examine GWR method of smoke PM2.5 estimations and associations with health outcomes using a time-stratified case-crossover study design.
-Pharmacy outcomes for a patient with a primary diagnosis of all pharmacy records from the state, and it just so happens we also have APAC data for people that were billed for a cardiopulmonary event that we can link to the pharmacy claim, and their date of admission (index time) were identified.
-We then created counterfactual observations for each patient on the same day of the week for the entire wildfire season (May 1st to Sep 30th, 2013).
-For patients who have a index time before May 1, 2013 or after September 30, 2013, their referent observations before or after these dates will be excluded as I will not be able to assign estimates of PM2.5 to these referent observations.
-
-1. Data cleaning for health outcomes
-
-There are about 77 million claims in the total data set, of which 23 million are pharmacy records denoted by the NDC variable.
-So I choose the claims in Oregon State with ndc existing, which is about 23 million claims.
-Also, the POS (Place of service code) all show "pharmacy" type, comparing to the previous health disease outcomes we limit the "emergency visit" type or "urgent care" type.
-
-For ndc, I choose the people the person was billed for filling a beta2 agonist prescription, which we use as a proxy for use to form the sample.
-Because the ndc csv file has more information than what I need, I filter the ndc file by SQLite Studio (code and steps not attached) to a file only containing the beta 2 agonists medication information.
-
-After choosing the claims with beta 2 agonists, the data set is reduced to 232,665 observations.
-After first recorded pharmacy fill and limit the wildfire season dates (May 1st, 2013 to September 30th, 2013), the data set is the final data set used for analysis and contained 24,538 beta2 agonist records.
-
-```{r import filtered original data and make index, eval=FALSE, echo=FALSE}
-### Import whole data set ------------------------------------------------------
-read_path <- paste0("../../../data/data_original/gan_episodes_of_care.txt")
-start_time <- Sys.time()
-oregon_df <- fread(read_path, sep = "|", showProgress = T) 
-stop_time <-  Sys.time() - start_time 
-# time it took
-stop_time # 7 mins
-
-### Import NDC data set --------------------------------------------------------
-read_path <- paste0("../../../data/data_new/oregon_ndc.csv")
-oregon_ndc <- read_csv(read_path)
-dim(oregon_ndc) # 23330714 obs
-
-## Import Asthma NDC table
-read_path2 <- paste0("../../../data/data_new/ndc_2013_beta_2_agonists.csv")
-inhaler_ndc <- read_csv(read_path2, col_names = FALSE)
-
-colnames(inhaler_ndc) <- c("ndc_code", "brand_name", "generic_product_name", 
-                          "route", "category", "drug_id")
-
-# convert to vector
-ndc_codes <- as.vector(as.matrix(inhaler_ndc$ndc_code))
-
-oregon_beta <- oregon_ndc %>%
-  filter(ndc %in% ndc_codes) # 232665, unique personkey 83787
-
-oregon_beta_df <- oregon_beta %>% 
-  group_by(personkey) %>% 
-  arrange(personkey, fromdate, line) %>% 
-  mutate(num_visit = dense_rank(fromdate)) %>%
-  arrange(personkey, num_visit) %>% 
-  select(personkey, clmid, num_visit) %>% 
-  unique() %>% 
-  full_join(oregon_beta, by = c("personkey", "clmid")) %>% 
-  arrange(personkey, clmid, line, fromdate) %>% 
-  filter(row_number() == 1) %>%
-  # add new transverted from date
-  mutate(dates = as.Date(fromdate, "%m/%d/%Y")) # 83787 (same)
-
-oregon_df_ndc <- oregon_df %>%
-  filter(personkey %in% oregon_beta$personkey) # 8909892
-
-### Import ICD9 code
-icd9_key <- read_excel("../../../data/data_original/CMS32_DESC_LONG_SHORT_DX.xlsx") %>% 
-  # rename the terrible variable names
-  select(dx_code = 1, long_desc = 2, short_desc = 3)
-
-# sort by icd9 code add row variable 
-icd9_key$X <- NULL
-icd9_key <- arrange(icd9_key, dx_code) %>%
-  mutate(n = as.numeric(row.names(icd9_key)))
-
-### without ICD-9
-or_ndc_no_icd <- oregon_df_ndc %>%
-  filter(dx1=="*NULL*" &
-         dx2=="*NULL*" &
-         dx3=="*NULL*" &
-         dx4=="*NULL*" &
-         dx5=="*NULL*" &
-         dx6=="*NULL*" &
-         dx7=="*NULL*" &
-         dx8=="*NULL*" &
-         dx9=="*NULL*" &
-         dx10=="*NULL*" &
-         dx11=="" &
-         dx12=="" &
-         dx13=="" )
-
-length(or_ndc_no_icd$personkey) # 3656720 claims
-length(unique(or_ndc_no_icd$personkey)) # 83787
-
-or_no_ndc_no_icd <- oregon_df %>%
-  filter(dx1=="*NULL*" &
-         dx2=="*NULL*" &
-         dx3=="*NULL*" &
-         dx4=="*NULL*" &
-         dx5=="*NULL*" &
-         dx6=="*NULL*" &
-         dx7=="*NULL*" &
-         dx8=="*NULL*" &
-         dx9=="*NULL*" &
-         dx10=="*NULL*" &
-         dx11=="" &
-         dx12=="" &
-         dx13=="") %>%
-  filter(ndc=="" |
-         ndc=="*NULL*") # 1057022 missing in total data set
-
-or_ndc_all_miss <- oregon_df_ndc %>%
-  filter(dx1=="*NULL*" &
-         dx2=="*NULL*" &
-         dx3=="*NULL*" &
-         dx4=="*NULL*" &
-         dx5=="*NULL*" &
-         dx6=="*NULL*" &
-         dx7=="*NULL*" &
-         dx8=="*NULL*" &
-         dx9=="*NULL*" &
-         dx10=="*NULL*" &
-         dx11=="" &
-         dx12=="" &
-         dx13=="") %>%
-  filter(ndc=="" |
-         ndc=="*NULL*") # 20269 missing in total data set
-
-or_ndc_all_miss_id <- or_ndc_asthma_df %>%
-  filter(personkey %in% unique(or_ndc_copd_df$personkey)) %>% # 7279 person
-  select(personkey)
-
-oregon_ndc_dx <- oregon_ndc %>%
-    filter(dx1!="*NULL*" |
-         dx2!="*NULL*" |
-         dx3!="*NULL*" |
-         dx4!="*NULL*" |
-         dx5!="*NULL*" |
-         dx6!="*NULL*" |
-         dx7!="*NULL*" |
-         dx8!="*NULL*" |
-         dx9!="*NULL*" |
-         dx10!="*NULL*" |
-         dx11!="" |
-         dx12!="" |
-         dx13!="" ) # 0 claims
-
-### asthma
-which(icd9_key$dx_code == '49300') # start of asthma is row 5206
-which(icd9_key$dx_code == '49392') # end of asthma is row 5219
-
-# limit just to asthma code and just the diagnosis column
-icd9_check <- icd9_key %>%
-  filter(n >= 5206 & n <= 5219) 
-icd9_check
-
-asthma_icd9 <- filter(icd9_key, n >= 5206 & n <= 5219) %>%
-  select(dx_code)
-# convert to vector
-asthma_icd9 <- as.vector(as.matrix(asthma_icd9))
-
-# now can I make a new variable, asthma1, that indicates an asthma claim?
-or_ndc_asthma_df <- oregon_df_ndc %>%
-  filter(dx1 %in% asthma_icd9 |
-         dx2 %in% asthma_icd9 |
-         dx3 %in% asthma_icd9 |
-         dx4 %in% asthma_icd9 |
-         dx5 %in% asthma_icd9 |
-         dx6 %in% asthma_icd9 |
-         dx7 %in% asthma_icd9 |
-         dx8 %in% asthma_icd9 |
-         dx9 %in% asthma_icd9 |
-         dx10 %in% asthma_icd9 |
-         dx11 %in% asthma_icd9 |
-         dx12 %in% asthma_icd9 |
-         dx13 %in% asthma_icd9) # 375451, unique personkey: 45165
-
-### COPD
-copd_icd9 <- c('490', '4910','4911','49120','49121','49122','4918','4919', '4920',
-               '4928', '4940', '4941', '496')
-
-# now can I make a new variable, asthma1, that indicates an asthma claim?
-or_ndc_copd_df <- oregon_df_ndc %>%
-  filter(dx1 %in% copd_icd9 |
-         dx2 %in% copd_icd9 |
-         dx3 %in% copd_icd9 |
-         dx4 %in% copd_icd9 |
-         dx5 %in% copd_icd9 |
-         dx6 %in% copd_icd9 |
-         dx7 %in% copd_icd9 |
-         dx8 %in% copd_icd9 |
-         dx9 %in% copd_icd9 |
-         dx10 %in% copd_icd9 |
-         dx11 %in% copd_icd9 |
-         dx12 %in% copd_icd9 |
-         dx13 %in% copd_icd9) # 361738, unique personkey: 18626
-
-### asthma or copd
-or_or_asthma_copd <- oregon_df_ndc %>%
-  filter(dx1 %in% asthma_icd9 |
-         dx2 %in% asthma_icd9 |
-         dx3 %in% asthma_icd9 |
-         dx4 %in% asthma_icd9 |
-         dx5 %in% asthma_icd9 |
-         dx6 %in% asthma_icd9 |
-         dx7 %in% asthma_icd9 |
-         dx8 %in% asthma_icd9 |
-         dx9 %in% asthma_icd9 |
-         dx10 %in% asthma_icd9 |
-         dx11 %in% asthma_icd9 |
-         dx12 %in% asthma_icd9 |
-         dx13 %in% asthma_icd9 |
-         dx1 %in% copd_icd9 |
-         dx2 %in% copd_icd9 |
-         dx3 %in% copd_icd9 |
-         dx4 %in% copd_icd9 |
-         dx5 %in% copd_icd9 |
-         dx6 %in% copd_icd9 |
-         dx7 %in% copd_icd9 |
-         dx8 %in% copd_icd9 |
-         dx9 %in% copd_icd9 |
-         dx10 %in% copd_icd9 |
-         dx11 %in% copd_icd9 |
-         dx12 %in% copd_icd9 | 
-         dx13 %in% copd_icd9) # 722107 , unique personkey: 56512
-
-# 45165+ 18626 - 56512
-# [1] 7279 on both asthma and copd
-
-### Both asthma and copd
-length(unique(or_or_asthma_copd$personkey))
-length(unique(or_ndc_asthma_df$personkey))
-length(unique(or_ndc_copd_df$personkey))
-
-or_with_icd <- oregon_df_ndc %>%
-  filter(!dx1=="*NULL*" |
-         !dx2=="*NULL*" |
-         !dx3=="*NULL*" |
-         !dx4=="*NULL*" |
-         !dx5=="*NULL*" |
-         !dx6=="*NULL*" |
-         !dx7=="*NULL*" |
-         !dx8=="*NULL*" |
-         !dx9=="*NULL*" |
-         !dx10=="*NULL*" |
-         !dx11=="" |
-         !dx12=="" |
-         !dx13=="" )
-
-asthma_copd_both_id <- or_ndc_asthma_df %>%
-  filter(personkey %in% unique(or_ndc_copd_df$personkey)) %>% # 7279 person
-  select(personkey) %>%
-  unique() %>%
-  mutate(index="both")
-
-### Only asthma
-asthma_only_id <- or_ndc_asthma_df %>%
-  filter(!personkey %in% asthma_copd_both_id$personkey) %>% # 37886 person
-  select(personkey) %>%
-  unique() %>%
-  mutate(index="asthma")
-
-### Only copd
-copd_only_id <- or_ndc_copd_df %>%
-  filter(!personkey %in% asthma_copd_both_id$personkey) %>% # 11347 person
-  select(personkey) %>%
-  unique() %>%
-  mutate(index="copd")
-
-### Other
-other_disease_id <- oregon_df_ndc %>%
-  filter(!personkey %in% or_or_asthma_copd$personkey) %>% # 27275 person
-  select(personkey) %>%
-  unique() %>%
-  mutate(index="other")
-
-id_index <- bind_rows(asthma_copd_both_id, asthma_only_id, copd_only_id, other_disease_id)
-
-## Join the index with beta 2 ndc claims
-oregon_beta_index <- oregon_beta %>%
-  left_join(id_index, by="personkey")
-summary(as.factor(oregon_beta_index$index))
-
-write_path3 <- paste0('../../../data/data_new/',
-                     'oregon_beta_ndc_index.csv')
-write_csv(oregon_beta_index, write_path3)
-
-## Join with original data and indicator for 4 categories
-oregon_ndc_index <- oregon_df_ndc %>%
-  left_join(id_index, by="personkey")
-summary(as.factor(oregon_ndc_index$index))
-
-write_path4 <- paste0('../../../data/data_new/',
-                     'oregon_beta_ndc_index_full.csv')
-write_csv(oregon_ndc_index, write_path4)
-
-### Check
-or_ndc_asthma_copd_sum <- oregon_ndc_index %>%
-  mutate(asthma1 = ifelse(dx1 %in% asthma_icd9, 1, 0),
-         asthma2 = ifelse(dx2 %in% asthma_icd9, 1, 0),
-         asthma3 = ifelse(dx3 %in% asthma_icd9, 1, 0),
-         asthma4 = ifelse(dx4 %in% asthma_icd9, 1, 0),
-         asthma5 = ifelse(dx5 %in% asthma_icd9, 1, 0),
-         asthma6 = ifelse(dx6 %in% asthma_icd9, 1, 0),
-         asthma7 = ifelse(dx7 %in% asthma_icd9, 1, 0),
-         asthma8 = ifelse(dx8 %in% asthma_icd9, 1, 0),
-         asthma9 = ifelse(dx9 %in% asthma_icd9, 1, 0),
-         asthma10 = ifelse(dx10 %in% asthma_icd9, 1, 0),
-         asthma11 = ifelse(dx11 %in% asthma_icd9, 1, 0),
-         asthma12 = ifelse(dx12 %in% asthma_icd9, 1, 0),
-         asthma13 = ifelse(dx13 %in% asthma_icd9, 1, 0),
-         # sum up the asthma indicators
-         asthma_sum = (asthma1 + asthma2 + asthma3 + asthma4 + asthma5 + asthma6 + asthma7 + asthma8 + asthma9 + asthma10 + asthma11 + asthma12 + asthma13),
-         asthma_dx = ifelse(asthma_sum>0, asthma_sum, 0)
-         ) %>% # end of mutate asthma
-  mutate(copd1 = ifelse(dx1 %in% copd_icd9, 1, 0),
-         copd2 = ifelse(dx2 %in% copd_icd9, 1, 0),
-         copd3 = ifelse(dx3 %in% copd_icd9, 1, 0),
-         copd4 = ifelse(dx4 %in% copd_icd9, 1, 0),
-         copd5 = ifelse(dx5 %in% copd_icd9, 1, 0),
-         copd6 = ifelse(dx6 %in% copd_icd9, 1, 0),
-         copd7 = ifelse(dx7 %in% copd_icd9, 1, 0),
-         copd8 = ifelse(dx8 %in% copd_icd9, 1, 0),
-         copd9 = ifelse(dx9 %in% copd_icd9, 1, 0),
-         copd10 = ifelse(dx10 %in% copd_icd9, 1, 0),
-         copd11 = ifelse(dx11 %in% copd_icd9, 1, 0),
-         copd12 = ifelse(dx12 %in% copd_icd9, 1, 0),
-         copd13 = ifelse(dx13 %in% copd_icd9, 1, 0),
-         # sum up the pneumonia indicators
-         copd_sum = (copd1 + copd2 + copd3 + copd4 +copd5 + copd6 + copd7 + copd8 + copd9 +copd10 + copd11 + copd12 + copd13),
-         copd_dx = ifelse(copd_sum>0, copd_sum, 0)
-  ) #end of mutate of copd
-
-or_ndc_check <- or_ndc_asthma_copd_sum %>%
-  select(personkey, index, asthma_dx, copd_dx)
-summary(as.factor(or_ndc_check$asthma_dx))
-
-or_asthma <- or_ndc_asthma_copd_sum %>%
-  filter(index=="asthma")
-or_copd <- or_ndc_asthma_copd_sum %>%
-  filter(index=="copd")
-or_both <- or_ndc_asthma_copd_sum %>%
-  filter(index=="both")
-or_other <- or_ndc_asthma_copd_sum %>%
-  filter(index=="other")
-
-summary(as.factor(or_other$dx1))
-
-## check for one sample
-summary(as.factor(or_asthma$personkey))
-summary(as.factor(or_copd$personkey))
-summary(as.factor(or_both$personkey))
-summary(as.factor(or_other$personkey))
-
-or_asthma_sample <- or_asthma %>%
-  filter(personkey == 10101244) 
-which(or_asthma_sample$asthma_dx!=0)
-
-or_copd_sample  <- or_copd %>%
-  filter(personkey == 13564804)
-
-or_both_sample  <- or_both %>%
-  filter(personkey == 12979399)
-
-or_other_sample  <- or_other %>%
-  filter(personkey == 10289787)
-
-or_asthma_miss <- or_asthma %>%
-  filter(dx1=="*NULL*" &
-         dx2=="*NULL*" &
-         dx3=="*NULL*" &
-         dx4=="*NULL*" &
-         dx5=="*NULL*" &
-         dx6=="*NULL*" &
-         dx7=="*NULL*" &
-         dx8=="*NULL*" &
-         dx9=="*NULL*" &
-         dx10=="*NULL*" &
-         dx11=="" &
-         dx12=="" &
-         dx13=="") %>%
-  filter(ndc=="" |
-         ndc=="*NULL*") # 89482 claims missing in asthma data set
-
-or_copd_miss <- or_copd %>%
-  filter(dx1=="*NULL*" &
-         dx2=="*NULL*" &
-         dx3=="*NULL*" &
-         dx4=="*NULL*" &
-         dx5=="*NULL*" &
-         dx6=="*NULL*" &
-         dx7=="*NULL*" &
-         dx8=="*NULL*" &
-         dx9=="*NULL*" &
-         dx10=="*NULL*" &
-         dx11=="" &
-         dx12=="" &
-         dx13=="") %>%
-  filter(ndc=="" |
-         ndc=="*NULL*") # 18941 claims missing in copd data set
-
-or_both_miss <- or_both %>%
-  filter(dx1=="*NULL*" &
-         dx2=="*NULL*" &
-         dx3=="*NULL*" &
-         dx4=="*NULL*" &
-         dx5=="*NULL*" &
-         dx6=="*NULL*" &
-         dx7=="*NULL*" &
-         dx8=="*NULL*" &
-         dx9=="*NULL*" &
-         dx10=="*NULL*" &
-         dx11=="" &
-         dx12=="" &
-         dx13=="") %>%
-  filter(ndc=="" |
-         ndc=="*NULL*") # 17963 claims missing in both data set
-
-or_other_miss <- or_other %>%
-  filter(dx1=="*NULL*" &
-         dx2=="*NULL*" &
-         dx3=="*NULL*" &
-         dx4=="*NULL*" &
-         dx5=="*NULL*" &
-         dx6=="*NULL*" &
-         dx7=="*NULL*" &
-         dx8=="*NULL*" &
-         dx9=="*NULL*" &
-         dx10=="*NULL*" &
-         dx11=="" &
-         dx12=="" &
-         dx13=="") %>%
-  filter(ndc=="" |
-         ndc=="*NULL*") # 46971 claims missing in other data set (5848 personkey)
-
-or_other_not_miss <- or_other %>%
-  filter(dx1!="*NULL*" |
-         dx2!="*NULL*" |
-         dx3!="*NULL*" |
-         dx4!="*NULL*" |
-         dx5!="*NULL*" |
-         dx6!="*NULL*" |
-         dx7!="*NULL*" |
-         dx8!="*NULL*" |
-         dx9!="*NULL*" |
-         dx10!="*NULL*" |
-         dx11!="" |
-         dx12!="" |
-         dx13!="" |
-         ndc!="" |
-         ndc!="*NULL*") # 20269 missing in total data set
-
-length(unique(or_other_not_miss$personkey)) # [1] 27275
-
-dim(or_asthma_miss)
-dim(or_copd_miss)
-dim(or_both_miss)
-dim(or_other_miss)
-
-sum(or_ndc_all_miss_id %in% unique(or_other$personkey))
-
-# write_path <- paste0('../../../data/data_new/',
-#                      'oregon_asthma_sample.csv')
-# write_csv(or_asthma_sample, write_path)
-
-# write_path2 <- paste0('../../../data/data_new/',
-#                      'oregon_ndc_with_index.csv')
-
-# write_csv(or_ndc_asthma_copd_sum, write_path2)
-
-```
-
-```{r choose ndc, eval=FALSE, echo=FALSE}
-
-# limit to the unique personkey different days' first visit
-oregon_asthma_df <- oregon_asthma %>% 
-  group_by(personkey) %>% 
-  arrange(personkey, fromdate, line) %>% 
-  mutate(num_visit = dense_rank(fromdate)) %>%
-  arrange(personkey, num_visit) %>% 
-  select(personkey, clmid, num_visit) %>% 
-  unique() %>% 
-  full_join(oregon_asthma, by = c("personkey", "clmid")) %>% 
-  arrange(personkey, clmid, line, fromdate) %>% 
-  filter(row_number() == 1) %>%
-  # add new transverted from date
-  mutate(dates = as.Date(fromdate, "%m/%d/%Y")) %>%
-  filter(dates >= '2013-05-01' & 
-         dates <= '2013-09-30') # 24538
-
-write_path2 <- paste0("../../../data/data_new/oregon_asthma_ndc.csv")
-write_csv(oregon_asthma_df, write_path2)
-
-```
-
-2. Exposure methods for smoke model
-
-For our research question, we focus on the GWR method (geographically weighted ridge regression).
-For Geo-Weighted Regression model, with 'smk' designator, I subtracted off the 'Background' estimates of smoke, which I believe are the monthly averages of PM2.5 for a given grid.
-However, the 'Background' is the mean of total period from May 1st to Sep 30th 2013 so I have to assign the mean for every day.
-
-3. Analytic method
-
-We use the conditional logistic regression model using the survival package in R.
-Each conditional logistic regression model accounts for the subject, and adjusts for temperature from the WRF-Chem model.
-The conditional logistic regression model \beta represents the change in risk of an pharmacy event associated with a short-term unit increase in exposure, and can be calculated as an average difference between exposure at the index time and a weighted average of exposure at all times in the referent window.
-
-```{r import original data, eval=FALSE, echo=FALSE}
-### Import whole data set ------------------------------------------------------
-read_path <- paste0("../../../data/data_original/gan_episodes_of_care.txt")
-start_time <- Sys.time()
-oregon_df <- fread(read_path, sep = "|", showProgress = T) 
-stop_time <-  Sys.time() - start_time 
-# time it took
-stop_time # 7 mins
-
-## Basic cleaning
-start_time <- Sys.time()
-oregon_df2 <- oregon_df %>% 
-  filter((!ndc=="")&(!ndc=="*NULL*")) %>% 
-  # filter State is Oregon
-  filter(STATE=="OR")
-stop_time <-  Sys.time() - start_time 
-# time it took
-stop_time # 40 secs
-
-write_path <- paste0('../../../data/data_new/',
-                     'oregon_ndc.csv')
-write_csv(oregon_df2, write_path)
-```
-
-### Time-Stratified Case-Crossover (Population-Weighted PM~2.5~ Assigned by Zip Code)
-*referents same day of week within fire season May - September*
-
-The time-stratified casecross-over design uses subjects as their own controls.
-For each pharmacy claim, I also created counterfactual observations on the the same day of the week over the study period for each subject to indicate the outcome for this day is "1".
-Then I make the dates week by week before this date, and after this date in the whole wildfire season, and indicate these dates' outcomes are "0".
-Then I can use these outcomes for the following model building.
-
-```{r casecrossover, eval=FALSE, echo=FALSE}
-### Casecrossover study
-read_path3 <- paste0("../../../data/data_new/medication/oregon_asthma_ndc.csv")
-oregon_asthma_df <- read_csv(read_path3)
-
-outcome_id <- oregon_asthma_df %>%
-  # arrange with dates
-  arrange(dates) %>%
-  mutate(id = seq(1, nrow(.), by = 1))
-
-# create dataset to populate
-id_date_df <- data_frame()
-
-# begin second loop to create counterfactual observations for each case subject
-for (k in 1:nrow(outcome_id)){
-
-  # find the replicate times of weeks
-  dates_l <- outcome_id[[k,74]] 
-  n1 <- 0
-  d=as.Date("2013-05-01")
-  i=1
-  while (dates_l >= "2013-05-01"){
-    dates_l <- dates_l - 7
-    d[i] = dates_l
-    i = i+1
-    n1 = n1+1
-  }
-  d[1:n1-1] # shows character(0) when the first week
-  n1-1
-
-  dates_l <- outcome_id[[k,74]] 
-  n2=0
-  e=as.Date("2013-09-30")
-  j=1
-  while (dates_l <= "2013-09-30"){
-    dates_l <- dates_l + 7
-    e[j]=dates_l
-    j=j+1
-    n2 = n2 + 1
-  }
-  e[1:n2-1] # shows character(0) when the last week
-  n2-1
-
-  # replicate covariates length of counterfactual dates
-  # and make conuterfactual dates
-  if (n1==1){
-    cov_df <- do.call("bind_rows", replicate(n1+n2-1, outcome_id[k,],simplify = F))
-    cov_df$dates <- c(outcome_id[[k,74]], e[1:n2-1])
-  } else if (n2==1){
-    cov_df <- do.call("bind_rows", replicate(n1+n2-1, outcome_id[k,],simplify = F))
-    cov_df$dates <- c(outcome_id[[k,74]], d[1:(n1-1)])
-  }else{
-    cov_df <- do.call("bind_rows", replicate(n1+n2-1, outcome_id[k,],simplify = F))
-    cov_df$dates <- c(outcome_id[[k,74]], d[1:(n1-1)], e[1:n2-1])
-  }
-
-  # bind unique id and date of the year with covariates
-  id_date <- bind_cols(cov_df)
-  # iteration which binds rows of unique ids
-  id_date_df <- bind_rows(id_date_df, id_date)
-}
-
-outcome_casecross <- id_date_df %>%
-  mutate(outcome = ifelse(dates == as.Date(fromdate, "%m/%d/%Y"), 1, 0)) %>%
-  arrange(id, dates) # order by id and date
-
-outcome_casecross <- id_date_df %>%
-  mutate(outcome = ifelse(dates == as.Date(fromdate, "%m/%d/%Y"), 1, 0)) %>%
-  arrange(personkey, dates) # order by id and date
-
-write_path2 <- paste0("../../../data/data_new/oregon_ndc_casecrossover.csv")
-write_csv(outcome_casecross, write_path2)
-
-```
-
-To get the time stratified data set, I create the lag variables that take smoke values from the previous days for zip codes, join the zipcode level populatoin-weighted pm to the data set, and create age, sex, month and season index.
-
-```{r time stratified zip for casecrossover, eval=FALSE, echo=FALSE}
-### Time stratified ------------------------------------------------------------
-# read in zipcode level populatoin-weighted pm
-read_path5 <- paste0('C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire_new/data/Oregon_PM/zip_pm_to_merge_with_acap.csv')
-
-zip_smoke <- read_csv(read_path5) # 63801 rows
-
-# descriptives of the two smoke datasets
-summary(zip_smoke)
-
-# Zipcode PM2.5 estimates
-# create lag variables that take smoke values from n previous days for zipcodes
-zip_smoke_w_lag <- zip_smoke %>% arrange(ZIPCODE, date) %>%
-  # group by zipcode
-  group_by(ZIPCODE) %>% 
-  # wrf
-  mutate(wrf_f_pm_lag1 = lag(wrf_f_pm, 1, order_by = ZIPCODE), 
-         wrf_f_pm_lag2 = lag(wrf_f_pm, 2, order_by = ZIPCODE),
-         wrf_f_pm_lag3 = lag(wrf_f_pm, 3, order_by = ZIPCODE),
-         wrf_f_pm_lag4 = lag(wrf_f_pm, 4, order_by = ZIPCODE),
-         wrf_f_pm_lag5 = lag(wrf_f_pm, 5, order_by = ZIPCODE),
-         wrf_f_pm_lag6 = lag(wrf_f_pm, 6, order_by = ZIPCODE),
-         wrf_f_pm_lag7 = lag(wrf_f_pm, 7, order_by = ZIPCODE),
-         # wrf no fire lag
-         wrf_nf_pm_lag1 = lag(wrf_nf_pm, 1, order_by = ZIPCODE),
-         wrf_nf_pm_lag2 = lag(wrf_nf_pm, 2, order_by = ZIPCODE),
-         wrf_nf_pm_lag3 = lag(wrf_nf_pm, 3, order_by = ZIPCODE),
-         wrf_nf_pm_lag4 = lag(wrf_nf_pm, 4, order_by = ZIPCODE),
-         wrf_nf_pm_lag5 = lag(wrf_nf_pm, 5, order_by = ZIPCODE),
-         wrf_nf_pm_lag6 = lag(wrf_nf_pm, 6, order_by = ZIPCODE),
-         wrf_nf_pm_lag7 = lag(wrf_nf_pm, 7, order_by = ZIPCODE),
-         # wrf_smk_pm
-         wrf_smk_pm_lag1 = lag(wrf_smk_pm, 1, order_by = ZIPCODE),
-         wrf_smk_pm_lag2 = lag(wrf_smk_pm, 2, order_by = ZIPCODE),
-         wrf_smk_pm_lag3 = lag(wrf_smk_pm, 3, order_by = ZIPCODE),
-         wrf_smk_pm_lag4 = lag(wrf_smk_pm, 4, order_by = ZIPCODE),
-         wrf_smk_pm_lag5 = lag(wrf_smk_pm, 5, order_by = ZIPCODE),
-         wrf_smk_pm_lag6 = lag(wrf_smk_pm, 6, order_by = ZIPCODE),
-         wrf_smk_pm_lag7 = lag(wrf_smk_pm, 7, order_by = ZIPCODE),
-         # geo weighted pm
-         geo_wt_pm_lag1 = lag(geo_wt_pm, 1, order_by = ZIPCODE),
-         geo_wt_pm_lag2 = lag(geo_wt_pm, 2, order_by = ZIPCODE),
-         geo_wt_pm_lag3 = lag(geo_wt_pm, 3, order_by = ZIPCODE),
-         geo_wt_pm_lag4 = lag(geo_wt_pm, 4, order_by = ZIPCODE),
-         geo_wt_pm_lag5 = lag(geo_wt_pm, 5, order_by = ZIPCODE),
-         geo_wt_pm_lag6 = lag(geo_wt_pm, 6, order_by = ZIPCODE),
-         geo_wt_pm_lag7 = lag(geo_wt_pm, 7, order_by = ZIPCODE),
-         # krig pm
-         krig_pm_lag1 = lag(krig_pm, 1, order_by = ZIPCODE),
-         krig_pm_lag2 = lag(krig_pm, 2, order_by = ZIPCODE),
-         krig_pm_lag3 = lag(krig_pm, 3, order_by = ZIPCODE),
-         krig_pm_lag4 = lag(krig_pm, 4, order_by = ZIPCODE),
-         krig_pm_lag5 = lag(krig_pm, 5, order_by = ZIPCODE), 
-         krig_pm_lag6 = lag(krig_pm, 6, order_by = ZIPCODE),
-         krig_pm_lag7 = lag(krig_pm, 7, order_by = ZIPCODE), 
-         # background pm
-         background_pm_lag1 = lag(background_pm, 1, order_by = ZIPCODE),
-         background_pm_lag2 = lag(background_pm, 2, order_by = ZIPCODE),
-         background_pm_lag3 = lag(background_pm, 3, order_by = ZIPCODE),
-         background_pm_lag4 = lag(background_pm, 4, order_by = ZIPCODE),
-         background_pm_lag5 = lag(background_pm, 5, order_by = ZIPCODE), 
-         background_pm_lag6 = lag(background_pm, 6, order_by = ZIPCODE),
-         background_pm_lag7 = lag(background_pm, 7, order_by = ZIPCODE), 
-         # geo_smk_pm 
-         geo_smk_pm_lag1 = lag(geo_smk_pm, 1, order_by = ZIPCODE),
-         geo_smk_pm_lag2 = lag(geo_smk_pm, 2, order_by = ZIPCODE),
-         geo_smk_pm_lag3 = lag(geo_smk_pm, 3, order_by = ZIPCODE),
-         geo_smk_pm_lag4 = lag(geo_smk_pm, 4, order_by = ZIPCODE),
-         geo_smk_pm_lag5 = lag(geo_smk_pm, 5, order_by = ZIPCODE),
-         geo_smk_pm_lag6 = lag(geo_smk_pm, 6, order_by = ZIPCODE),
-         geo_smk_pm_lag7 = lag(geo_smk_pm, 7, order_by = ZIPCODE),
-         # krig smk pm
-         krig_smk_pm_lag1 = lag(krig_smk_pm, 1, order_by = ZIPCODE),
-         krig_smk_pm_lag2 = lag(krig_smk_pm, 2, order_by = ZIPCODE),
-         krig_smk_pm_lag3 = lag(krig_smk_pm, 3, order_by = ZIPCODE),
-         krig_smk_pm_lag4 = lag(krig_smk_pm, 4, order_by = ZIPCODE),
-         krig_smk_pm_lag5 = lag(krig_smk_pm, 5, order_by = ZIPCODE),
-         krig_smk_pm_lag6 = lag(krig_smk_pm, 6, order_by = ZIPCODE),
-         krig_smk_pm_lag7 = lag(krig_smk_pm, 7, order_by = ZIPCODE),
-         # temp
-         wrf_temp_lag1 = lag(wrf_temp, 1, order_by = ZIPCODE),
-         wrf_temp_lag2 = lag(wrf_temp, 2, order_by = ZIPCODE),
-         wrf_temp_lag3 = lag(wrf_temp, 3, order_by = ZIPCODE),
-         wrf_temp_lag4 = lag(wrf_temp, 4, order_by = ZIPCODE),
-         wrf_temp_lag5 = lag(wrf_temp, 5, order_by = ZIPCODE), 
-         wrf_temp_lag6 = lag(wrf_temp, 6, order_by = ZIPCODE),
-         wrf_temp_lag7 = lag(wrf_temp, 7, order_by = ZIPCODE)) %>% 
-  # ungroup by zip
-  ungroup(ZIPCODE) %>% 
-  # attach a zip indicator for each smoke variable
-  setNames(paste(colnames(.), "zip", sep="_"))  %>% 
-  # remove the '_zip' from the zipcode and date variable 
-  rename(ZIPCODE = ZIPCODE_zip, date = date_zip)
-
-read_path4 <- paste0("../data_new/medication/oregon_ndc_casecrossover.csv")
-# read_path4 <- paste0("../../../data/data_new/oregon_ndc_casecrossover.csv") # for server
-or_disease <- read_csv(read_path4)
-
-### try join
-colnames(or_disease)[24] <- c("ZIPCODE")
-colnames(or_disease)[74] <- c("date")
-
-outcome_casecross <- or_disease %>%
-  # indicator for male=0, female=1, unknown = 2
-  mutate(age = 2013-yob,
-         sex_ind =ifelse(gender == "F", 1, 
-                         ifelse(gender == "M", 0, 2)),
-         age_ind = ifelse(age < 15, 0,
-                          ifelse(age >= 15 & age < 65, 1,
-                                 ifelse(age >= 65 & age <=105, 2, NA)))
-  ) %>% # end of mutate 
-  # create variables
-  mutate(day = as.factor(weekdays(fromdate)),
-         day_admit = as.factor(weekdays(date)),
-         month_smk = month(fromdate),
-         month_admit = month(date),
-         season_smk = ifelse(fromdate >= "2013-03-20" &  
-                               fromdate <= "2013-06-21", "spring",
-                             ifelse(fromdate >= "2013-06-22" &  
-                                      fromdate <= "2013-09-22", "summer",
-                                    ifelse(fromdate >= "2013-09-23" & 
-                                             fromdate <= "2013-12-21", "fall", "other"))),
-         season_admit = ifelse(date >= "2013-03-20" &  
-                                 date <= "2013-06-21", "spring",
-                               ifelse(date >= "2013-06-22" &  
-                                        date <= "2013-09-22", "summer",
-                                      ifelse(date >= "2013-09-23" & 
-                                               date <= "2013-12-21", "fall", "other")))) %>%
-  # join with zip-level pm estimates
-  left_join(zip_smoke_w_lag, by = c("date", "ZIPCODE"))%>%
-  arrange(personkey, fromdate) # order by id and fromdate
-
-write_path4 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover_zip.csv")
-write_csv(outcome_casecross, write_path4)
-
-```
-
-### Time stratified for time series (Population-Weighted PM~2.5~ Assigned by County)
-
-I also make time stratified time series data set.
-The original data set only contains the zip code, but for time series, we want to research on the county level.
-Therefore, I assign each zip code with one county which I found online.
-There are 488 zipcodes in total (4 are missing but I use google to find it).
-This is a sensitivity analysis to confirm the casecross over results.
-
-```{r time stratified county for ndc ts, echo=FALSE, include=FALSE, eval=FALSE}
-read_path7 <- paste0("../data_new/medication/oregon_asthma_ndc.csv")
-or_ndc_claim <- read_csv(read_path7)
-
-or_ndc_inhaler <- or_ndc_claim %>%
-  select(personkey, num_visit, clmid, line, ZIP, dates) %>%
-  rename(ZIPCODE = ZIP)
-
-# Join data with names of Oregon counties 
-oregon_fips <- read_csv('../instructions/oregon_FIPS.csv')
-# remove the "County" character
-# factor(oregon_fips$county)
-oregon_fips$county <- gsub(" County", "", as.character(factor(oregon_fips$county)))
-
-oregon_fips <- oregon_fips %>%
-  mutate(st_county_fips = with(oregon_fips, paste0(st_code, fips)))
-
-read_path <- paste0('../instructions/oregon_zip_county.csv')
-or_zip_county <- read_csv(read_path)
-
-# or_zip_county$county_name[which(or_zip_county$county_name=="Hood.River")] <- "Hood River"
-
-or_zip_county <- or_zip_county %>%
-  select(zip, county_name) %>%
-  rename(county = county_name) %>%
-  rename(ZIPCODE = zip) %>%
-  full_join(oregon_fips, by = "county") %>%
-  select(ZIPCODE, county, state, st_county_fips) %>%
-  rename(fips = st_county_fips) # 484
-
-county_new <- c("Washington", "Douglas", "Lane", "Deschutes")
-
-zip_new <- or_ndc_inhaler %>%
-  filter(!(ZIPCODE %in% or_zip_county$ZIPCODE)) %>%
-  select(ZIPCODE) %>%
-  unique() %>%
-  arrange(ZIPCODE) %>%
-  mutate(county = county_new)
-
-# summary(as.factor(zip_new$ZIPCODE))
-# 97003 97471 97475 97703 
-#    48   138     5    16
-
-or_zip_county_new <- or_zip_county %>%
-  select(ZIPCODE, county) %>%
-  bind_rows(zip_new)
-
-write_path <- paste0("../data_new/county_data/zip_to_county_new.csv")
-write_csv(or_zip_county_new, write_path)
-
-read_path9 <- paste0("../data_new/county_data/or_census_pop_est_county.csv")
-or_pop <- read_csv(read_path9)
-
-or_pop <- or_pop %>%
-  rename(county = GEO_display_label, pop = respop72013) %>%
-  select(county, pop) 
-or_pop$county <- gsub(" County, Oregon", "", as.character(factor(or_pop$county)))
-
-### import county data
-read_path_county_pm <- paste0('C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire_new/data_new/county_data/or_county_pop_wt_pm.csv')
-
-county_smoke <- read_csv(read_path_county_pm) # 63801 rows
-
-# descriptives of the two smoke datasets
-summary(county_smoke)
-
-county_smoke$county[which(county_smoke$county=="Hood.River")] <- "Hood River"
-
-# Zipcode PM2.5 estimates
-# create lag variables that take smoke values from n previous days for zipcodes
-county_smoke_w_lag <- county_smoke %>% arrange(county, date) %>%
-  # group by zipcode
-  group_by(county) %>% 
-  # wrf
-  mutate(wrf_f_pm_lag1 = lag(wrf_pm, 1, order_by = county), 
-         wrf_f_pm_lag2 = lag(wrf_pm, 2, order_by = county),
-         wrf_f_pm_lag3 = lag(wrf_pm, 3, order_by = county),
-         wrf_f_pm_lag4 = lag(wrf_pm, 4, order_by = county),
-         wrf_f_pm_lag5 = lag(wrf_pm, 5, order_by = county),
-         wrf_f_pm_lag6 = lag(wrf_pm, 6, order_by = county),
-         wrf_f_pm_lag7 = lag(wrf_pm, 7, order_by = county),
-         # wrf no fire lag
-         wrf_nf_pm_lag1 = lag(wrf_nf_pm, 1, order_by = county),
-         wrf_nf_pm_lag2 = lag(wrf_nf_pm, 2, order_by = county),
-         wrf_nf_pm_lag3 = lag(wrf_nf_pm, 3, order_by = county),
-         wrf_nf_pm_lag4 = lag(wrf_nf_pm, 4, order_by = county),
-         wrf_nf_pm_lag5 = lag(wrf_nf_pm, 5, order_by = county),
-         wrf_nf_pm_lag6 = lag(wrf_nf_pm, 6, order_by = county),
-         wrf_nf_pm_lag7 = lag(wrf_nf_pm, 7, order_by = county),
-         # wrf_smk_pm
-         wrf_smk_pm_lag1 = lag(wrf_smk_pm, 1, order_by = county),
-         wrf_smk_pm_lag2 = lag(wrf_smk_pm, 2, order_by = county),
-         wrf_smk_pm_lag3 = lag(wrf_smk_pm, 3, order_by = county),
-         wrf_smk_pm_lag4 = lag(wrf_smk_pm, 4, order_by = county),
-         wrf_smk_pm_lag5 = lag(wrf_smk_pm, 5, order_by = county),
-         wrf_smk_pm_lag6 = lag(wrf_smk_pm, 6, order_by = county),
-         wrf_smk_pm_lag7 = lag(wrf_smk_pm, 7, order_by = county),
-         # geo weighted pm
-         geo_wt_pm_lag1 = lag(geo_wt_pm, 1, order_by = county),
-         geo_wt_pm_lag2 = lag(geo_wt_pm, 2, order_by = county),
-         geo_wt_pm_lag3 = lag(geo_wt_pm, 3, order_by = county),
-         geo_wt_pm_lag4 = lag(geo_wt_pm, 4, order_by = county),
-         geo_wt_pm_lag5 = lag(geo_wt_pm, 5, order_by = county),
-         geo_wt_pm_lag6 = lag(geo_wt_pm, 6, order_by = county),
-         geo_wt_pm_lag7 = lag(geo_wt_pm, 7, order_by = county),
-         # krig pm
-         krig_pm_lag1 = lag(krig_pm, 1, order_by = county),
-         krig_pm_lag2 = lag(krig_pm, 2, order_by = county),
-         krig_pm_lag3 = lag(krig_pm, 3, order_by = county),
-         krig_pm_lag4 = lag(krig_pm, 4, order_by = county),
-         krig_pm_lag5 = lag(krig_pm, 5, order_by = county), 
-         krig_pm_lag6 = lag(krig_pm, 6, order_by = county),
-         krig_pm_lag7 = lag(krig_pm, 7, order_by = county), 
-         # background pm
-         background_pm_lag1 = lag(background_pm, 1, order_by = county),
-         background_pm_lag2 = lag(background_pm, 2, order_by = county),
-         background_pm_lag3 = lag(background_pm, 3, order_by = county),
-         background_pm_lag4 = lag(background_pm, 4, order_by = county),
-         background_pm_lag5 = lag(background_pm, 5, order_by = county), 
-         background_pm_lag6 = lag(background_pm, 6, order_by = county),
-         background_pm_lag7 = lag(background_pm, 7, order_by = county), 
-         # geo_smk_pm 
-         geo_smk_pm_lag1 = lag(geo_smk_pm, 1, order_by = county),
-         geo_smk_pm_lag2 = lag(geo_smk_pm, 2, order_by = county),
-         geo_smk_pm_lag3 = lag(geo_smk_pm, 3, order_by = county),
-         geo_smk_pm_lag4 = lag(geo_smk_pm, 4, order_by = county),
-         geo_smk_pm_lag5 = lag(geo_smk_pm, 5, order_by = county),
-         geo_smk_pm_lag6 = lag(geo_smk_pm, 6, order_by = county),
-         geo_smk_pm_lag7 = lag(geo_smk_pm, 7, order_by = county),
-         # krig smk pm
-         krig_smk_pm_lag1 = lag(krig_smk_pm, 1, order_by = county),
-         krig_smk_pm_lag2 = lag(krig_smk_pm, 2, order_by = county),
-         krig_smk_pm_lag3 = lag(krig_smk_pm, 3, order_by = county),
-         krig_smk_pm_lag4 = lag(krig_smk_pm, 4, order_by = county),
-         krig_smk_pm_lag5 = lag(krig_smk_pm, 5, order_by = county),
-         krig_smk_pm_lag6 = lag(krig_smk_pm, 6, order_by = county),
-         krig_smk_pm_lag7 = lag(krig_smk_pm, 7, order_by = county),
-         # temp
-         wrf_temp_lag1 = lag(wrf_temp, 1, order_by = county),
-         wrf_temp_lag2 = lag(wrf_temp, 2, order_by = county),
-         wrf_temp_lag3 = lag(wrf_temp, 3, order_by = county),
-         wrf_temp_lag4 = lag(wrf_temp, 4, order_by = county),
-         wrf_temp_lag5 = lag(wrf_temp, 5, order_by = county), 
-         wrf_temp_lag6 = lag(wrf_temp, 6, order_by = county),
-         wrf_temp_lag7 = lag(wrf_temp, 7, order_by = county)) %>% 
-  # ungroup by zip
-  ungroup(county) %>% 
-  # attach a zip indicator for each smoke variable
-  setNames(paste(colnames(.), "county", sep="_"))  %>% 
-  # remove the '_zip' from the zipcode and date variable 
-  rename(county = county_county, date = date_county, fips = fips_county) %>%
-  select(-1)
-
-or_ndc_county_ts <- or_ndc_inhaler %>%
-  rename(date = dates) %>%
-  mutate(day = ifelse(weekdays(date) == "Monday" |
-                      weekdays(date) == "Tuesday" |
-                      weekdays(date) == "Wednesday"|
-                      weekdays(date) == "Thursday"|
-                      weekdays(date) == "Friday", "weekday",
-               ifelse(weekdays(date) == "Saturday" |
-                      weekdays(date) == "Sunday", "weekend", NA))) %>%
-  # join with county-level population
-  left_join(or_zip_county_new, by = "ZIPCODE") %>% # 321 missing 
-  filter(!is.na(county)) %>% # no missing value
-  left_join(or_pop, by = "county") %>%
-  left_join(county_smoke_w_lag, by = c("date", "county")) %>%
-  arrange(personkey, date) %>% # order by id and fromdate
-  group_by(county, date) %>% 
-  # sum up each primary diagnosis for each outcome for each day for each county
-  summarise(n_obs = n()) %>% 
-  # join daily county estimates of pm2.5
-  left_join(county_smoke_w_lag, by = c('date', 'county'))
-
-# summary(as.factor(or_ndc_county_ts$county))
-
-or_ndc_county_ts <- or_ndc_county_ts %>%
-  left_join(or_pop, by = "county") %>%
-  mutate(geo_smk_pm10 = geo_smk_pm_county/10) %>%
-  mutate(rate_per_100k = (n_obs/pop)*100000) %>%
-  mutate(day = ifelse(weekdays(date) == "Monday" |
-                      weekdays(date) == "Tuesday" |
-                      weekdays(date) == "Wednesday"|
-                      weekdays(date) == "Thursday"|
-                      weekdays(date) == "Friday", "weekday",
-               ifelse(weekdays(date) == "Saturday" |
-                      weekdays(date) == "Sunday", "weekend", NA)))
-
-or_ndc_county_ts$day[which(or_ndc_county_ts$date=="2013-05-27")] <- "weekend"
-or_ndc_county_ts$day[which(or_ndc_county_ts$date=="2013-07-04")] <- "weekend"
-or_ndc_county_ts$day[which(or_ndc_county_ts$date=="2013-09-02")] <- "weekend"
-
-write_path5 <- paste0("../data_new/medication/or_ndc_county_time_series.csv")
-write_csv(or_ndc_county_ts, write_path5)
-
-```
-
-```{r time stratified county smk for ts, echo=FALSE, include=FALSE, eval=FALSE}
-read_path_county_pm <- paste0("../data_new/county_data/or_county_pop_wt_pm.csv")
-county_smoke <- read_csv(read_path_county_pm) # 5508 rows
-
-# descriptives of the two smoke datasets
-summary(county_smoke)
-
-county_smoke_w_lag <- county_smoke %>% arrange(county, date) %>%
-  # group by zipcode
-  group_by(county) %>% 
-  # wrf
-  mutate(wrf_f_pm_lag1 = lag(wrf_pm, 1, order_by = county), 
-         wrf_f_pm_lag2 = lag(wrf_pm, 2, order_by = county),
-         wrf_f_pm_lag3 = lag(wrf_pm, 3, order_by = county),
-         wrf_f_pm_lag4 = lag(wrf_pm, 4, order_by = county),
-         wrf_f_pm_lag5 = lag(wrf_pm, 5, order_by = county),
-         wrf_f_pm_lag6 = lag(wrf_pm, 6, order_by = county),
-         wrf_f_pm_lag7 = lag(wrf_pm, 7, order_by = county),
-         # wrf no fire lag
-         wrf_nf_pm_lag1 = lag(wrf_nf_pm, 1, order_by = county),
-         wrf_nf_pm_lag2 = lag(wrf_nf_pm, 2, order_by = county),
-         wrf_nf_pm_lag3 = lag(wrf_nf_pm, 3, order_by = county),
-         wrf_nf_pm_lag4 = lag(wrf_nf_pm, 4, order_by = county),
-         wrf_nf_pm_lag5 = lag(wrf_nf_pm, 5, order_by = county),
-         wrf_nf_pm_lag6 = lag(wrf_nf_pm, 6, order_by = county),
-         wrf_nf_pm_lag7 = lag(wrf_nf_pm, 7, order_by = county),
-         # wrf_smk_pm
-         wrf_smk_pm_lag1 = lag(wrf_smk_pm, 1, order_by = county),
-         wrf_smk_pm_lag2 = lag(wrf_smk_pm, 2, order_by = county),
-         wrf_smk_pm_lag3 = lag(wrf_smk_pm, 3, order_by = county),
-         wrf_smk_pm_lag4 = lag(wrf_smk_pm, 4, order_by = county),
-         wrf_smk_pm_lag5 = lag(wrf_smk_pm, 5, order_by = county),
-         wrf_smk_pm_lag6 = lag(wrf_smk_pm, 6, order_by = county),
-         wrf_smk_pm_lag7 = lag(wrf_smk_pm, 7, order_by = county),
-         # geo weighted pm
-         geo_wt_pm_lag1 = lag(geo_wt_pm, 1, order_by = county),
-         geo_wt_pm_lag2 = lag(geo_wt_pm, 2, order_by = county),
-         geo_wt_pm_lag3 = lag(geo_wt_pm, 3, order_by = county),
-         geo_wt_pm_lag4 = lag(geo_wt_pm, 4, order_by = county),
-         geo_wt_pm_lag5 = lag(geo_wt_pm, 5, order_by = county),
-         geo_wt_pm_lag6 = lag(geo_wt_pm, 6, order_by = county),
-         geo_wt_pm_lag7 = lag(geo_wt_pm, 7, order_by = county),
-         # krig pm
-         krig_pm_lag1 = lag(krig_pm, 1, order_by = county),
-         krig_pm_lag2 = lag(krig_pm, 2, order_by = county),
-         krig_pm_lag3 = lag(krig_pm, 3, order_by = county),
-         krig_pm_lag4 = lag(krig_pm, 4, order_by = county),
-         krig_pm_lag5 = lag(krig_pm, 5, order_by = county), 
-         krig_pm_lag6 = lag(krig_pm, 6, order_by = county),
-         krig_pm_lag7 = lag(krig_pm, 7, order_by = county), 
-         # background pm
-         background_pm_lag1 = lag(background_pm, 1, order_by = county),
-         background_pm_lag2 = lag(background_pm, 2, order_by = county),
-         background_pm_lag3 = lag(background_pm, 3, order_by = county),
-         background_pm_lag4 = lag(background_pm, 4, order_by = county),
-         background_pm_lag5 = lag(background_pm, 5, order_by = county), 
-         background_pm_lag6 = lag(background_pm, 6, order_by = county),
-         background_pm_lag7 = lag(background_pm, 7, order_by = county), 
-         # geo_smk_pm 
-         geo_smk_pm_lag1 = lag(geo_smk_pm, 1, order_by = county),
-         geo_smk_pm_lag2 = lag(geo_smk_pm, 2, order_by = county),
-         geo_smk_pm_lag3 = lag(geo_smk_pm, 3, order_by = county),
-         geo_smk_pm_lag4 = lag(geo_smk_pm, 4, order_by = county),
-         geo_smk_pm_lag5 = lag(geo_smk_pm, 5, order_by = county),
-         geo_smk_pm_lag6 = lag(geo_smk_pm, 6, order_by = county),
-         geo_smk_pm_lag7 = lag(geo_smk_pm, 7, order_by = county),
-         # krig smk pm
-         krig_smk_pm_lag1 = lag(krig_smk_pm, 1, order_by = county),
-         krig_smk_pm_lag2 = lag(krig_smk_pm, 2, order_by = county),
-         krig_smk_pm_lag3 = lag(krig_smk_pm, 3, order_by = county),
-         krig_smk_pm_lag4 = lag(krig_smk_pm, 4, order_by = county),
-         krig_smk_pm_lag5 = lag(krig_smk_pm, 5, order_by = county),
-         krig_smk_pm_lag6 = lag(krig_smk_pm, 6, order_by = county),
-         krig_smk_pm_lag7 = lag(krig_smk_pm, 7, order_by = county),
-         # temp
-         wrf_temp_lag1 = lag(wrf_temp, 1, order_by = county),
-         wrf_temp_lag2 = lag(wrf_temp, 2, order_by = county),
-         wrf_temp_lag3 = lag(wrf_temp, 3, order_by = county),
-         wrf_temp_lag4 = lag(wrf_temp, 4, order_by = county),
-         wrf_temp_lag5 = lag(wrf_temp, 5, order_by = county), 
-         wrf_temp_lag6 = lag(wrf_temp, 6, order_by = county),
-         wrf_temp_lag7 = lag(wrf_temp, 7, order_by = county)) %>% 
-  # ungroup by zip
-  ungroup(county) %>% 
-  # attach a zip indicator for each smoke variable
-  setNames(paste(colnames(.), "county", sep="_"))  %>% 
-  # remove the '_zip' from the zipcode and date variable 
-  rename(county = county_county, date = date_county, fips = fips_county) %>%
-  select(-1)
-
-or_county_ts <- county_smoke_w_lag %>%
-  mutate(day = ifelse(weekdays(date) == "Monday" |
-                      weekdays(date) == "Tuesday" |
-                      weekdays(date) == "Wednesday"|
-                      weekdays(date) == "Thursday"|
-                      weekdays(date) == "Friday", "weekday",
-               ifelse(weekdays(date) == "Saturday" |
-                      weekdays(date) == "Sunday", "weekend", NA))) %>%
-  group_by(county, date) %>%
-  mutate(geo_smk_pm10 = geo_smk_pm_county/10)
-
-or_county_ts$day[which(or_county_ts$date=="2013-05-27")] <- "weekend"
-or_county_ts$day[which(or_county_ts$date=="2013-07-04")] <- "weekend"
-or_county_ts$day[which(or_county_ts$date=="2013-09-02")] <- "weekend"
-
-write_path5 <- paste0("../data_new/medication/or_county_time_series.csv")
-write_csv(or_county_ts, write_path5)
-
-```
-
-## 1. smoke exposure time series and map of smoke days in Oregon
-
-### Prepare (already done in previous time stratified work)
-
-1. Make 10 units' geo smoke variable
-2. Make county population variable (to calculate rate)
-3. Make weekend/weekday variables
-
-```{r time series, warning=FALSE, message=FALSE, echo=FALSE}
-read_path5 <- paste0("../data_new/medication/or_ndc_county_time_series.csv")
-or_ndc_county_ts <- read_csv(read_path5)
-
-plot_ts <- ggplot(or_ndc_county_ts, aes(y =rate_per_100k, x = date)) +
-           geom_jitter() +
-           ylab("Rate of 100k of Beta2 agonist fills") +
-           xlab("Year 2013") +
-           ggtitle("Oregon: 2013 Fire Season Pharmacy Fills")
-
-# print(plot_ts)
-
-# save figure
-ggsave("../plot_new/ndc_rate_ts.pdf", plot = plot_ts, 
-       width = 12, height = 8, units = "in")
-
-plot_county <- ggplot(or_ndc_county_ts, aes(y =rate_per_100k, x = date)) +
-               geom_jitter() + facet_wrap(~county) +
-               ylab("Rate of 100k of Beta2 agonist fills") +
-               ggtitle("Oregon: 2013 Fire Season Pharmacy Fills by County")
-
-plot_county_line <- ggplot(or_ndc_county_ts, aes(y =rate_per_100k, x = date)) +
-                    facet_wrap(~county) + geom_line()
-
-# print(plot_county)
-# print(plot_county_line)
-# save figure
-ggsave("../plot_new/ndc_county_rate_ts.pdf", plot = plot_county, 
-       width = 12, height = 8, units = "in")
-ggsave("../plot_new/ndc_county_rate_ts_line.pdf", plot = plot_county_line, 
-       width = 12, height = 8, units = "in")
-
-# plot_smk_ts <- ggplot(or_ndc_county_ts, aes(y =geo_smk_pm10, x = date)) +
-#           geom_jitter() +
-#           ylab("GWR smoke concentration (10ug/m^3)") +
-#           xlab("Year 2013")
-
-# print(plot_smk_ts)
-
-### check if the time series has the association with get smoke
-a <- glm(rate_per_100k ~ geo_smk_pm10, data = or_ndc_county_ts, family = "poisson")
-# summary(a)
-
-plot_health_smk <- ggplot(or_ndc_county_ts, aes(x = geo_smk_pm10, y = log(rate_per_100k+0.1))) +
-  geom_point() + 
-  # geom_point(aes(colour = c("red")), alpha = 0.2,  size = 1) + 
-  # scale_colour_manual(values = c("red", "blue")) +
-  # truncated y axis
-  #scale_y_continuous(limits = c(0, 1.5)) +
-  # facet_wrap(~county, scale  = "free") +
-  ylab("Log Beta 2 agonists Rate per 100,000 Persons") +
-  xlab("GWR Smoke PM2.5 per 10ug/m^3") +
-  ggtitle("Regression of Pharmacy Fills on GWR Smoke Concentration") +
-  theme_bw()
-
-# unnecessary
-# plot_county_health_smk <- ggplot(or_ndc_county_ts, aes(x = geo_smk_pm10, y = log(rate_per_100k+0.1))) +
-#   geom_point(aes(colour = county), alpha = 0.2,  size = 1) + 
-  # scale_colour_manual(values = c("red", "blue")) +
-  # truncated y axis
-  #scale_y_continuous(limits = c(0, 1.5)) +
-#   facet_wrap(~county, scale  = "free") +
-#   ylab("Log Outcome Rate per 100,000 Persons") +
-#   xlab("GWR Smoke PM2.5 per 10*10ug/m^3") +
-#   theme_bw()
-
-# plot(plot_health_smk)
-# plot(plot_county_health_smk)
-
-ggsave("../plot_new/ndc_rate_health_smk.pdf", plot = plot_health_smk, 
-       width = 12, height = 8, units = "in")
-
-# ggsave("../plot_new/ndc_county_rate_health_smk.pdf", plot = plot_county_health_smk, 
-#        width = 12, height = 8, units = "in")
-
-```
-
-### map of smoke days (4)
-
-```{r smoke days and map, warning=FALSE, message=FALSE, echo=FALSE}
-read_path6 <- paste0("../data_new/medication/or_county_time_series.csv")
-or_county_ts <- read_csv(read_path6)
-
-or_ts <- or_county_ts %>% 
-  mutate(smoke0 = ifelse(geo_smk_pm_county > 0, 1, 0),
-         smoke5 = ifelse(geo_smk_pm_county > 5, 1, 0),
-         smoke10 = ifelse(geo_smk_pm_county > 10, 1, 0),
-         smoke15 = ifelse(geo_smk_pm_county > 15, 1, 0), 
-         # lag variable
-         geo_smk_lag1 = lag(geo_smk_pm10, 1, order_by = county),
-         geo_smk_lag2 = lag(geo_smk_pm10, 2, order_by = county),
-         geo_smk_lag3 = lag(geo_smk_pm10, 3, order_by = county),
-         geo_smk_lag4 = lag(geo_smk_pm10, 4, order_by = county),
-         geo_smk_lag5 = lag(geo_smk_pm10, 5, order_by = county),
-         geo_smk_lag6 = lag(geo_smk_pm10, 6, order_by = county),
-         geo_smk_lag7 = lag(geo_smk_pm10, 7, order_by = county)) %>% 
-  ungroup()
-
-# smoke wave days (at least 2 consecutive days in the 98th percentile of PM)
-# quantile(or_ts$geo_wt_pm_county, probs = c(0.95,0.98,0.99, 0.999))
-# following Coco's smoke wave day, the 98th percentile for PM2.5 concentrations
-# would be >16.3
-
-smoke_wave <- or_ts %>% 
-  select(county, date, geo_wt_pm_county) %>% 
-  mutate(high_pm_day = ifelse(geo_wt_pm_county > 10, 1, 0)) %>% 
-  group_by(county) %>% 
-  # identify consecutive days
-  mutate(smoke_wave = ifelse(lag(high_pm_day, 
-    order_by = county)==1 & high_pm_day ==1, 1, 0)) %>% 
-  select(-high_pm_day, -geo_wt_pm_county)
-
-# merge in smokewave day
-or_ts <- or_ts %>% 
-  right_join(smoke_wave, by = c("county", "date"))
-
-# count of state outcome timeseries
-state_outcome_rate <- or_ts %>% 
-  group_by(date) %>% 
-  # remove columns I don't want to calculate
-  select(-fips, -county)
-
-# average of environmental measures, joined with outcomes
-state_smoke <- or_ts %>% 
-  group_by(date) %>% 
-  select(-fips, -county) %>% 
-  summarise_each(funs(mean, median, min, max), wrf_pm_county:wrf_temp_county) # %>% 
-  # full_join(state_outcome_rate, by = c("date"))
-
-# small multiples dataset
-or_smk <- or_ts %>% 
-  select(county, date, geo_smk_pm_county) %>% 
-  gather(key = smk_method, value = pm, -date, -county)
-
-or_smk_plot <- ggplot(or_smk, aes(x = date, y = pm)) +
-  geom_point(aes(colour = smk_method), alpha = 0.8, size = 0.8) +
-  facet_wrap(~county) +
-  ggtitle(expression("Oregon: 2013 Fire Season Smoke PM"[2.5])) +
-   theme(legend.position="none") +
-   # theme(axis.text.x=element_blank()) +
-  theme(panel.background = element_rect(fill = 'white', colour = 'white')) +
-  theme(strip.background = element_rect(colour = "white", fill = "white"))+
-  ylab(expression(paste("GWR Smoke PM"[2.5], " Concentration, µg/m^3" ))) +
-  xlab("May 1st to Sep 30th in 2013") +
-  theme(axis.line.x = element_line(color="black"),
-        axis.line.y = element_line(color="black"))+
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
-
-or_smk_plot
-
- # data wrangle ----
-#first step is to summarise the counts of smoke days in each county 
-or_smk_count <- or_ts %>% 
-  group_by(county) %>% 
-  # summing up smoke days
-  summarise(smk_day10 = sum(smoke10)) %>% 
-  # county needs to be lowercase to join to maps dataframe
-  mutate(county = tolower(county))
-
-# maps package will be loaded to pull map data and create a dataframe
-or_county_df <- map_data("county", "oregon")
-
-# join county counts to map df
-or_smk_df <- or_county_df %>% 
-  right_join(or_smk_count, by = c("subregion" = "county")) %>% 
-  # small multiples of different smoke cutoff methods
-  gather(key = smk_cut_method, value = smoke_days, -long:-subregion) %>% 
-  # filter out smoke day 0 since it's not very specific
-  filter(smk_cut_method != "smk_day0") %>% 
-  # rename smoke day cutoffs 
-  mutate(smk_cut_method = 
-    ifelse(smk_cut_method == "smk_day10", "Smoke PM2.5 > 10 ug/m^3", NA),
-    # preserve smoke method order for small multiple
-    smk_cut_method = factor(smk_cut_method, 
-      levels = c("Smoke PM2.5 > 10 ug/m^3")))
-
-# map ----
-
-# or_county <- or_smk_df %>%
-#   group_by(subregion) %>%
-#   summarise(ave_long = mean(long), ave_lat = mean(lat))
-
-# or_county$subregion <- gsub("(^|[[:space:]])([[:alpha:]])", "\\1\\U\\2", or_county$subregion, perl=TRUE)
-
-getLabelPoint <- # Returns a county-named list of label points
-  function(county) {Polygon(county[c('long', 'lat')])@labpt}
-
-centroids <- by(or_smk_df, or_smk_df$subregion, getLabelPoint)     # Returns list
-centroids <- do.call("rbind.data.frame", centroids)  # Convert to Data Frame
-names(centroids) <- c('long', 'lat')                 # Appropriate Header
-
-rownames(centroids) <- gsub("(^|[[:space:]])([[:alpha:]])", "\\1\\U\\2", rownames(centroids), perl=TRUE)
-
-centroids[4,] <- centroids[4,] + c(0, 0.02) # Clatsop
-centroids[5,] <- centroids[5,] + c(0.1, 0) # Columbia
-centroids[29,] <- centroids[29,] + c(-0.1, -0.03) # Tillamook
-centroids[34,] <- centroids[34,] + c(-0.12, -0.02) # Washington
-centroids[26,] <- centroids[26,] + c(0, 0.03) # Multnomah
-centroids[11,] <- centroids[11,] + c(-0.05, -0.12) # Gilliam
-centroids[24,] <- centroids[24,] + c(-0.2, 0) # Marion
-centroids[33,] <- centroids[33,] + c(0, -0.1) # Wasco
-centroids[28,] <- centroids[28,] + c(0, 0.1) # Sherman
-centroids[36,] <- centroids[36,] + c(-0.05, -0.08) # Yamhill
-centroids[8,] <- centroids[8,] + c(0.08, -0.2) # Yamhill
-centroids[17,] <- centroids[17,] + c(-0.05, 0.08) # Yamhill
-centroids[29,] <- centroids[29,] + c(-0.01, -0.1) # Sherman
-
-# smoke days where GWR smoke > 15
-smoke_map <- ggplot(or_smk_df, aes(x=long, y=lat, group=group)) +
-  # fill with number of smoke days
-  geom_polygon(aes(fill = smoke_days), alpha = 0.7) +
-  scale_fill_gradient("Smoke-Impacted Days",
-                      low = "#3964C3", high = "#dc2430") +
-  # add county path on top
-  geom_path() +
-  facet_wrap(~smk_cut_method) +
-   theme(strip.background = element_blank(),
-       strip.text.x = element_blank()) +
-  xlab("Longitude") +
-  ylab("Latitude") +
-  ggtitle("Oregon: Counties Affected by Smoke PM2.5 > 10µg/m^3") +
-  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    strip.background = element_rect(colour=NA, fill=NA)) +
-  geom_text_repel(data = centroids, aes(long, lat, label = rownames(centroids), group=NULL), size = 2.5)
-
-smoke_map1 <- ggplot(or_smk_df, aes(x=long, y=lat, group=group)) +
-  # fill with number of smoke days
-  geom_polygon(aes(fill = smoke_days), alpha = 0.7) +
-  scale_fill_gradient("Smoke-Impacted Days",
-                      low = "#3964C3", high = "#dc2430") +
-  # add county path on top
-  geom_path() +
-  facet_wrap(~smk_cut_method) +
-   theme(strip.background = element_blank(),
-       strip.text.x = element_blank()) +
-  xlab("Longitude") +
-  ylab("Latitude") +
-  ggtitle("Oregon: Counties Affected by Smoke PM2.5 > 10µg/m^3") +
-  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    strip.background = element_rect(colour=NA, fill=NA)) +
-  geom_text(data = centroids, aes(long, lat, label = rownames(centroids), group=NULL), size = 2)
-
-smoke_map1
-
-```
-
-## 2. descriptive table of total ndc inhalhers; and age and sex category; Plots
-
-### Descriptives
-
-For all of 2013, there were a total of 77 millon claim records in the APAC dataset.
-During the May 1st to September 30th wildfire season in Oregon state, there were 24217 pharmacy visits for the ndc of beta 2 agonists.
-The number of cases, as well as age-specific and sex-specific strate are shown below.
+Load data
 
@@
 
-# outcome name
-outcome_name <- "Beta - 2 agonist"
-
-# dataframe for analysis creation
-# bind columns back together 
-df_analysis <- inhaler_ndc_df %>%
-  filter(!is.na(geo_smk_pm_zip)) %>% 
-  # only look at outcomes
-  filter(outcome == 1) %>%
-  # add another row that makes sure there is a person <15 in the dataframe
-  # tricking xtabs to produce a 0 cell for the outcome for age <15
-  add_row(outcome = 0, age_ind = 0)
-
-# cross tabs
-outcome_n <- xtabs(~ outcome, df_analysis)
-cross_tab_age <- xtabs(~ outcome + age_ind, df_analysis)
-cross_tab_sex <- xtabs(~ outcome + sex_ind, df_analysis)
-# empty matrix
-point_estimates <- matrix(nrow = 1, ncol = 7, byrow = T)
-
-colnames(point_estimates) <- c("outcome", "n", "age_15", "age_15_65", 
-                               "age_65", "female", "male")
-
-# fill in the outcome name for the dataframe before the loop
-point_estimates[, 1] <- outcome_name
-# fill n
-point_estimates[, 2] <- outcome_n[2] # second element of the 1 dimension vector
-# age <15
-point_estimates[, 3] <- cross_tab_age[2, 1]
-# age 15 to 65
-point_estimates[, 4] <- cross_tab_age[2, 2]
-# age >65
-point_estimates[, 5] <- cross_tab_age[2, 3]
-# male == 0
-point_estimates[, 7] <- cross_tab_sex[1, 1]
-# female == 1
-point_estimates[, 6] <- cross_tab_sex[1, 2]
-
-# save point estimates as a dataframe
-point_est_df <- as_data_frame(point_estimates)
-
-# combine each outcome dataframe itteration in to a big dataset
-asthma_point_est_df <- point_est_df %>% 
-  # find proportions/percents for each strata in a row
-  mutate(age_15_pr = as.character(round((as.numeric(age_15)/as.numeric(n))*100,1)),
-         age_15_65_pr = as.character(round((as.numeric(age_15_65)/as.numeric(n))*100,1)),
-         age_65_pr = as.character(round((as.numeric(age_65)/as.numeric(n))*100,1)),
-         female_pr = as.character(round((as.numeric(female)/as.numeric(n))*100,1)),
-         male_pr = as.character(round((as.numeric(male)/as.numeric(n))*100,1))) %>% 
-  select(outcome, n, age_15, age_15_pr, age_15_65, age_15_65_pr, age_65,
-         age_65_pr, female, female_pr, male, male_pr)
-
-# str(asthma_point_est_df)
-
-tab <- htmlTable(txtRound(asthma_point_est_df, digits = 1), 
-           caption = "Total number of cases for pharmacy fills of beta-2 agonist observed from May 1st to September 30st, 2013",
-           # column headers
-           header = c("Medication", "Cases n", "Less than 15", "(%)", "15 to 65", "(%)", "Greater than 65",
-                      "(%)", "Female", "(%)", "Male", "(%)"),
-           # column spanner
-           cgroup = c("","Age Category", "Sex"), 
-           n.cgroup = c(2, 6, 4),
-           padding.rgroup = "&nbsp;&nbsp;",
-           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
-           align = "llccccc" # column alignment,
-            ) # end table
-
-print(tab)
-
 ```
-
-### Outcomes of four kinds
-
-```{r category, echo = F, message = F, warning = F, results='asis'}
-read_path <- paste0("../data_new/medication/oregon_beta_ndc_index.csv")
-inhaler_index <- read_csv(read_path)
-
-# dataframe to loop through
-inhaler_index_df <- data.frame(inhaler_index) %>%
-  select(personkey, index)
-
-cate <- inhaler_index_df %>%
-  filter(personkey %in% inhaler_ndc$personkey) %>%
-  unique()
-
-inhaler_ndc_index_df <- inhaler_ndc_df %>%
-  left_join(cate, by="personkey")
-
-asthma_df <- inhaler_ndc_index_df %>%
-  filter(index=="asthma")
-
-copd_df <- inhaler_ndc_index_df %>%
-  filter(index=="copd")
-
-both_df <- inhaler_ndc_index_df %>%
-  filter(index=="both")
-
-other_df <- inhaler_ndc_index_df %>%
-  filter(index=="other")
-
-df_list <- list(asthma_df, copd_df, both_df, other_df)
-
-outcome_list <- c('Asthma', 'COPD', 'Both', 'Other')
-
-method_list <- c('Geo-Weighted Smoke')
-
-# create an empty list to row bind dataframes together
-datalist <- list()
-
-# set list and do cross tabs to find values
-# Producing conditional logit model estimates loop 
-for(i in 1:length(df_list)){
-
-# dataframe to loop through
-  df_to_loop <- data.frame(df_list[i])
-  # indication of column
-  outcome <- colnames(df_to_loop[76])
-  # outcome name
-  outcome_name <- outcome_list[i]
-
-  # dataframe for analysis creation
-  # bind columns back together 
-  df_analysis <- df_to_loop %>%
-    filter(!is.na(wrf_smk_pm_zip)) %>% 
-    # only look at outcomes
-    filter(outcome == 1) %>%
-    # add another row that makes sure there is a person <15 in the dataframe
-    # tricking xtabs to produce a 0 cell for the outcome for age <15
-    add_row(outcome = 0, age_ind = 0)
-
-  # cross tabs
-  outcome_n <- xtabs(~ outcome, df_analysis)
-  cross_tab_age <- xtabs(~ outcome + age_ind, df_analysis)
-  cross_tab_sex <- xtabs(~ outcome + sex_ind, df_analysis)
-  # empty matrix
-  point_estimates <- matrix(nrow = 1, ncol = 7, byrow = T)
-
-  colnames(point_estimates) <- c("outcome", "n", "age_15", "age_15_65", 
-                                 "age_65", "female", "male")
-
-  # fill in the outcome name for the dataframe before the loop
-  point_estimates[, 1] <- outcome_name
-  # fill n
-  point_estimates[, 2] <- outcome_n[2] # second element of the 1 dimension vector
-  # age <15
-  point_estimates[, 3] <- cross_tab_age[2, 1]
-  # age 15 to 65
-  point_estimates[, 4] <- cross_tab_age[2, 2]
-  # age >65
-  point_estimates[, 5] <- cross_tab_age[2, 3]
-  # male == 0
-  point_estimates[, 7] <- cross_tab_sex[1, 1]
-  # female == 1
-  point_estimates[, 6] <- cross_tab_sex[1, 2]
-
-  # save point estimates as a dataframe
-  point_est_df <- as_data_frame(point_estimates)
-
-  # combine previous values in dataframe that has all outcome/methods comparisons
-  datalist[[i]] <- point_est_df
-
-} # end of loop
-
-# combine each outcome dataframe itteration in to a big dataset
-combined_point_est_df <- bind_rows(datalist) %>% 
-  # find proportions/percents for each strata in a row
-  mutate(age_15_pr = as.character(round((as.numeric(age_15)/as.numeric(n))*100,1)),
-         age_15_65_pr = as.character(round((as.numeric(age_15_65)/as.numeric(n))*100,1)),
-         age_65_pr = as.character(round((as.numeric(age_65)/as.numeric(n))*100,1)),
-         female_pr = as.character(round((as.numeric(female)/as.numeric(n))*100,1)),
-         male_pr = as.character(round((as.numeric(male)/as.numeric(n))*100,1))) %>% 
-  select(outcome, n, age_15, age_15_pr, age_15_65, age_15_65_pr, age_65,
-         age_65_pr, female, female_pr, male, male_pr)
-
-#str(combined_point_est_df)
-
-tab <- htmlTable(txtRound(combined_point_est_df, digits = 1), 
-           caption = "Number of cases for each outcome observed from May 1st to September 30st, 2013",
-           # column headers
-           header = c("Outcome", "Cases n", "Less than 15", "(%)", "15 to 65", "(%)", "Greater than 65",
-                      "(%)", "Female", "(%)", "Male", "(%)"),
-           # column spanner
-           cgroup = c("","Age Category", "Sex"), 
-           n.cgroup = c(2, 6, 4),
-           padding.rgroup = "&nbsp;&nbsp;",
-           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
-           align = "llccccc" # column alignment,
-            ) # end table
-
-print(tab)
-
-```
-
-### Overall plot with GWR method
-
-The overall plot shows that when smoke concentration increases 10ug/m^3, the risk to get beta-2 agonist fill will increase about 7.5% significantly.
-
-```{r overall, warning =F, echo = F, results='asis'} 
-# dataframe list
-
-method_list <- c('Geo-Weighted Smoke')
-
-# dataframe to loop through
-inhaler_ndc_df <- data.frame(inhaler_ndc)
-
-# outcome name
-outcome <- "Beta - 2 agonists"
-outcome_name <- "Beta-2 agonists"
-
-# data wrangling ----
-# Producing conditional logit model estimates loop
-
-# extract covariates from dataframe
-covariates_df <- inhaler_ndc_df[, c(1:26, 74:84)]
-
-# extract pm values and divide by 10 and ordered
-#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
-pm_estimates_df <- inhaler_ndc_df[, c(91, 93)]/10  # create 10 unit increases
-
-# new loop for age categories
-
-# empty matrix (12 x 10 matrix)
-point_estimates <- matrix(nrow = 1, ncol = 9, byrow = T)
-
-colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 
-                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
-
-# fill in the outcome namedataframe before method loop
-point_estimates[, 1] <- outcome_name
-
-# dataframe for analysis creation
-# bind columns back together 
-df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
-  # remove missing pm values
-  filter(!is.na(geo_smk_pm_zip)) %>% 
-  # the following code makes sure that the counterfactual values retained are 
-  # symetric in that number of obs before = number of obs after
-  mutate(obs_diff_admission = (fromdate - date)/7) 
-  # dataframe is already for the entire fire season, so I don't need to subset anymore
-
-# second loop to run a model for each pm estimation method
-j <- 38
-
-# set row number to fill
-row_n <- j-37
-
-# only run the model if the dataframe has observations
-if(nrow(df_analysis) != 0){
-  # conditional logistic regression model
-  mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
-
-  # populate matrix
-  point_estimates[row_n, 2] <- method_list[row_n]
-  point_estimates[row_n, 3] <- mod$n
-  point_estimates[row_n, 4] <- mod$nevent
-  # odds ratio
-  point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)
-
-  # 95% lower bound
-  point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
-  # 95% upper bound
-  point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
-                                        1.96*(summary(mod)$coefficient[1,3])), 3)
-  # standard error
-  point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
-  # p val
-  point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)
-
-  # create else statement that fills matrix with missing so I still have the row
-  # in the final dataframe
-  } else {point_estimates[row_n, 2] <- method_list
-          point_estimates[row_n, 3] <- 0
-          point_estimates[row_n, 4] <- 0
-          point_estimates[row_n, c(5:9)] <- 99 } # end 'if else' statement
-
-# save point estimates as a dataframe
-combined_point_est_df <- as_data_frame(point_estimates)
-
-wrf_geo_age <- combined_point_est_df %>% 
-  # subset columns I want to put in to the table
-  select(2, 4:7) %>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
-         lower95 = ifelse(lower95 == 99, "--", lower95),
-         upper95 = ifelse(upper95 == 99, "--", upper95))
-
-tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
-         caption = "Association between a 10 ug/m^3 in PM2.5 three smoke method and pharmacy outcomes",
-         # row group by outcome
-         rgroup = "Beta - 2 agonist",
-         n.rgroup = c(rep(1, 1)), # 6 rows for each age cat for each outcome
-         # column headers
-         header = c("Est Method", "Events",
-                    "OR&dagger;", "Lower", "Upper"),
-         # column spanner
-         cgroup = c("", "95% CI"), 
-         n.cgroup = c(3, 2),
-         padding.rgroup = "&nbsp;&nbsp;",
-         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
-         align = "llcccc", # column alignment,
-         tfoot="&dagger; Referent periods matched to events on same day of week within May to September fire season."
-         ) # end table
-
-print(tab)
-
-combined_point_est_df_new <- combined_point_est_df %>%
-  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
-                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
-  select(-n_events) %>%
-  rename(n_events = n_events_new)
-
-# ggplot of odds ratios, facet wrapped by outcomes -----
-# convert variables from character to either numeric or factor
-# factor preserves the order of the variable
-wrf_geo_age_plot <- combined_point_est_df_new %>% 
-  # filter columns to just geo-smk
- #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
-  # do not plot results with less than 15 events, create missing vals
-  # in mutate
-  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
-                             99, odds_ratio),
-         lower95 = ifelse(as.numeric(n_events) < 15, 
-                             99, lower95),
-         upper95 = ifelse(as.numeric(n_events) < 15, 
-                             99, upper95)) %>% 
-  # subset columns I want to put in to the table
-  select(1, 2, 9, 4:6) %>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
-         lower95 = ifelse(lower95 == 99, NA, lower95),
-         upper95 = ifelse(upper95 == 99, NA, upper95))
-
-# change characters to numeric and factor  
-wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
-                              levels = unique(wrf_geo_age_plot$outcome))
-
-wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
-                                levels = unique(wrf_geo_age_plot$pm_method))
-
-wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
-wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
-wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
-wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)
-
-## ggplot 
-  print_plot <- ggplot(wrf_geo_age_plot,
-    aes(x = pm_method, y = odds_ratio, color = pm_method), na.rm = T) +
-    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
-    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
-                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
-    # custom color 
-    scale_color_manual(name = "Smoke-Estimation Method", 
-                       values = c("red",  "blue", "#32115C"),
-                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
-    facet_wrap(~outcome, nrow = 3, scales = "free") +
-    geom_hline(yintercept = 1, linetype=2) +
-    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
-    ylab(expression(paste("Odds Ratio for 10Âµg/m"^3, " Increase in PM"[2.5]))) +
-    #ylim(0, 2) +
-    xlab('Overall') +
-    # plot theme
-    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    # strip element
-    strip.background = element_rect(colour=NA, fill=NA),
-    panel.border = element_rect(fill = NA, color = "black"),
-    # facet text size
-    strip.text = element_text(size = 10),
-    # axis element
-    #axis.text.x = element_blank(),
-    #axis.title.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.title.y = element_text(angle = 90),
-    # legend elements
-    legend.position = "bottom")
-    #legend.text = element_text(size = 8))
-
-  print(print_plot)
-  # save figure
-  ggsave("../plot_new/ndc_overall_gwr.pdf", plot = print_plot, 
-       width = 12, height = 8, units = "in")
-
-```
-
-The plot shows that four different categories
-
-```{r category plot, warning =F, echo = F, results='asis'}
-df_list <- list(asthma_df, copd_df, both_df, other_df)
-
-outcome_list <- c('Asthma', 'COPD', 'Both', 'Other')
-
-method_list <- c('Geo-Weighted Smoke')
-
-# create an empty list to row bind dataframes together
-datalist <- list()
-
-# data wrangling ----
-# Producing conditional logit model estimates loop 
-for(i in 1:length(df_list)){
-
-  # dataframe to loop through
-  df_to_loop <- data.frame(df_list[i])
-  # indication of column
-  outcome <- colnames(df_to_loop[75])
-  # outcome name
-  outcome_name <- outcome_list[i]
-
-  covariates_df <- df_to_loop[, c(1:26, 74:84)]
-
-  # extract pm values and divide by 10 and ordered
-  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
-  pm_estimates_df <- df_to_loop[, c(91, 93)]/10  # create 10 unit increases
-
-  # dataframe for analysis creation
-  # bind columns back together 
-  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
-    # remove missing pm values
-    filter(!is.na(geo_smk_pm_zip)) %>%
-
-    # join in cdc met data
-    # full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
-    # the following code makes sure that the counterfactual values retained are 
-    # symetric in that number of obs before = number of obs after
-    mutate(obs_diff_admission = (fromdate - date)/7) 
-  # dataframe is already for the entire fire season, so I don't need to subset anymore
-
-  # empty df for table
-  table_df <- data.frame()
-
-  # empty matrix
-  point_estimates <- matrix(nrow = 1, ncol = 9, byrow = T)
-
-  colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 'odds_ratio', 
-                                 'lower95', 'upper95', 'se', 'p_val')
-
-  # fill in the outcome name for the dataframe before the loop
-  point_estimates[, 1] <- outcome_name
-
-# set row number to fill
-row_n <- 1
-
-# only run the model if the dataframe has observations
-if(nrow(df_analysis) != 0){
-  # conditional logistic regression model
-  mod <- clogit(outcome ~ df_analysis[[38]] + wrf_temp_zip + strata(personkey), df_analysis)
-
-      # populate matrix
-    point_estimates[row_n, 2] <- method_list[row_n]
-    point_estimates[row_n, 3] <- mod$n
-    point_estimates[row_n, 4] <- mod$nevent
-    # odds ratio
-    point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)
-
-    # 95% lower bound
-    point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -
-                                             1.96*(summary(mod)$coefficient[1,3])), 3)
-    # 95% upper bound
-    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
-                                             1.96*(summary(mod)$coefficient[1,3])), 3)
-    # standard error
-    point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
-    # p val
-    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)
-
-    # save point estimates as a dataframe
-    point_est_df <- as_data_frame(point_estimates)      
-  # create else statement that fills matrix with missing so I still have the row
-  # in the final dataframe
-  } else {point_estimates[row_n, 2] <- method_list
-          point_estimates[row_n, 3] <- 0
-          point_estimates[row_n, 4] <- 0
-          point_estimates[row_n, c(5:9)] <- 99 } # end 'if else' statement
-
-  # combine previous values in dataframe that has all outcome/methods comparisons
-  datalist[[i]] <- point_est_df
-
-} # end of loop
-
-# combine each outcome dataframe itteration in to a big dataset
-combined_point_est_df <- bind_rows(datalist)
-
-# subset columns I want to put in to the table
-table_df <- combined_point_est_df %>% select(2, 3:7)
-
-tab <- htmlTable(txtRound(table_df, digits = 3, 1:3), 
-                 caption = "Association between a 10 ug/m^3 in PM2.5 and Health Outcomes",
-                 # row group by outcome
-                 rgroup = outcome_list,
-                 n.rgroup = c(rep(1, 4)), # 4 rows for each method for each outcome
-                 # column headers
-                 header = c("Method", "Obs.", "Events",
-                            "OR&dagger;", "Lower", "Upper"),
-                 # column spanner
-                 cgroup = c("", "95% CI"), 
-                 n.cgroup = c(4, 2),
-                 padding.rgroup = "&nbsp;&nbsp;",
-                 css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
-                 align = "llccccc", # column alignment,
-                 tfoot="&dagger; Adjusted for CDC temperature, relatively humidity, wind speed, and precipitation; accounting for subject.
-Time-stratified: referent periods matched to events on same day of week within July to October fire season."
-) # end table
-
-print(tab)
-
-# ggplot of odds ratios, facet wrapped by outcomes -----
-# convert variables from character to either numeric or factor
-# factor preserves the order of the variable
-combined_point_est_df$outcome <- factor(combined_point_est_df$outcome, 
-                                        levels = unique(combined_point_est_df$outcome))
-
-combined_point_est_df$pm_method <- factor(combined_point_est_df$pm_method, 
-                                          levels = unique(combined_point_est_df$pm_method))
-
-combined_point_est_df$n <- as.numeric(combined_point_est_df$n)
-combined_point_est_df$n_events <- as.numeric(combined_point_est_df$n_events)
-combined_point_est_df$odds_ratio <- as.numeric(combined_point_est_df$odds_ratio)
-combined_point_est_df$lower95 <- as.numeric(combined_point_est_df$lower95)
-combined_point_est_df$upper95 <- as.numeric(combined_point_est_df$upper95)
-combined_point_est_df$se <- as.numeric(combined_point_est_df$se)
-combined_point_est_df$p_val <- as.numeric(combined_point_est_df$p_val)
-
-# subset out global method
-combined_point_est_df <- combined_point_est_df
-
-## ggplot
-print_plot <- ggplot(combined_point_est_df, 
-                     aes(x = pm_method, y = odds_ratio, colour = pm_method)) +
-  # custom color 
-  scale_color_manual(name = "Smoke-Estimation Method", 
-                     values = c("red", "blue", "#32115C"),
-                     guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
-  geom_point() + #geom_text(vjust = 0, nudge_x = 0.3) +
-  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
-  facet_wrap(~outcome, nrow = 3) +
-  geom_hline(yintercept = 1, linetype=2) +
-  #    ggtitle('Association Between PM2.5 \n from Wildfire Smoke on Hospitalizations') +
-  ylab(expression(paste("Odds Ratio for 10Âµg/m"^3, " Increase in PM"[2.5]))) +
-  #xlab('Smoke Estimation Method') +
-  # plot theme
-  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-        panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        # strip element
-        strip.background = element_rect(colour=NA, fill=NA),
-        panel.border = element_rect(fill = NA, color = "black"),
-        # facet text size
-        #strip.text = element_text(size = 8),
-        # axis element
-        axis.text.x = element_blank(),
-        axis.title.x = element_blank(),
-        axis.ticks.x = element_blank(),
-        axis.title.y = element_text(angle = 90),
-        # legend elements
-        legend.position = "bottom")
-#legend.text = element_text(size = 8))
-
-print(print_plot)
-# save figure
-# ggsave("../plot_new/fig_both_er_ur.pdf", plot = print_plot, 
-#        width = 12, height = 8, units = "in")
-
-```
-
-## 3. Distributed Lag Model
-
-Although the casecrossover study shows a significant association of getting beta-2 agonist pharmacy fills and smoke concentration, we still wonder the association of the pharmacy fills with some lag days.
-For medication analysis, people could get the pharmacy fills one or some days later than the day they find they have the disease.
-Then we make 0 (the original day) to 7 lags (one week) to see association of pharmacy fills with different lagged days.
-
-### Distributed lag 0-7 for casecrossover
-
-When choosing the df of 5, the plot shows that the same day of fill with smoke exposure has the highest significant association.
-But other lags shows that pharmacy fills with lagged days do not has the significant association with smoke.
-And the association tends to decrease as lag goes.
 
@@
 
-```{r DLM for casecrossover df 5, echo = FALSE, message=FALSE, warning=FALSE}
-inhaler_geo_lag <- inhaler_ndc_lag_df %>%
-  select(159, 160:166) # GWR-method
-
-inhaler_temp_lag <- inhaler_ndc_lag_df %>%
-  select(93, 150:156) # GWR-method
-
-lag_max <- 7
-n1 <- length(inhaler_geo_lag[,1])
-n2 <- length(inhaler_temp_lag[,1])
-
-inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
-inhaler_temp_lag <- as.matrix(inhaler_temp_lag)
-
-B <- ns(0:lag_max, df =5, intercept=TRUE)
-
-# the nrow B should equal ncol(pm)
-# dim(B)
-
-inhaler_geo_lag_B <- inhaler_geo_lag%*%B
-inhaler_temp_lag_B <- inhaler_temp_lag%*%B
-
-fit <- clogit(outcome ~ inhaler_geo_lag_B + inhaler_temp_lag_B + strata(personkey), 
-              inhaler_ndc_lag_df)
-
-# coef(fit)
-# summary(fit)
-# AIC(fit)
-
-D <- cbind(B,B)
-
-# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
-dlparms <- c(1,2,3,4,5,6,7,8,9,10)
-DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms]) 
-DLvar <- D%*%vcov(fit)[dlparms,dlparms]%*%t(D)
-DLestimate$SE <- sqrt(diag(DLvar))
-DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
-DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
-
-DLestimate$estimate <- exp(DLestimate$estimate)
-DLestimate$lower <- exp(DLestimate$lower)
-DLestimate$upper <- exp(DLestimate$upper)
-
-# DLestimate
-
-DLestimate$lag <- 0:lag_max
-p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
-p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
-p <- p + geom_line(aes(x=lag), linetype=2, size=2)
-p <- p + theme_bw() + xlab("lagged days of events") + ylab("Odds Ratio for 10ug/m^3 Increase in PM2.5")
-p
-
-DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms])
-
-cumulative <- sum(DLestimate$estimate)
-# cumulative
-
-cumulative_se <- sqrt(sum(DLvar))
-# cumulative_se
-
-cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
-# cumulative_CI
-
-cum <- cbind(cumulative, cumulative_se, cumulative_CI[1], cumulative_CI[2])
-colnames(cum) <- c("Cumulative Est", "Cumulative SE", "Lower 95% CI", "Upper 95% CI")
-
-knitr::kable(cum, caption = paste0("Distributed lag 0-7 spline estimates, se and CI"))
-
-```
-
-```{r DLM for casecrossover temp, echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}
-read_path8 <- paste0("../data_new/county_data/zip_to_county_new.csv")
-or_zip_to_county <- read_csv(read_path8)
-
-read_path9 <- paste0("../data_new/county_data/or_census_pop_est_county.csv")
-or_pop <- read_csv(read_path9)
-
-or_pop <- or_pop %>%
-  rename(county = GEO_display_label, pop = respop72013) %>%
-  select(county, pop) 
-or_pop$county <- gsub(" County, Oregon", "", as.character(factor(or_pop$county)))
-
-inhaler_ndc_lag_df <- inhaler_ndc_df %>%
-  left_join(or_zip_to_county, by = "ZIPCODE") %>%
-  left_join(or_pop, by = "county") %>%
-  mutate(day = ifelse(weekdays(date) == "Monday" |
-                      weekdays(date) == "Tuesday" |
-                      weekdays(date) == "Wednesday"|
-                      weekdays(date) == "Thursday"|
-                      weekdays(date) == "Friday", "weekday",
-               ifelse(weekdays(date) == "Saturday" |
-                      weekdays(date) == "Sunday", "weekend", NA))) %>%
-  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
-         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
-         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
-         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
-         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
-         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
-         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
-         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)
-
-inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-05-27")] <- "weekend"
-inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-07-04")] <- "weekend"
-inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-09-02")] <- "weekend"
-
-which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_zip") # 159
-which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_lag1_zip") # 160
-which(colnames(inhaler_ndc_lag_df)=="wrf_temp_zip") # 93
-which(colnames(inhaler_ndc_lag_df)=="wrf_temp_lag1_zip") # 150
-
-inhaler_geo_lag <- inhaler_ndc_lag_df %>%
-  select(159, 160:166) # GWR-method
-
-inhaler_temp_lag <- inhaler_ndc_lag_df %>%
-  select(93, 150:156) # GWR-method
-
-lag_max <- 7
-n1 <- length(inhaler_geo_lag[,1])
-n2 <- length(inhaler_temp_lag[,1])
-
-inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
-inhaler_temp_lag <- as.matrix(inhaler_temp_lag)
-
-B <- ns(0:lag_max, df =3, intercept=TRUE)
-
-# the nrow B should equal ncol(pm)
-dim(B)
-
-inhaler_geo_lag_B <- inhaler_geo_lag%*%B
-inhaler_temp_lag_B <- inhaler_temp_lag%*%B
-
-fit <- clogit(outcome ~  inhaler_temp_lag_B + strata(personkey), 
-              inhaler_ndc_lag_df)
-
-coef(fit)
-summary(fit)
-AIC(fit)
-
-D <- cbind(B,B)
-
-# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
-dlparms <- c(1,2,3)
-DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms]) 
-DLvar <- B%*%vcov(fit)[dlparms,dlparms]%*%t(B)
-DLestimate$SE <- sqrt(diag(DLvar))
-DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
-DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
-
-DLestimate$estimate <- exp(DLestimate$estimate)
-DLestimate$lower <- exp(DLestimate$lower)
-DLestimate$upper <- exp(DLestimate$upper)
-
-DLestimate
-
-DLestimate$lag <- 0:lag_max
-p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
-p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
-p <- p + geom_line(aes(x=lag), linetype=2, size=2)
-p <- p + theme_bw()
-p
-
-DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms])
-
-cumulative <- sum(DLestimate$estimate)
-cumulative
-
-cumulative_se <- sqrt(sum(DLvar))
-cumulative_se
-
-cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
-cumulative_CI
-
-```
-
-```{r DLM for disease, }
-
-# set variable list
-var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
-              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
-              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
-
-start <- Sys.time()
-for(j in var_list) { # begin first loop of variable names (outcomes)
-
-  j <- "broken_arm"
-  # standard for statement without parallel computing
-  #for(j in 1:length(var_list)){
-  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
-  or_disease <- read_csv(read_path)
-
-  or_disease_df <- data.frame(or_disease)
-
-  or_disease_lag_df <- or_disease_df %>%
-  left_join(or_zip_to_county, by = "ZIPCODE") %>%
-  left_join(or_pop, by = "county") %>%
-  mutate(day = ifelse(weekdays(date) == "Monday" |
-                      weekdays(date) == "Tuesday" |
-                      weekdays(date) == "Wednesday"|
-                      weekdays(date) == "Thursday"|
-                      weekdays(date) == "Friday", "weekday",
-               ifelse(weekdays(date) == "Saturday" |
-                      weekdays(date) == "Sunday", "weekend", NA))) %>%
-  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
-         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
-         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
-         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
-         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
-         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
-         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
-         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)
-
-or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-05-27")] <- "weekend"
-or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-07-04")] <- "weekend"
-or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-09-02")] <- "weekend"
-
-which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
-which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
-which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
-which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151
-
-or_disease_geo_lag <- or_disease_lag_df %>%
-  select(174, 175:181) # GWR-method
-
-or_disease_temp_lag <- or_disease_lag_df %>%
-  select(94, 151:157) # GWR-method
-
-lag_max <- 7
-n1 <- length(or_disease_geo_lag[,1])
-n2 <- length(or_disease_temp_lag[,1])
-
-or_disease_geo_lag <- as.matrix(or_disease_geo_lag)
-or_disease_temp_lag <- as.matrix(or_disease_temp_lag)
-
-B <- ns(0:lag_max, df =5, intercept=TRUE)
-
-# the nrow B should equal ncol(pm)
-# dim(B)
-
-or_disease_geo_lag_B <- or_disease_geo_lag%*%B
-or_disease_geo_lag_B <- or_disease_temp_lag%*%B
-
-fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_geo_lag_B + strata(personkey), 
-              or_disease_lag_df)
-
-# coef(fit)
-# summary(fit)
-# AIC(fit)
-
-D <- cbind(B,B)
-
-# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
-dlparms <- c(1,2,3,4,5,6,7,8,9,10)
-DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms]) 
-DLvar <- D%*%vcov(fit)[dlparms,dlparms]%*%t(D)
-DLestimate$SE <- sqrt(diag(DLvar))
-DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
-DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
-
-DLestimate$estimate <- exp(DLestimate$estimate)
-DLestimate$lower <- exp(DLestimate$lower)
-DLestimate$upper <- exp(DLestimate$upper)
-
-# DLestimate
-
-DLestimate$lag <- 0:lag_max
-p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
-p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
-p <- p + geom_line(aes(x=lag), linetype=2, size=2)
-p <- p + theme_bw() + xlab("lagged days of events") + ylab("Odds Ratio for 10ug/m^3 Increase in PM2.5")
-p
-
-DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms])
-
-cumulative <- sum(DLestimate$estimate)
-# cumulative
-
-cumulative_se <- sqrt(sum(DLvar))
-# cumulative_se
-
-cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
-# cumulative_CI
-
-cum <- cbind(cumulative, cumulative_se, cumulative_CI[1], cumulative_CI[2])
-colnames(cum) <- c("Cumulative Est", "Cumulative SE", "Lower 95% CI", "Upper 95% CI")
-
-knitr::kable(cum, caption = paste0("Distributed lag 0-7 spline estimates, se and CI"))
-
-}
-
-```
-
-## Supplement
-
-### Age and Sex category (8)
-### Time-Stratified by Age Categories
-
-Following table and figure look at some outcomes stratified by age category.
-
-For some outcomes, such like Age > 65 in WRF-Chem method, the CI is the only not significant one in the plot.
-
-The plot shows that overall, the risk of pharmacy fills of beta-2 agonist will increase significantly as the smoke concentrate increases.
-For age group of 15-65, the risk is the highest, for 10 ug/m^3 increase of pm2.5, the risk of filling the beta-2 agonist will increase about 9%.
-For age groups which are less than 15 or more than 65, the risk is similar, both of them will increase about 5.6%.
-
-```{r time stratified age strata cdc met, warning =F, echo = F, results='asis'} 
-# dataframe list
-
-method_list <- c('Geo-Weighted Smoke')
-
-# dataframe to loop through
-inhaler_ndc_df <- data.frame(inhaler_ndc)
-
-# outcome name
-outcome <- "Beta - 2 agonists"
-outcome_name <- "Beta - 2 agonists"
-
-# age category list
-age_cat_list <- c(0,1,2)
-
-# create an empty list to row bind dataframes together
-datalist1 <- list()
-
-# data wrangling ----
-# Producing conditional logit model estimates loop
-
-# extract covariates from dataframe
-covariates_df <- inhaler_ndc_df[, c(1:26, 74:84)]
-
-# extract pm values and divide by 10 and ordered
-#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
-pm_estimates_df <- inhaler_ndc_df[, c(91, 93)]/10  # create 10 unit increases
-
-# new loop for age categories
-for(k in 0:2){
-  # empty matrix (12 x 10 matrix)
-  point_estimates <- matrix(nrow = 1, ncol = 10, byrow = T)
-
-  colnames(point_estimates) <- c('outcome', 'age_cat', 'pm_method', 'n', 'n_events', 
-                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
-
-  # fill in the outcome namedataframe before method loop
-  point_estimates[, 1] <- outcome_name
-  # repeat the age category 4 times for each pm method
-  point_estimates[, 2] <- k
-
-  # dataframe for analysis creation
-  # bind columns back together 
-  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
-    # remove missing pm values
-    filter(!is.na(geo_smk_pm_zip)) %>% 
-    # limit to specific age category
-    filter(age_ind == k) %>% 
-    # the following code makes sure that the counterfactual values retained are 
-    # symetric in that number of obs before = number of obs after
-    mutate(obs_diff_admission = (fromdate - date)/7) 
-    # dataframe is already for the entire fire season, so I don't need to subset anymore
-
- j <- 38
-
-    # variable to model 
-    var_name <- colnames(df_analysis[j])
-
-    # set row number to fill
-    row_n <- j-37
-
-    # only run the model if the dataframe has observations
-    if(nrow(df_analysis) != 0){
-    # conditional logistic regression model
-    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
-
-    # populate matrix
-    point_estimates[row_n, 3] <- method_list[row_n]
-    point_estimates[row_n, 4] <- mod$n
-    point_estimates[row_n, 5] <- mod$nevent
-    # odds ratio
-    point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
-
-    # 95% lower bound
-    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
-    # 95% upper bound
-    point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
-                                        1.96*(summary(mod)$coefficient[1,3])), 3)
-    # standard error
-    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
-    # p val
-    point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
-
-    # create else statement that fills matrix with missing so I still have the row
-    # in the final dataframe
-    } else {point_estimates[row_n, 3] <- method_list[row_n]
-            point_estimates[row_n, 4] <- 0
-            point_estimates[row_n, 5] <- 0
-            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
-     # end methods loop
-
-  # save point estimates as a dataframe
-  point_est_df <- as_data_frame(point_estimates)
-  datalist1[[k+1]] <- point_est_df
-} # end age category loop
-
-# bind rows of age category estimates together
-combined_point_est_df <- bind_rows(datalist1)
-
-wrf_geo_age <- combined_point_est_df %>% 
-  # filter columns to just geo-smk
-  # filter(pm_method == "GWR Smoke" |
-    #      pm_method == "WRF-Chem Smoke") %>% 
-  mutate(age_cat2 = ifelse(age_cat == 0, "Less than 15", 
-                    ifelse(age_cat == 1, "15-65",
-                    ifelse(age_cat == 2, "Greater than 65", NA)))) %>% 
-  # subset columns I want to put in to the table
-  select(3, 11, 5:8) %>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
-         lower95 = ifelse(lower95 == 99, "--", lower95),
-         upper95 = ifelse(upper95 == 99, "--", upper95))
-
-tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
-         caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by age",
-         # row group by outcome
-         rgroup = "Beta - 2 agonists",
-         n.rgroup = c(rep(3, 1)), # 6 rows for each age cat for each outcome
-         # column headers
-         header = c("Est Method", "Age Group", "Events",
-                    "OR&dagger;", "Lower", "Upper"),
-         # column spanner
-         cgroup = c("", "95% CI"), 
-         n.cgroup = c(4, 2),
-         padding.rgroup = "&nbsp;&nbsp;",
-         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
-         align = "llcccc", # column alignment,
-         tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within May to September fire season."
-         ) # end table
-
-print(tab)
-
-combined_point_est_df_new <- combined_point_est_df %>%
-  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
-                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
-  select(-n_events) %>%
-  rename(n_events = n_events_new)
-
-# ggplot of odds ratios, facet wrapped by outcomes -----
-# convert variables from character to either numeric or factor
-# factor preserves the order of the variable
-wrf_geo_age_plot <- combined_point_est_df_new %>% 
-  # filter columns to just geo-smk
- #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
-  # do not plot results with less than 15 events, create missing vals
-  # in mutate
-  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
-                             99, odds_ratio),
-         lower95 = ifelse(as.numeric(n_events) < 15, 
-                             99, lower95),
-         upper95 = ifelse(as.numeric(n_events) < 15, 
-                             99, upper95),
-         age_cat2 = ifelse(age_cat == 0, "<15", 
-                    ifelse(age_cat == 1, "15-65",
-                    ifelse(age_cat == 2, ">65", NA)))) %>% 
-  # subset columns I want to put in to the table
-  select(1, 3, 11, 10, 5:7) %>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
-         lower95 = ifelse(lower95 == 99, NA, lower95),
-         upper95 = ifelse(upper95 == 99, NA, upper95))
-
-# change characters to numeric and factor  
-wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
-                              levels = unique(wrf_geo_age_plot$outcome))
-
-wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
-                                levels = unique(wrf_geo_age_plot$pm_method))
-
-wrf_geo_age_plot$age_cat2 <- factor(wrf_geo_age_plot$age_cat2,
-                                levels = unique(wrf_geo_age_plot$age_cat2))
-
-wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
-wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
-wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
-wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)
-
-## ggplot 
-  print_plot <- ggplot(wrf_geo_age_plot,
-    aes(x = age_cat2, y = odds_ratio, color = pm_method), na.rm = T) +
-    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
-    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
-                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
-    # custom color 
-    scale_color_manual(name = "Smoke-Estimation Method", 
-                       values = c("red",  "blue", "#32115C"),
-                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
-    facet_wrap(~outcome, nrow = 3, scales = "free") +
-    geom_hline(yintercept = 1, linetype=2) +
-    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
-    ylab(expression(paste("Odds Ratio for 10Âµg/m"^3, " Increase in PM"[2.5]))) +
-    #ylim(0, 2) +
-    xlab('Age Category ') +
-    # plot theme
-    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    # strip element
-    strip.background = element_rect(colour=NA, fill=NA),
-    panel.border = element_rect(fill = NA, color = "black"),
-    # facet text size
-    strip.text = element_text(size = 10), 
-    # axis element
-    #axis.text.x = element_blank(),
-    #axis.title.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.title.y = element_text(angle = 90),
-    # legend elements
-    legend.position = "bottom")
-    #legend.text = element_text(size = 8))
-
-  print(print_plot)
-  # save figure
-  ggsave("../plot_new/ndc_age_fig_gwr.pdf", plot = print_plot, 
-       width = 12, height = 8, units = "in")
-
-```
-
-### Seasonal Time-Stratified by Sex
-
-The association with smoke PM~2.5~ and health outcomes stratified by sex is shown below.
-
-All the CI do not include 1, which means the pharmacy filling (beta 2 agonists) increases about 5% when smoke increase 10 units.
-So the risk for getting asthma will increase about 5% ~ 8% when the smoke PM2.5 increase 10 units.
-Also, the risk for female is much higher than that for male.
-
-These results show that there might be effect modification by sex, where females might be more likely to fill a beta-2 agonist on the same day it's really smoky for the GWR estimates.
-
-```{r sex strata cdc met adj, warning = F, echo = F, results='asis'} 
-method_list <- c('Geo-Weighted Smoke')
-
-# dataframe to loop through
-inhaler_ndc_df <- data.frame(inhaler_ndc)
-
-# outcome name
-outcome <- "Beta - 2 agonists"
-outcome_name <- "Beta - 2 agonists"
-
-# create an empty list to row bind dataframes together
-datalist1 <- list()
-
-# sex category list
-sex_strata_list <- c(0,1)
-
-# data wrangling ----
-# Producing conditional logit model estimates loop
-
-# extract covariates from dataframe
-covariates_df <- inhaler_ndc_df[, c(1:26, 74:84)]
-
-# extract pm values and divide by 10 and ordered
-#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
-pm_estimates_df <- inhaler_ndc_df[, c(91, 93)]/10  # create 10 unit increases
-
-# new loop for sex categories
-for(k in 0:1){
-
-  # empty matrix (12 x 10 matrix)
-  point_estimates <- matrix(nrow = 1, ncol = 10, byrow = T)
-
-  colnames(point_estimates) <- c('outcome', 'sex', 'pm_method', 'n', 'n_events', 
-                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
-
-  # fill in the outcome namedataframe before method loop
-  point_estimates[, 1] <- outcome_name
-  # repeat the sex category 4 times for each pm method
-  point_estimates[, 2] <- k
-
-  # dataframe for analysis creation
-  # bind columns back together 
-  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
-    # remove missing pm values
-    filter(!is.na(geo_smk_pm_zip)) %>% 
-    # limit to specific sex category
-    filter(sex_ind == k) %>% 
-    # the following code makes sure that the counterfactual values retained are 
-    # symetric in that number of obs before = number of obs after
-    mutate(obs_diff_admission = (fromdate - date)/7) 
-    # dataframe is already for the entire fire season, so I don't need to subset anymore
-
-  j <- 38
-    # variable to model 
-    var_name <- colnames(df_analysis[j])
-
-    # set row number to fill
-    row_n <- j-37
-
-    # only run the model if the dataframe has observations
-    if(nrow(df_analysis) != 0){
-    # conditional logistic regression model
-    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)
-
-    # populate matrix
-    point_estimates[row_n, 3] <- method_list[row_n]
-    point_estimates[row_n, 4] <- mod$n
-    point_estimates[row_n, 5] <- mod$nevent
-    # odds ratio
-    point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
-
-    # 95% lower bound
-    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
-                                      1.96*(summary(mod)$coefficient[1,3])), 3)
-    # 95% upper bound
-    point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
-                                      1.96*(summary(mod)$coefficient[1,3])), 3)
-    # standard error
-    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
-    # p val
-    point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
-
-    # create else statement that fills matrix with missing so I still have the row
-    # in the final dataframe
-    } else {point_estimates[row_n, 3] <- method_list[row_n]
-            point_estimates[row_n, 4] <- 0
-            point_estimates[row_n, 5] <- 0
-            point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
-    # end methods loop
-
-  # save point estimates as a dataframe
-  point_est_df <- as_data_frame(point_estimates)
-
-  # combine previous values in dataframe that has all outcome/methods comparisons
-  datalist1[[k+1]] <- point_est_df
-} # end sex category loop
-
-# bind rows of sex category estimates together
-combined_point_est_df <- bind_rows(datalist1)
-
-wrf_geo_sex <- combined_point_est_df %>% 
-  # filter columns to just geo-smk
-  # filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
-  mutate(sex_cat = ifelse(sex == 0, "Male", 
-               ifelse(sex == 1, "Female", NA))) %>% 
-  # subset columns I want to put in to the table
-  select(3, 11, 5:8)%>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
-         lower95 = ifelse(lower95 == 99, "--", lower95),
-         upper95 = ifelse(upper95 == 99, "--", upper95))
-
-  tab <- htmlTable(txtRound(wrf_geo_sex, digits = 3, 1:3), 
-           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and pharmacy outcomes stratified by sex",
-           # row group by outcome
-           rgroup = "Beta - 2 agonists",
-           n.rgroup = c(rep(2, 1)), # 2 rows for each sex strata for each outcome
-           # column headers
-           header = c("Est Method", "Sex Strata", "Events",
-                      "OR&dagger;", "Lower", "Upper"),
-           # column spanner
-           cgroup = c("", "95% CI"), 
-           n.cgroup = c(4, 2),
-           padding.rgroup = "&nbsp;&nbsp;",
-           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
-           align = "llcccc", # column alignment,
-           tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within July to October fire season."
-            ) # end table
-
-  print(tab)
-
-# ggplot of odds ratios, facet wrapped by outcomes -----
-# convert variables from character to either numeric or factor
-# factor preserves the order of the variable
-wrf_geo_sex_plot <- combined_point_est_df %>% 
-  # filter columns to just geo-smk
- #  filter(pm_method == "GWR Smoke" | 
-   #       pm_method == "WRF-Chem Smoke") %>% 
-  mutate(sex_cat = ifelse(sex == 0, "Male", 
-               ifelse(sex == 1, "Female", NA))) %>%
-  # subset columns I want to put in to the table
-  select(1, 3, 11, 5:8) %>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
-         lower95 = ifelse(lower95 == 99, NA, lower95),
-         upper95 = ifelse(upper95 == 99, NA, upper95))
-
-# change characters to numeric and factor  
-wrf_geo_sex_plot$outcome <- factor(wrf_geo_sex_plot$outcome,
-                              levels = unique(wrf_geo_sex_plot$outcome))
-
-wrf_geo_sex_plot$pm_method <- factor(wrf_geo_sex_plot$pm_method,
-                                levels = unique(wrf_geo_sex_plot$pm_method))
-
-wrf_geo_sex_plot$sex_cat <- factor(wrf_geo_sex_plot$sex_cat,
-                                levels = unique(wrf_geo_sex_plot$sex_cat))
-
-wrf_geo_sex_plot$n_events <- as.numeric(wrf_geo_sex_plot$n_events)
-wrf_geo_sex_plot$odds_ratio <- as.numeric(wrf_geo_sex_plot$odds_ratio)
-wrf_geo_sex_plot$lower95 <- as.numeric(wrf_geo_sex_plot$lower95)
-wrf_geo_sex_plot$upper95 <- as.numeric(wrf_geo_sex_plot$upper95)
-
-## ggplot
-  print_plot <- ggplot(wrf_geo_sex_plot,
-    aes(x = sex_cat, y = odds_ratio, color = pm_method), na.rm = T) +
-    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
-    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
-                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
-    # custom color 
-    scale_color_manual(name = "Smoke-Estimation Method", 
-                       values = c("red", "blue", "#32115C"),
-                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
-    facet_wrap(~outcome, nrow = 3, scales = "free") +
-    geom_hline(yintercept = 1, linetype=2) +
-    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
-    ylab(expression(paste("Odds Ratio for 10Âµg/m"^3, " Increase in PM"[2.5]))) +
-    #ylim(0, 2) +
-    xlab('Sex') +
-    # plot theme
-    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    # strip element
-    strip.background = element_rect(colour=NA, fill=NA),
-    panel.border = element_rect(fill = NA, color = "black"),
-    # facet text size
-    strip.text = element_text(size = 10),
-    # axis element
-    #axis.text.x = element_blank(),
-    #axis.title.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.title.y = element_text(angle = 90),
-    # legend elements
-    legend.position = "bottom")
-    #legend.text = element_text(size = 8))
-
-  print(print_plot)
-  # save figure
-  ggsave("../plot_new/ndc_sex_fig_gwr.pdf", plot = print_plot, 
-       width = 12, height = 8, units = "in")
-
-```
-
-### Find the smallest of AIC with df from 3 to 6 (10)
+### Find the smallest of AIC with df from 3 to 6
 
@@
 
-```{r, message=FALSE, echo=FALSE, warning=FALSE}
+```{r, message=FALSE, echo = FALSE, warning=FALSE}
 
@@
 ```
-
-### Distributed lag -7 to 0 for casecrossover
-
-Then we make the reversed lag to check with the same df of 5. The plot shows the same result as previous analysis.
-
-```{r DLM for casecrossover lag -7 to 0 df 5, echo = FALSE, message=FALSE, warning=FALSE}
-inhaler_geo_lag <- inhaler_ndc_lag_df %>%
-  select(166:160, 159) # GWR-method
-
-inhaler_temp_lag <- inhaler_ndc_lag_df %>%
-  select(156:150, 93) # GWR-method
-
-lag_max <- 0
-n1 <- length(inhaler_geo_lag[,1])
-n2 <- length(inhaler_temp_lag[,1])
-
-inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
-inhaler_temp_lag <- as.matrix(inhaler_temp_lag)
-
-B <- ns(-7:0, df =5, intercept=TRUE)
-
-# the nrow B should equal ncol(pm)
-# dim(B)
-
-inhaler_geo_lag_B <- inhaler_geo_lag%*%B
-inhaler_temp_lag_B <- inhaler_temp_lag%*%B
-
-fit <- clogit(outcome ~ inhaler_geo_lag_B + inhaler_temp_lag_B + strata(personkey), 
-              inhaler_ndc_lag_df)
-
-# coef(fit)
-# summary(fit)
-# AIC(fit)
-
-D <- cbind(B,B)
-
-# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
-dlparms <- c(1,2,3,4,5,6,7,8,9,10)
-DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms]) 
-DLvar <- D%*%vcov(fit)[dlparms,dlparms]%*%t(D)
-DLestimate$SE <- sqrt(diag(DLvar))
-DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
-DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
-
-DLestimate$estimate <- exp(DLestimate$estimate)
-DLestimate$lower <- exp(DLestimate$lower)
-DLestimate$upper <- exp(DLestimate$upper)
-
-# DLestimate
-
-DLestimate$lag <- -7:0
-p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
-p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
-p <- p + geom_line(aes(x=lag), linetype=2, size=2)
-p <- p + theme_bw() + xlab("lagged days of events") + ylab("Odds Ratio for 10ug/m^3 Increase in PM2.5")
-p
-
-DLestimate<- data.frame(estimate=D%*%coef(fit)[dlparms])
-
-cumulative <- sum(DLestimate$estimate)
-# cumulative
-
-cumulative_se <- sqrt(sum(DLvar))
-# cumulative_se
-
-cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
-# cumulative_CI
-
-cum <- cbind(cumulative, cumulative_se, cumulative_CI[1], cumulative_CI[2])
-colnames(cum) <- c("Cumulative Est", "Cumulative SE", "Lower 95% CI", "Upper 95% CI")
-
-knitr::kable(cum, caption = paste0("Distributed lag -7 to 0 spline estimates, se and CI"))
-
-```
-
-```{r time stratified zip for casecrossover lag -7 to 0 prepare, echo=FALSE, eval=FALSE}
-### Time stratified ------------------------------------------------------------
-# read in zipcode level populatoin-weighted pm
-read_path5 <- paste0('C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire_new/data/Oregon_PM/zip_pm_to_merge_with_acap.csv')
-
-zip_smoke <- read_csv(read_path5) # 63801 rows
-
-# descriptives of the two smoke datasets
-summary(zip_smoke)
-
-# Zipcode PM2.5 estimates
-# create lag variables that take smoke values from n previous days for zipcodes
-zip_smoke_w_lag <- zip_smoke %>% arrange(ZIPCODE, date) %>%
-  # group by zipcode
-  group_by(ZIPCODE) %>% 
-  # wrf
-  mutate(wrf_f_pm_lag1 = lag(wrf_f_pm, 1, order_by = ZIPCODE), 
-         wrf_f_pm_lag2 = lag(wrf_f_pm, 2, order_by = ZIPCODE),
-         wrf_f_pm_lag3 = lag(wrf_f_pm, 3, order_by = ZIPCODE),
-         wrf_f_pm_lag4 = lag(wrf_f_pm, 4, order_by = ZIPCODE),
-         wrf_f_pm_lag5 = lag(wrf_f_pm, 5, order_by = ZIPCODE),
-         wrf_f_pm_lag6 = lag(wrf_f_pm, 6, order_by = ZIPCODE),
-         wrf_f_pm_lag7 = lag(wrf_f_pm, 7, order_by = ZIPCODE),
-         # wrf no fire lag
-         wrf_nf_pm_lag1 = lag(wrf_nf_pm, 1, order_by = ZIPCODE),
-         wrf_nf_pm_lag2 = lag(wrf_nf_pm, 2, order_by = ZIPCODE),
-         wrf_nf_pm_lag3 = lag(wrf_nf_pm, 3, order_by = ZIPCODE),
-         wrf_nf_pm_lag4 = lag(wrf_nf_pm, 4, order_by = ZIPCODE),
-         wrf_nf_pm_lag5 = lag(wrf_nf_pm, 5, order_by = ZIPCODE),
-         wrf_nf_pm_lag6 = lag(wrf_nf_pm, 6, order_by = ZIPCODE),
-         wrf_nf_pm_lag7 = lag(wrf_nf_pm, 7, order_by = ZIPCODE),
-         # wrf_smk_pm
-         wrf_smk_pm_lag1 = lag(wrf_smk_pm, 1, order_by = ZIPCODE),
-         wrf_smk_pm_lag2 = lag(wrf_smk_pm, 2, order_by = ZIPCODE),
-         wrf_smk_pm_lag3 = lag(wrf_smk_pm, 3, order_by = ZIPCODE),
-         wrf_smk_pm_lag4 = lag(wrf_smk_pm, 4, order_by = ZIPCODE),
-         wrf_smk_pm_lag5 = lag(wrf_smk_pm, 5, order_by = ZIPCODE),
-         wrf_smk_pm_lag6 = lag(wrf_smk_pm, 6, order_by = ZIPCODE),
-         wrf_smk_pm_lag7 = lag(wrf_smk_pm, 7, order_by = ZIPCODE),
-         # geo weighted pm
-         geo_wt_pm_lag1 = lag(geo_wt_pm, 1, order_by = ZIPCODE),
-         geo_wt_pm_lag2 = lag(geo_wt_pm, 2, order_by = ZIPCODE),
-         geo_wt_pm_lag3 = lag(geo_wt_pm, 3, order_by = ZIPCODE),
-         geo_wt_pm_lag4 = lag(geo_wt_pm, 4, order_by = ZIPCODE),
-         geo_wt_pm_lag5 = lag(geo_wt_pm, 5, order_by = ZIPCODE),
-         geo_wt_pm_lag6 = lag(geo_wt_pm, 6, order_by = ZIPCODE),
-         geo_wt_pm_lag7 = lag(geo_wt_pm, 7, order_by = ZIPCODE),
-         # krig pm
-         krig_pm_lag1 = lag(krig_pm, 1, order_by = ZIPCODE),
-         krig_pm_lag2 = lag(krig_pm, 2, order_by = ZIPCODE),
-         krig_pm_lag3 = lag(krig_pm, 3, order_by = ZIPCODE),
-         krig_pm_lag4 = lag(krig_pm, 4, order_by = ZIPCODE),
-         krig_pm_lag5 = lag(krig_pm, 5, order_by = ZIPCODE), 
-         krig_pm_lag6 = lag(krig_pm, 6, order_by = ZIPCODE),
-         krig_pm_lag7 = lag(krig_pm, 7, order_by = ZIPCODE), 
-         # background pm
-         background_pm_lag1 = lag(background_pm, 1, order_by = ZIPCODE),
-         background_pm_lag2 = lag(background_pm, 2, order_by = ZIPCODE),
-         background_pm_lag3 = lag(background_pm, 3, order_by = ZIPCODE),
-         background_pm_lag4 = lag(background_pm, 4, order_by = ZIPCODE),
-         background_pm_lag5 = lag(background_pm, 5, order_by = ZIPCODE), 
-         background_pm_lag6 = lag(background_pm, 6, order_by = ZIPCODE),
-         background_pm_lag7 = lag(background_pm, 7, order_by = ZIPCODE), 
-         # geo_smk_pm 
-         geo_smk_pm_lag1 = lag(geo_smk_pm, 1, order_by = ZIPCODE),
-         geo_smk_pm_lag2 = lag(geo_smk_pm, 2, order_by = ZIPCODE),
-         geo_smk_pm_lag3 = lag(geo_smk_pm, 3, order_by = ZIPCODE),
-         geo_smk_pm_lag4 = lag(geo_smk_pm, 4, order_by = ZIPCODE),
-         geo_smk_pm_lag5 = lag(geo_smk_pm, 5, order_by = ZIPCODE),
-         geo_smk_pm_lag6 = lag(geo_smk_pm, 6, order_by = ZIPCODE),
-         geo_smk_pm_lag7 = lag(geo_smk_pm, 7, order_by = ZIPCODE),
-         # krig smk pm
-         krig_smk_pm_lag1 = lag(krig_smk_pm, 1, order_by = ZIPCODE),
-         krig_smk_pm_lag2 = lag(krig_smk_pm, 2, order_by = ZIPCODE),
-         krig_smk_pm_lag3 = lag(krig_smk_pm, 3, order_by = ZIPCODE),
-         krig_smk_pm_lag4 = lag(krig_smk_pm, 4, order_by = ZIPCODE),
-         krig_smk_pm_lag5 = lag(krig_smk_pm, 5, order_by = ZIPCODE),
-         krig_smk_pm_lag6 = lag(krig_smk_pm, 6, order_by = ZIPCODE),
-         krig_smk_pm_lag7 = lag(krig_smk_pm, 7, order_by = ZIPCODE),
-         # temp
-         wrf_temp_lag1 = lag(wrf_temp, 1, order_by = ZIPCODE),
-         wrf_temp_lag2 = lag(wrf_temp, 2, order_by = ZIPCODE),
-         wrf_temp_lag3 = lag(wrf_temp, 3, order_by = ZIPCODE),
-         wrf_temp_lag4 = lag(wrf_temp, 4, order_by = ZIPCODE),
-         wrf_temp_lag5 = lag(wrf_temp, 5, order_by = ZIPCODE), 
-         wrf_temp_lag6 = lag(wrf_temp, 6, order_by = ZIPCODE),
-         wrf_temp_lag7 = lag(wrf_temp, 7, order_by = ZIPCODE),
-         # geo lead
-         geo_smk_pm_lag_m1 = lead(geo_smk_pm, 1, order_by = ZIPCODE),
-         geo_smk_pm_lag_m2 = lead(geo_smk_pm, 2, order_by = ZIPCODE),
-         geo_smk_pm_lag_m3 = lead(geo_smk_pm, 3, order_by = ZIPCODE),
-         geo_smk_pm_lag_m4 = lead(geo_smk_pm, 4, order_by = ZIPCODE),
-         geo_smk_pm_lag_m5 = lead(geo_smk_pm, 5, order_by = ZIPCODE),
-         geo_smk_pm_lag_m6 = lead(geo_smk_pm, 6, order_by = ZIPCODE),
-         geo_smk_pm_lag_m7 = lead(geo_smk_pm, 7, order_by = ZIPCODE),
-         # temp lead
-         wrf_temp_lag_m1 = lead(wrf_temp, 1, order_by = ZIPCODE),
-         wrf_temp_lag_m2 = lead(wrf_temp, 2, order_by = ZIPCODE),
-         wrf_temp_lag_m3 = lead(wrf_temp, 3, order_by = ZIPCODE),
-         wrf_temp_lag_m4 = lead(wrf_temp, 4, order_by = ZIPCODE),
-         wrf_temp_lag_m5 = lead(wrf_temp, 5, order_by = ZIPCODE), 
-         wrf_temp_lag_m6 = lead(wrf_temp, 6, order_by = ZIPCODE),
-         wrf_temp_lag_m7 = lead(wrf_temp, 7, order_by = ZIPCODE)) %>% 
-  # ungroup by zip
-  ungroup(ZIPCODE) %>% 
-  # attach a zip indicator for each smoke variable
-  setNames(paste(colnames(.), "zip", sep="_"))  %>% 
-  # remove the '_zip' from the zipcode and date variable 
-  rename(ZIPCODE = ZIPCODE_zip, date = date_zip)
-
-read_path4 <- paste0("../data_new/medication/oregon_ndc_casecrossover.csv")
-# read_path4 <- paste0("../../../data/data_new/oregon_ndc_casecrossover.csv") # for server
-or_disease <- read_csv(read_path4)
-
-### try join
-colnames(or_disease)[24] <- c("ZIPCODE")
-colnames(or_disease)[74] <- c("date")
-
-outcome_casecross <- or_disease %>%
-  # indicator for male=0, female=1, unknown = 2
-  mutate(age = 2013-yob,
-         sex_ind =ifelse(gender == "F", 1, 
-                         ifelse(gender == "M", 0, 2)),
-         age_ind = ifelse(age < 15, 0,
-                          ifelse(age >= 15 & age < 65, 1,
-                                 ifelse(age >= 65 & age <=105, 2, NA)))
-  ) %>% # end of mutate 
-  # create variables
-  mutate(day = as.factor(weekdays(fromdate)),
-         day_admit = as.factor(weekdays(date)),
-         month_smk = month(fromdate),
-         month_admit = month(date),
-         season_smk = ifelse(fromdate >= "2013-03-20" &  
-                               fromdate <= "2013-06-21", "spring",
-                             ifelse(fromdate >= "2013-06-22" &  
-                                      fromdate <= "2013-09-22", "summer",
-                                    ifelse(fromdate >= "2013-09-23" & 
-                                             fromdate <= "2013-12-21", "fall", "other"))),
-         season_admit = ifelse(date >= "2013-03-20" &  
-                                 date <= "2013-06-21", "spring",
-                               ifelse(date >= "2013-06-22" &  
-                                        date <= "2013-09-22", "summer",
-                                      ifelse(date >= "2013-09-23" & 
-                                               date <= "2013-12-21", "fall", "other")))) %>%
-  # join with zip-level pm estimates
-  left_join(zip_smoke_w_lag, by = c("date", "ZIPCODE"))%>%
-  arrange(personkey, fromdate) # order by id and fromdate
-
-write_path4 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover_zip_lag15.csv")
-write_csv(outcome_casecross, write_path4)
-
-```
-
-```{r time stratified county for ts lag -7 to 0 prepare, echo=FALSE, include=FALSE, eval=FALSE}
-read_path7 <- paste0("../data_new/medication/oregon_asthma_ndc.csv")
-or_ndc_claim <- read_csv(read_path7)
-
-or_ndc_inhaler <- or_ndc_claim %>%
-  select(personkey, num_visit, clmid, line, ZIP, dates) %>%
-  rename(ZIPCODE = ZIP)
-
-# Join data with names of Oregon counties 
-oregon_fips <- read_csv('../instructions/oregon_FIPS.csv')
-# remove the "County" character
-# factor(oregon_fips$county)
-oregon_fips$county <- gsub(" County", "", as.character(factor(oregon_fips$county)))
-
-oregon_fips <- oregon_fips %>%
-  mutate(st_county_fips = with(oregon_fips, paste0(st_code, fips)))
-
-### Join county with zip
-# read_path <- paste0('../data_new/update/or_zip_county_prop.csv') 
-# or_zip_county <- read_csv(read_path)
-
-read_path <- paste0('../instructions/oregon_zip_county.csv')
-or_zip_county <- read_csv(read_path)
-
-# or_zip_county$county_name[which(or_zip_county$county_name=="Hood.River")] <- "Hood River"
-
-or_zip_county <- or_zip_county %>%
-  select(zip, county_name) %>%
-  rename(county = county_name) %>%
-  rename(ZIPCODE = zip) %>%
-  full_join(oregon_fips, by = "county") %>%
-  select(ZIPCODE, county, state, st_county_fips) %>%
-  rename(fips = st_county_fips) # 484
-
-county_new <- c("Washington", "Douglas", "Lane", "Deschutes")
-
-zip_new <- or_ndc_inhaler %>%
-  filter(!(ZIPCODE %in% or_zip_county$ZIPCODE)) %>%
-  select(ZIPCODE) %>%
-  unique() %>%
-  arrange(ZIPCODE) %>%
-  mutate(county = county_new)
-
-# summary(as.factor(zip_new$ZIPCODE))
-# 97003 97471 97475 97703 
-#    48   138     5    16
-
-or_zip_county_new <- or_zip_county %>%
-  select(ZIPCODE, county) %>%
-  bind_rows(zip_new)
-
-write_path <- paste0("../data_new/county_data/zip_to_county_new.csv")
-write_csv(or_zip_county_new, write_path)
-
-read_path9 <- paste0("../data_new/county_data/or_census_pop_est_county.csv")
-or_pop <- read_csv(read_path9)
-
-or_pop <- or_pop %>%
-  rename(county = GEO_display_label, pop = respop72013) %>%
-  select(county, pop) 
-or_pop$county <- gsub(" County, Oregon", "", as.character(factor(or_pop$county)))
-
-### import county data
-read_path_county_pm <- paste0('C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire_new/data_new/county_data/or_county_pop_wt_pm.csv')
-
-county_smoke <- read_csv(read_path_county_pm) # 63801 rows
-
-# descriptives of the two smoke datasets
-summary(county_smoke)
-
-county_smoke$county[which(county_smoke$county=="Hood.River")] <- "Hood River"
-
-# Zipcode PM2.5 estimates
-# create lag variables that take smoke values from n previous days for zipcodes
-county_smoke_w_lag <- county_smoke %>% arrange(county, date) %>%
-  # group by zipcode
-  group_by(county) %>% 
-  # wrf
-  mutate(wrf_f_pm_lag1 = lag(wrf_pm, 1, order_by = county), 
-         wrf_f_pm_lag2 = lag(wrf_pm, 2, order_by = county),
-         wrf_f_pm_lag3 = lag(wrf_pm, 3, order_by = county),
-         wrf_f_pm_lag4 = lag(wrf_pm, 4, order_by = county),
-         wrf_f_pm_lag5 = lag(wrf_pm, 5, order_by = county),
-         wrf_f_pm_lag6 = lag(wrf_pm, 6, order_by = county),
-         wrf_f_pm_lag7 = lag(wrf_pm, 7, order_by = county),
-         # wrf no fire lag
-         wrf_nf_pm_lag1 = lag(wrf_nf_pm, 1, order_by = county),
-         wrf_nf_pm_lag2 = lag(wrf_nf_pm, 2, order_by = county),
-         wrf_nf_pm_lag3 = lag(wrf_nf_pm, 3, order_by = county),
-         wrf_nf_pm_lag4 = lag(wrf_nf_pm, 4, order_by = county),
-         wrf_nf_pm_lag5 = lag(wrf_nf_pm, 5, order_by = county),
-         wrf_nf_pm_lag6 = lag(wrf_nf_pm, 6, order_by = county),
-         wrf_nf_pm_lag7 = lag(wrf_nf_pm, 7, order_by = county),
-         # wrf_smk_pm
-         wrf_smk_pm_lag1 = lag(wrf_smk_pm, 1, order_by = county),
-         wrf_smk_pm_lag2 = lag(wrf_smk_pm, 2, order_by = county),
-         wrf_smk_pm_lag3 = lag(wrf_smk_pm, 3, order_by = county),
-         wrf_smk_pm_lag4 = lag(wrf_smk_pm, 4, order_by = county),
-         wrf_smk_pm_lag5 = lag(wrf_smk_pm, 5, order_by = county),
-         wrf_smk_pm_lag6 = lag(wrf_smk_pm, 6, order_by = county),
-         wrf_smk_pm_lag7 = lag(wrf_smk_pm, 7, order_by = county),
-         # geo weighted pm
-         geo_wt_pm_lag1 = lag(geo_wt_pm, 1, order_by = county),
-         geo_wt_pm_lag2 = lag(geo_wt_pm, 2, order_by = county),
-         geo_wt_pm_lag3 = lag(geo_wt_pm, 3, order_by = county),
-         geo_wt_pm_lag4 = lag(geo_wt_pm, 4, order_by = county),
-         geo_wt_pm_lag5 = lag(geo_wt_pm, 5, order_by = county),
-         geo_wt_pm_lag6 = lag(geo_wt_pm, 6, order_by = county),
-         geo_wt_pm_lag7 = lag(geo_wt_pm, 7, order_by = county),
-         # krig pm
-         krig_pm_lag1 = lag(krig_pm, 1, order_by = county),
-         krig_pm_lag2 = lag(krig_pm, 2, order_by = county),
-         krig_pm_lag3 = lag(krig_pm, 3, order_by = county),
-         krig_pm_lag4 = lag(krig_pm, 4, order_by = county),
-         krig_pm_lag5 = lag(krig_pm, 5, order_by = county), 
-         krig_pm_lag6 = lag(krig_pm, 6, order_by = county),
-         krig_pm_lag7 = lag(krig_pm, 7, order_by = county), 
-         # background pm
-         background_pm_lag1 = lag(background_pm, 1, order_by = county),
-         background_pm_lag2 = lag(background_pm, 2, order_by = county),
-         background_pm_lag3 = lag(background_pm, 3, order_by = county),
-         background_pm_lag4 = lag(background_pm, 4, order_by = county),
-         background_pm_lag5 = lag(background_pm, 5, order_by = county), 
-         background_pm_lag6 = lag(background_pm, 6, order_by = county),
-         background_pm_lag7 = lag(background_pm, 7, order_by = county), 
-         # geo_smk_pm 
-         geo_smk_pm_lag1 = lag(geo_smk_pm, 1, order_by = county),
-         geo_smk_pm_lag2 = lag(geo_smk_pm, 2, order_by = county),
-         geo_smk_pm_lag3 = lag(geo_smk_pm, 3, order_by = county),
-         geo_smk_pm_lag4 = lag(geo_smk_pm, 4, order_by = county),
-         geo_smk_pm_lag5 = lag(geo_smk_pm, 5, order_by = county),
-         geo_smk_pm_lag6 = lag(geo_smk_pm, 6, order_by = county),
-         geo_smk_pm_lag7 = lag(geo_smk_pm, 7, order_by = county),
-         # krig smk pm
-         krig_smk_pm_lag1 = lag(krig_smk_pm, 1, order_by = county),
-         krig_smk_pm_lag2 = lag(krig_smk_pm, 2, order_by = county),
-         krig_smk_pm_lag3 = lag(krig_smk_pm, 3, order_by = county),
-         krig_smk_pm_lag4 = lag(krig_smk_pm, 4, order_by = county),
-         krig_smk_pm_lag5 = lag(krig_smk_pm, 5, order_by = county),
-         krig_smk_pm_lag6 = lag(krig_smk_pm, 6, order_by = county),
-         krig_smk_pm_lag7 = lag(krig_smk_pm, 7, order_by = county),
-         # temp
-         wrf_temp_lag1 = lag(wrf_temp, 1, order_by = county),
-         wrf_temp_lag2 = lag(wrf_temp, 2, order_by = county),
-         wrf_temp_lag3 = lag(wrf_temp, 3, order_by = county),
-         wrf_temp_lag4 = lag(wrf_temp, 4, order_by = county),
-         wrf_temp_lag5 = lag(wrf_temp, 5, order_by = county), 
-         wrf_temp_lag6 = lag(wrf_temp, 6, order_by = county),
-         wrf_temp_lag7 = lag(wrf_temp, 7, order_by = county),
-          # geo lead
-         geo_smk_pm_lag_m1 = lead(geo_smk_pm, 1, order_by = county),
-         geo_smk_pm_lag_m2 = lead(geo_smk_pm, 2, order_by = county),
-         geo_smk_pm_lag_m3 = lead(geo_smk_pm, 3, order_by = county),
-         geo_smk_pm_lag_m4 = lead(geo_smk_pm, 4, order_by = county),
-         geo_smk_pm_lag_m5 = lead(geo_smk_pm, 5, order_by = county),
-         geo_smk_pm_lag_m6 = lead(geo_smk_pm, 6, order_by = county),
-         geo_smk_pm_lag_m7 = lead(geo_smk_pm, 7, order_by = county),
-         # temp lead
-         wrf_temp_lag_m1 = lead(wrf_temp, 1, order_by = county),
-         wrf_temp_lag_m2 = lead(wrf_temp, 2, order_by = county),
-         wrf_temp_lag_m3 = lead(wrf_temp, 3, order_by = county),
-         wrf_temp_lag_m4 = lead(wrf_temp, 4, order_by = county),
-         wrf_temp_lag_m5 = lead(wrf_temp, 5, order_by = county), 
-         wrf_temp_lag_m6 = lead(wrf_temp, 6, order_by = county),
-         wrf_temp_lag_m7 = lead(wrf_temp, 7, order_by = county)) %>% 
-  # ungroup by zip
-  ungroup(county) %>% 
-  # attach a zip indicator for each smoke variable
-  setNames(paste(colnames(.), "county", sep="_"))  %>% 
-  # remove the '_zip' from the zipcode and date variable 
-  rename(county = county_county, date = date_county, fips = fips_county) %>%
-  select(-1)
-
-or_ndc_county_ts <- or_ndc_inhaler %>%
-  rename(date = dates) %>%
-  mutate(day = ifelse(weekdays(date) == "Monday" |
-                      weekdays(date) == "Tuesday" |
-                      weekdays(date) == "Wednesday"|
-                      weekdays(date) == "Thursday"|
-                      weekdays(date) == "Friday", "weekday",
-               ifelse(weekdays(date) == "Saturday" |
-                      weekdays(date) == "Sunday", "weekend", NA))) %>%
-  # join with county-level population
-  left_join(or_zip_county_new, by = "ZIPCODE") %>% # 321 missing 
-  filter(!is.na(county)) %>% # no missing value
-  left_join(or_pop, by = "county") %>%
-  left_join(county_smoke_w_lag, by = c("date", "county")) %>%
-  arrange(personkey, date) %>% # order by id and fromdate
-  group_by(county, date) %>% 
-  # sum up each primary diagnosis for each outcome for each day for each county
-  summarise(n_obs = n()) %>% 
-  # join daily county estimates of pm2.5
-  left_join(county_smoke_w_lag, by = c('date', 'county'))
-
-# summary(as.factor(or_ndc_county_ts$county))
-
-or_ndc_county_ts <- or_ndc_county_ts %>%
-  left_join(or_pop, by = "county") %>%
-  mutate(geo_smk_pm10 = geo_smk_pm_county/10) %>%
-  mutate(rate_per_100k = (n_obs/pop)*100000) %>%
-  mutate(day = ifelse(weekdays(date) == "Monday" |
-                      weekdays(date) == "Tuesday" |
-                      weekdays(date) == "Wednesday"|
-                      weekdays(date) == "Thursday"|
-                      weekdays(date) == "Friday", "weekday",
-               ifelse(weekdays(date) == "Saturday" |
-                      weekdays(date) == "Sunday", "weekend", NA)))
-
-or_ndc_county_ts$day[which(or_ndc_county_ts$date=="2013-05-27")] <- "weekend"
-or_ndc_county_ts$day[which(or_ndc_county_ts$date=="2013-07-04")] <- "weekend"
-or_ndc_county_ts$day[which(or_ndc_county_ts$date=="2013-09-02")] <- "weekend"
-
-write_path5 <- paste0("../data_new/medication/or_ndc_county_time_series_lag15.csv")
-write_csv(or_ndc_county_ts, write_path5)
-
-```
-
-```{r time stratified zip for casecrossover lag -7 to 0 df, echo=FALSE, include=FALSE, eval=FALSE}
-read_path10 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover_zip_lag15.csv")
-inhaler_ndc_lag15 <- read_csv(read_path10)
-# asthma_ndc <- fread(read_path6, showProgress = T) # faster, but seems error in overall plot analysis
-
-# dataframe to loop through
-inhaler_ndc_lag15_df <- data.frame(inhaler_ndc_lag15)
-
-inhaler_ndc_lag_15_df <- inhaler_ndc_lag15_df %>%
-  mutate(geo_smk_pm10_lag_m1_zip =geo_smk_pm_lag_m1_zip/10,
-         geo_smk_pm10_lag_m2_zip =geo_smk_pm_lag_m2_zip/10,
-         geo_smk_pm10_lag_m3_zip =geo_smk_pm_lag_m3_zip/10,
-         geo_smk_pm10_lag_m4_zip =geo_smk_pm_lag_m4_zip/10,
-         geo_smk_pm10_lag_m5_zip =geo_smk_pm_lag_m5_zip/10,
-         geo_smk_pm10_lag_m6_zip =geo_smk_pm_lag_m6_zip/10,
-         geo_smk_pm10_lag_m7_zip =geo_smk_pm_lag_m7_zip/10) %>%
-  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
-         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
-         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
-         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
-         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
-         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
-         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
-         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)
-
-# which(colnames(inhaler_ndc_lag_15_df)=="geo_smk_pm10_zip") # 178
-# which(colnames(inhaler_ndc_lag_15_df)=="geo_smk_pm10_lag_m1_zip") # 171
-# which(colnames(inhaler_ndc_lag_15_df)=="geo_smk_pm10_lag1_zip") # 179
-
-# which(colnames(inhaler_ndc_lag_15_df)=="wrf_temp_zip") # 93
-# which(colnames(inhaler_ndc_lag_15_df)=="wrf_temp_lag1_zip") # 150
-# which(colnames(inhaler_ndc_lag_15_df)=="wrf_temp_lag_m1_zip") # 164
-
-inhaler_geo_lag <- inhaler_ndc_lag_15_df %>%
-  select(177:171, 178) # GWR-method
-
-inhaler_temp_lag <- inhaler_ndc_lag_15_df %>%
-  select(170:164, 93) # temp
-
-inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
-inhaler_temp_lag <- as.matrix(inhaler_temp_lag)
-
-# create fit matrix
-fit_mat <- matrix(NA, nrow=4, ncol=5)
-colnames(fit_mat) <- c("outcome", "exposure", "fit", "df", "aic")
-
-# fill outcome name in first column
-fit_mat[,1] <- "Beta - 2 agonists"
-# fill in exposure (geo smoke)
-fit_mat[,2] <- "geo_smk10"
-# fit type
-fit_mat[,3] <- "ns"
-
-# 2 df will not work for a spline, going from 3 to 6 df
-  for(h in 3:6){ # start loop for df
-
-    # fitting a natural spline for lag with 5 knots?
-    fit_mat[h-2, 4] <- degree_freedom <- h
-
-    # define basis b using natural spline function
-    B <- ns(-7:0, df = degree_freedom, intercept = T)
-
-    inhaler_geo_lag_B <- inhaler_geo_lag%*%B
-    inhaler_temp_lag_B <- inhaler_temp_lag%*%B
-
-    fit <- clogit(outcome ~ inhaler_geo_lag_B + inhaler_temp_lag_B + strata(personkey), 
-              inhaler_ndc_lag_15_df)
-
-    AIC(fit)
-
-    # fill AIC
-    fit_mat[h-2,5] <- round(AIC(fit),5)
-
-    } # end model fit loop
-
-# now that we have a range of df/knots, I want to filter to the minimum aic
-# for each outcome and print out that value in a table
-
-fit_mat
-
-lag_spline_best_fit <- as_tibble(fit_mat) %>%  
-  slice(which.min(aic)) # df of 3 has the smallest AIC
-
-# kable
-knitr::kable(lag_spline_best_fit, caption = paste0("Distributed lag spline",
-  "degree of freedom best fit by AIC for mixed model approach"))
-
-```
-
-### 1. For Overall Plot (Unconstrained lag model)
-
-```{r overall lags, warning =F, echo = F, results='asis', eval=FALSE} 
-# dataframe list
-method_list <- c('Geo-Weighted Smoke')
-
-lag_list <- c("Lag 1", "Lag 2", "Lag 3", "Lag 4", "Lag 5", "Lag 6", "Lag 7")
-
-# dataframe to loop through
-inhaler_ndc <- data.frame(inhaler_ndc) 
-#  filter(!is.na(geo_smk_pm_lag1_zip))
-
-# outcome name
-outcome <- "Beta - 2 agonists"
-
-# data wrangling ----
-# Producing conditional logit model estimates loop
-
-# extract covariates from dataframe
-covariates_df <- inhaler_ndc[, c(1:26, 74:79)] # 94:156 is lag
-
-# extract pm values and divide by 10 and ordered
-# which(colnames(inhaler_ndc)=="geo_smk_pm_lag1_zip") # 136
-# which(colnames(inhaler_ndc)=="wrf_temp_lag1_zip") # 150
-pm_estimates_df <- inhaler_ndc[, 136:142]/10  # create 10 unit increases
-temp_estimates_df <- inhaler_ndc[, 150:156]  # temperature
-
-# new loop for age categories
-
-# empty matrix (12 x 10 matrix)
-point_estimates <- matrix(nrow = 7, ncol = 10, byrow = T)
-
-colnames(point_estimates) <- c('outcome', 'pm_method',  'lag', 'n', 'n_events', 
-                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')
-
-# fill in the outcome namedataframe before method loop
-point_estimates[, 1] <- outcome
-
-for (i in 1:7){
-geo_lag <- i + 32 
-temp_lag <- i + 39
-
-# dataframe for analysis creation
-# bind columns back together 
-df_analysis <- cbind(covariates_df, pm_estimates_df, temp_estimates_df)
-
-df_analysis <- df_analysis %>% 
-  # remove missing pm values
-  filter(!is.na(df_analysis[geo_lag])) %>% 
-  # the following code makes sure that the counterfactual values retained are 
-  # symetric in that number of obs before = number of obs after
-  mutate(obs_diff_admission = (fromdate - date)/7) 
-  # dataframe is already for the entire fire season, so I don't need to subset anymore
-
-# second loop to run a model for each pm estimation method
-
-  # variable to model 
-  var_name <- colnames(df_analysis[geo_lag])
-
-  # set row number to fill
-  row_n <- i
-
-  # only run the model if the dataframe has observations
-  if(nrow(df_analysis) != 0){
-  # conditional logistic regression model
-  mod <- clogit(outcome ~ df_analysis[[geo_lag]] + df_analysis[[temp_lag]] + strata(personkey), df_analysis)
-
-  # populate matrix
-  point_estimates[row_n, 2] <- method_list
-  point_estimates[row_n, 3] <- lag_list[row_n]
-  point_estimates[row_n, 4] <- mod$n
-  point_estimates[row_n, 5] <- mod$nevent
-  # odds ratio
-  point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
-
-  # 95% lower bound
-  point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
-  # 95% upper bound
-  point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
-                                        1.96*(summary(mod)$coefficient[1,3])), 3)
-  # standard error
-  point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
-  # p val
-  point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)
-
-  # create else statement that fills matrix with missing so I still have the row
-  # in the final dataframe
-  } else {point_estimates[row_n, 3] <- method_list[row_n]
-          point_estimates[row_n, 4] <- 0
-          point_estimates[row_n, 5] <- 0
-          point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
-
-}
-# save point estimates as a dataframe
-combined_point_est_df <- as_data_frame(point_estimates)
-
-wrf_geo_lag <- combined_point_est_df %>% 
-  select(3, 5:8) %>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
-         lower95 = ifelse(lower95 == 99, "--", lower95),
-         upper95 = ifelse(upper95 == 99, "--", upper95))
-
-tab <- htmlTable(txtRound(wrf_geo_lag, digits = 3, 1:3), 
-         caption = "Association between a 10 ug/m^3 in PM2.5 three smoke method and pharmacy outcomes",
-         # row group by outcome
-         rgroup = "Beta - 2 agonists",
-         n.rgroup = c(rep(1, 1)), # 6 rows for each age cat for each outcome
-         # column headers
-         header = c("Lag", "Events",
-                    "OR&dagger;", "Lower", "Upper"),
-         # column spanner
-         cgroup = c("", "95% CI"), 
-         n.cgroup = c(3, 2),
-         padding.rgroup = "&nbsp;&nbsp;",
-         css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
-         align = "llcccc", # column alignment,
-         tfoot="&dagger; Referent periods matched to events on same day of week within May to September fire season."
-         ) # end table
-
-print(tab)
-
-combined_point_est_df_new <- combined_point_est_df %>%
-  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
-                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
-  select(-n_events) %>%
-  rename(n_events = n_events_new)
-
-# ggplot of odds ratios, facet wrapped by outcomes -----
-# convert variables from character to either numeric or factor
-# factor preserves the order of the variable
-wrf_geo_lag_plot <- combined_point_est_df_new %>% 
-  # filter columns to just geo-smk
- #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
-  # do not plot results with less than 15 events, create missing vals
-  # in mutate
-  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
-                             99, odds_ratio),
-         lower95 = ifelse(as.numeric(n_events) < 15, 
-                             99, lower95),
-         upper95 = ifelse(as.numeric(n_events) < 15, 
-                             99, upper95)) %>% 
-  # subset columns I want to put in to the table
-  select(1, 3, 10, 5:7) %>% 
-  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
-         lower95 = ifelse(lower95 == 99, NA, lower95),
-         upper95 = ifelse(upper95 == 99, NA, upper95))
-
-# change characters to numeric and factor  
-wrf_geo_lag_plot$outcome <- factor(wrf_geo_lag_plot$outcome,
-                              levels = unique(wrf_geo_lag_plot$outcome))
-
-wrf_geo_lag_plot$lag <- factor(wrf_geo_lag_plot$lag,
-                                levels = unique(wrf_geo_lag_plot$lag))
-
-wrf_geo_lag_plot$n_events <- as.numeric(wrf_geo_lag_plot$n_events)
-wrf_geo_lag_plot$odds_ratio <- as.numeric(wrf_geo_lag_plot$odds_ratio)
-wrf_geo_lag_plot$lower95 <- as.numeric(wrf_geo_lag_plot$lower95)
-wrf_geo_lag_plot$upper95 <- as.numeric(wrf_geo_lag_plot$upper95)
-
-## ggplot 
-  print_plot <- ggplot(wrf_geo_lag_plot,
-    aes(x = lag, y = odds_ratio, color = lag), na.rm = T) +
-    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
-    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
-                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
-    # custom color 
-    scale_color_manual(name = "Geo-smk Lag", 
-                       values = rep(c("black"), 7),
-                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
-    facet_wrap(~outcome, nrow = 7, scales = "free") +
-    geom_hline(yintercept = 1, linetype=2) +
-    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
-    ylab(expression(paste("Odds Ratio for 10Âµg/m"^3, " Increase in PM"[2.5]))) +
-    #ylim(0, 2) +
-    xlab('Smoke Estimation Method') +
-    # plot theme
-    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    # strip element
-    strip.background = element_rect(colour=NA, fill=NA),
-    panel.border = element_rect(fill = NA, color = "black"),
-    # facet text size
-    strip.text = element_text(size = 10),
-    # axis element
-    #axis.text.x = element_blank(),
-    #axis.title.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.title.y = element_text(angle = 90),
-    # legend elements
-    legend.position = "bottom")
-    #legend.text = element_text(size = 8))
-
-  print(print_plot)
-  # save figure
-  ggsave("../plot_new/ndc_health_lag.pdf", plot = print_plot, 
-       width = 12, height = 8, units = "in")
-
-```
-
-### 2. Unconstrained lag for time series
-
-```{r unconstrained ts lag, message=FALSE, echo=FALSE, warning=FALSE, eval=FALSE}
-or_ndc_county_ts_df <- or_ndc_county_ts %>%
-  select(1:4, pop, rate_per_100k, day, geo_smk_pm_county,
-         geo_smk_pm_lag1_county:geo_smk_pm_lag7_county, wrf_temp_county,
-         wrf_temp_lag1_county:wrf_temp_lag7_county) %>%
-  mutate(geo_smk_pm10_county = geo_smk_pm_county/10,
-         geo_smk_pm10_lag1_county =geo_smk_pm_lag1_county/10,
-         geo_smk_pm10_lag2_county =geo_smk_pm_lag2_county/10,
-         geo_smk_pm10_lag3_county =geo_smk_pm_lag3_county/10,
-         geo_smk_pm10_lag4_county =geo_smk_pm_lag4_county/10,
-         geo_smk_pm10_lag5_county =geo_smk_pm_lag5_county/10,
-         geo_smk_pm10_lag6_county =geo_smk_pm_lag6_county/10,
-         geo_smk_pm10_lag7_county =geo_smk_pm_lag7_county/10) 
-  # filter(!is.na(geo_smk_pm10_lag7_zip))
-
-# create empty dataframe
-rr_df <- data.frame(matrix(nrow = 7, ncol = 5))
-colnames(rr_df) <- c("method", "lag", "rel_risk", "lower_95", "upper_95")
-
-# vector of outcomes
-outcome_list <- "n_obs"
-
-# methods
-rr_df[,1] <- c("Mixed-Poisson")
-
-# populate outcome column
-rr_df[,2] <- c("Lag 1", "Lag 2", "Lag 3", "Lag 4", "Lag 5", "Lag 6", "Lag 7")
-
-# which(colnames(or_ndc_county_ts_df)=="geo_smk_pm10_lag1_county") #25
-# which(colnames(or_ndc_county_ts_df)=="wrf_temp_lag1_county") #17
-
-for (i in 1:7){
-
-  geo <- i + 24
-  tem <- i +14
-
-## mixed
-smk_est_mp <- tidy(glmer(as.formula(paste0(outcome_list, 
-    "~ or_ndc_county_ts_df[[geo]] + or_ndc_county_ts_df[[tem]] + (1|county) + day + offset(log(pop))")), 
-    or_ndc_county_ts_df, family="poisson"))[2, ]
-
-rr_mp <- c(round(exp(smk_est_mp[1,2]),5),
-         round(exp(smk_est_mp[1,2]-1.96*smk_est_mp[1,3]),5),
-         round(exp(smk_est_mp[1,2]+1.96*smk_est_mp[1,3]),5))
-
-rr_df[i,3:5] <- rr_mp
-}
-
-kable(rr_df, caption = "Unconstrained lag for time series")
-
-plot_rr <- ggplot(rr_df, aes(x=lag, y = rel_risk)) +
-  geom_point() +
-  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
-  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
-  theme(
-   panel.background = element_rect(fill = 'white', colour = 'black'),
-   panel.grid.major = element_blank(),
-   panel.grid.minor = element_blank(),
-   axis.text.x = element_text(angle = 90, vjust = .75)) +
-  ylab(expression(paste("Odds Ratio for 10Âµg/m"^3, " Increase in PM"[2.5]))) +
-  xlab("Outcome") +
-  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")
-
-print(plot_rr)
-
-ggsave("../plot_new/ndc_rr_geo_lag.pdf", plot = plot_rr, 
-       width = 12, height = 8, units = "in")
-
-```
-
-### 3. Distributed lag check for time series
-
-(not shown)
-
-```{r DLM ts, echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}
-which(colnames(or_ndc_county_ts_df)=="wrf_temp_lag1_county") # 17
-which(colnames(or_ndc_county_ts_df)=="geo_smk_pm10_lag1_county") # 25
-
-geo_ts_lag <- or_ndc_county_ts_df %>%
-  select(24:31) # GWR-method
-tem_ts_lag <- or_ndc_county_ts_df %>%
-  select(16:23) # temp
-
-lag_max <- 7
-
-geo_ts_lag <- as.matrix(geo_ts_lag)
-tem_ts_lag <- as.matrix(tem_ts_lag)
-
-B <- ns(0:lag_max, df =3, intercept=TRUE) # or 3?
-
-# the nrow B should equal ncol(pm)
-dim(B)
-dim(geo_ts_lag)
-dim(tem_ts_lag)
-
-geo_ts_lag_B <- geo_ts_lag%*%B
-
-tem_ts_lag_B <- tem_ts_lag%*%B
-
-## mixed model
-mod <-glmer(n_obs ~  geo_ts_lag_B + tem_ts_lag_B + day + (1|county) + offset(log(pop)),   
-            or_ndc_county_ts_df, family="poisson", control = glmerControl(optimizer = "bobyqa"))
-
-summary(mod)
-# check AIC
-AIC(mod)
-
-D <- cbind(B, B)
-# output distributed lag beta parameters
-dlparms <- mod@beta[2:7]
-# estimate distributed lag values for each day
-dl_estimates <- data.frame(estimate = D %*% dlparms)
-
-# covariance matrix for knots (need to convert to matrix object)
-cov_mat <- as.matrix(vcov(mod)[2:7,2:7])
-
-# estimate variance of splines
-dl_var <- D%*%cov_mat%*%t(D)
-
-# calculate standard error for each lag value
-dl_estimates$stderr <- sqrt(diag(dl_var))
-
-# calculate lower and upper bounds
-dl_estimates$lower_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(1-0.975,
-  df=df.residual(mod))  
-dl_estimates$upper_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(0.975,
-  df=df.residual(mod))
-
-# estimate relative risk and 95%CI from beta
-inhaler_dl_rr <- dl_estimates %>% 
-  mutate(estimate = exp(estimate),
-         lower_bound = exp(lower_bound),
-         upper_bound = exp(upper_bound),
-         lag = 0:7)
-
-# estimate cumulative effect
-cumulative <- sum(dl_estimates$estimate)
-cumulative
-
-# estimate cumulative effect stnd error
-cumulative_se <- sqrt(sum(dl_var))
-cumulative_se
-
-# estimate cumulative CI
-cumulative_ci <- cumulative+cumulative_se*qt(c(1-0.975,0.975), 
-                                             df=df.residual(mod))
-cumulative_ci
-
-exp(cumulative)
-exp(cumulative_ci)
-
-# plot 
-p <- ggplot(inhaler_dl_rr, aes(x=lag, y=estimate, 
-                              ymin=lower_bound, ymax=upper_bound)) +
-  geom_ribbon(alpha = 0.5) +
-  geom_line(size=2)
-
-p
-```
-
-```{r DLM ts temp, echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}
-which(colnames(or_ndc_county_ts_df)=="wrf_temp_lag1_county") # 17
-which(colnames(or_ndc_county_ts_df)=="geo_smk_pm10_lag1_county") # 25
-
-geo_ts_lag <- or_ndc_county_ts_df %>%
-  select(24:31) # GWR-method
-
-lag_max <- 7
-
-geo_ts_lag <- as.matrix(geo_ts_lag)
-tem_ts_lag <- as.matrix(tem_ts_lag)
-
-B <- ns(0:lag_max, df =3, intercept=TRUE) # or 3?
-
-# the nrow B should equal ncol(pm)
-dim(B)
-dim(tem_ts_lag)
-
-tem_ts_lag_B <- tem_ts_lag%*%B
-
-## mixed model
-mod <-glmer(n_obs ~ tem_ts_lag_B + day + (1|county) + offset(log(pop)),   
-            or_ndc_county_ts_df, family="poisson", control = glmerControl(optimizer = "bobyqa"))
-
-summary(mod)
-# check AIC
-AIC(mod)
-
-# output distributed lag beta parameters
-dlparms <- mod@beta[2:4]
-# estimate distributed lag values for each day
-dl_estimates <- data.frame(estimate = B %*% dlparms)
-
-# covariance matrix for knots (need to convert to matrix object)
-cov_mat <- as.matrix(vcov(mod)[2:4,2:4])
-
-# estimate variance of splines
-dl_var <- B%*%cov_mat%*%t(B)
-
-# calculate standard error for each lag value
-dl_estimates$stderr <- sqrt(diag(dl_var))
-
-# calculate lower and upper bounds
-dl_estimates$lower_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(1-0.975,
-  df=df.residual(mod))  
-dl_estimates$upper_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(0.975,
-  df=df.residual(mod))
-
-# estimate relative risk and 95%CI from beta
-inhaler_dl_rr <- dl_estimates %>% 
-  mutate(estimate = exp(estimate),
-         lower_bound = exp(lower_bound),
-         upper_bound = exp(upper_bound),
-         lag = 0:7)
-
-# estimate cumulative effect
-cumulative <- sum(dl_estimates$estimate)
-cumulative
-
-# estimate cumulative effect stnd error
-cumulative_se <- sqrt(sum(dl_var))
-cumulative_se
-
-# estimate cumulative CI
-cumulative_ci <- cumulative+cumulative_se*qt(c(1-0.975,0.975), 
-                                             df=df.residual(mod))
-cumulative_ci
-
-exp(cumulative)
-exp(cumulative_ci)
-
-# plot 
-p <- ggplot(inhaler_dl_rr, aes(x=lag, y=estimate, 
-                              ymin=lower_bound, ymax=upper_bound)) +
-  geom_ribbon(alpha = 0.5) +
-  geom_line(size=2)
-
-p
-```
-
-### 4. Distributed lag -7 to 0 for time series
-
-(not shown)
-
-```{r DLM for casecrossover lag -7 to 0 temp, echo = FALSE, message=FALSE, warning=FALSE, eval = FALSE}
-
-inhaler_temp_lag <- inhaler_ndc_lag_15_df %>%
-  select(170:164, 93, 150:156) # temp
-
-lag_max <- 7
-
-inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
-inhaler_temp_lag <- as.matrix(inhaler_temp_lag)
-
-B <- ns(-7:lag_max, df =3, intercept=TRUE)
-
-# the nrow B should equal ncol(pm)
-dim(B)
-
-inhaler_temp_lag_B <- inhaler_temp_lag%*%B
-
-fit <- clogit(outcome ~  inhaler_temp_lag_B + strata(personkey), 
-              inhaler_ndc_lag_15_df)
-
-coef(fit)
-summary(fit)
-AIC(fit)
-
-# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
-dlparms <- c(1,2,3)
-DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms]) 
-DLvar <- B%*%vcov(fit)[dlparms,dlparms]%*%t(B)
-DLestimate$SE <- sqrt(diag(DLvar))
-DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
-DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96
-
-DLestimate$estimate <- exp(DLestimate$estimate)
-DLestimate$lower <- exp(DLestimate$lower)
-DLestimate$upper <- exp(DLestimate$upper)
-
-DLestimate
-
-DLestimate$lag <- -7:lag_max
-p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
-p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
-p <- p + geom_line(aes(x=lag), linetype=2, size=2)
-p <- p + theme_bw()
-p
-
-DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms])
-
-cumulative <- sum(DLestimate$estimate)
-cumulative
-
-cumulative_se <- sqrt(sum(DLvar))
-cumulative_se
-
-cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
-cumulative_CI
-
-```
-
-```{r DLM ts -7 to 0, echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}
-read_path11 <- paste0("../data_new/medication/or_ndc_county_time_series_lag15.csv")
-or_ndc_county_ts_lag15 <- read_csv(read_path11)
-
-or_ndc_county_ts_df <- or_ndc_county_ts_lag15 %>%
-  select(1:4, pop, rate_per_100k, day,  geo_smk_pm_lag_m1_county:geo_smk_pm_lag_m7_county, geo_smk_pm_county,
-         geo_smk_pm_lag1_county:geo_smk_pm_lag7_county, wrf_temp_lag_m1_county:wrf_temp_lag_m7_county, 
-         wrf_temp_county, wrf_temp_lag1_county:wrf_temp_lag7_county) %>%
-  mutate(geo_smk_pm10_county = geo_smk_pm_county/10,
-         geo_smk_pm10_lag1_county =geo_smk_pm_lag1_county/10,
-         geo_smk_pm10_lag2_county =geo_smk_pm_lag2_county/10,
-         geo_smk_pm10_lag3_county =geo_smk_pm_lag3_county/10,
-         geo_smk_pm10_lag4_county =geo_smk_pm_lag4_county/10,
-         geo_smk_pm10_lag5_county =geo_smk_pm_lag5_county/10,
-         geo_smk_pm10_lag6_county =geo_smk_pm_lag6_county/10,
-         geo_smk_pm10_lag7_county =geo_smk_pm_lag7_county/10,
-         # lag -7 to -1
-         geo_smk_pm10_lag_m1_county =geo_smk_pm_lag_m1_county/10,
-         geo_smk_pm10_lag_m2_county =geo_smk_pm_lag_m2_county/10,
-         geo_smk_pm10_lag_m3_county =geo_smk_pm_lag_m3_county/10,
-         geo_smk_pm10_lag_m4_county =geo_smk_pm_lag_m4_county/10,
-         geo_smk_pm10_lag_m5_county =geo_smk_pm_lag_m5_county/10,
-         geo_smk_pm10_lag_m6_county =geo_smk_pm_lag_m6_county/10,
-         geo_smk_pm10_lag_m7_county =geo_smk_pm_lag_m7_county/10)
-
-which(colnames(or_ndc_county_ts_df)=="wrf_temp_lag1_county") # 31
-which(colnames(or_ndc_county_ts_df)=="wrf_temp_county") # 30
-which(colnames(or_ndc_county_ts_df)=="wrf_temp_lag_m1_county") # 23
-which(colnames(or_ndc_county_ts_df)=="geo_smk_pm10_lag1_county") # 39
-which(colnames(or_ndc_county_ts_df)=="geo_smk_pm10_lag_m1_county") # 46
-which(colnames(or_ndc_county_ts_df)=="geo_smk_pm10_county") # 38
-
-geo_ts_lag <- or_ndc_county_ts_df %>%
-  select(52:46, 38, 39:45) # GWR-method
-tem_ts_lag <- or_ndc_county_ts_df %>%
-  select(29:23, 30, 31:37) # temp
-
-lag_max <- 7
-
-geo_ts_lag <- as.matrix(geo_ts_lag)
-tem_ts_lag <- as.matrix(tem_ts_lag)
-
-B <- ns(-7:lag_max, df =3, intercept=TRUE) # or 3?
-
-# the nrow B should equal ncol(pm)
-dim(B)
-dim(geo_ts_lag)
-dim(tem_ts_lag)
-
-geo_ts_lag_B <- geo_ts_lag%*%B
-
-tem_ts_lag_B <- tem_ts_lag%*%B
-
-## mixed model
-mod <-glmer(n_obs ~  geo_ts_lag_B + tem_ts_lag_B + day + (1|county) + offset(log(pop)),   
-            or_ndc_county_ts_df, family="poisson", control = glmerControl(optimizer = "bobyqa"))
-
-summary(mod)
-# check AIC
-AIC(mod)
-
-D <- cbind(B, B)
-# output distributed lag beta parameters
-dlparms <- mod@beta[2:7]
-# estimate distributed lag values for each day
-dl_estimates <- data.frame(estimate = D %*% dlparms)
-
-# covariance matrix for knots (need to convert to matrix object)
-cov_mat <- as.matrix(vcov(mod)[2:7,2:7])
-
-# estimate variance of splines
-dl_var <- D%*%cov_mat%*%t(D)
-
-# calculate standard error for each lag value
-dl_estimates$stderr <- sqrt(diag(dl_var))
-
-# calculate lower and upper bounds
-dl_estimates$lower_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(1-0.975,
-  df=df.residual(mod))  
-dl_estimates$upper_bound <- dl_estimates$estimate+dl_estimates$stderr*qt(0.975,
-  df=df.residual(mod))
-
-# estimate relative risk and 95%CI from beta
-inhaler_dl_rr <- dl_estimates %>% 
-  mutate(estimate = exp(estimate),
-         lower_bound = exp(lower_bound),
-         upper_bound = exp(upper_bound),
-         lag = -7:7)
-
-# estimate cumulative effect
-cumulative <- sum(dl_estimates$estimate)
-cumulative
-
-# estimate cumulative effect stnd error
-cumulative_se <- sqrt(sum(dl_var))
-cumulative_se
-
-# estimate cumulative CI
-cumulative_ci <- cumulative+cumulative_se*qt(c(1-0.975,0.975), 
-                                             df=df.residual(mod))
-cumulative_ci
-
-exp(cumulative)
-exp(cumulative_ci)
-
-# plot 
-p <- ggplot(inhaler_dl_rr, aes(x=lag, y=estimate, 
-                              ymin=lower_bound, ymax=upper_bound)) +
-  geom_ribbon(alpha = 0.5) +
-  geom_line(size=2)
-
-p
-
-```

### 2017-11-06 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
testing new git password so I added minor text edit.
Updated smoke map.

@@
 ##2. Wildfire smoke descriptive characteristics
-Map of smoke count and smoke time series will go here.
+
+Importing ZIP code-level population-weighted PM~2.5~ daily data.
+
+```{r smoke data import and wrangle}
+# pm path
+pm_path <- paste0("./data/pm/2013-oregon_county_pm25.csv")
+# county pm
+county_pm_df <- read_csv(pm_path)
+
+# estimate county smoke 
+county_smk_count <- county_pm_df %>% 
+  # binary smoke variable at >10 ug/m^3
+  mutate(gwr_smk10 = ifelse(geo_smk_pm >= 10, 1, 0)) %>% 
+  group_by(county) %>% 
+  summarise(smk_day10 = sum(gwr_smk10)) %>% 
+  # lower case county name and remove "." to match spatial df
+  mutate(subregion = tolower(stringr::str_replace(county, "[.]", " ")))
+
+# use the map function from the maps package to get the polygon data
+county_poly <- maps:::map("county", "oregon", plot=FALSE, fill = TRUE)
+# find centroids
+county_centroids <- maps:::apply.polygon(county_poly, maps:::centroid.polygon)
+
+# create a data frame for graphing out of the centroids of each polygon
+county_centroids <- county_centroids[!is.na(names(county_centroids))]
+centroid_array <- Rreduce(rbind, county_centroids)
+
+dimnames(centroid_array) <- list(gsub("[^,]*,", "", names(county_centroids)),
+                                 c("long", "lat"))
+county_text <- as.data.frame(centroid_array)
+county_text$county <- rownames(label_df)
+```
+
+### Number of days in Oregon where smoke PM~2.5~ exceeded 10 ug/m^3.
+
+```{r smoke days map}
+
+smoke_map <- ggplot(or_county_spatial_df, aes(x=long,y=lat, group=group)) +
+  # fill with number of smoke days
+  geom_polygon(aes(fill =smk_day10)) +
+  scale_fill_gradient(expression("Smoke Days > 10 µg/m"^3),
+    low= "#3B9AB2", high="#F21A00", 
+    guide = guide_colorbar(direction="horizontal", title.position = "top",
+                           title.hjust = 0.5, barwidth = 20)) +
+  # add county path on top
+  geom_path(colour="#EBCC2A") +
+  # add county text 
+  geom_text(data = county_text, aes(x=long, y=lat, label = county, group=NULL), 
+    colour = "white", size = 2.5) +
+  xlab("Longitude") +
+  ylab("Latitude") +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    strip.text = element_text(size = 13),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    strip.background = element_rect(colour=NA, fill=NA),
+    legend.position = "bottom")
+
+smoke_map
+```
+
+Most of the heavy smoke days were in the south-west part of Oregon.
+I've added county names to map but some text aren't well placed.
+I should probably use centroid, but then I'll have to import a shapefile to do this.
+I'll fix this later; I plan to add in fire locations.
+
+### Time-series of geo-weighted regression estimates of smoke PM~2.5~ by county
 

### 2017-11-07 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
general functions for distributed lag work; will need to tweak.

--- 2017-11-06 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
@@
 ```{r setup}
-library(tidyverse)
+library(tidyverse) # general data wrangle
+library(survival) # conditional logistic regression
 ```
@@
 ##2. Wildfire smoke descriptive characteristics
-Map of smoke count and smoke time series will go here.
+
+Importing ZIP code-level population-weighted PM~2.5~ daily data.
+
+```{r smoke data import and wrangle}
+# pm path
+pm_path <- paste0("./data/pm/2013-oregon_county_pm25.csv")
+# county pm
+county_pm_df <- read_csv(pm_path) %>% 
+  mutate(county = stringr::str_replace(county, "[.]", ""))
+
+# estimate county smoke 
+county_smk_count <- county_pm_df %>% 
+  # binary smoke variable at >10 ug/m^3
+  mutate(gwr_smk10 = ifelse(geo_smk_pm >= 10, 1, 0)) %>% 
+  group_by(county) %>% 
+  summarise(smk_day10 = sum(gwr_smk10)) %>% 
+  # lower case county name and remove "." to match spatial df
+  mutate(subregion = tolower(stringr::str_replace(county, "[.]", " ")))
+
+# extract oregon map data
+or_county_spatial_df <- map_data("county", "oregon") %>% 
+  right_join(county_smk_count, by = "subregion")
+
+# use the map function from the maps package to get the polygon data
+county_poly <- maps::map("county", "oregon", plot=FALSE, fill = TRUE)
+# find centroids
+county_centroids <- maps::apply.polygon(county_poly, maps:::centroid.polygon)
+
+# create a data frame for graphing out of the centroids of each polygon
+county_centroids <- county_centroids[!is.na(names(county_centroids))]
+centroid_array <- Reduce(rbind, county_centroids)
+
+county_text <- reduce(county_centroids, rbind) %>% 
+  as_data_frame() %>% 
+  rename(long = V1, lat = V2) %>% 
+  mutate(county = stringr::str_sub(
+    stringr::str_to_title(names(county_centroids)), start=8L)) 
+```
+
+### Number of days in Oregon where smoke PM~2.5~ exceeded 10 ug/m^3.
+
+```{r smoke days map}
+smoke_map <- ggplot(or_county_spatial_df, aes(x=long,y=lat, group=group)) +
+  # fill with number of smoke days
+  geom_polygon(aes(fill =smk_day10)) +
+  scale_fill_gradient(expression("Smoke Days > 10 µg/m"^3),
+    low= "#3B9AB2", high="#F21A00", 
+    guide = guide_colorbar(direction="horizontal", title.position = "top",
+                           title.hjust = 0.5, barwidth = 20)) +
+  # add county path on top
+  geom_path(colour="#EBCC2A") +
+  # add county text 
+  geom_text(data = county_text, aes(x=long, y=lat, label = county, group=NULL), 
+    colour = "white", size = 2) +
+  xlab("Longitude") +
+  ylab("Latitude") +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    strip.text = element_text(size = 8),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    strip.background = element_rect(colour=NA, fill=NA),
+    legend.position = "bottom")
+
+smoke_map
+```
+
+Most of the heavy smoke days were in the south-west part of Oregon.
+I've added county names to map but some text aren't well placed.
+I should probably use centroid, but then I'll have to import a shapefile to do this.
+I'll fix this later; I plan to add in fire locations.
+
+### Time-series of geo-weighted regression estimates of smoke PM~2.5~ by county
+
+Using the geofacet package to organize small-multiples by geography.
+
+*Note 11/7/17: Not sure how much I love the layout of this Oregon geofacet grid, but we'll see what people think.
+I like that geofacet lets you look at the two maps together to understand the exposure pattern, but too much whitespace.
+However, with this orientation, I believe you could cut out the smoke days figure.*
+
+```{r geofacet time series of pm}
+# use or_counties_grid from geo_facet 
+or_grid <- geofacet::or_counties_grid1
+# small multiples plot
+plot <- ggplot(county_pm_df, aes(x=date, y= geo_smk_pm)) + 
+  geom_point(color = "#00A08A", size = 0.5) +
+  scale_x_date(date_labels = "%m") +
+  geofacet::facet_geo(~county, grid = or_grid) +
+  ylab(expression("Smoke PM2.5 µg/m"^3)) +
+  xlab("Month in 2013") +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    strip.text = element_text(size = 8),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    strip.background = element_rect(colour=NA, fill=NA),
+    legend.position = "bottom")
+# plot 
+plot  
+```
+
+Elevated levels of smoke in late July to early August in Curry, Josephine, Jackson, and Klamath counties.
 
@@
 # name list elements in order I'd like to present in document
-outcome_names <- c("SABA", "Respiratory", "Asthma", "COPD", "Pneumonia", 
-  "Acute Bronchitis", "Cardiovascular Disease", "Arrhythmia", 
-  "Cerebrovascular Disease", "Heart Failure", "Ischemic Heart Disease", 
-  "Myocardial Infarction")
+# commented out for now.
+# outcome_names <- c("SABA", "Respiratory", "Asthma", "COPD", "Pneumonia", 
+#   "Acute Bronchitis", "Cardiovascular Disease", "Arrhythmia", 
+#   "Cerebrovascular Disease", "Heart Failure", "Ischemic Heart Disease", 
+#   "Myocardial Infarction")
+outcome_names <- c("Asthma", "CVD")
 
@@
 
+For this project, we use the time-stratified case-crossover study design in the same way we did for the 2012 Wenatchee-Complex fire project.
+Briefly, we identify admissions dates for ED or urgent care visits, and pharmacy fill dates for SABA, using billing records.
+
+In the Wenatchee-Complex fire project, we looked at the same-day relationship between smoke exposure and health outcomes.
+For the Douglas-complex fire, we will evaluate a 0 to 7 day distributed lag association, first comparing the overall cumulative effect over these 7 days, and then get in to the lagged effects.
+
+*Note: 11/7/17: Jingyang will provide a csv of the best-fit knots for the distributed lag splines based on AIC.*
+
+*Note: 11/7/17: Jingyang, can I use the lagged days in the case-crossover dataframes?
+Please confrim that the value for lag day 1 is the day before, lag day 2 is from 2 days before, etc..., and not the week before, two weeks before, etc...
+What does the _m denoter mean too?
+Jingyang says it's not necissary; we will remove creating this variable in the data management scripts.
+Do we use the patid or personkey identifier for strata?
+check.*
+
+Basis function also included for WRF temp. We will need to find the best fit of spline for temp seperately from PM. For asthma, it looks like 2 DF is best.
+
+Making a custom distributed lag funciton.
+
+```{r distributed lag function}
+# custom distributed lag function ----
+distributed_lag <- function(data, pm_df, temp_df, cumulative=T){
+  # limit to complete cases
+  complete_data <- data %>% 
+    # filter to complete case for smoke
+    filter(!is.na(geo_smk_pm_lag7_zip))
+
+  # output matrix of gwr values
+  pm_matrix <- complete_data %>% 
+    select(contains("geo_smk_pm")) %>%
+    # remove anything with m in it; will go back in original code and not create 
+    # this to begin with
+    select(-contains("_m")) %>% 
+    # divide exposure values by 10 units to interpret on 10 ug/m^3 scale
+    mutate_all(funs(./10)) %>% 
+    # convert to matrix
+    as.matrix()
+
+  # output temp matrix
+  temp_matrix <- complete_data %>% 
+    select(contains("wrf_temp")) %>%
+    # remove anything with m in it; will go back in original code and not create 
+    # this to begin with
+    select(-contains("_m")) %>% 
+    # convert to matrix
+    as.matrix()
+
+  # calculation of basis ----
+  # define basis using natural spline function from "splines"" package
+  pm_b <- splines::ns(0:7, df=pm_df, intercept=T)
+  # create pm basis
+  pm_basis <- pm_matrix %*% pm_b
+  # create temp basis
+  temp_b <- splines::ns(0:7, df=temp_df, intercept=T)
+  temp_basis <- temp_matrix %*% temp_b
+
+  # fit model with basis ----
+  mod <- clogit(outcome ~ pm_basis + temp_basis + strata(personkey),
+    data = complete_data)
+
+  # calculate estimates ----
+  # output pm basis parameters
+  dl_parms <- broom::tidy(mod) %>% 
+    filter(stringr::str_detect(term, "pm")) %>% 
+    select(estimate) %>% 
+    as_vector()
+  # estimate distributed lag values for each day
+  dl_estimates <- data.frame(estimate = pm_b %*% dl_parms)
+  # covariance matrix for knots
+  cov_matrix <- as.matrix(vcov(mod))[1:3,1:3]
+
+  # estimate variance of spline
+  dl_var <- pm_b %*% cov_matrix %*% t(pm_b)
+  # estimate standard error
+  dl_estimates$stderr <- sqrt(diag(dl_var))
+
+  # calculate lower and upper bound
+  dl_estimates$lower95 <- dl_estimates$estimate-(dl_estimates$stderr*1.96)
+  dl_estimates$upper95 <- dl_estimates$estimate+(dl_estimates$stderr*1.96)
+
+    if(cumulative==T) {
+      type <- "cumulative"  
+      # cumulative outcome and 95CI
+      estimate <- sum(dl_estimates$estimate)
+      # stderr cumulative effect
+      estimate_se <- sqrt(sum(dl_var))
+      # cumulative 95CI
+      lower95 <- estimate-(estimate_se*1.96)
+      upper95 <- estimate+(estimate_se*1.96)
+      # return dataframe
+      return_estimate <- data_frame(type, estimate, lower95, upper95) %>% 
+        mutate_if(is.numeric, exp)
+      return(return_estimate)
+    }
+    else { # if cumulative is not true, or false, return distributed lag est
+      # return dataframe
+      return_estimate <- dl_estimates %>% 
+        mutate_if(is.numeric, exp) %>% 
+        mutate(type = "lag", 
+               time = as.numeric(rownames(.))-1) %>% 
+        select(type, time, estimate, lower95, upper95)
+       # return estimate  
+      return(return_estimate)
+      }
+      # end of if
+} # end of function
+
+# test function
+
+test <- outcome_df_list %>% 
+  map_dfr(~distributed_lag(., pm_df=3, temp_df=2, cumulative=T)) %>% 
+  bind_cols(test, data_frame(outcome_names)) %>%
+
+test2 <- outcome_df_list %>%
+
+map_dfr
+
+distributed_lag(asthma_df, pm_df=3, temp_df=2, cumulative=F)
+
+test2
+```
+
 ##5. Distributed lag results
@@
 ##6. Stratified cumulative results
+
+Stratify cumulative effect by sex and age group.
 

### 2017-11-13 r_scripts/analysis/oregon_cvd_pulm_results.Rmd

@@
 # find centroids
-county_centroids <- maps::apply.polygon(county_poly, maps:::centroid.polygon)
+county_centroids <- maps:::apply.polygon(county_poly, maps:::centroid.polygon)
 
@@
 # i can't find a lot of the files
-#resp_path <- paste0("./"
-
 # asthma
-asthma_path <- paste0("../../data/health/2017-oregon_casecross_asthma.csv")
+asthma_path <- paste0("./data/health/2017-oregon_casecross_asthma.csv")
 asthma_df <- read_csv(asthma_path)
@@
 # cvd 
-cvd_path <- paste0("../../data/health/2017-oregon_casecross_cvd.csv")
+cvd_path <- paste0("./data/health/2017-oregon_casecross_cvd.csv")
 cvd_df <- read_csv(cvd_path)
@@
 
-Making a custom distributed lag funciton.
+Making a custom distributed lag function. Consider adding additional options like specifying number of lag days. That would require a reworking of how the case-crossovers are set up (kind of).
 
 ```{r distributed lag function}
-# custom distributed lag function ----
+# custom distributed lag function 
 distributed_lag <- function(data, pm_df, temp_df, cumulative=T){
@@
         mutate(type = "lag", 
-               time = as.numeric(rownames(.))-1) %>% 
+               time = (as.numeric(rownames(.))-1)/-1) %>% 
         select(type, time, estimate, lower95, upper95)
@@
       # end of if
-} # end of function
+} 
+```
 
-# test function
+Using the custom function to estimate cumulative effects for the list of dataframes. It will work on an individual dataframe, but I built it to use with the purrr map functions.
 
-test <- outcome_df_list %>% 
+*Note for Jingyang 11/8/17: purrr map functions are the tidyverse equivalents of apply-type functions*
+
+```{r cumulative association}
+# use custom function to create a dataframe of cumulative values
+# using purrr::map_dfr function
+# purrr::map_dfr will work if it's just the cumulative, but won't work for lagged
+cumulative_results <- outcome_df_list %>% 
   map_dfr(~distributed_lag(., pm_df=3, temp_df=2, cumulative=T)) %>% 
-  bind_cols(test, data_frame(outcome_names)) %>%
+  bind_cols(., data_frame(outcome_names)) %>% 
+  select(outcome_names, type, estimate, lower95, upper95)
 
-test2 <- outcome_df_list %>%
+# will put this table eventually
+head(cumulative_results)
 
-map_dfr
+# ggplot ----
+plot <- ggplot(cumulative_results, aes(x=outcome_names, y = estimate)) +
+  geom_point(colour = "darkblue") +
+  geom_text(aes(label = round(estimate,2)), hjust = -0.25) +
+  geom_errorbar(aes(ymin=lower95, ymax=upper95), colour= "darkblue", width = 0.2) +
+  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
+  theme(
+   panel.background = element_rect(fill = 'white', colour = 'black'),
+   panel.grid.major = element_blank(),
+   panel.grid.minor = element_blank()) +
+  ylab(expression("Cumulative odds ratio for a 10 ug/m^3 increase smoke PM"[2.5])) +
+  xlab("Outcomes")
 
-distributed_lag(asthma_df, pm_df=3, temp_df=2, cumulative=F)
+plot
+```
 
-test2
-```
+This is generally how I'd like the plots to look, but I plan to add color grouping to SABA, respiratory, and CVD groupings.
+Also considering reducing the lag time evaluated since I think we are getting in to variance issues a week out with this dataset (as evidence, large error around asthma outcome relative to other studies).
+
+I will also come back and tweak color scheme and labels.
 
@@
 
+That same function can be used to generate detailed estimates of the association over time. In this case, we'll look at the same day to seven days prior the event.
+
+```{r distributed lag}
+# this approach will work that creates a list of dataframes with map then
+# binds the lists together in one dataframe
+dl_results <- outcome_df_list %>% 
+  purrr::map(~distributed_lag(., pm_df=3, temp_df=2, cumulative=F)) %>% 
+  plyr::rbind.fill() %>% 
+  bind_cols(., data_frame(rep(outcome_names, each=8))) %>% 
+  rename(outcome_names = `rep(outcome_names, each = 8)`) %>% 
+  select(outcome_names, type, time, estimate, lower95, upper95)
+
+# results will come in a format to make small multiples I think
+head(dl_results)
+
+# small multiples DL plot ----
+dl_plot <- ggplot(dl_results, aes(x=time, y=estimate)) +
+  geom_line(colour = "#046C9A", size = 1) +
+  geom_ribbon(aes(ymin = lower95, ymax = upper95), fill = "#ABDDDE", alpha = 0.5) + 
+  scale_y_continuous(limits = c(0.9, 1.15)) +
+  scale_x_continuous(breaks = c(seq(-7, 0, by=1))) +
+  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
+  facet_wrap(~outcome_names, scales = "free_x") +
+  ylab(expression("Odd ratio for a 10 ug/m^3 increase smoke PM"[2.5])) +
+  xlab("Lagged Days") +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    strip.text = element_text(size = 10),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    strip.background = element_rect(colour=NA, fill=NA))
+
+dl_plot
+```
+
+DL plot shows an effect with asthma even though the cumulative effect size is similar in magnitude to what we've seen before, but the 95%CIs overlap 1.
+I think this is a power issue trying to look at lagged effects for more than a week.
+
+Worth discussing with the group.
+
 ##6. Stratified cumulative results
 
+*Jingyang to complete using function*
 Stratify cumulative effect by sex and age group.
@@
 ##7. SABA cumulative effect by underlying respiratory diagnosis
+*Jingyang to complete using function*
 
@@
 2. Remove r code files that are not useful. There are too many files with similar code inside.
-3. Variables in the casecross over datasets raise some questions. For example, what is the difference from the varialbes geo_smk_pm and
+3. Variables in the casecross over datasets raise some questions.

### 2017-11-13 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
Update oregon_cvd_pulm_results.Rmd

@@
 # SABA case-crossover will be imported first
-# i can't find a lot of the files
+
+# SABA
+saba_path <- paste0("./data/health/2013-oregon_casecross_saba.csv")
+saba_df <- read_csv(saba_path)
+
+# respiratory
+respiratory_path <- paste0("./data/health/2013-oregon_casecross_respiratory.csv")
+respiratory_df <- read_csv(respiratory_path)
+
 # asthma
-asthma_path <- paste0("./data/health/2017-oregon_casecross_asthma.csv")
+asthma_path <- paste0("./data/health/2013-oregon_casecross_asthma.csv")
 asthma_df <- read_csv(asthma_path)
 
-# cvd 
-cvd_path <- paste0("./data/health/2017-oregon_casecross_cvd.csv")
+# COPD
+copd_path <- paste0("./data/health/2013-oregon_casecross_copd.csv")
+copd_df <- read_csv(copd_path)
+
+# pneumonia
+pneumonia_path <- paste0("./data/health/2013-oregon_casecross_pneumonia.csv")
+pneumonia_df <- read_csv(pneumonia_path)
+
+# acute bronchitis
+acute_bronch_path <- paste0("./data/health/2013-oregon_casecross_acute_bronch.csv")
+acute_bronch_df <- read_csv(arrhythmia_path)
+
+# Cardiovascular Disease
+cvd_path <- paste0("./data/health/2013-oregon_casecross_cvd.csv")
 cvd_df <- read_csv(cvd_path)
 # had a warning message for cvd dataframe import; should check on this
+
+# arrhythmia
+arrhythmia_path <- paste0("./data/health/2013-oregon_casecross_arrhythmia.csv")
+arrhythmia_df <- read_csv(arrhythmia_path)
+
+# Cerebrovascular Disease
+cerbrovas_dis_path <- paste0("./data/health/2013-oregon_casecross_cerbrovas_dis.csv")
+cerbrovas_dis_df <- read_csv(cerbrovas_dis_path)
+
+# Heart Failure
+heart_failure_path <- paste0("./data/health/2013-oregon_casecross_heart_failure.csv")
+heart_failure_df <- read_csv(heart_failure_path)
+
+# Ischemic Heart Disease
+isch_heart_dis_path <- paste0("./data/health/2013-oregon_casecross_isch_heart_dis.csv")
+isch_heart_dis_df <- read_csv(isch_heart_dis_path)
+
+# Myocardial Infarction
+myocardial_infarc_path <- paste0("./data/health/2013-oregon_casecross_myocardial_infarc.csv")
+myocardial_infarc_df <- read_csv(myocardial_infarc_path)
 
@@
 # bind dataframes in a list
-outcome_df_list <- list(asthma_df, cvd_df)
+outcome_df_list <- list(saba_df, respiratory_df, asthma_df, copd_df, pneumonia_df, 
+                        acute_bronch_df, cvd_df, arrhythmia_df, cerbrovas_dis_df, 
+                        heart_failure_df, isch_heart_dis_df, myocardial_infarc_df)
 
@@
 # commented out for now.
-# outcome_names <- c("SABA", "Respiratory", "Asthma", "COPD", "Pneumonia", 
-#   "Acute Bronchitis", "Cardiovascular Disease", "Arrhythmia", 
-#   "Cerebrovascular Disease", "Heart Failure", "Ischemic Heart Disease", 
-#   "Myocardial Infarction")
-outcome_names <- c("Asthma", "CVD")
+outcome_names <- c("SABA", "Respiratory", "Asthma", "COPD", "Pneumonia", 
+                   "Acute Bronchitis", "Cardiovascular Disease", "Arrhythmia", 
+                   "Cerebrovascular Disease", "Heart Failure", "Ischemic Heart Disease", 
+                   "Myocardial Infarction")
 
-names(outcome_df_list) <- c( 'Asthma', 'CVD')
+names(outcome_df_list) <- c('SABA', 'Respiratory', 'Asthma', 'COPD', 'Pneumonia', 
+                            'Acute Bronchitis', 'Cardiovascular Disease', 'Arrhythmia', 
+                            'Cerebrovascular Disease', 'Heart Failure', 
+                            'Ischemic Heart Disease', 'Myocardial Infarction')
+
+# try athma
+outcome_df_list <- list(asthma_df)
+
+outcome_names <- c("Asthma")
+
+names(outcome_df_list) <- c('Asthma')
+
 ```

### 2017-11-14 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
fixed table 1 function to handle 0 events in under 15 subgroup for certain cvd outcomes.

--- 2017-11-13 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
@@
 
+*Note 2017-11-14: It may be easiest to make a set of casecross over functions and put them in a package rather than import the timestratified casecross overs each time.*
+
 ```{r health data import, message=F, warning=F}
-# importing case-crossover dataframes 
-# SABA case-crossover will be imported first
-# i can't find a lot of the files
-# asthma
-asthma_path <- paste0("./data/health/2017-oregon_casecross_asthma.csv")
-asthma_df <- read_csv(asthma_path)
+# path for health dataframe
+health_path <- "./data/health/"
+# importing time-stratified case-crossover dataframe
+casecross_files <- list.files(path=health_path, pattern="*casecross*")
 
-# cvd 
-cvd_path <- paste0("./data/health/2017-oregon_casecross_cvd.csv")
-cvd_df <- read_csv(cvd_path)
-# had a warning message for cvd dataframe import; should check on this
+# reorder file names in way I'd like it to be analyzed
+# also ditching broken arm
+casecross_order <- casecross_files[c(13,12,3,6,11,1,7,2,5,8,9,10)]
 
-# import all other dataframes
+# extract the name of the dataframe after study type
+file_names <- stringr::str_extract(casecross_order, 
+  pattern="(?<=casecross_)\\w+")
 
-# I think purrr would be a good way to loop through a list of outcome dataframes
-# bind dataframes in a list
-outcome_df_list <- list(asthma_df, cvd_df)
+# read in all the files in to a list
+outcome_df_list <- casecross_order %>%  
+  purrr::map(~read_csv(paste0(health_path, .)))
+
+# name dataframes
+names(outcome_df_list) <- file_names
 
 # name list elements in order I'd like to present in document
-# commented out for now.
-# outcome_names <- c("SABA", "Respiratory", "Asthma", "COPD", "Pneumonia", 
-#   "Acute Bronchitis", "Cardiovascular Disease", "Arrhythmia", 
-#   "Cerebrovascular Disease", "Heart Failure", "Ischemic Heart Disease", 
-#   "Myocardial Infarction")
-outcome_names <- c("Asthma", "CVD")
-
-names(outcome_df_list) <- c( 'Asthma', 'CVD')
+outcome_names <- c("SABA", "Respiratory", "Asthma", "COPD", "Pneumonia", 
+  "Acute Bronchitis", "Cardiovascular Disease", "Arrhythmia", 
+  "Cerebrovascular Disease", "Heart Failure", "Ischemic Heart Disease", 
+  "Myocardial Infarction")
 ```
@@
 *Warning message:In rbind(names(probs), probs_f) :number of columns of result is not a multiple of vector length (arg 1)*
+
+*Also warnings on mixed data types of NULL and numeric; another reason to revise and create timestratified casecross over functions*.
 
@@
     summarise(n = n()) %>%
+    # this will fix the 0 categories of certain CVD outcomes in children
+    add_row(age_ind = 0, n = 0) %>% 
+    arrange(age_ind) %>% 
+    group_by(age_ind) %>% 
+    summarise(n = sum(n)) %>% 
     mutate(age_cat = ifelse(age_ind == 0, "age_under15",
@@
 # outcome names
-outcome_names <- names(outcome_df_list) %>% 
-  data_frame() %>% 
-  rename(outcome = ".")
+outcome <- outcome_names %>% 
+  as_tibble() %>%
+  rename(outcome = value)
 
@@
 # descriptive table, bind outcome names and values
-descriptive_table <- bind_cols(outcome_names, table_vals)
+descriptive_table <- bind_cols(outcome, table_vals)
 
@@
   bind_cols(., data_frame(outcome_names)) %>% 
-  select(outcome_names, type, estimate, lower95, upper95)
+  select(outcome_names, type, estimate, lower95, upper95) %>% 
+  # preserve order of names
+  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))
 
@@
   geom_text(aes(label = round(estimate,2)), hjust = -0.25) +
-  geom_errorbar(aes(ymin=lower95, ymax=upper95), colour= "darkblue", width = 0.2) +
+  geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                colour= "darkblue", width = 0.2) +
   geom_hline(yintercept = 1, linetype = 2, colour = "red") +
@@
    panel.grid.major = element_blank(),
-   panel.grid.minor = element_blank()) +
-  ylab(expression("Cumulative odds ratio for a 10 ug/m^3 increase smoke PM"[2.5])) +
+   panel.grid.minor = element_blank(),
+   axis.text.x = element_text(angle=90, hjust = 1)) +
+  ylab(expression(paste0("Cumulative odds ratio for a 10 ug/m"^3, 
+    " increase smoke PM"[2.5]))) +
   xlab("Outcomes")
@@
   geom_line(colour = "#046C9A", size = 1) +
-  geom_ribbon(aes(ymin = lower95, ymax = upper95), fill = "#ABDDDE", alpha = 0.5) + 
+  geom_ribbon(aes(ymin = lower95, ymax = upper95), 
+              fill = "#ABDDDE", alpha = 0.5) + 
   scale_y_continuous(limits = c(0.9, 1.15)) +

### 2017-11-14 r_scripts/analysis/model_dl_spline_knots.Rmd

@@
 # fill outcome name in first column
-fit_mat[,1] <- "Beta - 2 agonists"
+fit_mat[,1] <- "SABA"
 # fill in exposure (geo smoke)
@@
 
+fit_mat_saba <- as_data_frame(fit_mat)
+
 # now that we have a range of df/knots, I want to filter to the minimum aic
@@
 
-kable(fit_mat, caption="Distributed lag spline of 3-6 degree of freedom fit by AIC")
+# kable(fit_mat, caption="Distributed lag spline of 3-6 degree of freedom fit by AIC")
 
-lag_spline_best_fit <- as_tibble(fit_mat) %>%  
-  slice(which.min(aic)) # df of 3 has the smallest AIC
+# lag_spline_best_fit <- as_tibble(fit_mat) %>%  
+#   slice(which.min(aic)) # df of 3 has the smallest AIC
 
 # kable
-knitr::kable(lag_spline_best_fit, caption = paste0("Distributed lag spline ",
-  "degree of freedom best fit by AIC"))
+# knitr::kable(lag_spline_best_fit, caption = paste0("Distributed lag spline ",
+#   "degree of freedom best fit by AIC"))
 ```
+
+```{r diseases, message=FALSE, echo=FALSE, warning=FALSE}
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+
+datalist <- list()
+
+  # j <- c('respiratory')
+
+  # standard for statement without parallel computing
+  for(j in var_list){
+  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
+  or_disease <- read_csv(read_path)
+
+  or_disease_df <- data.frame(or_disease)
+
+  or_disease_lag_df <- or_disease_df %>%
+  left_join(or_zip_to_county, by = "ZIPCODE") %>%
+  left_join(or_pop, by = "county") %>%
+  mutate(day = ifelse(weekdays(date) == "Monday" |
+                      weekdays(date) == "Tuesday" |
+                      weekdays(date) == "Wednesday"|
+                      weekdays(date) == "Thursday"|
+                      weekdays(date) == "Friday", "weekday",
+               ifelse(weekdays(date) == "Saturday" |
+                      weekdays(date) == "Sunday", "weekend", NA))) %>%
+  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
+         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
+         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
+         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
+         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
+         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
+         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
+         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)
+
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-05-27")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-07-04")] <- "weekend"
+or_disease_lag_df$day[which(or_disease_lag_df$date=="2013-09-02")] <- "weekend"
+
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
+# which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
+# which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
+# which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151
+
+or_disease_geo_lag <- or_disease_lag_df %>%
+  select(174, 175:181) # GWR-method
+
+or_disease_temp_lag <- or_disease_lag_df %>%
+  select(94, 151:157) # GWR-method
+
+lag_max <- 7
+
+or_disease_geo_lag <- as.matrix(or_disease_geo_lag)
+or_disease_temp_lag <- as.matrix(or_disease_temp_lag)
+
+# create fit matrix
+fit_mat <- matrix(NA, nrow=4, ncol=5)
+colnames(fit_mat) <- c("outcome", "exposure", "fit", "df", "aic")
+
+# fill outcome name in first column
+fit_mat[,1] <- j
+# fill in exposure (geo smoke)
+fit_mat[,2] <- "geo_smk10"
+# fit type
+fit_mat[,3] <- "ns"
+
+# 2 df will not work for a spline, going from 3 to 6 df
+  for(h in 3:6){ # start loop for df
+
+    # fitting a natural spline for lag with 5 knots?
+    fit_mat[h-2, 4] <- degree_freedom <- h
+
+    # define basis b using natural spline function
+    B <- ns(0:7, df = degree_freedom, intercept = T)
+
+    # multiply lagged pm matrix by basis
+    or_disease_geo_lag_B <- or_disease_geo_lag%*%B
+    or_disease_temp_lag_B <- or_disease_temp_lag%*%B
+
+    fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_temp_lag_B + strata(personkey), 
+              or_disease_lag_df)
+
+    AIC(fit)
+
+    # fill AIC
+    fit_mat[h-2,5] <- round(AIC(fit),5)
+
+    } # end model fit loop
+i <- match(j, var_list)
+datalist[[i]] <- as_data_frame(fit_mat)
+  }
+
+combine <- bind_rows(datalist)
+
+combine_saba <- bind_rows(fit_mat_saba, combine)
+
+# now that we have a range of df/knots, I want to filter to the minimum aic
+# for each outcome and print out that value in a table
+
+kable(combine_saba, caption="Distributed lag spline of 3-6 degree of freedom fit by AIC")
+
+write_path <- paste0('../data_new/medication/model_dl_spline_knots.csv')
+write_csv(combine_saba, write_path)
+
+```

### 2017-11-14 r_scripts/analysis/oregon_cvd_pulm_results.Rmd

@@
 
-##4. Distributed lag cumulative effect results
+##4. Same-day association results
+
+*Jingyang, add same day associations here using purrr map functions here.
+I want to compare them to the cumulative associations.
+You do not need to make an elaborate function like I did for the distributed lag.
+You should just be able to use clogit.
+Include the same day GWR smoke estimate and same day temperature as a covariate.*
+
+##5. Distributed lag cumulative effect results
 
@@
 
-# will put this table eventually
-head(cumulative_results)
+# will put this table eventually and split out plot
+#head(cumulative_results)
 
@@
    axis.text.x = element_text(angle=90, hjust = 1)) +
-  ylab(expression(paste0("Cumulative odds ratio for a 10 ug/m"^3, 
-    " increase smoke PM"[2.5]))) +
+  ylab(expression("Cumulative odds ratio: 10 ug/m^3 increase smoke PM"[2.5])) +
   xlab("Outcomes")
@@
 
-##5. Distributed lag results
+##6. Distributed lag results
 
@@
   rename(outcome_names = `rep(outcome_names, each = 8)`) %>% 
-  select(outcome_names, type, time, estimate, lower95, upper95)
+  select(outcome_names, type, time, estimate, lower95, upper95) %>% 
+  # preserve order of names
+  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))
 
 # results will come in a format to make small multiples I think
-head(dl_results)
+#head(dl_results)
 
@@
               fill = "#ABDDDE", alpha = 0.5) + 
-  scale_y_continuous(limits = c(0.9, 1.15)) +
+  scale_y_continuous(limits = c(0.8, 1.2)) +
   scale_x_continuous(breaks = c(seq(-7, 0, by=1))) +
   geom_hline(yintercept = 1, linetype = 2, colour = "red") +
-  facet_wrap(~outcome_names, scales = "free_x") +
+  facet_wrap(~outcome_names) +
   ylab(expression("Odd ratio for a 10 ug/m^3 increase smoke PM"[2.5])) +
@@
 
-##6. Stratified cumulative results
+##7. Stratified cumulative results
 
@@
 
-##7. SABA cumulative effect by underlying respiratory diagnosis
+##8. SABA cumulative effect by underlying respiratory diagnosis
 *Jingyang to complete using function*

### 2017-11-15 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
Same day association, function, table and plot.

@@
 
+*Note from Jingyang on Nov 14:*
+*Just a little improvement advice for plot: Because the number (which labels the last disease's odds ratio) can only be seen one digital, I suggest to add more margin to the right side.
+Also for the DLM graph.*
+
+```{r same day association fuction}
+same_day_fun <- function(data, pm_df, temp_df){
+
+  # limit to complete cases
+  complete_data <- data %>% 
+    # filter to complete case for smoke
+    filter(!is.na(geo_smk_pm_zip))
+
+  # output matrix of gwr values
+  pm_matrix <- complete_data %>% 
+    select(geo_smk_pm_zip) %>%
+    # divide exposure values by 10 units to interpret on 10 ug/m^3 scale
+    mutate_all(funs(./10)) %>% 
+    # convert to matrix
+    as.matrix()
+
+  # output temp matrix
+  temp_matrix <- complete_data %>% 
+    select(wrf_temp_zip) %>%
+    # convert to matrix
+    as.matrix()
+
+  # fit model with basis ----
+  mod <- clogit(outcome ~ pm_matrix + temp_matrix + strata(personkey),
+    data = complete_data)
+
+  # odds ratio
+  sd_estimates <- round(exp(summary(mod)$coefficient[1,1]), 3) # sd stands for same day
+
+  # 95% lower bound
+  sd_lower <- round(exp((summary(mod)$coefficient[1,1]) -
+                                    1.96*(summary(mod)$coefficient[1,3])), 3)
+  # 95% upper bound
+  sd_upper <- round(exp((summary(mod)$coefficient[1,1]) +
+                                    1.96*(summary(mod)$coefficient[1,3])), 3)
+  # standard error
+  sd_stderr <- round(summary(mod)$coefficient[1,3], 4)
+  # p val
+  sd_pvalue <- round(summary(mod)$coefficient[1,5], 4)
+
+  return_estimate <- data_frame(sd_estimates, sd_lower, sd_upper) 
+  return(return_estimate) 
+} 
+```
+
+```{r same day association table}
+sameday_cumulative_results <- outcome_df_list %>% 
+  map_dfr(~same_day_fun(., pm_df=3, temp_df=2)) %>% 
+  bind_cols(., data_frame(outcome_names)) %>% 
+  select(outcome_names, sd_estimates, sd_lower, sd_upper) %>% 
+  # preserve order of names
+  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))
+
+knitr::kable(sameday_cumulative_results, 
+  caption = "Same day odds ratio and CI of pharmacy/ED/urgent care visits")
+```
+
+```{r same day association graph}
+# ggplot ----
+plot <- ggplot(sameday_cumulative_results, aes(x=outcome_names, y = sd_estimates)) +
+  geom_point(colour = "darkblue") +
+  geom_text(aes(label = round(sd_estimates,2)), hjust = -0.25) +
+  geom_errorbar(aes(ymin=sd_lower, ymax=sd_upper), 
+                colour= "darkblue", width = 0.2) +
+  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
+  theme(
+   panel.background = element_rect(fill = 'white', colour = 'black'),
+   panel.grid.major = element_blank(),
+   panel.grid.minor = element_blank(),
+   axis.text.x = element_text(angle=90, hjust = 1)) +
+  ylab(expression("Same day odds ratio: 10 ug/m^3 increase smoke PM"[2.5])) +
+  xlab("Outcomes")
+
+plot
+```
+
 ##5. Distributed lag cumulative effect results

### 2017-11-15 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
Updated some of the script and modified Jingyang's code slightly to match some of my naming conventions.

--- 2017-11-14 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
@@
 
+*Note from Jingyang on Nov 14:*
+*Just a little improvement advice for plot: Because the number (which labels the last disease's odds ratio) can only be seen one digital, I suggest to add more margin to the right side.
+Also for the DLM graph.*
+
+Create function to calculate same-day association (odds ratio) and 95% confidence interval between increasing smoke and outcomes.
+
+```{r same day association fuction}
+same_day_fun <- function(data){
+
+  # limit to complete cases
+  complete_data <- data %>% 
+    # filter to complete case for smoke
+    filter(!is.na(geo_smk_pm_zip)) %>% 
+    # create 10 unit smk variable 
+    mutate(geo_smk10 = geo_smk_pm_zip/10)
+
+  # fit model with basis ----
+  mod <- clogit(outcome ~ geo_smk10 + wrf_temp_zip + strata(personkey),
+    data = complete_data)
+
+  # odds ratio
+  estimates <- round(exp(summary(mod)$coefficient[1,1]), 3) # sd stands for same day
+
+  # 95% lower bound
+  lower95 <- round(exp((summary(mod)$coefficient[1,1]) -
+                                    1.96*(summary(mod)$coefficient[1,3])), 3)
+  # 95% upper bound
+  upper95 <- round(exp((summary(mod)$coefficient[1,1]) +
+                                    1.96*(summary(mod)$coefficient[1,3])), 3)
+  # standard error
+  sd_stderr <- round(summary(mod)$coefficient[1,3], 4)
+  # p val
+  sd_pvalue <- round(summary(mod)$coefficient[1,5], 4)
+
+  return_estimate <- data_frame(estimates, lower95, upper95) 
+  return(return_estimate) 
+} 
+```
+
+```{r same day association table}
+sameday_results <- outcome_df_list %>% 
+  map_dfr(~same_day_fun(.)) %>% 
+  bind_cols(., data_frame(outcome_names)) %>% 
+  select(outcome_names, estimates, lower95, upper95) %>% 
+  # preserve order of names
+  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))
+
+knitr::kable(sameday_results, 
+  caption = "Same day odds ratio and 95% CI of pharmacy/ED/urgent care visits")
+```
+
+Applying function to the list of outcome dataframes.
+
+```{r same day association graph}
+# ggplot ----
+plot <- ggplot(sameday_results, aes(x=outcome_names, y = estimates)) +
+  geom_point(colour = "#46ACC8") +
+  geom_text(aes(label = round(estimates,2)), hjust = -0.25) +
+  geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                colour= "#46ACC8", width = 0.2) +
+  geom_hline(yintercept = 1, linetype = 2, colour = "#B40F20") +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+   panel.grid.major = element_blank(),
+   panel.grid.minor = element_blank(),
+   axis.text.x = element_text(angle=90, hjust = 1)) +
+  ylab(expression("Same day odds ratio: 10 ug/m^3 increase smoke PM"[2.5])) +
+  xlab("Outcomes")
+
+plot
+```
+
 ##5. Distributed lag cumulative effect results
@@
 plot <- ggplot(cumulative_results, aes(x=outcome_names, y = estimate)) +
-  geom_point(colour = "darkblue") +
+  geom_point(colour = "#00A08A") +
   geom_text(aes(label = round(estimate,2)), hjust = -0.25) +
   geom_errorbar(aes(ymin=lower95, ymax=upper95), 
-                colour= "darkblue", width = 0.2) +
-  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
-  theme(
-   panel.background = element_rect(fill = 'white', colour = 'black'),
+                colour= "#00A08A", width = 0.2) +
+  geom_hline(yintercept = 1, linetype = 2, colour = "#FF0000") +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
@@
 
-DL plot shows an effect with asthma even though the cumulative effect size is similar in magnitude to what we've seen before, but the 95%CIs overlap 1.
+DL plot shows an effect with asthma even though the cumulative effect size is similar in magnitude to what we've seen before, but the 95% CIs overlap 1.
 I think this is a power issue trying to look at lagged effects for more than a week.
+I've applied a general 3 degrees of freedom knot for PM~2.5~ and 2 degrees of freedom knot for temperature.
+I think temperature is okay to do this, but I'm still figuring out a way to build a best-fit knot in to the function.
 
@@
 
+##9. Conculstions
+So far, it looks like SABA and Asthma are the only associtions.
+Furthermore, the cumulative 0-7 day relationship between asthma and smoke has wide error bars and overlaps 1, suggesting the association could be null.
+I think it's more likely that this is because we introduce some noise on lag days 4-7.
+It's possible that an adjustment like weekend may help.
+It may also make sense to do something similar to Anna Rappold or others and shorten our lagged response.
+I also think this is a produce of sample size too.
+
 ### Comments:

### 2017-11-15 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
stuff

@@
 table1_fun <- function(data_list){
+  data_list <- data_list %>%
+    filter(!is.na(geo_smk_pm_zip))
+
   # output of total n
@@
 distributed_lag <- function(data, pm_df, temp_df, cumulative=T){
+
   # limit to complete cases
@@
 
+```{r same day stratified sex function}
+same_day_sex_fun <- function(data_list){
+  data_list <- data_list %>%
+    filter(!is.na(geo_smk_pm_zip)) %>%
+    mutate(geo_smk10 = geo_smk_pm_zip/10)
+
+  # Male
+  data_M <- data_list %>%
+    filter(sex_ind == 0)
+
+  # fit model with basis ----
+  mod_M <- clogit(outcome ~ geo_smk10 + wrf_temp_zip + strata(personkey),
+                  data = data_M)
+
+  # odds ratio
+  estimates_M <- round(exp(summary(mod_M)$coefficient[1,1]), 3) # sd stands for same day
+
+  # 95% lower bound
+  lower95_M <- round(exp((summary(mod_M)$coefficient[1,1]) -
+                                    1.96*(summary(mod_M)$coefficient[1,3])), 3)
+  # 95% upper bound
+  upper95_M <- round(exp((summary(mod_M)$coefficient[1,1]) +
+                                    1.96*(summary(mod_M)$coefficient[1,3])), 3)
+  # standard error
+  sd_stderr_M <- round(summary(mod_M)$coefficient[1,3], 4)
+  # p val
+  sd_pvalue_M <- round(summary(mod_M)$coefficient[1,5], 4)
+
+  total_n_M <- data_M %>%
+    filter(outcome == 1) %>%
+    summarise(total_n = n())
+
+  # Female
+  data_F <- data_list %>%
+    filter(sex_ind == 1)
+
+  # fit model with basis ----
+  mod_F <- clogit(outcome ~ geo_smk10 + wrf_temp_zip + strata(personkey),
+                  data = data_F)
+
+  # odds ratio
+  estimates_F <- round(exp(summary(mod_F)$coefficient[1,1]), 3) # sd stands for same day
+
+  # 95% lower bound
+  lower95_F <- round(exp((summary(mod_F)$coefficient[1,1]) -
+                                    1.96*(summary(mod_F)$coefficient[1,3])), 3)
+  # 95% upper bound
+  upper95_F <- round(exp((summary(mod_F)$coefficient[1,1]) +
+                                    1.96*(summary(mod_F)$coefficient[1,3])), 3)
+  # standard error
+  sd_stderr_F <- round(summary(mod_F)$coefficient[1,3], 4)
+  # p val
+  sd_pvalue_F <- round(summary(mod_F)$coefficient[1,5], 4)
+
+  total_n_F <- data_F %>%
+    filter(outcome == 1) %>%
+    summarise(total_n = n())
+
+  ### try ---
+  gender <- rbind(c("Male"), c("Female"))
+  total_n <- rbind(total_n_M, total_n_F)
+  estimate <- rbind(estimates_M, estimates_F)
+  lower95 <- rbind(lower95_M, lower95_F)
+  upper95 <- rbind(upper95_M, upper95_F)
+
+  return_estimate <- tibble(total_n_F, estimates_F, lower95_F, upper95_F)
+
+  return(return_estimate)
+}
+
+same_day_sex <- outcome_df_list %>%
+  purrr::map(~same_day_sex_fun(.)) %>%
+  plyr::rbind.fill() # %>%
+#   bind_cols(., data_frame(rep(outcome_names, each=2))) %>%
+#   rename(outcome_names = `rep(outcome_names, each = 2)`) %>%
+#   select(outcome_names, return_estimate) %>%
+  # preserve order of names
+#   mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))
+
+```
+
+```{r r same day stratified sex plot}
+dl_results <- outcome_df_list %>%
+  purrr::map(~distributed_lag(.)) %>%
+  plyr::rbind.fill() %>%
+  bind_cols(., data_frame(rep(outcome_names, each=2))) %>%
+  rename(outcome_names = `rep(outcome_names, each = 2)`) %>%
+  select(outcome_names, gender, total_n, estimate, lower95, upper95) %>%
+  # preserve order of names
+  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))
+
+```
+
+```{r same day stratified age function}
+
+```
+
 ##8. SABA cumulative effect by underlying respiratory diagnosis
@@
 
-##9. Conculstions
+```{r saba same day assciation function}
+
+```
+
+##9. Conculsions
 So far, it looks like SABA and Asthma are the only associtions.

### 2017-11-16 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
Updated code to estimate for strata.

--- 2017-11-15 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
@@
 
-**Note to Jingyang 11/3/17: I saw all the packages you used.
-How many do you actually need to run these analyses?
-It also looks like you need to run this markdown file on the server?
-I would like to avoid that and I can help you do this.**
-
-Packages used: tidyverse
+Packages used: tidyverse, survival.
 ```{r setup}
@@
 
-```{r descriptive table}
+```{r descriptive table function}
 # creating a row calculation function
@@
     filter(outcome == 1) %>%
-    summarise(total_n = n())
+    # adding in a filter to remove NAs
+    filter(!is.na(geo_smk_pm_zip)) %>% 
+    summarise(total_n = n()) %>%
 
@@
 } # end function
+```
 
+Table 1.
+
+```{r descritive table}
 # outcome names
@@
 
-*Jingyang, add same day associations here using purrr map functions here.
-I want to compare them to the cumulative associations.
-You do not need to make an elaborate function like I did for the distributed lag.
-You should just be able to use clogit.
-Include the same day GWR smoke estimate and same day temperature as a covariate.*
-
 *Note from Jingyang on Nov 14:*
@@
 Also for the DLM graph.*
+
+*Note to Jingyang on Nov 16:*
+*I still don't understand why the case n would be different from table 1 to the number of cases used in the models below?
+You should have identified all the cases in the time period of May to October, and then assigned them an exposure value.
+Referent values will be assigned at any other time point during that time as well.
+There should be no missing data in outcomes or exposure.
+You will have to tell me why cases are being included in the time-stratified dataframes, but excluded in analysis.*
 
@@
 same_day_fun <- function(data){
-
   # limit to complete cases
@@
 
-  # fit model with basis ----
+    # fit conditional logistic regression model ----
   mod <- clogit(outcome ~ geo_smk10 + wrf_temp_zip + strata(personkey),
@@
   # odds ratio
-  estimates <- round(exp(summary(mod)$coefficient[1,1]), 3) # sd stands for same day
-
+  estimates <- round(exp(summary(mod)$coefficient[1,1]), 3) 
   # 95% lower bound
@@
                                     1.96*(summary(mod)$coefficient[1,3])), 3)
-  # standard error
-  sd_stderr <- round(summary(mod)$coefficient[1,3], 4)
-  # p val
-  sd_pvalue <- round(summary(mod)$coefficient[1,5], 4)
 
@@
 ```
+
+Apply same-day function to each outcome dataframe.
 
@@
 
-##7. Stratified cumulative results
+##7. Stratified same-day results
 
+*Note to Jingyang: Nov 16*
+*I realized there is a better way to calculate stratified results than modifying the code. See below for example. There is a way to do it in one chunk rather than 2, but I was running in to errors.*
+
+### 7.a Sex
+
+Presenting same-day stratified results because the computations are easier.
+I'll use it to decide if we should present cumulative effects.
+We can use map and filter to subset to group of interest and apply our same-day function.
+We then bind the two strata together.
+
+```{r same day association stratified by sex}
+# there is probably a better way to do this, but this will work.
+sameday_results_f <- outcome_df_list %>% 
+  # using gender because I forget which sex_ind is which
+  # filter to females
+  purrr::map(~filter(., gender == "F")) %>%  
+  map_dfr(~same_day_fun(.)) %>% 
+  bind_cols(., data_frame(outcome_names)) %>% 
+  mutate(sex = as.factor("Female")) %>% 
+  select(outcome_names, sex, estimates, lower95, upper95) %>% 
+  # preserve order of names
+  mutate(outcome_names = parse_factor(outcome_names, levels = outcome))
+
+sameday_results_m <- outcome_df_list %>% 
+  # filter to males
+  purrr::map(~filter(., gender == "M")) %>%  
+  map_dfr(~same_day_fun(.)) %>% 
+  bind_cols(., data_frame(outcome_names)) %>% 
+  mutate(sex = as.factor("Male")) %>% 
+  select(outcome_names, sex, estimates, lower95, upper95) %>% 
+  # preserve order of names
+  mutate(outcome_names = parse_factor(outcome_names, levels = outcome))
+
+# bind sex dataframes together
+sameday_results <- rbind(sameday_results_f, sameday_results_m)
+
+# print table
+knitr::kable(sameday_results, 
+  caption = "Same day odds ratio and 95% CI of pharmacy/ED/urgent care visits")
+```
+
+Plot.
+```{r sex strata plot}
+plot <- ggplot(data=sameday_results, aes(x=sex, y = estimates, color=sex)) +
+  scale_color_manual(name = "Sex", values = c("#00A08A", "#046C9A"),
+    guide = guide_legend(direction="horizontal", title.position = "top",
+                         title.hjust = 0.5)) +
+  geom_point() +
+  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
+  facet_wrap(~outcome_names) +
+  geom_hline(yintercept = 1, linetype = 2, colour = "#FF0000") +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    legend.position = "bottom",
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black")) +
+  ylab(expression("Odds ratio: 10 ug/m^3 increase smoke PM"[2.5])) +
+  xlab("Sex")
+
+plot  
+```
+
+### 7.b Age
 *Jingyang to complete using function*
+Color palette to use for the 3 age categories. c("#899DA4", "#35274A" "#F2300F").
+
+##8. SABA same-day results by underlying respiratory diagnosis
+*Jingyang to complete using function*
+Color palette to use for the 4 age categories.
+c("#E2D200" "#46ACC8" "#E58601" "#B40F20")
+
+##9. Stratified cumulative results
+Considering adding.
+
 Stratify cumulative effect by sex and age group.
 
-##8. SABA cumulative effect by underlying respiratory diagnosis
-*Jingyang to complete using function*
+##10. SABA cumulative effect by underlying respiratory diagnosis
+Considering adding.
 
-##9. Conculstions
+##11. Conculsions
 So far, it looks like SABA and Asthma are the only associtions.
@@
 3. Variables in the casecross over datasets raise some questions.
+4. Consider a cumulative effect that adds the preceding day and so on, i.e. 0, 0+1, 0+1+2, ...

### 2017-11-16 r_scripts/analysis/oregon_cvd_pulm_results.Rmd

--- 2017-11-15 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
@@
+<<<<<<< HEAD
 ---
@@
 3. Variables in the casecross over datasets raise some questions.
+
+=======
+---
+title: "Douglas-complex fire smoke and cardiopulmonary morbidity"
+author: "Ryan Gan"
+date: "11/3/2017"
+output: html_document
+---
+
+##1. Introduction
+
+In the summer of 2013, the Douglas-Complex fires occured in southwest Oregon.
+Oregonians in this part of the state were at risk of exposure to extreme levels of particulate matter (PM).
+This project aim is to determine if there is an association to smoke from the Douglas-Complex fires and acute cardiopulmonary morbidity in the state of Oregon.
+
+Data came from the Oregon All Payer All Claims Database (APAC) in the year 2013.
+The APAC records the health care billing data for Oregon's insured populations.
+APAC include individual billing records for both diagnoses codes (International Classification of Diseases, Clinical Modification (ICD-9-CM) diagnoses codes) and pharmacy codes (National Drug Codes (NDC)).
+
+Our previous research that found an association with wildfire smoke and respiratory outcomes in Washington state in 2012 wildfire season using a novel estimate of smoke concentration, geographically weighted ridge regression (GWR) guided the methodological approaches used in this project.
+As Oregon contains pharmacy records, we evaluate the association between smoke and respiratory rescue medications (beta 2 agonists) (abbreviate to SABA).
+
+*Research question*
+We evaluated the association between smoke concentrations using the GWR method and cardiopulmonary morbidity, including ED/urgent care visits and SABA fills in Oregon state during the 2013 wildfire season.
+
+This markdown document contains the code and results that were used to address this research question.
+
+Packages used: tidyverse, survival.
+```{r setup}
+library(tidyverse) # general data wrangle
+library(survival) # conditional logistic regression
+```
+
+##2. Wildfire smoke descriptive characteristics
+
+Importing ZIP code-level population-weighted PM~2.5~ daily data.
+
+```{r smoke data import and wrangle}
+# pm path
+pm_path <- paste0("./data/pm/2013-oregon_county_pm25.csv")
+# county pm
+county_pm_df <- read_csv(pm_path) %>% 
+  mutate(county = stringr::str_replace(county, "[.]", ""))
+
+# estimate county smoke 
+county_smk_count <- county_pm_df %>% 
+  # binary smoke variable at >10 ug/m^3
+  mutate(gwr_smk10 = ifelse(geo_smk_pm >= 10, 1, 0)) %>% 
+  group_by(county) %>% 
+  summarise(smk_day10 = sum(gwr_smk10)) %>% 
+  # lower case county name and remove "." to match spatial df
+  mutate(subregion = tolower(stringr::str_replace(county, "[.]", " ")))
+
+# extract oregon map data
+or_county_spatial_df <- map_data("county", "oregon") %>% 
+  right_join(county_smk_count, by = "subregion")
+
+# use the map function from the maps package to get the polygon data
+county_poly <- maps::map("county", "oregon", plot=FALSE, fill = TRUE)
+# find centroids
+county_centroids <- maps:::apply.polygon(county_poly, maps:::centroid.polygon)
+
+# create a data frame for graphing out of the centroids of each polygon
+county_centroids <- county_centroids[!is.na(names(county_centroids))]
+centroid_array <- Reduce(rbind, county_centroids)
+
+county_text <- reduce(county_centroids, rbind) %>% 
+  as_data_frame() %>% 
+  rename(long = V1, lat = V2) %>% 
+  mutate(county = stringr::str_sub(
+    stringr::str_to_title(names(county_centroids)), start=8L)) 
+```
+
+### Number of days in Oregon where smoke PM~2.5~ exceeded 10 ug/m^3.
+
+```{r smoke days map}
+smoke_map <- ggplot(or_county_spatial_df, aes(x=long,y=lat, group=group)) +
+  # fill with number of smoke days
+  geom_polygon(aes(fill =smk_day10)) +
+  scale_fill_gradient(expression("Smoke Days > 10 µg/m"^3),
+    low= "#3B9AB2", high="#F21A00", 
+    guide = guide_colorbar(direction="horizontal", title.position = "top",
+                           title.hjust = 0.5, barwidth = 20)) +
+  # add county path on top
+  geom_path(colour="#EBCC2A") +
+  # add county text 
+  geom_text(data = county_text, aes(x=long, y=lat, label = county, group=NULL), 
+    colour = "white", size = 2) +
+  xlab("Longitude") +
+  ylab("Latitude") +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    strip.text = element_text(size = 8),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    strip.background = element_rect(colour=NA, fill=NA),
+    legend.position = "bottom")
+
+smoke_map
+```
+
+Most of the heavy smoke days were in the south-west part of Oregon.
+I've added county names to map but some text aren't well placed.
+I should probably use centroid, but then I'll have to import a shapefile to do this.
+I'll fix this later; I plan to add in fire locations.
+
+### Time-series of geo-weighted regression estimates of smoke PM~2.5~ by county
+
+Using the geofacet package to organize small-multiples by geography.
+
+*Note 11/7/17: Not sure how much I love the layout of this Oregon geofacet grid, but we'll see what people think.
+I like that geofacet lets you look at the two maps together to understand the exposure pattern, but too much whitespace.
+However, with this orientation, I believe you could cut out the smoke days figure.*
+
+```{r geofacet time series of pm}
+# use or_counties_grid from geo_facet 
+or_grid <- geofacet::or_counties_grid1
+# small multiples plot
+plot <- ggplot(county_pm_df, aes(x=date, y= geo_smk_pm)) + 
+  geom_point(color = "#00A08A", size = 0.5) +
+  scale_x_date(date_labels = "%m") +
+  geofacet::facet_geo(~county, grid = or_grid) +
+  ylab(expression("Smoke PM2.5 µg/m"^3)) +
+  xlab("Month in 2013") +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    strip.text = element_text(size = 8),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    strip.background = element_rect(colour=NA, fill=NA),
+    legend.position = "bottom")
+# plot 
+plot  
+```
+
+Elevated levels of smoke in late July to early August in Curry, Josephine, Jackson, and Klamath counties.
+
+##3. Descriptive table of number of health outcomes
+
+Importing casecross over dataframes.
+**Note to Jingyang 11/3/17: Same issue as above.
+How many of these files needs to be uploaded to produce results?
+A lot of this seems redundant or inefficient.
+I cannot find a lot of outcome files I need.
+I also need the SABA casecrossover dataframe.
+Listing example of two casecross over dataframes for now.
+Let's talk about this.**
+
+*Note 2017-11-14: It may be easiest to make a set of casecross over functions and put them in a package rather than import the timestratified casecross overs each time.*
+
+```{r health data import, message=F, warning=F}
+# path for health dataframe
+health_path <- "./data/health/"
+# importing time-stratified case-crossover dataframe
+casecross_files <- list.files(path=health_path, pattern="*casecross*")
+
+# reorder file names in way I'd like it to be analyzed
+# also ditching broken arm
+casecross_order <- casecross_files[c(13,12,3,6,11,1,7,2,5,8,9,10)]
+
+# extract the name of the dataframe after study type
+file_names <- stringr::str_extract(casecross_order, 
+  pattern="(?<=casecross_)\\w+")
+
+# read in all the files in to a list
+outcome_df_list <- casecross_order %>%  
+  purrr::map(~read_csv(paste0(health_path, .)))
+
+# name dataframes
+names(outcome_df_list) <- file_names
+
+# name list elements in order I'd like to present in document
+outcome_names <- c("SABA", "Respiratory", "Asthma", "COPD", "Pneumonia", 
+  "Acute Bronchitis", "Cardiovascular Disease", "Arrhythmia", 
+  "Cerebrovascular Disease", "Heart Failure", "Ischemic Heart Disease", 
+  "Myocardial Infarction")
+```
+
+CVD warning message on import. Have Jingyang check. Maybe why proportion of CVD vs respiratory so different in Oregon?
+*Warning message:In rbind(names(probs), probs_f) :number of columns of result is not a multiple of vector length (arg 1)*
+
+*Also warnings on mixed data types of NULL and numeric; another reason to revise and create timestratified casecross over functions*.
+
+Descriptive table will go here. Jingyang, I'll have you add the other dataframes to analyze. I provided an example of Asthma and CVD.
+
+```{r descriptive table function}
+# creating a row calculation function
+table1_fun <- function(data_list){
+  # output of total n
+  total_n <- data_list %>% 
+    filter(outcome == 1) %>%
+    # adding in a filter to remove NAs
+    filter(!is.na(geo_smk_pm_zip)) %>% 
+    summarise(total_n = n()) %>%
+
+  # output row of summary numbers
+  age_cat_n <- data_list %>% 
+    filter(outcome == 1) %>%
+    group_by(age_ind) %>% 
+    summarise(n = n()) %>%
+    # this will fix the 0 categories of certain CVD outcomes in children
+    add_row(age_ind = 0, n = 0) %>% 
+    arrange(age_ind) %>% 
+    group_by(age_ind) %>% 
+    summarise(n = sum(n)) %>% 
+    mutate(age_cat = ifelse(age_ind == 0, "age_under15",
+                    ifelse(age_ind == 1, "age_15to65", "age_over65")),
+           # preserve factor order
+           age_cat = parse_factor(age_cat, levels = age_cat)) %>% 
+    select(-age_ind) %>% 
+    spread(age_cat, n)
+  # sex n
+  sex_n <- data_list %>% 
+    filter(outcome == 1) %>% 
+    group_by(gender) %>% 
+    summarise(n = n()) %>% 
+    filter(gender != "U") %>% 
+    spread(gender, n)
+  # bind all rows together
+  outcome_n_row <- bind_cols(total_n, age_cat_n, sex_n) %>% 
+    mutate(perc_15 = round(age_under15/total_n,2), 
+      perc_15to65 = round(age_15to65/total_n,2),
+      perc_65 = round(age_over65/total_n,2),
+      perc_F = round(F/total_n,2),
+      perc_M = round(M/total_n,2)) %>% 
+    select(total_n, age_under15, perc_15, age_15to65, perc_15to65, 
+           age_over65, perc_65, F, perc_F, M, perc_M)
+return(outcome_n_row)
+} # end function
+```
+
+Table 1.
+
+```{r descritive table}
+# outcome names
+outcome <- outcome_names %>% 
+  as_tibble() %>%
+  rename(outcome = value)
+
+# apply custom function to generate descriptive table
+table_vals <- outcome_df_list %>% 
+  # map to dataframe
+  map_dfr(table1_fun) 
+# descriptive table, bind outcome names and values
+descriptive_table <- bind_cols(outcome, table_vals)
+
+# output table
+knitr::kable(descriptive_table, 
+  caption = "Descriptive charactersitics of ED/urgent care visits")
+```
+
+##4. Same-day association results
+
+*Note from Jingyang on Nov 14:*
+*Just a little improvement advice for plot: Because the number (which labels the last disease's odds ratio) can only be seen one digital, I suggest to add more margin to the right side.
+Also for the DLM graph.*
+
+*Note to Jingyang on Nov 16:*
+*I still don't understand why the case n would be different from table 1 to the number of cases used in the models below?
+You should have identified all the cases in the time period of May to October, and then assigned them an exposure value.
+Referent values will be assigned at any other time point during that time as well.
+There should be no missing data in outcomes or exposure.
+You will have to tell me why cases are being included in the time-stratified dataframes, but excluded in analysis.*
+
+Create function to calculate same-day association (odds ratio) and 95% confidence interval between increasing smoke and outcomes.
+
+```{r same day association fuction}
+same_day_fun <- function(data){
+  # limit to complete cases
+  complete_data <- data %>% 
+    # filter to complete case for smoke
+    filter(!is.na(geo_smk_pm_zip)) %>% 
+    # create 10 unit smk variable 
+    mutate(geo_smk10 = geo_smk_pm_zip/10)
+
+    # fit conditional logistic regression model ----
+  mod <- clogit(outcome ~ geo_smk10 + wrf_temp_zip + strata(personkey),
+    data = complete_data)
+
+  # odds ratio
+  estimates <- round(exp(summary(mod)$coefficient[1,1]), 3) 
+  # 95% lower bound
+  lower95 <- round(exp((summary(mod)$coefficient[1,1]) -
+                                    1.96*(summary(mod)$coefficient[1,3])), 3)
+  # 95% upper bound
+  upper95 <- round(exp((summary(mod)$coefficient[1,1]) +
+                                    1.96*(summary(mod)$coefficient[1,3])), 3)
+
+  return_estimate <- data_frame(estimates, lower95, upper95) 
+  return(return_estimate) 
+} 
+```
+
+Apply same-day function to each outcome dataframe.
+
+```{r same day association table}
+sameday_results <- outcome_df_list %>% 
+  map_dfr(~same_day_fun(.)) %>% 
+  bind_cols(., data_frame(outcome_names)) %>% 
+  select(outcome_names, estimates, lower95, upper95) %>% 
+  # preserve order of names
+  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))
+
+knitr::kable(sameday_results, 
+  caption = "Same day odds ratio and 95% CI of pharmacy/ED/urgent care visits")
+```
+
+Applying function to the list of outcome dataframes.
+
+```{r same day association graph}
+# ggplot ----
+plot <- ggplot(sameday_results, aes(x=outcome_names, y = estimates)) +
+  geom_point(colour = "#46ACC8") +
+  geom_text(aes(label = round(estimates,2)), hjust = -0.25) +
+  geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                colour= "#46ACC8", width = 0.2) +
+  geom_hline(yintercept = 1, linetype = 2, colour = "#B40F20") +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+   panel.grid.major = element_blank(),
+   panel.grid.minor = element_blank(),
+   axis.text.x = element_text(angle=90, hjust = 1)) +
+  ylab(expression("Same day odds ratio: 10 ug/m^3 increase smoke PM"[2.5])) +
+  xlab("Outcomes")
+
+plot
+```
+
+##5. Distributed lag cumulative effect results
+
+For this project, we use the time-stratified case-crossover study design in the same way we did for the 2012 Wenatchee-Complex fire project.
+Briefly, we identify admissions dates for ED or urgent care visits, and pharmacy fill dates for SABA, using billing records.
+
+In the Wenatchee-Complex fire project, we looked at the same-day relationship between smoke exposure and health outcomes.
+For the Douglas-complex fire, we will evaluate a 0 to 7 day distributed lag association, first comparing the overall cumulative effect over these 7 days, and then get in to the lagged effects.
+
+*Note: 11/7/17: Jingyang will provide a csv of the best-fit knots for the distributed lag splines based on AIC.*
+
+*Note: 11/7/17: Jingyang, can I use the lagged days in the case-crossover dataframes?
+Please confrim that the value for lag day 1 is the day before, lag day 2 is from 2 days before, etc..., and not the week before, two weeks before, etc...
+What does the _m denoter mean too?
+Jingyang says it's not necissary; we will remove creating this variable in the data management scripts.
+Do we use the patid or personkey identifier for strata?
+check.*
+
+Basis function also included for WRF temp. We will need to find the best fit of spline for temp seperately from PM. For asthma, it looks like 2 DF is best.
+
+Making a custom distributed lag function. Consider adding additional options like specifying number of lag days. That would require a reworking of how the case-crossovers are set up (kind of).
+
+```{r distributed lag function}
+# custom distributed lag function 
+distributed_lag <- function(data, pm_df, temp_df, cumulative=T){
+  # limit to complete cases
+  complete_data <- data %>% 
+    # filter to complete case for smoke
+    filter(!is.na(geo_smk_pm_lag7_zip))
+
+  # output matrix of gwr values
+  pm_matrix <- complete_data %>% 
+    select(contains("geo_smk_pm")) %>%
+    # remove anything with m in it; will go back in original code and not create 
+    # this to begin with
+    select(-contains("_m")) %>% 
+    # divide exposure values by 10 units to interpret on 10 ug/m^3 scale
+    mutate_all(funs(./10)) %>% 
+    # convert to matrix
+    as.matrix()
+
+  # output temp matrix
+  temp_matrix <- complete_data %>% 
+    select(contains("wrf_temp")) %>%
+    # remove anything with m in it; will go back in original code and not create 
+    # this to begin with
+    select(-contains("_m")) %>% 
+    # convert to matrix
+    as.matrix()
+
+  # calculation of basis ----
+  # define basis using natural spline function from "splines"" package
+  pm_b <- splines::ns(0:7, df=pm_df, intercept=T)
+  # create pm basis
+  pm_basis <- pm_matrix %*% pm_b
+  # create temp basis
+  temp_b <- splines::ns(0:7, df=temp_df, intercept=T)
+  temp_basis <- temp_matrix %*% temp_b
+
+  # fit model with basis ----
+  mod <- clogit(outcome ~ pm_basis + temp_basis + strata(personkey),
+    data = complete_data)
+
+  # calculate estimates ----
+  # output pm basis parameters
+  dl_parms <- broom::tidy(mod) %>% 
+    filter(stringr::str_detect(term, "pm")) %>% 
+    select(estimate) %>% 
+    as_vector()
+  # estimate distributed lag values for each day
+  dl_estimates <- data.frame(estimate = pm_b %*% dl_parms)
+  # covariance matrix for knots
+  cov_matrix <- as.matrix(vcov(mod))[1:3,1:3]
+
+  # estimate variance of spline
+  dl_var <- pm_b %*% cov_matrix %*% t(pm_b)
+  # estimate standard error
+  dl_estimates$stderr <- sqrt(diag(dl_var))
+
+  # calculate lower and upper bound
+  dl_estimates$lower95 <- dl_estimates$estimate-(dl_estimates$stderr*1.96)
+  dl_estimates$upper95 <- dl_estimates$estimate+(dl_estimates$stderr*1.96)
+
+    if(cumulative==T) {
+      type <- "cumulative"  
+      # cumulative outcome and 95CI
+      estimate <- sum(dl_estimates$estimate)
+      # stderr cumulative effect
+      estimate_se <- sqrt(sum(dl_var))
+      # cumulative 95CI
+      lower95 <- estimate-(estimate_se*1.96)
+      upper95 <- estimate+(estimate_se*1.96)
+      # return dataframe
+      return_estimate <- data_frame(type, estimate, lower95, upper95) %>% 
+        mutate_if(is.numeric, exp)
+      return(return_estimate)
+    }
+    else { # if cumulative is not true, or false, return distributed lag est
+      # return dataframe
+      return_estimate <- dl_estimates %>% 
+        mutate_if(is.numeric, exp) %>% 
+        mutate(type = "lag", 
+               time = (as.numeric(rownames(.))-1)/-1) %>% 
+        select(type, time, estimate, lower95, upper95)
+       # return estimate  
+      return(return_estimate)
+      }
+      # end of if
+} 
+```
+
+Using the custom function to estimate cumulative effects for the list of dataframes. It will work on an individual dataframe, but I built it to use with the purrr map functions.
+
+*Note for Jingyang 11/8/17: purrr map functions are the tidyverse equivalents of apply-type functions*
+
+```{r cumulative association}
+# use custom function to create a dataframe of cumulative values
+# using purrr::map_dfr function
+# purrr::map_dfr will work if it's just the cumulative, but won't work for lagged
+cumulative_results <- outcome_df_list %>% 
+  map_dfr(~distributed_lag(., pm_df=3, temp_df=2, cumulative=T)) %>% 
+  bind_cols(., data_frame(outcome_names)) %>% 
+  select(outcome_names, type, estimate, lower95, upper95) %>% 
+  # preserve order of names
+  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))
+
+# will put this table eventually and split out plot
+#head(cumulative_results)
+
+# ggplot ----
+plot <- ggplot(cumulative_results, aes(x=outcome_names, y = estimate)) +
+  geom_point(colour = "#00A08A") +
+  geom_text(aes(label = round(estimate,2)), hjust = -0.25) +
+  geom_errorbar(aes(ymin=lower95, ymax=upper95), 
+                colour= "#00A08A", width = 0.2) +
+  geom_hline(yintercept = 1, linetype = 2, colour = "#FF0000") +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+   panel.grid.major = element_blank(),
+   panel.grid.minor = element_blank(),
+   axis.text.x = element_text(angle=90, hjust = 1)) +
+  ylab(expression("Cumulative odds ratio: 10 ug/m^3 increase smoke PM"[2.5])) +
+  xlab("Outcomes")
+
+plot
+```
+
+This is generally how I'd like the plots to look, but I plan to add color grouping to SABA, respiratory, and CVD groupings.
+Also considering reducing the lag time evaluated since I think we are getting in to variance issues a week out with this dataset (as evidence, large error around asthma outcome relative to other studies).
+
+I will also come back and tweak color scheme and labels.
+
+##6. Distributed lag results
+
+That same function can be used to generate detailed estimates of the association over time. In this case, we'll look at the same day to seven days prior the event.
+
+```{r distributed lag}
+# this approach will work that creates a list of dataframes with map then
+# binds the lists together in one dataframe
+dl_results <- outcome_df_list %>% 
+  purrr::map(~distributed_lag(., pm_df=3, temp_df=2, cumulative=F)) %>% 
+  plyr::rbind.fill() %>% 
+  bind_cols(., data_frame(rep(outcome_names, each=8))) %>% 
+  rename(outcome_names = `rep(outcome_names, each = 8)`) %>% 
+  select(outcome_names, type, time, estimate, lower95, upper95) %>% 
+  # preserve order of names
+  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))
+
+# results will come in a format to make small multiples I think
+#head(dl_results)
+
+# small multiples DL plot ----
+dl_plot <- ggplot(dl_results, aes(x=time, y=estimate)) +
+  geom_line(colour = "#046C9A", size = 1) +
+  geom_ribbon(aes(ymin = lower95, ymax = upper95), 
+              fill = "#ABDDDE", alpha = 0.5) + 
+  scale_y_continuous(limits = c(0.8, 1.2)) +
+  scale_x_continuous(breaks = c(seq(-7, 0, by=1))) +
+  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
+  facet_wrap(~outcome_names) +
+  ylab(expression("Odd ratio for a 10 ug/m^3 increase smoke PM"[2.5])) +
+  xlab("Lagged Days") +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    strip.text = element_text(size = 10),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    strip.background = element_rect(colour=NA, fill=NA))
+
+dl_plot
+```
+
+DL plot shows an effect with asthma even though the cumulative effect size is similar in magnitude to what we've seen before, but the 95% CIs overlap 1.
+I think this is a power issue trying to look at lagged effects for more than a week.
+I've applied a general 3 degrees of freedom knot for PM~2.5~ and 2 degrees of freedom knot for temperature.
+I think temperature is okay to do this, but I'm still figuring out a way to build a best-fit knot in to the function.
+
+Worth discussing with the group.
+
+##7. Stratified same-day results
+
+*Note to Jingyang: Nov 16*
+*I realized there is a better way to calculate stratified results than modifying the code. See below for example. There is a way to do it in one chunk rather than 2, but I was running in to errors.*
+
+### 7.a Sex
+
+Presenting same-day stratified results because the computations are easier.
+I'll use it to decide if we should present cumulative effects.
+We can use map and filter to subset to group of interest and apply our same-day function.
+We then bind the two strata together.
+
+```{r same day association stratified by sex}
+# there is probably a better way to do this, but this will work.
+sameday_results_f <- outcome_df_list %>% 
+  # using gender because I forget which sex_ind is which
+  # filter to females
+  purrr::map(~filter(., gender == "F")) %>%  
+  map_dfr(~same_day_fun(.)) %>% 
+  bind_cols(., data_frame(outcome_names)) %>% 
+  mutate(sex = as.factor("Female")) %>% 
+  select(outcome_names, sex, estimates, lower95, upper95) %>% 
+  # preserve order of names
+  mutate(outcome_names = parse_factor(outcome_names, levels = outcome))
+
+sameday_results_m <- outcome_df_list %>% 
+  # filter to males
+  purrr::map(~filter(., gender == "M")) %>%  
+  map_dfr(~same_day_fun(.)) %>% 
+  bind_cols(., data_frame(outcome_names)) %>% 
+  mutate(sex = as.factor("Male")) %>% 
+  select(outcome_names, sex, estimates, lower95, upper95) %>% 
+  # preserve order of names
+  mutate(outcome_names = parse_factor(outcome_names, levels = outcome))
+
+# bind sex dataframes together
+sameday_results <- rbind(sameday_results_f, sameday_results_m)
+
+# print table
+knitr::kable(sameday_results, 
+  caption = "Same day odds ratio and 95% CI of pharmacy/ED/urgent care visits")
+```
+
+Plot.
+```{r sex strata plot}
+plot <- ggplot(data=sameday_results, aes(x=sex, y = estimates, color=sex)) +
+  scale_color_manual(name = "Sex", values = c("#00A08A", "#046C9A"),
+    guide = guide_legend(direction="horizontal", title.position = "top",
+                         title.hjust = 0.5)) +
+  geom_point() +
+  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
+  facet_wrap(~outcome_names) +
+  geom_hline(yintercept = 1, linetype = 2, colour = "#FF0000") +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    legend.position = "bottom",
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black")) +
+  ylab(expression("Odds ratio: 10 ug/m^3 increase smoke PM"[2.5])) +
+  xlab("Sex")
+
+plot  
+```
+
+### 7.b Age
+*Jingyang to complete using function*
+Color palette to use for the 3 age categories. c("#899DA4", "#35274A" "#F2300F").
+
+##8. SABA same-day results by underlying respiratory diagnosis
+*Jingyang to complete using function*
+Color palette to use for the 4 age categories.
+c("#E2D200" "#46ACC8" "#E58601" "#B40F20")
+
+##9. Stratified cumulative results
+Considering adding.
+
+Stratify cumulative effect by sex and age group.
+
+##10. SABA cumulative effect by underlying respiratory diagnosis
+Considering adding.
+
+##11. Conculsions
+So far, it looks like SABA and Asthma are the only associtions.
+Furthermore, the cumulative 0-7 day relationship between asthma and smoke has wide error bars and overlaps 1, suggesting the association could be null.
+I think it's more likely that this is because we introduce some noise on lag days 4-7.
+It's possible that an adjustment like weekend may help.
+It may also make sense to do something similar to Anna Rappold or others and shorten our lagged response.
+I also think this is a produce of sample size too.
+
+### Comments:
+1. File structure needs to be cleaned up. There are way too many subdirectories.
+2. Remove r code files that are not useful. There are too many files with similar code inside.
+3. Variables in the casecross over datasets raise some questions.
+4. Consider a cumulative effect that adds the preceding day and so on, i.e. 0, 0+1, 0+1+2, ...
+
+>>>>>>> upstream/development

### 2017-11-16 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
Fixed bugs and able to knit results document.

--- 2017-11-16 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
@@
 library(survival) # conditional logistic regression
+
+# knitr options
+knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
+                      echo=FALSE, warning=FALSE, message=FALSE)
 ```
@@
 # pm path
-pm_path <- paste0("./data/pm/2013-oregon_county_pm25.csv")
+pm_path <- paste0("../../data/pm/2013-oregon_county_pm25.csv")
 # county pm
@@
 # path for health dataframe
-health_path <- "./data/health/"
+health_path <- "../../data/health/"
 # importing time-stratified case-crossover dataframe
@@
 
-```{r descriptive table function}
+```{r descriptive table}
 # creating a row calculation function
@@
     filter(!is.na(geo_smk_pm_zip)) %>% 
-    summarise(total_n = n()) %>%
+    summarise(total_n = n())
 
@@
   # map to dataframe
-  map_dfr(table1_fun) 
+  map_dfr(~table1_fun(.))
+
 # descriptive table, bind outcome names and values
@@
 dl_results <- outcome_df_list %>% 
-  purrr::map(~distributed_lag(., pm_df=3, temp_df=2, cumulative=F)) %>% 
+  purrr::map(~distributed_lag(., pm_df=3,temp_df=2,cumulative=F)) %>% 
   plyr::rbind.fill() %>% 
@@
   # preserve order of names
-  mutate(outcome_names = parse_factor(outcome_names, levels = outcome))
+  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))
 
@@
 knitr::kable(sameday_results, 
-  caption = "Same day odds ratio and 95% CI of pharmacy/ED/urgent care visits")
+  caption = "Same day odds ratio and 95% CI of pharmacy/ED/urgent care visits stratified by sex.")
 ```
@@
 ### 7.b Age
-*Jingyang to complete using function*
-Color palette to use for the 3 age categories. c("#899DA4", "#35274A" "#F2300F").
+
+Stratified by age category.
+
+```{r same day association stratified by age category}
+# subset outcomes evaluated for under 15
+under15_outcome <- outcome_names[1:7]
+
+# subset dataframes that won't have convergence issues
+sameday_results_under15 <- outcome_df_list[1:7] %>% 
+  # filter to children
+  purrr::map(~filter(., age_ind == 0)) %>%  
+  map_dfr(~same_day_fun(.)) %>% 
+  bind_cols(., data_frame(under15_outcome)) %>% 
+  rename(outcome_names = under15_outcome) %>% 
+  mutate(age_cat = as.factor("Under 15")) %>% 
+  select(outcome_names, age_cat, estimates, lower95, upper95) %>% 
+  # preserve order of names
+  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))
+
+# age category 1
+sameday_results_15to65 <- outcome_df_list %>% 
+  # filter to males
+  purrr::map(~filter(., age_ind == "1")) %>%  
+  map_dfr(~same_day_fun(.)) %>% 
+  bind_cols(., data_frame(outcome_names)) %>% 
+  mutate(age_cat = as.factor("15 to 65")) %>% 
+  select(outcome_names, age_cat, estimates, lower95, upper95) %>% 
+  # preserve order of names
+  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))
+
+# age category 2
+sameday_results_over65 <- outcome_df_list %>% 
+  purrr::map(~filter(., age_ind == "2")) %>%  
+  map_dfr(~same_day_fun(.)) %>% 
+  bind_cols(., data_frame(outcome_names)) %>% 
+  mutate(age_cat = as.factor("Over 65")) %>% 
+  select(outcome_names, age_cat, estimates, lower95, upper95) %>% 
+  # preserve order of names
+  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))
+
+# bind sex dataframes together
+sameday_results <- rbind(sameday_results_under15, sameday_results_15to65,
+                         sameday_results_over65)
+
+# print table
+knitr::kable(sameday_results, 
+  caption = "Same day odds ratio and 95% CI of pharmacy/ED/urgent care visits stratified by age category.")
+```
+
+Stratified by age plot.
+```{r age strata plot}
+plot <- ggplot(data=sameday_results, 
+               aes(x=age_cat, y = estimates, color=age_cat)) +
+  scale_color_manual(name = "Age Category", 
+    values = c("#899DA4", "#35274A", "#F2300F"),
+    guide = guide_legend(direction="horizontal", title.position = "top",
+                         title.hjust = 0.5)) +
+  geom_point() +
+  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
+  facet_wrap(~outcome_names, scales = "free_y") +
+  geom_hline(yintercept = 1, linetype = 2, colour = "#FF0000") +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    legend.position = "bottom",
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black")) +
+  ylab(expression("Odds ratio: 10 ug/m^3 increase smoke PM"[2.5])) +
+  xlab("Age Category")
+
+plot  
+```
+
+Under 15 unstable for CVD; consider removing. Heart failure is interesting for the over 65 group. That may agree with what Anna Rappold has seen.
 
 ##8. SABA same-day results by underlying respiratory diagnosis
-*Jingyang to complete using function*
-Color palette to use for the 4 age categories.
-c("#E2D200" "#46ACC8" "#E58601" "#B40F20")
+
+```{r saba underlying dx}
+# read in the index 
+saba_index <- read_csv("../../data/health/oregon_saba_index.csv") %>% 
+    mutate(underlying_dx = as.factor(ifelse(index == "copd",
+    stringr::str_to_upper(index),
+    stringr::str_to_title(index)))) %>% 
+    select(-index)
+
+# index vector
+underlying_dx <- unique(as.factor(saba_index$underlying_dx)) %>% 
+  data_frame() %>% 
+  rename(underlying_dx = ".") %>% 
+  mutate(underlying_dx = ) %>% 
+  # need to arrange since slice is going alphabetical
+  arrange(underlying_dx)
+
+# join index to saba
+# extract saba dataframe from list
+saba <- outcome_df_list[[1]] %>% 
+  left_join(saba_index, by = "personkey")
+
+# run function on each group
+sameday_results <- saba %>% 
+  split(.$underlying_dx) %>% 
+  map_dfr(~same_day_fun(.)) %>% 
+  bind_cols(., underlying_dx) %>% 
+  select(4,1:3)
+
+# table
+knitr::kable(sameday_results, 
+  caption = "Same day odds ratio and 95% CI for a SABA fill, stratified by underlying dx.")
+```
+
+*Jingyang, did I ask you to code "other" if a dx was missing? Maybe we should change other to "unknown or other" since I doubt we were able to link an underlying dx to everyone with a fill.*
+
+Plot of SABA underlying Dx categories.
+
+```{r plot saba underlying dx}
+plot <- ggplot(sameday_results, 
+   aes(x=underlying_dx, y = estimates, color = underlying_dx)) +
+   scale_color_manual(name = "Underlying Diagnosis", 
+    values = c("#0B775E","#46ACC8","#E58601","#B40F20"),
+    guide = guide_legend(direction="horizontal", title.position = "top",
+                         title.hjust = 0.5)) +
+  geom_point() +
+  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
+  geom_hline(yintercept = 1, linetype = 2, colour = "#FF0000") +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+   panel.grid.major = element_blank(),
+   panel.grid.minor = element_blank(),
+   axis.title.x = element_blank(),
+   legend.position = "bottom") +
+  ylab(expression("Odds Ratio: 10 ug/m^3 increase smoke PM"[2.5])) 
+# plot
+plot
+```
 
@@
 Considering adding.
-
-Stratify cumulative effect by sex and age group.
 
@@
 
-### Comments:
+### Issues/Concerns (in order of time they popped in my head):
 1. File structure needs to be cleaned up. There are way too many subdirectories.
@@
 4. Consider a cumulative effect that adds the preceding day and so on, i.e. 0, 0+1, 0+1+2, ...
+5. Plots need to have standard labels, ordering, scales, etc.

### 2017-11-16 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
updated setup chunk.
fixed geo_facet, hood river.

@@
 Packages used: tidyverse, survival.
-```{r setup}
+```{r setup, echo =F}
 library(tidyverse) # general data wrangle
@@
 # use or_counties_grid from geo_facet 
-or_grid <- geofacet::or_counties_grid1
+or_grid <- geofacet::or_counties_grid1 %>% 
+  mutate(name = ifelse(name == "HoodRiver", "Hood River", name))
+
 # small multiples plot

### 2017-11-17 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
Fixed color schemes.

@@
 Packages used: tidyverse, survival.
-```{r setup, echo =F}
+```{r setup, echo =F, warning=F, message=F}
 library(tidyverse) # general data wrangle
@@
 # knitr options
-knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
+knitr::opts_chunk$set(fig.width=12, fig.height=10, fig.path='Figs/',
                       echo=FALSE, warning=FALSE, message=FALSE)
@@
 
-Importing ZIP code-level population-weighted PM~2.5~ daily data.
+Smoke exposure is represented by ZIP code-level population-weighted PM~2.5~ estimated using geographically-weighted ridge regression (GWR), where we have subtracted off the background median PM~2.5~ for the month where no smoke was present overhead in the atmospheric column using HMS.
+It was estimated every day for the wildfire season of May to October, 2013.
+
+We've decided to represent descriptive characteristics of smoke by county-level as we can represent the location and temporal variation of smoke exposure better than we could at the ZIP code level.
 
@@
   mutate(county = stringr::str_replace(county, "[.]", ""))
+
+# fire locations
+fire_locations <- read_csv("../../data/pm/2013-oregon_firelocations.csv") %>% 
+  mutate(area = sqrt(total_acres/1000))
 
@@
 
+We summarize smoke exposure two ways: as a count of days where smoke PM~2.5~ is > 10 ug/m^3 (to increase the specificity to identify areas impacted by smoke) at the county level, and time-series of county-population weighted PM~2.5~ over May to October 2013.
+
 ### Number of days in Oregon where smoke PM~2.5~ exceeded 10 ug/m^3.
@@
   geom_polygon(aes(fill =smk_day10)) +
-  scale_fill_gradient(expression("Smoke Days > 10 µg/m"^3),
-    low= "#3B9AB2", high="#F21A00", 
+  scale_fill_gradient2(expression("Smoke Days > 10 µg/m"^3),
+    low = "white", mid= "#00dbde", high="#6a3093", midpoint = 15,
     guide = guide_colorbar(direction="horizontal", title.position = "top",
@@
   # add county path on top
-  geom_path(colour="#EBCC2A") +
+  geom_path(colour="#4ac29a") +
+  # adding fire locations in
+  geom_point(data=fire_locations, 
+    aes(x=lon, y=lat, size = area,
+    shape = "Fire Locations"), shape = 17, color = "#ff0084", alpha = 0.7,
+    show.legend = FALSE, inherit.aes = FALSE) +
   # add county text 
-  geom_text(data = county_text, aes(x=long, y=lat, label = county, group=NULL), 
-    colour = "white", size = 2) +
+  geom_text(data = county_text, 
+    aes(x=long, y=lat, label = county, group=NULL), 
+    colour = "#33001b", size = 2.5) +
   xlab("Longitude") +
@@
   theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    strip.text = element_text(size = 8),
     panel.grid.major = element_blank(),
@@
 plot <- ggplot(county_pm_df, aes(x=date, y= geo_smk_pm)) + 
-  geom_point(color = "#00A08A", size = 0.5) +
+  geom_point(color = "#6a3093", size = 0.5) +
   scale_x_date(date_labels = "%m") +
@@
 
-Importing casecross over dataframes.
-**Note to Jingyang 11/3/17: Same issue as above.
-How many of these files needs to be uploaded to produce results?
-A lot of this seems redundant or inefficient.
-I cannot find a lot of outcome files I need.
-I also need the SABA casecrossover dataframe.
-Listing example of two casecross over dataframes for now.
-Let's talk about this.**
-
-*Note 2017-11-14: It may be easiest to make a set of casecross over functions and put them in a package rather than import the timestratified casecross overs each time.*
+Importing time-stratified case-crossover dataframes.
+Like the Washington study, we used the time-stratified case-crossover where admissions/fills are identified by date, and then referent observations are created for each observation (assumed to be a unique person) from May to October 2013.
+We argue that this is a good referent period that allows for enough referent periods, without introducing seasonal variation in exposure/disease.
+We also subtract off background smoke.
 
@@
 
-CVD warning message on import. Have Jingyang check. Maybe why proportion of CVD vs respiratory so different in Oregon?
-*Warning message:In rbind(names(probs), probs_f) :number of columns of result is not a multiple of vector length (arg 1)*
+CVD warning message on import; I think it's caused by varialbes with mixed character/integer (common in claims data).
+They don't have any bearing on analyses (but I find them annoying and plan to fix).
 
@@
 
-Descriptive table will go here. Jingyang, I'll have you add the other dataframes to analyze. I provided an example of Asthma and CVD.
+Descriptive counts of the number of ED/urgent care or pharmacy fills observed between May to October, 2013 in the state of Oregon, and who were assigned a smoke value.
 
-```{r descriptive table}
+```{r descriptive table function}
 # creating a row calculation function
@@
     # adding in a filter to remove NAs
-    filter(!is.na(geo_smk_pm_zip)) %>% 
+    filter(!is.na(geo_smk_pm_zip) & gender != "U") %>% 
     summarise(total_n = n())
@@
 
-Table 1.
+Table 1 total, and stratified by age categories and sex.
 
@@
 knitr::kable(descriptive_table,
-  caption = "Descriptive charactersitics of ED/urgent care visits")
+  caption = paste0("Descriptive charactersitics SABA fills",
+                   " and of ED/urgent care visits"))
 ```
@@
 ##4. Same-day association results
+For this project, we use the time-stratified case-crossover study design in the same way we did for the 2012 Wenatchee-Complex fire project.
+Briefly, we identify admissions dates for ED or urgent care visits, and pharmacy fill dates for SABA, using billing records.
 
-*Note from Jingyang on Nov 14:*
-*Just a little improvement advice for plot: Because the number (which labels the last disease's odds ratio) can only be seen one digital, I suggest to add more margin to the right side.
-Also for the DLM graph.*
+In the Wenatchee-Complex fire project, we looked at the same-day relationship between smoke exposure and health outcomes.
+For the Douglas-complex fire, we will evaluate a 0 to 7 day distributed lag association, first comparing the overall cumulative effect over these 7 days, and then get in to the lagged effects.
 
-*Note to Jingyang on Nov 16:*
-*I still don't understand why the case n would be different from table 1 to the number of cases used in the models below?
-You should have identified all the cases in the time period of May to October, and then assigned them an exposure value.
-Referent values will be assigned at any other time point during that time as well.
-There should be no missing data in outcomes or exposure.
-You will have to tell me why cases are being included in the time-stratified dataframes, but excluded in analysis.*
-
-Create function to calculate same-day association (odds ratio) and 95% confidence interval between increasing smoke and outcomes.
+In this section of the document, we look at the association between GWR smoke PM~2.5~ concentration on the same day of admission or fill using conditional logistic regression.
+All models are adjusted for temperature on the same day.
 
@@
 
-Apply same-day function to each outcome dataframe.
+Table of odds ratios (OR) and 95% confidence intervals (95%CI) that represent the same-day association between an increase in 10 ug/m^3 increase in smoke PM~2.5~ and the odds (can be interpreted as risk due to low prevalence) of the outcome.
 
@@
   # preserve order of names
-  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))
+  mutate(outcome_names = parse_factor(outcome_names, levels=outcome_names)) %>%
+  rename(odds_ratio = estimates)
 
@@
 
-Applying function to the list of outcome dataframes.
+Plot of the ORs and 95%CIs in the table. I've grouped by cardiovascular and respiratory classes of outcomes.
 
 ```{r same day association graph}
+class <- as_tibble(c(rep("Respiratory", times = 6), 
+  rep("Cardiovascular", times = 6))) %>% 
+  rename(class = value)
+# add a class indicator to table
+sameday_results_w_class <- sameday_results %>% bind_cols(class)
+  sameday_results_w_class
+
 # ggplot ----
-plot <- ggplot(sameday_results, aes(x=outcome_names, y = estimates)) +
-  geom_point(colour = "#46ACC8") +
-  geom_text(aes(label = round(estimates,2)), hjust = -0.25) +
-  geom_errorbar(aes(ymin=lower95, ymax=upper95), 
-                colour= "#46ACC8", width = 0.2) +
+plot <- ggplot(data=sameday_results_w_class, 
+  aes(x=outcome_names, y = odds_ratio, group = class, color = class)) +
+  geom_point() +
+  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
+  scale_color_manual(name = "Disease Group", values = c("#4ac29a", "#7303c0"), 
+    guide = guide_legend(direction="horizontal", title.position = "top",
+                         title.hjust = 0.5)) +
   geom_hline(yintercept = 1, linetype = 2, colour = "#B40F20") +
   theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    legend.position = "bottom",
     panel.grid.major = element_blank(),
     panel.grid.minor = element_blank(),
-   axis.text.x = element_text(angle=90, hjust = 1)) +
+    axis.text.x = element_text(angle=45, hjust = 1)) +
   ylab(expression("Same day odds ratio: 10 ug/m^3 increase smoke PM"[2.5])) +
-  xlab("Outcomes")
+  xlab(NULL) # removed x axis name, not sure i want to keep it that way
 
@@
 
-For this project, we use the time-stratified case-crossover study design in the same way we did for the 2012 Wenatchee-Complex fire project.
-Briefly, we identify admissions dates for ED or urgent care visits, and pharmacy fill dates for SABA, using billing records.
-
-In the Wenatchee-Complex fire project, we looked at the same-day relationship between smoke exposure and health outcomes.
-For the Douglas-complex fire, we will evaluate a 0 to 7 day distributed lag association, first comparing the overall cumulative effect over these 7 days, and then get in to the lagged effects.
-
 *Note: 11/7/17: Jingyang will provide a csv of the best-fit knots for the distributed lag splines based on AIC.*
 
-*Note: 11/7/17: Jingyang, can I use the lagged days in the case-crossover dataframes?
-Please confrim that the value for lag day 1 is the day before, lag day 2 is from 2 days before, etc..., and not the week before, two weeks before, etc...
-What does the _m denoter mean too?
-Jingyang says it's not necissary; we will remove creating this variable in the data management scripts.
-Do we use the patid or personkey identifier for strata?
-check.*
-
-Basis function also included for WRF temp. We will need to find the best fit of spline for temp seperately from PM. For asthma, it looks like 2 DF is best.
-
-Making a custom distributed lag function. Consider adding additional options like specifying number of lag days. That would require a reworking of how the case-crossovers are set up (kind of).
+We set up basis functions using natural splines for smoke exposure 0 to 7 days prior to admission/fill.We've applied 3 DF to smoke splines as I think it's flexible enough to allow varying shapes and fits well, but we are working on including best-fit by AIC in to models.
+Basis function also included for WRF temp; it looks like 2 DF is best.
 
@@
 
-*Note for Jingyang 11/8/17: purrr map functions are the tidyverse equivalents of apply-type functions*
-
 ```{r cumulative association}
+class <- as_tibble(c(rep("Respiratory", times = 6), 
+  rep("Cardiovascular", times = 6))) %>% 
+  rename(class = value)
 # use custom function to create a dataframe of cumulative values
@@
 
-# will put this table eventually and split out plot
-#head(cumulative_results)
+# add a class indicator to table
+cumulative_results_w_class <- cumulative_results %>% bind_cols(class)
+  cumulative_results_w_class
 
 # ggplot ----
-plot <- ggplot(cumulative_results, aes(x=outcome_names, y = estimate)) +
-  geom_point(colour = "#00A08A") +
-  geom_text(aes(label = round(estimate,2)), hjust = -0.25) +
-  geom_errorbar(aes(ymin=lower95, ymax=upper95), 
-                colour= "#00A08A", width = 0.2) +
+plot <- ggplot(cumulative_results_w_class, 
+    aes(x=outcome_names, y = estimate, group=class, color=class)) +
+  geom_point() +
+  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
+  scale_color_manual(name = "Disease Group", values = c("#4ac29a", "#7303c0"), 
+    guide = guide_legend(direction="horizontal", title.position = "top",
+                         title.hjust = 0.5)) +
   geom_hline(yintercept = 1, linetype = 2, colour = "#FF0000") +
   theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    legend.position = "bottom",
     panel.grid.major = element_blank(),
     panel.grid.minor = element_blank(),
-   axis.text.x = element_text(angle=90, hjust = 1)) +
+    axis.text.x = element_text(angle=45, hjust = 1)) +
   ylab(expression("Cumulative odds ratio: 10 ug/m^3 increase smoke PM"[2.5])) +
-  xlab("Outcomes")
+  xlab(NULL)
 
@@
 
-This is generally how I'd like the plots to look, but I plan to add color grouping to SABA, respiratory, and CVD groupings.
 Also considering reducing the lag time evaluated since I think we are getting in to variance issues a week out with this dataset (as evidence, large error around asthma outcome relative to other studies).
-
-I will also come back and tweak color scheme and labels.
+Alternatively, we could add the previous day 0, 0+1, 0+1+2,...
+and so on.
+Discuss with group.
 
@@
 # bind sex dataframes together
-sameday_results <- rbind(sameday_results_f, sameday_results_m)
+sameday_results <- rbind(sameday_results_f, sameday_results_m) %>% 
+  rename(odds_ratio = estimates)
 
@@
 
-Plot.
+Plot of sex-stratified results.
+
 ```{r sex strata plot}
-plot <- ggplot(data=sameday_results, aes(x=sex, y = estimates, color=sex)) +
-  scale_color_manual(name = "Sex", values = c("#00A08A", "#046C9A"),
+plot <- ggplot(data=sameday_results, aes(x=sex, y = odds_ratio, color=sex)) +
+  scale_color_manual(name = "Sex", values = c("#43c6ac", "#191654"),
     guide = guide_legend(direction="horizontal", title.position = "top",
@@
   ylab(expression("Odds ratio: 10 ug/m^3 increase smoke PM"[2.5])) +
-  xlab("Sex")
+  xlab(NULL)
 
@@
 ```
+
+No obvious sex differences in admission/fill due to increasing wildfire smoke exposure.
 
@@
 sameday_results <- rbind(sameday_results_under15, sameday_results_15to65,
-                         sameday_results_over65)
+                         sameday_results_over65) %>% 
+  rename(odds_ratio = estimates)
 
@@
 Stratified by age plot.
+
 ```{r age strata plot}
 plot <- ggplot(data=sameday_results, 
-               aes(x=age_cat, y = estimates, color=age_cat)) +
+               aes(x=age_cat, y = odds_ratio, color=age_cat)) +
   scale_color_manual(name = "Age Category", 
-    values = c("#899DA4", "#35274A", "#F2300F"),
+    values = c("#0082c8", "#4ac29a", "#6a3093"),
     guide = guide_legend(direction="horizontal", title.position = "top",
@@
   ylab(expression("Odds ratio: 10 ug/m^3 increase smoke PM"[2.5])) +
-  xlab("Age Category")
+  xlab(NULL)
 
@@
 
-Under 15 unstable for CVD; consider removing. Heart failure is interesting for the over 65 group. That may agree with what Anna Rappold has seen.
+Under 15 unstable for CVD; consider removing.
+Heart failure is interesting for the over 65 group.
+That may agree with what Anna Rappold has seen.
+May be some difference in SABA fills by age category, but probably not; rough evaluation says no as the CIs overlap.
 
@@
    scale_color_manual(name = "Underlying Diagnosis", 
-    values = c("#0B775E","#46ACC8","#E58601","#B40F20"),
+    values = c("#7303c0","#12d8fa", "#00467f", "#f80759"),
     guide = guide_legend(direction="horizontal", title.position = "top",
@@
 
+Interesting thing here is that COPD as underlying diagnosis is not nearly as pronounced as any of the other underlying diagnoses.
+Could be a sample size issue.
+We'll have to check.
+Perhaps there are few people with COPD only that are prescribed SABA.
+
 ##9. Stratified cumulative results
-Considering adding.
+Considering adding. Splines likely hard to fit for smaller groups.
 
 ##10. SABA cumulative effect by underlying respiratory diagnosis
-Considering adding.
+Considering adding. Splines likely hard to fit for smaller groups.
 

### 2017-11-17 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
updated some language in the markdown file.

@@
 # knitr options
-knitr::opts_chunk$set(fig.width=12, fig.height=10, fig.path='Figs/',
+knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='Figs/',
                       echo=FALSE, warning=FALSE, message=FALSE)
@@
 
-Most of the heavy smoke days were in the south-west part of Oregon.
-I've added county names to map but some text aren't well placed.
-I should probably use centroid, but then I'll have to import a shapefile to do this.
-I'll fix this later; I plan to add in fire locations.
+Most of the heavy smoke days were in the south-west part of Oregon. Triangles are fires; had them in the legend, but the legend looked weird so I took them out.
 
@@
 sameday_results_w_class <- sameday_results %>% bind_cols(class)
-  sameday_results_w_class
 
@@
 cumulative_results_w_class <- cumulative_results %>% bind_cols(class)
-  cumulative_results_w_class
 

### 2017-11-20 r_scripts/data_management/oregon_basic_data_wrangle.R
removing old scripts to rebuild repo.

@@ -1,112 +1,0 @@
-# ------------------------------------------------------------------------------
-# Title: Basic data wrangle and cleaning of Oregon 2013 claims data
-# Author: Ryan Gan
-# Date Created: December 30, 2016
-# R Version: 3.3.2
-# ------------------------------------------------------------------------------
-
-# Note: Datasets are pretty large. I had to come up with some different ways of
-# processing these data. I checked the number of rows in the command line and
-# found there were 77,069,321 rows. 
-
-# load libraries ---------------------------------------------------------------
-library(dplyr)
-library(data.table)
-
-# import data ------------------------------------------------------------------
-
-# Read in first 100000 lines for exploration.
-
-#read_path <- paste0("./data/gan_episodes_of_care.txt")
-#oregon_df <- fread(read_path, sep = "|", nrows = 200000, showProgress = T)
-#fwrite(oregon_df, "./data/oregon_epis_care_reduced.csv")
-
-# I wrote a reduced dataframe of 200,000 rows to get an idea of data structure
-
-# import the reduced dataframe
-start_time <- Sys.time()
-
-oregon_df <- fread("./data/oregon_epis_care_reduced.csv", sep = ",", 
-                   colClasses = rep("character", 72))
-
-stop_time <-  Sys.time() - start_time 
-# time it took
-stop_time
-
-#oregon_df[1:10, 64]
-
-# Define read paths ------------ -----------------------------------------------
-# import r list of outcome-specific icd9 codes
-load_path <- paste0("./data/outcome_list.RData")
-load(load_path)
-
-# path and file name of data
-read_path <- paste0("./data/oregon_epis_care_reduced.csv")
-
-# now try to work the two chunks together to identify records that contain icd9s
-# of a certain outcome and then output those rows in chunks of data.
-
-# I need a dataframe of start row and length to read after
-n_start<- seq(from = 1, to = 200000, by = 10000) 
-
-#n_start<- seq(from = 1, to = 100, by = 5)   
-# create dataframe
-n_read <- data_frame(n_start) %>% # turn the sequence in to a dataframe
-  mutate(n_length = 10000) # set length to read 
-
-# read first line and get column names
-df_col_names <- fread(read_path, nrows = 0, header = T, 
-                      colClasses = rep("character", 72))
-
[…]

### 2017-11-20 r_scripts/data_management/oregon_casecrossover.R

@@ -1,214 +1,0 @@
-# ------------------------------------------------------------------------------
-# Title: Case crossover for diseases (Oregon State)
-# Author: Jingyang Liu
-# Date Created: Feb 24, 2017
-# R version: 3.3.2
-# ------------------------------------------------------------------------------
-
-#Creating case crossover dataframes
-library(tidyverse)
-library(data.table)
-library(readxl)
-getwd()
-setwd("C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire/data_disease")
-
-### ---------------------------------------------------------------------
-var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
-              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
-              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
-
-start <- Sys.time()
-for (i in var_list){
-  read_path <- paste('oregon', i, 'claims_2013.csv', sep='_')
-  or_disease <- read_csv(read_path)
-    
-  or_disease <- or_disease %>% 
-    group_by(personkey) %>% 
-    arrange(personkey, fromdate, line) %>% 
-    mutate(num_visit = dense_rank(fromdate)) %>%
-    arrange(personkey, num_visit) %>% 
-    select(personkey, clmid, num_visit) %>% 
-    unique() %>% 
-    full_join(or_disease, by = c("personkey", "clmid")) %>% 
-    arrange(personkey, clmid, line, fromdate) %>% 
-    filter(row_number() == 1) %>%
-    # add new transverted from date
-    mutate(dates = as.Date(fromdate, "%m/%d/%Y")) %>%
-    filter(dates >= '2013-05-01' & 
-           dates <= '2013-09-30') 
-  
-  # Create a permanent case-cross over dataset
-  file_name <- paste('oregon', i, 'may_to_sep_claim.csv', sep = '_')
-  
-  # write permanent dataset
-  write_csv(or_disease, paste0("../data_new/", file_name))
-  
-}
-total_time <- Sys.time() - start
-total_time # Time difference of 5.154874 mins
-
-setwd("../data_new/update")
-
-
-
-## Case crossover for each disease
-
-start <- Sys.time()
[…]

### 2017-11-20 r_scripts/data_management/oregon_casecrossover_server.R

@@ -1,105 +1,0 @@
-# ------------------------------------------------------------------------------
-# Title: Case crossover for diseases (Oregon State) for server
-# Author: Jingyang Liu
-# Date Created: May 3, 2017
-# R version: 3.3.2
-# ------------------------------------------------------------------------------
-
-## Case crossover for each disease
-## We can get the unique personkey with the first visit from data_new/casecrossover/ file
-
-# library(tidyverse)
-library(data.table)
-library(readr)
-library(dplyr)
-
-getwd() # [1] "/home/jyliu/wildfire/local_git_repo/oregon_wildfire/r_scripts"
-setwd("../../../data/data_new/casecrossover")
-
-var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
-              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
-              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
-
-
-start <- Sys.time()
-for(m in var_list){ # begin first loop of variable names (outcomes)
-
-  read_path <- paste('oregon', m, 'may_to_sep_claim.csv', sep='_')
-  disease <- read_csv( read_path)
-  
-  outcome_id <- disease %>%
-    # arrange with dates
-    arrange(dates) %>%
-    mutate(id = seq(1, nrow(.), by = 1))  # create subject id
-   
-  # create dataset to populate
-  id_date_df <- data_frame()
-  
-  
-  # begin second loop to create counterfactual observations for each case subject
-  for (k in 1:nrow(outcome_id)){
-    
-    # find the replicate times of weeks
-    dates_l <- outcome_id[[k,74]] 
-    n1 <- 0
-    d=as.Date("2013-05-01")
-    i=1
-    while (dates_l >= "2013-05-01"){
-      dates_l <- dates_l - 7
-      d[i] = dates_l
-      i = i+1
-      n1 = n1+1
-    }
-    d[1:n1-1] # shows character(0) when the first week
-    n1-1
-    
-    dates_l <- outcome_id[[k,74]] 
-    n2=0
-    e=as.Date("2013-09-30")
-    j=1
-    while (dates_l <= "2013-09-30"){
-      dates_l <- dates_l + 7
-      e[j]=dates_l
-      j=j+1
-      n2 = n2 + 1
-    }
-    e[1:n2-1] # shows character(0) when the last week
[…]

### 2017-11-20 r_scripts/data_management/oregon_data_wrangle_separate_server.R

@@ -1,106 +1,0 @@
-# ------------------------------------------------------------------------------
-# Title: Basic data wrangling  for 2013 oregon data
-# Author: Jingyang Liu
-# Date Created: Mar 03, 2016
-# R Version: 3.3.2
-# ------------------------------------------------------------------------------
-
-library(dplyr)
-library(data.table)
-library(readxl) 
-library(readr)
-
-# import data ------------------------------------------------------------------
-
-read_path <- paste0("../../../data/data_original/gan_episodes_of_care.txt")
-start_time <- Sys.time()
-oregon_df <- fread(read_path, sep = "|", showProgress = T) 
-stop_time <-  Sys.time() - start_time 
-# time it took
-stop_time # 7 mins
-
-# basic cleaning ---------------------------------------------------------------
-start_time <- Sys.time()
-oregon_df <- oregon_df %>% filter(!is.na(dx1)) %>% 
-  # filter State is Oregon
-  filter(STATE=="OR") %>%
-  # filter urgent or ED visits (pos = 20 or 23)
-  filter(pos=="20"|pos=="23") 
-stop_time <-  Sys.time() - start_time 
-# time it took
-stop_time # 2.5 mins
-
-write_path <- paste0('../../../data/data_original/',
-                     'oregon_epis_care_reduced.csv')
-write_csv(oregon_df, write_path)
-
-
-
-# Define read paths ------------ -----------------------------------------------
-# import r list of outcome-specific icd9 codes
-# 1514642 obs
-load_path <- paste0("../../../data/data_original/outcome_list.RData")
-load(load_path)
-
-# path and file name of data
-read_path <- paste0("../../../data/data_original/oregon_epis_care_reduced.csv")
-start_time <- Sys.time()
-oregon_df <- fread(read_path, sep = ",", 
-                   colClasses = rep("character", 72)) 
-stop_time <-  Sys.time() - start_time 
-# time it took
-stop_time # 14 sec
-
-
-
-# now try to work the two chunks together to identify records that contain icd9s
-# of a certain outcome and then output those rows in chunks of data.
-
-# I need a dataframe of start row and length to read after
-n_start<- seq(from = 1, to = length(oregon_df$personkey), by = 10000) 
-
-# create dataframe
-n_read <- data_frame(n_start) %>% # turn the sequence in to a dataframe
-  mutate(n_length = ifelse(row_number()!=n(), 10000,
[…]

### 2017-11-20 r_scripts/data_management/outcome_icd9_codes.R

@@ -1,77 +1,0 @@
-# ------------------------------------------------------------------------------
-# Title: Basic data wrangle and cleaning of Oregon 2013 claims data
-# Author: Ryan Gan
-# Date Created: December 30, 2016
-# R Version: 3.3.2
-# ------------------------------------------------------------------------------
-
-# The purpose of this code is to easily group the ICD-9 codes that make up
-# a specific outcome of interest using the ICD-9-CM version 32 list
-
-# Data files can be found here: 
-# https://www.cms.gov/Medicare/Coding/ICD9ProviderDiagnosticCodes/codes.html
-
-# load library -----------------------------------------------------------------
-library(tidyverse)
-library(readxl) # read excel files
-
-
-# read in icd9 data and create lists of vectors of icd9 codes ----
-# read in xlsx icd9 codes and description
-icd9_df <- read_excel("./data/CMS32_DESC_LONG_SHORT_DX.xlsx") %>% 
-  # rename the terrible variable names
-  select(dx_code = 1, long_desc = 2, short_desc = 3)
-
-# look at the first 6 rows of the icd9 codes; comes in sorted
-head(icd9_df)
-glimpse(icd9_df)
-
-# Overall goal is to make a list of vectors for each outcome that contains the
-# character strings of icd9 codes that make up the outcome of interest.
-# Some outcomes like COPD, where icd9 codes are not sequential, will not work
-# and will need to be entered by hand.
-
-# Create a dataframe of the outcome name, start, and stop rows
-outcome_name <- c("respiratory", "asthma", "pneumonia", "acute_bronch",
-                  "cvd", "arrhythmia", "cerbrovas_dis", "heart_failure",
-                  "isch_heart_dis", "myocardial_infarc", "broken_arm")
-# first icd9 code of outcome
-start_icd9 <- c("460", "49300", "4800", "4660", "390", "4270", "430", "4280",
-                "41000", "41000", "81300")
-
-# last icd9 code of outcome
-stop_icd9 <- c("5199", "49392", "486", "46619", "4599", "4279", "4389", "4289",
-               "4139", "41092", "81393")
-
-# dataframe of outcomes and start stop of icd9 codes
-outcome_icd9_range <- data_frame(outcome_name, start_icd9, stop_icd9)
-
-# list of icd9 codes -----------------------------------------------------------
[…]

### 2017-11-20 r_scripts/smoke/or_county_pop_wt_estimate_2013.R

@@ -1,354 +1,0 @@
-#-------------------------------------------------------------------------------
-#     Title: Oregon population-weighted smoke PM2.5 by day and county
-#     Author: Jingyang Liu                                                                
-#     Date Created: Apr 2017   Date Modified: May 02, 2017                                                 
-#     R version: 3.3.3                                                       
-#-------------------------------------------------------------------------------
-
-# Note: This is general code that could be submitted in batches to handle 
-#       multiple imputs (dataframes)
-
-# Note 8/8/16: I think the matrix approach actually runs pretty quick; doesn't
-#              need to be submitted in batches
-
-# load libraries ---------------------------------------------------------------
-library(tidyverse)
-library(data.table)
-
-# Setting Working Directory ----------------------------------------------------
-dir <- paste0("../data/Oregon_PM/")
-setwd(dir)
-getwd()
-list.files()
-
-# Import and prepare smoke PM 2.5 data for loop --------------------------------
-
-# These files never change and can take place outside the loop
-# Import population weight csv 
-pop_path <- paste0('./popdensity_oregon.csv')
-pop_grid <- read_csv(pop_path)
-
-# rename missing column
-colnames(pop_grid) <- c('WRFGRID_ID', 'Longitude', 'Latitude', 'pop_density')
-summary(pop_grid)
-
-# calculate the estimated population for each grid (multiply density by grid area)
-population_grid <- pop_grid %>% 
-  mutate(pop_n = pop_density *(15^2)) ##??
-summary(population_grid)
-
-
-# Import the file that contains the proportion of each grid that overlaps a county
-county_grid_path <- paste0('../../data_new/county_data/or_county_wrf_prop.csv')
-county_grid_proportion <- read_csv(county_grid_path)
-
-
-# Importing files that I would like to loop through
-# Geo weighted regresssion (Atmos model of choice)
-geo_wt_path <- paste0("oregon_gwrPM_2013.csv")
-geo_wt <- read_csv(geo_wt_path)
-
-# Kriging 
-krig_path <- paste0("oregon_krigedPM_2013.csv")
-krig <- read_csv(krig_path)
-
-# Background PM2.5
[…]

### 2017-11-20 r_scripts/smoke/or_smoke_jul_oct_time_stratified_case_crossover.R

@@ -1,182 +1,0 @@
-# ------------------------------------------------------------------------------
-# Title: Oregon Case-crossover dataframe creation (only patients with single obs)
-#        and joining with and zip-level PM2.5 values. 
-# Author: Jingyang Liu
-# Date Created: 04/05/2017     
-# R version: 3.3.3
-# ------------------------------------------------------------------------------
-
-# Notes ----
-# Notes 8/16/16: I updated smoke data with population weighted averages for zip
-# This code also assumed only one visit, so I remove the people with multiple
-# admissions for the outcome
-
-# Notes 10/3/16: I'm retaining the county assignment of the observation so I
-# can join the county population weighted averages as well for comparison
-# and to analyze morbidity similar to Rish
-
-# Libraries ----
-# load libraries
-#library(tidyverse) # tidyverse not on the server
-# if tidyverse is not available, use dplyr and readr
-library(dplyr) 
-library(readr)
-library(lubridate) # working with date
-# parallel computing libraries
-# library(foreach) 
-# library(doParallel)
-
-# Read in smoke data created in loop -------------------------------------------
-
-# read in zipcode level populatoin-weighted pm
-read_path <- paste0('C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire_new/data/Oregon_PM/zip_pm_to_merge_with_acap.csv')
-
-zip_smoke <- read_csv(read_path) # 63801 rows
-
-# descriptives of the two smoke datasets
-summary(zip_smoke)
-
-# Zipcode PM2.5 estimates
-# create lag variables that take smoke values from n previous days for zipcodes
-zip_smoke_w_lag <- zip_smoke %>% arrange(ZIPCODE, date) %>%
-  # group by zipcode
-  group_by(ZIPCODE) %>% 
-  # wrf
-  mutate(wrf_f_pm_lag1 = lag(wrf_f_pm, 1, order_by = ZIPCODE), 
-         wrf_f_pm_lag2 = lag(wrf_f_pm, 2, order_by = ZIPCODE),
-         wrf_f_pm_lag3 = lag(wrf_f_pm, 3, order_by = ZIPCODE),
[…]

### 2017-11-20 r_scripts/smoke/oregon_proportion_county_wrf.R

@@ -1,146 +1,0 @@
-# ------------------------------------------------------------------------------
-# Title: Oregon dataframe of proportion of each county in each WRF-Grid cell 
-# Author: Jingyang Liu
-# Date Created: Apr 27, 2017
-# R version: 3.3.2
-# ------------------------------------------------------------------------------
-
-library(rgdal) # package for shape files
-library(sp)
-library(rgeos) # rgeos package contains the intersect and area commands I need
-library(dplyr)
-# library(data.table)
-library(readr)
-
-setwd("C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire_new/")
-
-### First step: join the zipcode of health data with county
-
-colnames(or_zip_county) <- c("ZIP", "county")
-
-read_path <- paste0('./instructions/oregon_zip_county.csv')
-or_zip_county <- read_csv(read_path)
-
-# or_zip_county$county_name[which(or_zip_county$county_name=="Hood.River")] <- "Hood River"
-
-or_zip_county <- or_zip_county %>%
-  select(zip, county_name) %>%
-  rename(county = county_name) %>%
-  rename(ZIP = zip) %>%
-  select(ZIP, county) 
-
-county_new <- c("Washington", "Douglas", "Lane", "Deschutes")
-
-zip_new <- data.frame(c(97003, 97471, 97475, 97703))
-colnames(zip_new) <- "ZIP"
-
-zip_new <- zip_new %>%
-  mutate(county = county_new)
-
-or_zip_county_new <- or_zip_county %>%
-  bind_rows(zip_new) # 488 zip codes
-
-grid_dir <- paste0('./shapefile/oregon_new_grid/oregon_new_grid.shp')
-smoke_grid <- readOGR(dsn = grid_dir, layer = 'oregon_new_grid') 
-
-county_dir <- paste0('./shapefile/tl_2013_us_county/tl_2013_us_county.shp')
-us_county_2013 <- readOGR(dsn = county_dir, layer = 'tl_2013_us_county')
-or_county_2013 <- us_county_2013[us_county_2013$STATEFP =="41",]
-
-nad83 <- '+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0'
-proj4string(or_county_2013) <- nad83
-
-plot(or_county_2013,border="red")
-plot(smoke_grid, add=T)
-
-### Choose one sample for testing ----------------------------------------------
[…]

### 2017-11-20 r_scripts/smoke/oregon_proportion_zip_server.R

@@ -1,304 +1,0 @@
-# ------------------------------------------------------------------------------
-# Title: Oregon dataframe of proportion of each zipcode in each WRF-Grid cell 
-# Author: Jingyang Liu
-# Date Created: Mar 10, 2017
-# R version: 3.3.2
-# ------------------------------------------------------------------------------
-
-library(rgdal) # package for shape files
-library(sp)
-library(rgeos) # rgeos package contains the intersect and area commands I need
-library(dplyr)
-library(data.table)
-library(readxl)
-# parallel computing libraries
-# library(foreach) 
-library(doParallel)
-library(parallel)
-library(readr)
-
-getwd()
-# "/home/jyliu/wildfire/local_git_repo/oregon_wildfire/r_scripts"
-
-# Import Shapefiles  -----------------------------------------------------------
-# WRF Grid
-grid_dir <- paste0('../../../data/data_original/oregon_new_grid.shp')
-
-smoke_grid <- readOGR(dsn = grid_dir, layer = 'oregon_new_grid') 
-
-# summary(smoke_grid) 
-# summary(smoke_grid@data$WRFGRID_ID)
-
-# Zipcode shapefile
-shp_dir <- paste0('../../../data/data_original/', 
-                 'tl_2013_us_zcta510/tl_2013_us_zcta510.shp')
-
-us_zip_2013 <- readOGR(dsn = shp_dir, layer = 'tl_2013_us_zcta510')
-summary(us_zip_2013)
-# plot(us_zip_2013)
-
-
-### whole oregon zip code
-read_path <- paste0('../../../data/data_original/oregon_zipcode.csv')
-or_zip <- read_csv(read_path)
-names(or_zip) <- c('zip','type','city','county','area')
-
-# limit to just Oregon state zipcodes
-or_zip_map <- us_zip_2013[us_zip_2013$ZCTA5CE10 %in% or_zip$zip,]
-
-# output zipcodes from washington zipcode map to bind values to
-# plot map to check
-plot(or_zip_map)
-
-# saving wash shapefile ----
-# save shapefile to use in future work
-summary(or_zip_map)
-or_zip <- as.character(sort(or_zip_map@data$ZCTA5CE10))
-
-or_zip_df <- data.frame(or_zip)
-
-write_path <- paste0('../../../data/data_new/',
-                     'or_zip.csv')
[…]

### 2017-11-20 r_scripts/analysis/oregon_cvd_dl_over65_results.Rmd
started analysis looking at cvd dl outcomes in the over 65 group.

---
title: Distributed lag effects of smoke on cardiovascular disease in persons over 65
author: "Ryan Gan"
date: "2017-11-20"
output: html_document
---

### 2017-11-21 r_scripts/analysis/oregon_cvd_dl_over65_results.Rmd
worked on over 65 cvd outcomes; may incorporate in to overall results.

@@
 ---
+
+## Overview
+
+Extensive look at possible distributed lag relationships for Cardiovascular endpoints in the over 65 age group.
+
+Packages used: tidyverse, survival. Setting markdown options.
+
+```{r setup, echo =F, warning=F, message=F}
+library(tidyverse) # general data wrangle
+library(survival) # conditional logistic regression
+
+# knitr options
+knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='Figs/',
+                      echo=FALSE, warning=FALSE, message=FALSE)
+```
+
+Import CVD data.
+
+```{r health data import}
+# path for health dataframe
+health_path <- "../../data/health/"
+# importing time-stratified case-crossover dataframe
+casecross_files <- list.files(path=health_path, pattern="*casecross*")
+
+# reorder file names in way I'd like it to be analyzed and only importing cvd
+# also ditching broken arm
+casecross_order <- casecross_files[c(7,2,5,8,9,10)]
+
+# extract the name of the dataframe after study type
+file_names <- stringr::str_extract(casecross_order, 
+  pattern="(?<=casecross_)\\w+")
+
+# read in all the files in to a list
+outcome_df_list <- casecross_order %>%  
+  purrr::map(~read_csv(paste0(health_path, .)))
+
+# name dataframes
+names(outcome_df_list) <- file_names
+
+# name list elements in order I'd like to present in document
+outcome_names <- c("Cardiovascular Disease", "Arrhythmia", 
+  "Cerebrovascular Disease", "Heart Failure", "Ischemic Heart Disease", 
+  "Myocardial Infarction")
+```
+
+```{r distributed lag function}
+# custom distributed lag function 
+distributed_lag <- function(data, pm_df, temp_df, cumulative=T){
+  # i want to remove this part eventually ------
+  # limit to complete cases
+  complete_data <- data %>% 
+    # filter to complete case for smoke
+    filter(!is.na(geo_smk_pm_lag7_zip))
+
+  # output matrix of gwr values
+  pm_matrix <- complete_data %>% 
+    select(contains("geo_smk_pm")) %>%
+    # remove anything with m in it; will go back in original code and not create 
+    # this to begin with
+    select(-contains("_m")) %>% 
+    # divide exposure values by 10 units to interpret on 10 ug/m^3 scale
+    mutate_all(funs(./10)) %>% 
+    # convert to matrix
+    as.matrix()
+
+  # output temp matrix
+  temp_matrix <- complete_data %>% 
+    select(contains("wrf_temp")) %>%
+    # remove anything with m in it; will go back in original code and not create 
+    # this to begin with
+    select(-contains("_m")) %>% 
+    # convert to matrix
+    as.matrix()
+  # part to remove------
+
+  # calculation of basis ----
+  # define basis using natural spline function from "splines"" package
+  pm_b <- splines::ns(0:(ncol(pm_matrix)-1), df=pm_df, intercept=T)
+  # create pm basis
+  pm_basis <- pm_matrix %*% pm_b
+  # create temp basis
+  temp_b <- splines::ns(0:(ncol(temp_matrix)-1), df=temp_df, intercept=T)
+  temp_basis <- temp_matrix %*% temp_b
+
+  # fit model with basis ----
+  mod <- clogit(outcome ~ pm_basis + temp_basis + strata(personkey),
+    data = complete_data)
+
+  # section to figure out model fit
+  # 3 df
+  # AIC(mod)
+
+  # calculate estimates ----
+  # output pm basis parameters
+  dl_parms <- broom::tidy(mod) %>% 
+    filter(stringr::str_detect(term, "pm")) %>% 
+    select(estimate) %>% 
+    as_vector()
+  # estimate distributed lag values for each day
+  estimate <-  pm_b %*% dl_parms
+  # covariance matrix for knots 
+  # fix the matrix
+  cov_matrix <- as.matrix(vcov(mod))[1:pm_df,1:pm_df]
+
+  # estimate variance of spline
+  variance <- pm_b %*% cov_matrix %*% t(pm_b)
+  # estimate standard error
+  se <- sqrt(diag(dl_var))
+
+  # calculate lower and upper bound
+  lower95 <- estimate+(se*qnorm(1-0.975))
+  upper95 <- estimate+(se*qnorm(0.975))
+
+  # time variable
+  time <- ((rep(1:length(estimate))-1))
+
+    if(cumulative==T) {
+      type <- rep("cumulative", times = length(estimate))
+      # sequential cumulative estimate
+      cumulative_estimate <- sapply(seq_along(estimate), function(x){
+        sum(estimate[1:x])
+      })
+
+      # stderr cumulative effect
+      cumulative_se <- sapply(seq_along(estimate), function(x){
+        sqrt(sum(dl_var[1:x,1:x]))
+      })
+
+      # cumulative 95CI
+      lower95 <- cumulative_estimate+(cumulative_se*qnorm(1-0.975))
+      upper95 <- cumulative_estimate+(cumulative_se*qnorm(0.975))
+
+      # return dataframe
+      return_estimate <- data.frame(type, time, exp(cumulative_estimate), 
+                                    exp(lower95), exp(upper95)) 
+      # assign column names
+      colnames(return_estimate) <- c("type","time","estimate","lower95","upper95")
+
+      return(return_estimate)
+    }
+    else { # if cumulative is not true, or false, return distributed lag est
+      type <- rep("lag", times = length(estimate))
+
+      # return dataframe
+      return_estimate <- data.frame(type, time, exp(estimate), 
+                                    exp(lower95), exp(upper95)) 
+      # assign column names
+      colnames(return_estimate) <- c("type","time","estimate","lower95","upper95") 
+       # return estimate  
+      return(return_estimate)
+      }
+      # end of if
+} 
+```
+
+## Heart failure in over 65 group
+
+```{r subset data}
+hf65 <- outcome_df_list[[4]] %>% 
+  filter(age_ind == 2)
+
+cumulative <- distributed_lag(hf65, pm_df = 3, temp_df = 2, cumulative = T)
+cumulative
+
+#plot 
+cumulative_plot <- ggplot(cumulative, aes(x=time, y=estimate)) +
+  geom_line(colour = "#046C9A", size = 1) +
+  geom_ribbon(aes(ymin = lower95, ymax = upper95), 
+              fill = "#ABDDDE", alpha = 0.5) + 
+  scale_y_continuous(limits = c(0.9, 1.25)) +
+  scale_x_continuous(breaks = c(seq(0,7, by=1))) +
+  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
+  ylab(expression("Odd ratio for a 10 ug/m^3 increase smoke PM"[2.5])) +
+  xlab("Lagged Days") +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    strip.text = element_text(size = 10),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    strip.background = element_rect(colour=NA, fill=NA))
+
+cumulative_plot
+
+dl <- distributed_lag(hf65, pm_df = 3, temp_df = 2, cumulative = F)
+dl
+
+dl_plot <- ggplot(dl, aes(x=time, y=estimate)) +
+  geom_line(colour = "#046C9A", size = 1) +
+  geom_ribbon(aes(ymin = lower95, ymax = upper95), 
+              fill = "#ABDDDE", alpha = 0.5) + 
+  scale_y_continuous(limits = c(0.9, 1.25)) +
+  scale_x_continuous(breaks = c(seq(0,7, by=1))) +
+  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
+  ylab(expression("Odd ratio for a 10 ug/m^3 increase smoke PM"[2.5])) +
+  xlab("Lagged Days") +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    strip.text = element_text(size = 10),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    strip.background = element_rect(colour=NA, fill=NA))
+
+dl_plot
+
+```
+
+## Myocardial infarction in over 65 group

### 2017-11-21 r_scripts/analysis/oregon_cvd_dl_over65_results.Rmd
adding function to find spline of best fit.

@@
 
+  # if degrees of freedom for pm is defined
+  if(length(pm_df)==1){
+
   # calculation of basis ----
@@
     data = complete_data)
+
+  }
+
+  # if degrees of freedom is a vector > 1
+  pm_df <- c(2:5)
+
+  if(pm_df > 1){
+  # create temp basis
+    temp_b <- splines::ns(0:(ncol(temp_matrix)-1), df=temp_df, intercept=T)
+    temp_basis <- temp_matrix %*% temp_b
+  # create list of lengh pm_df
+  mod_aic <- sapply(pm_df, function(x){
+    pm_b <- splines::ns(0:(ncol(pm_matrix)-1), df=x, intercept=T)
+    # create pm basis
+    pm_basis <- pm_matrix %*% pm_b
+    # run model
+    mod <- clogit(outcome ~ pm_basis + temp_basis + strata(personkey),
+    data = complete_data)
+    aic <- AIC(mod)
+    return(aic)
+  })
+
+  # find minimum aic
+  pm_df_aic <- data.frame(pm_df, mod_aic)
+  min_df_aic <- as.numeric(subset(pm_df_aic, mod_aic == min(mod_aic), 
+                                  select=pm_df))
+  # run final best fit mod
+  pm_b <- splines::ns(0:(ncol(pm_matrix)-1), df=min_df_aic, intercept=T)
+  # create pm basis
+  pm_basis <- pm_matrix %*% pm_b
+  # run model
+  mod <- clogit(outcome ~ pm_basis + temp_basis + strata(personkey),
+    data = complete_data)
+  }
 
@@
 ## Myocardial infarction in over 65 group
+
+```{r mi data}
+mi65 <- outcome_df_list[[6]] %>% 
+  filter(age_ind == 2)
+
+cumulative <- distributed_lag(mi65, pm_df = 2, temp_df = 2, cumulative = T)
+cumulative
+
+#plot 
+cumulative_plot <- ggplot(cumulative, aes(x=time, y=estimate)) +
+  geom_line(colour = "#046C9A", size = 1) +
+  geom_ribbon(aes(ymin = lower95, ymax = upper95), 
+              fill = "#ABDDDE", alpha = 0.5) + 
+  scale_y_continuous(limits = c(0.7, 1.2)) +
+  scale_x_continuous(breaks = c(seq(0,7, by=1))) +
+  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
+  ylab(expression("Odd ratio for a 10 ug/m^3 increase smoke PM"[2.5])) +
+  xlab("Lagged Days") +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    strip.text = element_text(size = 10),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    strip.background = element_rect(colour=NA, fill=NA))
+
+cumulative_plot
+
+dl <- distributed_lag(mi65, pm_df = 3, temp_df = 2, cumulative = F)
+dl
+
+dl_plot <- ggplot(dl, aes(x=time, y=estimate)) +
+  geom_line(colour = "#046C9A", size = 1) +
+  geom_ribbon(aes(ymin = lower95, ymax = upper95), 
+              fill = "#ABDDDE", alpha = 0.5) + 
+  scale_y_continuous(limits = c(0.9, 1.25)) +
+  scale_x_continuous(breaks = c(seq(0,7, by=1))) +
+  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
+  ylab(expression("Odd ratio for a 10 ug/m^3 increase smoke PM"[2.5])) +
+  xlab("Lagged Days") +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    strip.text = element_text(size = 10),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    strip.background = element_rect(colour=NA, fill=NA))
+
+dl_plot
+
+```

### 2017-11-22 r_scripts/data_wrangle/2013_oregon_data_wrangle.R
filter original data; separate dx data set and saba data set

@@ -1,0 +1,146 @@
+# ------------------------------------------------------------------------------
+# Title: Basic data wrangling  for 2013 Oregon data and seperating by ICD9 lists
+# Author: Jingyang Liu
+# Date Created: Mar 03, 2016
+# Date Checked: Nov 20, 2017
+# R Version: 3.3.2
+# ------------------------------------------------------------------------------
+
+# This file is for 
+# Also it separates the data into 12 disease files due to the ICD9 lists.
+
+# This runs on server.
+
+# library
+library(dplyr)
+library(data.table)
+library(readxl) 
+library(readr)
+
+# import original data ---------------------------------------------------------
+read_path <- paste0("../../../data/data_original/gan_episodes_of_care.txt")
+start_time <- Sys.time()
+oregon_df <- fread(read_path, sep = "|", showProgress = T) 
+stop_time <-  Sys.time() - start_time 
+# time it took
+stop_time # 7 mins
+
+# basic cleaning ---------------------------------------------------------------
+
+oregon_df <- oregon_df %>% 
+  # filter(!is.na(dx1)) %>% 
+  # filter State is Oregon
+  filter(STATE=="OR") 
+# filter urgent or ED visits (pos = 20 or 23)
+# filter(pos=="20"|pos=="23") 
+
+write_path <- paste0('../../../data/data_original/',
+                     '2013_oregon_epis_care.csv')
+write_csv(oregon_df, write_path)
+
+
+
+# Define read paths ------------ -----------------------------------------------
+# import r list of outcome-specific icd9 codes
+load_path <- paste0("../../../data/data_original/outcome_list.RData")
+load(load_path)
+
+# path and file name of data
+read_path <- paste0("../../../data/data_original/2013_oregon_epis_care.csv")
+start_time <- Sys.time()
+oregon_df <- fread(read_path, sep = ",", 
+                   colClasses = rep("character", 72)) 
+stop_time <-  Sys.time() - start_time 
+# time it took
+stop_time # 20 min
+
+
+## make SABA data set(ndc is beta-2-agonist; personkey) ------------------------
+read_path2 <- paste0("../../../data/data_original/2013_ndc_beta2_agonists.csv")
+ndc_df <- read_csv(read_path2, col_names = FALSE)
+
+colnames(ndc_df) <- c("ndc_code", "brand_name", "generic_product_name", 
+                      "route", "category", "drug_id")
+
+# convert to vector
[…]

### 2017-11-22 r_scripts/analysis/oregon_cvd_dl_over65_results.Rmd
created a fit function to find AIC of degrees of freedom for spline fit. Next step is to store this information in the list of dataframes.

--- 2017-11-21 r_scripts/analysis/oregon_cvd_dl_over65_results.Rmd
@@
 
-```{r distributed lag function}
+Creating a fit function to find the degrees of freedom for the pm spline
+
+```{r distributed lag functions}
+
+# fit function
+dl_fit <- function(data, pm_df, temp_df){
+  # i want to remove this part eventually ------
+  # limit to complete cases
+  complete_data <- data %>% 
+    # filter to complete case for smoke
+    filter(!is.na(geo_smk_pm_lag7_zip))
+
+  # output matrix of gwr values
+  pm_matrix <- complete_data %>% 
+    select(contains("geo_smk_pm")) %>%
+    # remove anything with m in it; will go back in original code and not create 
+    # this to begin with
+    select(-contains("_m")) %>% 
+    # divide exposure values by 10 units to interpret on 10 ug/m^3 scale
+    mutate_all(funs(./10)) %>% 
+    # convert to matrix
+    as.matrix()
+
+  # output temp matrix
+  temp_matrix <- complete_data %>% 
+    select(contains("wrf_temp")) %>%
+    # remove anything with m in it; will go back in original code and not create 
+    # this to begin with
+    select(-contains("_m")) %>% 
+    # convert to matrix
+    as.matrix()
+  # end part to remove------
+
+    # create temp basis
+    temp_b <- splines::ns(0:(ncol(temp_matrix)-1), df=temp_df, intercept=T)
+    temp_basis <- temp_matrix %*% temp_b
+
+      # create list of lengh pm_df
+      mod_aic <- sapply(pm_df, function(x){
+        pm_b <- splines::ns(0:(ncol(pm_matrix)-1), df=x, intercept=T)
+        # create pm basis
+        pm_basis <- pm_matrix %*% pm_b
+        # run model
+        mod <- clogit(outcome ~ pm_basis + temp_basis + strata(personkey),
+          data = complete_data)
+        aic <- AIC(mod)
+        return(aic)
+      }) # end sapply
+
+    # find minimum aic
+    pm_df_aic <- data.frame(pm_df, mod_aic)
+    min_df_aic <- subset(pm_df_aic, mod_aic == min(mod_aic))
+    return(min_df_aic)
+  } # end fit option
+
+dl_fit(data = hf65, pm_df = c(3:5), temp_df=2)
+
 # custom distributed lag function 
@@
     as.matrix()
-  # part to remove------
+  # end part to remove------
 
@@
   temp_basis <- temp_matrix %*% temp_b
-
   # fit model with basis ----
@@
     data = complete_data)
-
-  # section to figure out model fit
-  # 3 df
-  # AIC(mod)
 
@@
   cov_matrix <- as.matrix(vcov(mod))[1:pm_df,1:pm_df]
-
   # estimate variance of spline
@@
   se <- sqrt(diag(dl_var))
-
   # calculate lower and upper bound
@@
   upper95 <- estimate+(se*qnorm(0.975))
-
   # time variable
@@
       })
-
       # stderr cumulative effect
@@
       })
-
       # cumulative 95CI
@@
       upper95 <- cumulative_estimate+(cumulative_se*qnorm(0.975))
-
       # return dataframe
@@
       return(return_estimate)
-    }
-    else { # if cumulative is not true, or false, return distributed lag est
+    } else { # if cumulative is not true, or false, return distributed lag est
     type <- rep("lag", times = length(estimate))
-
     # return dataframe
@@
     return(return_estimate)
-      }
-      # end of if
+    } # end of if
 } 
@@
 
-cumulative <- distributed_lag(hf65, pm_df = 3, temp_df = 2, cumulative = T)
+cumulative <- distributed_lag(hf65, pm_df = 4, temp_df = 2, cumulative = T)
 cumulative

### 2017-11-28 r_scripts/data_wrangle/2013_oregon_diagnosis_separate.R
separate data with ICD9 codes, and make saba data set

@@ -1,0 +1,147 @@
+# ------------------------------------------------------------------------------
+# Title: Basic data wrangling  for 2013 Oregon data and seperating by ICD9 lists
+# Author: Jingyang Liu
+# Date Created: Mar 03, 2016
+# Date Checked: Nov 20, 2017
+# R Version: 3.3.2
+# ------------------------------------------------------------------------------
+
+# This file is for 
+# Also it separates the data into 12 disease files due to the ICD9 lists.
+
+# This runs on server.
+
+# library
+library(dplyr)
+library(data.table)
+library(readxl) 
+library(readr)
+
+# import original data ---------------------------------------------------------
+read_path <- paste0("../../../data/data_original/gan_episodes_of_care.txt")
+start_time <- Sys.time()
+oregon_df <- fread(read_path, sep = "|", showProgress = T) 
+stop_time <-  Sys.time() - start_time 
+# time it took
+stop_time # 7 mins
+
+# basic cleaning ---------------------------------------------------------------
+
+oregon_df <- oregon_df %>% 
+  # filter(!is.na(dx1)) %>% 
+  # filter State is Oregon
+  filter(STATE=="OR") 
+# filter urgent or ED visits (pos = 20 or 23)
+# filter(pos=="20"|pos=="23") 
+
+write_path <- paste0('../../../data/data_original/',
+                     '2013_oregon_epis_care.csv')
+write_csv(oregon_df, write_path)
+
+
+
+# Define read paths ------------ -----------------------------------------------
+# import r list of outcome-specific icd9 codes
+load_path <- paste0("../../../data/data_original/outcome_list.RData")
+load(load_path)
+
+# path and file name of data
+read_path <- paste0("../../../data/data_original/2013_oregon_epis_care.csv")
+start_time <- Sys.time()
+oregon_df <- fread(read_path, sep = ",", 
+                   colClasses = rep("character", 72)) 
+stop_time <-  Sys.time() - start_time 
+# time it took
+stop_time # 20 min
+
+
+## make SABA data set(ndc is beta-2-agonist; personkey) ------------------------
+read_path2 <- paste0("../../../data/data_original/2013_ndc_beta2_agonists.csv")
+ndc_df <- read_csv(read_path2, col_names = FALSE)
+
+colnames(ndc_df) <- c("ndc_code", "brand_name", "generic_product_name", 
+                      "route", "category", "drug_id")
+
+# convert to vector
[…]

### 2017-11-28 r_scripts/data_wrangle/2013_oregon_case_crossover.R
case crossover study for all data sets

@@ -1,0 +1,154 @@
+# ------------------------------------------------------------------------------
+# Title: Case crossover for diseases (Oregon State) for server
+# Author: Jingyang Liu
+# Date Created: Nov 27, 2017
+# R version: 3.3.2
+# ------------------------------------------------------------------------------
+
+#Creating case crossover dataframes
+library(dplyr) 
+library(data.table)
+library(readxl)
+library(readr)
+getwd()
+setwd("../../../data/data_new/health/")
+
+### ---------------------------------------------------------------------
+var_list <- c('saba', 'respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+n <- NA
+
+start <- Sys.time()
+for (i in var_list){
+  read_path <- paste('2013_oregon', i, 'claims.csv', sep='_')
+  or_disease <- read_csv(read_path)
+  
+  or_disease1 <- or_disease %>% 
+    group_by(personkey) %>% 
+    arrange(personkey, fromdate, line) %>% 
+    mutate(num_visit = dense_rank(fromdate)) %>%
+    select(personkey, fromdate, num_visit, clmid, line) %>%
+    arrange(personkey, num_visit) %>% 
+    select(personkey, clmid, num_visit) %>% 
+    filter(num_visit>1)
+  
+  or_disease2 <-  or_disease %>% 
+    group_by(personkey) %>% 
+    arrange(personkey, fromdate, line) %>% 
+    mutate(num_visit = dense_rank(fromdate)) %>%
+    select(personkey, fromdate, num_visit, clmid, line) %>%
+    arrange(personkey, num_visit) %>% 
+    select(personkey, clmid, num_visit) %>%
+    filter(!(personkey %in% or_disease1$personkey)) %>%
+    unique() %>% 
+    left_join(or_disease, by = c("personkey", "clmid")) %>% 
+    arrange(personkey, fromdate, clmid, line) %>% 
+    filter(row_number() == 1) %>%
+    # add new transverted from date
+    mutate(dates = as.Date(fromdate, "%m/%d/%Y")) %>%
+    filter(dates >= '2013-05-01' & 
+             dates <= '2013-09-30') 
+  
[…]

### 2017-11-28 r_scripts/data_wrangle/2013_oregon_case_crossover.R
delete

@@ -1,154 +1,0 @@
-# ------------------------------------------------------------------------------
-# Title: Case crossover for diseases (Oregon State) for server
-# Author: Jingyang Liu
-# Date Created: Nov 27, 2017
-# R version: 3.3.2
-# ------------------------------------------------------------------------------
-
-#Creating case crossover dataframes
-library(dplyr) 
-library(data.table)
-library(readxl)
-library(readr)
-getwd()
-setwd("../../../data/data_new/health/")
-
-### ---------------------------------------------------------------------
-var_list <- c('saba', 'respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
-              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
-              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
-n <- NA
-
-start <- Sys.time()
-for (i in var_list){
-  read_path <- paste('2013_oregon', i, 'claims.csv', sep='_')
-  or_disease <- read_csv(read_path)
-  
-  or_disease1 <- or_disease %>% 
-    group_by(personkey) %>% 
-    arrange(personkey, fromdate, line) %>% 
-    mutate(num_visit = dense_rank(fromdate)) %>%
-    select(personkey, fromdate, num_visit, clmid, line) %>%
-    arrange(personkey, num_visit) %>% 
-    select(personkey, clmid, num_visit) %>% 
-    filter(num_visit>1)
-  
-  or_disease2 <-  or_disease %>% 
-    group_by(personkey) %>% 
-    arrange(personkey, fromdate, line) %>% 
-    mutate(num_visit = dense_rank(fromdate)) %>%
-    select(personkey, fromdate, num_visit, clmid, line) %>%
-    arrange(personkey, num_visit) %>% 
-    select(personkey, clmid, num_visit) %>%
-    filter(!(personkey %in% or_disease1$personkey)) %>%
-    unique() %>% 
-    left_join(or_disease, by = c("personkey", "clmid")) %>% 
-    arrange(personkey, fromdate, clmid, line) %>% 
-    filter(row_number() == 1) %>%
-    # add new transverted from date
-    mutate(dates = as.Date(fromdate, "%m/%d/%Y")) %>%
-    filter(dates >= '2013-05-01' & 
-             dates <= '2013-09-30') 
-  
[…]

### 2017-11-28 r_scripts/data_wrangle/2013_oregon_data_wrangle.R

@@ -1,146 +1,0 @@
-# ------------------------------------------------------------------------------
-# Title: Basic data wrangling  for 2013 Oregon data and seperating by ICD9 lists
-# Author: Jingyang Liu
-# Date Created: Mar 03, 2016
-# Date Checked: Nov 20, 2017
-# R Version: 3.3.2
-# ------------------------------------------------------------------------------
-
-# This file is for 
-# Also it separates the data into 12 disease files due to the ICD9 lists.
-
-# This runs on server.
-
-# library
-library(dplyr)
-library(data.table)
-library(readxl) 
-library(readr)
-
-# import original data ---------------------------------------------------------
-read_path <- paste0("../../../data/data_original/gan_episodes_of_care.txt")
-start_time <- Sys.time()
-oregon_df <- fread(read_path, sep = "|", showProgress = T) 
-stop_time <-  Sys.time() - start_time 
-# time it took
-stop_time # 7 mins
-
-# basic cleaning ---------------------------------------------------------------
-
-oregon_df <- oregon_df %>% 
-  # filter(!is.na(dx1)) %>% 
-  # filter State is Oregon
-  filter(STATE=="OR") 
-# filter urgent or ED visits (pos = 20 or 23)
-# filter(pos=="20"|pos=="23") 
-
-write_path <- paste0('../../../data/data_original/',
-                     '2013_oregon_epis_care.csv')
-write_csv(oregon_df, write_path)
-
-
-
-# Define read paths ------------ -----------------------------------------------
-# import r list of outcome-specific icd9 codes
-load_path <- paste0("../../../data/data_original/outcome_list.RData")
-load(load_path)
-
-# path and file name of data
-read_path <- paste0("../../../data/data_original/2013_oregon_epis_care.csv")
-start_time <- Sys.time()
-oregon_df <- fread(read_path, sep = ",", 
-                   colClasses = rep("character", 72)) 
-stop_time <-  Sys.time() - start_time 
-# time it took
-stop_time # 20 min
-
-
-## make SABA data set(ndc is beta-2-agonist; personkey) ------------------------
-read_path2 <- paste0("../../../data/data_original/2013_ndc_beta2_agonists.csv")
-ndc_df <- read_csv(read_path2, col_names = FALSE)
-
-colnames(ndc_df) <- c("ndc_code", "brand_name", "generic_product_name", 
-                      "route", "category", "drug_id")
-
-# convert to vector
[…]

### 2017-11-28 r_scripts/data_wrangle/2013_oregon_timestratified_casecross.R

@@ -1,201 +1,0 @@
-# ------------------------------------------------------------------------------
-# Title: Oregon Case-crossover dataframe creation (only patients with single obs)
-#        and joining with and zip-level PM2.5 values. 
-# Author: Jingyang Liu
-# Date Created: Nov 27 2017     
-# R version: 3.3.3
-# ------------------------------------------------------------------------------
-
-# Notes ----
-# Notes 8/16/16: I updated smoke data with population weighted averages for zip
-# This code also assumed only one visit, so I remove the people with multiple
-# admissions for the outcome
-
-# Notes 10/3/16: I'm retaining the county assignment of the observation so I
-# can join the county population weighted averages as well for comparison
-# and to analyze morbidity similar to Rish
-
-# Libraries ----
-# load libraries
-#library(tidyverse) # tidyverse not on the server
-# if tidyverse is not available, use dplyr and readr
-library(dplyr) 
-library(readr)
-library(lubridate) # working with date
-# parallel computing libraries
-# library(foreach) 
-# library(doParallel)
-
-# Read in smoke data created in loop -------------------------------------------
-
-# read in zipcode level populatoin-weighted pm
-read_path5 <- paste0('../data/smoke/zip_pm_to_merge_with_acap.csv')
-zip_smoke <- read_csv(read_path5) # 63801 rows
-
-# descriptives of the two smoke datasets
-summary(zip_smoke)
-
-# Zipcode PM2.5 estimates
-# create lag variables that take smoke values from n previous days for zipcodes
-zip_smoke_w_lag <- zip_smoke %>% arrange(ZIPCODE, date) %>%
-  # group by zipcode
-  group_by(ZIPCODE) %>% 
-  # wrf
-  mutate(# wrf_f_pm_lag1 = lag(wrf_f_pm, 1, order_by = ZIPCODE), 
-    # wrf_f_pm_lag2 = lag(wrf_f_pm, 2, order_by = ZIPCODE),
[…]

### 2017-11-28 r_scripts/data_wrangle/2013_oregon_diagnosis_separate.R
delete

@@ -1,147 +1,0 @@
-# ------------------------------------------------------------------------------
-# Title: Basic data wrangling  for 2013 Oregon data and seperating by ICD9 lists
-# Author: Jingyang Liu
-# Date Created: Mar 03, 2016
-# Date Checked: Nov 20, 2017
-# R Version: 3.3.2
-# ------------------------------------------------------------------------------
-
-# This file is for 
-# Also it separates the data into 12 disease files due to the ICD9 lists.
-
-# This runs on server.
-
-# library
-library(dplyr)
-library(data.table)
-library(readxl) 
-library(readr)
-
-# import original data ---------------------------------------------------------
-read_path <- paste0("../../../data/data_original/gan_episodes_of_care.txt")
-start_time <- Sys.time()
-oregon_df <- fread(read_path, sep = "|", showProgress = T) 
-stop_time <-  Sys.time() - start_time 
-# time it took
-stop_time # 7 mins
-
-# basic cleaning ---------------------------------------------------------------
-
-oregon_df <- oregon_df %>% 
-  # filter(!is.na(dx1)) %>% 
-  # filter State is Oregon
-  filter(STATE=="OR") 
-# filter urgent or ED visits (pos = 20 or 23)
-# filter(pos=="20"|pos=="23") 
-
-write_path <- paste0('../../../data/data_original/',
-                     '2013_oregon_epis_care.csv')
-write_csv(oregon_df, write_path)
-
-
-
-# Define read paths ------------ -----------------------------------------------
-# import r list of outcome-specific icd9 codes
-load_path <- paste0("../../../data/data_original/outcome_list.RData")
-load(load_path)
-
-# path and file name of data
-read_path <- paste0("../../../data/data_original/2013_oregon_epis_care.csv")
-start_time <- Sys.time()
-oregon_df <- fread(read_path, sep = ",", 
-                   colClasses = rep("character", 72)) 
-stop_time <-  Sys.time() - start_time 
-# time it took
-stop_time # 20 min
-
-
-## make SABA data set(ndc is beta-2-agonist; personkey) ------------------------
-read_path2 <- paste0("../../../data/data_original/2013_ndc_beta2_agonists.csv")
-ndc_df <- read_csv(read_path2, col_names = FALSE)
-
-colnames(ndc_df) <- c("ndc_code", "brand_name", "generic_product_name", 
-                      "route", "category", "drug_id")
-
-# convert to vector
[…]

### 2017-11-28 r_scripts/data_wrangle/asthma_cohort.R
delete not neccessary codes

@@ -7,7 +7,6 @@
 # ------------------------------------------------------------------------------
 
 # This file is for filtering saba outcomes.
-# Also it separates the data into 12 disease files due to the ICD9 lists.
 # And combine a asthma at-risk data set with ICD9 of asthma or saba.
 # original data sets and instrunction files such as index are stored in data_original file,
 # the new data sets we made are stored in data_new, which contain health and smoke
@@ -25,18 +24,12 @@ library(readr)
 ### Prepare work
 
 # import original data ---------------------------------------------------------
-read_path <- paste0("../../../data/data_original/gan_episodes_of_care.txt") # 77million
+read_path <- paste0("../../../data/data_original/gan_episodes_of_care.txt") # 77 million
 start_time <- Sys.time()
 oregon_df <- fread(read_path, sep = "|", showProgress = T) 
 stop_time <-  Sys.time() - start_time 
 # time it took
 stop_time # 7 mins
-
-
-# basic cleaning ---------------------------------------------------------------
-oregon_df <- oregon_df %>% 
-  # filter State is Oregon
-  filter(STATE=="OR") 
 
 write_path <- paste0('../../../data/data_original/',
                      '2013_oregon_epis_care.csv') # csv file, for future use
@@ -44,17 +37,7 @@ write_csv(oregon_df, write_path)
 ### ----------------------------------------------------------------------------
 
 
-### Make data set
-
-# path and file name of data
-read_path <- paste0("../../../data/data_original/2013_oregon_epis_care.csv")
-start_time <- Sys.time()
-oregon_df <- fread(read_path, sep = ",", 
-                   colClasses = rep("character", 72)) 
-stop_time <-  Sys.time() - start_time 
-# time it took
-stop_time # 14 min
-
+### Make saba index
 
 # the ndc beta-2 agonist file are made in SQLite previously.
 read_path2 <- paste0("../../../data/data_original/2013_ndc_beta2_agonists.csv")
@@ -74,11 +57,24 @@ oregon_saba_index <- ndc_in_oregon_df %>%
   select(personkey) %>%
   unique() # unique personkey, 109483
 
-saba_df <- oregon_df %>%
-  filter(personkey %in% oregon_saba_index$personkey) # 11358660, ndc only data set
+write_path <- paste0('../../../data/data_new/health/2013_oregon_saba_index.csv')
+write_csv(oregon_saba_index, write_csv)
[…]

### 2017-11-28 r_scripts/data_wrangle/asthma_cohort.R
delete cleaning code and extra about making saba data set; write saba index; make asthma at risk data set

@@ -1,3 +1,4 @@
+gitpull
 # ------------------------------------------------------------------------------
 # Title: Basic data wrangling  for 2013 Oregon data and seperating by ICD9 lists
 # Author: Jingyang Liu
@@ -7,8 +8,7 @@
 # ------------------------------------------------------------------------------
 
 # This file is for filtering saba outcomes.
-# Also it separates the data into 12 disease files due to the ICD9 lists.
-# And combine a asthma at-risk data set with ICD9 of asthma or saba.
+# And combine an asthma at-risk data set with ICD9 of asthma or saba.
 # original data sets and instrunction files such as index are stored in data_original file,
 # the new data sets we made are stored in data_new, which contain health and smoke
 # data sets. Now the outcomes are all stored in data_new/health.
@@ -32,29 +32,13 @@ stop_time <-  Sys.time() - start_time
 # time it took
 stop_time # 7 mins
 
-
-# basic cleaning ---------------------------------------------------------------
-oregon_df <- oregon_df %>% 
-  # filter State is Oregon
-  filter(STATE=="OR") 
-
 write_path <- paste0('../../../data/data_original/',
                      '2013_oregon_epis_care.csv') # csv file, for future use
 write_csv(oregon_df, write_path)
 ### ----------------------------------------------------------------------------
 
 
-### Make data set
-
-# path and file name of data
-read_path <- paste0("../../../data/data_original/2013_oregon_epis_care.csv")
-start_time <- Sys.time()
-oregon_df <- fread(read_path, sep = ",", 
-                   colClasses = rep("character", 72)) 
-stop_time <-  Sys.time() - start_time 
-# time it took
-stop_time # 14 min
-
+### Make saba index
 
 # the ndc beta-2 agonist file are made in SQLite previously.
 read_path2 <- paste0("../../../data/data_original/2013_ndc_beta2_agonists.csv")
@@ -74,11 +58,23 @@ oregon_saba_index <- ndc_in_oregon_df %>%
   select(personkey) %>%
   unique() # unique personkey, 109483
 
-saba_df <- oregon_df %>%
-  filter(personkey %in% oregon_saba_index$personkey) # 11358660, ndc only data set
+write_path <- paste0('../../../data/data_new/health/2013_oregon_saba_index.csv')
+write_csv(oregon_saba_index, write_csv)
[…]

### 2017-11-28 r_scripts/analysis/oregon_cvd_dl_over65_results.Rmd
tested out function.
figured out fit loop. Finalizing results tomorrow.

@@
 
-Import CVD data.
+Importing the time-stratified case-crossover dataframes for the cardiovascular endpoints.
 
@@
 # read in all the files in to a list
-outcome_df_list <- casecross_order %>%  
-  purrr::map(~read_csv(paste0(health_path, .)))
+cvd_over65_list <- casecross_order %>%  
+  purrr::map(~read_csv(paste0(health_path, .))) %>% 
+  # filter to over 65 group
+  purrr::map(~filter(., age >= 65))
 
 # name dataframes
-names(outcome_df_list) <- file_names
+names(cvd_over65_list) <- file_names
 
@@
 
-```{r distributed lag functions}
-
+```{r custom distributed lag fit function}
 # fit function
@@
   } # end fit option
+```
 
-dl_fit(data = hf65, pm_df = c(3:5), temp_df=2)
-
+```{r custom distributed lag function}
 # custom distributed lag function 
@@
 
+## Finding smoke distributed lag fit based on AIC
+
+This chunk of code uses the dl_fit function I created to find the degrees of freedom for the natural spline that has the lowest AIC for degrees of freem from 2 to 5 for the smoke lag function.
+I set temperature degrees of freedom to 2.
+
+```{r smk aic fit}
+fit_df <- cvd_over65_list %>% 
+  purrr::map_df(~dl_fit(., pm_df = c(2:5), temp_df =2)) %>% 
+  cbind(names(cvd_over65_list), .)
+
+# add degrees of freedom as a slot to the outcomes list
+degfreedom_list <- as.list(fit_df$pm_df)
+```
+
+## Distributed Lag Cumulative Effect
+
+Calculating the cumulative effect sequentially (0, 0+1, 0+1+2,....)
+
+```{r cumulative effect smooth}
+cumulative_cvd <- purrr::map2(.x = cvd_over65_list, .y = degfreedom_list, 
+  ~distributed_lag(data = .x, pm_df = .y, temp_df = 2, cumulative = T)) %>% 
+  plyr::rbind.fill() %>% 
+  bind_cols(., data_frame(rep(outcome_names, each=8))) %>% 
+  rename(outcome_names = `rep(outcome_names, each = 8)`) %>% 
+  select(outcome_names, type, time, estimate, lower95, upper95) %>% 
+  # preserve order of names
+  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))
+```
+
 ## Heart failure in over 65 group
@@
 
-cumulative <- distributed_lag(hf65, pm_df = 4, temp_df = 2, cumulative = T)
+cumulative <- distributed_lag(hf65, pm_df = 2, temp_df = 2, cumulative = T)
 cumulative
@@
 dl_plot
-
 ```
@@
   filter(age_ind == 2)
+
+check <- mi65 %>% filter(personkey == 1782923) %>% 
+  select(personkey, outcome, dx1, date, fromdate)
 

### 2017-11-29 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
worked on over 65 cvd outcomes; may incorporate in to overall results.
update from upstream/development

@@
 
-Packages used: tidyverse, survival.
+Packages used: idyverse, survival.
 ```{r setup, echo =F, warning=F, message=F}
@@
 Elevated levels of smoke in late July to early August in Curry, Josephine, Jackson, and Klamath counties.
+
+Jeff and Bonne suggest only presenting time-series of 4 counties affected, and send this figure to supplement.
 
@@
 
-    # fit conditional logistic regression model ----
+    # fit model with basis ----
   mod <- clogit(outcome ~ geo_smk10 + wrf_temp_zip + strata(personkey),
@@
   # odds ratio
-  estimates <- round(exp(summary(mod)$coefficient[1,1]), 3) 
+  estimates <- round(exp(summary(mod)$coefficient[1,1]), 3) # sd stands for same day
   # 95% lower bound
@@
   ylab(expression("Cumulative odds ratio: 10 ug/m^3 increase smoke PM"[2.5])) +
-  xlab(NULL)
+  xlab("Outcomes")
 
@@
 5. Plots need to have standard labels, ordering, scales, etc.
+6. I'd like to know n-cases and even effect estimates for SABA and asthma in the counties most impacted by smoke.

### 2017-11-29 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
correct the package name
update with space

--- 2017-11-17 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
@@
   ylab(expression("Cumulative odds ratio: 10 ug/m^3 increase smoke PM"[2.5])) +
-  xlab(NULL)
+  xlab("NULL")
 

### 2017-11-29 r_scripts/data_wrangle/asthma_cohort.R
working on code that subsets asthma and saba events.

@@ -1,0 +1,27 @@
+# ------------------------------------------------------------------------------
+# Title: Creation of asthma and beta2 agonist cohort
+# Author: Ryan Gan
+# Date Created: 2017-11-29
+# R Version: 3.4
+# ------------------------------------------------------------------------------
+
+# working with a example txt file of 10,000 rows to get a feel for data
+
+# load library tidyverse
+library(tidyverse)
+
+# read in dataframe ----
+read_path <- paste0("./data/health/oregon_subset.txt")
+# read in with fread
+oregon_df <- data.table::fread(read_path, sep = "|",
+  colClasses = c(rep("character", 72)), stringsAsFactors = FALSE)
+
+head(oregon_df)
+tail(oregon_df)
+
+# set *NULL* and blank to NA
+oregon_df <- oregon_df %>% 
+  # replace *NULL* with NA
+  mutate_all(funs(replace(., .== "*NULL*" | . == "", NA)))
+
+

### 2017-11-29 r_scripts/data_wrangle/oregon_casecrossover.R
case crossover for diagnosis data

@@ -1,0 +1,106 @@
+# ------------------------------------------------------------------------------
+# Title: Case crossover for diseases (Oregon State) for server
+# Author: Jingyang Liu
+# Date Created: Nov 27, 2017
+# R version: 3.3.2
+# ------------------------------------------------------------------------------
+
+#Creating case crossover dataframes
+library(dplyr) 
+library(data.table)
+library(readxl)
+library(readr)
+getwd()
+setwd("../../../data/data_new/health/")
+
+### ---------------------------------------------------------------------
+var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
+              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
+              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
+
+start <- Sys.time()
+for (m in var_list){
+  m = "broken_arm"
+  read_path <- paste('2013_oregon', m, 'claims.csv', sep='_')
+  
+  disease <- read_csv(read_path) %>%
+    select(personkey, clmid, line, gender, yob, race, ethn, STATE, ZIP, 
+           MSA, fromdate, todate, pos, dx1, px1) %>%
+    mutate(age = 2017 - yob, 
+           dx_outcome = m) %>%
+    arrange(personkey, clmid, line, fromdate)
+  
+  outcome_id <- disease %>%
+    # arrange with dates
+    arrange(fromdate) %>%
+    mutate(id = seq(1, nrow(.), by = 1))  # create subject id
+  
+  # create dataset to populate
+  id_date_df <- data_frame()
+  
+  
+  # begin second loop to create counterfactual observations for each case subject
+  for (k in 1:nrow(outcome_id)){
+    
+    # find the replicate times of weeks
+    dates_l <- outcome_id[[k,11]] 
+    n1 <- 0
+    d=as.Date("2013-01-01")
+    i=1
+    while (dates_l >= "2013-01-01"){
+      dates_l <- dates_l - 7
+      d[i] = dates_l
+      i = i+1
+      n1 = n1+1
+    }
+    d[1:n1-1] # shows character(0) when the first week
+    n1-1
+    
+    dates_l <- outcome_id[[k,11]] 
+    n2=0
+    e=as.Date("2013-12-31")
+    j=1
+    while (dates_l <= "2013-12-31"){
+      dates_l <- dates_l + 7
[…]

### 2017-11-30 r_scripts/data_wrangle/icd9_ndc_outcome.R
icd9 and ndc making

@@ -1,0 +1,147 @@
+
+# ------------------------------------------------------------------------------
+# Title: Basic data wrangling  for 2013 Oregon data and seperating by ICD9 lists
+# Author: Jingyang Liu
+# Date Created: Mar 03, 2016
+# Date Checked: Nov 20, 2017
+# R Version: 3.3.2
+# ------------------------------------------------------------------------------
+
+# This file is for 
+# Also it separates the data into 12 disease files due to the ICD9 lists.
+
+# This runs on server.
+
+# library
+library(dplyr)
+library(data.table)
+library(readxl) 
+library(readr)
+
+# import original data ---------------------------------------------------------
+read_path <- paste0("../../../data/data_original/gan_episodes_of_care.txt")
+start_time <- Sys.time()
+oregon_df <- fread(read_path, sep = "|", showProgress = T) 
+stop_time <-  Sys.time() - start_time 
+# time it took
+stop_time # 7 mins
+
+# basic cleaning ---------------------------------------------------------------
+
+oregon_df <- oregon_df %>% 
+  # filter(!is.na(dx1)) %>% 
+  # filter State is Oregon
+  filter(STATE=="OR") 
+# filter urgent or ED visits (pos = 20 or 23)
+# filter(pos=="20"|pos=="23") 
+
+write_path <- paste0('../../../data/data_original/',
+                     '2013_oregon_epis_care.csv')
+write_csv(oregon_df, write_path)
+
+
+
+# Define read paths ------------ -----------------------------------------------
+# import r list of outcome-specific icd9 codes
+load_path <- paste0("../../../data/data_original/outcome_list.RData")
+load(load_path)
+
+# path and file name of data
+read_path <- paste0("../../../data/data_original/2013_oregon_epis_care.csv")
+start_time <- Sys.time()
+oregon_df <- fread(read_path, sep = ",", 
+                   colClasses = rep("character", 72)) 
+stop_time <-  Sys.time() - start_time 
+# time it took
+stop_time # 20 min
+
+
+## make SABA data set(ndc is beta-2-agonist; personkey) ------------------------
+read_path2 <- paste0("../../../data/data_original/2013_ndc_beta2_agonists.csv")
+ndc_df <- read_csv(read_path2, col_names = FALSE)
+
+colnames(ndc_df) <- c("ndc_code", "brand_name", "generic_product_name", 
+                      "route", "category", "drug_id")
+
+# convert to vector
[…]

### 2017-11-30 r_scripts/data_wrangle/icd9_ndc_outcome.R
make icd9 and saba data sets

@@ -26,18 +26,88 @@ stop_time <-  Sys.time() - start_time
 # time it took
 stop_time # 7 mins
 
-# basic cleaning ---------------------------------------------------------------
 
-oregon_df <- oregon_df %>% 
-  # filter(!is.na(dx1)) %>% 
-  # filter State is Oregon
-  filter(STATE=="OR") 
-# filter urgent or ED visits (pos = 20 or 23)
-# filter(pos=="20"|pos=="23") 
+## make SABA data set(ndc is beta-2-agonist; personkey) ------------------------
+read_path2 <- paste0("../../../data/data_original/2014-hedis_asthma_ndc.xlsx")
+beta2_ndc <- read_excel(read_path2)
 
-write_path <- paste0('../../../data/data_original/',
-                     '2013_oregon_epis_care.csv')
-write_csv(oregon_df, write_path)
+saba_ndc <- beta2_ndc %>%
+  filter(category == "short-acting inhaled beta-2 agonists")
+
+ndc_in_oregon_df <- oregon_df %>%
+  filter(ndc %in% saba_ndc$ndc_code) # 463265, unique personkey 164975
+
+oregon_saba_index <- ndc_in_oregon_df %>%
+  select(personkey) %>%
+  unique() # 164975
+
+saba_df <- oregon_df %>%
+  filter(personkey %in% oregon_saba_index$personkey) # 16254091
+
+write_path2 <- paste0('../../../data/data_new/health/2013_oregon_saba_index.csv')
+write_csv(oregon_saba_index, write_path2)
+
+write_path3 <- paste0('../../../data/data_new/health/2013_oregon_saba_claims.csv')
+write_csv(saba_df, write_path3)
+
+# ------------------------------------------------------------------------------
+
+# read in icd9 data and create lists of vectors of icd9 codes ----
+# read in xlsx icd9 codes and description
+icd9_df <- read_excel("../../../data/data_original/CMS32_DESC_LONG_SHORT_DX.xlsx") %>% 
+  # rename the terrible variable names
+  select(dx_code = 1, long_desc = 2, short_desc = 3)
+
+# look at the first 6 rows of the icd9 codes; comes in sorted
+head(icd9_df)
+glimpse(icd9_df)
+
+# Overall goal is to make a list of vectors for each outcome that contains the
+# character strings of icd9 codes that make up the outcome of interest.
+# Some outcomes like COPD, where icd9 codes are not sequential, will not work
+# and will need to be entered by hand.
+
[…]

### 2017-11-30 r_scripts/data_wrangle/asthma_cohort.R
copy Ryan's file to avoid conflicts

@@ -1,0 +1,27 @@
+# ------------------------------------------------------------------------------
+# Title: Creation of asthma and beta2 agonist cohort
+# Author: Ryan Gan
+# Date Created: 2017-11-29
+# R Version: 3.4
+# ------------------------------------------------------------------------------
+
+# working with a example txt file of 10,000 rows to get a feel for data
+
+# load library tidyverse
+library(tidyverse)
+
+# read in dataframe ----
+read_path <- paste0("./data/health/oregon_subset.txt")
+# read in with fread
+oregon_df <- data.table::fread(read_path, sep = "|",
+                               colClasses = c(rep("character", 72)), stringsAsFactors = FALSE)
+
+head(oregon_df)
+tail(oregon_df)
+
+# set *NULL* and blank to NA
+oregon_df <- oregon_df %>% 
+  # replace *NULL* with NA
+  mutate_all(funs(replace(., .== "*NULL*" | . == "", NA)))
+
+

### 2017-11-30 r_scripts/data_wrangle/asthma_cohort.R
space

@@ -14,7 +14,7 @@ library(tidyverse)
 read_path <- paste0("./data/health/oregon_subset.txt")
 # read in with fread
 oregon_df <- data.table::fread(read_path, sep = "|",
-                               colClasses = c(rep("character", 72)), stringsAsFactors = FALSE)
+  colClasses = c(rep("character", 72)), stringsAsFactors = FALSE)
 
 head(oregon_df)
 tail(oregon_df)

### 2017-11-30 r_scripts/data_wrangle/outcome_icd9_codes.R
ICD9 outcome vectors making

@@ -1,0 +1,74 @@
+# ------------------------------------------------------------------------------
+# Title: Basic icd9 list for 2013 oregon data
+# Author: Jingyang Liu
+# Date Created: Mar 03, 2016
+# R Version: 3.3.2
+# ------------------------------------------------------------------------------
+
+# The purpose of this code is to easily group the ICD-9 codes that make up
+# a specific outcome of interest using the ICD-9-CM version 32 list
+
+# Data files can be found here: 
+# https://www.cms.gov/Medicare/Coding/ICD9ProviderDiagnosticCodes/codes.html
+
+# load library -----------------------------------------------------------------
+library(tidyverse)
+library(readxl) # read excel files
+
+
+# read in icd9 data and create lists of vectors of icd9 codes ----
+# read in xlsx icd9 codes and description
+icd9_df <- read_excel("./instructions/CMS32_DESC_LONG_SHORT_DX.xlsx") %>% 
+  # rename the terrible variable names
+  select(dx_code = 1, long_desc = 2, short_desc = 3)
+
+# look at the first 6 rows of the icd9 codes; comes in sorted
+head(icd9_df)
+glimpse(icd9_df)
+
+# Overall goal is to make a list of vectors for each outcome that contains the
+# character strings of icd9 codes that make up the outcome of interest.
+# Some outcomes like COPD, where icd9 codes are not sequential, will not work
+# and will need to be entered by hand.
+
+# Create a dataframe of the outcome name, start, and stop rows
+outcome_name <- c("respiratory", "asthma", "pneumonia", "acute_bronch",
+                  "cvd", "arrhythmia", "cerbrovas_dis", "heart_failure",
+                  "isch_heart_dis", "myocardial_infarc", "broken_arm")
+# first icd9 code of outcome
+start_icd9 <- c("460", "49300", "4800", "4660", "390", "4270", "430", "4280",
+                "41000", "41000", "81300")
+
+# last icd9 code of outcome
+stop_icd9 <- c("5199", "49392", "486", "46619", "4599", "4279", "4389", "4289",
+               "4139", "41092", "81393")
+
+# dataframe of outcomes and start stop of icd9 codes
+outcome_icd9_range <- data_frame(outcome_name, start_icd9, stop_icd9)
+
+# list of icd9 codes -----------------------------------------------------------
+outcome_icd9_list <- apply(outcome_icd9_range, 1, 
+                           function(x) { # start of function
+                             icd9_df %>% 
[…]

### 2017-11-30 r_scripts/data_wrangle/icd9_ndc_outcome.R
delete RData making part

@@ -52,64 +52,6 @@ write_csv(saba_df, write_path3)
 
 # ------------------------------------------------------------------------------
 
-# read in icd9 data and create lists of vectors of icd9 codes ----
-# read in xlsx icd9 codes and description
-icd9_df <- read_excel("../../../data/data_original/CMS32_DESC_LONG_SHORT_DX.xlsx") %>% 
-  # rename the terrible variable names
-  select(dx_code = 1, long_desc = 2, short_desc = 3)
-
-# look at the first 6 rows of the icd9 codes; comes in sorted
-head(icd9_df)
-glimpse(icd9_df)
-
-# Overall goal is to make a list of vectors for each outcome that contains the
-# character strings of icd9 codes that make up the outcome of interest.
-# Some outcomes like COPD, where icd9 codes are not sequential, will not work
-# and will need to be entered by hand.
-
-# Create a dataframe of the outcome name, start, and stop rows
-outcome_name <- c("respiratory", "asthma", "pneumonia", "acute_bronch",
-                  "cvd", "arrhythmia", "cerbrovas_dis", "heart_failure",
-                  "isch_heart_dis", "myocardial_infarc", "broken_arm")
-# first icd9 code of outcome
-start_icd9 <- c("460", "49300", "4800", "4660", "390", "4270", "430", "4280",
-                "41000", "41000", "81300")
-
-# last icd9 code of outcome
-stop_icd9 <- c("5199", "49392", "486", "46619", "4599", "4279", "4389", "4289",
-               "4139", "41092", "81393")
-
-# dataframe of outcomes and start stop of icd9 codes
-outcome_icd9_range <- data_frame(outcome_name, start_icd9, stop_icd9)
-
-# list of icd9 codes -----------------------------------------------------------
-outcome_icd9_list <- apply(outcome_icd9_range, 1, 
-                           function(x) { # start of function
-                             icd9_df %>% 
-                               slice(which(icd9_df$dx_code == x[2]): which(icd9_df$dx_code == x[3])) %>%
-                               select(1) %>%
-                               as_vector()
-                           } )
-
-# assign list names
-names(outcome_icd9_list) <- outcome_name
-
-# Add COPD ICD9 codes to the list
-# COPD, ICD9 490 to 492, 494, and 496 
-copd_icd9 <- c('490', '4910','4911','49120','49121','49122','4918','4919', '4920',
-               '4928', '4940', '4941', '496') 
-
-# append copd list to icd9 list
[…]

### 2017-11-30 r_scripts/data_wrangle/outcome_icd9_codes.R
add saba into .RData

@@ -12,13 +12,15 @@
 # https://www.cms.gov/Medicare/Coding/ICD9ProviderDiagnosticCodes/codes.html
 
 # load library -----------------------------------------------------------------
-library(tidyverse)
+library(dplyr)
+library(purrr)
+library(data.table)
 library(readxl) # read excel files
 
 
 # read in icd9 data and create lists of vectors of icd9 codes ----
 # read in xlsx icd9 codes and description
-icd9_df <- read_excel("./instructions/CMS32_DESC_LONG_SHORT_DX.xlsx") %>% 
+icd9_df <- read_excel("../../../data/data_original/CMS32_DESC_LONG_SHORT_DX.xlsx") %>% 
   # rename the terrible variable names
   select(dx_code = 1, long_desc = 2, short_desc = 3)
 
@@ -69,6 +71,19 @@ outcome_icd9_list <- c(outcome_icd9_list, list(copd = copd_icd9))
 # I don't know if not having atomic/flat values for each element of the list
 # will matter, but I need to try it out before I'll know
 
+## saba lists ------------------------------------------------------------------
+read_path2 <- paste0("../../../data/data_original/2014-hedis_asthma_ndc.xlsx")
+beta2_ndc <- read_excel(read_path2)
+
+saba_ndc <- beta2_ndc %>%
+  filter(category == "short-acting inhaled beta-2 agonists")
+
+saba_code <- saba_ndc %>%
+  select(ndc_code) %>%
+  as_vector()
+  
+outcome_icd9_list <- c(outcome_icd9_list, list(saba = saba_code))
+
 # save R file to use in other scripts
-save_path <- paste0("./data/outcome_list.RData")
+save_path <- paste0("../../../data/data_original/outcome_list.RData")
 save(outcome_icd9_list, file = save_path)

### 2017-11-30 r_scripts/data_wrangle/icd9_ndc_outcome.R
make saba data set from .RDate

@@ -26,30 +26,9 @@ stop_time <-  Sys.time() - start_time
 # time it took
 stop_time # 7 mins
 
-
-## make SABA data set(ndc is beta-2-agonist; personkey) ------------------------
-read_path2 <- paste0("../../../data/data_original/2014-hedis_asthma_ndc.xlsx")
-beta2_ndc <- read_excel(read_path2)
-
-saba_ndc <- beta2_ndc %>%
-  filter(category == "short-acting inhaled beta-2 agonists")
-
-ndc_in_oregon_df <- oregon_df %>%
-  filter(ndc %in% saba_ndc$ndc_code) # 463265, unique personkey 164975
-
-oregon_saba_index <- ndc_in_oregon_df %>%
-  select(personkey) %>%
-  unique() # 164975
-
-saba_df <- oregon_df %>%
-  filter(personkey %in% oregon_saba_index$personkey) # 16254091
-
-write_path2 <- paste0('../../../data/data_new/health/2013_oregon_saba_index.csv')
-write_csv(oregon_saba_index, write_path2)
-
-write_path3 <- paste0('../../../data/data_new/health/2013_oregon_saba_claims.csv')
-write_csv(saba_df, write_path3)
-
+# filter Oregon State
+oregon_df <- oregon_df %>% 
+  filter(STATE=="OR") 
 # ------------------------------------------------------------------------------
 
 
@@ -57,10 +36,6 @@ write_csv(saba_df, write_path3)
 # import r list of outcome-specific icd9 codes
 load_path <- paste0("../../../data/data_original/outcome_list.RData")
 load(load_path)
-
-oregon_df <- oregon_df %>% 
-  filter(STATE=="OR") 
-
 
 # now try to work the two chunks together to identify records that contain icd9s
 # of a certain outcome and then output those rows in chunks of data.
@@ -83,7 +58,7 @@ col_names <- colnames(df_col_names) # output column names
 # loop to output dataframes of outcomes ----------------------------------------
 start_time <- Sys.time() # start clock time
 
-for(i in 1:length(outcome_icd9_list)) { # start loop
+for(i in 1:(length(outcome_icd9_list)-1)) { # start loop
   
   outcome_name <- names(outcome_icd9_list)[i] # output the name of the list/outcome
   icd9_vector <- outcome_icd9_list[[i]] # output the icd9 codes of the list
[…]

### 2017-11-30 r_scripts/data_wrangle/asthma_cohort.R
finished oregon asthma cohort code.

@@ -5,7 +5,6 @@
 # R Version: 3.4
 # ------------------------------------------------------------------------------
 
-# working with a example txt file of 10,000 rows to get a feel for data
 
 # load library tidyverse
 library(tidyverse)
@@ -16,12 +15,54 @@ read_path <- paste0("./data/health/oregon_subset.txt")
 oregon_df <- data.table::fread(read_path, sep = "|",
   colClasses = c(rep("character", 72)), stringsAsFactors = FALSE)
 
-head(oregon_df)
-tail(oregon_df)
-
 # set *NULL* and blank to NA
+# i may be able to integrate this to the first part
 oregon_df <- oregon_df %>% 
   # replace *NULL* with NA
   mutate_all(funs(replace(., .== "*NULL*" | . == "", NA)))
+
+# read in Rdata icd9 outcome vectors and limit to asthma
+outcome_icd9_list<- load("./data/health/outcome_list.RData")
+# subset to asthma icd9
+asthma_icd9 <- pluck(outcome_icd9_list, "asthma")
+
+beta2_ndc <- read_csv("./data/health/2014-hedis_asthma_code.csv")
+
+# limit to short-acting beta2 agonist vector
+saba_ndc <- beta2_ndc %>% 
+  filter(category == "short-acting inhaled beta-2 agonists") %>% 
+  select(ndc_code) %>% 
+  as_vector()
+
+# filter oregon epidsodes of care to dx of asthma or ndc of saba ------
+# find ids with at least one dx of asthma
+asthma_id <- oregon_df %>% 
+  filter_at(vars(contains("dx")), any_vars(. %in% asthma_icd9)) %>% 
+  select(personkey) %>% 
+  as_vector() 
+
+saba_id <- oregon_df %>% 
+  filter(ndc %in% saba_ndc) %>% 
+  select(personkey) %>% 
+  as_vector()
+
+# bind two id vectors together
+asthma_saba_personkey <- unique(c(asthma_id, saba_id))
+
+# filter large dataset to persons with asthma or saba fills ------
+# creating persons at risk (with asthma or a saba fill) for adverse events
+asthma_saba_at_risk <- oregon_df %>% 
+  filter(personkey %in% asthma_saba_personkey) %>% 
+  # make a indicator if visit is asthma hosp, saba fill, or neither
+  mutate(visit_type = ifelse(dx1 %in% asthma_icd9, "asthma", 
[…]

### 2017-11-30 r_scripts/data_wrangle/oregon_casecrossover.R
delete not necessary for now

@@ -1,106 +1,0 @@
-# ------------------------------------------------------------------------------
-# Title: Case crossover for diseases (Oregon State) for server
-# Author: Jingyang Liu
-# Date Created: Nov 27, 2017
-# R version: 3.3.2
-# ------------------------------------------------------------------------------
-
-#Creating case crossover dataframes
-library(dplyr) 
-library(data.table)
-library(readxl)
-library(readr)
-getwd()
-setwd("../../../data/data_new/health/")
-
-### ---------------------------------------------------------------------
-var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
-              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
-              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')
-
-start <- Sys.time()
-for (m in var_list){
-  m = "broken_arm"
-  read_path <- paste('2013_oregon', m, 'claims.csv', sep='_')
-  
-  disease <- read_csv(read_path) %>%
-    select(personkey, clmid, line, gender, yob, race, ethn, STATE, ZIP, 
-           MSA, fromdate, todate, pos, dx1, px1) %>%
-    mutate(age = 2017 - yob, 
-           dx_outcome = m) %>%
-    arrange(personkey, clmid, line, fromdate)
-  
-  outcome_id <- disease %>%
-    # arrange with dates
-    arrange(fromdate) %>%
-    mutate(id = seq(1, nrow(.), by = 1))  # create subject id
-  
-  # create dataset to populate
-  id_date_df <- data_frame()
-  
-  
-  # begin second loop to create counterfactual observations for each case subject
-  for (k in 1:nrow(outcome_id)){
-    
-    # find the replicate times of weeks
-    dates_l <- outcome_id[[k,11]] 
-    n1 <- 0
-    d=as.Date("2013-01-01")
-    i=1
-    while (dates_l >= "2013-01-01"){
-      dates_l <- dates_l - 7
-      d[i] = dates_l
-      i = i+1
-      n1 = n1+1
-    }
-    d[1:n1-1] # shows character(0) when the first week
-    n1-1
-    
-    dates_l <- outcome_id[[k,11]] 
-    n2=0
-    e=as.Date("2013-12-31")
-    j=1
-    while (dates_l <= "2013-12-31"){
-      dates_l <- dates_l + 7
[…]

### 2017-11-30 r_scripts/data_wrangle/asthma_cohort.R
updated outcomes list using .Rdata file jingyang provided.

@@ -22,17 +22,12 @@ oregon_df <- oregon_df %>%
   mutate_all(funs(replace(., .== "*NULL*" | . == "", NA)))
 
 # read in Rdata icd9 outcome vectors and limit to asthma
-outcome_icd9_list<- load("./data/health/outcome_list.RData")
+load("./data/health/outcome_list.RData")
+
 # subset to asthma icd9
 asthma_icd9 <- pluck(outcome_icd9_list, "asthma")
-
-beta2_ndc <- read_csv("./data/health/2014-hedis_asthma_code.csv")
-
-# limit to short-acting beta2 agonist vector
-saba_ndc <- beta2_ndc %>% 
-  filter(category == "short-acting inhaled beta-2 agonists") %>% 
-  select(ndc_code) %>% 
-  as_vector()
+# subset to ndc saba
+saba_ndc <- pluck(outcome_icd9_list, "saba")
 
 # filter oregon epidsodes of care to dx of asthma or ndc of saba ------
 # find ids with at least one dx of asthma
@@ -60,6 +55,7 @@ asthma_saba_at_risk <- oregon_df %>%
 # save file
 write_path <- paste0("./data/health/asthma_saba_cohort.csv")
 write_csv(asthma_saba_at_risk, write_path)
+
 
 
 

### 2017-11-30 r_scripts/data_wrangle/asthma_cohort.R
updated asthma cohort script.

@@ -1,11 +1,9 @@
-<<<<<<< HEAD
 # ------------------------------------------------------------------------------
 # Title: Creation of asthma and beta2 agonist cohort
 # Author: Ryan Gan
 # Date Created: 2017-11-29
 # R Version: 3.4
 # ------------------------------------------------------------------------------
-
 
 # load library tidyverse
 library(tidyverse)
@@ -54,7 +52,7 @@ asthma_saba_at_risk <- oregon_df %>%
     ifelse(ndc %in% saba_ndc, "saba", "other")))
 
 # save file
-write_path <- paste0("./data/health/asthma_saba_cohort.csv")
+write_path <- paste0("./data/health/2013-asthma_saba_cohort.csv")
 write_csv(asthma_saba_at_risk, write_path)
 
 
@@ -63,68 +61,3 @@ write_csv(asthma_saba_at_risk, write_path)
 
 
 
-=======
-
-# ------------------------------------------------------------------------------
-# Title: Creation of asthma and beta2 agonist cohort
-# Author: Ryan Gan
-# Date Created: 2017-11-29
-# R Version: 3.4
-# ------------------------------------------------------------------------------
-
-
-# load library tidyverse
-library(tidyverse)
-
-# read in dataframe ----
-read_path <- paste0("./data/health/oregon_subset.txt")
-# read in with fread
-oregon_df <- data.table::fread(read_path, sep = "|",
-  colClasses = c(rep("character", 72)), stringsAsFactors = FALSE)
-
-# set *NULL* and blank to NA
-# i may be able to integrate this to the first part
-oregon_df <- oregon_df %>% 
-  # replace *NULL* with NA
-  mutate_all(funs(replace(., .== "*NULL*" | . == "", NA)))
-
-# read in Rdata icd9 outcome vectors and limit to asthma
-outcome_icd9_list<- load("./data/health/outcome_list.RData")
-# subset to asthma icd9
-asthma_icd9 <- pluck(outcome_icd9_list, "asthma")
-
-beta2_ndc <- read_csv("./data/health/2014-hedis_asthma_code.csv")
-
-# limit to short-acting beta2 agonist vector
-saba_ndc <- beta2_ndc %>% 
-  filter(category == "short-acting inhaled beta-2 agonists") %>% 
-  select(ndc_code) %>% 
-  as_vector()
-
[…]

### 2017-12-02 r_scripts/analysis/agu_outcomes_map.Rmd
data for agu.

---
title: "AGU Map"
author: "Ryan Gan"
date: "12/1/2017"
output: html_document
---

```{r setup, echo=F, warning=F, message=F}
library(tidyverse) # general data wrangle

# knitr options
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

## Document Purpose

Map for Kate and Jeff for AGU to demonstrate variability in health outcomes.

```{r data wrangle}
# saba timestratifed to counts for fills with a icd9 as well
saba <- read_csv("./data/health/2013-oregon_casecross_saba.csv") %>% 
  # get rid of referent values
  filter(outcome == 1)

# time sereis of saba
saba_ts <- saba %>% 
  group_by(MSA, date) %>% 
  summarise(saba_n = n()) %>% 
  mutate(metroarea = case_when(
    MSA == 13460 ~ "Bend-Redmond",
    MSA == 18700 ~ "Corvallis",
    MSA == 21660 ~ "Eugene-Springfield",
    MSA == 32780 ~ "Medford",
    MSA == 38900 ~ "Portland-Vancouver-Beaverton",
    MSA == 41420 ~ "Salem",
    MSA == 41999 ~ "Not in MSA"))
```

### 2017-12-04 r_scripts/analysis/agu_saba_timeseries.Rmd
renamed map file to time-series. finished saba timeseries.

---
title: "AGU Timeseries"
author: "Ryan Gan"
date: "12/1/2017"
output: html_document
---

```{r setup, echo=F, warning=F, message=F}
library(tidyverse) # general data wrangle

# knitr options
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

## Document Purpose

Map for Kate and Jeff for AGU to demonstrate variability in health outcomes.

```{r data wrangle}
# saba timestratifed to counts for fills with a icd9 as well
saba <- read_csv("../../data/health/2013-oregon_casecross_saba.csv") %>% 
  # get rid of referent values
  filter(outcome == 1)

# time sereis of saba
saba_ts <- saba %>% 
  group_by(MSA, date) %>% 
  summarise(saba_n = n()) %>% 
  mutate(metroarea = case_when(
    MSA == 13460 ~ "Bend-Redmond",
    MSA == 18700 ~ "Corvallis",
    MSA == 21660 ~ "Eugene-Springfield",
    MSA == 32780 ~ "Medford",
    MSA == 38900 ~ "Portland-Vancouver-Beaverton",
    MSA == 41420 ~ "Salem",
    MSA == 41999 ~ "Not in MSA"),
    day = weekdays(date),
    weekend = ifelse(day %in% c("Saturday", "Sunday"), "Weekend", "Weekday"))

# calulate median weekend and weekday value
saba_center_stats <- saba_ts %>% 
  # group by MSA and weekend
  group_by(metroarea, weekend) %>% 
  summarise(saba_median = median(saba_n), saba_mean = mean(saba_n),
            saba_sd = sd(saba_n))

# read county pm time sereis
# pm path
pm_path <- paste0("../../data/pm/2013-oregon_county_pm25.csv")
# county pm
county_pm_df <- read_csv(pm_path) %>% 
  mutate(county = stringr::str_replace(county, "[.]", ""))
```

Plot of SABA counts by metro areas in Oregon for 2013 wildfire season (May through September). I couldn't easily get counts in a county out of the data as we had to code county based on reported ZIP code and we never added this as a covariate to the dataset (I never explicitly asked Jingyang to do this; I plan to, but it will probably take a bit).

But, I think MSA works pretty well to illustrate variability in the SABA health outcome that may be linked with smoke.

```{r msa plot}
plot <- ggplot(saba_ts, 
    aes(x=date, y=saba_n, group = weekend, color = weekend)) +
  geom_point() +
  geom_hline(data=saba_center_stats, aes(yintercept=saba_mean, group=weekend,
    color=weekend)) +
  scale_color_manual(name = "Days of Week", values = c("#ec38bc", "#7303c0")) +
  facet_wrap(~metroarea) +
  ylab("SABA Count") +
  xlab("Date") +
  theme_minimal()

plot
```

I've produced plots of SABA counts by MSA area. If you look at the major MSA areas, you'll see that in the larger MSAs like Portland, there is a periodic seasonal increase in spring SABA counts that fall off in summer and increase again around fall. This is the same type of trend seen for asthma. Some of the other more populated counties begin to exhibit this shape too, but not as pronounced as the more populated areas. If you look at Medford (the county impacted to the greatest extent by the Douglas-Complex fire, there is a little increase around the time of the fire (started: July 26th, contained: September 3rd 2013). If you look at some of the other MSA trends, they do not have this little peak around this time period.

Note I had a MSA code that I was not able to link with an MSA so I'm guessing it was not in a MSA; data key was not explicit about this but I'm pretty sure it's for fills for people not in MSA. Also note that there are some weekdays that have similar fill-counts as weekends; these are Memorial Day, July 4th, and Labor Day holidays.

Limiting to MSA timeseries to Medford MSA, which is in Jackson county.

```{r medford}
medford_center_stats <- saba_center_stats %>% 
  filter(metroarea == "Medford")

medford_ts <- saba_ts %>% 
  filter(metroarea == "Medford") %>% 
  full_join(medford_center_stats, by = c("metroarea", "weekend")) %>% 
  mutate(diff_mean = saba_n - saba_mean,
    saba_z = (saba_n-saba_mean)/saba_sd, 
    diff_med = saba_n - saba_median,
    fire_period = ifelse(date >= "2013-07-26" & date <= "2013-09-03",50,NA))
```

Medford MSA SABA count time-series plot for Weekend and Workweek fills. Dotted verticle lines represent the start date of the Douglas-Complex fires and the date it was considered 100% contained.

```{r medford count timeseries}
# medford time series
ts_plot <- ggplot(medford_ts, 
  aes(x=date, y=saba_n, group=weekend, color=weekend)) +
  geom_point() +
  geom_hline(data=medford_center_stats, aes(yintercept=saba_mean, color=weekend)) +
  scale_color_manual(name = "Days of Week", values = c("#ec38bc", "#7303c0")) +
  geom_vline(xintercept = as.numeric(as.Date(c("2013-07-26", "2013-09-03"), 
      "%Y-%m-%d")),
    linetype=3, color = "#191654") +
  ylim(0, 50) +
  ylab("SABA Count") +
  xlab("Date") +
  #ggtitle("Medford SABA Counts by Date") +
  theme_minimal()

ts_plot
```

Plotting the Z-score of SABA counts (observed - mean)/standard deviation. Z-score accounted for weekend/weekday difference

```{r saba z}
# z-score plot
z_plot <-ggplot(medford_ts, 
  aes(x=date, y=saba_z, group=weekend, color=weekend)) +
  geom_point() +
  geom_hline(yintercept=0, color="red") +
  scale_color_manual(name = "Days of Week", values = c("#ec38bc", "#7303c0")) +
  geom_vline(xintercept = as.numeric(as.Date(c("2013-07-26", "2013-09-03"), 
      "%Y-%m-%d")),
    linetype=3, color = "#191654") +
  ylim(-3, 3) +
  ylab("SABA Z-Score") +
  xlab("Date") +
  #ggtitle("Medford SABA Counts by Date") +
  theme_minimal()
z_plot
```

Limiting population-weighted PM~2.5~ time-series to Jackson county. Jackson county population-weighted geographically-weighted ridge regression PM~2.5~ time-series.

```{r gwr smk timeseries}
# jackson plot
jackson_pm <- county_pm_df %>% 
  filter(county == "Jackson") %>% 
  mutate(smk10 = ifelse(geo_wt_pm > 10, 1, 0),
    fire_period = ifelse(date >= "2013-07-26" & date <= "2013-09-03",1,0))

# plot 
pm_plot <- ggplot(jackson_pm, aes(x=date, y=geo_wt_pm)) +
  geom_point(color = "#000046") +
  geom_vline(xintercept = as.numeric(as.Date(c("2013-07-26", "2013-09-03"), 
      "%Y-%m-%d")),
    linetype=3, color = "#191654") +
  ylab(expression(paste("PM"[2.5], "  ", mu, "g/", m^3, sep=""))) +
  xlab("Date") +
  #ggtitle("Medford SABA Counts by Date") +
  theme_minimal()

pm_plot  
```

### 2017-12-04 bash_scripts/r_submission_script.sh
adding bash script folder with setup script for ozone cluster

@@ -1,0 +1,17 @@
+#!/bin/bash
+#PBS -N ryan_rjob
+#PBS -l nodes=1:ppn=1
+#PBS -W group_list=pierce_group
+#PBS -q batch
+
+# set command directory
+cd /home/ryangan/local_git_repo/oregon_wildfire/
+# print directory
+pwd
+
+# export R library
+export R_LIB=~/R/x86_64-pc-linux-gnu-library/3.3
+
+# run rscript
+Rscript --vanilla ./r_scripts/data_wrangle/asthma_cohort.R
+

### 2017-12-05 r_scripts/analysis/agu_saba_timeseries.Rmd
Made figure for jeff and kate.

@@
     day = weekdays(date),
-    weekend = ifelse(day %in% c("Saturday", "Sunday"), "Weekend", "Weekday"))
+    weekend = ifelse(day %in% c("Saturday", "Sunday"), "Weekend", "Weekday"),
+    week_n = lubridate::floor_date(date, unit = "week"))
 
@@
 
+### Figure 1: Time-series of short-acting beta-2 agonists over the metropolitan statistical areas for the state of Oregon.
+
 ```{r msa plot}
@@
   xlab("Date") +
-  theme_minimal()
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_line(colour="grey"),
+    #panel.grid.minor = element_blank(),
+    legend.position = "bottom",
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"))
 
@@
 
+### Figure 2: Time-series of SABA limited to Medford MSA.
+
 ```{r medford}
@@
 
+### Figure 3: Time-series of SABA Z-Score.
+
 ```{r medford count timeseries}
@@
   #ggtitle("Medford SABA Counts by Date") +
-  theme_minimal()
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    legend.position = "bottom")
 
@@
 
+### Figure 4: Truncated SABA count
+
+I'm truncating to July to October.
+
+```{r trunc saba ts}
+jul_oct_medford <- medford_ts %>% 
+  filter(date >= "2013-07-01" & date <= "2013-10-01")
+
+jul_oct_central_vals <- medford_ts %>% 
+  group_by(weekend) %>% 
+  summarise(saba_med = median(saba_n))
+
+# medford time series
+saba_plot <- ggplot(jul_oct_medford, 
+  aes(x=date, y=saba_n, group=weekend, color=weekend)) +
+  geom_point() +
+  #geom_hline(data=jul_oct_central_vals , aes(yintercept=saba_med, color=weekend)) +
+  scale_color_manual(name = "Days of Week", values = c("#ec38bc", "#7303c0")) +
+  geom_vline(xintercept = as.numeric(as.Date(c("2013-07-26", "2013-09-03"), 
+      "%Y-%m-%d")),
+    linetype=3, color = "#191654") +
+  ylim(0, 40) +
+  ylab("SABA Count") +
+  xlab("Date") +
+  #ggtitle("Medford SABA Counts by Date") +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    legend.position = "bottom",
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"))
+
+saba_plot
+```
+
+### Figure 5: Time-series of GWR PM~2.5~ values.
+
 ```{r gwr smk timeseries}
@@
   mutate(smk10 = ifelse(geo_wt_pm > 10, 1, 0),
-    fire_period = ifelse(date >= "2013-07-26" & date <= "2013-09-03",1,0))
+    fire_period = ifelse(date >= "2013-07-26" & date <= "2013-09-03",1,0)) %>% 
+  filter(date >= "2013-07-01" & date <= "2013-10-01")
 
@@
   #ggtitle("Medford SABA Counts by Date") +
-  theme_minimal()
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank())
 
@@
 ```
+
+### Figure 6: Binding truncated SABA counts and PM~2.5~
+
+```{r trunc plot}
+cowplot::plot_grid(pm_plot, saba_plot,
+  #labels = c("PM2.5 Exposure", "Rescue Inhaler"), 
+  nrow = 2, align = "v")
+```

### 2017-12-05 r_scripts/data_wrangle/asthma_cohort.R
testing server script; failing to load .Rdata

@@ -9,10 +9,16 @@
 library(tidyverse)
 
 # read in dataframe ----
-read_path <- paste0("./data/health/gan_episodes_of_care.txt")
+getwd()
+#read_path <- paste0("./data/health/gan_episodes_of_care.txt")
+# test path
+read_path <- paste0("./data/health/oregon_subset.txt")
 # read in with fread
 oregon_df <- data.table::fread(read_path, sep = "|",
   colClasses = c(rep("character", 72)), stringsAsFactors = FALSE)
+
+# print first 6 rows as check
+head(oregon_df)
 
 # set *NULL* and blank to NA
 # i may be able to integrate this to the first part
@@ -22,6 +28,8 @@ oregon_df <- oregon_df %>%
 
 # read in Rdata icd9 outcome vectors and limit to asthma
 load("./data/health/outcome_list.RData")
+# print outcomes list as check
+outcome_list
 
 # subset to asthma icd9
 asthma_icd9 <- pluck(outcome_icd9_list, "asthma")
@@ -40,8 +48,15 @@ saba_id <- oregon_df %>%
   select(personkey) %>% 
   as_vector()
 
+# print 6 obs of ids
+head(asthma_id)
+tail(saba_id)
+
 # bind two id vectors together
 asthma_saba_personkey <- unique(c(asthma_id, saba_id))
+
+head(asthma_saba_personkey)
+tail(asthma_saba_personkey)
 
 # filter large dataset to persons with asthma or saba fills ------
 # creating persons at risk (with asthma or a saba fill) for adverse events
@@ -51,10 +66,12 @@ asthma_saba_at_risk <- oregon_df %>%
   mutate(visit_type = ifelse(dx1 %in% asthma_icd9, "asthma", 
     ifelse(ndc %in% saba_ndc, "saba", "other")))
 
+head(asthma_saba_at_risk)
+
 # save file
 write_path <- paste0("./data/health/2013-asthma_saba_cohort.csv")
 write_csv(asthma_saba_at_risk, write_path)
-
+list.files()
 
 
 

### 2017-12-05 r_scripts/data_wrangle/asthma_cohort.R
fixed error in server code.

@@ -15,21 +15,17 @@ getwd()
 read_path <- paste0("./data/health/oregon_subset.txt")
 # read in with fread
 oregon_df <- data.table::fread(read_path, sep = "|",
-  colClasses = c(rep("character", 72)), stringsAsFactors = FALSE)
+  colClasses = c(rep("character", 72)), stringsAsFactors = FALSE) %>% 
+  # replace *NULL* with NA
+  mutate_all(funs(replace(., .== "*NULL*" | . == "", NA)))
 
 # print first 6 rows as check
 head(oregon_df)
 
-# set *NULL* and blank to NA
-# i may be able to integrate this to the first part
-oregon_df <- oregon_df %>% 
-  # replace *NULL* with NA
-  mutate_all(funs(replace(., .== "*NULL*" | . == "", NA)))
-
 # read in Rdata icd9 outcome vectors and limit to asthma
 load("./data/health/outcome_list.RData")
 # print outcomes list as check
-outcome_list
+class(outcome_icd9_list)
 
 # subset to asthma icd9
 asthma_icd9 <- pluck(outcome_icd9_list, "asthma")

### 2017-12-05 r_scripts/data_wrangle/asthma_cohort.R
fixed list files check.

@@ -67,7 +67,7 @@ head(asthma_saba_at_risk)
 # save file
 write_path <- paste0("./data/health/2013-asthma_saba_cohort.csv")
 write_csv(asthma_saba_at_risk, write_path)
-list.files()
+list.files("./data/health/")
 
 
 

### 2017-12-05 r_scripts/data_wrangle/asthma_cohort.R
updated path to full data and ready to test on server.

@@ -10,9 +10,11 @@ library(tidyverse)
 
 # read in dataframe ----
 getwd()
-#read_path <- paste0("./data/health/gan_episodes_of_care.txt")
+
+# path 
+read_path <- paste0("./data/health/gan_episodes_of_care.txt")
 # test path
-read_path <- paste0("./data/health/oregon_subset.txt")
+#read_path <- paste0("./data/health/oregon_subset.txt")
 # read in with fread
 oregon_df <- data.table::fread(read_path, sep = "|",
   colClasses = c(rep("character", 72)), stringsAsFactors = FALSE) %>% 
@@ -65,8 +67,10 @@ asthma_saba_at_risk <- oregon_df %>%
 head(asthma_saba_at_risk)
 
 # save file
-write_path <- paste0("./data/health/2013-asthma_saba_cohort.csv")
+write_path <- paste0("./data/health/2013-oregon_asthma_cohort.csv")
 write_csv(asthma_saba_at_risk, write_path)
+
+# list files as final check
 list.files("./data/health/")
 
 

### 2017-12-05 bash_scripts/r_submission_script.sh
added unlimited memmory command to bash script to prevent system kill

@@ -3,11 +3,14 @@
 #PBS -l nodes=1:ppn=1
 #PBS -W group_list=pierce_group
 #PBS -q batch
+#PBS -M rgan@colostate.edu
+#PBS -m abe
+
+# set memory to unlimited
+ulimit -s unlimited
 
 # set command directory
 cd /home/ryangan/local_git_repo/oregon_wildfire/
-# print directory
-pwd
 
 # export R library
 export R_LIB=~/R/x86_64-pc-linux-gnu-library/3.3

### 2017-12-05 r_scripts/data_wrangle/asthma_cohort.R
Spliting out asthma cohort in to two parts; first to identify those with asthnma, second will be to subset out those with asthma.

@@ -5,22 +5,39 @@
 # R Version: 3.4
 # ------------------------------------------------------------------------------
 
-# load library tidyverse
+# note 2017-12-01: memory issues with this process; going to have to chunk the 
+# reads in to smaller parts
+# oregon txt file has 77069321 rows including header
+
+# load library tidyverse -----
 library(tidyverse)
 
-# read in dataframe ----
-getwd()
-
 # path 
-read_path <- paste0("./data/health/gan_episodes_of_care.txt")
+#read_path <- paste0("./data/health/gan_episodes_of_care.txt")
 # test path
-#read_path <- paste0("./data/health/oregon_subset.txt")
+read_path <- paste0("./data/health/oregon_subset.txt")
+
+# read sections dataframe -------
+# n_start <- seq(from=1, to=77069321, by=200000)
+n_start <- seq(from=1, to=10000, by = 100)
+n_length <- rep(100, each=length(n_start))
+# make a dataframe to apply to 
+n_read <- data.frame(cbind(n_start,n_length))
+# I will use fread function from data.table to read in chunks at a time and
+# it doesn't matter if the read length is not exact
+
+# read first line and get column names
+df_col_names <- data.table::fread(read_path, nrows = 0, header = T, 
+                      colClasses = rep("character", 72))
+
+col_names <- colnames(df_col_names) # output column names
+
 # read in with fread
 oregon_df <- data.table::fread(read_path, sep = "|",
-  colClasses = c(rep("character", 72)), stringsAsFactors = FALSE) %>% 
-  # replace *NULL* with NA
-  mutate_all(funs(replace(., .== "*NULL*" | . == "", NA)))
+  colClasses = c(rep("character", 72)), stringsAsFactors = FALSE) 
 
+
+# already written ------
 # print first 6 rows as check
 head(oregon_df)
 

### 2017-12-05 r_scripts/data_wrangle/asthma_personkey_identifier.R

@@ -1,0 +1,80 @@
+
+# load libraries -----
+# parallel pacakges
+library(parallel) # package for parallel computations
+# tidyverse
+library(tidyverse)
+
+# assign read_path for oregon data and load asthma icd9 -----
+read_path <- paste0("./data/health/gan_episodes_of_care.txt")
+# test path
+#read_path <- paste0("./data/health/oregon_subset.txt")
+
+# read in Rdata icd9 outcome vectors and limit to asthma icd9s and saba ndcs
+load("./data/health/outcome_list.RData")
+# subset to asthma icd9
+asthma_icd9 <- pluck(outcome_icd9_list, "asthma")
+# print asthma_icd9 as a check
+print(asthma_icd9)
+
+# read sections dataframe -------
+# there are 77069321 rows (including header) in the Oregon APAC
+# I am going to read 200000 rows at a time in parallel to find unique persons
+# with an asthma icd9 code in any of the dx variables
+# define number of lines
+read_length <- 200000
+n_start <- seq(from=1, to=77069321, by = read_length)
+
+# read first line and get column names
+df_col_names <- data.table::fread(read_path, nrows = 0, header = T, 
+                                  colClasses = rep("character", 72))
+col_names <- colnames(df_col_names) # output column names
+
+# set up parallel computing ----
+# find number of processing nodes (cores)
+cores <- detectCores() # should be 16 per node
+# make cluster/node
+cl <- makeCluster(cores)
+
+# load packages on each processor of the node/cluster
+clusterCall(cl, function() library(tidyverse))
+# export read path to each core
+clusterExport(cl, c("read_path", "read_length", "n_start", "col_names"), 
+              envir = .GlobalEnv)
+
+start_time <- proc.time()
+# run in parallel -----
+# read in rows of the oregon file
+personkey_list <- parSapply(n_start, function(x){
+  df <- data.table::fread(read_path, sep = "|", header = F,
+                          # read set number of rows and skip what has already been read
+                          nrows=read_length, skip=x, 
+                          colClasses = rep("character", 72))
+  # assign column names
+  colnames(df) <- col_names
+  # filter df to subset of asthma icd9s in any dx column
+  asthma_id <- df %>% 
+    filter_at(vars(contains("dx")), any_vars(. %in% asthma_icd9)) %>% 
+    select(personkey) %>% 
+    as_vector() 
+  # return vector of personkeys with asthma ids
[…]

### 2017-12-05 bash_scripts/r_submission_script.sh
updated number of processors per node to 16.

@@ -1,6 +1,6 @@
 #!/bin/bash
 #PBS -N ryan_rjob
-#PBS -l nodes=1:ppn=1
+#PBS -l nodes=1:ppn=16
 #PBS -W group_list=pierce_group
 #PBS -q batch
 #PBS -M rgan@colostate.edu

### 2017-12-05 bash_scripts/r_submission_script.sh
updated file to execute.

@@ -16,5 +16,5 @@ cd /home/ryangan/local_git_repo/oregon_wildfire/
 export R_LIB=~/R/x86_64-pc-linux-gnu-library/3.3
 
 # run rscript
-Rscript --vanilla ./r_scripts/data_wrangle/asthma_cohort.R
+Rscript --vanilla ./r_scripts/data_wrangle/asthma_personkey_identifier.R
 

### 2017-12-05 r_scripts/data_wrangle/asthma_personkey_identifier.R
forgot to register cluster in parSapply function

@@ -45,7 +45,7 @@ clusterExport(cl, c("read_path", "read_length", "n_start", "col_names"),
 start_time <- proc.time()
 # run in parallel -----
 # read in rows of the oregon file
-personkey_list <- parSapply(n_start, function(x){
+personkey_list <- parSapply(cl, n_start, function(x){
   df <- data.table::fread(read_path, sep = "|", header = F,
                           # read set number of rows and skip what has already been read
                           nrows=read_length, skip=x, 

### 2017-12-05 r_scripts/data_wrangle/asthma_personkey_identifier.R
forgot to register asthma_icd9 to cluster.

@@ -39,8 +39,9 @@ cl <- makeCluster(cores)
 # load packages on each processor of the node/cluster
 clusterCall(cl, function() library(tidyverse))
 # export read path to each core
-clusterExport(cl, c("read_path", "read_length", "n_start", "col_names"), 
-              envir = .GlobalEnv)
+clusterExport(cl, 
+  c("read_path", "read_length", "n_start", "col_names", "asthma_icd9"), 
+  envir = .GlobalEnv)
 
 start_time <- proc.time()
 # run in parallel -----

### 2017-12-05 r_scripts/data_wrangle/asthma_personkey_identifier.R
forgot to change personkey_id to personkey_list...

@@ -71,7 +71,7 @@ stop_time
 
 # create csv of personkey IDs with at least one diagnosis of asthma ----
 # unlist the list of personkeys with asthma and find the unique observations
-asthma_personkey <- as_data_frame(unique(unlist(personkey_id))) %>% 
+asthma_personkey <- as_data_frame(unique(unlist(personkey_list))) %>% 
   rename(personkey = value)
 
 # print first keys

### 2017-12-06 r_scripts/data_wrangle/asthma_cohort.R
updated asthma_cohort script to run in parallel.

@@ -1,94 +1,94 @@
 # ------------------------------------------------------------------------------
-# Title: Creation of asthma and beta2 agonist cohort
+# Title: Creation asthma cohort using Oregon APAC
 # Author: Ryan Gan
-# Date Created: 2017-11-29
+# Date Created: 2017-12-05
 # R Version: 3.4
 # ------------------------------------------------------------------------------
 
-# note 2017-12-01: memory issues with this process; going to have to chunk the 
-# reads in to smaller parts
-# oregon txt file has 77069321 rows including header
-
 # load library tidyverse -----
+# parallel pacakges
+library(parallel) 
+# tidyverse
 library(tidyverse)
 
-# path 
-#read_path <- paste0("./data/health/gan_episodes_of_care.txt")
-# test path
+# setup ----
+
+# read vector of personkeys with asthma
+vector_path <- paste0("./data/health/2013-oregon_asthma_personkey.csv")
+asthma_personkey <- data.table::fread(vector_path, 
+                                      colClasses = "character")
+# convert to vector
+asthma_personkey_v <- asthma_personkey %>% 
+  as_vector()
+
+# read path for APAC dataframe
 read_path <- paste0("./data/health/oregon_subset.txt")
 
 # read sections dataframe -------
-# n_start <- seq(from=1, to=77069321, by=200000)
-n_start <- seq(from=1, to=10000, by = 100)
-n_length <- rep(100, each=length(n_start))
-# make a dataframe to apply to 
-n_read <- data.frame(cbind(n_start,n_length))
-# I will use fread function from data.table to read in chunks at a time and
-# it doesn't matter if the read length is not exact
+# there are 77069321 rows (including header) in the Oregon APAC
+# I am going to read 200000 rows at a time in parallel to find unique persons
+# with an asthma icd9 code in any of the dx variables
+# define number of lines
+read_length <- 200000
+n_start <- seq(from=1, to=77069321, by = read_length)
+
+# test set
+# read_length <- 100
+# n_start <- seq(from=1, to=10000, by = read_length)
 
 # read first line and get column names
 df_col_names <- data.table::fread(read_path, nrows = 0, header = T, 
-                      colClasses = rep("character", 72))
-
+                                  colClasses = rep("character", 72))
[…]

### 2017-12-06 r_scripts/data_wrangle/asthma_personkey_identifier.R

@@ -1,3 +1,9 @@
+# ------------------------------------------------------------------------------
+# Title: Finding unique persons with a diagnosis of asthma in Oregon APAC
+# Author: Ryan Gan
+# Date Created: 2017-12-05
+# R Version: 3.4
+# ------------------------------------------------------------------------------
 
 # load libraries -----
 # parallel pacakges
@@ -7,8 +13,6 @@ library(tidyverse)
 
 # assign read_path for oregon data and load asthma icd9 -----
 read_path <- paste0("./data/health/gan_episodes_of_care.txt")
-# test path
-#read_path <- paste0("./data/health/oregon_subset.txt")
 
 # read in Rdata icd9 outcome vectors and limit to asthma icd9s and saba ndcs
 load("./data/health/outcome_list.RData")

### 2017-12-06 bash_scripts/r_submission_script.sh
changed bash script to run asthma_cohort script.

@@ -16,5 +16,5 @@ cd /home/ryangan/local_git_repo/oregon_wildfire/
 export R_LIB=~/R/x86_64-pc-linux-gnu-library/3.3
 
 # run rscript
-Rscript --vanilla ./r_scripts/data_wrangle/asthma_personkey_identifier.R
+Rscript --vanilla ./r_scripts/data_wrangle/asthma_cohort.R
 

### 2017-12-06 r_scripts/data_wrangle/asthma_cohort.R
fixed read path.

@@ -12,7 +12,6 @@ library(parallel)
 library(tidyverse)
 
 # setup ----
-
 # read vector of personkeys with asthma
 vector_path <- paste0("./data/health/2013-oregon_asthma_personkey.csv")
 asthma_personkey <- data.table::fread(vector_path, 
@@ -22,7 +21,9 @@ asthma_personkey_v <- asthma_personkey %>%
   as_vector()
 
 # read path for APAC dataframe
-read_path <- paste0("./data/health/oregon_subset.txt")
+read_path <- paste0("./data/health/gan_episodes_of_care.txt")
+# test path
+#read_path <- paste0("./data/health/oregon_subset.txt")
 
 # read sections dataframe -------
 # there are 77069321 rows (including header) in the Oregon APAC
@@ -30,7 +31,7 @@ read_path <- paste0("./data/health/oregon_subset.txt")
 # with an asthma icd9 code in any of the dx variables
 # define number of lines
 read_length <- 200000
-n_start <- seq(from=1, to=77069321, by = read_length)
+n_start <- seq(from=1, to=77069321, by=read_length)
 
 # test set
 # read_length <- 100
@@ -89,9 +90,4 @@ head(asthma_cohort[,1:10])
 # save as csv ----
 write_path <- paste0("./data/health/2013-oregon_asthma_cohort.csv")
 data.table::fwrite(asthma_cohort, write_path)
-
-
-
-
-
 

### 2017-12-06 r_scripts/analysis/asthma_cohort_descriptive_epi.Rmd
started asthma cohort exploration.

--- 2017-11-22 r_scripts/analysis/oregon_cvd_dl_over65_results.Rmd
@@
 ---
-title: Distributed lag effects of smoke on cardiovascular disease in persons over 65
+title: "Asthma Cohort Descriptive Epidemiology"
 author: "Ryan Gan"
-date: "2017-11-20"
+date: "2017-12-06"
 output: html_document
@@
 
-Extensive look at possible distributed lag relationships for Cardiovascular endpoints in the over 65 age group.
-
-Packages used: tidyverse, survival. Setting markdown options.
+Descriptive statistics of the assembled asthma cohort.
+This cohort is made up of all billing records for procedures or pharamacy fills for anyone in the Oregon APAC data that had been billed for a diagnosis of asthma at any visit in 2013.
 
@@
 ```
-
-Import CVD data.
-
-```{r health data import}
-# path for health dataframe
-health_path <- "../../data/health/"
-# importing time-stratified case-crossover dataframe
-casecross_files <- list.files(path=health_path, pattern="*casecross*")
-
-# reorder file names in way I'd like it to be analyzed and only importing cvd
-# also ditching broken arm
-casecross_order <- casecross_files[c(7,2,5,8,9,10)]
-
-# extract the name of the dataframe after study type
-file_names <- stringr::str_extract(casecross_order, 
-  pattern="(?<=casecross_)\\w+")
-
-# read in all the files in to a list
-outcome_df_list <- casecross_order %>%  
-  purrr::map(~read_csv(paste0(health_path, .)))
-
-# name dataframes
-names(outcome_df_list) <- file_names
-
-# name list elements in order I'd like to present in document
-outcome_names <- c("Cardiovascular Disease", "Arrhythmia", 
-  "Cerebrovascular Disease", "Heart Failure", "Ischemic Heart Disease", 
-  "Myocardial Infarction")
-```
-
-Creating a fit function to find the degrees of freedom for the pm spline
-
-```{r distributed lag functions}
-
-# fit function
-dl_fit <- function(data, pm_df, temp_df){
-  # i want to remove this part eventually ------
-  # limit to complete cases
-  complete_data <- data %>% 
-    # filter to complete case for smoke
-    filter(!is.na(geo_smk_pm_lag7_zip))
-
-  # output matrix of gwr values
-  pm_matrix <- complete_data %>% 
-    select(contains("geo_smk_pm")) %>%
-    # remove anything with m in it; will go back in original code and not create 
-    # this to begin with
-    select(-contains("_m")) %>% 
-    # divide exposure values by 10 units to interpret on 10 ug/m^3 scale
-    mutate_all(funs(./10)) %>% 
-    # convert to matrix
-    as.matrix()
-
-  # output temp matrix
-  temp_matrix <- complete_data %>% 
-    select(contains("wrf_temp")) %>%
-    # remove anything with m in it; will go back in original code and not create 
-    # this to begin with
-    select(-contains("_m")) %>% 
-    # convert to matrix
-    as.matrix()
-  # end part to remove------
-
-    # create temp basis
-    temp_b <- splines::ns(0:(ncol(temp_matrix)-1), df=temp_df, intercept=T)
-    temp_basis <- temp_matrix %*% temp_b
-
-      # create list of lengh pm_df
-      mod_aic <- sapply(pm_df, function(x){
-        pm_b <- splines::ns(0:(ncol(pm_matrix)-1), df=x, intercept=T)
-        # create pm basis
-        pm_basis <- pm_matrix %*% pm_b
-        # run model
-        mod <- clogit(outcome ~ pm_basis + temp_basis + strata(personkey),
-          data = complete_data)
-        aic <- AIC(mod)
-        return(aic)
-      }) # end sapply
-
-    # find minimum aic
-    pm_df_aic <- data.frame(pm_df, mod_aic)
-    min_df_aic <- subset(pm_df_aic, mod_aic == min(mod_aic))
-    return(min_df_aic)
-  } # end fit option
-
-dl_fit(data = hf65, pm_df = c(3:5), temp_df=2)
-
-# custom distributed lag function 
-distributed_lag <- function(data, pm_df, temp_df, cumulative=T){
-  # i want to remove this part eventually ------
-  # limit to complete cases
-  complete_data <- data %>% 
-    # filter to complete case for smoke
-    filter(!is.na(geo_smk_pm_lag7_zip))
-
-  # output matrix of gwr values
-  pm_matrix <- complete_data %>% 
-    select(contains("geo_smk_pm")) %>%
-    # remove anything with m in it; will go back in original code and not create 
-    # this to begin with
-    select(-contains("_m")) %>% 
-    # divide exposure values by 10 units to interpret on 10 ug/m^3 scale
-    mutate_all(funs(./10)) %>% 
-    # convert to matrix
-    as.matrix()
-
-  # output temp matrix
-  temp_matrix <- complete_data %>% 
-    select(contains("wrf_temp")) %>%
-    # remove anything with m in it; will go back in original code and not create 
-    # this to begin with
-    select(-contains("_m")) %>% 
-    # convert to matrix
-    as.matrix()
-  # end part to remove------
-
-  # calculation of basis ----
-  # define basis using natural spline function from "splines"" package
-  pm_b <- splines::ns(0:(ncol(pm_matrix)-1), df=pm_df, intercept=T)
-  # create pm basis
-  pm_basis <- pm_matrix %*% pm_b
-  # create temp basis
-  temp_b <- splines::ns(0:(ncol(temp_matrix)-1), df=temp_df, intercept=T)
-  temp_basis <- temp_matrix %*% temp_b
-  # fit model with basis ----
-  mod <- clogit(outcome ~ pm_basis + temp_basis + strata(personkey),
-    data = complete_data)
-
-  # calculate estimates ----
-  # output pm basis parameters
-  dl_parms <- broom::tidy(mod) %>% 
-    filter(stringr::str_detect(term, "pm")) %>% 
-    select(estimate) %>% 
-    as_vector()
-  # estimate distributed lag values for each day
-  estimate <-  pm_b %*% dl_parms
-  # covariance matrix for knots 
-  # fix the matrix
-  cov_matrix <- as.matrix(vcov(mod))[1:pm_df,1:pm_df]
-  # estimate variance of spline
-  variance <- pm_b %*% cov_matrix %*% t(pm_b)
-  # estimate standard error
-  se <- sqrt(diag(dl_var))
-  # calculate lower and upper bound
-  lower95 <- estimate+(se*qnorm(1-0.975))
-  upper95 <- estimate+(se*qnorm(0.975))
-  # time variable
-  time <- ((rep(1:length(estimate))-1))
-
-    if(cumulative==T) {
-      type <- rep("cumulative", times = length(estimate))
-      # sequential cumulative estimate
-      cumulative_estimate <- sapply(seq_along(estimate), function(x){
-        sum(estimate[1:x])
-      })
-      # stderr cumulative effect
-      cumulative_se <- sapply(seq_along(estimate), function(x){
-        sqrt(sum(dl_var[1:x,1:x]))
-      })
-      # cumulative 95CI
-      lower95 <- cumulative_estimate+(cumulative_se*qnorm(1-0.975))
-      upper95 <- cumulative_estimate+(cumulative_se*qnorm(0.975))
-      # return dataframe
-      return_estimate <- data.frame(type, time, exp(cumulative_estimate), 
-                                    exp(lower95), exp(upper95)) 
-      # assign column names
-      colnames(return_estimate) <- c("type","time","estimate","lower95","upper95")
-
-      return(return_estimate)
-    } else { # if cumulative is not true, or false, return distributed lag est
-    type <- rep("lag", times = length(estimate))
-    # return dataframe
-    return_estimate <- data.frame(type, time, exp(estimate), 
-                                  exp(lower95), exp(upper95)) 
-    # assign column names
-    colnames(return_estimate) <- c("type","time","estimate","lower95","upper95") 
-    # return estimate  
-    return(return_estimate)
-    } # end of if
-} 
-```
-
-## Heart failure in over 65 group
-
-```{r subset data}
-hf65 <- outcome_df_list[[4]] %>% 
-  filter(age_ind == 2)
-
-cumulative <- distributed_lag(hf65, pm_df = 4, temp_df = 2, cumulative = T)
-cumulative
-
-#plot 
-cumulative_plot <- ggplot(cumulative, aes(x=time, y=estimate)) +
-  geom_line(colour = "#046C9A", size = 1) +
-  geom_ribbon(aes(ymin = lower95, ymax = upper95), 
-              fill = "#ABDDDE", alpha = 0.5) + 
-  scale_y_continuous(limits = c(0.9, 1.25)) +
-  scale_x_continuous(breaks = c(seq(0,7, by=1))) +
-  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
-  ylab(expression("Odd ratio for a 10 ug/m^3 increase smoke PM"[2.5])) +
-  xlab("Lagged Days") +
-  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    strip.text = element_text(size = 10),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    strip.background = element_rect(colour=NA, fill=NA))
-
-cumulative_plot
-
-dl <- distributed_lag(hf65, pm_df = 3, temp_df = 2, cumulative = F)
-dl
-
-dl_plot <- ggplot(dl, aes(x=time, y=estimate)) +
-  geom_line(colour = "#046C9A", size = 1) +
-  geom_ribbon(aes(ymin = lower95, ymax = upper95), 
-              fill = "#ABDDDE", alpha = 0.5) + 
-  scale_y_continuous(limits = c(0.9, 1.25)) +
-  scale_x_continuous(breaks = c(seq(0,7, by=1))) +
-  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
-  ylab(expression("Odd ratio for a 10 ug/m^3 increase smoke PM"[2.5])) +
-  xlab("Lagged Days") +
-  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    strip.text = element_text(size = 10),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    strip.background = element_rect(colour=NA, fill=NA))
-
-dl_plot
-
-```
-
-## Myocardial infarction in over 65 group
-
-```{r mi data}
-mi65 <- outcome_df_list[[6]] %>% 
-  filter(age_ind == 2)
-
-cumulative <- distributed_lag(mi65, pm_df = 2, temp_df = 2, cumulative = T)
-cumulative
-
-#plot 
-cumulative_plot <- ggplot(cumulative, aes(x=time, y=estimate)) +
-  geom_line(colour = "#046C9A", size = 1) +
-  geom_ribbon(aes(ymin = lower95, ymax = upper95), 
-              fill = "#ABDDDE", alpha = 0.5) + 
-  scale_y_continuous(limits = c(0.7, 1.2)) +
-  scale_x_continuous(breaks = c(seq(0,7, by=1))) +
-  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
-  ylab(expression("Odd ratio for a 10 ug/m^3 increase smoke PM"[2.5])) +
-  xlab("Lagged Days") +
-  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    strip.text = element_text(size = 10),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    strip.background = element_rect(colour=NA, fill=NA))
-
-cumulative_plot
-
-dl <- distributed_lag(mi65, pm_df = 3, temp_df = 2, cumulative = F)
-dl
-
-dl_plot <- ggplot(dl, aes(x=time, y=estimate)) +
-  geom_line(colour = "#046C9A", size = 1) +
-  geom_ribbon(aes(ymin = lower95, ymax = upper95), 
-              fill = "#ABDDDE", alpha = 0.5) + 
-  scale_y_continuous(limits = c(0.9, 1.25)) +
-  scale_x_continuous(breaks = c(seq(0,7, by=1))) +
-  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
-  ylab(expression("Odd ratio for a 10 ug/m^3 increase smoke PM"[2.5])) +
-  xlab("Lagged Days") +
-  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    strip.text = element_text(size = 10),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    strip.background = element_rect(colour=NA, fill=NA))
-
-dl_plot
-
-```

### 2017-12-07 r_scripts/analysis/asthma_cohort_descriptive_epi.Rmd
working on descriptives for asthma cohort

@@
 library(tidyverse) # general data wrangle
-library(survival) # conditional logistic regression
+
+# load outcomes vectors
+load("./data/health/outcome_list.RData")
+# subset asthma and saba vector
+asthma_icd9 <- pluck(outcome_icd9_list, "asthma")
+saba_ndc <- pluck(outcome_icd9_list, "saba")
 
@@
 ```
+
+Read cohort dataset (reading in a small chunk for now.)
+
+```{r cohort_import}
+# read path
+read_path <- "./data/health/2013-oregon_asthma_cohort.csv"
+# read in asthma cohort (reading a million rows for now)
+asthma_cohort <- data.table::fread(read_path, 
+  # read options
+  header = T, nrows = 10000, colClasses = rep("character", 72)) %>% 
+  # replace *NULL* with NA
+  mutate_all(funs(replace(., .== "*NULL*" | . == "", NA))) %>% 
+  # create visit type indicator 
+  mutate(visit_type = as.factor(
+    case_when((dx1 %in% asthma_icd9) ~ "dx_asthma_primary",
+    # asthma but not as a primary
+    !(dx1 %in% asthma_icd9) & (dx2 %in% asthma_icd9 | dx3 %in% asthma_icd9 | 
+      dx4 %in% asthma_icd9 | dx5 %in% asthma_icd9 | dx6 %in% asthma_icd9 | 
+      dx7 %in% asthma_icd9 | dx8 %in% asthma_icd9 | dx9 %in% asthma_icd9 | 
+      dx10 %in% asthma_icd9 | dx11 %in% asthma_icd9 | dx12 %in% asthma_icd9 | 
+      dx13 %in% asthma_icd9) ~ "dx_asthma_not_primary",
+    # other dx 
+    !(dx1 %in% asthma_icd9 | dx2 %in% asthma_icd9 | dx3 %in% asthma_icd9 | 
+      dx4 %in% asthma_icd9 | dx5 %in% asthma_icd9 | dx6 %in% asthma_icd9 | 
+      dx7 %in% asthma_icd9 | dx8 %in% asthma_icd9 | dx9 %in% asthma_icd9 | 
+      dx10 %in% asthma_icd9 | dx11 %in% asthma_icd9 | dx12 %in% asthma_icd9 | 
+      dx13 %in% asthma_icd9) & is.na(ndc) ~ "dx_other",
+    # ndc code
+    (ndc %in% saba_ndc) ~ "pharm_saba",
+    !is.na(ndc) & !(ndc %in% saba_ndc) ~ "pharm_other")))
+```

### 2017-12-07 r_scripts/analysis/asthma_cohort_descriptive_epi.Rmd
working on summary statistics.

--- 2017-12-06 r_scripts/analysis/asthma_cohort_descriptive_epi.Rmd
@@
 
-## Overview
+### Document Purpose
 
 Descriptive statistics of the assembled asthma cohort.
-This cohort is made up of all billing records for procedures or pharamacy fills for anyone in the Oregon APAC data that had been billed for a diagnosis of asthma at any visit in 2013.
+This cohort is made up of all billing records for procedures or pharamacy fills related to cariopulmonary outcomes for anyone in the Oregon APAC data that had been billed for a diagnosis of asthma at any visit in 2013.
 
@@
 library(tidyverse) # general data wrangle
-library(survival) # conditional logistic regression
+
+# load outcomes vectors
+load("./data/health/outcome_list.RData")
+# subset asthma and saba vector
+asthma_icd9 <- pluck(outcome_icd9_list, "asthma")
+saba_ndc <- pluck(outcome_icd9_list, "saba")
+
+# read in place of service csv file
+place_of_service <- read_csv("./data/health/2013-oregon_pos.csv")
 
@@
 ```
+
+Read cohort dataset (reading in a small chunk for now.)
+
+```{r cohort_import}
+# read path
+read_path <- "./data/health/2013-oregon_asthma_cohort.csv"
+# read in asthma cohort (reading a million rows for now)
+asthma_cohort <- data.table::fread(read_path, 
+  # read options
+  header = T, nrows = 10000, colClasses = rep("character", 72)) %>% 
+  # replace *NULL* with NA
+  mutate_all(funs(replace(., .== "*NULL*" | . == "", NA))) %>% 
+  # create visit type indicator 
+  mutate(visit_type = as.factor(
+    case_when((dx1 %in% asthma_icd9) ~ "dx_asthma_primary",
+    # asthma but not as a primary
+    !(dx1 %in% asthma_icd9) & (dx2 %in% asthma_icd9 | dx3 %in% asthma_icd9 | 
+      dx4 %in% asthma_icd9 | dx5 %in% asthma_icd9 | dx6 %in% asthma_icd9 | 
+      dx7 %in% asthma_icd9 | dx8 %in% asthma_icd9 | dx9 %in% asthma_icd9 | 
+      dx10 %in% asthma_icd9 | dx11 %in% asthma_icd9 | dx12 %in% asthma_icd9 | 
+      dx13 %in% asthma_icd9) ~ "dx_asthma_not_primary",
+    # other dx 
+    !(dx1 %in% asthma_icd9 | dx2 %in% asthma_icd9 | dx3 %in% asthma_icd9 | 
+      dx4 %in% asthma_icd9 | dx5 %in% asthma_icd9 | dx6 %in% asthma_icd9 | 
+      dx7 %in% asthma_icd9 | dx8 %in% asthma_icd9 | dx9 %in% asthma_icd9 | 
+      dx10 %in% asthma_icd9 | dx11 %in% asthma_icd9 | dx12 %in% asthma_icd9 | 
+      dx13 %in% asthma_icd9) & is.na(ndc) ~ "dx_other",
+    # ndc code
+    (ndc %in% saba_ndc) ~ "pharm_saba",
+    !is.na(ndc) & !(ndc %in% saba_ndc) ~ "pharm_other"))) %>% 
+  # convert date variables to dates
+  mutate_at(vars(contains("date")), funs(as.Date(.,format="%Y-%m-%d"))) %>% 
+  mutate(los = as.numeric(los)) %>% 
+  # join in place of service text
+  left_join(place_of_service, by = "pos") %>% 
+  # simpler place of service grouping
+  mutate(pos_simple = as.factor(case_when(pos == "01" ~ "Pharmacy",
+    pos == "03" ~ "School",
+    pos == "11" ~ "Office",
+    pos == "20" ~ "Urgent Care", 
+    pos == "21" ~ "Inpatient Hospital",
+    pos == "22" ~ "Outpatient Hospital",
+    pos == "23" ~ "Emergency Room Hospital",
+    pos == "41" | pos == "42" ~ "Ambulance",
+    pos == "71" | pos == "72" ~ "Public Health Clinic",
+    pos == "49" ~ "Independent Clinic", 
+    TRUE ~ "Other")))
+```
+
+### Observation Types
+
+Description of number of observation types.
+I have not yet characterized the individuals; these are summaries of all observations.
+I've categorized observations in the previous step in to five categories as follows:
+- dx_asthma_primary: This is an observation where I've identified a medical procedure in the primary dx category that is a ICD-9 code of asthma. 
+- dx_asthma_not_primary: This is an observation where the primary medical procedure is not asthma, but at least one other medical procedure was for asthma.
+- dx_other: This is an observation that is a medical procedure, but the primary or any other procedure was not asthma.
+- pharm_saba: This ia an observation where a short-acting beta-2 agonist (SABA) prescription was billed
+
+```{r vis_type_summary}
+visit_type_summary <- asthma_cohort %>% 
+  group_by(visit_type) %>% 
+  summarise(n = n()) %>% 
+  mutate(percent = round((n/nrow(asthma_cohort)),3)*100)
+
+# kable table
+knitr::kable(visit_type_summary, caption = paste0("Visit type number and",
+                                                  " percentage."))
+```
+
+It appears that only a small percentage of visits for persons identified to have asthma were billed for asthma procedures. Same goes for pharmacy fills.
+
+#### Place of Service
+
+Evaluating the number of observations and percentages for place of service.
+Here is a table of all the places of service.
+A good majority of serv ices take place in the "office" location, which is a location where health professionals provide health care (Drs'.
+Office).
+
+```{r service_table}
+n_pos <- asthma_cohort %>% 
+  group_by(service_place) %>% 
+  summarise(n = n()) %>% 
+  mutate(percent = round((n/nrow(asthma_cohort)),3)*100)
+
+knitr::kable(n_pos, caption="Place of service look-up table")
+```
+
+There are a lot of visit types that will rarely be coded and won't be informative to us, so I've grouped them up on what I think would be a reasonable representation of major places of visits relevant to the study question.
+I've limited to a subset of places of services and grouped some together, but I'm open to discussion.
+I've removed the pharmacy records for this plot since the place of service is just pharmacy and it's not informative.
+
+```{r pos_visit_type}
+pos_visit_type <- asthma_cohort %>% 
+  group_by(pos_simple, visit_type) %>% 
+  summarise(n_obs = n()) %>% 
+  left_join(select(visit_type_summary, -percent), by = "visit_type") %>% 
+  mutate(percent = round(n_obs/n, 3)*100) %>% 
+  filter(visit_type != "pharm_saba" & visit_type != "pharm_other")
+
+plot <- ggplot(pos_visit_type, aes(x=pos_simple, y=percent)) +
+  geom_col(fill="#355c7d") +
+  facet_wrap(~visit_type) +
+  theme_bw()+
+  theme(axis.text.x=element_text(angle=90,hjust=1, vjust=0.5))
+
+plot
+```
+
+#### Distribution of Dates
+
+Plotting the date distributions by visit type. Looks like multimodal distributions of visits types by date, where asthma as a primary or not primary has peaks in spring and fall. Same with SABA fills.
+```{r vis_type_date}
+plot <- ggplot(asthma_cohort, aes(x=todate)) +
+  geom_density(color = "#19547b") +
+  facet_wrap(~visit_type) + 
+  theme_minimal()
+
+plot
+```
+
+#### Length of Stay
+Plotting the length of stay.
+Pharmacy fills do not have any distribution of length of stay, but it was easier to just plot it then remove it :).
+Heavy right-skew for all diagnosis types.
+This is likely elderly people in places of service where they may actually stay for the year.
+It could also be an error.
+But most stays are no more than a month.
+This would change depending on age or place of service.
+
+```{r vis_type_los}
+plot <- ggplot(asthma_cohort, aes(x=los)) +
+  geom_density(color = "#19547b") +
+  facet_wrap(~visit_type) + 
+  theme_minimal()
+
+plot
+```
+
+### Individual-Level Descriptives
+
+#### Unique Persons
+
+As this is a cohort of persons that were identified to have a diagnosis of asthma, I'd like to know a little more about them as a group.
+First step is to count up the number of unique persons.
+First is a count of unique personkey ids, which represents the number of unique persons in this cohort.
+
+```{r n_persons}
+# length of unique person ids
+n_persons <- length(unique(asthma_cohort$personkey))
+# represents the number of persons
+n_persons
+```
+
+Summary statistics and distribution of visit type per person. Includes both pharmacy fills and medical visits, so the distribution will be right-skewed.
+
+```{r summary_persons}
+person_obs <- asthma_cohort %>% 
+  group_by(personkey, visit_type) %>% 
+  summarise(n_obs = n())
+
+summary_type <- person_obs %>% 
+  group_by(visit_type) %>% 
+  summarise(n = n(), mean_n = mean(n_obs), median_n = median(n_obs), 
+            min_n = min(n_obs), max_n = max(n_obs))
+
+# table of summary types
+knitr::kable(summary_type, caption = "Summary of observations by visit type.")
+```
+
+Plot of observation number distribution by visit type.
+```{r obs_dist_plot}
+plot <- ggplot(person_obs, aes(x=n_obs)) +
+  geom_density(color = "#19547b") +
+  facet_wrap(~visit_type, scales = "free_x") +
+  theme_minimal()
+plot
+```
+
+I think the observations that are dx_other or pharm_other make the dataset too large and for the analyses I'm considering, may have no practical value. But they are probably worth retaining for now.

### 2017-12-07 r_scripts/analysis/asthma_cohort_descriptive_epi.Rmd
filtered only to oregon.

@@
   header = T, nrows = 10000, colClasses = rep("character", 72)) %>% 
+  # filter to oregon
+  filter(STATE == "OR") %>% 
   # replace *NULL* with NA

### 2017-12-08 r_scripts/analysis/asthma_cohort_descriptive_epi.Rmd
updated agu figure for kate and updated asthma summary statistics for cohort.

--- 2017-12-07 r_scripts/analysis/asthma_cohort_descriptive_epi.Rmd
@@
 # load outcomes vectors
-load("./data/health/outcome_list.RData")
+load("../../data/health/outcome_list.RData")
 # subset asthma and saba vector
@@
 # read in place of service csv file
-place_of_service <- read_csv("./data/health/2013-oregon_pos.csv")
+place_of_service <- read_csv("../../data/health/2013-oregon_pos.csv")
 
@@
 # read path
-read_path <- "./data/health/2013-oregon_asthma_cohort.csv"
+read_path <- "../../data/health/2013-oregon_asthma_cohort.csv"
 # read in asthma cohort (reading a million rows for now)
@@
   # read options
-  header = T, nrows = 10000, colClasses = rep("character", 72)) %>% 
+  header = T, #nrows = 10000, # removing nrows read 
+  colClasses = rep("character", 72)) %>% 
+  # filter to oregon
+  filter(STATE == "OR") %>% 
   # replace *NULL* with NA
@@
   mutate_at(vars(contains("date")), funs(as.Date(.,format="%Y-%m-%d"))) %>%
-  mutate(los = as.numeric(los)) %>% 
+  # adding and converting some variables
+  mutate(los = as.numeric(los),
+         # create age variable
+         age = 2013 - as.numeric(yob),
+         # code in race text to numeric race
+         race_factor = case_when(race == 1 ~ "White",
+                                 race == 2 ~ "Black",
+                                 race == 3 ~ "American Indian",
+                                 race == 4 ~ "Asian",
+                                 race == 5 ~ "Pacific Islander",
+                                 race == 6 ~ "Other", 
+                                 race == 7 ~ "Two or more",
+                                 race == 9 ~ "Unknown"),
+         ethnicity = case_when(ethn == "1" ~ "Hispanic",
+                               ethn == "2" ~ "Not Hispanic",
+                               ethn == "3" ~ "Unknown")) %>% 
   # join in place of service text
@@
 
-I think the observations that are dx_other or pharm_other make the dataset too large and for the analyses I'm considering, may have no practical value. But they are probably worth retaining for now.
+#### Age Distribution
+
+To understand the distribution of demographic factors, I'm going to subset the first (by date) diagnosis of asthma visit observed to look at age, gender, and race/ethnicity distributions.
+When you subset out the NDC and other diagnoses types, the cohort is reduced drastically.
+I believe this would be fine and actually create a more well-defined cohort if it's subsetted to just asthma.
+
+Here is a count of number of persons with either a primary or other diagnosis of asthma in the cohort dataset. 
+```{r age_n}
+asthma_demographic <- asthma_cohort %>% 
+  filter(visit_type == "dx_asthma_primary" | 
+         visit_type == "dx_asthma_not_primary") %>% 
+  group_by(personkey) %>% 
+  arrange(todate) %>% 
+  filter(row_number()==1)
+
+nrow(asthma_demographic)
+```
+
+Summary statistics of age in persons with a medical visit related to asthma.
+```{r age_sum}
+summary(asthma_demographic$age)
+```
+
+Age distribution plot in persons with a medical visit related to asthma..
+```{r age_dist}
+# asthma age distribution plot
+plot <- ggplot(asthma_demographic, aes(x=age)) +
+  geom_density(color = "#19547b") +
+  theme_minimal()
+plot
+```
+
+#### Gender Distribution
+
+Gender n and percentage.
+```{r gender_stat}
+# count
+xtabs(~gender, asthma_demographic) 
+# percentage
+round(prop.table(xtabs(~gender, asthma_demographic)),3)*100
+```
+
+#### Race/Ethnicity Distribution
+
+Race n and percentage. A lot of persons did not report race.
+```{r race_stat}
+xtabs(~race_factor, asthma_demographic)
+round(prop.table(xtabs(~race_factor, asthma_demographic)),3)*100
+```
+
+Ethnicity n and percentage. Same deal. A lot of persons did not report ethnicity.
+```{r ethn_stat}
+xtabs(~ethnicity, asthma_demographic)
+round(prop.table(xtabs(~ethnicity, asthma_demographic)),3)*100
+```
+
+### Conclusions
+
+Overall, I think this cohort approach could work.I think the observations that are dx_other or pharm_other make the dataset too large, and may have no practical value for the analyses I'm considering.
+But they are probably worth retaining for now.
+
+Some ideas are we could further limit the study population to those with at least 2 visits of asthma to further increase our specificity that a person has asthma.
+We could then look at the range of severity of outcomes in these people, where the most severe would be an inpatient hospital stay for a primary diagnosis of asthma.
+Second-most severe could be an ER/urgent care admission or ambulance visit.
+Least-severe could be a SABA fill in a cohort time-to-event manner.
+I was thinking we could start time at either the start of the wildfire period or July, and could include people who have asthma dx for some time that year, but have been without a medical bill for at least a month prior to start of study date.
+
+For this cohort approach, I'd like to try out VanderWeele's E-value measure, which estimates the impact unmeasured confounders would have to have to explain the association.
+
+We could also look at time-series and case-cross over study designs too like Sheryl suggested.
+I think this is a good idea too, and hopefully these results would agree with our general hypotheses that smoke is bad.

### 2017-12-08 r_scripts/analysis/asthma_cohort_descriptive_epi.Rmd
added MSA in to descriptive summary.

@@
     (ndc %in% saba_ndc) ~ "pharm_saba",
-    !is.na(ndc) & !(ndc %in% saba_ndc) ~ "pharm_other"))) %>% 
-  # convert date variables to dates
-  mutate_at(vars(contains("date")), funs(as.Date(.,format="%Y-%m-%d"))) %>%
-  # adding and converting some variables
-  mutate(los = as.numeric(los),
+    !is.na(ndc) & !(ndc %in% saba_ndc) ~ "pharm_other")),
+    # adding names of metro area
+    metroarea = case_when(
+      MSA == 13460 ~ "Bend-Redmond",
+      MSA == 18700 ~ "Corvallis",
+      MSA == 21660 ~ "Eugene-Springfield",
+      MSA == 32780 ~ "Medford",
+      MSA == 38900 ~ "Portland-Vancouver-Beaverton",
+      MSA == 41420 ~ "Salem",
+      MSA == 41999 ~ "Not in MSA"),
+    los = as.numeric(los),
          # create age variable
@@
                                ethn == "3" ~ "Unknown")) %>% 
+  # convert date variables to dates
+  mutate_at(vars(contains("date")), funs(as.Date(.,format="%Y-%m-%d"))) %>%
   # join in place of service text
@@
 
+#### Metropolitan Statistical Area
+
+MSA n and percentage.
+```{r msa}
+xtabs(~metroarea, asthma_demographic)
+round(prop.table(xtabs(~metroarea, asthma_demographic)),3)*100
+```
+
 ### Conclusions

### 2017-12-08 r_scripts/analysis/markdown_server_execute.R
added server execution script.

@@ -1,0 +1,9 @@
+# ------------------------------------------------------------------------------
+# Title: Script to execute markdown files on a server
+# Author: Ryan Gan
+# Date Created: 2017-12-08
+# ------------------------------------------------------------------------------
+
+# call R markdown file
+rmarkdown::render('./r_scripts/analysis/asthma_cohort_descriptive_epi.Rmd')
+

### 2017-12-08 r_scripts/data_wrangle/asthma_cohort_fire_season.R
Wrote script to create a smaller asthma cohort that will be easier to analyze.

@@ -1,0 +1,110 @@
+# ------------------------------------------------------------------------------
+# Title: Subsetting asthma cohort to fire season 
+# Author: Ryan Gan
+# Date Created: 2017-12-08
+# ------------------------------------------------------------------------------
+
+# Description: This script subsets the asthma cohort to further reduce the 
+# number of observations to only asthma or saba observations to reduce the 
+# number of events that need to be evaluated. I also add/modify this dataset
+
+
+# library ----
+library(tidyverse)
+
+# read asthma cohort dataset and affiliated files ------
+# read path
+read_path <- "../../data/health/2013-oregon_asthma_cohort.csv"
+
+# load outcomes vectors
+load("../../data/health/outcome_list.RData")
+# subset asthma and saba vector
+asthma_icd9 <- pluck(outcome_icd9_list, "asthma")
+saba_ndc <- pluck(outcome_icd9_list, "saba")
+
+# read in place of service csv file
+place_of_service <- read_csv("./data/health/2013-oregon_pos.csv")
+
+# read in asthma cohort (reading a million rows for now)
+asthma_cohort <- data.table::fread(read_path, 
+  # read options
+  header = T, nrows = 100000, # removing nrows read 
+  colClasses = rep("character", 72)) %>% 
+  # filter to oregon
+  filter(STATE == "OR") %>% 
+  # replace *NULL* with NA
+  mutate_all(funs(replace(., .== "*NULL*" | . == "", NA))) %>% 
+  # create visit type indicator 
+  mutate(visit_type = as.factor(
+    case_when((dx1 %in% asthma_icd9) ~ "dx_asthma_primary",
+      # asthma but not as a primary
+      !(dx1 %in% asthma_icd9) & (dx2 %in% asthma_icd9 | dx3 %in% asthma_icd9 | 
+      dx4 %in% asthma_icd9 | dx5 %in% asthma_icd9 | dx6 %in% asthma_icd9 | 
+      dx7 %in% asthma_icd9 | dx8 %in% asthma_icd9 | dx9 %in% asthma_icd9 | 
+      dx10 %in% asthma_icd9 | dx11 %in% asthma_icd9 | dx12 %in% asthma_icd9 | 
+      dx13 %in% asthma_icd9) ~ "dx_asthma_not_primary",
+      # other dx 
[…]

### 2017-12-08 r_scripts/data_wrangle/asthma_cohort_fire_season.R
updated to read the entire dataset.

@@ -23,12 +23,12 @@ asthma_icd9 <- pluck(outcome_icd9_list, "asthma")
 saba_ndc <- pluck(outcome_icd9_list, "saba")
 
 # read in place of service csv file
-place_of_service <- read_csv("./data/health/2013-oregon_pos.csv")
+place_of_service <- read_csv("../../data/health/2013-oregon_pos.csv")
 
 # read in asthma cohort (reading a million rows for now)
-asthma_cohort <- data.table::fread(read_path, 
+asthma_fire_season_cohort <- data.table::fread(read_path, 
   # read options
-  header = T, nrows = 100000, # removing nrows read 
+  header = T, #nrows = 100000, # removing nrows read after testing
   colClasses = rep("character", 72)) %>% 
   # filter to oregon
   filter(STATE == "OR") %>% 
@@ -97,14 +97,20 @@ asthma_cohort <- data.table::fread(read_path,
   # filter to asthma or saba observations
   filter(visit_type == "dx_asthma_primary" | 
          visit_type == "dx_asthma_not_primary" |
-         visit_type == "pharm_saba") %>% 
-  # coding outcomes classified
-  mutate(outcome_class = as.factor(case_when(
-    visit_type == "dx_asthma_primary" & pos == 23 ~ "asthma_emergency",
-    visit_type == "dx_asthma_primary" & pos == 20 ~ "asthma_urgentcare",
-    visit_type == "dx_asthma_primary" & pos == 21 ~ "asthma_inpatient",
-    visit_type == "dx_asthma_primary" & pos == 41 ~ "asthma_ambulance",
-    visit_type == "pharm_saba" ~ "inhaler_fill")))
+         visit_type == "pharm_saba") 
 
-summary(asthma_cohort$outcome_class)
-summary(asthma_cohort$pos_simple)
+# write csv ----
+write_path <- paste0("./data/health/2013-oregon_asthma_fireseason_cohort.csv")
+data.table::fwrite(asthma_fire_season_cohort, write_path)
+
+
+# saving this code for prep in analysis script
+  # # coding outcomes classified
+  # mutate(outcome_class = as.factor(case_when(
+  #   visit_type == "dx_asthma_primary" & pos == 23 ~ "asthma_emergency",
+  #   visit_type == "dx_asthma_primary" & pos == 20 ~ "asthma_urgentcare",
[…]

### 2017-12-08 bash_scripts/r_submission_script.sh
updated bash script to run asthma_cohort_fireseason.

@@ -1,6 +1,6 @@
 #!/bin/bash
 #PBS -N ryan_rjob
-#PBS -l nodes=1:ppn=16
+#PBS -l nodes=1:ppn=1
 #PBS -W group_list=pierce_group
 #PBS -q batch
 #PBS -M rgan@colostate.edu
@@ -16,5 +16,5 @@ cd /home/ryangan/local_git_repo/oregon_wildfire/
 export R_LIB=~/R/x86_64-pc-linux-gnu-library/3.3
 
 # run rscript
-Rscript --vanilla ./r_scripts/data_wrangle/asthma_cohort.R
+Rscript --vanilla ./r_scripts/data_wrangle/asthma_cohort_fireseason.R
 

### 2017-12-08 r_scripts/data_wrangle/asthma_cohort_fireseason.R
changed path.

@@ -14,16 +14,16 @@ library(tidyverse)
 
 # read asthma cohort dataset and affiliated files ------
 # read path
-read_path <- "../../data/health/2013-oregon_asthma_cohort.csv"
+read_path <- "./data/health/2013-oregon_asthma_cohort.csv"
 
 # load outcomes vectors
-load("../../data/health/outcome_list.RData")
+load("./data/health/outcome_list.RData")
 # subset asthma and saba vector
 asthma_icd9 <- pluck(outcome_icd9_list, "asthma")
 saba_ndc <- pluck(outcome_icd9_list, "saba")
 
 # read in place of service csv file
-place_of_service <- read_csv("../../data/health/2013-oregon_pos.csv")
+place_of_service <- read_csv("./data/health/2013-oregon_pos.csv")
 
 # read in asthma cohort (reading a million rows for now)
 asthma_fire_season_cohort <- data.table::fread(read_path, 

### 2017-12-11 r_scripts/analysis/asthma_cohort_analytic_epi.Rmd
working on analytic section.

---
title: "Asthma Cohort Analytic Epidemiology"
author: "Ryan Gan"
date: "12/11/2017"
output: html_document
---

### Document Purpose

Descriptive statistics of the assembled asthma cohort. This cohort is made up of all billing records for procedures or pharamacy fills related to cariopulmonary outcomes for anyone in the Oregon APAC data that had been billed for a diagnosis of asthma at any visit in 2013. I have limited the dataframe further to include only asthma related diagnoses claims or SABA fills between May 1, 2013 to September 30, 2013.

```{r setup, echo =F, warning=F, message=F}
library(tidyverse) # general data wrangle

# load outcomes vectors
load("./data/health/outcome_list.RData")
# subset asthma and saba vector
asthma_icd9 <- pluck(outcome_icd9_list, "asthma")
saba_ndc <- pluck(outcome_icd9_list, "saba")

# knitr options
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

Read in PM~2.5~ data and create binary smoke variables.

Binary smoke variables are based on the following criteria:
- Geoweighted smoke is the ZIPCODE population-weighted geoweighted ridge-regression estimate of PM~2.5~, subtracting off the background PM ~2.5~ value (period median without smoke in the column). 
- Smoke0 through Smoke15 indicators are simply a cutoff value of geo_smoke above 0, 5, 10, and 15 ug/m^3. 
- Smoke Wave indicators are based on Coco Liu's smoke wave, where I've sent the indicator as two days where population-weighted ZIPCODE PM~2.5~ estimated via geoweighted ridge-regression is above the 98th percentile for Oregon over this period (which is roughly 15 ug/m^3).

I just noticed I haves some negative values in the GWR and Krig values, which cannot happen. Need to check in to this. For now, setting values below 0 to 1. It could have happened during the population-weighting.

```{r pm_import}
# read in pm2.5 data
pm_path <- "./data/pm/2013-oregon_zip_pm25.csv"
pm <- read_csv(pm_path) %>% 
  mutate(ZIPCODE = as.character(ZIPCODE),
    geo_wt_pm = ifelse(geo_wt_pm < 0, 1, geo_wt_pm),
  # creating binary smoke indicators 
    smoke0 = ifelse(geo_smk_pm > 0, 1, 0),
    smoke5 = ifelse(geo_smk_pm > 5, 1, 0),
    smoke10 = ifelse(geo_smk_pm > 10, 1, 0),
    smoke15 = ifelse(geo_smk_pm > 15, 1, 0)) %>% 
  # create smoke wave (98%tile for 2 days) 
  group_by(ZIPCODE) %>% 
  arrange(ZIPCODE, date) %>% 
  # 98% is 15 in Oregon
  mutate(high_pm_day = ifelse(geo_wt_pm >= 15, 1, 0),
    smoke_wave = ifelse(lag(high_pm_day, oreder_by = ZIPCODE)==1 & 
                        high_pm_day==1, 1, 0))
```

Read in asthma cohort data and join to PM~2.5~ data.

```{r cohort_import}
# read path
read_path <- "./data/health/2013-oregon_asthma_fireseason_cohort.csv"
# read in asthma cohort (reading a million rows for now)
asthma_cohort <- read_csv(read_path, col_types = cols(.default = "c")) %>% 
  mutate(date = as.Date(todate, format = "%Y-%m-%d")) %>% 
  rename(ZIPCODE = ZIP) %>% 
  left_join(pm, by = c("ZIPCODE", "date"))
```

### Smoke Distributions

Before I look at associations between smoke PM~2.5~ and outcomes in persons with asthma, I want to understand our PM~2.5~ data.

First thing I want to do is look at the distribution of geoweighted PM~2.5~ by the various smoke indicators. I've log-transformed PM~2.5~ values to fit a log-normal distribution.

#### Plot: GWR PM by smoke classifier distribution
```{r pm_distribution}
# small multiples dataframe
geo_smoke_pm <- pm %>% 
  select(date, ZIPCODE, geo_wt_pm, smoke0, 
         smoke5, smoke10, smoke15, smoke_wave) %>% 
  gather("smoke_classifier", "smoke", smoke0:smoke_wave) %>% 
  mutate(smoke = as.factor(smoke),
    smoke_classifier = parse_factor(smoke_classifier, 
      levels = c("smoke0", "smoke5", "smoke10", "smoke15", "smoke_wave")))

# plot
ggplot(geo_smoke_pm, aes(x=log(geo_wt_pm), group=smoke, fill=smoke)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~smoke_classifier) +
  theme_minimal()
```

It would be great to check these classifiers with areas we know were impacted by smoke to see how accurate each classifier is, but I think it would be fair to say that smoke10, smoke15, and smoke_wave all likely distinguish smoke events with relatively high sensitivity. There could be a whole paper on accuracy of these binary classifiers I think. I have thoughts on how to do this.

### Cross-Sectional Comparison of PM2.5

Before any advanced analyses, I'll take a look at the cross-sectional relationship between PM~2.5~ and potential outcomes.

#### Plot: GWR PM by smoke > 10 ug/m^3 by place of service for asthma as a primary diagnosis
```{r dist_pos}
# subset to asthma dx primary
asthma_primary <- asthma_cohort %>% 
  # filter to asthma primary and not in a pharmacy
  filter(visit_type == "dx_asthma_primary" & 
        !is.na(geo_wt_pm) & pos_simple != "Pharmacy")

# ggplot
ggplot(asthma_primary, aes(x=log(geo_wt_pm), group=as.factor(smoke10), 
    fill=as.factor(smoke10))) +
  geom_density(alpha = 0.5) +
  facet_wrap(~pos_simple) +
  theme_minimal()
```

There may be some difference in distributions depending on visit type. For example, ER visit or inpatient hospital visits.

#### Plot: GWR PM by smoke > 10 ug/m^3 by SABA (all SABA fills are at the pharmacy)
```{r dist_saba}
# subset to asthma dx primary
saba <- asthma_cohort %>% 
  # filter to asthma primary and not in a pharmacy
  filter(visit_type == "pharm_saba")

# ggplot
ggplot(saba, aes(x=log(geo_wt_pm), group=as.factor(smoke10), 
    fill=as.factor(smoke10))) +
  geom_density(alpha = 0.5) +
  theme_minimal()
```

SABA fills may have different distributions of PM~2.5~ as well.

```{r saba_pm, eval=F}
saba_bin <- saba %>% 
  mutate(bin = cut(geo_wt_pm, breaks = seq(0,220, by = 10)),
    geo_bin = as.numeric(stringr::str_sub(
      stringr::str_extract(bin, "[^,]*"), start = 2L))) %>% 
  group_by(geo_bin) %>% 
  summarise(n=n()) %>% 
  mutate(p = n/nrow(saba)) %>% 
  filter(geo_bin > 10)

ggplot(data=saba_bin, aes(x=geo_bin, y=p)) +
  geom_point()
# this wasn't a helpful figure (and took too much time)
```

### Time-Stratified Case-Crossover

```{r general tscc function}
ed_cases <- asthma_cohort %>% 
  # filter to asthma primary and not in a pharmacy
  filter(visit_type == "dx_asthma_primary" & pos == "23") %>% 
  # filter to one observation per person (first obs)
  arrange(personkey, date) %>% 
  group_by(personkey) %>% 
  filter(row_number()==1)

# time stratified function; sets up referent observations
ts_casecross <- function(data, start_date, end_date, id, 
                         admit_date, interval){
 data <- ed_cases 
 sym
 # extract id and admit date
 if(id == T){

 }

}

```

### 2017-12-13 r_scripts/analysis/asthma_cohort_analytic_epi.Rmd
updated function.

@@
 # time stratified function; sets up referent observations
-ts_casecross <- function(data, start_date, end_date, id, 
-                         admit_date, interval){
- data <- ed_cases 
- sym
- # extract id and admit date
- if(id == T){
+casecross <- function(data, id, admit_date, start_date, end_date, interval){
+  # if id value is given
+  if(is.character(id)){
+    # vector of ids
+  id_vec <- data[,id]
 
+    } else { # else if no value provided, create an id variable
+      id_vec <- seq(1, nrow(data), by=1)
+  }
+  # vector of admit dates joined to the id vector
+  admit_date <- data[,admit_date]
+  id_date <- cbind(id_vec, admit_date)
+
+  # create list of vectors of referent dates and admission date
+  referent_date_list <- apply(id_date, 1, function(x){
+    date_prior <- as.character(seq(as.Date(x[2]), as.Date(start_date), by = (-interval)))
+    date_posterior <- as.character(seq(as.Date(x[2]), as.Date(end_date), by = (interval)))
+    date_ref <- sort(unique(c(date_prior, date_posterior)))
+    id_rep <- rep(x[1], length(date_ref))
+    id_date_vec <- cbind(id_rep, date_ref)
+    # outcome
+    data_list$outcome <- data_list[data_list$date_vec]
+    return(id_date_vec)
+  }) # end apply
+    # timestrat dataframe
+  ts_data <- do.call(rbind, referent_date_list)
+
+  # stuff below still needs to be figured out.
+
+  names(referent_date_list) <- id_vec
+  head(referent_date_list)
+  test_list <- referent_date_list[1]
+
+  casecross_data <- sapply(test_list, function(x){
+    id_vec <- rep(names(x), lengths(x))
+    date_vec <- unlist(x)
+    cc_data <- cbind(id_vec, date_vec)
+    return(cc_data)
+  })
+
+  do.call
+  rep(names(test_list), lengths(test_list))
 }
 
-}
+lengths(test_list)
+
+expand.grid(height = seq(60, 80, 5), weight = seq(100, 300, 50),
+            sex = c("Male","Female"))
+
+test <- casecross(data = test_ed, id="personkey", admit_date = "todate", 
+                  start_date = "2013-05-01", end_date = "2013-10-1", interval = 7)
+
+  date_prior
+  date_posterior
 

### 2017-12-13 r_scripts/analysis/asthma_cohort_analytic_epi.Rmd
finished function. consider adding day of week in future iterations.

--- 2017-12-11 r_scripts/analysis/asthma_cohort_analytic_epi.Rmd
@@
 # time stratified function; sets up referent observations
-ts_casecross <- function(data, start_date, end_date, id, 
-                         admit_date, interval){
- data <- ed_cases 
- sym
- # extract id and admit date
- if(id == T){
+time.stratified <- function(data, id, covariate=F, admit_date, 
+                            start_date, end_date, interval){
+  # if id value is given
+  if(is.character(id)){
+    # vector of ids
+  id_vec <- data[,id]
+    } else { # else if no value provided, create an id variable
+      id_vec <- seq(1, nrow(data), by=1)
+  }
+  # vector of admit dates joined to the id vector
+  admit_date <- data[,admit_date]
+  id_date <- data.frame(id_vec, admit_date)
 
+  # create list of vectors of referent dates and admission date
+  referent_date_list <- apply(id_date, 1, function(x){
+    date_prior <- as.character(seq(as.Date(x[2]), as.Date(start_date), by = (-interval)))
+    date_posterior <- as.character(seq(as.Date(x[2]), as.Date(end_date), by = (interval)))
+    date <- sort(unique(c(date_prior, date_posterior)))
+    identifier <- rep(x[1], length(date))
+    # outcome
+    outcome <- ifelse(date %in% x[2], 1, 0)
+    id_date_vec <- cbind(identifier, date, outcome)
+    return(id_date_vec)
+  }) # end apply
+    # timestrat dataframe
+    ts_data <- do.call(rbind, referent_date_list)
+    rownames(ts_data) <- NULL
+    ts_data <- as.data.frame(ts_data)
+    # if covariates provided, join to the dataframe
+    if(is.character(covariate)){
+      cov_data <- data[,c(id,covariate)]
+      ts_data <- merge(ts_data, cov_data, by.x = "identifier", by.y = id[])
     }
-
+  # return dataframe
+  return(ts_data) 
 }

### 2017-12-18 r_scripts/analysis/asthma_cohort_causal.Rmd
started cohort script.

---
title: "Asthma Cohort"
author: "Ryan Gan"
date: "12/18/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2018-01-09 r_scripts/analysis/asthma_care_utilization.Rmd
added new file.

---
title: "Multiple visits in persons likely to have asthma"
author: "Ryan Gan"
date: "1/9/2018"
output: html_document
---

## Document Purpose

This document attempts to describe utilization of healthcare in persons who are likely to have asthma. It will eventually see if wildfire smoke may influence utilization.

First step is to read in the persons with asthma in the fireseason. I may end up doing this for the entire year, but that dataset is much larger and would take a bit.

```{r setup, echo = F, warning = F, message = F}
# libraries.
library(tidyverse)

# knitr options
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

### 2018-01-12 r_scripts/analysis/asthma_care_utilization.Rmd
I have no idea how to visualize or make sense of mulitple visits.

@@
 
+## Setup
+
 First step is to read in the persons with asthma in the fireseason. I may end up doing this for the entire year, but that dataset is much larger and would take a bit.
@@
 library(tidyverse)
+library(lme4)
 
 # knitr options
-knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='Figs/',
-                      echo=FALSE, warning=FALSE, message=FALSE)
+knitr::opts_chunk$set(fig.width=8, fig.height=6, 
+                      warning=FALSE, message=FALSE)
 ```
+
+Reading in ZIP population-weighted smoke data and joining to ZIPCODE of persons.
+
+```{r smoke_import_wrangle}
+# read in pm2.5 data
+pm_path <- "../../data/pm/2013-oregon_zip_pm25.csv"
+
+pm <- read_csv(pm_path) %>% 
+  mutate(ZIPCODE = as.character(ZIPCODE),
+    geo_wt_pm = ifelse(geo_wt_pm < 0, 1, geo_wt_pm),
+  # creating binary smoke indicators 
+    smoke0 = ifelse(geo_smk_pm > 0, 1, 0),
+    smoke5 = ifelse(geo_smk_pm > 5, 1, 0),
+    smoke10 = ifelse(geo_smk_pm > 10, 1, 0),
+    smoke15 = ifelse(geo_smk_pm > 15, 1, 0)) %>% 
+  # create smoke wave (98%tile for 2 days) 
+  group_by(ZIPCODE) %>% 
+  arrange(ZIPCODE, date) %>% 
+  # 98% is 15 in Oregon
+  mutate(high_pm_day = ifelse(geo_wt_pm >= 15, 1, 0),
+    smoke_wave = ifelse(lag(high_pm_day, oreder_by = ZIPCODE)==1 & 
+                        high_pm_day==1, 1, 0)) %>% 
+  rename(ZIP = ZIPCODE)
+```
+
+Reading in asthma cohort data for wildfire season and joinin to asthma cohot observations.
+
+```{r read_data}
+read_path <- paste0("./data/health/",
+  "2013-oregon_asthma_fireseason_cohort.csv")
+
+asthma_cohort <- read_csv(read_path, col_types = cols(.default = "c")) %>% 
+  mutate(date = as.Date(fromdate)) %>% 
+  left_join(pm, by = c("date", "ZIP"))
+```
+
+Find the person with the most observations over this timeperiod.
+
+```{r person_most_obs}
+# create indicator of number of observations
+n_persons <- asthma_cohort %>% 
+  group_by(personkey) %>% 
+  summarise(n_obs = n())
+
+# summary of number of persons
+summary(n_persons)
+
+# plot of distribution
+ggplot(n_persons, aes(x=n_obs)) +
+  geom_density()
+```
+
+92,487 unique personkey IDs, with a range of observations of 444. Most people have a reasonable number of observations over this time.
+
+Some observations are likely multiple lines per claim.
+444 visits for a single person (personkey = 10733403) over the 5 month fireseason would be highly unlikely.
+I want to count up the number of lines per claim per person.
+
+```{r person_claims}
+n_lines <- asthma_cohort %>% 
+  group_by(personkey, clmid) %>% 
+  summarise(n_lines = n())
+
+# summary of number of lines per claim
+summary(n_lines)
+```
+
+There are 265,326 unique claims.
+Most persons have maybe 1 to 3.
+One person has a unique claim with 182 (personkey = 10483893) lines associated with it.
+Distribution plot took too long so I didn't add it.
+My guess that it would show similar distribution to previous plot.
+
+Now I want to know the number of unique visits per person.
+
+```{r visits_per_person}
+n_visits <- n_lines %>% 
+  group_by(personkey) %>% 
+  summarise(n_visits = n())
+
+# summary of visits
+summary(n_visits)
+
+# density plot
+ggplot(n_visits, aes(x=n_visits)) +
+  geom_density()
+```
+
+Someone had 117 medical claims I'm interpreting as visits/fills over the time period (personkey = 1738926). Most people have 1 or 2 visits.
+
+Now I'd like to know how many medical visits and pharmacy fills.
+
+```{r claims_visits}
+n_visit_types <- asthma_cohort %>% 
+  # group by person and clmid
+  group_by(personkey, clmid) %>% 
+  # filter to first line of clmid
+  arrange(personkey, clmid, line) %>% 
+  # filter to only one claimline 
+  filter(row_number()==1) %>% 
+  # filter again to only asthma as primary or saba
+  filter(visit_type != "dx_asthma_not_primary") %>% 
+  # break initial grouping
+  ungroup() %>% 
+  # remove whitespaces from place of service
+  mutate(pos_simple = stringr::str_replace_all(pos_simple, " ", "")) %>% 
+  # regroup by person and service place
+  group_by(personkey, pos_simple) %>% 
+  # summarise
+  summarise(n_type = n()) %>% 
+  # spread
+  spread(pos_simple, n_type) %>% 
+  mutate_at(vars(Ambulance:UrgentCare), funs(ifelse(is.na(.), 0, .)))
+
+# glipse of first couple lines
+glimpse(n_visit_types)
+
+# summary of overall
+summary(n_visit_types)
+```
+
+67,738 persons.
+Most people have very few visits for any particular place of service (measure of severity here).
+I tried a couple plots, but the dataframe was too large and I don't think it will be helpful.
+
+## Wildfire smoke
+
+Adding in a second grouping variable, smoke > 15. Overwritting previous file.
+
+```{r}
+n_visit_types <- asthma_cohort %>% 
+  # group by person and clmid
+  group_by(personkey, clmid) %>% 
+  # filter to first line of clmid
+  arrange(personkey, clmid, line) %>% 
+  # filter to only one claimline 
+  filter(row_number()==1) %>% 
+  # filter again to only asthma as primary or saba
+  filter(visit_type != "dx_asthma_not_primary") %>% 
+  # break initial grouping
+  ungroup() %>% 
+  # remove whitespaces from place of service
+  mutate(pos_simple = stringr::str_replace_all(pos_simple, " ", "")) %>% 
+  # regroup by person and service place
+  group_by(personkey, pos_simple, smoke15) %>% 
+  # summarise
+  summarise(n_type = n()) %>% 
+  # spread
+  spread(pos_simple, n_type) %>% 
+  mutate_at(vars(Ambulance:UrgentCare), funs(ifelse(is.na(.), 0, .)))
+
+# checking a person that has two visits
+check <- filter(n_visit_types, personkey == "10001535")
+```
+
+I think rates would be more informative than counts alone. Denominator will be number of days no smoke, and number of days smoke for each ZIP code.
+
+I also would like to extract covariates for each person. I'll do that first.
+
+``` {r covariates}
+# extract vector of ids
+ids <- unique(n_visit_types$personkey)
+
+# extract covariates for each person
+covariates <- asthma_cohort %>%
+  # filter to only persons in id vector
+  filter(personkey %in% ids) %>% 
+  # group by personkey
+  group_by(personkey) %>% 
+  # sort by personkey and date
+  arrange(personkey, date) %>% 
+  # filter first row
+  filter(row_number()==1) %>% 
+  select(personkey, ZIP, gender, age) %>% 
+  ungroup() %>% 
+  mutate(gender = as.factor(gender),
+         age = as.numeric(age))
+```
+
+```{r days_smoke}
+# count up smoke days
+days_smoke <- pm %>% 
+  group_by(ZIP, smoke15) %>% 
+  summarise(n_days=n())
+
+# join smoke days to n_visits_types
+asthma_n_visits <- n_visit_types %>% 
+  # join to covariates
+  left_join(covariates, by = "personkey") %>% 
+  left_join(days_smoke, by = c("ZIP", "smoke15")) %>% 
+  # calculate rates
+  mutate_at(vars(Ambulance:UrgentCare), funs(rate10days = (./n_days)*10)) %>% 
+  # filter out the NAs
+  #need to return to why this might be? zipcoudes outside oregon maybe
+  filter(!is.na(n_days)) %>% 
+  filter(gender != "U")
+
+summary(asthma_n_visits)
+
+glimpse(asthma_n_visits)
+
+```
+
+Generalized Poisson model to see if the counts of ED visits a person might have might be greater on smoke periods compared to non-smoke periods.
+
+```{r poisson_mod}
+mod <- glmer(EmergencyRoomHospital ~ smoke15 + (1|personkey) + 
+          offset(n_days), family="poisson"(link=log),
+          data = asthma_n_visits)
+
+lm_mod2 <- lmer(EmergencyRoomHospital_rate10days ~ smoke15 + (1|personkey) +
+             gender + age, 
+             data = asthma_n_visits)
+
+summary(lm_mod)
+summary(lm_mod2)
+
+ggplot(asthma_n_visits, aes(x=log(EmergencyRoomHospital_rate10days*1000+0.1), 
+    group = as.factor(smoke15), color=as.factor(smoke15))) +
+  geom_density()
+```
+
+### Note for Sheryl: 2018-01-11
+
+Sheryl, is this kind of what you were thinking?
+I have counts of numbers of times a person was seen for an asthma primary diagnosis by place of service or number of SABA fills (represented by pharmacy).
+
+I could have expanded in different ways, but I think this would have been a huge dataframe. Also, I should note I may be undercounting because I only took the first line of unique claims.
+
+Next step; find an indicator of smoke >10 or >15? or try smoke wave? Then compare number of visits, or rate by indicator.

### 2018-04-16 r_scripts/data_wrangle/asthma_pos_timestrat.R
Creating timestratified case-crossover dataframes; wrote script for parallel computation.

@@ -1,0 +1,83 @@
+# ------------------------------------------------------------------------------
+# Title: Asthma place of service time-stratified case-crossover
+# Author: Ryan Gan
+# Date Created: 2018-04-16
+# ------------------------------------------------------------------------------
+
+# Description: This script subsets the asthma claims to specific service places
+# and creates time-stratified case-crossover dataframes
+
+
+# library ----
+library(tidyverse)
+library(case.crossover)
+library(parallel)
+
+# asthma place of service ----
+
+# read in place of service csv file
+place_of_service <- read_csv("./data/health/2013-oregon_pos.csv")
+
+pos_subset <- c("11", "20", "21", "22", "23", "41")
+
+# read in asthma icd9 vector
+load("./data/health/outcome_list.RData")
+
+asthma_icd9 <- outcome_icd9_list$asthma
+
+# read in asthma claims
+asthma_claims <- read_csv("./data/health/2013_oregon_asthma_claims.csv") %>% 
+  # filter to date of PM2.5 data (wildfire season)
+  filter(fromdate >= "2013-05-01" & fromdate <= "2013-09-30") %>% 
+  # filter to oregon
+  filter(STATE == "OR") %>% 
+  # filter to asthma claims of a primary diagnosis only
+  filter(dx1 %in% asthma_icd9) %>% 
+  # filter to place of service i want to analyze
+  filter(pos %in% pos_subset) %>% 
+  # left join to place of service names
+  left_join(place_of_service, by = "pos") %>% 
+  # split by list
+  split(.$pos)
+
+# transform pos_subset to list
+asthma_list <- as.list(asthma_claims)
+
+# set up cores
+cl <- makeCluster(6)
+
+# load packages on each processor of the node/cluster
+clusterCall(cl, function() c(library(tidyverse), library(case.crossover)))
+
+# export new set of objects to global objects to cluster
+clusterExport(cl, c("asthma_list"), 
+              envir = .GlobalEnv)
+
+# parallel creation of casecrossover list -----
+# start time
+start <- Sys.time()
+
+asthma_cc_pos_list <- parLapply(cl, asthma_list, function(x){
+    pos_df <- x %>% 
+    # find the first observation and first claim line of each person 
+    group_by(personkey) %>%
+    arrange(fromdate, line) %>%
+    filter(row_number()==1) %>% 
+    ungroup()
+    
[…]

### 2018-04-16 bash_scripts/asthma_cc_sub.sh
created bash script to execute job submission.

@@ -1,0 +1,20 @@
+#!/bin/bash
+#PBS -N casecross_job
+#PBS -l nodes=6:ppn=1
+#PBS -W group_list=pierce_group
+#PBS -q batch
+#PBS -M rgan@colostate.edu
+#PBS -m abe
+
+# set memory to unlimited
+ulimit -s unlimited
+
+# set command directory
+cd /home/ryangan/local_git_repo/oregon_wildfire/
+
+# export R library
+export R_LIB=~/R/x86_64-pc-linux-gnu-library/3.3
+
+# run rscript
+Rscript --vanilla ./r_scripts/data_wrangle/asthma_pos_timestrat.R
+

### 2018-04-16 r_scripts/data_wrangle/asthma_pos_timestrat.R
fixed the timestratified function in custom package and added age instead of yob.

@@ -57,16 +57,18 @@ clusterExport(cl, c("asthma_list"),
 # start time
 start <- Sys.time()
 
-asthma_cc_pos_list <- parLapply(cl, asthma_list, function(x){
+#asthma_cc_pos_list <- parLapply(cl, asthma_list, function(x){
+asthma_cc_pos_list <- lapply(asthma_list, function(x){
     pos_df <- x %>% 
     # find the first observation and first claim line of each person 
     group_by(personkey) %>%
     arrange(fromdate, line) %>%
     filter(row_number()==1) %>% 
-    ungroup()
+    ungroup() %>% 
+    mutate(age = 2013 - yob)
     
     ts_df <- time_stratified(data = pos_df, id = "personkey", 
-      covariate = c("patid", "gender", "yob", "MSA", "ZIP", "pos", "dx1",
+      covariate = c("patid", "gender", "age", "MSA", "ZIP", "pos", "dx1",
                   "service_place"), 
       admit_date = "fromdate", 
       start_date = "2013-05-01", end_date = "2013-09-30", interval = 7)
@@ -75,6 +77,7 @@ asthma_cc_pos_list <- parLapply(cl, asthma_list, function(x){
 stop <- Sys.time()
 time <- stop - start
 print(time)
+
 
 # close cores 
 stopCluster(cl)

### 2018-04-18 r_scripts/analysis/asthma_place_of_service_analysis.Rmd
Figured out non-linear distributed lag model.

---
title: "Association between wildfire smoke and asthma claims by place of service"
author: "Ryan Gan"
date: '2018-04-16'
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

# Introduction

This markdown file follows up on analyses evaluating emergency department (ED) visits and the association between same-day wildfire smoke (WFS) exposure and the association with cardiopulmonary ED claims (proxy for a visit). Since we only observed an association with asthma-related ED visits, we are going to evaluate the impact of WFS exposure on other asthma-related visits and short-acting beta 2 agonist (SABA) fills. The primary aim of these analyses are to help us understand the extent of the impact of WFS on asthma.

```{r setup}
# libraries
library(tidyverse) # for ggplot and general data wrangle
library(survival) # for clogit
library(dlnm) # set up basis functions

# knitr options
knitr::opts_chunk$set(fig.width = 8, fig.height = 6)
```

## Methods

General methods overview. I will offer

### Outcome Definition

In addition to asthma ED visits, we will also evaluate place of service visits for urgent care visits, inpatient visits, outpatient visits, office visits, and ambulance visits.

First observation over this time period for the patient for visits at a medical facility, but we allow all observations for SABA fills.

We are going to use the same approach of the time-stratified case-crossover design for each event. We will limit the study period and referent period for each event to the wildfire season of May 1^st^, 2013 to September 30^th^, 2013.

```{r cc_list}
# load asthma casecross place of service list
load("../../data/health/asthma_cc_pos_list.RData")
# extract names
vars_keep <- names(asthma_cc_pos_list[[4]])
```

Reading in SABA time-stratified case-crossover, cleaning and adding to the asthma casecross list. I've assigned the NDC code to dx1 even though that's not technically what it is.

```{r saba_df}
saba <- read_csv("../../data/health/2013-oregon_casecross_saba.csv") %>% 
  rename(identifier = personkey, ZIP = ZIPCODE) %>% 
  mutate(service_place = "SABA Fill",
         dx1 = ndc) %>% 
  # keeping only vars in the other casecrossover dataframes
  select(vars_keep) %>% 
  # turn in to list
  list()

# assign name
names(saba) <- "saba"
```

```{r list_subset}
# adding in age category to each dataframe in list 
# (i should have done this earlier)
asthma_pos_list <- c(asthma_cc_pos_list, saba) %>% 
  map(., function(df){
    return_df <- df %>% 
      # remove unknown gender
      filter(gender != "U") %>% 
      mutate(age_cat = case_when(age < 15 ~ "Age < 15",
                                age >= 15 & age < 65 ~ "Age 15-65",
                                age >= 65 ~ "Age > 65"),
              metroarea = case_when(MSA == 13460 ~ "Bend-Redmond",
                                MSA == 18700 ~ "Corvallis",
                                MSA == 21660 ~ "Eugene-Springfield",
                                MSA == 32780 ~ "Medford",
                                MSA == 38900 ~ "Portland-Vancouver-Beaverton",
                                MSA == 41420 ~ "Salem",
                                MSA == 41999 ~ "Not in MSA"))
  })

# removing older list to save some space
rm(asthma_cc_pos_list)
```

Estimate the number of observations over the time period.

```{r descriptive_n_percent}
counts <- asthma_pos_list %>% 
  map_dfr(., function(df){
    # subset to outcome observations only
    out_df <- df %>%
      filter(outcome == 1)

    # place of service
    service_place <- as.character(unique(df$service_place))

    # total count
    n <- out_df %>% 
      summarise(n = n()) %>% 
      mutate(strata = "Total") %>% 
      select(strata, n)

    total <- as.numeric(n$n)
    # sex counts
    sex_n <- out_df %>% 
      group_by(gender) %>% 
      summarise(n = n()) %>% 
      rename(strata = gender) %>% 
      mutate(strata = ifelse(strata == "M", "Male", "Female"))
    # age cat counts
    age_n <- out_df %>% 
      group_by(age_cat) %>% 
      summarise(n = n()) %>% 
      rename(strata = age_cat)

    # bind in rows and spread
    counts <- bind_rows(n, sex_n, age_n) %>% 
      mutate(strata = factor(strata, levels = strata)) %>% 
      spread(strata, n) %>% 
      cbind(service_place, .) %>% 
      mutate_at(3:7, funs(perc = round(./Total,3)*100)) %>% 
      mutate(service_place = as.character(service_place))

  })

# counts of observations; supressing n lower than 20
n_tab <- counts %>% select(1:7) %>% 
  mutate_at(2:7, funs(ifelse(. < 20, NA, .)))
# print n tab
n_tab
```

Percentage breakdown of place of service distribution by strata.

```{r percent}
perc_tab <- counts %>% select(1, 8:12)
# print n tab
perc_tab
```

### PM~2.5~

Creating lagged structure just in case.

```{r lag_fun}
# defining a lag function
funlag <- function(var, n=6){
  var <- enquo(var)
  indices <- seq_len(n)
  map( indices, ~quo(lag(!!var, !!.x)) ) %>% 
    set_names(sprintf("%s_lag%d", rlang::quo_text(var), indices))
}
```

Importing ZIP code population-weighted PM~2.5~ data.

```{r pm}
# read in pm2.5
pm <- read_csv("../../data/pm/2013-oregon_zip_pm25.csv") %>% 
  # create 10 unit variables
  mutate(geo_pm10 = geo_wt_pm/10,
         geo_smk10 = geo_smk_pm/10) %>% 
  # apply lag function to estimate 5 day lag
  arrange(ZIPCODE, date) %>% 
  group_by(ZIPCODE) %>%
  mutate(., !!!funlag(geo_pm10, 5), !!!funlag(geo_smk10, 5), 
         !!!funlag(wrf_temp,5)) %>% 
  rename(ZIP = ZIPCODE)
```

Joining PM~2.5~ with each dataframe in asthma place of service list.

```{r join_pm}
asthma_pos_list <- asthma_pos_list %>% 
  map(., function(df){
    return_df <- df %>% 
      # join with pm on date and zip
      left_join(pm, by = c("date", "ZIP"))
  })
```

## Total Same-Day Association with GWR Smoke for All Events

```{r smk_association_all_obs}
sameday_result <- asthma_pos_list %>% 
  map_dfr(., function(df){
    service_place <- as.character(df$service_place[1])
    # complete case
    df <- df %>% filter(!is.na(geo_smk10))
    # model
    mod <- clogit(outcome ~ geo_smk10 + wrf_temp + strata(identifier), 
                  data = df)
    # n events
    n_events <- mod$nevent
    # odds ratio and 95% CI
    estimates <- broom::tidy(mod) %>% 
      filter(term == "geo_smk10") %>% 
      select(term, estimate, conf.low, conf.high) %>% 
      mutate_at(2:4, funs(round(exp(.),3))) %>% 
      cbind(service_place, n_events, .) %>% 
      mutate(service_place = as.character(service_place))
  }) %>% 
  mutate(strata = "Total") %>% 
  select(strata, service_place:conf.high)

# table
sameday_result
```

## Sex-Specific Same-Day Association

Feeding in asthma place of service list, applying a series of map functions to first nest the dataframes by sex, then run the clogit model, and finally to extract the results. I think this approach is faster than a loop approach or other methods.

```{r sameday_sex_strata}
sex_result <- asthma_pos_list %>% 
  map_dfr(., function(df){
    service_place <- as.character(df$service_place[1])
    # complete case
    sex_df <- df %>% filter(!is.na(geo_smk10)) %>% 
      # nest strata
      group_by(gender) %>% 
      nest() %>% 
      mutate(model = map(data, ~ clogit(outcome ~ geo_smk10 + wrf_temp +
                                          strata(identifier), data = .)),
             est_df = map(model, function(mod){
                  # n events
                  n_events <- mod$nevent
                  # odds ratio and 95% CI
                  estimates <- broom::tidy(mod) %>% 
                    filter(term == "geo_smk10") %>% 
                    select(term, estimate, conf.low, conf.high) %>% 
                    mutate_at(2:4, funs(round(exp(.),3))) %>% 
                    cbind(service_place, n_events, .) %>% 
                    mutate(service_place = as.character(service_place))})) %>% 
      select(-data, -model) %>% 
      unnest(est_df)
  }) %>% 
  mutate(strata = ifelse(gender == "M", "Male", "Female")) %>% 
  select(strata, service_place:conf.high)

# table
sex_result
```

## Age-Specific Same-Day Association

Repeating the same procedure as above, but for each age category.

```{r age_sameday}
age_results <- asthma_pos_list %>% 
  map_dfr(., function(df){
    service_place <- as.character(df$service_place[1])
    # complete case
    sex_df <- df %>% filter(!is.na(geo_smk10)) %>% 
      # nest strata
      nest(-age_cat) %>% 
      mutate(model = map(data, ~ clogit(outcome ~ geo_smk10 + wrf_temp +
                                          strata(identifier), data = .)),
             est_df = map(model, function(mod){
                  # n events
                  n_events <- mod$nevent
                  # odds ratio and 95% CI
                  estimates <- broom::tidy(mod) %>% 
                    filter(term == "geo_smk10") %>% 
                    select(term, estimate, conf.low, conf.high) %>% 
                    mutate_at(2:4, funs(round(exp(.),3))) %>% 
                    cbind(service_place, n_events, .) %>% 
                    mutate(service_place = as.character(service_place))})) %>% 
      select(-data, -model) %>% 
      unnest(est_df)
  }) %>% 
  rename(strata = age_cat) %>% 
  select(strata, service_place:conf.high) %>% 
  filter(n_events > 20) %>% 
  filter(conf.high < 5)

# print table
age_results
```

Binding all results together and creating a small-multiples plot.

```{r all_results}
results <- bind_rows(sameday_result, sex_result, age_results) %>% 
  rename(odds_ratio = estimate, lower95 = conf.low, upper95 = conf.high) %>% 
  mutate(strata = case_when(strata == "Total" ~ "All Strata",
                            TRUE ~ strata),
         # shorten names of pos
         service_place = case_when(service_place=="Ambulance-Land"~"Ambulance",
          service_place == "Emergency Room – Hospital" ~ "Emergency Room",
          service_place == "Urgent Care Facility" ~ "Urgent Care",
          TRUE ~ service_place),
         # set order of strata
          strata = factor(strata, 
            levels = c("All Strata", "Female", "Male", "Age < 15", 
                       "Age 15-65", "Age > 65")),
    # set order of service place
          service_place = factor(service_place, 
            levels = c("SABA Fill", "Office", "Outpatient Hospital", 
                       "Urgent Care", "Ambulance", "Emergency Room", 
                       "Inpatient Hospital")))
```

Plotting results.

```{r plot}
plot <- ggplot(results, aes(x = strata, y = odds_ratio)) +
  geom_point(color = "blue") + 
  geom_errorbar(aes(ymin=lower95, ymax=upper95), color = "blue",width = 0.2) +
  geom_hline(yintercept = 1, linetype=2) +
  facet_wrap(~service_place, scales = "free_y") +
  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
  xlab("Stratum") +
   # plot theme
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # strip element
    strip.background = element_rect(colour=NA, fill=NA),
    panel.border = element_rect(fill = NA, color = "black"),
    # facet text size
    strip.text = element_text(size = 10),
    # axis element (tilt)
    axis.text.x = element_text(angle = 45, hjust =1),
    #axis.title.x = element_blank(),
    #axis.ticks.x = element_blank(),
    axis.title.y = element_text(angle = 90))

# print plot
plot
```

I don't really like this plot because of the wide intervals. I also don't think there is really any evidence of effect modification by strata.

I think I'll just present the all strata estimate.

```{r total_results}
# filter to total results
total_results <- results %>% 
  filter(strata == "All Strata")

# plot total results
total_plot <- ggplot(total_results, aes(x = service_place, y = odds_ratio, 
                                        group = service_place)) +
  geom_point(color = "blue") + 
  geom_errorbar(aes(ymin=lower95, ymax=upper95), color = "blue",width = 0.2) +
  geom_hline(yintercept = 1, linetype=2) +
  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
  xlab("Place of Service") +
   # plot theme
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # axis element (tilt)
    axis.text.x = element_text(angle = 45, hjust =1),
    #axis.title.x = element_blank(),
    #axis.ticks.x = element_blank(),
    axis.title.y = element_text(angle = 90))

total_plot
```

I will probably color-code this. Interesting take-away would be that all types of asthma events were associated with a 10 ug/m^3^ increase in PM~2.5~ we are attributed to smoke. The only events not associated would be urgent care and ambulance visits. However, ambulance visits goes in the direction I'd expect if smoke were to be hazardous. Urgent care shows and inverse association, but it's not associated. SABA fill is any SABA fill, while we only take the first visit for each person, over the time period, which may or may not be appropriate. I can't decide.

## Medford-Specific Results

Limiting to MSA of Medford to see if I see signals with smoke for the specific places of service. I've excluded ambulance and urgent care since their 95% CI CIs were wide, suggesting poor model fits.

```{r medford}
# excluding ambulance since it won't run for medford
medford_result <- asthma_pos_list[c(1,3:5,7)] %>% 
  map_dfr(., function(df){
    service_place <- as.character(df$service_place[1])
    # complete case
    df <- df %>% filter(!is.na(geo_smk10)) %>% 
      filter(metroarea == "Medford")
    # model
    mod <- clogit(outcome ~ geo_smk10 + wrf_temp + strata(identifier), 
                  data = df)
    # n events
    n_events <- mod$nevent
    # odds ratio and 95% CI
    estimates <- broom::tidy(mod) %>% 
      filter(term == "geo_smk10") %>% 
      select(term, estimate, conf.low, conf.high) %>% 
      mutate_at(2:4, funs(round(exp(.),3))) %>% 
      cbind(service_place, n_events, .) %>% 
      mutate(service_place = as.character(service_place))
  }) %>% 
  mutate(strata = "Total") %>% 
  select(strata, service_place:conf.high) %>% 
  rename(odds_ratio = estimate, lower95 = conf.low, upper95 = conf.high) %>% 
  mutate(MSA = "Medford",
         service_place = case_when(service_place=="Ambulance-Land"~"Ambulance",
          service_place == "Emergency Room – Hospital" ~ "Emergency Room",
          TRUE ~ service_place),
    # set order of service place
          service_place = factor(service_place, 
            levels = c("SABA Fill", "Office", "Outpatient Hospital", 
                       "Emergency Room", "Inpatient Hospital")))

# table
medford_result
```

## Other MSAs (not Medford or outside of MSA)

```{r other_msa}
other_msa_result <- asthma_pos_list[c(1,3:5,7)] %>% 
  map_dfr(., function(df){
    service_place <- as.character(df$service_place[1])
    # complete case
    df <- df %>% filter(!is.na(geo_smk10)) %>% 
      filter(metroarea != "Medford") %>% 
      filter(metroarea != "Not in MSA")
    # model
    mod <- clogit(outcome ~ geo_smk10 + wrf_temp + strata(identifier), 
                  data = df)
    # n events
    n_events <- mod$nevent
    # odds ratio and 95% CI
    estimates <- broom::tidy(mod) %>% 
      filter(term == "geo_smk10") %>% 
      select(term, estimate, conf.low, conf.high) %>% 
      mutate_at(2:4, funs(round(exp(.),3))) %>% 
      cbind(service_place, n_events, .) %>% 
      mutate(service_place = as.character(service_place))
  }) %>% 
  mutate(strata = "Total") %>% 
  select(strata, service_place:conf.high) %>% 
  rename(odds_ratio = estimate, lower95 = conf.low, upper95 = conf.high) %>% 
  mutate(MSA = "Other",
         service_place = case_when(service_place=="Ambulance-Land"~"Ambulance",
          service_place == "Emergency Room – Hospital" ~ "Emergency Room",
          TRUE ~ service_place),
    # set order of service place
          service_place = factor(service_place, 
            levels = c("SABA Fill", "Office", "Outpatient Hospital", 
                       "Emergency Room", "Inpatient Hospital")))

# table
other_msa_result
```

## Plot of Same-Day Association by MSA

```{r medford_plot}
msa_results <- bind_rows(medford_result, other_msa_result)
# plot total results
med_plot <- ggplot(msa_results, aes(x = service_place, y = odds_ratio)) +
  geom_point(color = "purple") + 
  geom_errorbar(aes(ymin=lower95, ymax=upper95), color = "purple",width = 0.2) +
  geom_hline(yintercept = 1, linetype=2) +
  facet_wrap(~MSA) +
  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
  xlab("Place of Service") +
   # plot theme
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # strip element
    strip.background = element_rect(colour=NA, fill=NA),
    panel.border = element_rect(fill = NA, color = "black"),
    # facet text size
    strip.text = element_text(size = 10),
    # axis element (tilt)
    axis.text.x = element_text(angle = 45, hjust =1),
    #axis.title.x = element_blank(),
    #axis.ticks.x = element_blank(),
    axis.title.y = element_text(angle = 90))

med_plot
```

For people who reported living in Medford, I see strong, but small relative increases in the likelihood for a SABA fill and an office visits (and probably inpatient hospitalizations as well). When compared to people living in other MSAs during this period, there is a stronger signal for SABA fills and office visits, but the invervals are much wider even though the number of events is much more than Medford. A couple interesting thoughts: Medford was the only MSA really impacted by smoke during this time, where I'd guess that any signal for other MSAs during this time for smoke are likely because our definition of PM~2.5~ due to smoke is pretty crude as it's just the difference of the daily specific ZIP code estimate minus the estimated seasonal background for that ZIP code. Hence, it's probably picking up some elevated PM~2.5~ days.

## Non-Linear Distributed Lag

Preparing PM~2.5~ values, assigning an MSA by ZIP code.

```{r pm_msa}
# making a zipcode and msa key
zip_msa <- asthma_pos_list[[1]] %>% 
  select(MSA, metroarea, ZIP) %>% 
  unique()
# ungrouping pm and adding in MSA for color-coding
pm_msa <- ungroup(pm) %>% 
  left_join(zip_msa, by = "ZIP") %>% 
  #filter out missing metroareas
  filter(!is.na(metroarea))
```

Checking distribution of PM~2.5~ by each ZIP code in each MSA I was able to code. There is a heavy right tail for Medford and ZIPs not in an MSA (which probably capture some rural areas around the southern part of the state. If you let the scales vary, the distribution of PM~2.5~ values suggests to me that Medford was impacted by smoke, as evidenced by the heavy right tail. Other MSAs have distributions that are more like standard distributions for ambient air pollution not impacted by smoke or other extreme events.

```{r pm_dist}
# distribution of pm values is log-normal
ggplot(pm_msa, aes(x=geo_wt_pm, group = ZIP, color=metroarea)) +
  geom_density() +
  facet_wrap(~metroarea, scales = "free") +
  theme_minimal()
```

Looking at time-series of PM~2.5~ values by MSA.I plotted the mean seasonal estimate for each ZIP in each MSA as a red line for each MSA.

```{r pm_ts}
# estimating seasonal background
bg_pm <- pm_msa %>% 
  group_by(date, metroarea) %>% 
  summarise(mean_pm = mean(background_pm), med_pm = median(background_pm))

# plot
ggplot(pm_msa, aes(x=date, y=geo_wt_pm, group=ZIP, color = metroarea)) +
  geom_point() +
  geom_line(data=bg_pm, aes(x=date, y=mean_pm, group=metroarea), color = "red") +
  facet_wrap(~metroarea) +
  theme_minimal()

```

After looking at these plots, I find it hard to overlook the simple, but limited approach of just subtracting off the difference of the PM~2.5~ estimate from the background estimate.

I'm going to revist the non-linear distributed lag approach. I think we can then estimate a couple different lagged and cumulative effect over a 0 to 5 day prior event period. We could then estimate the function at 10 ug/m^3^ ("normal" air pollution), 20 ug/m^3^ (most likely light smoke), and 30 ug/m^3^ (smoke), and maybe 50 ug/m^3^ (heavy smoke).

Trying out the distributed lag non-linear package again. I'm going to limit to SABA time-stratified case-crossover dataframe in Medford to simplify.

```{r saba_medford}
# subset medfore saba
saba_medford <- asthma_pos_list$saba %>% 
  filter(metroarea == "Medford") %>% 
  # remove missing values for lag
  filter(!is.na(geo_pm10_lag5))
```

Set up. I'm going to try a 3 knot ns spline for the lag funciton and a 2 knot spline for exposure response to PM~2.5~.

```{r}
# extract geo_10 lagged matrix in same order as dataframe
pm_lag_mat <- saba_medford %>% 
  select(contains("geo_pm10")) %>% 
  as.matrix()
# view matrix
head(pm_lag_mat)

# crossbasis
cb_pm <- crossbasis(pm_lag_mat, lag = 5, argvar = list("ns", df = 2),
                    arglag = list(fun="ns", knots=c(2,4), intercept = T))

# clogit model
cb_mod <- clogit(outcome ~ cb_pm + strata(identifier), data = saba_medford)

# predict
cb_pred <- crosspred(basis = cb_pm, model = cb_mod, cen = 6, at = seq(5,30,5))

```

Cumulative effect at different PM~2.5~ levels.

```{r}
# predictor variable
pm_levels <- cb_pred$predvar
# cumulative rel risk/odds ratio from clogit mod
odds_ratio <- cb_pred$allRRfit
odds_ratio

plot(pm_levels, odds_ratio)
```

## Next Steps/Thoughts

- Have Jingyang focus the comparison of monthly rates by MSA to outcomes of SABA monthly fill rate, office visit, and inpatient visit (this might be interesting).

- Is it worth evaluating a cumulative distributed lag effect of a couple days based on Ana Rappold's new study with CVD?

- Should we try a further refinement of classifying smoke exposure, or even not transform PM~2.5~ and really focus on the Medford vs. other MSAs and just make the argument that any elevated PM~2.5~ during this time in Oregon is likely driven by smoke (particularly in Medford)?

### 2018-05-10 r_scripts/analysis/msa_month_count.Rmd
added script for msa plots.

---
title: "Counts of Asthma ED visits and SABA fills by MSA and Month"
author: "Ryan Gan"
date: "2018-05-10"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Introduction

Looking and different counts and rates for asthma ED visits and SABA fills in Oregon by metropolitan statistical area (MSA) and month. I hypothesize that the Douglas-Complex fire that took place in the south-west part of Oregon impacted the MSA of Medford,

```{r libraries}
library(tidyverse)
```

```{r read_counts}
# ed counts
ed_count <- read_csv("../../data/health/msa_month_er_counts.csv") %>% 
  rename(n_ed = n_visit)
# saba counts
saba_count <- read_csv("../../data/health/msa_month_saba_counts.csv") %>% 
  rename(n_saba = n_visit) %>% 
  mutate(rate_per1000 = (n_saba/POPESTIMATE2013)*1000,
         msa_name = gsub('(.*)-.*', '\\1', 
                      gsub('(.*)-.*', '\\1', gsub('(.*),.*','\\1',NAME))),
         month_name = factor(month.abb[month], 
            levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul",
                       "Aug", "Sep", "Oct", "Nov", "Dec")))
saba_count
```

Barchart of SABA counts by MSA by month.

```{r saba_hist}
# saba histogram
saba_bar_plot <- ggplot(saba_count, aes(x = msa_name, group = msa_name,
                                        fill = msa_name)) +
  geom_bar(aes(weight = rate_per1000)) +
  scale_fill_manual("Metro Area", values = c("#3c1053", "#7f00ff", "#e100ff", 
                                             "#4389a2", "#0ed2f7", "#74ebd5")) +
  facet_wrap(~month_name) +
  ylab("SABA Fill Rate per 1,000 Persons") +
  xlab("Metropolitan Statistical Area") +
  theme_minimal() +
  theme(axis.text.x=element_text(angle = 45, hjust = 1))

saba_bar_plot
```

Alternate of month by metro area.

```{r saba_month_plot}
# saba histogram second version
saba_bar_plot2 <- ggplot(saba_count, aes(x = month_name, group = msa_name,
                                        fill = msa_name)) +
  geom_bar(aes(weight = rate_per1000)) +
  scale_fill_manual("Metro Area", values = c("#3c1053", "#7f00ff", "#e100ff", 
                                             "#4389a2", "#0ed2f7", "#74ebd5")) +
  facet_wrap(~msa_name) +
  ylab("SABA Fill Rate per 1,000 Persons") +
  xlab("Metropolitan Statistical Area") +
  theme_minimal() +
  theme(axis.text.x=element_text(angle = 45, hjust = 1))

saba_bar_plot2
```

### 2018-05-16 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
updated
saved some plots as png.

--- 2017-11-17 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
@@
 smoke_map
+# save figure
+ggsave(filename = "./smoke_map.png", plot = smoke_map)
 ```
@@
 # use or_counties_grid from geo_facet 
-or_grid <- geofacet::or_counties_grid1 %>% 
+or_grid <- geofacet::us_or_counties_grid1 %>% 
   mutate(name = ifelse(name == "HoodRiver", "Hood River", name))

### 2018-05-18 r_scripts/analysis/oregon_part_c.Rmd
update the oregon saba $ asthma er data set visits by MSA by month

---
title: "Oregon average visit rates for saba & asthma emergency data set for metrapolitan per month"
author: "Jingyang Liu"
date: "April 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
# knitr options
knitr::opts_chunk$set(fig.width = 8, fig.height = 6)
knitr::opts_chunk$set(echo=F, warning=F, message=F)
```

## Overview

This is for ER asthma claims and saba data set.To check the if the counts are affected by specific month or by MSA.

1. basically making a time-series summary table by month and by MSA
make summary data table of each metropolitan's each month's visiting number. There are "Bend-Redmond", "Corvallis", "Eugene-Springfield", "Medford", "Portland-Vancouver-Beaverton" and "Salem". We may focus on "Medford" and during the wildfire season's months. (June to August is expected?)  
(can also explore the possibility of calculating average number of events per person like Sheryl suggested, but Ryan's guess is most people will have 0 events)  
(We will have inpatient hospitalization and office visit rates during this time in addition to SABA and ER fills)

2. running Poisson regression. (or check with Quasi-Poisson)
Then run this as generalized regression model with Poission distribution. But would also consider the family of quasipoisson rather than poisson since Ryan think we might violate the assumption of the poisson regression where mu = variance. And the poisson model indeed has overdispersion.

### Prepare asthma emergency room visits data set

**Library:**  
tidyverse is for data cleaning, also includes ggplot and readr.  
data.table is for quick reading files.  
htmlTable is for writing table.  
multcomp is for multiple comparison. Use it for different groups' contrasts in model analysis

```{r library, message=FALSE, warning=FALSE}
library(tidyverse)
library(data.table)
library(htmlTable)
library(multcomp)
library(broom)
# library(contrast)
```

Then load the data set from saba data set and asthma claims data set. 
Filter pos of ER and non-missing MSA data set (not in MSA was removed). 
And import the population for metropolitan in 2013. 20k observations in total.

Use the population table from 
https://www.census.gov/data/datasets/2017/demo/popest/total-metro-and-micro-statistical-areas.html#tables  
(The table offers Portland-Vancouver-Hillsboro area population, not separate.)

```{r load ER data set and zip smoke, message=FALSE, warning=FALSE}

## saba data set
read_path_ndc <- paste0("../data_new/health/2013_oregon_saba_ndc_claims.csv")
saba_df <- read_csv(read_path_ndc)
## asthma data set
read_path_asthma <- paste0("../data_new/health/2013_oregon_asthma_claims.csv")
asthma_df <- read_csv(read_path_asthma)

# (https://www.oregonmetro.gov/news/portland-region-grows-235-million-residents-census-estimates-newcomers-leading-way but it lacks the Portland.)
# read_path_pop <- paste0("../instructions/PEP_2013_PEPANNCHG.US24PR/PEP_2013_PEPANNCHG.US24PR_with_ann.csv")
# pop_df <- read_csv(read_path_pop)

## metropolitan population table
read_path_metro_pop <- paste0("../instructions/cbsa-est2017-alldata.csv")
metro_pop_df <- read_csv(read_path_metro_pop)

# filter Oregon
or_metro_pop <- metro_pop_df %>%
  filter(grepl('OR', metro_pop_df$NAME)) 
# filter useful variables, remove redundant information
or_msa_pop <- or_metro_pop %>%
  filter(grepl('Metropolitan', or_metro_pop$LSAD)) %>%
  dplyr::select(CBSA, NAME, POPESTIMATE2013) %>% # if not add dplyr, it shows error, maybe conflict with other packages
  rename(MSA = CBSA)

## filter ER pos
asthma_er <- asthma_df %>%
  filter(pos == 23)

# inpatient and office
# asthma_inp <- asthma_df %>%
#   filter(pos == 21)
# asthma_office <- asthma_df %>%
#   filter(pos == 11)

```

### Prepare SABA data set from ACAP

Saba data set contains most observations (463k observations). Filter the non missing MSA observations. Join with metropolitan's population in 2013.

The plots show that pharmacy visiting numbers are large on Jan and Oct

From the total visits plot, the Portland have most visits, which can be refered from the population table since it has most population.

For pharmacy in Medford during 2013. June is lowest while July and August shows pretty sharp increase. (and July to August has the sharpest slope, which may indicate the wildfire's effect.)

```{r saba, message=FALSE, warning=FALSE}
pharmacy_df <- saba_df %>%
  mutate(age = 2013 - year) %>%
  mutate(age_cat = case_when(age < 15 ~ "Age < 15",
                                age >= 15 & age < 65 ~ "Age 15-65",
                                age >= 65 ~ "Age > 65"),
              metroarea = case_when(MSA == 13460 ~ "Bend-Redmond",
                                MSA == 18700 ~ "Corvallis",
                                MSA == 21660 ~ "Eugene-Springfield",
                                MSA == 32780 ~ "Medford",
                                MSA == 38900 ~ "Portland-Vancouver-Beaverton",
                                MSA == 41420 ~ "Salem")) %>%
  mutate(month = month(fromdate)) %>%
  inner_join(or_msa_pop[,c(1,3)], by = "MSA") # 352,150 visits, 131,725 people

# summary(as.factor(total_df$metroarea))

## by MSA and month
summary_table_pharmacy <- pharmacy_df %>%
  group_by(MSA, month) %>%
  summarise(n_visit = n()) %>%
  inner_join(or_msa_pop, by = "MSA") %>%
  mutate(rate = n_visit/POPESTIMATE2013*100)

p1 <- ggplot(data = summary_table_pharmacy, aes(x = month, y = n_visit)) +
  geom_line(aes(color = NAME), size = 1) +
  geom_point(aes(color = NAME), size = 1) +
  scale_x_continuous(breaks=c(2,4,6,8,10,12)) +
  ggtitle("Pharmacy visits during year 2013 in different metropolitan") +
  xlab("Month") + ylab("Visiting number")

p2 <- ggplot(data = summary_table_pharmacy, aes(x = month, y = n_visit/POPESTIMATE2013)) +
  geom_line(aes(color = NAME), size = 1) +
  geom_point(aes(color = NAME), size = 1) +
  scale_x_continuous(breaks=c(2,4,6,8,10,12)) +
  ggtitle("Pharmacy visiting rates of population during year 2013 in different metropolitan") +
  xlab("Month") + ylab("Visiting rates")

```

```{r not for now, echo=FALSE, message=FALSE, warning=FALSE, eval = FALSE, include=FALSE}

summary_table_pharmacy2 <- pharmacy_df %>%
  group_by(personkey, MSA, month) %>%
  summarise(n_visit = n()) %>%
  inner_join(or_msa_pop, by = "MSA") %>%
  mutate(metroarea = case_when(MSA == 13460 ~ "Bend-Redmond",
                                MSA == 18700 ~ "Corvallis",
                                MSA == 21660 ~ "Eugene-Springfield",
                                MSA == 32780 ~ "Medford",
                                MSA == 38900 ~ "Portland-Vancouver-Beaverton",
                                MSA == 41420 ~ "Salem")) %>%
  select(-NAME) %>% # 274,213 visits, 1306487 are blank
  data.frame()

```

```{r message=FALSE, warning=FALSE}

# change the numeric variable as factor
summary_table_pharmacy$month <- as.factor(summary_table_pharmacy$month)
summary_table_pharmacy$MSA <- as.factor(summary_table_pharmacy$MSA)

summary_medford <- summary_table_pharmacy %>%
  filter(MSA == 32780)
# write_csv(summary_medford, "../data_new/health/medford_summary.csv")

fit2 <- glm(n_visit ~ month, family = poisson, data = summary_medford)
# summary(fit2)

# htmlTable(summary_table_pharmacy,
#          header =  paste(c("MSA", "Month", "n_visit", "Name", "Population_2013", "Rate")))

# ggplot(data = summary_medford, aes(x = as.numeric(month), y = n_visit)) +
#   geom_line(size = 1) +
#   geom_point(size = 1) +
#   scale_x_continuous(breaks=c(2,4,6,8,10,12)) +
#   ggtitle("Pharmacy visits during the year in different metropolitan") +
#   xlab("Month") + ylab("Visiting number") +
#   geom_text(aes(label=n_visit), vjust=-0.5)

```

Looking and different counts and rates for asthma ED visits and SABA fills in Oregon by metropolitan statistical area (MSA) and month.   
Ryan hypothesize that the Douglas-Complex fire that took place in the south-west part of Oregon impacted the MSA of Medford,

```{r message=FALSE, warning=FALSE}

asthma_er_df <- asthma_er %>%
  mutate(age = 2013 - year) %>%
  mutate(age_cat = case_when(age < 15 ~ "Age < 15",
                                age >= 15 & age < 65 ~ "Age 15-65",
                                age >= 65 ~ "Age > 65"),
              metroarea = case_when(MSA == 13460 ~ "Bend-Redmond",
                                MSA == 18700 ~ "Corvallis",
                                MSA == 21660 ~ "Eugene-Springfield",
                                MSA == 32780 ~ "Medford",
                                MSA == 38900 ~ "Portland-Vancouver-Beaverton",
                                MSA == 41420 ~ "Salem")) %>%
  mutate(month = month(fromdate)) %>%
  inner_join(or_msa_pop[,c(1,3)], by = "MSA") # 352,150 visits, 131,725 people

# summary(as.factor(total_df$metroarea))

## by MSA and month
summary_table_er <- asthma_er_df %>%
  group_by(MSA, month) %>%
  summarise(n_visit = n()) %>%
  inner_join(or_msa_pop, by = "MSA") %>%
  mutate(rate = n_visit/POPESTIMATE2013*100)

p1 <- ggplot(data = summary_table_er, aes(x = month, y = n_visit)) +
  geom_line(aes(color = NAME), size = 1) +
  geom_point(aes(color = NAME), size = 1) +
  scale_x_continuous(breaks=c(2,4,6,8,10,12)) +
  ggtitle("Emergency visits during year 2013 in different metropolitan") +
  xlab("Month") + ylab("Visiting number")

p2 <- ggplot(data = summary_table_er, aes(x = month, y = n_visit/POPESTIMATE2013)) +
  geom_line(aes(color = NAME), size = 1) +
  geom_point(aes(color = NAME), size = 1) +
  scale_x_continuous(breaks=c(2,4,6,8,10,12)) +
  ggtitle("Emergency visiting rates of population during year 2013 in different metropolitan") +
  xlab("Month") + ylab("Visiting rates")
```

```{r}
ed_count <- summary_table_er %>% 
  rename(n_ed = n_visit) %>% 
  mutate(rate_per10k = (n_ed/POPESTIMATE2013)*10000,
         msa_name = gsub('(.*)-.*', '\\1', 
                      gsub('(.*)-.*', '\\1', gsub('(.*),.*','\\1',NAME))),
         month_name = factor(month.abb[month], 
            levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul",
                       "Aug", "Sep", "Oct", "Nov", "Dec")))
# saba counts
saba_count <- summary_table_pharmacy %>% 
  rename(n_saba = n_visit) %>% 
  mutate(rate_per1000 = (n_saba/POPESTIMATE2013)*1000,
         msa_name = gsub('(.*)-.*', '\\1', 
                      gsub('(.*)-.*', '\\1', gsub('(.*),.*','\\1',NAME))),
         month_name = factor(month.abb[month], 
            levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul",
                       "Aug", "Sep", "Oct", "Nov", "Dec")))
saba_count

```

1. Barchart of SABA counts by MSA by month.

Medford in August has higher visits.

```{r}
# saba histogram
saba_bar_plot <- ggplot(saba_count, aes(x = msa_name, group = msa_name,
                                        fill = msa_name)) +
  geom_bar(aes(weight = rate_per1000)) +
  scale_fill_manual("Metro Area", values = c("#3c1053", "#7f00ff", "#e100ff", 
                                             "#4389a2", "#0ed2f7", "#74ebd5")) +
  facet_wrap(~month_name) +
  ylab("SABA Fill Rate per 1,000 Persons") +
  xlab("Metropolitan Statistical Area") +
  theme_minimal() +
  theme(axis.text.x=element_text(angle = 45, hjust = 1))

saba_bar_plot
```

2. Alternate of month by metro area.

The visiting number is higher in October.

```{r}
# saba histogram second version
saba_bar_plot2 <- ggplot(saba_count, aes(x = month_name, group = msa_name,
                                        fill = msa_name)) +
  geom_bar(aes(weight = rate_per1000)) +
  scale_fill_manual("Metro Area", values = c("#3c1053", "#7f00ff", "#e100ff", 
                                             "#4389a2", "#0ed2f7", "#74ebd5")) +
  facet_wrap(~msa_name) +
  ylab("SABA Fill Rate per 1,000 Persons") +
  xlab("Metropolitan Statistical Area") +
  theme_minimal() +
  theme(axis.text.x=element_text(angle = 45, hjust = 1))

saba_bar_plot2
```

3. Emergency room rates.

Medford has lower visiting number in August.

```{r}
# saba histogram second version
er_bar_plot2 <- ggplot(ed_count, aes(x = msa_name, group = msa_name,
                                        fill = msa_name)) +
  geom_bar(aes(weight = rate_per10k)) +
  scale_fill_manual("Metro Area", values = c("#3c1053", "#7f00ff", "#e100ff", 
                                             "#4389a2", "#0ed2f7", "#74ebd5")) +
  facet_wrap(~month_name) +
  ylab("Emergency Room Visits per 10,000 Persons") +
  xlab("Metropolitan Statistical Area") +
  theme_minimal() +
  theme(axis.text.x=element_text(angle = 45, hjust = 1))

er_bar_plot2
```

### 1. SABA visits Interaction model

Research Question:  
Null hypothesis: There is no difference of visiting number in each MSA during wildfire season (July to August and August to September).  
Alternative hypothesis: There is difference of visiting number in each MSA during wildfire season (July to August and August to September).   
Guess for Medford: Jul to Aug will be increase. Aug to Sep will be decrease since smoke occurs most in August.

#### For different MSA:
The order is Bend, Corvallis, Eugene, Medford, Portland and Salem.  
contrast statement:     
1. Each MSA's August visits - This MSA's July visits  
2. Each MSA's September visits - This MSA's August visits

```{r message=FALSE, warning=FALSE}

summary_table_pharmacy$group <- factor(paste0(summary_table_pharmacy$month, summary_table_pharmacy$NAME))
X <- model.matrix(~ 0 + group, data=summary_table_pharmacy)
# colnames(X)

fit <- glm(n_visit ~ 0 + month*NAME, offset = log(POPESTIMATE2013), family = poisson, data = summary_table_pharmacy)
# summary(fit)

fit_group <- glm(n_visit ~ 0 + group, offset = log(POPESTIMATE2013), family = poisson, data = summary_table_pharmacy)
# summary(fit_group)

fit_no <- glm(n_visit ~ month + NAME + month:NAME - 1, offset = log(POPESTIMATE2013), family = poisson, data = summary_table_pharmacy)
# summary(fit_no)

contr_msa_jul_to_sep <- rbind(
  "Bend: Aug vs Jul" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0), # 8-7
  "Bend: Sep vs Aug" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0),

  "Corvallis: Aug vs Jul" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0), # 8-7
  "Corvallis: Sep vs Aug" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0),

  "Eugene: Aug vs Jul" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0), # 8-7
  "Eugene: Sep vs Aug" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0),

  "Medford: Aug vs Jul" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0), # 8-7
  "Medford: Sep vs Aug" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0),

  "Portland: Aug vs Jul" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0), # 8-7
  "Portland: Sep vs Aug" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0),

  "Salem: Aug vs Jul" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0), # 8-7
  "Salem: Sep vs Aug" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1))

s1 <- glht(fit_group, contr_msa_jul_to_sep)
summary(s1)

df1 <- as.data.frame(broom::tidy(summary(s1)))

df1 <- df1 %>%
  dplyr::mutate(lower_ci = estimate - 1.96*std.error,
         upper_ci = estimate + 1.96*std.error) %>%
  dplyr::select(lhs, estimate, std.error, lower_ci, upper_ci, p.value)

df1[-1] <- round(df1[-1], 4)

htmlTable(df1,
          caption = "SABA MSA's Visits from July to September",
          header = c("Contrast", "Estimate", "SE", "Lower", "Upper", "p_value"),
          # n.rgroup = c(rep(3, 7)), # 6 rows for each age cat for each outcome
           # column spanner
           cgroup = c("", "95% CI", ""), 
           n.cgroup = c(3, 2, 1),
           padding.rgroup = "&nbsp;&nbsp;",
           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
           align = "llcccc")
```

The result shows:  
Bend: significant increasing visits during Jul to Aug with exp(0.12) = 1.127 times increasing rate.  
Corvallis: Not showing difference.  
Eugene: signigicant decreasing visits during Aug to Sep with 15% decreasing rate.  
Medford: significant increasing visits during Jul to Aug with 12.7% increasing rate. and signigicant decreasing visits during Aug to Sep with 14% decreasing rate.  
Portland: signigicant decreasing visits during Aug to Sep with 10.5% decreasing rate.  
Salem: signigicant decreasing visits during Aug to Sep with 12.7% decreasing rate.

So guess is true.

#### For August' visits difference in different MSAs
(although the change rate is more important)

Research Question:  
Null hypothesis: There is no difference of visiting number between Medford and other MSA in August.   
Alternative hypothesis: There is difference of visiting number between Medford and other MSA in August.  
Guess: Since Medford is the wildfire center in August, it may show more visiting number than other MSAs.

contrast statement:
Medford's August visits - Other MSA's August visits

```{r}
contr_aug_msa_medford <- rbind(
  "Aug: Medford vs Bend" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,1,0,0,0,0,0,0,0,0), 
  "Aug: Medford vs Corvallis" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,1,0,0,0,0,0,0,0,0),
  "Aug: Medford vs Eugene" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,1,0,0,0,0,0,0,0,0), 
  "Aug: Medford vs Portland" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,-1,0,0,0,0,0,0,0),
  "Aug: Medford vs Salem" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,-1,0,0,0,0,0,0))

s2 <- summary(glht(fit_group, contr_aug_msa_medford))
summary(s2)

df2 <- as.data.frame(broom::tidy(summary(s2)))

df2 <- df2 %>%
  dplyr::mutate(lower_ci = estimate - 1.96*std.error,
         upper_ci = estimate + 1.96*std.error) %>%
  dplyr::select(lhs, estimate, std.error, lower_ci, upper_ci, p.value)

df2[-1] <- round(df2[-1], 4)

htmlTable(df2,
          caption = "SABA Medford's Visits vs Other MSA's Visits in August",
          header = c("Contrast", "Estimate", "SE", "Lower", "Upper", "p_value"),
          # n.rgroup = c(rep(3, 7)), # 6 rows for each age cat for each outcome
           # column spanner
           cgroup = c("", "95% CI", ""), 
           n.cgroup = c(3, 2, 1),
           padding.rgroup = "&nbsp;&nbsp;",
           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
           align = "llcccc")
```

The result shows:  
All Medford's visits in Aug is larger than that in other MSAs. And all are significant except the Salem. 
For example, the estimate of Medford vs Portland is 0.408, which means: (Medford's visiting rate) / (Portland's visiting rate) is exp(0.408) = 1.5 (actual is 1.007/0.6693) is also 1.5

So guess is basically true.

### 2. Emergency Room visits Interaction Model

But the emergency data set shows different rerults with pharmacy data set.

For Medford, the number in August is pretty small, while in September is pretty large.

#### For different MSA

Research Question:  
Null hypothesis: There is no difference of visiting number in each MSA during wildfire season (July to August and August to September).  
Alternative hypothesis: There is difference of visiting number in each MSA during wildfire season (July to August and August to September).

contrast statement:   
1. Each MSA's August visits - This MSA's July visits  
2. Each MSA's September visits - This MSA's August visits

```{r message=FALSE, warning=FALSE}

# change the numeric variable as factor
summary_table_er$month <- as.factor(summary_table_er$month)
summary_table_er$MSA <- as.factor(summary_table_er$MSA)

summary_table_er$group <- factor(paste0(summary_table_er$month, summary_table_er$NAME))
X <- model.matrix(~ 0 + group, data=summary_table_er)
# colnames(X)

fit <- glm(n_visit ~ 0 + month*NAME, offset = log(POPESTIMATE2013), family = poisson, data = summary_table_er)
# summary(fit)

fit_group <- glm(n_visit ~ 0 + group, offset = log(POPESTIMATE2013), family = poisson, data = summary_table_er)
# summary(fit_group)

fit_no <- glm(n_visit ~ month + NAME + month:NAME - 1, offset = log(POPESTIMATE2013), family = poisson, data = summary_table_er)
# summary(fit_no)

contr_msa_jul_to_sep <- rbind(
  "Bend: Aug vs Jul" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0), # 8-7
  "Bend: Sep vs Aug" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0),

  "Corvallis: Aug vs Jul" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0), # 8-7
  "Corvallis: Sep vs Aug" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0),

  "Eugene: Aug vs Jul" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0), # 8-7
  "Eugene: Sep vs Aug" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0),

  "Medford: Aug vs Jul" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0), # 8-7
  "Medford: Sep vs Aug" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0),

  "Portland: Aug vs Jul" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0), # 8-7
  "Portland: Sep vs Aug" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0),

  "Salem: Aug vs Jul" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0), # 8-7
  "Salem: Sep vs Aug" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1))

s3 <- summary(glht(fit_group, contr_msa_jul_to_sep))
summary(s3)

df3 <- as.data.frame(broom::tidy(summary(s3)))

df3 <- df3 %>%
  dplyr::mutate(lower_ci = estimate - 1.96*std.error,
         upper_ci = estimate + 1.96*std.error) %>%
  dplyr::select(lhs, estimate, std.error, lower_ci, upper_ci, p.value)

df3[-1] <- round(df3[-1], 4)

htmlTable(df3,
          caption = "Asthma ER MSA's Visits from July to September",
          header = c("Contrast", "Estimate", "SE", "Lower", "Upper", "p_value"),
          # n.rgroup = c(rep(3, 7)), # 6 rows for each age cat for each outcome
           # column spanner
           cgroup = c("", "95% CI", ""), 
           n.cgroup = c(3, 2, 1),
           padding.rgroup = "&nbsp;&nbsp;",
           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
           align = "llcccc")
```

The result shows:  
Bend: significant increasing visits during Aug to Sep with exp(0.12) = 2.27 times increasing rate.  
Corvallis: Not showing difference.  
Eugene: significant decreasing visits during Jul to Aug and significant decreasing visits.  
Medford: significant increasing visits during Aug to Sep. 
Portland: signigicant incresing visits during Aug to Sep.  
Salem: signigicant decreasing visits during Aug to Sep with 12.7% decreasing rate.

So this result shows pretty different with SABA data set, especially in Medford, which shows opposite trends.

#### For August' visits difference in different MSAs

Research Question:  
Null hypothesis: There is no difference of visiting number between Medford and other MSA in August.   
Alternative hypothesis: There is difference of visiting number between Medford and other MSA in August.  
Guess: Since Medford is the wildfire center in August, it may show more visiting number than other MSAs. (But ER visits are different from SABA)

contrast statement:
Medford's August visits - Other MSA's August visits

```{r}
contr_aug_msa_medford <- rbind(
  "Aug: Medford vs Bend" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,1,0,0,0,0,0,0,0,0), 
  "Aug: Medford vs Corvallis" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,1,0,0,0,0,0,0,0,0),
  "Aug: Medford vs Eugene" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,1,0,0,0,0,0,0,0,0), 
  "Aug: Medford vs Portland" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,-1,0,0,0,0,0,0,0),
  "Aug: Medford vs Salem" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,-1,0,0,0,0,0,0))

s4 <- summary(glht(fit_group, contr_aug_msa_medford))
summary(s4)

df4 <- as.data.frame(broom::tidy(summary(s4)))

df4 <- df4 %>%
  dplyr::mutate(lower_ci = estimate - 1.96*std.error,
         upper_ci = estimate + 1.96*std.error) %>%
  dplyr::select(lhs, estimate, std.error, lower_ci, upper_ci, p.value)

df4[-1] <- round(df4[-1], 4)

htmlTable(df4,
          caption = "Asthma ER Medford's Visits vs Other MSA's visits in August",
          header = c("Contrast", "Estimate", "SE", "Lower", "Upper", "p_value"),
          # n.rgroup = c(rep(3, 7)), # 6 rows for each age cat for each outcome
           # column spanner
           cgroup = c("", "95% CI", ""), 
           n.cgroup = c(3, 2, 1),
           padding.rgroup = "&nbsp;&nbsp;",
           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
           align = "llcccc")

```

The result shows:  
The Medford's visits in August is larger than that in Bend, Corvallis and Portland, but none of them is significant. And it is significantly smaller than that in Eugene and Salem in August.

### Overall

SABA data set shows Medford's visits in August is higher than other MSA while ER visits data set shows that it is not higher.

SABA data set shows Medford's visits in August is higher than than in July and September while ER visits shows the visits in September is higher than that in August.

### 2018-05-30 r_scripts/analysis/oregon_wfs_ed_results.Rmd
renamed part_a and removed some code. Changed figure colors and added back in arrhythmia.

---
title: "Oregon GWR smoke association with Cardiopulmonary Emergency Room visits"
author: "Jingyang Liu"
date: "April 9, 2018"
output: html_document
---

## Overview

This project researches on the association between geographically weighted regression (GWR) smoke PM~2.5~ and cardiopulmonary (cardiovascular (CVD) and respiratory health outcomes) diseases for Oregon State in wildfire season (May to September) in 2013. 
This markdown document focuses on the zip code-level analysis. 
The analysis includes total analysis, age strata and sex strata.

## Methods Description

In these research, we mainly examine GWR PM~2.5~ estimations and associations with health outcomes using a time-stratified case-crossover study design. The data set contains all claims of Oregon for total 2013 year, so we choose only the emegency room (hospital) visits for this analysis. Health outcomes for a patient with a primary diagnosis of cardiopulmonary health outcomes and their date of admission  (index time) were identified. We then created counterfactual observations for each patient for the same day of the week for the entire wildfire season (MAy 1st to Sep 30th, 2013). We further limited our analyses to claims that were coded as emergency or urgent visits to eliminate bias from patients going in for elective/planned procedures.  For patients who have a index time closer to May 1, 2013 or closer to September 30, 2013, their referent observations before or after these dates will be excluded as we will not be able to assign estimates of PM~2.5~ to these referent observations. However, this is not a big issue for a time-stratified design as other referent values throughout the time period average out.

The **cardiopulmonary health outcomes** are all respiratory, asthma, COPD, cardiovascular disease, heart failure, and ischemic heart disease and cerebrovascular disease.

As for **exposure methods**, we use Geo-Weighted Regression and subtracted off the 'Background' estimates, which are the monthly averages of PM~2.5~ for a given grid.

The **analytic method** is the conditional logistic regression model using the *survival* package in R. This regression is controlling the time-invariant subject-specific counfounding like age and sex. Each conditional logistic regression model accounts for the subject, and adjusts for temperature from the WRF-Chem model (Weather Research and Forecasting model coupled with Chemistry). The conditional logistic regression model \beta represents the change in risk of an event associated with a short-term 10 units' increase in exposure, and can be calculated as an average difference between exposure at the index time and a weighted average of exposure at all times.

**Library:**  
tidyverse is for data cleaning, also includes ggplot and readr.  
data.table is for quick reading files.  
survival is for running conditional logistic regression model.  
htmlTable is for writing table.

```{r library, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(data.table)
library(survival)
library(htmlTable)
```

**Time-Stratified Case-Crossover Data Set (same day of week within wildfire season May - September)**

The data set is from the case crossover data set, which is made before. To save the running time, the data set is wildfire season and people's first visit for same date. Then filter the Emergency postal of service and non-missing gender visits.

```{r ER data making, include = F, eval = FALSE}
### Limit to only ER visits
## pos = 23
var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
              'cerbrovas_dis', 'myocardial_infarc')

N <- NA
n <- 1

for (i in var_list){
  read_path <- paste('../../data/health/or', i, 
                     'may_to_sep_time_strat_casecross.csv', sep = "_")
  or_disease <- read_csv(read_path)

  or_disease <- or_disease %>% 
    filter(pos=="23") %>%
    # filter the unknown gender
    filter(gender!= "U")

  # Create a permanent case-cross over dataset
  file_name <- paste('../data_new/casecrossover/ER/or', i,
                     'may_to_sep_time_strat_casecross_er.csv', sep = "_")

  # write permanent dataset
  write_csv(or_disease, paste0(file_name))
  N[n] = length(unique(or_disease$personkey))
  n = n+1
}

# check the number of people in each disease
num_person <- data.frame(rbind(var_list, N))
```

Then load the Emergency data set.

```{r load ER data set and zip smoke, echo=FALSE, message=FALSE, warning=FALSE}
## Analysis
path <- paste0("../../data/health/")

resp_casecross <- read_csv(paste(path, 
                                 "or_respiratory_may_to_sep_time_strat_casecross_er.csv", 
                                 sep = "/"))

# asthma
asthma_casecross <- read_csv(paste(path, 
                                   "or_asthma_may_to_sep_time_strat_casecross_er.csv", 
                                   sep = "/"))

# copd 
copd_casecross <- read_csv(paste(path, 
                                 "or_copd_may_to_sep_time_strat_casecross_er.csv", 
                                 sep = "/"))

# cvd
cvd_casecross <- read_csv(paste(path, 
                                "or_cvd_may_to_sep_time_strat_casecross_er.csv",
                                sep="/"))
# cerebral vascular
cereb_vas_casecross <- read_csv(paste(path, 
                                      "or_cerbrovas_dis_may_to_sep_time_strat_casecross_er.csv",
                                      sep="/"))

# arrythmia
arrhythmia_casecross <- read_csv(paste(path, 
                               "or_arrhythmia_may_to_sep_time_strat_casecross_er.csv", 
                               sep="/"))

# heart failure
hf_casecross <- read_csv(paste(path, 
                               "or_heart_failure_may_to_sep_time_strat_casecross_er.csv", 
                               sep="/"))
# ischemic heart disease
ihd_casecross <- read_csv(paste(path, 
                                "or_isch_heart_dis_may_to_sep_time_strat_casecross_er.csv",
                                sep="/"))

```

### Descriptive Table

This table shows number of cardiopulmonary visits for each outcome observed from May 1st to September 30st in 2013. Sex and Age strata for each outcome are also shown in this table. Female and Male show similar proportion in each outcome. As for age groups, adult group shows most proportion in asthma outcome than other age groups, and also it is the more likely disease adults tend to have. Older people tend to get cardiovascular disease while children tend to get respiratory diseases, which makes sense.

*Missing smoke value data is already removed for table and next analysis, although only a few.*

```{r descriptive table, echo = F, warning = F, results='asis'}

# dataframe list
df_list <- list(resp_casecross, asthma_casecross, copd_casecross, 
  cvd_casecross, arrhythmia_casecross, cereb_vas_casecross, hf_casecross, 
  ihd_casecross)

outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 
                  'Cardiovascular Disease', 'Arrhythmia', 
                  'Cerebrovascular Disease', 'Heart Failure',
                  'Ischemic Heart Disease')

method_list <- c('Geo-Weighted Smoke')

# create an empty list to row bind dataframes together
datalist <- list()

# set list and do cross tabs to find values
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){
# dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- colnames(df_to_loop[76])
  # outcome name
  outcome_name <- outcome_list[i]

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- df_to_loop %>%
    filter(!is.na(geo_smk_pm_zip)) %>% 
    # only look at outcomes
    filter(outcome == 1) %>%
    # add another row that makes sure there is a person <15 in the dataframe
    # tricking xtabs to produce a 0 cell for the outcome for age <15
    add_row(outcome = 0, age_ind = 0)

  # cross tabs
  outcome_n <- xtabs(~ outcome, df_analysis)
  cross_tab_age <- xtabs(~ outcome + age_ind, df_analysis)
  cross_tab_sex <- xtabs(~ outcome + sex_ind, df_analysis)
  # empty matrix
  point_estimates <- matrix(nrow = 1, ncol = 7, byrow = T)

  colnames(point_estimates) <- c("outcome", "n", "age_15", "age_15_65", 
                                 "age_65", "female", "male")

  # fill in the outcome name for the dataframe before the loop
  point_estimates[, 1] <- outcome_name
  # fill n
  point_estimates[, 2] <- outcome_n[2] # second element of the 1 dimension vector
  # age <15
  point_estimates[, 3] <- cross_tab_age[2, 1]
  # age 15 to 65
  point_estimates[, 4] <- cross_tab_age[2, 2]
  # age >65
  point_estimates[, 5] <- cross_tab_age[2, 3]
  # male == 0
  point_estimates[, 7] <- cross_tab_sex[1, 1]
  # female == 1
  point_estimates[, 6] <- cross_tab_sex[1, 2]

  # save point estimates as a dataframe
  point_est_df <- as_data_frame(point_estimates)

  # combine previous values in dataframe that has all outcome/methods comparisons
  datalist[[i]] <- point_est_df

} # end of loop

# combine each outcome dataframe itteration in to a big dataset
tab1 <- bind_rows(datalist) %>% 
  # find proportions/percents for each strata in a row
  mutate(age_15_pr = as.character(round((as.numeric(age_15)/as.numeric(n))*100,1)),
         age_15_65_pr = as.character(round((as.numeric(age_15_65)/as.numeric(n))*100,1)),
         age_65_pr = as.character(round((as.numeric(age_65)/as.numeric(n))*100,1)),
         female_pr = as.character(round((as.numeric(female)/as.numeric(n))*100,1)),
         male_pr = as.character(round((as.numeric(male)/as.numeric(n))*100,1))) %>% 
  select(outcome, n, age_15, age_15_pr, age_15_65, age_15_65_pr, age_65,
         age_65_pr, female, female_pr, male, male_pr)

knitr::kable(tab1, caption = "Descriptive Table")
```

### Total Outcome Analysis

The following table shows the total outcome without strata.

Since some event numbers in "less than 15" group are too small, then we filter the visits which are no more than 20, and add them to "15-65" group, and set the events, OR, and CI are 0. (But the descriptive table above shows real data summary)

```{r echo = F, warning = F, results='asis'}
# data wrangling ----
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){

  # dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- colnames(df_to_loop[76])
  # outcome name
  outcome_name <- outcome_list[i]

  # extract covariates from dataframe
  covariates_df <- df_to_loop[, c(1:26, 71, 74, 76:85, 92, 94)] %>%
    mutate(geo_smk_pm_zip10 = geo_smk_pm_zip/10)

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- covariates_df %>% 
    # remove missing pm values
    filter(!is.na(geo_smk_pm_zip10)) %>% 
    mutate(obs_diff_admission = (fromdate - date)/7) 
  # dataframe is already for the entire fire season, so I don't need to subset anymore

  # empty df for table
  table_df <- data.frame()

  # empty matrix
  point_estimates <- matrix(nrow = 1, ncol = 9, byrow = T)

  colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 'odds_ratio', 
                                 'lower95', 'upper95', 'se', 'p_val')

  # fill in the outcome name for the dataframe before the loop
  point_estimates[, 1] <- outcome_name

  # second loop to run a model for each pm estimation method

  j <- 39

    # variable to model 
    var_name <- colnames(df_analysis[j])

    # conditional logistic regression model
    mod <- clogit(outcome ~ geo_smk_pm_zip10 + wrf_temp_zip + strata(personkey),
                  df_analysis)

    # populate matrix
    row_n <- j-38

    point_estimates[row_n, 2] <- method_list[row_n]
    point_estimates[row_n, 3] <- mod$n
    point_estimates[row_n, 4] <- mod$nevent
    # odds ratio
    point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)

    # 95% lower bound
    point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -
                                             1.96*(summary(mod)$coefficient[1,3])), 3)
    # 95% upper bound
    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
                                             1.96*(summary(mod)$coefficient[1,3])), 3)
    # standard error
    point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
    # p val
    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)

    # save point estimates as a dataframe
    point_est_df <- as_data_frame(point_estimates)

  # combine previous values in dataframe that has all outcome/methods comparisons
  datalist[[i]] <- point_est_df

} # end of loop

# combine each outcome dataframe itteration in to a big dataset
ed_results <- bind_rows(datalist)

knitr::kable(ed_results, caption = "GWR and ED admissions results")
```

### Time-Stratified by Age Categories

Following table and figure shows some outcomes stratified by age category. Note that kids <=15 likely does not have many outcomes like COPD or MI. I have restricted the dataset to just the geo smoke method to reduce the number of rows/points on the tables and figures to make it easier to read for now.

For some outcomes, there may be some evidence of effect modificaiton, like COPD, where >65 age group is at significant risk. However, the point estimates are relatively similar for the 15 to 65 strata, just not significant. Some error bars are not shown as they are too wide; usually <15 age category where there are only a handful of outcomes.

```{r time stratified age strata cdc met, warning =F, echo = F, results='asis'} 
# age category list
age_cat_list <- c(0,1,2)

# create an empty list to row bind dataframes together
datalist1 <- list()
datalist2 <- list()

var_list <- c('respiratory', 'asthma', 'copd', 
              'cvd', 'isch_heart_dis',  'heart_failure', 
              'cerbrovas_dis')

# data wrangling ----
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){

# dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- var_list[i]
  # outcome name
  outcome_name <- outcome_list[i]

  # extract covariates from dataframe
  covariates_df <- df_to_loop[, c(1:26, 71, 74, 76:85, 92, 94)] %>%
    mutate(geo_smk_pm_zip10 = geo_smk_pm_zip/10)

  # new loop for age categories
  for(k in 0:2){

    # empty matrix (12 x 10 matrix)
    point_estimates <- matrix(nrow = 1, ncol = 10, byrow = T)

    colnames(point_estimates) <- c('outcome', 'age_cat', 'pm_method', 'n', 'n_events', 
                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')

    # fill in the outcome namedataframe before method loop
    point_estimates[, 1] <- outcome_name
    # repeat the age category 4 times for each pm method
    point_estimates[, 2] <- k

    # dataframe for analysis creation
    # bind columns back together 
    df_analysis <- covariates_df %>% 
      # remove missing pm values
      filter(!is.na(geo_smk_pm_zip10)) %>% 
      # limit to specific age category
      filter(age_ind == k) %>% 
      # the following code makes sure that the counterfactual values retained are 
      # symetric in that number of obs before = number of obs after
      mutate(obs_diff_admission = (fromdate - date)/7) 
      # dataframe is already for the entire fire season, so I don't need to subset anymore

  # second loop to run a model for each pm estimation method
    j <- 39

      # variable to model 
      var_name <- colnames(df_analysis[j])

      # set row number to fill
      row_n <- j-38

      # only run the model if the dataframe has observations
      if(nrow(df_analysis) != 0){
      # conditional logistic regression model
      mod <- clogit(outcome ~ geo_smk_pm_zip10 + wrf_temp_zip + strata(personkey), df_analysis)

      # populate matrix
      point_estimates[row_n, 3] <- method_list[row_n]
      point_estimates[row_n, 4] <- mod$n
      point_estimates[row_n, 5] <- mod$nevent
      # odds ratio
      point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)

      # 95% lower bound
      point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
                                        1.96*(summary(mod)$coefficient[1,3])), 3)
      # 95% upper bound
      point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
                                        1.96*(summary(mod)$coefficient[1,3])), 3)
      # standard error
      point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
      # p val
      point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)

      # create else statement that fills matrix with missing so I still have the row
      # in the final dataframe
      } else {point_estimates[row_n, 3] <- method_list[row_n]
              point_estimates[row_n, 4] <- 0
              point_estimates[row_n, 5] <- 0
              point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement

    # save point estimates as a dataframe
    point_est_df <- as_data_frame(point_estimates)

    # combine previous values in dataframe that has all outcome/methods comparisons
    datalist1[[k+1]] <- point_est_df
  } # end age category loop

  # bind rows of age category estimates together
 age_est_df <- as.data.frame(bind_rows(datalist1))
 age_est_df[,5] <- as.numeric(age_est_df[,5])

 if(age_est_df[1,5] <= 20){
   age_est_df[2,5] <- age_est_df[2,5] + age_est_df[1,5]
   age_est_df[1,5] <- 0
   age_est_df[1,6] <- 0
   age_est_df[1,7] <- 0
   age_est_df[1,8] <- 0
 }

 # populate second dataframe list
 datalist2[[i]] <- age_est_df

} # end of outcome loop

# combine each outcome dataframe itteration in to a big dataset
ed_age_results <- bind_rows(datalist2) %>% 
  mutate(age_cat2 = ifelse(age_cat == 0, "Less than 15", 
                    ifelse(age_cat == 1, "15-65",
                    ifelse(age_cat == 2, "Greater than 65", NA)))) %>% 
  # subset columns I want to put in to the table
  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
         lower95 = ifelse(lower95 == 99, "--", lower95),
         upper95 = ifelse(upper95 == 99, "--", upper95)) %>% 
  filter(n_events >= 50)

knitr::kable(ed_age_results, caption = "Age Stratified Results")
```

### Seasonal Time-Stratified by Sex

Association with smoke PM~2.5~ and health outcomes stratified by sex. May be some evidence of effect modification by sex, particularly where males appear to be at greater risk for certain CVD outcomes with increasing smoke. However, only marginally significant.

```{r sex strata cdc met adj, warning = F, echo = F, results='asis'} 
# sex category list
sex_strata_list <- c(0,1)

# create an empty list to row bind dataframes together
datalist1 <- list()
datalist2 <- list()

# data wrangling ----
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){

  # dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- var_list[i]
  # outcome name
  outcome_name <- outcome_list[i]

  covariates_df <- df_to_loop[, c(1:26, 71, 74, 76:85, 92, 94)] %>%
    mutate(geo_smk_pm_zip10 = geo_smk_pm_zip/10)

  # new loop for sex categories
  for(k in 0:1){

    # empty matrix (12 x 10 matrix)
    point_estimates <- matrix(nrow = 1, ncol = 10, byrow = T)

    colnames(point_estimates) <- c('outcome', 'sex', 'pm_method', 'n', 'n_events', 
                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')

    # fill in the outcome namedataframe before method loop
    point_estimates[, 1] <- outcome_name
    # repeat the sex category 4 times for each pm method
    point_estimates[, 2] <- k

    # dataframe for analysis creation
    # bind columns back together 
    df_analysis <- covariates_df %>% 
      # remove missing pm values
      filter(!is.na(geo_smk_pm_zip10)) %>% 
      # limit to specific sex category
      filter(sex_ind == k) %>% 
      # the following code makes sure that the counterfactual values retained are 
      # symetric in that number of obs before = number of obs after
      mutate(obs_diff_admission = (fromdate - date)/7) 
      # dataframe is already for the entire fire season, so I don't need to subset anymore

  # second loop to run a model for each pm estimation method
   j <- 39

      # variable to model 
      var_name <- colnames(df_analysis[j])

      # set row number to fill
      row_n <- j-38

      # only run the model if the dataframe has observations
      if(nrow(df_analysis) != 0){
      # conditional logistic regression model
      mod <- clogit(outcome ~ geo_smk_pm_zip10 + wrf_temp_zip + strata(personkey), df_analysis)

      # populate matrix
      point_estimates[row_n, 3] <- method_list[row_n]
      point_estimates[row_n, 4] <- mod$n
      point_estimates[row_n, 5] <- mod$nevent
      # odds ratio
      point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)

      # 95% lower bound
      point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
                                        1.96*(summary(mod)$coefficient[1,3])), 3)
      # 95% upper bound
      point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
                                        1.96*(summary(mod)$coefficient[1,3])), 3)
      # standard error
      point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
      # p val
      point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)

      # create else statement that fills matrix with missing so I still have the row
      # in the final dataframe
      } else {point_estimates[row_n, 3] <- method_list[row_n]
              point_estimates[row_n, 4] <- 0
              point_estimates[row_n, 5] <- 0
              point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement

    # save point estimates as a dataframe
    point_est_df <- as_data_frame(point_estimates)

    # combine previous values in dataframe that has all outcome/methods comparisons
    datalist1[[k+1]] <- point_est_df
  } # end sex category loop

  # bind rows of sex category estimates together
 sex_est_df <- bind_rows(datalist1)

 # populate second dataframe list
 datalist2[[i]] <- sex_est_df

} # end of outcome loop

# combine each outcome dataframe itteration in to a big dataset
ed_sex_results <- bind_rows(datalist2) %>% 
  mutate(sex_cat = ifelse(sex == 0, "Male", 
               ifelse(sex == 1, "Female", NA))) %>%
  # subset columns I want to put in to the table
  select(1, 3, 11, 5:8)

# table
knitr::kable(ed_sex_results, caption = "Sex-stratified results")
```

### Plots for Total, Sex Strata and Age Strata

The plot for all these strata, total, sex and age.

```{r warning = F, echo = F, results='asis', echo =FALSE, warning = F, message=FALSE}
# strata name in same order as files
strata <- c("age","total", "sex")

age_results <- ed_age_results %>% 
  mutate(strata = case_when(age_cat == 0 ~ "Age < 15",
                            age_cat == 1 ~ "Age 15-65",
                            age_cat == 2 ~ "Age > 65"),
         flag = ifelse(age_cat == 0 & (outcome %in% c("COPD",
                  "Cardiovascular Disease", "Ischemic Heart Disease", 
                  "Heart Failure", "Cerebrovascular Disease")), 1, 0)) %>% 
    # filter out rare events that have unstable ORs
  filter(flag == 0) %>% 
  select(outcome, strata, odds_ratio:upper95)

# total strata
total_results <- ed_results %>% 
  mutate(strata = "All Strata") %>% 
  select(outcome, strata, odds_ratio:upper95)

# sex strata
sex_results <- ed_sex_results %>% 
  mutate(strata = sex_cat) %>% 
  select(outcome, strata, odds_ratio:upper95)

# rowbind
results <- bind_rows(total_results, sex_results, age_results) %>% 
  # filter to certain outcomes
  filter(outcome %in% c("All Respiratory", "Asthma", "COPD", 
                        "Cardiovascular Disease", "Arrhythmia",
                        "Ischemic Heart Disease",
                        "Heart Failure", "Cerebrovascular Disease")) %>% 
  mutate_at(vars(odds_ratio:upper95), as.numeric) %>% 
  # set order of strata levels
  mutate(outcome = factor(outcome, levels = c("All Respiratory", "Asthma", "COPD", 
                        "Cardiovascular Disease", "Arrhythmia",
                        "Ischemic Heart Disease",
                        "Heart Failure", "Cerebrovascular Disease")),
         strata = factor(strata, levels = c("All Strata", "Female", "Male",
                                            "Age < 15", "Age 15-65", "Age > 65")),
         # group cvd and respiratory outcomes
         Group = as.factor(if_else(outcome %in% c("All Respiratory", "Asthma", "COPD"), 
                         "Respiratory", "Cardiovascular Disease")))

result_plot <- ggplot(results, aes(x = outcome, y = odds_ratio, color = Group)) +
  geom_point() + 
  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
  geom_hline(yintercept = 1, linetype=2) +
  scale_color_manual(values = c("#0ed2f7", "#3c1053")) +
  facet_wrap(~strata) +
  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
  xlab("Outcome") +
  # plot theme
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        strip.text = element_text(size = 10),
        # axis element (tilt)
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 0.7),
        #axis.title.x = element_blank(),
        #axis.ticks.x = element_blank(),
        axis.title.y = element_text(angle = 90))

result_plot

# print plot
ggsave(filename = "./r_scripts/analysis/Figs/cardiopulmonary_er_result.tiff", 
       plot = result_plot, 
       units="in", width=8, height=6, dpi=300)

```

### Result Summary: Time-Stratified Case-Crossover Design (same day of week within the season)

#### 1. Total
We found a  10 $\mu$g/m^3^ increase in PM~2.5~ was significantly associated with an 10.1% increase for asthma oucomes using GWR smoke method (OR: 1.101, 95% CI: 1.052-1.153).

#### 2. Sex
For Female and Male, there are both significant association with asthma and GWR smoke method. Male: (OR: 1.107, CI: 1.019	- 1.202) Female: (OR: 1.098,	CI:1.039 - 1.160).

#### 3. Age
Only age 15-65 has significant association with asthma and GWR smoke method. OR: 1.124, CI:	1.064	to 1.188.

*A general interpretation is as follows:*  
A 10 $\mu$g/m^3^ increase in PM~2.5~ attributed to wildfire smoke was associated with a X% increase in the risk for a health event. Where X% is the odds ratio - 1 (e.g. OR of 1.20 - 1 = 20% increase). 
Note that the odds ratio approximates the relative risk in this case due to the rare outcome assumption in that the calcuation for an odds ratio is similar to risk ratio calcuation.

We see that the exposure method is significantly associated with an increase in asthma with Geo-Weighted Regression, those are associated with COPD. I believe this is the most justifiable design approach that would average any extreme values of PM~2.5~ across the wildfire season for the vector of referent values for a particular subject.

### 2018-05-30 r_scripts/analysis/oregon_wfs_ed_results.Rmd
fixed markdown file to comment out file save.

@@
 # print plot
-ggsave(filename = "./r_scripts/analysis/Figs/cardiopulmonary_er_result.tiff", 
-       plot = result_plot, 
-       units="in", width=8, height=6, dpi=300)
+#ggsave(filename = "./r_scripts/analysis/Figs/cardiopulmonary_er_result.tiff", 
+#       plot = result_plot, 
+#       units="in", width=8, height=6, dpi=300)
 

### 2018-05-30 r_scripts/analysis/oregon_asthma_er_saba_msa_count_results.Rmd
modified and renamed part c to just import saved count by msa and month datasets rather than recreate whole thing.

--- 2018-05-18 r_scripts/analysis/oregon_part_c.Rmd
@@
 
-```{r load ER data set and zip smoke, message=FALSE, warning=FALSE}
+*Ryan: Setting Eval to False and importing timeseries rather than recreating.*
+```{r load ER data set and zip smoke, message=FALSE, warning=FALSE, eval=F}
 
@@
 
-```{r saba, message=FALSE, warning=FALSE}
+```{r saba, message=FALSE, warning=FALSE, eval=F}
 pharmacy_df <- saba_df %>%
@@
 
+```{r read_counts}
+saba_count <- read_csv("../../data/health/saba_month_counts.csv")
+ed_count <- read_csv("../../data/health/asthma_er_month_counts.csv")
+```
+
 Medford in August has higher visits.
@@
 ```{r message=FALSE, warning=FALSE}
-
-summary_table_pharmacy$group <- factor(paste0(summary_table_pharmacy$month, summary_table_pharmacy$NAME))
-X <- model.matrix(~ 0 + group, data=summary_table_pharmacy)
+saba_count
+saba_count$group <- factor(paste0(saba_count$month, saba_count$NAME))
+X <- model.matrix(~ 0 + group, data=saba_count)
 # colnames(X)
 
-fit <- glm(n_visit ~ 0 + month*NAME, offset = log(POPESTIMATE2013), family = poisson, data = summary_table_pharmacy)
+fit <- glm(n_saba ~ 0 + month*NAME, offset = log(POPESTIMATE2013), 
+           family = poisson, data = saba_count)
 # summary(fit)
 
-fit_group <- glm(n_visit ~ 0 + group, offset = log(POPESTIMATE2013), family = poisson, data = summary_table_pharmacy)
+fit_group <- glm(n_saba ~ 0 + group, offset = log(POPESTIMATE2013), 
+                 family = poisson, data = saba_count)
 # summary(fit_group)
-
-fit_no <- glm(n_visit ~ month + NAME + month:NAME - 1, offset = log(POPESTIMATE2013), family = poisson, data = summary_table_pharmacy)
-# summary(fit_no)
 
@@
 # change the numeric variable as factor
-summary_table_er$month <- as.factor(summary_table_er$month)
-summary_table_er$MSA <- as.factor(summary_table_er$MSA)
+ed_count$month <- as.factor(ed_count$month)
+ed_count$MSA <- as.factor(ed_count$MSA)
 
-summary_table_er$group <- factor(paste0(summary_table_er$month, summary_table_er$NAME))
-X <- model.matrix(~ 0 + group, data=summary_table_er)
+ed_count$group <- factor(paste0(ed_count$month, ed_count$NAME))
+X <- model.matrix(~ 0 + group, data=ed_count)
 # colnames(X)
 
-fit <- glm(n_visit ~ 0 + month*NAME, offset = log(POPESTIMATE2013), family = poisson, data = summary_table_er)
+fit <- glm(n_ed ~ 0 + month*NAME, 
+           offset = log(POPESTIMATE2013), family = poisson, 
+           data = ed_count)
 # summary(fit)
 
-fit_group <- glm(n_visit ~ 0 + group, offset = log(POPESTIMATE2013), family = poisson, data = summary_table_er)
+fit_group <- glm(n_ed ~ 0 + group, offset = log(POPESTIMATE2013), 
+                 family = poisson, data = summary_table_er)
 # summary(fit_group)

### 2018-06-02 r_scripts/analysis/asthma_place_of_service_analysis.Rmd
adding jupyter notebook to edit msa counts plots

@@
   # keeping only vars in the other casecrossover dataframes
-  select(vars_keep) %>% 
+  dplyr::select(vars_keep, ndc) %>% 
   # turn in to list
@@
       mutate(strata = "Total") %>% 
-      select(strata, n)
+      dplyr::select(strata, n)
 
@@
 # counts of observations; supressing n lower than 20
-n_tab <- counts %>% select(1:7) %>% 
+n_tab <- counts %>% dplyr::select(1:7) %>% 
   mutate_at(2:7, funs(ifelse(. < 20, NA, .)))
@@
 ```{r percent}
-perc_tab <- counts %>% select(1, 8:12)
+perc_tab <- counts %>% dplyr::select(1, 8:12)
 # print n tab
@@
       filter(term == "geo_smk10") %>% 
-      select(term, estimate, conf.low, conf.high) %>% 
+      dplyr::select(term, estimate, conf.low, conf.high) %>% 
       mutate_at(2:4, funs(round(exp(.),3))) %>% 
@@
   mutate(strata = "Total") %>% 
-  select(strata, service_place:conf.high)
+  dplyr::select(strata, service_place:conf.high)
 
@@
                     filter(term == "geo_smk10") %>% 
-                    select(term, estimate, conf.low, conf.high) %>% 
+                    dplyr::select(term, estimate, conf.low, conf.high) %>% 
                     mutate_at(2:4, funs(round(exp(.),3))) %>% 
@@
                     mutate(service_place = as.character(service_place))})) %>% 
-      select(-data, -model) %>% 
+      dplyr::select(-data, -model) %>% 
       unnest(est_df)
@@
   mutate(strata = ifelse(gender == "M", "Male", "Female")) %>% 
-  select(strata, service_place:conf.high)
+  dplyr::select(strata, service_place:conf.high)
 
@@
   rename(strata = age_cat) %>% 
-  select(strata, service_place:conf.high) %>% 
+  dplyr::select(strata, service_place:conf.high) %>%
    filter(n_events > 20) %>%
@@
                        "Urgent Care", "Ambulance", "Emergency Room", 
-                       "Inpatient Hospital")))
+                       "Inpatient Hospital"))) %>% 
+  # remove ambulance since it has so few observations
+  filter(service_place %in% c("SABA Fill", "Office", "Urgent Care",
+                              "Outpatient Hospital", "Inpatient Hospital"))
 ```
@@
 plot <- ggplot(results, aes(x = strata, y = odds_ratio)) +
-  geom_point(color = "blue") + 
-  geom_errorbar(aes(ymin=lower95, ymax=upper95), color = "blue",width = 0.2) +
+  geom_point(color = "#7f00ff") + 
+  geom_errorbar(aes(ymin=lower95, ymax=upper95), color = "#7f00ff",width = 0.2) +
   geom_hline(yintercept = 1, linetype=2) +
-  facet_wrap(~service_place, scales = "free_y") +
+  facet_wrap(~service_place) +
   ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
@@
     # axis element (tilt)
-    axis.text.x = element_text(angle = 45, hjust =1),
+    axis.text.x = element_text(angle = 45, vjust =0.9),
     #axis.title.x = element_blank(),
@@
 plot
+
+# save plot
+#ggsave("./Figs/asthma_pos_results_plot.tiff", plot, 
+#       units="in", width=8, height=6, dpi=300)
 ```
@@
                                         group = service_place)) +
-  geom_point(color = "blue") + 
-  geom_errorbar(aes(ymin=lower95, ymax=upper95), color = "blue",width = 0.2) +
+  geom_point(color = "#7f00ff") + 
+  geom_errorbar(aes(ymin=lower95, ymax=upper95), color = "#7f00ff",width = 0.2) +
   geom_hline(yintercept = 1, linetype=2) +
@@
     # axis element (tilt)
-    axis.text.x = element_text(angle = 45, hjust =1),
+    axis.text.x = element_text(angle = 45, hjust = 0.9, vjust = 0.7),
     #axis.title.x = element_blank(),
@@
 total_plot
+
+#ggsave("./Figs/total_pos_results_plot.tiff", total_plot, 
+#       units="in", width=6, height=4, dpi=300)
 ```
@@
   mutate(strata = "Total") %>% 
-  select(strata, service_place:conf.high) %>% 
+  dplyr::select(strata, service_place:conf.high) %>% 
   rename(odds_ratio = estimate, lower95 = conf.low, upper95 = conf.high) %>% 
@@
 
-## Non-Linear Distributed Lag
-
-Preparing PM~2.5~ values, assigning an MSA by ZIP code.
-
-```{r pm_msa}
-# making a zipcode and msa key
-zip_msa <- asthma_pos_list[[1]] %>% 
-  select(MSA, metroarea, ZIP) %>% 
-  unique()
-# ungrouping pm and adding in MSA for color-coding
-pm_msa <- ungroup(pm) %>% 
-  left_join(zip_msa, by = "ZIP") %>% 
-  #filter out missing metroareas
-  filter(!is.na(metroarea))
-```
-
-Checking distribution of PM~2.5~ by each ZIP code in each MSA I was able to code.
-There is a heavy right tail for Medford and ZIPs not in an MSA (which probably capture some rural areas around the southern part of the state.
-If you let the scales vary, the distribution of PM~2.5~ values suggests to me that Medford was impacted by smoke, as evidenced by the heavy right tail.
-Other MSAs have distributions that are more like standard distributions for ambient air pollution not impacted by smoke or other extreme events.
-
-```{r pm_dist}
-# distribution of pm values is log-normal
-ggplot(pm_msa, aes(x=geo_wt_pm, group = ZIP, color=metroarea)) +
-  geom_density() +
-  facet_wrap(~metroarea, scales = "free") +
-  theme_minimal()
-```
-
-Looking at time-series of PM~2.5~ values by MSA.I plotted the mean seasonal estimate for each ZIP in each MSA as a red line for each MSA.
-
-```{r pm_ts}
-# estimating seasonal background
-bg_pm <- pm_msa %>% 
-  group_by(date, metroarea) %>% 
-  summarise(mean_pm = mean(background_pm), med_pm = median(background_pm))
-
-# plot
-ggplot(pm_msa, aes(x=date, y=geo_wt_pm, group=ZIP, color = metroarea)) +
-  geom_point() +
-  geom_line(data=bg_pm, aes(x=date, y=mean_pm, group=metroarea), color = "red") +
-  facet_wrap(~metroarea) +
-  theme_minimal()
-
-```
-
-After looking at these plots, I find it hard to overlook the simple, but limited approach of just subtracting off the difference of the PM~2.5~ estimate from the background estimate.
-
-I'm going to revist the non-linear distributed lag approach.
-I think we can then estimate a couple different lagged and cumulative effect over a 0 to 5 day prior event period.
-We could then estimate the function at 10 ug/m^3^ ("normal" air pollution), 20 ug/m^3^ (most likely light smoke), and 30 ug/m^3^ (smoke), and maybe 50 ug/m^3^ (heavy smoke).
-
-Trying out the distributed lag non-linear package again. I'm going to limit to SABA time-stratified case-crossover dataframe in Medford to simplify.
-
-```{r saba_medford}
-# subset medfore saba
-saba_medford <- asthma_pos_list$saba %>% 
-  filter(metroarea == "Medford") %>% 
-  # remove missing values for lag
-  filter(!is.na(geo_pm10_lag5))
-```
-
-Set up. I'm going to try a 3 knot ns spline for the lag funciton and a 2 knot spline for exposure response to PM~2.5~.
-
-```{r}
-# extract geo_10 lagged matrix in same order as dataframe
-pm_lag_mat <- saba_medford %>% 
-  select(contains("geo_pm10")) %>% 
-  as.matrix()
-# view matrix
-head(pm_lag_mat)
-
-# crossbasis
-cb_pm <- crossbasis(pm_lag_mat, lag = 5, argvar = list("ns", df = 2),
-                    arglag = list(fun="ns", knots=c(2,4), intercept = T))
-
-# clogit model
-cb_mod <- clogit(outcome ~ cb_pm + strata(identifier), data = saba_medford)
-
-# predict
-cb_pred <- crosspred(basis = cb_pm, model = cb_mod, cen = 6, at = seq(5,30,5))
-
-```
-
-Cumulative effect at different PM~2.5~ levels.
-
-```{r}
-# predictor variable
-pm_levels <- cb_pred$predvar
-# cumulative rel risk/odds ratio from clogit mod
-odds_ratio <- cb_pred$allRRfit
-odds_ratio
-
-plot(pm_levels, odds_ratio)
-```
-
 ## Next Steps/Thoughts

### 2018-06-04 r_scripts/analysis/oregon_asthma_er_saba_msa_count_results.Rmd
Updated gitignore to include jupyter notebook checkpoings.

@@
 
-```{r message=FALSE, warning=FALSE}
+```{r message=FALSE, warning=FALSE, eval =F}
 
@@
 
-```{r}
+```{r eval=F}
 ed_count <- summary_table_er %>% 

### 2018-06-04 r_scripts/analysis/asthma_place_of_service_analysis.Rmd
Added scripts for exposure plots.

@@
   mutate(strata = ifelse(gender == "M", "Male", "Female")) %>% 
-  dplyr::select(strata, service_place:conf.high)
+  dplyr::select(strata, service_place:conf.high) %>% 
+  filter(n_events > 30)
 
@@
   dplyr::select(strata, service_place:conf.high) %>%
-   filter(n_events > 20) %>%
+   filter(n_events > 30) %>%
    filter(conf.high < 5)
@@
          service_place = case_when(service_place=="Ambulance-Land"~"Ambulance",
-          service_place == "Emergency Room – Hospital" ~ "Emergency Room",
+          service_place == "Emergency Room – Hospital" ~ "Emergency Department",
           service_place == "Urgent Care Facility" ~ "Urgent Care",
@@
             levels = c("All Strata", "Female", "Male", "Age < 15", 
-                       "Age 15-65", "Age > 65")),
+                       "Age 15-65", "Age > 65"))) %>% 
     # set order of service place
-          service_place = factor(service_place, 
-            levels = c("SABA Fill", "Office", "Outpatient Hospital", 
-                       "Urgent Care", "Ambulance", "Emergency Room", 
-                       "Inpatient Hospital"))) %>% 
+          # service_place = factor(service_place, 
+          #   levels = c("SABA Fill", "Office", "Outpatient Hospital", 
+          #              "Urgent Care", "Ambulance", "Emergency Room", 
+          #              "Inpatient Hospital"))) %>% 
   # remove ambulance since it has so few observations
-  filter(service_place %in% c("SABA Fill", "Office", "Urgent Care",
-                              "Outpatient Hospital", "Inpatient Hospital"))
+  filter(service_place %in% c("SABA Fill", "Office", "Urgent Care", "Ambulance",
+                              "Emergency Department", "Outpatient Hospital", 
+                              "Inpatient Hospital"))
 ```
@@
 plot <- ggplot(results, aes(x = strata, y = odds_ratio)) +
-  geom_point(color = "#7f00ff") + 
-  geom_errorbar(aes(ymin=lower95, ymax=upper95), color = "#7f00ff",width = 0.2) +
+  geom_point(color = "#3c1053") + 
+  geom_errorbar(aes(ymin=lower95, ymax=upper95), color = "#3c1053",width = 0.2) +
   geom_hline(yintercept = 1, linetype=2) +
-  facet_wrap(~service_place) +
-  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+  facet_wrap(~service_place, scales = "free_y", ncol = 2) +
+  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in WFS PM"[2.5]))) +
   xlab("Stratum") +
@@
     # axis element (tilt)
-    axis.text.x = element_text(angle = 45, vjust =0.9),
+    axis.text.x = element_text(angle = 45, hjust = 0.9, vjust = 0.7),
     #axis.title.x = element_blank(),

### 2018-06-04 r_scripts/analysis/wfs_exposure_figures.Rmd
Added scripts for exposure plots.

--- 2018-05-16 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
@@
 ---
-title: "Douglas-complex fire smoke and cardiopulmonary morbidity"
+title: "Douglas-Complex Fire PM2.5"
 author: "Ryan Gan"
-date: "11/3/2017"
+date: "6/4/2018"
 output: html_document
@@
 
-Data came from the Oregon All Payer All Claims Database (APAC) in the year 2013.
-The APAC records the health care billing data for Oregon's insured populations.
-APAC include individual billing records for both diagnoses codes (International Classification of Diseases, Clinical Modification (ICD-9-CM) diagnoses codes) and pharmacy codes (National Drug Codes (NDC)).
-
-Our previous research that found an association with wildfire smoke and respiratory outcomes in Washington state in 2012 wildfire season using a novel estimate of smoke concentration, geographically weighted ridge regression (GWR) guided the methodological approaches used in this project.
-As Oregon contains pharmacy records, we evaluate the association between smoke and respiratory rescue medications (beta 2 agonists) (abbreviate to SABA).
-
-*Research question*
-We evaluated the association between smoke concentrations using the GWR method and cardiopulmonary morbidity, including ED/urgent care visits and SABA fills in Oregon state during the 2013 wildfire season.
-
-This markdown document contains the code and results that were used to address this research question.
-
 Packages used: tidyverse, survival.
@@
 library(survival) # conditional logistic regression
+library(geofacet)
 
@@
     stringr::str_to_title(names(county_centroids)), start=8L))
+
+# import MSA location
+msa_path <- paste0("../../data/pm/msa_coordinate.csv")
+msa_df <- read_csv(msa_path)
+
+msa_location <- msa_df %>%
+  filter(grepl('OR', msa_df$MSA)) %>%
+  filter(MSA != "Albany, OR" &
+         MSA != "Grants Pass, OR")
+
+# create msa label data frame
+msa_text <- msa_location %>%
+  select(MSA, Longitude, Latitude)
+
+msa_text$MSA <- gsub(",.*", "", msa_text$MSA)
+
+msa_text <- msa_text %>% 
+  mutate(metro_area = case_when(MSA == 'Portland-Vancouver-Hillsboro' ~ 'Portland',
+                                MSA == 'Bend-Redmond' ~ 'Bend', 
+                                T ~ MSA))
 ```
@@
   scale_fill_gradient2(expression("Smoke Days > 10 µg/m"^3),
-    low = "white", mid= "#00dbde", high="#6a3093", midpoint = 15,
+                       low = "white", mid= "#00dbde", high="#6a3093", 
+                       midpoint = 15,
     guide = guide_colorbar(direction="horizontal", title.position = "top",
@@
   geom_point(data=fire_locations, 
-    aes(x=lon, y=lat, size = area,
-    shape = "Fire Locations"), shape = 17, color = "#ff0084", alpha = 0.7,
-    show.legend = FALSE, inherit.aes = FALSE) +
+             aes(x=lon, y=lat, size = sqrt(area/2)), 
+             color = "red", shape = 17, alpha = 0.7, 
+             show.legend = F, inherit.aes = FALSE) +
   # add county text 
   geom_text(data = county_text, 
-    aes(x=long, y=lat, label = county, group=NULL), 
-    colour = "#33001b", size = 2.5) +
+            aes(x=long, y=lat, label = county, group="county"), 
+            color = "#33001b", size = 2.5, inherit.aes = F) +
+  # add metropolitan
+  geom_point(data = msa_location, 
+    aes(x=Longitude, y=Latitude), 
+    size = 1, color = '#ff0084', shape = 19, alpha = 0.7,
+    show.legend = F, inherit.aes = FALSE) +
+  # add metro text
+  geom_text(data = msa_text, 
+    aes(x=Longitude + c(0.15,0.1,0.1,-0.05,0.2,0.1), 
+        y=Latitude + c(0.2,0.1,0.1,-0.1,0.1,0.1), 
+        label = metro_area, group="MSA"), 
+    color = "#ff0084", size = 3) +
   xlab("Longitude") +
@@
 smoke_map
+
 # save figure
-ggsave(filename = "./smoke_map.png", plot = smoke_map)
+ggsave(filename = "./r_scripts/analysis/Figs/smoke_map.tiff", plot = smoke_map,
+       units="in", width=6, height=5, dpi=300)
 ```
@@
 geofacet::facet_geo(~county, grid = or_grid) +
-  ylab(expression("Smoke PM2.5 µg/m"^3)) +
+ylab(expression("Smoke PM2.5 ?g/m"^3)) +
 xlab("Month in 2013") +
@@
 plot
+# save figure
+ggsave(filename = "./county_smoke_timeseries.png", plot = plot)
 ```
-
-Elevated levels of smoke in late July to early August in Curry, Josephine, Jackson, and Klamath counties.
-
-Jeff and Bonne suggest only presenting time-series of 4 counties affected, and send this figure to supplement.
-
-##3. Descriptive table of number of health outcomes
-
-Importing time-stratified case-crossover dataframes.
-Like the Washington study, we used the time-stratified case-crossover where admissions/fills are identified by date, and then referent observations are created for each observation (assumed to be a unique person) from May to October 2013.
-We argue that this is a good referent period that allows for enough referent periods, without introducing seasonal variation in exposure/disease.
-We also subtract off background smoke.
-
-```{r health data import, message=F, warning=F}
-# path for health dataframe
-health_path <- "../../data/health/"
-# importing time-stratified case-crossover dataframe
-casecross_files <- list.files(path=health_path, pattern="*casecross*")
-
-# reorder file names in way I'd like it to be analyzed
-# also ditching broken arm
-casecross_order <- casecross_files[c(13,12,3,6,11,1,7,2,5,8,9,10)]
-
-# extract the name of the dataframe after study type
-file_names <- stringr::str_extract(casecross_order, 
-  pattern="(?<=casecross_)\\w+")
-
-# read in all the files in to a list
-outcome_df_list <- casecross_order %>%  
-  purrr::map(~read_csv(paste0(health_path, .)))
-
-# name dataframes
-names(outcome_df_list) <- file_names
-
-# name list elements in order I'd like to present in document
-outcome_names <- c("SABA", "Respiratory", "Asthma", "COPD", "Pneumonia", 
-  "Acute Bronchitis", "Cardiovascular Disease", "Arrhythmia", 
-  "Cerebrovascular Disease", "Heart Failure", "Ischemic Heart Disease", 
-  "Myocardial Infarction")
-```
-
-CVD warning message on import; I think it's caused by varialbes with mixed character/integer (common in claims data).
-They don't have any bearing on analyses (but I find them annoying and plan to fix).
-
-*Also warnings on mixed data types of NULL and numeric; another reason to revise and create timestratified casecross over functions*.
-
-Descriptive counts of the number of ED/urgent care or pharmacy fills observed between May to October, 2013 in the state of Oregon, and who were assigned a smoke value.
-
-```{r descriptive table function}
-# creating a row calculation function
-table1_fun <- function(data_list){
-  # output of total n
-  total_n <- data_list %>% 
-    filter(outcome == 1) %>%
-    # adding in a filter to remove NAs
-    filter(!is.na(geo_smk_pm_zip) & gender != "U") %>% 
-    summarise(total_n = n())
-
-  # output row of summary numbers
-  age_cat_n <- data_list %>% 
-    filter(outcome == 1) %>%
-    group_by(age_ind) %>% 
-    summarise(n = n()) %>%
-    # this will fix the 0 categories of certain CVD outcomes in children
-    add_row(age_ind = 0, n = 0) %>% 
-    arrange(age_ind) %>% 
-    group_by(age_ind) %>% 
-    summarise(n = sum(n)) %>% 
-    mutate(age_cat = ifelse(age_ind == 0, "age_under15",
-                    ifelse(age_ind == 1, "age_15to65", "age_over65")),
-           # preserve factor order
-           age_cat = parse_factor(age_cat, levels = age_cat)) %>% 
-    select(-age_ind) %>% 
-    spread(age_cat, n)
-  # sex n
-  sex_n <- data_list %>% 
-    filter(outcome == 1) %>% 
-    group_by(gender) %>% 
-    summarise(n = n()) %>% 
-    filter(gender != "U") %>% 
-    spread(gender, n)
-  # bind all rows together
-  outcome_n_row <- bind_cols(total_n, age_cat_n, sex_n) %>% 
-    mutate(perc_15 = round(age_under15/total_n,2), 
-      perc_15to65 = round(age_15to65/total_n,2),
-      perc_65 = round(age_over65/total_n,2),
-      perc_F = round(F/total_n,2),
-      perc_M = round(M/total_n,2)) %>% 
-    select(total_n, age_under15, perc_15, age_15to65, perc_15to65, 
-           age_over65, perc_65, F, perc_F, M, perc_M)
-return(outcome_n_row)
-} # end function
-```
-
-Table 1 total, and stratified by age categories and sex.
-
-```{r descritive table}
-# outcome names
-outcome <- outcome_names %>%
-  as_tibble() %>%
-  rename(outcome = value)
-
-# apply custom function to generate descriptive table
-table_vals <- outcome_df_list %>%
-  # map to dataframe
-  map_dfr(~table1_fun(.))
-
-# descriptive table, bind outcome names and values
-descriptive_table <- bind_cols(outcome, table_vals)
-
-# output table
-knitr::kable(descriptive_table,
-  caption = paste0("Descriptive charactersitics SABA fills",
-                   " and of ED/urgent care visits"))
-```
-
-##4. Same-day association results
-For this project, we use the time-stratified case-crossover study design in the same way we did for the 2012 Wenatchee-Complex fire project.
-Briefly, we identify admissions dates for ED or urgent care visits, and pharmacy fill dates for SABA, using billing records.
-
-In the Wenatchee-Complex fire project, we looked at the same-day relationship between smoke exposure and health outcomes.
-For the Douglas-complex fire, we will evaluate a 0 to 7 day distributed lag association, first comparing the overall cumulative effect over these 7 days, and then get in to the lagged effects.
-
-In this section of the document, we look at the association between GWR smoke PM~2.5~ concentration on the same day of admission or fill using conditional logistic regression.
-All models are adjusted for temperature on the same day.
-
-```{r same day association fuction}
-same_day_fun <- function(data){
-  # limit to complete cases
-  complete_data <- data %>% 
-    # filter to complete case for smoke
-    filter(!is.na(geo_smk_pm_zip)) %>% 
-    # create 10 unit smk variable 
-    mutate(geo_smk10 = geo_smk_pm_zip/10)
-
-    # fit conditional logistic regression model ----
-  mod <- clogit(outcome ~ geo_smk10 + wrf_temp_zip + strata(personkey),
-    data = complete_data)
-
-  # odds ratio
-  estimates <- round(exp(summary(mod)$coefficient[1,1]), 3) 
-  # 95% lower bound
-  lower95 <- round(exp((summary(mod)$coefficient[1,1]) -
-                                    1.96*(summary(mod)$coefficient[1,3])), 3)
-  # 95% upper bound
-  upper95 <- round(exp((summary(mod)$coefficient[1,1]) +
-                                    1.96*(summary(mod)$coefficient[1,3])), 3)
-
-  return_estimate <- data_frame(estimates, lower95, upper95) 
-  return(return_estimate) 
-} 
-```
-
-Table of odds ratios (OR) and 95% confidence intervals (95%CI) that represent the same-day association between an increase in 10 ug/m^3 increase in smoke PM~2.5~ and the odds (can be interpreted as risk due to low prevalence) of the outcome.
-
-```{r same day association table}
-sameday_results <- outcome_df_list %>% 
-  map_dfr(~same_day_fun(.)) %>% 
-  bind_cols(., data_frame(outcome_names)) %>% 
-  select(outcome_names, estimates, lower95, upper95) %>% 
-  # preserve order of names
-  mutate(outcome_names = parse_factor(outcome_names, levels=outcome_names)) %>%
-  rename(odds_ratio = estimates)
-
-knitr::kable(sameday_results, 
-  caption = "Same day odds ratio and 95% CI of pharmacy/ED/urgent care visits")
-```
-
-Plot of the ORs and 95%CIs in the table. I've grouped by cardiovascular and respiratory classes of outcomes.
-
-```{r same day association graph}
-class <- as_tibble(c(rep("Respiratory", times = 6), 
-  rep("Cardiovascular", times = 6))) %>% 
-  rename(class = value)
-# add a class indicator to table
-sameday_results_w_class <- sameday_results %>% bind_cols(class)
-
-# ggplot ----
-plot <- ggplot(data=sameday_results_w_class, 
-  aes(x=outcome_names, y = odds_ratio, group = class, color = class)) +
-  geom_point() +
-  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
-  scale_color_manual(name = "Disease Group", values = c("#4ac29a", "#7303c0"), 
-    guide = guide_legend(direction="horizontal", title.position = "top",
-                         title.hjust = 0.5)) +
-  geom_hline(yintercept = 1, linetype = 2, colour = "#B40F20") +
-  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    legend.position = "bottom",
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    axis.text.x = element_text(angle=45, hjust = 1)) +
-  ylab(expression("Same day odds ratio: 10 ug/m^3 increase smoke PM"[2.5])) +
-  xlab(NULL) # removed x axis name, not sure i want to keep it that way
-
-plot
-```
-
-##5. Distributed lag cumulative effect results
-
-*Note: 11/7/17: Jingyang will provide a csv of the best-fit knots for the distributed lag splines based on AIC.*
-
-We set up basis functions using natural splines for smoke exposure 0 to 7 days prior to admission/fill.We've applied 3 DF to smoke splines as I think it's flexible enough to allow varying shapes and fits well, but we are working on including best-fit by AIC in to models.
-Basis function also included for WRF temp; it looks like 2 DF is best.
-
-```{r distributed lag function}
-# custom distributed lag function 
-distributed_lag <- function(data, pm_df, temp_df, cumulative=T){
-  # limit to complete cases
-  complete_data <- data %>% 
-    # filter to complete case for smoke
-    filter(!is.na(geo_smk_pm_lag7_zip))
-
-  # output matrix of gwr values
-  pm_matrix <- complete_data %>% 
-    select(contains("geo_smk_pm")) %>%
-    # remove anything with m in it; will go back in original code and not create 
-    # this to begin with
-    select(-contains("_m")) %>% 
-    # divide exposure values by 10 units to interpret on 10 ug/m^3 scale
-    mutate_all(funs(./10)) %>% 
-    # convert to matrix
-    as.matrix()
-
-  # output temp matrix
-  temp_matrix <- complete_data %>% 
-    select(contains("wrf_temp")) %>%
-    # remove anything with m in it; will go back in original code and not create 
-    # this to begin with
-    select(-contains("_m")) %>% 
-    # convert to matrix
-    as.matrix()
-
-  # calculation of basis ----
-  # define basis using natural spline function from "splines"" package
-  pm_b <- splines::ns(0:7, df=pm_df, intercept=T)
-  # create pm basis
-  pm_basis <- pm_matrix %*% pm_b
-  # create temp basis
-  temp_b <- splines::ns(0:7, df=temp_df, intercept=T)
-  temp_basis <- temp_matrix %*% temp_b
-
-  # fit model with basis ----
-  mod <- clogit(outcome ~ pm_basis + temp_basis + strata(personkey),
-    data = complete_data)
-
-  # calculate estimates ----
-  # output pm basis parameters
-  dl_parms <- broom::tidy(mod) %>% 
-    filter(stringr::str_detect(term, "pm")) %>% 
-    select(estimate) %>% 
-    as_vector()
-  # estimate distributed lag values for each day
-  dl_estimates <- data.frame(estimate = pm_b %*% dl_parms)
-  # covariance matrix for knots
-  cov_matrix <- as.matrix(vcov(mod))[1:3,1:3]
-
-  # estimate variance of spline
-  dl_var <- pm_b %*% cov_matrix %*% t(pm_b)
-  # estimate standard error
-  dl_estimates$stderr <- sqrt(diag(dl_var))
-
-  # calculate lower and upper bound
-  dl_estimates$lower95 <- dl_estimates$estimate-(dl_estimates$stderr*1.96)
-  dl_estimates$upper95 <- dl_estimates$estimate+(dl_estimates$stderr*1.96)
-
-    if(cumulative==T) {
-      type <- "cumulative"  
-      # cumulative outcome and 95CI
-      estimate <- sum(dl_estimates$estimate)
-      # stderr cumulative effect
-      estimate_se <- sqrt(sum(dl_var))
-      # cumulative 95CI
-      lower95 <- estimate-(estimate_se*1.96)
-      upper95 <- estimate+(estimate_se*1.96)
-      # return dataframe
-      return_estimate <- data_frame(type, estimate, lower95, upper95) %>% 
-        mutate_if(is.numeric, exp)
-      return(return_estimate)
-    }
-    else { # if cumulative is not true, or false, return distributed lag est
-      # return dataframe
-      return_estimate <- dl_estimates %>% 
-        mutate_if(is.numeric, exp) %>% 
-        mutate(type = "lag", 
-               time = (as.numeric(rownames(.))-1)/-1) %>% 
-        select(type, time, estimate, lower95, upper95)
-       # return estimate  
-      return(return_estimate)
-      }
-      # end of if
-} 
-```
-
-Using the custom function to estimate cumulative effects for the list of dataframes. It will work on an individual dataframe, but I built it to use with the purrr map functions.
-
-```{r cumulative association}
-class <- as_tibble(c(rep("Respiratory", times = 6), 
-  rep("Cardiovascular", times = 6))) %>% 
-  rename(class = value)
-# use custom function to create a dataframe of cumulative values
-# using purrr::map_dfr function
-# purrr::map_dfr will work if it's just the cumulative, but won't work for lagged
-cumulative_results <- outcome_df_list %>% 
-  map_dfr(~distributed_lag(., pm_df=3, temp_df=2, cumulative=T)) %>% 
-  bind_cols(., data_frame(outcome_names)) %>% 
-  select(outcome_names, type, estimate, lower95, upper95) %>% 
-  # preserve order of names
-  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))
-
-# add a class indicator to table
-cumulative_results_w_class <- cumulative_results %>% bind_cols(class)
-
-# ggplot ----
-plot <- ggplot(cumulative_results_w_class, 
-    aes(x=outcome_names, y = estimate, group=class, color=class)) +
-  geom_point() +
-  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
-  scale_color_manual(name = "Disease Group", values = c("#4ac29a", "#7303c0"), 
-    guide = guide_legend(direction="horizontal", title.position = "top",
-                         title.hjust = 0.5)) +
-  geom_hline(yintercept = 1, linetype = 2, colour = "#FF0000") +
-  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    legend.position = "bottom",
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    axis.text.x = element_text(angle=45, hjust = 1)) +
-  ylab(expression("Cumulative odds ratio: 10 ug/m^3 increase smoke PM"[2.5])) +
-  xlab(NULL)
-
-plot
-```
-
-Also considering reducing the lag time evaluated since I think we are getting in to variance issues a week out with this dataset (as evidence, large error around asthma outcome relative to other studies).
-Alternatively, we could add the previous day 0, 0+1, 0+1+2,...
-and so on.
-Discuss with group.
-
-##6. Distributed lag results
-
-That same function can be used to generate detailed estimates of the association over time. In this case, we'll look at the same day to seven days prior the event.
-
-```{r distributed lag}
-# this approach will work that creates a list of dataframes with map then
-# binds the lists together in one dataframe
-dl_results <- outcome_df_list %>% 
-  purrr::map(~distributed_lag(., pm_df=3,temp_df=2,cumulative=F)) %>% 
-  plyr::rbind.fill() %>% 
-  bind_cols(., data_frame(rep(outcome_names, each=8))) %>% 
-  rename(outcome_names = `rep(outcome_names, each = 8)`) %>% 
-  select(outcome_names, type, time, estimate, lower95, upper95) %>% 
-  # preserve order of names
-  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))
-
-# results will come in a format to make small multiples I think
-#head(dl_results)
-
-# small multiples DL plot ----
-dl_plot <- ggplot(dl_results, aes(x=time, y=estimate)) +
-  geom_line(colour = "#046C9A", size = 1) +
-  geom_ribbon(aes(ymin = lower95, ymax = upper95), 
-              fill = "#ABDDDE", alpha = 0.5) + 
-  scale_y_continuous(limits = c(0.8, 1.2)) +
-  scale_x_continuous(breaks = c(seq(-7, 0, by=1))) +
-  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
-  facet_wrap(~outcome_names) +
-  ylab(expression("Odd ratio for a 10 ug/m^3 increase smoke PM"[2.5])) +
-  xlab("Lagged Days") +
-  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    strip.text = element_text(size = 10),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    strip.background = element_rect(colour=NA, fill=NA))
-
-dl_plot
-```
-
-DL plot shows an effect with asthma even though the cumulative effect size is similar in magnitude to what we've seen before, but the 95% CIs overlap 1.
-I think this is a power issue trying to look at lagged effects for more than a week.
-I've applied a general 3 degrees of freedom knot for PM~2.5~ and 2 degrees of freedom knot for temperature.
-I think temperature is okay to do this, but I'm still figuring out a way to build a best-fit knot in to the function.
-
-Worth discussing with the group.
-
-##7. Stratified same-day results
-
-*Note to Jingyang: Nov 16*
-*I realized there is a better way to calculate stratified results than modifying the code. See below for example. There is a way to do it in one chunk rather than 2, but I was running in to errors.*
-
-### 7.a Sex
-
-Presenting same-day stratified results because the computations are easier.
-I'll use it to decide if we should present cumulative effects.
-We can use map and filter to subset to group of interest and apply our same-day function.
-We then bind the two strata together.
-
-```{r same day association stratified by sex}
-# there is probably a better way to do this, but this will work.
-sameday_results_f <- outcome_df_list %>% 
-  # using gender because I forget which sex_ind is which
-  # filter to females
-  purrr::map(~filter(., gender == "F")) %>%  
-  map_dfr(~same_day_fun(.)) %>% 
-  bind_cols(., data_frame(outcome_names)) %>% 
-  mutate(sex = as.factor("Female")) %>% 
-  select(outcome_names, sex, estimates, lower95, upper95) %>% 
-  # preserve order of names
-  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))
-
-sameday_results_m <- outcome_df_list %>% 
-  # filter to males
-  purrr::map(~filter(., gender == "M")) %>%  
-  map_dfr(~same_day_fun(.)) %>% 
-  bind_cols(., data_frame(outcome_names)) %>% 
-  mutate(sex = as.factor("Male")) %>% 
-  select(outcome_names, sex, estimates, lower95, upper95) %>% 
-  # preserve order of names
-  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))
-
-# bind sex dataframes together
-sameday_results <- rbind(sameday_results_f, sameday_results_m) %>% 
-  rename(odds_ratio = estimates)
-
-# print table
-knitr::kable(sameday_results, 
-  caption = "Same day odds ratio and 95% CI of pharmacy/ED/urgent care visits stratified by sex.")
-```
-
-Plot of sex-stratified results.
-
-```{r sex strata plot}
-plot <- ggplot(data=sameday_results, aes(x=sex, y = odds_ratio, color=sex)) +
-  scale_color_manual(name = "Sex", values = c("#43c6ac", "#191654"),
-    guide = guide_legend(direction="horizontal", title.position = "top",
-                         title.hjust = 0.5)) +
-  geom_point() +
-  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
-  facet_wrap(~outcome_names) +
-  geom_hline(yintercept = 1, linetype = 2, colour = "#FF0000") +
-  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    legend.position = "bottom",
-    # strip element
-    strip.background = element_rect(colour=NA, fill=NA),
-    panel.border = element_rect(fill = NA, color = "black")) +
-  ylab(expression("Odds ratio: 10 ug/m^3 increase smoke PM"[2.5])) +
-  xlab(NULL)
-
-plot  
-```
-
-No obvious sex differences in admission/fill due to increasing wildfire smoke exposure.
-
-### 7.b Age
-
-Stratified by age category.
-
-```{r same day association stratified by age category}
-# subset outcomes evaluated for under 15
-under15_outcome <- outcome_names[1:7]
-
-# subset dataframes that won't have convergence issues
-sameday_results_under15 <- outcome_df_list[1:7] %>% 
-  # filter to children
-  purrr::map(~filter(., age_ind == 0)) %>%  
-  map_dfr(~same_day_fun(.)) %>% 
-  bind_cols(., data_frame(under15_outcome)) %>% 
-  rename(outcome_names = under15_outcome) %>% 
-  mutate(age_cat = as.factor("Under 15")) %>% 
-  select(outcome_names, age_cat, estimates, lower95, upper95) %>% 
-  # preserve order of names
-  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))
-
-# age category 1
-sameday_results_15to65 <- outcome_df_list %>% 
-  # filter to males
-  purrr::map(~filter(., age_ind == "1")) %>%  
-  map_dfr(~same_day_fun(.)) %>% 
-  bind_cols(., data_frame(outcome_names)) %>% 
-  mutate(age_cat = as.factor("15 to 65")) %>% 
-  select(outcome_names, age_cat, estimates, lower95, upper95) %>% 
-  # preserve order of names
-  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))
-
-# age category 2
-sameday_results_over65 <- outcome_df_list %>% 
-  purrr::map(~filter(., age_ind == "2")) %>%  
-  map_dfr(~same_day_fun(.)) %>% 
-  bind_cols(., data_frame(outcome_names)) %>% 
-  mutate(age_cat = as.factor("Over 65")) %>% 
-  select(outcome_names, age_cat, estimates, lower95, upper95) %>% 
-  # preserve order of names
-  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))
-
-# bind sex dataframes together
-sameday_results <- rbind(sameday_results_under15, sameday_results_15to65,
-                         sameday_results_over65) %>% 
-  rename(odds_ratio = estimates)
-
-# print table
-knitr::kable(sameday_results, 
-  caption = "Same day odds ratio and 95% CI of pharmacy/ED/urgent care visits stratified by age category.")
-```
-
-Stratified by age plot.
-
-```{r age strata plot}
-plot <- ggplot(data=sameday_results, 
-               aes(x=age_cat, y = odds_ratio, color=age_cat)) +
-  scale_color_manual(name = "Age Category", 
-    values = c("#0082c8", "#4ac29a", "#6a3093"),
-    guide = guide_legend(direction="horizontal", title.position = "top",
-                         title.hjust = 0.5)) +
-  geom_point() +
-  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
-  facet_wrap(~outcome_names, scales = "free_y") +
-  geom_hline(yintercept = 1, linetype = 2, colour = "#FF0000") +
-  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    legend.position = "bottom",
-    # strip element
-    strip.background = element_rect(colour=NA, fill=NA),
-    panel.border = element_rect(fill = NA, color = "black")) +
-  ylab(expression("Odds ratio: 10 ug/m^3 increase smoke PM"[2.5])) +
-  xlab(NULL)
-
-plot  
-```
-
-Under 15 unstable for CVD; consider removing.
-Heart failure is interesting for the over 65 group.
-That may agree with what Anna Rappold has seen.
-May be some difference in SABA fills by age category, but probably not; rough evaluation says no as the CIs overlap.
-
-##8. SABA same-day results by underlying respiratory diagnosis
-
-```{r saba underlying dx}
-# read in the index 
-saba_index <- read_csv("../../data/health/oregon_saba_index.csv") %>% 
-    mutate(underlying_dx = as.factor(ifelse(index == "copd",
-    stringr::str_to_upper(index),
-    stringr::str_to_title(index)))) %>% 
-    select(-index)
-
-# index vector
-underlying_dx <- unique(as.factor(saba_index$underlying_dx)) %>% 
-  data_frame() %>% 
-  rename(underlying_dx = ".") %>% 
-  mutate(underlying_dx = ) %>% 
-  # need to arrange since slice is going alphabetical
-  arrange(underlying_dx)
-
-# join index to saba
-# extract saba dataframe from list
-saba <- outcome_df_list[[1]] %>% 
-  left_join(saba_index, by = "personkey")
-
-# run function on each group
-sameday_results <- saba %>% 
-  split(.$underlying_dx) %>% 
-  map_dfr(~same_day_fun(.)) %>% 
-  bind_cols(., underlying_dx) %>% 
-  select(4,1:3)
-
-# table
-knitr::kable(sameday_results, 
-  caption = "Same day odds ratio and 95% CI for a SABA fill, stratified by underlying dx.")
-```
-
-*Jingyang, did I ask you to code "other" if a dx was missing? Maybe we should change other to "unknown or other" since I doubt we were able to link an underlying dx to everyone with a fill.*
-
-Plot of SABA underlying Dx categories.
-
-```{r plot saba underlying dx}
-plot <- ggplot(sameday_results, 
-   aes(x=underlying_dx, y = estimates, color = underlying_dx)) +
-   scale_color_manual(name = "Underlying Diagnosis", 
-    values = c("#7303c0","#12d8fa", "#00467f", "#f80759"),
-    guide = guide_legend(direction="horizontal", title.position = "top",
-                         title.hjust = 0.5)) +
-  geom_point() +
-  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
-  geom_hline(yintercept = 1, linetype = 2, colour = "#FF0000") +
-  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    axis.title.x = element_blank(),
-    legend.position = "bottom") +
-  ylab(expression("Odds Ratio: 10 ug/m^3 increase smoke PM"[2.5])) 
-# plot
-plot
-```
-
-Interesting thing here is that COPD as underlying diagnosis is not nearly as pronounced as any of the other underlying diagnoses.
-Could be a sample size issue.
-We'll have to check.
-Perhaps there are few people with COPD only that are prescribed SABA.
-
-##9. Stratified cumulative results
-Considering adding. Splines likely hard to fit for smaller groups.
-
-##10. SABA cumulative effect by underlying respiratory diagnosis
-Considering adding. Splines likely hard to fit for smaller groups.
-
-##11. Conculsions
-So far, it looks like SABA and Asthma are the only associtions.
-Furthermore, the cumulative 0-7 day relationship between asthma and smoke has wide error bars and overlaps 1, suggesting the association could be null.
-I think it's more likely that this is because we introduce some noise on lag days 4-7.
-It's possible that an adjustment like weekend may help.
-It may also make sense to do something similar to Anna Rappold or others and shorten our lagged response.
-I also think this is a produce of sample size too.
-
-### Issues/Concerns (in order of time they popped in my head):
-1. File structure needs to be cleaned up. There are way too many subdirectories.
-2. Remove r code files that are not useful. There are too many files with similar code inside.
-3. Variables in the casecross over datasets raise some questions.
-4. Consider a cumulative effect that adds the preceding day and so on, i.e. 0, 0+1, 0+1+2, ...
-5. Plots need to have standard labels, ordering, scales, etc.
-6. I'd like to know n-cases and even effect estimates for SABA and asthma in the counties most impacted by smoke.

### 2018-06-06 r_scripts/analysis/wfs_exposure_figures.Rmd
Updated exposure figures to go in paper.

@@
  ggsave(filename = "./r_scripts/analysis/Figs/smoke_map.tiff", plot = smoke_map,
-       units="in", width=6, height=5, dpi=300)
+        units="in", width=7, height=6, dpi=300)
 ```
@@
   geom_point(color = "#6a3093", size = 0.5) +
-scale_x_date(date_labels = "%m") +
+  #scale_x_date(date_breaks = "1 month", labels = "%b") +
   geofacet::facet_geo(~county, grid = or_grid) +
-ylab(expression("Smoke PM2.5 ?g/m"^3)) +
-xlab("Month in 2013") +
+  ylab(expression(paste("Wildfire Smoke PM"[2.5], " µg/m"^3))) +
+  xlab("2013 Wildfire Season") +
   theme(panel.background = element_rect(fill = 'white', colour = 'black'),
@@
     strip.background = element_rect(colour=NA, fill=NA),
-legend.position = "bottom")
+    legend.position = "bottom",
+    axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0))
 # plot 
@@
 # save figure
-ggsave(filename = "./county_smoke_timeseries.png", plot = plot)
+# ggsave(filename = "./r_scripts/analysis/Figs/county_smoke_timeseries.tiff", 
+#        plot = plot, units="in", width=6, height=5, dpi=300)
 ```

### 2018-06-07 r_scripts/analysis/msa_asthma_timeseries.Rmd
updated msa smoke and asthma care plot.

---
title: "MSA Asthma Place of Service Timeseries"
author: "Ryan Gan"
date: "6/7/2018"
output: html_document
---

Libraries.

```{r library}
# libraries
library(tidyverse) # for ggplot and general data wrangle
```

Loading case-crossover dataframe list and SABA fill list.

```{r load_data}
# load asthma casecross place of service list
load("../../data/health/asthma_cc_pos_list.RData")
# extract names
vars_keep <- names(asthma_cc_pos_list[[4]])
```

Load SABA DF.

```{r saba_df}
saba = read_csv("../../data/health/2013-oregon_casecross_saba.csv") %>% 
  rename(identifier = personkey, ZIP = ZIPCODE) %>% 
  mutate(service_place = "SABA Fill",
         dx1 = ndc) %>% 
  # keeping only vars in the other casecrossover dataframes
  dplyr::select(vars_keep, ndc) %>% 
  # turn in to list
  list()

# assign name
names(saba) = "saba"
```

Load Jingyang's mothly count to extract populations for MSA.
```{r msa_pop}
# read saba and asthma ed count# read  
population <- read_csv("./data/health/saba_month_counts.csv") %>% 
  select(msa_name, POPESTIMATE2013) %>% 
  rename(metroarea = msa_name,
         pop = POPESTIMATE2013) %>% 
  unique()
```

Subset to only outcomes and add metroarea MSA.

```{r list_subset}
# adding in age category to each dataframe in list 
# (i should have done this earlier)
asthma_pos_count <- c(asthma_cc_pos_list, saba) %>% 
  map_dfr(., function(df){
    return_df <- df %>% 
      # remove unknown gender
      filter(outcome ==1 ) %>% 
      mutate(metroarea = case_when(MSA == 13460 ~ "Bend",
                                MSA == 18700 ~ "Corvallis",
                                MSA == 21660 ~ "Eugene",
                                MSA == 32780 ~ "Medford",
                                MSA == 38900 ~ "Portland",
                                MSA == 41420 ~ "Salem",
                                MSA == 41999 ~ "Not in MSA"),
        service_place = case_when(service_place == "Emergency Room – Hospital" ~ 
                                    "Emergency Department", 
          TRUE ~ service_place)) %>% 
      # group by date, metroarea, and service place
      group_by(date, metroarea, service_place) %>% 
      summarise(n_events = sum(outcome))
  })
```

```{r medford_plot}
check_ts <- asthma_pos_count %>% 
  filter(metroarea == 'Medford')

ggplot(check_ts, aes(x=date, y=n_events, group = metroarea, color=metroarea)) +
  geom_point() +
  facet_wrap(~service_place, scales = "free_y")
```

Weekly sum.

```{r weekly_count}
asthma_weekly_count <- asthma_pos_count %>% 
  mutate(week_date = lubridate::ceiling_date(date, unit = 'week')) %>% 
  group_by(week_date, metroarea, service_place) %>% 
  summarise(n_events = sum(n_events)) %>% 
  left_join(population, by = 'metroarea') %>% 
  mutate(rate_100k = (n_events/pop)*100000) %>% 
  filter(service_place != 'Ambulance-Land') %>% 
  filter(week_date != '2013-10-06') %>% 
  filter(!is.na(pop))
```

Weekly count plot.I actually like this plot more I think.

```{r pos_plot}
pos_plot = ggplot(asthma_weekly_count, 
  aes(x=week_date, y=rate_100k, group = service_place, color=service_place)) +
  geom_point(size = .7) +
  scale_color_manual('Care Utilization', 
    values = c("#3c1053", "#7f00ff", "#e100ff", "#00b09b", "#0ed2f7", "#c31432")) +
  facet_wrap(~metroarea, ncol=1) +
    ylab("Weekly Rate per 100,000 Persons") +
    xlab("Date") +
    ggtitle('B. Asthma Care Utilization') +
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        strip.text = elemnt_text(vjust= -1),
        # facet text size
        strip.text = element_text(size = 10),
        axis.title.y = element_text(angle = 90))

pos_plot
```

Read PM2.5 and create MSA PM

```{r}
# read saba
path = paste0('../../data/health/or_respiratory_may_to_sep_time_strat_casecross_er.csv')

# read respiratory casecross and find unique zipcodes by MSA values
zip_msa = read_csv(path, col_types = cols(.default = "c")) %>% 
    select(ZIPCODE, MSA) %>% 
    unique()

# read zip pm; join with msa
msa_pm = read_csv('./data//pm/2013-oregon_zip_pm25.csv', 
                  col_types = cols(ZIPCODE = 'c')) %>% 
    # join with unique MSA vector
    left_join(zip_msa, by = 'ZIPCODE') %>%
    # mutate
    mutate(ZIPCODE = as.factor(ZIPCODE),
        MSA = as.factor(MSA),
        # assign metro name to number
        metroarea = case_when(MSA == 13460 ~ "Bend",
                              MSA == 18700 ~ "Corvallis",
                              MSA == 21660 ~ "Eugene",
                              MSA == 32780 ~ "Medford",
                              MSA == 38900 ~ "Portland",
                              MSA == 41420 ~ "Salem"))  %>% 
    # filter to zips in an MSA only
    filter(!is.na(metroarea)) 
```

Plot of PM2.5 by MSA.

```{r pm_plot}
# plot of pm by msa
pm_msa_plot = ggplot(data = msa_pm, aes(x = date, y = geo_smk_pm, group = ZIPCODE)) +
    geom_line(color = '#e94057') +
    # facet by metro area
    facet_wrap(~metroarea, ncol = 1) +
    # y label
    ylab(expression(paste("Wildfire Smoke PM"[2.5]," 10µg/m"^3))) +
    # x label
    xlab("Date") +
    ggtitle(expression(paste('A. Wildfire Smoke PM'[2.5]))) +
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        strip.text = element_text(size = 10),
        axis.title.y = element_text(angle = 90))

pm_msa_plot
```

### 2018-06-08 r_scripts/analysis/msa_asthma_timeseries.Rmd
fixed some things.

@@
   aes(x=week_date, y=rate_100k, group = service_place, color=service_place)) +
-  geom_point(size = .7) +
+  geom_point(size = 1.25) +
   scale_color_manual('Care Utilization', 
@@
     theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-        panel.grid.major = element_blank(),
+        panel.grid.major = element_line(color = 'grey', linetype = 'dotted',
+                                        size = 0.5),
         panel.grid.minor = element_blank(),
@@
         panel.border = element_rect(fill = NA, color = "black"),
-        strip.text = elemnt_text(vjust= -1),
         # facet text size
         strip.text = element_text(size = 10),
-        axis.title.y = element_text(angle = 90))
+        axis.title.y = element_text(angle = 90),
+        legend.key = element_blank())
 
@@
 
-```{r}
+```{r msa_pm}
 # read saba
@@
 ```
+
+Bind plots together.
+
+```{r final_plot}
+# save plot
+#tiff('./r_scripts/analysis/Figs/msa_smoke_asthma.tiff', width = 600, height = 650)
+gridExtra::grid.arrange(pm_msa_plot, pos_plot, ncol=2, widths = c(1,1))
+#dev.off()
+```
+
+Saving dataframe to plot in python.
+
+```{r save_df}
+# write asthma weekly count
+#write_csv(asthma_weekly_count, './data/health/asthma_weekly_count.csv')
+# write pm msa
+#write_csv(msa_pm, './data/pm/msa_pm.csv')
+```

### 2018-06-08 r_scripts/analysis/msa_asthma_timeseries.Rmd
Couldn't figure out overlap.

@@
 
-```{r save_df}
+```{r save_df, eval=F}
 # write asthma weekly count
@@
 ```
+
+Overlap plot.
+
+```{r eval=F}
+grid::grid.newpage()
+vp1 <- grid::viewport(width = 1, height = 1, x = 0.5, y = 0.5)
+vp2 <- grid::viewport(width = 1, height = 1, x = 0.5, y = 0.5)
+print(pm_msa_plot, vp = vp1)
+print(pos_plot, vp = vp2)
+```

### 2018-06-11 r_scripts/analysis/asthma_place_of_service_analysis.Rmd
updated some stuff.

@@
 ```{r other_msa}
-other_msa_result <- asthma_pos_list[c(1,3:5,7)] %>% 
+msa <- unique(asthma_pos_list[[1]]$metroarea)
+
+msa_result <- asthma_pos_list[c(1,3:5,7)] %>% 
   map_dfr(., function(df){
+
     service_place <- as.character(df$service_place[1])
     # complete case
-    df <- df %>% filter(!is.na(geo_smk10)) %>% 
-      filter(metroarea != "Medford") %>% 
-      filter(metroarea != "Not in MSA")
-    # model
+    pos_list <- df %>% filter(!is.na(geo_smk10)) %>% 
+      filter(metroarea != "Not in MSA") %>% 
+      split(.$metroarea) %>% 
+      map_dfr(., function(df2){
+        # find msa
+        msa <- as.character(df2$metroarea[[1]])
+        # runm odel
         mod <- clogit(outcome ~ geo_smk10 + wrf_temp + strata(identifier), 
-                  data = df)
+                      data = df2)
       # n events
@@
         mutate_at(2:4, funs(round(exp(.),3))) %>% 
-      cbind(service_place, n_events, .) %>% 
-      mutate(service_place = as.character(service_place))
+        cbind(service_place, msa, n_events, .) %>% 
+        mutate(service_place = as.character(service_place),
+               msa = as.character(msa))})
   }) %>% 
@@
 # table
-other_msa_result
+msa_result
 ```
@@
 ```{r medford_plot}
-msa_results <- bind_rows(medford_result, other_msa_result)
+
 # plot total results
-med_plot <- ggplot(msa_results, aes(x = service_place, y = odds_ratio)) +
+med_plot <- ggplot(msa_result, aes(x = service_place, y = odds_ratio)) +
   geom_point(color = "purple") + 
@@
   geom_hline(yintercept = 1, linetype=2) +
-  facet_wrap(~MSA) +
+  facet_wrap(~msa, scales = 'free_y') +
   ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +

### 2018-06-11 r_scripts/analysis/msa_asthma_timeseries.Rmd
updated some stuff.

@@
 ```{r pos_plot}
-pos_plot = ggplot(asthma_weekly_count, 
+pos_plot <- ggplot(asthma_weekly_count, 
   aes(x=week_date, y=rate_100k, group = service_place, color=service_place)) +
@@
 # read zip pm; join with msa
-msa_pm = read_csv('./data//pm/2013-oregon_zip_pm25.csv', 
+msa_pm <- read_csv('./data//pm/2013-oregon_zip_pm25.csv', 
                   col_types = cols(ZIPCODE = 'c')) %>% 
@@
 # plot of pm by msa
-pm_msa_plot = ggplot(data = msa_pm, aes(x = date, y = geo_smk_pm, group = ZIPCODE)) +
+pm_msa_plot <- ggplot(data = msa_pm, aes(x = date, y = geo_smk_pm, group = ZIPCODE)) +
     geom_line(color = '#e94057') +
@@
 
+### Monthly Rate and Sum by MSA
+
+```{r monthly_counts_msa}
+monthly_msa_counts <- asthma_pos_count %>% 
+  mutate(month = lubridate::month(date)) %>% 
+  group_by(month, metroarea, service_place) %>% 
+  summarise(n_events = sum(n_events)) %>% 
+  left_join(population, by = 'metroarea') %>% 
+  mutate(rate_100k = (n_events/pop)*100000,
+         month_name = factor(case_when(month == 5 ~ 'May',
+                                       month == 6 ~ 'Jun',
+                                       month == 7 ~ 'Jul',
+                                       month == 8 ~ 'Aug', 
+                                       month == 9 ~ 'Sep'),
+                        levels = c('May', 'Jun', 'Jul', 'Aug', 'Sep'))) %>% 
+  filter(service_place != 'Ambulance-Land') %>% 
+  filter(service_place != 'Urgent Care Facility') %>% 
+  filter(!is.na(pop)) %>% 
+  arrange(service_place, metroarea, month)
+```
+
+Plot montly rate by MSA.
+
+ED Rate
+
+```{r ed_plot}
+ed_count <- monthly_msa_counts %>% filter(service_place == 'Emergency Department')
+
+# plot
+ed_plot <- ggplot(ed_count, aes(x = month_name, group = metroarea,
+                                        fill = metroarea)) +
+  geom_bar(aes(weight = rate_100k)) +
+  scale_fill_manual("Metro Area", values = c("#3c1053", "#7f00ff", "#e100ff", 
+                                             "#4389a2", "#0ed2f7", "#74ebd5"),
+                    guide = F) +
+  facet_wrap(~metroarea, ncol =1) +
+  ylab("Monthly Rate per 100,000 Persons") +
+  xlab("Date") +
+  ggtitle('A. ED') +
+  theme_minimal() +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+        panel.grid.major = element_line(color = 'grey', linetype = 'dotted',
+                                        size = 0.5),
+        panel.grid.minor = element_blank(),
+        # strip element
+        strip.background = element_rect(colour=NA, fill=NA),
+        panel.border = element_rect(fill = NA, color = "black"),
+        # facet text size
+        strip.text = element_text(size = 10),
+        axis.title.y = element_text(angle = 90),
+        legend.key = element_blank(),
+        axis.text.x=element_text(angle = 45, hjust = 1))
+
+ed_plot
+```
+
+Outpatient.
+
+```{r outpat}
+out_count <- monthly_msa_counts %>% 
+  filter(service_place == 'Outpatient Hospital')
+# plot
+out_plot <- ggplot(out_count, aes(x = month_name, group = metroarea,
+                                        fill = metroarea)) +
+  geom_bar(aes(weight = rate_100k)) +
+  scale_fill_manual("Metro Area", values = c("#3c1053", "#7f00ff", "#e100ff", 
+                                             "#4389a2", "#0ed2f7", "#74ebd5"),
+                    guide = F) +
+  facet_wrap(~metroarea, ncol =1) +
+  xlab("Date") +
+  ggtitle('B. Outpatient') +
+  theme_minimal() +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+        panel.grid.major = element_line(color = 'grey', linetype = 'dotted',
+                                        size = 0.5),
+        panel.grid.minor = element_blank(),
+        # strip element
+        strip.background = element_rect(colour=NA, fill=NA),
+        panel.border = element_rect(fill = NA, color = "black"),
+        # facet text size
+        strip.text = element_text(size = 10),
+        axis.title.y = element_blank(),
+        legend.key = element_blank(),
+        axis.text.x=element_text(angle = 45, hjust = 1))
+
+out_plot
+```
+
+Inpatient.
+
+```{r outpat}
+inpat_count <- monthly_msa_counts %>% 
+  filter(service_place == 'Inpatient Hospital')
+# plot
+inpat_plot <- ggplot(inpat_count, aes(x = month_name, group = metroarea,
+                                        fill = metroarea)) +
+  geom_bar(aes(weight = rate_100k)) +
+  scale_fill_manual("Metro Area", values = c("#3c1053", "#7f00ff", "#e100ff", 
+                                             "#4389a2", "#0ed2f7", "#74ebd5"),
+                    guide = F) +
+  facet_wrap(~metroarea, ncol =1) +
+  ylab("Monthly Rate per 100,000 Persons") +
+  xlab("Date") +
+  ggtitle('C. Inpatient') +
+  theme_minimal() +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+        panel.grid.major = element_line(color = 'grey', linetype = 'dotted',
+                                        size = 0.5),
+        panel.grid.minor = element_blank(),
+        # strip element
+        strip.background = element_rect(colour=NA, fill=NA),
+        panel.border = element_rect(fill = NA, color = "black"),
+        # facet text size
+        strip.text = element_text(size = 10),
+        axis.title.y = element_blank(),
+        axis.text.x=element_text(angle = 45, hjust = 1))
+
+inpat_plot
+```
+
+Office.
+
+```{r office}
+off_count <- monthly_msa_counts %>% 
+  filter(service_place == 'Office')
+# plot
+off_plot <- ggplot(off_count, aes(x = month_name, group = metroarea,
+                                        fill = metroarea)) +
+  geom_bar(aes(weight = rate_100k)) +
+  scale_fill_manual("Metro Area", values = c("#3c1053", "#7f00ff", "#e100ff", 
+                                             "#4389a2", "#0ed2f7", "#74ebd5"),
+                    guide = F) +
+  facet_wrap(~metroarea, ncol =1) +
+  ylab("Monthly Rate per 100,000 Persons") +
+  xlab("Date") +
+  ggtitle('D. Office') +
+  theme_minimal() +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+        panel.grid.major = element_line(color = 'grey', linetype = 'dotted',
+                                        size = 0.5),
+        panel.grid.minor = element_blank(),
+        # strip element
+        strip.background = element_rect(colour=NA, fill=NA),
+        panel.border = element_rect(fill = NA, color = "black"),
+        # facet text size
+        strip.text = element_text(size = 10),
+        axis.title.y = element_blank(),
+        legend.key = element_blank(),
+        axis.text.x=element_text(angle = 45, hjust = 1))
+
+off_plot
+```
+
+SABA fill.
+
+```{r saba}
+saba_count <- monthly_msa_counts %>% 
+  filter(service_place == 'SABA Fill')
+# plot
+saba_plot <- ggplot(saba_count, aes(x = month_name, group = metroarea,
+                                        fill = metroarea)) +
+  geom_bar(aes(weight = rate_100k)) +
+  scale_fill_manual("Metro Area", values = c("#3c1053", "#7f00ff", "#e100ff", 
+                                             "#4389a2", "#0ed2f7", "#74ebd5")) +
+  facet_wrap(~metroarea, ncol =1) +
+  ylab("Monthly Rate per 100,000 Persons") +
+  xlab("Date") +
+  ggtitle('E. SABA') +
+  theme_minimal() +
+  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+        panel.grid.major = element_line(color = 'grey', linetype = 'dotted',
+                                        size = 0.5),
+        panel.grid.minor = element_blank(),
+        # strip element
+        strip.background = element_rect(colour=NA, fill=NA),
+        panel.border = element_rect(fill = NA, color = "black"),
+        # facet text size
+        strip.text = element_text(size = 10),
+        axis.title.y = element_blank(),
+        legend.key = element_blank(),
+        axis.text.x=element_text(angle = 45, hjust = 1))
+
+saba_plot
+```
+
+Join all plots.
+
+```{r msa_rate_plot}
+# save plot
+#tiff('./r_scripts/analysis/Figs/msa_month_rate.tiff', width = 800, height = 600)
+gridExtra::grid.arrange(ed_plot, out_plot, inpat_plot, off_plot, saba_plot, 
+                        ncol=5, widths = c(0.2, 0.2, 0.2, 0.2, 0.3))
+#dev.off()
+```
+
+Compare August to July and September for each MSA.
+
+Model
+```{r rate_compare}
+metro <- unique(monthly_msa_counts$metroarea)
+service_place <- unique(monthly_msa_counts$service_place)
+month <- seq(5,9,by=1)
+# expand grid
+month_msa <- expand.grid(metro, service_place, month) %>% 
+  rename(metroarea = Var1, service_place = Var2, month = Var3)
+
+summary(outcome_list)
+# create outcome list
+outcome_list <- month_msa %>% 
+  left_join(monthly_msa_counts, by = c("metroarea", "service_place", "month")) %>% 
+  mutate(n_events = ifelse(is.na(n_events), 0, n_events),
+         pop = ifelse(is.na(pop), 85934, pop),
+         month_name = factor(case_when(month == 5 ~ 'May',
+            month == 6 ~ 'Jun',
+            month == 7 ~ 'Jul',
+            month == 8 ~ 'Aug', 
+            month == 9 ~ 'Sep'),
+            levels = c('May', 'Jun', 'Jul', 'Aug', 'Sep'))) %>% 
+  split(.$service_place)
+
+# function list
+compare_df <- outcome_list %>% 
+  map_dfr(., function(df){
+    df <- outcome_list[[5]]
+    serv_pl <- unique(df$service_place)
+    serv_pl
+    # model
+    mod <- glm(n_events ~ 0 + metroarea*month_name + offset(log(pop)), 
+                family = poisson, data = df)
+
+    summary(mod)
+    contr_msa_jul_to_sep <- rbind(
+      #'Bend: Aug to Jul' = c(0,0,0,0,0,0,0,-1,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
+      #'Bend: Aug to Sept' = c(0,0,0,0,0,0,0,0,-1,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
+      #'Cor: Aug to Jul' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0,0,0,0,0,0,0,0),
+      #'Cor: Aug to Sept' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0,0,0),
+      #'Eug: Aug to Jul' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0,0,0,0,0,0,0),
+      #'Eug: Aug to Sept' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0,0),
+      'Med: Aug to Jul' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0,0,0,0,0,0),
+      'Med: Aug to Sept' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0),
+      'Port vs Med: Aug to Jul' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-.5,-.5,0,0,0,.5,.5,0,0,0,0,0,0)
+      #'Port: Aug to Sept' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0),
+      #'Sal: Aug to Jul' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0,0,0,0),
+      #'Sal: Aug to Sept' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1)
+      )
+
+    # multicomp comparison of contrasts
+    jul_sep_contrast <- multcomp::glht(mod, contr_msa_jul_to_sep)
+    # summary of adjusted
+    compare_df <- data.frame(broom::tidy(
+      summary(jul_sep_contrast, test = multcomp::adjusted('none')))) %>% 
+      mutate(service_place = serv_pl) %>% 
+      select(service_place, lhs:p.value)
+  })
+compare_df
+```
+
+Test change by MSA.
+
+```{r}
+print(saba_count)
+```
+
+Fitting a time-series model to predict the observed across MSAs. I need the spline package.
+
+```{r spline}
+library(splines)
+```
+
+I want to map across each outcome and metroarea to fit splines I think. I'll first try an interaction spline by MSA for saba fills.
+
+```{r weekly_pm}
+weekly_pm = msa_pm %>% 
+  mutate(week_date = lubridate::ceiling_date(date, unit = 'week')) %>% 
+  group_by(metroarea, week_date) %>% 
+  summarise(mu_gwr_smk = mean(geo_smk_pm), mu_gwr = mean(geo_wt_pm))
+
+summary(weekly_pm)
+```
+
+```{r}
+glimpse(asthma_weekly_count)
+
+# vector of outcomes
+out_vec = unique(asthma_weekly_count$service_place)
+
+# list of metroareas
+msa_list = asthma_weekly_count %>% 
+  split(., list(.$metroarea,.$service_place))
+
+# map to get estiamtes of counts for each 
+exp_n = msa_list %>% 
+  map_dfr(., function(data){
+    service_place = as.character(data$service_place[1])
+    metroarea = as.character(data$metroarea[1])
+    # vector of dates
+    week_date = as.character(data$week_date)
+    # spline with 4 degrees of freedom fits msas pretty well
+    spl = ns(data$week_date, df = 4)
+    # fit spline to counts
+    mod = glm(n_events ~  spl, family = 'poisson'(link='identity'), 
+          data = data)
+    # estimated outcomes
+    est = predict(mod, newdata=data, type = 'response') 
+    # dataframe of estimates by msa and place of service
+    est_df = data.frame(cbind(service_place, metroarea, week_date, est)) %>% 
+      mutate(week_date = as.Date(week_date, format = "%Y-%m-%d"),
+             est = as.numeric(as.character(est)),
+             metroarea = as.character(metroarea),
+             service_place = as.character(service_place)) 
+  }) %>% 
+  left_join(asthma_weekly_count, ., 
+            by = c('metroarea', 'service_place', 'week_date')) %>% 
+  mutate(diff = est - n_events,
+         est_rate = (est/pop)*100000,
+         diff_rate = rate_100k - est_rate)
+
+# plot of expected over observed
+ggplot(data=exp_n, aes(x=week_date, y=diff_rate, 
+    group = service_place, color = service_place)) +
+  geom_point() +                 
+  facet_wrap(~metroarea)
+```
+
 Saving dataframe to plot in python.
@@
 #write_csv(msa_pm, './data/pm/msa_pm.csv')
+#write_csv(monthly_msa_counts, './data/monthly_counts.csv')
 ```
@@
 ```
+
+```{r}
+exp(-0.295)
+```

### 2018-06-12 r_scripts/analysis/msa_asthma_timeseries.Rmd
updated standardized plot. Too much noise to put in the paper.

@@
 
+# find mean and standard deviation for each difference rate by place of service
+# and metroarea
+normalize_val <- exp_n %>% 
+  group_by(metroarea, service_place) %>% 
+  summarise(mu_diff = mean(diff_rate), sd_diff = sd(diff_rate))
+
+# join sd to diff and calcualte normalized difference
+estimated_n <- exp_n %>% 
+  left_join(normalize_val, by = c("metroarea", "service_place")) %>% 
+  mutate(diff_rate_sd = (diff_rate - mu_diff)/sd_diff)
+
 # plot of expected over observed
-ggplot(data=exp_n, aes(x=week_date, y=diff_rate, 
+ggplot(data=estimated_n, aes(x=week_date, y=diff_rate_sd, 
     group = service_place, color = service_place)) +

### 2018-06-14 r_scripts/analysis/msa_asthma_timeseries.Rmd
updated asthma time series.

@@
   map_dfr(., function(df){
-    df <- outcome_list[[5]]
+
     serv_pl <- unique(df$service_place)
@@
 
-    summary(mod)
     contr_msa_jul_to_sep <- rbind(
@@
       'Med: Aug to Sept' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0),
-      'Port vs Med: Aug to Jul' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-.5,-.5,0,0,0,.5,.5,0,0,0,0,0,0)
+      'M v B: Aug to Jul' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,1,0,-1,0,0,1,0,0,0,0,0,0,0),
+      'M v C: Aug to Jul' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,1,0,0,-1,0,1,0,0,0,0,0,0,0),
+      'M v E: Aug to Jul' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,1,0,0,0,-1,1,0,0,0,0,0,0,0),
+      'M v P: Aug to Jul' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,1,0,0,0,-1,1,0,0,0,0,0,0),
+      'M v S: Aug to Jul' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,1,0,0,-1,0,1,0,0,0,0,0)
       #'Port: Aug to Sept' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0),

### 2018-06-19 r_scripts/analysis/asthma_place_of_service_analysis.Rmd
updated lag associaiton.

@@
 
+## Lagged Associations
+
+```{r lag_association}
+# itterate over each geosmk
+pm_vec <- c('geo_smk10', paste0('geo_smk10_lag', seq(1,5,1)))
+temp_vec <- c('wrf_temp', paste0('wrf_temp_lag', seq(1,5,1)))
+
+# map lag
+lag_result <- asthma_pos_list %>% 
+  map_dfr(., function(df){
+    service_place <- as.character(df$service_place[1])
+    # complete case up to lag day 5
+    df <- df %>% filter(!is.na(geo_smk10_lag5))
+    # lag map
+    result <- map2_dfr(pm_vec, temp_vec, function(.x, .y){
+      # model
+      mod <- clogit(as.formula(paste0('outcome ~', 
+             paste(.x, .y, 'strata(identifier)', sep = '+'))),
+             data = df)
+      # n events
+      n_events <- mod$nevent
+      # lag var
+      lag_var <- pm_vec
+      # odds ratio and 95% CI
+        estimates <- broom::tidy(mod) %>% 
+          #filter(term == lag_var) %>% 
+          dplyr::select(term, estimate, conf.low, conf.high) %>% 
+          mutate_at(2:4, funs(round(exp(.),3))) %>% 
+          cbind(service_place, n_events, .) %>% 
+          mutate(service_place = as.character(service_place))
+    }) %>% filter(grepl('geo_smk10', term))
+  }) %>% 
+  mutate(strata = "Total",
+         Lag = case_when(term == 'geo_smk10' ~ 'Lag 0',
+                         term == 'geo_smk10_lag1' ~ 'Lag 1',
+                         term == 'geo_smk10_lag2' ~ 'Lag 2',
+                         term == 'geo_smk10_lag3' ~ 'Lag 3',
+                         term == 'geo_smk10_lag4' ~ 'Lag 4',
+                         term == 'geo_smk10_lag5' ~ 'Lag 5'))
+
+# table
+lag_result
+```
+
+Plot for lag.
+
+```{r lag_plot}
+glimpse(lag_result)
+
+lag_result <- lag_result %>% 
+  mutate(service_place = ifelse(service_place == 'Emergency Room – Hospital',
+                                'Emergency Department', service_place))
+
+lag_plot <- ggplot(lag_result, aes(x = Lag, y = estimate)) +
+  geom_point(color = "#3c1053") + 
+  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), color = "#3c1053",width = 0.2) +
+  geom_hline(yintercept = 1, linetype=2) +
+  facet_wrap(~service_place, scales = "free_y", ncol = 2) +
+  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in WFS PM"[2.5]))) +
+  xlab("Lag") +
+   # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10),
+    # axis element (tilt)
+    axis.text.x = element_text(angle = 45, hjust = 0.9, vjust = 0.7),
+    #axis.title.x = element_blank(),
+    #axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90))
+
+lag_plot
+# save plot
+# ggsave("./Figs/asthma_pos_lag_plot.tiff", lag_plot, 
+#        units="in", width=8, height=6, dpi=300)
+```
+
 ## Medford-Specific Results

### 2018-07-09 r_scripts/analysis/msa_asthma_timeseries.Rmd
updated contrast estimates to figure out diff in diff.

@@
 ```{r saba_df}
-saba = read_csv("../../data/health/2013-oregon_casecross_saba.csv") %>% 
+saba <- read_csv("./data/health/2013-oregon_casecross_saba.csv") %>% 
   rename(identifier = personkey, ZIP = ZIPCODE) %>% 
@@
 population <- read_csv("./data/health/saba_month_counts.csv") %>% 
-  select(msa_name, POPESTIMATE2013) %>% 
+  dplyr::select(msa_name, POPESTIMATE2013) %>% 
   rename(metroarea = msa_name,
@@
   unique()
+
 ```
@@
 # read saba
-path = paste0('../../data/health/or_respiratory_may_to_sep_time_strat_casecross_er.csv')
+path = paste0('./data/health/or_respiratory_may_to_sep_time_strat_casecross_er.csv')
 
@@
 zip_msa = read_csv(path, col_types = cols(.default = "c")) %>% 
-    select(ZIPCODE, MSA) %>% 
+    dplyr::select(ZIPCODE, MSA) %>% 
     unique()
@@
     contr_msa_jul_to_sep <- rbind(
-      #'Bend: Aug to Jul' = c(0,0,0,0,0,0,0,-1,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
-      #'Bend: Aug to Sept' = c(0,0,0,0,0,0,0,0,-1,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
-      #'Cor: Aug to Jul' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0,0,0,0,0,0,0,0),
-      #'Cor: Aug to Sept' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0,0,0),
-      #'Eug: Aug to Jul' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0,0,0,0,0,0,0),
-      #'Eug: Aug to Sept' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0,0),
       'Med: Aug to Jul' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0,0,0,0,0,0),
@@
       'M v S: Aug to Jul' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,1,0,0,-1,0,1,0,0,0,0,0)
-      #'Port: Aug to Sept' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0),
-      #'Sal: Aug to Jul' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0,0,0,0),
-      #'Sal: Aug to Sept' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1)
       )
@@
       mutate(service_place = serv_pl) %>% 
-      select(service_place, lhs:p.value)
+      dplyr::select(service_place, lhs:p.value)
   })
@@
 
-Test change by MSA.
+Difference in difference test model.
+```{r diff_in_diff}
+saba_count <- monthly_msa_counts %>% 
+  # filter to SABA fill
+  filter(service_place == 'SABA Fill') %>% 
+  # filter to portland and medford to test
+  filter(metroarea %in% c('Portland', 'Medford'))
+
+# diff in rate is 49.24 from july to aug for medford and -2.47 for portland
+
+# running linear model to figure out rate
+mod <- glm(n_events ~ 0 + metroarea*month_name, 
+              family = poisson(link='identity'), data = saba_count)
+
+summary(mod)
+```
+
+Contrast on model.
+
+```{r diff_in_diff_contrast}
+city_cont <- rbind(
+      'Med: Aug to Jul' = c(0,0,0,-1,1,0,0,0,0,0),
+      'Port: Aug to Jul' = c(0,0,0,0,0,0,0,-1,1,0)
+      )
+
+# multicomp comparison of contrasts
+within_contrast <- multcomp::glht(mod, city_cont)
+
+summary(within_contrast)
+
+city_cont
+diff_cont <- rbind(
+  'diff' = (city_cont['Med: Aug to Jul',] - city_cont['Port: Aug to Jul', ])
+  )
+
+diff <- multcomp::glht(mod, diff_cont)
+# expecting 261 difference
+summary(diff)
+```
+
+261 diff.
+
+City difference in difference from July to August in cities compared to Medford.
+
+```{r diff_in_diff_final}
+# function list
+diff_df <- outcome_list %>% 
+  map_dfr(., function(df){
+
+    serv_pl <- unique(df$service_place)
+    serv_pl
+    # model
+    mod <- glm(n_events ~ 0 + metroarea*month_name + offset(log(pop)), 
+                family = poisson, data = df)
+
+    city_cont <- rbind(
+      'Med: Jul to Aug' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0,0,0,0,0,0),
+      'Bend: Jul to Aug' = c(0,0,0,0,0,0,0,-1,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
+      'Cor: Jul to Aug' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0,0,0,0,0,0,0,0),
+      'Eug: Jul to Aug' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0,0,0,0,0,0,0),
+      'Port: Jul to Aug' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0,0,0,0,0),
+      'Salem: Jul to Aug' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0,0,0,0)
+      )
+
+    diff_cont <- rbind(
+      'Med to Port diff' = (city_cont['Med: Jul to Aug',] - city_cont['Port: Jul to Aug', ]),
+      'Med to Bend diff' = (city_cont['Med: Jul to Aug',] - city_cont['Bend: Jul to Aug', ]),
+      'Med to Cor diff' = (city_cont['Med: Jul to Aug',] - city_cont['Cor: Jul to Aug', ]),
+      'Med to Eug diff' = (city_cont['Med: Jul to Aug',] - city_cont['Eug: Jul to Aug', ]),
+      'Med to Sal diff' = (city_cont['Med: Jul to Aug',] - city_cont['Salem: Jul to Aug', ])
+    )
+
+    # multicomp comparison of contrasts
+    diff_contrast <- multcomp::glht(mod, diff_cont)
+
+    # summary of adjusted
+    diff_df <- data.frame(broom::tidy(
+      summary(diff_contrast, test = multcomp::adjusted('none')))) %>% 
+      mutate(service_place = serv_pl) %>% 
+      dplyr::select(service_place, lhs:p.value)
+  })
+easy_read_p <- diff_df %>% 
+  mutate(p = round(p.value, 3))
+easy_read_p
+```
 
 ```{r}
-print(saba_count)
+
 ```

### 2018-08-12 r_scripts/analysis/msa_asthma_timeseries.Rmd
deleted useless slot.
updated something.

@@
 # load asthma casecross place of service list
-load("../../data/health/asthma_cc_pos_list.RData")
+load("./data/health/asthma_cc_pos_list.RData")
 # extract names
@@
 
-```{r}
-
-```
-
 Fitting a time-series model to predict the observed across MSAs. I need the spline package.

### 2018-09-04 r_scripts/analysis/asthma_place_of_service_analysis.Rmd
updated main result plot for job talk

@@
 sameday_result
+```
+
+Plot for job talk presentation.
+
+```{r total_strata_plot}
+# filter out amublance since the values are skewed
+plot_df <- sameday_result %>% filter(service_place != 'Ambulance-Land')
+
+plot <- ggplot(plot_df, aes(x = service_place, y = estimate)) +
+  geom_point(color = "#3c1053") + 
+  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
+                color = "#3c1053",width = 0.2) +
+  geom_hline(yintercept = 1, linetype=2) +
+  ylab(expression(paste("Odds Ratio for 10µg/m"^3, 
+                        " Increase in WFS PM"[2.5]))) +
+  xlab("Place of Service") +
+   # plot theme
+    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    # strip element
+    strip.background = element_rect(colour=NA, fill=NA),
+    panel.border = element_rect(fill = NA, color = "black"),
+    # facet text size
+    strip.text = element_text(size = 10),
+    # axis element (tilt)
+    axis.text.x = element_text(angle = 45, hjust = 0.9, vjust = 0.7),
+    #axis.title.x = element_blank(),
+    #axis.ticks.x = element_blank(),
+    axis.title.y = element_text(angle = 90))
+
+# print plot
+plot
+#ggsave("./Figs/asthma_pos_total.tiff", plot, 
+#       units="in", width=8, height=6, dpi=300)
 ```

### 2018-09-17 r_scripts/analysis/wfs_exposure_figures.Rmd
Updated scripts for paper.

@@
   # binary smoke variable at >10 ug/m^3
-  mutate(gwr_smk10 = ifelse(geo_smk_pm >= 10, 1, 0)) %>% 
+  mutate(gwr_smk10 = ifelse(geo_smk_pm >= 10, 1, 0),
+         gwr_smk15 = ifelse(geo_smk_pm >= 15, 1, 0)) %>% 
   group_by(county) %>% 
-  summarise(smk_day10 = sum(gwr_smk10)) %>% 
+  summarise(smk_day10 = sum(gwr_smk10), smk_day15 = sum(gwr_smk15)) %>% 
   # lower case county name and remove "." to match spatial df
@@
 
-We summarize smoke exposure two ways: as a count of days where smoke PM~2.5~ is > 10 ug/m^3 (to increase the specificity to identify areas impacted by smoke) at the county level, and time-series of county-population weighted PM~2.5~ over May to October 2013.
+We summarize smoke exposure two ways: as a count of days where smoke PM~2.5~ is > 15 ug/m^3 (to increase the specificity to identify areas impacted by smoke) at the county level, and time-series of county-population weighted PM~2.5~ over May to October 2013.
 
-### Number of days in Oregon where smoke PM~2.5~ exceeded 10 ug/m^3.
+### Number of days in Oregon where smoke PM~2.5~ exceeded 15 ug/m^3.
 
@@
   # fill with number of smoke days
-  geom_polygon(aes(fill =smk_day10)) +
-  scale_fill_gradient2(expression("Smoke Days > 10 µg/m"^3),
+  geom_polygon(aes(fill = smk_day15)) +
+  scale_fill_gradient2(expression("Smoke Days > 15 µg/m"^3),
                        low = "white", mid= "#00dbde", high="#6a3093", 
-                       midpoint = 15,
-    guide = guide_colorbar(direction="horizontal", title.position = "top",
-                          title.hjust = 0.5, barwidth = 20)) +
+                       midpoint = 10,
+    guide = guide_colorbar(direction = "horizontal", title.position = "top",
+                          title.vjust = 2, title.hjust = 0.5, barwidth = 15)) +
   # add county path on top
@@
              color = "red", shape = 17, alpha = 0.7, 
-             show.legend = F, inherit.aes = FALSE) +
+             show.legend = T, inherit.aes = FALSE) +
+  # legend for firsts for fires
+  scale_size('Wildfire Acres Burned', breaks = c(0.8, 1, 1.6), 
+             labels = c(1600, 4000, 26000),
+             guide = guide_legend(title.position = "top", direction = "horizontal",
+                                  title.hjust = 0.5, barwidth = 10)) +
   # add county text 
@@
         strip.background = element_rect(colour=NA, fill=NA),
-        legend.position = "bottom")
+        legend.position = "bottom",
+        legend.key = element_rect(fill = NA, colour = NA, size = 0.25))
 

### 2018-09-17 r_scripts/data_wrangle/asthma_saba_msa_timeseries.R

@@ -1,0 +1,125 @@
+# ------------------------------------------------------------------------------
+# Title: Creation of asthma and saba fill Oregon and MSA time series
+# Author: Ryan Gan
+# Date Created: 2018-09-12
+# ------------------------------------------------------------------------------
+
+# Script purpose is to create time series data frames for Oregon, and each
+# Oregon MSA for analysis
+
+# load tidyverse library
+library(tidyverse)
+
+# I am able to create the time series I need using the asthma fireseason cohort
+# dataset created. This contains all primary diagnoses of asthma and saba fills
+# during the study period.
+
+# read asthma cohort (n = 550610)
+asthma_c <- read_csv('./data/health/2013-oregon_asthma_fireseason_cohort.csv',
+                     col_types = cols(.default = "c")) %>% 
+  # filter to just primary diagnosis or saba fill
+  filter(visit_type != 'dx_asthma_not_primary')
+
+# find zipcodes in each metroarea
+zip <- asthma_c %>% 
+  select(ZIP, MSA) %>% 
+  rename(ZIPCODE = ZIP) %>% 
+  unique()
+
+# read zip pm; join with msa
+msa_pm = read_csv('./data/pm/2013-oregon_zip_pm25.csv', 
+                  col_types = cols(ZIPCODE = 'c')) %>% 
+  # join with unique MSA vector
+  left_join(zip, by = 'ZIPCODE') %>%
+  # mutate
+  mutate(ZIPCODE = as.character(ZIPCODE),
+         MSA = as.factor(MSA),
+         # assign metro name to number
+         metroarea = case_when(MSA == 13460 ~ "Bend",
+                               MSA == 18700 ~ "Corvallis",
+                               MSA == 21660 ~ "Eugene",
+                               MSA == 32780 ~ "Medford",
+                               MSA == 38900 ~ "Portland",
+                               MSA == 41420 ~ "Salem"))  %>% 
+  # filter to zips in an MSA only
+  filter(!is.na(metroarea))
+
+# It's possible we could duplicate claimids; not sure I want to count these
+claim_count <- asthma_c %>% 
+  group_by(clmid) %>% 
+  summarize(count = n())
+# looking at an example where a person had multiple asthma inhaler fills with
+# same claimid. It looks like they are all filled on unique dates, so I'm going
[…]

### 2018-09-18 r_scripts/data_wrangle/asthma_pos_allvis_timestrat.R
created a list of all events casecrossover dataframes

@@ -1,0 +1,107 @@
+# ------------------------------------------------------------------------------
+# Title: Asthma place of service time-stratified case-crossover using all visits
+# Author: Ryan Gan
+# Date Created: 2018-09-18
+# ------------------------------------------------------------------------------
+
+# Description: This script subsets the asthma claims to specific service places
+# and creates time-stratified case-crossover dataframes
+
+
+# library ----
+library(tidyverse)
+library(case.crossover)
+library(parallel)
+
+# asthma place of service ----
+
+# read in place of service csv file
+place_of_service <- read_csv("./data/health/2013-oregon_pos.csv")
+
+pos_subset <- c("11", "20", "21", "22", "23", "41")
+
+# read in asthma icd9 vector
+load("./data/health/outcome_list.RData")
+
+asthma_icd9 <- outcome_icd9_list$asthma
+
+# read in asthma claims
+asthma_claims <- read_csv('./data/health/2013-oregon_asthma_fireseason_cohort.csv',
+                          col_types = cols(.default = "c")) %>% 
+  # filter to just primary diagnosis or saba fill
+  filter(visit_type != 'dx_asthma_not_primary')%>% 
+  # filter to date of PM2.5 data (wildfire season)
+  filter(fromdate >= "2013-05-01" & fromdate <= "2013-09-30") %>% 
+  # filter to oregon
+  filter(STATE == "OR") %>%
+  # filter to places of service
+  filter(pos_simple %in% c('Ambulance', 'Emergency Room Hospital', 
+                           'Inpatient Hospital', 'Office', 'Outpatient Hospital',
+                           'Pharmacy', 'Urgent Care')) %>% 
+  # group by person id and date
+  group_by(personkey, service_place, fromdate) %>% 
+  filter(row_number()==1) %>% 
+  mutate(date = as.Date(fromdate), 
+         ZIPCODE = as.character(ZIP))
+
+
+# count of each place of service for each patient
+service_utilization <- asthma_claims %>% 
+  group_by(personkey, pos_simple) %>% 
+  summarize(count = n())
+
+# write event counts
+event_stats <- service_utilization %>% 
+  group_by(pos_simple) %>% 
+  summarize(n = n(), total_vis = sum(count), mean_vis = mean(count), 
+            med_vis = median(count), min_vis = min(count), max_vis = max(count))
+
[…]

### 2018-09-18 r_scripts/data_wrangle/asthma_pos_allvis_timestrat.R
updated case cross list creation.

@@ -46,15 +46,15 @@ asthma_claims <- read_csv('./data/health/2013-oregon_asthma_fireseason_cohort.cs
 
 
 # count of each place of service for each patient
-service_utilization <- asthma_claims %>% 
-  group_by(personkey, pos_simple) %>% 
-  summarize(count = n())
-
-# write event counts
-event_stats <- service_utilization %>% 
-  group_by(pos_simple) %>% 
-  summarize(n = n(), total_vis = sum(count), mean_vis = mean(count), 
-            med_vis = median(count), min_vis = min(count), max_vis = max(count))
+# service_utilization <- asthma_claims %>% 
+#   group_by(personkey, pos_simple) %>% 
+#   summarize(count = n())
+# 
+# # write event counts
+# event_stats <- service_utilization %>% 
+#   group_by(pos_simple) %>% 
+#   summarize(n = n(), total_vis = sum(count), mean_vis = mean(count), 
+#             med_vis = median(count), min_vis = min(count), max_vis = max(count))
 
 # write_csv(event_stats, './data/health/asthma_saba_event_sum_stats.csv')
 
@@ -81,7 +81,7 @@ clusterExport(cl, c("asthma_list"),
 start <- Sys.time()
 
 #asthma_cc_pos_list <- parLapply(cl, asthma_list, function(x){
-asthma_cc_pos_list <- lapply(asthma_list[[1]], function(x){
+asthma_cc_pos_list <- lapply(asthma_list, function(x){
     pos_df <- x %>% 
     # find the first observation and first claim line of each person 
     group_by(clmid) %>%

### 2018-09-18 bash_scripts/asthma_cc_sub.sh
updated bash script for job submission

@@ -16,5 +16,5 @@ cd /home/ryangan/local_git_repo/oregon_wildfire/
 export R_LIB=~/R/x86_64-pc-linux-gnu-library/3.3
 
 # run rscript
-Rscript --vanilla ./r_scripts/data_wrangle/asthma_pos_timestrat.R
+Rscript --vanilla ./r_scripts/data_wrangle/asthma_pos_allvis_timestrat.R
 

### 2018-09-18 r_scripts/data_wrangle/asthma_pos_allvis_timestrat.R
updated casecross list.

@@ -15,16 +15,6 @@ library(parallel)
 
 # asthma place of service ----
 
-# read in place of service csv file
-place_of_service <- read_csv("./data/health/2013-oregon_pos.csv")
-
-pos_subset <- c("11", "20", "21", "22", "23", "41")
-
-# read in asthma icd9 vector
-load("./data/health/outcome_list.RData")
-
-asthma_icd9 <- outcome_icd9_list$asthma
-
 # read in asthma claims
 asthma_claims <- read_csv('./data/health/2013-oregon_asthma_fireseason_cohort.csv',
                           col_types = cols(.default = "c")) %>% 
@@ -80,17 +70,17 @@ clusterExport(cl, c("asthma_list"),
 # start time
 start <- Sys.time()
 
-#asthma_cc_pos_list <- parLapply(cl, asthma_list, function(x){
-asthma_cc_pos_list <- lapply(asthma_list, function(x){
-    pos_df <- x %>% 
-    # find the first observation and first claim line of each person 
-    group_by(clmid) %>%
-    arrange(fromdate, line) %>%
-    ungroup() 
+asthma_cc_pos_list <- parLapply(cl, asthma_list, function(x){
+#asthma_cc_pos_list <- lapply(asthma_list, function(x){
+  pos_df <- x %>% 
+      # find the first observation and first claim line of each person 
+      group_by(clmid) %>%
+      arrange(fromdate, line) %>%
+      ungroup() 
     
-    ts_df <- time_stratified(data = pos_df, id = "personkey", 
-      covariate = c("patid", "gender", "age", "MSA", "ZIP", "pos", "dx1",
-                  "service_place"), 
+    ts_df <- time_stratified(data = x, id = "clmid", 
+      covariate = c("personkey", "gender", "age", "MSA", "ZIP", 
+                    "pos", "dx1", "service_place"), 
       admit_date = "fromdate", 
       start_date = "2013-05-01", end_date = "2013-09-30", interval = 7)
 })

### 2018-09-24 r_scripts/analysis/asthma_pos_sensitivity_results.Rmd
Oregon sensitivity analyses.

---
title: "Oregon Sensitivty Analysis"
author: "Ryan Gan"
date: "9/19/2018"
output: html_document
---

```{r setup, include=F}
library(tidyverse)
```

## Introduction

This document contains results for sensitivity analyses of the time-stratified case-crossover designs in the Oregon paper.

In the paper, our original method was to identify the first observations/visits for each patient for asthma care utilization. We then created time-stratified case-crossover dataframes for each asthma care utilization outcome, where the referent period was over the entire study period from May 1st to September 30th 2013. This scenario results in an independence of events across subjects.

There was concern on how limiting the analyses to only the first visit within the time period may result in biased estimates. Based on this concern, I have produced some summary statistics of the estimated number of visits/events per patient during the study period as well as two sensitivity analyses: including all events with a referent period over the entire study period, and including all events with a referent period of only the same month as the event. These scenario assumes independence across events and does not really account for multiple events per subject.

### Table of number of persons, events, and summary statistics of events.

```{r read_stats, include=F}
visits <- read_csv('../../data/health/asthma_saba_event_sum_stats.csv')
```

For some of the more severe outcomes like ED oor ambulance, the numbers between patients and number of events are pretty similar. With inpatient hosptial admissions, I'm worried I might have double counted claims. I tried my best to get to a single claim, but it's possible the same event was billed on multiple days. I think I'd need to account for length of stay to get this to a single event. I am not as worried about this with other places of service since they generally don't require overnight stay.

```{r sumstats_tab, echo = F, message=F}
knitr::kable(visits, caption = 'Patients and number of events')
```

## Sensitivity Results

Table of sen

```{r new_results, include = F}
sen_results <- read_csv('../../data/health/asthma_care_results_sensitivity.csv') %>% 
  select(-n_events)
```

### Table of Sensitivity Results

Table has results for all events per patient (assuming independence between events) with refrent period for the entire wildfire season and just for the month of the event.

```{r results_tab, echo = F}
knitr::kable(sen_results, caption = 'Sensitivity Results')
```

### Paper Results

These are the results in the paper where only the first event per subject is included.

```{r old_results, echo = F}
pos <- unique(sen_results$pos)
estimate <- c(1.259, 1.089, 1.072, 1.065, 1.050, 1.077, 0.809)
conf.low <- c(0.752, 1.043, 0.995, 1.029, 1.038, 1.065, 0.552)
conf.high <- c(2.107, 1.136, 1.154, 1.103, 1.063, 1.088, 1.187)

old_results <- data.frame(pos, estimate, conf.low, conf.high) %>% 
  mutate(scenario = 'First Event: Wildfire Season')

knitr::kable(old_results, caption = 'Paper Results')
```

### Plot of Results Side by Side

I'm excluding ambulance again since the monthly referent period is extremely unstable.

```{r append_results, echo=F}
plot_results <- sen_results %>% 
  mutate(scenario = paste0('All Events: ', ref_period)) %>% 
  select(-term, -ref_period) %>% 
  bind_rows(old_results) %>% 
  filter(pos != 'Ambulance') %>% 
  mutate(pos = case_when(pos == 'Emergency Room Hospital' ~ 'Emergency Department', 
                         pos == 'Pharmacy' ~ 'SABA Fill',
                         TRUE ~ pos))

ggplot(data=plot_results, aes(x=scenario, y = estimate, color = scenario, group = scenario)) +
  geom_point() +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), width = 0.2) +
  geom_hline(yintercept = 1, linetype=2) +
  facet_wrap(~pos, scales = 'free_y') +
  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
  xlab("Outcome") +
  # plot theme
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        strip.text = element_text(size = 10),
        # axis element (tilt)
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 0.7),
        #axis.title.x = element_blank(),
        #axis.ticks.x = element_blank(),
        axis.title.y = element_text(angle = 90),
        legend.position = 'bottom')
```

In regard to the sensitivity of my decision to use the first visit for each patient in the time series to avoid the scenario for accounting for multiple events in the same time period, I would say some results are more robust than others. Particuarly emergency department and urgent care estimates are pretty much the same despite the event case definition and referent period. Some thoughts on this are that these events are not as common compared to the other asthma care measures, and for peeople that may have multiple events, it's not more than 2.

As for SABA fills, office visits, and outpatient hospital visits, I see some difference here between the estimates. I would say it's more likely that some people have multiple events, and thus a greater chance of events occuring on referent dates. I would still say that it's these estimates are some what robust in that at least the odds ratio is in the same direction.

Inpatient hosptial visits changes the most and I think it's because it's possible that new claims for the same outcome may occur on the next day when someone is admitted. I would argue that this is effectively the same event, but it was difficult to split out these events. I would say probably the most 'accurate' is probably the first visit scenario since it does not count the same events twice. This can be fixed, but I will need to go back in to the original data and fully process each claim and event and figure out how to not count events that overlap in date admitted and date dischared.

### Identifying Overlap in Events

```{r asthma_cohort, include=F}
asthma_cohort <- read_csv("./data/health/2013-oregon_asthma_fireseason_cohort.csv")
```

Limiting to unique claims and visits first.

```{r unique_vis}
# this will get me to unique asthma visits
unique_vis <- asthma_cohort %>% 
  # filter by asthma visits and to paid claims to get rid of denied claims
  # for the same event
  filter(visit_type != 'dx_asthma_not_primary') %>% 
  filter(clmstatus == 'P') %>% 
  group_by(personkey, clmid, line, fromdate) %>% 
  slice(which.min(line)) %>% 
  group_by(personkey, clmid, fromdate) %>% 
  filter(row_number()==1)
```

For each unique patient and outcome, find the day of first admission and the number of additional events.

```{r}
first_vis <- unique_vis %>% 
  arrange(personkey, pos_simple, fromdate) %>% 
  group_by(personkey, pos_simple) %>% 
  filter(row_number()==1) %>% 
  filter(pos_simple %in% c('Ambulance', 'Emergency Room Hospital', 
                           'Inpatient Hospital', 'Office', 'Outpatient Hospital',
                           'Pharmacy', 'Urgent Care')) %>% 
  select(personkey, clmid, fromdate, pos_simple) %>%
  mutate(first_vis_day = lubridate::wday(fromdate, label = T)) %>% 
  rename(first_vis_date = fromdate)
```

Find additional admission dates for each subject and the days they were admitted.

```{r}
other_vis <- unique_vis %>% 
  ungroup() %>% 
  # filter by place of service
  filter(pos_simple %in% c('Ambulance', 'Emergency Room Hospital', 
                           'Inpatient Hospital', 'Office', 'Outpatient Hospital',
                           'Pharmacy', 'Urgent Care')) %>% 
  # find other visit day
  mutate(other_vis_day = lubridate::wday(fromdate, label = T)) %>% 
  select(personkey, pos_simple, other_vis_day) %>% 
  group_by(personkey, pos_simple, other_vis_day) %>% 
  summarize(n = n()) %>% 
  spread(other_vis_day, n)
```

Visit type.

```{r}
visit_counts <- first_vis %>% 
  left_join(other_vis, by = c('personkey', 'pos_simple'))
```

```{r}
glimpse(first_vis)
```

```{r}
check <- unique_vis %>% filter(pos_simple == 'Ambulance' & personkey == '8686197')
```

Check that each claim is only one count.

```{r claim_count, include = F}
check_count <- unique_vis %>% 
  group_by(personkey, clmid) %>% 
  summarize(n = n())
```

Check person with multiple inpatient asthma hosptializations. I have a feeling that the todate is not distinguishing the same event. They are the same as the fromdate.

```{r}
check <- filter(unique_vis, personkey == '10243608') %>% 
  select(personkey, clmid, line, fromdate, todate, paydate, visit_type, pos_simple, metroarea)
```

Checking inpatient todate.

```{r inpatient_length_count}
inpat_count <- unique_vis %>% 
  filter(pos_simple == 'Inpatient Hospital') %>% 
  mutate(days_stay = todate  - fromdate) %>% 
  select(personkey, clmid, line, fromdate, todate, paydate, 
         visit_type, pos_simple, metroarea, days_stay) %>% 
  group_by(days_stay) %>%
  summarize(n = n())
```

```{r}
ggplot(inpat_count, aes(x=days_stay, y=n)) +
  geom_point()
```

### 2018-09-27 r_scripts/analysis/asthma_pos_sensitivity_results.Rmd
Changed some labels.

@@
   mutate(scenario = paste0('All Events: ', ref_period)) %>% 
+  filter(scenario != 'All Events: Month') %>% 
   select(-term, -ref_period) %>% 
@@
                          pos == 'Pharmacy' ~ 'SABA Fill',
-                         TRUE ~ pos))
+                         TRUE ~ pos)) %>% 
+  rename(Scenario = scenario)
 
-ggplot(data=plot_results, aes(x=scenario, y = estimate, color = scenario, group = scenario)) +
+ggplot(data=plot_results, aes(x=Scenario, y = estimate, color = Scenario, group = Scenario)) +
   geom_point() +
@@
   facet_wrap(~pos, scales = 'free_y') +
-  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
+  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in WFS PM"[2.5]))) +
   xlab("Outcome") +
@@
   group_by(personkey, pos_simple, other_vis_day) %>% 
+  summarize(n = n()) #%>% 
+  #spread(other_vis_day, n)
+```
+
+Summary stats per person
+```{r}
+sum_stats <- other_vis %>% 
+  group_by(personkey, pos_simple) %>% 
   summarize(n = n()) %>% 
-  spread(other_vis_day, n)
+  group_by(pos_simple) %>% 
+  summarize(n_person = n(), n_visit = sum(n))
+```
+
+```{r}
+persons <- asthma_cohort %>% 
+  group_by(personkey) %>% 
+  summarize(n = n())
 ```

### 2019-03-23 r_scripts/analysis/asthma_place_of_service_analysis.Rmd
updated figure and results

@@
 ```{r age_sameday}
+
 age_results <- asthma_pos_list %>% 
@@
     # complete case
-    sex_df <- df %>% filter(!is.na(geo_smk10)) %>% 
+    age_df <- df %>% filter(!is.na(geo_smk10)) %>% 
       # nest strata
@@
    filter(n_events > 30) %>%
-   filter(conf.high < 5)
+   filter(conf.high < 5) %>% 
+  mutate(strata = paste0(strata, ' Yrs'))
 
@@
   rename(odds_ratio = estimate, lower95 = conf.low, upper95 = conf.high) %>% 
-  mutate(strata = case_when(strata == "Total" ~ "All Strata",
+  mutate(strata = case_when(strata == "Total" ~ "All Subjects",
                             TRUE ~ strata),
@@
           strata = factor(strata, 
-            levels = c("All Strata", "Female", "Male", "Age < 15", 
-                       "Age 15-65", "Age > 65"))) %>% 
+            levels = c("All Subjects", "Female", "Male", "Age < 15 Yrs", 
+                       "Age 15-65 Yrs", "Age > 65 Yrs"))) %>% 
     # set order of service place
@@
 ```{r plot}
+scaleFUN <- function(x) sprintf("%.2f", x)
+
 plot <- ggplot(results, aes(x = strata, y = odds_ratio)) +
@@
   geom_hline(yintercept = 1, linetype=2) +
-  facet_wrap(~service_place, scales = "free_y", ncol = 2) +
+  facet_wrap(~service_place, scales = 'free_y', ncol = 2) +
+  scale_y_continuous(label = scaleFUN) + 
   ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in WFS PM"[2.5]))) +
-  xlab("Stratum") +
+  xlab("All Subjects and Specific Subgroups") +
    # plot theme
@@
     # axis element (tilt)
-    axis.text.x = element_text(angle = 45, hjust = 0.9, vjust = 0.7),
+    axis.text.x = element_text(angle = 45, hjust = 0.95, vjust = 0.95),
     #axis.title.x = element_blank(),

### 2019-03-23 r_scripts/analysis/msa_asthma_timeseries.Rmd
updated figure and results

@@
 ```{r saba_df}
-saba <- read_csv("./data/health/2013-oregon_casecross_saba.csv") %>% 
+saba <- read_csv("../../data/health/2013-oregon_casecross_saba.csv") %>% 
   rename(identifier = personkey, ZIP = ZIPCODE) %>% 

### 2019-09-28 r_scripts/analysis/asthma_place_of_service_analysis.Rmd
updated some results and text.

@@
 
-I don't really like this plot because of the wide intervals. I also don't think there is really any evidence of effect modification by strata.
-
-I think I'll just present the all strata estimate.
-
-```{r total_results}
-# filter to total results
-total_results <- results %>% 
-  filter(strata == "All Strata")
-
-# plot total results
-total_plot <- ggplot(total_results, aes(x = service_place, y = odds_ratio, 
-                                        group = service_place)) +
-  geom_point(color = "#7f00ff") + 
-  geom_errorbar(aes(ymin=lower95, ymax=upper95), color = "#7f00ff",width = 0.2) +
-  geom_hline(yintercept = 1, linetype=2) +
-  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
-  xlab("Place of Service") +
-   # plot theme
-    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    # axis element (tilt)
-    axis.text.x = element_text(angle = 45, hjust = 0.9, vjust = 0.7),
-    #axis.title.x = element_blank(),
-    #axis.ticks.x = element_blank(),
-    axis.title.y = element_text(angle = 90))
-
-total_plot
-
-#ggsave("./Figs/total_pos_results_plot.tiff", total_plot, 
-#       units="in", width=6, height=4, dpi=300)
-```
-
 I will probably color-code this.
@@
 ```{r medford}
+names(asthma_pos_list)
+head(asthma_pos_list$`20`$geo_smk_pm)
 # excluding ambulance since it won't run for medford
-medford_result <- asthma_pos_list[c(1,3:5,7)] %>% 
+medford_result <- asthma_pos_list[c(1:5,7)] %>% 
   map_dfr(., function(df){
@@
     df <- df %>% filter(!is.na(geo_smk10)) %>% 
-      filter(metroarea == "Medford")
+      filter(metroarea == "Medford") %>% 
+      # create binary cutoff of >15
+      mutate(smk15 = ifelse(geo_smk_pm > 15, 1, 0))
     # model
-    mod <- clogit(outcome ~ geo_smk10 + wrf_temp + strata(identifier), 
+    mod <- clogit(outcome ~ smk15 + wrf_temp + strata(identifier), 
                   data = df)
@@
     estimates <- broom::tidy(mod) %>% 
-      filter(term == "geo_smk10") %>% 
+      filter(term == "smk15") %>% 
       dplyr::select(term, estimate, conf.low, conf.high) %>% 
@@
             levels = c("SABA Fill", "Office", "Outpatient Hospital", 
-                       "Emergency Room", "Inpatient Hospital")))
+                       "Emergency Room", "Inpatient Hospital", 
+                       "Urgent Care Facility")))
 
@@
 
-msa_result <- asthma_pos_list[c(1,3:5,7)] %>% 
+msa_result <- asthma_pos_list[c(1:5,7)] %>% 
   map_dfr(., function(df){
@@
 Hence, it's probably picking up some elevated PM~2.5~ days.
-
-## Next Steps/Thoughts
-
-- Have Jingyang focus the comparison of monthly rates by MSA to outcomes of SABA monthly fill rate, office visit, and inpatient visit (this might be interesting).
-
-- Is it worth evaluating a cumulative distributed lag effect of a couple days based on Ana Rappold's new study with CVD?
-
-- Should we try a further refinement of classifying smoke exposure, or even not transform PM~2.5~ and really focus on the Medford vs.
-other MSAs and just make the argument that any elevated PM~2.5~ during this time in Oregon is likely driven by smoke (particularly in Medford)?
